### 2017-06-27 paper/inExDMA.tex
add the paper directory

\documentclass[sigconf,anonymous,review]{acmart}
\usepackage[normalem]{ulem}

\usepackage{amssymb} % needed for: various symbols
\usepackage{amsmath} % needed for: math symbols
% http://tex.stackexchange.com/questions/229355/algorithm-algorithmic-algorithmicx-algorithm2e-algpseudocode-confused
\usepackage{algorithm} % this used as an environment for the algorithm code (linke figure and table)
\usepackage{algorithmicx} % this is used to write the pseudo code
\usepackage[noend]{algpseudocode} % this is for the layout of the pseudo code
\usepackage{tikz} % needed for: tikz pics
% \usepackage[inline]{enumitem}

\usepackage{graphicx}  
\usepackage{subfigure}

\begin{document}

\title{VIPOS: Virtualized Intermittently Powered Runtime Environment}

% \author{
% 	\IEEEauthorblockN{Amjad Yousef Majid, Przemys{\l}aw Pawe{\l}czak}
% 	\IEEEauthorblockA{TU Delft, Mekelweg\,4, 2628\,CD Delft, NL\\
% 	\{a.y.majid, p.pawelczak\}@tudelft.nl}
% }

\maketitle

\section{abstract}
Enabling battery-free devices is a mandatory step towards the environment friendly IoT. However, removing these large energy reservoirs (the batteries) makes the devices operate on an intermittent power supply, which make sustaining any long computation very challenging. correspondingly, there are two proposed approaches to enable long-running computations on intermittenly powered devices (IPDs): (i) Checkpointing based approach, where the volatile state of a program is frequently saved to the non-volatile memory and; (ii)  Task based approach, where a programmer splits the code into small idempotent modules. The Task based approach shows a better performance than the checkpointing approach. However, the task based method requires that the energy to execute any task must not exceed the maximum limit of the energy buffer. This limitation makes this solution a customized one that suits a particular energy buffer size.

VIPOS enables code portability by means of virtualizing the task based model. If a code is written for a very small energy buffer, VIPOS is able to dynamically merge these tasks and constructs a bigger virtual task to improve the performance, reduce energy consumption and provides mean for code portability.

\section{Introduction}
\label{sec:intro}
	Intermittently powered devices (IPDs) are battery-less devices that utilize the ambient energy to sense, compute and, communicate. For example, Wireless Identification and Sensing platform WISP \cite{WISP} uses the RF signal power to drive its computation and communication. Because of the reliance of TPDs on intermittent power supply, the harvested energy when it is available, their programs are susceptible to a very frequent power interrupts, in the order of tens of millisconds~\cite{}. Therefore, these devices require a different software execution model that complies with the nature of a discontinuous power supply.

	The intermittent (discontinuous) execution model defines a program execution as cumulative discrete process. The main difference between the intermittent and the conventional (or continuous) execution models is that, a power failure is seen by the continuous model as an \emph{exception} that may reset the progress of a program to its beginning. Whereas, in intermittent execution a power failure is regarded as a temporary \emph{pause} to the execution that may result in some progress degradation. Generally we can classify the intermittent execution model into: 
	\begin{itemize}
		\item \emph{Sequential Execution Model}:
			Under this model a program is seen as one big idempotent operational region that has one common context. Generally, The progress of the program is saved and updated by means of checkpointing---where all the program context (e.g. CPU registers, the stack and the global variables) is saved to a non-volatile memory. Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos, HarvOS, Hibernus}. The benefit of this model is that it does not require code modification by a programmer. However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{xxxx}; and (iii) it restricts the IPDs to run only a single application~\cite{InOS}. 
		\item \emph{Modular Intermittent Execution Model}:
			At the heart of this model is the concept of an idempotent task. The idempotent task is C function that does not have arguments and does not return a value. This task uses a well defined interface to interact with the non-volatile memory. Therefore, it tolerates arbitrary number of power interrupts. This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an IPD by interleaving their tasks~\cite{InOS}. However, it obviously requires code modification---for example, if an algorithm is written according to the continuous execution model it has to be splitted, by a programmer, into small tasks to run under the modular intermittent execution model.
	\end{itemize}

Despite that the superiority of the Modular Intermittent Execution model, it is still a static approach that completely depends on a programmer's estimation which is mostly result in a sub-optimal code devision. Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.

% Energy source

% The intermittent execution is cumulative discrete process. The intermittent execution model is only able to execute a few number of instructions, as compared to the conventional (continuous) model, before its progress is terminated. Therefore, the intermittent execution model adapts an execution progress state saving mechanism. This mechanism is realized either by means of a checkpoint~\cite{}, where all the context of the program is saved into non-volatile memory. Or

% This progress state saving mechanism is normally injected into a program either by a compiler~\cite{}. A compiler injects trigger pointers to checkpoint (save) the context of the program into the non-volatile memory.  or it is added by a programmer~\cite{chain}. [compare these two approaches]...

% We define virtualization within the context of intermittent execution as utilizing the volatile memory instead of non-volatile when the intermittent execution model attempts to access the non-volatile memory.

\subsection{Problem Statement}
How to approach the optimal task size given an unknown energy buffer size and environment conditions? 
	 \begin{enumerate}
		 \item How to eliminate the need for the hardware support while enabling the Modular Intermittent Execution model to dynamically adapt its task size?
		 \item How to enable code portability under the Modular Intermittent execution model? 
		 \item How to reduce the energy consumption of executing a task? 
	\end{enumerate}

\subsection{Why not Adapting Current Approaches}
Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes.

	\begin{figure}[t]
	    \centering
	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
	    \caption{Task flow control of Chain}
	    \label{fig:DynamicChain}
	\end{figure}

\noindent\textbf{Chain: Tasks and Channels for Reliable Intermittent Programs } \\
Some of the main problems in modifying Chain to support dynamic task's size are:

\begin{itemize}
	\item \emph{Multi-versioning system:} Chain defines for each task its own input that is not shared with other tasks. This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables. 
	\item \emph{Direct FRAM system:} Chain tasks directly access FRAM which is more energy expensive Than SRAM. As result of being multi-versioning system that directly interacts with FRAM, Chain has a significant energy consumption overhead. 
	\item \emph{Irreversible tasks transitions:} According to the Chain programming model, each task calls the next one. The transition is done through functions calls and manually clearing the stack, to prevent the stack overflow problem. This approach make all the transitions firm ones which prevent task merging.
	\item \emph{All channels are required:} Fig.~\ref{fig:DynamicChainSeq} shows that if the task control flow is a circular with no branches then task merging is possible. Moreover, all the intra-merged-tasks channel can be committed to SRAM instead of FRAM to save energy. However, this execution path is only a special. If the general case (see Fig.~\ref{fig:DynamicChainRan}) is considered then leaving out the intra-merged-task channels might result in data inconsistency---If an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready which result in an incorrect execution.

\end{itemize}

\noindent\textbf{Alpaca: Intermittent Execution without Checkpoints} \\
Some of the main problems in modifying Alpaca to support dynamic task's size are:

\begin{itemize}
	\item \emph{Privatization:} The core concept of Alpaca is privatization. Privatization depends completely on detecting the variables that have a Write-after-read dependency within the scope of a task. Therefore, if tasks are on-demand merged the scope of these dependencies are changed and the static analysis to privatize these variables is not valid anymore and data consistency can not be preserved.

	\item \emph{Irreversible tasks transitions:} Similar to Chain problem.

\end{itemize}

\noindent\textbf{Virtualizing Modular Intermittent Execution Model}

		An Intermittently executed program, as seen by the modular intermittent execution model, is a chain of tasks with a firm transition from one to another. 
		Virtualizing this model means softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption. Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance. Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses. This can be done by having a temporary copy of a global variable in the volatile memory (similar to the catching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.

		In this paper we are introducing VIPOS a runtime library the implement the Virtualed Intermittent Execution Model. VIPOS adapts two methods for data protection. One relies on a virtual buffer and the other uses the Direct Memory Access (DMA). VIPOS is able to approach the optimal tasks devision during the runtime. VIPOS reduces an application execution time by XX and the energy by XX.

\subsection{Contributions}
	 \begin{enumerate}
		 \item We have developed an algorithm that is able to determine the size of a virtual task based on the history of the execution of an application. 
		 \item We enable efficient code portability by requiring that a real task to small and merging them during the execution.   
	\end{enumerate}

 \section{Preliminary Results}
 \label{sec:prelResults}

\begin{figure}[t]
    \centering
         \subfigure[The energy cost of FRAM/SRAM writes operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
         \subfigure[The energy cost of FRAM/SRAM read operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
    \caption{The energy cost of accessing volatile/non-volatile memory}
    \label{fig:framEnergy}
\end{figure}

Since the Intermittent execution models rely on the non-volatile memory to enable long-running operation, it is important to characterize the energy consumption of accessing this type of memory as compare to the volatile one.

Four applications are developed to measure the energy consumption of \emph{Accessing} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{xxx}. The interference free debugger (EDB)~\cite{xxx} is used as the energy measurement tool. The EDB probes the energy buffer before and after accessing the FRAM/SRAM 1600 times, This large number of FRAM access is used to increase the reliability of the results generated by the used measurement tool (e.g. reducing the effect of the quantization error).  The energy buffer is charged, using the EDB, to $\approx$ 2.45 volts \textbf{only} at the beginning of the execution and of the programs. The write operations were performed by writing literal values, random numbers, to the memory. However, a read operation must be followed by a write operation, therefore, we chose to write a read value to SRAM. Fig.~\ref{fig:framEnergy} shows that an FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We would like to mentioned that TI has already mentioned in this document~\cite{xxxx} that FRAM write access is more energy expensive than SRAM write access. However, TI does not quantify the energy overhead.}. Therefore, reducing FRAM access is desirable when developing  energy limited software solution

\section{VIPOS: Virtualized Intermittently Powered Runtime Environment}
VIPOS is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs. It is able, on-demand, to merge tasks to reduce the number of commits to the non-volatile memory.

\subsection{VIPOS: Data/Memory Protection Methods}
VIPOS data consistency preservation methods are designed with goal of minimal interaction between the CPU and FRAM to reduce energy consumption (see Section~\ref{sec:prelResults}). These data/memory protection methods enables VIPOS to tolerate arbirary number of power failures.

	\subsubsection{Virtualized Operational Buffer}

		\begin{algorithm}[t]
			\caption{Virtualized Operational Buffer}
			\label{algo:virtuBufWrite}
			\scriptsize
			%\small
			\begin{algorithmic}[1]
				\State $var \in \text{\{global variables\}} $ 
				\State \label{lst:virtuBufWrite:line:begin}\Call{Virtual Task()}{} 
				\While { \textit{executing} } \Comment{Execution stage}
					\State $var$  $\rightarrow$ \textsf{volatile buffer} \label{lst:virtuBufWrite:line:output}
					\If { $var$  in \textsf{ volatile buffer} }				\label{lst:virtuBufWrite:line:inputBegin}
							\State $var$  $\leftarrow$  \textsf{volatile buffer} 
						\Else 
							\State $var$  $\leftarrow$  \textsf{FRAM}		\label{lst:virtuBufWrite:line:inputEnd}
						\EndIf

					\If {power interrupts}
						\State back to \ref{lst:virtuBufWrite:line:begin}
					\EndIf
				\EndWhile

				\While{  $\textsf{volatile buffer}\not=\emptyset$  } \label{lst:virtuBufWrite:line:commitBegin} \Comment{ First phase commit}
					\State \textsf{volatile buffer} $\rightarrow$  \textsf{persistent buffer}
					\If {power interrupts}
						\State discard \textsf{persistent buffer}
						\State back to  \ref{lst:virtuBufWrite:line:begin} 
						\State   \label{lst:virtuBufWrite:line:commitEnd}
					\EndIf
				\EndWhile

				\While{ $\textsf{persistent buffer}\not=\emptyset$ } \label{lst:virtuBufWrite:line:SecCommitBegin} \Comment{Second phase commit}
					\State \textsf{persistent buffer} $\rightarrow$ FRAM 
					\If {power interrupts}
						\State Continue
					\EndIf
				\EndWhile 
				\State     \label{lst:virtuBufWrite:line:SecCommitEnd}
				\State return
			\end{algorithmic}
		\end{algorithm}

		Since accessing FRAM is more energy expencive and slower than accessing SRAM, preventing frequent access to FRAM (i.e. during looping operations) is desireable. Therefore, the first proposed protection method utilizes a volatile buffer that holds temporary all the outputs of a virtual task (see Algorithm~\ref{algo:virtuBufWrite} line~\ref{lst:virtuBufWrite:line:output}). Consequently, a virtual task must first attempt to read a global variable from the volatile buffer before trying to obtain the value from the non-volatile memory (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:inputBegin}-\ref{lst:virtuBufWrite:line:inputEnd}) to ensure a correct execution progress and the consistency of the memory. Once the execution of a virtual task is done, the first phase of the commit process is started by copying the volatile buffer to a persistent buffer (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:commitBegin}-\ref{lst:virtuBufWrite:line:commitEnd}). If the power is interrupted the persistent buffer must be discarded and the execution must start again from the beginning of the virtual task. In another words, the first phase commit must be performed atomically to preserve the consistency of the output of a virtual task. The second phase commit is a power failure immune precess that is responsible for distributing the global variables to their final locations and make the non-volatile memory consistent and synchronized with the computation progress (see Algorithm~\ref{algo:virtuBufWrite}lines~\ref{lst:virtuBufWrite:line:SecCommitBegin}-\ref{lst:virtuBufWrite:line:SecCommitEnd}).

	\paragraph{Virtualized Operational Buffer: Implementation}

		We implement a reference implementation of VIPOS that uses virtualized Operational buffer to protect the data against power failure. 
		\paragraph{Virtual Buffer}
			We realized the virtaulzed buffer as a volatile hashed table of linked lists. The hashing technique was chosen to reduce the buffer searching time and the linked lists are used to prevent data loss when there is a conflict between multiple variables---If two variables have the same hash value they will occupy the same cell, however, with linked list a new node will be created for each variable to resolve the conflict and protect the data. The hashing function is based on the observation that the virtual addresses of the memory cells have approximately a flat distribution over the memory addressing space and the fact that each entry, of the virtual table, has the address and the value of a variable. As such, by using the least significant bits as an index to access the virtual buffer the hash function distributes its inputs uniformly in the virtual buffer. Accordingly, the linked lists search time, on average, is reduced.

		\paragraph{Persistent Buffer}
			The persistent buffer is implemented as static array of tuples, where each tuple holds  the address and the value of a variable. This data structure is very suitable for the second phase commit where each element has to be committed to its finial location. However, the complexity of committing this buffer is of size $O(N)$, where $N$ equals the length of the buffer. Therefore, the size of the buffer can have dramatic effect on the performance of VIPOS---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious. On the other hand, if the size of the buffer is large VIPOS will experience a significant performance degradation. However, by observing the type of the information that this buffer holds, namely memory addresses, we can, on average, reduce the complexity of committing this buffer by defining the \emph{effective size} of the buffer to be the size of the buffer up to the first cell that holds an invalid memory address. Moreover, since this buffer is in FRAM which is much bigger than SRAM its size can be relatively big.

		\paragraph{Virtualized Operational Buffer: delay analysis}[initial results]

		\begin{figure}[t]
			\centering
			\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
			\caption{VIPOS based on virtualized Operational Buffer. It runs the decompression application with different buffer configurarions. EPB: Enhanced Persistent Buffer. NR: no read operation from the buffer. NRW: not read or write operations from the buffer.}
			\label{fig:virtOperationalBuf}
		\end{figure}

			To analyze the performance of VIPOS versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment. We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}. The results presented in Fig.~\ref{fig:virtOperationalBuf} show the VIPOS has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.

			All the experiments were done with a persistent buffer of size 100. However, the bar labeled with "EPB/16" in  Fig.~\ref{fig:virtOperationalBuf} shows the effect of the enhanced commit process that uses the buffer effective size instead of the buffer maximum size. We see that the execution time is reduced to $\approx$80 ms. Furthermore, we quantized the overhead of the read operations from the buffer, by eliminating this operation, and the bar labeled with "NR" shows that it cost  $\approx$10 ms. Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to VIPOS best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.

\subsubsection{ Direct Memory Access Based Protection Method (better name is needed)}
%
	\begin{figure}[t]
		\centering
         \subfigure[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize.eps} }
	     \subfigure[The energy needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
		\caption{Time and energy consumption of moving a block of data from SRAM to FRAM}
		\label{fig:dmaTimeEnergy}
	\end{figure}
%
	Since accessing/updating buffers is more costly than directly accessing FRAM, at least without specialized hardware~\cite{Clank}, we introduce the second method to apply the virtualization principle to the Modular Execution Model. This method utilizing the Direct Memory Access (DMA) Module~\cite{DMA}. As Fig.~\ref{fig:dmaTimeEnergy} shows that DMA is much more efficient in transferring a block of data than the conventional (pointer based) way from the energy and time perspectives.

	This method combines a number of techniques to preserve the persistence and consistency of the data. It backs up the status of the global variables in the non-volatile memory, and double buffers them to guarantee their consistency. Furthermore, on each power up it populates the global variables in the volatile memory from the most recent back up in the non-volatile memory to enable a faster task execution since the task does not need to interact with non-volatile memory which is generally slower and more energy expensive.

\subsection{VIPOS: The Virtualizing Engine}
\subsubsection{Power Interrupt Immune Scheduler}
It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures. VIPOS provides an API to enable a programmer to have a full control over the execution flow of the program, i.e. (un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.

\subsubsection{Task Merging Algorithms}

\paragraph{Fixed Virtual Task Size} [better name needed]

	\begin{algorithm}[t]
		\caption{Fixed virtual Task size}
		\label{algo:fixVirtTask}
		\scriptsize
		%\small
		\begin{algorithmic}[1]
			\State $VT \subset \text{\{VIPOS Tasks\}} $  \Comment{$VT:$ Virtual Task}
			\State VTS : VT size
			\State MVTS: maximum VT size
			\vspace{0.1cm}

			\While {$True$}
				\State $VT \leftarrow VT_{next}$
				\vspace{0.1cm}
				\While {execute $VT$} 
					\If { $\text{power failed twice}$ }				
							\State $VTS--$  
							\State $ MVTS = VTS $
						\EndIf
				\EndWhile

				\vspace{0.1cm}
				\If {$ \text{All tasks executed}$}
					\If{$VTS < MVTS$}
					\State $VTS++$
					\EndIf
				\EndIf
			\EndWhile
		\end{algorithmic}
	\end{algorithm}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.7\columnwidth]{figures/virtualTaskSize.eps}
	\caption{Size of the virtual task versus the execution time of a dummy application that contains 12 empty tasks.}
	\label{fig:virtualTaskSize}
\end{figure}

VIPOS is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
Fig.~\ref{fig:virtualTaskSize} show the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).

Remark: A less obvious benefit of visualization is that it can enable a secure computation. By increasing the size of the virtual task, the energy in the super-capacitor will not be enough to finish the execution of the virtual task. Therefore, without the interrogator being close to the intermittently powered device and charging rate is not negligible as compared to the discharging rate this computation will not be finished. As such, virtualization can add a layer of security to the computation.

\section{Results}
We compare the performance of VIPOS against a state-of-the-art intermittent execution approach. For this purpose, we used two different applications: (i) A data decompression application, which utilizes Huffman decoding technique to decompress 100 byte of data.; (ii) A discrete Fourier transform (DFT) application that uses two different resolution (4 and 8 bites) to analyze a randomly generated signal.

\subsection{Continuous Power Supply}

\begin{figure}[t]
	\centering
	\includegraphics[width=0.7\columnwidth]{figures/executionTime.eps}
	\caption{Time to complation of VIPOS versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
	\label{fig:IPOSPerformance}
\end{figure}

Fig.~\ref{fig:IPOSPerformance} shows that VIPOS requires a shorter execution time to execute an application under the MIEM. The size of the virtual task that VIPOS uses equals One real task, in other words, no visualization is applied.

\subsection{VIPOS Operational Buffer Versus VIPOS DMA}[place holder]
			\begin{figure}[t]
				\centering
				\includegraphics[width=0.7\columnwidth]{figures/ipos_virtualizedBuffer_overhead.eps}
				\caption{Virtualized Oprational buffer overhead.}
				\label{fig:virtualBuffer}
			\end{figure}

\subsection{Intermittent Power Supply}
...

\bibliographystyle{plain}
\bibliography{references}

\end{document}

### 2017-06-27 paper/presentation/class/tikz_defs.tex
add the paper directory

\def\BayStationPosition{(0,0)}
%
\def\SensorOnePosition{(-1.2,	-0.3)}
\def\SensorTwoPosition{(-1.5,	0.3)}
\def\SensorThreePosition{(-1.3,	1.1)}
\def\SensorFourPosition{(-0.3,	1.6)}
\def\SensorFivePosition{(0.3,	2.2)}
\def\SensorSixPosition{(0.7,	1.7)}
\def\SensorSevenPosition{(1.7,	1.8)}
\def\SensorEightPosition{(1.6,	0.8)}
\def\SensorNinePosition{(2.1,	0.3)}
\def\SensorTenPosition{(1.8,	-0.4)}
\def\SensorElevenPosition{(1.0,	-0.7)}
\def\SensorTwelvePosition{(1.0,	-1.4)}
\def\SensorThirteenPosition{(0.0,	-2.2)}
\def\SensorFourteenPosition{(-0.3,	-1.4)}
\def\SensorFifteenPosition{(-1.3,	-1.9)}

\tikzset
{
	SensorNodeStyle/.style =
	{
		% -- shape properties --
		circle,									% shape
%		rounded corners	= 3mm,					% kind of corner (and radius of the roundness)
%		minimum height	= 0.15\textwidth,		% | minimum size of the node
%		minimum width	= 0.9\textwidth,		% |
		minimum size	= 4.5mm,				%
		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		%
		% -- colours properties --
		% filling: [ trasparent | monocolored | shaded]; decomment what you prefer
%		%										% transparent (all commented)
		fill			= green!10,				% monocolored
%		top color		= white,				% | filling of the node
%		bottom color	= red!50!black!20,		% |
		text			= black,				% colour of the fonts
		draw			= black,				% colour of the border
		%
		% -- fonts --
		font			= \scriptsize,				% shape of the font (or dimension, like \tiny)
		text centered,							% text alignment [text centered | text badly centered | text justified | text ragged | text badly ragged]
		inner xsep		= 0mm,					% minimum distance between text and borders along x dimension
		inner ysep		= 0mm,					% minimum distance between text and borders along y dimension
		text height		= 0.2cm,
		text depth		= 0.12cm,
	}
}

\tikzset
{
	BayStationStyle/.style =
	{
		% -- shape properties --
		rectangle,									% shape
		rounded corners	= 1mm,					% kind of corner (and radius of the roundness)
%		minimum height	= 0.15\textwidth,		% | minimum size of the node
%		minimum width	= 0.9\textwidth,		% |
		minimum size	= 4.5mm,				%
		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		%
		% -- colours properties --
		% filling: [ trasparent | monocolored | shaded]; decomment what you prefer
%		%										% transparent (all commented)
		fill			= black!30,				% monocolored
%		top color		= white,				% | filling of the node
%		bottom color	= red!50!black!20,		% |
		text			= black,				% colour of the fonts
		draw			= black,				% colour of the border
		%
		% -- fonts --
		font			= \scriptsize,				% shape of the font (or dimension, like \tiny)
		text centered,							% text alignment [text centered | text badly centered | text justified | text ragged | text badly ragged]
%		text height		= 1mm,					% ! minimum size of the text	% NOT WORKING
%		text depth		= 1mm,					% !
		inner xsep		= 0mm,					% minimum distance between text and borders along x dimension
		inner ysep		= 0mm					% minimum distance between text and borders along y dimension
	}
}

\tikzset
{
	FittingStyle/.style =
	{
		% -- shape properties --
		shape = ellipse,							% shape
%		rounded corners	= 1mm,					% kind of corner (and radius of the roundness)
%		minimum height	= 0.15\textwidth,		% | minimum size of the node
%		minimum width	= 0.9\textwidth,		% |
%		minimum size	= 4.5mm,				%
%		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		%
		% -- colours properties --
		% filling: [ trasparent | monocolored | shaded]; decomment what you prefer
%		%										% transparent (all commented)
%		fill			= black!30,				% monocolored
%		top color		= white,				% | filling of the node
%		bottom color	= red!50!black!20,		% |
%		text			= black,				% colour of the fonts
		draw			= red,				% colour of the border
		%
		% -- fonts --
%		font			= \scriptsize,				% shape of the font (or dimension, like \tiny)
%		text centered,							% text alignment [text centered | text badly centered | text justified | text ragged | text badly ragged]
%		text height		= 1mm,					% ! minimum size of the text	% NOT WORKING
%		text depth		= 1mm,					% !
		inner xsep		= 0mm,					% minimum distance between text and borders along x dimension
		inner ysep		= 0mm					% minimum distance between text and borders along y dimension
	}
}

\tikzset
{
	GenericNodeStyle/.style =
	{
		% -- shape properties --
		shape = rectangle,							% shape
		rounded corners	= 3mm,					% kind of corner (and radius of the roundness)
		minimum height	= 1cm,					% | minimum size of the node
		minimum width	= 2cm,					% |
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		%
		% -- colours properties --
		% filling: [ trasparent | monocolored | shaded]; decomment what you prefer
%		%										% transparent (all commented)
		fill			= green!10!white,				% monocolored
%		top color		= white,				% | filling of the node
%		bottom color	= red!50!black!20,		% |
%		text			= black,				% colour of the fonts
		draw			= green,				% colour of the border
		%
		% -- fonts --
%		font			= \scriptsize,				% shape of the font (or dimension, like \tiny)
%		text centered,							% text alignment [text centered | text badly centered | text justified | text ragged | text badly ragged]
%		text height		= 1mm,					% ! minimum size of the text	% NOT WORKING
%		text depth		= 1mm,					% !
		inner xsep		= 3mm,					% minimum distance between text and borders along x dimension
		inner ysep		= 3mm					% minimum distance between text and borders along y dimension
	}
}

\tikzset
{
	NormalNodeStyle/.style =
	{
		shape = circle,							% shape
		minimum size	= 20,					%
		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		text			= black,				% colour of the fonts
		draw			= black,				% colour of the border
		font			= \small,				% shape of the font (or dimension, like \tiny)
		text centered,							% text alignment
		inner xsep		= 0,					% minimum distance between text and borders along x dimension
		inner ysep		= 0						% minimum distance between text and borders along y dimension
	}
}

\tikzset
{
	BridgeNodeStyle/.style =
	{
		circle,									% shape
		minimum size	= 20,					%
		rotate			= 0,					% angle of rotation
		scale			= 1.0,					% scaling factor
		thick,									% thickness of the border
		fill			= black!30,				% monocolored
		text			= black,				% colour of the fonts
		draw			= black,				% colour of the border
		font			= \small,				% shape of the font (or dimension, like \tiny)
		text centered,							% text alignment 
		inner xsep		= 0,					% minimum distance between text and borders along x dimension
		inner ysep		= 0						% minimum distance between text and borders along y dimension
	}
}

\tikzstyle{sNormalBlockStyle} =
[
	draw,
	rectangle,
	rounded corners	= 0.1cm,
	fill			= blue!20,
	minimum height	= 3em,
	minimum width	= 6em,
]

\tikzstyle{sSumBlockStyle} =
[
	shape			= circle,
	draw,
	fill			= blue!20,
]

\tikzstyle{sArrowsStyle} =
[
	thick,
	color	= black,
	-latex
]

\tikzstyle{sLinesStyle} =
[
	thick,
	color	= black,
	-
]

\tikzstyle{sTextBlockStyle} =
[
	draw,
	rectangle,
	drop shadow,
	rounded corners	= 0.1cm,
	fill			= blue!10,
	thick,
	inner xsep		= 0.2cm,		% minimum distance between text and borders along x dimension
	inner ysep		= 0.2cm			% minimum distance between text and borders along y dimension
]

\tikzset
{
	AlertText/.style =
	{
		starburst,						% shape
		drop shadow,
		starburst points = 30,		% | shape parameters
		starburst point height = 0.4cm,	% |
		rotate			= 0,			% angle of rotation
		thick,							% thickness of the border
		fill			= tudOrange!10,		% monocolored
		text			= black,	% colour of the fonts
		draw			= tudTurquoise,			% colour of the border
		font			= \large,		% shape of the font (or dimension, like 
		%\tiny)
		text centered,					% text alignment 
	}
}

\tikzset
{
	WarningText/.style =
	{
		rectangle,						% shape
		rounded corners	= 0.6cm,		%
		drop shadow,
		rotate			= 0,			% angle of rotation
		scale			= 1.0,			% scaling factor
		thick,							% thickness of the border
		fill			= tudCyan!10,		% monocolored
		text			= black,	% colour of the fonts
		draw			= tudCyan,			% colour of the border
		font			= \large,		% shape of the font (or dimension, like 
		%\tiny)
		text centered,					% text alignment 
		inner xsep		= 0.4cm,		% minimum distance between text and 
		%borders along x dimension
		inner ysep		= 0.4cm			% minimum distance between text and 
		%borders along y dimension
	}
}

\tikzfading	% DO NOT CHANGE THE ORDER OF THE COLORS otherwise it will not work
[
	name			= middle,
	top color		= transparent!100,
	bottom color	= transparent!100,
	middle color	= transparent!00,
]

\newcommand{\AlertBox}[2]{
	\begin{tikzpicture}[overlay]
	\node [AlertText] at (#1) {#2};
	\end{tikzpicture}
}

\newcommand{\WarningBox}[2]{
	\begin{tikzpicture}[overlay]
	\node [PropertyText] at (#1) {#2};
	\end{tikzpicture}
}

### 2017-06-27 paper/presentation/images/chain-virtualization-II.tex
add the paper directory

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning} 
\usetikzlibrary{fit}

\tikzset{
	%Define standard arrow tip
	>=stealth',
	%Define style for boxes
	punkt/.style={
		circle,
		rounded corners,
		draw=black, very thick,
		text width=2.5em,
		minimum height=1em,
		text centered},
	% Define arrow style
	pil/.style={
		->,
		thick,
		shorten <=2pt,
		shorten >=2pt,}
}

\begin{document}

	\begin{tikzpicture}[node distance=1cm, auto,]

	\pgfdeclarelayer{bg}    % declare background layer
	\pgfsetlayers{bg,main}  % set the order of the layers (main is the standard layer)

	\node[punkt] (task1) {Task (1)};
	\node[punkt, below =0.7cm of task1] (task2) {Task (2)};
	\node[punkt, below =0.7cm of task2] (task3) {Task (3)};
	\node[punkt, below =0.7cm of task3] (task4) {Task (4)};
	\node[punkt, right =0.9cm of task2] (task5) {Task (5)};
	\node[punkt, right =2cm of task4] (task6) {Task (6)};

	\path[every node/.style={font=\sffamily\small}]
	(task1.west) edge [->, dotted, thick, bend right] node {Channel (1)} (task2.west)
	(task1.east) edge [->, thick,red, bend left] node {Channel (4)} (task5.north)
	(task2.west) edge [->, dotted, thick, bend right] node {Channel (2)} (task3.west)
	(task2.east) edge [->, thick,red, bend left] node {Channel (5)} (task6.north)
	(task3.west) edge [->, very thick, blue, bend right] node {Channel (3)} (task4.west)
	(task6.west) edge [->, thick,green, bend left] node[xshift=1.6cm,yshift=0.5cm] {Channel (6)} (task3.east);

	\begin{pgfonlayer}{bg}    % select the background layer
		\node[draw,thick,dotted,red,minimum width=7em, fill=blue!10, fit=(task1) (task2) (task3)] (Virtualized) {};
		\draw[<->] ([xshift=-1.7cm]Virtualized.north) --node[sloped, anchor=south] {Virtualization Window} ([xshift=-1.7cm]Virtualized.south);
	\end{pgfonlayer}

	\end{tikzpicture}

\end{document}

### 2017-06-27 paper/presentation/images/chain-virtualization.tex
add the paper directory

--- 2017-06-27 paper/presentation/images/chain-virtualization-II.tex
@@
 	\node[punkt, below =0.7cm of task3] (task4) {Task (4)};
-	\node[punkt, right =0.9cm of task2] (task5) {Task (5)};
-	\node[punkt, right =2cm of task4] (task6) {Task (6)};
 
@@
 	(task1.west) edge [->, dotted, thick, bend right] node {Channel (1)} (task2.west)
-	(task1.east) edge [->, thick,red, bend left] node {Channel (4)} (task5.north)
 	(task2.west) edge [->, dotted, thick, bend right] node {Channel (2)} (task3.west)
-	(task2.east) edge [->, thick,red, bend left] node {Channel (5)} (task6.north)
-	(task3.west) edge [->, very thick, blue, bend right] node {Channel (3)} (task4.west)
-	(task6.west) edge [->, thick,green, bend left] node[xshift=1.6cm,yshift=0.5cm] {Channel (6)} (task3.east);
+	(task3.west) edge [->, very thick, blue, bend right] node {Channel (3)} (task4.west);
 
@@
 		\node[draw,thick,dotted,red,minimum width=7em, fill=blue!10, fit=(task1) (task2) (task3)] (Virtualized) {};
-		\draw[<->] ([xshift=-1.7cm]Virtualized.north) --node[sloped, anchor=south] {Virtualization Window} ([xshift=-1.7cm]Virtualized.south);
+		\draw[<->] ([xshift=1.5cm]Virtualized.north) --node[sloped, anchor=south] {Virtualization Window} ([xshift=1.5cm]Virtualized.south);
 	\end{pgfonlayer}

### 2017-06-27 paper/presentation/preamble.tex
add the paper directory

% Slide Preamble
\documentclass[svgnames]{beamer}
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage{calc}
\usepackage[absolute,overlay]{textpos}
\usepackage{graphicx}
\usepackage{subfig}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{listings}
\usepackage{standalone}
\usepackage{comment}
\usepackage{bbding}
\usepackage[plain]{algorithm}
\usepackage{caption}
\usepackage[noend]{algpseudocode} % needed for: algorithm pseudocode
\usepackage{MnSymbol,wasysym}

\setbeamertemplate{navigation symbols}{} % remove navigation symbols
\mode<presentation>{\usepackage{class/beamerthemetud}}

% BIB SETTINGS
\usepackage[backend=bibtex,firstinits=true,maxnames=30,maxcitenames=20,url=false,style=authoryear]{biblatex}
\bibliography{references}
\setlength\bibitemsep{0.3cm} % space between entries in the reference list
\renewcommand{\bibfont}{\normalfont\scriptsize}
\setbeamerfont{footnote}{size=\tiny}
\renewcommand{\cite}[1]{\footnote<.->[frame]{\fullcite{#1}}}

\usepackage{tikz,tcolorbox}
\usetikzlibrary{decorations.pathreplacing} 
\usetikzlibrary{arrows,decorations.pathreplacing,decorations.pathmorphing,decorations.footprints,
	fadings,calc,trees,mindmap,shadows,decorations.text,patterns,positioning,shapes,matrix,fit}
\usepackage{stmaryrd}
\input{class/tikz_defs}

### 2017-06-27 paper/presentation/presentation.tex
add the paper directory

\include{preamble}

\begin{document}

\include{title}

{
	\setbeamertemplate{footline}{\usebeamertemplate*{footline}}
}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Virtualization --- Motivation}}
	\begin{alertblock}{\textbf{FRAM Write}}
		\begin{itemize}
			\item Almost twice more expensive than SRAM write.
		\end{itemize}	
	\end{alertblock}

	\centering
	\includegraphics[scale=0.5]{images/fram_write.eps}

	\begin{block}{\textbf{Problem Statement}}
		\begin{itemize}
			\item How to \textbf{postpone} FRAM access (commit) operations to reduce energy overhead? 	
			\begin{itemize}
				\item The gained energy will be used for \textbf{more computation} 
				\item E.g. more tasks will be executed in task-based programs --- in turn, the total program execution time will be reduced	
				\begin{itemize}
					\item Portability: different super-capacitor sizes will be used efficiently	
				\end{itemize}
			\end{itemize}		
		\end{itemize}	
	\end{block}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Task-based Control Flow}}

	\begin{itemize}
		\item e.g. Chain\cite{colin2016chain} 
		\begin{itemize}
			\item no checkpointing -- keep track of the current task 
			\item pipelined channels -- maintain the consistency of 
			FRAM
			\item seperate inputs from outputs -- idempotency
		\end{itemize}

	\end{itemize}

	\begin{columns}

		\column{.40\textwidth}
		\centering
		\includegraphics[scale=0.10]{images/chain.png}\\

		\column{.56\textwidth}
		\begin{block}{\textbf{Drawbacks}}
			\begin{itemize}
				\item outputs are committed \textbf{at the end} of each task
				\begin{itemize}
					\item \textbf{Commit} operation, in turn FRAM access (in particular write), is \textbf{expensive}
				\end{itemize}
				\item \textbf{multi-versioning} of variables --- memory overhead
				\begin{itemize}	
					\item increased number of FRAM operations
				\end{itemize}
			\end{itemize}	
		\end{block}

	\end{columns}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Details - Virtualization of Chain -I}}

	\begin{columns}

		\column{.70\textwidth}
		\begin{itemize}
			\item If all output channels are connected to and the transition path is only to the \textbf{next task}, virtualization is \textbf{easy}! 
			\item Example: no need to commit \texttt{Channel (1)} and \texttt{Channel (2)} since their values are already consumed to output \texttt{Channel (3)}.
		\end{itemize}

		\column{.40\textwidth}
		\center
		\includegraphics[scale=0.50]{images/chain-virtualization.pdf}
	\end{columns}

		\begin{block}{\textbf{Strategy}}
			\begin{itemize}
				\item Postpone commit inside the \textbf{virtualization window} using corresponding \textbf{soft-channels} in SRAM
				\item \textbf{Hard-commit} only the channel input of the last task 
				\item If power-interrupted, go to the first task of the window
			\end{itemize}	
		\end{block}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Details - Virtualization of Chain -II}}

	\begin{columns}

		\column{.77\textwidth}
		\begin{itemize}
			\item What if the channel connections and transition paths are \textbf{divergent}? 
			\begin{itemize}
				\item Example: \texttt{Channel (3)}, \texttt{Channel (4)} and \texttt{Channel (5)} \textbf{must} committed to FRAM ---\textbf{hard-channels}
				\item \texttt{Channel (1)} and \texttt{Channel (2)} are in SRAM---\textbf{soft-channels}
				\item \texttt{Channel (6)} should be \textbf{read} from FRAM---hard-channel
			\end{itemize}
		\end{itemize}

		\column{.35\textwidth}
		\includegraphics[scale=0.50]{images/chain-virtualization-II.pdf}
	\end{columns}

	\begin{block}{\textbf{Drawbacks}}
		\begin{itemize}
			\item The static \textbf{task-graph} and the \textbf{dynamic} transition path should be used at \textbf{run-time} to decide which channels need to be hard or soft
		\end{itemize}	
	\end{block}

	\begin{alertblock}{\textbf{Not Feasible}}
	\begin{itemize}
		\item Due to \textbf{loops} in task graph, e.g. self-channels, and \textbf{multi-versioning} ruins virtualization.
	\end{itemize}	
\end{alertblock}

\end{frame}
%---------------------------------------------------------------------------

\begin{frame}{\textbf{Case Study - Virtualization of ALPACA -I}}

		\centering
		\includegraphics[scale=0.3]{images/alpaca-privatization.png} \\
Only variables with W-A-R dependencies are protected in ALPACA.

	\begin{block}{\textbf{Challanges}}
		\begin{itemize}
			\item Variables without W-A-R dependencies are directly read from/written to FRAM 
			\item \textbf{Solution:} Use \textbf{SRAM buffer} and commit later---Reduced FRAM access
		\end{itemize}	
	\end{block}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Case Study - Virtualization of ALPACA -II}}

	\begin{columns}

	\column{.65\textwidth}
	\begin{itemize}
		\item Merged tasks might introduce new W-A-R dependencies ---at \textbf{run-time}!
		\item How to \textbf{dynamically} keep-track of the privitization of the W-A-R dependencies? ---\textbf{too much overhead!}
		\begin{itemize}
			\item Alpaca does not privatize \texttt{x}, \texttt{y} and \texttt{z}; they are directly written to FRAM. 
			\item We need software implementation of Clank :)
		\end{itemize}
	\end{itemize}

	\column{.45\textwidth}
	\includegraphics[scale=0.7]{images/alpaca-war.pdf}
\end{columns}

\begin{alertblock}{\textbf{Conclusions}}
	\begin{itemize}
		\item We need to \textbf{eliminate} privatization so that tasks can operate on SRAM efficiently.
		\item Not Feasible with ALPACA, requires fundamental design changes.
	\end{itemize}	
\end{alertblock}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Virtualization with SRAM Buffer-I}}
	\begin{columns}

	\column{.41\textwidth}
	\begin{itemize}
		\item Version \textbf{task-shared} variables in FRAM
		\item A \textbf{fixed-sized} SRAM buffer for virtual operations
		\item Two-phase commit at the end of the \textbf{virtual task}.
	\end{itemize}

		\column{.70\textwidth}
\includegraphics[scale=0.5]{images/sram-buffer.pdf}	
\end{columns}

	\begin{block}{\textbf{Overhead}}
		\begin{itemize}
			\item \texttt{get} and \texttt{set} operations require \textbf{search} in SRAM cache
			\item Example: The presented task in Alpaca will have 4 FRAM write operations - We will have 6 + SRAM search overhead!
		\end{itemize}	
	\end{block}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Virtualization with SRAM Buffers-II}}
	\begin{columns}

		\column{.50\textwidth}
		\begin{itemize}
			\item We did not observe any benefit, even worse!
			\item CPU is the main burden in energy consumption --- Search operations dominate FRAM write!
		\end{itemize}

		\column{.50\textwidth}
		\includegraphics[scale=0.3]{images/sram-sw-buffer.pdf}	
	\end{columns}

	\begin{alertblock}{\textbf{Problem}}
	\begin{itemize}
		\item How to eliminate SRAM Cache search operation? 
		\item How to minimize commit overhead?
	\end{itemize}	
	\end{alertblock}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Virtualization with DMA Support-I}}
	\begin{columns}

		\column{.60\textwidth}
		\begin{itemize}
			\item Exploit hardware that is found in almost every CPU architecture
			\item Copying block of words from/to SRAM-FRAM and FRAM-FRAM is very \textbf{cheap} and \textbf{scalable} with DMA!
		\end{itemize}

		\column{.42\textwidth}
		\includegraphics[scale=0.25]{images/dmaSize.eps}\\			
		\includegraphics[scale=0.25]{images/energyConsumptionDMA_SW.eps}	
	\end{columns}

	\begin{alertblock}{\textbf{Problem}}
		\begin{itemize}
			\item How to eliminate SRAM Cache search operation? 
			\item How to minimize commit overhead?
		\end{itemize}	
	\end{alertblock}

\end{frame}
%---------------------------------------------------------------------------

\begin{frame}{\textbf{Virtualization with DMA Support-II}}
	\begin{columns}

		\column{.41\textwidth}
		\begin{itemize}
			\item Version all \textbf{task-shared} variables in FRAM
			\item Bring all to SRAM buffer for virtual operations
			\item Two-phase commit at the end of the \textbf{virtual task}.
		\end{itemize}

		\column{.70\textwidth}
		\includegraphics[scale=0.5]{images/dma.pdf}	
	\end{columns}

	\begin{block}{\textbf{Feasibility}}
		\begin{itemize}
			\item Perform all these operations using DMA!
			\item All task-shared variables have a copy in SRAM, \textbf{no search}!
		\end{itemize}	
	\end{block}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Virtualization with DMA Support-III}}
	\center
	\includegraphics[scale=0.27]{images/executionTime-eps-converted-to}	
	\includegraphics[scale=0.27]{images/virtualTaskSize-eps-converted-to}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{Virtualization with DMA Support-Limitations-I}}

	\begin{itemize}
			\item The size of the task-shared variables, in turn the size of SRAM Cache is bounded with the \textbf{SRAM size} in the platform.
			\begin{itemize}
				\item The number of channels is directly limited with the FRAM size in Chain
				\item The size of the commit list in Alpaca, i.e. the number of variables with W-A-R dependency inside a task, is limited with the FRAM.
			\end{itemize}
	\end{itemize}

\centering
\begin{block}{}
	\centering
	\begin{tabular}{ l c | c | c }
	& VIPOS & Chain & Alpaca \\
	\hline
	\textbf{Limited with} & SRAM-size & FRAM-size & FRAM-size
\end{tabular}
\end{block}

\end{frame}
%---------------------------------------------------------------------------

\begin{frame}{\textbf{Virtualization with DMA Support-Limitations-II}}

\begin{itemize}

		\item The cost of start-up and commit operations increase with \textbf{the size of the task-shared variables}
		\begin{itemize}
			\item But this comes with the elimination of the search operations in SRAM cache and selective commit
			\item DMA cost increases linearly but very slowly.
		\end{itemize}
\end{itemize}	
All these limitations are paid in order to benefit from virtualization.

\end{frame}
%---------------------------------------------------------------------------

\begin{frame}{\textbf{What is next?}}

	\begin{itemize}
		\item How to get rid of flush in/out all task-shared global variables?
		\begin{itemize}
			\item Very difficult, how?
		\end{itemize} 
		\item Development and evaluation of several task-merging algorithms, i.e. \textbf{ virtualization scheduler}
		\begin{itemize}
			\item This will be one of the core sections of the paper!
		\end{itemize}
		\item Automatic "taskification"
		\begin{itemize}
			\item This will be one of the core sections of the paper!
		\end{itemize}
	\end{itemize}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{The structure of the article-I}}

	\begin{itemize}
		\item \textbf{Introduction}
		\begin{itemize}
			\item Motivation---FRAM access, in turn commit cost, need for the virtualization
			\item Problem statement
			\item Challenges---Why it is difficult to modify state of the art solutions to support virtualization?
			\item Our solution and contributions---Sketch of our solution and attractive results
		\end{itemize}
		\item \textbf{Challenges}
	\begin{itemize}
		\item Intermittent programming with checkpointing and task-based control flow
		\item Stress the problem of dynamic merging of the tasks, its feasibility with the state of the art solutions, e.g. Chain, InOS, Alpaca...
	\end{itemize}

		\item \textbf{Virtualization: Dynamic Merging of Tasks}
		\begin{itemize}
			\item Introduce the concept, requirements
			\item Design objectives
			\item Introduce the infrastructure here
		\end{itemize}
	\end{itemize}

\end{frame}

%---------------------------------------------------------------------------

\begin{frame}{\textbf{The structure of the article-II}}

	\begin{itemize}
		\item \textbf{Virtualization Scheduler: The Algorithms}
		\begin{itemize}
			\item Introduce the task-merging strategies that work on the infrastructre in the previous section
			\item Analysis of the algorithms
			\item Maybe some simulations
		\end{itemize}
		\item \textbf{Implementation and Evaluation}
		\begin{itemize}
			\item Introduce the implementation details
			\item Test-bed setup
			\item Experimental results
			\item Simulations
		\end{itemize}
		\item \textbf{Related Work}
		\item \textbf{Conclusions}
	\end{itemize}

\end{frame}

\end{document}

### 2017-06-27 paper/experiments/energyMeasurements/DMA_Energy_cost/ext/edb-console/edb.py

@@ -1,0 +1,588 @@
+#!/usr/bin/python3
+
+import sys
+import traceback
+import os
+import time
+import atexit
+import readline # loading this causes raw_input to offer a rich prompt
+import argparse
+import re
+
+from pyedb import edb
+
+parser = argparse.ArgumentParser(
+            description="EDB Console")
+parser.add_argument('--command', '-c',
+            help="Run the given command and exit")
+parser.add_argument('--stdio', '-o',
+            help="File to which to pipe std I/O data relayed from target (default: console)")
+args = parser.parse_args()
+
+monitor = None
+active_mode = False
+
+if args.stdio is not None:
+    CONSOLE_FILE = open(args.stdio, "w")
+else:
+    CONSOLE_FILE = sys.stdout
+
+def to_int(s):
+    if s.startswith("0x"):
+        return int(s, 16)
+    else:
+        return int(s)
+
+def match_keyword(part, words):
+    match = None
+    for word in words:
+        if word.startswith(part):
+            if match is not None:
+                raise Exception("Ambiguous keyword: " + part)
+            match = word
+    return match
+
+def print_interrupt_context(context):
+    print("Interrupted: %r " % context.type, end='')
+    if context.type == "ASSERT":
+        print("line: %r" % context.id, end='')
+    elif context.type == "BOOT":
+        pass
+    else:
+        print("id: %r" % context.id, end='')
+    if context.saved_vcap is not None:
+        print(" Vcap_saved = %.4f" % context.saved_vcap)
+    else:
+        print()
+
+def print_watchpoint_event(event):
+    print("Watchpoint: id: %r time: %.6f s Vcap = %.4f" % \
+            (event.id, event.timestamp, event.vcap))
+
+def init_watchpoint_log(fout):
+    fout.write("id,timestamp,vcap\n")
+def log_watchpoint_event(fout, event):
+    fout.write("%u,%.6f,%.4f\n" %  (event.id, event.timestamp, event.vcap))
+
+class Command:
+    def __init__(self, name, parser, handler):
+        self.name = name
+        self.parser = parser
+        self.handler = handler
+        self.description = parser.description
+
+CMD_HANDLER_FUNC_PREFIX = "cmd_"
+CMD_PARSER_FUNC_PREFIX = "parser_"
+
+class CommandNotFoundException(Exception):
+    def __init__(self, name):
+        self.name = name
+
[…]

### 2017-06-27 paper/experiments/energyMeasurements/DMA_Energy_cost/ext/edb-server/ext/libedbserver/src/delay.h

@@ -1,0 +1,10 @@
+#ifndef DELAY_H
+#define DELAY_H
+
+#include "config.h"
+
+#include <stdint.h>
+
+void delay_kcycles(uint16_t delay_kcycles);
+
+#endif // DELAY_H

### 2017-06-27 paper/experiments/energyMeasurements/DMA_Energy_cost/ext/edb-server/ext/libedbserver/src/include/libedbserver/codepoint.h

@@ -1,0 +1,32 @@
+#ifndef CODEPOINT_H
+#define CODEPOINT_H
+
+#include <stdint.h>
+#include <stdbool.h>
+
+#include "host_comm.h"
+
+extern uint16_t code_energy_breakpoints; // exposed for comparator ISR
+
+void set_external_breakpoint_pin_state(uint16_t bitmask, bool state);
+
+unsigned toggle_breakpoint(breakpoint_type_t type, unsigned index,
+                           uint16_t energy_level, comparator_ref_t cmp_ref,
+                           bool enable);
+unsigned toggle_watchpoint(unsigned index, bool enable, bool vcap_snapshot);
+
+void enable_watchpoints();
+void disable_watchpoints();
+
+void watchpoints_start_stream();
+void watchpoints_stop_stream();
+
+void init_watchpoint_event_bufs();
+void send_watchpoint_events();
+
+void handle_codepoint(unsigned index);
+
+typedef void (watchpoint_callback_t)(unsigned index, uint16_t vcap);
+void edb_set_watchpoint_callback(watchpoint_callback_t *cb);
+
+#endif // CODEPOINT_H

### 2017-06-27 paper/experiments/energyMeasurements/DMA_Energy_cost/ext/edb-server/ext/libedbserver/src/include/libedbserver/host_comm.h

@@ -1,0 +1,268 @@
+#ifndef HOST_COMM_H
+#define HOST_COMM_H
+
+#include <stdint.h>
+
+/**
+ * @defgroup    UART_PROTOCOL  UART interface
+ * @brief       UART usage
+ * @details     The UART message structure looks like this:\n
+ *              | Byte                 | Name                  | Description                            |
+ *              | -------------------- | --------------------- | -------------------------------------- |
+ *              | 0                    | UART identifier       | Identifies the source of the message   |
+ *              | 1                    | Message descriptor    | Identifies the message                 |
+ *              | 2                    | Length                | Length of the upcoming data            |
+ *              | 3 to (2 + length)    | Data                  | Message data (optional)                |
+ *
+ * @{
+ */
+
+/**
+ * @brief Indentifies the source of message to be the debugger
+ */
+#define UART_IDENTIFIER_USB                     0xF0
+
+/**
+ * @brief       Message descriptors sent from the computer to
+ *              the WISP monitor over the USB interface.
+ */
+typedef enum {
+    USB_CMD_SENSE                           = 0x01, //!< Get ADC12 reading of Vcap
+    USB_CMD_STREAM_BEGIN                    = 0x02, //!< Start streaming ADC12 readings continuously
+    USB_CMD_STREAM_END                      = 0x03, //!< Stop stream ADC12 readings continuously
+    USB_CMD_SET_VCAP                        = 0x04, //!< Inject charge until Vcap has this ADC12 reading, length should be 2
+    USB_CMD_SET_VBOOST                      = 0x05, //!< Inject charge until Vboost has this ADC12 reading, length should be 2
+    USB_CMD_SET_VREG                        = 0x06, //!< Inject charge until Vreg has this ADC12 reading, length should be 2
+    USB_CMD_SET_VRECT                       = 0x07, //!< Inject charge until Vrect has this ADC12 reading, length should be 2
+    USB_CMD_RELEASE_POWER                   = 0x08, //!< Release the hold on power state now
+    USB_CMD_ENTER_ACTIVE_DEBUG              = 0x09, //!< Enter active debug mode by triggering the WISP's port interrupt
[…]

### 2017-06-27 paper/experiments/energyMeasurements/DMA_Energy_cost/ext/edb-server/ext/libedbserver/src/main.c

@@ -1,0 +1,1406 @@
+#include <msp430.h>
+#include <stdint.h>
+#include <stdbool.h>
+#include <stdlib.h>
+
+#include <libmsp/periph.h>
+#include <libmsp/clock.h>
+#include <libmsp/watchdog.h>
+#include <libedb/target_comm.h>
+
+#include "pin_assign.h"
+#include "dev_console.h"
+#include "host_comm.h"
+#include "host_comm_impl.h"
+#include "target_comm_impl.h"
+#include "adc.h"
+#include "uart.h"
+#include "i2c.h"
+#include "systick.h"
+#include "rfid.h"
+#include "rfid_decoder.h"
+#include "minmax.h"
+#include "config.h"
+#include "error.h"
+#include "main_loop.h"
+#include "params.h"
+#include "charge.h"
+#include "comparator.h"
+#include "codepoint.h"
+#include "tether.h"
+#include "interrupt.h"
+#include "sched.h"
+#include "delay.h"
+
+#ifdef CONFIG_PWM_CHARGING
+#include "pwm.h"
+#endif
+
+/* @brief State-keeping flags used by debug mode implementation
+ * @details These flags are stored in the MSB of the bitmask word that also
+ * stores debug mode feature flags (defined in libedb/target_comm.h).
+ * */
+#define DEBUG_MODE_NESTED           0x10 //!< debug mode within an active debug mode (e.g. assert in energy guard)
+
+/**
+ * @defgroup    LOG_DEFINES     Logging flags
+ * @brief       Flags to set in a bit mask to check which ADC readings to log
+ * @{
+ */
+#define LOG_VCAP                    0x01 //!< Logging Vcap
+#define LOG_VBOOST                  0x02 //!< Logging Vboost
+#define LOG_VREG                    0x04 //!< Logging Vreg
+#define LOG_VRECT                   0x08 //!< Logging Vrect
+#define LOG_VINJ					0x10 //!< Logging Vinj
+/** @} End LOG_DEFINES */
+
+#define STREAM_REPLY_MAX_LEN (1 /* num chans */ + ADC_MAX_CHANNELS * sizeof(uint16_t))
+
[…]

### 2017-06-27 paper/experiments/energyMeasurements/DMA_Energy_cost/ext/edb-server/ext/libmspsoftuart/src/uart.c

@@ -1,0 +1,268 @@
+/******************************************************************************
+ * Software UART example for MSP430.
+ *
+ * Stefan Wendler
+ * sw@kaltpost.de
+ * http://gpio.kaltpost.de
+ *
+ * ****************************************************************************
+ * The original code is taken form Nicholas J. Conn example:
+ *
+ * http://www.msp430launchpad.com/2010/08/half-duplex-software-uart-on-launchpad.html
+ *
+ * Origial Description from Nicholas:
+ *
+ * Half Duplex Software UART on the LaunchPad
+ *
+ * Description: This code provides a simple Bi-Directional Half Duplex
+ * Software UART. The timing is dependant on SMCLK, which
+ * is set to 1MHz. The transmit function is based off of
+ * the example code provided by TI with the LaunchPad.
+ * This code was originally created for "NJC's MSP430
+ * LaunchPad Blog".
+ *
+ * Author: Nicholas J. Conn - http://msp430launchpad.com
+ * Email: webmaster at msp430launchpad.com
+ * Date: 08-17-10
+ * ****************************************************************************
+ * Includes also improvements from Joby Taffey and fixes from Colin Funnell
+ * as posted here:
+ *
+ * http://blog.hodgepig.org/tag/msp430/
+ ******************************************************************************/
+
+#include <msp430.h>
+#include <stdbool.h>
+#include <stdint.h>
+
+#include <libmsp/periph.h>
+
+#include "printf.h"
+
+#define CONFIG_RX (defined(PORT_SOFTUART_RXD) && defined(PIN_SOFTUART_RXD))
+#define CONFIG_TX (defined(PORT_SOFTUART_TXD) && defined(PIN_SOFTUART_TXD))
+
+#if !CONFIG_TX
+#error RX-only configuration not supported
+#endif // !CONFIG_TX
+
+#define TIMER_SOFTUART CONCAT(TIMER_SOFTUART_TYPE, TIMER_SOFTUART_IDX)
+
[…]

### 2017-06-27 paper/experiments/energyMeasurements/DMA_Energy_cost/ext/libwispbase/src/wired/uart.c

@@ -1,0 +1,327 @@
+/**
+ * @file uart.c
+ * @author Aaron Parks -- Framework + TX logic
+ * @author Ivar in 't Veen -- RX logic
+ * @brief UART module for transmitting/receiving data using the USCI_A0 peripheral
+ */
+
+#include <msp430.h>
+#include "uart.h"
+#include "globals.h"
+
+/**
+ * State variables for the UART module
+ */
+typedef struct {
+    uint8_t isTxBusy; // Is the module currently in the middle of a transmit operation?
+    uint8_t* txPtr; // Pointer to the next byte to be transmitted
+    uint16_t txBytesRemaining; // Number of bytes left to send
+
+    uint8_t isRxBusy; // Is the module currently in the middle of a receive operation?
+    uint8_t* rxPtr; // Pointer to the next byte to be received
+    uint16_t rxBytesRemaining; // Maximum number of bytes left to receive
+} uart_sm_t;
+
+static volatile uart_sm_t UART_SM;
+
+/**
+ * Configure the eUSCI_A0 module in UART mode and prepare for UART transmission.
+ *
+ * @todo Currently assumes an 8MHz SMCLK. Make robust to clock frequency changes by using 32k ACLK.
+ */
+void UART_init(void) {
+
+    // Configure USCI_A0 for UART mode
+    UCA0CTLW0 = UCSWRST;                      // Put eUSCI in reset
+    UCA0CTLW0 |= UCSSEL__SMCLK;               // CLK = SMCLK
+
+    // Baud Rate calculation
+    // 8000000/(16*9600) = 52.083
+    // Fractional portion = 0.083
+    // User's Guide Table 21-4: UCBRSx = 0x04
+    // UCBRFx = int ( (52.083-52)*16) = 1
+
+//#define UART_BAUDRATE 9600
+#define UART_BAUDRATE 115200
+
+//#define UART_CLOCK 1000000
+//#define UART_CLOCK 4000000
+#define UART_CLOCK 8000000
+
+#if UART_CLOCK == 1000000
+#if UART_BAUDRATE == 9600
+    UCA0BR0 = 52;                             // 8000000/16/9600
+    UCA0BR1 = 0x00;
+    UCA0MCTLW = UCOS16 | UCBRF_1;
+#endif // UART_BAUDRATE
+#elif UART_CLOCK == 4000000
+#if UART_BAUDRATE == 9600
[…]

### 2017-06-27 paper/experiments/energyMeasurements/DMA_Energy_cost/python/DMA_SW_energy.py

@@ -1,0 +1,119 @@
+#!/home/amjed/anaconda3/bin/python
+import numpy as np 
+import matplotlib.pyplot as plt 
+from matplotlib import style
+import copy
+style.use('seaborn-muted')
+
+def remOutLiers(voltagesIn):
+	voltages = copy.deepcopy( voltagesIn )
+	voltages = np.array(voltages).astype(np.float) 
+	stdVolt_3 = np.std( voltages ) * 2
+	meanVolt = np.mean( voltages )
+	voltages = list(voltages)
+	# print('3 * std = ', stdVolt_3)
+	for i, volt in enumerate(voltages):
+		if np.abs(volt - meanVolt) > stdVolt_3:
+			print('del', voltages[i] )
+			del voltages[i] 
+
+	return voltages
+
+
+def voltList(file ):
+	"""
+		@input: file name 
+		@output: list of voltages
+
+		voltList extract the voltages values 
+		from a the EDB output
+	 """
+	voltages=[]
+	for line in open(file):
+		# get the voltages and remove the `\n` char
+		voltages.append( (line.split(',')[-1])[:-1] ) 
+	
+	return(voltages)
+
+def statistics(voltages, numLines):
+	""" 
+		@input: voltages list, number instructions between each voltage probes
+		@output: mean and standard deviation 
+	"""
+	diffs=[]
+	list_len = len(voltages)
+	for i in range(0,list_len, 2):
+		diffs.append( (( (float(voltages[i])/numLines - float(voltages[i+1])/numLines ) \
+								 **2) * 10e-6 * 0.5 ) )
+	diffsClean = remOutLiers(diffs)
+	return(np.mean(diffsClean), np.std(diffsClean))
+
+def plot(xAxis, voltMeans, symb, labels):
+	plt.rcParams['axes.linewidth'] =  2
+	plt.rcParams['axes.labelsize'] =  16
+	plt.rcParams['xtick.labelsize'] =  16
+	plt.rcParams['ytick.labelsize'] =  16
+	plt.rcParams['lines.linewidth'] =  2
+	f = plt.figure(figsize=(8,4))
+	i=0
+	for val in voltMeans:
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_read_cost/ext/edb-server/ext/libedbserver/src/include/libedbserver/uart.h

@@ -1,0 +1,443 @@
+#ifndef UART_H
+#define UART_H
+
+#include <stdint.h>
+
+#include <libedb/target_comm.h>
+#include <libmsp/clock.h>
+
+#define CONFIG_UART_CLOCK_FREQ CONFIG_SMCLK_FREQ
+
+// UART baudrate specification:
+//
+// Not UCOS16:
+//
+// N = SMCLK / BAUD
+// BR0 = LSB(floor(N))
+// BR1 = MSB(floor(N))
+// BRS = (floor(N) - N) * 8
+//
+// UCOS16:
+//
+// N = SMCLK / BAUD
+// BR0 = LSB(floor(N/16))
+// BR1 = MSB(floor(N/16))
+// BRF = (floor(N/16) - N/16) * 16
+// UCOS16 = 1
+
+#if CONFIG_UART_CLOCK_FREQ == 24576000
+
+#if CONFIG_USB_UART_BAUDRATE == 1000000
+
+#ifdef CONFIG_USB_UART_UCOS16
+
+// N/16 = 24.576
+#define CONFIG_USB_UART_BAUDRATE_BR0 0x01
+#define CONFIG_USB_UART_BAUDRATE_BR1 0x00
+#define CONFIG_USB_UART_BAUDRATE_BRF 9
+#define CONFIG_USB_UART_BAUDRATE_UCOS16 1
+
+#endif // CONFIG_USB_UART_UCOS16
+
+#elif CONFIG_USB_UART_BAUDRATE == 921600
+
+#ifdef CONFIG_USB_UART_UCOS16
+// N/16 = 1.66666...
+#define CONFIG_USB_UART_BAUDRATE_BR0 0x01
+#define CONFIG_USB_UART_BAUDRATE_BR1 0x00
+#define CONFIG_USB_UART_BAUDRATE_BRF 10
+#define CONFIG_USB_UART_BAUDRATE_UCOS16 1
+#endif // CONFIG_USB_UART_UCOS16
+
+#elif CONFIG_USB_UART_BAUDRATE == 576000
+
+#ifdef CONFIG_USB_UART_UCOS16
+// N/16 = 2.66666...
+#define CONFIG_USB_UART_BAUDRATE_BR0 0x02
+#define CONFIG_USB_UART_BAUDRATE_BR1 0x00
+#define CONFIG_USB_UART_BAUDRATE_BRF 10
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_read_cost/ext/edb-server/ext/libedbserver/src/params.h

@@ -1,0 +1,16 @@
+#ifndef PARAMS_H
+#define PARAMS_H
+
+#include <stdint.h>
+
+#include "host_comm.h"
+
+extern uint16_t param_test;
+extern uint16_t param_target_boot_voltage_dl;
+extern uint16_t param_target_boot_latency_kcycles;
+extern uint16_t param_num_watchpoint_events_buffered;
+
+return_code_t set_param(param_t param, uint8_t *buf);
+unsigned get_param(param_t param, uint8_t *buf);
+
+#endif

### 2017-06-27 paper/experiments/energyMeasurements/Fram_read_cost/python/fram_sram_read_energy_consumption.py

@@ -1,0 +1,87 @@
+#!/home/amjed/anaconda3/bin/python
+import numpy as np 
+import matplotlib.pyplot as plt 
+from matplotlib import style
+
+style.use('seaborn-muted')
+
+numberOfReadLines = 1600;
+
+#----Get the volatge values of the cap-----#
+fram_voltages=[]
+sram_voltages=[]
+fram_cntr=0
+sram_cntr=0
+for line in open("../outputs/fram.txt"):
+	# get the voltages and remove the `\n` char
+	fram_voltages.append( (line.split(',')[-1])[:-1] ) 
+	fram_cntr+=1
+
+for line in open("../outputs/sram.txt"):
+	# get the voltages and remove the `\n` char
+	sram_voltages.append( (line.split(',')[-1])[:-1] )
+	sram_cntr+=1
+
+#print(fram_voltages)	 
+#print(sram_voltages)
+#exit()
+
+
+#-----SRAM/FRAM write costs------#
+v_list_len= min(fram_cntr, sram_cntr)
+fram_diffs=[]
+sram_diffs=[]
+for i in range(0,v_list_len, 2):
+	fram_diffs.append( ( ( (float(fram_voltages[i])/numberOfReadLines )**2 - (float(fram_voltages[i+1])/numberOfReadLines ) **2 ) * 10e-6 * 0.5 ) ) 
+	sram_diffs.append( ( ( (float(sram_voltages[i])/numberOfReadLines )**2 - (float(sram_voltages[i+1])/numberOfReadLines ) **2 ) * 10e-6 * 0.5 ) ) 
+
+# print(fram_diffs)
+# print(sram_diffs)
+# exit()
+
+print (sum(fram_diffs) /len(fram_diffs))
+print (sum(sram_diffs) /len(sram_diffs))
+
+
+#------------Line fitting------------#
+fit_order=3
+xVals = range(len(fram_diffs))
+fram_z =  np.polyfit( xVals, fram_diffs, fit_order)
+sram_z =  np.polyfit( xVals, sram_diffs, fit_order)
+fram_estYVals = np.polyval(fram_z, xVals)
+sram_estYVals = np.polyval(sram_z, xVals)
+
+
+#--------------Plotting--------------#
+
+plt.rcParams['axes.labelsize'] =  12
+plt.rcParams['lines.linewidth'] =  1.5
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_read_cost/python/generate_array.py

@@ -1,0 +1,25 @@
+#!/usr/python3
+# this script is used to generate an array for the experiment of 
+# FRAm energy cost.
+
+import random 
+import numpy as np 
+
+arraySize = 400
+listLen =  1600
+
+indices = np.random.choice(arraySize,arraySize)
+
+nv_num_f = open("nv_num.txt", "w")
+vo_num_f = open("vo_num.txt", "w")
+
+read_nv_num_f = open("read_nv_num.txt", "w")
+read_vo_num_f = open("read_vo_num.txt", "w")
+
+for i in range(arraySize):
+	nv_num_f.write("nv_num[{}] = {};\n".format(i, random.randrange(65000) ))
+	vo_num_f.write("vo_num[{}] = {};\n".format(i, random.randrange(65000) ))
+
+for i in range(listLen):
+	read_nv_num_f.write ("vo_num2[{}] = nv_num[{}];\n".format(indices[i%arraySize], indices[ indices[i%arraySize] ]) )
+	read_vo_num_f.write ("vo_num2[{}] = vo_num[{}];\n".format(indices[i%arraySize], indices[ indices[i%arraySize] ]) )

### 2017-06-27 paper/experiments/energyMeasurements/Fram_read_cost/src/main.c

@@ -1,0 +1,2062 @@
+/*
+ Author: Amjad Yousef Majid
+ Date: 30/May/2017
+*/
+#include <msp430.h>
+#include <stdint.h>
+#include <stdbool.h>
+#include <stdlib.h>
+
+#include <libmsp/periph.h>
+#include <libmsp/clock.h>
+#include <libmsp/watchdog.h>
+#include <libmsp/gpio.h>
+
+
+#ifdef CONFIG_EDB
+#include <libedb/edb.h>
+#endif
+
+#ifdef TEST_EIF_PRINTF
+#include <libio/printf.h>
+#endif // TEST_EIF_PRINTF
+
+#include "pins.h"
+
+
+
+volatile unsigned int i, j;
+volatile unsigned int  vo_num[400] = {0};
+volatile unsigned int  vo_num2[400] = {0};
+
+
+static void init_hw()
+{
+    msp_watchdog_disable();
+    msp_gpio_unlock();
+    msp_clock_setup();
+}
+
+
+int main() {
+init_hw();  
+#ifdef CONFIG_EDB
+    edb_init();
+#endif
+
+vo_num[0] = 7833;
+vo_num[1] = 3516;
+vo_num[2] = 38908;
+vo_num[3] = 61578;
+vo_num[4] = 9481;
+vo_num[5] = 62372;
+vo_num[6] = 39678;
+vo_num[7] = 43210;
+vo_num[8] = 37565;
+vo_num[9] = 64062;
+vo_num[10] = 16336;
+vo_num[11] = 18089;
+vo_num[12] = 41906;
+vo_num[13] = 5503;
+vo_num[14] = 21267;
+vo_num[15] = 33938;
+vo_num[16] = 54792;
+vo_num[17] = 31088;
+vo_num[18] = 6550;
+vo_num[19] = 59163;
+vo_num[20] = 60391;
+vo_num[21] = 59576;
+vo_num[22] = 55305;
+vo_num[23] = 39295;
+vo_num[24] = 1764;
+vo_num[25] = 35078;
+vo_num[26] = 57155;
+vo_num[27] = 11349;
+vo_num[28] = 10011;
+vo_num[29] = 28523;
+vo_num[30] = 12918;
+vo_num[31] = 39206;
+vo_num[32] = 2777;
+vo_num[33] = 24431;
+vo_num[34] = 35971;
+vo_num[35] = 52107;
+vo_num[36] = 161;
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_read_cost/src/nv_main.c

@@ -1,0 +1,2063 @@
+/*
+ Author: Amjad Yousef Majid
+ Date: 30/May/2017
+*/
+#include <msp430.h>
+#include <stdint.h>
+#include <stdbool.h>
+#include <stdlib.h>
+
+#include <libmsp/periph.h>
+#include <libmsp/clock.h>
+#include <libmsp/watchdog.h>
+#include <libmsp/gpio.h>
+
+
+#ifdef CONFIG_EDB
+#include <libedb/edb.h>
+#endif
+
+#ifdef TEST_EIF_PRINTF
+#include <libio/printf.h>
+#endif // TEST_EIF_PRINTF
+
+#include "pins.h"
+
+#define __nv __attribute__((section(".nv_vars")))
+
+volatile __nv unsigned int nv_num[400] = {0} ;
+
+volatile unsigned int i, j;
+volatile unsigned int  vo_num2[400] = {0};
+
+
+static void init_hw()
+{
+    msp_watchdog_disable();
+    msp_gpio_unlock();
+    msp_clock_setup();
+}
+
+
+int main() {
+init_hw();  
+#ifdef CONFIG_EDB
+    edb_init();
+#endif
+
+nv_num[0] = 8545;
+nv_num[1] = 15671;
+nv_num[2] = 49007;
+nv_num[3] = 42610;
+nv_num[4] = 45885;
+nv_num[5] = 20828;
+nv_num[6] = 963;
+nv_num[7] = 27524;
+nv_num[8] = 51583;
+nv_num[9] = 3351;
+nv_num[10] = 33963;
+nv_num[11] = 8762;
+nv_num[12] = 10093;
+nv_num[13] = 60462;
+nv_num[14] = 8514;
+nv_num[15] = 11504;
+nv_num[16] = 17629;
+nv_num[17] = 26425;
+nv_num[18] = 19995;
+nv_num[19] = 64204;
+nv_num[20] = 33644;
+nv_num[21] = 22068;
+nv_num[22] = 53532;
+nv_num[23] = 33635;
+nv_num[24] = 31211;
+nv_num[25] = 5843;
+nv_num[26] = 45928;
+nv_num[27] = 27514;
+nv_num[28] = 41902;
+nv_num[29] = 17058;
+nv_num[30] = 37243;
+nv_num[31] = 18625;
+nv_num[32] = 17755;
+nv_num[33] = 59690;
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_write_cost/ext/edb-server/ext/libedbserver/src/host_comm_impl.c

@@ -1,0 +1,107 @@
+#include <stdint.h>
+#include <stdbool.h>
+#include <string.h>
+
+#include <msp430.h>
+
+#include <libmsp/periph.h>
+
+#include "config.h"
+#include "pin_assign.h"
+#include "error.h"
+#include "uart.h"
+#include "params.h"
+#include "host_comm.h"
+
+#include "host_comm_impl.h"
+
+/**
+ * @brief Message payload pointer in a buffer for messages to host
+ * @details This buffer is used exclusively by main loop, so it is
+ *          shared only in the sense of being multi-plexed in time, i.e. it is
+ *          never used concurrently but to threads of control.
+ */
+static uint8_t host_msg_buf[HOST_MSG_BUF_SIZE];
+static uint8_t * const host_msg_payload = &host_msg_buf[UART_MSG_HEADER_SIZE];
+
+// Uses the main loop host_msg_buf
+static inline void send_msg_to_host(unsigned descriptor, unsigned payload_len)
+{
+    // Out-of-bound writes already happen before we get here, but the payload
+    // len should be in a register and this is not a function call (inline), so
+    // this check should be robust even if memory got a little corrupted.
+    ASSERT(ASSERT_HOST_MSG_BUF_OVERFLOW, payload_len <= HOST_MSG_BUF_SIZE - UART_MSG_HEADER_SIZE);
+
+    UART_send_msg_to_host(descriptor, payload_len, host_msg_buf);
+    UART_end_transmission();
+}
+
+void send_voltage(uint16_t voltage)
+{
+    unsigned payload_len = 0;
+
+    UART_begin_transmission();
+
+    host_msg_payload[payload_len++] = voltage & 0xFF;
+    host_msg_payload[payload_len++] = (voltage >> 8) & 0xFF;
+
+    send_msg_to_host(USB_RSP_VOLTAGE, payload_len);
+}
+
+void send_return_code(unsigned code)
+{
+    unsigned payload_len = 0;
+    UART_begin_transmission();
+    host_msg_payload[payload_len++] = code;
+    send_msg_to_host(USB_RSP_RETURN_CODE, payload_len);
+}
+
+void send_interrupt_context(interrupt_context_t *int_context)
+{
+    unsigned payload_len = 0;
+
+    UART_begin_transmission();
+
+    host_msg_payload[payload_len++] = int_context->type;
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_write_cost/ext/edb-server/ext/libedbserver/src/include/libedbserver/host_comm_impl.h

@@ -1,0 +1,21 @@
+#ifndef HOST_COMM_IMPL_H
+#define HOST_COMM_IMPL_H
+
+#include <stdint.h>
+
+#include "host_comm.h"
+#include "interrupt.h"
+
+#define HOST_MSG_BUF_SIZE       64 // buffer for UART messages (to host) for main loop
+
+// TODO: prefix names with host_comm
+
+void send_voltage(uint16_t voltage);
+void send_return_code(unsigned code);
+void send_interrupt_context(interrupt_context_t *int_context);
+void send_param(param_t param);
+void send_echo(uint8_t value);
+void forward_msg_to_host(unsigned descriptor, uint8_t *buf, unsigned len);
+
+#endif
+

### 2017-06-27 paper/experiments/energyMeasurements/Fram_write_cost/ext/edb-server/ext/libedbserver/src/include/libedbserver/pin_assign.h

@@ -1,0 +1,361 @@
+#ifndef PIN_ASSIGN_H
+#define PIN_ASSIGN_H
+
+/**
+ * @defgroup    PORTS   Pin assignments
+ * @{
+ */
+
+#if defined(BOARD_EDB) || defined(BOARD_EDB_1_1)
+
+#define PORT_LED                                J //!< GPIO port for LEDs
+#ifdef BOARD_EDB_1_1
+#define PIN_LED_ORANGE                          0
+#define PIN_LED_YELLOW                          1
+#endif // BOARD_EDB_1_1
+#define PIN_LED_GREEN                           2
+#define PIN_LED_RED                             3
+
+#define PORT_LED_MAIN_LOOP                      PORT_LED
+#define PIN_LED_MAIN_LOOP                       PIN_LED_GREEN
+
+#define PORT_LED_DEBUG_MODE                     PORT_LED
+#ifdef BOARD_EDB_1_1
+#define PIN_LED_DEBUG_MODE                      PIN_LED_ORANGE
+#else // !BOARD_EDB_1_1
+#define PIN_LED_DEBUG_MODE                      PIN_LED_RED
+#endif // !BOARD_EDB_1_1
+
+#define PORT_LED_BOOT							PORT_LED
+#ifdef BOARD_EDB_1_1
+#define PIN_LED_BOOT							PIN_LED_YELLOW
+#else // !BOARD_EDB_1_1
+#define PIN_LED_BOOT							PIN_LED_RED
+#endif // !BOARD_EDB_1_1
+
+#define PORT_UART_USB                           3  //!< GPIO port with UART connected to FTDI FT232R USB
+#define PIN_UART_USB_TX                         3 //!< TX pin for UART to host USB
+#define PIN_UART_USB_RX                         4 //!< RX pin for UART to host USB
+
+#define PORT_UART_TARGET                        4  //!< GPIO port with UART connected to target device
+#define PIN_UART_TARGET_TX                      4 //!< TX pin for UART to target device
+#define PIN_UART_TARGET_RX                      5 //!< RX pin for UART to target device
+
+#define PORT_VSENSE                             6 //!< GPIO port for voltage sense lines
+#define PIN_VCAP                                1 //!< P6.1: ADC input Vcap
+#define PIN_VBOOST                              2 //!< P6.2: ADC input Vboost
+#define PIN_VREG                                3 //!< P6.3: ADC input Vreg
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_write_cost/ext/edb-server/ext/libedbserver/src/uart.c

@@ -1,0 +1,507 @@
+#include <stdint.h>
+#include <msp430.h>
+
+#include <libmsp/periph.h>
+#include <libedb/target_comm.h>
+
+#include "host_comm.h"
+#include "minmax.h"
+#include "pin_assign.h"
+#include "config.h"
+#include "error.h"
+#include "main_loop.h"
+#include "dma.h"
+
+#include "uart.h"
+
+// TODO: rename "usb" to "host"
+
+volatile unsigned host_uart_status = 0;
+
+#ifdef UART_HOST
+static uartBuf_t usbRx = { .head = 0, .tail = 0 };
+#endif // UART_HOST
+
+#ifdef UART_TARGET
+static uartBuf_t wispRx = { .head = 0, .tail = 0 };
+static uartBuf_t wispTx = { .head = 0, .tail = 0 };
+#endif // UART_TARGET
+
+/**
+ * @brief       Determine the length of the circular buffer
+ * @param       buf        Pointer to the circular buffer
+ * @return      Length of the buffer
+ */
+static inline unsigned uartBuf_len(uartBuf_t *buf) {
+    int diff = buf->tail - buf->head;
+    if(diff >= 0) {
+        return diff;
+    } else {
+        return (diff + UART_BUF_MAX_LEN_WITH_TAIL);
+    }
+}
+
+void UART_setup(unsigned interface)
+{
+    switch(interface)
+    {
+#ifdef UART_HOST
+    case UART_INTERFACE_USB:
+        GPIO(PORT_UART_USB, SEL) |= BIT(PIN_UART_USB_TX) | BIT(PIN_UART_USB_RX);
+
+        UART(UART_HOST, CTL1) |= UCSWRST; // put state machine in reset
+        UART(UART_HOST, CTL1) |= UCSSEL__SMCLK;
+#ifdef CONFIG_ABORT_ON_HOST_UART_ERROR
+        UART(UART_HOST, CTL1) |= UCRXEIE;
+#endif
+
+        UART(UART_HOST, BR0) = CONFIG_USB_UART_BAUDRATE_BR0;
+        UART(UART_HOST, BR1) = CONFIG_USB_UART_BAUDRATE_BR1;
+        UART(UART_HOST, MCTL) |= 0
+#ifdef CONFIG_USB_UART_BAUDRATE_UCOS16
+            | UCOS16
+#endif
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_write_cost/ext/edb-server/ext/libmspsoftuart/src/include/libmspsoftuart/printf.h

@@ -1,0 +1,46 @@
+/******************************************************************************
+ * Software UART example for MSP430.
+ *
+ * Stefan Wendler
+ * sw@kaltpost.de
+ * http://gpio.kaltpost.de
+ *
+ * ****************************************************************************
+ * The original code is taken form Nicholas J. Conn example:
+ *
+ * http://www.msp430launchpad.com/2010/08/half-duplex-software-uart-on-launchpad.html
+ *
+ * Origial Description from Nicholas:
+ *
+ * Half Duplex Software UART on the LaunchPad
+ *
+ * Description: This code provides a simple Bi-Directional Half Duplex
+ * Software UART. The timing is dependant on SMCLK, which
+ * is set to 1MHz. The transmit function is based off of
+ * the example code provided by TI with the LaunchPad.
+ * This code was originally created for "NJC's MSP430
+ * LaunchPad Blog".
+ *
+ * Author: Nicholas J. Conn - http://msp430launchpad.com
+ * Email: webmaster at msp430launchpad.com
+ * Date: 08-17-10
+ *****************************************************************************
+ * Includes also improvements from Joby Taffey and fixes from Colin Funnell
+ * as posted here:
+ *
+ * http://blog.hodgepig.org/tag/msp430/
+ *****************************************************************************/
+
+#ifndef __UART_H
+#define __UART_H
+
+#include <stdbool.h>
+#include <stdint.h>
+
+/**
+ * Initialize soft UART backend for printf
+ */
+void mspsoftuart_init(void);
+
+#endif
+

### 2017-06-27 paper/experiments/energyMeasurements/Fram_write_cost/ext/edb-server/src/main.c

@@ -1,0 +1,96 @@
+#include <msp430.h>
+#include <stdint.h>
+#include <stdbool.h>
+
+#include <libmsp/periph.h>
+#include <libmsp/clock.h>
+#include <libmsp/watchdog.h>
+#include <libio/log.h>
+#include <libedbserver/edb.h>
+#include <libedbserver/pin_assign.h>
+
+static inline void pin_setup()
+{
+    // Uncomment this if R3 is not populated since in that case pin is unconnected
+    // GPIO(PORT_CONT_POWER, DIR) |= BIT(PIN_CONT_POWER);
+    // GPIO(PORT_CONT_POWER, OUT) &= ~BIT(PIN_CONT_POWER);
+
+#ifdef CONFIG_DEBUG_MODE_LED
+    GPIO(PORT_LED_DEBUG_MODE, OUT) &= ~BIT(PIN_LED_DEBUG_MODE);
+    GPIO(PORT_LED_DEBUG_MODE, DIR) |= BIT(PIN_LED_DEBUG_MODE);
+#endif // CONFIG_DEBUG_MODE_LED
+
+#ifdef CONFIG_MAIN_LOOP_LED
+    GPIO(PORT_LED_MAIN_LOOP, OUT) &= ~BIT(PIN_LED_MAIN_LOOP);
+    GPIO(PORT_LED_MAIN_LOOP, DIR) |= BIT(PIN_LED_MAIN_LOOP);
+#endif // CONFIG_MAIN_LOOP_LED
+
+#ifdef CONFIG_BOOT_LED
+    GPIO(PORT_LED_BOOT, OUT) &= ~BIT(PIN_LED_BOOT);
+    GPIO(PORT_LED_BOOT, DIR) |= BIT(PIN_LED_BOOT);
+#endif // CONFIG_BOOT_LED
+}
+
+int main(void)
+{
+#ifdef CONFIG_WATCHDOG
+	msp_watchdog_enable(CONFIG_WDT_BITS);
+#else // !CONFIG_WATCHDOG
+	msp_watchdog_disable();
+#endif // !CONFIG_WATCHDOG
+
+    pin_setup();
+
+#ifdef CONFIG_BOOT_LED
+    GPIO(PORT_LED_BOOT, OUT) |= BIT(PIN_LED_BOOT);
+#endif // CONFIG_BOOT_LED
+
+    msp_clock_setup(); // set up unified clock system
+
+#ifdef CONFIG_CLOCK_TEST_MODE
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_write_cost/python/fram_sram_write_energy_consumption.py

@@ -1,0 +1,87 @@
+#!/home/amjed/anaconda3/bin/python
+import numpy as np 
+import matplotlib.pyplot as plt 
+from matplotlib import style
+
+style.use('seaborn-muted')
+
+numberOfReadLines = 1600;
+
+#----Get the volatge values of the cap-----#
+fram_voltages=[]
+sram_voltages=[]
+fram_cntr=0
+sram_cntr=0
+for line in open("../outputs/fram.txt"):
+	# get the voltages and remove the `\n` char
+	fram_voltages.append( (line.split(',')[-1])[:-1] ) 
+	fram_cntr+=1
+
+for line in open("../outputs/sram.txt"):
+	# get the voltages and remove the `\n` char
+	sram_voltages.append( (line.split(',')[-1])[:-1] )
+	sram_cntr+=1
+
+#print(fram_voltages)	 
+#print(sram_voltages)
+#exit()
+
+
+#-----SRAM/FRAM write costs------#
+v_list_len= min(fram_cntr, sram_cntr)
+fram_diffs=[]
+sram_diffs=[]
+for i in range(0,v_list_len, 2):
+	fram_diffs.append( ( ( (float(fram_voltages[i])/numberOfReadLines )**2 - (float(fram_voltages[i+1])/numberOfReadLines ) **2 ) * 10e-6 * 0.5 ) ) 
+	sram_diffs.append( ( ( (float(sram_voltages[i])/numberOfReadLines )**2 - (float(sram_voltages[i+1])/numberOfReadLines ) **2 ) * 10e-6 * 0.5 ) ) 
+
+# print(fram_diffs[:5])
+# print(sram_diffs[:5])
+# exit()
+
+# print (sum(fram_diffs) /len(fram_diffs))
+# print (sum(sram_diffs) /len(sram_diffs))
+
+
+#------------Line fitting------------#
+fit_order=3
+xVals = range(len(fram_diffs))
+fram_z =  np.polyfit( xVals, fram_diffs, fit_order)
+sram_z =  np.polyfit( xVals, sram_diffs, fit_order)
+fram_estYVals = np.polyval(fram_z, xVals)
+sram_estYVals = np.polyval(sram_z, xVals)
+
+
+#--------------Plotting--------------#
+
+plt.rcParams['axes.labelsize'] =  12
+plt.rcParams['lines.linewidth'] =  1.5
[…]

### 2017-06-27 paper/experiments/energyMeasurements/Fram_write_cost/python/generate_array.py

@@ -1,0 +1,18 @@
+#!/usr/python3
+# this script is used to generate an array for the experiment of 
+# FRAm energy cost.
+
+import random 
+import numpy as np 
+
+arraySize = 400
+listLen =  1600
+
+indices = np.random.choice(arraySize,arraySize)
+
+nv_num_f = open("nv_num.txt", "w")
+vo_num_f = open("vo_num.txt", "w")
+
+for i in range(listLen):
+	nv_num_f.write("nv_num[{}] = {};\n".format(indices[i%arraySize], random.randrange(65000) ))
+	vo_num_f.write("vo_num[{}] = {};\n".format(indices[i%arraySize], random.randrange(65000) ))

### 2017-06-27 paper/experiments/runtime/SRC/Chain_code/DFT/main.c

@@ -1,0 +1,327 @@
+/*
+	This is a DFT implemention in chain
+
+	@ Author: Amjad Yousef Majid
+	@ Date  : 6/April/2017
+
+*/
+#include <msp430.h>
+#include <stdint.h>
+#include <stdbool.h>
+#include <stdio.h>
+
+#include <math.h>    /// using the normal c math library   "-lm" must be added to the Makefile
+
+#include <libchain/chain.h>
+#include <stdlib.h>
+
+#define SIZE 8
+#define PI2 6.28
+
+#ifdef DEBUG
+  float x[SIZE];              		// discrete-time signal, x
+  float x_chain[SIZE];              // discrete-time signal, x
+  float Xre[SIZE], Xim[SIZE]; 		// DFT of x (real and imaginary parts)
+  float P[SIZE]; 
+  float p_deb[SIZE]={0};
+ void dft_c() {
+ 	/*
+		The dft_c function is taken from 
+		https://batchloaf.wordpress.com/2013/12/07/simple-dft-in-c/
+ 	*/
+    int n, k; // iSIZEdices for time and frequency domains
+
+    // Generate random discrete-time signal x in range (-1,+1)
+    srand(0);
+    for (n = 0; n < SIZE; ++n)
+      x[n] = ((2.0 * rand()) / RAND_MAX) - 1.0;
+
+    // Calculate DFT of x using brute force
+    for (k = 0; k < SIZE; ++k) {
+      // Real part of X[k]
+      Xre[k] = 0;
+      for (n = 0; n < SIZE; ++n)
+        Xre[k] += x[n] * cos(n * k * PI2 / SIZE);
+
+      // Imaginary part of X[k]
+      Xim[k] = 0;
+      for (n = 0; n < SIZE; ++n)
+        Xim[k] -= x[n] * sin(n * k * PI2 / SIZE);
+
+      // Power at kth frequency bin
+      P[k] = Xre[k] * Xre[k] + Xim[k] * Xim[k];
+    }
+
+  }
+
+#endif
+
+void init()
+{
+	WDTCTL = WDTPW | WDTHOLD;	// Stop watchdog timer
+	PM5CTL0 &= ~LOCKLPM5;
+    __enable_interrupt();
[…]

### 2017-06-29 paper/paper.tex
partial intro draft plus todo macro plus ref fixes to make it all build

--- 2017-06-27 paper/inExDMA.tex
@@
 % \usepackage[inline]{enumitem}
+\usepackage{xcolor}
+\usepackage{xspace}
+\usepackage{natbib}
 
@@
 \usepackage{subfigure}
+\newcommand{\TODO}[1]{\bf \em \textcolor{red}{TODO: #1}\xspace}
 
@@
 
-\title{VIPOS: Virtualized Intermittently Powered Runtime Environment}
+\title{VIPOS: Virtualized Intermittently Powered Runtime Environment (``VIPRE''?)}
 
@@
 \label{sec:intro}
+
+Advances in processor efficiency along with the development of
+energy-harvesting power systems has created a new category of devices that
+require neither a battery nor a tethered power supply.  These devices operate
+entirely using energy extracted from their environment, such as RF energy,
+photovoltaics, and vibration.  Incorporating compute, storage, sensing, and
+communication hardware, such energy-independent devices are a promising
+candidate as the underlying technology for emerging internet of things (IoT)
+applications, in- and on-body medical applications, and emerging applications,
+like energy-harvesting nano-satellites.
+
+While promising, energy-harvesting devices present a unique design challenge in
+that they operate only {\em intermittently} when energy is available.   An
+energy-harvesting device buffers energy from its environment, e.g., in a
+capacitor. When a threshold quantity of energy accumulates, the device begins
+operating.  Harvestable energy sources typically produce relatively very little
+power compared to a platform's operating power level.  A devices operates for a
+brief period of time and when buffered energy is depeleted, shuts down and
+begins recharging to later operate again.  Charge and discharge times vary by
+device and some devices~\cite{wisp} may fail 10 to 100 times per second.
+
+Software in an energy-harvesting system operates according to the {\em
+intermittent execution model}~\cite{dino}, with bursts of execution that are
+interrupted by failure periods. When power fails, a device loses its volatile
+state (i.e., registers, stack, SRAM) and retains its non-volatile state (i.e.,
+FRAM, Flash). While capturing periodic checkpoints~\cite{mementos,tictpl,quickrecall}
+and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve
+execution progress, failures can leave volatile and non-volatile state inconsistent,
+leading to unrecoverable failures~\cite{mspcdino,edb,edbtoppicks}.
+
+\TODO{Don't forget all the references in this section!}
+New programming and execution models~\cite{dino,ratchet,chain,alpaca} and hardware
+mechanisms~\cite{clank, idetic} help avoid data inconsistency
+
+\section{Old Introduction}
 	Intermittently powered devices (IPDs) are battery-less devices that utilize the ambient energy to sense, compute and, communicate.
-For example, Wireless Identification and Sensing platform WISP \cite{WISP} uses the RF signal power to drive its computation and communication.
+For example, Wireless Identification and Sensing platform WISP \cite{wisp} uses the RF signal power to drive its computation and communication.
 Because of the reliance of TPDs on intermittent power supply, the harvested energy when it is available, their programs are susceptible to a very frequent power interrupts, in the order of tens of millisconds~\cite{}.
@@
 CPU registers, the stack and the global variables) is saved to a non-volatile memory.
-Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos, HarvOS, Hibernus}.
+Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos, harvOS, hibernus}.
 The benefit of this model is that it does not require code modification by a programmer.
-However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{xxxx}; and (iii) it restricts the IPDs to run only a single application~\cite{InOS}.
+However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{xxxx}; and (iii) it restricts the IPDs to run only a single application~\cite{inos}.
 		\item \emph{Modular Intermittent Execution Model}:
@@
 Therefore, it tolerates arbitrary number of power interrupts.
-This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an IPD by interleaving their tasks~\cite{InOS}.
+This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an IPD by interleaving their tasks~\cite{inos}.
 However, it obviously requires code modification---for example, if an algorithm is written according to the continuous execution model it has to be splitted, by a programmer, into small tasks to run under the modular intermittent execution model.
@@
 Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes. 
-
-	\begin{figure}[t]
-	    \centering
-	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
-	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
-	    \caption{Task flow control of Chain}
-	    \label{fig:DynamicChain}
-	\end{figure}
+%
+%	\begin{figure}[t]
+%	    \centering
+%	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
+%	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
+%	    \caption{Task flow control of Chain}
+%	    \label{fig:DynamicChain}
+%	\end{figure}
 
@@
 %
-	Since accessing/updating buffers is more costly than directly accessing FRAM, at least without specialized hardware~\cite{Clank}, we introduce the second method to apply the virtualization principle to the Modular Execution Model.
-This method utilizing the Direct Memory Access (DMA) Module~\cite{DMA}.
+	Since accessing/updating buffers is more costly than directly accessing FRAM, at least without specialized hardware~\cite{clank}, we introduce the second method to apply the virtualization principle to the Modular Execution Model.
+This method utilizing the Direct Memory Access (DMA) Module.
 As Fig.~\ref{fig:dmaTimeEnergy} shows that DMA is much more efficient in transferring a block of data than the conventional (pointer based) way from the energy and time perspectives.
@@
 
-\bibliographystyle{plain}
-\bibliography{references}
+\bibliographystyle{abbrvnat}
+\bibliography{bib}
 

### 2017-06-29 paper/paper.tex
rough draft of the intro

@@
 \usepackage{subfigure}
-\newcommand{\TODO}[1]{\bf \em \textcolor{red}{TODO: #1}\xspace}
+\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
+\newcommand{\sys}{NAME\xspace}
 
@@
 \title{VIPOS: Virtualized Intermittently Powered Runtime Environment (``VIPRE''?)}
-
 % \author{
@@
 \section{abstract}
+\TODO{Need to re-write this to make it consistent with the story after we have the story in place}
 Enabling battery-free devices is a mandatory step towards the environment friendly IoT.
@@
 \label{sec:intro}
+\TODO{VIPER: Virtualized Intermittent Programming model, Execution model, and Runtime environment}
+\TODO{PREVIEW: Programming and Runtime Environment with Virtualized Intermittent Execution Windows}
 
@@
 
-\TODO{Don't forget all the references in this section!}
-New programming and execution models~\cite{dino,ratchet,chain,alpaca} and hardware
-mechanisms~\cite{clank, idetic} help avoid data inconsistency
+\TODO{Don't forget to include all the references in this section!}
+
+%FRAM costs more than FRAM -- need to figure out how to execute out of SRAM
+New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
+architecture mechanisms~\cite{clank, idetic, nvp} help avoid data
+inconsistency.  New architectures require costly hardware changes and are
+inapplicable to today's systems.  New programming models and compilers force
+programmers to use new memory models~\cite{chain,ratchet}, which may inhibit
+adoption.  New memory models may also increase the amount of non-volatile
+memory accesses by requring non-volatile versioning~\cite{chain} or completely
+precluding the use of volatile memory~\cite{ratchet}.  Increasing pressure on
+non-volatile memory or precluding the use of volatile memory may degrade
+efficiency; e.g., FRAM has a higher access energy and latency than
+SRAM~\cite{alpaca,wisp}.  Commercially available devices are often provisioned
+with much less SRAM than FRAM~\cite{wolverine,wisp} and effectively leveraging
+the relatively high efficiency of SRAM is a key challenge for task-based
+intermittent systems.
+
+\TODO{concrete cost multiples here?}
+
+%Task sizing is hard -- how to chop without risking inefficiency or non-term
+Several recent programming efforts~\cite{alpaca,chain} proposed {\em
+task-based} programming models, that require the programmer to statically
+decompose their program into tasks.  A task is like a function with no callers
+that can include arbitrary computation and, upon completion, is guaranteed to
+have executed {\em atomically} and to have logically {\em committed} memory
+updates.  The programmer expresses the flow of control from one task to another
+and task control-flow may be input-dependent.  An important programming
+challenge in task-based models is that the length of execution of a task is
+limited by the total amount of energy that a device can buffer.  Assuming that
+input power is negligible compared to operating power, a task will never be
+able to complete if its execution consumes more energy than the system can
+buffer.  Transitioning from one task to the next imposes a cost and excessively
+small tasks may be inefficient.  Appropriately sizing tasks to ensure progress
+and high performance is a key challenge for task-based intermittent systems.
+
+In this work, we develop \sys: a new programming and execution model that uses
+{\em volatile memory virtualization} to leverage efficient volatile memory and
+and {\em dynamic task coalescing} to efficiently execute tasks without
+exceeding available energy.  \sys's memory virtualization mechanism uses a
+device's SRAM as the working memory during a task's execution, populating the
+SRAM with a page of data from FRAM as necessary at the start of the task.  At
+the end of the task, the SRAM is committed back to FRAM using efficient DMA
+block copies.  \TODO{Brandon sez: Is the next sentence what the system does?
+Or are we doing paging?} When a task accesses data outside of a page in SRAM,
+the task must fetch data from FRAM.  We explore two strategies. The first
+strategy is {\em demand paging}, which swaps the SRAM page with a new page from
+FRAM, buffering the swapped-out page until commit.  The second strategy is {\em
+buffered direct access} \TODO{better name}, which directly reads and writes
+FRAM relying on dynamic double-buffering to ensure memory is consistent at
+commit.
+
+\sys's dynamic task coalescing mechanism allows the programmer or a compiler to
+intuitively decompose their program into small tasks that amortize fixed
+per-task overheads, yet present no risk of exceeding device energy capacity. As
+such a decomposition executes, \sys {\em coalesces} dynamically consecutive
+tasks. Coalescing elides the commits of coalesced tasks by buffering multiple
+tasks' updates in an FRAM commit buffer.  Periodically, as the span of the
+coalesced task grows, \sys ends coalescing and commits the state of the
+coalesced task.  If a power failure interrupts a sequence of coalesced tasks,
+\sys adaptively reduces the number of tasks in that sequence that it will
+coalesce, committing sooner in future executions.  Consequently, \sys's
+coalescing mechanism allows a program to execute efficiently across a range of
+energy buffer sizes, avoiding transition overheads in larger buffers, and
+ensuring progress in smaller buffers. \TODO{articulate this last point better}
+
+We evaluated our system on a collection of benchmark programs taken from the
+literature, and compare directly to prior work, showing that \sys has high
+performance and can flexible target a variety of platforms without recompiling
+the code.  We further demonstrate the efficiency and versatility of \sys with
+an end-to-end evaluation of a sensing and data processing application deployed
+in a lab environment \TODO{blah blah blah eval}.
+
+\TODO{Bullet contribs if there's space}
 

### 2017-06-29 paper/abstract.tex
modularized.  next thing to do is do some outlining for the first few sections

--- 2017-06-27 paper/inExDMA.tex
@@
-\documentclass[sigconf,anonymous,review]{acmart}
-\usepackage[normalem]{ulem}
-
-\usepackage{amssymb} % needed for: various symbols
-\usepackage{amsmath} % needed for: math symbols
-% http://tex.stackexchange.com/questions/229355/algorithm-algorithmic-algorithmicx-algorithm2e-algpseudocode-confused
-\usepackage{algorithm} % this used as an environment for the algorithm code (linke figure and table)
-\usepackage{algorithmicx} % this is used to write the pseudo code
-\usepackage[noend]{algpseudocode} % this is for the layout of the pseudo code
-\usepackage{tikz} % needed for: tikz pics
-% \usepackage[inline]{enumitem}
-
-\usepackage{graphicx}  
-\usepackage{subfigure}
-
-\begin{document}
-
-\title{VIPOS: Virtualized Intermittently Powered Runtime Environment}
-
-% \author{
-% 	\IEEEauthorblockN{Amjad Yousef Majid, Przemys{\l}aw Pawe{\l}czak}
-% 	\IEEEauthorblockA{TU Delft, Mekelweg\,4, 2628\,CD Delft, NL\\
-% 	\{a.y.majid, p.pawelczak\}@tudelft.nl}
-% }
-
-\maketitle
-
-\section{abstract}
+\TODO{Need to re-write this to make it consistent with the story after we have the story in place}
 Enabling battery-free devices is a mandatory step towards the environment friendly IoT.
@@
 If a code is written for a very small energy buffer, VIPOS is able to dynamically merge these tasks and constructs a bigger virtual task to improve the performance, reduce energy consumption and provides mean for code portability.
-
-\section{Introduction}
-\label{sec:intro}
-	Intermittently powered devices (IPDs) are battery-less devices that utilize the ambient energy to sense, compute and, communicate.
-For example, Wireless Identification and Sensing platform WISP \cite{WISP} uses the RF signal power to drive its computation and communication.
-Because of the reliance of TPDs on intermittent power supply, the harvested energy when it is available, their programs are susceptible to a very frequent power interrupts, in the order of tens of millisconds~\cite{}.
-Therefore, these devices require a different software execution model that complies with the nature of a discontinuous power supply.
-
-	The intermittent (discontinuous) execution model defines a program execution as cumulative discrete process.
-The main difference between the intermittent and the conventional (or continuous) execution models is that, a power failure is seen by the continuous model as an \emph{exception} that may reset the progress of a program to its beginning.
-Whereas, in intermittent execution a power failure is regarded as a temporary \emph{pause} to the execution that may result in some progress degradation.
-Generally we can classify the intermittent execution model into:
-	\begin{itemize}
-		\item \emph{Sequential Execution Model}:
-			Under this model a program is seen as one big idempotent operational region that has one common context.
-Generally, The progress of the program is saved and updated by means of checkpointing---where all the program context (e.g.
-CPU registers, the stack and the global variables) is saved to a non-volatile memory.
-Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos, HarvOS, Hibernus}.
-The benefit of this model is that it does not require code modification by a programmer.
-However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{xxxx}; and (iii) it restricts the IPDs to run only a single application~\cite{InOS}.
-		\item \emph{Modular Intermittent Execution Model}:
-			At the heart of this model is the concept of an idempotent task.
-The idempotent task is C function that does not have arguments and does not return a value.
-This task uses a well defined interface to interact with the non-volatile memory.
-Therefore, it tolerates arbitrary number of power interrupts.
-This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an IPD by interleaving their tasks~\cite{InOS}.
-However, it obviously requires code modification---for example, if an algorithm is written according to the continuous execution model it has to be splitted, by a programmer, into small tasks to run under the modular intermittent execution model.
-	\end{itemize}
-
-Despite that the superiority of the Modular Intermittent Execution model, it is still a static approach that completely depends on a programmer's estimation which is mostly result in a sub-optimal code devision.
-Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.
-
-% Energy source
-
-% The intermittent execution is cumulative discrete process.
-The intermittent execution model is only able to execute a few number of instructions, as compared to the conventional (continuous) model, before its progress is terminated.
-Therefore, the intermittent execution model adapts an execution progress state saving mechanism.
-This mechanism is realized either by means of a checkpoint~\cite{}, where all the context of the program is saved into non-volatile memory.
-Or
-
-% This progress state saving mechanism is normally injected into a program either by a compiler~\cite{}.
-A compiler injects trigger pointers to checkpoint (save) the context of the program into the non-volatile memory.
-or it is added by a programmer~\cite{chain}.
-[compare these two approaches]...
-
-% We define virtualization within the context of intermittent execution as utilizing the volatile memory instead of non-volatile when the intermittent execution model attempts to access the non-volatile memory.
-
-\subsection{Problem Statement}
-How to approach the optimal task size given an unknown energy buffer size and environment conditions? 
-	 \begin{enumerate}
-		 \item How to eliminate the need for the hardware support while enabling the Modular Intermittent Execution model to dynamically adapt its task size?
-		 \item How to enable code portability under the Modular Intermittent execution model? 
-		 \item How to reduce the energy consumption of executing a task? 
-	\end{enumerate}
-
-\subsection{Why not Adapting Current Approaches}
-Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes.
-
-	\begin{figure}[t]
-	    \centering
-	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
-	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
-	    \caption{Task flow control of Chain}
-	    \label{fig:DynamicChain}
-	\end{figure}
-
-\noindent\textbf{Chain: Tasks and Channels for Reliable Intermittent Programs } \\
-Some of the main problems in modifying Chain to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Multi-versioning system:} Chain defines for each task its own input that is not shared with other tasks.
-This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
-	\item \emph{Direct FRAM system:} Chain tasks directly access FRAM which is more energy expensive Than SRAM.
-As result of being multi-versioning system that directly interacts with FRAM, Chain has a significant energy consumption overhead.
-	\item \emph{Irreversible tasks transitions:} According to the Chain programming model, each task calls the next one.
-The transition is done through functions calls and manually clearing the stack, to prevent the stack overflow problem.
-This approach make all the transitions firm ones which prevent task merging.
-	\item \emph{All channels are required:} Fig.~\ref{fig:DynamicChainSeq} shows that if the task control flow is a circular with no branches then task merging is possible.
-Moreover, all the intra-merged-tasks channel can be committed to SRAM instead of FRAM to save energy.
-However, this execution path is only a special.
-If the general case (see Fig.~\ref{fig:DynamicChainRan}) is considered then leaving out the intra-merged-task channels might result in data inconsistency---If an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready which result in an incorrect execution.
-
-\end{itemize}
-
-\noindent\textbf{Alpaca: Intermittent Execution without Checkpoints} \\
-Some of the main problems in modifying Alpaca to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Privatization:} The core concept of Alpaca is privatization.
-Privatization depends completely on detecting the variables that have a Write-after-read dependency within the scope of a task.
-Therefore, if tasks are on-demand merged the scope of these dependencies are changed and the static analysis to privatize these variables is not valid anymore and data consistency can not be preserved.
-
-	\item \emph{Irreversible tasks transitions:} Similar to Chain problem.
-
-\end{itemize}
-
-\noindent\textbf{Virtualizing Modular Intermittent Execution Model}
-
-		An Intermittently executed program, as seen by the modular intermittent execution model, is a chain of tasks with a firm transition from one to another. 
-		Virtualizing this model means softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
-Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
-Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
-This can be done by having a temporary copy of a global variable in the volatile memory (similar to the catching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
-
-		In this paper we are introducing VIPOS a runtime library the implement the Virtualed Intermittent Execution Model.
-VIPOS adapts two methods for data protection.
-One relies on a virtual buffer and the other uses the Direct Memory Access (DMA).
-VIPOS is able to approach the optimal tasks devision during the runtime.
-VIPOS reduces an application execution time by XX and the energy by XX.
-
-\subsection{Contributions}
-	 \begin{enumerate}
-		 \item We have developed an algorithm that is able to determine the size of a virtual task based on the history of the execution of an application. 
-		 \item We enable efficient code portability by requiring that a real task to small and merging them during the execution.   
-	\end{enumerate}
-
- \section{Preliminary Results}
- \label{sec:prelResults}
-
-\begin{figure}[t]
-    \centering
-         \subfigure[The energy cost of FRAM/SRAM writes operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
-         \subfigure[The energy cost of FRAM/SRAM read operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
-    \caption{The energy cost of accessing volatile/non-volatile memory}
-    \label{fig:framEnergy}
-\end{figure}
-
-Since the Intermittent execution models rely on the non-volatile memory to enable long-running operation, it is important to characterize the energy consumption of accessing this type of memory as compare to the volatile one.
-
-Four applications are developed to measure the energy consumption of \emph{Accessing} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{xxx}.
-The interference free debugger (EDB)~\cite{xxx} is used as the energy measurement tool.
-The EDB probes the energy buffer before and after accessing the FRAM/SRAM 1600 times, This large number of FRAM access is used to increase the reliability of the results generated by the used measurement tool (e.g.
-reducing the effect of the quantization error).
-The energy buffer is charged, using the EDB, to $\approx$ 2.45 volts \textbf{only} at the beginning of the execution and of the programs.
-The write operations were performed by writing literal values, random numbers, to the memory.
-However, a read operation must be followed by a write operation, therefore, we chose to write a read value to SRAM.
-Fig.~\ref{fig:framEnergy} shows that an FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We would like to mentioned that TI has already mentioned in this document~\cite{xxxx} that FRAM write access is more energy expensive than SRAM write access.
-However, TI does not quantify the energy overhead.}.
-Therefore, reducing FRAM access is desirable when developing  energy limited software solution
-
-\section{VIPOS: Virtualized Intermittently Powered Runtime Environment}
-VIPOS is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs.
-It is able, on-demand, to merge tasks to reduce the number of commits to the non-volatile memory.
-
-\subsection{VIPOS: Data/Memory Protection Methods}
-VIPOS data consistency preservation methods are designed with goal of minimal interaction between the CPU and FRAM to reduce energy consumption (see Section~\ref{sec:prelResults}).
-These data/memory protection methods enables VIPOS to tolerate arbirary number of power failures.
-
-	\subsubsection{Virtualized Operational Buffer}
-
-		\begin{algorithm}[t]
-			\caption{Virtualized Operational Buffer}
-			\label{algo:virtuBufWrite}
-			\scriptsize
-			%\small
-			\begin{algorithmic}[1]
-				\State $var \in \text{\{global variables\}} $ 
-				\State \label{lst:virtuBufWrite:line:begin}\Call{Virtual Task()}{} 
-				\While { \textit{executing} } \Comment{Execution stage}
-					\State $var$  $\rightarrow$ \textsf{volatile buffer} \label{lst:virtuBufWrite:line:output}
-					\If { $var$  in \textsf{ volatile buffer} }				\label{lst:virtuBufWrite:line:inputBegin}
-							\State $var$  $\leftarrow$  \textsf{volatile buffer} 
-						\Else 
-							\State $var$  $\leftarrow$  \textsf{FRAM}		\label{lst:virtuBufWrite:line:inputEnd}
-						\EndIf
-
-					\If {power interrupts}
-						\State back to \ref{lst:virtuBufWrite:line:begin}
-					\EndIf
-				\EndWhile
-
-				\While{  $\textsf{volatile buffer}\not=\emptyset$  } \label{lst:virtuBufWrite:line:commitBegin} \Comment{ First phase commit}
-					\State \textsf{volatile buffer} $\rightarrow$  \textsf{persistent buffer}
-					\If {power interrupts}
-						\State discard \textsf{persistent buffer}
-						\State back to  \ref{lst:virtuBufWrite:line:begin} 
-						\State   \label{lst:virtuBufWrite:line:commitEnd}
-					\EndIf
-				\EndWhile
-
-				\While{ $\textsf{persistent buffer}\not=\emptyset$ } \label{lst:virtuBufWrite:line:SecCommitBegin} \Comment{Second phase commit}
-					\State \textsf{persistent buffer} $\rightarrow$ FRAM 
-					\If {power interrupts}
-						\State Continue
-					\EndIf
-				\EndWhile 
-				\State     \label{lst:virtuBufWrite:line:SecCommitEnd}
-				\State return
-			\end{algorithmic}
-		\end{algorithm}
-
-		Since accessing FRAM is more energy expencive and slower than accessing SRAM, preventing frequent access to FRAM (i.e.
-during looping operations) is desireable.
-Therefore, the first proposed protection method utilizes a volatile buffer that holds temporary all the outputs of a virtual task (see Algorithm~\ref{algo:virtuBufWrite} line~\ref{lst:virtuBufWrite:line:output}).
-Consequently, a virtual task must first attempt to read a global variable from the volatile buffer before trying to obtain the value from the non-volatile memory (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:inputBegin}-\ref{lst:virtuBufWrite:line:inputEnd}) to ensure a correct execution progress and the consistency of the memory.
-Once the execution of a virtual task is done, the first phase of the commit process is started by copying the volatile buffer to a persistent buffer (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:commitBegin}-\ref{lst:virtuBufWrite:line:commitEnd}).
-If the power is interrupted the persistent buffer must be discarded and the execution must start again from the beginning of the virtual task.
-In another words, the first phase commit must be performed atomically to preserve the consistency of the output of a virtual task.
-The second phase commit is a power failure immune precess that is responsible for distributing the global variables to their final locations and make the non-volatile memory consistent and synchronized with the computation progress (see Algorithm~\ref{algo:virtuBufWrite}lines~\ref{lst:virtuBufWrite:line:SecCommitBegin}-\ref{lst:virtuBufWrite:line:SecCommitEnd}).
-
-	\paragraph{Virtualized Operational Buffer: Implementation}
-
-		We implement a reference implementation of VIPOS that uses virtualized Operational buffer to protect the data against power failure. 
-		\paragraph{Virtual Buffer}
-			We realized the virtaulzed buffer as a volatile hashed table of linked lists.
-The hashing technique was chosen to reduce the buffer searching time and the linked lists are used to prevent data loss when there is a conflict between multiple variables---If two variables have the same hash value they will occupy the same cell, however, with linked list a new node will be created for each variable to resolve the conflict and protect the data.
-The hashing function is based on the observation that the virtual addresses of the memory cells have approximately a flat distribution over the memory addressing space and the fact that each entry, of the virtual table, has the address and the value of a variable.
-As such, by using the least significant bits as an index to access the virtual buffer the hash function distributes its inputs uniformly in the virtual buffer.
-Accordingly, the linked lists search time, on average, is reduced.
-
-		\paragraph{Persistent Buffer}
-			The persistent buffer is implemented as static array of tuples, where each tuple holds  the address and the value of a variable.
-This data structure is very suitable for the second phase commit where each element has to be committed to its finial location.
-However, the complexity of committing this buffer is of size $O(N)$, where $N$ equals the length of the buffer.
-Therefore, the size of the buffer can have dramatic effect on the performance of VIPOS---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious.
-On the other hand, if the size of the buffer is large VIPOS will experience a significant performance degradation.
-However, by observing the type of the information that this buffer holds, namely memory addresses, we can, on average, reduce the complexity of committing this buffer by defining the \emph{effective size} of the buffer to be the size of the buffer up to the first cell that holds an invalid memory address.
-Moreover, since this buffer is in FRAM which is much bigger than SRAM its size can be relatively big.
-
-		\paragraph{Virtualized Operational Buffer: delay analysis}[initial results]
-
-		\begin{figure}[t]
-			\centering
-			\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
-			\caption{VIPOS based on virtualized Operational Buffer.
-It runs the decompression application with different buffer configurarions.
-EPB: Enhanced Persistent Buffer.
-NR: no read operation from the buffer.
-NRW: not read or write operations from the buffer.}
-			\label{fig:virtOperationalBuf}
-		\end{figure}
-
-			To analyze the performance of VIPOS versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
-We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}.
-The results presented in Fig.~\ref{fig:virtOperationalBuf} show the VIPOS has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.
-
-			All the experiments were done with a persistent buffer of size 100.
-However, the bar labeled with "EPB/16" in  Fig.~\ref{fig:virtOperationalBuf} shows the effect of the enhanced commit process that uses the buffer effective size instead of the buffer maximum size.
-We see that the execution time is reduced to $\approx$80 ms.
-Furthermore, we quantized the overhead of the read operations from the buffer, by eliminating this operation, and the bar labeled with "NR" shows that it cost  $\approx$10 ms.
-Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to VIPOS best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.
-
-\subsubsection{ Direct Memory Access Based Protection Method (better name is needed)}
-%
-	\begin{figure}[t]
-		\centering
-         \subfigure[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize.eps} }
-	     \subfigure[The energy needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
-		\caption{Time and energy consumption of moving a block of data from SRAM to FRAM}
-		\label{fig:dmaTimeEnergy}
-	\end{figure}
-%
-	Since accessing/updating buffers is more costly than directly accessing FRAM, at least without specialized hardware~\cite{Clank}, we introduce the second method to apply the virtualization principle to the Modular Execution Model.
-This method utilizing the Direct Memory Access (DMA) Module~\cite{DMA}.
-As Fig.~\ref{fig:dmaTimeEnergy} shows that DMA is much more efficient in transferring a block of data than the conventional (pointer based) way from the energy and time perspectives.
-
-	This method combines a number of techniques to preserve the persistence and consistency of the data.
-It backs up the status of the global variables in the non-volatile memory, and double buffers them to guarantee their consistency.
-Furthermore, on each power up it populates the global variables in the volatile memory from the most recent back up in the non-volatile memory to enable a faster task execution since the task does not need to interact with non-volatile memory which is generally slower and more energy expensive.
-
-\subsection{VIPOS: The Virtualizing Engine}
-\subsubsection{Power Interrupt Immune Scheduler}
-It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures.
-VIPOS provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
-(un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
-
-\subsubsection{Task Merging Algorithms}
-
-\paragraph{Fixed Virtual Task Size} [better name needed]
-
-	\begin{algorithm}[t]
-		\caption{Fixed virtual Task size}
-		\label{algo:fixVirtTask}
-		\scriptsize
-		%\small
-		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{VIPOS Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\State MVTS: maximum VT size
-			\vspace{0.1cm}
-
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
-				\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-							\State $ MVTS = VTS $
-						\EndIf
-				\EndWhile
-
-				\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\If{$VTS < MVTS$}
-					\State $VTS++$
-					\EndIf
-				\EndIf
-			\EndWhile
-		\end{algorithmic}
-	\end{algorithm}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{figures/virtualTaskSize.eps}
-	\caption{Size of the virtual task versus the execution time of a dummy application that contains 12 empty tasks.}
-	\label{fig:virtualTaskSize}
-\end{figure}
-
-VIPOS is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
-Fig.~\ref{fig:virtualTaskSize} show the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
-
-Remark: A less obvious benefit of visualization is that it can enable a secure computation.
-By increasing the size of the virtual task, the energy in the super-capacitor will not be enough to finish the execution of the virtual task.
-Therefore, without the interrogator being close to the intermittently powered device and charging rate is not negligible as compared to the discharging rate this computation will not be finished.
-As such, virtualization can add a layer of security to the computation.
-
-\section{Results}
-We compare the performance of VIPOS against a state-of-the-art intermittent execution approach.
-For this purpose, we used two different applications: (i) A data decompression application, which utilizes Huffman decoding technique to decompress 100 byte of data.; (ii) A discrete Fourier transform (DFT) application that uses two different resolution (4 and 8 bites) to analyze a randomly generated signal.
-
-\subsection{Continuous Power Supply}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{figures/executionTime.eps}
-	\caption{Time to complation of VIPOS versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
-	\label{fig:IPOSPerformance}
-\end{figure}
-
-Fig.~\ref{fig:IPOSPerformance} shows that VIPOS requires a shorter execution time to execute an application under the MIEM.
-The size of the virtual task that VIPOS uses equals One real task, in other words, no visualization is applied.
-
-\subsection{VIPOS Operational Buffer Versus VIPOS DMA}[place holder]
-			\begin{figure}[t]
-				\centering
-				\includegraphics[width=0.7\columnwidth]{figures/ipos_virtualizedBuffer_overhead.eps}
-				\caption{Virtualized Oprational buffer overhead.}
-				\label{fig:virtualBuffer}
-			\end{figure}
-
-\subsection{Intermittent Power Supply}
-...
-
-\bibliographystyle{plain}
-\bibliography{references}
-
-\end{document}

### 2017-06-29 paper/intro.tex
modularized.  next thing to do is do some outlining for the first few sections

--- 2017-06-29 paper/paper.tex
@@
-\documentclass[sigconf,anonymous,review]{acmart}
-\usepackage[normalem]{ulem}
-
-\usepackage{amssymb} % needed for: various symbols
-\usepackage{amsmath} % needed for: math symbols
-% http://tex.stackexchange.com/questions/229355/algorithm-algorithmic-algorithmicx-algorithm2e-algpseudocode-confused
-\usepackage{algorithm} % this used as an environment for the algorithm code (linke figure and table)
-\usepackage{algorithmicx} % this is used to write the pseudo code
-\usepackage[noend]{algpseudocode} % this is for the layout of the pseudo code
-\usepackage{tikz} % needed for: tikz pics
-% \usepackage[inline]{enumitem}
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{natbib}
-
-\usepackage{graphicx}  
-\usepackage{subfigure}
-\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
-\newcommand{\sys}{NAME\xspace}
-
-\begin{document}
-
-\title{VIPOS: Virtualized Intermittently Powered Runtime Environment (``VIPRE''?)}
-% \author{
-% 	\IEEEauthorblockN{Amjad Yousef Majid, Przemys{\l}aw Pawe{\l}czak}
-% 	\IEEEauthorblockA{TU Delft, Mekelweg\,4, 2628\,CD Delft, NL\\
-% 	\{a.y.majid, p.pawelczak\}@tudelft.nl}
-% }
-
-\maketitle
-
-\section{abstract}
-\TODO{Need to re-write this to make it consistent with the story after we have the story in place}
-Enabling battery-free devices is a mandatory step towards the environment friendly IoT.
-However, removing these large energy reservoirs (the batteries) makes the devices operate on an intermittent power supply, which make sustaining any long computation very challenging.
-correspondingly, there are two proposed approaches to enable long-running computations on intermittenly powered devices (IPDs): (i) Checkpointing based approach, where the volatile state of a program is frequently saved to the non-volatile memory and; (ii)  Task based approach, where a programmer splits the code into small idempotent modules.
-The Task based approach shows a better performance than the checkpointing approach.
-However, the task based method requires that the energy to execute any task must not exceed the maximum limit of the energy buffer.
-This limitation makes this solution a customized one that suits a particular energy buffer size.
-
-VIPOS enables code portability by means of virtualizing the task based model.
-If a code is written for a very small energy buffer, VIPOS is able to dynamically merge these tasks and constructs a bigger virtual task to improve the performance, reduce energy consumption and provides mean for code portability.
-
-\section{Introduction}
-\label{sec:intro}
-\TODO{VIPER: Virtualized Intermittent Programming model, Execution model, and Runtime environment}
-\TODO{PREVIEW: Programming and Runtime Environment with Virtualized Intermittent Execution Windows}
-
 Advances in processor efficiency along with the development of
@@
 \TODO{Bullet contribs if there's space}
-
-\section{Old Introduction}
-	Intermittently powered devices (IPDs) are battery-less devices that utilize the ambient energy to sense, compute and, communicate.
-For example, Wireless Identification and Sensing platform WISP \cite{wisp} uses the RF signal power to drive its computation and communication.
-Because of the reliance of TPDs on intermittent power supply, the harvested energy when it is available, their programs are susceptible to a very frequent power interrupts, in the order of tens of millisconds~\cite{}.
-Therefore, these devices require a different software execution model that complies with the nature of a discontinuous power supply.
-
-	The intermittent (discontinuous) execution model defines a program execution as cumulative discrete process.
-The main difference between the intermittent and the conventional (or continuous) execution models is that, a power failure is seen by the continuous model as an \emph{exception} that may reset the progress of a program to its beginning.
-Whereas, in intermittent execution a power failure is regarded as a temporary \emph{pause} to the execution that may result in some progress degradation.
-Generally we can classify the intermittent execution model into:
-	\begin{itemize}
-		\item \emph{Sequential Execution Model}:
-			Under this model a program is seen as one big idempotent operational region that has one common context.
-Generally, The progress of the program is saved and updated by means of checkpointing---where all the program context (e.g.
-CPU registers, the stack and the global variables) is saved to a non-volatile memory.
-Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos, harvOS, hibernus}.
-The benefit of this model is that it does not require code modification by a programmer.
-However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{xxxx}; and (iii) it restricts the IPDs to run only a single application~\cite{inos}.
-		\item \emph{Modular Intermittent Execution Model}:
-			At the heart of this model is the concept of an idempotent task.
-The idempotent task is C function that does not have arguments and does not return a value.
-This task uses a well defined interface to interact with the non-volatile memory.
-Therefore, it tolerates arbitrary number of power interrupts.
-This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an IPD by interleaving their tasks~\cite{inos}.
-However, it obviously requires code modification---for example, if an algorithm is written according to the continuous execution model it has to be splitted, by a programmer, into small tasks to run under the modular intermittent execution model.
-	\end{itemize}
-
-Despite that the superiority of the Modular Intermittent Execution model, it is still a static approach that completely depends on a programmer's estimation which is mostly result in a sub-optimal code devision.
-Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.
-
-% Energy source
-
-% The intermittent execution is cumulative discrete process.
-The intermittent execution model is only able to execute a few number of instructions, as compared to the conventional (continuous) model, before its progress is terminated.
-Therefore, the intermittent execution model adapts an execution progress state saving mechanism.
-This mechanism is realized either by means of a checkpoint~\cite{}, where all the context of the program is saved into non-volatile memory.
-Or
-
-% This progress state saving mechanism is normally injected into a program either by a compiler~\cite{}.
-A compiler injects trigger pointers to checkpoint (save) the context of the program into the non-volatile memory.
-or it is added by a programmer~\cite{chain}.
-[compare these two approaches]...
-
-% We define virtualization within the context of intermittent execution as utilizing the volatile memory instead of non-volatile when the intermittent execution model attempts to access the non-volatile memory.
-
-\subsection{Problem Statement}
-How to approach the optimal task size given an unknown energy buffer size and environment conditions? 
-	 \begin{enumerate}
-		 \item How to eliminate the need for the hardware support while enabling the Modular Intermittent Execution model to dynamically adapt its task size?
-		 \item How to enable code portability under the Modular Intermittent execution model? 
-		 \item How to reduce the energy consumption of executing a task? 
-	\end{enumerate}
-
-\subsection{Why not Adapting Current Approaches}
-Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes. 
-%
-%	\begin{figure}[t]
-%	    \centering
-%	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
-%	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
-%	    \caption{Task flow control of Chain}
-%	    \label{fig:DynamicChain}
-%	\end{figure}
-
-\noindent\textbf{Chain: Tasks and Channels for Reliable Intermittent Programs } \\
-Some of the main problems in modifying Chain to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Multi-versioning system:} Chain defines for each task its own input that is not shared with other tasks.
-This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
-	\item \emph{Direct FRAM system:} Chain tasks directly access FRAM which is more energy expensive Than SRAM.
-As result of being multi-versioning system that directly interacts with FRAM, Chain has a significant energy consumption overhead.
-	\item \emph{Irreversible tasks transitions:} According to the Chain programming model, each task calls the next one.
-The transition is done through functions calls and manually clearing the stack, to prevent the stack overflow problem.
-This approach make all the transitions firm ones which prevent task merging.
-	\item \emph{All channels are required:} Fig.~\ref{fig:DynamicChainSeq} shows that if the task control flow is a circular with no branches then task merging is possible.
-Moreover, all the intra-merged-tasks channel can be committed to SRAM instead of FRAM to save energy.
-However, this execution path is only a special.
-If the general case (see Fig.~\ref{fig:DynamicChainRan}) is considered then leaving out the intra-merged-task channels might result in data inconsistency---If an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready which result in an incorrect execution.
-
-\end{itemize}
-
-\noindent\textbf{Alpaca: Intermittent Execution without Checkpoints} \\
-Some of the main problems in modifying Alpaca to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Privatization:} The core concept of Alpaca is privatization.
-Privatization depends completely on detecting the variables that have a Write-after-read dependency within the scope of a task.
-Therefore, if tasks are on-demand merged the scope of these dependencies are changed and the static analysis to privatize these variables is not valid anymore and data consistency can not be preserved.
-
-	\item \emph{Irreversible tasks transitions:} Similar to Chain problem.
-
-\end{itemize}
-
-\noindent\textbf{Virtualizing Modular Intermittent Execution Model}
-
-		An Intermittently executed program, as seen by the modular intermittent execution model, is a chain of tasks with a firm transition from one to another. 
-		Virtualizing this model means softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
-Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
-Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
-This can be done by having a temporary copy of a global variable in the volatile memory (similar to the catching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
-
-		In this paper we are introducing VIPOS a runtime library the implement the Virtualed Intermittent Execution Model.
-VIPOS adapts two methods for data protection.
-One relies on a virtual buffer and the other uses the Direct Memory Access (DMA).
-VIPOS is able to approach the optimal tasks devision during the runtime.
-VIPOS reduces an application execution time by XX and the energy by XX.
-
-\subsection{Contributions}
-	 \begin{enumerate}
-		 \item We have developed an algorithm that is able to determine the size of a virtual task based on the history of the execution of an application. 
-		 \item We enable efficient code portability by requiring that a real task to small and merging them during the execution.   
-	\end{enumerate}
-
- \section{Preliminary Results}
- \label{sec:prelResults}
-
-\begin{figure}[t]
-    \centering
-         \subfigure[The energy cost of FRAM/SRAM writes operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
-         \subfigure[The energy cost of FRAM/SRAM read operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
-    \caption{The energy cost of accessing volatile/non-volatile memory}
-    \label{fig:framEnergy}
-\end{figure}
-
-Since the Intermittent execution models rely on the non-volatile memory to enable long-running operation, it is important to characterize the energy consumption of accessing this type of memory as compare to the volatile one.
-
-Four applications are developed to measure the energy consumption of \emph{Accessing} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{xxx}.
-The interference free debugger (EDB)~\cite{xxx} is used as the energy measurement tool.
-The EDB probes the energy buffer before and after accessing the FRAM/SRAM 1600 times, This large number of FRAM access is used to increase the reliability of the results generated by the used measurement tool (e.g.
-reducing the effect of the quantization error).
-The energy buffer is charged, using the EDB, to $\approx$ 2.45 volts \textbf{only} at the beginning of the execution and of the programs.
-The write operations were performed by writing literal values, random numbers, to the memory.
-However, a read operation must be followed by a write operation, therefore, we chose to write a read value to SRAM.
-Fig.~\ref{fig:framEnergy} shows that an FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We would like to mentioned that TI has already mentioned in this document~\cite{xxxx} that FRAM write access is more energy expensive than SRAM write access.
-However, TI does not quantify the energy overhead.}.
-Therefore, reducing FRAM access is desirable when developing  energy limited software solution
-
-\section{VIPOS: Virtualized Intermittently Powered Runtime Environment}
-VIPOS is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs.
-It is able, on-demand, to merge tasks to reduce the number of commits to the non-volatile memory.
-
-\subsection{VIPOS: Data/Memory Protection Methods}
-VIPOS data consistency preservation methods are designed with goal of minimal interaction between the CPU and FRAM to reduce energy consumption (see Section~\ref{sec:prelResults}).
-These data/memory protection methods enables VIPOS to tolerate arbirary number of power failures.
-
-	\subsubsection{Virtualized Operational Buffer}
-
-		\begin{algorithm}[t]
-			\caption{Virtualized Operational Buffer}
-			\label{algo:virtuBufWrite}
-			\scriptsize
-			%\small
-			\begin{algorithmic}[1]
-				\State $var \in \text{\{global variables\}} $ 
-				\State \label{lst:virtuBufWrite:line:begin}\Call{Virtual Task()}{} 
-				\While { \textit{executing} } \Comment{Execution stage}
-					\State $var$  $\rightarrow$ \textsf{volatile buffer} \label{lst:virtuBufWrite:line:output}
-					\If { $var$  in \textsf{ volatile buffer} }				\label{lst:virtuBufWrite:line:inputBegin}
-							\State $var$  $\leftarrow$  \textsf{volatile buffer} 
-						\Else 
-							\State $var$  $\leftarrow$  \textsf{FRAM}		\label{lst:virtuBufWrite:line:inputEnd}
-						\EndIf
-
-					\If {power interrupts}
-						\State back to \ref{lst:virtuBufWrite:line:begin}
-					\EndIf
-				\EndWhile
-
-				\While{  $\textsf{volatile buffer}\not=\emptyset$  } \label{lst:virtuBufWrite:line:commitBegin} \Comment{ First phase commit}
-					\State \textsf{volatile buffer} $\rightarrow$  \textsf{persistent buffer}
-					\If {power interrupts}
-						\State discard \textsf{persistent buffer}
-						\State back to  \ref{lst:virtuBufWrite:line:begin} 
-						\State   \label{lst:virtuBufWrite:line:commitEnd}
-					\EndIf
-				\EndWhile
-
-				\While{ $\textsf{persistent buffer}\not=\emptyset$ } \label{lst:virtuBufWrite:line:SecCommitBegin} \Comment{Second phase commit}
-					\State \textsf{persistent buffer} $\rightarrow$ FRAM 
-					\If {power interrupts}
-						\State Continue
-					\EndIf
-				\EndWhile 
-				\State     \label{lst:virtuBufWrite:line:SecCommitEnd}
-				\State return
-			\end{algorithmic}
-		\end{algorithm}
-
-		Since accessing FRAM is more energy expencive and slower than accessing SRAM, preventing frequent access to FRAM (i.e.
-during looping operations) is desireable.
-Therefore, the first proposed protection method utilizes a volatile buffer that holds temporary all the outputs of a virtual task (see Algorithm~\ref{algo:virtuBufWrite} line~\ref{lst:virtuBufWrite:line:output}).
-Consequently, a virtual task must first attempt to read a global variable from the volatile buffer before trying to obtain the value from the non-volatile memory (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:inputBegin}-\ref{lst:virtuBufWrite:line:inputEnd}) to ensure a correct execution progress and the consistency of the memory.
-Once the execution of a virtual task is done, the first phase of the commit process is started by copying the volatile buffer to a persistent buffer (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:commitBegin}-\ref{lst:virtuBufWrite:line:commitEnd}).
-If the power is interrupted the persistent buffer must be discarded and the execution must start again from the beginning of the virtual task.
-In another words, the first phase commit must be performed atomically to preserve the consistency of the output of a virtual task.
-The second phase commit is a power failure immune precess that is responsible for distributing the global variables to their final locations and make the non-volatile memory consistent and synchronized with the computation progress (see Algorithm~\ref{algo:virtuBufWrite}lines~\ref{lst:virtuBufWrite:line:SecCommitBegin}-\ref{lst:virtuBufWrite:line:SecCommitEnd}).
-
-	\paragraph{Virtualized Operational Buffer: Implementation}
-
-		We implement a reference implementation of VIPOS that uses virtualized Operational buffer to protect the data against power failure. 
-		\paragraph{Virtual Buffer}
-			We realized the virtaulzed buffer as a volatile hashed table of linked lists.
-The hashing technique was chosen to reduce the buffer searching time and the linked lists are used to prevent data loss when there is a conflict between multiple variables---If two variables have the same hash value they will occupy the same cell, however, with linked list a new node will be created for each variable to resolve the conflict and protect the data.
-The hashing function is based on the observation that the virtual addresses of the memory cells have approximately a flat distribution over the memory addressing space and the fact that each entry, of the virtual table, has the address and the value of a variable.
-As such, by using the least significant bits as an index to access the virtual buffer the hash function distributes its inputs uniformly in the virtual buffer.
-Accordingly, the linked lists search time, on average, is reduced.
-
-		\paragraph{Persistent Buffer}
-			The persistent buffer is implemented as static array of tuples, where each tuple holds  the address and the value of a variable.
-This data structure is very suitable for the second phase commit where each element has to be committed to its finial location.
-However, the complexity of committing this buffer is of size $O(N)$, where $N$ equals the length of the buffer.
-Therefore, the size of the buffer can have dramatic effect on the performance of VIPOS---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious.
-On the other hand, if the size of the buffer is large VIPOS will experience a significant performance degradation.
-However, by observing the type of the information that this buffer holds, namely memory addresses, we can, on average, reduce the complexity of committing this buffer by defining the \emph{effective size} of the buffer to be the size of the buffer up to the first cell that holds an invalid memory address.
-Moreover, since this buffer is in FRAM which is much bigger than SRAM its size can be relatively big.
-
-		\paragraph{Virtualized Operational Buffer: delay analysis}[initial results]
-
-		\begin{figure}[t]
-			\centering
-			\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
-			\caption{VIPOS based on virtualized Operational Buffer.
-It runs the decompression application with different buffer configurarions.
-EPB: Enhanced Persistent Buffer.
-NR: no read operation from the buffer.
-NRW: not read or write operations from the buffer.}
-			\label{fig:virtOperationalBuf}
-		\end{figure}
-
-			To analyze the performance of VIPOS versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
-We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}.
-The results presented in Fig.~\ref{fig:virtOperationalBuf} show the VIPOS has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.
-
-			All the experiments were done with a persistent buffer of size 100.
-However, the bar labeled with "EPB/16" in  Fig.~\ref{fig:virtOperationalBuf} shows the effect of the enhanced commit process that uses the buffer effective size instead of the buffer maximum size.
-We see that the execution time is reduced to $\approx$80 ms.
-Furthermore, we quantized the overhead of the read operations from the buffer, by eliminating this operation, and the bar labeled with "NR" shows that it cost  $\approx$10 ms.
-Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to VIPOS best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.
-
-\subsubsection{ Direct Memory Access Based Protection Method (better name is needed)}
-%
-	\begin{figure}[t]
-		\centering
-         \subfigure[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize.eps} }
-	     \subfigure[The energy needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
-		\caption{Time and energy consumption of moving a block of data from SRAM to FRAM}
-		\label{fig:dmaTimeEnergy}
-	\end{figure}
-%
-	Since accessing/updating buffers is more costly than directly accessing FRAM, at least without specialized hardware~\cite{clank}, we introduce the second method to apply the virtualization principle to the Modular Execution Model.
-This method utilizing the Direct Memory Access (DMA) Module.
-As Fig.~\ref{fig:dmaTimeEnergy} shows that DMA is much more efficient in transferring a block of data than the conventional (pointer based) way from the energy and time perspectives.
-
-	This method combines a number of techniques to preserve the persistence and consistency of the data.
-It backs up the status of the global variables in the non-volatile memory, and double buffers them to guarantee their consistency.
-Furthermore, on each power up it populates the global variables in the volatile memory from the most recent back up in the non-volatile memory to enable a faster task execution since the task does not need to interact with non-volatile memory which is generally slower and more energy expensive.
-
-\subsection{VIPOS: The Virtualizing Engine}
-\subsubsection{Power Interrupt Immune Scheduler}
-It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures.
-VIPOS provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
-(un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
-
-\subsubsection{Task Merging Algorithms}
-
-\paragraph{Fixed Virtual Task Size} [better name needed]
-
-	\begin{algorithm}[t]
-		\caption{Fixed virtual Task size}
-		\label{algo:fixVirtTask}
-		\scriptsize
-		%\small
-		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{VIPOS Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\State MVTS: maximum VT size
-			\vspace{0.1cm}
-
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
-				\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-							\State $ MVTS = VTS $
-						\EndIf
-				\EndWhile
-
-				\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\If{$VTS < MVTS$}
-					\State $VTS++$
-					\EndIf
-				\EndIf
-			\EndWhile
-		\end{algorithmic}
-	\end{algorithm}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{figures/virtualTaskSize.eps}
-	\caption{Size of the virtual task versus the execution time of a dummy application that contains 12 empty tasks.}
-	\label{fig:virtualTaskSize}
-\end{figure}
-
-VIPOS is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
-Fig.~\ref{fig:virtualTaskSize} show the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
-
-Remark: A less obvious benefit of visualization is that it can enable a secure computation.
-By increasing the size of the virtual task, the energy in the super-capacitor will not be enough to finish the execution of the virtual task.
-Therefore, without the interrogator being close to the intermittently powered device and charging rate is not negligible as compared to the discharging rate this computation will not be finished.
-As such, virtualization can add a layer of security to the computation.
-
-\section{Results}
-We compare the performance of VIPOS against a state-of-the-art intermittent execution approach.
-For this purpose, we used two different applications: (i) A data decompression application, which utilizes Huffman decoding technique to decompress 100 byte of data.; (ii) A discrete Fourier transform (DFT) application that uses two different resolution (4 and 8 bites) to analyze a randomly generated signal.
-
-\subsection{Continuous Power Supply}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{figures/executionTime.eps}
-	\caption{Time to complation of VIPOS versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
-	\label{fig:IPOSPerformance}
-\end{figure}
-
-Fig.~\ref{fig:IPOSPerformance} shows that VIPOS requires a shorter execution time to execute an application under the MIEM.
-The size of the virtual task that VIPOS uses equals One real task, in other words, no visualization is applied.
-
-\subsection{VIPOS Operational Buffer Versus VIPOS DMA}[place holder]
-			\begin{figure}[t]
-				\centering
-				\includegraphics[width=0.7\columnwidth]{figures/ipos_virtualizedBuffer_overhead.eps}
-				\caption{Virtualized Oprational buffer overhead.}
-				\label{fig:virtualBuffer}
-			\end{figure}
-
-\subsection{Intermittent Power Supply}
-...
-
-\bibliographystyle{abbrvnat}
-\bibliography{bib}
-
-\end{document}

### 2017-06-29 paper/paper.tex
modularized.  next thing to do is do some outlining for the first few sections

--- 2017-06-29 paper/paper.tex
@@
 
+\begin{abstract}
+\input{abstract}
+\end{abstract}
+
 \maketitle
-
-\section{abstract}
-Enabling battery-free devices is a mandatory step towards the environment friendly IoT.
-However, removing these large energy reservoirs (the batteries) makes the devices operate on an intermittent power supply, which make sustaining any long computation very challenging.
-correspondingly, there are two proposed approaches to enable long-running computations on intermittenly powered devices (IPDs): (i) Checkpointing based approach, where the volatile state of a program is frequently saved to the non-volatile memory and; (ii)  Task based approach, where a programmer splits the code into small idempotent modules.
-The Task based approach shows a better performance than the checkpointing approach.
-However, the task based method requires that the energy to execute any task must not exceed the maximum limit of the energy buffer.
-This limitation makes this solution a customized one that suits a particular energy buffer size.
-
-VIPOS enables code portability by means of virtualizing the task based model.
-If a code is written for a very small energy buffer, VIPOS is able to dynamically merge these tasks and constructs a bigger virtual task to improve the performance, reduce energy consumption and provides mean for code portability.
 
@@
 \label{sec:intro}
+\TODO{VIPER: Virtualized Intermittent Programming model, Execution model, and Runtime environment}
+\TODO{PREVIEW: Programming and Runtime Environment with Virtualized Intermittent Execution Windows}
 
-Advances in processor efficiency along with the development of
-energy-harvesting power systems has created a new category of devices that
-require neither a battery nor a tethered power supply.  These devices operate
-entirely using energy extracted from their environment, such as RF energy,
-photovoltaics, and vibration.  Incorporating compute, storage, sensing, and
-communication hardware, such energy-independent devices are a promising
-candidate as the underlying technology for emerging internet of things (IoT)
-applications, in- and on-body medical applications, and emerging applications,
-like energy-harvesting nano-satellites.
-
-While promising, energy-harvesting devices present a unique design challenge in
-that they operate only {\em intermittently} when energy is available.   An
-energy-harvesting device buffers energy from its environment, e.g., in a
-capacitor. When a threshold quantity of energy accumulates, the device begins
-operating.  Harvestable energy sources typically produce relatively very little
-power compared to a platform's operating power level.  A devices operates for a
-brief period of time and when buffered energy is depeleted, shuts down and
-begins recharging to later operate again.  Charge and discharge times vary by
-device and some devices~\cite{wisp} may fail 10 to 100 times per second.
-
-Software in an energy-harvesting system operates according to the {\em
-intermittent execution model}~\cite{dino}, with bursts of execution that are
-interrupted by failure periods. When power fails, a device loses its volatile
-state (i.e., registers, stack, SRAM) and retains its non-volatile state (i.e.,
-FRAM, Flash). While capturing periodic checkpoints~\cite{mementos,tictpl,quickrecall}
-and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve
-execution progress, failures can leave volatile and non-volatile state inconsistent,
-leading to unrecoverable failures~\cite{mspcdino,edb,edbtoppicks}.
-
-\TODO{Don't forget all the references in this section!}
-New programming and execution models~\cite{dino,ratchet,chain,alpaca} and hardware
-mechanisms~\cite{clank, idetic} help avoid data inconsistency
+\input{intro}
 

### 2017-07-03 paper/iposDataProtection.tex
minor changes

--- 2017-06-29 paper/paper.tex
@@
-\documentclass[sigconf,anonymous,review]{acmart}
-\usepackage[normalem]{ulem}
-
-\usepackage{amssymb} % needed for: various symbols
-\usepackage{amsmath} % needed for: math symbols
-% http://tex.stackexchange.com/questions/229355/algorithm-algorithmic-algorithmicx-algorithm2e-algpseudocode-confused
-\usepackage{algorithm} % this used as an environment for the algorithm code (linke figure and table)
-\usepackage{algorithmicx} % this is used to write the pseudo code
-\usepackage[noend]{algpseudocode} % this is for the layout of the pseudo code
-\usepackage{tikz} % needed for: tikz pics
-% \usepackage[inline]{enumitem}
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{natbib}
-
-\usepackage{graphicx}  
-\usepackage{subfigure}
-\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
-\newcommand{\sys}{NAME\xspace}
-
-\begin{document}
-
-\title{VIPOS: Virtualized Intermittently Powered Runtime Environment (``VIPRE''?)}
-% \author{
-% 	\IEEEauthorblockN{Amjad Yousef Majid, Przemys{\l}aw Pawe{\l}czak}
-% 	\IEEEauthorblockA{TU Delft, Mekelweg\,4, 2628\,CD Delft, NL\\
-% 	\{a.y.majid, p.pawelczak\}@tudelft.nl}
-% }
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:intro}
-\TODO{VIPER: Virtualized Intermittent Programming model, Execution model, and Runtime environment}
-\TODO{PREVIEW: Programming and Runtime Environment with Virtualized Intermittent Execution Windows}
-
-\input{intro}
-
-\section{Old Introduction}
-	Intermittently powered devices (IPDs) are battery-less devices that utilize the ambient energy to sense, compute and, communicate.
-For example, Wireless Identification and Sensing platform WISP \cite{wisp} uses the RF signal power to drive its computation and communication.
-Because of the reliance of TPDs on intermittent power supply, the harvested energy when it is available, their programs are susceptible to a very frequent power interrupts, in the order of tens of millisconds~\cite{}.
-Therefore, these devices require a different software execution model that complies with the nature of a discontinuous power supply.
-
-	The intermittent (discontinuous) execution model defines a program execution as cumulative discrete process.
-The main difference between the intermittent and the conventional (or continuous) execution models is that, a power failure is seen by the continuous model as an \emph{exception} that may reset the progress of a program to its beginning.
-Whereas, in intermittent execution a power failure is regarded as a temporary \emph{pause} to the execution that may result in some progress degradation.
-Generally we can classify the intermittent execution model into:
-	\begin{itemize}
-		\item \emph{Sequential Execution Model}:
-			Under this model a program is seen as one big idempotent operational region that has one common context.
-Generally, The progress of the program is saved and updated by means of checkpointing---where all the program context (e.g.
-CPU registers, the stack and the global variables) is saved to a non-volatile memory.
-Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos, harvOS, hibernus}.
-The benefit of this model is that it does not require code modification by a programmer.
-However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{xxxx}; and (iii) it restricts the IPDs to run only a single application~\cite{inos}.
-		\item \emph{Modular Intermittent Execution Model}:
-			At the heart of this model is the concept of an idempotent task.
-The idempotent task is C function that does not have arguments and does not return a value.
-This task uses a well defined interface to interact with the non-volatile memory.
-Therefore, it tolerates arbitrary number of power interrupts.
-This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an IPD by interleaving their tasks~\cite{inos}.
-However, it obviously requires code modification---for example, if an algorithm is written according to the continuous execution model it has to be splitted, by a programmer, into small tasks to run under the modular intermittent execution model.
-	\end{itemize}
-
-Despite that the superiority of the Modular Intermittent Execution model, it is still a static approach that completely depends on a programmer's estimation which is mostly result in a sub-optimal code devision.
-Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.
-
-% Energy source
-
-% The intermittent execution is cumulative discrete process.
-The intermittent execution model is only able to execute a few number of instructions, as compared to the conventional (continuous) model, before its progress is terminated.
-Therefore, the intermittent execution model adapts an execution progress state saving mechanism.
-This mechanism is realized either by means of a checkpoint~\cite{}, where all the context of the program is saved into non-volatile memory.
-Or
-
-% This progress state saving mechanism is normally injected into a program either by a compiler~\cite{}.
-A compiler injects trigger pointers to checkpoint (save) the context of the program into the non-volatile memory.
-or it is added by a programmer~\cite{chain}.
-[compare these two approaches]...
-
-% We define virtualization within the context of intermittent execution as utilizing the volatile memory instead of non-volatile when the intermittent execution model attempts to access the non-volatile memory.
-
-\subsection{Problem Statement}
-How to approach the optimal task size given an unknown energy buffer size and environment conditions? 
-	 \begin{enumerate}
-		 \item How to eliminate the need for the hardware support while enabling the Modular Intermittent Execution model to dynamically adapt its task size?
-		 \item How to enable code portability under the Modular Intermittent execution model? 
-		 \item How to reduce the energy consumption of executing a task? 
-	\end{enumerate}
-
-\subsection{Why not Adapting Current Approaches}
-Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes. 
-%
-%	\begin{figure}[t]
-%	    \centering
-%	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
-%	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
-%	    \caption{Task flow control of Chain}
-%	    \label{fig:DynamicChain}
-%	\end{figure}
-
-\noindent\textbf{Chain: Tasks and Channels for Reliable Intermittent Programs } \\
-Some of the main problems in modifying Chain to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Multi-versioning system:} Chain defines for each task its own input that is not shared with other tasks.
-This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
-	\item \emph{Direct FRAM system:} Chain tasks directly access FRAM which is more energy expensive Than SRAM.
-As result of being multi-versioning system that directly interacts with FRAM, Chain has a significant energy consumption overhead.
-	\item \emph{Irreversible tasks transitions:} According to the Chain programming model, each task calls the next one.
-The transition is done through functions calls and manually clearing the stack, to prevent the stack overflow problem.
-This approach make all the transitions firm ones which prevent task merging.
-	\item \emph{All channels are required:} Fig.~\ref{fig:DynamicChainSeq} shows that if the task control flow is a circular with no branches then task merging is possible.
-Moreover, all the intra-merged-tasks channel can be committed to SRAM instead of FRAM to save energy.
-However, this execution path is only a special.
-If the general case (see Fig.~\ref{fig:DynamicChainRan}) is considered then leaving out the intra-merged-task channels might result in data inconsistency---If an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready which result in an incorrect execution.
-
-\end{itemize}
-
-\noindent\textbf{Alpaca: Intermittent Execution without Checkpoints} \\
-Some of the main problems in modifying Alpaca to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Privatization:} The core concept of Alpaca is privatization.
-Privatization depends completely on detecting the variables that have a Write-after-read dependency within the scope of a task.
-Therefore, if tasks are on-demand merged the scope of these dependencies are changed and the static analysis to privatize these variables is not valid anymore and data consistency can not be preserved.
-
-	\item \emph{Irreversible tasks transitions:} Similar to Chain problem.
-
-\end{itemize}
-
-\noindent\textbf{Virtualizing Modular Intermittent Execution Model}
-
-		An Intermittently executed program, as seen by the modular intermittent execution model, is a chain of tasks with a firm transition from one to another. 
-		Virtualizing this model means softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
-Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
-Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
-This can be done by having a temporary copy of a global variable in the volatile memory (similar to the catching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
-
-		In this paper we are introducing VIPOS a runtime library the implement the Virtualed Intermittent Execution Model.
-VIPOS adapts two methods for data protection.
-One relies on a virtual buffer and the other uses the Direct Memory Access (DMA).
-VIPOS is able to approach the optimal tasks devision during the runtime.
-VIPOS reduces an application execution time by XX and the energy by XX.
-
-\subsection{Contributions}
-	 \begin{enumerate}
-		 \item We have developed an algorithm that is able to determine the size of a virtual task based on the history of the execution of an application. 
-		 \item We enable efficient code portability by requiring that a real task to small and merging them during the execution.   
-	\end{enumerate}
-
- \section{Preliminary Results}
- \label{sec:prelResults}
+\subsection{VIPOS: Data/Memory Protection Methods}
 
@@
 	\centering
-         \subfigure[The energy cost of FRAM/SRAM writes operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
-         \subfigure[The energy cost of FRAM/SRAM read operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
-    \caption{The energy cost of accessing volatile/non-volatile memory}
-    \label{fig:framEnergy}
+	\includegraphics[width=0.8\columnwidth]{figures/iposCommitSize}
+	\caption{...}
+	\label{fig:virtOperationalBuf}
 \end{figure}
 
-Since the Intermittent execution models rely on the non-volatile memory to enable long-running operation, it is important to characterize the energy consumption of accessing this type of memory as compare to the volatile one.
-
-Four applications are developed to measure the energy consumption of \emph{Accessing} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{xxx}.
-The interference free debugger (EDB)~\cite{xxx} is used as the energy measurement tool.
-The EDB probes the energy buffer before and after accessing the FRAM/SRAM 1600 times, This large number of FRAM access is used to increase the reliability of the results generated by the used measurement tool (e.g.
-reducing the effect of the quantization error).
-The energy buffer is charged, using the EDB, to $\approx$ 2.45 volts \textbf{only} at the beginning of the execution and of the programs.
-The write operations were performed by writing literal values, random numbers, to the memory.
-However, a read operation must be followed by a write operation, therefore, we chose to write a read value to SRAM.
-Fig.~\ref{fig:framEnergy} shows that an FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We would like to mentioned that TI has already mentioned in this document~\cite{xxxx} that FRAM write access is more energy expensive than SRAM write access.
-However, TI does not quantify the energy overhead.}.
-Therefore, reducing FRAM access is desirable when developing  energy limited software solution
-
-\section{VIPOS: Virtualized Intermittently Powered Runtime Environment}
-VIPOS is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs.
-It is able, on-demand, to merge tasks to reduce the number of commits to the non-volatile memory.
-
-\subsection{VIPOS: Data/Memory Protection Methods}
 VIPOS data consistency preservation methods are designed with goal of minimal interaction between the CPU and FRAM to reduce energy consumption (see Section~\ref{sec:prelResults}).
-These data/memory protection methods enables VIPOS to tolerate arbirary number of power failures.
+These data/memory protection methods enables VIPOS to tolerate arbitrary number of power failures.
 
@@
 
-		Since accessing FRAM is more energy expencive and slower than accessing SRAM, preventing frequent access to FRAM (i.e.
-during looping operations) is desireable.
+		Since accessing FRAM is more energy expensive and slower than accessing SRAM, preventing frequent access to FRAM (i.e.
+during looping operations) is desirable.
 Therefore, the first proposed protection method utilizes a volatile buffer that holds temporary all the outputs of a virtual task (see Algorithm~\ref{algo:virtuBufWrite} line~\ref{lst:virtuBufWrite:line:output}).
@@
 
+			Despite the fact that the linked lists reduce the average complexity of searching the virtual buffer significantly, they introduce a number of drawbacks: (i) the linked list data structure introduce a non-negligible memory overhead while it relies on a limited memory section, namely the heap; (ii) traversing the linked list is relatively slow.
+Therefore, we re-implemented the virtual buffer, after observing that most of the global variables have the same most significant bits, as a bigger hashed table that does not allow entries conflict.
+To eliminate the need for searching the hashed table linearly while committing its entries, we add a new buffer that holds the indices of the occupied cells of the hashed table.
+This table is of the same length as the hashed table.
+We can reason about the benefit of this buffer as follows: if we assume that the global variables are evenly divided between the tasks, then this buffer reduces the commit time by a factor equal to the number of tasks.
+
 		\paragraph{Persistent Buffer}
@@
 
-		\paragraph{Virtualized Operational Buffer: delay analysis}[initial results]
+		% \paragraph{Virtualized Operational Buffer: delay analysis}[initial results]
 
-		\begin{figure}[t]
-			\centering
-			\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
-			\caption{VIPOS based on virtualized Operational Buffer.
+		% \begin{figure}[t]
+		% 	\centering
+		% 	\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
+		% 	\caption{VIPOS based on virtualized Operational Buffer.
 It runs the decompression application with different buffer configurarions.
@@
 NRW: not read or write operations from the buffer.}
-			\label{fig:virtOperationalBuf}
-		\end{figure}
+		% 	\label{fig:virtOperationalBuf}
+		% \end{figure}
 
-			To analyze the performance of VIPOS versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
+		% 	To analyze the performance of VIPOS versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
 We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}.
@@
 
-			All the experiments were done with a persistent buffer of size 100.
+		% 	All the experiments were done with a persistent buffer of size 100.
 However, the bar labeled with "EPB/16" in  Fig.~\ref{fig:virtOperationalBuf} shows the effect of the enhanced commit process that uses the buffer effective size instead of the buffer maximum size.
@@
 		\centering
-         \subfigure[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize.eps} }
+         \subfigure[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize_time.eps} }
 	     \subfigure[The energy needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
@@
 Furthermore, on each power up it populates the global variables in the volatile memory from the most recent back up in the non-volatile memory to enable a faster task execution since the task does not need to interact with non-volatile memory which is generally slower and more energy expensive.
-
-\subsection{VIPOS: The Virtualizing Engine}
-\subsubsection{Power Interrupt Immune Scheduler}
-It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures.
-VIPOS provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
-(un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
-
-\subsubsection{Task Merging Algorithms}
-
-\paragraph{Fixed Virtual Task Size} [better name needed]
-
-	\begin{algorithm}[t]
-		\caption{Fixed virtual Task size}
-		\label{algo:fixVirtTask}
-		\scriptsize
-		%\small
-		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{VIPOS Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\State MVTS: maximum VT size
-			\vspace{0.1cm}
-
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
-				\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-							\State $ MVTS = VTS $
-						\EndIf
-				\EndWhile
-
-				\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\If{$VTS < MVTS$}
-					\State $VTS++$
-					\EndIf
-				\EndIf
-			\EndWhile
-		\end{algorithmic}
-	\end{algorithm}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{figures/virtualTaskSize.eps}
-	\caption{Size of the virtual task versus the execution time of a dummy application that contains 12 empty tasks.}
-	\label{fig:virtualTaskSize}
-\end{figure}
-
-VIPOS is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
-Fig.~\ref{fig:virtualTaskSize} show the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
-
-Remark: A less obvious benefit of visualization is that it can enable a secure computation.
-By increasing the size of the virtual task, the energy in the super-capacitor will not be enough to finish the execution of the virtual task.
-Therefore, without the interrogator being close to the intermittently powered device and charging rate is not negligible as compared to the discharging rate this computation will not be finished.
-As such, virtualization can add a layer of security to the computation.
-
-\section{Results}
-We compare the performance of VIPOS against a state-of-the-art intermittent execution approach.
-For this purpose, we used two different applications: (i) A data decompression application, which utilizes Huffman decoding technique to decompress 100 byte of data.; (ii) A discrete Fourier transform (DFT) application that uses two different resolution (4 and 8 bites) to analyze a randomly generated signal.
-
-\subsection{Continuous Power Supply}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.7\columnwidth]{figures/executionTime.eps}
-	\caption{Time to complation of VIPOS versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
-	\label{fig:IPOSPerformance}
-\end{figure}
-
-Fig.~\ref{fig:IPOSPerformance} shows that VIPOS requires a shorter execution time to execute an application under the MIEM.
-The size of the virtual task that VIPOS uses equals One real task, in other words, no visualization is applied.
-
-\subsection{VIPOS Operational Buffer Versus VIPOS DMA}[place holder]
-			\begin{figure}[t]
-				\centering
-				\includegraphics[width=0.7\columnwidth]{figures/ipos_virtualizedBuffer_overhead.eps}
-				\caption{Virtualized Oprational buffer overhead.}
-				\label{fig:virtualBuffer}
-			\end{figure}
-
-\subsection{Intermittent Power Supply}
-...
-
-\bibliographystyle{abbrvnat}
-\bibliography{bib}
-
-\end{document}

### 2017-07-03 paper/paper.tex
minor changes

@@
     \centering
-         \subfigure[The energy cost of FRAM/SRAM writes operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
-         \subfigure[The energy cost of FRAM/SRAM read operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
+         \subfigure[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
+         \subfigure[The energy cost of FRAM/SRAM \textbf{read} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
     \caption{The energy cost of accessing volatile/non-volatile memory}
@@
 
-\subsection{VIPOS: Data/Memory Protection Methods}
-VIPOS data consistency preservation methods are designed with goal of minimal interaction between the CPU and FRAM to reduce energy consumption (see Section~\ref{sec:prelResults}).
-These data/memory protection methods enables VIPOS to tolerate arbirary number of power failures.
-
-	\subsubsection{Virtualized Operational Buffer}
-
-		\begin{algorithm}[t]
-			\caption{Virtualized Operational Buffer}
-			\label{algo:virtuBufWrite}
-			\scriptsize
-			%\small
-			\begin{algorithmic}[1]
-				\State $var \in \text{\{global variables\}} $ 
-				\State \label{lst:virtuBufWrite:line:begin}\Call{Virtual Task()}{} 
-				\While { \textit{executing} } \Comment{Execution stage}
-					\State $var$  $\rightarrow$ \textsf{volatile buffer} \label{lst:virtuBufWrite:line:output}
-					\If { $var$  in \textsf{ volatile buffer} }				\label{lst:virtuBufWrite:line:inputBegin}
-							\State $var$  $\leftarrow$  \textsf{volatile buffer} 
-						\Else 
-							\State $var$  $\leftarrow$  \textsf{FRAM}		\label{lst:virtuBufWrite:line:inputEnd}
-						\EndIf
-
-					\If {power interrupts}
-						\State back to \ref{lst:virtuBufWrite:line:begin}
-					\EndIf
-				\EndWhile
-
-				\While{  $\textsf{volatile buffer}\not=\emptyset$  } \label{lst:virtuBufWrite:line:commitBegin} \Comment{ First phase commit}
-					\State \textsf{volatile buffer} $\rightarrow$  \textsf{persistent buffer}
-					\If {power interrupts}
-						\State discard \textsf{persistent buffer}
-						\State back to  \ref{lst:virtuBufWrite:line:begin} 
-						\State   \label{lst:virtuBufWrite:line:commitEnd}
-					\EndIf
-				\EndWhile
-
-				\While{ $\textsf{persistent buffer}\not=\emptyset$ } \label{lst:virtuBufWrite:line:SecCommitBegin} \Comment{Second phase commit}
-					\State \textsf{persistent buffer} $\rightarrow$ FRAM 
-					\If {power interrupts}
-						\State Continue
-					\EndIf
-				\EndWhile 
-				\State     \label{lst:virtuBufWrite:line:SecCommitEnd}
-				\State return
-			\end{algorithmic}
-		\end{algorithm}
-
-		Since accessing FRAM is more energy expencive and slower than accessing SRAM, preventing frequent access to FRAM (i.e.
-during looping operations) is desireable.
-Therefore, the first proposed protection method utilizes a volatile buffer that holds temporary all the outputs of a virtual task (see Algorithm~\ref{algo:virtuBufWrite} line~\ref{lst:virtuBufWrite:line:output}).
-Consequently, a virtual task must first attempt to read a global variable from the volatile buffer before trying to obtain the value from the non-volatile memory (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:inputBegin}-\ref{lst:virtuBufWrite:line:inputEnd}) to ensure a correct execution progress and the consistency of the memory.
-Once the execution of a virtual task is done, the first phase of the commit process is started by copying the volatile buffer to a persistent buffer (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:commitBegin}-\ref{lst:virtuBufWrite:line:commitEnd}).
-If the power is interrupted the persistent buffer must be discarded and the execution must start again from the beginning of the virtual task.
-In another words, the first phase commit must be performed atomically to preserve the consistency of the output of a virtual task.
-The second phase commit is a power failure immune precess that is responsible for distributing the global variables to their final locations and make the non-volatile memory consistent and synchronized with the computation progress (see Algorithm~\ref{algo:virtuBufWrite}lines~\ref{lst:virtuBufWrite:line:SecCommitBegin}-\ref{lst:virtuBufWrite:line:SecCommitEnd}).
-
-	\paragraph{Virtualized Operational Buffer: Implementation}
-
-		We implement a reference implementation of VIPOS that uses virtualized Operational buffer to protect the data against power failure. 
-		\paragraph{Virtual Buffer}
-			We realized the virtaulzed buffer as a volatile hashed table of linked lists.
-The hashing technique was chosen to reduce the buffer searching time and the linked lists are used to prevent data loss when there is a conflict between multiple variables---If two variables have the same hash value they will occupy the same cell, however, with linked list a new node will be created for each variable to resolve the conflict and protect the data.
-The hashing function is based on the observation that the virtual addresses of the memory cells have approximately a flat distribution over the memory addressing space and the fact that each entry, of the virtual table, has the address and the value of a variable.
-As such, by using the least significant bits as an index to access the virtual buffer the hash function distributes its inputs uniformly in the virtual buffer.
-Accordingly, the linked lists search time, on average, is reduced.
-
-		\paragraph{Persistent Buffer}
-			The persistent buffer is implemented as static array of tuples, where each tuple holds  the address and the value of a variable.
-This data structure is very suitable for the second phase commit where each element has to be committed to its finial location.
-However, the complexity of committing this buffer is of size $O(N)$, where $N$ equals the length of the buffer.
-Therefore, the size of the buffer can have dramatic effect on the performance of VIPOS---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious.
-On the other hand, if the size of the buffer is large VIPOS will experience a significant performance degradation.
-However, by observing the type of the information that this buffer holds, namely memory addresses, we can, on average, reduce the complexity of committing this buffer by defining the \emph{effective size} of the buffer to be the size of the buffer up to the first cell that holds an invalid memory address.
-Moreover, since this buffer is in FRAM which is much bigger than SRAM its size can be relatively big.
-
-		\paragraph{Virtualized Operational Buffer: delay analysis}[initial results]
-
-		\begin{figure}[t]
-			\centering
-			\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
-			\caption{VIPOS based on virtualized Operational Buffer.
-It runs the decompression application with different buffer configurarions.
-EPB: Enhanced Persistent Buffer.
-NR: no read operation from the buffer.
-NRW: not read or write operations from the buffer.}
-			\label{fig:virtOperationalBuf}
-		\end{figure}
-
-			To analyze the performance of VIPOS versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
-We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}.
-The results presented in Fig.~\ref{fig:virtOperationalBuf} show the VIPOS has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.
-
-			All the experiments were done with a persistent buffer of size 100.
-However, the bar labeled with "EPB/16" in  Fig.~\ref{fig:virtOperationalBuf} shows the effect of the enhanced commit process that uses the buffer effective size instead of the buffer maximum size.
-We see that the execution time is reduced to $\approx$80 ms.
-Furthermore, we quantized the overhead of the read operations from the buffer, by eliminating this operation, and the bar labeled with "NR" shows that it cost  $\approx$10 ms.
-Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to VIPOS best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.
-
-\subsubsection{ Direct Memory Access Based Protection Method (better name is needed)}
-%
-	\begin{figure}[t]
-		\centering
-         \subfigure[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize.eps} }
-	     \subfigure[The energy needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
-		\caption{Time and energy consumption of moving a block of data from SRAM to FRAM}
-		\label{fig:dmaTimeEnergy}
-	\end{figure}
-%
-	Since accessing/updating buffers is more costly than directly accessing FRAM, at least without specialized hardware~\cite{clank}, we introduce the second method to apply the virtualization principle to the Modular Execution Model.
-This method utilizing the Direct Memory Access (DMA) Module.
-As Fig.~\ref{fig:dmaTimeEnergy} shows that DMA is much more efficient in transferring a block of data than the conventional (pointer based) way from the energy and time perspectives.
-
-	This method combines a number of techniques to preserve the persistence and consistency of the data.
-It backs up the status of the global variables in the non-volatile memory, and double buffers them to guarantee their consistency.
-Furthermore, on each power up it populates the global variables in the volatile memory from the most recent back up in the non-volatile memory to enable a faster task execution since the task does not need to interact with non-volatile memory which is generally slower and more energy expensive.
+\input{iposDataProtection}
 
@@
 
+	\begin{algorithm}[t]
+		\caption{Opportunistic virtual Task size}
+		\label{algo:fixVirtTask}
+		\scriptsize
+		%\small
+		\begin{algorithmic}[1]
+			\State $VT \subset \text{\{VIPOS Tasks\}} $  \Comment{$VT:$ Virtual Task}
+			\State VTS : VT size
+			\vspace{0.1cm}
+
+			\While {$True$}
+				\State $VT \leftarrow VT_{next}$
+				\vspace{0.1cm}
+				\While {execute $VT$} 
+					\If { $\text{power failed twice}$ }				
+							\State $VTS--$  
+						\EndIf
+				\EndWhile
+
+				\vspace{0.1cm}
+				\If {$ \text{All tasks executed}$}
+					\State $VTS++$
+				\EndIf
+			\EndWhile
+		\end{algorithmic}
+	\end{algorithm}
+
 \begin{figure}[t]
 	\centering
-	\includegraphics[width=0.7\columnwidth]{figures/virtualTaskSize.eps}
+	\includegraphics[width=0.8\columnwidth]{figures/virtualTaskSize.eps}
 	\caption{Size of the virtual task versus the execution time of a dummy application that contains 12 empty tasks.}
@@
 	\centering
-	\includegraphics[width=0.7\columnwidth]{figures/executionTime.eps}
+	\includegraphics[width=0.8\columnwidth]{figures/ipos_chain}
 	\caption{Time to complation of VIPOS versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
@@
 The size of the virtual task that VIPOS uses equals One real task, in other words, no visualization is applied.
-
-\subsection{VIPOS Operational Buffer Versus VIPOS DMA}[place holder]
-			\begin{figure}[t]
-				\centering
-				\includegraphics[width=0.7\columnwidth]{figures/ipos_virtualizedBuffer_overhead.eps}
-				\caption{Virtualized Oprational buffer overhead.}
-				\label{fig:virtualBuffer}
-			\end{figure}
 

### 2017-07-03 DMA/ipos-gcc-dma/ipos/dataProtec.c

@@ -7,7 +7,7 @@
 
 
  // THIS IS ALSO DEFINED IN THE LINKER FILE
-#define BLOCKSIZE 0x64           // The block size of the transfered data in WORDS
+#define BLOCKSIZE 128           // The block size of the transfered data in WORDS
 #define ROM_2   (0xFF7F - BLOCKSIZE - BLOCKSIZE)     //we need to duplicate the number of bytes, since the transfer is in words
 #define ROM_1   (0xFF7F - BLOCKSIZE - BLOCKSIZE- BLOCKSIZE - BLOCKSIZE- 0x2)  // DMA transfer more than specified, therefore we added offset
 #define RAM_VARS_ADDR  0x2000

### 2017-07-03 DMA/largeDataCommit/main.c

@@ -2,8 +2,10 @@
 #include <msp430fr5969.h>
 #include <ipos.h>
 
-#define ARRSIZE 50
-__v volatile uint16_t data[ARRSIZE] = {0};
+//#define ARRSIZE 50
+#define ARRSIZE 128
+#define FIX_ARRSIZE 128
+__v volatile uint16_t data[FIX_ARRSIZE] = {0};
 
 void init();
 void task0();
@@ -20,6 +22,7 @@ void init()
     PM5CTL0 &= ~LOCKLPM5;       // Lock LPM5.
 
     P3DIR |=BIT5;
+    P4DIR |=BIT0;
 }
 
 /********************************
@@ -76,6 +79,14 @@ void task4()
 }
 void task5()
 {
+
+    if(data[1] == data[50])
+    {
+        P4OUT |= BIT0;
+        __delay_cycles(5000);
+        P4OUT &= ~BIT0;
+    }
+
     volatile uint16_t i;
     for (i = 0; i < ARRSIZE ; i++)
     {

### 2017-07-03 experiments/IPOS-performance_vs_commit-size/IPOS-performance_vs_commit-size.py

@@ -1,0 +1,139 @@
+"""
+IPOS versions: performance
+    Metric : time
+    Task   : manipulate and commit a certain number of variables
+
+@ Author  : Amjad Yousef Majid
+@ Gmail ID: amjad.y.majid 
+@ Date    : 2/July/2017 
+"""
+
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+
+def executionTime(fh, diffs):
+    numbers = []
+    for num in fh:
+        if num[-2:-1] == '1':
+            #print(num)
+            numbers.append( float(num.split(',')[0] ) )
+    aDiff = 0
+    for i in range(0, len(numbers)-1, 2):
+        aDiff += (numbers[i+1] - numbers[i])
+    aDiff *=1000  # make the diff in milliseconds
+    
+    
+    diffs.append(aDiff / int(len(numbers)/2) )
+#     print(diffs[:10])
+#     exit()
+    fh.close()
+
+data_points =7
+
+dma_diffs=[]
+ipos_ml_diffs =[]
+ipos_pl_diffs =[]
+
+indx = [1, 2, 4, 8, 16, 32, 64, 128]
+
+for f_idx in range(data_points):
+    fh = open('../DMA_128words/{}var.csv'.format(indx[f_idx]) , 'r')
+    executionTime(fh, dma_diffs)
+    fh = open('../ipos-ml/{}var.csv'.format(indx[f_idx]) , 'r')
+    executionTime(fh, ipos_ml_diffs)
+    if f_idx != 7: 
+        fh = open('../ipos-pl/{}var.csv'.format(indx[f_idx]) , 'r')
+        executionTime(fh, ipos_pl_diffs)
+        
+
+# ipos_pl_diffs.append(0)
+    
+#set figure linewidth
+# plt.rcParams['axes.linewidth']=2
+
+#set graph line width
+# plt.rcParams['lines.linewidth'] = 2
+
+#set font size for titles 
+#plt.rcParams['axes.titlesize'] = 16
+
+#set font size for labels on axes
+plt.rcParams['axes.labelsize'] = 16
+
+#set size of numbers on x-axis
+plt.rcParams['xtick.labelsize'] = 14
+#set size of numbers on y-axis
+plt.rcParams['ytick.labelsize'] = 14
+
+#set size of ticks on x-axis
+# plt.rcParams['xtick.major.size'] = 7
[…]

### 2017-07-03 experiments/energyMeasurements/DMA_Energy_cost/python/DMA_SW_energy.py

@@ -1,0 +1,120 @@
+#!/home/amjed/anaconda3/bin/python
+import numpy as np 
+import matplotlib.pyplot as plt 
+from matplotlib import style
+import copy
+style.use('seaborn-muted')
+
+def remOutLiers(voltagesIn):
+	voltages = copy.deepcopy( voltagesIn )
+	voltages = np.array(voltages).astype(np.float) 
+	stdVolt_3 = np.std( voltages ) * 2
+	meanVolt = np.mean( voltages )
+	voltages = list(voltages)
+	# print('3 * std = ', stdVolt_3)
+	for i, volt in enumerate(voltages):
+		if np.abs(volt - meanVolt) > stdVolt_3:
+			print('del', voltages[i] )
+			del voltages[i] 
+
+	return voltages
+
+
+def voltList(file ):
+	"""
+		@input: file name 
+		@output: list of voltages
+
+		voltList extract the voltages values 
+		from a the EDB output
+	 """
+	voltages=[]
+	for line in open(file):
+		# get the voltages and remove the `\n` char
+		voltages.append( (line.split(',')[-1])[:-1] ) 
+	
+	return(voltages)
+
+def statistics(voltages, numLines):
+	""" 
+		@input: voltages list, number instructions between each voltage probes
+		@output: mean and standard deviation 
+	"""
+	diffs=[]
+	list_len = len(voltages)
+	for i in range(0,list_len, 2):
+		diffs.append( (( (float(voltages[i])/numLines - float(voltages[i+1])/numLines ) \
+								 **2) * 10e-6 * 0.5 ) )
+	diffsClean = remOutLiers(diffs)
+	return(np.mean(diffsClean), np.std(diffsClean))
+
+def plot(xAxis, voltMeans, symb, labels):
+	plt.rcParams['axes.linewidth'] =  2
+	plt.rcParams['axes.labelsize'] =  16
+	plt.rcParams['xtick.labelsize'] =  16
+	plt.rcParams['ytick.labelsize'] =  16
+	plt.rcParams['lines.linewidth'] =  2
+	f = plt.figure(figsize=(8,4))
+	i=0
+	colors= ['g','b']
+	for val in voltMeans:
[…]

### 2017-07-03 experiments/pythonScript/ipos_versions_compare.py

@@ -1,129 +1,0 @@
-"""
-Performance: IPOS versus Chain
-    Metric : time
-    Task   : Decompressing 100 bytes of data
-    App    : data_decompression
-
-
-@ Author  : Amjad Yousef Majid
-@ Gmail ID: amjad.y.majid 
-@ Date    : 30/June/2017 
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.ticker as ticker
-
-
-def executionTime(fh):
-    """
-    Input: file handler 
-    return: the average execution time of a program
-    """
-    numbers = []
-    for num in fh:
-        if num[-2:-1] == '1':
-            #print(num)
-            numbers.append( float(num.split(',')[0] ) )
-    aDiff = 0
-    for i in range(0, len(numbers)-1, 2):
-        aDiff += (numbers[i+1] - numbers[i])
-#     aDiff *=1000  # make the diff in milliseconds
-
-    time = (aDiff / int(len(numbers)/2) )
-    fh.close()
-    return time
-
-def figureSetting():
-    """
-    Set a figure params
-    """
-    #set figure linewidth
-    plt.rcParams['axes.linewidth']=2
-
-    #set font size for labels on axes
-    plt.rcParams['axes.labelsize'] = 16
-
-    #set size of numbers on x-axis
-    plt.rcParams['xtick.labelsize'] = 16
-    #set size of numbers on y-axis
-    plt.rcParams['ytick.labelsize'] = 16
-
-    #set size of ticks on x-axis
-    plt.rcParams['xtick.major.size'] = 7
-    #set size of ticks on y-axis
-    plt.rcParams['ytick.major.size'] = 7
-
-    #set size of markers, e.g., circles representing points
-    #set numpoints for legend
-    plt.rcParams['legend.numpoints'] = 1
-
-    
-    
-decomp_dma      = open('../IPOS-versions-performance/data_decompression/dma_decomp_100B.csv')
-decomp_vob_ml   = open('../IPOS-versions-performance/data_decompression/vob_memLimitted_100B.csv')
-decomp_vob_pl   = open('../IPOS-versions-performance/data_decompression/vob_perforLimitted_100B.csv')
-
-sel_dma         = open('../IPOS-versions-performance/selectionSort/dma_sort_20num.csv')
[…]

### 2017-07-03 experiments/runtime/SRC/IPOS-DMA_vs_Chain.py

@@ -93,7 +93,7 @@ data = [ [chain_dd_time, ipos_dd_time],
 
 data = np.transpose(data)
 hatches = ['+', '/']
-colors = ['0.6', '0.8']
+colors = ['#edf8b1', '#7fcdbb']
 # colors = ['c', 'm']
 labels = ["Chain", "IPOS"]
 

### 2017-07-03 paper/experiments/energyMeasurements/DMA_Energy_cost/python/DMA_SW_energy.py

@@ -1,119 +1,0 @@
-#!/home/amjed/anaconda3/bin/python
-import numpy as np 
-import matplotlib.pyplot as plt 
-from matplotlib import style
-import copy
-style.use('seaborn-muted')
-
-def remOutLiers(voltagesIn):
-	voltages = copy.deepcopy( voltagesIn )
-	voltages = np.array(voltages).astype(np.float) 
-	stdVolt_3 = np.std( voltages ) * 2
-	meanVolt = np.mean( voltages )
-	voltages = list(voltages)
-	# print('3 * std = ', stdVolt_3)
-	for i, volt in enumerate(voltages):
-		if np.abs(volt - meanVolt) > stdVolt_3:
-			print('del', voltages[i] )
-			del voltages[i] 
-
-	return voltages
-
-
-def voltList(file ):
-	"""
-		@input: file name 
-		@output: list of voltages
-
-		voltList extract the voltages values 
-		from a the EDB output
-	 """
-	voltages=[]
-	for line in open(file):
-		# get the voltages and remove the `\n` char
-		voltages.append( (line.split(',')[-1])[:-1] ) 
-	
-	return(voltages)
-
-def statistics(voltages, numLines):
-	""" 
-		@input: voltages list, number instructions between each voltage probes
-		@output: mean and standard deviation 
-	"""
-	diffs=[]
-	list_len = len(voltages)
-	for i in range(0,list_len, 2):
-		diffs.append( (( (float(voltages[i])/numLines - float(voltages[i+1])/numLines ) \
-								 **2) * 10e-6 * 0.5 ) )
-	diffsClean = remOutLiers(diffs)
-	return(np.mean(diffsClean), np.std(diffsClean))
-
-def plot(xAxis, voltMeans, symb, labels):
-	plt.rcParams['axes.linewidth'] =  2
-	plt.rcParams['axes.labelsize'] =  16
-	plt.rcParams['xtick.labelsize'] =  16
-	plt.rcParams['ytick.labelsize'] =  16
-	plt.rcParams['lines.linewidth'] =  2
-	f = plt.figure(figsize=(8,4))
-	i=0
-	for val in voltMeans:
[…]

### 2017-07-03 paper/experiments/energyMeasurements/Fram_read_cost/python/fram_sram_read_energy_consumption.py

@@ -1,87 +1,0 @@
-#!/home/amjed/anaconda3/bin/python
-import numpy as np 
-import matplotlib.pyplot as plt 
-from matplotlib import style
-
-style.use('seaborn-muted')
-
-numberOfReadLines = 1600;
-
-#----Get the volatge values of the cap-----#
-fram_voltages=[]
-sram_voltages=[]
-fram_cntr=0
-sram_cntr=0
-for line in open("../outputs/fram.txt"):
-	# get the voltages and remove the `\n` char
-	fram_voltages.append( (line.split(',')[-1])[:-1] ) 
-	fram_cntr+=1
-
-for line in open("../outputs/sram.txt"):
-	# get the voltages and remove the `\n` char
-	sram_voltages.append( (line.split(',')[-1])[:-1] )
-	sram_cntr+=1
-
-#print(fram_voltages)	 
-#print(sram_voltages)
-#exit()
-
-
-#-----SRAM/FRAM write costs------#
-v_list_len= min(fram_cntr, sram_cntr)
-fram_diffs=[]
-sram_diffs=[]
-for i in range(0,v_list_len, 2):
-	fram_diffs.append( ( ( (float(fram_voltages[i])/numberOfReadLines )**2 - (float(fram_voltages[i+1])/numberOfReadLines ) **2 ) * 10e-6 * 0.5 ) ) 
-	sram_diffs.append( ( ( (float(sram_voltages[i])/numberOfReadLines )**2 - (float(sram_voltages[i+1])/numberOfReadLines ) **2 ) * 10e-6 * 0.5 ) ) 
-
-# print(fram_diffs)
-# print(sram_diffs)
-# exit()
-
-print (sum(fram_diffs) /len(fram_diffs))
-print (sum(sram_diffs) /len(sram_diffs))
-
-
-#------------Line fitting------------#
-fit_order=3
-xVals = range(len(fram_diffs))
-fram_z =  np.polyfit( xVals, fram_diffs, fit_order)
-sram_z =  np.polyfit( xVals, sram_diffs, fit_order)
-fram_estYVals = np.polyval(fram_z, xVals)
-sram_estYVals = np.polyval(sram_z, xVals)
-
-
-#--------------Plotting--------------#
-
-plt.rcParams['axes.labelsize'] =  12
-plt.rcParams['lines.linewidth'] =  1.5
[…]

### 2017-07-03 paper/experiments/energyMeasurements/Fram_write_cost/python/fram_sram_write_energy_consumption.py

@@ -1,87 +1,0 @@
-#!/home/amjed/anaconda3/bin/python
-import numpy as np 
-import matplotlib.pyplot as plt 
-from matplotlib import style
-
-style.use('seaborn-muted')
-
-numberOfReadLines = 1600;
-
-#----Get the volatge values of the cap-----#
-fram_voltages=[]
-sram_voltages=[]
-fram_cntr=0
-sram_cntr=0
-for line in open("../outputs/fram.txt"):
-	# get the voltages and remove the `\n` char
-	fram_voltages.append( (line.split(',')[-1])[:-1] ) 
-	fram_cntr+=1
-
-for line in open("../outputs/sram.txt"):
-	# get the voltages and remove the `\n` char
-	sram_voltages.append( (line.split(',')[-1])[:-1] )
-	sram_cntr+=1
-
-#print(fram_voltages)	 
-#print(sram_voltages)
-#exit()
-
-
-#-----SRAM/FRAM write costs------#
-v_list_len= min(fram_cntr, sram_cntr)
-fram_diffs=[]
-sram_diffs=[]
-for i in range(0,v_list_len, 2):
-	fram_diffs.append( ( ( (float(fram_voltages[i])/numberOfReadLines )**2 - (float(fram_voltages[i+1])/numberOfReadLines ) **2 ) * 10e-6 * 0.5 ) ) 
-	sram_diffs.append( ( ( (float(sram_voltages[i])/numberOfReadLines )**2 - (float(sram_voltages[i+1])/numberOfReadLines ) **2 ) * 10e-6 * 0.5 ) ) 
-
-# print(fram_diffs[:5])
-# print(sram_diffs[:5])
-# exit()
-
-# print (sum(fram_diffs) /len(fram_diffs))
-# print (sum(sram_diffs) /len(sram_diffs))
-
-
-#------------Line fitting------------#
-fit_order=3
-xVals = range(len(fram_diffs))
-fram_z =  np.polyfit( xVals, fram_diffs, fit_order)
-sram_z =  np.polyfit( xVals, sram_diffs, fit_order)
-fram_estYVals = np.polyval(fram_z, xVals)
-sram_estYVals = np.polyval(sram_z, xVals)
-
-
-#--------------Plotting--------------#
-
-plt.rcParams['axes.labelsize'] =  12
-plt.rcParams['lines.linewidth'] =  1.5
[…]

### 2017-07-03 paper/experiments/runtime/SRC/Chain_code/DFT/main.c

@@ -1,327 +1,0 @@
-/*
-	This is a DFT implemention in chain
-
-	@ Author: Amjad Yousef Majid
-	@ Date  : 6/April/2017
-
-*/
-#include <msp430.h>
-#include <stdint.h>
-#include <stdbool.h>
-#include <stdio.h>
-
-#include <math.h>    /// using the normal c math library   "-lm" must be added to the Makefile
-
-#include <libchain/chain.h>
-#include <stdlib.h>
-
-#define SIZE 8
-#define PI2 6.28
-
-#ifdef DEBUG
-  float x[SIZE];              		// discrete-time signal, x
-  float x_chain[SIZE];              // discrete-time signal, x
-  float Xre[SIZE], Xim[SIZE]; 		// DFT of x (real and imaginary parts)
-  float P[SIZE]; 
-  float p_deb[SIZE]={0};
- void dft_c() {
- 	/*
-		The dft_c function is taken from 
-		https://batchloaf.wordpress.com/2013/12/07/simple-dft-in-c/
- 	*/
-    int n, k; // iSIZEdices for time and frequency domains
-
-    // Generate random discrete-time signal x in range (-1,+1)
-    srand(0);
-    for (n = 0; n < SIZE; ++n)
-      x[n] = ((2.0 * rand()) / RAND_MAX) - 1.0;
-
-    // Calculate DFT of x using brute force
-    for (k = 0; k < SIZE; ++k) {
-      // Real part of X[k]
-      Xre[k] = 0;
-      for (n = 0; n < SIZE; ++n)
-        Xre[k] += x[n] * cos(n * k * PI2 / SIZE);
-
-      // Imaginary part of X[k]
-      Xim[k] = 0;
-      for (n = 0; n < SIZE; ++n)
-        Xim[k] -= x[n] * sin(n * k * PI2 / SIZE);
-
-      // Power at kth frequency bin
-      P[k] = Xre[k] * Xre[k] + Xim[k] * Xim[k];
-    }
-
-  }
-
-#endif
-
-void init()
-{
-	WDTCTL = WDTPW | WDTHOLD;	// Stop watchdog timer
-	PM5CTL0 &= ~LOCKLPM5;
-    __enable_interrupt();
[…]

### 2017-07-03 virtOperatBuffer/memoryLimited/ipos-gcc-vob/include/operationalBuffer.h

@@ -14,7 +14,7 @@ typedef struct wb_node {
 }wb_node;
 
 
-#define TABLESIZE 0x007f // My hashing function applies a masking technique
+#define TABLESIZE 0x00ff // My hashing function applies a masking technique
 #define TEMPBUFSIZE_RAWS TABLESIZE
 #define TEMPBUFSIZE_COLS 2
 

### 2017-07-03 virtOperatBuffer/memoryLimited/largeDataCommit/main.c

@@ -2,8 +2,10 @@
 #include <msp430fr5969.h>
 #include <ipos.h>
 
-#define ARRSIZE 50
-__nv volatile uint16_t data[ARRSIZE] = {0};
+#define ARRSIZE 128
+#define FIX_ARRSIZE 128
+__nv volatile uint16_t data[FIX_ARRSIZE] = {0};
+
 
 void init();
 void task0();

### 2017-07-03 virtOperatBuffer/performanceLimited/largeDataCommit/main.c

@@ -2,8 +2,11 @@
 #include <msp430fr5969.h>
 #include <ipos.h>
 
-#define ARRSIZE 50
-__nv volatile uint16_t data[ARRSIZE] = {0};
+//#define ARRSIZE 50
+#define ARRSIZE 1
+#define FIX_ARRSIZE 128
+__nv volatile uint16_t data[FIX_ARRSIZE] = {0};
+
 
 void init();
 void task0();

### 2017-07-06 paper/intro.tex
put Brandon's propose section structure in the paper; initial corrections in Brandon's intro, added all author names in comments

@@
-Advances in processor efficiency along with the development of
-energy-harvesting power systems has created a new category of devices that
-require neither a battery nor a tethered power supply.  These devices operate
-entirely using energy extracted from their environment, such as RF energy,
-photovoltaics, and vibration.  Incorporating compute, storage, sensing, and
-communication hardware, such energy-independent devices are a promising
-candidate as the underlying technology for emerging internet of things (IoT)
-applications, in- and on-body medical applications, and emerging applications,
-like energy-harvesting nano-satellites.
+Advances in processor efficiency along with the development of energy-harvesting power systems has created a new category of devices that require neither a battery nor a tethered power supply.
+These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{}, photovoltaics~\cite{}, and vibration~\cite{}.
+Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising candidate as the underlying technology for emerging Internet of Things (IoT) applications, in- and on-body medical applications~\cite{}, and emerging applications, like energy-harvesting nano-satellites~\cite{}.
 
@@
 power compared to a platform's operating power level. A devices operates for a
-brief period of time and when buffered energy is depeleted, shuts down and
+brief period of time and when buffered energy is depleted, shuts down and
 begins recharging to later operate again. Charge and discharge times vary by
-device and some devices~\cite{wisp} may fail 10 to 100 times per second.
+device and some devices~\cite{wisp} may fail 10 to 100 times per second~\cite{}.
 
@@
 interrupted by failure periods. When power fails, a device loses its volatile
-state (i.e., registers, stack, SRAM) and retains its non-volatile state (i.e.,
-FRAM, Flash). While capturing periodic checkpoints~\cite{mementos,tictpl,quickrecall}
+state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e.,
+FRAM, Flash. While capturing periodic checkpoints~\cite{mementos,tictpl,quickrecall}
 and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve
-execution progress, failures can leave volatile and non-volatile state inconsistent,
-leading to unrecoverable failures~\cite{mspcdino,edb,edbtoppicks}.
+execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb,edbtoppicks}.
 
@@
 New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
-architecture mechanisms~\cite{clank, idetic, nvp} help avoid data
+architecture mechanisms~\cite{clank,idetic,nvp} help avoid data
 inconsistency. New architectures require costly hardware changes and are
@@
 adoption. New memory models may also increase the amount of non-volatile
-memory accesses by requring non-volatile versioning~\cite{chain} or completely
+memory accesses by requiring non-volatile versioning~\cite{chain} or completely
 precluding the use of volatile memory~\cite{ratchet}. Increasing pressure on
@@
 an end-to-end evaluation of a sensing and data processing application deployed
-in a lab environment \TODO{blah blah blah eval}.
+in a lab environment. \TODO{blah blah blah eval}.
 

### 2017-07-06 paper/paper.tex

@@
 \documentclass[sigconf,anonymous,review]{acmart}
+
 \usepackage[normalem]{ulem}
-
-\usepackage{amssymb} % needed for: various symbols
-\usepackage{amsmath} % needed for: math symbols
-% http://tex.stackexchange.com/questions/229355/algorithm-algorithmic-algorithmicx-algorithm2e-algpseudocode-confused
-\usepackage{algorithm} % this used as an environment for the algorithm code (linke figure and table)
-\usepackage{algorithmicx} % this is used to write the pseudo code
-\usepackage[noend]{algpseudocode} % this is for the layout of the pseudo code
+\usepackage{amssymb}
+\usepackage{amsmath}
+\usepackage{algorithm} % environment for the algorithm code (like figure and table)
+\usepackage{algorithmicx} % write the pseudo code
+\usepackage[noend]{algpseudocode} % layout of the pseudo code
 \usepackage{tikz} % needed for: tikz pics
-% \usepackage[inline]{enumitem}
 \usepackage{xcolor}
@@
 \usepackage{natbib}
-
 \usepackage{graphicx}
 \usepackage{subfigure}
+%\usepackage[inline]{enumitem}
+
 \newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
-\newcommand{\sys}{NAME\xspace}
+\newcommand{\sys}{VIPRE\xspace}
+\newcommand{\sysfull}{Virtualized Intermittently-Powered Runtime Environment\xspace}
 
@@
 
-\title{VIPOS: Virtualized Intermittently Powered Runtime Environment (``VIPRE''?)}
-% \author{
-% 	\IEEEauthorblockN{Amjad Yousef Majid, Przemys{\l}aw Pawe{\l}czak}
-% 	\IEEEauthorblockA{TU Delft, Mekelweg\,4, 2628\,CD Delft, NL\\
-% 	\{a.y.majid, p.pawelczak\}@tudelft.nl}
-% }
+\title{VIPRE: Virtualized Intermittently-Powered Runtime Environment}
+
+%Amjad Yousef Majid, Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m, Przemys{\l}aw Pawe{\l}czak,
+%TU Delft, Mekelweg\,4, 2628\,CD Delft, NL
+%\{a.y.majid, k.s.yildirim, p.pawelczak\}@tudelft.nl
+%
+%Kiwan Maeng, Alexei Collin, Brandon Lucia
+%Carnegie Mellon University, 4720 Forbes Avenue, Pittsburgh, PA 15213, USA
+%\{kmaeng, acolin, blucia\}@kmaeng@andrew.cmu.edu
 
@@
 \label{sec:intro}
-\TODO{VIPER: Virtualized Intermittent Programming model, Execution model, and Runtime environment}
-\TODO{PREVIEW: Programming and Runtime Environment with Virtualized Intermittent Execution Windows}
+
+%\TODO{VIPER: Virtualised Intermittent Programming model, Execution model, and Runtime environment}
+%\TODO{PREVIEW: Programming and Runtime Environment with Virtualised Intermittent Execution Windows}
 
 \input{intro}
+
+\section{Computation on Intermittent Devices: Background}
+\label{sec:background}
+
+Performing complex operations on battery-less energy harvesting embedded platforms is challenging from many points of view.
+Here we will briefly provide a background on energy harvesting for embedded systems.
+
+\subsection{Energy Harvesting: Background}
+
+\subsection{(Task-based) Intermittent Computing: Background}
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=0.25\columnwidth]{figures/}
+	\caption{Background image.}
+	\label{fig:}
+\end{figure}
+
+Elaborating on the ``key challenge'' paragraphs near the end of the intro. We will need a picture.
+
+\section{\sysfull}
+\label{sec:overeall_system}
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=0.25\columnwidth]{figures/}
+	\caption{\sys top-level description.}
+	\label{fig:}
+\end{figure}
+
+Describe the overall system, the programming model, and the semantics/execution model. This should include a discussion of memory virtualization and task coalescing.
+
+\section{Memory Virtualization}
+\label{sec:memory_virtulaization}
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=0.25\columnwidth]{figures/}
+	\caption{Memory virtualization architecture.}
+	\label{fig:}
+\end{figure}
+
+\section{Task Coalescing}
+\label{sec:memory_virtulaization}
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=0.25\columnwidth]{figures/}
+	\caption{Task coalescing architecture.}
+	\label{fig:}
+\end{figure}
+
+\section{Discussion}
+\label{sec:discussion}
+
+\section{Methodology and Evaluation}
+\label{sec:methodology_evaluation}
+
+\section{Related Work}
+\label{sec:related_work}
+
+\section{Conclusions}
+\label{sec:conclusions}
+
+This paper presented \sys: \sysfull.
+
+%%%%%
 

### 2017-07-06 paper/paper.tex
first notes to the background section (not finished, many placeholders)

@@
 Performing complex operations on battery-less energy harvesting embedded platforms is challenging from many points of view.
-Here we will briefly provide a background on energy harvesting for embedded systems.
+Here we shall briefly provide a background on computation with energy harvesting embedded systems.
 
 \subsection{Energy Harvesting: Background}
+
+Supplying power to tiny embedded computers using batteries alone is not sustainable.
+Batteries are a great pollutant.
+A solution is to replace batteries with smaller, but more environment friendly capacitors and supply power directly from the ambient energy sources.
+
+What platforms enable intermittent computation: WISP~\cite{} (with its variants such as WISPCam~\cite{}), Moo~\cite{}, ambient backscatter~\cite{}, or computation RFID commercial platforms such as~\cite{}.
+In case of the above platforms, the main source of energy is electromagnetic radiation in the radio frequency range.
+
+Application include long-term battery-less monitoring.
+One example is the distributed monitoring of moisture of large plant fields during vegetation~\cite{}.
+Other applications include battery-less image capture and processing~\cite{}, distributed low-power networking~\cite{}.
+
+Energy provided from the ambient is not stable and difficult to predict accurately.
+Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation.
+Energy breaks happen every hundred of ms~\cite{}.
+
+Solution to the problem is to divide the program into smaller pieces that guarantee execution within a discharge region and track the state of the program in-between the program part execution times.
 
@@
 	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Background image.}
-	\label{fig:}
+	\caption{Two problems with fixed-size task in intermittent execution.
+(a) task underestimation: task executed at device with capacitor size $X$, will not execute at device with capacitor $X\gg Y$; (b) task overestimation: with stable energy source task tracking causes unnecessary overhead.}
+	\label{fig:fixed_task_problem}
 \end{figure}
 
-Elaborating on the ``key challenge'' paragraphs near the end of the intro. We will need a picture.
+Various capacitor sizes of intermittent device result in a program being executed
+
+Other forms of intermittent platforms will share the same problem.
+One of them considers actuation platforms, which will be powered directly from energy harvesting sources~\cite{}.
+will share the energy storage with computing platform.
+However, power supply for actuators is order of magnitude larger than for computation.
+Therefore, although capacitor is large enough to perform computation alone, energy to power actuators takes precedence leaving not much space for computing.
+
+\TODO{Elaborate on the ``key challenge'' paragraphs near the end of the intro.}
 

### 2017-07-06 paper/paper.tex
ACM conference-required information added to the paper

@@
 
+\setcopyright{rightsretained}
+
+\acmDOI{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
+
+\acmISBN{978-1-xxxx-xxxx-x/17/11}
+
+\acmConference[ASPLOS]{The 23rd ACM International Conference on Architectural Support for Programming Languages and Operating Systems}{March 24--28, 2017}{Williamsburg, VA, USA}
+
+\acmYear{2017}
+
+\copyrightyear{2017}
+
+\acmPrice{15.00}
+
 \begin{document}
 
-\title{VIPRE: Virtualized Intermittently-Powered Runtime Environment}
+\title[VIPRE: Virtualized Intermittently-Powered Runtime Environment]{VIPRE: Virtualized Intermittently-Powered Runtime Environment}
 
-%Amjad Yousef Majid, Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m, Przemys{\l}aw Pawe{\l}czak,
-%TU Delft, Mekelweg\,4, 2628\,CD Delft, NL
-%\{a.y.majid, k.s.yildirim, p.pawelczak\}@tudelft.nl
-%
-%Kiwan Maeng, Alexei Collin, Brandon Lucia
-%Carnegie Mellon University, 4720 Forbes Avenue, Pittsburgh, PA 15213, USA
-%\{kmaeng, acolin, blucia\}@kmaeng@andrew.cmu.edu
+%Note: order of authors placed randomly
+
+\author{Kiwan Maeng}
+%\authornote{}
+%\orcid{}
+\affiliation{%
+	\institution{Carnegie Mellon University}
+	\streetaddress{4720 Forbes Avenue}
+	\city{Pittsburgh} 
+	\state{PA} 
+	\postcode{15213}
+}
+\email{kmaeng@andrew.cmu.edu}
+
+\author{Alexei Colin}
+%\authornote{}
+%\orcid{}
+\affiliation{%
+	\institution{Carnegie Mellon University}
+	\streetaddress{4720 Forbes Avenue}
+	\city{Pittsburgh} 
+	\state{PA} 
+	\postcode{15213}
+}
+\email{acolin@andrew.cmu.edu}
+
+\author{Brandon Lucia}
+%\authornote{}
+%\orcid{}
+\affiliation{%
+	\institution{Carnegie Mellon University}
+	\streetaddress{4720 Forbes Avenue}
+	\city{Pittsburgh} 
+	\state{PA} 
+	\postcode{15213}
+}
+\email{blucia@andrew.cmu.edu}
+
+\author{Amjad Yousef Majid}
+%\authornote{}
+%\orcid{}
+\affiliation{%
+	\institution{Delft University of Technology}
+	\streetaddress{Mekelweg 4}
+	\city{Delft, The Netherlands} 
+	\state{Zuid Holland} 
+	\postcode{2628\,CD}
+}
+\email{a.y.majid@tudelft.nl}
+
+\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
+%\authornote{}
+%\orcid{}
+\affiliation{%
+	\institution{Delft University of Technology}
+	\streetaddress{Mekelweg 4}
+	\city{Delft, The Netherlands} 
+	\state{Zuid Holland} 
+	\postcode{2628\,CD}
+}
+\email{k.s.yildirim@tudelft.nl}
+
+\author{Przemys{\l}aw Pawe{\l}czak}
+%\authornote{}
+%\orcid{}
+\affiliation{%
+	\institution{Delft University of Technology}
+	\streetaddress{Mekelweg 4}
+	\city{Delft, The Netherlands} 
+	\state{Zuid Holland} 
+	\postcode{2628\,CD}
+}
+\email{p.pawelczak@tudelft.nl}
+
+\renewcommand{\shortauthors}{A. Y. Majid et al.}
 
@@
 \end{abstract}
+
+% The code below should be generated by the tool at 
+%http://dl.acm.org/ccs.cfm Please copy and paste the code instead of the 
+%example below.
+
+\begin{CCSXML}
+	<ccs2012>
+	<concept>
+	<concept_id>10010520.10010553.10010562.10010564</concept_id>
+	<concept_desc>Computer systems organization~Embedded 
+	software</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010520.10010575.10010578</concept_id>
+	<concept_desc>Computer systems organization~Availability</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
+	<concept_desc>Software and its engineering~Scheduling</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
+	<concept_desc>Software and its engineering~Communications 
+	management</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10011007.10010940.10010971.10010564</concept_id>
+	<concept_desc>Software and its engineering~Embedded 
+	software</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Computer systems organization~Embedded software}
+\ccsdesc[300]{Computer systems organization~Availability}
+\ccsdesc[300]{Software and its engineering~Scheduling}
+\ccsdesc[300]{Software and its engineering~Communications management}
+\ccsdesc[300]{Software and its engineering~Embedded software}
+
+\keywords{Energy harvesting, transient operation, operating system}
 

### 2017-07-07 paper/compiler.tex
initial draft for the compiler

\section{Compiler Support for Automatic Task Generation}
\label{sec:compiler}

\sys additionally provides a compiler support for automatic task
generation, eliminating programmer's burden of writing a task-based
program from scratch. The compiler gets a plain-C code as an input
and generates a binary that is rewritten in task-based programming
style that can run on \sys without additional programming effort. 
Automatic task generation compiler is often inefficient than human-written
code for \sys because if unbounded loop or pointer is present, the
compiler makes a conservative decision. \xxx{If the performance is critical,
the programmer may want to manually write the task-based code by himself.
Otherwise, the programmer can rely on the compiler for simplicity.}

The compiler consists of two part: analysis pass and transform pass.
Analysis pass analyze the code structure and decides where to put
task boundaries. Transform pass converts the code into task-style
program and pepper it with necessary \sys keywords so that it can run
on \sys system. It is implemented using LLVM.

\subsection{Analysis Pass}
\label{sec:analysis}

\sys compiler decides the task size based on the number of LLVM IR instructions.
That is, it tries to make every task to contain roughly the same amount of LLVM IR
instructions. Intuitively, the number of LLVM IR instructions have a correlation with
the total workload done by the task. However, the number of LLVM IR instructions is 
a crude approach of estimating workload of a task since actual machine instructions are
what count and it is not exactly the same as LLVM IR instruction. Also, different instructions
have different power usage, use different cycles, etc. However, the size of each
task need not be exactly the same and it is sufficient to make sure that all of the
tasks are small enough to fit in to the energy budget. Counting LLVM IR instruction is sufficient
in this perspective.

\sys compiler first builds a CFG graph where each basic block is a vertice and branch
from one basic block to another is a unidirectional edge. Each vertices has a
{\em price}, which is the number of LLVM IR inside the vertice. 
When forming a task, each task should contain a basic block so that the sum of price of the 
blocks within each tasks should be roughly the same. Price makes the workload done by
each tasks roughly the same and make no task exceed the energy bound.

Each edge has a {\em penalty}, which indicates the penalty of placing a task boundary at
the specific edge. Intuitively, when a task boundary is placed on a frequently visited edge
(such as a back edge of a loop) the system suffers from higher runtime overhead.
There is no way of statically estimating the number of visit of each edges, so we initally set every
edge penalty as 1. When analyzing loops, penalty of edges within loops are multiplied by loop count.
\xxx{TODO: change the penalty to be 1 or 1*loopcount}

Ater constructing a graph and assigning price and penalty, the compiler analyzes loops.

except those inside a loop which is set as the loop count. Unbounded loops are conservatively considered very large.

\subsection{Transform Pass}
\label{sec:transform}

### 2017-07-09 paper/intro.tex
Paragraph 1 revised with 10 new references added

@@
-Advances in processor efficiency along with the development of energy-harvesting power systems has created a new category of devices that require neither a battery nor a tethered power supply.
-These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{}, photovoltaics~\cite{}, and vibration~\cite{}.
-Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising candidate as the underlying technology for emerging Internet of Things (IoT) applications, in- and on-body medical applications~\cite{}, and emerging applications, like energy-harvesting nano-satellites~\cite{}.
+Advances in processor efficiency along with the development of energy-harvesting power systems has created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
+These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaic components~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
+Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising candidate as the underlying technology for emerging Internet of Things (IoT) applications~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical applications, and emerging applications, like energy-harvesting nano-satellites~\cite{kicksat}.
 

### 2017-07-09 paper/abstract.tex
Paragraph 1 revised with 10 new references added

@@
-\TODO{Need to re-write this to make it consistent with the story after we have the story in place}
+%\TODO{Need to re-write this to make it consistent with the story after we have the story in place}
+
+\todo{Re-write the abstract once the paper is ready}{Brandon}
+
 Enabling battery-free devices is a mandatory step towards the environment friendly IoT.

### 2017-07-09 paper/intro.tex
References and text up to paragraph 3 (intro) fixed

@@
 These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaic components~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
-Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising candidate as the underlying technology for emerging Internet of Things (IoT) applications~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical applications, and emerging applications, like energy-harvesting nano-satellites~\cite{kicksat}.
+Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising underlying technology for applications in emerging areas such as Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
 
 While promising, energy-harvesting devices present a unique design challenge in
-that they operate only {\em intermittently} when energy is available. An
+that they operate only {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}. An
 energy-harvesting device buffers energy from its environment, e.g., in a
-capacitor. When a threshold quantity of energy accumulates, the device begins
-operating. Harvestable energy sources typically produce relatively very little
-power compared to a platform's operating power level. A devices operates for a
-brief period of time and when buffered energy is depleted, shuts down and
-begins recharging to later operate again. Charge and discharge times vary by
-device and some devices~\cite{wisp} may fail 10 to 100 times per second~\cite{}.
+capacitor~\cite[Fig.
+3]{gorlatova_tmc_2013},~\cite[Fig.
+1]{gunduz_commag_2014}.
+When a threshold quantity of energy accumulates, the device begins operating.
+Harvestable energy sources typically produce relatively very little power compared to a platform's operating power level---compare for instance~\cite[Table III and V]{prasad_comst_2014} with~\cite[Table I]{carrano_cst_2014}.
+A devices operates for a brief period of time and when buffered energy is depleted, shuts down and begins recharging to later operate again.
+As an example, inter-task time may even exceed 20\,s for wireless energy-harvesting battery-less vivo digestive tract monitoring system~\cite[Fig.
+3c]{nadeau_naturebio_2017}.
+Charge and discharge times vary by device and some of them~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite[Fig.
+1]{tan_infocom_2016},~\cite[Fig.
+1]{mementos}.
+
+\todo{Put exact reference for "tictpl"}{Brandon}
 
 Software in an energy-harvesting system operates according to the {\em
-intermittent execution model}~\cite{dino}, with bursts of execution that are
-interrupted by failure periods. When power fails, a device loses its volatile
-state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e.,
-FRAM, Flash. While capturing periodic checkpoints~\cite{mementos,tictpl,quickrecall}
-and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve
-execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb,edbtoppicks}.
-
-\TODO{Don't forget to include all the references in this section!}
+intermittent execution model}~\cite[Sec.
+2]{dino}~\cite[Fig.
+2]{lucia_snapl_2017}, with bursts of execution that are interrupted by failure periods.
+When power fails, a device loses its volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
+While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb}.
 
 %FRAM costs more than FRAM -- need to figure out how to execute out of SRAM
+
 New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
-architecture mechanisms~\cite{clank,idetic,nvp} help avoid data
+architecture mechanisms~\cite{hicks_isca_2017,idetic,nvp} help avoid data
 inconsistency. New architectures require costly hardware changes and are

### 2017-07-10 paper/intro.tex
All references of the intro cleaned and missing added
Paragraph 4 ready

@@
 These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaic components~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
-Incorporating compute, storage, sensing, and communication hardware, such energy-independent devices are a promising underlying technology for applications in emerging areas such as Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
+Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such energy-independent devices are a promising underlying technology for applications in emerging areas such as Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
 
@@
 Harvestable energy sources typically produce relatively very little power compared to a platform's operating power level---compare for instance~\cite[Table III and V]{prasad_comst_2014} with~\cite[Table I]{carrano_cst_2014}.
-A devices operates for a brief period of time and when buffered energy is depleted, shuts down and begins recharging to later operate again.
+A devices operates for a brief period of time and when buffered energy is depleted, shuts down and begins recharging to operate again later.
 As an example, inter-task time may even exceed 20\,s for wireless energy-harvesting battery-less vivo digestive tract monitoring system~\cite[Fig.
@@
 1]{tan_infocom_2016},~\cite[Fig.
-1]{mementos}.
+1]{mementos},~\cite[Fig.
+3]{nvp}.
 
 \todo{Put exact reference for "tictpl"}{Brandon}
+\todo{Check if the list of authors for Alpaca is correct}{Brandon}
 
@@
 2]{dino}~\cite[Fig.
-2]{lucia_snapl_2017}, with bursts of execution that are interrupted by failure periods.
+2]{lucia_snapl_2017}, with bursts of execution interrupted by failure periods.
 When power fails, a device loses its volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
@@
 New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
-architecture mechanisms~\cite{hicks_isca_2017,idetic,nvp} help avoid data
-inconsistency. New architectures require costly hardware changes and are
-inapplicable to today's systems. New programming models and compilers force
-programmers to use new memory models~\cite{chain,ratchet}, which may inhibit
-adoption. New memory models may also increase the amount of non-volatile
-memory accesses by requiring non-volatile versioning~\cite{chain} or completely
-precluding the use of volatile memory~\cite{ratchet}. Increasing pressure on
-non-volatile memory or precluding the use of volatile memory may degrade
-efficiency; e.g., FRAM has a higher access energy and latency than
-SRAM~\cite{alpaca,wisp}. Commercially available devices are often provisioned
-with much less SRAM than FRAM~\cite{wolverine,wisp} and effectively leveraging
-the relatively high efficiency of SRAM is a key challenge for task-based
-intermittent systems.
-
-\TODO{concrete cost multiples here?}
+architecture mechanisms~\cite{hicks_isca_2017,idetic,nvp} for intermittently-powered devices help avoid data inconsistency.
+Nonetheless, both approaches---models and architectures---posses critical limitations.
+New architectures require costly hardware changes and are inapplicable to today's systems.
+For instance, they require dedicated idempotency violation control hardware buffers~\cite[Fig.
+3]{hicks_isca_2017} or dedicated on-processor logic like on demand register backup~\cite[Fig.
+9]{nvp}.
+Then, new programming models and compilers force programmers to use new memory models~\cite{chain,ratchet}, which may inhibit widespread adoption by programmers accustomed to classical (C) program writing.
+New memory models may also increase the amount of non-volatile memory accesses by requiring non-volatile versioning~\cite{chain} or completely precluding the use of volatile memory~\cite{ratchet}.
+Increasing pressure on non-volatile memory or precluding the use of volatile memory may degrade efficiency; e.g., FRAM has a higher access energy and latency than SRAM~\cite[Fig.
+4]{nvp}---which will be also shown quantitatively in this paper.
+Commercially available devices are often provisioned with much less SRAM than FRAM, e.g.
+up to 64 FRAM/SRAM ratio of~\cite{wolverine}, and effectively leveraging the relatively high efficiency of SRAM is a key challenge for task-based intermittent systems.
 
 %Task sizing is hard -- how to chop without risking inefficiency or non-term
+
 Several recent programming efforts~\cite{alpaca,chain} proposed {\em

### 2017-07-10 paper/intro.tex

@@
 3c]{nadeau_naturebio_2017}.
-Charge and discharge times vary by device and some of them~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite[Fig.
+On top of that, charge and discharge times vary by device and some of them~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite[Fig.
 1]{tan_infocom_2016},~\cite[Fig.
@@
 
-New programming and execution models~\cite{dino,ratchet,chain,alpaca} and
-architecture mechanisms~\cite{hicks_isca_2017,idetic,nvp} for intermittently-powered devices help avoid data inconsistency.
-Nonetheless, both approaches---models and architectures---posses critical limitations.
-New architectures require costly hardware changes and are inapplicable to today's systems.
+There are two fundamental approaches for dealing with data inconsistency for intermittently-powered devices: (i) programming and execution models~\cite{dino,ratchet,chain,alpaca} and (ii) architecture mechanisms~\cite{hicks_isca_2017,idetic,nvp}.
+Nonetheless, they both posses critical limitations.
+New architectures require costly hardware changes and are inapplicable to today's embedded off-the-shelf systems.
 For instance, they require dedicated idempotency violation control hardware buffers~\cite[Fig.
@@
 up to 64 FRAM/SRAM ratio of~\cite{wolverine}, and effectively leveraging the relatively high efficiency of SRAM is a key challenge for task-based intermittent systems.
+Despite the limitations, we conjecture that programming and execution models will be a {\em de facto choice for sustaining data consistency} of intermittently-powered devices.
+Simply put, programming model does not require any hardware investment and is dependent only upon the programmer's skills.
 
@@
 
-Several recent programming efforts~\cite{alpaca,chain} proposed {\em
-task-based} programming models, that require the programmer to statically
-decompose their program into tasks. A task is like a function with no callers
-that can include arbitrary computation and, upon completion, is guaranteed to
-have executed {\em atomically} and to have logically {\em committed} memory
-updates. The programmer expresses the flow of control from one task to another
-and task control-flow may be input-dependent. An important programming
-challenge in task-based models is that the length of execution of a task is
-limited by the total amount of energy that a device can buffer. Assuming that
-input power is negligible compared to operating power, a task will never be
-able to complete if its execution consumes more energy than the system can
-buffer. Transitioning from one task to the next imposes a cost and excessively
-small tasks may be inefficient. Appropriately sizing tasks to ensure progress
-and high performance is a key challenge for task-based intermittent systems.
+Two recent programming and execution model~\cite{alpaca,chain} proposed {\em task-based} paradigm, that require the programmer to statically decompose their program into tasks.
+A task resembles a function with no callers.
+Task can include arbitrary computation and, upon completion, is guaranteed to have executed {\em atomically} and to have logically {\em committed} memory updates.
+The programmer expresses the flow of control from one task to another and task control-flow may be input-dependent~\cite[Fig.
+4]{chain}.
+An important programming challenge in task-based models is that the length of execution of a task is limited by the total amount of energy that a device can buffer.
+Assuming that input power is negligible compared to operating power, a task will never be able to complete if its execution consumes more energy than the system can buffer.
+Transitioning from one task to the next imposes a cost and excessively small tasks may be inefficient.
+Appropriately sizing tasks to ensure progress and high performance is a key challenge for task-based intermittent systems.
 
-In this work, we develop \sys: a new programming and execution model that uses
-{\em volatile memory virtualization} to leverage efficient volatile memory and
-and {\em dynamic task coalescing} to efficiently execute tasks without
-exceeding available energy. \sys's memory virtualization mechanism uses a
-device's SRAM as the working memory during a task's execution, populating the
-SRAM with a page of data from FRAM as necessary at the start of the task. At
-the end of the task, the SRAM is committed back to FRAM using efficient DMA
-block copies. \TODO{Brandon sez: Is the next sentence what the system does?
-Or are we doing paging?} When a task accesses data outside of a page in SRAM,
-the task must fetch data from FRAM. We explore two strategies. The first
-strategy is {\em demand paging}, which swaps the SRAM page with a new page from
-FRAM, buffering the swapped-out page until commit. The second strategy is {\em
-buffered direct access} \TODO{better name}, which directly reads and writes
-FRAM relying on dynamic double-buffering to ensure memory is consistent at
+In this work, we develop \sys: a new programming and execution model that uses two new runtime features: (i) {\em volatile memory virtualization} to leverage efficient volatile memory and (ii) {\em dynamic task coalescing} to efficiently execute tasks without exceeding available energy.
+\sys's memory virtualization mechanism uses a device's SRAM as the working memory during a task's execution, populating the SRAM with a page of data from FRAM as necessary at the start of the task.
+At the end of the task, the SRAM is committed back to FRAM using efficient DMA block copies~\cite{kistler_micro_2006}.
+When a task accesses data outside of a page in SRAM, the task must fetch data from FRAM.
+
+\todo{articulate this last point better}{Przemek}
+
+With this paper we provide the following list of contributions.
+
+\begin{enumerate}
+
+\item {\bf Evaluation of data swapping strategies:} We propose and explore two data swapping strategies for intermittently-powered systems based on FRAM/SRAM architecture.
+The first strategy is {\em demand paging}, which swaps the SRAM page with a new page from FRAM, buffering the swapped-out page until commit.
+The second strategy is {\em buffered direct access}, which directly reads and writes FRAM relying on dynamic double-buffering to ensure memory is consistent at
 commit.
 
-\sys's dynamic task coalescing mechanism allows the programmer or a compiler to
-intuitively decompose their program into small tasks that amortize fixed
-per-task overheads, yet present no risk of exceeding device energy capacity. As
-such a decomposition executes, \sys {\em coalesces} dynamically consecutive
-tasks. Coalescing elides the commits of coalesced tasks by buffering multiple
-tasks' updates in an FRAM commit buffer. Periodically, as the span of the
-coalesced task grows, \sys ends coalescing and commits the state of the
-coalesced task. If a power failure interrupts a sequence of coalesced tasks,
-\sys adaptively reduces the number of tasks in that sequence that it will
-coalesce, committing sooner in future executions. Consequently, \sys's
-coalescing mechanism allows a program to execute efficiently across a range of
+\todo{Once paging is fully implemented--update the paragraph}{Przemek}
+\todo{Provide numerical results for this section}{Przemek}
+\todo{Find better name for demand paging}{Brandon}
+
+\item {\bf Task coalescing mechanism:} \sys's core contribution is the introduction of dynamic task coalescing mechanism.
+It allows the programmer or a compiler to intuitively decompose the program into small tasks that amortize fixed per-task overheads, yet present no risk of exceeding device energy capacity.
+As such a decomposition executes, \sys {\em coalesces} dynamically consecutive tasks.
+Coalescing elides the commits of coalesced tasks by buffering multiple tasks' updates in an FRAM commit buffer.
+Periodically, as the span of the coalesced task grows, \sys ends coalescing and commits the state of the coalesced task.
+If a power failure interrupts a sequence of coalesced tasks, \sys adaptively reduces the number of tasks in that sequence that it will
+coalesce, committing sooner in future executions. Consequently, \sys's coalescing mechanism allows a program to execute efficiently across a range of
 energy buffer sizes, avoiding transition overheads in larger buffers, and
-ensuring progress in smaller buffers. \TODO{articulate this last point better}
+ensuring progress in smaller buffers.
 
-We evaluated our system on a collection of benchmark programs taken from the
-literature, and compare directly to prior work, showing that \sys has high
-performance and can flexible target a variety of platforms without recompiling
-the code. We further demonstrate the efficiency and versatility of \sys with
-an end-to-end evaluation of a sensing and data processing application deployed
-in a lab environment. \TODO{blah blah blah eval}.
+\todo{Provide numerical results for this section}{Przemek}
 
-\TODO{Bullet contribs if there's space}
+\item {\bf Extensive \sys evaluation:} We evaluated \sys on a collection of benchmark programs~\cite[Sec.
+5]{chain}, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexible target a variety of platforms without recompiling the code.
+We further demonstrate the efficiency and versatility of \sys with an end-to-end evaluation of a sensing and data processing application deployed in a lab environment.
+
+\todo{add detailed text about evaluation results}{Przemek}
+
+\end{enumerate}

### 2017-07-10 paper/intro.tex
Second pass on the Intro ready

@@
 
-Two recent programming and execution model~\cite{alpaca,chain} proposed {\em task-based} paradigm, that require the programmer to statically decompose their program into tasks.
+Two recent works~\cite{alpaca,chain} proposed {\em task-based} programming and execution paradigm, that require a programmer to statically decompose a program into tasks.
 A task resembles a function with no callers.
@@
 4]{chain}.
-An important programming challenge in task-based models is that the length of execution of a task is limited by the total amount of energy that a device can buffer.
+An important programming challenge in task-based models is that the length of execution of a task is limited by the total amount of energy that a device can buffer and {\em does not change after compilation}.
 Assuming that input power is negligible compared to operating power, a task will never be able to complete if its execution consumes more energy than the system can buffer.
-Transitioning from one task to the next imposes a cost and excessively small tasks may be inefficient.
-Appropriately sizing tasks to ensure progress and high performance is a key challenge for task-based intermittent systems.
+On the other hand, transitioning from one task to the next imposes a cost and excessively small tasks may be inefficient.
 
-In this work, we develop \sys: a new programming and execution model that uses two new runtime features: (i) {\em volatile memory virtualization} to leverage efficient volatile memory and (ii) {\em dynamic task coalescing} to efficiently execute tasks without exceeding available energy.
-\sys's memory virtualization mechanism uses a device's SRAM as the working memory during a task's execution, populating the SRAM with a page of data from FRAM as necessary at the start of the task.
+This brings the following research question: {\bf how to approach the appropriate task size to ensure progress and high performance given an unknown energy buffer size and energy arrival process?} More specifically, we ask two questions: (i) how to eliminate the need for hardware support while enabling the intermittent execution model to dynamically adapt task size {\em after compilation}, and (ii) how to reduce the energy consumption caused by frequent memory swapping operations during task traversals?
+
+To answer these, in this work we develop {\bf \sys}: a new programming and execution model that utilizes two new runtime features: {\em dynamic task coalescing} to efficiently execute tasks without exceeding available energy (answering question (i)) and  {\em volatile memory virtualization} to leverage efficient volatile memory (answering question (ii)).
+\sys's novel memory virtualization mechanism uses a device's SRAM as the working memory during a task's execution, populating the SRAM with a page of data from FRAM as necessary at the start of the task.
 At the end of the task, the SRAM is committed back to FRAM using efficient DMA block copies~\cite{kistler_micro_2006}.
@@
 With this paper we provide the following list of contributions.
-
+%
 \begin{enumerate}
-
-\item {\bf Evaluation of data swapping strategies:} We propose and explore two data swapping strategies for intermittently-powered systems based on FRAM/SRAM architecture.
-The first strategy is {\em demand paging}, which swaps the SRAM page with a new page from FRAM, buffering the swapped-out page until commit.
-The second strategy is {\em buffered direct access}, which directly reads and writes FRAM relying on dynamic double-buffering to ensure memory is consistent at
-commit.
-
-\todo{Once paging is fully implemented--update the paragraph}{Przemek}
-\todo{Provide numerical results for this section}{Przemek}
-\todo{Find better name for demand paging}{Brandon}
-
+%
 \item {\bf Task coalescing mechanism:} \sys's core contribution is the introduction of dynamic task coalescing mechanism.
@@
 ensuring progress in smaller buffers.
-
+%
 \todo{Provide numerical results for this section}{Przemek}
-
+%
+\item {\bf Evaluation of data swapping strategies:} We propose and explore two data swapping strategies for intermittently-powered systems based on FRAM/SRAM architecture.
+The first strategy is {\em demand paging}, which swaps the SRAM page with a new page from FRAM, buffering the swapped-out page until commit.
+The second strategy is {\em buffered direct access}, which directly reads and writes FRAM relying on dynamic double-buffering to ensure memory is consistent at
+commit. 
+%
+\todo{Once paging is fully implemented--update the paragraph}{Przemek}
+\todo{Provide numerical results for this section}{Przemek}
+\todo{Find better name for demand paging}{Brandon}
+%
 \item {\bf Extensive \sys evaluation:} We evaluated \sys on a collection of benchmark programs~\cite[Sec.
@@
 We further demonstrate the efficiency and versatility of \sys with an end-to-end evaluation of a sensing and data processing application deployed in a lab environment.
-
+%
 \todo{add detailed text about evaluation results}{Przemek}
-
+%
 \end{enumerate}

### 2017-07-10 paper/paper.tex
Paragraph 1 revised with 10 new references added
Second pass on the Intro ready

@@
 \newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
+
+\newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}
+
 \newcommand{\sys}{VIPRE\xspace}
@@
 
-\subsection{Problem Statement}
-How to approach the optimal task size given an unknown energy buffer size and environment conditions? 
-	 \begin{enumerate}
-		 \item How to eliminate the need for the hardware support while enabling the Modular Intermittent Execution model to dynamically adapt its task size?
-		 \item How to enable code portability under the Modular Intermittent execution model? 
-		 \item How to reduce the energy consumption of executing a task? 
-	\end{enumerate}
-
 \subsection{Why not Adapting Current Approaches}

### 2017-07-10 paper/background.tex
background.tex added

--- 2017-07-06 paper/paper.tex
@@
-\documentclass[sigconf,anonymous,review]{acmart}
-
-\usepackage[normalem]{ulem}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{algorithm} % environment for the algorithm code (like figure and table)
-\usepackage{algorithmicx} % write the pseudo code
-\usepackage[noend]{algpseudocode} % layout of the pseudo code
-\usepackage{tikz} % needed for: tikz pics
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{natbib}
-\usepackage{graphicx}
-\usepackage{subfigure}
-%\usepackage[inline]{enumitem}
-
-\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
-\newcommand{\sys}{VIPRE\xspace}
-\newcommand{\sysfull}{Virtualized Intermittently-Powered Runtime Environment\xspace}
-
-\begin{document}
-
-\title{VIPRE: Virtualized Intermittently-Powered Runtime Environment}
-
-%Amjad Yousef Majid, Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m, Przemys{\l}aw Pawe{\l}czak,
-%TU Delft, Mekelweg\,4, 2628\,CD Delft, NL
-%\{a.y.majid, k.s.yildirim, p.pawelczak\}@tudelft.nl
-%
-%Kiwan Maeng, Alexei Collin, Brandon Lucia
-%Carnegie Mellon University, 4720 Forbes Avenue, Pittsburgh, PA 15213, USA
-%\{kmaeng, acolin, blucia\}@kmaeng@andrew.cmu.edu
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}
-\label{sec:intro}
-
-%\TODO{VIPER: Virtualised Intermittent Programming model, Execution model, and Runtime environment}
-%\TODO{PREVIEW: Programming and Runtime Environment with Virtualised Intermittent Execution Windows}
-
-\input{intro}
-
-\section{Computation on Intermittent Devices: Background}
-\label{sec:background}
-
 Performing complex operations on battery-less energy harvesting embedded platforms is challenging from many points of view.
@@
 \TODO{Elaborate on the ``key challenge'' paragraphs near the end of the intro.}
-
-\section{\sysfull}
-\label{sec:overeall_system}
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{\sys top-level description.}
-	\label{fig:}
-\end{figure}
-
-Describe the overall system, the programming model, and the semantics/execution model. This should include a discussion of memory virtualization and task coalescing.
-
-\section{Memory Virtualization}
-\label{sec:memory_virtulaization}
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Memory virtualization architecture.}
-	\label{fig:}
-\end{figure}
-
-\section{Task Coalescing}
-\label{sec:memory_virtulaization}
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Task coalescing architecture.}
-	\label{fig:}
-\end{figure}
-
-\section{Discussion}
-\label{sec:discussion}
-
-\section{Methodology and Evaluation}
-\label{sec:methodology_evaluation}
-
-\section{Related Work}
-\label{sec:related_work}
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-This paper presented \sys: \sysfull.
-
-%%%%%
-
-\section{Old Introduction}
-	Intermittently powered devices (IPDs) are battery-less devices that utilize the ambient energy to sense, compute and, communicate.
-For example, Wireless Identification and Sensing platform WISP \cite{wisp} uses the RF signal power to drive its computation and communication.
-Because of the reliance of TPDs on intermittent power supply, the harvested energy when it is available, their programs are susceptible to a very frequent power interrupts, in the order of tens of millisconds~\cite{}.
-Therefore, these devices require a different software execution model that complies with the nature of a discontinuous power supply.
-
-	The intermittent (discontinuous) execution model defines a program execution as cumulative discrete process.
-The main difference between the intermittent and the conventional (or continuous) execution models is that, a power failure is seen by the continuous model as an \emph{exception} that may reset the progress of a program to its beginning.
-Whereas, in intermittent execution a power failure is regarded as a temporary \emph{pause} to the execution that may result in some progress degradation.
-Generally we can classify the intermittent execution model into:
-	\begin{itemize}
-		\item \emph{Sequential Execution Model}:
-			Under this model a program is seen as one big idempotent operational region that has one common context.
-Generally, The progress of the program is saved and updated by means of checkpointing---where all the program context (e.g.
-CPU registers, the stack and the global variables) is saved to a non-volatile memory.
-Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos, harvOS, hibernus}.
-The benefit of this model is that it does not require code modification by a programmer.
-However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{xxxx}; and (iii) it restricts the IPDs to run only a single application~\cite{inos}.
-		\item \emph{Modular Intermittent Execution Model}:
-			At the heart of this model is the concept of an idempotent task.
-The idempotent task is C function that does not have arguments and does not return a value.
-This task uses a well defined interface to interact with the non-volatile memory.
-Therefore, it tolerates arbitrary number of power interrupts.
-This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an IPD by interleaving their tasks~\cite{inos}.
-However, it obviously requires code modification---for example, if an algorithm is written according to the continuous execution model it has to be splitted, by a programmer, into small tasks to run under the modular intermittent execution model.
-	\end{itemize}
-
-Despite that the superiority of the Modular Intermittent Execution model, it is still a static approach that completely depends on a programmer's estimation which is mostly result in a sub-optimal code devision.
-Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.
-
-% Energy source
-
-% The intermittent execution is cumulative discrete process.
-The intermittent execution model is only able to execute a few number of instructions, as compared to the conventional (continuous) model, before its progress is terminated.
-Therefore, the intermittent execution model adapts an execution progress state saving mechanism.
-This mechanism is realized either by means of a checkpoint~\cite{}, where all the context of the program is saved into non-volatile memory.
-Or
-
-% This progress state saving mechanism is normally injected into a program either by a compiler~\cite{}.
-A compiler injects trigger pointers to checkpoint (save) the context of the program into the non-volatile memory.
-or it is added by a programmer~\cite{chain}.
-[compare these two approaches]...
-
-% We define virtualization within the context of intermittent execution as utilizing the volatile memory instead of non-volatile when the intermittent execution model attempts to access the non-volatile memory.
-
-\subsection{Problem Statement}
-How to approach the optimal task size given an unknown energy buffer size and environment conditions? 
-	 \begin{enumerate}
-		 \item How to eliminate the need for the hardware support while enabling the Modular Intermittent Execution model to dynamically adapt its task size?
-		 \item How to enable code portability under the Modular Intermittent execution model? 
-		 \item How to reduce the energy consumption of executing a task? 
-	\end{enumerate}
-
-\subsection{Why not Adapting Current Approaches}
-Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes. 
-%
-%	\begin{figure}[t]
-%	    \centering
-%	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
-%	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
-%	    \caption{Task flow control of Chain}
-%	    \label{fig:DynamicChain}
-%	\end{figure}
-
-\noindent\textbf{Chain: Tasks and Channels for Reliable Intermittent Programs } \\
-Some of the main problems in modifying Chain to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Multi-versioning system:} Chain defines for each task its own input that is not shared with other tasks.
-This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
-	\item \emph{Direct FRAM system:} Chain tasks directly access FRAM which is more energy expensive Than SRAM.
-As result of being multi-versioning system that directly interacts with FRAM, Chain has a significant energy consumption overhead.
-	\item \emph{Irreversible tasks transitions:} According to the Chain programming model, each task calls the next one.
-The transition is done through functions calls and manually clearing the stack, to prevent the stack overflow problem.
-This approach make all the transitions firm ones which prevent task merging.
-	\item \emph{All channels are required:} Fig.~\ref{fig:DynamicChainSeq} shows that if the task control flow is a circular with no branches then task merging is possible.
-Moreover, all the intra-merged-tasks channel can be committed to SRAM instead of FRAM to save energy.
-However, this execution path is only a special.
-If the general case (see Fig.~\ref{fig:DynamicChainRan}) is considered then leaving out the intra-merged-task channels might result in data inconsistency---If an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready which result in an incorrect execution.
-
-\end{itemize}
-
-\noindent\textbf{Alpaca: Intermittent Execution without Checkpoints} \\
-Some of the main problems in modifying Alpaca to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Privatization:} The core concept of Alpaca is privatization.
-Privatization depends completely on detecting the variables that have a Write-after-read dependency within the scope of a task.
-Therefore, if tasks are on-demand merged the scope of these dependencies are changed and the static analysis to privatize these variables is not valid anymore and data consistency can not be preserved.
-
-	\item \emph{Irreversible tasks transitions:} Similar to Chain problem.
-
-\end{itemize}
-
-\noindent\textbf{Virtualizing Modular Intermittent Execution Model}
-
-		An Intermittently executed program, as seen by the modular intermittent execution model, is a chain of tasks with a firm transition from one to another. 
-		Virtualizing this model means softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
-Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
-Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
-This can be done by having a temporary copy of a global variable in the volatile memory (similar to the catching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
-
-		In this paper we are introducing VIPOS a runtime library the implement the Virtualed Intermittent Execution Model.
-VIPOS adapts two methods for data protection.
-One relies on a virtual buffer and the other uses the Direct Memory Access (DMA).
-VIPOS is able to approach the optimal tasks devision during the runtime.
-VIPOS reduces an application execution time by XX and the energy by XX.
-
-\subsection{Contributions}
-	 \begin{enumerate}
-		 \item We have developed an algorithm that is able to determine the size of a virtual task based on the history of the execution of an application. 
-		 \item We enable efficient code portability by requiring that a real task to small and merging them during the execution.   
-	\end{enumerate}
-
- \section{Preliminary Results}
- \label{sec:prelResults}
-
-\begin{figure}[t]
-    \centering
-         \subfigure[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
-         \subfigure[The energy cost of FRAM/SRAM \textbf{read} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
-    \caption{The energy cost of accessing volatile/non-volatile memory}
-    \label{fig:framEnergy}
-\end{figure}
-
-Since the Intermittent execution models rely on the non-volatile memory to enable long-running operation, it is important to characterize the energy consumption of accessing this type of memory as compare to the volatile one.
-
-Four applications are developed to measure the energy consumption of \emph{Accessing} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{xxx}.
-The interference free debugger (EDB)~\cite{xxx} is used as the energy measurement tool.
-The EDB probes the energy buffer before and after accessing the FRAM/SRAM 1600 times, This large number of FRAM access is used to increase the reliability of the results generated by the used measurement tool (e.g.
-reducing the effect of the quantization error).
-The energy buffer is charged, using the EDB, to $\approx$ 2.45 volts \textbf{only} at the beginning of the execution and of the programs.
-The write operations were performed by writing literal values, random numbers, to the memory.
-However, a read operation must be followed by a write operation, therefore, we chose to write a read value to SRAM.
-Fig.~\ref{fig:framEnergy} shows that an FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We would like to mentioned that TI has already mentioned in this document~\cite{xxxx} that FRAM write access is more energy expensive than SRAM write access.
-However, TI does not quantify the energy overhead.}.
-Therefore, reducing FRAM access is desirable when developing  energy limited software solution
-
-\section{VIPOS: Virtualized Intermittently Powered Runtime Environment}
-VIPOS is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs.
-It is able, on-demand, to merge tasks to reduce the number of commits to the non-volatile memory.
-
-\input{iposDataProtection}
-
-\subsection{VIPOS: The Virtualizing Engine}
-\subsubsection{Power Interrupt Immune Scheduler}
-It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures.
-VIPOS provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
-(un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
-
-\subsubsection{Task Merging Algorithms}
-
-\paragraph{Fixed Virtual Task Size} [better name needed]
-
-	\begin{algorithm}[t]
-		\caption{Fixed virtual Task size}
-		\label{algo:fixVirtTask}
-		\scriptsize
-		%\small
-		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{VIPOS Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\State MVTS: maximum VT size
-			\vspace{0.1cm}
-
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
-				\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-							\State $ MVTS = VTS $
-						\EndIf
-				\EndWhile
-
-				\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\If{$VTS < MVTS$}
-					\State $VTS++$
-					\EndIf
-				\EndIf
-			\EndWhile
-		\end{algorithmic}
-	\end{algorithm}
-
-	\begin{algorithm}[t]
-		\caption{Opportunistic virtual Task size}
-		\label{algo:fixVirtTask}
-		\scriptsize
-		%\small
-		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{VIPOS Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\vspace{0.1cm}
-
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
-				\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-						\EndIf
-				\EndWhile
-
-				\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\State $VTS++$
-				\EndIf
-			\EndWhile
-		\end{algorithmic}
-	\end{algorithm}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{figures/virtualTaskSize.eps}
-	\caption{Size of the virtual task versus the execution time of a dummy application that contains 12 empty tasks.}
-	\label{fig:virtualTaskSize}
-\end{figure}
-
-VIPOS is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
-Fig.~\ref{fig:virtualTaskSize} show the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
-
-Remark: A less obvious benefit of visualization is that it can enable a secure computation.
-By increasing the size of the virtual task, the energy in the super-capacitor will not be enough to finish the execution of the virtual task.
-Therefore, without the interrogator being close to the intermittently powered device and charging rate is not negligible as compared to the discharging rate this computation will not be finished.
-As such, virtualization can add a layer of security to the computation.
-
-\section{Results}
-We compare the performance of VIPOS against a state-of-the-art intermittent execution approach.
-For this purpose, we used two different applications: (i) A data decompression application, which utilizes Huffman decoding technique to decompress 100 byte of data.; (ii) A discrete Fourier transform (DFT) application that uses two different resolution (4 and 8 bites) to analyze a randomly generated signal.
-
-\subsection{Continuous Power Supply}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{figures/ipos_chain}
-	\caption{Time to complation of VIPOS versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
-	\label{fig:IPOSPerformance}
-\end{figure}
-
-Fig.~\ref{fig:IPOSPerformance} shows that VIPOS requires a shorter execution time to execute an application under the MIEM.
-The size of the virtual task that VIPOS uses equals One real task, in other words, no visualization is applied.
-
-\subsection{Intermittent Power Supply}
-...
-
-\bibliographystyle{abbrvnat}
-\bibliography{bib}
-
-\end{document}

### 2017-07-10 paper/paper.tex
background.tex added

@@
 
-Performing complex operations on battery-less energy harvesting embedded platforms is challenging from many points of view.
-Here we shall briefly provide a background on computation with energy harvesting embedded systems.
-
-\subsection{Energy Harvesting: Background}
-
-Supplying power to tiny embedded computers using batteries alone is not sustainable.
-Batteries are a great pollutant.
-A solution is to replace batteries with smaller, but more environment friendly capacitors and supply power directly from the ambient energy sources.
-
-What platforms enable intermittent computation: WISP~\cite{} (with its variants such as WISPCam~\cite{}), Moo~\cite{}, ambient backscatter~\cite{}, or computation RFID commercial platforms such as~\cite{}.
-In case of the above platforms, the main source of energy is electromagnetic radiation in the radio frequency range.
-
-Application include long-term battery-less monitoring.
-One example is the distributed monitoring of moisture of large plant fields during vegetation~\cite{}.
-Other applications include battery-less image capture and processing~\cite{}, distributed low-power networking~\cite{}.
-
-Energy provided from the ambient is not stable and difficult to predict accurately.
-Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation.
-Energy breaks happen every hundred of ms~\cite{}.
-
-Solution to the problem is to divide the program into smaller pieces that guarantee execution within a discharge region and track the state of the program in-between the program part execution times.
-
-\subsection{(Task-based) Intermittent Computing: Background}
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Two problems with fixed-size task in intermittent execution.
-(a) task underestimation: task executed at device with capacitor size $X$, will not execute at device with capacitor $X\gg Y$; (b) task overestimation: with stable energy source task tracking causes unnecessary overhead.}
-	\label{fig:fixed_task_problem}
-\end{figure}
-
-Various capacitor sizes of intermittent device result in a program being executed
-
-Other forms of intermittent platforms will share the same problem.
-One of them considers actuation platforms, which will be powered directly from energy harvesting sources~\cite{}.
-will share the energy storage with computing platform.
-However, power supply for actuators is order of magnitude larger than for computation.
-Therefore, although capacitor is large enough to perform computation alone, energy to power actuators takes precedence leaving not much space for computing.
-
-\TODO{Elaborate on the ``key challenge'' paragraphs near the end of the intro.}
+\input{background}
 

### 2017-07-10 paper/background.tex
first paragraph of background ready

@@
 Supplying power to tiny embedded computers using batteries alone is not sustainable.
-Batteries are a great pollutant.
-A solution is to replace batteries with smaller, but more environment friendly capacitors and supply power directly from the ambient energy sources.
+European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
+Since not all batteries are properly recycled, they constitute a potential environmental hazard.
+A potent solution is to replace batteries with smaller, but more environment friendly (super-)capacitors (promising millions of charge/discharge cycles~\cite[Sec.
+I]{ongaro_pwre_2012}) and to supply them directly from the ambient energy sources.
+The limitation of the capacitor as storage is lower energy density than battery, making effective capacitor discharge times orders of magnitude smaller than for battery.
 
-What platforms enable intermittent computation: WISP~\cite{} (with its variants such as WISPCam~\cite{}), Moo~\cite{}, ambient backscatter~\cite{}, or computation RFID commercial platforms such as~\cite{}.
+Many platforms already enable intermittent computation: WISP~\cite{} (with its variants such as WISPCam~\cite{}), Moo~\cite{moo}, ambient backscatter~\cite{liu_sigcomm_2013}, or computation RFID commercial platforms such as~\cite{}.
 In case of the above platforms, the main source of energy is electromagnetic radiation in the radio frequency range.
+Naturally, other forms of energy harvesting exist, The reader is refered to extensive surveys~\cite{}.
 
@@
 One example is the distributed monitoring of moisture of large plant fields during vegetation~\cite{}.
-Other applications include battery-less image capture and processing~\cite{}, distributed low-power networking~\cite{}.
+Other applications include battery-less image capture and processing~\cite{}, distributed low-power networking~\cite{} or implantable sensors~\cite{rodriguez_tbcs_2015}.
 

### 2017-07-11 paper/background.tex
Expanding background section

@@
 
-Many platforms already enable intermittent computation: WISP~\cite{} (with its variants such as WISPCam~\cite{}), Moo~\cite{moo}, ambient backscatter~\cite{liu_sigcomm_2013}, or computation RFID commercial platforms such as~\cite{}.
+Many platforms already enable intermittent computation: WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, ambient backscatter~\cite{liu_sigcomm_2013}, or commercial computation RFID platforms such as~\cite{medusa_farsens_2017}.
 In case of the above platforms, the main source of energy is electromagnetic radiation in the radio frequency range.
-Naturally, other forms of energy harvesting exist, The reader is refered to extensive surveys~\cite{}.
+Naturally, other forms of energy harvesting sources exist: temperature gradient, light/sun radiation, vibrations, (micro-)motions, fluid flow or body fluids (blood, gastric acid).
+The reader is referred to example extensive surveys discussing energy harvesting for low-power embedded systems in detail~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016}.
 
@@
 One example is the distributed monitoring of moisture of large plant fields during vegetation~\cite{}.
-Other applications include battery-less image capture and processing~\cite{}, distributed low-power networking~\cite{} or implantable sensors~\cite{rodriguez_tbcs_2015}.
+Other applications include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, distributed low-power networking~\cite{}, insect monitoring~\cite{thomas_jbcs_2012} or implantable sensors~\cite{rodriguez_tbcs_2015}.
 
@@
 Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation.
-Energy breaks happen every hundred of ms~\cite{}.
+Energy breaks happen every hundred of ms~\cite{soyata_csm_2016,}.
 

### 2017-07-11 paper/background.tex
Expanding background section

--- 2017-07-10 paper/background.tex
@@
 Supplying power to tiny embedded computers using batteries alone is not sustainable.
-Batteries are a great pollutant.
-A solution is to replace batteries with smaller, but more environment friendly capacitors and supply power directly from the ambient energy sources.
+European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
+Since not all batteries are recycled, they constitute a potential environmental hazard.
+A potent solution is to replace batteries with smaller, but environment friendly {(super-)capacitors} (promising millions of charge/discharge cycles~\cite[Sec.
+I]{ongaro_pwre_2012}) and to supply them {\em directly from the ambient energy sources}---creating a truly self-sustainable computing device.
+The limitation of the capacitor as storage is its lower than battery energy density, making effective capacitor discharge times orders of magnitude smaller than for a battery.
 
-What platforms enable intermittent computation: WISP~\cite{} (with its variants such as WISPCam~\cite{}), Moo~\cite{}, ambient backscatter~\cite{}, or computation RFID commercial platforms such as~\cite{}.
-In case of the above platforms, the main source of energy is electromagnetic radiation in the radio frequency range.
+Many platforms enable intermittent computation already.
+Those include computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
+Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
+In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
+Naturally, other forms of energy harvesting sources exist, including temperature gradient, light/sun radiation, vibrations, (micro-)motions, and body fluid flow (blood, gastric acid).
+The reader is referred to example extensive surveys discussing energy harvesting for low-power embedded systems~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016}.
 
@@
 One example is the distributed monitoring of moisture of large plant fields during vegetation~\cite{}.
-Other applications include battery-less image capture and processing~\cite{}, distributed low-power networking~\cite{}.
+Other applications include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, insect monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}  and digestible~\cite{nadeau_naturebio_2017} sensors.
+
+\begin{table}
+	\begin{tabular}{|c|c|}
+	\hline
+	Platform Name & Storage capacitor size \\
+	\hline \hline
+	NeuralWISP~\cite{} & 100\,$\mu$F \\
+	BioImpedance~\cite{} & 20\,$\mu$F \\
+	WISPCam~\cite{} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
+	NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
+	WISP~\cite{} & 47\,$\mu$F \\
+	\hline
+	\end{tabular} 
+\caption{Comparison of storage sizes for various battery-less platforms.}
+\label{table:capacitor}
+\end{table}
+
+%DragonFly~\cite{} Not reported
+%Ambient Backscatter~\cite{,} Not reported
+%Fasense~\cite{} Not reported
+%Battery-less phone~\cite{} Not reported
 
@@
 Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation.
-Energy breaks happen every hundred of ms~\cite{}.
+Moreover, each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}.
+Energy breaks happen every hundred of milliseconds~\cite{soyata_csm_2016}.
 
-Solution to the problem is to divide the program into smaller pieces that guarantee execution within a discharge region and track the state of the program in-between the program part execution times.
+Solution to the problem is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between the program part execution times.
 
@@
 
+We can write the problem formally as
+%
+\begin{equation}
+\underset{i}{\arg\,\min} \sum_{i}m_i+t_i \text{~subject to~} \forall_i mi+t_i>e
+\end{equation}
+
 Other forms of intermittent platforms will share the same problem.

### 2017-07-11 paper/background.tex
Expanding background section

@@
 	\hline
-	Platform Name & Storage capacitor size \\
+	Platform name & Storage capacitor size \\
 	\hline \hline
-	NeuralWISP~\cite{} & 100\,$\mu$F \\
-	BioImpedance~\cite{} & 20\,$\mu$F \\
-	WISPCam~\cite{} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
+	NeuralWISP~\cite{holleman_biocas_2008} & 100\,$\mu$F \\
+	BioImpedance~\cite{rodriguez_tbcs_2015} & 20\,$\mu$F \\
+	WISPCam~\cite{naderiparizi_rfid_2015} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
 	NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
-	WISP~\cite{} & 47\,$\mu$F \\
+	Moo~\cite{moo} & X\,$mu$F \\
+	WISP~\cite{wisp5} & 47\,$\mu$F \\
+	X~\cite{nadeau_naturebio_2017}~\cite{} & X\,$mu$F\\
 	\hline
 	\end{tabular} 
-\caption{Comparison of storage sizes for various battery-less platforms.}
+\caption{Comparison of storage sizes for various battery-less platforms.
+Values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.}
 \label{table:capacitor}
@@
 
-We can write the problem formally as
+We can write the problem of maximizing all task sizes formally as
 %
 \begin{equation}
-\underset{i}{\arg\,\min} \sum_{i}m_i+t_i \text{~subject to~} \forall_i mi+t_i>e
+\underset{i}{\arg\,\max} \sum_{i}m_i+t_i \text{~subject to~} \forall_i mi+t_i>e,
 \end{equation}
+%
+where $e$ is the required minimum energy to perform a task, $i$ is the total number of tasks, $m_i$ and $t_i$ are the cost of task traversal (memory copying time of task variables) and time to complete one task, respectively.
 

### 2017-07-11 paper/background.tex
Expanding background section (part 1 - harvesting - kind of ready)

@@
 
-Many platforms enable intermittent computation already.
-Those include computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
+Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible.
+These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, insect monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}  and digestible~\cite{nadeau_naturebio_2017} sensors.
+
+Many platforms enable intermittent, battery-less computation.
+For instance, computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
 Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
 In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
-Naturally, other forms of energy harvesting sources exist, including temperature gradient, light/sun radiation, vibrations, (micro-)motions, and body fluid flow (blood, gastric acid).
+Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid).
 The reader is referred to example extensive surveys discussing energy harvesting for low-power embedded systems~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016}.
-
-Application include long-term battery-less monitoring.
-One example is the distributed monitoring of moisture of large plant fields during vegetation~\cite{}.
-Other applications include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, insect monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}  and digestible~\cite{nadeau_naturebio_2017} sensors.
 
@@
 	\hline
-	Platform Name & Storage capacitor size \\
+	Platform name & Storage capacitor size \\
 	\hline \hline
-	NeuralWISP~\cite{} & 100\,$\mu$F \\
-	BioImpedance~\cite{} & 20\,$\mu$F \\
-	WISPCam~\cite{} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
+	Moo~\cite{moo} & 0.1\,F \\
+	WISPCam~\cite{naderiparizi_rfid_2015} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
 	NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
-	WISP~\cite{} & 47\,$\mu$F \\
+	NeuralWISP~\cite{holleman_biocas_2008} & 100\,$\mu$F \\
+	WISP~\cite{wisp5} & 47\,$\mu$F \\
+	{\em BioImpedance} sensor~\cite{rodriguez_tbcs_2015} & 20\,$\mu$F \\
+	{\em Ingestible} sensor~\cite{nadeau_naturebio_2017} & 220\,$\mu$F\\
 	\hline
 	\end{tabular} 
-\caption{Comparison of storage sizes for various battery-less platforms.}
+\caption{Comparison of {\em default} energy storage sizes for various battery-less platforms.
+Observe a huge variation in storage capacity.
+We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.}
 \label{table:capacitor}
@@
 
-%DragonFly~\cite{} Not reported
-%Ambient Backscatter~\cite{,} Not reported
-%Fasense~\cite{} Not reported
-%Battery-less phone~\cite{} Not reported
-
-Energy provided from the ambient is not stable and difficult to predict accurately.
+The biggest technical problem with energy provided from the ambient to battery-less platforms is that harvested energy is not stable and difficult to predict accurately.
 Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation.
-Moreover, each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}.
-Energy breaks happen every hundred of milliseconds~\cite{soyata_csm_2016}.
-
-Solution to the problem is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between the program part execution times.
+Energy breaks happen every hundred of milliseconds, refer again to e.g.~\cite[Fig.
+1]{mementos}.
+This causes each program running on such platform to restart.
+Solution to the problem is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between execution times.
 
@@
 
-Various capacitor sizes of intermittent device result in a program being executed
+Moreover, each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}, which causes different charge/discharge periods.
+This result in different execution times.
+In the worst case the size of the capacitor might be too small to execute one task written by the programmer.
 
@@
 
+%prior systems do fine grained task division (virtualization) {clank,aplaca}
+
 \TODO{Elaborate on the ``key challenge'' paragraphs near the end of the intro.}

### 2017-07-14 paper/paging.tex
paging related things

One of the main objectives of VIPRE run-time is to restrict virtual tasks to interface only with the volatile memory, since volatile memory access is relatively cheaper non-volatile memory access. To this end, the persistent variables a virtual task operates on should be brought from non-volatile memory to a fixed-sized \emph{volatile buffer} and when the virtual task is finished the modified variables in this buffer should be committed back to their original locations in non-volatile memory. However, this is not a trivial task due to the following facts:

\paragraph{I-Dynamic Access Overhead} The persistent variables accessed and/or modified by a virtual task might not be known in advance due to the dynamic program flow. Therefore, the  volatile buffer should keep track of the persistent variables accessed by any virtual task at runtime. When a persistent variable is read or written, first the volatile copy of the corresponding variable is \emph{searched} in this buffer. If it is found, the read or write operation is performed on the volatile copy. Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations. Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead and might kill the benefit of task virtualization.

\paragraph{II-Two Phase Commit Overhead}  For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each virtual task by means of a \emph{two-phase commit} approach.  However, two-phase commit is also very expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed with CPU intervention.

\paragraph{III-Main Memory Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support. In order to eliminate run-time buffer \emph{search} overhead, \emph{all} persistent variables can be brought from non-volatile memory to the volatile buffer at system start-up---very efficiently and faster thanks to Direct memory access (DMA) that eliminates CPU intervention for \emph{block} memory-memory copy/move operations. Two-phase commit is also very efficient in this case since DMA can be utilized to move volatile buffer as a block to non-volatile locations. However, bringing all persistent variables from non-volatile memory to the volatile buffer suffers from the \emph{memory fit problem}: the volatile memory is a scarce resource and may not be large enough to hold all persistent variables.

Considering these facts, we designed a \emph{paging-based} virtual memory in software that (i) partitions contiguously allocated persistent variables into pages; (ii) implements the volatile \emph{page buffer} that holds a single page---solves the aforementioned memory-fit problem; (ii) swaps the page in volatile memory with a new page in non-volatile memory efficiently by utilizing DMA-based memory block-copy operations; (iii) buffers the swapped volatile page until commit to enable task virtualization; (iv) implements two-phase commit of the modified pages to preserve the consistency of non-volatile memory; (v) and performs two-phase commit operations using DMA for the sake of efficiency. We explain the details of our design in the following subsections.

\subsection{Address Translation and Variable Access}

Our virtual memory system provides an interface to the applications which is composed of two C macros: \texttt{RVAR(var)} and \texttt{WVAR(var,val)} that accept the name of the persistent variable \texttt{var} as an input argument. These macros obtain the \emph{physical address} of the corresponding persistent variable in non-volatile memory and translates it into the \emph{virtual address} in volatile memory which is composed of a \emph{page tag} and \emph{offset}. For the sake of simplicity and efficiency, we implemented this translation in a trivial way: the first byte of the physical address is considered to be the page tag and the rest is the offset within the corresponding page.

\begin{figure}
	\centering
	%\includegraphics[width=0.25\columnwidth]{figures/}
	\caption{Address translation}
	\label{fig:address-translation}
\end{figure}

\begin{algorithm}[t]
	\caption{\texttt{RWAR(var)} pseudo-code}
	\label{algo:rwar}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
		\State $\texttt{tag}\leftarrow \texttt{getTag(var)}$ 
		\If { \texttt{tag} != \texttt{CrntPagTag} }	\Comment{Check if the page is in page buffer}
		\State	\texttt{PageFault(tag)} \Comment{Page is not in the page buffer, bring it}
		\EndIf
				\State $\texttt{offset}\leftarrow \texttt{getOffset(var)}$ 		
		\State \texttt{return pageBuf[offset]}  \Comment{Return directly from page buffer}
	\end{algorithmic}
\end{algorithm}

After address translation and obtaining the virtual address, it is required to check if the corresponding page is already in volatile page buffer. To this end, our system maintains a \texttt{CrntPagTag} variable that holds the tag of the page currently in the page buffer. If the tag of the virtual address and \texttt{CrntPagTag} are equal, this indicates that page buffer holds the required page and the offset of the virtual address is used to read from/write to the location in the volatile page buffer immediately. If the required page is not in the page buffer, a \emph{page fault} routine is executed as we discuss in the next subsection. Observe that the computational complexity of the whole steps pertaining to address translation and variable access is $\mathcal{O}(1)$---introducing minimal overhead to the system to boost up benefits of volatile-only memory interfacing. The pseudo-code of \texttt{RWAR} is presented in Algorithm~\ref{algo:rwar}.

\subsection{Page Faults and Swapping}

\begin{algorithm}[t]
	\caption{\texttt{PageFault(tag)} pseudo-code}
	\label{algo:rwar}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
		\If { \texttt{isModified(CrntPagTag)} }	\Comment{Check if the active page modified}
		\State commit current page to \texttt{pagesTemp}
		\EndIf
		\If { tag is in \texttt{pagesTemp} }	\Comment{Check if the requested page has been modified}
		\State bring from \texttt{pagesTemp} to \texttt{pageBuff} 
		\Else
		\State bring from original to \texttt{pageBuff} 
		\EndIf 
	\end{algorithmic}
\end{algorithm}

If the tag of the virtual address and \texttt{CrntPagTag} are not equal, the active page in page buffer should be swapped with the requested page. In this situation, we have the following two cases:
\paragraph{Modified Active Page:} If the contents of the page buffer is modified, we commit it to an intermediate page buffer in non-volatile memory, namely to \texttt{pagesTemp}. If it is not modified, there is no need to commit the active page.

\paragraph{Modified Requested Page:}  If the requested page is previously modified, it should be brought from \texttt{pagesTemp} rather than its original location in non-volatile memory.

After the swapping, the \texttt{CrntPagTag} is set to the tag of the requested page. It should be noted that DMA is used to introduce minimal overhead during page swapping. The pseudo-code of the page fault handling is presented in Algorithm~\ref{algo:pagefault}.

\subsection{Page Committing}

Using a two-phase commit by utilizing DMA.

### 2017-07-17 paper/background.tex
ASPLOS latex template replaced with a correct one/bib file updated following the style

@@
 \subsection{Energy Harvesting: Background}
+\label{sec:background_harvesting}
 
@@
 
+\subsection{Memory Consistency}
+\label{sec:background_consistency}
+
+Naturally, content of a volatile memory (i.e.
+run-time stack, global and local variables) of the microcontroller and program registers are erased after each power failure.
+Only what resides in a non-volatile memory, e.g.
+FRAM, persists after system reboot.
+A solution to sustain the progress of the program from the last power failure, instead from the beginning (the \emph{Sisyphus} effect~\cite[Sec.
+2]{mementos})is a periodic checkpointing of a program state~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
+The periodicity, and a position of a checkpoint, is the core of the contribution behind each of the proposed checkpoining solution.
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=0.25\columnwidth]{figures/code_demo_incosistency}
+	%	for (byte = 0; byte < nBytes; ++byte){\\
+	%	checkpoint();\\
+		%data = reflectData(message\[byte\]) $^$ (remainder >> (WIDTH - 8));\\
+		%remainder = crcTable[data] $^$ (remainder << 8);}
+	\caption{Code demonstrating WAR and volatile memory checkpointing.}
+	\label{fig:code_demo_incosistency}
+\end{figure}
+
+Fig.~\ref{fig:code_demo_incosistency} demonstrates the problem through example.
+
 \subsection{(Task-based) Intermittent Computing: Background}
@@
 
-\TODO{Elaborate on the ``key challenge'' paragraphs near the end of the intro.}
+%Q1: Does Program Counter change size with task (checkpoint) size in Chain? In other words: bigger Chain task=bigger Program Counter?
+%Q2: Does Register File (RF) includes Program Counter (PC)? In other words: is always PC>RF in terms of memory size?
+
+%N_{checkpoints} \times e^{2\times N_{shared}}
+
+\todo{Elaborate on the ``key challenge'' paragraphs near the end of the intro}{Przemek}
+\todo{Define the Task and Task based execution model}{Przemek}

### 2017-07-17 paper/intro.tex
ASPLOS latex template replaced with a correct one/bib file updated following the style

@@
 
-\todo{Put exact reference for "tictpl"}{Brandon}
-\todo{Check if the list of authors for Alpaca is correct}{Brandon}
+\todo{Put reference for "tictpl" somewhere}{Brandon}
 
@@
 It allows the programmer or a compiler to intuitively decompose the program into small tasks that amortize fixed per-task overheads, yet present no risk of exceeding device energy capacity.
-As such a decomposition executes, \sys {\em coalesces} dynamically consecutive tasks.
+As such a decomposition executes, \sys~{\em coalesces dynamically} consecutive tasks.
 Coalescing elides the commits of coalesced tasks by buffering multiple tasks' updates in an FRAM commit buffer.
 Periodically, as the span of the coalesced task grows, \sys ends coalescing and commits the state of the coalesced task.
-If a power failure interrupts a sequence of coalesced tasks, \sys adaptively reduces the number of tasks in that sequence that it will
-coalesce, committing sooner in future executions. Consequently, \sys's coalescing mechanism allows a program to execute efficiently across a range of
-energy buffer sizes, avoiding transition overheads in larger buffers, and
-ensuring progress in smaller buffers.
+If a power failure interrupts a sequence of coalesced tasks, \sys adaptively reduces the number of tasks in that sequence that it will coalesce, committing sooner in future executions.
+Consequently, \sys's coalescing mechanism allows a program to execute efficiently across a range of energy buffer sizes, avoiding transition overheads in larger buffers, and ensuring progress in smaller buffers.
 %
 \todo{Provide numerical results for this section}{Przemek}
+\todo{Mention automatic task generation}{Przemek}
 %
@@
 \todo{Provide numerical results for this section}{Przemek}
-\todo{Find better name for demand paging}{Brandon}
 %

### 2017-07-17 paper/paper.tex
first paragraph of background ready
Expanding background section
paging related things
change the name, this first name include run-time and our system include run time and compile time components
ASPLOS latex template replaced with a correct one/bib file updated following the style

@@
-\documentclass[sigconf,anonymous,review]{acmart}
+%\documentclass[sigconf,anonymous,review]{acmart}
+\documentclass[pageno]{jpaper}
+
+%replace XXX with the submission number you are given from the ASPLOS submission site.
+\newcommand{\asplossubmissionnumber}{XXX}
 
@@
 \usepackage{natbib}
-\usepackage{graphicx}
-\usepackage{subfigure}
+%\usepackage{graphicx}
+%\usepackage{subfigure}
 %\usepackage[inline]{enumitem}
@@
 \newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
+\newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}
+\newcommand{\sys}{VIPE\xspace}
+\newcommand{\sysfull}{Virtualized Intermittently-Powered Executor\xspace}
 
-\newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}
+%\setcopyright{rightsretained}
 
-\newcommand{\sys}{VIPRE\xspace}
-\newcommand{\sysfull}{Virtualized Intermittently-Powered Runtime Environment\xspace}
+%\acmDOI{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
 
-\setcopyright{rightsretained}
+%\acmISBN{978-1-xxxx-xxxx-x/17/11}
 
-\acmDOI{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
+%\acmConference[ASPLOS]{The 23rd ACM International Conference on Architectural Support for Programming Languages and Operating Systems}{March 24--28, 2017}{Williamsburg, VA, USA}
 
-\acmISBN{978-1-xxxx-xxxx-x/17/11}
+%\acmYear{2017}
 
-\acmConference[ASPLOS]{The 23rd ACM International Conference on Architectural Support for Programming Languages and Operating Systems}{March 24--28, 2017}{Williamsburg, VA, USA}
+%\copyrightyear{2017}
 
-\acmYear{2017}
-
-\copyrightyear{2017}
-
-\acmPrice{15.00}
+%\acmPrice{15.00}
 
@@
 
-\title[VIPRE: Virtualized Intermittently-Powered Runtime Environment]{VIPRE: Virtualized Intermittently-Powered Runtime Environment}
+\title{VIPE: Virtualized Intermittently-Powered Executor}
 
@@
 
-\author{Kiwan Maeng}
+%\author{Kiwan Maeng}
 %\authornote{}
 %\orcid{}
-\affiliation{%
-	\institution{Carnegie Mellon University}
-	\streetaddress{4720 Forbes Avenue}
-	\city{Pittsburgh} 
-	\state{PA} 
-	\postcode{15213}
-}
-\email{kmaeng@andrew.cmu.edu}
+%\affiliation{%
+%	\institution{Carnegie Mellon University}
+%	\streetaddress{4720 Forbes Avenue}
+%\city{Pittsburgh} 
+%	\state{PA} 
+%	\postcode{15213}
+%}
+%\email{kmaeng@andrew.cmu.edu}
 
-\author{Alexei Colin}
+%\author{Alexei Colin}
 %\authornote{}
 %\orcid{}
-\affiliation{%
-	\institution{Carnegie Mellon University}
-	\streetaddress{4720 Forbes Avenue}
-	\city{Pittsburgh} 
-	\state{PA} 
-	\postcode{15213}
-}
-\email{acolin@andrew.cmu.edu}
+%\affiliation{%
+%	\institution{Carnegie Mellon University}
+%	\streetaddress{4720 Forbes Avenue}
+%	\city{Pittsburgh} 
+%	\state{PA} 
+%	\postcode{15213}
+%}
+%\email{acolin@andrew.cmu.edu}
 
-\author{Brandon Lucia}
+%\author{Brandon Lucia}
 %\authornote{}
 %\orcid{}
-\affiliation{%
-	\institution{Carnegie Mellon University}
-	\streetaddress{4720 Forbes Avenue}
-	\city{Pittsburgh} 
-	\state{PA} 
-	\postcode{15213}
-}
-\email{blucia@andrew.cmu.edu}
+%\affiliation{%
+%	\institution{Carnegie Mellon University}
+%	\streetaddress{4720 Forbes Avenue}
+%	\city{Pittsburgh} 
+%	\state{PA} 
+%	\postcode{15213}
+%}
+%\email{blucia@andrew.cmu.edu}
 
-\author{Amjad Yousef Majid}
+%\author{Amjad Yousef Majid}
 %\authornote{}
 %\orcid{}
-\affiliation{%
-	\institution{Delft University of Technology}
-	\streetaddress{Mekelweg 4}
-	\city{Delft, The Netherlands} 
-	\state{Zuid Holland} 
-	\postcode{2628\,CD}
-}
-\email{a.y.majid@tudelft.nl}
+%\affiliation{%
+%	\institution{Delft University of Technology}
+%	\streetaddress{Mekelweg 4}
+%	\city{Delft, The Netherlands} 
+%	\state{Zuid Holland} 
+%	\postcode{2628\,CD}
+%}
+%\email{a.y.majid@tudelft.nl}
 
-\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
+%\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
 %\authornote{}
 %\orcid{}
-\affiliation{%
-	\institution{Delft University of Technology}
-	\streetaddress{Mekelweg 4}
-	\city{Delft, The Netherlands} 
-	\state{Zuid Holland} 
-	\postcode{2628\,CD}
-}
-\email{k.s.yildirim@tudelft.nl}
+%\affiliation{%
+%	\institution{Delft University of Technology}
+%	\streetaddress{Mekelweg 4}
+%	\city{Delft, The Netherlands} 
+%	\state{Zuid Holland} 
+%	\postcode{2628\,CD}
+%}
+%\email{k.s.yildirim@tudelft.nl}
 
-\author{Przemys{\l}aw Pawe{\l}czak}
+%\author{Przemys{\l}aw Pawe{\l}czak}
 %\authornote{}
 %\orcid{}
-\affiliation{%
-	\institution{Delft University of Technology}
-	\streetaddress{Mekelweg 4}
-	\city{Delft, The Netherlands} 
-	\state{Zuid Holland} 
-	\postcode{2628\,CD}
-}
-\email{p.pawelczak@tudelft.nl}
+%\affiliation{%
+%	\institution{Delft University of Technology}
+%	\streetaddress{Mekelweg 4}
+%	\city{Delft, The Netherlands} 
+%	\state{Zuid Holland} 
+%	\postcode{2628\,CD}
+%}
+%\email{p.pawelczak@tudelft.nl}
 
-\renewcommand{\shortauthors}{A. Y. Majid et al.}
+%\renewcommand{\shortauthors}{A. Y. Majid et al.}
+
+\date{}
+\maketitle
+
+\thispagestyle{empty}
 
@@
 
-\begin{CCSXML}
-	<ccs2012>
-	<concept>
-	<concept_id>10010520.10010553.10010562.10010564</concept_id>
-	<concept_desc>Computer systems organization~Embedded 
-	software</concept_desc>
-	<concept_significance>500</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10010520.10010575.10010578</concept_id>
-	<concept_desc>Computer systems organization~Availability</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
-	<concept_desc>Software and its engineering~Scheduling</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
-	<concept_desc>Software and its engineering~Communications 
-	management</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	<concept>
-	<concept_id>10011007.10010940.10010971.10010564</concept_id>
-	<concept_desc>Software and its engineering~Embedded 
-	software</concept_desc>
-	<concept_significance>300</concept_significance>
-	</concept>
-	</ccs2012>
-\end{CCSXML}
+%\begin{CCSXML}
+%	<ccs2012>
+%	<concept>
+%	<concept_id>10010520.10010553.10010562.10010564</concept_id>
+%	<concept_desc>Computer systems organization~Embedded 
+%	software</concept_desc>
+%	<concept_significance>500</concept_significance>
+%	</concept>
+%	<concept>
+%	<concept_id>10010520.10010575.10010578</concept_id>
+%	<concept_desc>Computer systems organization~Availability</concept_desc>
+%	<concept_significance>300</concept_significance>
+%	</concept>
+%	<concept>
+%	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
+%	<concept_desc>Software and its engineering~Scheduling</concept_desc>
+%	<concept_significance>300</concept_significance>
+%	</concept>
+%	<concept>
+%	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
+%	<concept_desc>Software and its engineering~Communications 
+%	management</concept_desc>
+%	<concept_significance>300</concept_significance>
+%	</concept>
+%	<concept>
+%	<concept_id>10011007.10010940.10010971.10010564</concept_id>
+%	<concept_desc>Software and its engineering~Embedded 
+%	software</concept_desc>
+%	<concept_significance>300</concept_significance>
+%	</concept>
+%	</ccs2012>
+%\end{CCSXML}
 
-\ccsdesc[500]{Computer systems organization~Embedded software}
-\ccsdesc[300]{Computer systems organization~Availability}
-\ccsdesc[300]{Software and its engineering~Scheduling}
-\ccsdesc[300]{Software and its engineering~Communications management}
-\ccsdesc[300]{Software and its engineering~Embedded software}
+%\ccsdesc[500]{Computer systems organization~Embedded software}
+%\ccsdesc[300]{Computer systems organization~Availability}
+%\ccsdesc[300]{Software and its engineering~Scheduling}
+%\ccsdesc[300]{Software and its engineering~Communications management}
+%\ccsdesc[300]{Software and its engineering~Embedded software}
 
-\keywords{Energy harvesting, transient operation, operating system}
-
-\maketitle
+%\keywords{Energy harvesting, transient operation, operating system}
 
@@
 
-\section{Computation on Intermittent Devices: Background}
+\section{Intermittent Computation: Background}
 \label{sec:background}
@@
 \label{sec:memory_virtulaization}
-
+\input{paging}
 \begin{figure}
@@
 
-%%%%%
+%%%%%Old part of the paper - will be successfully moved section by section to a final one%%%%%
+
+\newpage
 
@@
     \centering
-         \subfigure[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
-         \subfigure[The energy cost of FRAM/SRAM \textbf{read} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
+         \subfloat[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
+         \subfloat[The energy cost of FRAM/SRAM \textbf{read} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
     \caption{The energy cost of accessing volatile/non-volatile memory}
@@
 
-\bibliographystyle{abbrvnat}
+\bibliographystyle{plain}
 \bibliography{bib}

### 2017-07-17 paper/iposDataProtection.tex
iposDataProtection error corrected

@@
 		\centering
-         \subfigure[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize_time.eps} }
-	     \subfigure[The energy needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
+         \subfloat[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize_time.eps} }
+	     \subfloat[The energy needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
 		\caption{Time and energy consumption of moving a block of data from SRAM to FRAM}

### 2017-07-19 paper/background.tex
added section on the type of computing HW

@@
 	\centering
+	\fbox{\texttt{text}}
 	%\includegraphics[width=0.25\columnwidth]{figures/code_demo_incosistency}
@@
 Fig.~\ref{fig:code_demo_incosistency} demonstrates the problem through example.
+
+\subsection{Assumption on the Type of Computing Hardware}
+\label{sec:background_hardware}
+
+We assume that we operate on the microcontroller (more generally, computing hardware) equipped with a mix of volatile and non-volatile memory~\cite{}, such as FLASH, Ferroelectric RAM or Magnetoresistive RAM.
+\sys does not assume to rely on specific memory type (either volatile or non-volatile), in contrary to earlier works targeting non-volatile processors~\cite{ratchet}.
+Actually, \sys will exploit the properties of both memory types to maximize performance metrics of the program being executed.
+We make \sys able to I/O operations, however we do not enable the scheduling of operations (which nota bene forms a foundation for a simple runtime).
+All in all, our hardware model follows in assumptions the two most relevant intermittently-powered execution environments~\cite{alpaca,chain}.
 

### 2017-07-19 paper/background.tex
corrections on section the type of computing HW

--- 2017-07-17 paper/background.tex
@@
 Fig.~\ref{fig:code_demo_incosistency} demonstrates the problem through example.
+
+\subsection{Assumption on the Type of Computing Hardware}
+\label{sec:background_hardware}
+
+We assume that we operate on the microcontroller (more generally, computing hardware) equipped with a mix of volatile and non-volatile memory, such as FLASH, Ferroelectric RAM or Magnetoresistive RAM.
+\sys does not assume to rely on specific memory type (either volatile or non-volatile), in contrary to earlier works targeting non-volatile processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
+Actually, \sys will exploit the properties of both memory types to maximize performance metrics of the program being executed.
+We make \sys able to I/O operations, however we do not enable the scheduling of operations (which \emph{nota bene} forms a foundation for a simple runtime).
+All in all, our hardware model follows in assumptions the two most relevant intermittently-powered execution environments~\cite{alpaca,chain}.
 

### 2017-07-23 paper/background.tex
WAR inconsistency section complete

@@
 I]{ongaro_pwre_2012}) and to supply them {\em directly from the ambient energy sources}---creating a truly self-sustainable computing device.
-The limitation of the capacitor as storage is its lower than battery energy density, making effective capacitor discharge times orders of magnitude smaller than for a battery.
+The limitation of the capacitor as a storage device is energy density that is inferior to a battery, making effective capacitor discharge times orders of magnitude smaller than for a battery.
 
 Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible.
-These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, insect monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}  and digestible~\cite{nadeau_naturebio_2017} sensors.
+These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
 
@@
 \begin{table}
+	\centering
 	\begin{tabular}{|c|c|}
@@
 	\centering
-	\fbox{\texttt{text}}
-	%\includegraphics[width=0.25\columnwidth]{figures/code_demo_incosistency}
-	%	for (byte = 0; byte < nBytes; ++byte){\\
-	%	checkpoint();\\
-		%data = reflectData(message\[byte\]) $^$ (remainder >> (WIDTH - 8));\\
-		%remainder = crcTable[data] $^$ (remainder << 8);}
-	\caption{Code demonstrating WAR and volatile memory checkpointing.}
+	\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
+	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 10) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
+	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
 	\label{fig:code_demo_incosistency}
@@
 
-Fig.~\ref{fig:code_demo_incosistency} demonstrates the problem through example.
+Nevertheless, regular checkpointing of program state to a non-volatile memory still does not guarantee consistency.
+This phenomenon has been discussed extensively in e.g.~\cite{chain,alpaca}.
+For the completeness of this exposition we will provide our own example demonstrating this problem, please refer to Fig.~\ref{fig:code_demo_incosistency} and compare with~\cite[Sec 2.3]{alpaca},~\cite[Sec.
+2.1]{chain}.
+We take an example of a fast procedure to calculate cyclic redundancy check (CRC) from~\cite{hicks_mibench2_2016}.
+Therein a CRC must be calculated for $\texttt{nBytes}$-long message $\texttt{msg}$ with specified initial reminder $\texttt{Rmnd}$ and constant, $\texttt{Rmnd}$ variable type-dependent, $\texttt{Wdth}$.
+Despite regular, fixed position checkpoints (line 3 in Fig.~\ref{fig:crc_example}) of all variables which are declared as non-volatile (denoted as \texttt{NV}) variable $\texttt{data}$ is calculated incorrectly.
+This is because of a write after read (WAR) hazard of variable $\texttt{Rmnd}$.
+WAR hazard is caused by consecutive write and read in-between two consecutive checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is written \emph{before} (line 10 in Fig.~\ref{fig:crc_example}) $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example}), or more specifically---updated before a new execution loop.
 

### 2017-07-24 paper/paper.tex
WAR inconsistency section complete
relocate the figure tex code

@@
 %replace XXX with the submission number you are given from the ASPLOS submission site.
-\newcommand{\asplossubmissionnumber}{XXX}
+\newcommand{\asplossubmissionnumber}{39}
 
@@
 \usepackage{natbib}
+\usepackage{tikz}
+\usepackage{listings}
+
 %\usepackage{graphicx}
@@
 
-		In this paper we are introducing VIPOS a runtime library the implement the Virtualed Intermittent Execution Model.
-VIPOS adapts two methods for data protection.
+		In this paper we are introducing \sys a runtime library the implement the Virtualed Intermittent Execution Model.
+\sys adapts two methods for data protection.
 One relies on a virtual buffer and the other uses the Direct Memory Access (DMA).
-VIPOS is able to approach the optimal tasks devision during the runtime.
-VIPOS reduces an application execution time by XX and the energy by XX.
+\sys is able to approach the optimal tasks devision during the runtime.
+\sys reduces an application execution time by XX and the energy by XX.
 
@@
     \centering
-         \subfloat[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }
+         \subfloat[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }\\
          \subfloat[The energy cost of FRAM/SRAM \textbf{read} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
@@
 
-\section{VIPOS: Virtualized Intermittently Powered Runtime Environment}
-VIPOS is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs.
+\section{\sysfull}
+\sys is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs.
 It is able, on-demand, to merge tasks to reduce the number of commits to the non-volatile memory.
@@
 
-\subsection{VIPOS: The Virtualizing Engine}
+\subsection{\sys: The Virtualizing Engine}
 \subsubsection{Power Interrupt Immune Scheduler}
 It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures.
-VIPOS provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
+\sys provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
 (un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
@@
 		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{VIPOS Tasks\}} $  \Comment{$VT:$ Virtual Task}
+			\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
 			\State VTS : VT size
@@
 
-VIPOS is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
+\sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
 Fig.~\ref{fig:virtualTaskSize} show the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
@@
 \section{Results}
-We compare the performance of VIPOS against a state-of-the-art intermittent execution approach.
+We compare the performance of \sys against a state-of-the-art intermittent execution approach.
 For this purpose, we used two different applications: (i) A data decompression application, which utilizes Huffman decoding technique to decompress 100 byte of data.; (ii) A discrete Fourier transform (DFT) application that uses two different resolution (4 and 8 bites) to analyze a randomly generated signal.
@@
 	\centering
-	\includegraphics[width=0.8\columnwidth]{figures/ipos_chain}
-	\caption{Time to complation of VIPOS versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
+	\includegraphics[width=0.8\columnwidth]{figures/chain_vipe}
+	\caption{Time to complation of \sys versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
 	\label{fig:IPOSPerformance}
@@
 
-Fig.~\ref{fig:IPOSPerformance} shows that VIPOS requires a shorter execution time to execute an application under the MIEM.
-The size of the virtual task that VIPOS uses equals One real task, in other words, no visualization is applied.
+Fig.~\ref{fig:IPOSPerformance} shows that \sys requires a shorter execution time to execute an application under the MIEM.
+The size of the virtual task that \sys uses equals One real task, in other words, no visualization is applied.
 

### 2017-07-24 paper/texFigCode/crc_example_war.tex
relocate the figure tex code

\documentclass[border={20pt 5pt 25pt 10pt}]{standalone} % Left Bottom Right Top
%\documentclass{standalone}

%\usepackage{verbatim}
%\usepackage{parcolumns}
%\usepackage{fancyvrb}

\usepackage{xparse}
\usepackage{listings}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath} %for \text{} macro
\usepackage{wasysym} %for \lightning symbol
\usepackage{amssymb} %for \circlearrowright symbol

\lstdefinestyle{mystyle}{
	tabsize          = 3,     						% tab space width (for "for" loop for instance)
	numbers          = left,  						% display line numbers on the left
	numberblanklines = false,
	%framesep         = 0pt,   						% expand outward
	%framerule        = 0pt, 						% expand outward
	frameround		 = ffff,
	frame 			 = single,
	commentstyle     = \color{ForestGreen},      			% comment color
	keywordstyle     = \bfseries\color{blue},       % keyword color
	stringstyle      = \color{DarkRed},  	  		% string color
	backgroundcolor  = \color{white}, 				% backgroundcolor color
	showstringspaces = true,             	 		% mark spaces in strings
	otherkeywords	 = {NV, signal,inos_task,ios_vars, uint8_t},
	firstnumber=1,
	language = C,
	escapechar = !,
	basicstyle = \ttfamily,
	columns = flexible,
	mathescape = true, %activates $ sign
	morecomment = [l]{//}
}

\begin{document}

\begin{lstlisting}[style = mystyle]
$\text{\textcolor{red}{checkpoint()}}$; //Iteration 1: correct execution of the loop body
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
Rmnd = crcTbl[data] ^ (Rmnd << 8);

$\text{\textcolor{red}{checkpoint()}}$; //Iteration 2: break before reminder calculation
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
$\text{\textcolor{red}{\lightning\hspace{0.5cm}\lightning\hspace{0.5cm}\lightning}\hspace{0.5cm}}\textcolor{red}{\circlearrowright}$ //Power interrupt and restart

$\text{\textcolor{red}{checkpoint()}}$; //Iteration 3: wrong/stale reminder used for data
data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
$\ldots$
\end{lstlisting}

\end{document}

### 2017-07-24 paper/iposDataProtection.tex
change VIPOS to \sys

@@
-\subsection{VIPOS: Data/Memory Protection Methods}
+\subsection{\sys: Data/Memory Protection Methods}
 
@@
 
-VIPOS data consistency preservation methods are designed with goal of minimal interaction between the CPU and FRAM to reduce energy consumption (see Section~\ref{sec:prelResults}).
-These data/memory protection methods enables VIPOS to tolerate arbitrary number of power failures.
+\sys data consistency preservation methods are designed with goal of minimal interaction between the CPU and FRAM to reduce energy consumption (see Section~\ref{sec:prelResults}).
+These data/memory protection methods enables \sys to tolerate arbitrary number of power failures.
 
@@
 
-		We implement a reference implementation of VIPOS that uses virtualized Operational buffer to protect the data against power failure. 
+		We implement a reference implementation of \sys that uses virtualized Operational buffer to protect the data against power failure. 
 		\paragraph{Virtual Buffer}
@@
 However, the complexity of committing this buffer is of size $O(N)$, where $N$ equals the length of the buffer.
-Therefore, the size of the buffer can have dramatic effect on the performance of VIPOS---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious.
-On the other hand, if the size of the buffer is large VIPOS will experience a significant performance degradation.
+Therefore, the size of the buffer can have dramatic effect on the performance of \sys---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious.
+On the other hand, if the size of the buffer is large \sys will experience a significant performance degradation.
 However, by observing the type of the information that this buffer holds, namely memory addresses, we can, on average, reduce the complexity of committing this buffer by defining the \emph{effective size} of the buffer to be the size of the buffer up to the first cell that holds an invalid memory address.
@@
 		% 	\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
-		% 	\caption{VIPOS based on virtualized Operational Buffer.
+		% 	\caption{\sys based on virtualized Operational Buffer.
 It runs the decompression application with different buffer configurarions.
@@
 
-		% 	To analyze the performance of VIPOS versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
+		% 	To analyze the performance of \sys versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
 We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}.
-The results presented in Fig.~\ref{fig:virtOperationalBuf} show the VIPOS has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.
+The results presented in Fig.~\ref{fig:virtOperationalBuf} show the \sys has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.
 
@@
 Furthermore, we quantized the overhead of the read operations from the buffer, by eliminating this operation, and the bar labeled with "NR" shows that it cost  $\approx$10 ms.
-Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to VIPOS best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.
+Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to \sys best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.
 

### 2017-07-25 paper/abstract.tex

@@
 
-VIPOS enables code portability by means of virtualizing the task based model.
-If a code is written for a very small energy buffer, VIPOS is able to dynamically merge these tasks and constructs a bigger virtual task to improve the performance, reduce energy consumption and provides mean for code portability.
+\sys enables code portability by means of virtualizing the task based model.
+If a code is written for a very small energy buffer, \sys is able to dynamically merge these tasks and constructs a bigger virtual task to improve the performance, reduce energy consumption and provides mean for code portability.

### 2017-07-25 paper/background.tex

@@
 This phenomenon has been discussed extensively in e.g.~\cite{chain,alpaca}.
-For the completeness of this exposition we will provide our own example demonstrating this problem, please refer to Fig.~\ref{fig:code_demo_incosistency} and compare with~\cite[Sec 2.3]{alpaca},~\cite[Sec.
+For the completeness of this exposition we will provide our own example demonstrating this problem.
+For this please refer to Fig.~\ref{fig:code_demo_incosistency} and compare with~\cite[Sec 2.3]{alpaca},~\cite[Sec.
 2.1]{chain}.
@@
 This is because of a write after read (WAR) hazard of variable $\texttt{Rmnd}$.
-WAR hazard is caused by consecutive write and read in-between two consecutive checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is written \emph{before} (line 10 in Fig.~\ref{fig:crc_example}) $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example}), or more specifically---updated before a new execution loop.
+WAR hazard is caused by consecutive write and read in-between two consecutive checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war}) $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more specifically---updated before a new execution loop.
 
@@
 
-\begin{figure}
+There is increasing amount of work that deals with protecting memory consistency.
+Table~\ref{table:chechpoint_comparison} provides a summary of various costs associated with most relevant task-based/checkpointing solutions.
+This table generalizes discussion from~\cite[Sec.
+2.4]{alpaca}.
+
+\begin{table}
 	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Two problems with fixed-size task in intermittent execution.
-(a) task underestimation: task executed at device with capacitor size $X$, will not execute at device with capacitor $X\gg Y$; (b) task overestimation: with stable energy source task tracking causes unnecessary overhead.}
-	\label{fig:fixed_task_problem}
-\end{figure}
+	\begin{tabular}{|c|c|c|c|}
+		\hline
+		{~} & CP & SZ & NV \\
+		\hline\hline
+		\!\!Mementos~\cite{mementos}\!\! & \!\!$N_2\leq N_1$\!\! & \!\!R + S\!\! & \!\!R + S\!\! \\
+		\!\!DINO~\cite{dino}\!\! & $N_2$\!\! & \!\!R + W\!\! & \!\!R + W\!\! \\
+		\!\!Chain~\cite{chain}, Alpaca~\cite{alpaca}\!\! & \!\!$N_2$\!\! & P\!\! & \!\!2W + P\!\!\\
+		\!\!Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017}\!\! & $N_3\geq N_2\geq N_1$\!\! & \!\!R\!\! & R\!\!\\
+		\hline 
+	\end{tabular}
+	\caption{Cost of various checkpointing methods; Symbols---R: register, S: stack, W: WAR-dependent variables, P: program counter, $N_1, N_2, N_3$: relative number of checkpoints for each method, CP: no.
+checkpoints, SZ: checkpoint size, NV: copy to non-volatile memory size.}
+	\label{table:chechpoint_comparison}
+\end{table}
+
+The energy cost of a single checkpoint is $w\text{CP(SZ+NV)}$, where $w$ is the cost of a single word copy into a non-volatile memory.
+Decrease of CP decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
+On the other hand, this increases task/checkpoint re-execution time.
 
@@
 
-We can write the problem of maximizing all task sizes formally as
+%prior systems do fine grained task division (virtualization) {clank,aplaca}
+
+Task is a programmer (or compiler) defined code that is executed based on consistent set of inputs, i.e.
+memory fetched from a specified memory location, producing a consistent output stored in a specified (the same or different) memory location.
+Syntactically, we can treat a task as a continuous sub-section of a large program encapsulated by (i) task specific keywords defining the beginning and the end of a task and (ii) message passing between tasks connected with each other.
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=\columnwidth]{figures/taskification_example}
+	\caption{Task flow example executed by \sys.\todo{Draw an example task diagram}{Przemek}}
+	\label{fig:task_flow_example}
+\end{figure}
+
+Let us look into how task-based programming is performed relying on an example provided in Fig.~\ref{fig:task_flow_example}.
+Each task is uniquely identified by a keyword.
+A separate part of code specifies a state machine than governs how tasks are connected with each other.
+This way a programmer specifies to which other task a current task needs to pass the result of its computation (and it can be either a new task or the task itself).
+Tasks are executed atomically.
+That is, task will be executed only once and will be considered as completed only when a task reaches a final instruction line informing a scheduler that a program is ready to enter into a new task.
+
+Just like Chain~\cite{chain} and Alpaca~\cite{alpaca}, \sys does not support concurrent programming, multi-threading (multi-task sequence) and does not consider priorities and on-demand scheduling.
+These features are extremely important to consider but go beyond the scope of this work.
+A comparative description of \sys is given in Table~\ref{table:feature_comparison}.
+
+Important and a strong assumption about task-based programming for intermittently-powered systems is that an energy availability for the largest possible task execution must be guaranteed.
+Task is an atomic operation and  must preserve execution order and memory manipulations.
+Task that follows such an assumption is equivalent to task being executed on a continuously-powered non-interrupted system.
+Whenever a task cannot fit into a single energy reservoir, it must be divided into tasks that will be executed within a single discharge cycle.
+
+\begin{table}
+	\begin{tabular}{|c|c|c|c|}
+		\hline
+		{~} & Alpaca~\cite{alpaca} & Chain~\cite{chain} & This work \\
+		\hline\hline
+		Task connections & in-task & in-task & outside \\
+		Task re-use & N & N & Y\\
+		Memory model & globals & globals & globals\\
+		Concurrency & N & N & N \\
+		Task blocking & N & N & Y \\
+		Energy-adaptive & N & N & Y \\
+		\hline
+	\end{tabular}
+	\caption{Comparison of existing task-based execution models against \sys.}
+	\label{table:feature_comparison}
+\end{table}
+
+\todo{Elaborate on the ``key challenge'' paragraphs near the end of the intro}{Przemek}
+
+%%%Notes%%%
+
+%We can write the problem of maximizing all task sizes formally as
 %
-\begin{equation}
-\underset{i}{\arg\,\max} \sum_{i}m_i+t_i \text{~subject to~} \forall_i mi+t_i>e,
-\end{equation}
+%\begin{equation}
+%\underset{i}{\arg\,\max} \sum_{i}m_i+t_i \text{~subject to~} \forall_i mi+t_i>e,
+%\end{equation}
 %
-where $e$ is the required minimum energy to perform a task, $i$ is the total number of tasks, $m_i$ and $t_i$ are the cost of task traversal (memory copying time of task variables) and time to complete one task, respectively.
+%where $e$ is the required minimum energy to perform a task, $i$ is the total number of tasks, $m_i$ and $t_i$ are the cost of task traversal (memory copying time of task variables) and time to complete one task, respectively.
 
-Other forms of intermittent platforms will share the same problem.
+%Two problems with fixed-size task in intermittent execution.
+(a) task underestimation: task executed at device with capacitor size $X$, will not execute at device with capacitor $X\gg Y$; (b) task overestimation: with stable energy source task tracking causes unnecessary overhead.
+
+%\begin{figure}
+%\centering
+%\includegraphics[width=\columnwidth]{figures/taskification_example}
+%\caption{Program divided into tasks.}
+%\label{fig:taskification_example}
+%\end{figure}
+
+%Other forms of intermittent platforms will share the same problem.
 One of them considers actuation platforms, which will be powered directly from energy harvesting sources~\cite{}.
@@
 Therefore, although capacitor is large enough to perform computation alone, energy to power actuators takes precedence leaving not much space for computing.
-
-%prior systems do fine grained task division (virtualization) {clank,aplaca}
-
-%Q1: Does Program Counter change size with task (checkpoint) size in Chain? In other words: bigger Chain task=bigger Program Counter?
-%Q2: Does Register File (RF) includes Program Counter (PC)? In other words: is always PC>RF in terms of memory size?
-
-%N_{checkpoints} \times e^{2\times N_{shared}}
-
-\todo{Elaborate on the ``key challenge'' paragraphs near the end of the intro}{Przemek}
-\todo{Define the Task and Task based execution model}{Przemek}

### 2017-07-25 paper/texFigCode/crc_example.tex

\documentclass[border={20pt 5pt 35pt 10pt}]{standalone} % Left Bottom Right Top
%\documentclass{standalone}

%\usepackage{verbatim}
%\usepackage{parcolumns}
%\usepackage{fancyvrb}

\usepackage{xparse}
\usepackage{listings}
\usepackage[dvipsnames]{xcolor}
\usepackage{amsmath} %for \text{} macro
\usepackage{wasysym} %for \lightning symbol
\usepackage{amssymb} %for \circlearrowright symbol

\lstdefinestyle{mystyle}{
	tabsize          = 3,     						% tab space width (for "for" loop for instance)
	numbers          = left,  						% display line numbers on the left
	numberblanklines = false,
	%framesep         = 0pt,   						% expand outward
	%framerule        = 0pt, 						% expand outward
	frameround		 = ffff,
	frame 			 = single,
	commentstyle     = \color{ForestGreen},      			% comment color
	keywordstyle     = \bfseries\color{blue},       % keyword color
	stringstyle      = \color{DarkRed},  	  		% string color
	backgroundcolor  = \color{white}, 				% backgroundcolor color
	showstringspaces = true,             	 		% mark spaces in strings
	otherkeywords	 = {NV, signal,inos_task,ios_vars, uint8_t},
	%numbers = none,
	firstnumber=1,
	language = C,
	escapechar = !,
	basicstyle = \ttfamily,
	columns = flexible,
	mathescape = true, %activates $ sign
	morecomment = [l]{//}
}

\begin{document}

 \begin{lstlisting}[style = mystyle]
NV int byte, nBytes, Rmnd, Wdth, char msg[], crcTbl[];
for (byte = 0; byte < nBytes; ++byte){
	$\text{\textcolor{red}{checkpoint()}}$; //Inserted checkpoint routine
	data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
	Rmnd = crcTbl[data] ^ (Rmnd << 8);}
\end{lstlisting}

\end{document}

### 2017-07-25 paper/background.tex
did some revision in the background

@@
-Performing complex operations on battery-less energy harvesting embedded platforms is challenging from many points of view.
-Here we shall briefly provide a background on computation with energy harvesting embedded systems.
+\sys addresses the problem of intermittently executing task-based computations
+on energy-harvesting hardware, without being limited by inflexible, static task
+decomposition and without being limited to fully non-volatile main memory.  In
+this section, we provide background on energy harvesting systems and the
+intermittent software execution model.   We then discuss the key limitations of
+existing task-based programming and execution models for intermittent computing
+that \sys addresses.
 
-\subsection{Energy Harvesting: Background}
+\subsection{Energy Harvesting Systems}
 \label{sec:background_harvesting}
 
-Supplying power to tiny embedded computers using batteries alone is not sustainable.
-European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
-Since not all batteries are recycled, they constitute a potential environmental hazard.
-A potent solution is to replace batteries with smaller, but environment friendly {(super-)capacitors} (promising millions of charge/discharge cycles~\cite[Sec.
-I]{ongaro_pwre_2012}) and to supply them {\em directly from the ambient energy sources}---creating a truly self-sustainable computing device.
-The limitation of the capacitor as a storage device is energy density that is inferior to a battery, making effective capacitor discharge times orders of magnitude smaller than for a battery.
+Energy harvesting devices are systems that operate using energy extracted from
+their environment, from sources such as radio waves (RF)~\cite{wisp}, and solar
+energy.  By operating entirely using environmental energy, energy-harvesting
+devices need not use a tethered power supply or battery.  Instead, devices
+typically collect energy into a (super-)capacitor, operate when sufficient
+energy has accumulated, and upon depleting the stored energy, turn off and wait
+to collect more energy.
 
-Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible.
-These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
+Batteryless operation has a number of important advantages, making intermittent
+computing an important domain.  Supplying power to the ``next
+billion''~\cite{citationneeded} embedded computers using batteries is not
+sustainable.  The European Commission estimates that more than 160 kilotons of
+consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
+Batteries constitute a potential environmental hazard because not all batteries
+are recycled.  Batteries are fragile and limited in their number of
+charge/discharge cycles, requiring costly maintenance that is possibly only
+through physical device access, which may be difficult if devices are deployed
+in inaccessible places (e.g., in space~\cite{kicksat}). By contrast,
+super-capacitors are durable, promising millions of charge/discharge
+cycles~\cite[Sec. I]{ongaro_pwre_2012} The main limitation in moving to
+capacitor-based energy storage is that capacitor energy density --- and
+consequently operating discharge time --- is orders of magnitude less than a
+battery.
 
-Many platforms enable intermittent, battery-less computation.
-For instance, computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
-Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
-In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
-Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid).
-The reader is referred to example extensive surveys discussing energy harvesting for low-power embedded systems~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016}.
+\TODO{I think this can be compressed a bit}
+Given current technology development, battery-less systems are best suited for
+very long-term sensing and monitoring where access to recharge is either
+prohibitive or impossible. These include battery-less image capture and
+processing~\cite{naderiparizi_rfid_2015}, animal
+monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}
+and digestible~\cite{nadeau_naturebio_2017} sensors.
+% 
+Many platforms enable intermittent, battery-less computation. For instance,
+computation RFIDs---open-source TI MSP430-based~\cite{wolverine}
+WISP~\cite{wisp5} (with its variants such as
+WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
+NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
+such as~\cite{medusa_farsens_2017}. Other intermittently-powered platforms
+include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or
+battery-less phone~\cite{talla_imwut_2017}. In all of the above, the main
+source of energy harvested is the electromagnetic radiation in the radio
+frequency range (ambient transmitters such as high power TV
+transmitters~\cite{liu_sigcomm_2013} or dedicated RFID
+antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
+Naturally, other forms of energy harvesting sources exist, including
+temperature gradient, (micro-)motions, light/sun radiation, vibrations, and
+body fluid flow (blood, gastric acid).  Several recent surveys discussing
+energy harvesting, low-power, embedded systems and intermittent computing at a
+high
+level~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016,lucia_snapl_2017}.
 
@@
 
-\subsection{Memory Consistency}
+\subsection{Intermittent Computing}
 \label{sec:background_consistency}

### 2017-07-25 paper/background.tex
Still working on a background section

@@
 1]{mementos}.
-This causes each program running on such platform to restart.
-Solution to the problem is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between execution times.
+This causes each program running on such platform to restart and requires a fundamentally new model for program execution:  \emph{intermittent} (discontinuous) execution model.
+In the conventional (or \emph{continuous}) execution model a power failure is seen as an \emph{exception} that may reset the progress of a program to its start.
+Whereas, in an intermittent execution model a power failure is regarded as a temporary \emph{pause} to the execution consequently prolonging execution time.
+A conceptual implementation of intermittent execution model is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between execution times.
+
+\subsection{Assumption on the Type of Computing Hardware}
+\label{sec:background_hardware}
+
+Before proceeding further with the background we need to make certain assumptions about the underlying hardware.
+We assume that we operate on the microcontroller (more generally, computing hardware) equipped with a mix of volatile and non-volatile memory, such as FLASH, Ferroelectric RAM or Magnetoresistive RAM.
+\sys does not assume to rely on specific memory type (either volatile or non-volatile), in contrary to earlier works targeting non-volatile processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
+Actually, \sys will exploit the properties of both memory types to maximize performance metrics of the program being executed.
+We make \sys able to I/O operations, however we do not enable the scheduling of operations (which \emph{nota bene} forms a foundation for a simple runtime).
+All in all, our hardware model follows in assumptions the two most relevant intermittently-powered execution environments~\cite{alpaca,chain}.
 
@@
 2]{mementos})is a periodic checkpointing of a program state~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
-The periodicity, and a position of a checkpoint, is the core of the contribution behind each of the proposed checkpoining solution.
+The periodicity, and a position of a checkpoint, as well as the way to combat the \emph{program state consistency} phenomenon (discussed extensively in e.g.~\cite{chain,alpaca}), is the core of the contribution behind each of the proposed checkpoining solution.
+For the completeness of this exposition we will provide our own example demonstrating this phenomenon, viz.
+please refer to Fig.~\ref{fig:code_demo_incosistency} and compare with~\cite[Sec 2.3]{alpaca},~\cite[Sec.
+2.1]{chain}.
 
@@
 
-Nevertheless, regular checkpointing of program state to a non-volatile memory still does not guarantee consistency.
-This phenomenon has been discussed extensively in e.g.~\cite{chain,alpaca}.
-For the completeness of this exposition we will provide our own example demonstrating this problem.
-For this please refer to Fig.~\ref{fig:code_demo_incosistency} and compare with~\cite[Sec 2.3]{alpaca},~\cite[Sec.
-2.1]{chain}.
-We take an example of a fast procedure to calculate cyclic redundancy check (CRC) from~\cite{hicks_mibench2_2016}.
+Let us take an example of a fast procedure to calculate cyclic redundancy check (CRC) from~\cite{hicks_mibench2_2016}.
 Therein a CRC must be calculated for $\texttt{nBytes}$-long message $\texttt{msg}$ with specified initial reminder $\texttt{Rmnd}$ and constant, $\texttt{Rmnd}$ variable type-dependent, $\texttt{Wdth}$.
@@
 
-\subsection{Assumption on the Type of Computing Hardware}
-\label{sec:background_hardware}
-
-We assume that we operate on the microcontroller (more generally, computing hardware) equipped with a mix of volatile and non-volatile memory, such as FLASH, Ferroelectric RAM or Magnetoresistive RAM.
-\sys does not assume to rely on specific memory type (either volatile or non-volatile), in contrary to earlier works targeting non-volatile processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
-Actually, \sys will exploit the properties of both memory types to maximize performance metrics of the program being executed.
-We make \sys able to I/O operations, however we do not enable the scheduling of operations (which \emph{nota bene} forms a foundation for a simple runtime).
-All in all, our hardware model follows in assumptions the two most relevant intermittently-powered execution environments~\cite{alpaca,chain}.
-
 \subsection{(Task-based) Intermittent Computing: Background}
 
-There is increasing amount of work that deals with protecting memory consistency.
+Generally we can classify the intermittent execution model into:
+
+\subsubsection{Sequential Execution Model}: Under this model a program is seen as one big idempotent operational region that has one common context.
+Generally, the progress of the program is saved and updated by means of checkpointing, refer to Section~\ref{sec:background_consistency}.
+Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos,mottola2017harvos,hibernus}.
+The benefit of this model is that it does not require code modification by a programmer.
+However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{}; and (iii) it restricts the intermittently-powered devices to run only a single application.
+
+\subsubsection{Modular Intermittent Execution Model}: At the heart of this model is the concept of an \emph{idempotent task}.
+The idempotent task is a C language function that does not have arguments and does not return a value.
+This task uses a well defined interface to interact with the non-volatile memory.
+Therefore, it tolerates arbitrary number of power interrupts.
+This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an intermittently-powered device by interleaving their tasks~\cite{}.
+However, it obviously requires code modification.
+For example, if an algorithm is written according to the continuous execution model it has to be split, by a programmer, into small tasks to run under the modular intermittent execution model.
+
+Both modular and intermittent models are present in the literature.
 Table~\ref{table:chechpoint_comparison} provides a summary of various costs associated with most relevant task-based/checkpointing solutions.
@@
 
-The energy cost of a single checkpoint is $w\text{CP(SZ+NV)}$, where $w$ is the cost of a single word copy into a non-volatile memory.
+\textbf{Cost of a checkpoint:} Whatever the approach, the energy cost of a single checkpoint is $w\text{CP(SZ+NV)}$, where $w$ is the cost of a single word copy into a non-volatile memory.
 Decrease of CP decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
@@
 
-Moreover, each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}, which causes different charge/discharge periods.
-This result in different execution times.
-In the worst case the size of the capacitor might be too small to execute one task written by the programmer.
+\textbf{Cost of a fixed task size:} A modular intermittent execution model completely depends on a programmer's estimation which mostly results in a sub-optimal code division.
+Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.
+Each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}, which causes different charge/discharge periods.
+This result in different execution times and in the worst case the size of the capacitor might be too small to execute one task written by the programmer.
 

### 2017-07-25 paper/paper.tex
title suggestion
Still working on a background section

@@
 \newcommand{\sysfull}{Virtualized Intermittently-Powered Executor\xspace}
+\newcommand{\sysfullalt}{Virtualized Intermittently-Powered Execution Runtime\xspace}
 
@@
 
-\title{VIPE: Virtualized Intermittently-Powered Executor}
+\title{\sysfull \\ OR \\ \sysfullalt}
 
@@
 
+\section{Compiler Support for Automatic Task Generation}
+\label{sec:compiler}
+
+\input{compiler}
+
 \section{Memory Virtualization}
 \label{sec:memory_virtulaization}
+
 \input{paging}
+
 \begin{figure}
@@
 \section{Old Introduction}
-	Intermittently powered devices (IPDs) are battery-less devices that utilize the ambient energy to sense, compute and, communicate.
-For example, Wireless Identification and Sensing platform WISP \cite{wisp} uses the RF signal power to drive its computation and communication.
-Because of the reliance of TPDs on intermittent power supply, the harvested energy when it is available, their programs are susceptible to a very frequent power interrupts, in the order of tens of millisconds~\cite{}.
-Therefore, these devices require a different software execution model that complies with the nature of a discontinuous power supply.
-
-	The intermittent (discontinuous) execution model defines a program execution as cumulative discrete process.
-The main difference between the intermittent and the conventional (or continuous) execution models is that, a power failure is seen by the continuous model as an \emph{exception} that may reset the progress of a program to its beginning.
-Whereas, in intermittent execution a power failure is regarded as a temporary \emph{pause} to the execution that may result in some progress degradation.
-Generally we can classify the intermittent execution model into:
-	\begin{itemize}
-		\item \emph{Sequential Execution Model}:
-			Under this model a program is seen as one big idempotent operational region that has one common context.
-Generally, The progress of the program is saved and updated by means of checkpointing---where all the program context (e.g.
-CPU registers, the stack and the global variables) is saved to a non-volatile memory.
-Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos, harvOS, hibernus}.
-The benefit of this model is that it does not require code modification by a programmer.
-However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{xxxx}; and (iii) it restricts the IPDs to run only a single application~\cite{inos}.
-		\item \emph{Modular Intermittent Execution Model}:
-			At the heart of this model is the concept of an idempotent task.
-The idempotent task is C function that does not have arguments and does not return a value.
-This task uses a well defined interface to interact with the non-volatile memory.
-Therefore, it tolerates arbitrary number of power interrupts.
-This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an IPD by interleaving their tasks~\cite{inos}.
-However, it obviously requires code modification---for example, if an algorithm is written according to the continuous execution model it has to be splitted, by a programmer, into small tasks to run under the modular intermittent execution model.
-	\end{itemize}
-
-Despite that the superiority of the Modular Intermittent Execution model, it is still a static approach that completely depends on a programmer's estimation which is mostly result in a sub-optimal code devision.
-Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.
 
@@
 \subsection{Why not Adapting Current Approaches}
+
 Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes. 

### 2017-07-26 paper/background.tex

@@
 
-\TODO{I think this can be compressed a bit}
+\todo{Compress text}{Brandon}
 Given current technology development, battery-less systems are best suited for
@@
 prohibitive or impossible. These include battery-less image capture and
-processing~\cite{naderiparizi_rfid_2015}, animal
-monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}
+processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}
 and digestible~\cite{nadeau_naturebio_2017} sensors.
-% 
+
 Many platforms enable intermittent, battery-less computation. For instance,
 computation RFIDs---open-source TI MSP430-based~\cite{wolverine}
-WISP~\cite{wisp5} (with its variants such as
-WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
+WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
 NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
@@
 antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
-Naturally, other forms of energy harvesting sources exist, including
-temperature gradient, (micro-)motions, light/sun radiation, vibrations, and
-body fluid flow (blood, gastric acid).  Several recent surveys discussing
-energy harvesting, low-power, embedded systems and intermittent computing at a
-high
-level~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016,lucia_snapl_2017}.
+Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid).
+Several recent surveys discussing energy harvesting, low-power, embedded systems and intermittent computing at a high level~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016,lucia_snapl_2017}.
 
@@
 Naturally, content of a volatile memory (i.e.
-run-time stack, global and local variables) of the microcontroller and program registers are erased after each power failure.
+run-time stack, global and local variables) and program registers of the microcontroller are erased after each power failure.
 Only what resides in a non-volatile memory, e.g.
 FRAM, persists after system reboot.
-A solution to sustain the progress of the program from the last power failure, instead from the beginning (the \emph{Sisyphus} effect~\cite[Sec.
-2]{mementos})is a periodic checkpointing of a program state~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
+A solution to sustain the progress of the program from the last power failure, instead from the beginning (also denoted as the \emph{Sisyphus} effect~\cite[Sec.
+2]{mementos}) is a periodic checkpointing of a program state~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
 The periodicity, and a position of a checkpoint, as well as the way to combat the \emph{program state consistency} phenomenon (discussed extensively in e.g.~\cite{chain,alpaca}), is the core of the contribution behind each of the proposed checkpoining solution.
@@
 
-Generally we can classify the intermittent execution model into:
+We can classify the intermittent execution models into two classes: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
+)
+\subsubsection{Baseline---Sequential Execution Model}
 
-\subsubsection{Sequential Execution Model}: Under this model a program is seen as one big idempotent operational region that has one common context.
-Generally, the progress of the program is saved and updated by means of checkpointing, refer to Section~\ref{sec:background_consistency}.
+Under this model a program for intermittently-powered device is seen as a one idempotent operational region with one common context.
+Generally, the progress of the program is saved and updated by means of checkpointing, refer again to Section~\ref{sec:background_consistency}.
 Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos,mottola2017harvos,hibernus}.
 The benefit of this model is that it does not require code modification by a programmer.
-However, it has a number of drawbacks: (i) It suffers from significant overhead \cite{chain}; (ii) the programmer should not access the non-volatile memory to guarantee the consistency of the memory~\cite{}; and (iii) it restricts the intermittently-powered devices to run only a single application.
+However it suffers from significant overhead (refer to experiment result in~\cite[Fig.
+3]{chain}, where it is concluded that each checkpoint causes overhead).
 
-\subsubsection{Modular Intermittent Execution Model}: At the heart of this model is the concept of an \emph{idempotent task}.
+\subsubsection{Alternative---Modular Execution Model}
+
+At the heart of this model is the concept of an \emph{idempotent task}.
 The idempotent task is a C language function that does not have arguments and does not return a value.
@@
 However, it obviously requires code modification.
-For example, if an algorithm is written according to the continuous execution model it has to be split, by a programmer, into small tasks to run under the modular intermittent execution model.
+For example, if an algorithm is written according to the continuous execution model it has to be split by a programmer, into small tasks to run under the modular intermittent execution model.
 
-Both modular and intermittent models are present in the literature.
-Table~\ref{table:chechpoint_comparison} provides a summary of various costs associated with most relevant task-based/checkpointing solutions.
-This table generalizes discussion from~\cite[Sec.
-2.4]{alpaca}.
+Despite the benefit of the modular model certain limitations are still present that we must overcome and which inspired us to design \sys. Let us look at these problems in detail.
 
@@
 
-\textbf{Cost of a checkpoint:} Whatever the approach, the energy cost of a single checkpoint is $w\text{CP(SZ+NV)}$, where $w$ is the cost of a single word copy into a non-volatile memory.
-Decrease of CP decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
+\paragraph{Problem 1---Idempotent task is also costly}
+
+The most important modular and intermittent models are compared in Table~\ref{table:chechpoint_comparison}\footnote{This table generalizes discussion from~\cite[Sec.
+2.4]{alpaca}.}.
+Although we concluded that sequential execution model introduces a significant overhead, a modular approach also introduces certain execution penalty.
+Whatever the approach, the energy cost of a single checkpoint/idempotent task is $w\text{CP(SZ+NV)}$, where $w$ is the cost of a single word copy into a non-volatile memory.
+
+\begin{figure}
+	\centering
+	\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
+	\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
+	\caption{Energy cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
+	\label{fig:framEnergy}
+\end{figure}
+
+We have executed an experiment to measure exactly cost of read/write operations to volatile and non-volatile memory.
+The result is given in Figure~\ref{fig:framEnergy}.
+Four applications were developed to measure the energy consumption of \emph{accessing} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{msp430datasheet}.
+The interference free debugger (EDB)~\cite{edb} is used as the energy measurement tool.
+The EDB probes the energy buffer before and after accessing the FRAM/SRAM 1600 times.
+This large number of FRAM access is used to increase the reliability of the results generated by the used measurement tool (e.g.
+reducing the effect of the quantization error).
+The energy buffer is charged, using the EDB, to $\approx$ 2.45\,V \textbf{only} at the beginning of the execution and of the programs.
+The write operations were performed by writing literal values, random numbers, to the memory.
+However, a read operation must be followed by a write operation.
+Therefore, we chose to write a read value to SRAM.
+Based on our measurements we conclude that a FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We would like to mention that TI has already mentioned in this document~\cite{} that FRAM write access is more energy expensive than SRAM write access.
+However, TI does not quantify the energy overhead.}.
+Therefore, reducing FRAM access is desirable when developing energy limited software solution---\emph{which is in sharp contrast to the main goal of all checkpoining solutions}.
+
+The above result reassures the intuitive observation that a decrease of CP (Table~\ref{table:chechpoint_comparison}) decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
 On the other hand, this increases task/checkpoint re-execution time.
+The ideal solution is a runtime that \emph{adaptively} changes task size during runtime, which brings us to a second problem.
 
-\textbf{Cost of a fixed task size:} A modular intermittent execution model completely depends on a programmer's estimation which mostly results in a sub-optimal code division.
+\paragraph{Problem 2---Modular execution models are not adaptive}
+
+A modular intermittent execution model completely depends on a programmer's estimation which mostly results in a sub-optimal code division.
 Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.
@@
 This result in different execution times and in the worst case the size of the capacitor might be too small to execute one task written by the programmer.
+This calls for the ability of the modular execution runtimes to \textbf{merge} multiple small tasks into a larger one or \textbf{split} a large task into a set of small ones on the fly (post-compilation time).
+Unfortunately the following features prevent us to perform this task using existing runtimes~\cite{chain,alpaca}.
 
-%prior systems do fine grained task division (virtualization) {clank,aplaca}
+\begin{figure}
+	\centering
+	%\includegraphics[width=\columnwidth]{figures/Dynamic_chain_sequence_example}
+	\caption{Dynamic task merging problem.\todo{Draw merging diagram}{Przemek/Amjad}}
+	\label{fig:DynamicChainSeq}
+\end{figure}
+
+\begin{itemize}
+	\item \textbf{Lack of multi-versioning system} All known modular execution models define its own input for each task that is not shared with other tasks.
+This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
+
+	\item \textbf{Irreversible tasks transitions} According to both Chain~\cite{chain} and Alpaca~\cite{alpaca} programming models, each task calls the next one.
+The transition is done through functions calls and manually clearing the stack, to prevent the stack overflow problem.
+This approach makes all the task transitions firm, preventing task merging.
+
+	\item \textbf{`All channels' requirement} Fig.~\ref{fig:DynamicChainSeq} shows that if the task control flow is a circular with no branches then task merging is possible.
+Moreover, all the intra-merged-tasks channel can be committed to SRAM instead of FRAM to save energy.
+However, this execution path is only a special one.
+If the general case (see Fig.~\ref{fig:DynamicChainRan}) is considered then leaving out the intra-merged-task channels might result in data inconsistency.
+Than is, if an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready, which result in an incorrect execution.
+
+	\item \textbf{Dependency on Privatization} Alpaca~\cite{alpaca} extended Chain~\cite{chain} with a concept of privatization, i.e.
+it protects task-shared variables from WAR dependency problem \emph{within} the scope of a task.
+Therefore, if tasks are on-demand merged the scope of these dependencies are changed and the static analysis to privatize these variables is not valid anymore and data consistency can not be preserved.
+\end{itemize}
+
+\subsection{Structuring a Task-based Intermittent Computing Program}
 

### 2017-07-26 paper/paper.tex

@@
 \newcommand{\sys}{VIPE\xspace}
-\newcommand{\sysfull}{Virtualized Intermittently-Powered Executor\xspace}
+\newcommand{\sysfull}{Virtualized Intermittently-Powered Execution\xspace}
 \newcommand{\sysfullalt}{Virtualized Intermittently-Powered Execution Runtime\xspace}
@@
 Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes.
-%
+
 %	\begin{figure}[t]
@@
 
-\begin{itemize}
-	\item \emph{Multi-versioning system:} Chain defines for each task its own input that is not shared with other tasks.
-This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
-	\item \emph{Direct FRAM system:} Chain tasks directly access FRAM which is more energy expensive Than SRAM.
-As result of being multi-versioning system that directly interacts with FRAM, Chain has a significant energy consumption overhead.
-	\item \emph{Irreversible tasks transitions:} According to the Chain programming model, each task calls the next one.
-The transition is done through functions calls and manually clearing the stack, to prevent the stack overflow problem.
-This approach make all the transitions firm ones which prevent task merging.
-	\item \emph{All channels are required:} Fig.~\ref{fig:DynamicChainSeq} shows that if the task control flow is a circular with no branches then task merging is possible.
-Moreover, all the intra-merged-tasks channel can be committed to SRAM instead of FRAM to save energy.
-However, this execution path is only a special.
-If the general case (see Fig.~\ref{fig:DynamicChainRan}) is considered then leaving out the intra-merged-task channels might result in data inconsistency---If an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready which result in an incorrect execution.
-
-\end{itemize}
-
 \noindent\textbf{Alpaca: Intermittent Execution without Checkpoints} \\
 Some of the main problems in modifying Alpaca to support dynamic task's size are:
-
-\begin{itemize}
-	\item \emph{Privatization:} The core concept of Alpaca is privatization.
-Privatization depends completely on detecting the variables that have a Write-after-read dependency within the scope of a task.
-Therefore, if tasks are on-demand merged the scope of these dependencies are changed and the static analysis to privatize these variables is not valid anymore and data consistency can not be preserved.
-
-	\item \emph{Irreversible tasks transitions:} Similar to Chain problem.
-
-\end{itemize}
 
@@
 	\end{enumerate}
-
- \section{Preliminary Results}
- \label{sec:prelResults}
-
-\begin{figure}[t]
-    \centering
-         \subfloat[The energy cost of FRAM/SRAM \textbf{writes} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_write.eps} }\\
-         \subfloat[The energy cost of FRAM/SRAM \textbf{read} operations.]{\includegraphics[width=0.48\columnwidth]{figures/fram_read.eps}}
-    \caption{The energy cost of accessing volatile/non-volatile memory}
-    \label{fig:framEnergy}
-\end{figure}
-
-Since the Intermittent execution models rely on the non-volatile memory to enable long-running operation, it is important to characterize the energy consumption of accessing this type of memory as compare to the volatile one.
-
-Four applications are developed to measure the energy consumption of \emph{Accessing} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{xxx}.
-The interference free debugger (EDB)~\cite{xxx} is used as the energy measurement tool.
-The EDB probes the energy buffer before and after accessing the FRAM/SRAM 1600 times, This large number of FRAM access is used to increase the reliability of the results generated by the used measurement tool (e.g.
-reducing the effect of the quantization error).
-The energy buffer is charged, using the EDB, to $\approx$ 2.45 volts \textbf{only} at the beginning of the execution and of the programs.
-The write operations were performed by writing literal values, random numbers, to the memory.
-However, a read operation must be followed by a write operation, therefore, we chose to write a read value to SRAM.
-Fig.~\ref{fig:framEnergy} shows that an FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We would like to mentioned that TI has already mentioned in this document~\cite{xxxx} that FRAM write access is more energy expensive than SRAM write access.
-However, TI does not quantify the energy overhead.}.
-Therefore, reducing FRAM access is desirable when developing  energy limited software solution
 

### 2017-07-27 paper/evaluation.tex
Slowly making a paper one thing - moved old written sections into main part/divided paper into separate .tex files/small fixes here and there

We compare the performance of \sys against the state-of-the-art intermittent execution approach---Chain~\cite{chain}.

The applications that were used in the evaluation are given in Table~\ref{table:benchmark_list}.

\begin{table}
	\begin{tabular}{|c|c|c|}
		\hline
		Name & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
		\hline\hline
		Temperature sensing & 388 & 721 \\ %53\%
		Cuckoo & 483 & 762 \\ %63\%
		RSA & 887 & 1233 \\ %71\%
		DFT & --- & --- \\ %Two resolutions: 4 and 8 Bytes
		Huffman & --- & --- \\ %Data decompression size: 100 bytes
		Selection sort & --- & --- \\
		\hline
	\end{tabular}
\caption{List of benchmarks used in \sys evaluation.}
\label{table:benchmark_list}
\end{table}

\subsection{Continuous Power Supply}
\label{sec:results_continuous_power}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/chain_vipe}
	\caption{Time to complete for a set of testing applications (refer to Table~\cite{table:benchmark_list}) using \sys versus Chain.}
	\label{fig:IPOSPerformance}
\end{figure}

Fig.~\ref{fig:IPOSPerformance} shows that \sys requires a shorter execution time to execute an application under the modular execution model. The size of the virtual task that \sys uses is one real task, i.e. no task coalescing is applied.

%Set of experiments

%- Compiler result: taskifying completion time, correctness, 
%- Task coalescing experiment: with WISP at 3 distances and with fixed power, for all apps, compared against chain, show amount of task merged, time to completion/inter-operation time, overhead (in terms of what?), energy consumed
%- Battery-less microphone: with \sys and without \sys, compared against Chain, 
%- Paging/memory experiment: ?

\subsection{Intermittent Power Supply}
\label{sec:intermittent_continuous_power}

### 2017-07-27 paper/relatedwork.tex
Slowly making a paper one thing - moved old written sections into main part/divided paper into separate .tex files/small fixes here and there

\section{Intermittenly-Powered Devices}

\section{Intermittent Execution Environments}

### 2017-07-27 paper/coalescing.tex
New files added

--- 2017-07-26 paper/paper.tex
@@
-%\documentclass[sigconf,anonymous,review]{acmart}
-\documentclass[pageno]{jpaper}
+\subsubsection{Task Merging Algorithms}
 
-%replace XXX with the submission number you are given from the ASPLOS submission site.
-\newcommand{\asplossubmissionnumber}{39}
+\subsubsection{Power Interrupt Immune Scheduler}
 
-\usepackage[normalem]{ulem}
-\usepackage{amssymb}
-\usepackage{amsmath}
-\usepackage{algorithm} % environment for the algorithm code (like figure and table)
-\usepackage{algorithmicx} % write the pseudo code
-\usepackage[noend]{algpseudocode} % layout of the pseudo code
-\usepackage{tikz} % needed for: tikz pics
-\usepackage{xcolor}
-\usepackage{xspace}
-\usepackage{natbib}
-\usepackage{tikz}
-\usepackage{listings}
-
-%\usepackage{graphicx}
-%\usepackage{subfigure}
-%\usepackage[inline]{enumitem}
-
-\newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
-\newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}
-\newcommand{\sys}{VIPE\xspace}
-\newcommand{\sysfull}{Virtualized Intermittently-Powered Execution\xspace}
-\newcommand{\sysfullalt}{Virtualized Intermittently-Powered Execution Runtime\xspace}
-
-%\setcopyright{rightsretained}
-
-%\acmDOI{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
-
-%\acmISBN{978-1-xxxx-xxxx-x/17/11}
-
-%\acmConference[ASPLOS]{The 23rd ACM International Conference on Architectural Support for Programming Languages and Operating Systems}{March 24--28, 2017}{Williamsburg, VA, USA}
-
-%\acmYear{2017}
-
-%\copyrightyear{2017}
-
-%\acmPrice{15.00}
-
-\begin{document}
-
-\title{\sysfull \\ OR \\ \sysfullalt}
-
-%Note: order of authors placed randomly
-
-%\author{Kiwan Maeng}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Carnegie Mellon University}
-%	\streetaddress{4720 Forbes Avenue}
-%\city{Pittsburgh} 
-%	\state{PA} 
-%	\postcode{15213}
-%}
-%\email{kmaeng@andrew.cmu.edu}
-
-%\author{Alexei Colin}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Carnegie Mellon University}
-%	\streetaddress{4720 Forbes Avenue}
-%	\city{Pittsburgh} 
-%	\state{PA} 
-%	\postcode{15213}
-%}
-%\email{acolin@andrew.cmu.edu}
-
-%\author{Brandon Lucia}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Carnegie Mellon University}
-%	\streetaddress{4720 Forbes Avenue}
-%	\city{Pittsburgh} 
-%	\state{PA} 
-%	\postcode{15213}
-%}
-%\email{blucia@andrew.cmu.edu}
-
-%\author{Amjad Yousef Majid}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Delft University of Technology}
-%	\streetaddress{Mekelweg 4}
-%	\city{Delft, The Netherlands} 
-%	\state{Zuid Holland} 
-%	\postcode{2628\,CD}
-%}
-%\email{a.y.majid@tudelft.nl}
-
-%\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Delft University of Technology}
-%	\streetaddress{Mekelweg 4}
-%	\city{Delft, The Netherlands} 
-%	\state{Zuid Holland} 
-%	\postcode{2628\,CD}
-%}
-%\email{k.s.yildirim@tudelft.nl}
-
-%\author{Przemys{\l}aw Pawe{\l}czak}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Delft University of Technology}
-%	\streetaddress{Mekelweg 4}
-%	\city{Delft, The Netherlands} 
-%	\state{Zuid Holland} 
-%	\postcode{2628\,CD}
-%}
-%\email{p.pawelczak@tudelft.nl}
-
-%\renewcommand{\shortauthors}{A. Y. Majid et al.}
-
-\date{}
-\maketitle
-
-\thispagestyle{empty}
-
-\begin{abstract}
-\input{abstract}
-\end{abstract}
-
-% The code below should be generated by the tool at 
-%http://dl.acm.org/ccs.cfm Please copy and paste the code instead of the 
-%example below.
-
-%\begin{CCSXML}
-%	<ccs2012>
-%	<concept>
-%	<concept_id>10010520.10010553.10010562.10010564</concept_id>
-%	<concept_desc>Computer systems organization~Embedded 
-%	software</concept_desc>
-%	<concept_significance>500</concept_significance>
-%	</concept>
-%	<concept>
-%	<concept_id>10010520.10010575.10010578</concept_id>
-%	<concept_desc>Computer systems organization~Availability</concept_desc>
-%	<concept_significance>300</concept_significance>
-%	</concept>
-%	<concept>
-%	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
-%	<concept_desc>Software and its engineering~Scheduling</concept_desc>
-%	<concept_significance>300</concept_significance>
-%	</concept>
-%	<concept>
-%	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
-%	<concept_desc>Software and its engineering~Communications 
-%	management</concept_desc>
-%	<concept_significance>300</concept_significance>
-%	</concept>
-%	<concept>
-%	<concept_id>10011007.10010940.10010971.10010564</concept_id>
-%	<concept_desc>Software and its engineering~Embedded 
-%	software</concept_desc>
-%	<concept_significance>300</concept_significance>
-%	</concept>
-%	</ccs2012>
-%\end{CCSXML}
-
-%\ccsdesc[500]{Computer systems organization~Embedded software}
-%\ccsdesc[300]{Computer systems organization~Availability}
-%\ccsdesc[300]{Software and its engineering~Scheduling}
-%\ccsdesc[300]{Software and its engineering~Communications management}
-%\ccsdesc[300]{Software and its engineering~Embedded software}
-
-%\keywords{Energy harvesting, transient operation, operating system}
-
-\section{Introduction}
-\label{sec:intro}
-
-%\TODO{VIPER: Virtualised Intermittent Programming model, Execution model, and Runtime environment}
-%\TODO{PREVIEW: Programming and Runtime Environment with Virtualised Intermittent Execution Windows}
-
-\input{intro}
-
-\section{Intermittent Computation: Background}
-\label{sec:background}
-
-\input{background}
-
-\section{\sysfull}
-\label{sec:overeall_system}
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{\sys top-level description.}
-	\label{fig:}
-\end{figure}
-
-Describe the overall system, the programming model, and the semantics/execution model. This should include a discussion of memory virtualization and task coalescing.
-
-\section{Compiler Support for Automatic Task Generation}
-\label{sec:compiler}
-
-\input{compiler}
-
-\section{Memory Virtualization}
-\label{sec:memory_virtulaization}
-
-\input{paging}
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Memory virtualization architecture.}
-	\label{fig:}
-\end{figure}
-
-\section{Task Coalescing}
-\label{sec:memory_virtulaization}
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Task coalescing architecture.}
-	\label{fig:}
-\end{figure}
-
-\section{Discussion}
-\label{sec:discussion}
-
-\section{Methodology and Evaluation}
-\label{sec:methodology_evaluation}
-
-\section{Related Work}
-\label{sec:related_work}
-
-\section{Conclusions}
-\label{sec:conclusions}
-
-This paper presented \sys: \sysfull.
-
-%%%%%Old part of the paper - will be successfully moved section by section to a final one%%%%%
-
-\newpage
-
-\section{Old Introduction}
-
-% Energy source
-
-% The intermittent execution is cumulative discrete process.
-The intermittent execution model is only able to execute a few number of instructions, as compared to the conventional (continuous) model, before its progress is terminated.
-Therefore, the intermittent execution model adapts an execution progress state saving mechanism.
-This mechanism is realized either by means of a checkpoint~\cite{}, where all the context of the program is saved into non-volatile memory.
-Or
-
-% This progress state saving mechanism is normally injected into a program either by a compiler~\cite{}.
-A compiler injects trigger pointers to checkpoint (save) the context of the program into the non-volatile memory.
-or it is added by a programmer~\cite{chain}.
-[compare these two approaches]...
-
-% We define virtualization within the context of intermittent execution as utilizing the volatile memory instead of non-volatile when the intermittent execution model attempts to access the non-volatile memory.
-
-\subsection{Why not Adapting Current Approaches}
-
-Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes.
-
-%	\begin{figure}[t]
-%	    \centering
-%	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
-%	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
-%	    \caption{Task flow control of Chain}
-%	    \label{fig:DynamicChain}
-%	\end{figure}
-
-\noindent\textbf{Chain: Tasks and Channels for Reliable Intermittent Programs } \\
-Some of the main problems in modifying Chain to support dynamic task's size are:
-
-\noindent\textbf{Alpaca: Intermittent Execution without Checkpoints} \\
-Some of the main problems in modifying Alpaca to support dynamic task's size are:
-
-\noindent\textbf{Virtualizing Modular Intermittent Execution Model}
-
-		An Intermittently executed program, as seen by the modular intermittent execution model, is a chain of tasks with a firm transition from one to another. 
-		Virtualizing this model means softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
-Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
-Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
-This can be done by having a temporary copy of a global variable in the volatile memory (similar to the catching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
-
-		In this paper we are introducing \sys a runtime library the implement the Virtualed Intermittent Execution Model.
-\sys adapts two methods for data protection.
-One relies on a virtual buffer and the other uses the Direct Memory Access (DMA).
-\sys is able to approach the optimal tasks devision during the runtime.
-\sys reduces an application execution time by XX and the energy by XX.
-
-\subsection{Contributions}
-	 \begin{enumerate}
-		 \item We have developed an algorithm that is able to determine the size of a virtual task based on the history of the execution of an application. 
-		 \item We enable efficient code portability by requiring that a real task to small and merging them during the execution.   
-	\end{enumerate}
-
-\section{\sysfull}
-\sys is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs.
-It is able, on-demand, to merge tasks to reduce the number of commits to the non-volatile memory.
-
-\input{iposDataProtection}
-
-\subsection{\sys: The Virtualizing Engine}
-\subsubsection{Power Interrupt Immune Scheduler}
 It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures.
@@
 
-\subsubsection{Task Merging Algorithms}
-
-\paragraph{Fixed Virtual Task Size} [better name needed]
+\paragraph{Fixed Virtual Task Size}
 
@@
 
+\begin{figure}
+	\centering
+	%\includegraphics[width=0.25\columnwidth]{figures/task_coalescing}
+	\caption{Task coalescing architecture.}
+	\label{fig:}
+\end{figure}
+
 \sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
@@
 As such, virtualization can add a layer of security to the computation.
-
-\section{Results}
-We compare the performance of \sys against a state-of-the-art intermittent execution approach.
-For this purpose, we used two different applications: (i) A data decompression application, which utilizes Huffman decoding technique to decompress 100 byte of data.; (ii) A discrete Fourier transform (DFT) application that uses two different resolution (4 and 8 bites) to analyze a randomly generated signal.
-
-\subsection{Continuous Power Supply}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{figures/chain_vipe}
-	\caption{Time to complation of \sys versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
-	\label{fig:IPOSPerformance}
-\end{figure}
-
-Fig.~\ref{fig:IPOSPerformance} shows that \sys requires a shorter execution time to execute an application under the MIEM.
-The size of the virtual task that \sys uses equals One real task, in other words, no visualization is applied.
-
-\subsection{Intermittent Power Supply}
-...
-
-\bibliographystyle{plain}
-\bibliography{bib}
-
-\end{document}

### 2017-07-27 paper/background.tex
Changes to text

@@
 
+\subsection{Main Idea: Virtualization of a Task}
+
+To combat above limitations we propose a new concept of \textbf{virtualization}: a modular execution model with softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
+Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
+Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
+This can be done by having a temporary copy of a global variable in the volatile memory (similar to the catching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
+
 \subsection{Structuring a Task-based Intermittent Computing Program}

### 2017-07-27 paper/compiler.tex
Still working on a background section
Changes to text

@@
-\section{Compiler Support for Automatic Task Generation}
-\label{sec:compiler}
-
-\sys additionally provides a compiler support for automatic task
-generation, eliminating programmer's burden of writing a task-based
-program from scratch. The compiler gets a plain-C code as an input
-and generates a binary that is rewritten in task-based programming
-style that can run on \sys without additional programming effort. 
-Automatic task generation compiler is often inefficient than human-written
-code for \sys because if unbounded loop or pointer is present, the
-compiler makes a conservative decision. \xxx{If the performance is critical,
-the programmer may want to manually write the task-based code by himself.
+\sys additionally provides a compiler support for automatic task generation, eliminating programmer's burden of writing a task-based program from scratch.
+The compiler gets a plain-C code as an input and generates a binary that is rewritten in task-based programming style that can run on \sys without additional programming effort.
+Automatic task generation compiler is often inefficient than human-written code for \sys because if unbounded loop or pointer is present, the compiler makes a conservative decision.
+\emph{If the performance is critical, the programmer may want to manually write the task-based code by himself.
 Otherwise, the programmer can rely on the compiler for simplicity.}
@@
 The compiler consists of two part: analysis pass and transform pass.
-Analysis pass analyze the code structure and decides where to put
-task boundaries. Transform pass converts the code into task-style
-program and pepper it with necessary \sys keywords so that it can run
-on \sys system. It is implemented using LLVM.
+Analysis pass analyze the code structure and decides where to put task boundaries.
+Transform pass converts the code into task-style program and pepper it with necessary \sys keywords so that it can run on \sys system.
+It is implemented using LLVM.
 
@@
 edge penalty as 1. When analyzing loops, penalty of edges within loops are multiplied by loop count.
-\xxx{TODO: change the penalty to be 1 or 1*loopcount}
+\todo{Change the penalty to be 1 or 1*loopcount}{Kiwan}
 

### 2017-07-27 paper/paper.tex
Changes to text

@@
 \usepackage{xspace}
-\usepackage{natbib}
 \usepackage{tikz}
 \usepackage{listings}
+\usepackage{flushend}
 
+%\usepackage{natbib}
 %\usepackage{graphicx}
@@
 \newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}
+
 \newcommand{\sys}{VIPE\xspace}
-\newcommand{\sysfull}{Virtualized Intermittently-Powered Execution\xspace}
-\newcommand{\sysfullalt}{Virtualized Intermittently-Powered Execution Runtime\xspace}
+
+%Title Options
+%\newcommand{\sysfull}{VIPER: Virtualised Intermittent Programming model, Execution model, and Runtime environment}
+%\newcommand{\sysfull}{VIPER: Virtualized Intermittently-Powered Execution Runtime}
+%\newcommand{\sysfull}{PREVIEW: Programming and Runtime Environment with Virtualised Intermittent Execution Windows}
+%\newcommand{\sysfull}{VIPE: Virtualized Intermittently-Powered Execution}
+\newcommand{\sysfull}{VIPER: Virtual Intermittently-Powered Execution on the Run}
 
@@
 
-\title{\sysfull \\ OR \\ \sysfullalt}
+\title{\sysfull}
 
@@
 \begin{abstract}
+
 \input{abstract}
+
 \end{abstract}
@@
 
-%\TODO{VIPER: Virtualised Intermittent Programming model, Execution model, and Runtime environment}
-%\TODO{PREVIEW: Programming and Runtime Environment with Virtualised Intermittent Execution Windows}
-
 \input{intro}
@@
 
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{\sys top-level description.}
-	\label{fig:}
-\end{figure}
-
-Describe the overall system, the programming model, and the semantics/execution model. This should include a discussion of memory virtualization and task coalescing.
+\input{systemdescription}
 
@@
 
+\input{iposdataprotection}
+
 \section{Memory Virtualization}
@@
 
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Memory virtualization architecture.}
-	\label{fig:}
-\end{figure}
+\section{Task Coalescing}
+\label{sec:task_coalescing}
 
-\section{Task Coalescing}
-\label{sec:memory_virtulaization}
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Task coalescing architecture.}
-	\label{fig:}
-\end{figure}
+\input{coalescing}
 
@@
 
+\input{discussion}
+
 \section{Methodology and Evaluation}
 \label{sec:methodology_evaluation}
+
+\input{evaluation}
 
@@
 
+\input{relatedwork}
+
 \section{Conclusions}
@@
 
-This paper presented \sys: \sysfull.
-
-%%%%%Old part of the paper - will be successfully moved section by section to a final one%%%%%
+\input{conclusions}
 
 \newpage
-
-\section{Old Introduction}
-
-% Energy source
 
@@
 
-% We define virtualization within the context of intermittent execution as utilizing the volatile memory instead of non-volatile when the intermittent execution model attempts to access the non-volatile memory.
-
-\subsection{Why not Adapting Current Approaches}
-
-Here we highlight the challenges of adapting the state-of-the-art proposed methods to enable the execution of dynamic task sizes.
-
-%	\begin{figure}[t]
-%	    \centering
-%	         \subfigure[Chain: Sequential task execution flow control.]{\includegraphics[width=0.35\columnwidth]{figures/dynamic-chain.pdf} \label{fig:DynamicChainSeq} } 
-%	         \subfigure[Chain: Random task execution flow control.]{\includegraphics[width=0.61\columnwidth]{figures/dynamic-chain-2.pdf} \label{fig:DynamicChainRan}}
-%	    \caption{Task flow control of Chain}
-%	    \label{fig:DynamicChain}
-%	\end{figure}
-
-\noindent\textbf{Chain: Tasks and Channels for Reliable Intermittent Programs } \\
-Some of the main problems in modifying Chain to support dynamic task's size are:
-
-\noindent\textbf{Alpaca: Intermittent Execution without Checkpoints} \\
-Some of the main problems in modifying Alpaca to support dynamic task's size are:
-
-\noindent\textbf{Virtualizing Modular Intermittent Execution Model}
-
-		An Intermittently executed program, as seen by the modular intermittent execution model, is a chain of tasks with a firm transition from one to another. 
-		Virtualizing this model means softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
-Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
-Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
-This can be done by having a temporary copy of a global variable in the volatile memory (similar to the catching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
-
-		In this paper we are introducing \sys a runtime library the implement the Virtualed Intermittent Execution Model.
-\sys adapts two methods for data protection.
-One relies on a virtual buffer and the other uses the Direct Memory Access (DMA).
-\sys is able to approach the optimal tasks devision during the runtime.
-\sys reduces an application execution time by XX and the energy by XX.
-
-\subsection{Contributions}
-	 \begin{enumerate}
-		 \item We have developed an algorithm that is able to determine the size of a virtual task based on the history of the execution of an application. 
-		 \item We enable efficient code portability by requiring that a real task to small and merging them during the execution.   
-	\end{enumerate}
-
-\section{\sysfull}
-\sys is runtime library that facilitates tasks navigation and preserves data/memory consistency of the IPDs.
-It is able, on-demand, to merge tasks to reduce the number of commits to the non-volatile memory.
-
-\input{iposDataProtection}
-
-\subsection{\sys: The Virtualizing Engine}
-\subsubsection{Power Interrupt Immune Scheduler}
-It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures.
-\sys provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
-(un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
-
-\subsubsection{Task Merging Algorithms}
-
-\paragraph{Fixed Virtual Task Size} [better name needed]
-
-	\begin{algorithm}[t]
-		\caption{Fixed virtual Task size}
-		\label{algo:fixVirtTask}
-		\scriptsize
-		%\small
-		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\State MVTS: maximum VT size
-			\vspace{0.1cm}
-
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
-				\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-							\State $ MVTS = VTS $
-						\EndIf
-				\EndWhile
-
-				\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\If{$VTS < MVTS$}
-					\State $VTS++$
-					\EndIf
-				\EndIf
-			\EndWhile
-		\end{algorithmic}
-	\end{algorithm}
-
-	\begin{algorithm}[t]
-		\caption{Opportunistic virtual Task size}
-		\label{algo:fixVirtTask}
-		\scriptsize
-		%\small
-		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\vspace{0.1cm}
-
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
-				\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-						\EndIf
-				\EndWhile
-
-				\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\State $VTS++$
-				\EndIf
-			\EndWhile
-		\end{algorithmic}
-	\end{algorithm}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{figures/virtualTaskSize.eps}
-	\caption{Size of the virtual task versus the execution time of a dummy application that contains 12 empty tasks.}
-	\label{fig:virtualTaskSize}
-\end{figure}
-
-\sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
-Fig.~\ref{fig:virtualTaskSize} show the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
-
-Remark: A less obvious benefit of visualization is that it can enable a secure computation.
-By increasing the size of the virtual task, the energy in the super-capacitor will not be enough to finish the execution of the virtual task.
-Therefore, without the interrogator being close to the intermittently powered device and charging rate is not negligible as compared to the discharging rate this computation will not be finished.
-As such, virtualization can add a layer of security to the computation.
-
-\section{Results}
-We compare the performance of \sys against a state-of-the-art intermittent execution approach.
-For this purpose, we used two different applications: (i) A data decompression application, which utilizes Huffman decoding technique to decompress 100 byte of data.; (ii) A discrete Fourier transform (DFT) application that uses two different resolution (4 and 8 bites) to analyze a randomly generated signal.
-
-\subsection{Continuous Power Supply}
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.8\columnwidth]{figures/chain_vipe}
-	\caption{Time to complation of \sys versus Chain. Decomp: Data decompression application. DFTx: Discrete Fourier Transform.}
-	\label{fig:IPOSPerformance}
-\end{figure}
-
-Fig.~\ref{fig:IPOSPerformance} shows that \sys requires a shorter execution time to execute an application under the MIEM.
-The size of the virtual task that \sys uses equals One real task, in other words, no visualization is applied.
-
-\subsection{Intermittent Power Supply}
-...
-
 \bibliographystyle{plain}

### 2017-07-27 paper/systemdescription.tex

\begin{figure}
	\centering
	%\includegraphics[width=0.25\columnwidth]{figures/}
	\caption{\sys top-level description.}
	\label{fig:}
\end{figure}

Describe the overall system, the programming model, and the semantics/execution model. This should include a discussion of memory virtualization and task coalescing.

### 2017-07-27 paper/evaluation.tex

@@
-We compare the performance of \sys against the state-of-the-art intermittent execution approach---Chain~\cite{chain}.
+We are now ready to present experimental results of \sys performance.
+We start with describing the hardware environment (Section~\ref{sec:results_hardware}) followed by the software environment (Section~\ref{sec:results_software}) and we concluded by the evaluation results (Section~\ref{sec:results_evaluation}).
 
-The applications that were used in the evaluation are given in Table~\ref{table:benchmark_list}.
+\subsection{Hardware Suite Description}
+\label{sec:results_hardware}
+
+In the experiments we used WISP\,5.1~\cite{wisp5,wisp} (with MSP430FR5969~\cite{wolverine} 64\,kB FRAM) or MSP430XXX launchpad~\cite{} (hardware type will be specified per experiment in Section~\ref{sec:results_evaluation}).
+WISP\,5.1 has been programmed using TI FET for experiments.
+FET was also used as a power supply in non-intermittent power experiment.
+To measure energy consumption of the applications (refer to Section~\ref{sec:results_software}) EDB~\cite{edb} was used.
+In experiments involving wireless power, WISP was powered through RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
+RFID reader is controlled by a PC running Ubuntu 10.4 executing open source Python-based \emph{sllurp} library~\cite{sllrp_github} enabling low-level RFID reader control.
+Transmitted power of the RFID reader was 30\,dBm at 915\,MHz center frequency.
+For distance controlled experiments, WISP was positioned at hollow paper tubes placed directly on a lying antenna.
+
+\subsection{Software Suite Description}
+\label{sec:results_software}
+
+We compare \sys against state of the art intermittent execution approach---Chain~\cite{chain}.
+We need to note that although another very relevant task-based runtime has been recently presented and referenced many times throughout this paper---Alpaca~\cite{alpaca}---we did not have the access to its code during the preparation of this manuscript.
+
+The applications that were used in the evaluation are given in Table~\ref{table:benchmark_list}. These benchmarks are the same ones as used in~\cite{chain,alpaca}.
 
@@
 	\end{tabular}
-\caption{List of benchmarks used in \sys evaluation.}
+\caption{List of benchmark programs used in \sys evaluation; SLOC: source lines of code.}
 \label{table:benchmark_list}
@@
 
-\subsection{Continuous Power Supply}
-\label{sec:results_continuous_power}
+\subsection{Evaluation}
+\label{sec:results_evaluation}
+
+\subsubsection{Compiler Performance}
+\label{sec:results_compiler}
+
+Compiler experiment: taskifying completion time of a code run, correctness (manual inspection---proof by example), comparison manual/automatic taskification - fixed power supply
+
+\todo{Provide structure to this section}{Kiwan}
+
+\subsubsection{}
+\label{sec:}
 
@@
 
-%Set of experiments
+Task coalescing experiment: with WISP at 3 distances and with fixed power, for all apps, compared against chain, show amount of task merged, time to completion, overhead (in terms of what?), energy consumed with fixed power
 
-%- Compiler result: taskifying completion time, correctness, 
-%- Task coalescing experiment: with WISP at 3 distances and with fixed power, for all apps, compared against chain, show amount of task merged, time to completion/inter-operation time, overhead (in terms of what?), energy consumed
-%- Battery-less microphone: with \sys and without \sys, compared against Chain, 
-%- Paging/memory experiment: ?
+\subsubsection{Memory Management}
+\label{sec:results_memory_management}
 
-\subsection{Intermittent Power Supply}
-\label{sec:intermittent_continuous_power}
+Paging/memory experiment: page swapping cost (execution time)---comparing against chain and full double buffering approach, show number of page swaps,
+
+\todo{Provide argument for page swapping using hardware support}{Brandon}
+
+\subsubsection{Case Study}
+\label{sec:case_study}
+
+Battery-less microphone: with \sys and without \sys, compared against Chain.

### 2017-07-28 paper/evaluation.tex
bibs have proper links

@@
 
-In the experiments we used WISP\,5.1~\cite{wisp5,wisp} (with MSP430FR5969~\cite{wolverine} 64\,kB FRAM) or MSP430XXX launchpad~\cite{} (hardware type will be specified per experiment in Section~\ref{sec:results_evaluation}).
-WISP\,5.1 has been programmed using TI FET for experiments.
+In all experiments we used the following computing hardware: WISP\,5.1~\cite{wisp5,wisp}, i.e.
+build around MSP430FR5969~\cite{wolverine} 64\,kB FRAM microcontroller, and TI's MSP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} (hardware type will be specified per experiment in Section~\ref{sec:results_evaluation}).
+WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet}.
 FET was also used as a power supply in non-intermittent power experiment.
-To measure energy consumption of the applications (refer to Section~\ref{sec:results_software}) EDB~\cite{edb} was used.
-In experiments involving wireless power, WISP was powered through RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
+To measure energy consumption of all benchmark applications (please refer to Section~\ref{sec:results_software}) Carnegie Mellon's EDB~\cite{edb} was used.
+In experiments involving wireless power, WISP was powered through an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
 RFID reader is controlled by a PC running Ubuntu 10.4 executing open source Python-based \emph{sllurp} library~\cite{sllrp_github} enabling low-level RFID reader control.
 Transmitted power of the RFID reader was 30\,dBm at 915\,MHz center frequency.
-For distance controlled experiments, WISP was positioned at hollow paper tubes placed directly on a lying antenna.
+For distance controlled experiments, WISP was positioned at varying length hollow paper tubes placed directly on a flat lying antenna.
+Code completion measurements have been performed using Salea~\cite{saleae} digital logic analyzer.
 

### 2017-07-28 paper/evaluation.tex
work on evalation section

@@
 We compare \sys against state of the art intermittent execution approach---Chain~\cite{chain}.
-We need to note that although another very relevant task-based runtime has been recently presented and referenced many times throughout this paper---Alpaca~\cite{alpaca}---we did not have the access to its code during the preparation of this manuscript.
+We are not comparing \sys against DINO~\cite{dino} or Mementos~\cite{mementos} (Refer to Section~\ref{sec:background_consistency}), as (i) Chain is the only modular execution model available to which we can compare \sys, and (ii) Chain proves already its superiority over sequential model counterparts (i.e.
+DINO and Mementos).
 
-The applications that were used in the evaluation are given in Table~\ref{table:benchmark_list}. These benchmarks are the same ones as used in~\cite{chain,alpaca}.
+We need to note that although another very relevant task-based runtime has been recently presented and referenced many times throughout this paper---Alpaca~\cite{alpaca} (written by the same group as Chain)---except the recently accepted manuscript we did not have the access to Alpaca's code during the preparation of this paper.
+Therefore Alpaca was excluded from the evaluation.
 
-\begin{table}
+\task{Check if the line of argument is solid for the above paragraphs about not selecting DINO, Memenos and Aplaca}{Brandon}
+
+The applications that were used in the evaluation are given in Table~\ref{table:benchmark_list}.
+These benchmarks are the same ones as used in~\cite{chain,alpaca}.
+In addition to Chain benchmark programs new applications were used for benchmarking: \emph{DFT}, \emph{Huffman compressor}, and \emph{selection sort}.
+
+\begin{table}[t]
 	\begin{tabular}{|c|c|c|}
 		\hline
-		Name & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
+		Nickname & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
 		\hline\hline
-		Temperature sensing & 388 & 721 \\ %53\%
+		Cold chain & 388 & 721 \\ %53\%
 		Cuckoo & 483 & 762 \\ %63\%
 		RSA & 887 & 1233 \\ %71\%
+		Bitcount & 351 & 588 \\ %59\%
 		DFT & --- & --- \\ %Two resolutions: 4 and 8 Bytes
@@
 
-\subsection{Evaluation}
+\todo{finalize describing all applications once all experiments are done}{Przemek}
+
+\subsection{Evaluation Results}
 \label{sec:results_evaluation}
@@
 
-Compiler experiment: taskifying completion time of a code run, correctness (manual inspection---proof by example), comparison manual/automatic taskification - fixed power supply
+We start with evaluating the efficiency of a \sys compiler.
 
-\todo{Provide structure to this section}{Kiwan}
+Expected results:
+\begin{itemize}
+	\item comparison between manual (Chain code) and automatic taskification: in terms of code SLOCs, number of tasks, completion time (Chain), memory use (other metrics required?); 
+	\item assessment of correctness (through manual inspection only?).
+\end{itemize}
 
-\subsubsection{}
-\label{sec:}
+\todo{Provide results and text to this section}{Kiwan}
+
+\subsubsection{Task Coalescing Performance}
+\label{sec:results_coalescing}
+
+We now proceed with the evaluation of task coalescing mechanism of \sys.
+
+Expected results:
+\begin{itemize}
+	\item For all apps: show amount of task merged, time to completion, overhead (in terms of what?), energy consumed.
+All experiments for fixed power, at 3 distances with wireless power, compared against Chain
+\end{itemize}
+
+\todo{Provide results and text to this section}{Amjad}
+
+\subsubsection{Memory Management Performance}
+\label{sec:results_memory_management}
 
@@
 	\includegraphics[width=\columnwidth]{figures/chain_vipe}
-	\caption{Time to complete for a set of testing applications (refer to Table~\cite{table:benchmark_list}) using \sys versus Chain.}
+	\caption{Time to complete for a set of testing applications (refer to Table~\ref{table:benchmark_list}) using \sys versus Chain (no coalescing applied, fixed power supply).
+\todo{Figure placeholder---will be replaced with final one}{Amjad}}
 	\label{fig:IPOSPerformance}
@@
 
-Fig.~\ref{fig:IPOSPerformance} shows that \sys requires a shorter execution time to execute an application under the modular execution model.
-The size of the virtual task that \sys uses is one real task, i.e.
-no task coalescing is applied.
-
-Task coalescing experiment: with WISP at 3 distances and with fixed power, for all apps, compared against chain, show amount of task merged, time to completion, overhead (in terms of what?), energy consumed with fixed power
-
-\subsubsection{Memory Management}
-\label{sec:results_memory_management}
-
-Paging/memory experiment: page swapping cost (execution time)---comparing against chain and full double buffering approach, show number of page swaps,
+Expected results:
+\begin{itemize}
+	\item Paging/memory experiment: page swapping cost (execution time)---comparing against chain and full double buffering approach, show number of page swaps,
+\end{itemize}
 
@@
 
-Battery-less microphone: with \sys and without \sys, compared against Chain.
+The final experiment assessing usefulness of \sys is its use in real application: battery-less microphone.
+
+Expected results:
+\begin{itemize}
+	\item Compare execution time: \sys compared against Chain
+\end{itemize}
+
+\todo{Provide results and text to this section---if time allows (this result is not critical)}{Amjad/Przemek}

### 2017-07-28 paper/evaluation.tex
Evalation section complete for now (will resume after experiments done)

@@
 
-\task{Check if the line of argument is solid for the above paragraphs about not selecting DINO, Memenos and Aplaca}{Brandon}
+\todo{Check if the line of argument is solid for the above paragraphs about not selecting DINO, Memenos and Aplaca}{Brandon}
 
@@
 \begin{itemize}
-	\item comparison between manual (Chain code) and automatic taskification: in terms of code SLOCs, number of tasks, completion time (Chain), memory use (other metrics required?); 
+	\item comparison between manual (Chain code) and automatic taskification: in terms of code SLOCs, number of tasks, completion time (Chain), memory use (other metrics required?)
 	\item assessment of correctness (through manual inspection only?).
@@
 \begin{itemize}
-	\item Paging/memory experiment: page swapping cost (execution time)---comparing against chain and full double buffering approach, show number of page swaps,
+	\item page swapping cost in terms of execution time: compare against Chain and full double buffering approach; show number of page swaps---do the experiment for all applications
 \end{itemize}
@@
 
-The final experiment assessing usefulness of \sys is its use in real application: battery-less microphone.
+The final experiment assessing usefulness of \sys's use in real application.
+For this we have a implemented an battery-less wirelessly-powered sound detector as an example.
+To the best of our knowledge, this is the first case study that demonstrates a intermittently-powered system supplied by harvested energy at runtime, which continuously interacts with real-life signals.
+
+The application aims at detecting a specific tone in the audio signal (in the implementation: 1\,kHz) based on on-the-fly audio measurement.
+For this we have connected an analog MEMS microphone~\cite{microphone} output to AUX3 port of WISP\,5.1 (while the microphone was powered directly by WISP, which was then wirelessly power-supplied by the RFID reader).
+Audio signal is sampled by the microcontroller's ADC and passed to the FFT function (internal function of the TI's Digital Signal Processing library~\cite{ti_dsp}) to find a tone above a predefined threshold.
+The output of the detection is signalled by the WISP on-board LED.
+The video of demonstrating the working system is available via \url{https://link_to_a_video}.
 
@@
 \begin{itemize}
-	\item Compare execution time: \sys compared against Chain
+	\item Compare execution time: \sys against Chain for X\,s long audio signal detection at two antenna distances
+	\item Make a video of the demonstration 
 \end{itemize}
 
+\todo{Check if we really need this application}{Brandon}
 \todo{Provide results and text to this section---if time allows (this result is not critical)}{Amjad/Przemek}

### 2017-07-28 paper/abstract.tex
Asbtract swiftly rewritten - ready for Brandon

@@
-%\TODO{Need to re-write this to make it consistent with the story after we have the story in place}
-
+Enabling battery-free devices is a mandatory step towards an environment-friendly Internet of Things (IoT).
+However, removing the batteries requires IoT to operate on an harvested, ergo intermittent, power supply.
+This make sustaining long computation very challenging.
+Correspondingly, there are two approaches to enable long-running computations on intermittenly-powered IoT: (i) checkpointing, where the volatile state of a program is frequently saved to the non-volatile memory, and (ii) task-based approach, where a programmer splits the code into small idempotent modules.
+Results so far suggest that task-based approach performs better than checkpointing.
+However, the use of tasks requires that the energy to execute any task must not exceed the maximum limit of the energy buffer.
+This makes the code unefficient (when executed on a large energy buffer than indended) and potentially completely disables the program (in the opposite case).
+Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing.
+If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve the performance.
 \todo{Re-write the abstract once the paper is ready}{Brandon}
-
-Enabling battery-free devices is a mandatory step towards the environment friendly IoT.
-However, removing these large energy reservoirs (the batteries) makes the devices operate on an intermittent power supply, which make sustaining any long computation very challenging.
-correspondingly, there are two proposed approaches to enable long-running computations on intermittenly powered devices (IPDs): (i) Checkpointing based approach, where the volatile state of a program is frequently saved to the non-volatile memory and; (ii)  Task based approach, where a programmer splits the code into small idempotent modules.
-The Task based approach shows a better performance than the checkpointing approach.
-However, the task based method requires that the energy to execute any task must not exceed the maximum limit of the energy buffer.
-This limitation makes this solution a customized one that suits a particular energy buffer size.
-
-\sys enables code portability by means of virtualizing the task based model.
-If a code is written for a very small energy buffer, \sys is able to dynamically merge these tasks and constructs a bigger virtual task to improve the performance, reduce energy consumption and provides mean for code portability.

### 2017-07-28 paper/intro.tex
Intro revised - ready for Brandon to re-read it

@@
 3]{nvp}.
-
+%
 \todo{Put reference for "tictpl" somewhere}{Brandon}
@@
 While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb}.
-
-%FRAM costs more than FRAM -- need to figure out how to execute out of SRAM
 
@@
 
-%Task sizing is hard -- how to chop without risking inefficiency or non-term
-
 Two recent works~\cite{alpaca,chain} proposed {\em task-based} programming and execution paradigm, that require a programmer to statically decompose a program into tasks.
@@
 
-To answer these, in this work we develop {\bf \sys}: a new programming and execution model that utilizes two new runtime features: {\em dynamic task coalescing} to efficiently execute tasks without exceeding available energy (answering question (i)) and  {\em volatile memory virtualization} to leverage efficient volatile memory (answering question (ii)).
+To answer the above, in this work we develop {\bf \sys}: a new programming and execution model that utilizes two new runtime features: {\em dynamic task coalescing} to efficiently execute tasks without exceeding available energy (answering question (i)) and  {\em volatile memory virtualization} to leverage efficient volatile memory (answering question (ii)).
 \sys's novel memory virtualization mechanism uses a device's SRAM as the working memory during a task's execution, populating the SRAM with a page of data from FRAM as necessary at the start of the task.
@@
 When a task accesses data outside of a page in SRAM, the task must fetch data from FRAM.
-
+%
 \todo{articulate this last point better}{Przemek}
@@
 \todo{Provide numerical results for this section}{Przemek}
-\todo{Mention automatic task generation}{Przemek}
+
+\item {\bf Compiler-support automatic task generation:} second \sys's core contribution is the ability to split standard C programs into \sys's compatible tasks, effectively circumventing the need for labour-intensive manual task generation of the state of the art solutions~\cite{chain,alpaca}.
+%
+\todo{Elaborate on this more and provide numerical results}{Przemek}
 %
@@
 %
-\item {\bf Extensive \sys evaluation:} We evaluated \sys on a collection of benchmark programs~\cite[Sec.
-5]{chain}, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexible target a variety of platforms without recompiling the code.
-We further demonstrate the efficiency and versatility of \sys with an end-to-end evaluation of a sensing and data processing application deployed in a lab environment.
+\item {\bf Extensive \sys evaluation:} We evaluated \sys end-to-end on a collection of existing benchmark programs~\cite[Sec.
+5]{chain} and new ones (including realistic case study of battery-less embedded sound detector) in a laboratory environment, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexible target a variety of platforms without recompiling the code.
 %
 \todo{add detailed text about evaluation results}{Przemek}
+\todo{decide if we want to make this a contribution paragraph or not (then, put it as a closing paragraph)}{Brandon}
 %

### 2017-07-28 DMA/ipos-gcc-dma/ipos/psched.c
add a second method for the scheduler to do relative jump (initial test is done), VIPER still uses the old method

@@ -5,10 +5,28 @@
 #define COMMITTING      1
 #define COMMIT_FINISH   0
 
-#define JUMP2()  unsigned int totJumpSize = (*(__current_task_virtual + BLOCK_OFFSET_PT) + jump_by)\
-                    if(totJumpSize > __totNumTask)\
-                    {  __current_task_virtual  =  (unsigned int*) (*(__current_task_virtual+ ( (__totNumTask - totJumSize)+(__totNumTask - totJumSize) )  ) ) } \
-                    else{__current_task_virtual  =  (unsigned int*) (*(__current_task_virtual+ ( jump_by+jump_by  ) )  }
+
+#define JUMP2()  { \
+    if(__jump !=1)  \
+    {   \
+            __current_task_virtual = (unsigned int *) *(__current_task_virtual + NEXT_OFFSET_PT) ;     /* soft transition */  \
+    }else{  \
+        unsigned int totJumpSize = (*(__current_task_virtual + BLOCK_OFFSET_PT) + __jump_by);   \
+            if(totJumpSize > __totNumTask)  \
+            {   \
+                int dis  = totJumpSize - __totNumTask;  \
+                while( dis  > __totNumTask) \
+                {   \
+                    dis = dis - __totNumTask;   \
+                }   \
+                dis = dis - (totJumpSize - __jump_by);  \
+                __current_task_virtual  += (dis + dis + dis) ;  \
+            }else{  \
+                __current_task_virtual  += ( __jump_by+__jump_by  + __jump_by )  ;  \
+                }   \
+            __jump = 0; \
+        }   \
+}
 
 #define JUMP();    if(__jump !=1){   \
                         __current_task_virtual  =  \
@@ -42,8 +60,10 @@ volatile unsigned int __taskCounter = 1;
 __nv volatile unsigned int __temp_taskCounter = 0;
 __nv volatile unsigned int __totalTaskCounter = 0;
 
+//TODO I do not think these two variables need to be persistent
 __nv volatile unsigned int __jump = 0;
 __nv volatile unsigned int __jump_by = 0;
+
 volatile unsigned int __jump_cnt = 0;
 
 __nv unsigned int __reboot_state[2] = { 0 };    //virtual Task size control
@@ -60,6 +80,29 @@ void os_exit_critical()
 {
     __bic_SR_register(GIE);
 }
+
+
+//void JUMP2()  {
+//    if(__jump !=1)
+//    {
[…]

### 2017-07-28 paper/background.tex
Background revised almost

--- 2017-07-26 paper/background.tex
@@
-\sys addresses the problem of intermittently executing task-based computations
-on energy-harvesting hardware, without being limited by inflexible, static task
-decomposition and without being limited to fully non-volatile main memory.  In
-this section, we provide background on energy harvesting systems and the
-intermittent software execution model.   We then discuss the key limitations of
-existing task-based programming and execution models for intermittent computing
-that \sys addresses.
+\sys addresses the problem of intermittently executing task-based computations on energy-harvesting hardware, without being limited by inflexible, static task decomposition and without being limited to fully non-volatile main memory.
+In this section, we provide background on energy harvesting systems and the intermittent software execution model.
+We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys addresses.
 
@@
 
-Energy harvesting devices are systems that operate using energy extracted from
-their environment, from sources such as radio waves (RF)~\cite{wisp}, and solar
-energy.  By operating entirely using environmental energy, energy-harvesting
-devices need not use a tethered power supply or battery.  Instead, devices
-typically collect energy into a (super-)capacitor, operate when sufficient
-energy has accumulated, and upon depleting the stored energy, turn off and wait
-to collect more energy.
+Energy harvesting devices are systems that operate using energy extracted from their environment, from sources such as radio waves (RF)~\cite{wisp}, and solar energy.
+By operating entirely using environmental energy, energy-harvesting devices need not use a tethered power supply or battery.
+Instead, devices typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.
 
-Batteryless operation has a number of important advantages, making intermittent
-computing an important domain.  Supplying power to the ``next
-billion''~\cite{citationneeded} embedded computers using batteries is not
-sustainable.  The European Commission estimates that more than 160 kilotons of
-consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
-Batteries constitute a potential environmental hazard because not all batteries
-are recycled.  Batteries are fragile and limited in their number of
-charge/discharge cycles, requiring costly maintenance that is possibly only
-through physical device access, which may be difficult if devices are deployed
-in inaccessible places (e.g., in space~\cite{kicksat}). By contrast,
-super-capacitors are durable, promising millions of charge/discharge
-cycles~\cite[Sec. I]{ongaro_pwre_2012} The main limitation in moving to
-capacitor-based energy storage is that capacitor energy density --- and
-consequently operating discharge time --- is orders of magnitude less than a
-battery.
+Batteryless operation has a number of important advantages, making intermittent computing an important research domain.
+Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable.
+The European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
+Batteries constitute a potential environmental hazard because not all batteries are recycled.
+Batteries are fragile and limited in their number of charge/discharge cycles, requiring costly maintenance that is possibly only through physical device access, which may be difficult if devices are deployed in inaccessible places, e.g.
+in space~\cite{kicksat}.
+By contrast, super-capacitors are durable, promising millions of charge/discharge cycles~\cite[Sec.
+I]{ongaro_pwre_2012}.
+The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently: operating discharge time---is orders of magnitude less than a battery.
+\todo{Compress text}{Brandon}
 
-\todo{Compress text}{Brandon}
-Given current technology development, battery-less systems are best suited for
-very long-term sensing and monitoring where access to recharge is either
-prohibitive or impossible. These include battery-less image capture and
-processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}
-and digestible~\cite{nadeau_naturebio_2017} sensors.
+Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible.
+These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
 
-Many platforms enable intermittent, battery-less computation. For instance,
-computation RFIDs---open-source TI MSP430-based~\cite{wolverine}
-WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
-NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
-such as~\cite{medusa_farsens_2017}. Other intermittently-powered platforms
-include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or
-battery-less phone~\cite{talla_imwut_2017}. In all of the above, the main
-source of energy harvested is the electromagnetic radiation in the radio
-frequency range (ambient transmitters such as high power TV
-transmitters~\cite{liu_sigcomm_2013} or dedicated RFID
-antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
+Many platforms enable intermittent, battery-less computation.
+For instance, computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
+Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
+In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio
+frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
 Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid).
@@
 	\centering
+	\footnotesize
 	\begin{tabular}{|c|c|}
@@
 A conceptual implementation of intermittent execution model is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between execution times.
-
-\subsection{Assumption on the Type of Computing Hardware}
-\label{sec:background_hardware}
-
-Before proceeding further with the background we need to make certain assumptions about the underlying hardware.
-We assume that we operate on the microcontroller (more generally, computing hardware) equipped with a mix of volatile and non-volatile memory, such as FLASH, Ferroelectric RAM or Magnetoresistive RAM.
-\sys does not assume to rely on specific memory type (either volatile or non-volatile), in contrary to earlier works targeting non-volatile processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
-Actually, \sys will exploit the properties of both memory types to maximize performance metrics of the program being executed.
-We make \sys able to I/O operations, however we do not enable the scheduling of operations (which \emph{nota bene} forms a foundation for a simple runtime).
-All in all, our hardware model follows in assumptions the two most relevant intermittently-powered execution environments~\cite{alpaca,chain}.
 
@@
 
-We can classify the intermittent execution models into two classes: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
+Intermittent execution models combatting WAR can be classified into: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
 )
@@
 3]{chain}, where it is concluded that each checkpoint causes overhead).
+\todo{I use the word `checkpoint' for normal checkpoint and task boundary interchangeably---check if it does not cause confusion}{Brandon}
 
@@
 Therefore, it tolerates arbitrary number of power interrupts.
-This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an intermittently-powered device by interleaving their tasks~\cite{}.
+This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an intermittently-powered device by interleaving their tasks.
 However, it obviously requires code modification.
-For example, if an algorithm is written according to the continuous execution model it has to be split by a programmer, into small tasks to run under the modular intermittent execution model.
+For example, if an algorithm is written according to the continuous execution model it has to be split by a programmer into small tasks to run under the modular intermittent execution model.
 
-Despite the benefit of the modular model certain limitations are still present that we must overcome and which inspired us to design \sys. Let us look at these problems in detail.
+Despite the benefit of the modular model certain limitations are still present that must be overcome and which inspired us to design \sys. Let us look at these problems in detail.
 
@@
 	\centering
+	\footnotesize
 	\begin{tabular}{|c|c|c|c|}
@@
 
-We have executed an experiment to measure exactly cost of read/write operations to volatile and non-volatile memory.
+We have executed an experiment to measure the cost of read/write operations of $w$ to volatile and non-volatile memory.
 The result is given in Figure~\ref{fig:framEnergy}.
-Four applications were developed to measure the energy consumption of \emph{accessing} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{msp430datasheet}.
+Four applications were developed that \emph{access} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{msp430datasheet}.
 The interference free debugger (EDB)~\cite{edb} is used as the energy measurement tool.
@@
 Therefore, we chose to write a read value to SRAM.
-Based on our measurements we conclude that a FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We would like to mention that TI has already mentioned in this document~\cite{} that FRAM write access is more energy expensive than SRAM write access.
-However, TI does not quantify the energy overhead.}.
+Based on our measurements we conclude that a FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We remark that TI already mentions~\cite{ti_fram_faq} that FRAM write access is more energy expensive than SRAM write access.
+However, TI does not quantify this energy overhead.}.
 Therefore, reducing FRAM access is desirable when developing energy limited software solution---\emph{which is in sharp contrast to the main goal of all checkpoining solutions}.
@@
 On the other hand, this increases task/checkpoint re-execution time.
-The ideal solution is a runtime that \emph{adaptively} changes task size during runtime, which brings us to a second problem.
+The ideal solution is a runtime that \emph{adaptively} changes task size during runtime, which however exposes a second problem.
 
@@
 A modular intermittent execution model completely depends on a programmer's estimation which mostly results in a sub-optimal code division.
-Moreover, this model can only consider a single hardware configuration and it does not take environment changing into considerations.
+Moreover, this model can only consider a single hardware configuration and it does not take environment variability into considerations.
 Each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}, which causes different charge/discharge periods.
-This result in different execution times and in the worst case the size of the capacitor might be too small to execute one task written by the programmer.
+This result in different execution times.
+In the worst case the size of the capacitor might be too small to execute one task written by the programmer.
 This calls for the ability of the modular execution runtimes to \textbf{merge} multiple small tasks into a larger one or \textbf{split} a large task into a set of small ones on the fly (post-compilation time).
@@
 \begin{itemize}
-	\item \textbf{Lack of multi-versioning system} All known modular execution models define its own input for each task that is not shared with other tasks.
-This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
+	\item \textbf{Irreversible tasks transitions} According to both Chain~\cite{chain} and Alpaca~\cite{alpaca} programming models, each task calls the next one.
+The transition is done through functions calls and manually clearing the stack to prevent the stack overflow problem.
+This approach makes all the task transitions firm, preventing task merging.
 
-	\item \textbf{Irreversible tasks transitions} According to both Chain~\cite{chain} and Alpaca~\cite{alpaca} programming models, each task calls the next one.
-The transition is done through functions calls and manually clearing the stack, to prevent the stack overflow problem.
-This approach makes all the task transitions firm, preventing task merging.
+	\item \textbf{Dependency on privatization} Alpaca~\cite{alpaca} extended Chain~\cite{chain} with a concept of privatization, i.e.
+it protects task-shared variables from WAR dependency problem \emph{within} the scope of a task.
+Therefore, if tasks are on-demand merged, the scope of these dependencies are changed and the static analysis to privatize these variables is not valid any more and the data consistency can not be preserved.
 
@@
 However, this execution path is only a special one.
-If the general case (see Fig.~\ref{fig:DynamicChainRan}) is considered then leaving out the intra-merged-task channels might result in data inconsistency.
+If the general case (see again Fig.~\ref{fig:DynamicChainSeq}) is considered then leaving out the intra-merged-task channels might result in data inconsistency.
 Than is, if an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready, which result in an incorrect execution.
 
-	\item \textbf{Dependency on Privatization} Alpaca~\cite{alpaca} extended Chain~\cite{chain} with a concept of privatization, i.e.
-it protects task-shared variables from WAR dependency problem \emph{within} the scope of a task.
-Therefore, if tasks are on-demand merged the scope of these dependencies are changed and the static analysis to privatize these variables is not valid anymore and data consistency can not be preserved.
+	\item \textbf{Lack of multi-versioning system} All known modular execution models define its own input for each task that is not shared with other tasks.
+This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
+\todo{Rewrite this paragraph: still unclear}{Przemek}
 \end{itemize}
+
+\subsection{Main Idea: Tasks Coalescing}
+
+To combat the above limitations we propose a new concept of \textbf{virtualization}: a modular execution model with softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
+Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
+Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
+This can be done by having a temporary copy of a global variable in the volatile memory (similar to the caching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
+
+\subsection{Assumption on the Type of Computing Hardware}
+\label{sec:background_hardware}
+
+Before proceeding further with the background we need to make certain assumptions about the underlying hardware.
+We assume that we operate on the microcontroller (more generally, computing hardware) equipped with a mix of volatile and non-volatile memory, such as FLASH, Ferroelectric RAM or Magnetoresistive RAM.
+We do not assume to rely on specific memory type (either volatile or non-volatile), in contrary to earlier works targeting non-volatile processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
+Actually, we will exploit the properties of both memory types to maximize performance metrics of the program being executed.
+We enable to I/O operations, however we do not enable the scheduling of operations (which \emph{nota bene} forms a foundation for a simple runtime).
+All in all, our hardware model follows in assumptions the two most relevant intermittently-powered execution environments~\cite{alpaca,chain}.
+\todo{Verify if this section is not out of place}{Brandon}
 
@@
 \begin{table}
+	\centering
+	\footnotesize
 	\begin{tabular}{|c|c|c|c|}
@@
 \todo{Elaborate on the ``key challenge'' paragraphs near the end of the intro}{Przemek}
-
-%%%Notes%%%
 

### 2017-07-29 paper/background.tex

@@
 
-Task is a programmer (or compiler) defined code that is executed based on consistent set of inputs, i.e.
+\todo{Please check this is the right way for this section: it reads as it should be in Section 3 describing \sys}{Brandon} Task is a programmer (or compiler) defined code that is executed based on consistent set of inputs, i.e.
 memory fetched from a specified memory location, producing a consistent output stored in a specified (the same or different) memory location.
@@
 A comparative description of \sys is given in Table~\ref{table:feature_comparison}.
+\todo{Decide if we need such table}{Brandon}
 

### 2017-07-30 paper/compiler.tex
Cleaned up compiler section; provided writing tasks for Kiwan

@@
-\sys additionally provides a compiler support for automatic task generation, eliminating programmer's burden of writing a task-based program from scratch.
-The compiler gets a plain-C code as an input and generates a binary that is rewritten in task-based programming style that can run on \sys without additional programming effort.
-Automatic task generation compiler is often inefficient than human-written code for \sys because if unbounded loop or pointer is present, the compiler makes a conservative decision.
-\emph{If the performance is critical, the programmer may want to manually write the task-based code by himself.
-Otherwise, the programmer can rely on the compiler for simplicity.}
+A unique feature of \sys is a compiler support for \emph{automatic task generation}.
+\sys literally eliminates the programmer's burden of writing a task-based program from scratch.
+In essence, the \sys compiler obtains a plain C code as an input and generates a binary file that is rewritten in task-based programming style that \sys executes.
 
-The compiler consists of two part: analysis pass and transform pass.
-Analysis pass analyze the code structure and decides where to put task boundaries.
-Transform pass converts the code into task-style program and pepper it with necessary \sys keywords so that it can run on \sys system.
-It is implemented using LLVM.
+The compiler consists of two parts called \emph{Passes}:
+%
+\begin{itemize}
+	\item \textbf{Analysis Pass}, which analyzes the code structure and decides where to put task boundaries;
+	\item \textbf{Transform Pass}, which converts the code into task-style program and fills it with necessary \sys keywords so that it can run on \sys.
+\end{itemize}
+%
+We shall now proceed with describing these two parts in detail.
 
-\subsection{Analysis Pass}
-\label{sec:analysis}
+\subsection{Compiler Analysis Pass}
+\label{sec:compiler_analysis_pass}
 
-\sys compiler decides the task size based on the number of LLVM IR instructions.
-That is, it tries to make every task to contain roughly the same amount of LLVM IR
-instructions. Intuitively, the number of LLVM IR instructions have a correlation with
-the total workload done by the task. However, the number of LLVM IR instructions is 
-a crude approach of estimating workload of a task since actual machine instructions are
-what count and it is not exactly the same as LLVM IR instruction. Also, different instructions
-have different power usage, use different cycles, etc. However, the size of each
-task need not be exactly the same and it is sufficient to make sure that all of the
-tasks are small enough to fit in to the energy budget. Counting LLVM IR instruction is sufficient
-in this perspective.
+In the first pass \sys compiler decides on the task size.
+\sys' task size metric is defined by the number of LLVM Intermediate Representation (IR) instructions.
+That is, \sys tries to make every task to have roughly the same amount of LLVM IR instructions.
 
-\sys compiler first builds a CFG graph where each basic block is a vertice and branch
-from one basic block to another is a unidirectional edge. Each vertices has a
-{\em price}, which is the number of LLVM IR inside the vertice. 
-When forming a task, each task should contain a basic block so that the sum of price of the 
-blocks within each tasks should be roughly the same. Price makes the workload done by
-each tasks roughly the same and make no task exceed the energy bound.
+\begin{figure}
+	\centering
+	%\includegraphics[width=\columnwidth]{figures/compiler_graph}
+	\caption{\sys compiler flow diagram.\todo{Draw compiler flow graph diagram}{Kiwan}}
+	\label{fig:compiler_graph}
+\end{figure}
 
-Each edge has a {\em penalty}, which indicates the penalty of placing a task boundary at
-the specific edge. Intuitively, when a task boundary is placed on a frequently visited edge
-(such as a back edge of a loop) the system suffers from higher runtime overhead.
-There is no way of statically estimating the number of visit of each edges, so we initally set every
-edge penalty as 1. When analyzing loops, penalty of edges within loops are multiplied by loop count.
-\todo{Change the penalty to be 1 or 1*loopcount}{Kiwan}
+\sys compiler first builds a control flow graph where (i) vertice is a \emph{each basic block} and (ii) a unidirectional edge is a \emph{branch from one basic block to another}, see Figure~\cite{fig:compiler_graph}.
+Each vertice has a {\em price}, which is the number of LLVM IRs inside the vertice.
+When forming a task, each task should contain a basic block so that the sum of price of all blocks within each tasks should be roughly the same \todo{Provide more details on this algorithm}{Kiwan}.
+Price makes the workload done by each tasks roughly the same and make no task exceed the energy bound.\todo{Specify how do you decide on the minimum task size/energy reservoir}{Kiwan}
 
-Ater constructing a graph and assigning price and penalty, the compiler analyzes loops.
+Each edge has a {\em penalty}, which indicates the penalty of placing a task boundary at the specific edge.
+Intuitively, when a task boundary is placed on a frequently visited edge (such as a back edge of a loop) the system suffers from higher runtime overhead.
+As we are unaware of a way to statically estimate the number of visit at each edge, we therefore initially set every edge penalty to one.
+When analyzing loops, penalty of edges within loops are multiplied by loop count \todo{Change the penalty to be one or 1$\times$loopcount}{Kiwan}.
+After constructing a graph and assigning price and penalty, the compiler analyzes loops \todo{Explain what do you mean by analyzing loops}{Kiwan}.
+Except those inside a loop which is set as the loop count \todo{Explain this statement}{Kiwan}.
+Unbounded loops are conservatively considered as very large \todo{Explain this statement}{Kiwan}.
 
-except those inside a loop which is set as the loop count. Unbounded loops are conservatively considered very large.
+\subsection{Compiler Transform Pass}
+\label{sec:compiler_transform_pass}
 
-\subsection{Transform Pass}
-\label{sec:transform}
+Compiler transform pass is also implemented using LLVM. \todo{Fill in this section with details}{Kiwan}
+
+\subsection{Compiler Design Decisions}
+\label{sec:compiler_limitations}
+
+To be able to implement \sys compiler certain critical assumptions had to be made.
+
+\begin{enumerate}
+	\item \textbf{Conservative handling of unbounded code sections:} When a C program contains an unbounded loop or pointer, the \sys compiler makes a conservative decision.
+\todo{Expand on this statement (define `conservative decision') by providing example and more details on this operation}{Kiwan}.
+This means that for performance-critical implementations, the programmer is advised to manually split the code into tasks.
+	%
+	\item \textbf{LLVM IR instruction as a workload metric:} Intuitively, the number of LLVM IR instructions correlates with the total workload performed by the task.
+However, we need to be aware that is nevertheless a crude approach of estimating task workload---actual machine instructions are what counts and they are not exactly equal to the LLVM IR instruction.
+Also, different instructions have different power usage, use different cycles, etc.
+However, the size of each task need not be exactly the same and it is sufficient to make sure that all of the tasks are small enough to fit in to the energy budget.
+Therefore we conjecture that from the \emph{implementation simplicity perspective} of \sys's compiler counting the LLVM IR instruction is sufficient.
+\end{enumerate}
+
+In the light of this design decisions we will assess the compiler overhead in Section~\ref{sec:results_compiler}.

### 2017-07-30 paper/coalescing.tex
add a second method for the scheduler to do relative jump (initial test is done), VIPER still uses the old method

@@
+\sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks.
+
 \subsubsection{Task Merging Algorithms}
@@
 
-It utilizes a persistent circular buffer (persistent linked-list) to keep the state of a program across power failures.
+\input{relativeJumpAlgo.tex} %Task jumping algorithm
+
+It utilizes a persistent circular buffer (persistent linked list) to keep the state of a program across power failures.
 \sys provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
@@
 
-	\begin{algorithm}[t]
-		\caption{Fixed virtual Task size}
+	\begin{algorithm}
+		\caption{Task coalescing mechanism}
 		\label{algo:fixVirtTask}
@@
 			\State VTS : VT size
-			\State MVTS: maximum VT size
+			\State MVTS: maximum VT size \Comment{line added in fixed size approach}
 			\vspace{0.1cm}
@@
 							\State $VTS--$  
-							\State $ MVTS = VTS $
+							\State $ MVTS = VTS $ \Comment{line added in fixed size approach}
 						\EndIf
@@
 				\If {$ \text{All tasks executed}$}
-					\If{$VTS < MVTS$}
+					\If{$VTS < MVTS$} \Comment{line added in fixed size approach}
 					\State $VTS++$
@@
 
-	\begin{algorithm}[t]
-		\caption{Opportunistic virtual Task size}
-		\label{algo:fixVirtTask}
-		\scriptsize
-		%\small
-		\begin{algorithmic}[1]
-			\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\vspace{0.1cm}
-
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
-				\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-						\EndIf
-				\EndWhile
-
-				\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\State $VTS++$
-				\EndIf
-			\EndWhile
-		\end{algorithmic}
-	\end{algorithm}
-
-\begin{figure}[t]
+\begin{figure}
 	\centering
@@
 
-\sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks. 
-Fig.~\ref{fig:virtualTaskSize} show the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
+The question remains, is the best strategy to coalesce all tasks into a singe one, as the energy becomes fully available, i.e.
+from energy harvesting to continuous power environment.
+To answer this question we have implemented a simple program composed of $x$ empty tasks.
+Figure~\ref{fig:virtualTaskSize} shows the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
 
-Remark: A less obvious benefit of visualization is that it can enable a secure computation.
-By increasing the size of the virtual task, the energy in the super-capacitor will not be enough to finish the execution of the virtual task.
-Therefore, without the interrogator being close to the intermittently powered device and charging rate is not negligible as compared to the discharging rate this computation will not be finished.
-As such, virtualization can add a layer of security to the computation.
+%\begin{algorithm}
+%	\caption{Opportunistic virtual Task size}
+%	\label{algo:fixVirtTask}
+%	\scriptsize
+%	%\small
+%	\begin{algorithmic}[1]
+%		\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
+%		\State VTS : VT size
+%		\vspace{0.1cm}
+%		
+%		\While {$True$}
+%		\State $VT \leftarrow VT_{next}$
+%		\vspace{0.1cm}
+%		\While {execute $VT$} 
+%		\If { $\text{power failed twice}$ }				
+%		\State $VTS--$  
+%		\EndIf
+%		\EndWhile
+%		
+%		\vspace{0.1cm}
+%		\If {$ \text{All tasks executed}$}
+%		\State $VTS++$
+%		\EndIf
+%		\EndWhile
+%	\end{algorithmic}
+%\end{algorithm}

### 2017-07-30 paper/systemdescription.tex

@@
 	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{\sys top-level description.}
-	\label{fig:}
+	\caption{\sys top-level view.}
+	\label{fig:system_overview}
 \end{figure}
 
-Describe the overall system, the programming model, and the semantics/execution model. This should include a discussion of memory virtualization and task coalescing.
+\sys as a whole is characterized by the following features:
+
+\begin{itemize}
+	\item \textbf{Task coalescing}: details will be provided in Section~\ref{sec:task_coalescing}
+	\item \textbf{Memory virtualization}: details will be provided in Section~\ref{sec:memory_virtulaization}
+	\item \textbf{Compiler support}: details will be provided in Section~\ref{sec:compiler}
+\end{itemize}
+
+A complete logical structure of \sys is given in Figure~\ref{fig:system_overview}. Before describing each feature of \sys in detail we need to introduce a programming and semantics/execution model.
+
+\subsection{Programming Model}
+\label{sec:overview_programming_model}
+
+\subsection{Semantics and Execution Model}
+\label{sec:overview_semantics}

### 2017-07-31 paper/intro.tex
pass over intro with lots of edits.  I think the contribs list is about 400% too long.  I would opt to have 2 single-column lines per contrib and relegate the details to sections 3 and 4

@@
 3]{nvp}.
-%
-\todo{Put reference for "tictpl" somewhere}{Brandon}
 
 Software in an energy-harvesting system operates according to the {\em
-intermittent execution model}~\cite[Sec.
-2]{dino}~\cite[Fig.
-2]{lucia_snapl_2017}, with bursts of execution interrupted by failure periods.
-When power fails, a device loses its volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
-While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb}.
+intermittent execution model}~\cite{dino,lucia_snapl_2017}, with bursts of
+execution interrupted by failure periods.  When power fails, a device loses its
+volatile state, i.e., registers, stack, SRAM, and retains its non-volatile
+state, i.e., FRAM, Flash. While capturing periodic
+checkpoints~\cite{mementos,quickrecall} and sleep
+scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve execution
+progress, failures can leave volatile and non-volatile state inconsistent,
+leading to unrecoverable failures~\cite{mspcdino,edb}.
 
-There are two fundamental approaches for dealing with data inconsistency for intermittently-powered devices: (i) programming and execution models~\cite{dino,ratchet,chain,alpaca} and (ii) architecture mechanisms~\cite{hicks_isca_2017,idetic,nvp}.
-Nonetheless, they both posses critical limitations.
-New architectures require costly hardware changes and are inapplicable to today's embedded off-the-shelf systems.
-For instance, they require dedicated idempotency violation control hardware buffers~\cite[Fig.
-3]{hicks_isca_2017} or dedicated on-processor logic like on demand register backup~\cite[Fig.
-9]{nvp}.
-Then, new programming models and compilers force programmers to use new memory models~\cite{chain,ratchet}, which may inhibit widespread adoption by programmers accustomed to classical (C) program writing.
-New memory models may also increase the amount of non-volatile memory accesses by requiring non-volatile versioning~\cite{chain} or completely precluding the use of volatile memory~\cite{ratchet}.
-Increasing pressure on non-volatile memory or precluding the use of volatile memory may degrade efficiency; e.g., FRAM has a higher access energy and latency than SRAM~\cite[Fig.
-4]{nvp}---which will be also shown quantitatively in this paper.
-Commercially available devices are often provisioned with much less SRAM than FRAM, e.g.
-up to 64 FRAM/SRAM ratio of~\cite{wolverine}, and effectively leveraging the relatively high efficiency of SRAM is a key challenge for task-based intermittent systems.
-Despite the limitations, we conjecture that programming and execution models will be a {\em de facto choice for sustaining data consistency} of intermittently-powered devices.
-Simply put, programming model does not require any hardware investment and is dependent only upon the programmer's skills.
+There are two main approaches for dealing with data inconsistency for
+intermittently-powered devices: (i) programming and execution
+models~\cite{dino,ratchet,chain,alpaca} and (ii) architecture
+mechanisms~\cite{hicks_isca_2017,idetic,nvp,tictpl}. While both approaches represent
+progress, each has critical limitations. New architectures require costly
+hardware changes and are inapplicable to today's embedded off-the-shelf
+systems~\cite[Fig.  3]{hicks_isca_2017}~\cite[Fig. 9]{nvp}. New programming
+models and compilers force programmers to change behavior~\cite{chain,ratchet},
+inhibiting adoption.  New memory models may require more accesses to
+non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude
+the use of volatile memory~\cite{ratchet}.  Effective use of both volatile and
+non-volatile memory is important for efficiency, because volatile memory has
+higher access energy and latency than SRAM~\cite[Fig. 4]{nvp}, and generality,
+because devices often have much more FRAM than SRAM (e.g., 64 times
+more~\cite{wolverine}). Hardware is effective, but inapplicable to existing
+devices, making programming and execution models essential to the success of
+intermittent computing. 
+%Despite the limitations, we conjecture that programming
+%and execution models will be a {\em de facto choice for sustaining data
+%consistency} of intermittently-powered devices. Simply put, programming model
+%does not require any hardware investment and is dependent only upon the
+%programmer's skills.
 
-Two recent works~\cite{alpaca,chain} proposed {\em task-based} programming and execution paradigm, that require a programmer to statically decompose a program into tasks.
-A task resembles a function with no callers.
-Task can include arbitrary computation and, upon completion, is guaranteed to have executed {\em atomically} and to have logically {\em committed} memory updates.
-The programmer expresses the flow of control from one task to another and task control-flow may be input-dependent~\cite[Fig.
-4]{chain}.
-An important programming challenge in task-based models is that the length of execution of a task is limited by the total amount of energy that a device can buffer and {\em does not change after compilation}.
-Assuming that input power is negligible compared to operating power, a task will never be able to complete if its execution consumes more energy than the system can buffer.
-On the other hand, transitioning from one task to the next imposes a cost and excessively small tasks may be inefficient.
+Recent work~\cite{alpaca,chain} proposed {\em task-based} programming and
+execution models in which the programmer statically decomposes their program
+into a collection of tasks.  Tasks can include arbitrary computation and, upon
+completion, are guaranteed to have executed {\em atomically}, despite
+arbitrarily-timed power failures. 
+%
+The programmer explicitly expresses task-to-task control flow, which may be
+conditionally input-dependent~\cite[Fig. 4]{chain}. An important programming
+challenge presented by a task-based model is that the length of a software task's execution
+is limited by the fixed, total amount of energy that a device can buffer in
+hardware.  A task's code is static, but the duration of its execution
+may be input-dependent and is difficult to predict.
+%
+Assuming that input power is negligible compared to operating
+power~\cite{wisp}, a task will never be able to complete if its execution
+consumes more energy than the system can buffer, creating the potential for
+long tasks to be irreparably non-terminating.  On the other hand, transitioning
+from one task to another imposes a run time cost and excessively small tasks
+are inefficient.
 
-This brings the following research question: {\bf how to approach the appropriate task size to ensure progress and high performance given an unknown energy buffer size and energy arrival process?} More specifically, we ask two questions: (i) how to eliminate the need for hardware support while enabling the intermittent execution model to dynamically adapt task size {\em after compilation}, and (ii) how to reduce the energy consumption caused by frequent memory swapping operations during task traversals?
+The key research question addressed by this work is to determine how to define
+and execute tasks efficiently and avoid non-termination, given an unknown
+energy buffer size and energy arrival process. In particular, we ask (i) how to
+use software support to dynamically adapt the effective size of a task, while
+still respecting programmer-specified task atomicity, and (ii) how to minimize
+run time and energy consumption while automatically maintaining memory
+consistency during execution?
 
-To answer the above, in this work we develop {\bf \sys}: a new programming and execution model that utilizes two new runtime features: {\em dynamic task coalescing} to efficiently execute tasks without exceeding available energy (answering question (i)) and  {\em volatile memory virtualization} to leverage efficient volatile memory (answering question (ii)).
-\sys's novel memory virtualization mechanism uses a device's SRAM as the working memory during a task's execution, populating the SRAM with a page of data from FRAM as necessary at the start of the task.
-At the end of the task, the SRAM is committed back to FRAM using efficient DMA block copies~\cite{kistler_micro_2006}.
-When a task accesses data outside of a page in SRAM, the task must fetch data from FRAM.
+\TODO{This is the main pitch statement.  we should all agree that this is what
+\sys does.} In this work we develop {\bf \sys}: a new programming and execution
+model that efficiently executes tasks and avoids non-termination across a range
+of energy buffer sizes, without burdening the programmer.
+%
+To accomplish this goal, \sys introduces three new capabilities: {\em automatic
+task compilation} that compiles arbitrary C code into tasks, {\em dynamic task
+coalescing} that efficiently executes tasks without exceeding available energy,
+and {\em volatile memory virtualization} that efficiently leverages both
+volatile and non-volatile memory.  
+%
+\sys's compiler transforms arbitrary C code into a graph of tasks that reflects
+the original program's control- and data-flow constraints.  The compiler
+identifies data shared by multiple of the newly created tasks and instruments
+reads and writes of those data so that \sys's memory virtualization mechanism
+keeps those data consistent.
+%
+\sys's memory virtualization mechanism uses SRAM as working memory, that \sys
+dynamically populates with pages of data from FRAM on demand during a task's
+execution.  When a task ends, \sys commits dirty pages to FRAM using efficient
+DMA block copies, ensuring task atomicity and memory consistency.
+%
+\sys's task coalescing mechanism executes multiple statically defined tasks
+together as a single dynamic task. A programmer can specify small tasks that
+will execute on a device with a small energy buffer. Coalescing tasks allows
+running the same tasks on a device with a larger energy buffer, while avoiding
+run time overheads associated with ending one task and beginning another.  
+%
+
+%~\cite{kistler_micro_2006} This reference will confuse people familiar with DMA.  DMA is well known I think we should cite the MCU datasheet, or use no cite. 
 %
@@
 
-With this paper we provide the following list of contributions.
+To summarize, this work makes the following contributions:
 %
+\TODO{I think these were all much too long, given the depth of the intro and background already.  I am cutting to the barest minimum.}
 \begin{enumerate}

### 2017-07-31 paper/background.tex
added some edits to section 2.  still needs  alot, especially in the task section

@@
-\sys addresses the problem of intermittently executing task-based computations on energy-harvesting hardware, without being limited by inflexible, static task decomposition and without being limited to fully non-volatile main memory.
-In this section, we provide background on energy harvesting systems and the intermittent software execution model.
-We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys addresses.
+\In this section, we provide background on energy harvesting systems and the
+intermittent software execution model. We then discuss the key limitations of
+existing task-based execution models for intermittent computing that \sys
+addresses.
 
@@
 
-Energy harvesting devices are systems that operate using energy extracted from their environment, from sources such as radio waves (RF)~\cite{wisp}, and solar energy.
-By operating entirely using environmental energy, energy-harvesting devices need not use a tethered power supply or battery.
-Instead, devices typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.
+Energy harvesting devices operate using energy extracted from their
+environment, from sources such as radio waves (RF)~\cite{wisp} and solar
+energy. Energy-harvesting devices need not use a tethered power supply or
+battery and instead, typically collect energy into a (super-)capacitor, operate
+when sufficient energy has accumulated, and upon depleting the stored energy,
+turn off and wait to collect more energy. 
+%
+Batteryless operation has a number of important advantages, making intermittent
+computing an important research domain. Supplying power to
+billions~\cite{gartner_iot} of embedded computers using batteries is not
+sustainable. The European Commission estimates that more than 160 kilotons of
+consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
+Batteries are an environmental risk, are fragile, are limited in their number
+of charge/discharge cycles, and may require costly physical maintenance that is
+difficult or impossible deployed (e.g. in space~\cite{kicksat}). By contrast,
+super-capacitors are durable, promising millions of charge/discharge
+cycles~\cite[Sec. I]{ongaro_pwre_2012}. The main limitation in moving to
+capacitor-based energy storage is that capacitor energy density---and
+consequently operating discharge time---is orders of magnitude less than a
+battery.
 
-Batteryless operation has a number of important advantages, making intermittent computing an important research domain.
-Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable.
-The European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
-Batteries constitute a potential environmental hazard because not all batteries are recycled.
-Batteries are fragile and limited in their number of charge/discharge cycles, requiring costly maintenance that is possibly only through physical device access, which may be difficult if devices are deployed in inaccessible places, e.g.
-in space~\cite{kicksat}.
-By contrast, super-capacitors are durable, promising millions of charge/discharge cycles~\cite[Sec.
-I]{ongaro_pwre_2012}.
-The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently: operating discharge time---is orders of magnitude less than a battery.
-\todo{Compress text}{Brandon}
-
-Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible.
-These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
+%Given current technology development, battery-less systems are best suited for
+%very long-term sensing and monitoring where access to recharge is either
+%prohibitive or impossible. These include battery-less image capture and
+%processing~\cite{naderiparizi_rfid_2015}, animal
+%monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}
+%and digestible~\cite{nadeau_naturebio_2017} sensors.
 
@@
 Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
-In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio
-frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
+In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
 Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid).
@@
 
-The biggest technical problem with energy provided from the ambient to battery-less platforms is that harvested energy is not stable and difficult to predict accurately.
-Combined with the fact that energy supply is small, there is little leeway to store enough of energy to guarantee prolonged periods of computation.
-Energy breaks happen every hundred of milliseconds, refer again to e.g.~\cite[Fig.
-1]{mementos}.
-This causes each program running on such platform to restart and requires a fundamentally new model for program execution: \emph{intermittent} (discontinuous) execution model.
-In the conventional (or \emph{continuous}) execution model a power failure is seen as an \emph{exception} that may reset the progress of a program to its start.
-Whereas, in an intermittent execution model a power failure is regarded as a temporary \emph{pause} to the execution consequently prolonging execution time.
-A conceptual implementation of intermittent execution model is to divide the program into smaller parts that guarantee execution within a discharge region and track the state of the program in-between execution times.
-
 \subsection{Intermittent Computing}
 \label{sec:background_consistency}
+Energy-harvesting devices execute software according to the {\em intermittent
+execution model}~\cite{dino,chain,alpaca,ratchet}.  Physically, a device
+charges until a threshold, operates briefly until its energy is depleted, shuts
+down recharges, and repeats the cycle.  Software running on the device executes
+{\em intermittently} because energy is only available between when the
+capacitor reaches its threshold and when it is depleted. An intermittent
+execution is composed of operating periods interspersed with power failures.
+The frequency of power failures in an intermittent execution depends on the
+size of the device's energy storage buffer: a larger buffer allows longer
+operating periods.  Energy-harvesters provide input power that is orders of
+magnitude less than a device's operating power, making recharging negligible
+during operation.
 
-Naturally, content of a volatile memory (i.e.
-run-time stack, global and local variables) and program registers of the microcontroller are erased after each power failure.
-Only what resides in a non-volatile memory, e.g.
-FRAM, persists after system reboot.
-A solution to sustain the progress of the program from the last power failure, instead from the beginning (also denoted as the \emph{Sisyphus} effect~\cite[Sec.
-2]{mementos}) is a periodic checkpointing of a program state~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
-The periodicity, and a position of a checkpoint, as well as the way to combat the \emph{program state consistency} phenomenon (discussed extensively in e.g.~\cite{chain,alpaca}), is the core of the contribution behind each of the proposed checkpoining solution.
-For the completeness of this exposition we will provide our own example demonstrating this phenomenon, viz.
-please refer to Fig.~\ref{fig:code_demo_incosistency} and compare with~\cite[Sec 2.3]{alpaca},~\cite[Sec.
-2.1]{chain}.
+Software has different behavior when executed intermittently than when executed
+with continuous power.  Each power failure clears volatile memory (registers,
+stack, and global variables). Non-volatile memory (e.g., FRAM) persists across
+power failures. On a power failure, control flows back to a prior point in the
+execution. By default, after a power failure control flows to the beginning of
+{\tt main()}; a host of recent work proposed variations of the intermittent
+execution model that restart execution differently. Some systems restore to a
+checkpoint of execution context (i.e., registers and stack) stored in
+non-volatile memory~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
 
@@
 
-Let us take an example of a fast procedure to calculate cyclic redundancy check (CRC) from~\cite{hicks_mibench2_2016}.
-Therein a CRC must be calculated for $\texttt{nBytes}$-long message $\texttt{msg}$ with specified initial reminder $\texttt{Rmnd}$ and constant, $\texttt{Rmnd}$ variable type-dependent, $\texttt{Wdth}$.
-Despite regular, fixed position checkpoints (line 3 in Fig.~\ref{fig:crc_example}) of all variables which are declared as non-volatile (denoted as \texttt{NV}) variable $\texttt{data}$ is calculated incorrectly.
-This is because of a write after read (WAR) hazard of variable $\texttt{Rmnd}$.
-WAR hazard is caused by consecutive write and read in-between two consecutive checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war}) $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more specifically---updated before a new execution loop.
+Data may become inconsistent if an attempt to execute some computation {\em
+writes} to a non-volatile variable, then power fails, then a second attempt to
+re-execute the same computation {\em reads} the value written in the first
+attempt.  Figure~\ref{fig:code_demo_incosistency} illustrates how state can
+become inconsistent in an intermittent execution using cyclic redundancy check
+(CRC) code from MIBench~\cite{hicks_mibench2_2016}. The code computes CRC for
+an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder
+$\texttt{Rmnd}$. Even with programmer defined checkpoints (i.e., line 3 in
+Fig.~\ref{fig:crc_example}) of registers, stack, and non-volatile variables
+(labeled \texttt{NV}), the code may compute  $\texttt{data}$ incorrectly.  The
+code includes a write after read (WAR) hazard of variable $\texttt{Rmnd}$.  WAR
+hazard is caused by consecutive write and read in-between two consecutive
+checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is
+written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war})
+$\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more
+specifically---updated before a new execution loop.
 
-\subsection{(Task-based) Intermittent Computing: Background}
+As the figure illustrates, checkpoint-based systems risk violating the
+consistency of volatile and non-volatile memory~\cite{dino}. To maintain memory
+consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of
+non-volatile variables with the checkpointed execution context. Task-based
+systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail
+next, ensure state is consistent by using programmer, compiler, and runtime
+support.
 
+\subsection{Task-based Intermittent Computing}
+\TODO{This sequential / modular split seems pretty odd to me.
+I think we can just talk about tasks in the sense of ratchet, alpaca, and chain.
+Anything with static (programmer-written or compiler inserted) boundaries applies}
 Intermittent execution models combatting WAR can be classified into: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
 )
-\subsubsection{Baseline---Sequential Execution Model}
 
@@
 \todo{I use the word `checkpoint' for normal checkpoint and task boundary interchangeably---check if it does not cause confusion}{Brandon}
-
-\subsubsection{Alternative---Modular Execution Model}
 

### 2017-07-31 paper/coalescing.tex

@@
-\sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks.
+\sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks to improve execution time at stable energy supply environment.
 
-\subsubsection{Task Merging Algorithms}
+\begin{figure}
+	\centering
+	%\includegraphics[width=\columnwidth]{figures/task_coalescing}
+	\caption{Task coalescing architecture.}
+	\label{fig:}
+\end{figure}
 
-\subsubsection{Power Interrupt Immune Scheduler}
+\subsection{Task Merging Algorithm}
 
-\input{relativeJumpAlgo.tex} %Task jumping algorithm
-
-It utilizes a persistent circular buffer (persistent linked list) to keep the state of a program across power failures.
-\sys provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
-(un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
-
-\paragraph{Fixed Virtual Task Size}
-
-	\begin{algorithm}
-		\caption{Task coalescing mechanism}
-		\label{algo:fixVirtTask}
+\begin{algorithm}[t]
+	\caption{\sys task coalescing mechanism}
+	\label{algorithm:coalescing_algorithm}
 	\scriptsize
-		%\small
 	\begin{algorithmic}[1]
-			\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
-			\State VTS : VT size
-			\State MVTS: maximum VT size \Comment{line added in fixed size approach}
+		\State $\text{VT} \subset \{\sys~\text{tasks set}\}$  \Comment{Virtual task}
+		\State $|\text{VT}|$ \Comment{VT size \todo{Define unit}{Amjad}}
+		\State $\text{VT}_{\max}$ \Comment{Maximum VT size (With upper bound on coalescing)}
 		\vspace{0.1cm}
 
-			\While {$True$}
-				\State $VT \leftarrow VT_{next}$
+		\While {True}
+		\State $\text{VT} \leftarrow \text{VT}_{\text{next}}$ \Comment{\todo{Define $\text{VT}_{\text{next}}$}{Amjad}}
 		\vspace{0.1cm}
-				\While {execute $VT$} 
-					\If { $\text{power failed twice}$ }				
-							\State $VTS--$  
-							\State $ MVTS = VTS $ \Comment{line added in fixed size approach}
+		\While {execute VT} 
+		\If { power failed $n$ times during task execution} \Comment{\todo{Define how it is detected}{Amjad}}				
+		\State $|\text{VT}|:=|\text{VT}|-1$
+		\State $\text{VT}_{\max} := |\text{VT}|$ \Comment{With upper bound on coalescing}
 		\EndIf
@@
 		\vspace{0.1cm}
-				\If {$ \text{All tasks executed}$}
-					\If{$VTS < MVTS$} \Comment{line added in fixed size approach}
-					\State $VTS++$
+		\If {All tasks executed} \Comment{\todo{Define how it is detected and when}{Amjad}}
+		\If{$|\text{VT}| < \text{VT}_{\max}$} \Comment{line added in fixed size approach}
+		\State $|\text{VT}|:=|\text{VT}|+1$
 		\EndIf
@@
 
+The complete algorithm is provided in Algorithm~\ref{algorithm:coalescing_algorithm}. \todo{Fill in with missing information: scheduler description, implementation details, discussion}{Przemek/Amjad}
+
+\subsection{Power Interrupt Immune Scheduler}
+
+\input{relativeJumpAlgo.tex} %Task jumping algorithm
+
+It utilizes a persistent circular buffer (i.e.
+persistent linked list) to keep the state of a program across power failures.
+\sys provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
+(un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
+\todo{Expand this section}{Amjad}
+
 \begin{figure}
 	\centering
-	\includegraphics[width=0.8\columnwidth]{figures/virtualTaskSize.eps}
-	\caption{Size of the virtual task versus the execution time of a dummy application that contains 12 empty tasks.}
+	\includegraphics[width=\columnwidth]{figures/virtualTaskSize.eps}
+	\caption{Execution time of a dummy application containing twelve empty tasks as a function of coalescing factor.
+We observe three times improvement from coalescing until four virtual tasks, while we observe a diminishing return from task coalescing beyond four.
+\todo{Improve the figure: make it narrower, larger font, remove legend, do not capitalize "Task" in X axis, add Y axis label (Execution time (ms))}{Amjad}}
 	\label{fig:virtualTaskSize}
@@
 
-\begin{figure}
-	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/task_coalescing}
-	\caption{Task coalescing architecture.}
-	\label{fig:}
-\end{figure}
-
-The question remains, is the best strategy to coalesce all tasks into a singe one, as the energy becomes fully available, i.e.
-from energy harvesting to continuous power environment.
+The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e.
+from energy harvesting to continuous power environment) is the best one?.
 To answer this question we have implemented a simple program composed of $x$ empty tasks.
-Figure~\ref{fig:virtualTaskSize} shows the benefit of virtualizing the Modular Intermittent Execution Mode (MIEM) in the best case scenario (continuous power supply).
+We execute the same program by increasing coalescing factor by one while the device on which the program was executed was running on a stable power supply (refer to Section~\ref{sec:methodology_evaluation} for details).
+The result is presented in Figure~\ref{fig:virtualTaskSize}.
+We observe a significant increase in execution time as task are coalescent from one to four (three times execution time improvement).
+However, as the coalescing increases beyond four, we observe diminishing returns from coalescing in terms of execution time.
+\todo{Decide whether to place this figure in Evaluation section or leave it here}{Przemek}
 

### 2017-07-31 paper/discussion.tex
Added paragraph to discussion

\subsection{Task Coalescing}
\label{sec:discussion_coalescing}

We speculate that a non-obvious benefit of task coalescing is a secure computation support. That is, if the size of the virtual task increases infinitely, it will eventually exceed the maximum size allowed by the energy reservoir. Therefore the program will never finish executing the virtual task. Unless the intermittently-powered device will be physically close to the harvestable energy source and charging rate is larger than the discharging rate, the computation will never complete. This way we entangle \emph{physical proximity} of the device with \emph{success rate} of the computation.

### 2017-07-31 paper/background.tex
Corrected error in background.tex

@@
-\In this section, we provide background on energy harvesting systems and the
-intermittent software execution model. We then discuss the key limitations of
-existing task-based execution models for intermittent computing that \sys
+In this section, we provide background on energy harvesting systems and the intermittent software execution model.
+We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys
 addresses.
@@
 
-Energy harvesting devices operate using energy extracted from their
-environment, from sources such as radio waves (RF)~\cite{wisp} and solar
-energy. Energy-harvesting devices need not use a tethered power supply or
-battery and instead, typically collect energy into a (super-)capacitor, operate
-when sufficient energy has accumulated, and upon depleting the stored energy,
-turn off and wait to collect more energy. 
-%
-Batteryless operation has a number of important advantages, making intermittent
-computing an important research domain. Supplying power to
-billions~\cite{gartner_iot} of embedded computers using batteries is not
-sustainable. The European Commission estimates that more than 160 kilotons of
-consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
-Batteries are an environmental risk, are fragile, are limited in their number
-of charge/discharge cycles, and may require costly physical maintenance that is
-difficult or impossible deployed (e.g. in space~\cite{kicksat}). By contrast,
-super-capacitors are durable, promising millions of charge/discharge
-cycles~\cite[Sec. I]{ongaro_pwre_2012}. The main limitation in moving to
-capacitor-based energy storage is that capacitor energy density---and
-consequently operating discharge time---is orders of magnitude less than a
-battery.
+Energy harvesting devices operate using energy extracted from their environment, from sources such as radio waves (RF) and solar energy.
+Energy-harvesting devices need not use a tethered power supply or battery and instead, typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.
 
-%Given current technology development, battery-less systems are best suited for
-%very long-term sensing and monitoring where access to recharge is either
-%prohibitive or impossible. These include battery-less image capture and
-%processing~\cite{naderiparizi_rfid_2015}, animal
-%monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015}
-%and digestible~\cite{nadeau_naturebio_2017} sensors.
+Batteryless operation has a number of important advantages, making intermittent computing an important research domain.
+Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable.
+The European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
+Batteries are an environmental risk, are fragile, are limited in their number of charge/discharge cycles, and may require costly physical maintenance that is difficult or impossible deployed (e.g.
+in space~\cite{kicksat}).
+By contrast, super-capacitors are durable, promising millions of charge/discharge cycles~\cite[Sec.
+I]{ongaro_pwre_2012}.
+The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently operating discharge time---is orders of magnitude less than a battery.
+
+%Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible.
+These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
 
@@
 \label{sec:background_consistency}
-Energy-harvesting devices execute software according to the {\em intermittent
-execution model}~\cite{dino,chain,alpaca,ratchet}.  Physically, a device
-charges until a threshold, operates briefly until its energy is depleted, shuts
-down recharges, and repeats the cycle.  Software running on the device executes
-{\em intermittently} because energy is only available between when the
-capacitor reaches its threshold and when it is depleted. An intermittent
-execution is composed of operating periods interspersed with power failures.
-The frequency of power failures in an intermittent execution depends on the
-size of the device's energy storage buffer: a larger buffer allows longer
-operating periods.  Energy-harvesters provide input power that is orders of
-magnitude less than a device's operating power, making recharging negligible
-during operation.
 
-Software has different behavior when executed intermittently than when executed
-with continuous power.  Each power failure clears volatile memory (registers,
-stack, and global variables). Non-volatile memory (e.g., FRAM) persists across
-power failures. On a power failure, control flows back to a prior point in the
-execution. By default, after a power failure control flows to the beginning of
-{\tt main()}; a host of recent work proposed variations of the intermittent
-execution model that restart execution differently. Some systems restore to a
-checkpoint of execution context (i.e., registers and stack) stored in
-non-volatile memory~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
+Energy-harvesting devices execute software according to the {\em intermittent execution model}~\cite{dino,chain,alpaca,ratchet}.
+Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
+Software running on the device executes {\em intermittently} because energy is only available between when the capacitor reaches its threshold and when it is depleted.
+An intermittent execution is composed of operating periods interspersed with power failures.
+The frequency of power failures in an intermittent execution depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods.
+Energy-harvesters provide input power that is orders of magnitude less than a device's operating power, making recharging negligible during operation.
+
+Software has different behavior when executed intermittently than when executed with continuous power.
+Each power failure clears volatile memory (registers, stack, and global variables).
+Non-volatile memory (e.g., FRAM) persists across power failures.
+On a power failure, control flows back to a prior point in the execution.
+By default, after a power failure control flows to the beginning of {\tt main()}; a host of recent work proposed variations of the intermittent execution model that restart execution differently.
+Some systems restore to a checkpoint of execution context (i.e., registers and stack) stored in non-volatile memory~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
 
@@
 
-Data may become inconsistent if an attempt to execute some computation {\em
-writes} to a non-volatile variable, then power fails, then a second attempt to
-re-execute the same computation {\em reads} the value written in the first
-attempt.  Figure~\ref{fig:code_demo_incosistency} illustrates how state can
-become inconsistent in an intermittent execution using cyclic redundancy check
-(CRC) code from MIBench~\cite{hicks_mibench2_2016}. The code computes CRC for
-an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder
-$\texttt{Rmnd}$. Even with programmer defined checkpoints (i.e., line 3 in
-Fig.~\ref{fig:crc_example}) of registers, stack, and non-volatile variables
-(labeled \texttt{NV}), the code may compute  $\texttt{data}$ incorrectly.  The
-code includes a write after read (WAR) hazard of variable $\texttt{Rmnd}$.  WAR
-hazard is caused by consecutive write and read in-between two consecutive
-checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is
-written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war})
-$\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more
-specifically---updated before a new execution loop.
+Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation {\em reads} the value written in the first
+attempt.
+Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
+The code computes CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
+Even with programmer defined checkpoints (i.e., line 3 in Fig.~\ref{fig:crc_example}) of registers, stack, and non-volatile variables (labelled \texttt{NV}), the code may compute $\texttt{data}$ incorrectly.
+The code includes a write after read (WAR) hazard of variable $\texttt{Rmnd}$.
+WAR hazard is caused by consecutive write and read in-between two consecutive checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war}) $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more specifically---updated before a new execution loop.
 
-As the figure illustrates, checkpoint-based systems risk violating the
-consistency of volatile and non-volatile memory~\cite{dino}. To maintain memory
-consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of
-non-volatile variables with the checkpointed execution context. Task-based
-systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail
-next, ensure state is consistent by using programmer, compiler, and runtime
-support.
+As the Figure~\ref{fig:code_demo_incosistency} illustrates, checkpoint-based systems risk violating the consistency of volatile and non-volatile memory~\cite{dino}.
+To maintain memory consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of non-volatile variables with the checkpointed execution context.
+Task-based systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail next, ensure state is consistent by using programmer, compiler, and runtime support.
 
 \subsection{Task-based Intermittent Computing}
-\TODO{This sequential / modular split seems pretty odd to me.
+\label{section:background_task_computing}
+
+%\TODO{This sequential/modular split seems pretty odd to me.
 I think we can just talk about tasks in the sense of ratchet, alpaca, and chain.
 Anything with static (programmer-written or compiler inserted) boundaries applies}
+
 Intermittent execution models combatting WAR can be classified into: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
-)
+\todo{Remove division per sequential/modular split}{Przemek}
 

### 2017-07-31 paper/evaluation.tex
placeholder for figures in evaluation

@@
 \begin{itemize}
-	\item For all apps: show amount of task merged, time to completion, overhead (in terms of what?), energy consumed.
-All experiments for fixed power, at 3 distances with wireless power, compared against Chain
+	\item performance for each application comparing bars for Chain, no coalescing/hand-tasked, coalescing/hand-tasked, no-coalescing/compiler, coalescing/compiler.
+Performance depicted, normalized to one of the configurations (probably Chain).
+This plot will cover most of the eval dimensions that we care about.
+	\item Sensitivity plot---show amount of task merged, time to completion, overhead (in terms of task-size/cap-size sweep (remember we can vary either with approximately the same effect)), energy consumed.
+All experiments for fixed power, at 3 distances with wireless power, compared against Chain for 3 task/cap size configurations
 \end{itemize}
+
+%Point for decision after main figure is ready
+
+%Pull the compiler stuff into its own plot that compares compiler to hand-tasked, maybe only for the coalescing fig.
+This reduces bars in plot 1, but runs the risk of making the compiler sound secondary, which we do not want to do.
+
+%*Report only Chain, coalescing/hand-tasked, coalescing/compiler in the main plot.
+This says that coalescing is the "default configuration" for VIPER.
+Later, in a second figure, we compare coalescing to no coalescing for hand-tasked and compiler as a characterization plot.
 
@@
 \begin{itemize}
-	\item page swapping cost in terms of execution time: compare against Chain and full double buffering approach; show number of page swaps---do the experiment for all applications
+	\item Sensitivity plot---normalized page swapping cost ("bathtub" curve) in terms of execution time: compare against Chain and full double buffering approach; show number of page swaps---do the experiment for all applications
 \end{itemize}

### 2017-07-31 paper/intro.tex

@@
 
-While promising, energy-harvesting devices present a unique design challenge in
-that they operate only {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}. An
-energy-harvesting device buffers energy from its environment, e.g., in a
-capacitor~\cite[Fig.
+While promising, energy-harvesting devices present a unique design challenge in that they operate only {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}. An
+energy-harvesting device buffers energy from its environment, e.g., in a capacitor~\cite[Fig.
 3]{gorlatova_tmc_2013},~\cite[Fig.
@@
 
-Software in an energy-harvesting system operates according to the {\em
-intermittent execution model}~\cite{dino,lucia_snapl_2017}, with bursts of
-execution interrupted by failure periods.  When power fails, a device loses its
-volatile state, i.e., registers, stack, SRAM, and retains its non-volatile
-state, i.e., FRAM, Flash. While capturing periodic
-checkpoints~\cite{mementos,quickrecall} and sleep
-scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve execution
-progress, failures can leave volatile and non-volatile state inconsistent,
-leading to unrecoverable failures~\cite{mspcdino,edb}.
+Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with bursts of execution interrupted by failure periods.
+When power fails, a device loses its volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
+While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb}.
 
-There are two main approaches for dealing with data inconsistency for
-intermittently-powered devices: (i) programming and execution
-models~\cite{dino,ratchet,chain,alpaca} and (ii) architecture
-mechanisms~\cite{hicks_isca_2017,idetic,nvp,tictpl}. While both approaches represent
-progress, each has critical limitations. New architectures require costly
-hardware changes and are inapplicable to today's embedded off-the-shelf
-systems~\cite[Fig.  3]{hicks_isca_2017}~\cite[Fig. 9]{nvp}. New programming
-models and compilers force programmers to change behavior~\cite{chain,ratchet},
-inhibiting adoption.  New memory models may require more accesses to
-non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude
-the use of volatile memory~\cite{ratchet}.  Effective use of both volatile and
-non-volatile memory is important for efficiency, because volatile memory has
-higher access energy and latency than SRAM~\cite[Fig. 4]{nvp}, and generality,
-because devices often have much more FRAM than SRAM (e.g., 64 times
-more~\cite{wolverine}). Hardware is effective, but inapplicable to existing
-devices, making programming and execution models essential to the success of
-intermittent computing. 
-%Despite the limitations, we conjecture that programming
-%and execution models will be a {\em de facto choice for sustaining data
-%consistency} of intermittently-powered devices. Simply put, programming model
-%does not require any hardware investment and is dependent only upon the
-%programmer's skills.
+There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) programming and execution models~\cite{dino,ratchet,chain,alpaca} and (ii) architecture
+mechanisms~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
+While both approaches represent progress, each has critical limitations.
+New architectures require costly hardware changes and are inapplicable to today's embedded off-the-shelf systems~\cite[Fig.
+3]{hicks_isca_2017}~\cite[Fig.
+9]{nvp}.
+New programming models and compilers force programmers to change behavior~\cite{chain,ratchet}, inhibiting adoption.
+New memory models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude
+the use of volatile memory~\cite{ratchet}.
+Effective use of both volatile and non-volatile memory is important for efficiency, because volatile memory has higher access energy and latency than SRAM~\cite[Fig.
+4]{nvp}, and generality, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
+Hardware is effective, but inapplicable to existing devices, making programming and execution models essential to the success of intermittent computing.
 
-Recent work~\cite{alpaca,chain} proposed {\em task-based} programming and
-execution models in which the programmer statically decomposes their program
-into a collection of tasks.  Tasks can include arbitrary computation and, upon
-completion, are guaranteed to have executed {\em atomically}, despite
-arbitrarily-timed power failures. 
+Recent work~\cite{alpaca,chain} proposed {\em task-based} programming and execution models in which the programmer statically decomposes their program into a collection of tasks.
+Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
+The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent~\cite[Fig.
+4]{chain}.
+An important programming challenge presented by a task-based model is that the length of a software task's execution
+is limited by the fixed, total amount of energy that a device can buffer in hardware.
+A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict.
+Assuming that input power is negligible compared to operating power~\cite{wisp}, a task will never be able to complete if its execution consumes more energy than the system can buffer, creating the potential for long tasks to be irreparably non-terminating.
+On the other hand, transitioning from one task to another imposes a run time cost and excessively small tasks are inefficient.
+
+The key research question addressed by this work is to determine how to define and execute tasks efficiently and avoid non-termination, given an unknown energy buffer size and energy arrival process.
+In particular, we ask (i) how to use software support to dynamically adapt the effective size of a task, while still respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while automatically maintaining memory consistency during execution?
+
+\todo{This is the main pitch statement: we should all agree that this is what \sys does}{All} In this work we develop {\bf \sys}: a new programming and execution model that efficiently executes tasks and avoids non-termination across a range of energy buffer sizes, without burdening the programmer.
+To accomplish this goal, \sys introduces three new capabilities:
+
+\begin{itemize}
+	\item {\bf Automatic task compilation:} \sys's compiler transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints.
+The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent.
+\sys complier effectively circumvents the need for labour-intensive manual task generation of the state of the art solutions~\cite{chain,alpaca}.
+\todo{Elaborate on this more and provide numerical results}{Przemek}
 	%
-The programmer explicitly expresses task-to-task control flow, which may be
-conditionally input-dependent~\cite[Fig. 4]{chain}. An important programming
-challenge presented by a task-based model is that the length of a software task's execution
-is limited by the fixed, total amount of energy that a device can buffer in
-hardware.  A task's code is static, but the duration of its execution
-may be input-dependent and is difficult to predict.
+	\item {\bf Dynamic task coalescing:} \sys's memory virtualization mechanism uses SRAM as working memory, that \sys dynamically populates with pages of data from FRAM on demand during a task's execution.
+When a task ends, \sys commits dirty pages to FRAM using efficient DMA block copies, ensuring task atomicity and memory consistency.
+\todo{Once paging is fully implemented--update the paragraph}{Przemek}\todo{Provide numerical results for this section}{Przemek}
 	%
-Assuming that input power is negligible compared to operating
-power~\cite{wisp}, a task will never be able to complete if its execution
-consumes more energy than the system can buffer, creating the potential for
-long tasks to be irreparably non-terminating.  On the other hand, transitioning
-from one task to another imposes a run time cost and excessively small tasks
-are inefficient.
+	\item {\bf Volatile memory virtualization:} \sys's task coalescing mechanism executes multiple statically defined tasks together as a single dynamic task.
+A programmer can specify small tasks that will execute on a device with a small energy buffer.
+Coalescing tasks allows running the same tasks on a device with a larger energy buffer, while avoiding run time overheads associated with ending one task and beginning another.\todo{Provide numerical results for this section}{Przemek}
+\end{itemize}
 
-The key research question addressed by this work is to determine how to define
-and execute tasks efficiently and avoid non-termination, given an unknown
-energy buffer size and energy arrival process. In particular, we ask (i) how to
-use software support to dynamically adapt the effective size of a task, while
-still respecting programmer-specified task atomicity, and (ii) how to minimize
-run time and energy consumption while automatically maintaining memory
-consistency during execution?
+We evaluated \sys end-to-end on a collection of existing benchmark programs~\cite[Sec.
+5]{chain} and new ones (including realistic case study of battery-less embedded sound detector) in a laboratory environment, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexible target a variety of platforms without recompiling the code.
+\todo{add detailed text about evaluation results}{Przemek}
 
-\TODO{This is the main pitch statement.  we should all agree that this is what
-\sys does.} In this work we develop {\bf \sys}: a new programming and execution
-model that efficiently executes tasks and avoids non-termination across a range
-of energy buffer sizes, without burdening the programmer.
-%
-To accomplish this goal, \sys introduces three new capabilities: {\em automatic
-task compilation} that compiles arbitrary C code into tasks, {\em dynamic task
-coalescing} that efficiently executes tasks without exceeding available energy,
-and {\em volatile memory virtualization} that efficiently leverages both
-volatile and non-volatile memory.  
-%
-\sys's compiler transforms arbitrary C code into a graph of tasks that reflects
-the original program's control- and data-flow constraints.  The compiler
-identifies data shared by multiple of the newly created tasks and instruments
-reads and writes of those data so that \sys's memory virtualization mechanism
-keeps those data consistent.
-%
-\sys's memory virtualization mechanism uses SRAM as working memory, that \sys
-dynamically populates with pages of data from FRAM on demand during a task's
-execution.  When a task ends, \sys commits dirty pages to FRAM using efficient
-DMA block copies, ensuring task atomicity and memory consistency.
-%
-\sys's task coalescing mechanism executes multiple statically defined tasks
-together as a single dynamic task. A programmer can specify small tasks that
-will execute on a device with a small energy buffer. Coalescing tasks allows
-running the same tasks on a device with a larger energy buffer, while avoiding
-run time overheads associated with ending one task and beginning another.  
-%
+%\todo{I think these are all much too long, given the depth of the intro and background already. I suggest we cut to the barest minimum.}
 
-%~\cite{kistler_micro_2006} This reference will confuse people familiar with DMA.  DMA is well known I think we should cite the MCU datasheet, or use no cite. 
-%
-\todo{articulate this last point better}{Przemek}
-
-To summarize, this work makes the following contributions:
-%
-\TODO{I think these were all much too long, given the depth of the intro and background already.  I am cutting to the barest minimum.}
-\begin{enumerate}
-%
-\item {\bf Task coalescing mechanism:} \sys's core contribution is the introduction of dynamic task coalescing mechanism.
+%\sys's core contribution is the introduction of dynamic task coalescing mechanism.
 It allows the programmer or a compiler to intuitively decompose the program into small tasks that amortize fixed per-task overheads, yet present no risk of exceeding device energy capacity.
@@
 Consequently, \sys's coalescing mechanism allows a program to execute efficiently across a range of energy buffer sizes, avoiding transition overheads in larger buffers, and ensuring progress in smaller buffers.
-%
-\todo{Provide numerical results for this section}{Przemek}
 
-\item {\bf Compiler-support automatic task generation:} second \sys's core contribution is the ability to split standard C programs into \sys's compatible tasks, effectively circumventing the need for labour-intensive manual task generation of the state of the art solutions~\cite{chain,alpaca}.
-%
-\todo{Elaborate on this more and provide numerical results}{Przemek}
-%
-\item {\bf Evaluation of data swapping strategies:} We propose and explore two data swapping strategies for intermittently-powered systems based on FRAM/SRAM architecture.
+%We propose and explore two data swapping strategies for intermittently-powered systems based on FRAM/SRAM architecture.
 The first strategy is {\em demand paging}, which swaps the SRAM page with a new page from FRAM, buffering the swapped-out page until commit.
-The second strategy is {\em buffered direct access}, which directly reads and writes FRAM relying on dynamic double-buffering to ensure memory is consistent at
-commit. 
-%
-\todo{Once paging is fully implemented--update the paragraph}{Przemek}
-\todo{Provide numerical results for this section}{Przemek}
-%
-\item {\bf Extensive \sys evaluation:} We evaluated \sys end-to-end on a collection of existing benchmark programs~\cite[Sec.
-5]{chain} and new ones (including realistic case study of battery-less embedded sound detector) in a laboratory environment, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexible target a variety of platforms without recompiling the code.
-%
-\todo{add detailed text about evaluation results}{Przemek}
-\todo{decide if we want to make this a contribution paragraph or not (then, put it as a closing paragraph)}{Brandon}
-%
-\end{enumerate}
+The second strategy is {\em buffered direct access}, which directly reads and writes FRAM relying on dynamic double-buffering to ensure memory is consistent at commit.

### 2017-07-31 paper/background.tex
revised figure to have WAR bug

--- 2017-07-31 paper/background.tex
@@
-\In this section, we provide background on energy harvesting systems and the
+In this section, we provide background on energy harvesting systems and the
 intermittent software execution model. We then discuss the key limitations of
@@
 execution. By default, after a power failure control flows to the beginning of
-{\tt main()}; a host of recent work proposed variations of the intermittent
-execution model that restart execution differently. Some systems restore to a
-checkpoint of execution context (i.e., registers and stack) stored in
-non-volatile memory~\cite{mementos,hibernusplusplus,quickrecall,idetic}.
+{\tt main()}. Early work on intermittent computing preserve intermittent
+progress by periodically capturing and restoring to a non-volatile checkpoint
+of the volatile execution context (i.e., registers and
+stack)~\cite{mementos,hibernusplusplus,quickrecall,idetic}
 
@@
 
-Data may become inconsistent if an attempt to execute some computation {\em
-writes} to a non-volatile variable, then power fails, then a second attempt to
-re-execute the same computation {\em reads} the value written in the first
-attempt.  Figure~\ref{fig:code_demo_incosistency} illustrates how state can
-become inconsistent in an intermittent execution using cyclic redundancy check
-(CRC) code from MIBench~\cite{hicks_mibench2_2016}. The code computes CRC for
-an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder
-$\texttt{Rmnd}$. Even with programmer defined checkpoints (i.e., line 3 in
-Fig.~\ref{fig:crc_example}) of registers, stack, and non-volatile variables
-(labeled \texttt{NV}), the code may compute  $\texttt{data}$ incorrectly.  The
-code includes a write after read (WAR) hazard of variable $\texttt{Rmnd}$.  WAR
-hazard is caused by consecutive write and read in-between two consecutive
-checkpoints using $\texttt{checkpoint()}$ function: $\texttt{data}$ variable is
-written \emph{before} (line 10 in Fig.~\ref{fig:crc_example_war})
+Checkpoints of volatile state preserve intermittent progress, but do not ensure
+data consistency~\cite{dino,chain,ratchet}.  Data may become inconsistent if an
+attempt to execute some computation {\em writes} to a non-volatile variable,
+then power fails, then a second attempt to re-execute the same computation
+incorrectly {\em reads} the value written in the first attempt, rather than the
+variable's original value.  The situation occurs when code includes a write
+after read (WAR) dependence between operations that manipulate non-volatile
+variables~\cite{ratchet,dino,alpaca}.
+
+Figure~\ref{fig:code_demo_incosistency} illustrates how state can become
+inconsistent in an intermittent execution using cyclic redundancy check (CRC)
+code from MIBench~\cite{hicks_mibench2_2016}. The code computes CRC for an
+$\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
+Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may
+compute  $\texttt{data}$ incorrectly because  of a WAR dependence between the
+read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.  
+Figure~\ref{fig:crc_example_war} shows an execution of the code in which
+power fails during the second loop iteration
+
  $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more

### 2017-07-31 paper/texFigCode/crc_example_war.tex
revised figure to have WAR bug

@@
 \begin{lstlisting}[style = mystyle]
-$\text{\textcolor{red}{checkpoint()}}$; //Iteration 1: correct execution of the loop body
+$\text{\textcolor{red}{checkpoint()}}$; //Correct execution iteration 1 
 data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
 Rmnd = crcTbl[data] ^ (Rmnd << 8);
+$\text{\textcolor{red}{checkpoint()}}$; //Capture checkpoint before iteration 2
+data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
+Rmnd = crcTbl[data] ^ (Rmnd << 8);      
+$\text{\textcolor{red}{\lightning\hspace{0.5cm}\lightning\hspace{0.5cm}\lightning}\hspace{0.5cm}}$ //Power fails after Rmnd written, before checkpoint()
 
-$\text{\textcolor{red}{checkpoint()}}$; //Iteration 2: break before reminder calculation
+$\textcolor{red}{\circlearrowright}\text{\textcolor{red}{checkpoint()}}$; //Restart iter. 2 from checkpoint; $\text{\textcolor{red}{Rmnd is stale.}}$
 data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8)); 
-$\text{\textcolor{red}{\lightning\hspace{0.5cm}\lightning\hspace{0.5cm}\lightning}\hspace{0.5cm}}\textcolor{red}{\circlearrowright}$ //Power interrupt and restart
+$\textcolor{red}{\times}\text{\textcolor{red}{ERROR: data updated with stale value of Rmnd}}$
 
-$\text{\textcolor{red}{checkpoint()}}$; //Iteration 3: wrong/stale reminder used for data
-data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
 $\ldots$

### 2017-07-31 paper/background.tex

@@
-In this section, we provide background on energy harvesting systems and the
-intermittent software execution model. We then discuss the key limitations of
-existing task-based execution models for intermittent computing that \sys
+In this section, we provide background on energy harvesting systems and the intermittent software execution model.
+We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys
 addresses.
@@
 \subsection{Task-based Intermittent Computing}
-\TODO{This sequential / modular split seems pretty odd to me.
+\label{section:background_task_computing}
+
+%\TODO{This sequential/modular split seems pretty odd to me.
 I think we can just talk about tasks in the sense of ratchet, alpaca, and chain.
 Anything with static (programmer-written or compiler inserted) boundaries applies}
+
 Intermittent execution models combatting WAR can be classified into: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
-)
+\todo{Remove division per sequential/modular split}{Przemek}
 

### 2017-08-01 paper/background.tex

@@
 read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
-Figure~\ref{fig:crc_example_war} shows an execution of the code in which
-power fails during the second loop iteration
-
- $\texttt{Rmnd}$ is read (line 6 in Fig.~\ref{fig:crc_example_war}), or more
-specifically---updated before a new execution loop.
+Figure~\ref{fig:crc_example_war} shows an execution of the code in which power
+fails during the second loop iteration.  The problem arises because line 6 of
+the execution writes $\texttt{Rmnd}$ in the second iteration, but power fails
+before capturing the next checkpoint.  After restarting from the checkpoint,
+the update to $\texttt{data}$ at line 10 reads the value of $\texttt{Rmnd}$
+that was updated before the power failure, producing an incorrect result.
 

### 2017-08-01 paper/texFigCode/crc_example_war.tex

@@
 
-$\textcolor{red}{\circlearrowright}\text{\textcolor{red}{checkpoint()}}$; //Restart iter. 2 from checkpoint; $\text{\textcolor{red}{Rmnd is stale.}}$
+$\textcolor{red}{\circlearrowright}\text{\textcolor{red}{checkpoint()}}$; //Restart from checkpoint; $\text{\textcolor{red}{Rmnd remains updated.}}$
 data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8)); 
-$\textcolor{red}{\times}\text{\textcolor{red}{ERROR: data updated with stale value of Rmnd}}$
+$\textcolor{red}{\times}\text{\textcolor{red}{ERROR: data updated using incorrect Rmnd value}}$
 

### 2017-08-01 paper/figures/taskification_example.tex
placeholder for VIPER API figure

--- 2017-07-25 paper/texFigCode/crc_example.tex
@@
-\documentclass[border={20pt 5pt 35pt 10pt}]{standalone} % Left Bottom Right Top
+\documentclass[border={20pt 5pt 105pt 10pt}]{standalone} % Left Bottom Right Top
 %\documentclass{standalone}
@@
 	showstringspaces = true,             	 		% mark spaces in strings
-	otherkeywords	 = {NV, signal,inos_task,ios_vars, uint8_t},
+	otherkeywords	 = {NV, signal,inos_task,ios_vars, uint8_t, type},
 	%numbers = none,
@@
 \begin{lstlisting}[style = mystyle]
-NV int byte, nBytes, Rmnd, Wdth, char msg[], crcTbl[];
-for (byte = 0; byte < nBytes; ++byte){
-	$\text{\textcolor{red}{checkpoint()}}$; //Inserted checkpoint routine
-	data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
-	Rmnd = crcTbl[data] ^ (Rmnd << 8);}
+_p type pvar1, $\ldots$, _p type pvar$i$;
+
+void task$i$(void){type nvar1, $\ldots$, type nvar$i$
+task_body }
+$\ldots$
+int main(void){
+	init(); //system initialization
+	taskID taskName[] = {{task1}, {task2}, 
+		$\ldots$,{task$i$}}; /list of tasks for state machine
+	addTask(taskName); //create state machine
+	scheduler(); //schedule tasks
+	return 0;}
 \end{lstlisting}

### 2017-08-01 paper/background.tex
makefile patch that should build figs and background edits

@@
 of the volatile execution context (i.e., registers and
-stack)~\cite{mementos,hibernusplusplus,quickrecall,idetic}
+stack)~\cite{mementos,quickrecall}, sometimes using
+hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
 
@@
 Anything with static (programmer-written or compiler inserted) boundaries applies}
+Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the
+programmer to decompose their program into a collection tasks. A task is a
+function with no caller that contains arbitrary computation, sensing, and
+communication.  A programmer describes the sequence in which tasks may execute,
+creating a task graph. The flow of control from task to task happens at
+programmer demarcated points and may conditionally depend on program values.
+Task-based programming abstractions guarantee that tasks exhibit behavior as
+though they executed {\em atomically}, regardless of arbitrary power failures.
+Task-based runtime systems tolerate failures by repeatedly attempting to
+execute a task until it completes. Task-based runtime systems ensure
+task-atomic semantics by ensuring that repeated task executions are idempotent
+(i.e., have the same effect when repeatedly executed).  The key to idempotence
+is ensuring that updates to non-volatile state made by an interrupted, partial
+execution of a task are never visible to a subsequent (potentially completing)
+re-execution of the task.
 
-Intermittent execution models combatting WAR can be classified into: \emph{sequential execution model} (which historically formed a baseline for this stream of research) and \emph{modular execution model} (a potent alternative, on which we base the design decisions of \sys).
-\todo{Remove division per sequential/modular split}{Przemek}
+There are several run time strategies to ensuring task idempotence. One
+approach~\cite{alpaca} is to identify non-volatile data involved in WAR
+dependences in a task (like DINO and Ratchet did for
+checkpoints~\cite{dino,ratchet}), execute the task using buffered copies of
+those data, and commit the buffered copies before completing the task. Another
+strategy is to statically create multiple versions of non-volatile data shared
+by tasks and ensure by construction that no task reads and writes the same
+non-volatile memory location~\cite{chain}.  Regardless of the idempotence
+strategy, the key to task-based execution is that fixed, statically defined
+tasks execute as though atomic, completing after zero or more attempts.
 
-Under this model a program for intermittently-powered device is seen as a one idempotent operational region with one common context.
-Generally, the progress of the program is saved and updated by means of checkpointing, refer again to Section~\ref{sec:background_consistency}.
-Normally, the sequential model relies on a hardware assistant to measure the voltage level in the energy reservoir to place a checkpoint~\cite{mementos,mottola2017harvos,hibernus}.
-The benefit of this model is that it does not require code modification by a programmer.
-However it suffers from significant overhead (refer to experiment result in~\cite[Fig.
-3]{chain}, where it is concluded that each checkpoint causes overhead).
-\todo{I use the word `checkpoint' for normal checkpoint and task boundary interchangeably---check if it does not cause confusion}{Brandon}
-
-At the heart of this model is the concept of an \emph{idempotent task}.
-The idempotent task is a C language function that does not have arguments and does not return a value.
-This task uses a well defined interface to interact with the non-volatile memory.
-Therefore, it tolerates arbitrary number of power interrupts.
-This model, generally, produces less overhead~\cite{chain} and allows multiple applications to run on an intermittently-powered device by interleaving their tasks.
-However, it obviously requires code modification.
-For example, if an algorithm is written according to the continuous execution model it has to be split by a programmer into small tasks to run under the modular intermittent execution model.
+\TODO{Brandon left off here.}
 

### 2017-08-02 paper/intro.tex
cutting text in the intro

@@
 When a threshold quantity of energy accumulates, the device begins operating.
-Harvestable energy sources typically produce relatively very little power compared to a platform's operating power level---compare for instance~\cite[Table III and V]{prasad_comst_2014} with~\cite[Table I]{carrano_cst_2014}.
+Harvestable energy sources typically produce little power compared to a platform's operating power level---compare for instance~\cite[Table III and V]{prasad_comst_2014} with~\cite[Table I]{carrano_cst_2014}.
 A devices operates for a brief period of time and when buffered energy is depleted, shuts down and begins recharging to operate again later.

### 2017-08-02 DMA/sortAlgo/main.c
modifiy the coalescing algorithm to enable throttling up and down

@@ -36,8 +36,6 @@ __p unsigned int j = 1;
 unsigned int in_i, in_j, arr_i, arr_j;
 void task_inner_loop()
 {
-
-
     in_i = RVAR(i);
     in_j = RVAR( j);
     arr_i = RVAR( arr[ in_i ]);
@@ -66,6 +64,10 @@ void task_inner_loop()
 
 void task_outer_loop()
 {
+//    PMMCTL0|= PMMSWBOR;
+
+//    PMMCTL0 = PMMPW|PMMSWBOR;
+
     unsigned int in_i;
     in_i = RVAR(i);
 //    in_i_pt = PVAR(i);

### 2017-08-02 cem_delft/main.c

@@ -278,12 +278,14 @@ void task_print()
     for (i = 0; i < BLOCK_SIZE; ++i) {
         index_t index = RVAR(_v_compressed_data[i].letter);
     }
-    os_jump(0);
+//    os_jump(0);
 }
 
 void task_done()
 {
-    while (1);
+//    while (1);
+    P3OUT |=BIT5;
+    P3OUT &=~BIT5;
 }
 
 
@@ -294,6 +296,8 @@ void init()
       WDTCTL = WDTPW | WDTHOLD;   // Stop watchdog timer
     // Disable the GPIO power-on default high-impedance mode to activate previously configured port settings.
       PM5CTL0 &= ~LOCKLPM5;       // Lock LPM5.
+
+      P3DIR |=BIT5;
 
       CSCTL0_H = CSKEY >> 8;                // Unlock CS registers
   //    CSCTL1 = DCOFSEL_4 |  DCORSEL;                   // Set DCO to 16MHz

### 2017-08-02 develpment_workspace/paging/dataProtec.c

@@ -20,6 +20,7 @@ unsigned int CrntPagHeader = BIGEN_ROM;
 __nv unsigned int __persis_CrntPagHeader = BIGEN_ROM;
 __nv uint8_t pageCommit = 0;
 
+uint16_t dirtyPag=0;
 
 /*
  * Bring a page to its ROM buffer
@@ -196,6 +197,59 @@ void __pagsCommit()
 
     pageCommit = 0;  // enable sending the first page to the temp buffer
 
+}
+
+
+
+
+unsigned int __pageSwap_w(unsigned int * varAddr)
+{
+    //**// This page swap is caused by write operation //**//
+
+    //2// Find the requested page
+    unsigned int ReqPagTag;
+    unsigned int ReqPagTag_dirty = (unsigned int) varAddr;
+    unsigned int __temp_pagSize = BIGEN_ROM+PAG_SIZE ; // the upper limit of the first page
+    // TODO we are not checking if the var is not in any page !
+    unsigned char idx=0;
+
+    // Search the page in the buffer
+    while( (ReqPagTag = __pagsInTemp[idx]) != 0)
+    {
+        if ( (ReqPagTag_dirty >= ReqPagTag) && (ReqPagTag_dirty < (ReqPagTag+PAG_SIZE) ) )
+                {
+                    if(dirtyPag)
+                    {
+                        __sendPagTemp( CrntPagHeader );  // check the buffer before insertting to it
+                    }
+                    // we found the page
+                    __bringPagTemp( ReqPagTag );
+
+                    goto PAG_IN_TEMP;
+                }
+        idx++;
+    }
+
+    //TODO optimize this search (maybe with a switch statement )
+    // search the page in the ROM if it is not in the buffer
+    while( ! (ReqPagTag_dirty < __temp_pagSize) )  // if the var is not with the page
+    {
+        __temp_pagSize +=   PAG_SIZE;               // move to next page
+    }
+
+    ReqPagTag = __temp_pagSize-PAG_SIZE;
+
+    if(dirtyPag)
+    {
+        __sendPagTemp( CrntPagHeader );
+    }
+    __bringPagROM(ReqPagTag);
+
+PAG_IN_TEMP:
+
+     dirtyPag=1;   // we want to write to this page show it is dirty
+     CrntPagHeader = ReqPagTag;
+    return 0;
 }
 
 

### 2017-08-02 develpment_workspace/paging/dataProtec.h

@@ -50,10 +50,12 @@ void __bringPagTemp(unsigned int pagTag);
 void __bringPagROM(unsigned int pagTag);
 void __sendPagROM(unsigned int pagTag);
 unsigned int __pageSwap(unsigned int * varAddr);
+unsigned int __pageSwap_w(unsigned int * varAddr);
 void __pagsCommit();
 void __bringCrntPagROM();
 
 extern unsigned int CrntPagHeader;  // Holds the address of the first byte of a page
+extern uint16_t dirtyPag;
 
 // Memory access interface
 
@@ -90,13 +92,6 @@ extern unsigned int CrntPagHeader;  // Holds the address of the first byte of a 
                                     }
 
 
-#define RVAR(var)   (\
-                        (  __IS_VAR_IN_CRNT_PAG(var) ) ? \
-                        ( * __VAR_PT_IN_RAM(var) ):\
-                        ( *(  (__typeof__(var)*) ( (( __pageSwap(&(var)) +  __VAR_ADDR(var) ) - CrntPagHeader)  + RAM_PAG  )  )  )\
-                    )
-
-
 #define PVAR(var)   (\
                         (  __IS_VAR_IN_CRNT_PAG(var) ) ? \
                         ( __VAR_PT_IN_RAM(var) ):\
@@ -104,10 +99,21 @@ extern unsigned int CrntPagHeader;  // Holds the address of the first byte of a 
                     )
 
 
-#define PPVAR(wvar, rvar)  __typeof__(rvar) __var##__var = (*PVAR(rvar)); \
-                          (*PVAR(wvar)) = __var##__var
+#define RVAR(var)   ( * PVAR(var))
 
 
+
+#define PPVAR(wvar, rvar)  __temp_temp = (*PVAR( (rvar) )); \
+                          (*PVAR( (wvar) )) =  __temp_temp
+
+
+#define __WP(var)   (\
+                        ( ( __IS_VAR_IN_CRNT_PAG(var) ) ) ? \
+                        ( (dirtyPag = (uint16_t)__VAR_PT_IN_RAM(var)) ):\
+                        ( (  (__typeof__(var)*) ( (( __pageSwap_w(&(var)) +  __VAR_ADDR(var) ) - CrntPagHeader)  + RAM_PAG  )  )  )\
+                    )
+
+#define WP(var)   (*__WP(var))
 
 #endif /* INCLUDE_DATAPROTEC_H_ */
 

### 2017-08-02 develpment_workspace/pagingAnalysis/main.c

@@ -5,10 +5,10 @@
 __p uint32_t rVar_p2 = 0x22222222;
 __p uint8_t fillPage_2[999] ={0};
 
-__p uint32_t rVar_p1 = 0x11111111;
 __p uint8_t fillPage_1[999] ={0};
+__p uint32_t rVar_p1 = 0x11111111;
 
-
+uint32_t __temp_temp=0;
 
 
 volatile uint16_t res1 =0 ;
@@ -23,6 +23,44 @@ int main(void) {
 
     //******// Check the correctness of the address translation
     __bringCrntPagROM();            // bring the default page
+
+    WP(rVar_p1) = 1;
+    WP(rVar_p2) = 2;
+    __sendPagTemp( CrntPagHeader );
+    __pagsCommit();
+
+    WP(rVar_p1)++;
+
+    __sendPagTemp( CrntPagHeader );
+    __pagsCommit();
+    WP(rVar_p2) = 5;
+    WP(rVar_p1) = *PVAR(rVar_p2);
+
+    __sendPagTemp( CrntPagHeader );
+    __pagsCommit();
+
+    RVAR(rVar_p1)++;
+
+    RVAR(rVar_p1)++;
+    __sendPagTemp( CrntPagHeader );
+    __pagsCommit();
+
+
+    RVAR(rVar_p1) = 0;
+    RVAR(rVar_p1)+=8;
+    __sendPagTemp( CrntPagHeader );
+    __pagsCommit();
+
+
+    RVAR(rVar_p2) = 2;
+    RVAR(rVar_p1) = 0;
+    RVAR(rVar_p1) =     RVAR(rVar_p2) + 8;
+    __sendPagTemp( CrntPagHeader );
+    __pagsCommit();
+
+    WVAR(rVar_p1, rVar_p1+1 );
+    __sendPagTemp( CrntPagHeader );
+    __pagsCommit();
 
     P3OUT |= BIT5;
     PPVAR(rVar_p1, rVar_p2);          // write a value into the RAM page

### 2017-08-02 develpment_workspace/var_32bits/main.c

@@ -1,7 +1,7 @@
 #include <msp430.h> 
 #include <stdint.h>
 #include <ipos.h>
-
+#include "printf.h"
 
 __p uint16_t volatile arr16_1[512] = {0};
 __p uint16_t volatile arr16_2[512] = {0};
@@ -46,10 +46,9 @@ void init()
   P3DIR |=BIT5;
   P1DIR |=BIT0;
 }
-
+volatile uint16_t x = 90;
 int main(void) {
     init();
-
        taskId tasks[] = {  {task_test32_1,0},
                            {task_test32_2,0},
                            {task_test16_1,0},

### 2017-08-02 paper/compiler.tex
pushing initial draft

@@
 
-The compiler consists of two parts called \emph{Passes}:
+The compiler consists of two {\em passes} that runs on LLVM Intermediate Representation (IR). LLVM IR is a intermediate language that Clang, a popular compiler frontend for C, generates.
+The two passes are as follows:
 %
@@
 
-In the first pass \sys compiler decides on the task size.
-\sys' task size metric is defined by the number of LLVM Intermediate Representation (IR) instructions.
-That is, \sys tries to make every task to have roughly the same amount of LLVM IR instructions.
+In the first pass \sys compiler decides where to put task boundaries. By definition, a \sys task should adhere to the following correctness constraints (C1-2).
 
-\begin{figure}
-	\centering
-	%\includegraphics[width=\columnwidth]{figures/compiler_graph}
-	\caption{\sys compiler flow diagram.\todo{Draw compiler flow graph diagram}{Kiwan}}
-	\label{fig:compiler_graph}
-\end{figure}
+\begin{enumerate}[label={\bf C\arabic*:}]
+\item{No task should ever use up more energy than what the capacitor of the system can provide.} 
+\item{Task should be single-entry.}
+\end{enumerate}
 
-\sys compiler first builds a control flow graph where (i) vertice is a \emph{each basic block} and (ii) a unidirectional edge is a \emph{branch from one basic block to another}, see Figure~\cite{fig:compiler_graph}.
-Each vertice has a {\em price}, which is the number of LLVM IRs inside the vertice.
-When forming a task, each task should contain a basic block so that the sum of price of all blocks within each tasks should be roughly the same \todo{Provide more details on this algorithm}{Kiwan}.
-Price makes the workload done by each tasks roughly the same and make no task exceed the energy bound.\todo{Specify how do you decide on the minimum task size/energy reservoir}{Kiwan}
+Also for the system to be efficient, the compiler should try to optimize the following points (O1-3).
 
-Each edge has a {\em penalty}, which indicates the penalty of placing a task boundary at the specific edge.
-Intuitively, when a task boundary is placed on a frequently visited edge (such as a back edge of a loop) the system suffers from higher runtime overhead.
-As we are unaware of a way to statically estimate the number of visit at each edge, we therefore initially set every edge penalty to one.
-When analyzing loops, penalty of edges within loops are multiplied by loop count \todo{Change the penalty to be one or 1$\times$loopcount}{Kiwan}.
-After constructing a graph and assigning price and penalty, the compiler analyzes loops \todo{Explain what do you mean by analyzing loops}{Kiwan}.
-Except those inside a loop which is set as the loop count \todo{Explain this statement}{Kiwan}.
-Unbounded loops are conservatively considered as very large \todo{Explain this statement}{Kiwan}.
+\begin{enumerate}[label={\bf O\arabic*:}]
+\item{Tasks should be roughly the same size for efficient coalescing.}
+\item{The number of protected variables should be minimized.}
+\item{Task boundary should be visited as least as possible, because there is overhead on boundaries.}
+\end{enumerate}
+
+Analysis pass tries to meet constraints C1-2 and O1-3 by the following steps.
+
+\paragraph{Building Graph.}
+
+\sys compiler first builds a control flow graph where (i) each vertices is a \emph{basic block} and (ii) each unidirectional edge is a \emph{branch from one basic block to another} (Figure~\ref{fig:comp_2}).
+All function calls are inlined in the main function prior to building graph.
+
+Each vertice has a {\em price}, which represents the energy usage of the basic block.
+We used the number of LLVM IRs inside the vertice as a indicator to energy usage.
+When forming a task, each task should contain a basic block so that the sum of price of all blocks within each tasks should be roughly the same (O1) and does not exceed given energy budget (C1).
+For our implementation, the programmer should provide the energy budget in compile time as a form of number of LLVM IR instruction that the hardware can execute safely in one energy cycle.
+The number can be found by experiment.
+
+Each edge holds the information about the {\em live variables} across the edge, and if the edge is within any loop, the {\em loop count}.
+By this information, the compiler tries to minimize the number of protected variable (O2), because placing boundary at an edge means the live variables across that edge needs to be protected.
+Also the compiler avoids placing boundaries at an edge that is within a loop whose loopcount is big (O3).
+
+The system also calculate dominance relation between vertices. This information is used to ensure task's single-entry
+property (C2).
+
+\paragraph{Analyzing Loops.}
+
+After constructing the graph and assigning all the information discussed above to each vertices and edges, 
+the compiler analyze loops. The compiler identifies all loops and multiplies the loop count to the price of a vertices
+within the loop. This is because if the loop is inside a task, it will use up energy that is roughly multiplied by the 
+loop count compared to code that is outside a loop. If the resulting price of a loop is too large, the compiler
+places boundary at the back edge of the loop, and reverts the change to the price of the vertices inside. It means
+if the loop is too large to fit in one task, each iteration should be a seperate task execution.
+
+If we have an unbounded loop (such as {\tt while} statement), the compiler conservatively assumes the loop count to be very large.
+The programmer can prevent this by annotating maximum possible loop count of such unbounded loop if the programmer has the information.
+
+\paragraph{Chopping Into Tasks.}
+
+After initially placing boundaries with analyzing the loops, we place additional boundaries if some tasks are still too large.
+We start from the beginning of the main function, and traverse through the graph in a depth-first search manner. If we find
+any part of the graph that is too long to fit in the energy bound, it means we have to place boundary somewhere within. The subset of graph
+that needs boundary placement within is called a {\em boundary-candidate}.
+
+As discussed above, when placing boundary we take three optimization criteria into account: (i) tasks should be roughly the same size (O1),
+(ii) protected variable should be minimized (O2), and (iii) boundary within loop should be avoided. We introduce a heuristic that
+consider all these criteria and choose the local-optimum. The heuristic calculates the {\em score} of each edge inside the boundary-candidate
+as follows and choose the edge with the highest score:
+
+\begin{center}
+$score = \frac{w_{0}\times normIRcount - w_{1}\times normPvarChange}{loopCount}$
+\end{center}
+
+Here, $normIRcount$ is a normalized IR count that the task will end up having if we place boundary at the specific edge. Similarly, $normPvarChange$ is
+the number of protected variable that will be newly introduced by placing boundary at the edge. Both are normalized within the boundary-candidate, since
+usually IR count is much larger and has higher variance, thus has dominating effect if not normalized. It is better to have higher IR count within the task
+since it means the task size is similar to the size given by the programmer (C1). Change in protected variable is better to be small (C2), so we have minus sign
+for the $normPvarChange$. The effect of the two can be weighted by $w_{0}$ and $w_{0}$. The study of the effect of different weights will be presented in Section~{}.
+If the edge is within a loop, the score is divided by the number of loop count. It will result in not placing boundary within loops 
+unless other parts of the formula really pays off (C3). Note that it never make decision that will end up having more IR count than the programmer specified
+because at the beginning boundary-candidate is chosen to have less IR than the programmer-specified number.
+
+Figure~\ref{fig:comp_2} shows example of such boundary placement on graph. The first three vertices will become one task, and the rest will become another task (See Figure~\ref{fig:comp_3}).
+
+\paragraph{Ensuring Single Entry.}
+
+After placing boundary, the compiler ensures that the newly made task is single-entried by checking dominanace relation. That is, every vertice within
+a task should be dominated by the entry vertice of the task. If this rule is violated, the compiler places additional boundaries to make sure that every
+task is single-entried.
 
@@
 
-Compiler transform pass is also implemented using LLVM. \todo{Fill in this section with details}{Kiwan}
+\paragraph{Declaring and Migrating Codes to Tasks.}
+
+After the boundary placement is finished, the analysis pass passes the information to transform pass.
+The transform pass makes new task, traverse through the graph, and copies every code within the vertices it
+encounters to the new task until it meets the task boundary. At the task boundary, it makes another task and
+starts copying again, and repeat this action until every code is moved to form a task.
+
+\paragraph{Renaming Variables.}
+
+On copying the codes, the variables should be renamed. If a variable is alive across multiple tasks,
+it needs to become a global-protected variable, and every access to the variable should be redirected.
+If a variable's live range is local to the task, it does not need to go through such renaming.
+For example, in Figure~\ref{fig:comp_2}, boundary is place
+within the live range of variable {\tt b} (Note that the red cross overlaps the blue arrow). Thus variable {\tt b}
+needs to become a protected variable {\tt p\_b} and every access to {\tt b} should be redirected to {\tt p\_b} (Figure~\ref{fig:comp_3}).
+On the other hand, {\tt a}'s live range is always local to a single task. In such case, it need not become a protected variable.
+Note, however, still new declaration of {\tt a} should be inserted in {\tt task1} (Figure~\ref{fig:comp_3}).
+As an optimization, constants and read-only variable never becomes protected.
+
+\begin{figure}
+	\centering
+	\subfloat[Plain C code.]{\includegraphics[width=0.3\columnwidth]{figures/compiler1.pdf}\label{fig:comp_1}}
+	\subfloat[Boundary placement.]{\includegraphics[width=0.6\columnwidth]{figures/compiler2.pdf}\label{fig:comp_2}}
+	\subfloat[Code Transformation.]{\includegraphics[width=0.3\columnwidth]{figures/compiler3.pdf}\label{fig:comp_3}}
+	\caption{Example of how \sys compiler works.}
+	\label{fig:compiler_overview}
+\end{figure}
+
+\paragraph{Inserting os\_jump Calls.}
+
+After renaming is done, the compiler detects the broken {\tt branch} instructions which tries to branch to a
+basic block that is inside different task and change it to an {\tt os\_jump} call (Figure~\ref{fig:comp_3}).
+Actually {\tt os\_jump} call takes offset number as an argument, but in the figure we showed the name of the
+next task to be executed for readability. In reality the compiler statically calculates and passes the offset.
+
+\paragraph{Dealing With PHI Nodes.}
+
+Whenever more than two different control flow tries to write different value in the same memory location,
+LLVM makes a unique IR called {\em PHI Node} when reading the memory location~\cite{citationneeded}.
+However if there is a boundary between the memory write and the PHI Node, the compiler cannot work.
+Thus, on each of this occurance, we declare a special protected variable and make the write relocated to the
+protected variable. Then the PHI Node becomes simply read to the protected variable.
+
+\paragraph{Inserting API Call for Paging.}
+
+To run \sys, every read or write to a protected variable should call related API for page swapping and dirty page
+marking. The compiler conservatively inserts such calls before a {\em possible} read or write to the protected variable.
+We use the basic pointer alias feature provided by the LLVM library. We are being conservative, so it is always correct,
+albeit sometimes inefficient.
+
+\paragraph{Creating New Main Function and Cleaning Unnecessary Codes.}
+
+After task generation is done, the compiler generates new main function and peppers the system with
+necessary codes that is needed by the system (such as code for envoking {\tt os\_sceduler}).
+Also it removes codes that became dead (such as the body of a function call, for all the function calls are now inlined).
 
@@
 \begin{enumerate}
-	\item \textbf{Conservative handling of unbounded code sections:} When a C program contains an unbounded loop or pointer, the \sys compiler makes a conservative decision.
-\todo{Expand on this statement (define `conservative decision') by providing example and more details on this operation}{Kiwan}.
-This means that for performance-critical implementations, the programmer is advised to manually split the code into tasks.
+	\item \textbf{Conservative handling of unbounded code sections:} When a C program contains an unbounded loop, the \sys compiler makes a conservative decision of regarding its loop count to be very large.
+However, if a programmer annotates the expected loop count manually the overhead can be relieved.
+	\item \textbf{Conservative handling of pointers:} When a C program contains pointer, due to the limitation of pointer alias, the compiler makes the conservative decision.
+That is, it may consider a variable's live range longer than actual, or it may insert API call for page swap even when it is unneccesary.
+However, these overhead can be reduced by better pointer aliasing algorithm.
+	\item \textbf{Function call inlining:} \sys compiler inlines all function call prior to analysis for simplicity.
+However it can introduce code bloat.
+Better way would be instead convert large-frequently visited functions to a seperate task and make the function call a jump to the task.
+However we did not implemented such feature for simplicity of the system.
 	%
-	\item \textbf{LLVM IR instruction as a workload metric:} Intuitively, the number of LLVM IR instructions correlates with the total workload performed by the task.
-However, we need to be aware that is nevertheless a crude approach of estimating task workload---actual machine instructions are what counts and they are not exactly equal to the LLVM IR instruction.
-Also, different instructions have different power usage, use different cycles, etc.
-However, the size of each task need not be exactly the same and it is sufficient to make sure that all of the tasks are small enough to fit in to the energy budget.
+	\item \textbf{Using LLVM IR instruction count to estimate energy usage:} The compiler uses the number of LLVM IR count as an indicator for energy usage in trying to meet correctness constraint C1 and optimization constraint O1.
+Intuitively, the number of LLVM IR correlates with the energy usage since more instructions will lead to more energy usage.
+However, we need to be aware that is nevertheless a crude approach of estimating energy consumption---actual machine instructions are what counts and they are not exactly equal to the LLVM IR.
+Also, different instructions have different power usage, and even same instruction has different energy usage from time to time (for example cache hit and miss makes even the same {\tt load} instruction to use different amount of energy).
+
+		However, the energy usage of each task need not be exactly the same to what programmer indicated, and it is sufficient to make sure that all of the tasks are small enough to fit in to the energy budget.
 Therefore we conjecture that from the \emph{implementation simplicity perspective} of \sys's compiler counting the LLVM IR instruction is sufficient.
+The programmer can always ensure correctness constraint C1 by conservatively providing IR count in compile time that is empirically proven to be safe.
+Also, since the estimation of energy usage of each basic block is perpendicular to the compiler itself, it is simple to plug in a better energy model if there is one.
+	\item \textbf{Limitation of the boundary placement heuristics:} The boundary placement stategy is a heuristic that chooses the local optimum.
+However, selecting every local optimum might not be the global optimal solution.
+We can instead do a random search multiple times to find a global optima.
+Since current heuristic is already showing good result that is comparable to human-written code~\ref{}, we leave such optimization as a future work.
 \end{enumerate}
-
-In the light of this design decisions we will assess the compiler overhead in Section~\ref{sec:results_compiler}.

### 2017-08-03 paper/background.tex
Kiwan's compiler section revised - really nice text Kiwan

@@
 \label{sec:background_consistency}
-Energy-harvesting devices execute software according to the {\em intermittent
-execution model}~\cite{dino,chain,alpaca,ratchet}.  Physically, a device
-charges until a threshold, operates briefly until its energy is depleted, shuts
-down recharges, and repeats the cycle.  Software running on the device executes
-{\em intermittently} because energy is only available between when the
-capacitor reaches its threshold and when it is depleted. An intermittent
-execution is composed of operating periods interspersed with power failures.
-The frequency of power failures in an intermittent execution depends on the
-size of the device's energy storage buffer: a larger buffer allows longer
-operating periods.  Energy-harvesters provide input power that is orders of
-magnitude less than a device's operating power, making recharging negligible
-during operation.
 
-Software has different behavior when executed intermittently than when executed
-with continuous power.  Each power failure clears volatile memory (registers,
-stack, and global variables). Non-volatile memory (e.g., FRAM) persists across
-power failures. On a power failure, control flows back to a prior point in the
-execution. By default, after a power failure control flows to the beginning of
-{\tt main()}. Early work on intermittent computing preserve intermittent
-progress by periodically capturing and restoring to a non-volatile checkpoint
-of the volatile execution context (i.e., registers and
-stack)~\cite{mementos,quickrecall}, sometimes using
-hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
+Energy-harvesting devices execute software according to the {\em intermittent execution model}~\cite{dino,chain,alpaca,ratchet}.
+Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
+Software running on the device executes {\em intermittently} because energy is only available between when the
+capacitor reaches its threshold and when it is depleted.
+An intermittent execution is composed of operating periods interspersed with power failures.
+The frequency of power failures in an intermittent execution depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods.
+Energy-harvesters provide input power that is orders of magnitude less than a device's operating power, making recharging negligible during operation.
+
+Software has different behavior when executed intermittently than when executed with continuous power.
+Each power failure clears volatile memory (registers, stack, and global variables).
+Non-volatile memory (e.g., FRAM) persists across power failures.
+On a power failure, control flows back to a prior point in the execution.
+By default, after a power failure control flows to the beginning of {\tt main()}.
+Early work on intermittent computing preserve intermittent progress by periodically capturing and restoring to a non-volatile checkpoint of the volatile execution context (i.e., registers and stack)~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
 
@@
 
-Checkpoints of volatile state preserve intermittent progress, but do not ensure
-data consistency~\cite{dino,chain,ratchet}.  Data may become inconsistent if an
-attempt to execute some computation {\em writes} to a non-volatile variable,
-then power fails, then a second attempt to re-execute the same computation
-incorrectly {\em reads} the value written in the first attempt, rather than the
-variable's original value.  The situation occurs when code includes a write
-after read (WAR) dependence between operations that manipulate non-volatile
-variables~\cite{ratchet,dino,alpaca}.
+Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}.
+Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value.
+The situation occurs when code includes a write after read (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}.
 
-Figure~\ref{fig:code_demo_incosistency} illustrates how state can become
-inconsistent in an intermittent execution using cyclic redundancy check (CRC)
-code from MIBench~\cite{hicks_mibench2_2016}. The code computes CRC for an
-$\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
-Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may
-compute  $\texttt{data}$ incorrectly because  of a WAR dependence between the
-read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
+Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using cyclic redundancy check (CRC) code from MIBench~\cite{hicks_mibench2_2016}.
+The code computes CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
+Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of a WAR dependence between the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
 Figure~\ref{fig:crc_example_war} shows an execution of the code in which power
-fails during the second loop iteration.  The problem arises because line 6 of
-the execution writes $\texttt{Rmnd}$ in the second iteration, but power fails
-before capturing the next checkpoint.  After restarting from the checkpoint,
-the update to $\texttt{data}$ at line 10 reads the value of $\texttt{Rmnd}$
-that was updated before the power failure, producing an incorrect result.
+fails during the second loop iteration.
+The problem arises because line 6 of the execution writes $\texttt{Rmnd}$ in the second iteration, but power fails before capturing the next checkpoint.
+After restarting from the checkpoint, the update to $\texttt{data}$ at line 10 reads the value of $\texttt{Rmnd}$ that was updated before the power failure, producing an incorrect result.
 
-As the figure illustrates, checkpoint-based systems risk violating the
-consistency of volatile and non-volatile memory~\cite{dino}. To maintain memory
-consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of
-non-volatile variables with the checkpointed execution context. Task-based
-systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail
-next, ensure state is consistent by using programmer, compiler, and runtime
-support.
+As the figure illustrates, checkpoint-based systems risk violating the consistency of volatile and non-volatile memory~\cite{dino}.
+To maintain memory consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of non-volatile variables with the checkpointed execution context.
+Task-based
+systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail next, ensure state is consistent by using programmer, compiler, and runtime support.
 
@@
 Anything with static (programmer-written or compiler inserted) boundaries applies}
-Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the
-programmer to decompose their program into a collection tasks. A task is a
-function with no caller that contains arbitrary computation, sensing, and
-communication.  A programmer describes the sequence in which tasks may execute,
-creating a task graph. The flow of control from task to task happens at
-programmer demarcated points and may conditionally depend on program values.
-Task-based programming abstractions guarantee that tasks exhibit behavior as
-though they executed {\em atomically}, regardless of arbitrary power failures.
-Task-based runtime systems tolerate failures by repeatedly attempting to
-execute a task until it completes. Task-based runtime systems ensure
-task-atomic semantics by ensuring that repeated task executions are idempotent
-(i.e., have the same effect when repeatedly executed).  The key to idempotence
-is ensuring that updates to non-volatile state made by an interrupted, partial
-execution of a task are never visible to a subsequent (potentially completing)
-re-execution of the task.
 
-There are several run time strategies to ensuring task idempotence. One
-approach~\cite{alpaca} is to identify non-volatile data involved in WAR
-dependences in a task (like DINO and Ratchet did for
-checkpoints~\cite{dino,ratchet}), execute the task using buffered copies of
-those data, and commit the buffered copies before completing the task. Another
-strategy is to statically create multiple versions of non-volatile data shared
-by tasks and ensure by construction that no task reads and writes the same
-non-volatile memory location~\cite{chain}.  Regardless of the idempotence
-strategy, the key to task-based execution is that fixed, statically defined
-tasks execute as though atomic, completing after zero or more attempts.
+Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into a collection tasks.
+A task is a function with no caller that contains arbitrary computation, sensing, and communication.
+A programmer describes the sequence in which tasks may execute,
+creating a task graph.
+The flow of control from task to task happens at programmer demarcated points and may conditionally depend on program values.
+Task-based programming abstractions guarantee that tasks exhibit behavior as though they executed {\em atomically}, regardless of arbitrary power failures.
+Task-based runtime systems tolerate failures by repeatedly attempting to execute a task until it completes.
+Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent (i.e., have the same effect when repeatedly executed).
+The key to idempotence is ensuring that updates to non-volatile state made by an interrupted, partial execution of a task are never visible to a subsequent (potentially completing) re-execution of the task.
 
+There are several run time strategies to ensuring task idempotence.
+One approach, utilized by Alpaca~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{rachet} did for checkpoints), execute the task using buffered copies of those data, and commit the buffered copies before completing the task.
+Another strategy is to statically create multiple versions of non-volatile data shared by tasks and ensure by construction that no task reads and writes the same non-volatile memory location~\cite{chain}.
+Regardless of the idempotence strategy, the key to task-based execution is that fixed, statically defined tasks execute as though atomic, completing after zero or more attempts.
 \TODO{Brandon left off here.}

### 2017-08-03 paper/evaluation.tex
Kiwan's compiler section revised - really nice text Kiwan

@@
 \begin{itemize}
-	\item comparison between manual (Chain code) and automatic taskification: in terms of code SLOCs, number of tasks, completion time (Chain), memory use (other metrics required?)
+	\item comparison between manual (Chain code) and automatic taskification: in terms of code SLOCs, number of tasks, completion time (Chain), memory use (other metrics required?), show number of LLVM instructions per energy cycle
 	\item assessment of correctness (through manual inspection only?).
@@
 \label{sec:results_memory_management}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
+	\caption{Initial figure about page size overhead. \todo{Figure placeholder---will be replaced with final one}{Amjad}}
+	\label{fig:IPOSPerformance}
+\end{figure}
 

### 2017-08-03 DMA/emtyTasks/main.c
working on coalescing

@@ -133,10 +133,11 @@ void task22()
 }
 
 
-
+__nv volatile uint16_t dummy = 10;
 int main(int argc, char const *argv[])
 {
     init();
+    dummy = 20;
     taskId tasks[] = { {task1,0},
                        {task2,0},
                        {task3,0},

### 2017-08-03 DMA/ipos-gcc-dma/ipos/psched.c

@@ -1,274 +1,233 @@
 #include <psched.h>
 #include <msp430fr5969.h>
 
-//#define __KEY  0xAD
+
 #define COMMITTING      1
 #define COMMIT_FINISH   0
 
 #define JUMP2()  if(!__jump)  \
-                    {   \
-                            __current_task_virtual = (unsigned int *) *(__current_task_virtual + NEXT_OFFSET_PT) ;     /* soft transition */  \
-                    }else{  \
-                        unsigned int totJumpSize = (*(__current_task_virtual + BLOCK_OFFSET_PT) + __jump_by);   \
-                            if(totJumpSize > __totNumTask)  \
-                            {   \
-                                int dis  = totJumpSize - __totNumTask;  \
-                                while( dis  > __totNumTask) \
-                                {   \
-                                    dis = dis - __totNumTask;   \
-                                }   \
-                                dis = dis - (totJumpSize - __jump_by);  \
-                                __current_task_virtual  += (dis + dis + dis) ;  \
-                            }else{  \
-                                __current_task_virtual  += ( __jump_by+__jump_by  + __jump_by )  ;  \
-                                }   \
-                            __jump = 0; \
-                        }
+        {   \
+    __current_task_virtual = (unsigned int *) *(__current_task_virtual + NEXT_OFFSET_PT) ;     /* soft transition */  \
+        }else{  \
+            unsigned int totJumpSize = (*(__current_task_virtual + BLOCK_OFFSET_PT) + __jump_by);   \
+            if(totJumpSize > __totNumTask)  \
+            {   \
+                int dis  = totJumpSize - __totNumTask;  \
+                while( dis  > __totNumTask) \
+                {   \
+                    dis = dis - __totNumTask;   \
+                }   \
+                dis = dis - (totJumpSize - __jump_by);  \
+                __current_task_virtual  += (dis + dis + dis) ;  \
+            }else{  \
+                __current_task_virtual  += ( __jump_by+__jump_by  + __jump_by )  ;  \
+            }   \
+            __jump = 0; \
+        }
 
 #define JUMP();    if(!__jump){   \
-                        __current_task_virtual  =  \
-                        (unsigned int*) (*(__current_task_virtual + NEXT_OFFSET_PT)) ;     /* soft transition */ \
-                    }else{  \
-                        while(__jump_cnt < __jump_by)   \
-                            {   \
[…]

### 2017-08-03 paper/compiler.tex

--- 2017-07-30 paper/compiler.tex
@@
 
-The compiler consists of two parts called \emph{Passes}:
+The compiler consists of two {\em passes} that runs on LLVM Intermediate Representation (IR)\todo{Which version?}{Kiwan}---an intermediate language generated by Clang~\cite{clang_website}:
 %
-\begin{itemize}
+\begin{enumerate}
 	\item \textbf{Analysis Pass}, which analyzes the code structure and decides where to put task boundaries;
 	\item \textbf{Transform Pass}, which converts the code into task-style program and fills it with necessary \sys keywords so that it can run on \sys.
-\end{itemize}
+\end{enumerate}
 %
@@
 
-In the first pass \sys compiler decides on the task size.
-\sys' task size metric is defined by the number of LLVM Intermediate Representation (IR) instructions.
-That is, \sys tries to make every task to have roughly the same amount of LLVM IR instructions.
+In the first pass \sys compiler decides where to put task boundaries. By definition, a \sys task should adhere to the following correctness constraints:
+
+\begin{enumerate}[label={\bf C\arabic*:}]
+\item{No task should ever use up more energy than what the capacitor of the system can provide;} 
+\item{A task should be single-entry.}\todo{Define single entry}{Kiwan}
+\end{enumerate}
+%
+Also for the system to be efficient, the compiler should optimize the following points:
+%
+\begin{enumerate}[label={\bf O\arabic*:}]
+\item{Tasks should be roughly the same size for efficient coalescing;}
+\item{The number of protected variables should be minimized;}
+\item{Task boundary should be visited as least as possible due to overhead on boundaries.}\todo{Define overhead on boundaries}{Kiwan}
+\end{enumerate}
+
+Analysis pass tries to meet constraints \{C1, C2\} and \{O1, O2, O3\} by (i) \emph{graph building}, (ii) \emph{loop analysis}, (iii) \emph{task preparation}, and (iv) \emph{single entry assurance} procedures described below.
 
@@
 	\centering
-	%\includegraphics[width=\columnwidth]{figures/compiler_graph}
-	\caption{\sys compiler flow diagram.\todo{Draw compiler flow graph diagram}{Kiwan}}
-	\label{fig:compiler_graph}
+	\subfloat[Plain C code]{\includegraphics[width=0.3\columnwidth]{figures/compiler1.pdf}\label{fig:comp_1}}
+	\subfloat[Boundary placement]{\includegraphics[width=0.3\columnwidth]{figures/compiler2.pdf}\label{fig:comp_2}}
+	\subfloat[Code Transformation]{\includegraphics[width=0.3\columnwidth]{figures/compiler3.pdf}\label{fig:comp_3}}
+	\caption{Illustration of \sys compiler task division operation.\todo{Make figures of the same readable size}{Przemek}}
+	\label{fig:compiler_overview}
 \end{figure}
 
-\sys compiler first builds a control flow graph where (i) vertice is a \emph{each basic block} and (ii) a unidirectional edge is a \emph{branch from one basic block to another}, see Figure~\cite{fig:compiler_graph}.
-Each vertice has a {\em price}, which is the number of LLVM IRs inside the vertice.
-When forming a task, each task should contain a basic block so that the sum of price of all blocks within each tasks should be roughly the same \todo{Provide more details on this algorithm}{Kiwan}.
-Price makes the workload done by each tasks roughly the same and make no task exceed the energy bound.\todo{Specify how do you decide on the minimum task size/energy reservoir}{Kiwan}
+\textbf{Graph Building:} starting from a plain C code (see example in Figure~\ref{fig:comp_1}) \sys compiler first builds a control flow graph where (i) each vertices is a \emph{basic block} and (ii) each unidirectional edge is a \emph{branch from one basic block to another} (see example in Figure~\ref{fig:comp_2}).
+All function calls are inlined in the main function prior to building a graph.
 
-Each edge has a {\em penalty}, which indicates the penalty of placing a task boundary at the specific edge.
-Intuitively, when a task boundary is placed on a frequently visited edge (such as a back edge of a loop) the system suffers from higher runtime overhead.
-As we are unaware of a way to statically estimate the number of visit at each edge, we therefore initially set every edge penalty to one.
-When analyzing loops, penalty of edges within loops are multiplied by loop count \todo{Change the penalty to be one or 1$\times$loopcount}{Kiwan}.
-After constructing a graph and assigning price and penalty, the compiler analyzes loops \todo{Explain what do you mean by analyzing loops}{Kiwan}.
-Except those inside a loop which is set as the loop count \todo{Explain this statement}{Kiwan}.
-Unbounded loops are conservatively considered as very large \todo{Explain this statement}{Kiwan}.
+Each vertex has a {\em price}, which represents the energy usage of the basic block.
+We used the number of LLVM IRs inside a vertex as the energy use metric.
+When forming a task, each task should contain a basic block so that the sum of price of all blocks within each tasks should be roughly the same (viz.
+O1) and does not exceed given energy budget (viz.
+C1).
+For our implementation, the programmer should provide the energy budget in compile time defined in number of LLVM IR instructions that the hardware can execute safely in one energy cycle.
+This number must be found experimentally.
+
+Each edge holds the information about the {\em live variables} across the edge\todo{Define live variables}{Kiwan}, and if the edge is within any loop, the {\em loop count}.
+By this information, the compiler tries to minimize the number of protected variable (viz.
+O2), because placing boundary at an edge means the live variables across that edge need to be protected.
+Also the compiler avoids placing boundaries at an edge that is within a loop whose loop count is big (viz.
+O3).
+
+The compiler also calculates a dominance relation \todo{Define dominance relation}{Kiwan} between vertices. This information is used to ensure task's single-entry property (viz. C2).
+
+\textbf{Loop Analysis:} After constructing the graph and assigning the information on a price and live variables to each vertex and edge, respectively, the compiler begins with analyzing the program loops.
+The compiler first identifies all loops and then multiplies the loop count by the price of a vertices within the loop.
+The reason is, if the loop is inside a task it will use up energy that is roughly multiplied by the loop count compared to code that is outside of a loop.
+If the resulting price of a loop is too large\todo{how do you decide on the acceptable loop price?}{Kiwan}, the compiler places boundary at the back edge of the loop, and reverts the change to the price of the vertices inside.
+It means that if the loop is too large to fit in one task, each iteration should be a separate task execution.
+
+If a program has an unbounded loop (such as {\tt while} statement), the compiler conservatively assumes the loop count to be very large.
+The programmer can prevent this by annotating maximum possible loop count of such unbounded loop if the programmer has the information.\todo{'the' or 'this'?
+Does it mean programmed still needs to inspect the code?}{Kiwan}
+
+\textbf{Tasks Preparation:} After initial boundary placemen with loops analysis, the compiler places additional boundaries if some tasks are still too large.
+The \sys compiler starts from the beginning of the main function and traverse through the graph in a depth-first search manner.
+If any part of the graph is found that is too long to fit in the energy bound, it means a boundary must be placed somewhere within.
+The subset of graph that needs boundary placement within is called a {\em boundary-candidate}.
+
+As discussed above, when placing boundary we take three optimization criteria into account: \{O1, O2, O3\}.
+We now introduce a heuristic that consider all these criteria and choose the local optimum\todo{Optimum of what?}{Kiwan}.
+The heuristic calculates the {\em score}, $S$, of each edge inside the boundary candidate as
+%
+\begin{equation}
+S = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
+\end{equation}
+%
+where $N_{\text{IR}}$ is a normalized IR count that the task will end up having if the compiler places boundary at the specific edge, $N_{\text{PV}}$ is the number of protected variables that will be newly introduced by placing boundary at the edge, and choose the edge with the highest score, $N_{\text{loop}}$ is the loop count and $w_{0},w_{1}\in [0,1]$ are pre-selected weights before compilation\footnote{Naturally, different $w_{0}, w_{1}$ will have different effect on the compiler result.
+We will study this effect in Section~\ref{sec:results_compiler}.}.
+Both $N_{\text{IR}}$ and $N_{\text{PV}}$ are normalized within the boundary candidate, since usually IR count is much larger and has higher variance, thus has dominating effect if not normalized.
+It is better to have higher IR count within the task since it means the task size is similar to the size given by the programmer (viz.
+C1).
+On the other hand, change in protected variable is better to be small (viz.
+C2), so we penalize it by assigning a negative value to $N_{\text{PV}}$.
+If the edge is within a loop, the score is divided by $N_{\text{loop}}$.
+It will result in not placing boundary within loops unless other parts of the formula really pays off (viz.
+C3)\todo{This sentence is not clear: what "pays off" means in this context?}{Kiwan}.
+Note that complier never makes a decision that will end up in having more IR count than the programmer specified because at the beginning boundary candidate is chosen to have less IR than the programmer specified number.
+
+Figure~\ref{fig:comp_2} shows example of such boundary placement on graph. The first three vertices will become one task while the rest will become another (see Figure~\ref{fig:comp_3}).
+
+\textbf{Ensuring Single Entry:} After placing a boundary, the compiler ensures that the newly made task is single-entried by checking dominance relation.
+That is, every vertex within a task should be dominated by the entry vertex of the task.
+If this rule is violated, the compiler places additional boundaries to make sure that every task is single entried.
 
@@
 
-Compiler transform pass is also implemented using LLVM. \todo{Fill in this section with details}{Kiwan}
+In the compiler transform pass, a complier executes sub-tasks described below.
+
+\textbf{Declaring and Migrating Codes to Tasks:} After the boundary placement is finished, the analysis pass passes the information to transform pass.
+The transform pass makes new task, traverse through the graph, and copies every code within the vertices it encounters to the new task until it meets the task boundary.
+At the task boundary, it makes another task and starts copying again, and repeat this action until every code is moved to form a task.
+
+\textbf{Renaming Variables:} On copying the code, the variables should be renamed.
+If a variable is alive across multiple tasks, it needs to become a global-protected variable, and every access to the variable should be redirected.
+If a variable's live range is local to the task, it does not need to go through such renaming.
+For example, in Figure~\ref{fig:comp_2}, boundary is placed within the live range of variable {\tt b} (Note that the red cross overlaps the blue arrow).
+Thus variable {\tt b} needs to become a protected variable {\tt p\_b} and every access to {\tt b} should be redirected to {\tt p\_b} (Figure~\ref{fig:comp_3}).
+On the other hand, {\tt a}'s live range is always local to a single task.
+In such case, it need not become a protected variable.
+Note, however, still new declaration of {\tt a} should be inserted in {\tt task1} (Figure~\ref{fig:comp_3}).
+As an improvement to the complier, constants and read-only variable never becomes protected.
+
+\textbf{Inserting \texttt{os\_jump} Calls:} After renaming is done, the compiler detects the broken {\tt branch} instructions\todo{Define protected branch instruction}{Kiwan} which tries to branch to a basic block that is inside different task and change it to an {\tt os\_jump} call (Figure~\ref{fig:comp_3}).
+Note that in this figure we showed the name of the next task to be executed for readability, however {\tt os\_jump} call takes offset number as an argument.
+In reality the compiler statically calculates and passes the offset.
+
+\textbf{Dealing With PHI Nodes:} Whenever more than two different control flows try to write different value in the same memory location, LLVM makes a unique IR called {\em PHI Node} when reading the memory location~\cite{}.
+However if there is a boundary between the memory write and the PHI Node, the compiler will not work.
+Thus, on each of this occurrence, \sys compiler declares a special protected variable and makes the write relocated to the protected variable.
+Then the PHI Node becomes simply a read to a protected variable.
+
+\textbf{Inserting API Call for Paging:} To run \sys, every read or write to a protected variable should call related API for page swapping and dirty page marking.
+The compiler conservatively inserts such calls before a {\em possible} read or write to the protected variable.
+We use the basic pointer alias feature provided by the LLVM library.
+We are being conservative, so it is always correct, albeit not always efficient.\todo{We need to say more about the level of this inefficiency and how it is measured}{Kiwan}
+
+\textbf{Creating New Main Function and Cleaning Unnecessary Codes:} After task generation is done, the compiler generates new main function and peppers the system with necessary codes that is needed by \sys (such as code for evoking {\tt os\_sceduler}).
+Also it removes the code that became stale (such as the body of a function call, for all the function calls are now inlined).
 
@@
 \begin{enumerate}
-	\item \textbf{Conservative handling of unbounded code sections:} When a C program contains an unbounded loop or pointer, the \sys compiler makes a conservative decision.
-\todo{Expand on this statement (define `conservative decision') by providing example and more details on this operation}{Kiwan}.
-This means that for performance-critical implementations, the programmer is advised to manually split the code into tasks.
 	%
-	\item \textbf{LLVM IR instruction as a workload metric:} Intuitively, the number of LLVM IR instructions correlates with the total workload performed by the task.
-However, we need to be aware that is nevertheless a crude approach of estimating task workload---actual machine instructions are what counts and they are not exactly equal to the LLVM IR instruction.
-Also, different instructions have different power usage, use different cycles, etc.
-However, the size of each task need not be exactly the same and it is sufficient to make sure that all of the tasks are small enough to fit in to the energy budget.
+	\item \textbf{Conservative handling of unbounded code sections:} When a C program contains an unbounded loop, the \sys compiler makes a conservative decision of regarding its loop count to be very large.
+However, if a programmer annotates the expected loop count manually the overhead can be relieved.
+	%
+	\item \textbf{Conservative handling of pointers:} When a C program contains pointer, due to the limitation of pointer alias, the compiler makes the conservative decision.
+That is, it may consider a variable's live range longer than actual, or it may insert API call for page swap even when it is unnecessary.
+However, these overhead can be reduced by better pointer aliasing algorithm.
+	%
+	\item \textbf{Function call inlining:} \sys compiler inlines all function call prior to analysis for simplicity.
+However it can introduce code bloat.
+Insted, a better way would be to convert large-frequently visited functions to a separate task and make the function call a jump to the task.
+However we did not implemented such feature for simplicity of the system.
+	%
+	\item \textbf{LLVM IR instruction count to estimate the energy usage:} The compiler uses the number of LLVM IR count as an indicator for energy usage in trying to meet correctness constraint C1 and optimization constraint O1.
+Intuitively, the number of LLVM IR correlates with the energy usage since more instructions will lead to more energy usage.
+However, we need to be aware that is nevertheless a crude approach of estimating energy consumption---actual machine instructions are what counts and they are not exactly equal to the LLVM IR.
+Also, different instructions have different power usage, and even same instruction has different energy usage from time to time (for example cache hit and miss makes even the same {\tt load} instruction to use different amount of energy).
+However, the energy usage of each task need not be exactly the same to what programmer indicated, and it is sufficient to make sure that all of the tasks are small enough to fit in to the energy budget.
 Therefore we conjecture that from the \emph{implementation simplicity perspective} of \sys's compiler counting the LLVM IR instruction is sufficient.
+The programmer can always ensure correctness constraint C1 by conservatively providing IR count in compile time that is empirically proven to be safe.
+Also, since the estimation of energy usage of each basic block is perpendicular to the compiler itself, it is simple to plug in a better energy model if there is one.
+	%
+	\item \textbf{Limitation of the boundary placement heuristics:} The boundary placement strategy is a heuristic that chooses the local optimum.
+However, selecting every local optimum might not be the global optimal solution.
+We can instead do a random search multiple times to find a global optima.
+Since current heuristic is already showing good result that is comparable to human-written code, refer to Section~\ref{sec:results_coalescing}, we leave such optimization as a future work.
 \end{enumerate}
-
-In the light of this design decisions we will assess the compiler overhead in Section~\ref{sec:results_compiler}.

### 2017-08-03 paper/compiler.tex
Kiwan's compiler section revised - really nice text Kiwan

@@
 
-Each edge holds the information about the {\em live variables} across the edge\todo{Define live variables}{Kiwan}, and if the edge is within any loop, the {\em loop count}.
+Each edge holds the information about the {\em live variables} across the edge, and if the edge is within any loop, the {\em loop count}.
 By this information, the compiler tries to minimize the number of protected variable (viz.

### 2017-08-03 paper/compiler.tex
committing partial edit before quick phone call

@@
-A unique feature of \sys is a compiler support for \emph{automatic task generation}.
-\sys literally eliminates the programmer's burden of writing a task-based program from scratch.
-In essence, the \sys compiler obtains a plain C code as an input and generates a binary file that is rewritten in task-based programming style that \sys executes.
+\sys provides compiler support for \emph{automatic task generation} that simplifies programming writing task-based code by automatically translating C code into a task-based executable.
+Translating C code into task-based code automatically requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
+The \sys compiler first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary.
+After identifying where to put task boundaries, the \sys compiler transforms the imperative code into tasks, instrumenting protected memory accesses.
 
-The compiler consists of two {\em passes} that runs on LLVM Intermediate Representation (IR)\todo{Which version?}{Kiwan}---an intermediate language generated by Clang~\cite{clang_website}:
+%The compiler consists of two {\em passes} that runs on LLVM Intermediate Representation (IR)\todo{Which version?}{Kiwan}---an intermediate language generated by Clang~\cite{clang_website}:
 %
-\begin{enumerate}
-	\item \textbf{Analysis Pass}, which analyzes the code structure and decides where to put task boundaries;
-	\item \textbf{Transform Pass}, which converts the code into task-style program and fills it with necessary \sys keywords so that it can run on \sys.
-\end{enumerate}
+%\begin{enumerate}
+%	\item \textbf{Analysis Pass}, which analyzes the code structure and decides where to put task boundaries;
+%	\item \textbf{Transform Pass}, which converts the code into task-style program and fills it with necessary \sys keywords so that it can run on \sys.
+%\end{enumerate}
 %
-We shall now proceed with describing these two parts in detail.
+%We shall now proceed with describing these two parts in detail.
 
-\subsection{Compiler Analysis Pass}
+\subsection{Task Boundary Identification Analysis}
 \label{sec:compiler_analysis_pass}
 
-In the first pass \sys compiler decides where to put task boundaries. By definition, a \sys task should adhere to the following correctness constraints:
+\todo{Define single entry}{Kiwan}
+\TODO{Brandon: the term "single-entry" is well-known in the community to mean that there is a single control-flow edge entering the task.
+Thinking about this a bit, is there a simple reason why we cannot have multiple-entry tasks?
+A totally plausible task graph could be like A->B B->C C->B C->A; here, B has two predecessors.
+Is this fundamental?
+Or an analysis convenience?
+I've left it as "single-entry" for now}
 
-\begin{enumerate}[label={\bf C\arabic*:}]
-\item{No task should ever use up more energy than what the capacitor of the system can provide;} 
-\item{A task should be single-entry.}\todo{Define single entry}{Kiwan}
-\end{enumerate}
+\sys analyzes a program to determine where to put task boundaries to divide the
+program into tasks. A task is a single-entry, multiple-exit region of the CFG
+and in a correct task decomposition, no task should ever consume more energy
+than the device's energy buffer can store.  There are several factors that
+determine whether tasks will execute efficiently. A task decomposition should
+minimize the number of variables that are accessed by multiple tasks and, 
+consequently, must be protected by \sys's paging mechanism.  Tasks should
+be roughly the same size, simplifying \sys's dynamic reasoning about coalescing
+by eliminating the need to consider variable task cost at run time.  To minimize
+overhead introduced at task boundaries (i.e., commit), a task decomposition should 
+maximize task size, which minimizes the frequency of task transitions.
+
+%\begin{enumerate}[label={\bf C\arabic*:}]
+%\item{No task should ever use up more energy than what the capacitor of the system can provide;} \TODO{We do not (cannot, actually) guarantee this, so I think we can, at best, say that this is an optimization constraint.
+What do you think?}
+%\item{A task should be single-entry.}\todo{Define single entry}{Kiwan} 
+%\end{enumerate}
 %
-Also for the system to be efficient, the compiler should optimize the following points:
+%For a task-decomposition to be efficient, the compiler should optimize the following points:
 %
-\begin{enumerate}[label={\bf O\arabic*:}]
-\item{Tasks should be roughly the same size for efficient coalescing;}
-\item{The number of protected variables should be minimized;}
-\item{Task boundary should be visited as least as possible due to overhead on boundaries.}\todo{Define overhead on boundaries}{Kiwan}
-\end{enumerate}
+%\begin{enumerate}[label={\bf O\arabic*:}]
+%\item{Tasks should be roughly the same size for efficient coalescing;}
+%\item{The number of protected variables should be minimized;}
+%\item{Task boundary should be visited as least as possible due to overhead on boundaries.}\todo{Define overhead on boundaries}{Kiwan}
+%\end{enumerate}
 
-Analysis pass tries to meet constraints \{C1, C2\} and \{O1, O2, O3\} by (i) \emph{graph building}, (ii) \emph{loop analysis}, (iii) \emph{task preparation}, and (iv) \emph{single entry assurance} procedures described below.
+\sys meets these correctness and efficiency requirements with a four-part
+analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis},
+(iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.
 
@@
 
-\textbf{Graph Building:} starting from a plain C code (see example in Figure~\ref{fig:comp_1}) \sys compiler first builds a control flow graph where (i) each vertices is a \emph{basic block} and (ii) each unidirectional edge is a \emph{branch from one basic block to another} (see example in Figure~\ref{fig:comp_2}).
+\textbf{Graph Construction:} starting from a plain C code (see example in Figure~\ref{fig:comp_1}) \sys compiler first builds a control flow graph where (i) each vertices is a \emph{basic block} and (ii) each unidirectional edge is a \emph{branch from one basic block to another} (see example in Figure~\ref{fig:comp_2}).
 All function calls are inlined in the main function prior to building a graph.
 
-Each vertex has a {\em price}, which represents the energy usage of the basic block.
+Each vertex has a {\em cost}, which represents the energy usage of the basic block.
 We used the number of LLVM IRs inside a vertex as the energy use metric.
-When forming a task, each task should contain a basic block so that the sum of price of all blocks within each tasks should be roughly the same (viz.
+When forming a task, each task should contain a basic block so that the sum of cost of all blocks within each tasks should be roughly the same (viz.
 O1) and does not exceed given energy budget (viz.
@@
 
-\textbf{Loop Analysis:} After constructing the graph and assigning the information on a price and live variables to each vertex and edge, respectively, the compiler begins with analyzing the program loops.
-The compiler first identifies all loops and then multiplies the loop count by the price of a vertices within the loop.
+\textbf{Loop Analysis:} After constructing the graph and assigning the information on a cost and live variables to each vertex and edge, respectively, the compiler begins with analyzing the program loops.
+The compiler first identifies all loops and then multiplies the loop count by the cost of a vertices within the loop.
 The reason is, if the loop is inside a task it will use up energy that is roughly multiplied by the loop count compared to code that is outside of a loop.
-If the resulting price of a loop is too large\todo{how do you decide on the acceptable loop price?}{Kiwan}, the compiler places boundary at the back edge of the loop, and reverts the change to the price of the vertices inside.
+If the resulting cost of a loop is too large\todo{how do you decide on the acceptable loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop, and reverts the change to the cost of the vertices inside.
 It means that if the loop is too large to fit in one task, each iteration should be a separate task execution.

### 2017-08-03 paper/compiler.tex
more progress, but the text still needs a lot of work.

@@
 
-\sys meets these correctness and efficiency requirements with a four-part
-analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis},
-(iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.
+%\sys meets these correctness and efficiency requirements with a four-part
+%analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis},
+%(iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.
 
@@
 
-\textbf{Graph Construction:} starting from a plain C code (see example in Figure~\ref{fig:comp_1}) \sys compiler first builds a control flow graph where (i) each vertices is a \emph{basic block} and (ii) each unidirectional edge is a \emph{branch from one basic block to another} (see example in Figure~\ref{fig:comp_2}).
-All function calls are inlined in the main function prior to building a graph.
+The input to \sys is a standard C program (as shown in Figure~\ref{fig:comp_1})
+and the output of \sys's analysis is the CFG, with some vertices marked as {\em
+task boundaries}. 
+%
+The \sys compiler builds the program's CFG
+(as shown in Figure~\ref{fig:comp_2}).  To simplify interprocedural analysis,
+our prototype of \sys's analysis inlines all function calls into the main
+function before building a graph. 
+%
+\sys uses a cost-based task boundary placement analysis that
+considers per-block energy consumption estimates, inter-task variable liveness,
+and the cost of partitioning loops.  We first describe how \sys measures each of these costs 
+and then describe \sys's procedure for placing boundaries in the CFG.
 
-Each vertex has a {\em cost}, which represents the energy usage of the basic block.
-We used the number of LLVM IRs inside a vertex as the energy use metric.
-When forming a task, each task should contain a basic block so that the sum of cost of all blocks within each tasks should be roughly the same (viz.
-O1) and does not exceed given energy budget (viz.
-C1).
-For our implementation, the programmer should provide the energy budget in compile time defined in number of LLVM IR instructions that the hardware can execute safely in one energy cycle.
-This number must be found experimentally.
+\subsubsection{Task Cost Analysis}
+\sys's task cost analysis considers estimated task energy cost, variable
+liveness, and the impact of splitting loops across tasks.  The result of the
+task cost analysis determines where to place boundaries in the CFG.
 
-Each edge holds the information about the {\em live variables} across the edge, and if the edge is within any loop, the {\em loop count}.
-By this information, the compiler tries to minimize the number of protected variable (viz.
-O2), because placing boundary at an edge means the live variables across that edge need to be protected.
-Also the compiler avoids placing boundaries at an edge that is within a loop whose loop count is big (viz.
-O3).
+{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the
+CFG a {\em cost}, which is an estimate of the energy consumption of the block.
+Precisely modeling the energy cost of arbitrary C code is a difficult problem
+and beyond the scope of this work; we used the number of intermediate
+representation (IR) instructions in a block as a proxy for energy because
+energy cost is likely to tightly correspond  to block length.  Moreover, we
+assume as an input to the compiler the maximum number of operations that the
+device can execute without exhausting its buffered energy. This parameter is
+highly device-dependent, and we assume that a developer or device manufacturer
+determines the number empirically, as we did in this work through direct
+measurement.  \sys's analysis tries ensure that the estimated energy cost of different tasks
+is roughly the same and does not exceed given energy budget.
 
-The compiler also calculates a dominance relation \todo{Define dominance relation}{Kiwan} between vertices. This information is used to ensure task's single-entry property (viz. C2).
+{\noindent \bf Live variable cost.}
+Each CFG edge is associated with a set of {\em live variables} that are
+accessible in both of the blocks that define the edge.  Variables that are live
+across an edge between two tasks must be protected by \sys's memory
+virtualization mechanism and each live, inter-task variable increases the run
+time overhead for a boundary placement.  \sys's analysis tries to minimize the
+overhead of protected variables by placing boundaries to minimize the number of
+inter-task live variables.
 
-\textbf{Loop Analysis:} After constructing the graph and assigning the information on a cost and live variables to each vertex and edge, respectively, the compiler begins with analyzing the program loops.
-The compiler first identifies all loops and then multiplies the loop count by the cost of a vertices within the loop.
-The reason is, if the loop is inside a task it will use up energy that is roughly multiplied by the loop count compared to code that is outside of a loop.
-If the resulting cost of a loop is too large\todo{how do you decide on the acceptable loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop, and reverts the change to the cost of the vertices inside.
-It means that if the loop is too large to fit in one task, each iteration should be a separate task execution.
+{\noindent \bf Loop splitting cost.}
+and if the edge is within any loop, the {\em loop count}. 
+Also the compiler avoids placing boundaries at an edge that is
+within a loop whose loop count is big.
+
+After constructing the graph and assigning the information on a cost and live
+variables to each vertex and edge, respectively, the compiler begins with
+analyzing the program loops. The compiler first identifies all loops and then
+multiplies the loop count by the cost of a vertices within the loop. The reason
+is, if the loop is inside a task it will use up energy that is roughly
+multiplied by the loop count compared to code that is outside of a loop. If the
+resulting cost of a loop is too large\todo{how do you decide on the acceptable
+loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop,
+and reverts the change to the cost of the vertices inside. It means that if the
+loop is too large to fit in one task, each iteration should be a separate task
+execution.
 
@@
 
+\subsubsection{Task Boundary Placement}
 \textbf{Tasks Preparation:} After initial boundary placemen with loops analysis, the compiler places additional boundaries if some tasks are still too large.
@@
 
-\textbf{Ensuring Single Entry:} After placing a boundary, the compiler ensures that the newly made task is single-entried by checking dominance relation.
-That is, every vertex within a task should be dominated by the entry vertex of the task.
-If this rule is violated, the compiler places additional boundaries to make sure that every task is single entried.
+\textbf{Ensuring Single Entry:}
 
-\subsection{Compiler Transform Pass}
+The compiler also calculates a dominance relation \todo{Define dominance
+relation}{Kiwan} between vertices. This information is used to ensure task's
+single-entry property.  After placing a boundary, the compiler ensures that the
+newly made task is single-entried by checking dominance relation. That is,
+every vertex within a task should be dominated by the entry vertex of the task.
+If this rule is violated, the compiler places additional boundaries to make
+sure that every task is single entried.
+
+\subsection{Task Transformation}
 \label{sec:compiler_transform_pass}

### 2017-08-03 paper/paging.tex
paging related things
merged two files to write on paging

@@
 However, this is not a trivial task due to the following facts:
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=0.25\columnwidth]{figures/}
+	\caption{Memory virtualization architecture.}
+	\label{fig:}
+\end{figure}
 
@@
 	\caption{\texttt{PageFault(tag)} pseudo-code}
-	\label{algo:rwar}
+	\label{algo:pagefault}
 	\scriptsize
@@
 Using a two-phase commit by utilizing DMA.
+
+\subsection{\sys: Data/Memory Protection Methods}
+
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=0.8\columnwidth]{figures/iposCommitSize}
+	\caption{...}
+	\label{fig:virtOperationalBuf}
+\end{figure}
+
+\sys data consistency preservation methods are designed with goal of minimal interaction between the CPU and FRAM to reduce energy consumption (see Section~\ref{sec:prelResults}).
+These data/memory protection methods enables \sys to tolerate arbitrary number of power failures.
+
+\subsubsection{Virtualized Operational Buffer}
+
+\begin{algorithm}[t]
+	\caption{Virtualized Operational Buffer}
+	\label{algo:virtuBufWrite}
+	\scriptsize
+	%\small
+	\begin{algorithmic}[1]
+		\State $var \in \text{\{global variables\}} $ 
+		\State \label{lst:virtuBufWrite:line:begin}\Call{Virtual Task()}{} 
+		\While { \textit{executing} } \Comment{Execution stage}
+		\State $var$  $\rightarrow$ \textsf{volatile buffer} \label{lst:virtuBufWrite:line:output}
+		\If { $var$  in \textsf{ volatile buffer} }				\label{lst:virtuBufWrite:line:inputBegin}
+		\State $var$  $\leftarrow$  \textsf{volatile buffer} 
+		\Else 
+		\State $var$  $\leftarrow$  \textsf{FRAM}		\label{lst:virtuBufWrite:line:inputEnd}
+		\EndIf
+
+		\If {power interrupts}
+		\State back to \ref{lst:virtuBufWrite:line:begin}
+		\EndIf
+		\EndWhile
+
+		\While{  $\textsf{volatile buffer}\not=\emptyset$  } \label{lst:virtuBufWrite:line:commitBegin} \Comment{ First phase commit}
+		\State \textsf{volatile buffer} $\rightarrow$  \textsf{persistent buffer}
+		\If {power interrupts}
+		\State discard \textsf{persistent buffer}
+		\State back to  \ref{lst:virtuBufWrite:line:begin} 
+		\State   \label{lst:virtuBufWrite:line:commitEnd}
+		\EndIf
+		\EndWhile
+
+		\While{ $\textsf{persistent buffer}\not=\emptyset$ } \label{lst:virtuBufWrite:line:SecCommitBegin} \Comment{Second phase commit}
+		\State \textsf{persistent buffer} $\rightarrow$ FRAM 
+		\If {power interrupts}
+		\State Continue
+		\EndIf
+		\EndWhile 
+		\State     \label{lst:virtuBufWrite:line:SecCommitEnd}
+		\State return
+	\end{algorithmic}
+\end{algorithm}
+
+Since accessing FRAM is more energy expensive and slower than accessing SRAM, preventing frequent access to FRAM (i.e.
+during looping operations) is desirable.
+Therefore, the first proposed protection method utilizes a volatile buffer that holds temporary all the outputs of a virtual task (see Algorithm~\ref{algo:virtuBufWrite} line~\ref{lst:virtuBufWrite:line:output}).
+Consequently, a virtual task must first attempt to read a global variable from the volatile buffer before trying to obtain the value from the non-volatile memory (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:inputBegin}-\ref{lst:virtuBufWrite:line:inputEnd}) to ensure a correct execution progress and the consistency of the memory.
+Once the execution of a virtual task is done, the first phase of the commit process is started by copying the volatile buffer to a persistent buffer (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:commitBegin}-\ref{lst:virtuBufWrite:line:commitEnd}).
+If the power is interrupted the persistent buffer must be discarded and the execution must start again from the beginning of the virtual task.
+In another words, the first phase commit must be performed atomically to preserve the consistency of the output of a virtual task.
+The second phase commit is a power failure immune precess that is responsible for distributing the global variables to their final locations and make the non-volatile memory consistent and synchronized with the computation progress (see Algorithm~\ref{algo:virtuBufWrite}lines~\ref{lst:virtuBufWrite:line:SecCommitBegin}-\ref{lst:virtuBufWrite:line:SecCommitEnd}).
+
+\paragraph{Virtualized Operational Buffer: Implementation}
+
+We implement a reference implementation of \sys that uses virtualized Operational buffer to protect the data against power failure. 
+\paragraph{Virtual Buffer}
+We realized the virtaulzed buffer as a volatile hashed table of linked lists.
+The hashing technique was chosen to reduce the buffer searching time and the linked lists are used to prevent data loss when there is a conflict between multiple variables---If two variables have the same hash value they will occupy the same cell, however, with linked list a new node will be created for each variable to resolve the conflict and protect the data.
+The hashing function is based on the observation that the virtual addresses of the memory cells have approximately a flat distribution over the memory addressing space and the fact that each entry, of the virtual table, has the address and the value of a variable.
+As such, by using the least significant bits as an index to access the virtual buffer the hash function distributes its inputs uniformly in the virtual buffer.
+Accordingly, the linked lists search time, on average, is reduced.
+
+Despite the fact that the linked lists reduce the average complexity of searching the virtual buffer significantly, they introduce a number of drawbacks: (i) the linked list data structure introduce a non-negligible memory overhead while it relies on a limited memory section, namely the heap; (ii) traversing the linked list is relatively slow.
+Therefore, we re-implemented the virtual buffer, after observing that most of the global variables have the same most significant bits, as a bigger hashed table that does not allow entries conflict.
+To eliminate the need for searching the hashed table linearly while committing its entries, we add a new buffer that holds the indices of the occupied cells of the hashed table.
+This table is of the same length as the hashed table.
+We can reason about the benefit of this buffer as follows: if we assume that the global variables are evenly divided between the tasks, then this buffer reduces the commit time by a factor equal to the number of tasks.
+
+\paragraph{Persistent Buffer}
+The persistent buffer is implemented as static array of tuples, where each tuple holds  the address and the value of a variable.
+This data structure is very suitable for the second phase commit where each element has to be committed to its finial location.
+However, the complexity of committing this buffer is of size $O(N)$, where $N$ equals the length of the buffer.
+Therefore, the size of the buffer can have dramatic effect on the performance of \sys---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious.
+On the other hand, if the size of the buffer is large \sys will experience a significant performance degradation.
+However, by observing the type of the information that this buffer holds, namely memory addresses, we can, on average, reduce the complexity of committing this buffer by defining the \emph{effective size} of the buffer to be the size of the buffer up to the first cell that holds an invalid memory address.
+Moreover, since this buffer is in FRAM which is much bigger than SRAM its size can be relatively big.
+
+% \paragraph{Virtualized Operational Buffer: delay analysis}[initial results]
+
+% \begin{figure}[t]
+% 	\centering
+% 	\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
+% 	\caption{\sys based on virtualized Operational Buffer.
+It runs the decompression application with different buffer configurarions.
+EPB: Enhanced Persistent Buffer.
+NR: no read operation from the buffer.
+NRW: not read or write operations from the buffer.}
+% 	\label{fig:virtOperationalBuf}
+% \end{figure}
+
+% 	To analyze the performance of \sys versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
+We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}.
+The results presented in Fig.~\ref{fig:virtOperationalBuf} show the \sys has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.
+
+% 	All the experiments were done with a persistent buffer of size 100.
+However, the bar labeled with "EPB/16" in  Fig.~\ref{fig:virtOperationalBuf} shows the effect of the enhanced commit process that uses the buffer effective size instead of the buffer maximum size.
+We see that the execution time is reduced to $\approx$80 ms.
+Furthermore, we quantized the overhead of the read operations from the buffer, by eliminating this operation, and the bar labeled with "NR" shows that it cost  $\approx$10 ms.
+Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to \sys best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.
+
+\subsubsection{ Direct Memory Access Based Protection Method (better name is needed)}
+%
+\begin{figure}[t]
+	\centering
+	\subfloat[The time needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/dmaSize_time.eps} }
+	\subfloat[The energy needed to transfer a block of data using pointers versus Direct Memory Access (DMA).]{\includegraphics[width=0.48\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
+	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM}
+	\label{fig:dmaTimeEnergy}
+\end{figure}
+%
+Since accessing/updating buffers is more costly than directly accessing FRAM, at least without specialized hardware~\cite{clank}, we introduce the second method to apply the virtualization principle to the Modular Execution Model.
+This method utilizing the Direct Memory Access (DMA) Module.
+As Fig.~\ref{fig:dmaTimeEnergy} shows that DMA is much more efficient in transferring a block of data than the conventional (pointer based) way from the energy and time perspectives.
+
+This method combines a number of techniques to preserve the persistence and consistency of the data.
+It backs up the status of the global variables in the non-volatile memory, and double buffers them to guarantee their consistency.
+Furthermore, on each power up it populates the global variables in the volatile memory from the most recent back up in the non-volatile memory to enable a faster task execution since the task does not need to interact with non-volatile memory which is generally slower and more energy expensive.

### 2017-08-03 paper/paging.tex
First pass over paging (virtualization) - more edits tomorrow

--- 2017-07-14 paper/paging.tex
@@
-One of the main objectives of VIPRE run-time is to restrict virtual tasks to interface only with the volatile memory, since volatile memory access is relatively cheaper non-volatile memory access.
-To this end, the persistent variables a virtual task operates on should be brought from non-volatile memory to a fixed-sized \emph{volatile buffer} and when the virtual task is finished the modified variables in this buffer should be committed back to their original locations in non-volatile memory.
-However, this is not a trivial task due to the following facts:
+One of the main objectives of \sys run-time is to restrict virtual tasks to interface only with the volatile memory\todo{Do we define what a virtual task is and why is it important?}{Amjad/Sinan}, since volatile memory access is relatively cheaper than non-volatile memory access, as we already have shown experimentally in Section~\ref{section:background_task_computing} (Problem 1, Figure~\ref{fig:framEnergy}).
+Before dwelling into the details of virtualization of \sys, we need to shed more light on the rationale behind our approach.
 
-\paragraph{I-Dynamic Access Overhead} The persistent variables accessed and/or modified by a virtual task might not be known in advance due to the dynamic program flow.
+\subsection{Memory Virtualization versus Intermittence}
+\label{sec:virtualization_problems}
+
+The persistent variables a virtual task operates on should be brought from non-volatile memory to a fixed-sized \emph{volatile buffer}.
+When the virtual task is finished the modified variables in this buffer should be committed back to their original locations in non-volatile memory.
+We list three fundamental problems proving that this is however a non-trivial task.
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=\columnwidth]{figures/}
+	\caption{\sys memory virtualization architecture.\todo{Decide if we really need this figure (we already have a lot of figs)}{Brandon}}
+	\label{fig:virtualization_architecture}
+\end{figure}
+
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=\columnwidth]{figures/iposCommitSize}
+%	\caption{Commit size delay for three memory access methods.}
+%	\label{fig:virtOperationalBuf}
+%\end{figure}
+
+\textbf{Problem 1---Dynamic Access Overhead:} The persistent variables accessed and/or modified by a virtual task might not be known in advance due to the dynamic program flow.
 Therefore, the volatile buffer should keep track of the persistent variables accessed by any virtual task at runtime.
@@
 Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations.
-Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead and might kill the benefit of task virtualization.
+Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead and might kill the benefit of task virtualization.\todo{How to show it experimentally?}{Sinan} %refer to Figure~\ref{fig:virtOperationalBuf} for experimental evaluation.
 
-\paragraph{II-Two Phase Commit Overhead}  For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each virtual task by means of a \emph{two-phase commit} approach.
-However, two-phase commit is also very expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed with CPU intervention.
+\begin{figure}[t]
+	\centering
+	\subfloat[SRAM Double Buffer Copy]{\includegraphics[width=0.49\columnwidth]{figures/sram-buffer} }
+	\subfloat[DMA Copy]{\includegraphics[width=0.49\columnwidth]{figures/dma}}
+	\caption{.\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
+	\label{fig:sram_vs_dma}
+\end{figure}
 
-\paragraph{III-Main Memory Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
-In order to eliminate run-time buffer \emph{search} overhead, \emph{all} persistent variables can be brought from non-volatile memory to the volatile buffer at system start-up---very efficiently and faster thanks to Direct memory access (DMA) that eliminates CPU intervention for \emph{block} memory-memory copy/move operations.
+\textbf{Problem 2---Two Phase Commit Overhead:} For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each virtual task by means of a \emph{two-phase commit} approach.
+However, two-phase commit is also very expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed with CPU intervention, refer to Fig.~\ref{fig:sram_vs_dma}.
+
+\begin{figure}[t]
+	\centering
+	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
+	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
+	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: pointers versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
+	\label{fig:dmaTimeEnergy}
+\end{figure}
+
+\textbf{Problem 3---Main Memory Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
+In order to eliminate run-time buffer \emph{search} overhead (by not relying on specialized hardware such as e.g.
+in~\cite{hicks_isca_2017}), \emph{all} persistent variables can be brought from non-volatile memory to the volatile buffer at system start-up---very efficiently and faster thanks to Direct Memory Access (DMA) that eliminates CPU intervention for \emph{block} memory-memory copy/move operations, as shown experimentally in Figure~\ref{fig:dmaTimeEnergy}.
 Two-phase commit is also very efficient in this case since DMA can be utilized to move volatile buffer as a block to non-volatile locations.
@@
 
-Considering these facts, we designed a \emph{paging-based} virtual memory in software that (i) partitions contiguously allocated persistent variables into pages; (ii) implements the volatile \emph{page buffer} that holds a single page---solves the aforementioned memory-fit problem; (ii) swaps the page in volatile memory with a new page in non-volatile memory efficiently by utilizing DMA-based memory block-copy operations; (iii) buffers the swapped volatile page until commit to enable task virtualization; (iv) implements two-phase commit of the modified pages to preserve the consistency of non-volatile memory; (v) and performs two-phase commit operations using DMA for the sake of efficiency.
-We explain the details of our design in the following subsections.
+Considering these facts, \sys implements a \emph{paging-based} virtual memory in software that (i) partitions contiguously allocated persistent variables into pages; (ii) implements the volatile \emph{page buffer} that holds a single page---solves the aforementioned memory-fit problem; (ii) swaps the page in volatile memory with a new page in non-volatile memory efficiently by utilizing DMA-based memory block-copy operations; (iii) buffers the swapped volatile page until commit to enable task virtualization; (iv) implements two-phase commit of the modified pages to preserve the consistency of non-volatile memory; (v) and performs two-phase commit operations using DMA for the sake of efficiency.
+We explain the details of \sys virtualization design in the following sections.
 
@@
 	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
-	\caption{Address translation}
+	%\includegraphics[width=\columnwidth]{figures/}
+	\caption{Address translation.\todo{Draw figure}{Sinan}}
 	\label{fig:address-translation}
@@
 
+\todo{Section to be expanded}{Sinan}
+
 \begin{algorithm}[t]
 	\caption{\texttt{PageFault(tag)} pseudo-code}
-	\label{algo:rwar}
+	\label{algo:pagefault}
 	\scriptsize
@@
 If the tag of the virtual address and \texttt{CrntPagTag} are not equal, the active page in page buffer should be swapped with the requested page. In this situation, we have the following two cases:
+
 \paragraph{Modified Active Page:} If the contents of the page buffer is modified, we commit it to an intermediate page buffer in non-volatile memory, namely to \texttt{pagesTemp}.
@@
 
-Using a two-phase commit by utilizing DMA.
+Using a two-phase commit by utilizing DMA. \todo{Text to be written}{Sinan}
+
+%%% Old Text by Amjad %%%
+
+%\begin{algorithm}[t]
+%	\caption{Virtualized Operational Buffer}
+%	\label{algo:virtuBufWrite}
+%	\scriptsize
+%	%\small
+%	\begin{algorithmic}[1]
+%		\State $var \in \text{\{global variables\}} $ 
+%		\State \label{lst:virtuBufWrite:line:begin}\Call{Virtual Task()}{} 
+%		\While { \textit{executing} } \Comment{Execution stage}
+%		\State $var$  $\rightarrow$ \textsf{volatile buffer} \label{lst:virtuBufWrite:line:output}
+%		\If { $var$  in \textsf{ volatile buffer} }				\label{lst:virtuBufWrite:line:inputBegin}
+%		\State $var$  $\leftarrow$  \textsf{volatile buffer} 
+%		\Else 
+%		\State $var$  $\leftarrow$  \textsf{FRAM}		\label{lst:virtuBufWrite:line:inputEnd}
+%		\EndIf
+%		
+%		\If {power interrupts}
+%		\State back to \ref{lst:virtuBufWrite:line:begin}
+%		\EndIf
+%		\EndWhile
+%		
+%		\While{  $\textsf{volatile buffer}\not=\emptyset$  } \label{lst:virtuBufWrite:line:commitBegin} \Comment{ First phase commit}
+%		\State \textsf{volatile buffer} $\rightarrow$  \textsf{persistent buffer}
+%		\If {power interrupts}
+%		\State discard \textsf{persistent buffer}
+%		\State back to  \ref{lst:virtuBufWrite:line:begin} 
+%		\State   \label{lst:virtuBufWrite:line:commitEnd}
+%		\EndIf
+%		\EndWhile 
+%		
+%		\While{ $\textsf{persistent buffer}\not=\emptyset$ } \label{lst:virtuBufWrite:line:SecCommitBegin} \Comment{Second phase commit}
+%		\State \textsf{persistent buffer} $\rightarrow$ FRAM 
+%		\If {power interrupts}
+%		\State Continue
+%		\EndIf
+%		\EndWhile 
+%		\State     \label{lst:virtuBufWrite:line:SecCommitEnd}
+%		\State return
+%	\end{algorithmic}
+%\end{algorithm}
+
+%Since accessing FRAM is more energy expensive and slower than accessing SRAM, preventing frequent access to FRAM (i.e.
+during looping operations) is desirable.
+Therefore, the first proposed protection method utilizes a volatile buffer that holds temporary all the outputs of a virtual task (see Algorithm~\ref{algo:virtuBufWrite} line~\ref{lst:virtuBufWrite:line:output}).
+Consequently, a virtual task must first attempt to read a global variable from the volatile buffer before trying to obtain the value from the non-volatile memory (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:inputBegin}-\ref{lst:virtuBufWrite:line:inputEnd}) to ensure a correct execution progress and the consistency of the memory.
+Once the execution of a virtual task is done, the first phase of the commit process is started by copying the volatile buffer to a persistent buffer (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:commitBegin}-\ref{lst:virtuBufWrite:line:commitEnd}).
+If the power is interrupted the persistent buffer must be discarded and the execution must start again from the beginning of the virtual task.
+In another words, the first phase commit must be performed atomically to preserve the consistency of the output of a virtual task.
+The second phase commit is a power failure immune precess that is responsible for distributing the global variables to their final locations and make the non-volatile memory consistent and synchronized with the computation progress (see Algorithm~\ref{algo:virtuBufWrite}lines~\ref{lst:virtuBufWrite:line:SecCommitBegin}-\ref{lst:virtuBufWrite:line:SecCommitEnd}).
+
+%\paragraph{Virtualized Operational Buffer: Implementation}
+
+%We implement a reference implementation of \sys that uses virtualized Operational buffer to protect the data against power failure.
+
+%\paragraph{Virtual Buffer}
+
+%We realized the virtaulzed buffer as a volatile hashed table of linked lists.
+The hashing technique was chosen to reduce the buffer searching time and the linked lists are used to prevent data loss when there is a conflict between multiple variables---If two variables have the same hash value they will occupy the same cell, however, with linked list a new node will be created for each variable to resolve the conflict and protect the data.
+The hashing function is based on the observation that the virtual addresses of the memory cells have approximately a flat distribution over the memory addressing space and the fact that each entry, of the virtual table, has the address and the value of a variable.
+As such, by using the least significant bits as an index to access the virtual buffer the hash function distributes its inputs uniformly in the virtual buffer.
+Accordingly, the linked lists search time, on average, is reduced.
+
+%Despite the fact that the linked lists reduce the average complexity of searching the virtual buffer significantly, they introduce a number of drawbacks: (i) the linked list data structure introduce a non-negligible memory overhead while it relies on a limited memory section, namely the heap; (ii) traversing the linked list is relatively slow.
+Therefore, we re-implemented the virtual buffer, after observing that most of the global variables have the same most significant bits, as a bigger hashed table that does not allow entries conflict.
+To eliminate the need for searching the hashed table linearly while committing its entries, we add a new buffer that holds the indices of the occupied cells of the hashed table.
+This table is of the same length as the hashed table.
+We can reason about the benefit of this buffer as follows: if we assume that the global variables are evenly divided between the tasks, then this buffer reduces the commit time by a factor equal to the number of tasks.
+
+%\paragraph{Persistent Buffer}
+
+%The persistent buffer is implemented as static array of tuples, where each tuple holds  the address and the value of a variable.
+This data structure is very suitable for the second phase commit where each element has to be committed to its finial location.
+However, the complexity of committing this buffer is of size $O(N)$, where $N$ equals the length of the buffer.
+Therefore, the size of the buffer can have dramatic effect on the performance of \sys---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious.
+On the other hand, if the size of the buffer is large \sys will experience a significant performance degradation.
+However, by observing the type of the information that this buffer holds, namely memory addresses, we can, on average, reduce the complexity of committing this buffer by defining the \emph{effective size} of the buffer to be the size of the buffer up to the first cell that holds an invalid memory address.
+Moreover, since this buffer is in FRAM which is much bigger than SRAM its size can be relatively big.
+
+% \paragraph{Virtualized Operational Buffer: delay analysis}[initial results]
+
+% \begin{figure}[t]
+% 	\centering
+% 	\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
+% 	\caption{\sys based on virtualized Operational Buffer.
+It runs the decompression application with different buffer configurarions.
+EPB: Enhanced Persistent Buffer.
+NR: no read operation from the buffer.
+NRW: not read or write operations from the buffer.}
+% 	\label{fig:virtOperationalBuf}
+% \end{figure}
+
+% 	To analyze the performance of \sys versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
+We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}.
+The results presented in Fig.~\ref{fig:virtOperationalBuf} show the \sys has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.
+
+% 	All the experiments were done with a persistent buffer of size 100.
+However, the bar labeled with "EPB/16" in  Fig.~\ref{fig:virtOperationalBuf} shows the effect of the enhanced commit process that uses the buffer effective size instead of the buffer maximum size.
+We see that the execution time is reduced to $\approx$80 ms.
+Furthermore, we quantized the overhead of the read operations from the buffer, by eliminating this operation, and the bar labeled with "NR" shows that it cost  $\approx$10 ms.
+Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to \sys best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.

### 2017-08-04 paper/compiler.tex
done with an aggressive pass over compiler

@@
 
-The input to \sys is a standard C program (as shown in Figure~\ref{fig:comp_1})
-and the output of \sys's analysis is the CFG, with some vertices marked as {\em
-task boundaries}. 
+The input to the task boundary identification analysis is a standard C program
+(as shown in Figure~\ref{fig:comp_1}) and the output of \sys's analysis is the
+CFG, with some vertices marked as {\em task boundaries}. 
 %
-The \sys compiler builds the program's CFG
-(as shown in Figure~\ref{fig:comp_2}).  To simplify interprocedural analysis,
-our prototype of \sys's analysis inlines all function calls into the main
-function before building a graph. 
+The \sys compiler builds the program's CFG (as shown in
+Figure~\ref{fig:comp_2}).  To simplify interprocedural analysis, our prototype
+inlines all function calls into the main function in the CFG. 
 %
@@
 representation (IR) instructions in a block as a proxy for energy because
-energy cost is likely to tightly correspond  to block length.  Moreover, we
+energy cost is likely to closely correspond to block length.  Moreover, we
 assume as an input to the compiler the maximum number of operations that the
@@
 Each CFG edge is associated with a set of {\em live variables} that are
-accessible in both of the blocks that define the edge.  Variables that are live
-across an edge between two tasks must be protected by \sys's memory
-virtualization mechanism and each live, inter-task variable increases the run
-time overhead for a boundary placement.  \sys's analysis tries to minimize the
-overhead of protected variables by placing boundaries to minimize the number of
-inter-task live variables.
+accessible in both of the blocks that define the edge (i.e., predecessor and
+successor).  Variables that live across an edge between two tasks must be
+protected by \sys's memory virtualization mechanism and each live, inter-task
+variable increases the run time overhead for a boundary placement.  \sys's
+analysis tries to minimize the overhead of protected variables by placing
+boundaries to minimize the number of inter-task live variables.
 
-{\noindent \bf Loop splitting cost.}
-and if the edge is within any loop, the {\em loop count}. 
-Also the compiler avoids placing boundaries at an edge that is
-within a loop whose loop count is big.
+{\noindent \bf Loop cost.}
+Loops are a challenge for \sys because the analysis must account for the cost
+of multiple iterations of the loop body.  \sys assumes that most loops are
+annotated with a static loop bound, which is common in embedded code.  For
+loops that do not have a static bound, \sys conservatively assumes the maximum
+possible loop bound, based on the type of the loop induction variable.
 
-After constructing the graph and assigning the information on a cost and live
-variables to each vertex and edge, respectively, the compiler begins with
-analyzing the program loops. The compiler first identifies all loops and then
-multiplies the loop count by the cost of a vertices within the loop. The reason
-is, if the loop is inside a task it will use up energy that is roughly
-multiplied by the loop count compared to code that is outside of a loop. If the
-resulting cost of a loop is too large\todo{how do you decide on the acceptable
-loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop,
-and reverts the change to the cost of the vertices inside. It means that if the
-loop is too large to fit in one task, each iteration should be a separate task
-execution.
+The analysis tries to minimize the dynamic overhead of traversing task
+boundaries by avoiding placing a task boundary on a loop with a large bound,
+when possible. At the same time, the analysis must avoid placing a loop in a
+single task if the total energy cost of all of its iterations is likely to
+exceed the device's energy buffer.  \sys first multiplies the estimated energy
+cost of the blocks in the loop body by the loop bound to determine estimate the
+cost of the loop's execution.  If the loop's cost exceeds the device's energy
+buffer, \sys can place a task boundary on the loop's back-edge, converting each
+iteration into its own task.
 
-If a program has an unbounded loop (such as {\tt while} statement), the compiler conservatively assumes the loop count to be very large.
+%of  The reason
+%is, if the loop is inside a task it will use up energy that is roughly
+%multiplied by the loop count compared to code that is outside of a loop. If the
+%resulting cost of a loop is too large\todo{how do you decide on the acceptable
+%loop cost?}{Kiwan}, the compiler places boundary at the back edge of the loop,
+%and reverts the change to the cost of the vertices inside. It means that if the
+%loop is too large to fit in one task, each iteration should be a separate task
+%execution.
+
+%If a program has an unbounded loop (such as {\tt while} statement), the compiler conservatively assumes the loop count to be very large.
 The programmer can prevent this by annotating maximum possible loop count of such unbounded loop if the programmer has the information.\todo{'the' or 'this'?
@@
 \subsubsection{Task Boundary Placement}
-\textbf{Tasks Preparation:} After initial boundary placemen with loops analysis, the compiler places additional boundaries if some tasks are still too large.
-The \sys compiler starts from the beginning of the main function and traverse through the graph in a depth-first search manner.
-If any part of the graph is found that is too long to fit in the energy bound, it means a boundary must be placed somewhere within.
-The subset of graph that needs boundary placement within is called a {\em boundary-candidate}.
 
-As discussed above, when placing boundary we take three optimization criteria into account: \{O1, O2, O3\}.
-We now introduce a heuristic that consider all these criteria and choose the local optimum\todo{Optimum of what?}{Kiwan}.
-The heuristic calculates the {\em score}, $S$, of each edge inside the boundary candidate as
+After completing its cost analysis of a program, \sys inserts task boundaries
+into the CFG in two steps.  The first step is to insert task boundaries on the
+back-edges of all loops that have an energy cost greater than the device's
+energy buffer capacity, producing an initial boundary placement. The second
+step is to insert task boundaries at additional points in the code,
+sub-dividing tasks that have an estimated cost that exceeds the energy buffer's
+capacity.
+
+To sub-divide these tasks, \sys does depth-first traversals of all paths in the
+CFG, ignoring back-edges. Traversals start from the beginning of the main
+function, or from an existing task boundary.  \sys accumulates the cost of
+basic blocks during the traversal until reaching another task boundary or an
+exit node in the CFG.   At the end of a traversal, if the accumulated cost
+exceeds to device's energy capacity \sys must split the traversed path with a
+task boundary.  Each edge on the path is a candidate location for a boundary.
+
+The analysis uses a heuristic to compute a score for each candidate and places
+a boundary at the candidate with the highest score. The heuristic tries to
+avoid placing boundaries on loops, to minimize the number of live variables
+that must be protected across the new boundary, and to maximize the number of
+operations along the path.  A boundary candidate's score is computed as:
 %
 \begin{equation}
-S = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
+\text{score} = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
 \end{equation}
 %
-where $N_{\text{IR}}$ is a normalized IR count that the task will end up having if the compiler places boundary at the specific edge, $N_{\text{PV}}$ is the number of protected variables that will be newly introduced by placing boundary at the edge, and choose the edge with the highest score, $N_{\text{loop}}$ is the loop count and $w_{0},w_{1}\in [0,1]$ are pre-selected weights before compilation\footnote{Naturally, different $w_{0}, w_{1}$ will have different effect on the compiler result.
-We will study this effect in Section~\ref{sec:results_compiler}.}.
-Both $N_{\text{IR}}$ and $N_{\text{PV}}$ are normalized within the boundary candidate, since usually IR count is much larger and has higher variance, thus has dominating effect if not normalized.
-It is better to have higher IR count within the task since it means the task size is similar to the size given by the programmer (viz.
-C1).
-On the other hand, change in protected variable is better to be small (viz.
-C2), so we penalize it by assigning a negative value to $N_{\text{PV}}$.
-If the edge is within a loop, the score is divided by $N_{\text{loop}}$.
-It will result in not placing boundary within loops unless other parts of the formula really pays off (viz.
-C3)\todo{This sentence is not clear: what "pays off" means in this context?}{Kiwan}.
-Note that complier never makes a decision that will end up in having more IR count than the programmer specified because at the beginning boundary candidate is chosen to have less IR than the programmer specified number.
+where $N_{\text{IR}}$ is the length of the path being traversed up to the
+candidate, $N_{\text{PV}}$ is the number of variables live variables at the
+candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops
+(i.e., nested loops) containing the boundary candidate. \sys normalizes $N_{\text{IR}}$ to the
+total number of instructions along the path being traversed. The weights, $w_{0},
+w_{1}$ are between zero and one, and are hand-tuned parameters that define the
+relative importance of $N_{\text{IR}}$, $N_{\text{PV}}$, and $N_{\text{loop}}$.
+We will study the effect of weights in Section~\ref{sec:results_compiler}.
 
-Figure~\ref{fig:comp_2} shows example of such boundary placement on graph. The first three vertices will become one task while the rest will become another (see Figure~\ref{fig:comp_3}).
-
-\textbf{Ensuring Single Entry:}
-
-The compiler also calculates a dominance relation \todo{Define dominance
-relation}{Kiwan} between vertices. This information is used to ensure task's
-single-entry property.  After placing a boundary, the compiler ensures that the
-newly made task is single-entried by checking dominance relation. That is,
-every vertex within a task should be dominated by the entry vertex of the task.
-If this rule is violated, the compiler places additional boundaries to make
-sure that every task is single entried.
+After placing task boundaries, the compiler must ensure that all tasks are
+single-entry.  To do so, \sys computes the dominator tree of the CFG and checks
+that every vertex within a task is dominated by the entry vertex of the task.
+If this dominance condition is unmet for a CFG node, the compiler places
+an additional task boundary at that node, meeting the condition.
+%
+Figure~\ref{fig:comp_2} illustrates boundary placement: the first three
+vertices make up one task and the remainder become a second task.
 
@@
 
-In the compiler transform pass, a complier executes sub-tasks described below.
+After placing task boundaries during analysis, \sys's transformation analysis
+restructures the code into tasks defined by the placed boundaries. The analysis
+traverses the CFG starting from an arbitrary task boundary.  On reaching a
+subsequent boundary, the transformation analysis identifies all paths between
+the two task boundaries and copies those paths into a new task function.  The
+analysis continues until all paths are contained within a task.
 
-\textbf{Declaring and Migrating Codes to Tasks:} After the boundary placement is finished, the analysis pass passes the information to transform pass.
-The transform pass makes new task, traverse through the graph, and copies every code within the vertices it encounters to the new task until it meets the task boundary.
-At the task boundary, it makes another task and starts copying again, and repeat this action until every code is moved to form a task.
+After forming tasks, variables that are live beyond the end of the task become
+protected, which requires renaming those variables and updating references to
+renamed variable names.  
+%
+For example, Figure~\ref{fig:comp_2} shows a boundary placed within the live
+range of {\tt b}; {\tt b} becomes protected and accesses to {\tt b} should be
+redirected to {\tt p\_b} (as shown in Figure~\ref{fig:comp_3}).  {\tt a} is
+task-local and need not be protected, but only declared in the task. Constants
+and read-only variable are never protected.
+%
+When writes in multiple, newly formed tasks may update the same location (i.e.,
+at a PHI node in the CFG), \sys re-writes the accesses to refer to the same global
+name.   Subsequent reads simply read the protected variable.
 
-\textbf{Renaming Variables:} On copying the code, the variables should be renamed.
-If a variable is alive across multiple tasks, it needs to become a global-protected variable, and every access to the variable should be redirected.
-If a variable's live range is local to the task, it does not need to go through such renaming.
-For example, in Figure~\ref{fig:comp_2}, boundary is placed within the live range of variable {\tt b} (Note that the red cross overlaps the blue arrow).
-Thus variable {\tt b} needs to become a protected variable {\tt p\_b} and every access to {\tt b} should be redirected to {\tt p\_b} (Figure~\ref{fig:comp_3}).
-On the other hand, {\tt a}'s live range is always local to a single task.
-In such case, it need not become a protected variable.
-Note, however, still new declaration of {\tt a} should be inserted in {\tt task1} (Figure~\ref{fig:comp_3}).
-As an improvement to the complier, constants and read-only variable never becomes protected.
+After forming tasks, some branches in a task may refer to targets in another
+task. \sys patches the CFG to re-write those branches to instead be \transition
+statements that transition to the entry point of the other task. Such a
+``broken'' branch will always refer to a task's entry node because \sys ensures
+that all tasks have only one entry.
 
-\textbf{Inserting \texttt{os\_jump} Calls:} After renaming is done, the compiler detects the broken {\tt branch} instructions\todo{Define protected branch instruction}{Kiwan} which tries to branch to a basic block that is inside different task and change it to an {\tt os\_jump} call (Figure~\ref{fig:comp_3}).
-Note that in this figure we showed the name of the next task to be executed for readability, however {\tt os\_jump} call takes offset number as an argument.
-In reality the compiler statically calculates and passes the offset.
+The compiler re-writes accesses to protected variables to use \sys's memory
+virtualization API. Pointers pose a challenge to re-writing protected accesses
+because unprotected pointers may point to protected data.  Leveraging LLVM's
+{\tt basic} pointer alias analysis the compiler conservatively re-writes all
+accesses that {\em may} read or write protected data.
 
-\textbf{Dealing With PHI Nodes:} Whenever more than two different control flows try to write different value in the same memory location, LLVM makes a unique IR called {\em PHI Node} when reading the memory location~\cite{}.
-However if there is a boundary between the memory write and the PHI Node, the compiler will not work.
-Thus, on each of this occurrence, \sys compiler declares a special protected variable and makes the write relocated to the protected variable.
-Then the PHI Node becomes simply a read to a protected variable.
+%After task generation is done, the compiler generates new main function and
+%peppers the system with necessary codes that is needed by \sys (such as code
+%for evoking {\tt os\_sceduler}). Also it removes the code that became stale
+%(such as the body of a function call, for all the function calls are now
+%inlined).
 
-\textbf{Inserting API Call for Paging:} To run \sys, every read or write to a protected variable should call related API for page swapping and dirty page marking.
-The compiler conservatively inserts such calls before a {\em possible} read or write to the protected variable.
-We use the basic pointer alias feature provided by the LLVM library.
-We are being conservative, so it is always correct, albeit not always efficient.\todo{We need to say more about the level of this inefficiency and how it is measured}{Kiwan}
-
-\textbf{Creating New Main Function and Cleaning Unnecessary Codes:} After task generation is done, the compiler generates new main function and peppers the system with necessary codes that is needed by \sys (such as code for evoking {\tt os\_sceduler}).
-Also it removes the code that became stale (such as the body of a function call, for all the function calls are now inlined).
-
-\subsection{Compiler Design Decisions}
-\label{sec:compiler_limitations}
-
-To be able to implement \sys compiler certain critical assumptions had to be made.
-
-\begin{enumerate}
+%\subsection{Compiler Design Decisions}
+%\label{sec:compiler_limitations}
 %
-	\item \textbf{Conservative handling of unbounded code sections:} When a C program contains an unbounded loop, the \sys compiler makes a conservative decision of regarding its loop count to be very large.
+%To be able to implement \sys compiler certain critical assumptions had to be
+%made. 
+%
+%\begin{enumerate}
+%	%
+%	\item \textbf{Conservative handling of unbounded code sections:} When a C program contains an unbounded loop, the \sys compiler makes a conservative decision of regarding its loop count to be very large.
 However, if a programmer annotates the expected loop count manually the overhead can be relieved.
-	%
-	\item \textbf{Conservative handling of pointers:} When a C program contains pointer, due to the limitation of pointer alias, the compiler makes the conservative decision.
+%	%
+%	\item \textbf{Conservative handling of pointers:} When a C program contains pointer, due to the limitation of pointer alias, the compiler makes the conservative decision.
 That is, it may consider a variable's live range longer than actual, or it may insert API call for page swap even when it is unnecessary.
 However, these overhead can be reduced by better pointer aliasing algorithm.
-	%
-	\item \textbf{Function call inlining:} \sys compiler inlines all function call prior to analysis for simplicity.
+%	%
+%	\item \textbf{Function call inlining:} \sys compiler inlines all function call prior to analysis for simplicity.
 However it can introduce code bloat.
@@
 However we did not implemented such feature for simplicity of the system.
-	%
-	\item \textbf{LLVM IR instruction count to estimate the energy usage:} The compiler uses the number of LLVM IR count as an indicator for energy usage in trying to meet correctness constraint C1 and optimization constraint O1.
+%	%
+%	\item \textbf{LLVM IR instruction count to estimate the energy usage:} The compiler uses the number of LLVM IR count as an indicator for energy usage in trying to meet correctness constraint C1 and optimization constraint O1.
 Intuitively, the number of LLVM IR correlates with the energy usage since more instructions will lead to more energy usage.
@@
 Also, since the estimation of energy usage of each basic block is perpendicular to the compiler itself, it is simple to plug in a better energy model if there is one.
-	%
-	\item \textbf{Limitation of the boundary placement heuristics:} The boundary placement strategy is a heuristic that chooses the local optimum.
+%	%
+%	\item \textbf{Limitation of the boundary placement heuristics:} The boundary placement strategy is a heuristic that chooses the local optimum.
 However, selecting every local optimum might not be the global optimal solution.
@@
 Since current heuristic is already showing good result that is comparable to human-written code, refer to Section~\ref{sec:results_coalescing}, we leave such optimization as a future work.
-\end{enumerate}
+%\end{enumerate}

### 2017-08-04 paper/figures/sram-buffer.tex
done with an aggressive pass over compiler

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning} 
\usetikzlibrary{fit}

\tikzset{
	%Define standard arrow tip
	>=stealth',
	%Define style for boxes
	punkt/.style={
		rectangle,
		rounded corners,
		draw=black, very thick,
		text width=4.5em,
		minimum height=1.5em,
		text centered},
	% Define arrow style
	pil/.style={
		->,
		thick,
		shorten <=2pt,
		shorten >=2pt,}
}

\begin{document}

	\begin{tikzpicture}[node distance=1cm, auto,]

	\node[punkt,minimum height=3em,fill=orange!10] (task1) {y = x+1 
		\\ z = y+1 \\ x = z+1};
	\node[below = 0.1cm of task1] () {Task};

	\node[draw,very thick,dotted,minimum width=7em, fill=blue!5,minimum height=10em, right = 3cm of task1] (sram) {};
	\node[below = 0.1cm of sram] () {SRAM Cache};

	\node[punkt, below = -3cm of sram,fill=red!30] (x-sram) {x};
	\node[punkt, below = 0.3cm of x-sram,fill=blue!30] (y-sram) {y};
	\node[punkt, below = 0.3cm of y-sram,fill=blue!30] (z-sram) {z};

	\node[right = 4.5cm of sram] (ref) {};
	\node[draw,very thick,dotted,minimum width=7em, fill=red!5,minimum height=10em, above = 1cm of ref] (fram1) {};
	\node[below = 0.1cm of fram1] () {FRAM (original)};

	\node[draw,very thick,dotted,minimum width=7em, fill=green!5,minimum height=10em, below = 1cm of fram1] (fram2) {};
	\node[below = 0.1cm of fram2] () {FRAM (versioned)};

	\node[punkt, below = -3cm of fram1,fill=red!70] (x1-fram) {x};
	\node[punkt, below = 0.3cm of x1-fram,fill=blue!50] (y1-fram) {y};
	\node[punkt, below = 0.3cm of y1-fram,fill=blue!50] (z1-fram) {z};

	\node[punkt, below = -3cm of fram2,fill=blue!30] (x2-fram) {x};
	\node[punkt, below = 0.3cm of x2-fram,fill=blue!30] (y2-fram) {y};
	\node[punkt, below = 0.3cm of y2-fram,fill=blue!30] (z2-fram) {z};

	\path[every node/.style={font=\sffamily\small}]
	([yshift=0.3cm] task1.east) edge [->, dotted, thick, bend left] node {(1) get x} (x-sram.west)
	(x-sram.east) edge [->, dotted, thick, bend left] node[xshift=0.5cm] {(2) read x} (x1-fram.west)

	([yshift=0.3cm]task1.east) edge [->, dotted, thick, bend right] node[xshift=-0.5cm] {(3) set y} (y-sram.west)

	(task1.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(4) set z} (z-sram.west)
	([yshift=-0.4cm]task1.east) edge [->, dotted, thick, bend left] node[xshift=1.3cm,yshift=-0.4cm] {(5) set x} (x-sram.west)

	(x-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(6) commit-I x} (x2-fram.west)
	(z-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(7) commit-I z} (z2-fram.west)
	(y-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(8) commit-I y} (y2-fram.west)

	(x2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(9) commit-II x} (x1-fram.east)
	(y2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(10) commit-II y} (y1-fram.east)
	(z2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(11) commit-II z} (z1-fram.east)
	;

	\end{tikzpicture}

\end{document}

### 2017-08-04 paper/evaluation.tex
Text added describing result on page swaps (needs second iteration)

@@
 \begin{table}[t]
+	\centering
+	\footnotesize
 	\begin{tabular}{|c|c|c|}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
-	\caption{Initial figure about page size overhead. \todo{Figure placeholder---will be replaced with final one}{Amjad}}
-	\label{fig:IPOSPerformance}
-\end{figure}
-
-\begin{figure}
-	\centering
 	\includegraphics[width=\columnwidth]{figures/chain_vipe}
@@
 
-Expected results:
-\begin{itemize}
-	\item Sensitivity plot---normalized page swapping cost ("bathtub" curve) in terms of execution time: compare against Chain and full double buffering approach; show number of page swaps---do the experiment for all applications
-\end{itemize}
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
+	\caption{Page size overhead.\todo{Figure placeholder---will be replaced with final one}{Amjad}}
+	\label{fig:IPOSPerformance}
+	\label{fig:page_size}
+\end{figure}
 
-\todo{Provide argument for page swapping using hardware support}{Brandon}
+Figure~\ref{fig:page_size} presents the result on the effect of page swap size on the execution time of the programs executed by \sys.
+Overhead has been measured in the MCU clock cycles\todo{WISP?
+EDB?
+CCS?}{Amjad}.
+Result has been normalized \todo{To what?}{Amjad}.
+A subset of applications has been selected for evaluation.
+Additionally, page faults have been measured.
+\todo{Comparison with Chain and double buffering?}{Amjad} The result show clearly a bathtub-shape bar plot for each application, demonstrating that for each application there is always a page size that minimizes \emph{both} execution time and page swaps.
+For all applications (except for Sort) this value equals to 64 pages\todo{And one page is?}{Amjad}.
+The shape of each figure can be explained as follows.
+For small page swaps (<64) the execution time is dominated by the \emph{frequency} of page swaps (each swap takes time, as shown in Figure~\ref{fig:dmaTimeEnergy}).
+On the other hand, in the opposite case execution time is dominated by the time of swapping one large page.
+Another interesting observation is the non-symmetric shape of the bar plot.\todo{Why do we see non-symmetry}{Amjad}
 

### 2017-08-04 paper/conclusions.tex
Placeholder for Conclusions (will require Brandon's hand and eye later)

--- 2017-07-27 paper/relatedwork.tex
@@
-\section{Intermittenly-Powered Devices}
-
-\section{Intermittent Execution Environments}
+This paper presented \sys: a new task-based runtime for intermittently-powered devices that does not require any hardware support.
+\sys core new features are: (i) automatic task-generation---removing programmer intervention in the program re-composition to the intermittent domain \emph{completely}, (ii) memory virtualization---reducing the execution speeds by \todo{provide numbers}{Przemek} of broad set of benchmarks executed on popular energy-harvesting embedded platform, and (iii) (adaptive) task coalescing---enabling fastest execution \todo{provide numbers}{Przemek} despite changing energy arrivals and enabling re-use of the same code on intermittent systems with varying energy reservoirs.
+\sys, with its task creation/execution automation, responses to the need that all existing predecessors of \sys (chronologically, Mementos, DINO, Chain, Alpaca) strongly advocated.
+We believe that \sys is a next major step in enabling computing community to enter the intermittent-power domain without big learning curve investments.
+Next logical step in \sys design is the introduction of features needed to implement first fully-operational operating system for transiently-powered devices.\todo{find a more convincing reason for a follow-up work}{Brandon} \todo{Revise this section}{Brandon}

### 2017-08-04 paper/paging.tex
Text added describing result on page swaps (needs second iteration)
added a todo

@@
 
-\begin{figure}
-	\centering
-	%\includegraphics[width=\columnwidth]{figures/}
-	\caption{\sys memory virtualization architecture.\todo{Decide if we really need this figure (we already have a lot of figs)}{Brandon}}
-	\label{fig:virtualization_architecture}
-\end{figure}
+%?\begin{figure}
+%?	\centering
+%?	%\includegraphics[width=\columnwidth]{figures/}
+%?	\caption{\sys memory virtualization architecture.\todo{Decide if we really need this figure (we already have a lot of figs)}{Brandon}}
+%?	\label{fig:virtualization_architecture}
+%?\end{figure}
 
@@
 	\subfloat[DMA Copy]{\includegraphics[width=0.49\columnwidth]{figures/dma}}
-	\caption{.\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
+	\caption{.\todo{Explain HW/SW setup; make fonts larger, unify legends and axes.
+We only need to show the right side; this figure shows that we operate out of SRAM and page to FRAM; the copying mechanism is peripheral to the main idea.}{Amjad}}
 	\label{fig:sram_vs_dma}

### 2017-08-04 paper/compiler.tex
removed unnecessary todo about single-entry; text reflects why we need it now

@@
 \label{sec:compiler_analysis_pass}
-
-\todo{Define single entry}{Kiwan}
-\TODO{Brandon: the term "single-entry" is well-known in the community to mean that there is a single control-flow edge entering the task.
-Thinking about this a bit, is there a simple reason why we cannot have multiple-entry tasks?
-A totally plausible task graph could be like A->B B->C C->B C->A; here, B has two predecessors.
-Is this fundamental?
-Or an analysis convenience?
-I've left it as "single-entry" for now}
 

### 2017-08-04 paper/compiler.tex
added sensitivity result

--- 2017-08-04 paper/compiler.tex
@@
-\sys provides compiler support for \emph{automatic task generation} that simplifies programming writing task-based code by automatically translating C code into a task-based executable.
+\sys provides compiler support for \emph{automatic task generation} that simplifies writing task-based code by automatically translating C code into a task-based executable.
 Translating C code into task-based code automatically requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
@@
 \subsubsection{Task Boundary Placement}
+\label{sec:compiler_boundary}
 
@@
 relative importance of $N_{\text{IR}}$, $N_{\text{PV}}$, and $N_{\text{loop}}$.
-We will study the effect of weights in Section~\ref{sec:results_compiler}.
+We will study the effect of weights in Section~\ref{sec:compiler_heuristics}.
 
@@
 
+\subsection{Boundary Placement Heuristics}
+\label{sec:compiler_heuristics}
+
+By studing different weights discussed in Section~\ref{sec:compiler_boundary},
+we discovered that there is no one global weights that are optimal for every
+application, and more importantly, the weights do not matter much and
+result in less than 5\% run time difference. The result is mainly because
+no matter what the weight is, we place boundaries at obvious points (such as back-edge of a
+large loop) and we never place boudaries at obviously-bad points (such as
+inside a loop when the loop is not large). Since these obvious decision
+dominates the structure of tasks and heuristics only give minute perturbation,
+the result does not gets affected significantly.
+As an extreme, for small apps like CEM, only placing boundaries at obvious points are enough to
+determine the entire task structure. It never needs any additional heuristics.
+We used weight of 
+$w_{0} = 0.5$ and $w_{1} = 0.5$ for the experiment result in Section~\ref{sec:experiment}.
+
 %After task generation is done, the compiler generates new main function and

### 2017-08-04 paper/systemdescription.tex
VIPER block diagram placeholder add
placeholder for VIPER API figure
change or order of contributions in system description
added placeholdef for viper syntax in ststem description

@@
 	\centering
-	%\includegraphics[width=0.25\columnwidth]{figures/}
+	\includegraphics[width=\columnwidth]{figures/viper_block_diagram}
 	\caption{\sys top-level view.}
@@
 \begin{itemize}
+	\item \textbf{Compiler-supported automatic task generator}: details will be provided in Section~\ref{sec:compiler}
+	\item \textbf{Memory virtualization}: details will be provided in Section~\ref{sec:memory_virtulaization}
 	\item \textbf{Task coalescing}: details will be provided in Section~\ref{sec:task_coalescing}
-	\item \textbf{Memory virtualization}: details will be provided in Section~\ref{sec:memory_virtulaization}
-	\item \textbf{Compiler support}: details will be provided in Section~\ref{sec:compiler}
 \end{itemize}
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/taskification_example}
+	\caption{\sys API.}
+	\label{fig:system_overview}
+\end{figure}
+
 \subsection{Programming Model}
@@
 
+\begin{table}
+	\centering
+	\footnotesize
+	\begin{tabular}{|c|c|}
+		\hline
+		Keyword & Description\\
+		\hline\hline
+		\texttt{RP} & Read-protected variable\\
+		\texttt{WP} & Write-protected variable\\
+		\texttt{\_\_task\_\_ foo(){...}} & Task declaration\\
+		\texttt{addTask()} & Build internal data structure for executor \\
+		\texttt{run\_tasks()} & Task executor \\ %scheduler
+		\texttt{next\_task()} & Transition to a new task\\
+		%\texttt{PageFault()} & --- \\
+		%\texttt{P()} & ---\\
+		\hline
+	\end{tabular}
+\caption{Summary of \sys syntax: top---user-specific variables}
+\label{tab:viper_syntax}
+\end{table}
+
 \subsection{Semantics and Execution Model}

### 2017-08-04 paper/compiler.tex
added some specification about normalization

--- 2017-08-04 paper/compiler.tex
@@
 candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops
-(i.e., nested loops) containing the boundary candidate. \sys normalizes $N_{\text{IR}}$ to the
-total number of instructions along the path being traversed. The weights, $w_{0},
-w_{1}$ are between zero and one, and are hand-tuned parameters that define the
-relative importance of $N_{\text{IR}}$, $N_{\text{PV}}$, and $N_{\text{loop}}$.
-We will study the effect of weights in Section~\ref{sec:results_compiler}.
+(i.e., nested loops) containing the boundary candidate. \sys normalizes
+$N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates.
+The weights, $w_{0}, w_{1}$ are between zero and one, and are
+hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
+$N_{\text{PV}}$, and $N_{\text{loop}}$.  We empirically determined these weight
+values by measuring the runtime for different weight values across
+applications.  We found that varying weights varies runtime by only around 5\%
+in most cases and in our experimental evaluation, we used $w_{0} = w_{1} =
+0.5$.
 

### 2017-08-05 paper/background.tex

@@
 
-Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into a collection tasks.
-A task is a function with no caller that contains arbitrary computation, sensing, and communication.
-A programmer describes the sequence in which tasks may execute,
-creating a task graph.
-The flow of control from task to task happens at programmer demarcated points and may conditionally depend on program values.
-Task-based programming abstractions guarantee that tasks exhibit behavior as though they executed {\em atomically}, regardless of arbitrary power failures.
-Task-based runtime systems tolerate failures by repeatedly attempting to execute a task until it completes.
-Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent (i.e., have the same effect when repeatedly executed).
-The key to idempotence is ensuring that updates to non-volatile state made by an interrupted, partial execution of a task are never visible to a subsequent (potentially completing) re-execution of the task.
+Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the
+programmer to decompose their program into a collection tasks. A task is a
+function with no caller that contains arbitrary computation, sensing, and
+communication. A programmer describes the sequence in which tasks may execute,
+creating a task graph. The flow of control from task to task happens at
+programmer demarcated points and may conditionally depend on program values.
+Task-based programming abstractions guarantee that tasks exhibit behavior as
+though they executed {\em atomically}, regardless of arbitrary power failures.
+Task-based runtime systems tolerate failures by repeatedly attempting to
+execute a task until it completes. Task-based runtime systems ensure
+task-atomic semantics by ensuring that repeated task executions are idempotent
+(i.e., have the same effect when repeatedly executed). The key to idempotence
+is ensuring that updates to non-volatile state made by an interrupted, partial
+execution of a task are never visible to a subsequent (potentially completing)
+re-execution of the task.
 
-There are several run time strategies to ensuring task idempotence.
-One approach, utilized by Alpaca~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{rachet} did for checkpoints), execute the task using buffered copies of those data, and commit the buffered copies before completing the task.
-Another strategy is to statically create multiple versions of non-volatile data shared by tasks and ensure by construction that no task reads and writes the same non-volatile memory location~\cite{chain}.
-Regardless of the idempotence strategy, the key to task-based execution is that fixed, statically defined tasks execute as though atomic, completing after zero or more attempts.
-\TODO{Brandon left off here.}
-
-Despite the benefit of the modular model certain limitations are still present that must be overcome and which inspired us to design \sys. Let us look at these problems in detail.
+There are several run time strategies to ensuring task idempotence. One
+approach, utilized by Alpaca~\cite{alpaca}, is to identify non-volatile data
+involved in WAR dependences in a task (like DINO~\cite{dino} and
+Ratchet~\cite{rachet} did for checkpoints), execute the task using buffered
+copies of those data, and commit the buffered copies before completing the
+task. Another strategy is to statically create multiple versions of
+non-volatile data shared by tasks and ensure by construction that no task reads
+and writes the same non-volatile memory location~\cite{chain}. Regardless of
+the idempotence strategy, the key to task-based execution is that fixed,
+statically defined tasks execute as though atomic, completing after zero or
+more attempts.
 
@@
 	\footnotesize
-	\begin{tabular}{|c|c|c|c|}
+	\begin{tabular}{l|l}
+		Model & Data Copied to/from NVRAM\\
 		\hline
-		{~} & CP & SZ & NV \\
-		\hline\hline
-		\!\!Mementos~\cite{mementos}\!\! & \!\!$N_2\leq N_1$\!\! & \!\!R + S\!\! & \!\!R + S\!\! \\
-		\!\!DINO~\cite{dino}\!\! & $N_2$\!\! & \!\!R + W\!\! & \!\!R + W\!\! \\
-		\!\!Chain~\cite{chain}, Alpaca~\cite{alpaca}\!\! & \!\!$N_2$\!\! & P\!\! & \!\!2W + P\!\!\\
-		\!\!Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017}\!\! & $N_3\geq N_2\geq N_1$\!\! & \!\!R\!\! & R\!\!\\
-		\hline 
+		Mementos~\cite{mementos}                             & Reg. + Stack     \\
+		DINO~\cite{dino}                                     & Reg. + WAR vars. \\
+		Chain~\cite{chain}                                   & PC   + Channel data\\
+		Alpaca~\cite{alpaca}                                 & PC   + WAR vars. \\
+		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main mem.)             \\
 	\end{tabular}
-	\caption{Cost of various checkpointing methods; Symbols---R: register, S: stack, W: WAR-dependent variables, P: program counter, $N_1, N_2, N_3$: relative number of checkpoints for each method, CP: no.
-checkpoints, SZ: checkpoint size, NV: copy to non-volatile memory size.}
+	\caption{Memory consistency enforcement overheads of various intermittent execution models.
+Reg.
+includes the entire register file, PC is the program counter, Channel data are variables explicitly task-shared by the programmer, and WAR vars.
+are variables involved in WAR dependences.}
 	\label{table:chechpoint_comparison}
@@
 
-\paragraph{Problem 1---Idempotent task is also costly}
+\subsubsection{Costs of Task-based Models}
 
-The most important modular and intermittent models are compared in Table~\ref{table:chechpoint_comparison}\footnote{This table generalizes discussion from~\cite[Sec.
-2.4]{alpaca}.}.
-Although we concluded that sequential execution model introduces a significant overhead, a modular approach also introduces certain execution penalty.
-Whatever the approach, the energy cost of a single checkpoint/idempotent task is $w\text{CP(SZ+NV)}$, where $w$ is the cost of a single word copy into a non-volatile memory.
+Task-based intermittent execution models ensure that programs execute correctly
+despite arbitrary power failures and in doing so, incur several costs that are
+the motivation for \sys.
+
+\paragraph{Memory Consistency Enforcement Overhead}
+
+Intermittent execution models incur an overhead because they must checkpoint
+data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or
+privatize and commit data~\cite{alpaca}.
+%
+Table~\ref{table:chechpoint_comparison} compares the overheads for several
+recent intermittent execution models (generalizing discussion from~\cite[Sec.
+2.4]{alpaca}.)  The mechanism responsible for the overhead of each model
+varies, but all the bulk of overhead in all models is a manipulating
+non-volatile memory.  Checkpointing moves volatile memory and registers to and
+from non-volatile memory, channel accesses directly manipulate non-volatile
+data~\cite{chain}, privatization copies data from and commits data to
+non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet}
+require all of memory to be non-volatile.
 
@@
 
-We have executed an experiment to measure the cost of read/write operations of $w$ to volatile and non-volatile memory.
-The result is given in Figure~\ref{fig:framEnergy}.
-Four applications were developed that \emph{access} the FRAM and SRAM of the MSP430FR5969 microcontroller~\cite{msp430datasheet}.
-The interference free debugger (EDB)~\cite{edb} is used as the energy measurement tool.
-The EDB probes the energy buffer before and after accessing the FRAM/SRAM 1600 times.
-This large number of FRAM access is used to increase the reliability of the results generated by the used measurement tool (e.g.
-reducing the effect of the quantization error).
-The energy buffer is charged, using the EDB, to $\approx$ 2.45\,V \textbf{only} at the beginning of the execution and of the programs.
-The write operations were performed by writing literal values, random numbers, to the memory.
-However, a read operation must be followed by a write operation.
-Therefore, we chose to write a read value to SRAM.
-Based on our measurements we conclude that a FRAM access equals $\frac{3}{2}$ SRAM access~\footnote{We remark that TI already mentions~\cite{ti_fram_faq} that FRAM write access is more energy expensive than SRAM write access.
-However, TI does not quantify this energy overhead.}.
-Therefore, reducing FRAM access is desirable when developing energy limited software solution---\emph{which is in sharp contrast to the main goal of all checkpoining solutions}.
+To concretely understand the cost of these manipulations of non-volatile
+memory, we directly measured the energy consumption of volatile (SRAM) and
+non-volatile (FRAM) memory operations for the
+MSP430FR5969~\cite{msp430datasheet} microcontroller. We used EDB~\cite{edb} to
+monitor the voltage drop on an enegy buffering capacitor that powers the device
+across 1600 read and write accesses to each type of memory, randomizing
+locations to avoid caching effects.  Figure~\ref{fig:framEnergy} summarizes
+these measurements, illustrating that a non-volatile memory access consumes
+around 1.5x as much energy as a volatile memory access in this device.  In an
+intermittently-powered device, the increased energy corresponds to decreased
+run time between power failures.  The overheads imposed by relying on
+non-volatile memory motivate \sys, which uses memory virtualization to operate
+primarily using volatile SRAM.
 
-The above result reassures the intuitive observation that a decrease of CP (Table~\ref{table:chechpoint_comparison}) decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
-On the other hand, this increases task/checkpoint re-execution time.
-The ideal solution is a runtime that \emph{adaptively} changes task size during runtime, which however exposes a second problem.
+The above result reassures the intuitive observation that a decrease of CP
+(Table~\ref{table:chechpoint_comparison}) decreases the energy cost associated
+with each state preservation method and reduces operation cost which in most
+simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$,
+where PL is the cost of program execution without checkpointing. On the other
+hand, this increases task/checkpoint re-execution time. The ideal solution is a
+runtime that \emph{adaptively} changes task size during runtime, which however
+exposes a second problem.
 
-\paragraph{Problem 2---Modular execution models are not adaptive}
-
-A modular intermittent execution model completely depends on a programmer's estimation which mostly results in a sub-optimal code division.
-Moreover, this model can only consider a single hardware configuration and it does not take environment variability into considerations.
-Each battery-less platform has different storage sizes, refer to Table~\ref{table:capacitor}, which causes different charge/discharge periods.
-This result in different execution times.
-In the worst case the size of the capacitor might be too small to execute one task written by the programmer.
-This calls for the ability of the modular execution runtimes to \textbf{merge} multiple small tasks into a larger one or \textbf{split} a large task into a set of small ones on the fly (post-compilation time).
-Unfortunately the following features prevent us to perform this task using existing runtimes~\cite{chain,alpaca}.
+\paragraph{Fixed-size Tasks are Inflexible}
+\TODO{Brandon left off here - I've cut some bullets below that I did not understand.
+I think the key point here is that fixed task size only works for one cap size; inefficient with a larger cap, unimplementable with a smaller one.
+There might be an efficiency argument, but it's tough to make because we still require some memory consistency mechanism (in our case paged FRAM), which has an overhead.}
+A modular intermittent execution model completely depends on a programmer's
+estimation which mostly results in a sub-optimal code division. Moreover, this
+model can only consider a single hardware configuration and it does not take
+environment variability into considerations. Each battery-less platform has
+different storage sizes, refer to Table~\ref{table:capacitor}, which causes
+different charge/discharge periods. This result in different execution times.
+In the worst case the size of the capacitor might be too small to execute one
+task written by the programmer. This calls for the ability of the modular
+execution runtimes to \textbf{merge} multiple small tasks into a larger one or
+\textbf{split} a large task into a set of small ones on the fly
+(post-compilation time). Unfortunately the following features prevent us to
+perform this task using existing runtimes~\cite{chain,alpaca}.
 
@@
 
-\begin{itemize}
-	\item \textbf{Irreversible tasks transitions} According to both Chain~\cite{chain} and Alpaca~\cite{alpaca} programming models, each task calls the next one.
+	\textbf{Irreversible tasks transitions} According to both Chain~\cite{chain} and Alpaca~\cite{alpaca} programming models, each task calls the next one.
 The transition is done through functions calls and manually clearing the stack to prevent the stack overflow problem.
@@
 
-	\item \textbf{Dependency on privatization} Alpaca~\cite{alpaca} extended Chain~\cite{chain} with a concept of privatization, i.e.
+	\textbf{Dependency on privatization} Alpaca~\cite{alpaca} extended Chain~\cite{chain} with a concept of privatization, i.e.
 it protects task-shared variables from WAR dependency problem \emph{within} the scope of a task.
 Therefore, if tasks are on-demand merged, the scope of these dependencies are changed and the static analysis to privatize these variables is not valid any more and the data consistency can not be preserved.
-
-	\item \textbf{`All channels' requirement} Fig.~\ref{fig:DynamicChainSeq} shows that if the task control flow is a circular with no branches then task merging is possible.
-Moreover, all the intra-merged-tasks channel can be committed to SRAM instead of FRAM to save energy.
-However, this execution path is only a special one.
-If the general case (see again Fig.~\ref{fig:DynamicChainSeq}) is considered then leaving out the intra-merged-task channels might result in data inconsistency.
-Than is, if an application tries to access a merged task from the middle after a power interrupt then the input for that task is not ready, which result in an incorrect execution.
-
-	\item \textbf{Lack of multi-versioning system} All known modular execution models define its own input for each task that is not shared with other tasks.
-This design decision makes the benefit of merging tasks of less importance since each task is interacting with its own variables.
-\todo{Rewrite this paragraph: still unclear}{Przemek}
-\end{itemize}
 

### 2017-08-05 DMA/DFT/main.c
new coalescing algorithm

@@ -77,7 +77,7 @@ void dft_end() ;
 static void blinkLed(uint32_t wait );
 static void burn( uint32_t iters);
 
-#define SIZE 16
+#define SIZE 8
 #define PI2 6.28
 
 __p  float x[SIZE] ={0.0};              // discrete-time signal, x

### 2017-08-05 DMA/ipos-gcc-dma/include/global.h

@@ -28,7 +28,7 @@
 // function pointer
 typedef void (* funcPt)(void);
 
-extern unsigned int __task_address;
+extern unsigned int __virtualTaskAddr;
 extern volatile unsigned int  __totNumTask;
 extern unsigned int __persis_CrntPagHeader;
 extern unsigned int __pagsInTemp[];

### 2017-08-05 DMA/ipos-gcc-dma/ipos/pTCB.c

@@ -54,7 +54,7 @@ void os_addTasks(unsigned char numTasks, taskId tasks[]){
         }
         *(__head +(--cnt) ) =  __temp_head;   // link the tail of the linkedlist with the head
 
-        __task_address    =  (unsigned int) __head ;
+        __virtualTaskAddr    =  (unsigned int) __head ;
         funcBlocker = 0xAD;
     }
 }

### 2017-08-05 DMA/ipos-gcc-dma/ipos/psched.c

@@ -4,6 +4,7 @@
 //#define __KEY  0xAD
 #define COMMITTING      1
 #define COMMIT_FINISH   0
+#define VERTUTASK       10
 
 #define JUMP2()  if(!__jump)  \
                     {   \
@@ -25,48 +26,38 @@
                             __jump = 0; \
                         }
 
+
 #define JUMP();    if(!__jump){   \
-                        __current_task_virtual  =  \
-                        (unsigned int*) (*(__current_task_virtual + NEXT_OFFSET_PT)) ;     /* soft transition */ \
-                    }else{  \
-                        while(__jump_cnt < __jump_by)   \
-                            {   \
-                                __current_task_virtual  =  \
-                                (unsigned int*) (*(__current_task_virtual + NEXT_OFFSET_PT)) ;  /* soft transition */ \
-                                __jump_cnt++;     \
-                            }       \
-                        __jump = 0;     \
-                        __jump_cnt = 0;    \
-                    }
+        __realTask  =  \
+        (unsigned int*) (*(__realTask + NEXT_OFFSET_PT)) ;     /* soft transition */ \
+        }else{  \
+            while(__jump_cnt < __jump_by)   \
+            {   \
+                __realTask  =  \
+                (unsigned int*) (*(__realTask + NEXT_OFFSET_PT)) ;  /* soft transition */ \
+        __jump_cnt++;     \
+            }       \
+            __jump = 0;     \
+            __jump_cnt = 0;    \
+        }
 
-#define COALESCING 1
 
-#if COALESCING
-__nv volatile unsigned int __virtualTaskSize = 2;
-__nv volatile unsigned int __maxVirtualTaskSize = 20;
-     volatile unsigned int __taskCounter = 0;
-#else
-__nv volatile unsigned int __virtualTaskSize = 1;
-__nv volatile unsigned int __maxVirtualTaskSize = 1;
-     volatile unsigned int __taskCounter = 1;
-#endif
 
-__nv volatile unsigned int __temp_taskCounter = 0;
-__nv volatile unsigned int __totalTaskCounter = 0;
+    unsigned char __jump = 0;
+    unsigned int __jump_by = 0;
+    unsigned int __jump_cnt = 0;
 
-//TODO I do not think these two variables need to be persistent
-__nv volatile unsigned int __jump = 0;   //TODO we can eliminated and use only __jump_by
[…]

### 2017-08-05 experiments/coalescing/coalescing.py

@@ -57,50 +57,36 @@ def figureSetting():
     plt.rcParams['legend.numpoints'] = 1
 
     
-distances = ['20cm.csv', '40cm.csv', "50cm.csv", "cont_power.csv"]
+distances = ["cont.csv",'20cm.csv', '40cm.csv', "60cm.csv"]
 
 data =[]
 for dis in distances:
     row=[]
-    f_c = open("Sort/"+dis) 
-    f_v = open("temp-log/"+dis) 
+    f_c = open("coalescing/page_128/bc/"+dis) 
+    f_n = open("nocoalescing/page_128/bc/"+dis) 
     row.append( executionTime(f_c) )
-    row.append( executionTime(f_v) )
+    row.append( executionTime(f_n) )
     f_c.close()
-    f_v.close()
+    f_n.close()
     data.append(row)
 
 # print(data)
 # data = np.transpose(data)
 colors = ['0.2', '0.4', '0.6', '0']
-labels = ['20cm', '40cm', '50cm', 'cpwr']
-
-
-num_bars = len(data)
-gap = 0.1   
-bar_width = (1- gap) / num_bars            
+labels = ['20cm', '40cm', '60cm', 'cpwr']
+          
+print(data) 
 
 f = plt.figure(figsize=(8,4))
 figureSetting()                        # Set figure layout
 
 ax = plt.axes()
-ax.xaxis.set_major_locator(ticker.FixedLocator([0,1,2,3]) )
-ax.xaxis.set_major_formatter(ticker.FixedFormatter( ("Sort", "CEM") ))
+ax.xaxis.set_major_locator(ticker.FixedLocator([1,2,3,4]) )
+ax.xaxis.set_major_formatter(ticker.FixedFormatter( ("COAL", "NOCOAL") ))
 
 plt.ylabel('Seconds')
 
-
-for i, row in enumerate(data):    # enumerate returns a data unit (a row) and its index
-    print('row', row)
-    x = np.arange(len(row))      
-    plt.bar((x-0.35)+i *bar_width, row,    
-            width=bar_width, 
-            color=colors[i % len(colors)], 
-           label = labels[i% len(labels)], 
-           align='center')
-# plt.bar( data)
-# plt.bar(1, chain_dd_time, align='center', color="0.6", hatch='+', label="Chain")
-# plt.bar(2, ipos_dd_time, align='center', color="0.85", hatch='/', label="IPOS")
[…]

### 2017-08-06 paper/evaluation.tex

--- 2017-08-03 paper/evaluation.tex
@@
 build around MSP430FR5969~\cite{wolverine} 64\,kB FRAM microcontroller, and TI's MSP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} (hardware type will be specified per experiment in Section~\ref{sec:results_evaluation}).
-WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet}.
+WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet} with TI Code Composer Studio version X~\cite{}\todo{Report on CCS version; CMUs environment}{Kiwan/Amjad}.
 FET was also used as a power supply in non-intermittent power experiment.
@@
 
-Expected results:
-\begin{itemize}
-	\item Sensitivity plot---normalized page swapping cost ("bathtub" curve) in terms of execution time: compare against Chain and full double buffering approach; show number of page swaps---do the experiment for all applications
-\end{itemize}
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
+	\caption{Normalized page size overhead and page faults per benchmark.\todo{Figure placeholder---will be replaced with final one}{Amjad}}
+	\label{fig:IPOSPerformance}
+	\label{fig:page_size}
+\end{figure}
 
-\todo{Provide argument for page swapping using hardware support}{Brandon}
+Figure~\ref{fig:page_size} presents the result on the effect of page swap size on the execution time of the programs executed by \sys.
+Overhead has been measured in the MCU clock cycles with values provided by TI CCS.
+Result has been normalized per benchmark: for the overhead results---to the minimum obtained value and for page faults---to the third measured maximum value\todo{Check again}{Amjad}.
+A subset of applications has been selected for evaluation.
+Additionally, page faults have been measured.
+The result show clearly a bathtub-shape bar plot for each application, demonstrating that for each application there is always a page size that minimizes \emph{both} execution time and page swaps.
+For all applications (except for Sort) this value equals roughly to 64 bytes.
+The shape of each figure can be explained as follows.
+For small page sizes (<64) the execution time is dominated by the \emph{frequency} of page swaps (each swap takes non-zero time, as shown in Figure~\ref{fig:dmaTimeEnergy}).
+On the other hand, in the opposite case the execution time is dominated by the time of swapping of one large page from SRAM to FRAM.
+Another interesting observation is the non-symmetric shape of each bar plot.
+This is due the size of data manipulated by the program: minimum point of overhead equals roughly to the size of the variables manipulated by each benchmark.
+Worth noting is also large discrepancy between minimum and maximum overhead for each benchmark (e.g.
+for Cuckoo 8 (at 16 bytes) to 1 (at 64 bytes), BC: 1 (at 16 bytes) to $\approx$2 (at 1024 bytes)).
+\todo{Explain this phenomenon}{Amjad}
 

### 2017-08-06 paper/evaluation.tex
Software suite section expanded

@@
 Therefore Alpaca was excluded from the evaluation.
-
 \todo{Check if the line of argument is solid for the above paragraphs about not selecting DINO, Memenos and Aplaca}{Brandon}
 
-The applications that were used in the evaluation are given in Table~\ref{table:benchmark_list}.
-These benchmarks are the same ones as used in~\cite{chain,alpaca}.
-In addition to Chain benchmark programs new applications were used for benchmarking: \emph{DFT}, \emph{Huffman compressor}, and \emph{selection sort}.
+The applications that were used in the evaluation are: (i) \textbf{Bitcount}: counts bits in x y byte pseudo-random string using x algorithms.
+Each result is compared for correctness; (ii) \textbf{Cold chain}: emulates embedded temperature sensing engine: measurements are randomly generated and are compressed using LZW algorithm for x-entry dictionary and y-byte block; (iii) \textbf{Cuckoo}: implements cuckoo filter---x entry filter filled to y of capacity---program first stores a sequence of pseudo-random numbers, then queries the filter to recover it; (iv) \textbf{RSA}: encrypts predefined x byte long string with a y bytes long key\todo{were keys of 64 bytes tested?
+If yes: d they exhaust energy}{Amjad}.
+This set of benchmarks is the same as used in~\cite{chain,alpaca}\todo{Fill-in the missing data and verify its correctness}{Przemek/Amjad}.
+In addition to Chain benchmarks new applications were implemented: (v) \emph{DFT}, (vi)\emph{Huffman compressor}, and (vii) \emph{Selection Sort}.
+All applications are compared in Table~\ref{table:benchmark_list} in terms of source lines of code against Chain.
+\todo{finalize describing all applications once all experiments are done}{Przemek}
 
@@
 		\hline\hline
+		Bitcount & 351 & 588 \\ %59\%
 		Cold chain & 388 & 721 \\ %53\%
@@
 		RSA & 887 & 1233 \\ %71\%
-		Bitcount & 351 & 588 \\ %59\%
-		DFT & --- & --- \\ %Two resolutions: 4 and 8 Bytes
-		Huffman & --- & --- \\ %Data decompression size: 100 bytes
-		Selection sort & --- & --- \\
+		DFT & {} & {} \\ %Two resolutions: 4 and 8 Bytes
+		Huffman & {} & {} \\ %Data decompression size: 100 bytes
+		Selection sort & {} & {} \\
 		\hline
@@
 \end{table}
-
-\todo{finalize describing all applications once all experiments are done}{Przemek}
 

### 2017-08-06 paper/coalescing.tex
add some coalescing results

@@
 	\centering
-	%\includegraphics[width=\columnwidth]{figures/task_coalescing}
-	\caption{Task coalescing architecture.}
-	\label{fig:}
+	\includegraphics[width=\columnwidth]{figures/coalescing}
+	\caption{This figure shows the performance of \sys when it uses the dynamic tasks coalescing mechanism as compared to its performance when the virtual task size is fixed to one, on multiple distances.
+}
+	\label{fig:coalescing}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/averageVirtualTaskSize}
+	\caption{ This figure shows the average virtual task size (measured in real task unit) of the virtual task sizes that the coalescing algorithm used to accomplished the applications on a distance of 60cm.
+}
+	\label{fig:aveVirtuTaskSize}
 \end{figure}

### 2017-08-06 DMA/ipos-gcc-dma/ipos/psched.c

@@ -59,7 +59,7 @@ __nv unsigned int  __maxVirtualTaskSize = 0x7f;
      unsigned int *__realTask = NULL;
 __nv unsigned int  __realTaskCntr = VERTUTASK;
 
-__nv unsigned int pers_pt[1024]={0};
+__nv unsigned char * pers_pt= (unsigned char *)0xB000;
 __nv unsigned int pers_cnt = 0;
 
 void os_enter_critical()
@@ -108,7 +108,7 @@ void os_scheduler()
 #if COALESCING
 
         if(pers_cnt < 1024){
-            pers_pt[pers_cnt] = __virtualTaskCntr;  // Debugging
+            *(pers_pt+pers_cnt) = __virtualTaskCntr;  // Debugging
             pers_cnt++;
         }
 

### 2017-08-06 paper/background.tex
background groomed into shape

@@
 
-\paragraph{Fixed-size Tasks are Inflexible}
-\TODO{Brandon left off here - I've cut some bullets below that I did not understand.
+%\TODO{Brandon left off here - I've cut some bullets below that I did not understand.
 I think the key point here is that fixed task size only works for one cap size; inefficient with a larger cap, unimplementable with a smaller one.
 There might be an efficiency argument, but it's tough to make because we still require some memory consistency mechanism (in our case paged FRAM), which has an overhead.}
-A modular intermittent execution model completely depends on a programmer's
-estimation which mostly results in a sub-optimal code division. Moreover, this
-model can only consider a single hardware configuration and it does not take
-environment variability into considerations. Each battery-less platform has
-different storage sizes, refer to Table~\ref{table:capacitor}, which causes
-different charge/discharge periods. This result in different execution times.
-In the worst case the size of the capacitor might be too small to execute one
-task written by the programmer. This calls for the ability of the modular
-execution runtimes to \textbf{merge} multiple small tasks into a larger one or
-\textbf{split} a large task into a set of small ones on the fly
-(post-compilation time). Unfortunately the following features prevent us to
-perform this task using existing runtimes~\cite{chain,alpaca}.
+\paragraph{Fixed-size tasks are inflexible and inefficient} 
+A programmer must write each task with statically defined boundaries, which
+makes statically sized tasks inflexible and non-portable.  The programmer must
+ensure statically that a task's energy requirements match the device's energy
+storage capacity.  If a programmer writes a program that consumes more energy
+than the device can buffer, the task will never be able to complete using the
+buffered energy. The consequence of such tasks is a non-terminating program
+that repeatedly attempts to re-execute the task of excess energy cost.  If a
+programmer writes a program that consumes less energy than the device can
+buffer, the system may operate inefficiently. When a task and its successor
+both complete , the two tasks were interleaved by a task boundary that
+preserves progress and state (i.e., checkpoint or commit).  If the two tasks
+were not also interleaved by a power failure, then the computation done to
+preserve state was {\em unnecessary}, only incurring an overhead on the tasks'
+executions.   
+%
+Avoiding excessively costly, non-terminating tasks and short, high-overhead
+tasks is a programming challenge, given fixed hardware with a fixed energy
+buffering capacity.   Such tasks also make it extremely difficult to {\em port}
+code from  one device to another. A task that is excessively costly on a device
+with a small energy buffer may become a short, high-overhead task  on a device
+with a larger energy buffer and vice versa. The key problem is that the programmer
+writes tasks statically for an energy buffer of a fixed size.
+%
+\sys's {\em task coalescing} mechanism, which we describe in
+Section~\ref{sec:task_coalescing}, is motivated by the observation that systems
+with statically defined tasks may be inflexible and inefficient; \sys addresses
+the problem
 
-\begin{figure}
-	\centering
-	%\includegraphics[width=\columnwidth]{figures/Dynamic_chain_sequence_example}
-	\caption{Dynamic task merging problem.\todo{Draw merging diagram}{Przemek/Amjad}}
-	\label{fig:DynamicChainSeq}
-\end{figure}
-
-	\textbf{Irreversible tasks transitions} According to both Chain~\cite{chain} and Alpaca~\cite{alpaca} programming models, each task calls the next one.
-The transition is done through functions calls and manually clearing the stack to prevent the stack overflow problem.
-This approach makes all the task transitions firm, preventing task merging.
-
-	\textbf{Dependency on privatization} Alpaca~\cite{alpaca} extended Chain~\cite{chain} with a concept of privatization, i.e.
-it protects task-shared variables from WAR dependency problem \emph{within} the scope of a task.
-Therefore, if tasks are on-demand merged, the scope of these dependencies are changed and the static analysis to privatize these variables is not valid any more and the data consistency can not be preserved.
-
-\subsection{Main Idea: Tasks Coalescing}
-
-To combat the above limitations we propose a new concept of \textbf{virtualization}: a modular execution model with softening a number of transitions, by keeping the state of the execution progress in the volatile memory, to construct a bigger \emph{virtual task} to reduce energy consumption.
-Ideally, the size of the virtual task should match the length of a continuous interval of the intermittent power supply for the best performance.
-Another form of virtualization can be achieved by reducing the number of non-volatile memory accesses.
-This can be done by having a temporary copy of a global variable in the volatile memory (similar to the caching principle) and a task interacts with it and updates it before writing back the most recent value to the non-volatile memory during the commit process.
-
-\subsection{Assumption on the Type of Computing Hardware}
+\subsection{Hardware Assumptions}
 \label{sec:background_hardware}
 
-Before proceeding further with the background we need to make certain assumptions about the underlying hardware.
-We assume that we operate on the microcontroller (more generally, computing hardware) equipped with a mix of volatile and non-volatile memory, such as FLASH, Ferroelectric RAM or Magnetoresistive RAM.
-We do not assume to rely on specific memory type (either volatile or non-volatile), in contrary to earlier works targeting non-volatile processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
-Actually, we will exploit the properties of both memory types to maximize performance metrics of the program being executed.
-We enable to I/O operations, however we do not enable the scheduling of operations (which \emph{nota bene} forms a foundation for a simple runtime).
-All in all, our hardware model follows in assumptions the two most relevant intermittently-powered execution environments~\cite{alpaca,chain}.
-\todo{Verify if this section is not out of place}{Brandon}
-
-\subsection{Structuring a Task-based Intermittent Computing Program}
-
-\todo{Please check this is the right way for this section: it reads as it should be in Section 3 describing \sys}{Brandon} Task is a programmer (or compiler) defined code that is executed based on consistent set of inputs, i.e.
-memory fetched from a specified memory location, producing a consistent output stored in a specified (the same or different) memory location.
-Syntactically, we can treat a task as a continuous sub-section of a large program encapsulated by (i) task specific keywords defining the beginning and the end of a task and (ii) message passing between tasks connected with each other.
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=\columnwidth]{figures/taskification_example}
-	\caption{Task flow example executed by \sys.\todo{Draw an example task diagram}{Przemek}}
-	\label{fig:task_flow_example}
-\end{figure}
-
-Let us look into how task-based programming is performed relying on an example provided in Fig.~\ref{fig:task_flow_example}.
-Each task is uniquely identified by a keyword.
-A separate part of code specifies a state machine than governs how tasks are connected with each other.
-This way a programmer specifies to which other task a current task needs to pass the result of its computation (and it can be either a new task or the task itself).
-Tasks are executed atomically.
-That is, task will be executed only once and will be considered as completed only when a task reaches a final instruction line informing a scheduler that a program is ready to enter into a new task.
-
-Just like Chain~\cite{chain} and Alpaca~\cite{alpaca}, \sys does not support concurrent programming, multi-threading (multi-task sequence) and does not consider priorities and on-demand scheduling.
-These features are extremely important to consider but go beyond the scope of this work.
-A comparative description of \sys is given in Table~\ref{table:feature_comparison}.
-\todo{Decide if we need such table}{Brandon}
-
-Important and a strong assumption about task-based programming for intermittently-powered systems is that an energy availability for the largest possible task execution must be guaranteed.
-Task is an atomic operation and must preserve execution order and memory manipulations.
-Task that follows such an assumption is equivalent to task being executed on a continuously-powered non-interrupted system.
-Whenever a task cannot fit into a single energy reservoir, it must be divided into tasks that will be executed within a single discharge cycle.
-
-\begin{table}
-	\centering
-	\footnotesize
-	\begin{tabular}{|c|c|c|c|}
-		\hline
-		{~} & Alpaca~\cite{alpaca} & Chain~\cite{chain} & This work \\
-		\hline\hline
-		Task connections & in-task & in-task & outside \\
-		Task re-use & N & N & Y\\
-		Memory model & globals & globals & globals\\
-		Concurrency & N & N & N \\
-		Task blocking & N & N & Y \\
-		Energy-adaptive & N & N & Y \\
-		\hline
-	\end{tabular}
-	\caption{Comparison of existing task-based execution models against \sys.}
-	\label{table:feature_comparison}
-\end{table}
-
-\todo{Elaborate on the ``key challenge'' paragraphs near the end of the intro}{Przemek}
-
-%We can write the problem of maximizing all task sizes formally as
+\sys is designed for the demands of existing and future energy-harvesting
+platforms based around general purpose, commodity computing
+components~\cite{wisp,msp430datasheet}.  
 %
-%\begin{equation}
-%\underset{i}{\arg\,\max} \sum_{i}m_i+t_i \text{~subject to~} \forall_i mi+t_i>e,
-%\end{equation}
+We assume a device with a memory system that has fast, byte-addressable
+volatile and non-volatile memory; in particular, our target
+platform~\cite{wisp} is equpped with a mixture of SRAM and FRAM.
 %
-%where $e$ is the required minimum energy to perform a task, $i$ is the total number of tasks, $m_i$ and $t_i$ are the cost of task traversal (memory copying time of task variables) and time to complete one task, respectively.
-
-%Two problems with fixed-size task in intermittent execution.
-(a) task underestimation: task executed at device with capacitor size $X$, will not execute at device with capacitor $X\gg Y$; (b) task overestimation: with stable energy source task tracking causes unnecessary overhead.
-
-%\begin{figure}
-%\centering
-%\includegraphics[width=\columnwidth]{figures/taskification_example}
-%\caption{Program divided into tasks.}
-%\label{fig:taskification_example}
-%\end{figure}
-
-%Other forms of intermittent platforms will share the same problem.
-One of them considers actuation platforms, which will be powered directly from energy harvesting sources~\cite{}.
-will share the energy storage with computing platform.
-However, power supply for actuators is order of magnitude larger than for computation.
-Therefore, although capacitor is large enough to perform computation alone, energy to power actuators takes precedence leaving not much space for computing.
+Our implementation leverages hardware support for fast, bulk-copying between
+memories via DMA~\cite{msp430datasheet}.
+%
+We do not require a particular non-volatile memory technology, nor do we
+require architectural additions commodity
+processors.~\cite{su_date_2017,ratchet,quickrecall,nvp}. 
+%
+\sys supports I/O behavior similar to recent intermittent execution
+models~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe,
+asynchronous I/O.

### 2017-08-06 paper/evaluation.tex

@@
 \label{sec:results_coalescing}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalescing}
+	\caption{This figure shows the performance of \sys when it uses the dynamic tasks coalescing mechanism as compared to its performance when the virtual task size is fixed to one, on multiple distances.
+}
+	\label{fig:coalescing}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/averageVirtualTaskSize}
+	\caption{ This figure shows the average virtual task size (measured in real task unit) of the virtual task sizes that the coalescing algorithm used to accomplished the applications on a distance of 60cm.
+}
+	\label{fig:aveVirtuTaskSize}
+\end{figure}
 

### 2017-08-06 paper/systemdescription.tex
draft of system overview text

@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/viper_block_diagram}
+	\includegraphics[width=\columnwidth]{figures/viper_block_diagram.pdf}
 	\caption{\sys top-level view.}
@@
 
-\sys as a whole is characterized by the following features:
+\sys is a new programming and execution model for intermittent computing on
+energy-harvesting devices.  \sys addresses the challenges motivated in
+Section~\ref{sec:background} to make task-based intermittent programs {\em
+efficient}, {\em flexible}, and {\em programmable}.  \sys accomplishes this
+goal with a constellation of a new programming model, compiler, and run time
+software system support.  Figure~\ref{fig:system_overview} shows a block
+diagram overview of \sys.
 
-\begin{itemize}
-	\item \textbf{Compiler-supported automatic task generator}: details will be provided in Section~\ref{sec:compiler}
-	\item \textbf{Memory virtualization}: details will be provided in Section~\ref{sec:memory_virtulaization}
-	\item \textbf{Task coalescing}: details will be provided in Section~\ref{sec:task_coalescing}
-\end{itemize}
+To use \sys, a programmer first writes plain, imperative C code.  The
+programmer then has an option to {\em manually} or {\em automatically}
+translate their program into code for \sys's task-based programming  model.  We
+describe the programming primitives involved in either translation in
+Section~\ref{sec:overview:programming}.
+%
+To manually translate the programmer decomposes the program into tasks, and
+annotates memory accesses that manipulate data shared by multiple tasks.  To
+automatically translate, the programmer simply uses \sys's compiler (described
+in Section~\ref{sec:compiler}), which decomposes a program into tasks and
+annotates accesses to task-shared data. 
+%
+After translating to task-based code and compiling, the programmer has an
+executable \sys binary.
 
-A complete logical structure of \sys is given in Figure~\ref{fig:system_overview}. Before describing each feature of \sys in detail we need to introduce a programming and semantics/execution model.
+When the binary executes, \sys's runtime support affects its execution through
+its {\em virtualizing memory manager} and through its {\em coalescing task
+scheduler}.
+%
+The \sys binary executes its tasks correctly and with consistent memory state
+due to \sys's virtual memory mechanism.  Relying on programmer or compiler
+annotation of accesses to memory locations that might be shared beetween tasks,
+\sys's memory virtualization {\em pages} these data to SRAM from FRAM as they
+are needed.  Tasks operate, accessing SRAM only, demand-paging data in to
+accomodate an SRAM smaller than FRAM.  When a task completes, \sys commits
+modified, paged data back to its original FRAM location atomically ensuring
+that subsequent tasks page in correct memory state.
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/taskification_example}
-	\caption{\sys API.}
-	\label{fig:system_overview}
-\end{figure}
+\sys runs a lightweight task scheduler that uses {\em task coalescing} to make
+\sys's statically defined tasks efficiently use available energy.  By default,
+tasks run in sequence and each task commits as it completes, potentially
+incurring unnecessary overhead for commits between consecutive, non-failing
+tasks.  \sys dynamically coalesces consecutive tasks by deferring the commit of
+an earlier task until the completion of a later task.  \sys's memory management
+mechanism facilitates coalescing because \sys can defer an earlier task's
+commit by simply buffering modified, paged data as the later task executes.
+After coalescing a sequence of tasks, \sys commits all tasks' updates to FRAM,
+together, combining tasks updates to the same pages and minimizing the commit
+overhead.
 
@@
 \end{table}
-
-\subsection{Semantics and Execution Model}
-\label{sec:overview_semantics}

### 2017-08-07 paper/systemdescription.tex
added programming model overview

@@
 
+\sys primarily consists of its compiler and the run time software support for
+its execution model, but \sys also includes several lightweight programming
+model extensions that are similar to those in prior work~\cite{chain,alpaca}.
+Table~\ref{tab:viper_syntax} provides an overview of these extensions, which
+can be used manually by a programmer, or can be added to a program
+automatically, using \sys's compiler.  As in prior task-based intermittent
+programming models, a \sys program is decomposed into tasks, which are
+functions that are not allowed to have any direct callers.  The programmer
+identifies a task by adding the {\tt task} keyword to the function's
+declaration.  A programmer invokes one task, {\tt t} from another task {\tt u}
+using {\tt next\_task( t )}, which atomically commits the state of {\tt u} and
+transfers control to the entry point of {\tt t}.   A program has a single {\em
+origin} task that executes first during the program's first execution; after a
+power failure, in subsequent operating periods, \sys restarts from the task
+that was executing just before the failure.
+
+A \sys program manipulates {\em protected} state --- state that is accessed by
+more than one task --- using the {\tt RP} and {\tt WP} operations, which read
+and write (respectively) a protected variable. These operations convey to
+\sys's runtime system which data must be protected through paging by the memory
+manager.
+
 \begin{table}
@@
 		\hline\hline
+		\texttt{task foo(){...}} & Task declaration\\
+		\texttt{origin task foo(){...}} & Origin task declaration\\
+		\texttt{next\_task()} & Transition to a new task\\
 		\texttt{RP} & Read-protected variable\\
 		\texttt{WP} & Write-protected variable\\
-		\texttt{\_\_task\_\_ foo(){...}} & Task declaration\\
-		\texttt{addTask()} & Build internal data structure for executor \\
-		\texttt{run\_tasks()} & Task executor \\ %scheduler
-		\texttt{next\_task()} & Transition to a new task\\
-		%\texttt{PageFault()} & --- \\
-		%\texttt{P()} & ---\\
 		\hline
 	\end{tabular}
-\caption{Summary of \sys syntax: top---user-specific variables}
+\caption{Summary of \sys's programming model extensions}
 \label{tab:viper_syntax}

### 2017-08-07 paper/evaluation.tex
added text explaining result in Figure 9

--- 2017-08-06 paper/evaluation.tex
@@
 
-We now proceed with the evaluation of task coalescing mechanism of \sys.
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalescing}
+	\caption{Performance of \sys's  tasks coalescing mechanism compared to virtual task size equal to one (per application) at multiple WISP to RFID reader antenna distances.\todo{Replace "coal" with "coalescing", "No-coal" with "No coalescing"; use markers in addition to color; decide on uniform naming for apps (either all small or capitalized)}{Amjad}}
+	\label{fig:coalescing}
+\end{figure}
 
-Expected results:
+Experimental evaluation of task coalescing mechanism under intermittent power of \sys is given in Figure~\ref{fig:coalescing}.
+For each application we measured execution time (with and without task coalescing) at \{20, 40, 60\}\,cm distance between RFID antenna and WISP and when WISP was powered directly from FET.
+Each measurement was repeated x times per each distance/application.
+From the result we clearly see an enormous benefit of coalescing, especially at large WISP to RFID reader antenna distance ($\approx$56\% for cem, 100\% for sort and $\approx$400\% for bc).
+At each distance, \sys coalescing mechanism provided faster execution times compared to non-coalescing case.
+\todo{Discuss different shapes per app}{Amjad}
+
+Measured size of tasks for each application is given in Figure~\ref{fig:aveVirtuTaskSize}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/averageVirtualTaskSize}
+	\caption{Average virtual task size (measured in real task unit\todo{Define real task unit}{Amjad}) of the virtual task sizes that the coalescing algorithm used to accomplished the applications at a WISP to RFID reader distance of 60\,cm.\todo{Please rewrite this sentence}{Amjad}}
+	\label{fig:aveVirtuTaskSize}
+\end{figure}
+
+expected results:
 \begin{itemize}

### 2017-08-07 paper/coalescing.tex
move the coalescing results to the evaluation section
moved figure on task merging from coalescing to evaluation

--- 2017-07-31 paper/coalescing.tex
@@
 \sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks to improve execution time at stable energy supply environment.
-
-\begin{figure}
-	\centering
-	%\includegraphics[width=\columnwidth]{figures/task_coalescing}
-	\caption{Task coalescing architecture.}
-	\label{fig:}
-\end{figure}
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/virtualTaskSize.eps}
-	\caption{Execution time of a dummy application containing twelve empty tasks as a function of coalescing factor.
-We observe three times improvement from coalescing until four virtual tasks, while we observe a diminishing return from task coalescing beyond four.
-\todo{Improve the figure: make it narrower, larger font, remove legend, do not capitalize "Task" in X axis, add Y axis label (Execution time (ms))}{Amjad}}
-	\label{fig:virtualTaskSize}
-\end{figure}
-
-The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e.
-from energy harvesting to continuous power environment) is the best one?.
-To answer this question we have implemented a simple program composed of $x$ empty tasks.
-We execute the same program by increasing coalescing factor by one while the device on which the program was executed was running on a stable power supply (refer to Section~\ref{sec:methodology_evaluation} for details).
-The result is presented in Figure~\ref{fig:virtualTaskSize}.
-We observe a significant increase in execution time as task are coalescent from one to four (three times execution time improvement).
-However, as the coalescing increases beyond four, we observe diminishing returns from coalescing in terms of execution time.
-\todo{Decide whether to place this figure in Evaluation section or leave it here}{Przemek}
-
 %\begin{algorithm}

### 2017-08-07 paper/systemdescription.tex
Section 3 revised (good job Brandon)

@@
 
-\sys is a new programming and execution model for intermittent computing on
-energy-harvesting devices.  \sys addresses the challenges motivated in
-Section~\ref{sec:background} to make task-based intermittent programs {\em
-efficient}, {\em flexible}, and {\em programmable}.  \sys accomplishes this
-goal with a constellation of a new programming model, compiler, and run time
-software system support.  Figure~\ref{fig:system_overview} shows a block
-diagram overview of \sys.
+\sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
+\sys addresses the challenges motivated in Section~\ref{sec:background} to make task-based intermittent programs {\em efficient}, {\em flexible}, and {\em programmable}.
+\sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.
 
-To use \sys, a programmer first writes plain, imperative C code.  The
-programmer then has an option to {\em manually} or {\em automatically}
-translate their program into code for \sys's task-based programming  model.  We
-describe the programming primitives involved in either translation in
-Section~\ref{sec:overview:programming}.
-%
-To manually translate the programmer decomposes the program into tasks, and
-annotates memory accesses that manipulate data shared by multiple tasks.  To
-automatically translate, the programmer simply uses \sys's compiler (described
-in Section~\ref{sec:compiler}), which decomposes a program into tasks and
-annotates accesses to task-shared data. 
-%
-After translating to task-based code and compiling, the programmer has an
-executable \sys binary.
+\textbf{\sys Programming Model:} To use \sys, a programmer first writes plain, imperative C code.
+The programmer then has an option to {\em manually} or {\em automatically} translate the program into code for \sys's task-based programming model.
+To manually translate, the programmer decomposes the program into tasks, and annotates memory accesses that manipulate data shared by multiple tasks.
+To automatically translate, the programmer simply uses \sys's compiler (described in Section~\ref{sec:compiler}), which decomposes a program into tasks and annotates accesses to task-shared data.
+After translating to task-based code and compiling, the programmer has an executable \sys binary.
 
-When the binary executes, \sys's runtime support affects its execution through
-its {\em virtualizing memory manager} and through its {\em coalescing task
-scheduler}.
-%
-The \sys binary executes its tasks correctly and with consistent memory state
-due to \sys's virtual memory mechanism.  Relying on programmer or compiler
-annotation of accesses to memory locations that might be shared beetween tasks,
-\sys's memory virtualization {\em pages} these data to SRAM from FRAM as they
-are needed.  Tasks operate, accessing SRAM only, demand-paging data in to
-accomodate an SRAM smaller than FRAM.  When a task completes, \sys commits
-modified, paged data back to its original FRAM location atomically ensuring
-that subsequent tasks page in correct memory state.
+\sys consists of its compiler and the run time software support for its execution model and includes several lightweight programming model extensions similar to those in prior work~\cite{chain,alpaca}.
+Table~\ref{tab:viper_syntax} provides an overview of these extensions, which can be used manually by a programmer or added to a program automatically, using \sys's compiler.
+As in prior task-based intermittent programming models~\cite{chain,alpaca}, a \sys program is decomposed into tasks, which are functions that are not allowed to have any direct callers.
+The programmer identifies a task by adding the {\tt task} keyword to the function's declaration.
+A programmer invokes one task, {\tt t} from another task {\tt u} using {\tt next\_task( t )}, which atomically commits the state of {\tt u} and transfers control to the entry point of {\tt t}.
+A program has a single {\em origin} task that executes first during the program's first execution; after a power failure, in subsequent operating periods, \sys restarts from the task that was executing just before the failure.
 
-\sys runs a lightweight task scheduler that uses {\em task coalescing} to make
-\sys's statically defined tasks efficiently use available energy.  By default,
-tasks run in sequence and each task commits as it completes, potentially
-incurring unnecessary overhead for commits between consecutive, non-failing
-tasks.  \sys dynamically coalesces consecutive tasks by deferring the commit of
-an earlier task until the completion of a later task.  \sys's memory management
-mechanism facilitates coalescing because \sys can defer an earlier task's
-commit by simply buffering modified, paged data as the later task executes.
-After coalescing a sequence of tasks, \sys commits all tasks' updates to FRAM,
-together, combining tasks updates to the same pages and minimizing the commit
-overhead.
-
-\subsection{Programming Model}
-\label{sec:overview_programming_model}
-
-\sys primarily consists of its compiler and the run time software support for
-its execution model, but \sys also includes several lightweight programming
-model extensions that are similar to those in prior work~\cite{chain,alpaca}.
-Table~\ref{tab:viper_syntax} provides an overview of these extensions, which
-can be used manually by a programmer, or can be added to a program
-automatically, using \sys's compiler.  As in prior task-based intermittent
-programming models, a \sys program is decomposed into tasks, which are
-functions that are not allowed to have any direct callers.  The programmer
-identifies a task by adding the {\tt task} keyword to the function's
-declaration.  A programmer invokes one task, {\tt t} from another task {\tt u}
-using {\tt next\_task( t )}, which atomically commits the state of {\tt u} and
-transfers control to the entry point of {\tt t}.   A program has a single {\em
-origin} task that executes first during the program's first execution; after a
-power failure, in subsequent operating periods, \sys restarts from the task
-that was executing just before the failure.
-
-A \sys program manipulates {\em protected} state --- state that is accessed by
-more than one task --- using the {\tt RP} and {\tt WP} operations, which read
-and write (respectively) a protected variable. These operations convey to
-\sys's runtime system which data must be protected through paging by the memory
-manager.
+A \sys program manipulates {\em protected} state---accessed by more than one task---using the {\tt RP} and {\tt WP} operations, which read and write (respectively) a protected variable.
+These operations convey to \sys's runtime system which data must be protected through paging by the memory manager.
 
@@
 \end{table}
+
+\textbf{\sys Memory Manager:} When the binary executes, \sys's runtime support affects its execution through {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}) and {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}).
+The \sys binary executes its tasks correctly, with consistent memory state, due to \sys's virtual memory mechanism.
+Relying on programmer or compiler annotation of accesses to memory locations that might be shared between tasks, \sys's memory virtualization {\em pages} these data to SRAM from FRAM as they are needed.
+Tasks operate, accessing SRAM only, demand-paging data to accommodate a SRAM smaller than FRAM\todo{Sentence unclear}{Brandon}.
+When a task completes, \sys commits modified, paged data back to its original FRAM location atomically ensuring subsequent tasks page in correct memory state.
+
+\textbf{\sys Task Scheduler:} \sys runs a lightweight task scheduler that uses {\em task coalescing} to make \sys's statically defined tasks efficiently use available energy.
+By default, tasks run in sequence and each task commits as it completes, potentially incurring unnecessary overhead for commits between consecutive, non-failing tasks.
+\sys dynamically coalesces consecutive tasks by deferring the commit of an earlier task until the completion of a later task.
+\sys's memory management mechanism facilitates coalescing because \sys can defer an earlier task's commit by simply buffering modified, paged data as the later task executes.
+After coalescing a sequence of tasks, \sys commits all tasks' updates to FRAM, together, combining tasks updates to the same pages and minimizing the commit overhead.

### 2017-08-07 paper/background.tex
Background section revised (yaay)

@@
 In this section, we provide background on energy harvesting systems and the intermittent software execution model.
-We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys
-addresses.
+We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys addresses.
 
@@
 
-Energy harvesting devices operate using energy extracted from their environment, from sources such as radio waves (RF) and solar energy.
+Energy harvesting devices operate using energy extracted from their environment, from sources such as radio waves and solar energy.
 Energy-harvesting devices need not use a tethered power supply or battery and instead, typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.
@@
 Several recent surveys discussing energy harvesting, low-power, embedded systems and intermittent computing at a high level~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016,lucia_snapl_2017}.
+
+\subsection{Intermittent Computing}
+\label{sec:background_consistency}
+
+Energy-harvesting devices execute software according to the {\em intermittent execution model}~\cite{dino,chain,alpaca,ratchet}.
+Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
+Software running on the device executes {\em intermittently} because energy is only available between when the capacitor reaches its threshold and when it is depleted.
+An intermittent execution is composed of operating periods interspersed with power failures.
+The frequency of power failures in an intermittent execution depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods.
+Energy-harvesters provide input power that is orders of magnitude less than a device's operating power, making recharging negligible during operation.
+
+Software has different behavior when executed intermittently than when executed with continuous power.
+Each power failure clears volatile memory (registers, stack, and global variables).
+Non-volatile memory (e.g., FRAM) persists across power failures.
+On a power failure, control flows back to a prior point in the execution.
+By default, after a power failure control flows to the beginning of {\tt main()}.
+Early work on intermittent computing preserve intermittent progress by periodically capturing and restoring to a non-volatile checkpoint of the volatile execution context, i.e., registers and stack~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
+
+\begin{figure}
+	\centering
+	\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
+	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 10) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
+	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
+	\label{fig:code_demo_incosistency}
+\end{figure}
+
+Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}.
+Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value.
+The situation occurs when code includes a write after read (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}.
+
+Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using cyclic redundancy check (CRC) code from MIBench~\cite{hicks_mibench2_2016}.
+The code computes CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
+Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of a WAR dependence between the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
+Figure~\ref{fig:crc_example_war} shows an execution of the code in which power fails during the second loop iteration.
+The problem arises because line 6 of the execution writes $\texttt{Rmnd}$ in the second iteration, but power fails before capturing the next checkpoint.
+After restarting from the checkpoint, the update to $\texttt{data}$ at line 10 reads the value of $\texttt{Rmnd}$ that was updated before the power failure, producing an incorrect result.
+
+As the figure illustrates, checkpoint-based systems risk violating the consistency of volatile and non-volatile memory~\cite{dino}.
+To maintain memory consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of non-volatile variables with the checkpointed execution context.
+Task-based systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail next, ensure state is consistent by using programmer, compiler, and runtime support.
+
+\subsection{Task-based Intermittent Computing}
+\label{section:background_task_computing}
+
+Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into a collection tasks.
+A task is a function with no caller that contains arbitrary computation, sensing, and communication.
+A programmer describes the sequence in which tasks may execute, creating a task graph.
+The flow of control from task to task happens at programmer demarcated points and may conditionally depend on program values.
+Task-based programming abstractions guarantee that tasks exhibit behavior as though they executed {\em atomically}, regardless of arbitrary power failures.
+Task-based runtime systems tolerate failures by repeatedly attempting to execute a task until it completes.
+Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent (i.e., have the same effect when repeatedly executed).
+The key to idempotence is ensuring that updates to non-volatile state made by an interrupted, partial execution of a task are never visible to a subsequent (potentially completing)
+re-execution of the task.
+
+There are several run time strategies to ensuring task idempotence.
+One approach, utilized by Alpaca~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{rachet} did for checkpoints), execute the task using buffered copies of those data, and commit the buffered copies before completing the task.
+Another strategy is to statically create multiple versions of non-volatile data shared by tasks and ensure by construction that no task reads and writes the same non-volatile memory location~\cite{chain}.
+Regardless of the idempotence strategy, the key to task-based execution is that fixed, statically defined tasks execute as though atomic, completing after zero or more attempts.
+
+\begin{table}
+	\centering
+	\footnotesize
+	\begin{tabular}{|c|c|}
+		\hline
+		Model & Data Copied to/from NVRAM \\
+		\hline\hline
+		Mementos~\cite{mementos}                             & Reg. + Stack     \\
+		DINO~\cite{dino}                                     & Reg. + WAR variables \\
+		Chain~\cite{chain}                                   & PC   + Channel data\\
+		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
+		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
+		\hline
+	\end{tabular}
+	\caption{Memory consistency enforcement overheads of various intermittent execution models.
+\emph{Reg.} includes the entire register file, \emph{PC} is the program counter, \emph{Channel data} are variables explicitly task-shared by the programmer, and \emph{WAR variables} are variables involved in WAR dependences (\emph{NV}: non-volatile).}
+	\label{table:chechpoint_comparison}
+\end{table}
+
+\subsubsection{Costs of Task-based Models}
+
+Task-based intermittent execution models ensure that programs execute correctly despite arbitrary power failures and in doing so, incur several costs that are the motivation for \sys.
+
+\text{Memory Consistency Enforcement Overhead:} Intermittent execution models incur an overhead because they must checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
+Table~\ref{table:chechpoint_comparison} compares the overheads for several recent intermittent execution models (generalizing discussion from~\cite[Sec.
+2.4]{alpaca}.) The mechanism responsible for the overhead of each model varies, but all the bulk of overhead in all models is a manipulating non-volatile memory.
+Checkpointing moves volatile memory and registers to and from non-volatile memory, channel accesses directly manipulate non-volatile data~\cite{chain}, privatization copies data from and commits data to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} require all of memory to be non-volatile.
+
+\begin{figure}
+	\centering
+	\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
+	\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
+	\caption{Energy cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
+	\label{fig:framEnergy}
+\end{figure}
+
+To concretely understand the cost of these manipulations of non-volatile memory, we directly measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations for the MSP430FR5969~\cite{msp430datasheet} microcontroller.
+We used EDB~\cite{edb} to monitor the voltage drop on an energy buffering capacitor that powers the device across 1600 read and write accesses to each type of memory, randomizing locations to avoid caching effects.
+Figure~\ref{fig:framEnergy} summarizes these measurements, illustrating that a non-volatile memory access consumes around 1.5 times as much energy as a volatile memory access in this device.
+In an intermittently-powered device, the increased energy corresponds to decreased run time between power failures.
+The overheads imposed by relying on non-volatile memory motivate \sys, which uses memory virtualization to operate primarily using volatile SRAM.
+
+The above result reassures the intuitive observation that a decrease of checkpoint frequency (CP) decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
+On the other hand, this increases task/checkpoint re-execution time.
+The ideal solution is a runtime that \emph{adaptively} changes task size during runtime, which however exposes a second problem.
 
@@
 
-\subsection{Intermittent Computing}
-\label{sec:background_consistency}
-
-Energy-harvesting devices execute software according to the {\em intermittent execution model}~\cite{dino,chain,alpaca,ratchet}.
-Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
-Software running on the device executes {\em intermittently} because energy is only available between when the
-capacitor reaches its threshold and when it is depleted.
-An intermittent execution is composed of operating periods interspersed with power failures.
-The frequency of power failures in an intermittent execution depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods.
-Energy-harvesters provide input power that is orders of magnitude less than a device's operating power, making recharging negligible during operation.
-
-Software has different behavior when executed intermittently than when executed with continuous power.
-Each power failure clears volatile memory (registers, stack, and global variables).
-Non-volatile memory (e.g., FRAM) persists across power failures.
-On a power failure, control flows back to a prior point in the execution.
-By default, after a power failure control flows to the beginning of {\tt main()}.
-Early work on intermittent computing preserve intermittent progress by periodically capturing and restoring to a non-volatile checkpoint of the volatile execution context (i.e., registers and stack)~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
-
-\begin{figure}
-	\centering
-	\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
-	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 10) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
-	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
-	\label{fig:code_demo_incosistency}
-\end{figure}
-
-Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}.
-Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value.
-The situation occurs when code includes a write after read (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}.
-
-Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using cyclic redundancy check (CRC) code from MIBench~\cite{hicks_mibench2_2016}.
-The code computes CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
-Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of a WAR dependence between the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
-Figure~\ref{fig:crc_example_war} shows an execution of the code in which power
-fails during the second loop iteration.
-The problem arises because line 6 of the execution writes $\texttt{Rmnd}$ in the second iteration, but power fails before capturing the next checkpoint.
-After restarting from the checkpoint, the update to $\texttt{data}$ at line 10 reads the value of $\texttt{Rmnd}$ that was updated before the power failure, producing an incorrect result.
-
-As the figure illustrates, checkpoint-based systems risk violating the consistency of volatile and non-volatile memory~\cite{dino}.
-To maintain memory consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of non-volatile variables with the checkpointed execution context.
-Task-based
-systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail next, ensure state is consistent by using programmer, compiler, and runtime support.
-
-\subsection{Task-based Intermittent Computing}
-\label{section:background_task_computing}
-
-%\TODO{This sequential/modular split seems pretty odd to me.
-I think we can just talk about tasks in the sense of ratchet, alpaca, and chain.
-Anything with static (programmer-written or compiler inserted) boundaries applies}
-
-Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the
-programmer to decompose their program into a collection tasks. A task is a
-function with no caller that contains arbitrary computation, sensing, and
-communication. A programmer describes the sequence in which tasks may execute,
-creating a task graph. The flow of control from task to task happens at
-programmer demarcated points and may conditionally depend on program values.
-Task-based programming abstractions guarantee that tasks exhibit behavior as
-though they executed {\em atomically}, regardless of arbitrary power failures.
-Task-based runtime systems tolerate failures by repeatedly attempting to
-execute a task until it completes. Task-based runtime systems ensure
-task-atomic semantics by ensuring that repeated task executions are idempotent
-(i.e., have the same effect when repeatedly executed). The key to idempotence
-is ensuring that updates to non-volatile state made by an interrupted, partial
-execution of a task are never visible to a subsequent (potentially completing)
-re-execution of the task.
-
-There are several run time strategies to ensuring task idempotence. One
-approach, utilized by Alpaca~\cite{alpaca}, is to identify non-volatile data
-involved in WAR dependences in a task (like DINO~\cite{dino} and
-Ratchet~\cite{rachet} did for checkpoints), execute the task using buffered
-copies of those data, and commit the buffered copies before completing the
-task. Another strategy is to statically create multiple versions of
-non-volatile data shared by tasks and ensure by construction that no task reads
-and writes the same non-volatile memory location~\cite{chain}. Regardless of
-the idempotence strategy, the key to task-based execution is that fixed,
-statically defined tasks execute as though atomic, completing after zero or
-more attempts.
-
-\begin{table}
-	\centering
-	\footnotesize
-	\begin{tabular}{l|l}
-		Model & Data Copied to/from NVRAM\\
-		\hline
-		Mementos~\cite{mementos}                             & Reg. + Stack     \\
-		DINO~\cite{dino}                                     & Reg. + WAR vars. \\
-		Chain~\cite{chain}                                   & PC   + Channel data\\
-		Alpaca~\cite{alpaca}                                 & PC   + WAR vars. \\
-		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main mem.)             \\
-	\end{tabular}
-	\caption{Memory consistency enforcement overheads of various intermittent execution models.
-Reg.
-includes the entire register file, PC is the program counter, Channel data are variables explicitly task-shared by the programmer, and WAR vars.
-are variables involved in WAR dependences.}
-	\label{table:chechpoint_comparison}
-\end{table}
-
-\subsubsection{Costs of Task-based Models}
-
-Task-based intermittent execution models ensure that programs execute correctly
-despite arbitrary power failures and in doing so, incur several costs that are
-the motivation for \sys.
-
-\paragraph{Memory Consistency Enforcement Overhead}
-
-Intermittent execution models incur an overhead because they must checkpoint
-data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or
-privatize and commit data~\cite{alpaca}.
-%
-Table~\ref{table:chechpoint_comparison} compares the overheads for several
-recent intermittent execution models (generalizing discussion from~\cite[Sec.
-2.4]{alpaca}.)  The mechanism responsible for the overhead of each model
-varies, but all the bulk of overhead in all models is a manipulating
-non-volatile memory.  Checkpointing moves volatile memory and registers to and
-from non-volatile memory, channel accesses directly manipulate non-volatile
-data~\cite{chain}, privatization copies data from and commits data to
-non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet}
-require all of memory to be non-volatile.
-
-\begin{figure}
-	\centering
-	\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
-	\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
-	\caption{Energy cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
-	\label{fig:framEnergy}
-\end{figure}
-
-To concretely understand the cost of these manipulations of non-volatile
-memory, we directly measured the energy consumption of volatile (SRAM) and
-non-volatile (FRAM) memory operations for the
-MSP430FR5969~\cite{msp430datasheet} microcontroller. We used EDB~\cite{edb} to
-monitor the voltage drop on an enegy buffering capacitor that powers the device
-across 1600 read and write accesses to each type of memory, randomizing
-locations to avoid caching effects.  Figure~\ref{fig:framEnergy} summarizes
-these measurements, illustrating that a non-volatile memory access consumes
-around 1.5x as much energy as a volatile memory access in this device.  In an
-intermittently-powered device, the increased energy corresponds to decreased
-run time between power failures.  The overheads imposed by relying on
-non-volatile memory motivate \sys, which uses memory virtualization to operate
-primarily using volatile SRAM.
-
-The above result reassures the intuitive observation that a decrease of CP
-(Table~\ref{table:chechpoint_comparison}) decreases the energy cost associated
-with each state preservation method and reduces operation cost which in most
-simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$,
-where PL is the cost of program execution without checkpointing. On the other
-hand, this increases task/checkpoint re-execution time. The ideal solution is a
-runtime that \emph{adaptively} changes task size during runtime, which however
-exposes a second problem.
-
-%\TODO{Brandon left off here - I've cut some bullets below that I did not understand.
-I think the key point here is that fixed task size only works for one cap size; inefficient with a larger cap, unimplementable with a smaller one.
-There might be an efficiency argument, but it's tough to make because we still require some memory consistency mechanism (in our case paged FRAM), which has an overhead.}
-\paragraph{Fixed-size tasks are inflexible and inefficient} 
-A programmer must write each task with statically defined boundaries, which
-makes statically sized tasks inflexible and non-portable.  The programmer must
-ensure statically that a task's energy requirements match the device's energy
-storage capacity.  If a programmer writes a program that consumes more energy
-than the device can buffer, the task will never be able to complete using the
-buffered energy. The consequence of such tasks is a non-terminating program
-that repeatedly attempts to re-execute the task of excess energy cost.  If a
-programmer writes a program that consumes less energy than the device can
-buffer, the system may operate inefficiently. When a task and its successor
-both complete , the two tasks were interleaved by a task boundary that
-preserves progress and state (i.e., checkpoint or commit).  If the two tasks
-were not also interleaved by a power failure, then the computation done to
-preserve state was {\em unnecessary}, only incurring an overhead on the tasks'
-executions.   
-%
-Avoiding excessively costly, non-terminating tasks and short, high-overhead
-tasks is a programming challenge, given fixed hardware with a fixed energy
-buffering capacity.   Such tasks also make it extremely difficult to {\em port}
-code from  one device to another. A task that is excessively costly on a device
-with a small energy buffer may become a short, high-overhead task  on a device
-with a larger energy buffer and vice versa. The key problem is that the programmer
-writes tasks statically for an energy buffer of a fixed size.
-%
-\sys's {\em task coalescing} mechanism, which we describe in
-Section~\ref{sec:task_coalescing}, is motivated by the observation that systems
-with statically defined tasks may be inflexible and inefficient; \sys addresses
-the problem
+\textbf{Fixed-size Tasks are Inflexible and Inefficient:} A programmer must write each task with statically defined boundaries, which makes statically sized tasks inflexible and non-portable.
+The programmer must ensure statically that a task's energy requirements match the device's energy storage capacity.
+If a programmer writes a program that consumes more energy than the device can buffer, the task will never be able to complete using the buffered energy.
+The consequence of such tasks is a non-terminating program that repeatedly attempts to re-execute the task of excess energy cost.
+If a programmer writes a program that consumes less energy than the device can buffer, the system may operate inefficiently.
+When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).
+If the two tasks were not also interleaved by a power failure, then the computation done to preserve state was {\em unnecessary}, only incurring an overhead on the tasks' executions.
+Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity that might be \emph{different} for each transiently-powered device (see Table~\ref{table:capacitor}).
+Such tasks also make it extremely difficult to {\em port} code from  one device to another.
+A task that is excessively costly on a device with a small energy buffer may become a short, high-overhead task on a device with a larger energy buffer and vice versa.
+The key problem is that the programmer writes tasks statically for an energy buffer of a fixed size.
+\sys's {\em task coalescing} mechanism, which we describe in Section~\ref{sec:task_coalescing}, is motivated by the observation that systems with statically defined tasks may be inflexible and inefficient.
 
@@
 
-\sys is designed for the demands of existing and future energy-harvesting
-platforms based around general purpose, commodity computing
-components~\cite{wisp,msp430datasheet}.  
-%
-We assume a device with a memory system that has fast, byte-addressable
-volatile and non-volatile memory; in particular, our target
-platform~\cite{wisp} is equpped with a mixture of SRAM and FRAM.
-%
-Our implementation leverages hardware support for fast, bulk-copying between
-memories via DMA~\cite{msp430datasheet}.
-%
-We do not require a particular non-volatile memory technology, nor do we
-require architectural additions commodity
-processors.~\cite{su_date_2017,ratchet,quickrecall,nvp}. 
-%
-\sys supports I/O behavior similar to recent intermittent execution
-models~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe,
-asynchronous I/O.
+\sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
+We assume a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform~\cite{wisp} is equipped with a mixture of SRAM and FRAM.
+Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
+We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors.~\cite{su_date_2017,ratchet,quickrecall,nvp}.
+\sys supports I/O behavior similar to recent intermittent execution models~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.

### 2017-08-07 paper/systemdescription.tex
Compiler section revised (yaay)

@@
 
-\textbf{\sys Programming Model:} To use \sys, a programmer first writes plain, imperative C code.
+\noindent \textbf{\sys Programming Model:} To use \sys, a programmer first writes plain, imperative C code.
 The programmer then has an option to {\em manually} or {\em automatically} translate the program into code for \sys's task-based programming model.
@@
 
-\textbf{\sys Memory Manager:} When the binary executes, \sys's runtime support affects its execution through {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}) and {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}).
+\noindent \textbf{\sys Memory Manager:} When the binary executes, \sys's runtime support affects its execution through {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}) and {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}).
 The \sys binary executes its tasks correctly, with consistent memory state, due to \sys's virtual memory mechanism.
@@
 
-\textbf{\sys Task Scheduler:} \sys runs a lightweight task scheduler that uses {\em task coalescing} to make \sys's statically defined tasks efficiently use available energy.
+\noindent \textbf{\sys Task Scheduler:} \sys runs a lightweight task scheduler that uses {\em task coalescing} to make \sys's statically defined tasks efficiently use available energy.
 By default, tasks run in sequence and each task commits as it completes, potentially incurring unnecessary overhead for commits between consecutive, non-failing tasks.

### 2017-08-07 paper/texFigCode/sram-buffer.tex
paging

\documentclass{standalone}
\usepackage{tikz}
\usetikzlibrary{arrows,positioning} 
\usetikzlibrary{fit}
\usetikzlibrary{shadows}

\tikzset{
	%Define standard arrow tip
	>=stealth',
	%Define style for boxes
	punkt/.style={
		rectangle,
		rounded corners,
		draw=black, very thick,
		text width=4.5em,
		minimum height=1.5em,
		text centered},
	% Define arrow style
	pil/.style={
		->,
		thick,
		shorten <=2pt,
		shorten >=2pt,}
}

\begin{document}

	\begin{tikzpicture}[node distance=1cm, auto,]

	\node[punkt,minimum height=3em] (task1) {y = x+1 
		\\ z = y+1 \\ x = z+1};
	\node[below = 0.1cm of task1] () {\textbf{Task}};

	\node[draw,very thick,dotted,minimum width=7em, minimum height=7em, right = 3cm of task1] (sram) {};
	\node[below = 0.1cm of sram] () {\textbf{Volatile Buffer}};

	\node[punkt, below = -2.4cm of sram,fill=red!3] (x-sram) {x};
	\node[punkt, below = 0.1cm of x-sram,fill=black!5] (y-sram) {y};
	\node[punkt, below = 0.1cm of y-sram,fill=black!5] (z-sram) {z};

	\node[right = 1.8cm of sram] (ref) {};
	\node[draw,very thick,minimum width=7em, minimum height=15em, right = of ref] (fram1) {};

	%\node[draw,very thick,dotted,minimum width=7em, fill=green!5,minimum height=9em, below = 1cm of fram1] (fram2) {};

	\node[punkt, above = -0.8cm of fram1,fill=green!3] (x1-fram) {x};
	\node[punkt, below = 0.1cm of x1-fram,fill=black!10] (y1-fram) {y};
	\node[punkt, below = 0.1cm of y1-fram,fill=black!10] (z1-fram) {z};

	\node[punkt, below = -2.5cm of fram1,fill=black!5] (x2-fram) {x};
	\node[punkt, below = 0.1cm of x2-fram,fill=black!5] (y2-fram) {y};
	\node[punkt, below = 0.1cm of y2-fram,fill=black!5] (z2-fram) {z};

	\node[below = 0.1cm of z2-fram] (copies) {copies};	
	\node[below = 0.1cm of z1-fram] (original) {original};	
	\node[below = 0.7cm of z2-fram] () {\textbf{FRAM}};

	\path[every node/.style={font=\sffamily\small}]
	([yshift=0.3cm] task1.east) edge [->, bend left] node {(1) get x} (x-sram.west)
	(x-sram.east) edge [->, bend left] node[xshift=0.5cm] {(2) read x} (x1-fram.west)

	([yshift=0.3cm]task1.east) edge [->, bend right] node[xshift=-0.5cm] {(3) set y} (y-sram.west)

	(task1.east) edge [->, bend right] node[xshift=-0.4cm] {(4) set z} (z-sram.west)
	([yshift=-0.4cm]task1.east) edge [->, bend left] node[xshift=1.3cm,yshift=-0.4cm] {(5) set x} (x-sram.west)

	(x-sram.east) edge [->, bend right] node[xshift=-0.4cm] {(6) commit-I x} (x2-fram.west)
	(z-sram.east) edge [->, bend right] node[xshift=-0.4cm] {(7) commit-I z} (z2-fram.west)
	(y-sram.east) edge [->, bend right] node[xshift=-0.4cm] {(8) commit-I y} (y2-fram.west)

	(x2-fram.east) edge [->, bend right] node[xshift=2.5cm] {(9) commit-II x} (x1-fram.east)
	(y2-fram.east) edge [->, bend right] node[xshift=2.5cm] {(10) commit-II y} (y1-fram.east)
	(z2-fram.east) edge [->, bend right] node[xshift=2.5cm] {(11) commit-II z} (z1-fram.east)
	;

	\end{tikzpicture}

\end{document}

### 2017-08-07 paper/background.tex

@@
 
-\text{Memory Consistency Enforcement Overhead:} Intermittent execution models incur an overhead because they must checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
+\noindent \textbf{Memory Consistency Enforcement Overhead:} Intermittent execution models incur an overhead because they must checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
 Table~\ref{table:chechpoint_comparison} compares the overheads for several recent intermittent execution models (generalizing discussion from~\cite[Sec.
@@
 
-\textbf{Fixed-size Tasks are Inflexible and Inefficient:} A programmer must write each task with statically defined boundaries, which makes statically sized tasks inflexible and non-portable.
+\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} A programmer must write each task with statically defined boundaries, which makes statically sized tasks inflexible and non-portable.
 The programmer must ensure statically that a task's energy requirements match the device's energy storage capacity.

### 2017-08-07 paper/compiler.tex
merged heuristic eval section into compiler text

@@
 
-%The compiler consists of two {\em passes} that runs on LLVM Intermediate Representation (IR)\todo{Which version?}{Kiwan}---an intermediate language generated by Clang~\cite{clang_website}:
-%
-%\begin{enumerate}
-%	\item \textbf{Analysis Pass}, which analyzes the code structure and decides where to put task boundaries;
-%	\item \textbf{Transform Pass}, which converts the code into task-style program and fills it with necessary \sys keywords so that it can run on \sys.
-%\end{enumerate}
-%
-%We shall now proceed with describing these two parts in detail.
-
 \subsection{Task Boundary Identification Analysis}
@@
 
-\sys analyzes a program to determine where to put task boundaries to divide the
-program into tasks. A task is a single-entry, multiple-exit region of the CFG
-and in a correct task decomposition, no task should ever consume more energy
-than the device's energy buffer can store.  There are several factors that
-determine whether tasks will execute efficiently. A task decomposition should
-minimize the number of variables that are accessed by multiple tasks and, 
-consequently, must be protected by \sys's paging mechanism.  Tasks should
-be roughly the same size, simplifying \sys's dynamic reasoning about coalescing
-by eliminating the need to consider variable task cost at run time.  To minimize
-overhead introduced at task boundaries (i.e., commit), a task decomposition should 
-maximize task size, which minimizes the frequency of task transitions.
+\sys analyzes a program to determine where to put task boundaries to divide the program into tasks.
+A task is a single-entry, multiple-exit region of the CFG and in a correct task decomposition, no task should ever consume more energy than the device's energy buffer can store.
+There are several factors that determine whether tasks will execute efficiently.
+A task decomposition should minimize the number of variables that are accessed by multiple tasks and, consequently, must be protected by \sys's paging mechanism.
+Tasks should be roughly the same size, simplifying \sys's dynamic reasoning about coalescing by eliminating the need to consider variable task cost at run time.
+To minimize overhead introduced at task boundaries (i.e., commit), a task decomposition should maximize task size, which minimizes the frequency of task transitions.
 
@@
 
-%\sys meets these correctness and efficiency requirements with a four-part
-%analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis},
-%(iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.
+%\sys meets these correctness and efficiency requirements with a four-part analysis composed of (i) \emph{graph construction}, (ii) \emph{loop analysis}, (iii) \emph{task preparation}, and (iv) \emph{single entry assurance}.
 
@@
 
-The input to the task boundary identification analysis is a standard C program
-(as shown in Figure~\ref{fig:comp_1}) and the output of \sys's analysis is the
-CFG, with some vertices marked as {\em task boundaries}. 
-%
+The input to the task boundary identification analysis is a standard C program (as shown in Figure~\ref{fig:comp_1}) and the output of \sys's analysis is the CFG, with some vertices marked as {\em task boundaries}.
 The \sys compiler builds the program's CFG (as shown in
-Figure~\ref{fig:comp_2}).  To simplify interprocedural analysis, our prototype
-inlines all function calls into the main function in the CFG. 
-%
-\sys uses a cost-based task boundary placement analysis that
-considers per-block energy consumption estimates, inter-task variable liveness,
-and the cost of partitioning loops.  We first describe how \sys measures each of these costs 
-and then describe \sys's procedure for placing boundaries in the CFG.
+Figure~\ref{fig:comp_2}).
+To simplify inter-procedural analysis, our prototype inlines all function calls into the main function in the CFG.
+\sys uses a cost-based task boundary placement analysis that considers per-block energy consumption estimates, inter-task variable liveness, and the cost of partitioning loops.
+We first describe how \sys measures each of these costs and then describe \sys's procedure for placing boundaries in the CFG.
 
 \subsubsection{Task Cost Analysis}
-\sys's task cost analysis considers estimated task energy cost, variable
-liveness, and the impact of splitting loops across tasks.  The result of the
-task cost analysis determines where to place boundaries in the CFG.
 
-{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the
-CFG a {\em cost}, which is an estimate of the energy consumption of the block.
-Precisely modeling the energy cost of arbitrary C code is a difficult problem
-and beyond the scope of this work; we used the number of intermediate
-representation (IR) instructions in a block as a proxy for energy because
-energy cost is likely to closely correspond to block length.  Moreover, we
-assume as an input to the compiler the maximum number of operations that the
-device can execute without exhausting its buffered energy. This parameter is
-highly device-dependent, and we assume that a developer or device manufacturer
-determines the number empirically, as we did in this work through direct
-measurement.  \sys's analysis tries ensure that the estimated energy cost of different tasks
-is roughly the same and does not exceed given energy budget.
+\sys's task cost analysis considers estimated task energy cost, variable liveness, and the impact of splitting loops across tasks.
+The result of this analysis determines where to place boundaries in the CFG.
 
-{\noindent \bf Live variable cost.}
-Each CFG edge is associated with a set of {\em live variables} that are
-accessible in both of the blocks that define the edge (i.e., predecessor and
-successor).  Variables that live across an edge between two tasks must be
-protected by \sys's memory virtualization mechanism and each live, inter-task
-variable increases the run time overhead for a boundary placement.  \sys's
-analysis tries to minimize the overhead of protected variables by placing
-boundaries to minimize the number of inter-task live variables.
+{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block.
+Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we used the number of intermediate representation (IR) instructions in a block as a proxy for energy because energy cost is likely to closely correspond to block length.
+Moreover, we
+assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy.
+This parameter is highly device-dependent, and we assume that a developer or device manufacturer determines the number empirically, as we did in this work through direct measurement.
+\sys's analysis tries ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget.
 
-{\noindent \bf Loop cost.}
-Loops are a challenge for \sys because the analysis must account for the cost
-of multiple iterations of the loop body.  \sys assumes that most loops are
-annotated with a static loop bound, which is common in embedded code.  For
-loops that do not have a static bound, \sys conservatively assumes the maximum
-possible loop bound, based on the type of the loop induction variable.
+{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of {\em live variables} that are accessible in both of the blocks that define the edge (i.e., predecessor and successor).
+Variables that live across an edge between two tasks must be protected by \sys's memory virtualization mechanism and each live, inter-task variable increases the run time overhead for a boundary placement.
+\sys's analysis tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.
 
-The analysis tries to minimize the dynamic overhead of traversing task
-boundaries by avoiding placing a task boundary on a loop with a large bound,
-when possible. At the same time, the analysis must avoid placing a loop in a
-single task if the total energy cost of all of its iterations is likely to
-exceed the device's energy buffer.  \sys first multiplies the estimated energy
-cost of the blocks in the loop body by the loop bound to determine estimate the
-cost of the loop's execution.  If the loop's cost exceeds the device's energy
-buffer, \sys can place a task boundary on the loop's back-edge, converting each
-iteration into its own task.
+{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis must account for the cost of multiple iterations of the loop body. \sys assumes that most loops are
+annotated with a static loop bound, which is common in embedded code.
+For loops that do not have a static bound, \sys conservatively assumes the maximum possible loop bound, based on the type of the loop induction variable.
+
+The analysis tries to minimize the dynamic overhead of traversing task boundaries by avoiding placing a task boundary on a loop with a large bound, when possible.
+At the same time, the analysis must avoid placing a loop in a single task if the total energy cost of all of its iterations is likely to exceed the device's energy buffer.
+\sys first multiplies the estimated energy cost of the blocks in the loop body by the loop bound to determine estimate the cost of the loop's execution.
+If the loop's cost exceeds the device's energy buffer, \sys can place a task boundary on the loop's back-edge, converting each iteration into its own task.
 
@@
 
-After completing its cost analysis of a program, \sys inserts task boundaries
-into the CFG in two steps.  The first step is to insert task boundaries on the
-back-edges of all loops that have an energy cost greater than the device's
-energy buffer capacity, producing an initial boundary placement. The second
-step is to insert task boundaries at additional points in the code,
-sub-dividing tasks that have an estimated cost that exceeds the energy buffer's
-capacity.
+After completing program cost analysis, \sys inserts task boundaries into the CFG in two steps.
+The first step is to insert task boundaries on the back-edges of all loops that have an energy cost greater than the device's energy buffer capacity, producing an initial boundary placement.
+The second step is to insert task boundaries at additional points in the code, sub-dividing tasks that have an estimated cost that exceeds the energy buffer's capacity.
 
-To sub-divide these tasks, \sys does depth-first traversals of all paths in the
-CFG, ignoring back-edges. Traversals start from the beginning of the main
-function, or from an existing task boundary.  \sys accumulates the cost of
-basic blocks during the traversal until reaching another task boundary or an
-exit node in the CFG.   At the end of a traversal, if the accumulated cost
-exceeds to device's energy capacity \sys must split the traversed path with a
-task boundary.  Each edge on the path is a candidate location for a boundary.
+To sub-divide these tasks, \sys does depth-first traversals of all paths in the CFG, ignoring back-edges.
+Traversals start from the beginning of the main function, or from an existing task boundary.
+\sys accumulates the cost of basic blocks during the traversal until reaching another task boundary or an exit node in the CFG.
+At the end of a traversal, if the accumulated cost exceeds to device's energy capacity \sys must split the traversed path with a task boundary.
+Each edge on the path is a candidate location for a boundary.
 
-The analysis uses a heuristic to compute a score for each candidate and places
-a boundary at the candidate with the highest score. The heuristic tries to
-avoid placing boundaries on loops, to minimize the number of live variables
-that must be protected across the new boundary, and to maximize the number of
-operations along the path.  A boundary candidate's score is computed as:
+The analysis uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score.
+The heuristic tries to avoid placing boundaries on loops, to minimize the number of live variables that must be protected across the new boundary, and to maximize the number of operations along the path.
+A boundary candidate's score is computed as
 %
@@
 %
-where $N_{\text{IR}}$ is the length of the path being traversed up to the
-candidate, $N_{\text{PV}}$ is the number of variables live variables at the
-candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops
-(i.e., nested loops) containing the boundary candidate. \sys normalizes
-$N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates.
-The weights, $w_{0}, w_{1}$ are between zero and one, and are
-hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
-$N_{\text{PV}}$, and $N_{\text{loop}}$.  We empirically determined these weight
-values by measuring the runtime for different weight values across
-applications.  We found that varying weights varies runtime by only around 5\%
-in most cases and in our experimental evaluation, we used $w_{0} = w_{1} =
-0.5$.
+where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of variables live variables at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate.
+\sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates.
+The weights, $w_{0}, w_{1}$ are between zero and one, and are hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
+$N_{\text{PV}}$, and $N_{\text{loop}}$.
+We empirically determined these weight values by measuring the runtime for different weight values across applications.
+We found that varying weights varies runtime by only around 5\% in most cases and in our experimental evaluation, we used $w_{0} = w_{1} = 0.5$.
 
-After placing task boundaries, the compiler must ensure that all tasks are
-single-entry.  To do so, \sys computes the dominator tree of the CFG and checks
-that every vertex within a task is dominated by the entry vertex of the task.
-If this dominance condition is unmet for a CFG node, the compiler places
-an additional task boundary at that node, meeting the condition.
-%
-Figure~\ref{fig:comp_2} illustrates boundary placement: the first three
-vertices make up one task and the remainder become a second task.
+After placing task boundaries, the compiler must ensure that all tasks are single-entry.
+To do so, \sys computes the dominator tree of the CFG and checks that every vertex within a task is dominated by the entry vertex of the task.
+If this dominance condition is unmet for a CFG node, the compiler places an additional task boundary at that node, meeting the condition.
+Figure~\ref{fig:comp_2} illustrates boundary placement: the first three vertices make up one task and the remainder become a second task.
 
@@
 
-After placing task boundaries during analysis, \sys's transformation analysis
-restructures the code into tasks defined by the placed boundaries. The analysis
-traverses the CFG starting from an arbitrary task boundary.  On reaching a
-subsequent boundary, the transformation analysis identifies all paths between
-the two task boundaries and copies those paths into a new task function.  The
-analysis continues until all paths are contained within a task.
+After placing task boundaries during analysis, \sys's transformation analysis restructures the code into tasks defined by the placed boundaries.
+The analysis traverses the CFG starting from an arbitrary task boundary.
+On reaching a subsequent boundary, the transformation analysis identifies all paths between the two task boundaries and copies those paths into a new task function.
+The analysis continues until all paths are contained within a task.
 
-After forming tasks, variables that are live beyond the end of the task become
-protected, which requires renaming those variables and updating references to
-renamed variable names.  
-%
-For example, Figure~\ref{fig:comp_2} shows a boundary placed within the live
-range of {\tt b}; {\tt b} becomes protected and accesses to {\tt b} should be
-redirected to {\tt p\_b} (as shown in Figure~\ref{fig:comp_3}).  {\tt a} is
-task-local and need not be protected, but only declared in the task. Constants
-and read-only variable are never protected.
-%
-When writes in multiple, newly formed tasks may update the same location (i.e.,
-at a PHI node in the CFG), \sys re-writes the accesses to refer to the same global
-name.   Subsequent reads simply read the protected variable.
+After forming tasks, variables that are live beyond the end of the task become protected, which requires renaming those variables and updating references to renamed variable names.
+For example, Figure~\ref{fig:comp_2} shows a boundary placed within the live range of {\tt b}; {\tt b} becomes protected and accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in Figure~\ref{fig:comp_3}).
+{\tt a} is task-local and need not be protected, but only declared in the task.
+Constants and read-only variable are never protected.
+When writes in multiple, newly formed tasks may update the same location (i.e., at a PHI node in the CFG), \sys re-writes the accesses to refer to the same global name.
+Subsequent reads simply read the protected variable.
 
-After forming tasks, some branches in a task may refer to targets in another
-task. \sys patches the CFG to re-write those branches to instead be \transition
-statements that transition to the entry point of the other task. Such a
-``broken'' branch will always refer to a task's entry node because \sys ensures
-that all tasks have only one entry.
+After forming tasks, some branches in a task may refer to targets in another task.
+\sys patches the CFG to re-write those branches to instead be \transition statements that transition to the entry point of the other task.
+Such a ``broken'' branch will always refer to a task's entry node because \sys ensures that all tasks have only one entry.
 
-The compiler re-writes accesses to protected variables to use \sys's memory
-virtualization API. Pointers pose a challenge to re-writing protected accesses
-because unprotected pointers may point to protected data.  Leveraging LLVM's
-{\tt basic} pointer alias analysis the compiler conservatively re-writes all
-accesses that {\em may} read or write protected data.
+The compiler re-writes accesses to protected variables to use \sys's memory virtualization API.
+Pointers pose a challenge to re-writing protected accesses because unprotected pointers may point to protected data.
+Leveraging LLVM's {\tt basic} pointer alias analysis the compiler conservatively re-writes all accesses that {\em may} read or write protected data.
 

### 2017-08-07 paper/evaluation.tex

@@
-We are now ready to present experimental results of \sys performance.
+We present the experimental results of \sys performance.
 We start with describing the hardware environment (Section~\ref{sec:results_hardware}) followed by the software environment (Section~\ref{sec:results_software}) and we concluded by the evaluation results (Section~\ref{sec:results_evaluation}).
@@
 
-Measured size of tasks for each application is given in Figure~\ref{fig:aveVirtuTaskSize}.
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/virtualTaskSize.eps}
+	\caption{Execution time of a dummy application containing twelve empty tasks as a function of coalescing factor.
+We observe three times improvement from coalescing until four virtual tasks, while we observe a diminishing return from task coalescing beyond four.
+\todo{Improve the figure: make it narrower, larger font, remove legend, do not capitalize "Task" in X axis, add Y axis label (Execution time (ms))}{Amjad}}
+	\label{fig:virtualTaskSize}
+\end{figure}
+
+The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e.
+from energy harvesting to continuous power environment) is the best one?.
+To answer this question we have implemented a simple program composed of $x$ empty tasks.
+We execute the same program by increasing coalescing factor by one while the device on which the program was executed was running on a stable power supply (refer to Section~\ref{sec:methodology_evaluation} for details).
+The result is presented in Figure~\ref{fig:virtualTaskSize}.
+We observe a significant increase in execution time as task are coalescent from one to four (three times execution time improvement).
+However, as the coalescing increases beyond four, we observe diminishing returns from coalescing in terms of execution time.
+\todo{Decide whether to place this figure in Evaluation section or leave it here}{Przemek}
 
@@
 
-expected results:
+Measured size of tasks for each application is given in Figure~\ref{fig:aveVirtuTaskSize}.
+
+Expected results:
 \begin{itemize}

### 2017-08-07 paper/paging.tex

@@
-One of the main objectives of \sys run-time is to restrict virtual tasks to interface only with the volatile memory\todo{Do we define what a virtual task is and why is it important?}{Amjad/Sinan}, since volatile memory access is relatively cheaper than non-volatile memory access, as we already have shown experimentally in Section~\ref{section:background_task_computing} (Problem 1, Figure~\ref{fig:framEnergy}).
-Before dwelling into the details of virtualization of \sys, we need to shed more light on the rationale behind our approach.
+One of the main objectives of \sys run-time is to restrict tasks to interface only with the volatile memory, since volatile memory access is relatively cheaper than non-volatile memory access, as we already have shown experimentally in Section~\ref{section:background_task_computing} (Problem 1, Figure~\ref{fig:framEnergy}).
+This feature will also provide several advantages during task coalescing, as we will be dealing with in Section \ref{sec:task_coalescing}.
+Before dwelling into the details of memory virtualization of \sys, we need to shed more light on the rationale behind our approach.
 
-\subsection{Memory Virtualization versus Intermittence}
+\subsection{Intermittent Volatile-Only Memory Access}
 \label{sec:virtualization_problems}
 
-The persistent variables a virtual task operates on should be brought from non-volatile memory to a fixed-sized \emph{volatile buffer}.
-When the virtual task is finished the modified variables in this buffer should be committed back to their original locations in non-volatile memory.
+In order to enable volatile-only memory access, the persistent variables a task operates on should be brought from non-volatile memory to a fixed-sized \emph{volatile buffer} in main memory.
+During its execution, the active task reads and modifies only the contents of this buffer.
+When the task is finished its execution, the modified variables, i.e.
+memory locations, in this buffer should be committed back to their original locations in non-volatile memory.
 We list three fundamental problems proving that this is however a non-trivial task.
@@
 
-\textbf{Problem 1---Dynamic Access Overhead:} The persistent variables accessed and/or modified by a virtual task might not be known in advance due to the dynamic program flow.
-Therefore, the volatile buffer should keep track of the persistent variables accessed by any virtual task at runtime.
-When a persistent variable is read or written, first the volatile copy of the corresponding variable is \emph{searched} in this buffer.
-If it is found, the read or write operation is performed on the volatile copy.
-Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations.
-Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead and might kill the benefit of task virtualization.\todo{How to show it experimentally?}{Sinan} %refer to Figure~\ref{fig:virtOperationalBuf} for experimental evaluation.
-
 \begin{figure}[t]
 	\centering
-	\subfloat[SRAM Double Buffer Copy]{\includegraphics[width=0.49\columnwidth]{figures/sram-buffer} }
-	\subfloat[DMA Copy]{\includegraphics[width=0.49\columnwidth]{figures/dma}}
-	\caption{.\todo{Explain HW/SW setup; make fonts larger, unify legends and axes.
-We only need to show the right side; this figure shows that we operate out of SRAM and page to FRAM; the copying mechanism is peripheral to the main idea.}{Amjad}}
-	\label{fig:sram_vs_dma}
+	\includegraphics[width=\columnwidth]{figures/sram-buffer}
+	\caption{The interaction between the task, volatile buffer and the non-volatile memory (FRAM).
+Initially, the persistent variable \texttt{x} is not in the volatile buffer but \texttt{y} and \texttt{z} are.
+}
+	\label{fig:volatile-buffer}
 \end{figure}
 
-\textbf{Problem 2---Two Phase Commit Overhead:} For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each virtual task by means of a \emph{two-phase commit} approach.
-However, two-phase commit is also very expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed with CPU intervention, refer to Fig.~\ref{fig:sram_vs_dma}.
+\paragraph{Problem 1---Runtime Access Overhead:} The persistent variables read and/or modified by a task might not be known in advance due to the dynamic program flow.
+Therefore, the volatile buffer should keep track of the persistent variables at runtime, as presented in Fig.\,\ref{fig:volatile-buffer}.
+When a persistent variable is read or written---refer steps 1 and 3--5 in this figure, first the volatile copy of the corresponding variable is \emph{searched} in this buffer.
+If it is found, the read or write operation is performed on the volatile copy.
+Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations---refer step 2 in Fig.\,\ref{fig:volatile-buffer}.
+Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead and might kill the benefit of task coalescing as we present in Section \ref{sec:task_coalescing}.\todo{How to show it experimentally?}{Maybe we can show one of the results of Amjad?} %refer to Figure~\ref{fig:virtOperationalBuf} for experimental evaluation.
 
@@
 	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
-	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: pointers versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
+	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
 	\label{fig:dmaTimeEnergy}
@@
 
-\textbf{Problem 3---Main Memory Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
+\paragraph{Problem 2---Two Phase Commit Overhead:} For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each task by means of a \emph{two-phase commit} approach---refer Fig.\,\ref{fig:volatile-buffer} steps 6--8 for the first phase and 9--11 for the second step of the commit operation.
+However, two-phase commit is also expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed selectively with CPU intervention.
+Non-selective block memory copy operations using Direct Memory Access (DMA) that eliminates CPU intervention are considerably more efficient, refer to Fig.~\ref{fig:dmaTimeEnergy} for a comparison, but infeasible in this case.
+
+\paragraph{Problem 3---Main Memory Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
 In order to eliminate run-time buffer \emph{search} overhead (by not relying on specialized hardware such as e.g.
-in~\cite{hicks_isca_2017}), \emph{all} persistent variables can be brought from non-volatile memory to the volatile buffer at system start-up---very efficiently and faster thanks to Direct Memory Access (DMA) that eliminates CPU intervention for \emph{block} memory-memory copy/move operations, as shown experimentally in Figure~\ref{fig:dmaTimeEnergy}.
-Two-phase commit is also very efficient in this case since DMA can be utilized to move volatile buffer as a block to non-volatile locations.
+in~\cite{hicks_isca_2017}), \emph{all} persistent variables can be brought from non-volatile memory to the volatile buffer at system start-up---very efficiently and faster thanks to DMA for \emph{block} memory-memory copy/move operations.
+Two-phase commit is also very efficient in this case since DMA can be utilized to move volatile buffer as a block to the corresponding non-volatile locations.
 However, bringing all persistent variables from non-volatile memory to the volatile buffer suffers from the \emph{memory fit problem}: the volatile memory is a scarce resource and may not be large enough to hold all persistent variables.
 
-Considering these facts, \sys implements a \emph{paging-based} virtual memory in software that (i) partitions contiguously allocated persistent variables into pages; (ii) implements the volatile \emph{page buffer} that holds a single page---solves the aforementioned memory-fit problem; (ii) swaps the page in volatile memory with a new page in non-volatile memory efficiently by utilizing DMA-based memory block-copy operations; (iii) buffers the swapped volatile page until commit to enable task virtualization; (iv) implements two-phase commit of the modified pages to preserve the consistency of non-volatile memory; (v) and performs two-phase commit operations using DMA for the sake of efficiency.
-We explain the details of \sys virtualization design in the following sections.
+\subsection{Software-based Paging Mechanism}
+Considering these facts, \sys implements a \emph{paging-based} virtual memory in software that (i) partitions contiguously allocated persistent variables into pages; (ii) implements the volatile \emph{page buffer} that holds a single page---solves the aforementioned memory-fit problem; (iii) swaps the page in volatile memory with a new page in non-volatile memory efficiently by utilizing DMA-based memory block-copy operations; (iv) buffers the swapped volatile page until commit to enable task coalescing we describe in Section \ref{sec:task_coalescing}; (v) implements two-phase commit of the modified pages to preserve the consistency of non-volatile memory; (vi) and performs two-phase commit operations using DMA for the sake of efficiency.
+We explain the details of \sys paging mechanism in the following subsections.
 
-\subsection{Address Translation and Variable Access}
+\subsubsection{Address Translation and Variable Access}
 
 Our virtual memory system provides an interface to the applications which is composed of two C macros: \texttt{RVAR(var)} and \texttt{WVAR(var,val)} that accept the name of the persistent variable \texttt{var} as an input argument.
-These macros obtain the \emph{physical address} of the corresponding persistent variable in non-volatile memory and translates it into the \emph{virtual address} in volatile memory which is composed of a \emph{page tag} and \emph{offset}.
+These macros obtain the \emph{physical address} of the corresponding persistent variable in non-volatile memory and translates it into the \emph{virtual address} in page buffer \texttt{pageBuf} in volatile memory, which is composed of a \emph{page tag} and \emph{offset}.
 For the sake of simplicity and efficiency, we implemented this translation in a trivial way: the first byte of the physical address is considered to be the page tag and the rest is the offset within the corresponding page.
 
-\begin{figure}
-	\centering
-	%\includegraphics[width=\columnwidth]{figures/}
-	\caption{Address translation.\todo{Draw figure}{Sinan}}
-	\label{fig:address-translation}
-\end{figure}
-
 \begin{algorithm}[t]
-	\caption{\texttt{RWAR(var)} pseudo-code}
+	\caption{\texttt{RVAR(var)} pseudo-code}
 	\label{algo:rwar}
@@
 To this end, our system maintains a \texttt{CrntPagTag} variable that holds the tag of the page currently in the page buffer.
-If the tag of the virtual address and \texttt{CrntPagTag} are equal, this indicates that page buffer holds the required page and the offset of the virtual address is used to read from/write to the location in the volatile page buffer immediately.
+If the tag of the virtual address and \texttt{CrntPagTag} are equal, this indicates that page buffer holds the required page and the offset of the virtual address can be used to read from/write to the location in the volatile page buffer immediately.
 If the required page is not in the page buffer, a \emph{page fault} routine is executed as we discuss in the next subsection.
 Observe that the computational complexity of the whole steps pertaining to address translation and variable access is $\mathcal{O}(1)$---introducing minimal overhead to the system to boost up benefits of volatile-only memory interfacing.
-The pseudo-code of \texttt{RWAR} is presented in Algorithm~\ref{algo:rwar}.
+The pseudo-code of \texttt{RVAR}, which is quite similar to that of \texttt{WVAR}, is presented in Algorithm~\ref{algo:rwar}.
 
-\subsection{Page Faults and Swapping}
-
-\todo{Section to be expanded}{Sinan}
+\subsubsection{Page Faults and Page Swapping}
 
@@
 	\begin{algorithmic}[1]
-		\If { \texttt{isModified(CrntPagTag)} }	\Comment{Check if the active page modified}
-		\State commit current page to \texttt{pagesTemp}
+		\If {\texttt{isDirty(CrntPagTag)} }	\Comment{Check if the active page is modified}
+		\State \texttt{pageCommit(pagesTemp)} \Comment{Phase-I commit the active page}
 		\EndIf
-		\If { tag is in \texttt{pagesTemp} }	\Comment{Check if the requested page has been modified}
-		\State bring from \texttt{pagesTemp} to \texttt{pageBuff} 
+		\If {\texttt{isDirty(tag)}} \Comment{Check if the requested page has been modified}
+		\State {\texttt{getPage(tag,pagesTemp)}=false} \Comment{Get page from temporary buffer}
 		\Else
-		\State bring from original to \texttt{pageBuff} 
+		\State \texttt{getPage(tag,pagesOrg)}==false \Comment{Get page from original location}
 		\EndIf 
+		\State \texttt{CrntPagTag} = \texttt{tag} \Comment{Update tag}
 	\end{algorithmic}
@@
 
-If the tag of the virtual address and \texttt{CrntPagTag} are not equal, the active page in page buffer should be swapped with the requested page. In this situation, we have the following two cases:
+If the page tag of the variable \texttt{var} and \texttt{CrntPagTag} are not equal (e.g.
+Lines 2--3 of Algorithm\,\ref{algo:rwar}), the active page, i.e.
+the contents of the page buffer \texttt{pageBuf}, should be swapped with the page whose tag is equal to the \texttt{var}'s tag.
+The pseudo-code of the Algorithm \texttt{PageFault} that handles this operation is presented in Algorithm\,\ref{algo:pagefault}.
+This algorithm handles two fundamental cases:
 
-\paragraph{Modified Active Page:} If the contents of the page buffer is modified, we commit it to an intermediate page buffer in non-volatile memory, namely to \texttt{pagesTemp}.
-If it is not modified, there is no need to commit the active page.
+\paragraph{Modified Active Page:} If the task performed at least \texttt{WVAR} operation, then the page buffer \texttt{pageBuf} that holds the current active page is modified (Line 1 of Alg.\,\ref{algo:pagefault}).
+In order to keep the modified values of the variables, the active page is committed to an intermediate page buffer in non-volatile memory, namely to \texttt{pagesTemp}.
+This operation can be considered as first phase of the commit operation, as we present in the next subsection.
+If the active page is not modified, there is no need to commit it to the temporary page buffer.
 
-\paragraph{Modified Requested Page:}  If the requested page is previously modified, it should be brought from \texttt{pagesTemp} rather than its original location in non-volatile memory.
+\paragraph{Modified Requested Page:} After committing the active page to its temporary location, Algorithm\,\ref{algo:pagefault} checks if the requested page is modified previously by the same task (Line 3).
+If this is the case, the requested page should be be brought from \texttt{pagesTemp} rather than its original location in non-volatile memory (Line 4), so that the task accesses the modified values of the persistent variables.
+Otherwise, the requested page is brought from its original location, i.e.
+\texttt{pagesOrg} (Line 6).
+After the swapping operation, the \texttt{CrntPagTag} is set to the tag of the requested page (Line 7).
+It should be noted that DMA is used to memory block copy operations, introducing minimal overhead during page swapping.
 
-After the swapping, the \texttt{CrntPagTag} is set to the tag of the requested page.
-It should be noted that DMA is used to introduce minimal overhead during page swapping.
-The pseudo-code of the page fault handling is presented in Algorithm~\ref{algo:pagefault}.
+\subsubsection{Pages Final Committing}
 
-\subsection{Page Committing}
+After a task finishes its execution, all pages committed to \texttt{pagesTemp} should be committed to their original locations in \texttt{pagesOrg}.
+This is the second phase of the \emph{two-phase commit} operation.
+As mentioned previously, memory block copy operations for the second phase commit is also utilized by DMA.
 
-Using a two-phase commit by utilizing DMA. \todo{Text to be written}{Sinan}
-
-%%% Old Text by Amjad %%%
-
-%\begin{algorithm}[t]
-%	\caption{Virtualized Operational Buffer}
-%	\label{algo:virtuBufWrite}
-%	\scriptsize
-%	%\small
-%	\begin{algorithmic}[1]
-%		\State $var \in \text{\{global variables\}} $ 
-%		\State \label{lst:virtuBufWrite:line:begin}\Call{Virtual Task()}{} 
-%		\While { \textit{executing} } \Comment{Execution stage}
-%		\State $var$  $\rightarrow$ \textsf{volatile buffer} \label{lst:virtuBufWrite:line:output}
-%		\If { $var$  in \textsf{ volatile buffer} }				\label{lst:virtuBufWrite:line:inputBegin}
-%		\State $var$  $\leftarrow$  \textsf{volatile buffer} 
-%		\Else 
-%		\State $var$  $\leftarrow$  \textsf{FRAM}		\label{lst:virtuBufWrite:line:inputEnd}
-%		\EndIf
-%		
-%		\If {power interrupts}
-%		\State back to \ref{lst:virtuBufWrite:line:begin}
-%		\EndIf
-%		\EndWhile
-%		
-%		\While{  $\textsf{volatile buffer}\not=\emptyset$  } \label{lst:virtuBufWrite:line:commitBegin} \Comment{ First phase commit}
-%		\State \textsf{volatile buffer} $\rightarrow$  \textsf{persistent buffer}
-%		\If {power interrupts}
-%		\State discard \textsf{persistent buffer}
-%		\State back to  \ref{lst:virtuBufWrite:line:begin} 
-%		\State   \label{lst:virtuBufWrite:line:commitEnd}
-%		\EndIf
-%		\EndWhile 
-%		
-%		\While{ $\textsf{persistent buffer}\not=\emptyset$ } \label{lst:virtuBufWrite:line:SecCommitBegin} \Comment{Second phase commit}
-%		\State \textsf{persistent buffer} $\rightarrow$ FRAM 
-%		\If {power interrupts}
-%		\State Continue
-%		\EndIf
-%		\EndWhile 
-%		\State     \label{lst:virtuBufWrite:line:SecCommitEnd}
-%		\State return
-%	\end{algorithmic}
-%\end{algorithm}
-
-%Since accessing FRAM is more energy expensive and slower than accessing SRAM, preventing frequent access to FRAM (i.e.
-during looping operations) is desirable.
-Therefore, the first proposed protection method utilizes a volatile buffer that holds temporary all the outputs of a virtual task (see Algorithm~\ref{algo:virtuBufWrite} line~\ref{lst:virtuBufWrite:line:output}).
-Consequently, a virtual task must first attempt to read a global variable from the volatile buffer before trying to obtain the value from the non-volatile memory (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:inputBegin}-\ref{lst:virtuBufWrite:line:inputEnd}) to ensure a correct execution progress and the consistency of the memory.
-Once the execution of a virtual task is done, the first phase of the commit process is started by copying the volatile buffer to a persistent buffer (see Algorithm~\ref{algo:virtuBufWrite} lines~\ref{lst:virtuBufWrite:line:commitBegin}-\ref{lst:virtuBufWrite:line:commitEnd}).
-If the power is interrupted the persistent buffer must be discarded and the execution must start again from the beginning of the virtual task.
-In another words, the first phase commit must be performed atomically to preserve the consistency of the output of a virtual task.
-The second phase commit is a power failure immune precess that is responsible for distributing the global variables to their final locations and make the non-volatile memory consistent and synchronized with the computation progress (see Algorithm~\ref{algo:virtuBufWrite}lines~\ref{lst:virtuBufWrite:line:SecCommitBegin}-\ref{lst:virtuBufWrite:line:SecCommitEnd}).
-
-%\paragraph{Virtualized Operational Buffer: Implementation}
-
-%We implement a reference implementation of \sys that uses virtualized Operational buffer to protect the data against power failure.
-
-%\paragraph{Virtual Buffer}
-
-%We realized the virtaulzed buffer as a volatile hashed table of linked lists.
-The hashing technique was chosen to reduce the buffer searching time and the linked lists are used to prevent data loss when there is a conflict between multiple variables---If two variables have the same hash value they will occupy the same cell, however, with linked list a new node will be created for each variable to resolve the conflict and protect the data.
-The hashing function is based on the observation that the virtual addresses of the memory cells have approximately a flat distribution over the memory addressing space and the fact that each entry, of the virtual table, has the address and the value of a variable.
-As such, by using the least significant bits as an index to access the virtual buffer the hash function distributes its inputs uniformly in the virtual buffer.
-Accordingly, the linked lists search time, on average, is reduced.
-
-%Despite the fact that the linked lists reduce the average complexity of searching the virtual buffer significantly, they introduce a number of drawbacks: (i) the linked list data structure introduce a non-negligible memory overhead while it relies on a limited memory section, namely the heap; (ii) traversing the linked list is relatively slow.
-Therefore, we re-implemented the virtual buffer, after observing that most of the global variables have the same most significant bits, as a bigger hashed table that does not allow entries conflict.
-To eliminate the need for searching the hashed table linearly while committing its entries, we add a new buffer that holds the indices of the occupied cells of the hashed table.
-This table is of the same length as the hashed table.
-We can reason about the benefit of this buffer as follows: if we assume that the global variables are evenly divided between the tasks, then this buffer reduces the commit time by a factor equal to the number of tasks.
-
-%\paragraph{Persistent Buffer}
-
-%The persistent buffer is implemented as static array of tuples, where each tuple holds  the address and the value of a variable.
-This data structure is very suitable for the second phase commit where each element has to be committed to its finial location.
-However, the complexity of committing this buffer is of size $O(N)$, where $N$ equals the length of the buffer.
-Therefore, the size of the buffer can have dramatic effect on the performance of \sys---On one hand, if the size of the buffer is small the buffer overflow problem will be very serious.
-On the other hand, if the size of the buffer is large \sys will experience a significant performance degradation.
-However, by observing the type of the information that this buffer holds, namely memory addresses, we can, on average, reduce the complexity of committing this buffer by defining the \emph{effective size} of the buffer to be the size of the buffer up to the first cell that holds an invalid memory address.
-Moreover, since this buffer is in FRAM which is much bigger than SRAM its size can be relatively big.
-
-% \paragraph{Virtualized Operational Buffer: delay analysis}[initial results]
-
-% \begin{figure}[t]
-% 	\centering
-% 	\includegraphics[width=\columnwidth]{figures/virtual_buffer_size.eps}
-% 	\caption{\sys based on virtualized Operational Buffer.
-It runs the decompression application with different buffer configurarions.
-EPB: Enhanced Persistent Buffer.
-NR: no read operation from the buffer.
-NRW: not read or write operations from the buffer.}
-% 	\label{fig:virtOperationalBuf}
-% \end{figure}
-
-% 	To analyze the performance of \sys versus the sizes of the virtual buffer and the persistent buffer we setup the following experiment.
-We let IPOS to run the decompression application---until the data decompression is completely finished---and we capture the execution time using a logic analyzer~\cite{xxx}.
-The results presented in Fig.~\ref{fig:virtOperationalBuf} show the \sys has its best performance when the virtual buffer size is 16, where the execution time was $\approx$196 ms.
-
-% 	All the experiments were done with a persistent buffer of size 100.
-However, the bar labeled with "EPB/16" in  Fig.~\ref{fig:virtOperationalBuf} shows the effect of the enhanced commit process that uses the buffer effective size instead of the buffer maximum size.
-We see that the execution time is reduced to $\approx$80 ms.
-Furthermore, we quantized the overhead of the read operations from the buffer, by eliminating this operation, and the bar labeled with "NR" shows that it cost  $\approx$10 ms.
-Moreover, the overhead of the read and write operations is $\approx$34 ms as compared to \sys best performance, see the bar labeled with "NRW" in Fig.~\ref{fig:virtOperationalBuf}.
+\subsection{Non-Volatile Memory Consistency Issues}  
+During the execution of any task, modifications to the persistent variables will not be reflected to the original locations in the non-volatile memory, since the task interfaces with only the volatile page buffer \texttt{pageBuf} and the modifications are temporarily stored in the non-volatile \texttt{pagesTemp} buffer.
+The changes stored in the \texttt{pagesTmp} is committed to the original locations in \texttt{pagesOrg} after the task is finished its execution.
+Therefore, any power interruptions within the execution will not lead to inconsistencies in the \texttt{pagesOrg} when the task is re-executed after a power boot.
+As a consequence, the proposed paging mechanism keeps the consistency of the non-volatile memory thanks to the two-phase commit operation for the modified pages.
+\todo{Benefits of the paging to the task coalescing should be mentioned in the next section}{Amjad/Przemek/Sinan}

### 2017-08-07 paper/paging.tex
Revisions of virtualization section

@@
-One of the main objectives of \sys run-time is to restrict tasks to interface only with the volatile memory, since volatile memory access is relatively cheaper than non-volatile memory access, as we already have shown experimentally in Section~\ref{section:background_task_computing} (Problem 1, Figure~\ref{fig:framEnergy}).
+One of the main objectives of \sys run-time is to restrict tasks to interface only with the volatile memory, since volatile memory access is relatively cheaper, as we already have shown experimentally in Section~\ref{section:background_task_computing} (Problem 1, Figure~\ref{fig:framEnergy}).
 This feature will also provide several advantages during task coalescing, as we will be dealing with in Section \ref{sec:task_coalescing}.
-Before dwelling into the details of memory virtualization of \sys, we need to shed more light on the rationale behind our approach.
 
@@
 
-In order to enable volatile-only memory access, the persistent variables a task operates on should be brought from non-volatile memory to a fixed-sized \emph{volatile buffer} in main memory.
+In order to enable volatile-only memory access, the persistent variables a task operates on should be brought from non-volatile memory to a fixed-sized \emph{volatile buffer} in the main memory.
 During its execution, the active task reads and modifies only the contents of this buffer.
@@
 
-%?\begin{figure}
-%?	\centering
-%?	%\includegraphics[width=\columnwidth]{figures/}
-%?	\caption{\sys memory virtualization architecture.\todo{Decide if we really need this figure (we already have a lot of figs)}{Brandon}}
-%?	\label{fig:virtualization_architecture}
-%?\end{figure}
-
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=\columnwidth]{figures/iposCommitSize}
-%	\caption{Commit size delay for three memory access methods.}
-%	\label{fig:virtOperationalBuf}
-%\end{figure}
-
 \begin{figure}[t]
@@
 	\caption{The interaction between the task, volatile buffer and the non-volatile memory (FRAM).
-Initially, the persistent variable \texttt{x} is not in the volatile buffer but \texttt{y} and \texttt{z} are.
-}
+Initially, the persistent variable \texttt{x} is not in the volatile buffer but \texttt{y} and \texttt{z} are.}
 	\label{fig:volatile-buffer}
@@
 \paragraph{Problem 1---Runtime Access Overhead:} The persistent variables read and/or modified by a task might not be known in advance due to the dynamic program flow.
-Therefore, the volatile buffer should keep track of the persistent variables at runtime, as presented in Fig.\,\ref{fig:volatile-buffer}.
-When a persistent variable is read or written---refer steps 1 and 3--5 in this figure, first the volatile copy of the corresponding variable is \emph{searched} in this buffer.
+Therefore, the volatile buffer should keep track of the persistent variables at runtime, as presented in Figure~\ref{fig:volatile-buffer}.
+When a persistent variable is read or written (refer to steps 1 and 3--5 in Figure~\ref{fig:volatile-buffer}) first the volatile copy of the corresponding variable is \emph{searched} in this buffer.
 If it is found, the read or write operation is performed on the volatile copy.
-Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations---refer step 2 in Fig.\,\ref{fig:volatile-buffer}.
-Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead and might kill the benefit of task coalescing as we present in Section \ref{sec:task_coalescing}.\todo{How to show it experimentally?}{Maybe we can show one of the results of Amjad?} %refer to Figure~\ref{fig:virtOperationalBuf} for experimental evaluation.
+Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations (refer step 2 in Figure~\ref{fig:volatile-buffer}).
+Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead (refer to Figure~\ref{fig:dmaTimeEnergy} and pointer-based operation therein) and might nullify the benefit of task coalescing, a cores \sys feature.
 
@@
 
-\paragraph{Problem 2---Two Phase Commit Overhead:} For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each task by means of a \emph{two-phase commit} approach---refer Fig.\,\ref{fig:volatile-buffer} steps 6--8 for the first phase and 9--11 for the second step of the commit operation.
+\paragraph{Problem 2---Two Phase Commit Overhead:} For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each task by means of a \emph{two-phase commit}---refer Figure~\ref{fig:volatile-buffer} steps 6--8 for the first phase and 9--11 for the second step of the commit operation.
 However, two-phase commit is also expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed selectively with CPU intervention.
-Non-selective block memory copy operations using Direct Memory Access (DMA) that eliminates CPU intervention are considerably more efficient, refer to Fig.~\ref{fig:dmaTimeEnergy} for a comparison, but infeasible in this case.
+Non-selective block memory copy operations using Direct Memory Access (DMA) that eliminates CPU intervention are considerably more efficient, refer again to Figure~\ref{fig:dmaTimeEnergy} for a comparison, but infeasible in this case.
 
-\paragraph{Problem 3---Main Memory Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
+\paragraph{Problem 3---Memory Fit Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
 In order to eliminate run-time buffer \emph{search} overhead (by not relying on specialized hardware such as e.g.
@@
 \subsection{Software-based Paging Mechanism}
-Considering these facts, \sys implements a \emph{paging-based} virtual memory in software that (i) partitions contiguously allocated persistent variables into pages; (ii) implements the volatile \emph{page buffer} that holds a single page---solves the aforementioned memory-fit problem; (iii) swaps the page in volatile memory with a new page in non-volatile memory efficiently by utilizing DMA-based memory block-copy operations; (iv) buffers the swapped volatile page until commit to enable task coalescing we describe in Section \ref{sec:task_coalescing}; (v) implements two-phase commit of the modified pages to preserve the consistency of non-volatile memory; (vi) and performs two-phase commit operations using DMA for the sake of efficiency.
-We explain the details of \sys paging mechanism in the following subsections.
+
+Considering these facts, \sys implements a \emph{paging-based} virtual memory in software that (i) partitions contiguously allocated persistent variables into pages; (ii) implements the volatile \emph{page buffer} that holds a single page; (iii) swaps the page in volatile memory with a new page in non-volatile memory efficiently by utilizing DMA-based memory block-copy operations; (iv) buffers the swapped volatile page until commit to enable task coalescing which we describe in Section~\ref{sec:task_coalescing}; (v) implements two-phase commit of the modified pages to preserve the consistency of non-volatile memory; and (vi) performs this two-phase commit operations using DMA for the sake of efficiency.
+We now explain the details of \sys paging mechanism.
 
@@
 
-Our virtual memory system provides an interface to the applications which is composed of two C macros: \texttt{RVAR(var)} and \texttt{WVAR(var,val)} that accept the name of the persistent variable \texttt{var} as an input argument.
+\sys virtual memory system provides an interface to the applications which is composed of two C macros: \texttt{RVAR(var)} and \texttt{WVAR(var,val)} that accept the name of the persistent variable \texttt{var} as an input argument.
 These macros obtain the \emph{physical address} of the corresponding persistent variable in non-volatile memory and translates it into the \emph{virtual address} in page buffer \texttt{pageBuf} in volatile memory, which is composed of a \emph{page tag} and \emph{offset}.
@@
 		\State $\texttt{tag}\leftarrow \texttt{getTag(var)}$ 
-		\If { \texttt{tag} != \texttt{CrntPagTag} }	\Comment{Check if the page is in page buffer}
+		\If { \texttt{tag} != \texttt{CrntPagTag} }	\Comment{Check if the page is in the page buffer}
 		\State	\texttt{PageFault(tag)} \Comment{Page is not in the page buffer, bring it}
@@
 After address translation and obtaining the virtual address, it is required to check if the corresponding page is already in volatile page buffer.
-To this end, our system maintains a \texttt{CrntPagTag} variable that holds the tag of the page currently in the page buffer.
+\sys maintains a \texttt{CrntPagTag} variable that holds the tag of the page currently in the page buffer.
 If the tag of the virtual address and \texttt{CrntPagTag} are equal, this indicates that page buffer holds the required page and the offset of the virtual address can be used to read from/write to the location in the volatile page buffer immediately.
-If the required page is not in the page buffer, a \emph{page fault} routine is executed as we discuss in the next subsection.
+If the required page is not in the page buffer, a \emph{page fault} routine is executed as we discuss in the section below.
 Observe that the computational complexity of the whole steps pertaining to address translation and variable access is $\mathcal{O}(1)$---introducing minimal overhead to the system to boost up benefits of volatile-only memory interfacing.
@@
 		\If {\texttt{isDirty(CrntPagTag)} }	\Comment{Check if the active page is modified}
-		\State \texttt{pageCommit(pagesTemp)} \Comment{Phase-I commit the active page}
+		\State \texttt{pageCommit(pagesTemp)} \Comment{Phase I commit the active page}
 		\EndIf
 		\If {\texttt{isDirty(tag)}} \Comment{Check if the requested page has been modified}
-		\State {\texttt{getPage(tag,pagesTemp)}=false} \Comment{Get page from temporary buffer}
+		\State {\texttt{getPage(tag,pagesTemp)}=false} \Comment{Get page from the temporary buffer}
 		\Else
-		\State \texttt{getPage(tag,pagesOrg)}==false \Comment{Get page from original location}
+		\State \texttt{getPage(tag,pagesOrg)}=false \Comment{Get page from the original location}
 		\EndIf 
@@
 the contents of the page buffer \texttt{pageBuf}, should be swapped with the page whose tag is equal to the \texttt{var}'s tag.
-The pseudo-code of the Algorithm \texttt{PageFault} that handles this operation is presented in Algorithm\,\ref{algo:pagefault}.
-This algorithm handles two fundamental cases:
+The pseudo-code of the Algorithm \texttt{PageFault} that handles this operation is presented in Algorithm~\ref{algo:pagefault}.
+This algorithm handles two fundamental cases.
 
-\paragraph{Modified Active Page:} If the task performed at least \texttt{WVAR} operation, then the page buffer \texttt{pageBuf} that holds the current active page is modified (Line 1 of Alg.\,\ref{algo:pagefault}).
+\paragraph{Case 1---Modified Active Page:} If the task performed at least \texttt{WVAR} operation, then the page buffer \texttt{pageBuf} that holds the current active page is modified (Line 1 of Algorithm~\ref{algo:pagefault}).
 In order to keep the modified values of the variables, the active page is committed to an intermediate page buffer in non-volatile memory, namely to \texttt{pagesTemp}.
-This operation can be considered as first phase of the commit operation, as we present in the next subsection.
+This operation can be considered as first phase of the commit operation, as we present in the next section.
 If the active page is not modified, there is no need to commit it to the temporary page buffer.
 
-\paragraph{Modified Requested Page:} After committing the active page to its temporary location, Algorithm\,\ref{algo:pagefault} checks if the requested page is modified previously by the same task (Line 3).
+\paragraph{Case 2---Modified Requested Page:} After committing the active page to its temporary location, Algorithm~\ref{algo:pagefault} checks if the requested page is modified previously by the same task (Line 3).
 If this is the case, the requested page should be be brought from \texttt{pagesTemp} rather than its original location in non-volatile memory (Line 4), so that the task accesses the modified values of the persistent variables.
@@
 \subsection{Non-Volatile Memory Consistency Issues}
+
 During the execution of any task, modifications to the persistent variables will not be reflected to the original locations in the non-volatile memory, since the task interfaces with only the volatile page buffer \texttt{pageBuf} and the modifications are temporarily stored in the non-volatile \texttt{pagesTemp} buffer.
@@
 As a consequence, the proposed paging mechanism keeps the consistency of the non-volatile memory thanks to the two-phase commit operation for the modified pages.
-\todo{Benefits of the paging to the task coalescing should be mentioned in the next section}{Amjad/Przemek/Sinan}
+
+%\todo{Benefits of the paging to the task coalescing should be mentioned in the next section}{Amjad/Przemek/Sinan}

### 2017-08-07 paper/coalescing.tex
Section 3 revised (good job Brandon)
pseudo code of the new coalescing algo

@@
-\sys is able to virtually merge tasks to construct a bigger virtual task and commit the state of the virtual task instead of the individual real tasks to improve execution time at stable energy supply environment.
+\sys is able to dynamically merge real tasks to construct a bigger virtual task. This enables \sys to reduce the number of commits by a value equals to 
+$\frac{\text{VTS}-1}{\text{VTS}}$, 
+where VTS is virtual task size.
+Furthermore, \sys trys to estimiate the optimial virtual task size by relying on its recent execution history.
+The challenge in designing the \sys's coalescing algorithm is in the fact that it should not introduce as less overhread as possible, and able to respond to the changes in the environment.
 
@@
 	\caption{\sys task coalescing mechanism}
-	\label{algorithm:coalescing_algorithm}
+	\label{algo:coalescing}
 	\scriptsize
 	\begin{algorithmic}[1]
-		\State $\text{VT} \subset \{\sys~\text{tasks set}\}$  \Comment{Virtual task}
-		\State $|\text{VT}|$ \Comment{VT size \todo{Define unit}{Amjad}}
-		\State $\text{VT}_{\max}$ \Comment{Maximum VT size (With upper bound on coalescing)}
+		\State $\text{p}$  \Comment{ indicates persistent variables, initialized \textbf{only} during divice programming}
+		\State $\text{RT} \in \{\sys~\text{static tasks}\}$  \Comment{Real task}
+		\State $\text{VT} \subset \{~\text{RT's}\}$  \Comment{Virtual task}
+		\State $\text{RTC}$  \Comment{RT counter}
+		\State $\text{VTC}$  \Comment{VT counter}
+		% \State $|\text{VT}|$ \Comment{VT size measured in RT}
+		\State $\text{TVT}$ \Comment{temprary VT}
+		\State $\text{VT}_{\max}$ \Comment{Maximum VT size measured in RT (coalescing upper bound)}
+		\vspace{0.1cm}
+
+		\State $\text{p RTC} = x $ 
+		\State $\text{p VT}_{\max} = y$
+		\State $\text{p VT} = \text{RT}$ 
+		\State $\text{p TVT} = \text{VT} $ 
+		\State $\text{VTC} = 0 $ 
+		\vspace{0.1cm}
+
+		\Function{scheduler}{\null}
+
+			\State $\text{VTC} = \text{RTC/2} $  \label{algo:coalescing:executionHistory1}
+			\State $\text{RTC} = 0 $
+			\State $\text{RT} = \text{VT}$ \Comment{recover virtual state} \label{algo:coalescing:virtualProgressing1}
+			\If{commiting}
+				\State $\text{RT}=\text{TVT}$
+				\State $\texttt{goto} \ \ \ref{commitStage}$ \label{algo:coalescing:firmTransition1}
+			\EndIf
 			\vspace{0.1cm}
@@
 			\While {True}
-		\State $\text{VT} \leftarrow \text{VT}_{\text{next}}$ \Comment{\todo{Define $\text{VT}_{\text{next}}$}{Amjad}}
-		\vspace{0.1cm}
-		\While {execute VT} 
-		\If { power failed $n$ times during task execution} \Comment{\todo{Define how it is detected}{Amjad}}				
-		\State $|\text{VT}|:=|\text{VT}|-1$
-		\State $\text{VT}_{\max} := |\text{VT}|$ \Comment{With upper bound on coalescing}
+				\While{$\text{VTC}--$}
+					\Function{execute}{$\text{RT}$}
+					\EndFunction
+					\If{$\text{RTC} < \text{VT}_{\max}$}
+						\State $\text{RTC}++$  \label{algo:coalescing:realTaskCounter}
 					\EndIf
+					\State $\text{RT} \leftarrow \text{RT}_{next}$ \Comment{virtual progressing}
+				\EndWhile        \label{algo:coalescing:virtualProgressing2}
+
+				\State \Comment{virtual task is finished}
+				\State $\text{Data} \rightarrow \text{pTempBuf}$ \Comment{pTempBuf: persistent temp. buffer}
+				\State $\text{TVT} = \text{RT}$ 
+				\State $\text{commiting} = \text{True}$ \label{algo:coalescing:firmTransition2}
+				\State 
+				\State $\text{VT} = \text{TVT}$ \label{commitStage}
+				\State $\text{pTempBuf} \rightarrow \text{FRAM}$ 
+				\State $\text{VTC} = \text{RTC/2}$ 		\Comment{Set VT size} \label{algo:coalescing:executionHistory2}
+				\State $\text{commiting} = \text{False}$
+
 			\EndWhile
 
-		\vspace{0.1cm}
-		\If {All tasks executed} \Comment{\todo{Define how it is detected and when}{Amjad}}
-		\If{$|\text{VT}| < \text{VT}_{\max}$} \Comment{line added in fixed size approach}
-		\State $|\text{VT}|:=|\text{VT}|+1$
-		\EndIf
-		\EndIf
-		\EndWhile
+		\EndFunction
+
 	\end{algorithmic}
@@
 
-The complete algorithm is provided in Algorithm~\ref{algorithm:coalescing_algorithm}. \todo{Fill in with missing information: scheduler description, implementation details, discussion}{Przemek/Amjad}
+\sys's coalescing algorithm uses a single point back in the \sys's execution history to estimate the appropriate virtual task size.
+However, it takes a conservative approach by making the virtual task size equals to the half of the number of the successfully executed real tasks before the last power interrupt (see lines~\ref{algo:coalescing:executionHistory1},\ref{algo:coalescing:executionHistory2} of algorithm~\ref{algo:coalescing} ).
+We advocate the algorithm conservative approach by the following two reasons: (i) The power interrupts are not uniformly distributed over the tasks; and (ii) the re-execution penalty, in case of a power interrupt, grows linearly with the size of a virtual task.
+Furthermore, the algorithm considers the charging rate by allowing the real tasks counter (RTC), which represents the execution history between the last two power interrupts, to increase from one up to multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of algorithm~\ref{algo:coalescing}).
+As a consequence, the virtual task size changes rapidly (since it is half of the execution history) in response to the change in the charging rate.
+However, the algorithm sets a limit to the size of the virtual task for these two reasons: (i) having an infinity virtual task size, results in a certain computation progress loss; and (ii) the benefit of task merging is governed by diminishing return principle.
+
+The algorithm protects itself from power interrupts by going into three stages: (i) virtual progressing, where the operations are done on the volatile memory (see lines~\ref{algo:coalescing:virtualProgressing1}-\ref{algo:coalescing:virtualProgressing2} of algorithm~\ref{algo:coalescing} ); (ii) transition stage, where the volatile state is moved to a temporary persistent buffer.
+Until the end of the second stage if power interrupts, then all the virtual progress will be canceled and the virtual task will re-executed and re-initialized from a consistent input (see line~\ref{algo:coalescing:virtualProgressing1} of algorithm~\ref{algo:coalescing}); and (iii) persistent commit stage, upon entering this stage the algorithm will make a firm transition (see lines~\ref{algo:coalescing:firmTransition1},\ref{algo:coalescing:firmTransition2} of algorithm~\ref{algo:coalescing}) and it will not go back to the other stages unless all the data is committed from  the persistent buffer to non-volatile memory.
+Since the data is being moved in one direction and from a persistent location, power interrupt is tolerable at this stage.
 

### 2017-08-07 DMA/ipos-gcc-dma/ipos/psched.c

@@ -41,7 +41,7 @@
             __jump_cnt = 0;    \
         }
 
-
+#define DEBUG
 #define COALESCING 1
 
     unsigned char __jump = 0;
@@ -51,8 +51,9 @@
 __nv unsigned char  __commit_flag = 0;
 
 __nv unsigned int  __virtualTaskAddr = 0;    // Modified externally
-__nv unsigned int *__temp_virtualTAskAddr = NULL;
-__nv unsigned int  __virtualTaskSize = VERTUTASK;
+__nv unsigned int *__temp_virtualTaskAddr = NULL;
+
+//__nv unsigned int  __virtualTaskSize = VERTUTASK;
      unsigned int  __virtualTaskCntr = 0;
 __nv unsigned int  __maxVirtualTaskSize = 0x7f;
 
@@ -83,9 +84,8 @@ void os_scheduler()
 {
 
 # if COALESCING
-    __virtualTaskSize = __realTaskCntr;
-    __virtualTaskCntr = (__virtualTaskSize >> 1)+1 ;  // make half of the last execution history, your new virtual task size
-    __realTaskCntr = 0;
+    __virtualTaskCntr = (__realTaskCntr >> 1)+1 ;  // make half of the last execution history, your new virtual task size
+    __realTaskCntr = 0;    // on a power reboot the __realTaskCntr must be reseted
 #endif
 
     if (__commit_flag == COMMITTING)
@@ -93,7 +93,7 @@ void os_scheduler()
         // we need to do a page swap to bring the correct page from temp  or  ROM
         // this page swap must not send a page, and that must not happen since the dirtyPag is 0 at this stage
         __bringPersisCrntPag(__persis_CrntPagHeader );
-        __realTask = __temp_virtualTAskAddr;
+        __realTask = __temp_virtualTaskAddr;
         goto commit;
     }
 
@@ -107,11 +107,13 @@ void os_scheduler()
 
 #if COALESCING
 
+#if DEBUG
         if(pers_cnt < 1024){
             *(pers_pt+pers_cnt) = __virtualTaskCntr;  // Debugging
             pers_cnt++;
         }
 
+#endif
         for(; __virtualTaskCntr; __virtualTaskCntr-- )
         {
 #endif
@@ -123,10 +125,9 @@ void os_scheduler()
 
 # if COALESCING
             //last execution history (my next virtual task size)
+            // set a maximum virtual task size (64)
+            if(__realTaskCntr < __maxVirtualTaskSize){
             __realTaskCntr++;
[…]

### 2017-08-07 paper/coalescing.tex
Task coalescing first pass ready

@@
-\sys is able to dynamically merge real tasks to construct a bigger virtual task. This enables \sys to reduce the number of commits by a value equals to 
-$\frac{\text{VTS}-1}{\text{VTS}}$, 
-where VTS is virtual task size.
-Furthermore, \sys trys to estimiate the optimial virtual task size by relying on its recent execution history.
-The challenge in designing the \sys's coalescing algorithm is in the fact that it should not introduce as less overhread as possible, and able to respond to the changes in the environment.
+\sys is able to dynamically merge real tasks to construct a bigger virtual task.
+This enables \sys to reduce the number of commits by $\frac{\text{VTS}-1}{\text{VTS}}$, where VTS is virtual task size.
+Furthermore, \sys tries to estimate the optimal virtual task size by relying on its recent execution history.
+The challenge in designing \sys's coalescing algorithm is in introducing as less overhead as possible and able to respond to the changes in the environment as quickly as possible.
 
-\subsection{Task Merging Algorithm}
+\subsection{Task Coalescing Algorithm}
 
@@
 	\begin{algorithmic}[1]
-		\State $\text{p}$  \Comment{ indicates persistent variables, initialized \textbf{only} during divice programming}
+		\State $\text{p}$  \Comment{Indicates persistent variables, initialized \textbf{only} during device programming}
 		\State $\text{RT} \in \{\sys~\text{static tasks}\}$  \Comment{Real task}
@@
 		% \State $|\text{VT}|$ \Comment{VT size measured in RT}
-		\State $\text{TVT}$ \Comment{temprary VT}
+		\State $\text{TVT}$ \Comment{Temporary VT}
 		\State $\text{VT}_{\max}$ \Comment{Maximum VT size measured in RT (coalescing upper bound)}
@@
 			\State $\text{RTC} = 0 $
-			\State $\text{RT} = \text{VT}$ \Comment{recover virtual state} \label{algo:coalescing:virtualProgressing1}
+			\State $\text{RT} = \text{VT}$ \Comment{Recover virtual state} \label{algo:coalescing:virtualProgressing1}
 			\If{commiting}
@@
 					\EndIf
-					\State $\text{RT} \leftarrow \text{RT}_{next}$ \Comment{virtual progressing}
+					\State $\text{RT} \leftarrow \text{RT}_{next}$ \Comment{Virtual progressing}
 				\EndWhile        \label{algo:coalescing:virtualProgressing2}
 
-				\State \Comment{virtual task is finished}
-				\State $\text{Data} \rightarrow \text{pTempBuf}$ \Comment{pTempBuf: persistent temp. buffer}
+				\State \Comment{Virtual task is finished}
+				\State $\text{Data} \rightarrow \text{pTempBuf}$ \Comment{pTempBuf: persistent temporary buffer}
 				\State $\text{TVT} = \text{RT}$ 
@@
 \sys's coalescing algorithm uses a single point back in the \sys's execution history to estimate the appropriate virtual task size.
-However, it takes a conservative approach by making the virtual task size equals to the half of the number of the successfully executed real tasks before the last power interrupt (see lines~\ref{algo:coalescing:executionHistory1},\ref{algo:coalescing:executionHistory2} of algorithm~\ref{algo:coalescing} ).
-We advocate the algorithm conservative approach by the following two reasons: (i) The power interrupts are not uniformly distributed over the tasks; and (ii) the re-execution penalty, in case of a power interrupt, grows linearly with the size of a virtual task.
-Furthermore, the algorithm considers the charging rate by allowing the real tasks counter (RTC), which represents the execution history between the last two power interrupts, to increase from one up to multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of algorithm~\ref{algo:coalescing}).
+However, it takes a conservative approach by making the virtual task size equal to the half of the number of successfully executed real tasks before the last power interrupt (see lines~\ref{algo:coalescing:executionHistory1}, \ref{algo:coalescing:executionHistory2} of Algorithm~\ref{algo:coalescing}).
+We advocate for the algorithm conservative approach by the following two reasons: (i) the power interrupts are not uniformly distributed over the tasks; and (ii) the re-execution penalty, in case of a power interrupt, grows linearly with the size of a virtual task.
+Furthermore, the algorithm considers the charging rate by allowing the real tasks counter (RTC), which represents the execution history between the last two power interrupts, to increase from one up to multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of Algorithm~\ref{algo:coalescing}).
 As a consequence, the virtual task size changes rapidly (since it is half of the execution history) in response to the change in the charging rate.
-However, the algorithm sets a limit to the size of the virtual task for these two reasons: (i) having an infinity virtual task size, results in a certain computation progress loss; and (ii) the benefit of task merging is governed by diminishing return principle.
+However, the algorithm sets a limit to the size of the virtual task for these two reasons: (i) having an infinity virtual task size, results in a certain computation progress loss; and (ii) the benefit of task merging is governed by the diminishing return principle (which will be demonstrated experimentally in Section~\ref{sec:results_coalescing}).
 
-The algorithm protects itself from power interrupts by going into three stages: (i) virtual progressing, where the operations are done on the volatile memory (see lines~\ref{algo:coalescing:virtualProgressing1}-\ref{algo:coalescing:virtualProgressing2} of algorithm~\ref{algo:coalescing} ); (ii) transition stage, where the volatile state is moved to a temporary persistent buffer.
-Until the end of the second stage if power interrupts, then all the virtual progress will be canceled and the virtual task will re-executed and re-initialized from a consistent input (see line~\ref{algo:coalescing:virtualProgressing1} of algorithm~\ref{algo:coalescing}); and (iii) persistent commit stage, upon entering this stage the algorithm will make a firm transition (see lines~\ref{algo:coalescing:firmTransition1},\ref{algo:coalescing:firmTransition2} of algorithm~\ref{algo:coalescing}) and it will not go back to the other stages unless all the data is committed from  the persistent buffer to non-volatile memory.
+The algorithm protects itself from power interrupts by going into three stages: (i) virtual progressing, where the operations are done on the volatile memory (see lines~\ref{algo:coalescing:virtualProgressing1}--\ref{algo:coalescing:virtualProgressing2} of Algorithm~\ref{algo:coalescing} ); (ii) transition stage, where the volatile state is moved to a temporary persistent buffer.
+Until the end of the second stage if a power is interrupted, then all the virtual progress will be cancelled and the virtual task will re-executed and re-initialized from a consistent input (see line~\ref{algo:coalescing:virtualProgressing1} of Algorithm~\ref{algo:coalescing}); and (iii) persistent commit stage, upon entering this stage the algorithm will make a firm transition (see lines~\ref{algo:coalescing:firmTransition1}, \ref{algo:coalescing:firmTransition2} of Algorithm~\ref{algo:coalescing}) and it will not go back to the other stages unless all the data is committed from the persistent buffer to non-volatile memory.
 Since the data is being moved in one direction and from a persistent location, power interrupt is tolerable at this stage.

### 2017-08-08 paper/paging.tex

@@
-One of the main objectives of \sys run-time is to restrict tasks to interface only with the volatile memory, since volatile memory access is relatively cheaper, as we already have shown experimentally in Section~\ref{section:background_task_computing} (Problem 1, Figure~\ref{fig:framEnergy}).
-This feature will also provide several advantages during task coalescing, as we will be dealing with in Section \ref{sec:task_coalescing}.
+\sys introduces a dynamic memory virtualization mechanism that allows tasks to
+operate primarily through interaction with efficient, volatile SRAM, rather
+than non-volatile FRAM which, as we showed in Figure~\ref{fig:framEnergy}, is
+more costly to access than SRAM. \sys supports memory virtualization with a
+mechanism that moves pages of the address space from their location in  a
+large, non-volatile memory to a small, volatile memory that tasks use during
+their execution.
 
-\subsection{Intermittent Volatile-Only Memory Access}
-\label{sec:virtualization_problems}
+\subsection{Paged non-volatile virtual memory}
 
-In order to enable volatile-only memory access, the persistent variables a task operates on should be brought from non-volatile memory to a fixed-sized \emph{volatile buffer} in the main memory.
-During its execution, the active task reads and modifies only the contents of this buffer.
-When the task is finished its execution, the modified variables, i.e.
-memory locations, in this buffer should be committed back to their original locations in non-volatile memory.
-We list three fundamental problems proving that this is however a non-trivial task.
-
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=\columnwidth]{figures/sram-buffer}
-	\caption{The interaction between the task, volatile buffer and the non-volatile memory (FRAM).
-Initially, the persistent variable \texttt{x} is not in the volatile buffer but \texttt{y} and \texttt{z} are.}
-	\label{fig:volatile-buffer}
-\end{figure}
-
-\paragraph{Problem 1---Runtime Access Overhead:} The persistent variables read and/or modified by a task might not be known in advance due to the dynamic program flow.
-Therefore, the volatile buffer should keep track of the persistent variables at runtime, as presented in Figure~\ref{fig:volatile-buffer}.
-When a persistent variable is read or written (refer to steps 1 and 3--5 in Figure~\ref{fig:volatile-buffer}) first the volatile copy of the corresponding variable is \emph{searched} in this buffer.
-If it is found, the read or write operation is performed on the volatile copy.
-Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations (refer step 2 in Figure~\ref{fig:volatile-buffer}).
-Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead (refer to Figure~\ref{fig:dmaTimeEnergy} and pointer-based operation therein) and might nullify the benefit of task coalescing, a cores \sys feature.
-
-\begin{figure}[t]
-	\centering
-	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
-	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
-	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
-	\label{fig:dmaTimeEnergy}
-\end{figure}
-
-\paragraph{Problem 2---Two Phase Commit Overhead:} For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each task by means of a \emph{two-phase commit}---refer Figure~\ref{fig:volatile-buffer} steps 6--8 for the first phase and 9--11 for the second step of the commit operation.
-However, two-phase commit is also expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed selectively with CPU intervention.
-Non-selective block memory copy operations using Direct Memory Access (DMA) that eliminates CPU intervention are considerably more efficient, refer again to Figure~\ref{fig:dmaTimeEnergy} for a comparison, but infeasible in this case.
-
-\paragraph{Problem 3---Memory Fit Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
-In order to eliminate run-time buffer \emph{search} overhead (by not relying on specialized hardware such as e.g.
-in~\cite{hicks_isca_2017}), \emph{all} persistent variables can be brought from non-volatile memory to the volatile buffer at system start-up---very efficiently and faster thanks to DMA for \emph{block} memory-memory copy/move operations.
-Two-phase commit is also very efficient in this case since DMA can be utilized to move volatile buffer as a block to the corresponding non-volatile locations.
-However, bringing all persistent variables from non-volatile memory to the volatile buffer suffers from the \emph{memory fit problem}: the volatile memory is a scarce resource and may not be large enough to hold all persistent variables.
-
-\subsection{Software-based Paging Mechanism}
-
-Considering these facts, \sys implements a \emph{paging-based} virtual memory in software that (i) partitions contiguously allocated persistent variables into pages; (ii) implements the volatile \emph{page buffer} that holds a single page; (iii) swaps the page in volatile memory with a new page in non-volatile memory efficiently by utilizing DMA-based memory block-copy operations; (iv) buffers the swapped volatile page until commit to enable task coalescing which we describe in Section~\ref{sec:task_coalescing}; (v) implements two-phase commit of the modified pages to preserve the consistency of non-volatile memory; and (vi) performs this two-phase commit operations using DMA for the sake of efficiency.
-We now explain the details of \sys paging mechanism.
+\sys uses a page-based virtual memory scheme that allows tasks to primarily use
+SRAM instead of frequently accessing FRAM.  The paging scheme partitions the
+non-volatile memory address space into pages of {\em protected data}. As they
+are accessed by a task, pages of protected data are swapped into a {\em page
+buffer} that is stored in volatile memory and holds a single page. When a
+modified page is swapped out, it is buffered in a dedicated commit buffer for
+later commit.  \sys uses two-phase commit to write buffered, modifed pages back
+to their original memory locations atomically, ensuring data consistency even
+with power failures.
+%
+\sys's page-based memory model allows moving large blocks of data between
+non-volatile and volatile memory, which \sys does efficiently using common,
+hardware DMA block-copy operations that have higher bandwidth and avoid fixed,
+per-variable data movement costs.  \sys uses efficient DMA block copies when
+paging data in and out, as well as during two-phase commit.  \sys is unlike
+prior work~\cite{chain,alpaca} that moves one variable at a time, prohibiting
+efficient, DMA bulk transfers. 
+%
+Moreover, \sys's paging and buffering scheme is a key enabler of task
+coalescing, which we describe in Section~\ref{sec:task_coalescing}. Paging
+simplifies coalescing because deferring a commit of paged data at a task
+boundary entails leaving paged data buffered only, not requiring
+additional state management logic at task boundaries.
 
@@
 
-\sys virtual memory system provides an interface to the applications which is composed of two C macros: \texttt{RVAR(var)} and \texttt{WVAR(var,val)} that accept the name of the persistent variable \texttt{var} as an input argument.
-These macros obtain the \emph{physical address} of the corresponding persistent variable in non-volatile memory and translates it into the \emph{virtual address} in page buffer \texttt{pageBuf} in volatile memory, which is composed of a \emph{page tag} and \emph{offset}.
-For the sake of simplicity and efficiency, we implemented this translation in a trivial way: the first byte of the physical address is considered to be the page tag and the rest is the offset within the corresponding page.
+\sys virtualizes the protected, non-volatile data that a task can access
+through a modified memory interface.  The interface includes two operations:
+\texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and
+\texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory
+location {\tt var}.  
+%
+{\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in
+non-volatile memory and translates it into a \emph{virtual address} in the page
+buffer \texttt{pageBuf} in volatile memory.  In \sys, a virtual address is
+composed of a \emph{page tag} and a \emph{page offset}.  The page tag
+identifies in which page of non-volatile memory the memory location resides and
+the page offset identifies a particular byte of that page.  
+%
+How an $a$-bit memory address decomposes into a tag and an offset depends on
+the number (and size) of each page of protected memory.  To address {\tt p}
+pages of size {\tt s}, a tag comprises $log(p)$ bits, the offset comprises
+$log(s)$ bits and $log(p) + log(s) = a$.   For example, in a 16-bit address
+space (i.e., 16kB of protected memory) with 512 pages of 128 bytes each, a
+page's tag is the high-order nine bits of the address and a byte's page offset
+is the remaining 7 bits.
 
@@
 
-After address translation and obtaining the virtual address, it is required to check if the corresponding page is already in volatile page buffer.
-\sys maintains a \texttt{CrntPagTag} variable that holds the tag of the page currently in the page buffer.
-If the tag of the virtual address and \texttt{CrntPagTag} are equal, this indicates that page buffer holds the required page and the offset of the virtual address can be used to read from/write to the location in the volatile page buffer immediately.
-If the required page is not in the page buffer, a \emph{page fault} routine is executed as we discuss in the section below.
-Observe that the computational complexity of the whole steps pertaining to address translation and variable access is $\mathcal{O}(1)$---introducing minimal overhead to the system to boost up benefits of volatile-only memory interfacing.
-The pseudo-code of \texttt{RVAR}, which is quite similar to that of \texttt{WVAR}, is presented in Algorithm~\ref{algo:rwar}.
+After address translation, \sys attempts to access the protected variable's
+location in the volatile paging buffer.  \sys keeps track of the page tag for
+the page currently resident in the paging buffer in {\tt CrntPagTag}, which is
+stored in non-volatile memory. When a task accesses a variable, it compares the
+variable's page tag to {\tt CrntPagTag}.  If an accessed variable's page is
+equal to {\tt CrntPagTag} then the accessed variable's page is resident in the
+page buffer.  If the accessed variable's page tag is not equal to {\tt
+CrntPagTag} then the variable's page is not resident in the page buffer.
+%
+If the accessed variable's page is resident in the page buffer, an executing
+memory operation indexes the page buffer using the variable's page offset and
+manipulates that byte (i.e., either reading or writing).
+%
+If the accessed variable's page is not resident in the page buffer, the
+operation incurs a {\em page fault}. At a page fault, \sys must swap out the
+page that is resident int he page buffer and swap in the the variable's page.
+
+Algorithm~\ref{algo:rwar} shows pseudo-code of \texttt{RVAR} and we omit a
+listing for {\tt WVAR} because it is very similar except it writes the memory
+location and returns nothing.  If the page tag of the variable \texttt{var} and
+\texttt{CrntPagTag} are not equal (e.g. Lines 2--3), there is a page fault
+that requires a new page to be swapped in to the page buffer. 
+%Observe that the computational complexity of the whole steps pertaining to
+%address translation and variable access is $\mathcal{O}(1)$---introducing
+%minimal overhead to the system to boost up benefits of volatile-only memory
+%interfacing.
 
@@
 
-If the page tag of the variable \texttt{var} and \texttt{CrntPagTag} are not equal (e.g.
-Lines 2--3 of Algorithm\,\ref{algo:rwar}), the active page, i.e.
-the contents of the page buffer \texttt{pageBuf}, should be swapped with the page whose tag is equal to the \texttt{var}'s tag.
-The pseudo-code of the Algorithm \texttt{PageFault} that handles this operation is presented in Algorithm~\ref{algo:pagefault}.
-This algorithm handles two fundamental cases.
+A page fault on a memory access requires \sys to swap out whatever page of data
+is in the page buffer, preserving updates made to that page, and to copy the
+accessed page of protected data into the page buffer. 
+%
+Algorithm~\ref{algo:pagefault} shows the how \sys handles a page fault.  If
+both the page being swapped out and the page being swapped in are not {\em
+dirty} --- which means that they do not contain bytes modified by the executing
+task --- then handling a page fault simply requires copying the accessed page
+into the page buffer.  If either the {\em resident} page being swapped out or
+the {\em accessed} page being swapped in is dirty, \sys must take extra steps
+when handling a page fault.
 
-\paragraph{Case 1---Modified Active Page:} If the task performed at least \texttt{WVAR} operation, then the page buffer \texttt{pageBuf} that holds the current active page is modified (Line 1 of Algorithm~\ref{algo:pagefault}).
-In order to keep the modified values of the variables, the active page is committed to an intermediate page buffer in non-volatile memory, namely to \texttt{pagesTemp}.
-This operation can be considered as first phase of the commit operation, as we present in the next section.
-If the active page is not modified, there is no need to commit it to the temporary page buffer.
+\paragraph{Dirty resident page.} If a task modified any byte in the resident
+page (i.e.,  using \texttt{WVAR}), then the page is dirty. Before swapping in
+the accessed page, \sys must first preserve the updated values in the dirty
+page so that they can eventually be committed back to their locations in memory
+when the task ends.  To preserve dirty values, \sys copies the resident page to
+its dedicated memory region called \texttt{pagesTemp}, which is acts as a double buffer
+for each page of protected memory.  Dirty pages copied into {\tt pagesTemp}
+during a task remain there until the task completes, at which point \sys
+commits the modified pages from {\tt pagesTemp} back to memory.
 
-\paragraph{Case 2---Modified Requested Page:} After committing the active page to its temporary location, Algorithm~\ref{algo:pagefault} checks if the requested page is modified previously by the same task (Line 3).
-If this is the case, the requested page should be be brought from \texttt{pagesTemp} rather than its original location in non-volatile memory (Line 4), so that the task accesses the modified values of the persistent variables.
-Otherwise, the requested page is brought from its original location, i.e.
-\texttt{pagesOrg} (Line 6).
-After the swapping operation, the \texttt{CrntPagTag} is set to the tag of the requested page (Line 7).
-It should be noted that DMA is used to memory block copy operations, introducing minimal overhead during page swapping.
+\paragraph{Dirty accessed page.} If an accessed page is modified, \sys should
+swap it into the page buffer from its location in {\tt pagesTemp}, rather than
+swapping it in from its original location in memory. 
+%
+As Algorithm~\ref{algo:pagefault} shows in Line 3, \sys checks if the requested
+page was previously modified by the executing task. To track whether a page is
+modified, \sys stores a dirty bit with the page that is set when a task
+modifies the page.  If an accessed page's dirty bit is set, the page must be
+swapped in from \texttt{pagesTemp} rather than its original memory location
+(Line 4), furnishing the task with the page's modified values. If the accessed
+page is not dirty, \sys swaps it in from its original location in memory (Line
+6). After swapping the page in, \sys sets \texttt{CrntPagTag} to the tag of the
+accessed page (Line 7). \sys uses DMA block copies for all page swapping
+operations, leveraging the efficiency of hardware support.
 
-\subsubsection{Pages Final Committing}
+\subsubsection{Committing a page}
 
-After a task finishes its execution, all pages committed to \texttt{pagesTemp} should be committed to their original locations in \texttt{pagesOrg}.
-This is the second phase of the \emph{two-phase commit} operation.
-As mentioned previously, memory block copy operations for the second phase commit is also utilized by DMA.
+When a task completes, it must {\em commit} all (dirty) pages stored in
+\texttt{pagesTemp} back to their original memory locations. \sys maintains a
+global, non-volatile {\em commit bit} that indicates that a task is committing.  The
+commit bit is set before the first page is committed and the bit is unset after the
+last page is committed.  If power fails during commit, the commit bit remains set and
+\sys resumes committing before moving on to execute the next task.  To commit
+pages from {\tt pagesTemp}, \sys iterates over the pages' dirty bits and for
+each page with its dirty bit set, \sys DMA block-copies the page back to its
+original memory location, then clears its dirty bit.
 
-\subsection{Non-Volatile Memory Consistency Issues}
+\subsection{Paging Example}
+\todo{Redo this figure to actually reflect paging -- right now it isn't clear if x,y,z are in a page or not, it isn't clear when the page gets copied to pagesTemp, it isn't clear what state we add to the system to help with commit, etc (i.e., dirty bits, commit bit).
+Make
+this figure's labels consistent with the text: pagesTemp, RVAR, WVAR, explicit
+commit in task, Page Buffer}{whoever} 
+Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task
+using \sys's paging mechanism.  The task manipulates protected variables
+{\tt x}, {\tt y}, and {\tt z}, which are in the same page.  On accessing {\tt
+x}, its page is swapped into the page buffer.  On accessing {\tt y} and {\tt
+z}, the task need only access the volatile page buffer.  The task updates the
+page, leaving it dirty at the task's completion.  When the task completes, it
+first copies the page from the page buffer into the non-volatile {\tt
+pagesTemp} region before commit.  Then, \sys sets the commit bit, 
+atomically commits the page back to its original memory location, and unsets
+the commit bit.
 
-During the execution of any task, modifications to the persistent variables will not be reflected to the original locations in the non-volatile memory, since the task interfaces with only the volatile page buffer \texttt{pageBuf} and the modifications are temporarily stored in the non-volatile \texttt{pagesTemp} buffer.
-The changes stored in the \texttt{pagesTmp} is committed to the original locations in \texttt{pagesOrg} after the task is finished its execution.
-Therefore, any power interruptions within the execution will not lead to inconsistencies in the \texttt{pagesOrg} when the task is re-executed after a power boot.
-As a consequence, the proposed paging mechanism keeps the consistency of the non-volatile memory thanks to the two-phase commit operation for the modified pages.
+\begin{figure}[t]
+	\centering
+	\includegraphics[width=\columnwidth]{figures/sram-buffer}
+	\caption{The interaction between the task, volatile buffer and the non-volatile memory (FRAM).
+Initially, the persistent variable \texttt{x} is not in the volatile buffer but \texttt{y} and \texttt{z} are.}
+	\label{fig:volatile-buffer}
+\end{figure}
+
+\subsection{Memory Consistency}
+
+\sys's paging mechanism ensures that a task only ever executes using consistent
+protected data.  During task execution, modifications to protected data do not
+affect their original memory locations because a task reads and writes the
+volatile page buffer only, and modified pages are swapped out to
+\texttt{pagesTemp} until commit.  A power failure erases the contents of the
+page buffer, preventing a re-executing task from observing updates in the page
+buffer from a previous execution attempt.  On a reboot after a power failure,
+\sys clears all dirty bits in {\tt pagesTemp} {\em if the commit bit is unset}.
+Clearing these dirty bits ensures that all accesses to protected variables
+correctly access their original memory locations. The non-volatile commit bit
+persists across power failures and \sys repeatedly attempts to commit until it
+succeeds, despite arbitrary power failures.
+
+%One of the main objectives of \sys run-time is to restrict tasks to interface
+%only with the volatile memory, since volatile memory access is relatively
+%cheaper, as we already have shown experimentally in
+%Section~\ref{section:background_task_computing} (Problem 1,
+%Figure~\ref{fig:framEnergy}). This feature will also provide several advantages
+%during task coalescing, as we will be dealing with in Section
+%\ref{sec:task_coalescing}.
+
+%\subsection{Intermittent Volatile-Only Memory Access}
+%\label{sec:virtualization_problems}
+%
+%In order to enable volatile-only memory access, the persistent variables a task
+%operates on should be brought from non-volatile memory to a fixed-sized
+%\emph{volatile buffer} in the main memory. During its execution, the active
+%task reads and modifies only the contents of this buffer. When the task is
+%finished its execution, the modified variables, i.e. memory locations, in this
+%buffer should be committed back to their original locations in non-volatile
+%memory. We list three fundamental problems proving that this is however a
+%non-trivial task. 
+%
+%
+\subsection{Managing paging overhead}
+The efficiency of \sys's paging scheme relies on the fact that data in pages
+are contiguous, which makes them amenable to being manipulated by hardware
+assisted DMA bulk copy operations. To justify this design choice we
+experimentally evaluated the efficiency of using DMA block copies, compared to
+explicit, software copy loops.  We copied blocks of data of different size
+between memory regions on a WISP~\cite{wisp} using a hand-written loop and
+using DMA.  Figure~\ref{fig:dmaTimeEnergy} shows that for moving large blocks
+of data like a page, hardware DMA is considerably better than software in both
+run time and energy efficiency.  \sys's design moves data in pages, not
+individual words, leveraging the increased efficiency of DMA data movement.
+%\paragraph{Problem 1---Runtime Access Overhead:} The persistent variables read and/or modified by a task might not be known in advance due to the dynamic program flow.
+Therefore, the volatile buffer should keep track of the persistent variables at runtime, as presented in Figure~\ref{fig:volatile-buffer}.
+When a persistent variable is read or written (refer to steps 1 and 3--5 in Figure~\ref{fig:volatile-buffer}) first the volatile copy of the corresponding variable is \emph{searched} in this buffer.
+If it is found, the read or write operation is performed on the volatile copy.
+Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations (refer step 2 in Figure~\ref{fig:volatile-buffer}).
+Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead (refer to Figure~\ref{fig:dmaTimeEnergy} and pointer-based operation therein) and might nullify the benefit of task coalescing, a cores \sys feature.
+%
+\begin{figure}[t]
+	\centering
+	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
+	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
+	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
+	\label{fig:dmaTimeEnergy}
+\end{figure}
+%
+%\paragraph{Problem 2---Two Phase Commit Overhead:} For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each task by means of a \emph{two-phase commit}---refer Figure~\ref{fig:volatile-buffer} steps 6--8 for the first phase and 9--11 for the second step of the commit operation.
+However, two-phase commit is also expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed selectively with CPU intervention.
+Non-selective block memory copy operations using Direct Memory Access (DMA) that eliminates CPU intervention are considerably more efficient, refer again to Figure~\ref{fig:dmaTimeEnergy} for a comparison, but infeasible in this case.
+%
+%\paragraph{Problem 3---Memory Fit Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
+In order to eliminate run-time buffer \emph{search} overhead (by not relying on specialized hardware such as e.g.
+in~\cite{hicks_isca_2017}), \emph{all} persistent variables can be brought from non-volatile memory to the volatile buffer at system start-up---very efficiently and faster thanks to DMA for \emph{block} memory-memory copy/move operations.
+Two-phase commit is also very efficient in this case since DMA can be utilized to move volatile buffer as a block to the corresponding non-volatile locations.
+However, bringing all persistent variables from non-volatile memory to the volatile buffer suffers from the \emph{memory fit problem}: the volatile memory is a scarce resource and may not be large enough to hold all persistent variables.
 

### 2017-08-08 paper/paging.tex
Small corrections (typos/capitals) to paging

@@
-\sys introduces a dynamic memory virtualization mechanism that allows tasks to
-operate primarily through interaction with efficient, volatile SRAM, rather
-than non-volatile FRAM which, as we showed in Figure~\ref{fig:framEnergy}, is
-more costly to access than SRAM. \sys supports memory virtualization with a
-mechanism that moves pages of the address space from their location in  a
-large, non-volatile memory to a small, volatile memory that tasks use during
-their execution.
+\sys introduces a dynamic memory virtualization mechanism that allows tasks to operate primarily through interaction with efficient, volatile SRAM, rather than non-volatile FRAM which, as we showed in Figure~\ref{fig:framEnergy}, is more costly to access than SRAM.
+\sys supports memory virtualization with a mechanism that moves pages of the address space from their location in a large, non-volatile memory to a small, volatile memory that tasks use during their execution.
 
-\subsection{Paged non-volatile virtual memory}
+\subsection{Paged Non-Volatile Virtual Memory}
 
-\sys uses a page-based virtual memory scheme that allows tasks to primarily use
-SRAM instead of frequently accessing FRAM.  The paging scheme partitions the
-non-volatile memory address space into pages of {\em protected data}. As they
-are accessed by a task, pages of protected data are swapped into a {\em page
-buffer} that is stored in volatile memory and holds a single page. When a
-modified page is swapped out, it is buffered in a dedicated commit buffer for
-later commit.  \sys uses two-phase commit to write buffered, modifed pages back
-to their original memory locations atomically, ensuring data consistency even
-with power failures.
-%
-\sys's page-based memory model allows moving large blocks of data between
-non-volatile and volatile memory, which \sys does efficiently using common,
+\sys uses a page-based virtual memory scheme that allows tasks to primarily use SRAM instead of frequently accessing FRAM.
+The paging scheme partitions the non-volatile memory address space into pages of {\em protected data}.
+As they are accessed by a task, pages of protected data are swapped into a {\em page buffer} that is stored in volatile memory and holds a single page.
+When a modified page is swapped out, it is buffered in a dedicated commit buffer for later commit.
+\sys uses two-phase commit to write buffered, modified pages back to their original memory locations atomically, ensuring data consistency even with power failures.
+\sys's page-based memory model allows moving large blocks of data between non-volatile and volatile memory, which \sys does efficiently using common,
 hardware DMA block-copy operations that have higher bandwidth and avoid fixed, 
-per-variable data movement costs.  \sys uses efficient DMA block copies when
-paging data in and out, as well as during two-phase commit.  \sys is unlike
-prior work~\cite{chain,alpaca} that moves one variable at a time, prohibiting
-efficient, DMA bulk transfers. 
-%
-Moreover, \sys's paging and buffering scheme is a key enabler of task
-coalescing, which we describe in Section~\ref{sec:task_coalescing}. Paging
-simplifies coalescing because deferring a commit of paged data at a task
-boundary entails leaving paged data buffered only, not requiring
-additional state management logic at task boundaries.
+per-variable data movement costs.
+\sys uses efficient DMA block copies when paging data in and out, as well as during two-phase commit.
+\sys is unlike prior work~\cite{chain,alpaca} that moves one variable at a time, prohibiting efficient, DMA bulk transfers.
+Moreover, \sys's paging and buffering scheme is a key enabler of task coalescing, which we describe in Section~\ref{sec:task_coalescing}.
+Paging simplifies coalescing because deferring a commit of paged data at a task boundary entails leaving paged data buffered only, not requiring additional state management logic at task boundaries.
 
@@
 
-\sys virtualizes the protected, non-volatile data that a task can access
-through a modified memory interface.  The interface includes two operations:
-\texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and
-\texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory
-location {\tt var}.  
-%
-{\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in
-non-volatile memory and translates it into a \emph{virtual address} in the page
-buffer \texttt{pageBuf} in volatile memory.  In \sys, a virtual address is
-composed of a \emph{page tag} and a \emph{page offset}.  The page tag
-identifies in which page of non-volatile memory the memory location resides and
+\sys virtualizes the protected, non-volatile data that a task can access through a modified memory interface.
+The interface includes two operations: \texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and \texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory location {\tt var}.
+{\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory and translates it into a \emph{virtual address} in the page buffer \texttt{pageBuf} in volatile memory.
+In \sys, a virtual address is composed of a \emph{page tag} and a \emph{page offset}.
+The page tag identifies in which page of non-volatile memory the memory location resides and
 the page offset identifies a particular byte of that page.
-%
-How an $a$-bit memory address decomposes into a tag and an offset depends on
-the number (and size) of each page of protected memory.  To address {\tt p}
-pages of size {\tt s}, a tag comprises $log(p)$ bits, the offset comprises
-$log(s)$ bits and $log(p) + log(s) = a$.   For example, in a 16-bit address
-space (i.e., 16kB of protected memory) with 512 pages of 128 bytes each, a
+How an $a$-bit memory address decomposes into a tag and an offset depends on the number (and size) of each page of protected memory.
+To address {\tt p} pages of size {\tt s}, a tag comprises $log(p)$ bits, the offset comprises $log(s)$ bits and $log(p) + log(s) = a$.
+For example, in a 16-bit address space (i.e., 16\,kB of protected memory) with 512 pages of 128 bytes each, a
 page's tag is the high-order nine bits of the address and a byte's page offset
@@
 
-After address translation, \sys attempts to access the protected variable's
-location in the volatile paging buffer.  \sys keeps track of the page tag for
-the page currently resident in the paging buffer in {\tt CrntPagTag}, which is
-stored in non-volatile memory. When a task accesses a variable, it compares the
-variable's page tag to {\tt CrntPagTag}.  If an accessed variable's page is
-equal to {\tt CrntPagTag} then the accessed variable's page is resident in the
-page buffer.  If the accessed variable's page tag is not equal to {\tt
-CrntPagTag} then the variable's page is not resident in the page buffer.
-%
-If the accessed variable's page is resident in the page buffer, an executing
-memory operation indexes the page buffer using the variable's page offset and
-manipulates that byte (i.e., either reading or writing).
-%
-If the accessed variable's page is not resident in the page buffer, the
-operation incurs a {\em page fault}. At a page fault, \sys must swap out the
+After address translation, \sys attempts to access the protected variable's location in the volatile paging buffer.
+\sys keeps track of the page tag for the page currently resident in the paging buffer in {\tt CrntPagTag}, which is stored in non-volatile memory.
+When a task accesses a variable, it compares the variable's page tag to {\tt CrntPagTag}.
+If an accessed variable's page is equal to {\tt CrntPagTag} then the accessed variable's page is resident in the page buffer.
+If the accessed variable's page tag is not equal to {\tt CrntPagTag} then the variable's page is not resident in the page buffer.
+If the accessed variable's page is resident in the page buffer, an executing memory operation indexes the page buffer using the variable's page offset and manipulates that byte (i.e., either reading or writing).
+If the accessed variable's page is not resident in the page buffer, the operation incurs a {\em page fault}.
+At a page fault, \sys must swap out the
 page that is resident int he page buffer and swap in the the variable's page.
 
-Algorithm~\ref{algo:rwar} shows pseudo-code of \texttt{RVAR} and we omit a
-listing for {\tt WVAR} because it is very similar except it writes the memory
-location and returns nothing.  If the page tag of the variable \texttt{var} and
-\texttt{CrntPagTag} are not equal (e.g. Lines 2--3), there is a page fault
-that requires a new page to be swapped in to the page buffer. 
-%Observe that the computational complexity of the whole steps pertaining to
-%address translation and variable access is $\mathcal{O}(1)$---introducing
-%minimal overhead to the system to boost up benefits of volatile-only memory
-%interfacing.
+Algorithm~\ref{algo:rwar} shows pseudo-code of \texttt{RVAR} and we omit a listing for {\tt WVAR} as it is very similar except it writes the memory location and returns nothing.
+If the page tag of the variable \texttt{var} and \texttt{CrntPagTag} are not equal (e.g.
+Lines 2--3), there is a page fault that requires a new page to be swapped in to the page buffer.
 
@@
 
-A page fault on a memory access requires \sys to swap out whatever page of data
-is in the page buffer, preserving updates made to that page, and to copy the
-accessed page of protected data into the page buffer. 
-%
-Algorithm~\ref{algo:pagefault} shows the how \sys handles a page fault.  If
-both the page being swapped out and the page being swapped in are not {\em
-dirty} --- which means that they do not contain bytes modified by the executing
-task --- then handling a page fault simply requires copying the accessed page
-into the page buffer.  If either the {\em resident} page being swapped out or
-the {\em accessed} page being swapped in is dirty, \sys must take extra steps
-when handling a page fault.
+A page fault on a memory access requires \sys to swap out whatever page of data is in the page buffer, preserving updates made to that page, and to copy the accessed page of protected data into the page buffer.
+Algorithm~\ref{algo:pagefault} shows the how \sys handles a page fault.
+If both the page being swapped out and the page being swapped in are not {\em dirty}---which means that they do not contain bytes modified by the executing
+task---then handling a page fault simply requires copying the accessed page into the page buffer.
+If either the {\em resident} page being swapped out or the {\em accessed} page being swapped in is dirty, \sys must take extra steps when handling a page fault.
 
-\paragraph{Dirty resident page.} If a task modified any byte in the resident
-page (i.e.,  using \texttt{WVAR}), then the page is dirty. Before swapping in
-the accessed page, \sys must first preserve the updated values in the dirty
-page so that they can eventually be committed back to their locations in memory
-when the task ends.  To preserve dirty values, \sys copies the resident page to
-its dedicated memory region called \texttt{pagesTemp}, which is acts as a double buffer
-for each page of protected memory.  Dirty pages copied into {\tt pagesTemp}
-during a task remain there until the task completes, at which point \sys
-commits the modified pages from {\tt pagesTemp} back to memory.
+\paragraph{Dirty resident page.} If a task modified any byte in the resident page (i.e.,  using \texttt{WVAR}), then the page is dirty.
+Before swapping in the accessed page, \sys must first preserve the updated values in the dirty page so that they can eventually be committed back to their locations in memory when the task ends.
+To preserve dirty values, \sys copies the resident page to its dedicated memory region called \texttt{pagesTemp}, which is acts as a double buffer for each page of protected memory.
+Dirty pages copied into {\tt pagesTemp} during a task remain there until the task completes, at which point \sys commits the modified pages from {\tt pagesTemp} back to memory.
 
-\paragraph{Dirty accessed page.} If an accessed page is modified, \sys should
-swap it into the page buffer from its location in {\tt pagesTemp}, rather than
-swapping it in from its original location in memory. 
-%
-As Algorithm~\ref{algo:pagefault} shows in Line 3, \sys checks if the requested
-page was previously modified by the executing task. To track whether a page is
-modified, \sys stores a dirty bit with the page that is set when a task
-modifies the page.  If an accessed page's dirty bit is set, the page must be
-swapped in from \texttt{pagesTemp} rather than its original memory location
-(Line 4), furnishing the task with the page's modified values. If the accessed
-page is not dirty, \sys swaps it in from its original location in memory (Line
-6). After swapping the page in, \sys sets \texttt{CrntPagTag} to the tag of the
-accessed page (Line 7). \sys uses DMA block copies for all page swapping
-operations, leveraging the efficiency of hardware support.
+\paragraph{Dirty accessed page.} If an accessed page is modified, \sys should swap it into the page buffer from its location in {\tt pagesTemp}, rather than swapping it in from its original location in memory.
+As Algorithm~\ref{algo:pagefault} shows in Line 3, \sys checks if the requested page was previously modified by the executing task.
+To track whether a page is modified, \sys stores a dirty bit with the page that is set when a task
+modifies the page.
+If an accessed page's dirty bit is set, the page must be swapped in from \texttt{pagesTemp} rather than its original memory location (Line 4), furnishing the task with the page's modified values.
+If the accessed page is not dirty, \sys swaps it in from its original location in memory (Line 6).
+After swapping the page in, \sys sets \texttt{CrntPagTag} to the tag of the accessed page (Line 7).
+\sys uses DMA block copies for all page swapping operations, leveraging the efficiency of hardware support.
 
-\subsubsection{Committing a page}
+\subsubsection{Committing a Page}
 
-When a task completes, it must {\em commit} all (dirty) pages stored in
-\texttt{pagesTemp} back to their original memory locations. \sys maintains a
-global, non-volatile {\em commit bit} that indicates that a task is committing.  The
-commit bit is set before the first page is committed and the bit is unset after the
-last page is committed.  If power fails during commit, the commit bit remains set and
-\sys resumes committing before moving on to execute the next task.  To commit
-pages from {\tt pagesTemp}, \sys iterates over the pages' dirty bits and for
-each page with its dirty bit set, \sys DMA block-copies the page back to its
+When a task completes, it must {\em commit} all (dirty) pages stored in \texttt{pagesTemp} back to their original memory locations.
+\sys maintains a global, non-volatile {\em commit bit} that indicates that a task is committing.
+The commit bit is set before the first page is committed and the bit is unset after the last page is committed.
+If power fails during commit, the commit bit remains set and \sys resumes committing before moving on to execute the next task.
+To commit pages from {\tt pagesTemp}, \sys iterates over the pages' dirty bits and for each page with its dirty bit set, \sys DMA block-copies the page back to its
 original memory location, then clears its dirty bit.
@@
 \subsection{Paging Example}
+
 \todo{Redo this figure to actually reflect paging -- right now it isn't clear if x,y,z are in a page or not, it isn't clear when the page gets copied to pagesTemp, it isn't clear what state we add to the system to help with commit, etc (i.e., dirty bits, commit bit).
-Make
-this figure's labels consistent with the text: pagesTemp, RVAR, WVAR, explicit
+Make this figure's labels consistent with the text: pagesTemp, RVAR, WVAR, explicit
 commit in task, Page Buffer}{whoever}
-Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task
-using \sys's paging mechanism.  The task manipulates protected variables
-{\tt x}, {\tt y}, and {\tt z}, which are in the same page.  On accessing {\tt
-x}, its page is swapped into the page buffer.  On accessing {\tt y} and {\tt
-z}, the task need only access the volatile page buffer.  The task updates the
-page, leaving it dirty at the task's completion.  When the task completes, it
-first copies the page from the page buffer into the non-volatile {\tt
-pagesTemp} region before commit.  Then, \sys sets the commit bit, 
-atomically commits the page back to its original memory location, and unsets
-the commit bit.
+
+Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task using \sys's paging mechanism.
+The task manipulates protected variables {\tt x}, {\tt y}, and {\tt z}, which are in the same page.
+On accessing {\tt x}, its page is swapped into the page buffer.
+On accessing {\tt y} and {\tt z}, the task need only access the volatile page buffer.
+The task updates the page, leaving it dirty at the task's completion.
+When the task completes, it first copies the page from the page buffer into the non-volatile {\tt
+pagesTemp} region before commit. Then, \sys sets the commit bit, atomically commits the page back to its original memory location, and unsets the commit bit.
 
@@
 
-\sys's paging mechanism ensures that a task only ever executes using consistent
-protected data.  During task execution, modifications to protected data do not
-affect their original memory locations because a task reads and writes the
-volatile page buffer only, and modified pages are swapped out to
-\texttt{pagesTemp} until commit.  A power failure erases the contents of the
-page buffer, preventing a re-executing task from observing updates in the page
-buffer from a previous execution attempt.  On a reboot after a power failure,
+\sys's paging mechanism ensures that a task only ever executes using consistent protected data.
+During task execution, modifications to protected data do not affect their original memory locations because a task reads and writes the volatile page buffer only, and modified pages are swapped out to \texttt{pagesTemp} until commit.
+A power failure erases the contents of the page buffer, preventing a re-executing task from observing updates in the page buffer from a previous execution attempt.
+On a reboot after a power failure,
 \sys clears all dirty bits in {\tt pagesTemp} {\em if the commit bit is unset}.
-Clearing these dirty bits ensures that all accesses to protected variables
-correctly access their original memory locations. The non-volatile commit bit
-persists across power failures and \sys repeatedly attempts to commit until it
-succeeds, despite arbitrary power failures.
+Clearing these dirty bits ensures that all accesses to protected variables correctly access their original memory locations.
+The non-volatile commit bit persists across power failures and \sys repeatedly attempts to commit until it succeeds, despite arbitrary power failures.
 
-%One of the main objectives of \sys run-time is to restrict tasks to interface
-%only with the volatile memory, since volatile memory access is relatively
-%cheaper, as we already have shown experimentally in
-%Section~\ref{section:background_task_computing} (Problem 1,
-%Figure~\ref{fig:framEnergy}). This feature will also provide several advantages
-%during task coalescing, as we will be dealing with in Section
-%\ref{sec:task_coalescing}.
+\subsection{Managing Paging Overhead}
 
-%\subsection{Intermittent Volatile-Only Memory Access}
-%\label{sec:virtualization_problems}
-%
-%In order to enable volatile-only memory access, the persistent variables a task
-%operates on should be brought from non-volatile memory to a fixed-sized
-%\emph{volatile buffer} in the main memory. During its execution, the active
-%task reads and modifies only the contents of this buffer. When the task is
-%finished its execution, the modified variables, i.e. memory locations, in this
-%buffer should be committed back to their original locations in non-volatile
-%memory. We list three fundamental problems proving that this is however a
-%non-trivial task. 
-%
-%
-\subsection{Managing paging overhead}
-The efficiency of \sys's paging scheme relies on the fact that data in pages
-are contiguous, which makes them amenable to being manipulated by hardware
-assisted DMA bulk copy operations. To justify this design choice we
-experimentally evaluated the efficiency of using DMA block copies, compared to
-explicit, software copy loops.  We copied blocks of data of different size
-between memory regions on a WISP~\cite{wisp} using a hand-written loop and
-using DMA.  Figure~\ref{fig:dmaTimeEnergy} shows that for moving large blocks
-of data like a page, hardware DMA is considerably better than software in both
-run time and energy efficiency.  \sys's design moves data in pages, not
-individual words, leveraging the increased efficiency of DMA data movement.
-%\paragraph{Problem 1---Runtime Access Overhead:} The persistent variables read and/or modified by a task might not be known in advance due to the dynamic program flow.
-Therefore, the volatile buffer should keep track of the persistent variables at runtime, as presented in Figure~\ref{fig:volatile-buffer}.
-When a persistent variable is read or written (refer to steps 1 and 3--5 in Figure~\ref{fig:volatile-buffer}) first the volatile copy of the corresponding variable is \emph{searched} in this buffer.
-If it is found, the read or write operation is performed on the volatile copy.
-Otherwise, the persistent variable is read from the non-volatile memory and inserted in this buffer for future operations (refer step 2 in Figure~\ref{fig:volatile-buffer}).
-Unfortunately, we observed that \emph{search--bring} operations at each access to the persistent variables introduce considerable overhead (refer to Figure~\ref{fig:dmaTimeEnergy} and pointer-based operation therein) and might nullify the benefit of task coalescing, a cores \sys feature.
-%
+The efficiency of \sys's paging scheme relies on the fact that data in pages are contiguous, which makes them amenable to being manipulated by hardware assisted DMA bulk copy operations.
+To justify this design choice we experimentally evaluated the efficiency of using DMA block copies, compared to explicit, software copy loops.
+We copied blocks of data of different size between memory regions on a WISP~\cite{wisp} using a hand-written loop and using DMA (refer to Section~\ref{sec:results_hardware} for hardware setup details).
+Figure~\ref{fig:dmaTimeEnergy} shows that for moving large blocks of data like a page, hardware DMA is considerably better than software in both run time and energy efficiency.
+\sys's design moves data in pages, not individual words, leveraging the increased efficiency of DMA data movement.
+
 \begin{figure}[t]
@@
 \end{figure}
-%
-%\paragraph{Problem 2---Two Phase Commit Overhead:} For the consistency of the non-volatile memory, the volatile buffer should be committed at the end of each task by means of a \emph{two-phase commit}---refer Figure~\ref{fig:volatile-buffer} steps 6--8 for the first phase and 9--11 for the second step of the commit operation.
-However, two-phase commit is also expensive since the committed variables are not \emph{contiguous}---leading the fact that copying or moving of data between/within volatile and non-volatile memory should be always performed selectively with CPU intervention.
-Non-selective block memory copy operations using Direct Memory Access (DMA) that eliminates CPU intervention are considerably more efficient, refer again to Figure~\ref{fig:dmaTimeEnergy} for a comparison, but infeasible in this case.
-%
-%\paragraph{Problem 3---Memory Fit Constraints:} Persistent variables can be allocated \emph{contiguously} in non-volatile memory with compiler support.
-In order to eliminate run-time buffer \emph{search} overhead (by not relying on specialized hardware such as e.g.
-in~\cite{hicks_isca_2017}), \emph{all} persistent variables can be brought from non-volatile memory to the volatile buffer at system start-up---very efficiently and faster thanks to DMA for \emph{block} memory-memory copy/move operations.
-Two-phase commit is also very efficient in this case since DMA can be utilized to move volatile buffer as a block to the corresponding non-volatile locations.
-However, bringing all persistent variables from non-volatile memory to the volatile buffer suffers from the \emph{memory fit problem}: the volatile memory is a scarce resource and may not be large enough to hold all persistent variables.
-
-%\todo{Benefits of the paging to the task coalescing should be mentioned in the next section}{Amjad/Przemek/Sinan}

### 2017-08-08 paper/background.tex
reference correction and removed word ophran

@@
 
-As the figure illustrates, checkpoint-based systems risk violating the consistency of volatile and non-volatile memory~\cite{dino}.
-To maintain memory consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of non-volatile variables with the checkpointed execution context.
-Task-based systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail next, ensure state is consistent by using programmer, compiler, and runtime support.
+As the figure illustrates, checkpoint-based systems risk violating the volatile and non-volatile memory consistency~\cite{dino}.
+To maintain consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of non-volatile variables with the checkpointed execution context.
+Task-based systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail next, ensure state consistency by using programmer, compiler, and runtime support.
 
@@
 Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent (i.e., have the same effect when repeatedly executed).
-The key to idempotence is ensuring that updates to non-volatile state made by an interrupted, partial execution of a task are never visible to a subsequent (potentially completing)
-re-execution of the task.
+The key to idempotence is ensuring that updates to non-volatile state made by an interrupted, partial execution of a task are never visible to a subsequent (potentially completing) re-execution of the task.
 
 There are several run time strategies to ensuring task idempotence.
-One approach, utilized by Alpaca~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{rachet} did for checkpoints), execute the task using buffered copies of those data, and commit the buffered copies before completing the task.
+One approach, utilized by Alpaca~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{ratchet} did for checkpoints), execute the task using buffered copies of those data, and commit the buffered copies before completing the task.
 Another strategy is to statically create multiple versions of non-volatile data shared by tasks and ensure by construction that no task reads and writes the same non-volatile memory location~\cite{chain}.

### 2017-08-08 paper/relatedwork.tex
placed writing tasks in conclusions and related work
Related Work-subsection 1 ready

@@
-\section{Intermittenly-Powered Devices}
+\noindent \textbf{Intermittently-Powered Devices:} There is plentiful of research results available on energy harvesting, or more specifically wireless power provision, for embedded devices like sensors or microcontrollers.
+This paragraph is too short to cover this area extensively and we refer the reader to overview papers~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016} for in-depth discussion.
+With power consumption of electronics going down dramatically with every year, computing systems can be now powered completely from ambient energy.
+This opens new research areas such as radio frequency-powered computers~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}.
+Indeed, plenty of new embedded platforms appear that are powered solely from ambient (intentional or non-intentional) energy and include~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
+But there is a price tag associated to them: harvested energy is miniscule, unstable, difficult to predict ergo \emph{intermittent}.
+Therefore, to guarantee completion of computing, programming and runtime models are being actively researched that are specifically tailored to these intermittently-powered systems.
+\todo{Revise this section}{Brandon}
 
-\section{Intermittent Execution Environments}
+\noindent \textbf{Intermittent Execution Environments:} Review of intermittent runtime environments. \todo{Fill-in this section}{Przemek}
+
+\noindent \textbf{Memory Virtualization:} Review of memory virtualization environments (not necessarily for intermittent power devices). \todo{Fill-in this section}{Przemek}

### 2017-08-08 paper/relatedwork.tex
Related Work-subsection 2 placeholder

@@
 
-\noindent \textbf{Intermittent Execution Environments:} Review of intermittent runtime environments. \todo{Fill-in this section}{Przemek}
+\noindent \textbf{Intermittent Execution Environments:} Checkpoint runtimes include DINO~\cite{dino}, Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus}.
+The only available Task-based intermittent runtime systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}.
+\sys goes beyond what any task-based system performed by providing \emph{automatic} task creation and \emph{adaptivity} to energy state and energy storage \emph{obliviousness} of the compiled code.
+\todo{complete this section}{Przemek}
 

### 2017-08-08 paper/relatedwork.tex
Related Work-subsection 2 placeholder

@@
 
-\noindent \textbf{Intermittent Execution Environments:} Checkpoint runtimes include DINO~\cite{dino}, Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus}.
+\noindent \textbf{Intermittent Execution Environments:} Earlier work include Dewdrop~\cite{dewdrop} (splitting task such that they fill-in the capacitor).
+Similar, specific-application tailored system includes Wisent and Stork~\cite{stork,wisent}.
+Checkpoint runtimes include DINO~\cite{dino}, Quickrecall~\cite{quickrecall}, Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus} are checkpointing-based systems requiring hardware support to measure remaining storage energy---a feature prohibitive given typical high energy cost expenditure of such hardware.
 The only available Task-based intermittent runtime systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}.

### 2017-08-09 paper/evaluation.tex
compiler eval added

@@
 
-We start with evaluating the efficiency of a \sys compiler.
+\begin{table}[t]
+	\centering
+	\footnotesize
+	\begin{tabular}{|c|c|c|c|c|c|c|c|}
+		\hline
+		{} & \multicolumn{2}{l|}{{PVAR Size (Byte)}} & \multicolumn{2}{l|}{{\# Tasks}} & \multicolumn{2}{l|}{{\# Paging Code}} \\
+		\hline
+		Nickname & Hand & Comp. & Hand & Comp. & Hand & Comp.\\
+		\hline\hline
+		Bitcount & 22 & 22 & 10 & 15 & 81 & 93 \\
+		Cold chain & 3492 & 3242 & 12 & 9 & 92 & 123\\
+		Cuckoo & 282 & 288 & 15 & 6 & 90 & 76\\
+		RSA & 332 & 250 & 20 & 27 & 130 & 296\\
+		AR & 166 & 218 & 11 & 6 & 112 & 333\\
+		Sort & 104 & 104 & 4 & 2 & 70 & 23\\
+		\hline
+	\end{tabular}
+	\caption{Comparison between compiler-generated code (Comp.) and hand-written \sys code (Hand).}
+\label{table:compiler_result}
+\end{table}
 
-Expected results:
-\begin{itemize}
-	\item comparison between manual (Chain code) and automatic taskification: in terms of code SLOCs, number of tasks, completion time (Chain), memory use (other metrics required?), show number of LLVM instructions per energy cycle
-	\item assessment of correctness (through manual inspection only?).
-\end{itemize}
-
-\todo{Provide results and text to this section}{Kiwan}
+We start with evaluating the efficiency of a \sys compiler. Table~\ref{table:compiler_result} shows
+the difference in size of protected variable (PVAR), number of tasks, and number of inserted paging call in the code between hand-written \sys code
+and compiler-generated \sys code. All three values are better to be small. On average, compiler-generated
+code is only worse by 0.3\% in the size of protected variable. Regarding the number of tasks, compiler-generated code
+is better by roughly 16\%. The result indicate compiler is selecting boundaries better than programmer manually doing it.
+However, compiler should insert paging code conservatively when it {\em may} access protected variable, due to limitation of
+pointer alias. Thus, it results in compiler introducing 48\% more paging code than programmer-written code. The resulting run time
+difference is summarized in Figure~\ref{}. The performance is worse than programmer-written \sys code
+in overall, meaning the programmer should try to write code manually if performance is critical; otherwise compiler can be
+used to simplify the programmability.
 

### 2017-08-09 paper/evaluation.tex
added compile time

@@
 used to simplify the programmability.
+
+We also evaluated the compile time for automatic task generation. The result are in Table~\ref{table:compile_time}.
+Most of the compile time was due to pointer aliasing. We could not compile DFT application since it was using
+builtins for math functions that weren't provided by Clang MSP430 backend.
+
+\begin{table}[t]
+	\centering
+	\footnotesize
+	\begin{tabular}{|c|c|c|c|c|c|}
+		\hline
+		Bitcount & Cold chain & Cuckoo & RSA & AR & Sort \\
+		\hline\hline
+		3 & 2 & 6 & 86 & 34 & > 1 \\
+		\hline
+	\end{tabular}
+	\caption{Compile time in seconds.}
+\label{table:compile_time}
+\end{table}
 

### 2017-08-09 paper/evaluation.tex
added environment

@@
 build around MSP430FR5969~\cite{wolverine} 64\,kB FRAM microcontroller, and TI's MSP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} (hardware type will be specified per experiment in Section~\ref{sec:results_evaluation}).
-WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet} with TI Code Composer Studio version X~\cite{}\todo{Report on CCS version; CMUs environment}{Kiwan/Amjad}.
+WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet} with TI Code Composer Studio version X~\cite{}\todo{Report on CCS version}{Amjad}.
 FET was also used as a power supply in non-intermittent power experiment.
@@
 Code completion measurements have been performed using Salea~\cite{saleae} digital logic analyzer.
+
+We also had alternative setup where we used MSP-TS430RGZ48C launchpad for continuous power experiment.
+We used internal timer for measuring run time.
+For harvested energy, we used ThingMagic Astra-EX RFID reader with power level of 17.25-30dBm.
+We used EDB~\cite{edb} for run time measurement.
+Applications were cross compiled using GCC and Clang, so that we can use LLVM for \sys compiler.
 

### 2017-08-09 paper/evaluation.tex
Kiwan's compiler text revised (added questions for Kiwan to respond)

@@
 		\hline
-		{} & \multicolumn{2}{l|}{{PVAR Size (Byte)}} & \multicolumn{2}{l|}{{\# Tasks}} & \multicolumn{2}{l|}{{\# Paging Code}} \\
+		{} & \multicolumn{2}{l|}{{PVAR size (Byte)}} & \multicolumn{2}{c|}{{No. tasks}} & \multicolumn{2}{c|}{{No. paging code}} \\
 		\hline
-		Nickname & Hand & Comp. & Hand & Comp. & Hand & Comp.\\
+		Nickname & Hand & Comp. & Hand & Comp. & Hand & Comp. \\
 		\hline\hline
 		Bitcount & 22 & 22 & 10 & 15 & 81 & 93 \\
-		Cold chain & 3492 & 3242 & 12 & 9 & 92 & 123\\
-		Cuckoo & 282 & 288 & 15 & 6 & 90 & 76\\
-		RSA & 332 & 250 & 20 & 27 & 130 & 296\\
-		AR & 166 & 218 & 11 & 6 & 112 & 333\\
-		Sort & 104 & 104 & 4 & 2 & 70 & 23\\
+		Cold chain & 3492 & 3242 & 12 & 9 & 92 & 123 \\
+		Cuckoo & 282 & 288 & 15 & 6 & 90 & 76 \\
+		RSA & 332 & 250 & 20 & 27 & 130 & 296 \\
+		AR & 166 & 218 & 11 & 6 & 112 & 333 \\
+		Sort & 104 & 104 & 4 & 2 & 70 & 23 \\
 		\hline
 	\end{tabular}
-	\caption{Comparison between compiler-generated code (Comp.) and hand-written \sys code (Hand).}
+	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand).}
 \label{table:compiler_result}
@@
 
-We start with evaluating the efficiency of a \sys compiler. Table~\ref{table:compiler_result} shows
-the difference in size of protected variable (PVAR), number of tasks, and number of inserted paging call in the code between hand-written \sys code
-and compiler-generated \sys code. All three values are better to be small. On average, compiler-generated
-code is only worse by 0.3\% in the size of protected variable. Regarding the number of tasks, compiler-generated code
-is better by roughly 16\%. The result indicate compiler is selecting boundaries better than programmer manually doing it.
-However, compiler should insert paging code conservatively when it {\em may} access protected variable, due to limitation of
-pointer alias. Thus, it results in compiler introducing 48\% more paging code than programmer-written code. The resulting run time
-difference is summarized in Figure~\ref{}. The performance is worse than programmer-written \sys code
-in overall, meaning the programmer should try to write code manually if performance is critical; otherwise compiler can be
-used to simplify the programmability.
+We start with evaluating the efficiency of a \sys compiler.
+Table~\ref{table:compiler_result} shows the difference in the number of protected variables (PVAR), number of tasks, and number of inserted paging calls in the code between (i) hand-written \sys and (ii) compiler-generated \sys code.
+On average, compiler-generated code is only worse by 0.3\% in the number of protected variables.
+Regarding the number of tasks, compiler-generated code is better by roughly 16\%.
+The result indicate that \sys compiler is selecting boundaries better than the programmer doing it manually.
+However, \sys compiler should insert paging code conservatively when it {\em may} access protected variables, due to limitation of pointer alias.
+Thus, it results in \sys compiler introducing 48\% more paging code than programmer-written code (the results on run time for various \sys code versions is provide in Section~\ref{}).
+The performance is worse than programmer-written \sys code in overall, meaning the programmer should structure the code manually if the performance is critical; otherwise compiler can be used to simplify the programmability.
 
-We also evaluated the compile time for automatic task generation. The result are in Table~\ref{table:compile_time}.
-Most of the compile time was due to pointer aliasing. We could not compile DFT application since it was using
-builtins for math functions that weren't provided by Clang MSP430 backend.
+Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile DFT application since it was using built-ins for C math functions (i.e.
+\texttt{rand()}, \texttt{sinf()}, \texttt{cosf()}) that were not provided by the Clang MSP430 back-end.}.
+The result are given in Table~\ref{table:compile_time}.
+Most of the compile time was due to pointer aliasing.\todo{Comment on: whether these times are good or bad; why there is such a difference in time between all applications; how many runs you performed; what PC/OS you used to run the code; how the result was measured; why results are rounded (there are no decimals in your result); what is "Clang MSP430 back-end"}{Kiwan}
 
@@
 	\end{tabular}
-	\caption{Compile time in seconds.}
+	\caption{\sys compile time per application (seconds).}
 \label{table:compile_time}

### 2017-08-09 paper/relatedwork.tex
Related Work-subsection 2 ready (to be revised by Brandon)

--- 2017-08-08 paper/relatedwork.tex
@@
 But there is a price tag associated to them: harvested energy is miniscule, unstable, difficult to predict ergo \emph{intermittent}.
-Therefore, to guarantee completion of computing, programming and runtime models are being actively researched that are specifically tailored to these intermittently-powered systems.
+Therefore, to guarantee completion of computing, programming and runtime models (and measurement and support systems such as~\cite{hester_sensys_2014,hester_sensys_2015,edb} which are beyond the scope of this paper) are being actively researched that are specifically tailored to these intermittently-powered systems.
 \todo{Revise this section}{Brandon}
 
-\noindent \textbf{Intermittent Execution Environments:} Review of intermittent runtime environments. \todo{Fill-in this section}{Przemek}
+\noindent \textbf{Intermittent Execution Environments:} Earlier works on runtime environments for energy-harvesting (volatile memory-only) systems include Dewdrop~\cite{dewdrop} and Eon~\cite{sorber_sensys_2007}, in which it is ensured (by a programmer or a scheduler) that task will fit the available remaining energy.
+No intermittency and data consistency issues are considered.
+Similar, specific-application tailored system includes Wisent and Stork~\cite{stork,wisent}.
+DINO~\cite{dino} versions non-volatile memory and checkpoints volatile program state which is memory and energy demanding.
+Rachet~\cite{ratchet}, implementing checkpointing as well, assumes complete non-volatile hardware.
+Clang~\cite{hicks_isca_2017} assumes a dedicated hardware for idempotency violation detection.
+On-demand checkpoint runtimes include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall}, Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus} and HarvOS~\cite{mottola2017harvos}.
+They all require hardware support to measure remaining storage energy---a feature prohibitive given typical high energy cost expenditure of such hardware.
+The only available task-based intermittent runtime systems, free from hardware assumptions or heavy checkpointing, are Chain~\cite{chain} and Alpaca~\cite{alpaca}.
+\sys goes beyond what any task-based system performed by providing \emph{automatic} task creation and \emph{adaptivity} to energy state and energy storage \emph{obliviousness} of the compiled code.
+\todo{Revise this section}{Brandon}
 
-\noindent \textbf{Memory Virtualization:} Review of memory virtualization environments (not necessarily for intermittent power devices). \todo{Fill-in this section}{Przemek}
+\noindent \textbf{Memory Virtualization:} Review of memory virtualization/transactional memory environments (not necessarily for intermittent power devices). \todo{Fill-in this section}{Alexei}

### 2017-08-09 paper/intro.tex
Section I revised

@@
 These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaic components~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
-Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such energy-independent devices are a promising underlying technology for applications in emerging areas such as Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
+Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such energy-independent devices are a promising underlying technology for applications in emerging areas such as the Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
 
@@
 When a threshold quantity of energy accumulates, the device begins operating.
-Harvestable energy sources typically produce little power compared to a platform's operating power level---compare for instance~\cite[Table III and V]{prasad_comst_2014} with~\cite[Table I]{carrano_cst_2014}.
-A devices operates for a brief period of time and when buffered energy is depleted, shuts down and begins recharging to operate again later.
-As an example, inter-task time may even exceed 20\,s for wireless energy-harvesting battery-less vivo digestive tract monitoring system~\cite[Fig.
+Harvestable energy sources typically produce little power compared to a platform's operating power level.
+A device operates for a brief period of time and when buffered energy is depleted, shuts down and begins recharging to operate again later.
+As an example, inter-task time may even exceed 20\,s for a wireless energy-harvesting battery-less vivo digestive tract monitoring system~\cite[Fig.
 3c]{nadeau_naturebio_2017}.
-On top of that, charge and discharge times vary by device and some of them~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite[Fig.
+On top of that, charge and discharge times vary by device (different energy buffer sizes) and some of them~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite[Fig.
 1]{tan_infocom_2016},~\cite[Fig.
@@
 3]{nvp}.
+
+%---compare for instance~\cite[Table III and V]{prasad_comst_2014} with~\cite[Table I]{carrano_cst_2014}
 
@@
 9]{nvp}.
-New programming models and compilers force programmers to change behavior~\cite{chain,ratchet}, inhibiting adoption.
-New memory models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude
-the use of volatile memory~\cite{ratchet}.
-Effective use of both volatile and non-volatile memory is important for efficiency, because volatile memory has higher access energy and latency than SRAM~\cite[Fig.
-4]{nvp}, and generality, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
+New programming models and compilers require programmers to change behavior~\cite{chain,ratchet}, inhibiting adoption.
+New memory models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
+Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite[Fig.
+4]{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
 Hardware is effective, but inapplicable to existing devices, making programming and execution models essential to the success of intermittent computing.
 
-Recent work~\cite{alpaca,chain} proposed {\em task-based} programming and execution models in which the programmer statically decomposes their program into a collection of tasks.
+Recent work~\cite{alpaca,chain} proposed {\em task-based} programming and execution models in which programmers statically decompose their program into a collection of tasks.
 Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
@@
 4]{chain}.
-An important programming challenge presented by a task-based model is that the length of a software task's execution
-is limited by the fixed, total amount of energy that a device can buffer in hardware.
+An important programming challenge presented by a task-based model is that the length of a software task's execution is limited by the fixed, total amount of energy that a device can buffer in hardware.
 A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict.
@@
 
-The key research question addressed by this work is to determine how to define and execute tasks efficiently and avoid non-termination, given an unknown energy buffer size and energy arrival process.
-In particular, we ask (i) how to use software support to dynamically adapt the effective size of a task, while still respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while automatically maintaining memory consistency during execution?
+The key research question addressed by this work is to determine how to define and execute tasks efficiently and ensures non-termination, given an unknown energy buffer size and energy arrival process.
+In particular, we ask (i) how to use software support to dynamically adapt the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while automatically maintaining memory consistency during execution?
 
@@
 The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent.
-\sys complier effectively circumvents the need for labour-intensive manual task generation of the state of the art solutions~\cite{chain,alpaca}.
+\sys complier effectively circumvents the need for labour-intensive manual task generation of the state-of-the-art solutions~\cite{chain,alpaca}.
 \todo{Elaborate on this more and provide numerical results}{Przemek}
 	%
-	\item {\bf Dynamic task coalescing:} \sys's memory virtualization mechanism uses SRAM as working memory, that \sys dynamically populates with pages of data from FRAM on demand during a task's execution.
+	\item {\bf Dynamic task coalescing:} \sys's memory virtualization mechanism uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand during a task's execution.
 When a task ends, \sys commits dirty pages to FRAM using efficient DMA block copies, ensuring task atomicity and memory consistency.
@@
 We evaluated \sys end-to-end on a collection of existing benchmark programs~\cite[Sec.
-5]{chain} and new ones (including realistic case study of battery-less embedded sound detector) in a laboratory environment, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexible target a variety of platforms without recompiling the code.
+5]{chain} and new ones in a laboratory environment, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without recompiling the code.
 \todo{add detailed text about evaluation results}{Przemek}

### 2017-08-09 paper/intro.tex
Section I revised

@@
 
-\todo{This is the main pitch statement: we should all agree that this is what \sys does}{All} In this work we develop {\bf \sys}: a new programming and execution model that efficiently executes tasks and avoids non-termination across a range of energy buffer sizes, without burdening the programmer.
+In this work we develop {\bf \sys}: a new programming and execution model that efficiently executes tasks and avoids non-termination across a range of energy buffer sizes, without burdening the programmer.
 To accomplish this goal, \sys introduces three new capabilities:
@@
 The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent.
-\sys complier effectively circumvents the need for labour-intensive manual task generation of the state-of-the-art solutions~\cite{chain,alpaca}.
-\todo{Elaborate on this more and provide numerical results}{Przemek}
+\sys complier effectively circumvents the need for labour-intensive manual task generation of the state-of-the-art solutions~\cite{chain,alpaca} (spanning hours of programmer work) completing task compilation in the order of seconds.
 	%
-	\item {\bf Dynamic task coalescing:} \sys's memory virtualization mechanism uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand during a task's execution.
+	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism executes multiple statically-derived tasks together as a single dynamic task.
+A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer.
+Coalescing tasks allows running the same tasks on a device with a larger energy buffer, while avoiding run time overheads associated with ending one task and beginning another.\todo{Provide numerical results for this section}{Przemek}
+	%
+	\item {\bf Volatile memory virtualization:} \sys's memory virtualization mechanism uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand during a task's execution.
 When a task ends, \sys commits dirty pages to FRAM using efficient DMA block copies, ensuring task atomicity and memory consistency.
 \todo{Once paging is fully implemented--update the paragraph}{Przemek}\todo{Provide numerical results for this section}{Przemek}
-	%
-	\item {\bf Volatile memory virtualization:} \sys's task coalescing mechanism executes multiple statically defined tasks together as a single dynamic task.
-A programmer can specify small tasks that will execute on a device with a small energy buffer.
-Coalescing tasks allows running the same tasks on a device with a larger energy buffer, while avoiding run time overheads associated with ending one task and beginning another.\todo{Provide numerical results for this section}{Przemek}
 \end{itemize}
@@
 \todo{add detailed text about evaluation results}{Przemek}
-
-%\todo{I think these are all much too long, given the depth of the intro and background already. I suggest we cut to the barest minimum.}
 

### 2017-08-09 paper/background.tex
background section reduced (Sec 2.1)

@@
-In this section, we provide background on energy harvesting systems and the intermittent software execution model.
+Here, we provide background on energy harvesting systems and the intermittent software execution model.
 We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys addresses.
@@
 Energy harvesting devices operate using energy extracted from their environment, from sources such as radio waves and solar energy.
-Energy-harvesting devices need not use a tethered power supply or battery and instead, typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.
-
-Batteryless operation has a number of important advantages, making intermittent computing an important research domain.
+Energy-harvesting devices do not use a tethered power supply or battery and instead, typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.
+%
+%Batteryless operation has a number of important advantages, making intermittent computing an important research domain.
 Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable.
@@
 The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently operating discharge time---is orders of magnitude less than a battery.
-
+%
 %Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible.
 These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
+%
+Many platforms enable intermittent, battery-less energy harvesting-based computation.
+For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
+Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
 
-Many platforms enable intermittent, battery-less computation.
-For instance, computation RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
-Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
-In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
+%In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
 Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid).

### 2017-08-09 paper/paper.tex
Asbtract swiftly rewritten - ready for Brandon
added some edits to section 2.  still needs  alot, especially in the task section
remove iposDataProtection reference from main file
done with an aggressive pass over compiler
Placeholder for Conclusions (will require Brandon's hand and eye later)
Background section revised (yaay)
patched wrong section label
Revisions of virtualization section
Section I revised
title edit

@@
 \usepackage{listings}
-\usepackage{flushend}
+%Brandon: this broke the build for me
+%\usepackage[keeplastbox]{flushend}
+
+\hypersetup{colorlinks=false,linkcolor=blue,urlcolor=blue}
+\urlstyle{rm}
 
@@
 
-\newcommand{\sys}{VIPE\xspace}
+\newcommand{\sys}{Coala\xspace}
+\newcommand{\transition}{{\tt next\_task}\xspace}
 
-%Title Options
+%Title Options (Brandon: your call which one)
 %\newcommand{\sysfull}{VIPER: Virtualised Intermittent Programming model, Execution model, and Runtime environment}
@@
 %\newcommand{\sysfull}{VIPE: Virtualized Intermittently-Powered Execution}
-\newcommand{\sysfull}{VIPER: Virtual Intermittently-Powered Execution on the Run}
+\newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Execution with Adaptive Coalescing and Memory Virtualization}
 
@@
 \section{\sysfull}
-\label{sec:overeall_system}
+\label{sec:overall_system}
 
@@
 
-\section{Compiler Support for Automatic Task Generation}
+\section{\sys Automatic Task Generation}
 \label{sec:compiler}
@@
 
-\input{iposdataprotection}
-
-\section{Memory Virtualization}
+\section{\sys Memory Virtualization}
 \label{sec:memory_virtulaization}
@@
 
-\section{Task Coalescing}
+\section{\sys Task Coalescing}
 \label{sec:task_coalescing}
@@
 
-\section{Conclusions}
+\section{Conclusions and Future Work}
 \label{sec:conclusions}

### 2017-08-09 paper/abstract.tex

@@
 Enabling battery-free devices is a mandatory step towards an environment-friendly Internet of Things (IoT).
-However, removing the batteries requires IoT to operate on an harvested, ergo intermittent, power supply.
-This make sustaining long computation very challenging.
-Correspondingly, there are two approaches to enable long-running computations on intermittenly-powered IoT: (i) checkpointing, where the volatile state of a program is frequently saved to the non-volatile memory, and (ii) task-based approach, where a programmer splits the code into small idempotent modules.
-Results so far suggest that task-based approach performs better than checkpointing.
-However, the use of tasks requires that the energy to execute any task must not exceed the maximum limit of the energy buffer.
-This makes the code unefficient (when executed on a large energy buffer than indended) and potentially completely disables the program (in the opposite case).
-Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing.
+However, removing the batteries requires IoT to operate on a harvested, ergo intermittent, power supply.
+This makes sustaining long computations challenging.
+Basically, there are two approaches: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, and (ii) task-based approach, where a programmer splits the code into small idempotent sections.
+Results so far suggest that a task-based approach performs better than checkpointing.
+However, the use of tasks requires that the energy to execute any task must not exceed the capacity of the energy buffer.
+This makes the code inefficient (when executed on a larger energy buffer than intended) and potentially completely disables the program (in the opposite case).
+Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing and automatic compiler-supported task generation.
 If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve the performance.

### 2017-08-09 paper/background.tex

@@
 Energy-harvesting devices execute software according to the {\em intermittent execution model}~\cite{dino,chain,alpaca,ratchet}.
-Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
+%
+%Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
+%
 Software running on the device executes {\em intermittently} because energy is only available between when the capacitor reaches its threshold and when it is depleted.
@@
 By default, after a power failure control flows to the beginning of {\tt main()}.
-Early work on intermittent computing preserve intermittent progress by periodically capturing and restoring to a non-volatile checkpoint of the volatile execution context, i.e., registers and stack~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
+Early work on intermittent computing preserved intermittent progress by periodically capturing and restoring to a non-volatile checkpoint of the volatile execution context, i.e., registers and stack~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
 
@@
 
-Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using cyclic redundancy check (CRC) code from MIBench~\cite{hicks_mibench2_2016}.
-The code computes CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
+Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench~\cite{hicks_mibench2_2016}.
+The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
 Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of a WAR dependence between the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
@@
 Another strategy is to statically create multiple versions of non-volatile data shared by tasks and ensure by construction that no task reads and writes the same non-volatile memory location~\cite{chain}.
-Regardless of the idempotence strategy, the key to task-based execution is that fixed, statically defined tasks execute as though atomic, completing after zero or more attempts.
+Regardless of the idempotence strategy, the key to task-based execution is that fixed, statically-defined tasks execute as though atomic, completing after zero or more attempts.
 
@@
 		\hline \hline
-		Moo~\cite{moo} & 0.1\,F \\
+		%Moo~\cite{moo} & 0.1\,F \\
 		WISPCam~\cite{naderiparizi_rfid_2015} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
-		NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
+		%NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
 		NeuralWISP~\cite{holleman_biocas_2008} & 100\,$\mu$F \\
@@
 	\end{tabular} 
-	\caption{Comparison of {\em default} energy storage sizes for various battery-less platforms.
-Observe a huge variation in storage capacity.
-We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.}
+	\caption{Comparison of {\em default} energy storage sizes for example battery-less platforms. Observe a huge variation in storage capacity.}
+	%We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.
 	\label{table:capacitor}
@@
 
-\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} A programmer must write each task with statically defined boundaries, which makes statically sized tasks inflexible and non-portable.
+\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} A programmer must write each task with statically defined boundaries, which makes statically-sized tasks inflexible and non-portable.
 The programmer must ensure statically that a task's energy requirements match the device's energy storage capacity.
-If a programmer writes a program that consumes more energy than the device can buffer, the task will never be able to complete using the buffered energy.
-The consequence of such tasks is a non-terminating program that repeatedly attempts to re-execute the task of excess energy cost.
+If a programmer writes a task that consumes more energy than the device can buffer, the task will never be able to complete using the buffered energy.
+The consequence of such tasks is a non-progressing program that repeatedly attempts to re-execute the task of excess energy cost.
 If a programmer writes a program that consumes less energy than the device can buffer, the system may operate inefficiently.
-When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).
+When a task and its successor both complete, the two tasks were interrupted by a task boundary that preserves progress and state (i.e., checkpoint or commit).
 If the two tasks were not also interleaved by a power failure, then the computation done to preserve state was {\em unnecessary}, only incurring an overhead on the tasks' executions.
@@
 The key problem is that the programmer writes tasks statically for an energy buffer of a fixed size.
-\sys's {\em task coalescing} mechanism, which we describe in Section~\ref{sec:task_coalescing}, is motivated by the observation that systems with statically defined tasks may be inflexible and inefficient.
+\sys's {\em task coalescing} mechanism, which we will describe in Section~\ref{sec:task_coalescing}, is motivated by the observation that systems with statically-defined tasks may be inflexible and inefficient.
 
@@
 Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
-We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors.~\cite{su_date_2017,ratchet,quickrecall,nvp}.
+We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
 \sys supports I/O behavior similar to recent intermittent execution models~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.

### 2017-08-09 paper/evaluation.tex
Kiwan's compiler text revised (added questions for Kiwan to respond)

@@
 This set of benchmarks is the same as used in~\cite{chain,alpaca}\todo{Fill-in the missing data and verify its correctness}{Przemek/Amjad}.
-In addition to Chain benchmarks new applications were implemented: (v) \emph{DFT}, (vi)\emph{Huffman compressor}, and (vii) \emph{Selection Sort}.
+In addition to Chain benchmarks new applications were implemented: (v) \textbf{DFT}, (vi) \textbf{Huffman compressor}, and (vii) \textbf{Selection Sort}.
 All applications are compared in Table~\ref{table:benchmark_list} in terms of source lines of code against Chain.
@@
 	\end{tabular}
-	\caption{\sys compile time per application (seconds).}
+	\caption{\sys compile time per application (seconds).\todo{for Sort:">" or "<"?}{Kiwan}}
 \label{table:compile_time}

### 2017-08-09 paper/compiler.tex
Compiler section revised

@@
 \sys provides compiler support for \emph{automatic task generation} that simplifies writing task-based code by automatically translating C code into a task-based executable.
-Translating C code into task-based code automatically requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
-The \sys compiler first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary.
+Doing so requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
+The \sys compiler, \textcolor{red}{based on LLVM}, first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary.
 After identifying where to put task boundaries, the \sys compiler transforms the imperative code into tasks, instrumenting protected memory accesses.
@@
 Figure~\ref{fig:comp_2}).
-To simplify inter-procedural analysis, our prototype inlines all function calls into the main function in the CFG.
-\sys uses a cost-based task boundary placement analysis that considers per-block energy consumption estimates, inter-task variable liveness, and the cost of partitioning loops.
+To simplify inter-procedural analysis, our prototype inlines all function calls\todo{how about recursions and code duplication?}{Brandon/Kiwan} into the main function in the CFG.
+%
+%\sys uses a cost-based task boundary placement analysis that considers per-block energy consumption estimates, inter-task variable liveness, and the cost of partitioning loops.
+%
 We first describe how \sys measures each of these costs and then describe \sys's procedure for placing boundaries in the CFG.
@@
 {\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block.
-Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we used the number of intermediate representation (IR) instructions in a block as a proxy for energy because energy cost is likely to closely correspond to block length.
-Moreover, we
-assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy.
+Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we simply use the number of intermediate representation (IR) instructions in a block as a proxy for energy because energy cost.
+Moreover, we assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy.
 This parameter is highly device-dependent, and we assume that a developer or device manufacturer determines the number empirically, as we did in this work through direct measurement.
-\sys's analysis tries ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget.
+\sys's analysis tries to ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget (which is part of the boundary placement process described later).
 
-{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of {\em live variables} that are accessible in both of the blocks that define the edge (i.e., predecessor and successor).
+{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of {\em live variables} that are accessed in both\todo{Transitive closure?}{Brandon/Kiwan} of the blocks that define the edge (i.e., predecessor and successor).
 Variables that live across an edge between two tasks must be protected by \sys's memory virtualization mechanism and each live, inter-task variable increases the run time overhead for a boundary placement.
-\sys's analysis tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.
+\sys's tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.
 
 {\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis must account for the cost of multiple iterations of the loop body. \sys assumes that most loops are
-annotated with a static loop bound, which is common in embedded code.
-For loops that do not have a static bound, \sys conservatively assumes the maximum possible loop bound, based on the type of the loop induction variable.
+annotated with a static loop bound. For loops that do not have a static bound, \sys conservatively assumes the maximum possible loop bound, based on the type of the loop induction variable.
+
+%removed statement "which is common in embedded code"
 
@@
 
-The analysis uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score.
+\sys uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score.
 The heuristic tries to avoid placing boundaries on loops, to minimize the number of live variables that must be protected across the new boundary, and to maximize the number of operations along the path.

### 2017-08-09 paper/coalescing.tex
pass over the coalescing section

@@
-\sys is able to dynamically merge real tasks to construct a bigger virtual task.
-This enables \sys to reduce the number of commits by $\frac{\text{VTS}-1}{\text{VTS}}$, where VTS is virtual task size.
-Furthermore, \sys tries to estimate the optimal virtual task size by relying on its recent execution history.
-The challenge in designing \sys's coalescing algorithm is in introducing as less overhead as possible and able to respond to the changes in the environment as quickly as possible.
+\sys dynamically {\em coalesces} consecutively executing tasks at run time.
+When two tasks are coalesced, the first of the two tasks does not commit its
+updates to protected memory locations when it reaches a \transition statement.
+Instead, execution immediately continues from the beginning of the second of
+the two tasks.  At the end of the second of the two coalesced tasks, \sys
+commits the pages that were updated by either of the coalescing tasks.  The
+purpose of coalescing is to better amortize the run time overhead of committing
+updated pages when a task ends. If consecutive tasks access the same pages of
+data, the pages accessed by both tasks only need to be committed only once for
+the coalesced task rather than twice, one time for each of the tasks executed
+individually.
+
+\sys can coalesce an arbitrary number consecutive tasks.  As more tasks
+coalesce, their collective commit overhead amortizes better, potentially
+improving performance.  As more tasks coalesce, the amount of potentially {\em
+wasted work} also increases. If power fails during a long sequence of coalesced
+tasks the lack of an intervening commit requires execution to restart from the
+{\em first} task in the sequence, failing to preserve the progress of any of
+the tasks in the long sequence.
 
@@
 
-\sys's coalescing algorithm uses a single point back in the \sys's execution history to estimate the appropriate virtual task size.
-However, it takes a conservative approach by making the virtual task size equal to the half of the number of successfully executed real tasks before the last power interrupt (see lines~\ref{algo:coalescing:executionHistory1}, \ref{algo:coalescing:executionHistory2} of Algorithm~\ref{algo:coalescing}).
-We advocate for the algorithm conservative approach by the following two reasons: (i) the power interrupts are not uniformly distributed over the tasks; and (ii) the re-execution penalty, in case of a power interrupt, grows linearly with the size of a virtual task.
-Furthermore, the algorithm considers the charging rate by allowing the real tasks counter (RTC), which represents the execution history between the last two power interrupts, to increase from one up to multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of Algorithm~\ref{algo:coalescing}).
-As a consequence, the virtual task size changes rapidly (since it is half of the execution history) in response to the change in the charging rate.
-However, the algorithm sets a limit to the size of the virtual task for these two reasons: (i) having an infinity virtual task size, results in a certain computation progress loss; and (ii) the benefit of task merging is governed by the diminishing return principle (which will be demonstrated experimentally in Section~\ref{sec:results_coalescing}).
+\sys's coalescing algorithm uses a single point back in the \sys's execution
+history to estimate the appropriate virtual task size. However, it takes a
+conservative approach by making the virtual task size equal to the half of the
+number of successfully executed real tasks before the last power interrupt (see
+lines~\ref{algo:coalescing:executionHistory1},
+\ref{algo:coalescing:executionHistory2} of Algorithm~\ref{algo:coalescing}). We
+advocate for the algorithm conservative approach by the following two reasons:
+(i) the power interrupts are not uniformly distributed over the tasks; and (ii)
+the re-execution penalty, in case of a power interrupt, grows linearly with the
+size of a virtual task. Furthermore, the algorithm considers the charging rate
+by allowing the real tasks counter (RTC), which represents the execution
+history between the last two power interrupts, to increase from one up to
+multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of
+Algorithm~\ref{algo:coalescing}). As a consequence, the virtual task size
+changes rapidly (since it is half of the execution history) in response to the
+change in the charging rate. However, the algorithm sets a limit to the size of
+the virtual task for these two reasons: (i) having an infinity virtual task
+size, results in a certain computation progress loss; and (ii) the benefit of
+task merging is governed by the diminishing return principle (which will be
+demonstrated experimentally in Section~\ref{sec:results_coalescing}).
 

### 2017-08-09 paper/systemdescription.tex

@@
 \sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
-\sys addresses the challenges motivated in Section~\ref{sec:background} to make task-based intermittent programs {\em efficient}, {\em flexible}, and {\em programmable}.
+\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em efficient}, {\em flexible}, and {\em programmable}.
 \sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.
@@
 	\end{tabular}
-	\caption{Summary of \sys's programming model extensions}
+	\caption{Summary of \sys's programming model extensions.\todo{Revise this table (confusion with next\_task() function)}{Amjad/Przemek}}
 	\label{tab:viper_syntax}
@@
 
-\noindent \textbf{\sys Memory Manager:} When the binary executes, \sys's runtime support affects its execution through {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}) and {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}).
+\noindent \textbf{\sys Memory Manager:} When the binary executes, \sys's runtime support affects its execution through a {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}) and a {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}).
 The \sys binary executes its tasks correctly, with consistent memory state, due to \sys's virtual memory mechanism.

### 2017-08-09 paper/coalescing.tex
wrote the coalescing section.  we need the figure

@@
 {\em first} task in the sequence, failing to preserve the progress of any of
-the tasks in the long sequence.
+the tasks in the long sequence. \sys moderates the risk of coalescing and
+capitalizes on its benefit by dynamically, adaptively determining a coalescing
+factor that combines many tasks without suffering an excess of wasted work.
 
-\subsection{Task Coalescing Algorithm}
+The key challenge to coalescing tasks is determining how many tasks to coalesce
+before committing.  During execution, an executing task is either a {\em
+coalesced} task, which does not commit its state before transitioning or a {\em
+committing} task, which does commit its state before transitioning. \sys
+monitors a program's execution to adaptively determine $C$, the number of tasks
+to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks,
+followed by a committing task.
 
-\begin{algorithm}[t]
-	\caption{\sys task coalescing mechanism}
-	\label{algo:coalescing}
-	\scriptsize
-	\begin{algorithmic}[1]
-		\State $\text{p}$  \Comment{Indicates persistent variables, initialized \textbf{only} during device programming}
-		\State $\text{RT} \in \{\sys~\text{static tasks}\}$  \Comment{Real task}
-		\State $\text{VT} \subset \{~\text{RT's}\}$  \Comment{Virtual task}
-		\State $\text{RTC}$  \Comment{RT counter}
-		\State $\text{VTC}$  \Comment{VT counter}
-		% \State $|\text{VT}|$ \Comment{VT size measured in RT}
-		\State $\text{TVT}$ \Comment{Temporary VT}
-		\State $\text{VT}_{\max}$ \Comment{Maximum VT size measured in RT (coalescing upper bound)}
-		\vspace{0.1cm}
+\sys determines $C$ for a given operating based on the number of tasks
+completed during the last operational period.  \sys begins execution in a
+coalesced task with an infinite $C$ value.  With a continuous power supply, the
+system executes from the start with all tasks as coalesced tasks entirely
+avoiding the cost of committing state.  On intermittent power, the system
+executes a sequence of tasks that is interrupted by a power failure.  \sys
+explicitly tracks $S$, the number of tasks completed during the previous
+operating period.  \sys assigns $C$ to be equal to half of the value of $S$.
+If more tasks in total are able to complete in an operating period, \sys
+attempts to coalesce more tasks during the next operating period.  If fewer
+tasks complete, \sys attempts to coalesce fewer tasks during the next operating
+period. $S$ includes the count of both coalesced and committing tasks
+completed.  If no committing task completes (i.e., execution does not make
+progress), \sys updates $S$ and attempts to coalesce fewer tasks during the
+next operating period.
 
-		\State $\text{p RTC} = x $ 
-		\State $\text{p VT}_{\max} = y$
-		\State $\text{p VT} = \text{RT}$ 
-		\State $\text{p TVT} = \text{VT} $ 
-		\State $\text{VTC} = 0 $ 
-		\vspace{0.1cm}
+Figure~\ref{fig:coalescing} illustrates how \sys coalesces tasks.
+Figure~\ref{fig:coalescing}(a) shows a sequence of six tasks executing in \sys
+with no coalescing, incurring the cost of commit at each task transition.
+Figure~\ref{fig:coalescing}(b) shows the same sequence of tasks executing as
+coalesced groups of three tasks. Figure~\ref{fig:coalescing}(c) shows how \sys
+adapts the value of $C$ based on $S$.  \sys executes 12 tasks in the first
+operating period, assigning $S$ to six.  \sys completes only five of the six
+coalesced tasks in the next operating period and then power fails. \sys sets
+$S$ to two in the next operating period, coalescing pairs of tasks and
+successfully committing two pairs before power fails. Execution resumes at the
+start of the third pair of tasks with a coalescing factor of two.  If a larger
+number of tasks successfully complete in the next operating period (i.e., $S$
+is larger), $C$ correspondingly increases, allowing more aggressive coalescing.
 
-		\Function{scheduler}{\null}
+%
+%However, it takes a conservative approach by making the virtual task size equal
+%to the half of the number of successfully executed real tasks before the last
+%power interrupt (see lines~\ref{algo:coalescing:executionHistory1},
+%\ref{algo:coalescing:executionHistory2} of Algorithm~\ref{algo:coalescing}). 
+%
+%We
+%advocate for the algorithm conservative approach by the following two reasons:
+%(i) the power interrupts are not uniformly distributed over the tasks; and (ii)
+%the re-execution penalty, in case of a power interrupt, grows linearly with the
+%size of a virtual task. 
+%
+%Furthermore, the algorithm considers the charging rate
+%by allowing the real tasks counter (RTC), which represents the execution
+%history between the last two power interrupts, to increase from one up to
+%multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of
+%Algorithm~\ref{algo:coalescing}). 
+%
+%As a consequence, the virtual task size
+%changes rapidly (since it is half of the execution history) in response to the
+%change in the charging rate. 
+%
+%However, the algorithm sets a limit to the size of
+%the virtual task for these two reasons: (i) having an infinity virtual task
+%size, results in a certain computation progress loss; and (ii) the benefit of
+%task merging is governed by the diminishing return principle (which will be
+%demonstrated experimentally in Section~\ref{sec:results_coalescing}).
+%
+%The algorithm protects itself from power interrupts by going into three stages:
+%(i) virtual progressing, where the operations are done on the volatile memory
+%(see
+%lines~\ref{algo:coalescing:virtualProgressing1}--\ref{algo:coalescing:virtualProgressing2}
+%of Algorithm~\ref{algo:coalescing} ); 
+%(ii) transition stage, where the volatile
+%state is moved to a temporary persistent buffer. 
+%
+%Until the end of the second
+%stage if a power is interrupted, then all the virtual progress will be
+%cancelled and the virtual task will re-executed and re-initialized from a
+%consistent input (see line~\ref{algo:coalescing:virtualProgressing1} of
+%Algorithm~\ref{algo:coalescing}); 
+%
+%and (iii) persistent commit stage, upon
+%entering this stage the algorithm will make a firm transition (see
+%lines~\ref{algo:coalescing:firmTransition1},
+%\ref{algo:coalescing:firmTransition2} of Algorithm~\ref{algo:coalescing}) and
+%it will not go back to the other stages unless all the data is committed from
+%the persistent buffer to non-volatile memory. 
+%
+%Since the data is being moved in one direction and from a persistent location,
+%power interrupt is tolerable at this stage.
 
-			\State $\text{VTC} = \text{RTC/2} $  \label{algo:coalescing:executionHistory1}
-			\State $\text{RTC} = 0 $
-			\State $\text{RT} = \text{VT}$ \Comment{Recover virtual state} \label{algo:coalescing:virtualProgressing1}
-			\If{commiting}
-				\State $\text{RT}=\text{TVT}$
-				\State $\texttt{goto} \ \ \ref{commitStage}$ \label{algo:coalescing:firmTransition1}
-			\EndIf
-			\vspace{0.1cm}
+%\begin{algorithm}[t]
+%	\caption{\sys task coalescing mechanism}
+%	\label{algo:coalescing}
+%	\scriptsize
+%	\begin{algorithmic}[1]
+%		\State $\text{p}$  \Comment{Indicates persistent variables, initialized \textbf{only} during device programming}
+%		\State $\text{RT} \in \{\sys~\text{static tasks}\}$  \Comment{Real task}
+%		\State $\text{VT} \subset \{~\text{RT's}\}$  \Comment{Virtual task}
+%		\State $\text{RTC}$  \Comment{RT counter}
+%		\State $\text{VTC}$  \Comment{VT counter}
+%		% \State $|\text{VT}|$ \Comment{VT size measured in RT}
+%		\State $\text{TVT}$ \Comment{Temporary VT}
+%		\State $\text{VT}_{\max}$ \Comment{Maximum VT size measured in RT (coalescing upper bound)}
+%		\vspace{0.1cm}
+%		
+%		\State $\text{p RTC} = x $ 
+%		\State $\text{p VT}_{\max} = y$
+%		\State $\text{p VT} = \text{RT}$ 
+%		\State $\text{p TVT} = \text{VT} $ 
+%		\State $\text{VTC} = 0 $ 
+%		\vspace{0.1cm}
+%
+%		\Function{scheduler}{\null}
+%
+%			\State $\text{VTC} = \text{RTC/2} $  \label{algo:coalescing:executionHistory1}
+%			\State $\text{RTC} = 0 $
+%			\State $\text{RT} = \text{VT}$ \Comment{Recover virtual state} \label{algo:coalescing:virtualProgressing1}
+%			\If{commiting}
+%				\State $\text{RT}=\text{TVT}$
+%				\State $\texttt{goto} \ \ \ref{commitStage}$ \label{algo:coalescing:firmTransition1}
+%			\EndIf
+%			\vspace{0.1cm}
+%
+%			\While {True}
+%				\While{$\text{VTC}--$}
+%					\Function{execute}{$\text{RT}$}
+%					\EndFunction
+%					\If{$\text{RTC} < \text{VT}_{\max}$}
+%						\State $\text{RTC}++$  \label{algo:coalescing:realTaskCounter}
+%					\EndIf
+%					\State $\text{RT} \leftarrow \text{RT}_{next}$ \Comment{Virtual progressing}
+%				\EndWhile        \label{algo:coalescing:virtualProgressing2}
+%
+%				\State \Comment{Virtual task is finished}
+%				\State $\text{Data} \rightarrow \text{pTempBuf}$ \Comment{pTempBuf: persistent temporary buffer}
+%				\State $\text{TVT} = \text{RT}$ 
+%				\State $\text{commiting} = \text{True}$ \label{algo:coalescing:firmTransition2}
+%				\State 
+%				\State $\text{VT} = \text{TVT}$ \label{commitStage}
+%				\State $\text{pTempBuf} \rightarrow \text{FRAM}$ 
+%				\State $\text{VTC} = \text{RTC/2}$ 		\Comment{Set VT size} \label{algo:coalescing:executionHistory2}
+%				\State $\text{commiting} = \text{False}$ 
+%
+%			\EndWhile
+%
+%		\EndFunction
+%			
+%	\end{algorithmic}
+%\end{algorithm}
 
-			\While {True}
-				\While{$\text{VTC}--$}
-					\Function{execute}{$\text{RT}$}
-					\EndFunction
-					\If{$\text{RTC} < \text{VT}_{\max}$}
-						\State $\text{RTC}++$  \label{algo:coalescing:realTaskCounter}
-					\EndIf
-					\State $\text{RT} \leftarrow \text{RT}_{next}$ \Comment{Virtual progressing}
-				\EndWhile        \label{algo:coalescing:virtualProgressing2}
-
-				\State \Comment{Virtual task is finished}
-				\State $\text{Data} \rightarrow \text{pTempBuf}$ \Comment{pTempBuf: persistent temporary buffer}
-				\State $\text{TVT} = \text{RT}$ 
-				\State $\text{commiting} = \text{True}$ \label{algo:coalescing:firmTransition2}
-				\State 
-				\State $\text{VT} = \text{TVT}$ \label{commitStage}
-				\State $\text{pTempBuf} \rightarrow \text{FRAM}$ 
-				\State $\text{VTC} = \text{RTC/2}$ 		\Comment{Set VT size} \label{algo:coalescing:executionHistory2}
-				\State $\text{commiting} = \text{False}$
-
-			\EndWhile
-
-		\EndFunction
-
-	\end{algorithmic}
-\end{algorithm}
-
-\sys's coalescing algorithm uses a single point back in the \sys's execution
-history to estimate the appropriate virtual task size. However, it takes a
-conservative approach by making the virtual task size equal to the half of the
-number of successfully executed real tasks before the last power interrupt (see
-lines~\ref{algo:coalescing:executionHistory1},
-\ref{algo:coalescing:executionHistory2} of Algorithm~\ref{algo:coalescing}). We
-advocate for the algorithm conservative approach by the following two reasons:
-(i) the power interrupts are not uniformly distributed over the tasks; and (ii)
-the re-execution penalty, in case of a power interrupt, grows linearly with the
-size of a virtual task. Furthermore, the algorithm considers the charging rate
-by allowing the real tasks counter (RTC), which represents the execution
-history between the last two power interrupts, to increase from one up to
-multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of
-Algorithm~\ref{algo:coalescing}). As a consequence, the virtual task size
-changes rapidly (since it is half of the execution history) in response to the
-change in the charging rate. However, the algorithm sets a limit to the size of
-the virtual task for these two reasons: (i) having an infinity virtual task
-size, results in a certain computation progress loss; and (ii) the benefit of
-task merging is governed by the diminishing return principle (which will be
-demonstrated experimentally in Section~\ref{sec:results_coalescing}).
-
-The algorithm protects itself from power interrupts by going into three stages: (i) virtual progressing, where the operations are done on the volatile memory (see lines~\ref{algo:coalescing:virtualProgressing1}--\ref{algo:coalescing:virtualProgressing2} of Algorithm~\ref{algo:coalescing} ); (ii) transition stage, where the volatile state is moved to a temporary persistent buffer.
-Until the end of the second stage if a power is interrupted, then all the virtual progress will be cancelled and the virtual task will re-executed and re-initialized from a consistent input (see line~\ref{algo:coalescing:virtualProgressing1} of Algorithm~\ref{algo:coalescing}); and (iii) persistent commit stage, upon entering this stage the algorithm will make a firm transition (see lines~\ref{algo:coalescing:firmTransition1}, \ref{algo:coalescing:firmTransition2} of Algorithm~\ref{algo:coalescing}) and it will not go back to the other stages unless all the data is committed from the persistent buffer to non-volatile memory.
-Since the data is being moved in one direction and from a persistent location, power interrupt is tolerable at this stage.
-
-\subsection{Power Interrupt Immune Scheduler}
-
-\input{relativeJumpAlgo.tex} %Task jumping algorithm
-
-It utilizes a persistent circular buffer (i.e.
+%\subsection{Power Interrupt Immune Scheduler}
+%
+%\input{relativeJumpAlgo.tex} %Task jumping algorithm
+%
+%It utilizes a persistent circular buffer (i.e.
 persistent linked list) to keep the state of a program across power failures.
@@
 \todo{Expand this section}{Amjad}
-
+%
 %\begin{algorithm}

### 2017-08-10 paper/coalescing.tex
First pass over new coalescing section (preparing to draw a figure)

@@
 \sys dynamically {\em coalesces} consecutively executing tasks at run time.
-When two tasks are coalesced, the first of the two tasks does not commit its
-updates to protected memory locations when it reaches a \transition statement.
-Instead, execution immediately continues from the beginning of the second of
-the two tasks.  At the end of the second of the two coalesced tasks, \sys
-commits the pages that were updated by either of the coalescing tasks.  The
-purpose of coalescing is to better amortize the run time overhead of committing
-updated pages when a task ends. If consecutive tasks access the same pages of
-data, the pages accessed by both tasks only need to be committed only once for
-the coalesced task rather than twice, one time for each of the tasks executed
-individually.
+When two tasks are coalesced, the first of the two tasks does not commit its updates to protected memory locations when it reaches a \transition statement.
+Instead, execution immediately continues from the beginning of the second of the two tasks.
+At the end of the second of the two coalesced tasks, \sys commits the pages that were updated by either of the coalescing tasks.
+The purpose of coalescing is to better amortize the run time overhead of committing updated pages when a task ends.
+If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed only once for the coalesced task rather than twice, one time for each of the tasks executed individually.
 
-\sys can coalesce an arbitrary number consecutive tasks.  As more tasks
-coalesce, their collective commit overhead amortizes better, potentially
-improving performance.  As more tasks coalesce, the amount of potentially {\em
-wasted work} also increases. If power fails during a long sequence of coalesced
-tasks the lack of an intervening commit requires execution to restart from the
-{\em first} task in the sequence, failing to preserve the progress of any of
-the tasks in the long sequence. \sys moderates the risk of coalescing and
-capitalizes on its benefit by dynamically, adaptively determining a coalescing
-factor that combines many tasks without suffering an excess of wasted work.
+\sys can coalesce an arbitrary number consecutive tasks.
+As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance.
+As more tasks coalesce, the amount of potentially {\em wasted work} also increases.
+If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of
+the tasks in the long sequence.
+\sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, adaptively determining a coalescing factor that combines many tasks without suffering an excess of wasted work.
 
-The key challenge to coalescing tasks is determining how many tasks to coalesce
-before committing.  During execution, an executing task is either a {\em
-coalesced} task, which does not commit its state before transitioning or a {\em
-committing} task, which does commit its state before transitioning. \sys
-monitors a program's execution to adaptively determine $C$, the number of tasks
-to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks,
-followed by a committing task.
+The key challenge to coalescing tasks is determining how many tasks to coalesce before committing.
+During execution, an executing task is either a {\em coalesced} task, which does not commit its state before transitioning or a {\em committing} task, which does commit its state before transitioning.
+\sys monitors a program's execution to adaptively determine $C$, the number of tasks to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks, followed by a committing task.
 
-\sys determines $C$ for a given operating based on the number of tasks
-completed during the last operational period.  \sys begins execution in a
-coalesced task with an infinite $C$ value.  With a continuous power supply, the
-system executes from the start with all tasks as coalesced tasks entirely
-avoiding the cost of committing state.  On intermittent power, the system
-executes a sequence of tasks that is interrupted by a power failure.  \sys
-explicitly tracks $S$, the number of tasks completed during the previous
-operating period.  \sys assigns $C$ to be equal to half of the value of $S$.
-If more tasks in total are able to complete in an operating period, \sys
-attempts to coalesce more tasks during the next operating period.  If fewer
-tasks complete, \sys attempts to coalesce fewer tasks during the next operating
-period. $S$ includes the count of both coalesced and committing tasks
-completed.  If no committing task completes (i.e., execution does not make
-progress), \sys updates $S$ and attempts to coalesce fewer tasks during the
-next operating period.
+\sys determines $C$ for a given operating based on the number of tasks completed during the last operational period.
+\sys begins execution in a coalesced task with an infinite $C$ value.
+With a continuous power supply, the system executes from the start with all tasks as coalesced tasks entirely avoiding the cost of committing state.
+On intermittent power, the system executes a sequence of tasks that is interrupted by a power failure.
+\sys explicitly tracks $S$, the number of tasks completed during the previous operating period.
+\sys assigns $C$ to be equal to half of the value of $S$.
+If more tasks in total are able to complete in an operating period, \sys attempts to coalesce more tasks during the next operating period.
+If fewer tasks complete, \sys attempts to coalesce fewer tasks during the next operating period.
+$S$ includes the count of both coalesced and committing tasks completed.
+If no committing task completes, i.e., execution does not make progress, \sys updates $S$ and attempts to coalesce fewer tasks during the next operating period.
 
-Figure~\ref{fig:coalescing} illustrates how \sys coalesces tasks.
-Figure~\ref{fig:coalescing}(a) shows a sequence of six tasks executing in \sys
-with no coalescing, incurring the cost of commit at each task transition.
-Figure~\ref{fig:coalescing}(b) shows the same sequence of tasks executing as
-coalesced groups of three tasks. Figure~\ref{fig:coalescing}(c) shows how \sys
-adapts the value of $C$ based on $S$.  \sys executes 12 tasks in the first
-operating period, assigning $S$ to six.  \sys completes only five of the six
-coalesced tasks in the next operating period and then power fails. \sys sets
-$S$ to two in the next operating period, coalescing pairs of tasks and
-successfully committing two pairs before power fails. Execution resumes at the
-start of the third pair of tasks with a coalescing factor of two.  If a larger
-number of tasks successfully complete in the next operating period (i.e., $S$
-is larger), $C$ correspondingly increases, allowing more aggressive coalescing.
+\begin{figure}
+	\centering
+	%\subfloat[]{\includegraphics[width=\columnwidth]{figures/}\label{fig:}}\\
+	%\subfloat[]{\includegraphics[width=\columnwidth]{figures/}\label{fig:}}
+	\caption{\sys task coalescing process.}
+	\label{ffig:coalescing_example}
+\end{figure}
 
-%
+\nonindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
+Figure~\ref{fig:coalescing_example}(a) shows a sequence of six tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.
+Figure~\ref{fig:coalescing_example}(b) shows the same sequence of tasks executing as coalesced groups of three tasks.
+Figure~\ref{fig:coalescing_example}(c) shows how \sys adapts the value of $C$ based on $S$.
+\sys executes 12 tasks in the first operating period, assigning $S$ to six.
+\sys completes only five of the six coalesced tasks in the next operating period and then power fails.
+\sys sets $S$ to two in the next operating period, coalescing pairs of tasks and successfully committing two pairs before power fails.
+Execution resumes at the start of the third pair of tasks with a coalescing factor of two.
+If a larger number of tasks successfully complete in the next operating period, i.e., $S$ is larger, $C$ correspondingly increases, allowing more aggressive coalescing.
+
 %However, it takes a conservative approach by making the virtual task size equal

### 2017-08-10 paper/coalescing.tex
Covering Koen's comments

@@
 the tasks in the long sequence.
-\sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, adaptively determining a coalescing factor that combines many tasks without suffering an excess of wasted work.
+\sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, \emph{adaptively} determining a coalescing factor that combines many tasks without suffering an excess of wasted work.
 
@@
 
-\nonindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
+\noindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
 Figure~\ref{fig:coalescing_example}(a) shows a sequence of six tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.

### 2017-08-10 paper/evaluation.tex
Compiler section revised
Covering Koen's comments

@@
 We used EDB~\cite{edb} for run time measurement.
-Applications were cross compiled using GCC and Clang, so that we can use LLVM for \sys compiler.
+Applications were cross compiled using GCC and Clang, so that we can use LLVM for \sys compiler.\todo{Report on LLVM version}{Brandon}
 
@@
 
-\subsubsection{Task Coalescing Performance}
-\label{sec:results_coalescing}
+%\subsubsection{Task Coalescing Performance}
+%\label{sec:results_coalescing}
+
+\vspace{2cm}
+
+List of figures: 
+\begin{itemize}
+	\item Fig. 1: Viper only/per app/fixed power/(hand/complier)/no coalescing/clang [Resuls are ready by Kiwan]
+	\item Fig. 2: Viper only/per app/per distance (fixed/20/40/60)/(hand/complier)/coalescing [Resuls are ready by Kiwan]
+	\item Fig. 3a: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/clang [Resuls are ready by Kiwan]
+	\item Fig. 3b: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/gcc [Resuls are not ready: Kiwan needs to make them]
+	\item Fig. 4: exactly as Fig 12 in the paper (4 apps only)
+	\item Fig. 5: average virtual task size/gcc/fixed distance/per app 
+\end{itemize}
+
+\vspace{2cm}
 
@@
 The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e.
-from energy harvesting to continuous power environment) is the best one?.
+from energy harvesting to continuous power environment) is the best one?
 To answer this question we have implemented a simple program composed of $x$ empty tasks.

### 2017-08-10 paper/paging.tex
Covering Koen's comments

@@
 
-\sys uses a page-based virtual memory scheme that allows tasks to primarily use SRAM instead of frequently accessing FRAM.
-The paging scheme partitions the non-volatile memory address space into pages of {\em protected data}.
+\sys paging scheme partitions the non-volatile memory address space into pages of {\em protected data}.
 As they are accessed by a task, pages of protected data are swapped into a {\em page buffer} that is stored in volatile memory and holds a single page.
-When a modified page is swapped out, it is buffered in a dedicated commit buffer for later commit.
+When a modified page is swapped out, it is buffered in a dedicated commit buffer for later.
 \sys uses two-phase commit to write buffered, modified pages back to their original memory locations atomically, ensuring data consistency even with power failures.
 \sys's page-based memory model allows moving large blocks of data between non-volatile and volatile memory, which \sys does efficiently using common,
-hardware DMA block-copy operations that have higher bandwidth and avoid fixed, 
-per-variable data movement costs.
-\sys uses efficient DMA block copies when paging data in and out, as well as during two-phase commit.
-\sys is unlike prior work~\cite{chain,alpaca} that moves one variable at a time, prohibiting efficient, DMA bulk transfers.
+hardware DMA block-copy operations, unlike prior work~\cite{chain,alpaca} that moves one variable at a time.
 Moreover, \sys's paging and buffering scheme is a key enabler of task coalescing, which we describe in Section~\ref{sec:task_coalescing}.
@@
 In \sys, a virtual address is composed of a \emph{page tag} and a \emph{page offset}.
-The page tag identifies in which page of non-volatile memory the memory location resides and
-the page offset identifies a particular byte of that page.
-How an $a$-bit memory address decomposes into a tag and an offset depends on the number (and size) of each page of protected memory.
+The page tag identifies in which page of non-volatile memory the memory location resides and the page offset identifies a particular byte of that page.
+
+%How an $a$-bit memory address decomposes into a tag and an offset depends on the number (and size) of each page of protected memory.
 To address {\tt p} pages of size {\tt s}, a tag comprises $log(p)$ bits, the offset comprises $log(s)$ bits and $log(p) + log(s) = a$.
-For example, in a 16-bit address space (i.e., 16\,kB of protected memory) with 512 pages of 128 bytes each, a
-page's tag is the high-order nine bits of the address and a byte's page offset
-is the remaining 7 bits.
+For example, in a 16-bit address space (i.e., 16\,kB of protected memory) with 512 pages of 128 bytes each, a page's tag is the high-order nine bits of the address and a byte's page offset is the remaining 7 bits.
 
@@
 If an accessed variable's page is equal to {\tt CrntPagTag} then the accessed variable's page is resident in the page buffer.
-If the accessed variable's page tag is not equal to {\tt CrntPagTag} then the variable's page is not resident in the page buffer.
-If the accessed variable's page is resident in the page buffer, an executing memory operation indexes the page buffer using the variable's page offset and manipulates that byte (i.e., either reading or writing).
-If the accessed variable's page is not resident in the page buffer, the operation incurs a {\em page fault}.
-At a page fault, \sys must swap out the
-page that is resident int he page buffer and swap in the the variable's page.
+The executing memory operation indexes the page buffer using the variable's page offset and manipulates that byte (i.e., either reading or writing).
+If the accessed variable's page tag is not equal to {\tt CrntPagTag} then the variable's page is not resident in the page buffer; the operation incurs a {\em page fault}.
+At a page fault, \sys must swap out the page that is resident in the page buffer and swap in the the variable's page.
 
@@
 Algorithm~\ref{algo:pagefault} shows the how \sys handles a page fault.
-If both the page being swapped out and the page being swapped in are not {\em dirty}---which means that they do not contain bytes modified by the executing
-task---then handling a page fault simply requires copying the accessed page into the page buffer.
+If both the page being swapped out and the page being swapped in are not {\em dirty}---which means that they do not contain bytes modified by the executing task---then handling a page fault simply requires copying the accessed page into the page buffer.
 If either the {\em resident} page being swapped out or the {\em accessed} page being swapped in is dirty, \sys must take extra steps when handling a page fault.
@@
 As Algorithm~\ref{algo:pagefault} shows in Line 3, \sys checks if the requested page was previously modified by the executing task.
-To track whether a page is modified, \sys stores a dirty bit with the page that is set when a task
-modifies the page.
+To track whether a page is modified, \sys stores a dirty bit with the page that is set when a task modifies the page.
 If an accessed page's dirty bit is set, the page must be swapped in from \texttt{pagesTemp} rather than its original memory location (Line 4), furnishing the task with the page's modified values.
@@
 The task updates the page, leaving it dirty at the task's completion.
-When the task completes, it first copies the page from the page buffer into the non-volatile {\tt
-pagesTemp} region before commit. Then, \sys sets the commit bit, atomically commits the page back to its original memory location, and unsets the commit bit.
+When the task completes, \sys first copies the page from the page buffer into the non-volatile {\tt pagesTemp} region before commit.
+Then, \sys sets the commit bit, atomically commits the page back to its original memory location, and unsets the commit bit.
 

### 2017-08-10 paper/coalescing.tex
Covering Koen's comments

@@
 The purpose of coalescing is to better amortize the run time overhead of committing updated pages when a task ends.
-If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed only once for the coalesced task rather than twice, one time for each of the tasks executed individually.
+If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed once.
 
-\sys can coalesce an arbitrary number consecutive tasks.
+\sys can coalesce an arbitrary number of consecutive tasks.
 As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance.
@@
 
-\sys determines $C$ for a given operating based on the number of tasks completed during the last operational period.
+\sys determines $C$ for a given operation based on the number of tasks completed during the last operational period.
 \sys begins execution in a coalesced task with an infinite $C$ value.
@@
 
-\noindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
+\noindent\textbf{Task Coalescing Illustration:} 
+%
+%Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
 Figure~\ref{fig:coalescing_example}(a) shows a sequence of six tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.
 Figure~\ref{fig:coalescing_example}(b) shows the same sequence of tasks executing as coalesced groups of three tasks.
-Figure~\ref{fig:coalescing_example}(c) shows how \sys adapts the value of $C$ based on $S$.
+%
+Figure~\ref{fig:coalescing_example} shows how \sys adapts the value of $C$ based on $S$.
 \sys executes 12 tasks in the first operating period, assigning $S$ to six.

### 2017-08-10 paper/evaluation.tex
Covering Koen's comments

@@
-We present the experimental results of \sys performance.
+We present the experimental results of \sys's performance.
 We start with describing the hardware environment (Section~\ref{sec:results_hardware}) followed by the software environment (Section~\ref{sec:results_software}) and we concluded by the evaluation results (Section~\ref{sec:results_evaluation}).
@@
 
-In all experiments we used the following computing hardware: WISP\,5.1~\cite{wisp5,wisp}, i.e.
-build around MSP430FR5969~\cite{wolverine} 64\,kB FRAM microcontroller, and TI's MSP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} (hardware type will be specified per experiment in Section~\ref{sec:results_evaluation}).
-WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet} with TI Code Composer Studio version X~\cite{}\todo{Report on CCS version}{Amjad}.
+In all experiments we used the following computing hardware: WISP\,5.1~\cite{wisp5,wisp} build around MSP430FR5969~\cite{wolverine} 64\,kB FRAM microcontroller.
+WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet} with TI Code Composer Studio version 7.
 FET was also used as a power supply in non-intermittent power experiment.
 To measure energy consumption of all benchmark applications (please refer to Section~\ref{sec:results_software}) Carnegie Mellon's EDB~\cite{edb} was used.
-In experiments involving wireless power, WISP was powered through an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
-RFID reader is controlled by a PC running Ubuntu 10.4 executing open source Python-based \emph{sllurp} library~\cite{sllrp_github} enabling low-level RFID reader control.
+In experiments involving wireless power, WISP was powered through an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
+The RFID reader is controlled by a PC running Ubuntu 10.4 executing open source Python-based \emph{sllurp} library~\cite{sllrp_github}.
 Transmitted power of the RFID reader was 30\,dBm at 915\,MHz center frequency.
-For distance controlled experiments, WISP was positioned at varying length hollow paper tubes placed directly on a flat lying antenna.
-Code completion measurements have been performed using Salea~\cite{saleae} digital logic analyzer.
+For distance-controlled experiments, the WISP was positioned at varying length hollow paper tubes placed directly on a flat lying antenna.
+Code execution times have been measured using Salea~\cite{saleae} digital logic analyzer.
 
-We also had alternative setup where we used MSP-TS430RGZ48C launchpad for continuous power experiment.
+We also used an alternative setup where the MSP-TS430RGZ48C launchpad was for continuous power experiments.
 We used internal timer for measuring run time.
-For harvested energy, we used ThingMagic Astra-EX RFID reader with power level of 17.25-30dBm.
-We used EDB~\cite{edb} for run time measurement.
-Applications were cross compiled using GCC and Clang, so that we can use LLVM for \sys compiler.\todo{Report on LLVM version}{Brandon}
+For harvested energy, we used the ThingMagic Astra-EX RFID reader with power levels in the [17.25,30]\,dBm range.
+We used EDB~\cite{edb} for execution time measurement.
+Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler.\todo{Report on LLVM version}{Brandon}
 

### 2017-08-10 paper/compiler.tex
fixed some points due to red text

@@
 Figure~\ref{fig:comp_2}).
-To simplify inter-procedural analysis, our prototype inlines all function calls\todo{how about recursions and code duplication?}{Brandon/Kiwan} into the main function in the CFG.
+To simplify inter-procedural analysis, our prototype inlines all function calls into the main function in the CFG.
+Recursive function call cannot be handled, so the programmer has to rewrite it into a non-recursive call if there is one.
+The approach of always inlining function calls will introduce code duplication.
+If the code size is problematic, alternative approach is to make each function call a task boundary.
+We leave implementation of such alternative strategy as a future work.
 %
@@
 
-{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of {\em live variables} that are accessed in both\todo{Transitive closure?}{Brandon/Kiwan} of the blocks that define the edge (i.e., predecessor and successor).
-Variables that live across an edge between two tasks must be protected by \sys's memory virtualization mechanism and each live, inter-task variable increases the run time overhead for a boundary placement.
+{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of variable that {\em may} be {\em alive} across the edge (i.e., variables whose value may be consumed later on).
+Variables that live across an edge must be protected by \sys's memory virtualization mechanism and each protected, inter-task variable increases the run time overhead for a boundary placement.
 \sys's tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.

### 2017-08-10 paper/relatedwork.tex
finished the pass over related work and added the VM stuff

--- 2017-08-08 paper/relatedwork.tex
@@
-\noindent \textbf{Intermittently-Powered Devices:} There is plentiful of research results available on energy harvesting, or more specifically wireless power provision, for embedded devices like sensors or microcontrollers.
-This paragraph is too short to cover this area extensively and we refer the reader to overview papers~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016} for in-depth discussion.
-With power consumption of electronics going down dramatically with every year, computing systems can be now powered completely from ambient energy.
-This opens new research areas such as radio frequency-powered computers~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}.
-Indeed, plenty of new embedded platforms appear that are powered solely from ambient (intentional or non-intentional) energy and include~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
-But there is a price tag associated to them: harvested energy is miniscule, unstable, difficult to predict ergo \emph{intermittent}.
-Therefore, to guarantee completion of computing, programming and runtime models are being actively researched that are specifically tailored to these intermittently-powered systems.
-\todo{Revise this section}{Brandon}
+\noindent \textbf{Intermittently-Powered Devices:} There is a large body of
+research on energy harvesting and wirelessly powered embedded devices like
+sensors or microcontrollers. Several recent reviews provide a broad overview
+papers~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}
+in depth.  As computing power consumption continues to decrease, a new wave of
+{\em ambiently} powered systems is emerging, powered by radio
+waves~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}.
+Radio-powered
+devices~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}
+are realted to \sys because they are the technological foundation of an
+important category of intermittent computing devices. The emergence of
+intermittent computing hardware platforms has led to the development of tooling
+and instrumentation that support using these platforms for system
+building~\cite{hester_sensys_2014,hester_sensys_2015,edb,stork,wisent}.
 
-\noindent \textbf{Intermittent Execution Environments:} Checkpoint runtimes include DINO~\cite{dino}, Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus}.
-The only available Task-based intermittent runtime systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}.
-\sys goes beyond what any task-based system performed by providing \emph{automatic} task creation and \emph{adaptivity} to energy state and energy storage \emph{obliviousness} of the compiled code.
-\todo{complete this section}{Przemek}
+\noindent \textbf{Intermittent Execution Environments:} 
+The proliferation of energy-harvesting hardware led to a need for system
+support.  Early work developed runtime environments for energy-harvesting
+systems assuming simple, small computations that do not exceed a predictable
+burst of energy~\cite{dewdrop} or long power-cycles and priority-based
+scheduling to ensure important tasks execute often
+enough~\cite{sorber_sensys_2007}.  These early efforts did not consider
+intermittence or data consistency issues. 
+%
+Later efforts supported intermittent computing using dynamic, on-demand
+checkpoints include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall},
+Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus} and
+HarvOS~\cite{mottola2017harvos}. Dynamic checkpointing systems may checkpoint
+at any point, making it difficult to implement application-level atomicity
+requirements, such as collecting correlated sensor samples together.  Moreover,
+some of these systems require hardware support for a device to measure its
+stored energy, which requires additional hardware and may incur an energy cost. 
+Moreover, these systems do not checkpoint non-volatile state, compromising 
+memory consistency.
+%
+DINO~\cite{dino} selectively versions non-volatile memory and checkpoints
+volatile state, ensuring progress and memory consistency.
+Rachet~\cite{ratchet} assumes all memory is non-volatile and checkpoints
+volatile state.  Clang~\cite{hicks_isca_2017} takes a Ratchet-like approach
+that leverages dedicated hardware to eliminate checkpointing and versioning
+overhead.
 
-\noindent \textbf{Memory Virtualization:} Review of memory virtualization environments (not necessarily for intermittent power devices). \todo{Fill-in this section}{Przemek}
+Two task-based system developed in recent work are Chain~\cite{chain} and
+Alpaca~\cite{alpaca}.  Using static tasks, these systems eliminate the need to
+checkpoint volatile state.  Using novel, channel-based memory
+models~\cite{chain} or automatic privatization and redo-logging~\cite{alpaca}
+these systems avoid checkpointing overheads, do not require fully non-volatile
+memory, and require no special hardware.  Moreover, task-based models
+facilitate specifying application-level atomicity properties.
+
+\sys relates to earlier efforts on intermittent computing in their shared
+purpose of making it simple to write correct programs that run intermittently,
+but is distinct in its approach and mechanism.  \sys relates more closely to
+task-based intermittent programming execution models because \sys also relies
+on statically defined tasks to avoid the need to checkpoint volatile state.
+\sys's automatic compilation spares the programmer the work of writing
+task-based code that prior systems demand.  \sys's memory virtualization
+provides an alternative design for ensuring memory consistency that is
+optimized for bulk accesses to task-shared data with high locality.  \sys's
+task coalescing optimization ameliorates the key overheads associated with
+committing tasks present in these prior systems.
+
+{\noindent \bf Memory Virtualization}
+
+Prior work on operating system support for embedded systems and sensor motes
+has studied a variety of memory virtualization strategies that relate to \sys's
+memory virtualization mechanism.
+
+Several operating system efforts developed support for memory management.
+TinyOS and nesC support dynamic memory management~\cite{nesc}.  Later work
+extended the memory manager to support memory virtualization backed by flash
+memory~\cite{sensornetvm} and to ensure memory and type
+safety~\cite{tinyosmemorysafety}. SOS~\cite{sos}, Contiki~\cite{contiki}, and
+t-kernel~\cite{tkernel} also developed memory management abstractions that
+virtualize memory size and provide safe and indirect access.
+Mat\'e~\cite{mate} developed full virtual machines support for sensor nodes,
+virtualizing not just memory resources, but other state and peripheral
+resources.  These prior efforts to virtualize memory in embedded and sensor
+node systems are similar in their purpose and mechanism to \sys, but \sys is
+distinct in its goal.  The goal of \sys is to provide consistent, intermittent
+execution, leveraging the performance benefits of efficient bulk copying.  In
+contrast, these prior efforts were focused more on programmability and run time
+reliability properties provided by virtual memory.
+
+Another prior effort that is related to \sys is work on unbounded, page-based
+transactional memory and deterministic parallel runtime
+systems~\cite{pagebasedtm,grace}.  These prior efforts have a different
+mechanism and purpose than \sys -- ensuring that data are consistent and
+deterministically updated during concurrent executions.  What makes these
+efforts related is that they manage state required to ensure consistency at the
+granularity of pages to amortize checking and tracking costs.  \sys also tracks
+accesses at the granularity of a page and uses its paging mechanism to ensure
+consistency.  Moreover, \sys's paging implementation, which keeps a shadow page
+for each page to use during commit is very similar to the shadow paging scheme
+use for transactional commit~\cite{pagebasedtm}.

### 2017-08-10 paper/compiler.tex

@@
 {\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block.
-Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we simply use the number of intermediate representation (IR) instructions in a block as a proxy for energy because energy cost.
+Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we simply use the number of intermediate representation (IR) instructions in a block as a proxy for energy because more instructions mean more energy consumption.
 Moreover, we assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy.
@@
 {\noindent \bf Live variable cost.} Each CFG edge is associated with a set of variable that {\em may} be {\em alive} across the edge (i.e., variables whose value may be consumed later on).
-Variables that live across an edge must be protected by \sys's memory virtualization mechanism and each protected, inter-task variable increases the run time overhead for a boundary placement.
+Variables that live across an edge must be protected by \sys's memory virtualization mechanism when boundary is placed at the edge and the resulting inter-task variable increases the run time overhead.
 \sys's tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.
@@
 {\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis must account for the cost of multiple iterations of the loop body. \sys assumes that most loops are
-annotated with a static loop bound. For loops that do not have a static bound, \sys conservatively assumes the maximum possible loop bound, based on the type of the loop induction variable.
+annotated with a static loop bound. For loops that do not have a static bound, \sys conservatively assumes the loop will be too large to fit within a single task.
 
@@
 %
-where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of variables live variables at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate.
+where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of protected variables that will be newly introduced by placing boundary at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate.
 \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates.

### 2017-08-10 paper/evaluation.tex

@@
 We used internal timer for measuring run time.
-For harvested energy, we used the ThingMagic Astra-EX RFID reader with power levels in the [17.25,30]\,dBm range.
+For harvested energy, we used the ThingMagic Astra-EX RFID reader with power levels of 30\,dBm.
 We used EDB~\cite{edb} for execution time measurement.
-Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler.\todo{Report on LLVM version}{Brandon}
+Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler.
+Both for Clang and LLVM we used version 3.8.0.
 
@@
 
-Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile DFT application since it was using built-ins for C math functions (i.e.
-\texttt{rand()}, \texttt{sinf()}, \texttt{cosf()}) that were not provided by the Clang MSP430 back-end.}.
+Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile DFT application since the LLVM inserts mathematical built-in functions (i.e.
+\texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}.
+We compiled the apps on a 4-core virtual machine with 1GB memory that has Arch Linux, running on Intel Core i5-6200U (2.30GHz) processor.
 The result are given in Table~\ref{table:compile_time}.
-Most of the compile time was due to pointer aliasing.\todo{Comment on: whether these times are good or bad; why there is such a difference in time between all applications; how many runs you performed; what PC/OS you used to run the code; how the result was measured; why results are rounded (there are no decimals in your result); what is "Clang MSP430 back-end"}{Kiwan}
+We simply used {\em date} command provided by linux to measure the rough time.
+Most of the compile time was due to pointer aliasing, when calculating possible live range of each variables.
+Apps such as RSA whose code size is large and uses a lot of variables needed longer compilation time due to larger search space for pointer aliasing.
 
@@
 		\hline\hline
-		3 & 2 & 6 & 86 & 34 & > 1 \\
+		3 & 2 & 6 & 86 & 34 & < 1 \\
 		\hline
 	\end{tabular}
-	\caption{\sys compile time per application (seconds).\todo{for Sort:">" or "<"?}{Kiwan}}
+	\caption{\sys compile time per application (seconds).}
 \label{table:compile_time}

### 2017-08-10 paper/coalescing.tex
coalescing figures added

@@
 As more tasks coalesce, the amount of potentially {\em wasted work} also increases.
-If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of
-the tasks in the long sequence.
+If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of the tasks in the long sequence.
 \sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, \emph{adaptively} determining a coalescing factor that combines many tasks without suffering an excess of wasted work.
@@
 
-\sys determines $C$ for a given operation based on the number of tasks completed during the last operational period.
+\sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the last operational period.
 \sys begins execution in a coalesced task with an infinite $C$ value.
@@
 On intermittent power, the system executes a sequence of tasks that is interrupted by a power failure.
-\sys explicitly tracks $S$, the number of tasks completed during the previous operating period.
+\sys explicitly tracks $S$ during the previous operating period.
 \sys assigns $C$ to be equal to half of the value of $S$.
@@
 	\centering
-	%\subfloat[]{\includegraphics[width=\columnwidth]{figures/}\label{fig:}}\\
-	%\subfloat[]{\includegraphics[width=\columnwidth]{figures/}\label{fig:}}
+	\subfloat[Core of task coalescing: commit postponing]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
+	\subfloat[Adaptive task coalescing based on task failure history]{\includegraphics[width=\columnwidth]{figures/coalescing_flowchart}\label{fig:coalescing_example_flowchart}}
 	\caption{\sys task coalescing process.}
-	\label{ffig:coalescing_example}
+	\label{fig:coalescing_example}
 \end{figure}
 
-\noindent\textbf{Task Coalescing Illustration:} 
-%
-%Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
-Figure~\ref{fig:coalescing_example}(a) shows a sequence of six tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.
-Figure~\ref{fig:coalescing_example}(b) shows the same sequence of tasks executing as coalesced groups of three tasks.
-%
-Figure~\ref{fig:coalescing_example} shows how \sys adapts the value of $C$ based on $S$.
-\sys executes 12 tasks in the first operating period, assigning $S$ to six.
-\sys completes only five of the six coalesced tasks in the next operating period and then power fails.
-\sys sets $S$ to two in the next operating period, coalescing pairs of tasks and successfully committing two pairs before power fails.
+\noindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
+Figure~\ref{fig:coalescing_example_states} (left) shows a sequence of three tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.
+The same figure shows the same sequence of tasks executing as coalesced groups of three tasks $C=2$ (center) and $C=3$ (right).
+Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$.
+\sys executes eight tasks in the first operating period, assigning $C=4$.
+\sys completes only two of the four coalesced tasks in the next operating period and then power fails.
+\sys sets $C=(6+8+2)/2$ in the next operating period, coalescing pairs of tasks and successfully committing two pairs before power fails.
 Execution resumes at the start of the third pair of tasks with a coalescing factor of two.
-If a larger number of tasks successfully complete in the next operating period, i.e., $S$ is larger, $C$ correspondingly increases, allowing more aggressive coalescing.
+If a larger number of tasks successfully complete in the next operating period, i.e., $S$ is larger, $C$ correspondingly increases, allowing more aggressive coalescing.\todo{Rewrite this paragraph to reflect what is in the figure}{Przemek}
 

### 2017-08-10 paper/evaluation.tex
commented out section about batery-less FFT

@@
 
-\subsubsection{Case Study}
-\label{sec:case_study}
+%\subsubsection{Case Study}
+%\label{sec:case_study}
 
-The final experiment assessing usefulness of \sys's use in real application.
+%The final experiment assessing usefulness of \sys's use in real application.
 For this we have a implemented an battery-less wirelessly-powered sound detector as an example.
@@
 
-The application aims at detecting a specific tone in the audio signal (in the implementation: 1\,kHz) based on on-the-fly audio measurement.
+%The application aims at detecting a specific tone in the audio signal (in the implementation: 1\,kHz) based on on-the-fly audio measurement.
 For this we have connected an analog MEMS microphone~\cite{microphone} output to AUX3 port of WISP\,5.1 (while the microphone was powered directly by WISP, which was then wirelessly power-supplied by the RFID reader).
@@
 
-Expected results:
-\begin{itemize}
-	\item Compare execution time: \sys against Chain for X\,s long audio signal detection at two antenna distances
-	\item Make a video of the demonstration 
-\end{itemize}
+%Expected results:
+%\begin{itemize}
+%	\item Compare execution time: \sys against Chain for X\,s long audio signal detection at two antenna distances
+%	\item Make a video of the demonstration 
+%\end{itemize}
 
-\todo{Check if we really need this application}{Brandon}
-\todo{Provide results and text to this section---if time allows (this result is not critical)}{Amjad/Przemek}
+%\todo{Check if we really need this application}{Brandon}
+%\todo{Provide results and text to this section---if time allows (this result is not critical)}{Amjad/Przemek}

### 2017-08-10 paper/evaluation.tex
Made placeholders for figures in evaluation section

@@
 
-\vspace{2cm}
+%\vspace{2cm}
 
-List of figures: 
-\begin{itemize}
-	\item Fig. 1: Viper only/per app/fixed power/(hand/complier)/no coalescing/clang [Resuls are ready by Kiwan]
-	\item Fig. 2: Viper only/per app/per distance (fixed/20/40/60)/(hand/complier)/coalescing [Resuls are ready by Kiwan]
-	\item Fig. 3a: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/clang [Resuls are ready by Kiwan]
-	\item Fig. 3b: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/gcc [Resuls are not ready: Kiwan needs to make them]
-	\item Fig. 4: exactly as Fig 12 in the paper (4 apps only)
-	\item Fig. 5: average virtual task size/gcc/fixed distance/per app 
-\end{itemize}
+%List of figures: 
+%\begin{itemize}
+%	\item Fig. 1: Viper only/per app/fixed power/(hand/complier)/no coalescing/clang [Resuls are ready by Kiwan]
+%	\item Fig. 2: Viper only/per app/per distance (fixed/20/40/60)/(hand/complier)/coalescing [Resuls are ready by Kiwan]
+%	\item Fig. 3a: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/clang [Resuls are ready by Kiwan]
+%	\item Fig. 3b: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/gcc [Resuls are not ready: Kiwan needs to make them]
+%	\item Fig. 4: exactly as Fig 12 in the paper (4 apps only)
+%	\item Fig. 5: average virtual task size/gcc/fixed distance/per app 
+%\end{itemize}
 
-\vspace{2cm}
+%\vspace{2cm}
+
+\subsubsection{\sys Performance}
+\label{sec:result_coalescing}
 
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/virtualTaskSize.eps}
-	\caption{Execution time of a dummy application containing twelve empty tasks as a function of coalescing factor.
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\columnwidth]{figures/virtualTaskSize.eps}
+%	\caption{Execution time of a dummy application containing twelve empty tasks as a function of coalescing factor.
 We observe three times improvement from coalescing until four virtual tasks, while we observe a diminishing return from task coalescing beyond four.
 \todo{Improve the figure: make it narrower, larger font, remove legend, do not capitalize "Task" in X axis, add Y axis label (Execution time (ms))}{Amjad}}
-	\label{fig:virtualTaskSize}
-\end{figure}
+%	\label{fig:virtualTaskSize}
+%\end{figure}
 
-The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e.
+%The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e.
 from energy harvesting to continuous power environment) is the best one?
@@
 \todo{Decide whether to place this figure in Evaluation section or leave it here}{Przemek}
+
+\subsubsection{\sys Compiler Overhead}
+\label{sec:result_compiler time}
+
+\begin{figure}
+	\centering
+	%\includegraphics[width=\columnwidth]{figures/}
+	\caption{\sys compiler overhead.}
+	\label{fig:coalescing}
+\end{figure}
+
+\subsubsection{Virtual Task Size}
+\label{sec:result_task_size}
 
@@
 
-Expected results:
-\begin{itemize}
-	\item performance for each application comparing bars for Chain, no coalescing/hand-tasked, coalescing/hand-tasked, no-coalescing/compiler, coalescing/compiler.
+%\subsubsection{Completion Time}
+%\label{sec:result_completion}
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\columnwidth]{figures/chain_vipe}
+%	\caption{Time to complete for a set of testing applications (refer to Table~\ref{table:benchmark_list}) using \sys versus Chain (no coalescing applied, fixed power supply).
+\todo{Figure placeholder---will be replaced with final one}{Amjad}}
+%	\label{fig:IPOSPerformance}
+%\end{figure}
+
+%Expected results:
+%\begin{itemize}
+%	\item performance for each application comparing bars for Chain, no coalescing/hand-tasked, coalescing/hand-tasked, no-coalescing/compiler, coalescing/compiler.
 Performance depicted, normalized to one of the configurations (probably Chain).
 This plot will cover most of the eval dimensions that we care about.
-	\item Sensitivity plot---show amount of task merged, time to completion, overhead (in terms of task-size/cap-size sweep (remember we can vary either with approximately the same effect)), energy consumed.
+%	\item Sensitivity plot---show amount of task merged, time to completion, overhead (in terms of task-size/cap-size sweep (remember we can vary either with approximately the same effect)), energy consumed.
 All experiments for fixed power, at 3 distances with wireless power, compared against Chain for 3 task/cap size configurations
-\end{itemize}
+%\end{itemize}
 
@@
 
-\todo{Provide results and text to this section}{Amjad}
+%\todo{Provide results and text to this section}{Amjad}
 
@@
 \label{sec:results_memory_management}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/chain_vipe}
-	\caption{Time to complete for a set of testing applications (refer to Table~\ref{table:benchmark_list}) using \sys versus Chain (no coalescing applied, fixed power supply).
-\todo{Figure placeholder---will be replaced with final one}{Amjad}}
-	\label{fig:IPOSPerformance}
-\end{figure}
 

### 2017-08-11 paper/intro.tex
patched intro
pass to tighten intro

@@
 Advances in processor efficiency along with the development of energy-harvesting power systems has created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
-These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaic components~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
-Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such energy-independent devices are a promising underlying technology for applications in emerging areas such as the Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
+These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaics~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
+Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such devices are a promising technology for use in emerging areas such as the Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
 
-While promising, energy-harvesting devices present a unique design challenge in that they operate only {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}. An
-energy-harvesting device buffers energy from its environment, e.g., in a capacitor~\cite[Fig.
-3]{gorlatova_tmc_2013},~\cite[Fig.
-1]{gunduz_commag_2014}.
-When a threshold quantity of energy accumulates, the device begins operating.
-Harvestable energy sources typically produce little power compared to a platform's operating power level.
-A device operates for a brief period of time and when buffered energy is depleted, shuts down and begins recharging to operate again later.
-As an example, inter-task time may even exceed 20\,s for a wireless energy-harvesting battery-less vivo digestive tract monitoring system~\cite[Fig.
+Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}. An
+energy-harvesting device buffers energy in a capacitor~\cite{gorlatova_tmc_2013},~\cite{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating.
+Harvestable energy sources are low-power compared to a platform's operating level.
+A device operates briefly and when buffered energy is depleted, shuts down and recharges to operate again later.
+As an example, recharge time may be tens of seconds in RF powered medical device~\cite[Fig.
 3c]{nadeau_naturebio_2017}.
-On top of that, charge and discharge times vary by device (different energy buffer sizes) and some of them~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite[Fig.
-1]{tan_infocom_2016},~\cite[Fig.
-1]{mementos},~\cite[Fig.
-3]{nvp}.
+Charge and discharge times vary by device (different capacitor sizes) and some~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite{tan_infocom_2016},~\cite{mementos},~\cite{nvp}.
 
@@
 
-Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with bursts of execution interrupted by failure periods.
-When power fails, a device loses its volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
-While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help to preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{mspcdino,edb}.
+Software in an energy-harvesting system operates in the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with bursts of execution interrupted by failure periods.
+When power fails, a device loses volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
+While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{dino,edb}.
 
@@
 mechanisms~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
-While both approaches represent progress, each has critical limitations.
-New architectures require costly hardware changes and are inapplicable to today's embedded off-the-shelf systems~\cite[Fig.
-3]{hicks_isca_2017}~\cite[Fig.
-9]{nvp}.
+While both approaches represent progress, each has limitations.
+New architectures require costly hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}.
 New programming models and compilers require programmers to change behavior~\cite{chain,ratchet}, inhibiting adoption.
 New memory models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
-Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite[Fig.
-4]{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
-Hardware is effective, but inapplicable to existing devices, making programming and execution models essential to the success of intermittent computing.
+Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
 
-Recent work~\cite{alpaca,chain} proposed {\em task-based} programming and execution models in which programmers statically decompose their program into a collection of tasks.
+Recent work~\cite{alpaca,chain} proposed {\em task-based} programming and execution models in which programmers statically decompose a program into a collection of tasks.
 Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
-The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent~\cite[Fig.
-4]{chain}.
-An important programming challenge presented by a task-based model is that the length of a software task's execution is limited by the fixed, total amount of energy that a device can buffer in hardware.
+The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent~\cite{chain}.
+An key programming challenge in a task-based model is that the length of a software task's execution is limited by the fixed, total amount of energy that a device can buffer in hardware.
 A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict.
@@
 
-The key research question addressed by this work is to determine how to define and execute tasks efficiently and ensures non-termination, given an unknown energy buffer size and energy arrival process.
+The key research question addressed by this work is to determine how to define and execute tasks efficiently and ensure non-termination, with an unknown energy buffer size and energy arrival process.
 In particular, we ask (i) how to use software support to dynamically adapt the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while automatically maintaining memory consistency during execution?
@@
 The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent.
-\sys complier effectively circumvents the need for labour-intensive manual task generation of the state-of-the-art solutions~\cite{chain,alpaca} (spanning hours of programmer work) completing task compilation in the order of seconds.
+\sys's complier the labour-intensive process of writing tasks~\cite{chain,alpaca} (i.e., hours of programmer time).
 	%
-	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism executes multiple statically-derived tasks together as a single dynamic task.
+	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks together as a single task.
 A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer.
-Coalescing tasks allows running the same tasks on a device with a larger energy buffer, while avoiding run time overheads associated with ending one task and beginning another.\todo{Provide numerical results for this section}{Przemek}
+Coalescing tasks allows running the same tasks on a device with a larger energy buffer, while avoiding run time overheads associated with ending one task and beginning another.%\todo{Provide numerical results for this section}{Przemek}
 	%
-	\item {\bf Volatile memory virtualization:} \sys's memory virtualization mechanism uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand during a task's execution.
+	\item {\bf Volatile memory virtualization:} \sys's memory virtualization mechanism uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution.
 When a task ends, \sys commits dirty pages to FRAM using efficient DMA block copies, ensuring task atomicity and memory consistency.
-\todo{Once paging is fully implemented--update the paragraph}{Przemek}\todo{Provide numerical results for this section}{Przemek}
+%\todo{Once paging is fully implemented--update the paragraph}{Przemek}\todo{Provide numerical results for this section}{Przemek}
 \end{itemize}
 
-We evaluated \sys end-to-end on a collection of existing benchmark programs~\cite[Sec.
-5]{chain} and new ones in a laboratory environment, and compare directly to state-of-the art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without recompiling the code.
-\todo{add detailed text about evaluation results}{Przemek}
+We evaluated \sys end-to-end on a collection of existing benchmarks ~\cite{chain} and new programs, and compare directly to a state-of-the-art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without rewriting code.
+%\todo{add detailed text about evaluation results}{Przemek}
 

### 2017-08-11 paper/relatedwork.tex
Related work compressed/ready

--- 2017-08-08 paper/relatedwork.tex
@@
-\noindent \textbf{Intermittently-Powered Devices:} There is plentiful of research results available on energy harvesting, or more specifically wireless power provision, for embedded devices like sensors or microcontrollers.
-This paragraph is too short to cover this area extensively and we refer the reader to overview papers~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016} for in-depth discussion.
-With power consumption of electronics going down dramatically with every year, computing systems can be now powered completely from ambient energy.
-This opens new research areas such as radio frequency-powered computers~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}.
-Indeed, plenty of new embedded platforms appear that are powered solely from ambient (intentional or non-intentional) energy and include~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
-But there is a price tag associated to them: harvested energy is miniscule, unstable, difficult to predict ergo \emph{intermittent}.
-Therefore, to guarantee completion of computing, programming and runtime models are being actively researched that are specifically tailored to these intermittently-powered systems.
-\todo{Revise this section}{Brandon}
+{\noindent \textbf{Intermittently-Powered Devices:}} There is a large body of research on energy harvesting and wirelessly powered embedded devices.
+Several recent reviews provide a broad in-depth overview~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
+As computing power consumption continues to decrease, a new wave of {\em ambiently} powered systems is emerging, powered by e.g.
+radio waves~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}.
+Radio-powered devices~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}
+are related to \sys---they are the technological foundation of an
+important category of intermittent computing devices. The emergence of
+intermittent computing hardware platforms has led to the development of tooling
+and instrumentation for such systems
+building~\cite{hester_sensys_2014,hester_sensys_2015,edb,stork,wisent}.
 
-\noindent \textbf{Intermittent Execution Environments:} Earlier work include Dewdrop~\cite{dewdrop} (splitting task such that they fill-in the capacitor).
-Similar, specific-application tailored system includes Wisent and Stork~\cite{stork,wisent}.
-Checkpoint runtimes include DINO~\cite{dino}, Quickrecall~\cite{quickrecall}, Hibernus/Hibernus++~\cite{hibernus,hibernusplusplus} are checkpointing-based systems requiring hardware support to measure remaining storage energy---a feature prohibitive given typical high energy cost expenditure of such hardware.
-The only available Task-based intermittent runtime systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}.
-\sys goes beyond what any task-based system performed by providing \emph{automatic} task creation and \emph{adaptivity} to energy state and energy storage \emph{obliviousness} of the compiled code.
-\todo{complete this section}{Przemek}
+{\noindent \textbf{Intermittent Execution Environments:}} Early work on Energy-harvesting runtime systems assumed simple, small computations not exceeding a predictable burst of energy~\cite{dewdrop} or long power-cycles/priority-based scheduling~\cite{sorber_sensys_2007}.
+These efforts did not consider data consistency with intermittence.
+Works supporting intermittent computing using dynamic, on-demand checkpoints for consistency include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall}, Hibernus++~\cite{hibernusplusplus} and HarvOS~\cite{mottola2017harvos}.
+Dynamic checkpointing systems may checkpoint
+at any point, making it difficult to implement application-level atomicity
+requirements.
+Moreover, some systems require hardware to measure stored energy, incurring energy cost.
+Moreover, these systems do not checkpoint non-volatile state, compromising memory consistency.
+DINO~\cite{dino} selectively versions non-volatile memory and checkpoints volatile state, ensuring progress and memory consistency.
+Rachet~\cite{ratchet} assumes all memory is non-volatile and checkpoints volatile state.
+Clang~\cite{hicks_isca_2017} takes a Ratchet-like approach leveraging special hardware, eliminating checkpointing and versioning overhead.
 
-\noindent \textbf{Memory Virtualization:} Review of memory virtualization environments (not necessarily for intermittent power devices). \todo{Fill-in this section}{Przemek}
+Recent task-based systems are Chain~\cite{chain} and
+Alpaca~\cite{alpaca}.
+Using static tasks, they eliminate checkpointing volatile state.
+Using channel-based memory models~\cite{chain} or automatic privatization and redo-logging~\cite{alpaca} they avoid checkpointing overheads.
+Moreover, task-based models facilitate specifying application-level atomicity properties.
+
+\sys relates to the above efforts but is distinct in approach and mechanism.
+\sys relates closely to task-based models because \sys also relies on statically defined tasks to avoid checkpointing volatile state.
+\sys's automatic compilation spares the programmer's work of writing task-based code that prior systems demand.
+\sys's memory virtualization provides an alternative for ensuring memory consistency, optimized for bulk accesses to task-shared data with high locality.
+\sys's task coalescing optimization ameliorates the key overheads associated with prior systems.
+
+{{\noindent \bf Memory Virtualization:}} Prior work on embedded systems has studied a variety of memory virtualization strategies that relate to \sys.
+TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory management.
+Later work extended the memory manager to support memory virtualization backed by flash memory~\cite{sensornetvm} and to ensure memory and type safety~\cite{tinyosmemorysafety}.
+SOS~\cite{sos}, Contiki~\cite{contiki}, and
+t-kernel~\cite{tkernel} also developed memory management abstractions that
+virtualize memory size and provide safe and indirect access.
+Mat\'e~\cite{mate} developed full virtual machines support for sensor nodes, virtualizing not just memory resources, but other states and peripherals.
+The goal of \sys is to provide consistent, intermittent execution, leveraging the benefits of efficient bulk copying.
+In contrast, these prior efforts were focused more on programmability and runtime reliability properties provided by virtual memory.
+
+Another \sys-related prior effort is on unbounded, page-based transactional memory and deterministic parallel runtime systems~\cite{pagebasedtm,grace}.
+These works have a different mechanism and purpose than \sys---ensuring that data are consistent and
+deterministically updated during concurrent executions. What makes these
+efforts related is that they manage state to ensure consistency at the
+granularity of pages to amortize checking and tracking costs. \sys also tracks
+accesses at the granularity of a page and uses its paging mechanism to ensure
+consistency. Moreover, \sys's paging implementation, which keeps a shadow page
+for each page to use during commit, is similar to the shadow paging scheme
+use for transactional commit~\cite{pagebasedtm}.

### 2017-08-11 paper/conclusions.tex
conclusions compressed

--- 2017-07-27 paper/relatedwork.tex
@@
-\section{Intermittenly-Powered Devices}
+This paper presented \sys: a new task-based runtime for intermittently-powered devices that does not require any hardware support.
+\sys core new features are: (i) automatic task-generation---removing programmer intervention in the program re-composition to the intermittent domain \emph{completely}, (ii) memory virtualization---improving the execution speeds of broad set of benchmarks executed on popular energy-harvesting embedded platform, and (iii) (adaptive) task coalescing---enabling fast execution despite changing energy arrivals and enabling re-use of the same code on intermittent systems with varying energy reservoirs.
+\sys, with its task creation/execution automation, responses to the need that all existing predecessors of \sys advocated.
 
-\section{Intermittent Execution Environments}
+%We believe that \sys is a next major step in enabling computing community to enter the intermittent-power domain without big learning investments.
+
+%Next step in \sys design is the introduction of features needed to implement first fully-operational operating system for transiently-powered devices.

### 2017-08-11 paper/abstract.tex
abstract compressed/revised

@@
-Enabling battery-free devices is a mandatory step towards an environment-friendly Internet of Things (IoT).
-However, removing the batteries requires IoT to operate on a harvested, ergo intermittent, power supply.
-This makes sustaining long computations challenging.
-Basically, there are two approaches: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, and (ii) task-based approach, where a programmer splits the code into small idempotent sections.
+Computation consistency on intermittently-powered devices  can be ensured by: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) via tasks, where a programmer splits the code into small idempotent sections.
 Results so far suggest that a task-based approach performs better than checkpointing.
-However, the use of tasks requires that the energy to execute any task must not exceed the capacity of the energy buffer.
-This makes the code inefficient (when executed on a larger energy buffer than intended) and potentially completely disables the program (in the opposite case).
+However, existing tasks-based approaches do not adapt to changing energy levels and storage.
+This makes the code inefficient (when executed on a larger energy buffer than intended) or can potentially completely disable the program (in the opposite case).
 Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing and automatic compiler-supported task generation.
-If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve the performance.
-\todo{Re-write the abstract once the paper is ready}{Brandon}
+If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve execution times.

### 2017-08-11 paper/intro.tex
intro revise/shortened

@@
-Advances in processor efficiency along with the development of energy-harvesting power systems has created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
-These devices operate entirely using energy extracted from their environment, such as RF energy~\cite{rf_powered_computing_gollakota_2014}, photovoltaics~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
-Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such devices are a promising technology for use in emerging areas such as the Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
+Advances in processor efficiency along with the development of energy-harvesting systems has created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
+These devices operate using energy extracted from their environment, such as RF~\cite{rf_powered_computing_gollakota_2014}, photovoltaics~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
+Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such devices are a promising technology for use in the Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
 
@@
 
-%---compare for instance~\cite[Table III and V]{prasad_comst_2014} with~\cite[Table I]{carrano_cst_2014}
-
-Software in an energy-harvesting system operates in the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with bursts of execution interrupted by failure periods.
+Software in an energy-harvesting system operates in the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with execution interrupted by failure periods.
 When power fails, a device loses volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
@@
 
-There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) programming and execution models~\cite{dino,ratchet,chain,alpaca} and (ii) architecture
-mechanisms~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
+There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) programming and execution models~\cite{dino,ratchet,chain,alpaca} and (ii) architectures~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
 While both approaches represent progress, each has limitations.
-New architectures require costly hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}.
+New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}.
 New programming models and compilers require programmers to change behavior~\cite{chain,ratchet}, inhibiting adoption.
@@
 The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent~\cite{chain}.
-An key programming challenge in a task-based model is that the length of a software task's execution is limited by the fixed, total amount of energy that a device can buffer in hardware.
+An key challenge is that the length of a software task's execution is limited by the fixed, total amount of energy that a device can buffer in hardware.
 A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict.
-Assuming that input power is negligible compared to operating power~\cite{wisp}, a task will never be able to complete if its execution consumes more energy than the system can buffer, creating the potential for long tasks to be irreparably non-terminating.
+Assuming that input power is negligible compared to operating power~\cite{wisp}, a task will never be able to complete if its execution consumes more energy than the system can buffer.
 On the other hand, transitioning from one task to another imposes a run time cost and excessively small tasks are inefficient.
 
-The key research question addressed by this work is to determine how to define and execute tasks efficiently and ensure non-termination, with an unknown energy buffer size and energy arrival process.
-In particular, we ask (i) how to use software support to dynamically adapt the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while automatically maintaining memory consistency during execution?
+The key research question addressed here is how to define and execute tasks efficiently and ensure non-termination, with an unknown energy buffer size and energy arrival process.
+In particular, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?
 
-In this work we develop {\bf \sys}: a new programming and execution model that efficiently executes tasks and avoids non-termination across a range of energy buffer sizes, without burdening the programmer.
-To accomplish this goal, \sys introduces three new capabilities:
+In this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer.
+To accomplish this, \sys introduces new capabilities:
 
@@
 	%
-	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks together as a single task.
+	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task.
 A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer.
-Coalescing tasks allows running the same tasks on a device with a larger energy buffer, while avoiding run time overheads associated with ending one task and beginning another.%\todo{Provide numerical results for this section}{Przemek}
+Coalescing tasks allows running the same tasks on a device with a larger energy buffer, avoiding run time overheads associated with ending one task and beginning another.
 	%
-	\item {\bf Volatile memory virtualization:} \sys's memory virtualization mechanism uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution.
-When a task ends, \sys commits dirty pages to FRAM using efficient DMA block copies, ensuring task atomicity and memory consistency.
-%\todo{Once paging is fully implemented--update the paragraph}{Przemek}\todo{Provide numerical results for this section}{Przemek}
+	\item {\bf Volatile memory virtualization:} \sys's memory virtualization uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution.
+When a task ends, \sys commits dirty pages to FRAM using DMA block copies, ensuring task atomicity and memory consistency.
 \end{itemize}

### 2017-08-11 paper/background.tex
Bacground compressed

@@
-Here, we provide background on energy harvesting systems and the intermittent software execution model.
+%Here, we provide background on energy harvesting systems and the intermittent software execution model.
 We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys addresses.
@@
 
-Energy harvesting devices operate using energy extracted from their environment, from sources such as radio waves and solar energy.
+Energy harvesting devices operate using energy extracted from sources such as RF and solar energy.
 Energy-harvesting devices do not use a tethered power supply or battery and instead, typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.
@@
 
-\subsection{Intermittent Computing}
+\subsection{Energy Harvesting and Intermittent Computation}
 \label{sec:background_consistency}
 
-Energy-harvesting devices execute software according to the {\em intermittent execution model}~\cite{dino,chain,alpaca,ratchet}.
+%Energy-harvesting devices execute software according to the {\em intermittent execution model}.
+Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
 %
-%Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
-%
-Software running on the device executes {\em intermittently} because energy is only available between when the capacitor reaches its threshold and when it is depleted.
-An intermittent execution is composed of operating periods interspersed with power failures.
-The frequency of power failures in an intermittent execution depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods.
+Software running on energy harvesting device executes {\em intermittently} because energy is only available between when the capacitor reaches its threshold and when it is depleted.
+An intermittent execution is composed of periods interspersed with power failures~\cite{dino,chain,alpaca,ratchet}.
+The frequency of power failures depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods (but requires longer charge times).
 Energy-harvesters provide input power that is orders of magnitude less than a device's operating power, making recharging negligible during operation.
 
-Software has different behavior when executed intermittently than when executed with continuous power.
+Software behaves uniquely when executed intermittently.
 Each power failure clears volatile memory (registers, stack, and global variables).
@@
 By default, after a power failure control flows to the beginning of {\tt main()}.
-Early work on intermittent computing preserved intermittent progress by periodically capturing and restoring to a non-volatile checkpoint of the volatile execution context, i.e., registers and stack~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
+Early architectures preserved intermittent progress by periodically capturing and restoring to a non-volatile checkpoint of the volatile execution context, i.e., registers and stack~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
 
@@
 	\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
-	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 10) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
+	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
 	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
@@
 The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
-Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of a WAR dependence between the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
+%
+%Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of a WAR dependence between the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
 Figure~\ref{fig:crc_example_war} shows an execution of the code in which power fails during the second loop iteration.
 The problem arises because line 6 of the execution writes $\texttt{Rmnd}$ in the second iteration, but power fails before capturing the next checkpoint.
-After restarting from the checkpoint, the update to $\texttt{data}$ at line 10 reads the value of $\texttt{Rmnd}$ that was updated before the power failure, producing an incorrect result.
-
+After restarting from the checkpoint, the update to $\texttt{data}$ at line 9 reads the value of $\texttt{Rmnd}$ that was updated before the power failure, producing an incorrect result.
+%
 As the figure illustrates, checkpoint-based systems risk violating the volatile and non-volatile memory consistency~\cite{dino}.
@@
 
-\subsection{Task-based Intermittent Computing}
+\subsection{Task-based Intermittent Computation}
 \label{section:background_task_computing}
 
-Task-based intermittent execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into a collection tasks.
+Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into a collection tasks.
 A task is a function with no caller that contains arbitrary computation, sensing, and communication.
@@
 Task-based programming abstractions guarantee that tasks exhibit behavior as though they executed {\em atomically}, regardless of arbitrary power failures.
-Task-based runtime systems tolerate failures by repeatedly attempting to execute a task until it completes.
-Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent (i.e., have the same effect when repeatedly executed).
+Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent.
 The key to idempotence is ensuring that updates to non-volatile state made by an interrupted, partial execution of a task are never visible to a subsequent (potentially completing) re-execution of the task.
@@
 	\end{tabular}
-	\caption{Memory consistency enforcement overheads of various intermittent execution models.
+	\caption{Memory consistency enforcement overheads of intermittent execution models.
 \emph{Reg.} includes the entire register file, \emph{PC} is the program counter, \emph{Channel data} are variables explicitly task-shared by the programmer, and \emph{WAR variables} are variables involved in WAR dependences (\emph{NV}: non-volatile).}
@@
 
-Task-based intermittent execution models ensure that programs execute correctly despite arbitrary power failures and in doing so, incur several costs that are the motivation for \sys.
+Despite valuable features, task-based models incur costs that are the motivation for \sys.
 
@@
 	\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
-	\caption{Energy cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
+	\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
 	\label{fig:framEnergy}
@@
 
-To concretely understand the cost of these manipulations of non-volatile memory, we directly measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations for the MSP430FR5969~\cite{msp430datasheet} microcontroller.
+To concretely understand the cost of these manipulations of non-volatile memory, we directly measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations for the MSP430FR5969~\cite{msp430datasheet}.
 We used EDB~\cite{edb} to monitor the voltage drop on an energy buffering capacitor that powers the device across 1600 read and write accesses to each type of memory, randomizing locations to avoid caching effects.
@@
 
-The above result reassures the intuitive observation that a decrease of checkpoint frequency (CP) decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
+%The above result reassures the intuitive observation that a decrease of checkpoint frequency (CP) decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
 On the other hand, this increases task/checkpoint re-execution time.
@@
 	\end{tabular} 
-	\caption{Comparison of {\em default} energy storage sizes for example battery-less platforms. Observe a huge variation in storage capacity.}
+	\caption{{\em Default} energy storage sizes for example battery-less platforms. Observe a huge variation in storage capacity.}
 	%We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.
@@
 
-\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} A programmer must write each task with statically defined boundaries, which makes statically-sized tasks inflexible and non-portable.
-The programmer must ensure statically that a task's energy requirements match the device's energy storage capacity.
-If a programmer writes a task that consumes more energy than the device can buffer, the task will never be able to complete using the buffered energy.
+\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} If a programmer writes a task that consumes more energy than the device can buffer, the task will never be able to complete using the buffered energy.
 The consequence of such tasks is a non-progressing program that repeatedly attempts to re-execute the task of excess energy cost.
@@
 The key problem is that the programmer writes tasks statically for an energy buffer of a fixed size.
-\sys's {\em task coalescing} mechanism, which we will describe in Section~\ref{sec:task_coalescing}, is motivated by the observation that systems with statically-defined tasks may be inflexible and inefficient.
+\sys's {\em task coalescing} mechanism, which we will describe in Section~\ref{sec:task_coalescing}, is motivated by the above observations.
 
@@
 We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
-\sys supports I/O behavior similar to recent intermittent execution models~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.
+\sys supports I/O behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.

### 2017-08-11 paper/texFigCode/crc_example_war.tex
Bacground compressed

--- 2017-07-24 paper/texFigCode/crc_example_war.tex
@@
-\documentclass[border={20pt 5pt 25pt 10pt}]{standalone} % Left Bottom Right Top
+\documentclass[border={20pt 5pt 15pt 10pt}]{standalone} % Left Bottom Right Top
 %\documentclass{standalone}
@@
 \begin{lstlisting}[style = mystyle]
-$\text{\textcolor{red}{checkpoint()}}$; //Iteration 1: correct execution of the loop body
+$\text{\textcolor{red}{checkpoint()}}$; //Correct execution iteration 1 
 data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
 Rmnd = crcTbl[data] ^ (Rmnd << 8);
-
-$\text{\textcolor{red}{checkpoint()}}$; //Iteration 2: break before reminder calculation
+$\text{\textcolor{red}{checkpoint()}}$; //Capture checkpoint before iteration 2
 data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8));
-$\text{\textcolor{red}{\lightning\hspace{0.5cm}\lightning\hspace{0.5cm}\lightning}\hspace{0.5cm}}\textcolor{red}{\circlearrowright}$ //Power interrupt and restart
-
-$\text{\textcolor{red}{checkpoint()}}$; //Iteration 3: wrong/stale reminder used for data
+Rmnd = crcTbl[data] ^ (Rmnd << 8);      
+$\text{\textcolor{red}{\lightning\hspace{0.5cm}\lightning\hspace{0.5cm}\lightning}\hspace{0.5cm}}$ //Power fails after Rmnd written, before checkpoint()
+$\textcolor{red}{\circlearrowright}\text{\textcolor{red}{checkpoint()}}$; //Restart from checkpoint; $\text{\textcolor{red}{Rmnd remains updated}}$
 data = functReflectData(msg[byte]) ^ (Rmnd >> (Wdth - 8)); 
+$\textcolor{red}{\times}\text{\textcolor{red}{ERROR: data updated using incorrect Rmnd value}}$
 $\ldots$

### 2017-08-11 paper/evaluation.tex
Revisions on evaluation section: results added

@@
-We present the experimental results of \sys's performance.
+%We present the experimental results of \sys's performance.
 We start with describing the hardware environment (Section~\ref{sec:results_hardware}) followed by the software environment (Section~\ref{sec:results_software}) and we concluded by the evaluation results (Section~\ref{sec:results_evaluation}).
@@
 		\hline\hline
-		Bitcount & 351 & 588 \\ %59\%
-		Cold chain & 388 & 721 \\ %53\%
-		Cuckoo & 483 & 762 \\ %63\%
-		RSA & 887 & 1233 \\ %71\%
-		DFT & {} & {} \\ %Two resolutions: 4 and 8 Bytes
-		Huffman & {} & {} \\ %Data decompression size: 100 bytes
-		Selection sort & {} & {} \\
+		Bitcount (bc) & 351 & 588 \\ %59\%
+		Cold chain (cem) & 388 & 721 \\ %53\%
+		Cuckoo (cuckoo) & 483 & 762 \\ %63\%
+		RSA (rsa) & 887 & 1233 \\ %71\%
+		Activity (ar) & 483 & 762 \\ %63\%
+		DFT (dft) & {} & {} \\ %Two resolutions: 4 and 8 Bytes
+		Huffman (dd) & {} & {} \\ %Data decompression size: 100 bytes
+		Selection sort (sort) & {} & {} \\
 		\hline
@@
 
-%\vspace{2cm}
-
 %List of figures: 
@@
 
-%\vspace{2cm}
-
 \subsubsection{\sys Performance}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coalescing}
-	\caption{Performance of \sys's  tasks coalescing mechanism compared to virtual task size equal to one (per application) at multiple WISP to RFID reader antenna distances.\todo{Replace "coal" with "coalescing", "No-coal" with "No coalescing"; use markers in addition to color; decide on uniform naming for apps (either all small or capitalized)}{Amjad}}
+	\includegraphics[width=\columnwidth]{figures/coala_chain}
+	\caption{Performance of \sys's fixed power and at multiple WISP to RFID reader antenna distances, \{20, 40, 60\}\,cm, compared against Chain.}
 	\label{fig:coalescing}
@@
 \subsubsection{\sys Compiler Overhead}
-\label{sec:result_compiler time}
+\label{sec:result_compiler_time}
 
@@
 	\centering
-	%\includegraphics[width=\columnwidth]{figures/}
-	\caption{\sys compiler overhead.}
+	\includegraphics[width=\columnwidth]{figures/comp_user}
+	\caption{\sys compiler overhead measured per application.}
+	\label{fig:comp_user}
+\end{figure}
+
+Result is provided in Figure~\ref{fig:comp_user}.
+
+\subsubsection{\sys Coalescing Overhead}
+\label{sec:result_compiler_time}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalescing}
+	\caption{\sys coalescing overhead measured per application.}
 	\label{fig:coalescing}
@@
 
-\subsubsection{Virtual Task Size}
-\label{sec:result_task_size}
+Result is provided in Figure~\ref{fig:coalescing}.
 
@@
 	\includegraphics[width=\columnwidth]{figures/averageVirtualTaskSize}
-	\caption{Average virtual task size (measured in real task unit\todo{Define real task unit}{Amjad}) of the virtual task sizes that the coalescing algorithm used to accomplished the applications at a WISP to RFID reader distance of 60\,cm.\todo{Please rewrite this sentence}{Amjad}}
+	\caption{Average virtual task size.}
 	\label{fig:aveVirtuTaskSize}
@@
 
-%\subsubsection{Completion Time}
-%\label{sec:result_completion}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\columnwidth]{figures/chain_vipe}
-%	\caption{Time to complete for a set of testing applications (refer to Table~\ref{table:benchmark_list}) using \sys versus Chain (no coalescing applied, fixed power supply).
-\todo{Figure placeholder---will be replaced with final one}{Amjad}}
-%	\label{fig:IPOSPerformance}
-%\end{figure}
-
-%Expected results:
-%\begin{itemize}
-%	\item performance for each application comparing bars for Chain, no coalescing/hand-tasked, coalescing/hand-tasked, no-coalescing/compiler, coalescing/compiler.
-Performance depicted, normalized to one of the configurations (probably Chain).
-This plot will cover most of the eval dimensions that we care about.
-%	\item Sensitivity plot---show amount of task merged, time to completion, overhead (in terms of task-size/cap-size sweep (remember we can vary either with approximately the same effect)), energy consumed.
-All experiments for fixed power, at 3 distances with wireless power, compared against Chain for 3 task/cap size configurations
-%\end{itemize}
-
-%Point for decision after main figure is ready
-
-%Pull the compiler stuff into its own plot that compares compiler to hand-tasked, maybe only for the coalescing fig.
-This reduces bars in plot 1, but runs the risk of making the compiler sound secondary, which we do not want to do.
-
-%*Report only Chain, coalescing/hand-tasked, coalescing/compiler in the main plot.
-This says that coalescing is the "default configuration" for VIPER.
-Later, in a second figure, we compare coalescing to no coalescing for hand-tasked and compiler as a characterization plot.
-
-%\todo{Provide results and text to this section}{Amjad}
-
-\subsubsection{Memory Management Performance}
+\subsubsection{\sys Paging Performance}
 \label{sec:results_memory_management}

### 2017-08-11 paper/evaluation.tex
work on evaluation section

@@
 WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet} with TI Code Composer Studio version 7.
-FET was also used as a power supply in non-intermittent power experiment.
-To measure energy consumption of all benchmark applications (please refer to Section~\ref{sec:results_software}) Carnegie Mellon's EDB~\cite{edb} was used.
-In experiments involving wireless power, WISP was powered through an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
-The RFID reader is controlled by a PC running Ubuntu 10.4 executing open source Python-based \emph{sllurp} library~\cite{sllrp_github}.
+FET was also used as a power supply in non-intermittent power experiments.
+Energy consumption measurements provided in eariler sections were done with Carnegie Mellon's EDB~\cite{edb}.
+In experiments involving wireless power, the WISP was powered through an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
+The RFID reader is controlled by a PC running Ubuntu 10.4 executing an open source Python-based \emph{sllurp} library~\cite{sllrp_github}.
 Transmitted power of the RFID reader was 30\,dBm at 915\,MHz center frequency.
 For distance-controlled experiments, the WISP was positioned at varying length hollow paper tubes placed directly on a flat lying antenna.
-Code execution times have been measured using Salea~\cite{saleae} digital logic analyzer.
+Code execution times were measured using Salea~\cite{saleae} logic analyzer.
 
-We also used an alternative setup where the MSP-TS430RGZ48C launchpad was for continuous power experiments.
-We used internal timer for measuring run time.
-For harvested energy, we used the ThingMagic Astra-EX RFID reader with power levels of 30\,dBm.
-We used EDB~\cite{edb} for execution time measurement.
+We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used for continuous power experiments.
+We used internal timer for measuring run time\todo{how about EDB?}{Kiwan}.
+For harvested energy, we used the ThingMagic Astra-EX RFID reader set at 30\,dBm transmission power\todo{say what antenna are you using with what gain}{Kiwan}.
 Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler.
-Both for Clang and LLVM we used version 3.8.0.
+Version 3.8.0 was used for both Clang and LLVM.
 
@@
 
-We compare \sys against state of the art intermittent execution approach---Chain~\cite{chain}.
-We are not comparing \sys against DINO~\cite{dino} or Mementos~\cite{mementos} (Refer to Section~\ref{sec:background_consistency}), as (i) Chain is the only modular execution model available to which we can compare \sys, and (ii) Chain proves already its superiority over sequential model counterparts (i.e.
-DINO and Mementos).
-
-We need to note that although another very relevant task-based runtime has been recently presented and referenced many times throughout this paper---Alpaca~\cite{alpaca} (written by the same group as Chain)---except the recently accepted manuscript we did not have the access to Alpaca's code during the preparation of this paper.
-Therefore Alpaca was excluded from the evaluation.
-\todo{Check if the line of argument is solid for the above paragraphs about not selecting DINO, Memenos and Aplaca}{Brandon}
+We compare \sys against state-of-the-art task-based runtime Chain~\cite{chain}.
+We are not comparing \sys against DINO~\cite{dino} or Mementos~\cite{mementos} (Refer to Section~\ref{sec:background_consistency}), as (i) Chain is the only modular execution model available to which we can compare \sys, and (ii) Chain proved already its superiority over sequential model counterparts (i.e.
+DINO and Mementos)\footnote{We need to note that although another very relevant task-based runtime has been recently presented and referenced many times throughout this paper---Alpaca~\cite{alpaca} (written by the same group as Chain)---except the recently accepted manuscript we did not have the access to Alpaca's code during the preparation of this paper.
+Therefore Alpaca was excluded from the evaluation.}.
 
@@
 		\hline
-		Nickname & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
+		Nickname (abbreviation) & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
 		\hline\hline

### 2017-08-11 paper/evaluation.tex
work on evaluation section

@@
 
-The applications that were used in the evaluation are: (i) \textbf{Bitcount}: counts bits in x y byte pseudo-random string using x algorithms.
-Each result is compared for correctness; (ii) \textbf{Cold chain}: emulates embedded temperature sensing engine: measurements are randomly generated and are compressed using LZW algorithm for x-entry dictionary and y-byte block; (iii) \textbf{Cuckoo}: implements cuckoo filter---x entry filter filled to y of capacity---program first stores a sequence of pseudo-random numbers, then queries the filter to recover it; (iv) \textbf{RSA}: encrypts predefined x byte long string with a y bytes long key\todo{were keys of 64 bytes tested?
-If yes: d they exhaust energy}{Amjad}.
+The applications that were used in the evaluation are: (i) \textbf{Bitcount}: counts bits in byte pseudo-random string using varying algorithms.
+Each result is compared for correctness; (ii) \textbf{Cold chain}: emulates an embedded temperature sensing engine: measurements are randomly generated and are compressed using a LZW algorithm for x-entry dictionary and y-byte block; (iii) \textbf{Cuckoo}: implements cuckoo filter---x entry filter filled to y of capacity---the program first stores a sequence of pseudo-random numbers, then queries the filter to recover it; (iv) \textbf{RSA}: encrypts predefined x byte long string with a y bytes long key\todo{were keys of 64 bytes tested?
+If yes: do they exhaust energy}{Amjad}.
 This set of benchmarks is the same as used in~\cite{chain,alpaca}\todo{Fill-in the missing data and verify its correctness}{Przemek/Amjad}.
@@
 All applications are compared in Table~\ref{table:benchmark_list} in terms of source lines of code against Chain.
-\todo{finalize describing all applications once all experiments are done}{Przemek}
 
@@
 
-We start with evaluating the efficiency of a \sys compiler.
-Table~\ref{table:compiler_result} shows the difference in the number of protected variables (PVAR), number of tasks, and number of inserted paging calls in the code between (i) hand-written \sys and (ii) compiler-generated \sys code.
+We start with evaluating the efficiency of the \sys compiler Table~\ref{table:compiler_result} shows the difference in the number of protected variables (PVAR), number of tasks, and number of inserted paging calls\todo{calls or code?
+what is paging call?}{Kiwan} in the code between (i) hand-written \sys and (ii) compiler-generated \sys code.
 On average, compiler-generated code is only worse by 0.3\% in the number of protected variables.
-Regarding the number of tasks, compiler-generated code is better by roughly 16\%.
+Regarding the number of tasks, compiler-generated code is better by roughly 16\%\todo{except for cc and ar---why?}{Kiwan}.
 The result indicate that \sys compiler is selecting boundaries better than the programmer doing it manually.
-However, \sys compiler should insert paging code conservatively when it {\em may} access protected variables, due to limitation of pointer alias.
-Thus, it results in \sys compiler introducing 48\% more paging code than programmer-written code (the results on run time for various \sys code versions is provide in Section~\ref{}).
-The performance is worse than programmer-written \sys code in overall, meaning the programmer should structure the code manually if the performance is critical; otherwise compiler can be used to simplify the programmability.
+However, the \sys compiler should insert paging code conservatively when it {\em may} access protected variables, due to limitation of pointer aliasing.
+Thus, it results in the \sys compiler introducing 48\% more paging code than programmer-written code (the results on run time for various \sys code versions is provided in Section~\ref{sec:result_compiler_time}).
+The performance is worse than programmer-written \sys code in overall, meaning the programmer should structure the code manually if the performance is critical; otherwise the compiler can be used to simplify the programmability.
 
@@
 \texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}.
-We compiled the apps on a 4-core virtual machine with 1GB memory that has Arch Linux, running on Intel Core i5-6200U (2.30GHz) processor.
+We compiled the apps on a 4-core virtual machine with 1\,GB memory with Arch Linux, running on Intel Core i5-6200U (2.30\,GHz) processor.
 The result are given in Table~\ref{table:compile_time}.
-We simply used {\em date} command provided by linux to measure the rough time.
-Most of the compile time was due to pointer aliasing, when calculating possible live range of each variables.
-Apps such as RSA whose code size is large and uses a lot of variables needed longer compilation time due to larger search space for pointer aliasing.
+We used a {\em date}\todo{date or time?}{Kiwan} linux command to measure the rough time.
+Most of the compile time was due to pointer aliasing, when calculating possible live ranges of variables.
+Apps such as RSA whose code size is large and uses a lot of variables, need longer compilation time due to a larger search space for pointer aliasing.
 

### 2017-08-11 paper/evaluation.tex
work on evaluation section

@@
 
-The applications that were used in the evaluation are: (i) \textbf{Bitcount}: counts bits in byte pseudo-random string using varying algorithms.
-Each result is compared for correctness; (ii) \textbf{Cold chain}: emulates an embedded temperature sensing engine: measurements are randomly generated and are compressed using a LZW algorithm for x-entry dictionary and y-byte block; (iii) \textbf{Cuckoo}: implements cuckoo filter---x entry filter filled to y of capacity---the program first stores a sequence of pseudo-random numbers, then queries the filter to recover it; (iv) \textbf{RSA}: encrypts predefined x byte long string with a y bytes long key\todo{were keys of 64 bytes tested?
-If yes: do they exhaust energy}{Amjad}.
-This set of benchmarks is the same as used in~\cite{chain,alpaca}\todo{Fill-in the missing data and verify its correctness}{Przemek/Amjad}.
-In addition to Chain benchmarks new applications were implemented: (v) \textbf{DFT}, (vi) \textbf{Huffman compressor}, and (vii) \textbf{Selection Sort}.
+The applications that were used in the evaluation are: (i) \textbf{bc}: counts bits in byte pseudo-random string using varying algorithms.
+Each result is compared for correctness; (ii) \textbf{cem}: emulates an embedded temperature sensing engine: measurements are randomly generated and are compressed using a LZW algorithm for  fixed entry dictionary and fixed byte block; (iii) \textbf{cuckoo}: implements Cuckoo filter---the program first stores a sequence of pseudo-random numbers, then queries the filter to recover it; (iv) \textbf{rsa}: encrypts predefined string with a predefined key; (v) \textbf{ar} performs nearest neighbour classification using randomly generated sequence of three-axes accelerometer data.
+This set of benchmarks is the same as used in~\cite{chain,alpaca}.
+In addition to Chain benchmarks new applications were implemented: (vi) \textbf{dft}: discreet Fourier transform, (vii) \textbf{dd}: Huffman-based data decompression, and (viii) \textbf{sort}: selection sort algorithm.
 All applications are compared in Table~\ref{table:benchmark_list} in terms of source lines of code against Chain.
@@
 		\hline
-		Nickname (abbreviation) & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
+		Application & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
 		\hline\hline
-		Bitcount (bc) & 351 & 588 \\ %59\%
-		Cold chain (cem) & 388 & 721 \\ %53\%
-		Cuckoo (cuckoo) & 483 & 762 \\ %63\%
-		RSA (rsa) & 887 & 1233 \\ %71\%
-		Activity (ar) & 483 & 762 \\ %63\%
-		DFT (dft) & {} & {} \\ %Two resolutions: 4 and 8 Bytes
-		Huffman (dd) & {} & {} \\ %Data decompression size: 100 bytes
-		Selection sort (sort) & {} & {} \\
+		{bc} & 351 & 588 \\ %59\%
+		{cem} & 388 & 721 \\ %53\%
+		{cuckoo} & 483 & 762 \\ %63\%
+		{rsa} & 887 & 1233 \\ %71\%
+		{ar} & 483 & 762 \\ %63\%
+		{dft} & {} & {} \\ %Two resolutions: 4 and 8 Bytes
+		{dd} & {} & {} \\ %Data decompression size: 100 bytes
+		{sort} & {} & {} \\
 		\hline
@@
 
-\subsection{Evaluation Results}
+\subsection{\sys Evaluation Results}
 \label{sec:results_evaluation}
 
-\subsubsection{Compiler Performance}
+\subsubsection{\sys Compiler Performance}
 \label{sec:results_compiler}
@@
 		\hline\hline
-		Bitcount & 22 & 22 & 10 & 15 & 81 & 93 \\
-		Cold chain & 3492 & 3242 & 12 & 9 & 92 & 123 \\
-		Cuckoo & 282 & 288 & 15 & 6 & 90 & 76 \\
-		RSA & 332 & 250 & 20 & 27 & 130 & 296 \\
-		AR & 166 & 218 & 11 & 6 & 112 & 333 \\
-		Sort & 104 & 104 & 4 & 2 & 70 & 23 \\
+		bc & 22 & 22 & 10 & 15 & 81 & 93 \\
+		cem & 3492 & 3242 & 12 & 9 & 92 & 123 \\
+		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 \\
+		rsa & 332 & 250 & 20 & 27 & 130 & 296 \\
+		ar & 166 & 218 & 11 & 6 & 112 & 333 \\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 \\
 		\hline
 	\end{tabular}
-	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand).}
+	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand).\todo{Data for Huffman in Table 5 and 6 is missing}{Kiwan}}
 \label{table:compiler_result}
@@
 
-Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile DFT application since the LLVM inserts mathematical built-in functions (i.e.
+Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile the DFT application since LLVM inserts mathematical built-in functions (i.e.
 \texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}.
@@
 		\hline
-		Bitcount & Cold chain & Cuckoo & RSA & AR & Sort \\
+		bc & cem & cuckoo & rsa & ar & sort \\
 		\hline\hline
@@
 
-\subsubsection{\sys Performance}
+\subsubsection{\sys Runtime Performance}
 \label{sec:result_coalescing}
@@
 From the result we clearly see an enormous benefit of coalescing, especially at large WISP to RFID reader antenna distance ($\approx$56\% for cem, 100\% for sort and $\approx$400\% for bc).
-At each distance, \sys coalescing mechanism provided faster execution times compared to non-coalescing case.
+At each distance, \sys coalescing mechanism provided faster execution times compared to the non-coalescing case.
 \todo{Discuss different shapes per app}{Amjad}

### 2017-08-11 paper/texFigCode/sram-buffer.tex
page figure

@@
 
-	\node[punkt,minimum height=3em] (task1) {y = x+1 
-		\\ z = y+1 \\ x = z+1};
+	\node[punkt,minimum height=3em] (task1) {...\\z = y+1 \\ ... \\ print x \\ ... \\ commit};
 	\node[below = 0.1cm of task1] () {\textbf{Task}};
 
-	\node[draw,very thick,dotted,minimum width=7em, minimum height=7em, right = 3cm of task1] (sram) {};
-	\node[below = 0.1cm of sram] () {\textbf{Volatile Buffer}};
+	\node[draw,very thick,minimum width=7em, minimum height=7em, right = 3cm of task1] (sram) {};
+	\node[below = 0.1cm of sram] () {\texttt{pageBuf}};
 
-	\node[punkt, below = -2.4cm of sram,fill=red!3] (x-sram) {x};
-	\node[punkt, below = 0.1cm of x-sram,fill=black!5] (y-sram) {y};
-	\node[punkt, below = 0.1cm of y-sram,fill=black!5] (z-sram) {z};
+	\node[punkt,draw=none, below = -2cm of sram,fill=black!5] (y-sram) {y};
+	\node[punkt,draw=none, below = 0.1cm of y-sram,fill=black!5] (z-sram) {z};
 
-	\node[right = 1.8cm of sram] (ref) {};
-	\node[draw,very thick,minimum width=7em, minimum height=15em, right = of ref] (fram1) {};
+	\node[right = 0.5cm of sram] (ref) {};
+	\node[draw,very thick,minimum width=8em, minimum height=30em, right = of ref] (fram1) {};
 
-	%\node[draw,very thick,dotted,minimum width=7em, fill=green!5,minimum height=9em, below = 1cm of fram1] (fram2) {};
+	\node[draw,above = -2cm of fram1,minimum width=6em, minimum height=5em] (pagesOrg1) {};
+	\node[draw,below = 0 cm of pagesOrg1,minimum width=6em, minimum height=5em] (pagesOrg2) {};
+	\node[below = 0.1cm of pagesOrg2] (original) {\texttt{pagesOrg}};
 
-	\node[punkt, above = -0.8cm of fram1,fill=green!3] (x1-fram) {x};
-	\node[punkt, below = 0.1cm of x1-fram,fill=black!10] (y1-fram) {y};
-	\node[punkt, below = 0.1cm of y1-fram,fill=black!10] (z1-fram) {z};
+	\node[punkt, above = -0.8cm of pagesOrg1] (x1-fram) {x};
+	\node[punkt,draw=none, above =-0.8cm of pagesOrg2,fill=black!10] (y1-fram) {y};
 
-	\node[punkt, below = -2.5cm of fram1,fill=black!5] (x2-fram) {x};
-	\node[punkt, below = 0.1cm of x2-fram,fill=black!5] (y2-fram) {y};
-	\node[punkt, below = 0.1cm of y2-fram,fill=black!5] (z2-fram) {z};
+	\node[punkt,draw=none, below = 0.1cm of y1-fram,fill=black!10] (z1-fram) {z};
 
-	\node[below = 0.1cm of z2-fram] (copies) {copies};	
-	\node[below = 0.1cm of z1-fram] (original) {original};	
-	\node[below = 0.7cm of z2-fram] () {\textbf{FRAM}};
+	\node[draw,below = -2.7cm of fram1,minimum width=6em, minimum height=5em] (pagesTmp) {};
+	\node[punkt,draw=none, below = -2.5cm of fram1,fill=black!5] (y2-fram) {y};
+	\node[punkt,draw=none, below = 0.1cm of y2-fram,fill=black!5] (z2-fram) {z};
+
+	\node[below = 0.1cm of pagesTmp] (copies) {\texttt{pagesTmp}};
+
+	\node[below = 0cm of fram1] () {\textbf{FRAM}};
 
 	\path[every node/.style={font=\sffamily\small}]
-	([yshift=0.3cm] task1.east) edge [->, bend left] node {(1) get x} (x-sram.west)
-	(x-sram.east) edge [->, bend left] node[xshift=0.5cm] {(2) read x} (x1-fram.west)
+	([yshift=0.6cm] task1.east) edge [->, bend right,dotted] node[xshift=-0.5cm]  {(2) \texttt{WVAR(z)} } (z-sram.west)
 
-	([yshift=0.3cm]task1.east) edge [->, bend right] node[xshift=-0.5cm] {(3) set y} (y-sram.west)
+	([yshift=0.7cm]task1.east) edge [->, bend left,dotted,thick] node[xshift=-1cm] {(1) \texttt{RVAR(y)}} (y-sram.west)
 
-	(task1.east) edge [->, bend right] node[xshift=-0.4cm] {(4) set z} (z-sram.west)
-	([yshift=-0.4cm]task1.east) edge [->, bend left] node[xshift=1.3cm,yshift=-0.4cm] {(5) set x} (x-sram.west)
+	([yshift=-0.5cm]task1.east) edge [->, bend right,dotted,thick] node[below] { \shortstack{(3) \texttt{RVAR(x)}:\\ page fault}} ([yshift=-1cm]sram.west)
 
-	(x-sram.east) edge [->, bend right] node[xshift=-0.4cm] {(6) commit-I x} (x2-fram.west)
-	(z-sram.east) edge [->, bend right] node[xshift=-0.4cm] {(7) commit-I z} (z2-fram.west)
-	(y-sram.east) edge [->, bend right] node[xshift=-0.4cm] {(8) commit-I y} (y2-fram.west)
+	([xshift=1cm]sram.south) edge [->, bend right,dotted,thick] node[xshift=-1.4cm] {(4.1) page swap} (z2-fram.west)
 
-	(x2-fram.east) edge [->, bend right] node[xshift=2.5cm] {(9) commit-II x} (x1-fram.east)
-	(y2-fram.east) edge [->, bend right] node[xshift=2.5cm] {(10) commit-II y} (y1-fram.east)
-	(z2-fram.east) edge [->, bend right] node[xshift=2.5cm] {(11) commit-II z} (z1-fram.east)
-	;
+	(pagesOrg1.west) edge [->, bend right,dotted,thick] node[xshift=-1.4cm] {(4.2) page swap} (sram.north)
+
+	(pagesTmp.east) edge [->, bend right,dotted,thick] node{(6) \texttt{commit}} (pagesOrg2.east);
 

### 2017-08-11 paper/coalescing.tex
coalescing figure revised

@@
 
-The key challenge to coalescing tasks is determining how many tasks to coalesce before committing.
+The challenge to coalescing tasks is determining how many tasks to coalesce before committing.
 During execution, an executing task is either a {\em coalesced} task, which does not commit its state before transitioning or a {\em committing} task, which does commit its state before transitioning.
@@
 
-\noindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
+\noindent\textbf{Task Coalescing Implementation:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
 Figure~\ref{fig:coalescing_example_states} (left) shows a sequence of three tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.
@@
 Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$.
-\sys executes eight tasks in the first operating period, assigning $C=4$.
+\sys executes eight tasks in the first operating period, assigning the half of previous $C=8/2$.
 \sys completes only two of the four coalesced tasks in the next operating period and then power fails.

### 2017-08-11 paper/evaluation.tex
fixed some red text

@@
 We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used for continuous power experiments.
-We used internal timer for measuring run time\todo{how about EDB?}{Kiwan}.
-For harvested energy, we used the ThingMagic Astra-EX RFID reader set at 30\,dBm transmission power\todo{say what antenna are you using with what gain}{Kiwan}.
+We used internal timer for measuring run time.
+For harvested energy, we used the ThingMagic Astra-EX RFID antenna set at 30\,dBm transmission power.
+We used EDB~\cite{edb} to measure run time.
 Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler.

### 2017-08-11 paper/evaluation.tex
combined tables

--- 2017-08-11 paper/evaluation.tex
@@
 
-\begin{table}[t]
-	\centering
-	\footnotesize
-	\begin{tabular}{|c|c|c|}
-		\hline
-		Application & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
-		\hline\hline
-		{bc} & 351 & 588 \\ %59\%
-		{cem} & 388 & 721 \\ %53\%
-		{cuckoo} & 483 & 762 \\ %63\%
-		{rsa} & 887 & 1233 \\ %71\%
-		{ar} & 483 & 762 \\ %63\%
-		{dft} & {} & {} \\ %Two resolutions: 4 and 8 Bytes
-		{dd} & {} & {} \\ %Data decompression size: 100 bytes
-		{sort} & {} & {} \\
-		\hline
-	\end{tabular}
-\caption{List of benchmark programs used in \sys evaluation; SLOC: source lines of code.}
-\label{table:benchmark_list}
-\end{table}
+%\begin{table}[t]
+%	\centering
+%	\footnotesize
+%	\begin{tabular}{|c|c|c|}
+%		\hline
+%		Application & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
+%		\hline\hline
+%		{bc} & 351 & 588 \\ %59\%
+%		{cem} & 388 & 721 \\ %53\%
+%		{cuckoo} & 483 & 762 \\ %63\%
+%		{rsa} & 887 & 1233 \\ %71\%
+%		{ar} & 483 & 762 \\ %63\%
+%		{dft} & {} & {} \\ %Two resolutions: 4 and 8 Bytes
+%		{dd} & {} & {} \\ %Data decompression size: 100 bytes
+%		{sort} & {} & {} \\
+%		\hline
+%	\end{tabular}
+%\caption{List of benchmark programs used in \sys evaluation; SLOC: source lines of code.}
+%\label{table:benchmark_list}
+%\end{table}
 
@@
 	\footnotesize
-	\begin{tabular}{|c|c|c|c|c|c|c|c|}
+        \renewcommand{\tabcolsep}{1pt}
+	\begin{tabular}{l|cc|cc|cc|cc|c}
 		\hline
-		{} & \multicolumn{2}{l|}{{PVAR size (Byte)}} & \multicolumn{2}{c|}{{No. tasks}} & \multicolumn{2}{c|}{{No. paging code}} \\
+		{} & \multicolumn{2}{l|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
+		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
 		\hline
-		Nickname & Hand & Comp. & Hand & Comp. & Hand & Comp. \\
-		\hline\hline
-		bc & 22 & 22 & 10 & 15 & 81 & 93 \\
-		cem & 3492 & 3242 & 12 & 9 & 92 & 123 \\
-		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 \\
-		rsa & 332 & 250 & 20 & 27 & 130 & 296 \\
-		ar & 166 & 218 & 11 & 6 & 112 & 333 \\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 \\
+		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
+		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
+		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
+		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
+		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & & & $<$1\\
 		\hline
 	\end{tabular}
-	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand).\todo{Data for Huffman in Table 5 and 6 is missing}{Kiwan}}
+	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand).\todo{Data for dd, dft, sort missing}{Kiwan}}
 \label{table:compiler_result}
@@
 On average, compiler-generated code is only worse by 0.3\% in the number of protected variables.
-Regarding the number of tasks, compiler-generated code is better by roughly 16\%\todo{except for cc and ar---why?}{Kiwan}.
-The result indicate that \sys compiler is selecting boundaries better than the programmer doing it manually.
+Regarding the number of tasks, compiler-generated code is better by roughly 16\%.
+The results indicate that \sys compiler is selecting boundaries better than the programmer doing it manually.
 However, the \sys compiler should insert paging code conservatively when it {\em may} access protected variables, due to limitation of pointer aliasing.
@@
 We compiled the apps on a 4-core virtual machine with 1\,GB memory with Arch Linux, running on Intel Core i5-6200U (2.30\,GHz) processor.
-The result are given in Table~\ref{table:compile_time}.
-We used a {\em date}\todo{date or time?}{Kiwan} linux command to measure the rough time.
+The result are given in Table~\ref{table:compiler_result}.
+We used the {\tt time} linux command to measure the rough time.
 Most of the compile time was due to pointer aliasing, when calculating possible live ranges of variables.
@@
 
-\begin{table}[t]
-	\centering
-	\footnotesize
-	\begin{tabular}{|c|c|c|c|c|c|}
-		\hline
-		bc & cem & cuckoo & rsa & ar & sort \\
-		\hline\hline
-		3 & 2 & 6 & 86 & 34 & < 1 \\
-		\hline
-	\end{tabular}
-	\caption{\sys compile time per application (seconds).}
-\label{table:compile_time}
-\end{table}
-
+%\begin{table}[t]
+%	\centering
+%	\footnotesize
+%	\begin{tabular}{|c|c|c|c|c|c|}
+%		\hline
+%		bc & cem & cuckoo & rsa & ar & sort \\
+%		\hline\hline
+%		3 & 2 & 6 & 86 & 34 & < 1 \\
+%		\hline
+%	\end{tabular}
+%	\caption{\sys compile time per application (seconds).}
+%\label{table:compile_time}
+%\end{table}
+%
 %\subsubsection{Task Coalescing Performance}

### 2017-08-11 paper/coalescing.tex

@@
 Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$.
-\sys executes eight tasks in the first operating period, assigning the half of previous $C=8/2$.
-\sys completes only two of the four coalesced tasks in the next operating period and then power fails.
-\sys sets $C=(6+8+2)/2$ in the next operating period, coalescing pairs of tasks and successfully committing two pairs before power fails.
-Execution resumes at the start of the third pair of tasks with a coalescing factor of two.
-If a larger number of tasks successfully complete in the next operating period, i.e., $S$ is larger, $C$ correspondingly increases, allowing more aggressive coalescing.\todo{Rewrite this paragraph to reflect what is in the figure}{Przemek}
+\sys executes tasks in the first operating period, assigning the \emph{half of sum of all previous $S$ periods} (in the example case: $C=8/2$).
+\sys completes then coalesced tasks successfully (four tasks were executed) and in the next operating period, i.e.
+$C=(8+4)/2$.
+In the third task execution, power fails after the second task.
+Then \sys re-executes coalesced task since the previous failure and sets $C=(6+8+2)/2$ in the next operating period.
+If a larger number of tasks successfully complete in the next operating period, i.e., $S$ is larger, $C$ correspondingly increases, allowing more aggressive coalescing.
 

### 2017-08-11 paper/paging.tex

@@
 
-\todo{Redo this figure to actually reflect paging -- right now it isn't clear if x,y,z are in a page or not, it isn't clear when the page gets copied to pagesTemp, it isn't clear what state we add to the system to help with commit, etc (i.e., dirty bits, commit bit).
-Make this figure's labels consistent with the text: pagesTemp, RVAR, WVAR, explicit
-commit in task, Page Buffer}{whoever}
-
 Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task using \sys's paging mechanism.
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/sram-buffer}
+	\includegraphics[width=0.75\columnwidth]{figures/sram-buffer}
 	\caption{The interaction between the task, volatile buffer and the non-volatile memory (FRAM).

### 2017-08-11 paper/evaluation.tex
work on evaluation

@@
         \renewcommand{\tabcolsep}{1pt}
-	\begin{tabular}{l|cc|cc|cc|cc|c}
-		\hline
+	\begin{tabular}{|l|cc|cc|cc|cc|c|}
+		\hline\hline
 		{} & \multicolumn{2}{l|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
@@
 
-\subsubsection{\sys Coalescing Overhead}
+\subsubsection{\sys Coalescing Performance}
 \label{sec:result_compiler_time}
@@
 	\includegraphics[width=\columnwidth]{figures/coalescing}
-	\caption{\sys coalescing overhead measured per application.}
+	\caption{\sys coalescing performance measured per application at fixed power and various WISP to RFID antenna distances, \{20, 40, 60\}\,cm, assessed against non-coalescing \sys.}
 	\label{fig:coalescing}
@@
 
+We measured \sys coalescing performance.
 Result is provided in Figure~\ref{fig:coalescing}.
+Result clearly show that coalescing provides a benefit for all applications, providing even seven times improvement for bc application at 40\,cm.
 
@@
 	\includegraphics[width=\columnwidth]{figures/averageVirtualTaskSize}
-	\caption{Average virtual task size.}
+	\caption{Average \sys virtual task size at fixed power per application. Top (green): number of coalesced tasks; bottom (red): average size of real task.}
 	\label{fig:aveVirtuTaskSize}
@@
 
-Measured size of tasks for each application is given in Figure~\ref{fig:aveVirtuTaskSize}.
+We need to know how many tasks \sys can coalesce.
+We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and amount of tasks coalesced.
+Result is given in Figure~\ref{fig:aveVirtuTaskSize}.
+We see clearly that \sys manages to coalesce more tasks as individual tasks are small.
+As the size of individual task increases, e.g., as in the case of dft application, the number of coalesced tasks is also small.
+This result clearly shows that for the benefit of coalescing and for code portability, initial code should be split by \emph{as small tasks as possible}.
 

### 2017-08-11 paper/evaluation.tex
table 4 needs fix

--- 2017-08-11 paper/evaluation.tex
@@
 		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & & & $<$1\\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & & $<$1\\
 		\hline
 	\end{tabular}
-	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand).\todo{Data for dd, dft, sort missing}{Kiwan}}
+	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand). DFT does not compile with Clang}
 \label{table:compiler_result}

### 2017-08-11 paper/evaluation.tex
table 4 filled

@@
 		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & & $<$1\\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
 		\hline

### 2017-08-11 paper/evaluation.tex
added table content.

--- 2017-08-11 paper/evaluation.tex
@@
 		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & & $<$1\\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
+		dft* & - & - & - & - & - & - & 222 & 293 & -\\
+		dd &  &  &  &  &  &  &  & 287 & $ \\
 		\hline
 	\end{tabular}
-	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand). DFT does not compile with Clang}
+	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand). *DFT does not compile with Clang}
 \label{table:compiler_result}

### 2017-08-11 paper/evaluation.tex
fixed bug

--- 2017-08-11 paper/evaluation.tex
@@
 		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & & $<$1\\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
+		dft* & - & - & - & - & - & - & 222 & 293 & -\\
+		dd &  &  &  &  &  &  &  & 287 &  \\
 		\hline
 	\end{tabular}
-	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand). DFT does not compile with Clang}
+	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand). *DFT does not compile with Clang\todo{incomplete table}{Responsible:Kiwan}}
 \label{table:compiler_result}

### 2017-08-11 paper/evaluation.tex
added explanation

@@
 
-We start with evaluating the efficiency of the \sys compiler Table~\ref{table:compiler_result} shows the difference in the number of protected variables (PVAR), number of tasks, and number of inserted paging calls\todo{calls or code?
-what is paging call?}{Kiwan} in the code between (i) hand-written \sys and (ii) compiler-generated \sys code.
+We start with evaluating the efficiency of the \sys compiler Table~\ref{table:compiler_result} shows the difference in the number of protected variables (PVAR), number of tasks, and number of inserted paging macro ({\tt RVAR} and {\tt WVAR}) in the code between (i) hand-written \sys and (ii) compiler-generated \sys code.
 On average, compiler-generated code is only worse by 0.3\% in the number of protected variables.

### 2017-08-11 paper/background.tex
merge commit
stash pop merge

@@
 Energy harvesting devices operate using energy extracted from sources such as RF and solar energy.
-Energy-harvesting devices do not use a tethered power supply or battery and instead, typically collect energy into a (super-)capacitor, operate when sufficient energy has accumulated, and upon depleting the stored energy, turn off and wait to collect more energy.
+These devices elide tethered power or a battery, instead collecting energy into a capacitor, operating when sufficient energy accumulates, and upon depleting the energy, turning off and recharging.
 %
@@
 %
-Software running on energy harvesting device executes {\em intermittently} because energy is only available between when the capacitor reaches its threshold and when it is depleted.
-An intermittent execution is composed of periods interspersed with power failures~\cite{dino,chain,alpaca,ratchet}.
-The frequency of power failures depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods (but requires longer charge times).
-Energy-harvesters provide input power that is orders of magnitude less than a device's operating power, making recharging negligible during operation.
+Software running on an energy-harvesting device executes {\em intermittently} because buffered energy is only sometimes available.
+An intermittent execution is composed of operating periods interspersed with power failures~\cite{dino,chain,alpaca,ratchet}.
+The frequency of failures depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods.
+Energy-harvesters provide input power orders of magnitude less than operating power, making recharging negligible during operation.
 
-Software behaves uniquely when executed intermittently.
-Each power failure clears volatile memory (registers, stack, and global variables).
-Non-volatile memory (e.g., FRAM) persists across power failures.
-On a power failure, control flows back to a prior point in the execution.
-By default, after a power failure control flows to the beginning of {\tt main()}.
-Early architectures preserved intermittent progress by periodically capturing and restoring to a non-volatile checkpoint of the volatile execution context, i.e., registers and stack~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
+Intermittent execution is different from continuous execution.
+A power failure clears volatile  state (registers, stack, and globals) and non-volatile memory (e.g., FRAM) persists.
+At a failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
+Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
 
@@
 The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
+
+Despite checkpointing at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
+Figure~\ref{fig:crc_example_war} shows an intermittent execution.
+The second iteration writes $\texttt{Rmnd}$, but power fails before the checkpoint.
+After restarting, line 9 reads the updated value of $\texttt{Rmnd}$, producing an incorrect $\texttt{data}$ value.
 %
-%Even with the checkpoint at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of a WAR dependence between the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
-Figure~\ref{fig:crc_example_war} shows an execution of the code in which power fails during the second loop iteration.
-The problem arises because line 6 of the execution writes $\texttt{Rmnd}$ in the second iteration, but power fails before capturing the next checkpoint.
-After restarting from the checkpoint, the update to $\texttt{data}$ at line 9 reads the value of $\texttt{Rmnd}$ that was updated before the power failure, producing an incorrect result.
-%
-As the figure illustrates, checkpoint-based systems risk violating the volatile and non-volatile memory consistency~\cite{dino}.
-To maintain consistency, versioning systems~\cite{dino,ratchet} {\em version} a subset of non-volatile variables with the checkpointed execution context.
-Task-based systems~\cite{chain,alpaca}, which we focus on this work and discuss in detail next, ensure state consistency by using programmer, compiler, and runtime support.
+Checkpoint-based systems risk violating memory consistency~\cite{dino}.
+To maintain consistency, prior systems~\cite{dino,ratchet} {\em version} a subset of non-volatile data with the checkpoint.
+Task-based systems~\cite{chain,alpaca}, which we focus on, ensure consistency using programmer, compiler, and runtime support.
 
@@
 
-Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into a collection tasks.
-A task is a function with no caller that contains arbitrary computation, sensing, and communication.
-A programmer describes the sequence in which tasks may execute, creating a task graph.
-The flow of control from task to task happens at programmer demarcated points and may conditionally depend on program values.
-Task-based programming abstractions guarantee that tasks exhibit behavior as though they executed {\em atomically}, regardless of arbitrary power failures.
+Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into tasks.
+A task is a function with no caller containing arbitrary computation, sensing, and communication.
+A programmer describes task control-flow as a {\em task graph}.
+Task flow happens at programmer demarcated transition points that may be conditional on program values.
+Task-based programming abstractions guarantee that tasks execute {\em atomically}, regardless of power failures.
 Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent.
-The key to idempotence is ensuring that updates to non-volatile state made by an interrupted, partial execution of a task are never visible to a subsequent (potentially completing) re-execution of the task.
+The key to idempotence is ensuring that non-volatile updates made by an interrupted task are never visible to a future task execution.
 
 There are several run time strategies to ensuring task idempotence.
-One approach, utilized by Alpaca~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{ratchet} did for checkpoints), execute the task using buffered copies of those data, and commit the buffered copies before completing the task.
-Another strategy is to statically create multiple versions of non-volatile data shared by tasks and ensure by construction that no task reads and writes the same non-volatile memory location~\cite{chain}.
-Regardless of the idempotence strategy, the key to task-based execution is that fixed, statically-defined tasks execute as though atomic, completing after zero or more attempts.
+One approach~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{ratchet} did for checkpoints), execute the task using private copies of those data, and commit the private copies on task completion.
+Another way is to statically create multiple versions of non-volatile data shared by tasks and ensure that no task reads and writes the same version~\cite{chain}.
+Regardless of the strategy, task-based systems execute statically-defined tasks atomically, completing in one or more attempt.
 
@@
 
-Despite valuable features, task-based models incur costs that are the motivation for \sys.
+Task-based models incur costs that motivate \sys.
 
-\noindent \textbf{Memory Consistency Enforcement Overhead:} Intermittent execution models incur an overhead because they must checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
-Table~\ref{table:chechpoint_comparison} compares the overheads for several recent intermittent execution models (generalizing discussion from~\cite[Sec.
-2.4]{alpaca}.) The mechanism responsible for the overhead of each model varies, but all the bulk of overhead in all models is a manipulating non-volatile memory.
-Checkpointing moves volatile memory and registers to and from non-volatile memory, channel accesses directly manipulate non-volatile data~\cite{chain}, privatization copies data from and commits data to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} require all of memory to be non-volatile.
+\noindent \textbf{Memory Consistency Enforcement Overhead:} Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
+Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf.
+\cite[Sec.
+2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory.
+Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.
 
@@
 
-To concretely understand the cost of these manipulations of non-volatile memory, we directly measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations for the MSP430FR5969~\cite{msp430datasheet}.
-We used EDB~\cite{edb} to monitor the voltage drop on an energy buffering capacitor that powers the device across 1600 read and write accesses to each type of memory, randomizing locations to avoid caching effects.
-Figure~\ref{fig:framEnergy} summarizes these measurements, illustrating that a non-volatile memory access consumes around 1.5 times as much energy as a volatile memory access in this device.
-In an intermittently-powered device, the increased energy corresponds to decreased run time between power failures.
-The overheads imposed by relying on non-volatile memory motivate \sys, which uses memory virtualization to operate primarily using volatile SRAM.
+To understand these non-volatile memory overheads, we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations for the MSP430FR5969~\cite{msp430datasheet} MCU.
+We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device across 1600 read and write accesses to each memory type, randomizing to avoid caching.
+Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device.
+In an intermittent device, energy cost corresponds to decreased run time between power failures.
+The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.
 
@@
 
-\begin{table}
-	\centering
-	\footnotesize
-	\begin{tabular}{|c|c|}
-		\hline
-		Platform name & Storage capacitor size \\
-		\hline \hline
-		%Moo~\cite{moo} & 0.1\,F \\
-		WISPCam~\cite{naderiparizi_rfid_2015} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
-		%NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
-		NeuralWISP~\cite{holleman_biocas_2008} & 100\,$\mu$F \\
-		WISP~\cite{wisp5} & 47\,$\mu$F \\
-		{\em BioImpedance} sensor~\cite{rodriguez_tbcs_2015} & 20\,$\mu$F \\
-		{\em Ingestible} sensor~\cite{nadeau_naturebio_2017} & 220\,$\mu$F\\
-		\hline
-	\end{tabular} 
-	\caption{{\em Default} energy storage sizes for example battery-less platforms. Observe a huge variation in storage capacity.}
-	%We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.
-	\label{table:capacitor}
-\end{table}
+%\begin{table}
+%	\centering
+%	\footnotesize
+%	\begin{tabular}{|c|c|}
+%		\hline
+%		Platform name & Storage capacitor size \\
+%		\hline \hline
+%		%Moo~\cite{moo} & 0.1\,F \\
+%		WISPCam~\cite{naderiparizi_rfid_2015} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
+%		%NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
+%		NeuralWISP~\cite{holleman_biocas_2008} & 100\,$\mu$F \\
+%		WISP~\cite{wisp5} & 47\,$\mu$F \\
+%		{\em BioImpedance} sensor~\cite{rodriguez_tbcs_2015} & 20\,$\mu$F \\
+%		{\em Ingestible} sensor~\cite{nadeau_naturebio_2017} & 220\,$\mu$F\\
+%		\hline
+%	\end{tabular} 
+%	\caption{{\em Default} energy storage sizes for example battery-less platforms. Observe a huge variation in storage capacity.}
+%	%We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.
+%	\label{table:capacitor}
+%\end{table}
 
-\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} If a programmer writes a task that consumes more energy than the device can buffer, the task will never be able to complete using the buffered energy.
-The consequence of such tasks is a non-progressing program that repeatedly attempts to re-execute the task of excess energy cost.
-If a programmer writes a program that consumes less energy than the device can buffer, the system may operate inefficiently.
-When a task and its successor both complete, the two tasks were interrupted by a task boundary that preserves progress and state (i.e., checkpoint or commit).
-If the two tasks were not also interleaved by a power failure, then the computation done to preserve state was {\em unnecessary}, only incurring an overhead on the tasks' executions.
-Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity that might be \emph{different} for each transiently-powered device (see Table~\ref{table:capacitor}).
-Such tasks also make it extremely difficult to {\em port} code from one device to another.
-A task that is excessively costly on a device with a small energy buffer may become a short, high-overhead task on a device with a larger energy buffer and vice versa.
-The key problem is that the programmer writes tasks statically for an energy buffer of a fixed size.
-\sys's {\em task coalescing} mechanism, which we will describe in Section~\ref{sec:task_coalescing}, is motivated by the above observations.
+\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} If a
+programmer writes a task that consumes {\em more} energy than the device can
+buffer, the task will never complete using buffered energy. {\em preventing
+forward progress} by repeatedly re-executing. If a programmer writes tasks that
+consume far {\em less} energy than a device can buffer, the system may operate
+{\em inefficiently}. When a task and its successor both complete, the two tasks
+were interleaved by a task boundary that preserves progress and state (i.e.,
+checkpoint or commit). However, absent a power failure, the work to preserve
+state was {\em unnecessary}, incurring overhead on the tasks' executions.
+Avoiding excessively costly, non-terminating tasks and short, high-overhead
+tasks is a programming challenge, given fixed hardware with a fixed energy
+buffering capacity. Buffer sizes from prior work vary widely from $20\mu F$~\cite{rodriguez_tbcs_2015}
+to $0.1F$~\cite{moo}.
+Sizing tasks complicates 
+{\em porting} code from one device to another. An excessively costly task 
+on a device with a small energy buffer may be a relatively short,
+task on a device with a larger buffer. The
+key problem is that using existing systems, the programmer statically sizes tasks for a fixed energy buffer. \sys's {\em task coalescing}, described
+in Section~\ref{sec:task_coalescing}, is motivated by these observations.
 

### 2017-08-11 paper/evaluation.tex
fixed table name
work on evaluation

--- 2017-08-11 paper/evaluation.tex
@@
 	\begin{tabular}{|l|cc|cc|cc|cc|c|}
+		\hline
+		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
+		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
 		\hline\hline
-		{} & \multicolumn{2}{l|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
-		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
-		\hline
 		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
@@
 		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
-		dft* & - & - & - & - & - & - & 222 & 293 & -\\
-		dd &  &  &  &  &  &  &  & 287 & $ \\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & & & $<$1\\
 		\hline
 	\end{tabular}
-	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand). *DFT does not compile with Clang}
+	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand).\todo{Data for dd, dft, sort missing}{Kiwan}}
 \label{table:compiler_result}
@@
 	\includegraphics[width=\columnwidth]{figures/coalescing}
-	\caption{\sys coalescing performance measured per application at fixed power and various WISP to RFID antenna distances, \{20, 40, 60\}\,cm, assessed against non-coalescing \sys.}
+	\caption{\sys coalescing performance measured per application at fixed power and various WISP to RFID antenna distances (\{20, 40, 60\}\,cm) assessed against non-coalescing \sys.}
 	\label{fig:coalescing}
@@
 
-We measured \sys coalescing performance.
-Result is provided in Figure~\ref{fig:coalescing}.
-Result clearly show that coalescing provides a benefit for all applications, providing even seven times improvement for bc application at 40\,cm.
+An important benchmark is to measured the effectiveness of \sys coalescing.
+At fixed power and varying WISP to RFID antenna distances we compared \sys execution time per each application (tasked manually).
+Result, given in Figure~\ref{fig:coalescing}, clearly illustrates that coalescing provides a benefit for all the applications, providing even seven times improvement (bc application at 40\,cm).
+bc application is composed of small tasks, see Figure~\ref{fig:aveVirtuTaskSize}, which \sys can easily coalesce.
+On the other hand, for certain applications, e.g.
+cem, coalescing improvement is not significant.
+cem uses large data structure that expands over multiple pages, which does not leave room for other variables.
+Any interaction will induce large page swap which cause little improvement in \sys performance against non-coalescing case.
 
@@
 
-We need to know how many tasks \sys can coalesce.
-We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and amount of tasks coalesced.
+Then, we need to know how many tasks \sys can coalesce within a given execution scenario.
+We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and the amount of tasks coalesced.
 Result is given in Figure~\ref{fig:aveVirtuTaskSize}.
@@
 This result clearly shows that for the benefit of coalescing and for code portability, initial code should be split by \emph{as small tasks as possible}.
+This will help \sys to find the best possible virtual task size to maximize its runtime.
 

### 2017-08-11 paper/evaluation.tex

@@
 		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & & & $<$1\\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
+		dft* & - & - & - & - & - & - & 222 & 293 & -\\
+		dd &  &  &  &  &  &  &  & 287 &  \\
 		\hline
 	\end{tabular}
-	\caption{Comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Hand).\todo{Data for dd, dft, sort missing}{Kiwan}}
+	\caption{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns).
+*DFT does not compile with Clang\todo{incomplete table}{Responsible:Kiwan}}
 \label{table:compiler_result}

### 2017-08-11 paper/compiler.tex
added compiler figure

@@
 	\centering
-	\subfloat[Plain C code]{\includegraphics[width=0.3\columnwidth]{figures/compiler1.pdf}\label{fig:comp_1}}
-	\subfloat[Boundary placement]{\includegraphics[width=0.3\columnwidth]{figures/compiler2.pdf}\label{fig:comp_2}}
-	\subfloat[Code Transformation]{\includegraphics[width=0.3\columnwidth]{figures/compiler3.pdf}\label{fig:comp_3}}
-	\caption{Illustration of \sys compiler task division operation.\todo{Make figures of the same readable size}{Przemek}}
+	\includegraphics[width=.85\columnwidth]{figures/compiler.pdf}
+	\caption{\sys compiles code to use tasks.}
 	\label{fig:compiler_overview}
@@
 
-The input to the task boundary identification analysis is a standard C program (as shown in Figure~\ref{fig:comp_1}) and the output of \sys's analysis is the CFG, with some vertices marked as {\em task boundaries}.
+The input to the task boundary identification analysis is a standard C program (as shown in Figure~\ref{fig:compiler_overview}) and the output of \sys's analysis is the CFG, with some vertices marked as {\em task boundaries}.
 The \sys compiler builds the program's CFG (as shown in
-Figure~\ref{fig:comp_2}).
+Figure~\ref{fig:compiler_overview}).
 To simplify inter-procedural analysis, our prototype inlines all function calls into the main function in the CFG.
@@
 \begin{equation}
-\text{score} = \frac{w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}}{N_{\text{loop}}},\nonumber
+\text{score} = {\left(w_{0} N_{\text{IR}} - w_{1} N_{\text{PV}}\right)}{N_{\text{loop}}^{-1}},\nonumber
 \end{equation}
@@
 If this dominance condition is unmet for a CFG node, the compiler places an additional task boundary at that node, meeting the condition.
-Figure~\ref{fig:comp_2} illustrates boundary placement: the first three vertices make up one task and the remainder become a second task.
+Figure~\ref{fig:compiler_overview} illustrates boundary placement: the first three vertices make up one task and the remainder become a second task.
 
@@
 After forming tasks, variables that are live beyond the end of the task become protected, which requires renaming those variables and updating references to renamed variable names.
-For example, Figure~\ref{fig:comp_2} shows a boundary placed within the live range of {\tt b}; {\tt b} becomes protected and accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in Figure~\ref{fig:comp_3}).
+For example, Figure~\ref{fig:compiler_overview} shows a boundary placed within the live range of {\tt b}; {\tt b} becomes protected and accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in Figure~\ref{fig:compiler_overview}).
 {\tt a} is task-local and need not be protected, but only declared in the task.

### 2017-08-11 paper/evaluation.tex

@@
 Apps such as RSA whose code size is large and uses a lot of variables, need longer compilation time due to a larger search space for pointer aliasing.
+Nonetheless, the compile time is orders of magnitude shorter than manual task splitting by a programmer (in  order of hours).
+For this reason we believe that \sys complier will be a major step in enabling computing community to enter the intermittent-power domain without big time investments.
 
@@
 	\includegraphics[width=\columnwidth]{figures/coala_chain}
-	\caption{Performance of \sys's fixed power and at multiple WISP to RFID reader antenna distances, \{20, 40, 60\}\,cm, compared against Chain.}
+	\caption{Performance of \sys's fixed power and at multiple WISP to RFID reader antenna distances, \{20, 40, 60\}\,cm, compared against Chain (results are normalized).}
 	\label{fig:coalescing}
@@
 
-Experimental evaluation of task coalescing mechanism under intermittent power of \sys is given in Figure~\ref{fig:coalescing}.
+We evaluate \sys experimentally to see its performance compared against Chain.
+Result is given in Figure~\ref{fig:coalescing}.
 For each application we measured execution time (with and without task coalescing) at \{20, 40, 60\}\,cm distance between RFID antenna and WISP and when WISP was powered directly from FET.
-Each measurement was repeated x times per each distance/application.
-From the result we clearly see an enormous benefit of coalescing, especially at large WISP to RFID reader antenna distance ($\approx$56\% for cem, 100\% for sort and $\approx$400\% for bc).
-At each distance, \sys coalescing mechanism provided faster execution times compared to the non-coalescing case.
-\todo{Discuss different shapes per app}{Amjad}
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\columnwidth]{figures/virtualTaskSize.eps}
-%	\caption{Execution time of a dummy application containing twelve empty tasks as a function of coalescing factor.
-We observe three times improvement from coalescing until four virtual tasks, while we observe a diminishing return from task coalescing beyond four.
-\todo{Improve the figure: make it narrower, larger font, remove legend, do not capitalize "Task" in X axis, add Y axis label (Execution time (ms))}{Amjad}}
-%	\label{fig:virtualTaskSize}
-%\end{figure}
-
-%The question remains: does coalescing all program's tasks into a singe virtual task (as the energy becomes fully available, i.e.
-from energy harvesting to continuous power environment) is the best one?
-To answer this question we have implemented a simple program composed of $x$ empty tasks.
-We execute the same program by increasing coalescing factor by one while the device on which the program was executed was running on a stable power supply (refer to Section~\ref{sec:methodology_evaluation} for details).
-The result is presented in Figure~\ref{fig:virtualTaskSize}.
-We observe a significant increase in execution time as task are coalescent from one to four (three times execution time improvement).
-However, as the coalescing increases beyond four, we observe diminishing returns from coalescing in terms of execution time.
-\todo{Decide whether to place this figure in Evaluation section or leave it here}{Przemek}
+Each measurement was for a predefined duration (each application was looping continuously) each distance/application case.
+From the result we clearly see an enormous benefit of \sys, especially at large WISP to RFID reader antenna distance.
+At each distance, \sys coalescing mechanism provided faster execution times compared to Chain.
 

### 2017-08-11 paper/paging.tex
cut DMA fig

@@
 
-\subsection{Managing Paging Overhead}
-
-The efficiency of \sys's paging scheme relies on the fact that data in pages are contiguous, which makes them amenable to being manipulated by hardware assisted DMA bulk copy operations.
+%\subsection{Managing Paging Overhead}
+%
+%The efficiency of \sys's paging scheme relies on the fact that data in pages are contiguous, which makes them amenable to being manipulated by hardware assisted DMA bulk copy operations.
 To justify this design choice we experimentally evaluated the efficiency of using DMA block copies, compared to explicit, software copy loops.
@@
 \sys's design moves data in pages, not individual words, leveraging the increased efficiency of DMA data movement.
-
-\begin{figure}[t]
-	\centering
-	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
-	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
-	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
-	\label{fig:dmaTimeEnergy}
-\end{figure}
+%
+%\begin{figure}[t]
+%	\centering
+%	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
+%	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
+%	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
+%	\label{fig:dmaTimeEnergy}
+%\end{figure}

### 2017-08-11 paper/evaluation.tex

@@
 This set of benchmarks is the same as used in~\cite{chain,alpaca}.
-In addition to Chain benchmarks new applications were implemented: (vi) \textbf{dft}: discreet Fourier transform, (vii) \textbf{dd}: Huffman-based data decompression, and (viii) \textbf{sort}: selection sort algorithm.
-All applications are compared in Table~\ref{table:benchmark_list} in terms of source lines of code against Chain.
+In addition to Chain benchmarks new applications were implemented: (vi) \textbf{dft}: discreet Fourier transform, and (vii) \textbf{sort}: selection sort algorithm.
+All applications are compared in Table~\ref{table:compiler_result} in terms of source lines of code against Chain.
 
@@
 		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
-		dft* & - & - & - & - & - & - & 222 & 293 & -\\
-		dd &  &  &  &  &  &  &  & 287 &  \\
+		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
+		%dd &  &  &  &  &  &  &  & 287 &  \\
 		\hline
@@
 	\caption{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns).
-*DFT does not compile with Clang\todo{incomplete table}{Responsible:Kiwan}}
+$^\dagger$dft does not compile with Clang}
 \label{table:compiler_result}
@@
 	\includegraphics[width=\columnwidth]{figures/coala_chain}
-	\caption{Performance of \sys's fixed power and at multiple WISP to RFID reader antenna distances, \{20, 40, 60\}\,cm, compared against Chain (results are normalized).}
+	\caption{Performance of \sys applications (divided manually into tasks) at fixed power and at multiple WISP to RFID reader antenna distances, \{20, 40, 60\}\,cm, compared against Chain (results are normalized).}
 	\label{fig:coalescing}
@@
 Each measurement was for a predefined duration (each application was looping continuously) each distance/application case.
-From the result we clearly see an enormous benefit of \sys, especially at large WISP to RFID reader antenna distance.
-At each distance, \sys coalescing mechanism provided faster execution times compared to Chain.
+From the result we clearly see an enormous benefit of \sys, for applications composed of small tasks (bc, sort, cuckoo)---at all distances and at fixed power.
+As the individual task become larger (e.g.
+dft) or program uses large memory pages (e.g.
+cem, refer to Section~
+\ref{sec:result_compiler_time} for further explanation) \sys incurs overhead from memory virtualization that cause its performance to be comparable (or in a single cases, cem, worse) to Chain.
 
@@
 	\includegraphics[width=\columnwidth]{figures/comp_user}
-	\caption{\sys compiler overhead measured per application.}
+	\caption{\sys compiler code overhead measured in terms of execution time per application; WISP was powered continuously.}
 	\label{fig:comp_user}
@@
 
-Result is provided in Figure~\ref{fig:comp_user}.
+In Table~\ref{table:compiler_result} we have assessed the overhead induced by the \sys compiler.
+It is, nonetheless, important to find out how much does this overhead translates to increased execution time for each \sys application.
+The result of this investigation is provided in Figure~\ref{fig:comp_user}.
+We conclude that in most of the cases \sys compiler code provided comparable execution time to
 

### 2017-08-11 paper/intro.tex
added anonymized link

@@
 
-In this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer.
+In this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Available via \href{http://anonymized.link}{http://anonymized.link}.}.
 To accomplish this, \sys introduces new capabilities:
@@
 We evaluated \sys end-to-end on a collection of existing benchmarks ~\cite{chain} and new programs, and compare directly to a state-of-the-art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without rewriting code.
-%\todo{add detailed text about evaluation results}{Przemek}
 

### 2017-08-11 paper/evaluation.tex
small fixes
evaluation section corrections

@@
 We used internal timer for measuring run time.
-For harvested energy, we used the ThingMagic Astra-EX RFID antenna set at 30\,dBm transmission power.
+For harvested energy, we used the ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power.
 We used EDB~\cite{edb} to measure run time.
@@
 This set of benchmarks is the same as used in~\cite{chain,alpaca}.
-In addition to Chain benchmarks new applications were implemented: (vi) \textbf{dft}: discreet Fourier transform, and (vii) \textbf{sort}: selection sort algorithm.
+In addition to Chain benchmarks new applications were implemented: (vi) \textbf{sort}: selection sort algorithm, and (vii) \textbf{dft}: discreet Fourier transform.
 All applications are compared in Table~\ref{table:compiler_result} in terms of source lines of code against Chain.
@@
 
-Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile the DFT application since LLVM inserts mathematical built-in functions (i.e.
+Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile the dft application since LLVM inserts mathematical built-in functions (i.e.
 \texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}.
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coala_chain}
-	\caption{Performance of \sys applications (divided manually into tasks) at fixed power and at multiple WISP to RFID reader antenna distances, \{20, 40, 60\}\,cm, compared against Chain (results are normalized).}
-	\label{fig:coalescing}
+	\includegraphics[width=\columnwidth]{figures/coala_chain_clang}
+	\caption{Performance of \sys applications (divided manually into tasks) at fixed power (left-most bar per application) and at multiple WISP to RFID reader antenna distances, \{20, 40\}\,cm (center and right-most bar per application) compared against Chain (results are normalized).}
+	\label{fig:runtime}
 \end{figure}
 
-We evaluate \sys experimentally to see its performance compared against Chain.
-Result is given in Figure~\ref{fig:coalescing}.
-For each application we measured execution time (with and without task coalescing) at \{20, 40, 60\}\,cm distance between RFID antenna and WISP and when WISP was powered directly from FET.
+We evaluate \sys experimentally and compare its performance against Chain.
+Result is given in Figure~\ref{fig:runtime}.
+For each application we measured execution time (with and without task coalescing) at \{20, 40\}\,cm distance between RFID antenna and WISP, and when WISP was powered directly from FET.
 Each measurement was for a predefined duration (each application was looping continuously) each distance/application case.
@@
 	\includegraphics[width=\columnwidth]{figures/comp_user}
-	\caption{\sys compiler code overhead measured in terms of execution time per application; WISP was powered continuously.}
+	\caption{\sys compiler code overhead measured in terms of execution time per application (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
 	\label{fig:comp_user}
@@
 The result of this investigation is provided in Figure~\ref{fig:comp_user}.
-We conclude that in most of the cases \sys compiler code provided comparable execution time to
+We conclude that in most of the cases \sys compiler code provided comparable execution time to code split into tasks manually (except for ar and rsa applications).
 
@@
 	\includegraphics[width=\columnwidth]{figures/coalescing}
-	\caption{\sys coalescing performance measured per application at fixed power and various WISP to RFID antenna distances (\{20, 40, 60\}\,cm) assessed against non-coalescing \sys.}
+	\caption{\sys coalescing performance measured per application (left-most bar per application) and at multiple WISP to RFID reader antenna distances, \{20, 40\}\,cm (center and right-most bar per application) assessed against non-coalescing \sys.}
 	\label{fig:coalescing}
@@
 This result clearly shows that for the benefit of coalescing and for code portability, initial code should be split by \emph{as small tasks as possible}.
-This will help \sys to find the best possible virtual task size to maximize its runtime.
+This will help \sys to find the best possible virtual task size to minimise its runtime.
 
@@
 	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
-	\caption{Normalized page size overhead and page faults per benchmark.\todo{Figure placeholder---will be replaced with final one}{Amjad}}
+	\caption{Normalized page size overhead and page faults per benchmark.}
 	\label{fig:IPOSPerformance}
@@
 Overhead has been measured in the MCU clock cycles with values provided by TI CCS.
-Result has been normalized per benchmark: for the overhead results---to the minimum obtained value and for page faults---to the third measured maximum value\todo{Check again}{Amjad}.
+Result has been normalized per benchmark: for the overhead results---to the minimum obtained value and for page faults---to the third measured maximum value.
 A subset of applications has been selected for evaluation.
@@
 for Cuckoo 8 (at 16 bytes) to 1 (at 64 bytes), BC: 1 (at 16 bytes) to $\approx$2 (at 1024 bytes)).
-\todo{Explain this phenomenon}{Amjad}
 

### 2017-08-11 paper/paging.tex
evaluation section corrections

@@
 		\If {\texttt{isDirty(tag)}} \Comment{Check if the requested page has been modified}
-		\State {\texttt{getPage(tag,pagesTemp)}=false} \Comment{Get page from the temporary buffer}
+		\State {\texttt{getPage(tag,pagesTemp)}} \Comment{Get page from the temporary buffer}
 		\Else
-		\State \texttt{getPage(tag,pagesOrg)}=false \Comment{Get page from the original location}
+		\State \texttt{getPage(tag,pagesOrg)} \Comment{Get page from the original location}
 		\EndIf 

### 2017-08-11 paper/paging.tex
paging

--- 2017-08-11 paper/paging.tex
@@
 Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task using \sys's paging mechanism.
-The task manipulates protected variables {\tt x}, {\tt y}, and {\tt z}, which are in the same page.
-On accessing {\tt x}, its page is swapped into the page buffer.
-On accessing {\tt y} and {\tt z}, the task need only access the volatile page buffer.
-The task updates the page, leaving it dirty at the task's completion.
-When the task completes, \sys first copies the page from the page buffer into the non-volatile {\tt pagesTemp} region before commit.
-Then, \sys sets the commit bit, atomically commits the page back to its original memory location, and unsets the commit bit.
+The task manipulates the protected variables {\tt x}, {\tt y}, and {\tt z}.
+The variables {\tt y} and {\tt z} are in the same page and they are currently in the volatile page buffer---on accessing {\tt y} and {\tt z}, the task needs only to access \texttt{pageBuf}.
+Since variable \texttt{x} is kept in a different page in \texttt{pagesOrg}, accessing it creates a page fault.
+At this point, \sys swaps the active page in \texttt{pageBuf} by committing it into the non-volatile {\tt pagesTemp} region and copying the corresponding page from \texttt{pagesOrg} to \texttt{pageBuf}.
+When the task completes, \sys sets the commit bit, copies the dirty page from \texttt{pagesTmp} to its original location in \texttt{pagesOrg} and unsets the commit bit.
 

### 2017-08-11 paper/coalescing.tex

@@
 
-\noindent\textbf{Task Coalescing Implementation:} Figure~\ref{fig:coalescing_example} illustrates how \sys coalesces tasks.
-Figure~\ref{fig:coalescing_example_states} (left) shows a sequence of three tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.
+\noindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example_states} (left) shows a sequence of three tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.
 The same figure shows the same sequence of tasks executing as coalesced groups of three tasks $C=2$ (center) and $C=3$ (right).

### 2017-08-11 paper/intro.tex

--- 2017-08-11 paper/intro.tex
@@
 
-In this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer.
+In this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
 To accomplish this, \sys introduces new capabilities:

### 2017-08-11 paper/systemdescription.tex

@@
 	\end{tabular}
-	\caption{Summary of \sys's programming model extensions.\todo{Revise this table (confusion with next\_task() function)}{Amjad/Przemek}}
+	\caption{Summary of \sys's programming model extensions.}
 	\label{tab:viper_syntax}
@@
 Relying on programmer or compiler annotation of accesses to memory locations that might be shared between tasks, \sys's memory virtualization {\em pages} these data to SRAM from FRAM as they are needed.
-Tasks operate, accessing SRAM only, demand-paging data to accommodate a SRAM smaller than FRAM\todo{Sentence unclear}{Brandon}.
+Tasks operate, accessing SRAM only, demand-paging data to accommodate a SRAM smaller than FRAM.
 When a task completes, \sys commits modified, paged data back to its original FRAM location atomically ensuring subsequent tasks page in correct memory state.

### 2017-08-11 paper/evaluation.tex
added WISP info for CMU
fix typo

@@
 We used internal timer for measuring run time.
-For harvested energy, we used the ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power.
+For harvested energy, we used WISP\,5.0~\cite{wisp5,wisp} hardware and ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power.
 We used EDB~\cite{edb} to measure run time.
@@
 
-We start with evaluating the efficiency of the \sys compiler Table~\ref{table:compiler_result} shows the difference in the number of protected variables (PVAR), number of tasks, and number of inserted paging macro ({\tt RVAR} and {\tt WVAR}) in the code between (i) hand-written \sys and (ii) compiler-generated \sys code.
+We start with evaluating the efficiency of the \sys compiler.
+Table~\ref{table:compiler_result} shows the difference in the number of protected variables (PVAR), number of tasks, and number of inserted paging macro ({\tt RVAR} and {\tt WVAR}) in the code between (i) hand-written \sys and (ii) compiler-generated \sys code.
 On average, compiler-generated code is only worse by 0.3\% in the number of protected variables.

### 2017-08-11 paper/evaluation.tex
specified that compile time is in sec
2 figures replaced

@@
 We compiled the apps on a 4-core virtual machine with 1\,GB memory with Arch Linux, running on Intel Core i5-6200U (2.30\,GHz) processor.
-The result are given in Table~\ref{table:compiler_result}.
+The result in seconds are given in Table~\ref{table:compiler_result}.
 We used the {\tt time} linux command to measure the rough time.
@@
 The shape of each figure can be explained as follows.
-For small page sizes (<64) the execution time is dominated by the \emph{frequency} of page swaps (each swap takes non-zero time, as shown in Figure~\ref{fig:dmaTimeEnergy}).
+For small page sizes (<64) the execution time is dominated by the \emph{frequency} of page swaps.
 On the other hand, in the opposite case the execution time is dominated by the time of swapping of one large page from SRAM to FRAM.

### 2017-08-11 paper/coalescing.tex
added figure for coalescing.  need to fix text

@@
 	\centering
-	\subfloat[Core of task coalescing: commit postponing]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
-	\subfloat[Adaptive task coalescing based on task failure history]{\includegraphics[width=\columnwidth]{figures/coalescing_flowchart}\label{fig:coalescing_example_flowchart}}
-	\caption{\sys task coalescing process.}
+	\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
+	\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
+	\caption{\sys coalesces tasks to avoid overheads.}
 	\label{fig:coalescing_example}

### 2017-08-11 paper/evaluation.tex
evaluation updated (fig 10 into table)

@@
 
-\begin{figure}
+\begin{table}
 	\centering
-	\includegraphics[width=\columnwidth]{figures/averageVirtualTaskSize}
-	\caption{Average \sys virtual task size at fixed power per application. Top (green): number of coalesced tasks; bottom (red): average size of real task.}
-	\label{fig:aveVirtuTaskSize}
-\end{figure}
+	\footnotesize
+	\begin{tabular}{|r|c|c|c|}
+		\hline
+		{} & bc & dft & cem \\
+		\hline\hline
+		Avg. no. coalesced tasks & 56 & 3 & 26 \\
+		Avg. task size (MCU cycles) & 657 & 9680 & 722 \\
+		\hline
+	\end{tabular}
+\caption{Average \sys virtual task size (at fixed power) per manually task split application.}
+\label{tab:aveVirtuTaskSize}
+\end{table}
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\columnwidth]{figures/averageVirtualTaskSize}
+%	\caption{Average \sys virtual task size at fixed power per application. Top (green): number of coalesced tasks; bottom (red): average size of real task.}
+%	\label{fig:aveVirtuTaskSize}
+%\end{figure}
 
@@
 We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and the amount of tasks coalesced.
-Result is given in Figure~\ref{fig:aveVirtuTaskSize}.
+Result is given in Table~\ref{tab:aveVirtuTaskSize}.
 We see clearly that \sys manages to coalesce more tasks as individual tasks are small.

### 2017-08-11 paper/evaluation.tex
tweaks to system and evaluation

@@
 Result, given in Figure~\ref{fig:coalescing}, clearly illustrates that coalescing provides a benefit for all the applications, providing even seven times improvement (bc application at 40\,cm).
-bc application is composed of small tasks, see Figure~\ref{fig:aveVirtuTaskSize}, which \sys can easily coalesce.
+bc application is composed of small tasks, see Table~\ref{tab:aveVirtuTaskSize}, which \sys can easily coalesce.
 On the other hand, for certain applications, e.g.
@@
 	\end{tabular}
-\caption{Average \sys virtual task size (at fixed power) per manually task split application.}
+\caption{Average \sys virtual task size (at 60\,cm WISP to RFID antenna distance) per manually task split application.}
 \label{tab:aveVirtuTaskSize}

### 2017-08-12 paper/evaluation.tex
methodology patches

@@
 
-\subsection{Hardware Suite Description}
+\subsection{Methodology}
 \label{sec:results_hardware}
 
-In all experiments we used the following computing hardware: WISP\,5.1~\cite{wisp5,wisp} build around MSP430FR5969~\cite{wolverine} 64\,kB FRAM microcontroller.
-WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet} with TI Code Composer Studio version 7.
-FET was also used as a power supply in non-intermittent power experiments.
-Energy consumption measurements provided in eariler sections were done with Carnegie Mellon's EDB~\cite{edb}.
-In experiments involving wireless power, the WISP was powered through an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
-The RFID reader is controlled by a PC running Ubuntu 10.4 executing an open source Python-based \emph{sllurp} library~\cite{sllrp_github}.
-Transmitted power of the RFID reader was 30\,dBm at 915\,MHz center frequency.
-For distance-controlled experiments, the WISP was positioned at varying length hollow paper tubes placed directly on a flat lying antenna.
-Code execution times were measured using Salea~\cite{saleae} logic analyzer.
+We evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp} energy-harvesting
+platform build around a MSP430FR5969~\cite{wolverine} MCU with 64kB of FRAM.
+We used a TI Flash Emulation Tool (FET)~\cite{fet} as a power supply during
+experiments with continuous power.  For energy measurements, we used
+EDB~\cite{edb}.  In experiments with wireless power, we powered the WISP using
+an RFID reader (Impinj R1000 with firmware version
+3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ
+8\,dBic~\cite{atlas2015}. We ccontrolled the RFID reader from a workstation
+running Ubuntu 10.4 using \emph{sllurp}~\cite{sllrp_github}.  We set the
+transmitted power of the RFID reader to 30\,dBm at 915\,MHz center frequency.
+For distance-controlled experiments, we affixed the WISP at varying lengths to
+hollow paper tubes placed flat on a lying antenna. We did timing measurements
+using the Salea~\cite{saleae} logic analyzer.
 
-We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used for continuous power experiments.
-We used internal timer for measuring run time.
-For harvested energy, we used WISP\,5.0~\cite{wisp5,wisp} hardware and ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power.
-We used EDB~\cite{edb} to measure run time.
-Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler.
-Version 3.8.0 was used for both Clang and LLVM.
+We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used
+for continuous power experiments. We used internal timer for measuring run
+time. For harvested energy, we used WISP\,5.0~\cite{wisp5,wisp} hardware and
+ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power.
+We used EDB~\cite{edb} to measure run time. Applications were cross-compiled
+using GCC and Clang, so that we can use LLVM for \sys compiler. Version 3.8.0
+was used for both Clang and LLVM.
 

### 2017-08-12 paper/evaluation.tex

@@
 Result, given in Figure~\ref{fig:coalescing}, clearly illustrates that coalescing provides a benefit for all the applications, providing even seven times improvement (bc application at 40\,cm).
-bc application is composed of small tasks, see Table~\ref{tab:aveVirtuTaskSize}, which \sys can easily coalesce.
+bc application is composed of small tasks, which \sys can easily coalesce.
 On the other hand, for certain applications, e.g.
@@
 
-\begin{table}
-	\centering
-	\footnotesize
-	\begin{tabular}{|r|c|c|c|}
-		\hline
-		{} & bc & dft & cem \\
-		\hline\hline
-		Avg. no. coalesced tasks & 56 & 3 & 26 \\
-		Avg. task size (MCU cycles) & 657 & 9680 & 722 \\
-		\hline
-	\end{tabular}
-\caption{Average \sys virtual task size (at 60\,cm WISP to RFID antenna distance) per manually task split application.}
-\label{tab:aveVirtuTaskSize}
-\end{table}
+%\begin{table}
+%	\centering
+%	\footnotesize
+%	\begin{tabular}{|r|c|c|c|}
+%		\hline
+%		{} & bc & dft & cem \\
+%		\hline\hline
+%		Avg. no. coalesced tasks & 56 & 3 & 26 \\
+%		Avg. task size (MCU cycles) & 657 & 9680 & 722 \\
+%		\hline
+%	\end{tabular}
+%\caption{Average \sys virtual task size (at 60\,cm WISP to RFID antenna distance) per manually task split application.}
+%\label{tab:aveVirtuTaskSize}
+%\end{table}
 
@@
 
-Then, we need to know how many tasks \sys can coalesce within a given execution scenario.
+%Then, we need to know how many tasks \sys can coalesce within a given execution scenario.
 We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and the amount of tasks coalesced.

### 2017-08-12 paper/coalescing.tex

@@
 \sys dynamically {\em coalesces} consecutively executing tasks at run time.
-When two tasks are coalesced, the first of the two tasks does not commit its updates to protected memory locations when it reaches a \transition statement.
-Instead, execution immediately continues from the beginning of the second of the two tasks.
-At the end of the second of the two coalesced tasks, \sys commits the pages that were updated by either of the coalescing tasks.
-The purpose of coalescing is to better amortize the run time overhead of committing updated pages when a task ends.
-If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed once.
+When two tasks are coalesced, the first of the two tasks does not commit its
+updates to protected memory locations when it reaches a \transition statement.
+Instead, execution immediately continues from the beginning of the second of
+the two tasks. At the end of the second of the two coalesced tasks, \sys
+commits the pages that were updated by either of the coalescing tasks. The
+purpose of coalescing is to better amortize the run time overhead of committing
+updated pages when a task ends. If consecutive tasks access the same pages of
+data, the pages accessed by both tasks only need to be committed once.
 
-\sys can coalesce an arbitrary number of consecutive tasks.
-As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance.
-As more tasks coalesce, the amount of potentially {\em wasted work} also increases.
-If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of the tasks in the long sequence.
-\sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, \emph{adaptively} determining a coalescing factor that combines many tasks without suffering an excess of wasted work.
+\sys can coalesce an arbitrary number of consecutive tasks. As more tasks
+coalesce, their collective commit overhead amortizes better, potentially
+improving performance. As more tasks coalesce, the amount of potentially {\em
+wasted work} also increases. If power fails during a long sequence of coalesced
+tasks the lack of an intervening commit requires execution to restart from the
+{\em first} task in the sequence, failing to preserve the progress of any of
+the tasks in the long sequence. \sys moderates the risk of coalescing and
+capitalizes on its benefit by dynamically, \emph{adaptively} determining a
+coalescing factor that combines many tasks without suffering an excess of
+wasted work.
 
-The challenge to coalescing tasks is determining how many tasks to coalesce before committing.
-During execution, an executing task is either a {\em coalesced} task, which does not commit its state before transitioning or a {\em committing} task, which does commit its state before transitioning.
-\sys monitors a program's execution to adaptively determine $C$, the number of tasks to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks, followed by a committing task.
+The challenge to coalescing tasks is determining how many tasks to coalesce
+before committing. During execution, an executing task is either a {\em
+coalesced} task, which does not commit its state before transitioning or a {\em
+committing} task, which does commit its state before transitioning. \sys
+monitors a program's execution to adaptively determine $C$, the number of tasks
+to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks,
+followed by a committing task.
 
-\sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the last operational period.
-\sys begins execution in a coalesced task with an infinite $C$ value.
-With a continuous power supply, the system executes from the start with all tasks as coalesced tasks entirely avoiding the cost of committing state.
-On intermittent power, the system executes a sequence of tasks that is interrupted by a power failure.
-\sys explicitly tracks $S$ during the previous operating period.
-\sys assigns $C$ to be equal to half of the value of $S$.
-If more tasks in total are able to complete in an operating period, \sys attempts to coalesce more tasks during the next operating period.
-If fewer tasks complete, \sys attempts to coalesce fewer tasks during the next operating period.
-$S$ includes the count of both coalesced and committing tasks completed.
-If no committing task completes, i.e., execution does not make progress, \sys updates $S$ and attempts to coalesce fewer tasks during the next operating period.
+\sys determines $C$ for a given operation based on the number of tasks
+completed, $S$, during the current operational period. \sys begins execution in
+a coalesced task with $C$ initialized to half of the number of tasks executed
+during the previous operating period and $S$ initialized to $0$. The system
+executes $C-1$ coalesced tasks followed by one committing task. After each
+commit, \sys assigns $C$ to half of $S$.   If more tasks complete in an
+operating period, \sys attempts to coalesce more tasks. If fewer tasks
+complete, \sys attempts to coalesce fewer tasks. If no committing task
+completes due to a power failure, \sys updates $C$ on reboot to half of $S$
+(the number of tasks completed before the power failure) and continues.
 
@@
 
-\noindent\textbf{Task Coalescing Illustration:} Figure~\ref{fig:coalescing_example_states} (left) shows a sequence of three tasks executing in \sys with no coalescing, incurring the cost of commit at each task transition.
-The same figure shows the same sequence of tasks executing as coalesced groups of three tasks $C=2$ (center) and $C=3$ (right).
-Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$.
-\sys executes tasks in the first operating period, assigning the \emph{half of sum of all previous $S$ periods} (in the example case: $C=8/2$).
-\sys completes then coalesced tasks successfully (four tasks were executed) and in the next operating period, i.e.
-$C=(8+4)/2$.
-In the third task execution, power fails after the second task.
-Then \sys re-executes coalesced task since the previous failure and sets $C=(6+8+2)/2$ in the next operating period.
-If a larger number of tasks successfully complete in the next operating period, i.e., $S$ is larger, $C$ correspondingly increases, allowing more aggressive coalescing.
+\noindent\textbf{Task Coalescing Illustration:}
+Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks
+executing in \sys. At left, there is no coalescing and tasks incur the cost of
+commit at each transition. The center shows the same tasks executing as
+coalesced groups of two tasks $C=2$, avoiding one commit.  At right, \sys
+coalesces three tasks ($C=3$) avoiding two commits.
+Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value
+of $C$ based on $S$. \sys executes tasks in the first operating period, with
+$C$ equal to half the number of tasks completed in the prior execution period.
+In the example, $C$ starts at eight. After completing eight tasks, \sys commits
+and reduces $C$ by half to four.  \sys completes two of four tasks and power
+fails.  During the operating period, \sys executed a total of ten tasks (i.e.,
+$S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$).
+\sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$
+to half of $S$, which is three (after rounding up). \sys coalesces three tasks
+and again commits.  At this point, $S=8$; \sys assigns $C=4$ and execution
+continues.
 

### 2017-08-12 paper/evaluation.tex

@@
 
-\subsection{Methodology}
+\subsection{Hardware Suite Description}
 \label{sec:results_hardware}
 
-We evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp} energy-harvesting
-platform build around a MSP430FR5969~\cite{wolverine} MCU with 64kB of FRAM.
-We used a TI Flash Emulation Tool (FET)~\cite{fet} as a power supply during
-experiments with continuous power.  For energy measurements, we used
-EDB~\cite{edb}.  In experiments with wireless power, we powered the WISP using
-an RFID reader (Impinj R1000 with firmware version
-3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ
-8\,dBic~\cite{atlas2015}. We ccontrolled the RFID reader from a workstation
-running Ubuntu 10.4 using \emph{sllurp}~\cite{sllrp_github}.  We set the
-transmitted power of the RFID reader to 30\,dBm at 915\,MHz center frequency.
-For distance-controlled experiments, we affixed the WISP at varying lengths to
-hollow paper tubes placed flat on a lying antenna. We did timing measurements
-using the Salea~\cite{saleae} logic analyzer.
+In all experiments we used the following computing hardware: WISP\,5.1~\cite{wisp5,wisp} build around MSP430FR5969~\cite{wolverine} 64\,kB FRAM microcontroller.
+WISP\,5.1 has been re-programmed during experiments using TI's Flash Emulation Tool (FET)~\cite{fet} with TI Code Composer Studio version 7.
+FET was also used as a power supply in non-intermittent power experiments.
+Energy consumption measurements provided in eariler sections were done with Carnegie Mellon's EDB~\cite{edb}.
+In experiments involving wireless power, the WISP was powered through an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
+The RFID reader is controlled by a PC running Ubuntu 10.4 executing an open source Python-based \emph{sllurp} library~\cite{sllrp_github}.
+Transmitted power of the RFID reader was 30\,dBm at 915\,MHz center frequency.
+For distance-controlled experiments, the WISP was positioned at varying length hollow paper tubes placed directly on a flat lying antenna.
+Code execution times were measured using Salea~\cite{saleae} logic analyzer.
 
-We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used
-for continuous power experiments. We used internal timer for measuring run
-time. For harvested energy, we used WISP\,5.0~\cite{wisp5,wisp} hardware and
-ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power.
-We used EDB~\cite{edb} to measure run time. Applications were cross-compiled
-using GCC and Clang, so that we can use LLVM for \sys compiler. Version 3.8.0
-was used for both Clang and LLVM.
+We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used for continuous power experiments.
+We used internal timer for measuring run time.
+For harvested energy, we used WISP\,5.0~\cite{wisp5,wisp} hardware and ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power.
+We used EDB~\cite{edb} to measure run time.
+Applications were cross-compiled using GCC and Clang, so that we can use LLVM for \sys compiler.
+Version 3.8.0 was used for both Clang and LLVM.
 

### 2017-08-12 paper/evaluation.tex
partial commit to eval

--- 2017-08-12 paper/evaluation.tex
@@
 
-We also used an alternative setup where the MSP-TS430RGZ48C launchpad was used
-for continuous power experiments. We used internal timer for measuring run
-time. For harvested energy, we used WISP\,5.0~\cite{wisp5,wisp} hardware and
-ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power.
-We used EDB~\cite{edb} to measure run time. Applications were cross-compiled
-using GCC and Clang, so that we can use LLVM for \sys compiler. Version 3.8.0
-was used for both Clang and LLVM.
+For some experiments (noted below) we used an alternative setup where the
+MSP-TS430RGZ48C launchpad was used for continuous power experiments. We used
+internal timer for measuring run time. For harvested energy, we used
+WISP\,5.0~\cite{wisp5,wisp} hardware and ThingMagic Astra-EX integrated RFID
+reader set at 30\,dBm transmission power at 915\,MHz.  We used EDB~\cite{edb}
+to measure run time. Applications were cross-compiled using GCC and Clang,
+allowing us to use LLVM for the \sys compiler. We used version 3.8.0 of both
+Clang and LLVM.
 
@@
 We compare \sys against state-of-the-art task-based runtime Chain~\cite{chain}.
-We are not comparing \sys against DINO~\cite{dino} or Mementos~\cite{mementos} (Refer to Section~\ref{sec:background_consistency}), as (i) Chain is the only modular execution model available to which we can compare \sys, and (ii) Chain proved already its superiority over sequential model counterparts (i.e.
-DINO and Mementos)\footnote{We need to note that although another very relevant task-based runtime has been recently presented and referenced many times throughout this paper---Alpaca~\cite{alpaca} (written by the same group as Chain)---except the recently accepted manuscript we did not have the access to Alpaca's code during the preparation of this paper.
-Therefore Alpaca was excluded from the evaluation.}.
+We do not compare \sys to other, earlier systems, such as DINO~\cite{dino} or
+Mementos~\cite{mementos} (cf. Section~\ref{sec:background_consistency}), as
+Chain is the only task-based execution model available to which we can compare
+\sys.
 
-The applications that were used in the evaluation are: (i) \textbf{bc}: counts bits in byte pseudo-random string using varying algorithms.
-Each result is compared for correctness; (ii) \textbf{cem}: emulates an embedded temperature sensing engine: measurements are randomly generated and are compressed using a LZW algorithm for  fixed entry dictionary and fixed byte block; (iii) \textbf{cuckoo}: implements Cuckoo filter---the program first stores a sequence of pseudo-random numbers, then queries the filter to recover it; (iv) \textbf{rsa}: encrypts predefined string with a predefined key; (v) \textbf{ar} performs nearest neighbour classification using randomly generated sequence of three-axes accelerometer data.
-This set of benchmarks is the same as used in~\cite{chain,alpaca}.
-In addition to Chain benchmarks new applications were implemented: (vi) \textbf{sort}: selection sort algorithm, and (vii) \textbf{dft}: discreet Fourier transform.
-All applications are compared in Table~\ref{table:compiler_result} in terms of source lines of code against Chain.
+We evaluated \sys using benchmarks from the literature~\cite{chain,alpaca} and
+several new programs: (i) \textbf{bc}: counts bits in byte pseudo-random string
+using several algorithms, and compares results for correctness; (ii)
+\textbf{cem}, which emulates an embedded temperature sensing engine and in
+which measurements are randomly generated and compressed using LZW with a fixed
+entry dictionary and fixed byte block; (iii) \textbf{cuckoo}: which implements
+a Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random
+numbers, then queries the filter to recover the sequence; (iv) \textbf{rsa}:
+which encrypts a fixed string with a stored key; (v) \textbf{ar}: which
+performs nearest neighbour classification of randomly generated data modeling a
+three-axis accelerometer.  This set of benchmarks overlaps with those used
+in~\cite{chain,alpaca}. We implemented several new applications: (vi)
+\textbf{sort}: selection sort algorithm on a fixed sequence of numbers, and
+(vii) \textbf{dft}: discrete Fourier transform. All applications are compared
+in columns eight and nine of Table~\ref{table:compiler_result} showing the
+number of source lines of code compared to Chain.
 
@@
 \label{sec:results_evaluation}
+
+We evaluated \sys to show that it is easy to use and efficient across our
+benchmarks.  We evaluate \sys's ease of use by characterizing \sys's automatic
+task compiler.  We evaluate \sys efficiency by comparing its runtime
+performance to Chain.  We characterize \sys by studying how its performance
+varies due to coalescing and when using automatic task compilation versus
+manual task compilation.
 

### 2017-08-12 paper/evaluation.tex
pass over most of eval

@@
 
-\subsection{Software Suite Description}
+\subsection{Benchmarks and Baseline}
 \label{sec:results_software}
@@
 
-\subsubsection{\sys Compiler Performance}
-\label{sec:results_compiler}
-
-\begin{table}[t]
-	\centering
-	\footnotesize
-        \renewcommand{\tabcolsep}{1pt}
-	\begin{tabular}{|l|cc|cc|cc|cc|c|}
-		\hline
-		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
-		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
-		\hline\hline
-		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
-		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
-		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
-		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
-		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
-		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
-		%dd &  &  &  &  &  &  &  & 287 &  \\
-		\hline
-	\end{tabular}
-	\caption{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns).
-$^\dagger$dft does not compile with Clang}
-\label{table:compiler_result}
-\end{table}
-
-We start with evaluating the efficiency of the \sys compiler.
-Table~\ref{table:compiler_result} shows the difference in the number of protected variables (PVAR), number of tasks, and number of inserted paging macro ({\tt RVAR} and {\tt WVAR}) in the code between (i) hand-written \sys and (ii) compiler-generated \sys code.
-On average, compiler-generated code is only worse by 0.3\% in the number of protected variables.
-Regarding the number of tasks, compiler-generated code is better by roughly 16\%.
-The results indicate that \sys compiler is selecting boundaries better than the programmer doing it manually.
-However, the \sys compiler should insert paging code conservatively when it {\em may} access protected variables, due to limitation of pointer aliasing.
-Thus, it results in the \sys compiler introducing 48\% more paging code than programmer-written code (the results on run time for various \sys code versions is provided in Section~\ref{sec:result_compiler_time}).
-The performance is worse than programmer-written \sys code in overall, meaning the programmer should structure the code manually if the performance is critical; otherwise the compiler can be used to simplify the programmability.
-
-Additionally, we evaluated the compile time for automatic task generation per application\footnote{We could not compile the dft application since LLVM inserts mathematical built-in functions (i.e.
-\texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}.
-We compiled the apps on a 4-core virtual machine with 1\,GB memory with Arch Linux, running on Intel Core i5-6200U (2.30\,GHz) processor.
-The result in seconds are given in Table~\ref{table:compiler_result}.
-We used the {\tt time} linux command to measure the rough time.
-Most of the compile time was due to pointer aliasing, when calculating possible live ranges of variables.
-Apps such as RSA whose code size is large and uses a lot of variables, need longer compilation time due to a larger search space for pointer aliasing.
-Nonetheless, the compile time is orders of magnitude shorter than manual task splitting by a programmer (in  order of hours).
-For this reason we believe that \sys complier will be a major step in enabling computing community to enter the intermittent-power domain without big time investments.
-
 %\begin{table}[t]
@@
 
-We evaluate \sys experimentally and compare its performance against Chain.
-Result is given in Figure~\ref{fig:runtime}.
-For each application we measured execution time (with and without task coalescing) at \{20, 40\}\,cm distance between RFID antenna and WISP, and when WISP was powered directly from FET.
-Each measurement was for a predefined duration (each application was looping continuously) each distance/application case.
-From the result we clearly see an enormous benefit of \sys, for applications composed of small tasks (bc, sort, cuckoo)---at all distances and at fixed power.
-As the individual task become larger (e.g.
-dft) or program uses large memory pages (e.g.
-cem, refer to Section~
-\ref{sec:result_compiler_time} for further explanation) \sys incurs overhead from memory virtualization that cause its performance to be comparable (or in a single cases, cem, worse) to Chain.
+We evaluate \sys's run time performance and compare the results to the run time
+performance of Chain.  We show the results in Figure~\ref{fig:runtime}. For
+each application we measured its execution time on continuous power (the first
+two bars of each group) and on intermittent power at 20cm and 40 cm from the
+RFID antenna (the second two and third two bars of each group, respectively).
+The data show the average execution time of each application run repeatedly for 20 seconds, with runtimes normalized to \sys's average.
+The results show that \sys provides 
+a performance benefit compared to Chain for most applications.  The performance benefit
+is greatest for applications that have small tasks (bc,
+sort, cuckoo) at all distances on intermittent power and with fixed power. In applications that  
+have larger tasks (e.g. dft) or access more different pages of memory, \sys incurs
+overhead from memory virtualization that cause its performance to be comparable to
+(or in the single case of cem, worse than) Chain.
 
-\subsubsection{\sys Compiler Overhead}
+We evaluated the impact that \sys's dynamic task coalescing had on run time
+performance.  Figure~\ref{fig:coalescing} shows the run time performance of
+\sys's coalescing, compared to \sys running without coalescing enabled with
+data normalized to the run time performance of \sys with coalescing.  We show
+data from experiments with continuous power (the first two bars in each bar
+group), and with intermittent power, with the WISP at 20 cm and 40cm from the
+RFID reader (the second and third groups of two bars in each group).
+%
+The data illustrate that coalescing improves \sys's performance by eliminating
+the overhead of committing after each task.  The performance benefit varies,
+providing up to 4.5x performance improvement (for cuckoo filtering, at 40cm).
+The cuckoo application (like other top-performers, bc and sort) has small tasks
+that \sys can easily coalesce to eliminate many commits. In contrast, for some
+applications, e.g. cem, coalescing does not improve performance. The reason for
+cem's poor performance is that cem uses a large LZW dictionary structure that
+spans multiple pages, but has very little {\em locality}.  The absence of
+locality means that coalesced tasks do not access the same pages, and as a
+result, do not amortize commit overhead. The lack of locality causes
+performance to regress to be similar to the non-coalescing case.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalescing}
+	\caption{\sys coalescing performance measured per application on fixed power (left-most bar per application) and at 20 cm and 40 cm from the RFID reader (center and right-most two bars per application) compared to \sys without coalescing.}
+	\label{fig:coalescing}
+\end{figure}
+
+\subsubsection{Evaluating \sys's Compiler}
 \label{sec:result_compiler_time}
+\label{sec:results_compiler}
+
+We characterized the behavior of, and evaluated the efficiency of the \sys
+compiler.  Table~\ref{table:compiler_result} shows, for each application, for
+manually written benchmarks (Man.) and automatically compiled benchmarks
+(Comp.) the difference in the number of bytes of protected data (columns 2 and
+3), the number of tasks (columns 4 and 5), the number of \sys protected data
+API access calls (columns 6 and 7), the lines of code in each application for
+\sys and Chain (columns 8 and 9), and compilation time using \sys's compiler
+(column 10).
+
+\begin{table}[t]
+	\centering
+	\footnotesize
+        \renewcommand{\tabcolsep}{1pt}
+	\begin{tabular}{|l|cc|cc|cc|cc|c|}
+		\hline
+		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
+		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
+		\hline\hline
+		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
+		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
+		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
+		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
+		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
+		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
+		%dd &  &  &  &  &  &  &  & 287 &  \\
+		\hline
+	\end{tabular}
+	\caption{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns).
+$^\dagger$dft does not compile with Clang}
+\label{table:compiler_result}
+\end{table}
+
+On average, compiler-generated code introduces about 0.3\% more bytes of
+protected data, suggesting that liveness analysis captures natural boundaries
+in the use of variables also caught by a manual implementation.
+Compiler-generated code generates about 16\% fewer tasks for applications, on
+average.  The results indicate that the \sys compiler selects task boundaries
+differently than a programmer would manually, favoring fewer, larger tasks.
+The \sys compiler inserts protected data access operations conservatively when
+a variable {\em may} alias a protected variables, due to the limitations of
+pointer aliasing. Consequently, the \sys compiler introduces 48\% more
+protected data accesses than programmer-written code. The run time results in
+Section~\ref{sec:result_compiler_time} show that the increase in protected data
+accesses sometimes degrades performance. Overall, the data show that the
+compiler automatically generates tasks similarly to manually written tasks.
+Doing so, \sys's compiler reaps the static task benefits of \sys, but
+simplifies programming by sparing the programmer the job of task decomposition.
+
+We measured compile time for \sys's compiler's automatic task generation for
+each application\footnote{We could not compile the dft application since LLVM
+inserts mathematical built-in functions (i.e. \texttt{\_\_muldf3},
+\texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not
+provided by the LLVM MSP430 back-end.}. We compiled the applications on a
+4-core virtual machine with 1\,GB of memory running Arch Linux on an Intel Core
+i5-6200U at 2.30\,GHz. We collected results using the Linux system time
+collected in seconds. Most of the compile time was spent computing pointer
+alias analysis, while calculating the possible live ranges of variables.
+Applications such as RSA that have a larger code size and that manipulate more
+variables have a longer compilation time, due to a larger search space for
+pointer aliasing. Nonetheless, the compile time is orders of magnitude shorter
+than manually defining tasks, which requires hours of programmer time. The
+characterization of \sys's compiler suggests that the \sys compiler enables
+simpler task-based intermittent programmer for \sys.
+
+In addition to the characterization of the \sys compiler in Table~\ref{table:compiler_result},  
+we experimentally assessed the performance of the \sys compiler, compared to the performance
+of the manually written benchmarks, as written in prior work~\cite{alpaca,chain}. 
+%
+Figure~\ref{fig:comp_user} compares the performance of manually written
+benchmark code to the performance of code automatically compiled using \sys.
+The figure shows run time for both cases, normalized to the run time of the
+manually written program, running on continuous power.
+%
+The data show that in all cases except for ar and rsa, the run time performance
+of compiled code generated by the \sys compiler has comparable performance to
+manually written code.  With comparable performance in most cases, the data
+suggest that the \sys compiler not only makes programming simpler, but does so,
+in most cases, without a high run time overhead.  In the cases of both ar and
+rsa, the code manipulates a large amount of data that is stored in an array and
+passed into functions by reference.   We expect that the data being manipulated
+by reference is leading to a large amount of may-alias references in the
+compiler's analysis, which leads to a larger number of accesses to protected
+data in the automatically compiled code, compared to the manually written code.
 
@@
 \end{figure}
-
-In Table~\ref{table:compiler_result} we have assessed the overhead induced by the \sys compiler.
-It is, nonetheless, important to find out how much does this overhead translates to increased execution time for each \sys application.
-The result of this investigation is provided in Figure~\ref{fig:comp_user}.
-We conclude that in most of the cases \sys compiler code provided comparable execution time to code split into tasks manually (except for ar and rsa applications).
-
-\subsubsection{\sys Coalescing Performance}
-\label{sec:result_compiler_time}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/coalescing}
-	\caption{\sys coalescing performance measured per application (left-most bar per application) and at multiple WISP to RFID reader antenna distances, \{20, 40\}\,cm (center and right-most bar per application) assessed against non-coalescing \sys.}
-	\label{fig:coalescing}
-\end{figure}
-
-An important benchmark is to measured the effectiveness of \sys coalescing.
-At fixed power and varying WISP to RFID antenna distances we compared \sys execution time per each application (tasked manually).
-Result, given in Figure~\ref{fig:coalescing}, clearly illustrates that coalescing provides a benefit for all the applications, providing even seven times improvement (bc application at 40\,cm).
-bc application is composed of small tasks, which \sys can easily coalesce.
-On the other hand, for certain applications, e.g.
-cem, coalescing improvement is not significant.
-cem uses large data structure that expands over multiple pages, which does not leave room for other variables.
-Any interaction will induce large page swap which cause little improvement in \sys performance against non-coalescing case.
 

### 2017-08-12 paper/compiler.tex

@@
 Doing so requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
-The \sys compiler, \textcolor{red}{based on LLVM}, first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary.
+The \sys compiler, based on LLVM, first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary.
 After identifying where to put task boundaries, the \sys compiler transforms the imperative code into tasks, instrumenting protected memory accesses.

### 2017-08-12 paper/evaluation.tex
fixed typo
pagefault figs
eval text

@@
 characterization of \sys's compiler suggests that the \sys compiler enables
-simpler task-based intermittent programmer for \sys.
+simpler task-based intermittent programming for \sys.
 
@@
 
+We evaluated the effect that paging with different page sizes has on \sys's
+performance and we show data in Figure~\ref{fig:page_size}. We show two plots:
+the top plot shows the run time performance normalized to the best
+per-application performance, using pages of different sizes in \sys, indicated
+with the labels on top of each cem bar; the bottom plot shows the number of
+page faults for the same set of executions normalized to the maximum of each
+set.
+%
+We measured performance as MCU clock cycles using values read out by TI's CCS
+IDE. 
+%
+The results show a clear ``bathtub curve'' in the performance of each
+application, as the number of page faults varies. The data show that across
+application there is a page size that minimizes \emph{both} execution time.  We
+observe that the page size is not the same for each application, although 64
+byte pages perform well for all applications except sort. The increased rate of
+page faults is responsible for the overhead with small page sizes; smaller
+pages require accessing more different pages, leading to higher paging costs.
+With larger page sizes, the overhead is higher than with moderate pages, too.
+The explanation for these higher overheads is that large pages have a higher
+commit cost. Even if an application accesses few memory locations, \sys pages
+data at full page size, degrading performance by sometimes paging in and out
+unnecessary data.  The data reveal a reasonable default page size of 64 bytes
+and when developing an application, a developer should consider varying the
+page size to moderate poor performance.
+
 \begin{figure}
@@
 	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
-	\caption{Normalized page size overhead and page faults per benchmark.}
+	\includegraphics[width=\columnwidth]{figures/pagefault}
+	\caption{\sys's page size affects overhead and page fault count.}
 	\label{fig:IPOSPerformance}
@@
 \end{figure}
-
-Figure~\ref{fig:page_size} presents the result on the effect of page swap size on the execution time of the programs executed by \sys.
-Overhead has been measured in the MCU clock cycles with values provided by TI CCS.
-Result has been normalized per benchmark: for the overhead results---to the minimum obtained value and for page faults---to the third measured maximum value.
-A subset of applications has been selected for evaluation.
-Additionally, page faults have been measured.
-The result show clearly a bathtub-shape bar plot for each application, demonstrating that for each application there is always a page size that minimizes \emph{both} execution time and page swaps.
-For all applications (except for Sort) this value equals roughly to 64 bytes.
-The shape of each figure can be explained as follows.
-For small page sizes (<64) the execution time is dominated by the \emph{frequency} of page swaps.
-On the other hand, in the opposite case the execution time is dominated by the time of swapping of one large page from SRAM to FRAM.
-Another interesting observation is the non-symmetric shape of each bar plot.
-This is due the size of data manipulated by the program: minimum point of overhead equals roughly to the size of the variables manipulated by each benchmark.
-Worth noting is also large discrepancy between minimum and maximum overhead for each benchmark (e.g.
-for Cuckoo 8 (at 16 bytes) to 1 (at 64 bytes), BC: 1 (at 16 bytes) to $\approx$2 (at 1024 bytes)).
 

### 2017-10-30 paper/paper.tex
Covering Koen's comments
conclusions compressed
Bacground compressed
stash pop merge
\newpage removed
made article style PLDI complaint

@@
 %\documentclass[sigconf,anonymous,review]{acmart}
-\documentclass[pageno]{jpaper}
+%\documentclass[pageno]{jpaper}
+\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
@@
 \usepackage{xspace}
-\usepackage{tikz}
+%\usepackage{tikz}
 \usepackage{listings}
+%\usepackage{float}
 %Brandon: this broke the build for me
@@
 
+%%Added for sub-float package from jpaper.cls - will check later
+
+\RequirePackage[font={normalsize,sf,bf}]{caption}
+\RequirePackage[position=bottom]{subfig}
+\captionsetup[table]{aboveskip=0.5em, belowskip=0.5em}
+\captionsetup[figure]{aboveskip=0.5em, belowskip=0em}
+\captionsetup[subfloat]{font={small,sf}}
+
 \hypersetup{colorlinks=false,linkcolor=blue,urlcolor=blue}
 \urlstyle{rm}
+\linespread{0.98}
+
+\newlength{\sectionbelowskip}
+\newlength{\sectionaboveskip}
+\newlength{\subsectionbelowskip}
+\newlength{\subsectionaboveskip}
+\newlength{\paragraphaboveskip}
+
+% Defaults in sigplanconf (note that horizontal space after \paragraph heading will be different):
+%\setlength{\sectionbelowskip}{4pt}
+%\setlength{\sectionaboveskip}{10pt plus 3pt minus 2pt}
+%\setlength{\subsectionbelowskip}{4pt}
+%\setlength{\subsectionaboveskip}{8pt plus 2pt minus 2pt}
+%\setlength{\paragraphaboveskip}{6pt plus 2pt minus 2pt}
+
+% Slightly tighter than defaults:
+% \setlength{\sectionbelowskip}{3pt}
+% \setlength{\sectionaboveskip}{6pt plus 2pt minus 2pt}
+% \setlength{\subsectionbelowskip}{2pt}
+% \setlength{\subsectionaboveskip}{4pt plus 1pt minus 1pt}
+% \setlength{\paragraphaboveskip}{2pt plus 1pt minus 1pt}
+
+% Tighter than defaults:
+ \setlength{\sectionbelowskip}{1pt}
+ \setlength{\sectionaboveskip}{5pt plus 1.5pt minus 1.5pt}
+ \setlength{\subsectionbelowskip}{1pt}
+ \setlength{\subsectionaboveskip}{3pt plus 1pt minus 1pt}
+ \setlength{\paragraphaboveskip}{1pt plus 1pt minus 1pt}
 
@@
 %\newcommand{\sysfull}{VIPE: Virtualized Intermittently-Powered Execution}
-\newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Execution with Adaptive Coalescing and Memory Virtualization}
+\newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization}
 
@@
 
-\section{\sysfull}
+\section{\sys: System Overview}
 \label{sec:overall_system}
@@
 
-\section{Discussion}
-\label{sec:discussion}
+%\section{Discussion}
+%\label{sec:discussion}
 
-\input{discussion}
+%\input{discussion}
 
@@
 
-\section{Conclusions and Future Work}
+\section{Conclusions}
 \label{sec:conclusions}
@@
 
-\newpage
+%\newpage
 

### 2017-11-02 paper/paper.tex
Paper made PLDI 2018 template complaint

@@
-%\documentclass[sigconf,anonymous,review]{acmart}
-%\documentclass[pageno]{jpaper}
-\documentclass[sigplan,10pt,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[sigplan,10pt,review,anonymous]{acmart}
+\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
-%replace XXX with the submission number you are given from the ASPLOS submission site.
-\newcommand{\asplossubmissionnumber}{39}
+\acmConference[Submitted to PLDI'18]{ACM SIGPLAN Conference on Programming Language Design and Implementation}{June 18--22, 2018}{Philadelphia, PA, USA}
 
-\usepackage[normalem]{ulem}
+\acmYear{2018}
+\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+\setcopyright{none}
+
+%% Some recommended packages.
+%\usepackage{booktabs}   %% For formal tables: http://ctan.org/pkg/booktabs
+%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions http://ctan.org/pkg/subcaption
+
+%\usepackage[normalem]{ulem}
 \usepackage{amssymb}
@@
 \usepackage{xspace}
-%\usepackage{tikz}
 \usepackage{listings}
+
 %\usepackage{float}
-%Brandon: this broke the build for me
-%\usepackage[keeplastbox]{flushend}
-
-%%Added for sub-float package from jpaper.cls - will check later
-
-\RequirePackage[font={normalsize,sf,bf}]{caption}
+%\usepackage[keeplastbox]{flushend} %Brandon: this broke the build for me
+%Added for sub-float package from jpaper.cls - will check later
+%\RequirePackage[font={normalsize,sf,bf}]{caption}
 \RequirePackage[position=bottom]{subfig}
-\captionsetup[table]{aboveskip=0.5em, belowskip=0.5em}
-\captionsetup[figure]{aboveskip=0.5em, belowskip=0em}
+%\captionsetup[table]{aboveskip=0.5em, belowskip=0.5em}
+%\captionsetup[figure]{aboveskip=0.5em, belowskip=0em}
 \captionsetup[subfloat]{font={small,sf}}
@@
 
-%\usepackage{natbib}
-%\usepackage{graphicx}
-%\usepackage{subfigure}
-%\usepackage[inline]{enumitem}
-
 \newcommand{\TODO}[1]{{\bf \em \textcolor{red}{TODO: #1}\xspace}}
@@
 
-%Title Options (Brandon: your call which one)
-%\newcommand{\sysfull}{VIPER: Virtualised Intermittent Programming model, Execution model, and Runtime environment}
-%\newcommand{\sysfull}{VIPER: Virtualized Intermittently-Powered Execution Runtime}
-%\newcommand{\sysfull}{PREVIEW: Programming and Runtime Environment with Virtualised Intermittent Execution Windows}
-%\newcommand{\sysfull}{VIPE: Virtualized Intermittently-Powered Execution}
-\newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization}
-
-%\setcopyright{rightsretained}
-
-%\acmDOI{http://dx.doi.org/xx.xxxx/xxxxxxx.xxxxxxx}
-
-%\acmISBN{978-1-xxxx-xxxx-x/17/11}
-
-%\acmConference[ASPLOS]{The 23rd ACM International Conference on Architectural Support for Programming Languages and Operating Systems}{March 24--28, 2017}{Williamsburg, VA, USA}
-
-%\acmYear{2017}
-
-%\copyrightyear{2017}
-
-%\acmPrice{15.00}
+%newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization} %Old title
+\newcommand{\sysfull}{\sys: Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization}
 
@@
 %	\streetaddress{4720 Forbes Avenue}
-%\city{Pittsburgh} 
+%   \city{Pittsburgh} 
 %	\state{PA} 
@@
 
-\date{}
-\maketitle
-
-\thispagestyle{empty}
-
 \begin{abstract}
@@
 
+\maketitle
+
 \section{Introduction}
@@
 
-%\newpage
-
-% The intermittent execution is cumulative discrete process.
-The intermittent execution model is only able to execute a few number of instructions, as compared to the conventional (continuous) model, before its progress is terminated.
-Therefore, the intermittent execution model adapts an execution progress state saving mechanism.
-This mechanism is realized either by means of a checkpoint~\cite{}, where all the context of the program is saved into non-volatile memory.
-Or
-
-% This progress state saving mechanism is normally injected into a program either by a compiler~\cite{}.
-A compiler injects trigger pointers to checkpoint (save) the context of the program into the non-volatile memory.
-or it is added by a programmer~\cite{chain}.
-[compare these two approaches]...
-
 \bibliographystyle{plain}

### 2017-11-02 paper/abstract.tex
Abstract updated

@@
 This makes the code inefficient (when executed on a larger energy buffer than intended) or can potentially completely disable the program (in the opposite case).
-Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing and automatic compiler-supported task generation.
+Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing.
 If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve execution times.
+\todo{Provide concrete numbers for improvement}{Amjad} \todo{Provide concrete numbers for checkpointing improvement}{Amjad}

### 2017-11-04 paper/intro.tex
first structural changes to Intro

--- 2017-08-11 paper/intro.tex
@@
 Advances in processor efficiency along with the development of energy-harvesting systems has created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
-These devices operate using energy extracted from their environment, such as RF~\cite{rf_powered_computing_gollakota_2014}, photovoltaics~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
-Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such devices are a promising technology for use in the Internet of Things (IoT)~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
+These devices operate using ambient energy, such as radio frequency transmissions~\cite{rf_powered_computing_gollakota_2014}, light~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
+Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such devices are a promising technology for use in the Internet of Things~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
 
 Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}. An
-energy-harvesting device buffers energy in a capacitor~\cite{gorlatova_tmc_2013},~\cite{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating.
+energy-harvesting device buffers energy in a small storage, e.g.
+a capacitor~\cite{gorlatova_tmc_2013},~\cite{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating.
 Harvestable energy sources are low-power compared to a platform's operating level.
 A device operates briefly and when buffered energy is depleted, shuts down and recharges to operate again later.
-As an example, recharge time may be tens of seconds in RF powered medical device~\cite[Fig.
+As an example, recharge time may be tens of seconds in radio frequency powered medical device~\cite[Fig.
 3c]{nadeau_naturebio_2017}.
-Charge and discharge times vary by device (different capacitor sizes) and some~\cite{wisp} may fail $\approx$10 to $\approx$100 times per second~\cite{tan_infocom_2016},~\cite{mementos},~\cite{nvp}.
+Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\cite{tan_infocom_2016},~\cite{mementos},~\cite{nvp}.
 
-Software in an energy-harvesting system operates in the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with execution interrupted by failure periods.
+\begin{figure}
+	\begin{subfigure}[t]{.35\linewidth}
+		\centering\large Figure
+		\caption{\centering WISP~\cite{wisp} against an RFID antenna}\label{fig:1a}
+	\end{subfigure}%
+	\begin{subfigure}[t]{.65\linewidth}
+		\centering\large Figure
+		\caption{\centering Program X~\cite{hicks_mibench2_2016} (split into X and Y tasks~\cite{chain}) at two distances}\label{fig:1b}
+	\end{subfigure}
+	\caption{Impact of an inadequate task sizing on the computation speed of a moving transiently-powered device.}\label{fig:1}
+\end{figure}
+
+\textbf{Data Consistency in Intermittent Computing:} Software in an energy-harvesting system operates in the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with execution interrupted by failure periods.
 When power fails, a device loses volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
@@
 
-There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) programming and execution models~\cite{dino,ratchet,chain,alpaca} and (ii) architectures~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
+There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
 While both approaches represent progress, each has limitations.
 New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}.
-New programming models and compilers require programmers to change behavior~\cite{chain,ratchet}, inhibiting adoption.
-New memory models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
+New programming models and compilers require programmers to change behaviour~\cite{chain,ratchet}, inhibiting adoption.
+\todo{Solve: no logical connection between data consistency models and memory models}{Przemek} New memory models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
 Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
 
-Recent work~\cite{alpaca,chain} proposed {\em task-based} programming and execution models in which programmers statically decompose a program into a collection of tasks.
+\textbf{Static Task-based Model and its Inefficiency:} Recently two {\em task-based} programming and execution models~\cite{alpaca,chain} were proposed in which programmers statically decompose a program into a collection of tasks.
 Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
-The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent~\cite{chain}.
-An key challenge is that the length of a software task's execution is limited by the fixed, total amount of energy that a device can buffer in hardware.
+The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent.
+An key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware.
 A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict.
-Assuming that input power is negligible compared to operating power~\cite{wisp}, a task will never be able to complete if its execution consumes more energy than the system can buffer.
-On the other hand, transitioning from one task to another imposes a run time cost and excessively small tasks are inefficient.
+To illustrate this, we refer to Figure~\ref{fig:1} showing the execution time of two applications (the same application X, but divided into X and Y tasks as in~\cite{chain}) running on RFID antenna-powered Computational RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna distances (near---stable energy supply/far---high energy intermittency).
+At short distance: execution will takes too long caused by runtime cost of marshalling excessively small tasks.
+At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer.
 
-The key research question addressed here is how to define and execute tasks efficiently and ensure non-termination, with an unknown energy buffer size and energy arrival process.
+\textbf{Research Question and Contributions:} We need to know how to define and execute tasks efficiently and ensure non-termination, with an unknown energy buffer size and energy arrival process.
 In particular, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?
 
-In this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer.
+To address this problem, in this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
 To accomplish this, \sys introduces new capabilities:
@@
 \begin{itemize}
-	\item {\bf Automatic task compilation:} \sys's compiler transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints.
-The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent.
-\sys's complier the labour-intensive process of writing tasks~\cite{chain,alpaca} (i.e., hours of programmer time).
-	%
 	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task.
@@
 
+As a helpful feature, in this work we also address the problem of automatic task compilation: \sys's compiler transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints.
+The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent.
+\sys's complier eliminates the labour-intensive process of writing tasks~\cite{chain,alpaca} (i.e., hours of programmer time).
+
 We evaluated \sys end-to-end on a collection of existing benchmarks ~\cite{chain} and new programs, and compare directly to a state-of-the-art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without rewriting code.
-%\todo{add detailed text about evaluation results}{Przemek}
 

### 2017-11-04 paper/background.tex
commented out subfig references in background and coalescing

@@
 	\centering
-	\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
-	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
+	%\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
+	%\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
 	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
@@
 	\centering
-	\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
-	\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
+	%\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
+	%\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
 	\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}

### 2017-11-04 paper/coalescing.tex
commented out subfig references in background and coalescing

@@
 	\centering
-	\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
-	\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
+	%\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
+	%\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
 	\caption{\sys coalesces tasks to avoid overheads.}

### 2017-11-04 paper/intro.tex
Introduction rewritten, refliecting better the novelty and contribution

@@
 There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
-While both approaches represent progress, each has limitations.
-New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}.
-New programming models and compilers require programmers to change behaviour~\cite{chain,ratchet}, inhibiting adoption.
-\todo{Solve: no logical connection between data consistency models and memory models}{Przemek} New memory models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
-Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
+New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
+However, they have their own limitations that need to be addressed---the core one being inability to adapt to changing energy conditions.
 
-\textbf{Static Task-based Model and its Inefficiency:} Recently two {\em task-based} programming and execution models~\cite{alpaca,chain} were proposed in which programmers statically decompose a program into a collection of tasks.
+\textbf{Static Task-based Model and its Inefficiency:} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks.
 Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
 The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent.
-An key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware.
+A key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware.
 A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict.
@@
 At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer.
+This calls for at-runtime task division of any transiently-powered application.
 
-\textbf{Research Question and Contributions:} We need to know how to define and execute tasks efficiently and ensure non-termination, with an unknown energy buffer size and energy arrival process.
-In particular, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?
+\textbf{Data Inconsistency of Virtual Task:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task.
+This however induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}.
+Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively.
+All tasks have no WAR dependency and are consistent.
+Now, suppose three tasks have been virtualized into one, namely Task 4: {\tt{\{y=x+1; z=y+1; x=z+1;\}}} which has inherent WAR dependency (variable {\tt{x}}).
 
-To address this problem, in this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
-To accomplish this, \sys introduces new capabilities:
+\textbf{Trade-off with Virtual Memory Manipulation:} WAR dependencies cannot be complied at runtime for merged (virtual) tasks.
+Therefore virtual task execution model needs to manipulate merged tasks in-memory to guarantee data consistency.
+Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
+Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
+How to connect this trade-off with task virtualization remains an open question.
+
+\textbf{Research Question and Contributions:} Specifically, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?
+
+To address this question, in this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
+To accomplish this, \sys introduces two new capabilities:
 
@@
 	%
-	\item {\bf Volatile memory virtualization:} \sys's memory virtualization uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution.
+	\item {\bf Volatile memory virtualization on transient power:} Colascing through task virtualization requires memory virtualization.
+\sys's uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution.
 When a task ends, \sys commits dirty pages to FRAM using DMA block copies, ensuring task atomicity and memory consistency.
+To the best of our knowledge this is the first time when memory virtualization has been demonstrated on a transient power budget.
 \end{itemize}
 
-As a helpful feature, in this work we also address the problem of automatic task compilation: \sys's compiler transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints.
+We evaluated \sys end-to-end on a collection of existing benchmarks ~\cite{chain} and new programs, and compare directly to a state-of-the-art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without rewriting code.
+\todo{Provide concrete numeric results supporting this claim}{Amjad/Przemek}
+
+In this work we also address the problem of task construction burden imposed on the programmer.
+That is, we introduce \sys's compiler that transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints.
 The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent.
-\sys's complier eliminates the labour-intensive process of writing tasks~\cite{chain,alpaca} (i.e., hours of programmer time).
-
-We evaluated \sys end-to-end on a collection of existing benchmarks ~\cite{chain} and new programs, and compare directly to a state-of-the-art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without rewriting code.
+While \sys's complier is still \emph{limited in scope}, e.g.
+is unable to handle instruction jumps, normalized loops and recursions, it already eliminates the labour-intensive process of writing tasks~\cite{chain,alpaca}, i.e., hours of programmer time.
 

### 2017-11-05 DMA/ipos-gcc-dma/include/dataProtec.h
multipages results for 128 and 64 bytes page sizes

@@ -23,13 +23,14 @@
 
 uint8_t* __return_addr(uint8_t* var) ;
 uint8_t* __return_addr_wr(uint8_t* var) ;
-uint8_t* __return_addr_no_check(uint8_t* var) ;
-uint8_t* __return_addr_wr_no_check(uint8_t* var) ;
+uint8_t* __return_addr_no_check(unsigned int* var) ;
+uint8_t* __return_addr_wr_no_check(unsigned int* var) ;
 void __sendPagTemp(unsigned int pagTag);
 void __bringPagTemp(unsigned int pagTag);
 void __bringPagROM(unsigned int pagTag);
 void __sendPagROM(unsigned int pagTag);
 unsigned int __pageSwap(unsigned int * varAddr);
+void __pagesSwap(unsigned int * CurnPag, unsigned int * varAddr);
 void __bringPersisCrntPag(unsigned int curntPag);
 void __pagsCommit();
 void __bringCrntPagROM();

### 2017-11-05 DMA/ipos-gcc-dma/ipos/psched.c

@@ -99,15 +99,16 @@ void os_scheduler()
 
     if (__commit_flag == COMMITTING)
     {
-        // we need to do a page swap to bring the correct page from temp  or  ROM
-        // this page swap must not send a page, and that must not happen since the dirtyPag is 0 at this stage
-        __bringPersisCrntPag(__persis_CrntPagHeader );
-        __realTask = __temp_coalTskAddr;
-        goto commit;
+        // firm transition
+        __coalTskAddr = (unsigned int) __temp_coalTskAddr;
+
+        // Commit the dirty pages
+        __pagsCommit();
+        __commit_flag = COMMIT_FINISH;
     }
 
     // Recover the virtual state
-    __bringCrntPagROM();
+    __bringPagROM(__persis_CrntPagHeader);
     __realTask = (unsigned int *) __coalTskAddr;
 
 

### 2017-11-05 apps/bc/main.c

@@ -308,13 +308,13 @@ void task_AR_btbl_bitcount() {
        cp_reset();
 #endif
 
-    unsigned char * Ptr = (unsigned char *) &P(_v_seed) ;
+    unsigned char * Ptr = (unsigned char *) &RP(_v_seed) ;
     int Accu ;
 
-    Accu  = bits[ P(*Ptr++) ];
-    Accu += bits[ P(*Ptr++) ];
-    Accu += bits[ P(*Ptr++) ];
-    Accu += bits[ P(*Ptr) ];
+    Accu  = bits[ RP(*Ptr++) ];
+    Accu += bits[ RP(*Ptr++) ];
+    Accu += bits[ RP(*Ptr++) ];
+    Accu += bits[ RP(*Ptr) ];
     WP(_v_n_5)+= Accu;
     uint32_t tmp_seed = RP(_v_seed);
     WP(_v_seed) = tmp_seed + 13;

### 2017-11-05 apps/crc/main.c

@@ -165,10 +165,10 @@ unsigned int CCITT_Update(unsigned int init, unsigned int input)
 int main(void) {
     init();
 
-    taskId tasks[] = {  {initTask,1},
-                        {firstByte,2},
-                        {secondByte,3 },
-                        {task_finish, 4}
+    taskId tasks[] = {  {initTask,1, 1},
+                        {firstByte,2, 1},
+                        {secondByte,3, 1 },
+                        {task_finish, 4, 1}
     };
     //This function should be called only once
     os_addTasks(4, tasks );

### 2017-11-05 apps/cuckoo_delft/main.c

@@ -150,7 +150,7 @@ void task_init_array() {
 #endif
     unsigned i;
         for (i = 0; i < BUFFER_SIZE - 1; ++i) {
-            WP(_v_filter[i + P(_v_index)*(BUFFER_SIZE-1)]) = 0;
+            WP(_v_filter[i + RP(_v_index)*(BUFFER_SIZE-1)]) = 0;
         }
         ++WP(_v_index);
         if (RP(_v_index) == NUM_BUCKETS/(BUFFER_SIZE-1)) {
@@ -180,8 +180,8 @@ void task_generate_key()
     __cry = (RP(_v_key) + 1) * 17;
     WP(_v_key) = __cry;
 
-    if (P(_v_next_task) >= t_generate_key) {
-        os_jump(P(_v_next_task) - t_generate_key);
+    if (RP(_v_next_task) >= t_generate_key) {
+        os_jump(RP(_v_next_task) - t_generate_key);
     }
     else {
         os_jump(TASK_NUM - RP(_v_next_task) + t_generate_key);
@@ -236,7 +236,7 @@ void task_calc_indexes_index_2()
      __cry = RP(_v_index1) ^ fp_hash;
      WP(_v_index2) = __cry;
 
-    if (P(_v_next_task) >= t_calc_indexes_index_2) {
+    if (RP(_v_next_task) >= t_calc_indexes_index_2) {
         os_jump(RP(_v_next_task) - t_calc_indexes_index_2);
     }
     else {
@@ -272,7 +272,7 @@ void task_add()
     uint16_t __cry;
     uint16_t __cry_idx = RP(_v_index1);
     uint16_t __cry_idx2 = RP(_v_index2);
-    if (!P(_v_filter[__cry_idx])) {
+    if (!RP(_v_filter[__cry_idx])) {
 
         WP(_v_success) = true;
         __cry = RP(_v_fingerprint);
@@ -280,7 +280,7 @@ void task_add()
         os_jump(2);
         return;
     } else {
-        if (!P(_v_filter[__cry_idx2])) {
+        if (!RP(_v_filter[__cry_idx2])) {
 
             WP(_v_success) = true;
             __cry = RP(_v_fingerprint);
@@ -369,7 +369,7 @@ void task_insert_done()
     WP(_v_inserted_count) = __cry;
 
 
-    if (P(_v_insert_count) < NUM_INSERTS) {
+    if (RP(_v_insert_count) < NUM_INSERTS) {
[…]

### 2017-11-05 apps/sortAlgo/main.c

@@ -4,9 +4,9 @@
 #include "mspProfiler.h"
 #include "mspDebugger.h"
 
-#define CODEPROFILE 0
-#define RAISE_PIN 1
-#define PWR_INT_SIM 1
+//#define TSK_SIZ 0
+//#define RAISE_PIN 1
+//#define PWR_INT_SIM 1
 
 /*
 void sortAlgo(int arr[], int arrLen){
@@ -26,11 +26,19 @@ void task_finish();
 void task_outer_loop();
 void task_inner_loop();
 
-////// Global Piables
-__p unsigned int arr[] = {3,1,4,6,9,5,10,8,16,20,19,40,16,17,2,41,80,100,5,89,66,77,8,3,32,55,8,11,99,65,25,89,3,22,25,121,11,90,74,1,12,39,54,20,22,43,45,90,81,40};
-    unsigned int arr2[] = {3,1,4,6,9,5,10,8,16,20,19,40,16,17,2,41,80,100,5,89,66,77,8,3,32,55,8,11,99,65,25,89,3,22,25,121,11,90,74,1,12,39,54,20,22,43,45,90,81,40};
+////// Global variables
 
-    unsigned int arr_len = 50;
+__p unsigned int arr[] = {3,1,4,6,9,5,10,8,16,20,19,40,16,17,2,41,80,100,5,89,66,77,8,3,32,55,8,11,99,
+                          65,25,89,3,22,25,121,11,90,74,1,12,39,54,20,22,43,45,90,81,40,
+                          3,1,4,6,9,5,10,8,16,20,19,40,16,17,2,41,80,100,5,89,66,77,8,3,32,55,8,11,99,
+                          65,25,89,3,22,25,121,11,90,74,1,12,39,54,20,22,43,45,90,81,40};
+
+    unsigned int arr2[] = {3,1,4,6,9,5,10,8,16,20,19,40,16,17,2,41,80,100,5,89,66,77,8,3,32,55,8,11,99,
+                           65,25,89,3,22,25,121,11,90,74,1,12,39,54,20,22,43,45,90,81,40,
+                           3,1,4,6,9,5,10,8,16,20,19,40,16,17,2,41,80,100,5,89,66,77,8,3,32,55,8,11,99,
+                           65,25,89,3,22,25,121,11,90,74,1,12,39,54,20,22,43,45,90,81,40};
+
+    unsigned int arr_len = 100;
 __p unsigned int i =  0;
[…]

### 2017-11-05 experiments/newExperiments/scripts/ramPage_timeToCompletion.py

@@ -1,0 +1,80 @@
+#! /usr/bin/python
+
+
+"""
+    IPOS: performance 
+    Metric : time
+    Task   : execute apps
+
+
+@ Author  : Amjad Yousef Majid
+@ Gmail ID: amjad.y.majid 
+@ Date    : 17/October/2017 
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+from csf import *
+
+numPags =  ("single",  "double",  "triple" )   
+
+############### page size of 128 bytes ##############
+
+cem      = np.array( [ 3691384,       2081138,    2658959 ]) 
+cuckoo   = np.array( [ 443089,        513159,     622667  ]) 
+crc      = np.array( [ 82567,         84605,      86641   ]) 
+dft      = np.array( [ 26267755,      26010302,   26254522]) 
+rsa      = np.array( [ 2196414,       1771543,    2260314 ]) 
+bc       = np.array( [ 831246,        849276,     868791  ]) 
+sort     = np.array( [ 12242871,      7765337,    7920915 ]) 
+
+############### page size of 64 bytes ##############
+
+# cem      = np.array( [ 3249640,       1869253,    2205335 ]) 
+# cuckoo   = np.array( [ 382723,        440148,     552546  ]) 
+# crc      = np.array( [ 70087,         72061,      74033   ]) 
+# dft      = np.array( [ 26372167,      26306501,   25931295]) 
+# rsa      = np.array( [ 2277448,       2070869,    1815520 ]) 
+# bc       = np.array( [ 713362,        726396,     745847  ]) 
+# sort     = np.array( [ 13915418,      9599409,    8834494 ]) 
+
+
+cem     = cem / float( cem[0])
+cuckoo  = cuckoo / float( cuckoo[0])
+crc     = crc /float( crc[0])
+dft     = dft/ float( dft[0])
+rsa     = rsa / float( rsa[0])
+bc      = bc / float( bc[0])
+sort    = sort / float( sort[0])
+
+
+
+
+apps = [cem, cuckoo, crc, dft, rsa, bc, sort]
+appsName = ["cem", "cuckoo", "crc", "dft", "rsa", "bc", "sort"]
+
+figureSetting()
+f = plt.figure(figsize=(8,2.5))
+
+xlabelsPos=[]
+s = 0
+for i in range(7):
[…]

### 2017-11-05 paper/intro.tex
multipages results for 128 and 64 bytes page sizes
text for virtualization

@@
 
-\textbf{Static Task-based Model and its Inefficiency:} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks.
+\textbf{\replace{Static Task-based Model and its Inefficiency:}{The Inefficiency of Static Task-based Models}} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks.
 Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
@@
 At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer.
-This calls for at-runtime task division of any transiently-powered application.
+This calls for at-runtime \add{adaptive} task division of any transiently-powered application\add{---namely \emph{task virtualization}}.
 
-\textbf{Data Inconsistency of Virtual Task:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task.
-This however induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}.
-Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively.
-All tasks have no WAR dependency and are consistent.
-Now, suppose three tasks have been virtualized into one, namely Task 4: {\tt{\{y=x+1; z=y+1; x=z+1;\}}} which has inherent WAR dependency (variable {\tt{x}}).
+\begin{figure}
+	\includegraphics[width=0.6\columnwidth]{presentation/images/alpaca-war.pdf}
+	\caption{\label{fig:virtualization}\color{blue} Adaptive task division at run-time might lead to WAR dependencies that do not exist with static task division.}
+\end{figure}
+
+\textbf{\replace{Data Inconsistency of Virtual Task:}{The Challange Behind Task Virtualization}} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task.
+This however \replace{induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}}{might introduce new \emph{write-after-read (WAR)} dependencies at \emph{run-time} that can break the atomicity of the \emph{virtual task}.
+Consider the example depicted in Fig.~\ref{fig:virtualization}: Tasks 1--3 being executed consecutively.} \remove{Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively.} All tasks \replace{have no WAR dependency and are consistent}{in this figure are atomic since they do not have WAR dependency on the variables they are modifying, e.g.
+\emph{x} is only read and \emph{y} is only written within Task 1.} Now, suppose three tasks have been virtualized into one, namely Task 4: \replace{ \{y=x+1; z=y+1; x=z+1;\} which has inherent WAR dependency (variable \{\tt{x}\}).
+}{since now \emph{x} is first read and then written, a WAR dependency is introduced dynamically at run-time---unfortunately Task 4 is no more atomic since its re-execution will not always produce the same results.}
 

### 2017-11-05 paper/intro.tex
text for virtualization

--- 2017-11-04 paper/intro.tex
@@
 
-\textbf{Static Task-based Model and its Inefficiency:} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks.
+\textbf{\replace{Static Task-based Model and its Inefficiency:}{The Need for Adaptive Task Division:}} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks.
 Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
@@
 At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer.
-This calls for at-runtime task division of any transiently-powered application.
+This calls for at-runtime \add{adaptive} task division of any transiently-powered application\add{---namely \emph{task virtualization}}.
 
-\textbf{Data Inconsistency of Virtual Task:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task.
-This however induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}.
-Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively.
-All tasks have no WAR dependency and are consistent.
-Now, suppose three tasks have been virtualized into one, namely Task 4: {\tt{\{y=x+1; z=y+1; x=z+1;\}}} which has inherent WAR dependency (variable {\tt{x}}).
+\begin{figure}
+	\includegraphics[width=0.6\columnwidth]{presentation/images/alpaca-war.pdf}
+	\caption{\label{fig:virtualization}\color{blue} Adaptive task division at run-time might lead to WAR dependencies that do not exist with static task division.}
+\end{figure}
 
-\textbf{Trade-off with Virtual Memory Manipulation:} WAR dependencies cannot be complied at runtime for merged (virtual) tasks.
-Therefore virtual task execution model needs to manipulate merged tasks in-memory to guarantee data consistency.
-Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
+\textbf{\replace{Data Inconsistency of Virtual Task:}{The Challange Behind Task Virtualization:}} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task.
+This however \replace{induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}}{might introduce new \emph{write-after-read (WAR)} dependencies on the variables in non-volatile memory at \emph{run-time} that can break the atomicity of the \emph{virtual task}.
+Consider the example depicted in Fig.~\ref{fig:virtualization}: Tasks 1--3 being executed consecutively.} \remove{Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively.} All tasks \replace{have no WAR dependency and are consistent}{in this figure are atomic since they do not have WAR dependency on the persistent variables they are accessing, e.g.
+\emph{x} is only read and \emph{y} is only written within Task 1.} Now, suppose three tasks have been virtualized into one \add{at run-time}, namely Task 4: \replace{ \{y=x+1; z=y+1; x=z+1;\} which has inherent WAR dependency (variable \{\tt{x}\}).
+}{since now \emph{x} is first read and then written, a WAR dependency on \emph{x} is introduced dynamically at run-time---unfortunately Task 4 is \emph{no more atomic} since its re-execution will not always produce the same results.}
+
+\textbf{\replace{Trade-off with Virtual Memory Manipulation:}{Guaranteeing Atomicity of Virtual Tasks:}} \replace{WAR dependencies cannot be complied at runtime for merged (virtual) tasks.}{Even though WAR dependencies introduced by the static task division can be eliminated at compile time, e.g.~\cite{alpaca}, those introduced at run-time dynamically due to task virtualization cannot.} \replace{Therefore virtual task execution model needs to manipulate merged tasks in-memory to guarantee data consistency.}{Therefore, we require an \emph{execution model} that keeps each virtual task atomic by committing the modified persistent variables at the boundary defined dynamically at run-time---keeping the non-volatile memory unmodified upon a power-interrupt and preserving its consistency.}
+
+\TODO{Below is totally disconnected from the flow}
+{\color{red}Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
 Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
-How to connect this trade-off with task virtualization remains an open question.
+How to connect this trade-off with task virtualization remains an open question.}
 

### 2017-11-06 paper/intro.tex
Revised text in intro after Sinan's comments; remarks about Figure 1

--- 2017-11-04 paper/intro.tex
@@
 New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
-However, they have their own limitations that need to be addressed---the core one being inability to adapt to changing energy conditions.
+However, they have their own limitations that need to be addressed---the core one being \emph{inability to adapt} to changing energy conditions.
 
-\textbf{Static Task-based Model and its Inefficiency:} The latest two {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks.
+\textbf{The Need for Adaptive Task Division:} Recently proposed {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks.
 Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
@@
 At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer.
-This calls for at-runtime task division of any transiently-powered application.
+This calls for at-runtime adaptive task division of any transiently-powered application---namely \emph{task virtualization}.
 
-\textbf{Data Inconsistency of Virtual Task:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task.
-This however induces a problem of \emph{virtual task write-after-read (WAR) dependency violation}.
-Consider the following example of Task 1: {\tt{\{y=x+1;\}}}, Task 2: {\tt{\{z=y+1;\}}}, and Task 3 {\tt{\{x=z+1;\}}} being executed consecutively.
-All tasks have no WAR dependency and are consistent.
-Now, suppose three tasks have been virtualized into one, namely Task 4: {\tt{\{y=x+1; z=y+1; x=z+1;\}}} which has inherent WAR dependency (variable {\tt{x}}).
+\begin{figure}
+	\includegraphics[width=.5\columnwidth]{presentation/images/alpaca-war.pdf}
+	\caption{\label{fig:virtualization}Adaptive task division at run-time might lead to write-after-read dependencies (variable \ttfamily{x}) that do not exist with static task division.\todo{Box with Task 1,2,3 must be the same size as merged task; Merged task must be named as Task 4; x must be marked in text as having WAR; "Task X" name should be next to the task box; "Window" - small letter}{Sinan}}
+\end{figure}
 
-\textbf{Trade-off with Virtual Memory Manipulation:} WAR dependencies cannot be complied at runtime for merged (virtual) tasks.
-Therefore virtual task execution model needs to manipulate merged tasks in-memory to guarantee data consistency.
-Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
+\textbf{The Challenge of Task Virtualization:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task.
+This however might introduce \emph{write-after-read (WAR)} dependencies on the variables in non-volatile memory at \emph{run-time} that can break the atomicity of the \emph{virtual task}.
+Consider the example depicted in Fig.~\ref{fig:virtualization}: Tasks 1, 2 and 3 being executed consecutively.
+All tasks in this example are atomic since they do not have WAR dependency on the persistent variables they are accessing, e.g.
+\emph{x} is only read and \emph{y} is only written within Task 1.
+Now, suppose three tasks have been virtualized into a single one at run-time, namely Task 4: since \emph{x} is now first \emph{read} and then \emph{written}, a WAR dependency on \emph{x} is introduced dynamically at run-time---unfortunately Task 4 is \emph{no more atomic} since its re-execution will not always produce the same results.
+
+\textbf{Guaranteeing Atomicity of Virtual Tasks:} Although WAR dependencies introduced by the static task division can be eliminated at compile time, e.g.~\cite{alpaca}, those introduced dynamically at run-time, due to task virtualization, cannot.
+Therefore, we require a \emph{new execution model} that keeps each virtual task atomic by committing the modified persistent variables at the boundary defined dynamically at run-time---keeping the non-volatile memory unmodified upon a power-interrupt and preserving its consistency.
+
+%Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
 Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
@@
 
-To address this question, in this work we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
+To address this question, we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
 To accomplish this, \sys introduces two new capabilities:
@@
 	%
-	\item {\bf Volatile memory virtualization on transient power:} Colascing through task virtualization requires memory virtualization.
+	\item {\bf Volatile memory virtualization with transiently-powered device:} Colascing through task virtualization requires memory virtualization.
 \sys's uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution.

### 2017-11-06 paper/coalescing.tex
First changes to coalescing section

@@
 \sys dynamically {\em coalesces} consecutively executing tasks at run time.
-When two tasks are coalesced, the first of the two tasks does not commit its
-updates to protected memory locations when it reaches a \transition statement.
-Instead, execution immediately continues from the beginning of the second of
-the two tasks. At the end of the second of the two coalesced tasks, \sys
-commits the pages that were updated by either of the coalescing tasks. The
-purpose of coalescing is to better amortize the run time overhead of committing
-updated pages when a task ends. If consecutive tasks access the same pages of
-data, the pages accessed by both tasks only need to be committed once.
+When two tasks are coalesced, the first of the two tasks does not commit its updates to protected memory locations when it reaches a \transition statement.
+Instead, execution immediately continues from the beginning of the second of the two tasks.
+At the end of the second of the two coalesced tasks, \sys commits the pages that were updated by either of the coalescing tasks.
+The purpose of coalescing is to better amortize the run time overhead of committing updated pages when a task ends.
+If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed once.
 
-\sys can coalesce an arbitrary number of consecutive tasks. As more tasks
-coalesce, their collective commit overhead amortizes better, potentially
-improving performance. As more tasks coalesce, the amount of potentially {\em
-wasted work} also increases. If power fails during a long sequence of coalesced
-tasks the lack of an intervening commit requires execution to restart from the
-{\em first} task in the sequence, failing to preserve the progress of any of
-the tasks in the long sequence. \sys moderates the risk of coalescing and
-capitalizes on its benefit by dynamically, \emph{adaptively} determining a
-coalescing factor that combines many tasks without suffering an excess of
-wasted work.
+\textbf{Task Coalescing Problem:} \sys can coalesce an arbitrary number of consecutive tasks.
+As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance.
+As more tasks coalesce, the amount of potentially {\em wasted work} also increases.
+If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of the tasks in the long sequence.
+\sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, \emph{adaptively} determining a coalescing factor that combines many tasks without suffering an excess of wasted work.
 
-The challenge to coalescing tasks is determining how many tasks to coalesce
-before committing. During execution, an executing task is either a {\em
-coalesced} task, which does not commit its state before transitioning or a {\em
-committing} task, which does commit its state before transitioning. \sys
-monitors a program's execution to adaptively determine $C$, the number of tasks
-to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks,
+The challenge to coalescing tasks is determining how many tasks to coalesce before committing.
+During execution, an executing task is either a {\em coalesced} task, which does not commit its state before transitioning or a {\em committing} task, which does commit its state before transitioning.
+\sys monitors a program's execution to adaptively determine $C$, the number of tasks to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks,
 followed by a committing task.
 
-\sys determines $C$ for a given operation based on the number of tasks
-completed, $S$, during the current operational period. \sys begins execution in
-a coalesced task with $C$ initialized to half of the number of tasks executed
-during the previous operating period and $S$ initialized to $0$. The system
-executes $C-1$ coalesced tasks followed by one committing task. After each
-commit, \sys assigns $C$ to half of $S$.   If more tasks complete in an
-operating period, \sys attempts to coalesce more tasks. If fewer tasks
+\textbf{Task Coalescing Strategies:} There are infinite possible strategies for coalescing (infinite amount of $(C_U,C_D)$ pairs). We will consider the three fundamental ones:
+
+\begin{enumerate}
+	\item \textbf{$N$ up/$N$ down:} where task coalescing factor increases by $N$ at each successful virtual task execution and decreases by $N$, otherwise;
+	\item \textbf{History-driven:} where coalescing will consider history of power failures into deciding on $(C_U,C_D)$ values on the fly;
+	\item \textbf{Program profiling-driven:} where coalescing strategy takes into account \emph{expected size} of each task in the application to pick the best coalescing factor.
+\end{enumerate}
+
+\subsection{Operation of $N$ up/$N$ Down Coalescing:} \todo{To be written}{Przemek}
+
+\subsection{Operation of History-driven Coalescing}: \sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the current operational period.
+\sys begins execution in a coalesced task with $C$ initialized to half of the number of tasks executed during the previous operating period and $S$ initialized to $0$.
+The system executes $C-1$ coalesced tasks followed by one committing task.
+After each commit, \sys assigns $C$ to half of $S$.
+If more tasks complete in an operating period, \sys attempts to coalesce more tasks.
+If fewer tasks
 complete, \sys attempts to coalesce fewer tasks. If no committing task
@@
 (the number of tasks completed before the power failure) and continues.
-
-\begin{figure}
-	\centering
-	%\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
-	%\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
-	\caption{\sys coalesces tasks to avoid overheads.}
-	\label{fig:coalescing_example}
-\end{figure}
 
@@
 continues.
+
+\subsection{Operation of Program Profiling-driven Coalescing:} \todo{To be written}{Przemek}
+
+\begin{figure}
+	\centering
+	%\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
+	%\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
+	\caption{\sys coalesces tasks to avoid overheads.}
+	\label{fig:coalescing_example}
+\end{figure}
 

### 2017-11-12 paper/intro.tex
incorporate motivation discussion and shorten

@@
 
+%\begin{figure}
+%	\begin{subfigure}[t]{.35\linewidth}
+%		\centering\large Figure
+%		\caption{\centering WISP~\cite{wisp} against an RFID antenna}\label{fig:1a}
+%	\end{subfigure}%
+%	\begin{subfigure}[t]{.65\linewidth}
+%		\centering\large Figure
+%		\caption{\centering Program X~\cite{hicks_mibench2_2016} (split into X and Y tasks~\cite{chain}) at two distances}\label{fig:1b}
+%    \end{subfigure}
+%	\caption{Impact of an inadequate task sizing on the computation speed of a moving transiently-powered device.}\label{fig:1}
+%\end{figure}
+
 \begin{figure}
-	\begin{subfigure}[t]{.35\linewidth}
-		\centering\large Figure
-		\caption{\centering WISP~\cite{wisp} against an RFID antenna}\label{fig:1a}
-	\end{subfigure}%
-	\begin{subfigure}[t]{.65\linewidth}
-		\centering\large Figure
-		\caption{\centering Program X~\cite{hicks_mibench2_2016} (split into X and Y tasks~\cite{chain}) at two distances}\label{fig:1b}
-    \end{subfigure}
-	\caption{Impact of an inadequate task sizing on the computation speed of a moving transiently-powered device.}\label{fig:1}
+	%\includegraphics[width=.5\columnwidth]{presentation/images/alpaca-war.pdf}
+    \caption{\label{fig:coalesce}Task decomposition is required for
+intermittent execution but task transitions introduce overhead.}
 \end{figure}
 
-\textbf{Data Consistency in Intermittent Computing:} Software in an energy-harvesting system operates in the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with execution interrupted by failure periods.
+\textbf{Data Consistency in Intermittent Computing.} Software in an energy-harvesting system operates in the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with execution interrupted by failure periods.
 When power fails, a device loses volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
@@
 
-\textbf{The Need for Adaptive Task Division:} Recently proposed {\em task-based} programming and execution models~\cite{alpaca,chain} advocate for static decomposition of a program (by a programmer) into a collection of tasks.
-Tasks can include arbitrary computation and, upon completion, are guaranteed to have executed {\em atomically}, despite arbitrarily-timed power failures.
-The programmer explicitly expresses task-to-task control flow, which may be conditionally input-dependent.
-A key challenge is that the length of a software task's execution is \emph{limited by the fixed total amount of energy} that a device can buffer in hardware.
-A task's code is static, but the duration of its execution may be input-dependent and is difficult to predict.
-To illustrate this, we refer to Figure~\ref{fig:1} showing the execution time of two applications (the same application X, but divided into X and Y tasks as in~\cite{chain}) running on RFID antenna-powered Computational RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna distances (near---stable energy supply/far---high energy intermittency).
-At short distance: execution will takes too long caused by runtime cost of marshalling excessively small tasks.
-At far distance: program \emph{might never execute} if task execution consumes more energy than the system can buffer.
-This calls for at-runtime adaptive task division of any transiently-powered application---namely \emph{task virtualization}.
+\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
+task-based} programming and execution models~\cite{alpaca,chain} advocate for
+static decomposition of a program (by a programmer) into a collection of tasks.
+Tasks can include arbitrary computation and, upon completion, are guaranteed to
+have executed {\em atomically}, despite arbitrarily-timed power failures.
+The programmer explicitly expresses task-to-task control flow.
+%, which may be conditionally input-dependent.
+%
+Figure~\ref{fig:decomp} illustrates how the decomposition into tasks affects
+the time the application takes to complete.
+%
+Each task transition introduces the overhead of tracking and atomically
+committing the modifications to non-volatile memory by the task, to maintain
+consistency of program state.
+%
+Naively minimizing the number of task transitions, (i.e. the task count),
+\emph{at compile time} risks creating a task that requires more energy than the
+device can buffer in its storage capacitor.
+%
+Such a large task would fail to complete unless the environment provides
+sufficient incoming energy to supplement the energy in the capacitor.
+%
+To eliminate this risk, the programmer (or the compiler) must decompose the
+program conservatively into many small tasks, at the cost of higher task
+transition overhead.
 
-\begin{figure}
-	\includegraphics[width=.5\columnwidth]{presentation/images/alpaca-war.pdf}
-	\caption{\label{fig:virtualization}Adaptive task division at run-time might lead to write-after-read dependencies (variable \ttfamily{x}) that do not exist with static task division.\todo{Box with Task 1,2,3 must be the same size as merged task; Merged task must be named as Task 4; x must be marked in text as having WAR; "Task X" name should be next to the task box; "Window" - small letter}{Sinan}}
-\end{figure}
+One approach to decrease the task count, without assumming a minimum level of
+incoming power always available, is to predict task energy consumption and
+choose a decomposition specific to the device capacitor size.
+%
+Howerver, statically predicting energy of arbitrary input-dependent code with
+peripheral access is a problem without a general solution.
+%
+Furthermore, a static decomposition approach cannot produce decompositions
+portable across devices with different storage capacitorss.
+%
+In this paper, we propose a \emph{dynamic} task decomposition approach, that
+accepts any static decomposition and opportunistically \emph{coalesces} tasks
+at runtime when there is sufficient energy.
 
-\textbf{The Challenge of Task Virtualization:} Task-based model which allows for dynamic task division must implement \emph{task virtualization mechanism}: merging of atomic, data-consistent tasks into a single (larger) task.
-This however might introduce \emph{write-after-read (WAR)} dependencies on the variables in non-volatile memory at \emph{run-time} that can break the atomicity of the \emph{virtual task}.
-Consider the example depicted in Fig.~\ref{fig:virtualization}: Tasks 1, 2 and 3 being executed consecutively.
-All tasks in this example are atomic since they do not have WAR dependency on the persistent variables they are accessing, e.g.
-\emph{x} is only read and \emph{y} is only written within Task 1.
-Now, suppose three tasks have been virtualized into a single one at run-time, namely Task 4: since \emph{x} is now first \emph{read} and then \emph{written}, a WAR dependency on \emph{x} is introduced dynamically at run-time---unfortunately Task 4 is \emph{no more atomic} since its re-execution will not always produce the same results.
+%A key challenge is that the length of a software
+%task's execution is \emph{limited by the fixed total amount of energy} that a
+%device can buffer in hardware. A task's code is static, but the duration of its
+%execution may be input-dependent and is difficult to predict. To illustrate
+%this, we refer to Figure~\ref{fig:1} showing the execution time of two
+%applications (the same application X, but divided into X and Y tasks as
+%in~\cite{chain}) running on RFID antenna-powered Computational
+%RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna
+%distances (near---stable energy supply/far---high energy intermittency). At
+%short distance: execution will takes too long caused by runtime cost of
+%marshalling excessively small tasks. At far distance: program \emph{might never
+%execute} if task execution consumes more energy than the system can buffer.
+%This calls for at-runtime adaptive task division of any transiently-powered
+%application---namely \emph{task virtualization}.
 
-\textbf{Guaranteeing Atomicity of Virtual Tasks:} Although WAR dependencies introduced by the static task division can be eliminated at compile time, e.g.~\cite{alpaca}, those introduced dynamically at run-time, due to task virtualization, cannot.
-Therefore, we require a \emph{new execution model} that keeps each virtual task atomic by committing the modified persistent variables at the boundary defined dynamically at run-time---keeping the non-volatile memory unmodified upon a power-interrupt and preserving its consistency.
+\textbf{Task Coalescing.} To add support for dynamic task
+dvision to a task-based programming model, we propose a
+\emph{task coalescing} mechanism: merging of atomic tasks
+into a single (larger) task while preserving atomicity.
+Merging two atomic tasks might produce a non-atomic task,
+that may leave memory inconsistent after a partial
+execution. A merge breaks atomicity when the accesses to
+non-volatile memory from the previously atomic merged tasks
+form a \emph{write-after-read (WAR)} dependency.
+%
+%variables in non-volatile memory at \emph{run-time} that can
+%break the atomicity of the \emph{virtual task}. Consider the
+%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
+%2 and 3 being executed consecutively. All tasks in this
+%example are atomic since they do not have WAR dependency on
+%the persistent variables they are accessing, e.g. \emph{x}
+%is only read and \emph{y} is only written within Task 1.
+%Now, suppose three tasks have been virtualized into a single
+%one at run-time, namely Task 4: since \emph{x} is now first
+%\emph{read} and then \emph{written}, a WAR dependency on
+%\emph{x} is introduced dynamically at
+%run-time---unfortunately Task 4 is \emph{no more atomic}
+%since its re-execution will not always produce the same
+%results.
+%
+Although WAR dependencies introduced by the \emph{static}
+task division can be eliminated at compile time~\cite{alpaca}, those introduced \emph{dynamically} at
+run-time cannot.
+%
+To maintain memory consistency in merged tasks, we propose a
+mechanism for \emph{memory virtualization} that buffers the
+updates to non-volatile variables in volatile memory before
+committing them at the dynamic task boundary.
+%
+%Therefore, we require a \emph{new execution
+%model} that keeps each virtual task atomic by committing the
+%modified persistent variables at the boundary defined
+%dynamically at run-time---keeping the non-volatile memory
+%unmodified upon a power-interrupt and preserving its
+%consistency.
+%
+
+Task coallescing removes the optimization burden of choosing
+the best task decomposition from the programmer, because it
+accepts any decomposition and improves it dynamically.
+%
+To further reduce the programmer effort, we propose a
+compiler pass for \emph{automatic decomposition} of programs
+into (small) \emph{atomic} tasks.
+%
+The compiler identifies non-volatile variables shared across
+tasks as tasks are created, and instruments reads and writes
+of those variables, using memory virtualization, to keep
+the data consistent in the presence of power loss.
+%
+Despite being limited to a subset of the C language, the
+automatic task decomposition allowed us to port several
+applications to an intermittent platform with a moderate
+effort.
+
+We develop \sys%
+\footnote{To be released at \url{http://anonymized.link}}
+%
+: a task coallescing system for task-based
+programming models on intermittently-powered devices. The
+capabilities required for effective coallescing mark our
+contributions:
+
+\begin{itemize}
+\item Two adaptive coallescing strategies for selecting boundaries to coallesce based on past execution behavior
+\item Software memory virtualization for transiently-powered devices that preserves atomicity of coallesced tasks
+\item Compiler pass for automatic decomposition into small atomic tasks for adaptive coallescing at runtime
+\end{itemize}
 
@@
 
-\textbf{Research Question and Contributions:} Specifically, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?
+%\textbf{Research Question and Contributions:} Specifically, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?
 
-To address this question, we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
+%To address this question, we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
 To accomplish this, \sys introduces two new capabilities:
 
-\begin{itemize}
-	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task.
+%\begin{itemize}
+%	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task.
 A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer.
 Coalescing tasks allows running the same tasks on a device with a larger energy buffer, avoiding run time overheads associated with ending one task and beginning another.
-	%
-	\item {\bf Volatile memory virtualization with transiently-powered device:} Colascing through task virtualization requires memory virtualization.
+%	%
+%	\item {\bf Volatile memory virtualization with transiently-powered device:} Colascing through task virtualization requires memory virtualization.
 \sys's uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution.
@@
 To the best of our knowledge this is the first time when memory virtualization has been demonstrated on a transient power budget.
-\end{itemize}
+%\end{itemize}
 
-We evaluated \sys end-to-end on a collection of existing benchmarks ~\cite{chain} and new programs, and compare directly to a state-of-the-art task-based runtime~\cite{chain}, showing that \sys has high performance and can flexibly target a variety of platforms without rewriting code.
+We evaluated \sys end-to-end on a real energy-harvesting
+device~\cite{wisp} running existing benchmarks ~\cite{chain}
+and new programs. A comparison to a state-of-the-art
+task-based system~\cite{chain} showed that \sys recovers
+performance on conservative task decompositions missed by
+systems with static decomposition.
+%
+%and can flexibly target a variety of platforms without rewriting code.
+%
 \todo{Provide concrete numeric results supporting this claim}{Amjad/Przemek}
 
-In this work we also address the problem of task construction burden imposed on the programmer.
+%In this work we also address the problem of task construction burden imposed on the programmer.
 That is, we introduce \sys's compiler that transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints.

### 2017-11-12 paper/coalescing.tex

@@
 \subsection{Operation of Program Profiling-driven Coalescing:} \todo{To be written}{Przemek}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}
+	\caption{Result of benchmark applications (Section~ref{}) task size measurement, measured in the number of instruction cycles.
+We observe very large variations of task sizes per each benchmark.
+\todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
+\label{fig:task_profiling}
+\end{figure}
+
+Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}.
+Each benchmark was run for predefined period of time, on non-intermittent power, and at each task evocation, its execution time was measured.
+If a task was evoked more than once, longer execution time replaces the short one---this way we obtain conservative values of task sizes.
+At the end of task measurement, normalized values of task execution are assigned for each task, the shortest one being one task, the longest one X tasks.
 

### 2017-11-12 paper/texFigCode/alpaca-war.tex

--- 2017-06-27 paper/presentation/images/chain-virtualization.tex
@@
 	punkt/.style={
-		circle,
+		rectangle,
 		rounded corners,
 		draw=black, very thick,
-		text width=2.5em,
+		text width=4.5em,
 		minimum height=1em,
@@
 
-	\node[punkt] (task1) {Task (1)};
-	\node[punkt, below =0.7cm of task1] (task2) {Task (2)};
-	\node[punkt, below =0.7cm of task2] (task3) {Task (3)};
-	\node[punkt, below =0.7cm of task3] (task4) {Task (4)};
-
-	\path[every node/.style={font=\sffamily\small}]
-	(task1.west) edge [->, dotted, thick, bend right] node {Channel (1)} (task2.west)
-	(task2.west) edge [->, dotted, thick, bend right] node {Channel (2)} (task3.west)
-	(task3.west) edge [->, very thick, blue, bend right] node {Channel (3)} (task4.west);
+	\node[punkt] (task1) {y = x+1};
+	\node[below = 0.1cm of task1] (t1) {Task 1};
+	\node[punkt, below =0.7cm of task1] (task2) {z = y+1};
+	\node[below = 0.1cm of task2] (t2) {Task 2};
+	\node[punkt, below =0.7cm of task2] (task3) {x = z+1};
+	\node[below = 0.1cm of task3] (t3) {Task 3};
+	\node[punkt, blue, dotted, right = 2cm of task2, fill=red!10] (task4) {y = \textcolor{red}{x}+1 
+		\\ z = y+1 \\ \textcolor{red}{x} = z+1};
+	\node[below = 0.3cm of task4](t4) {Merged tasks};
 
 	\begin{pgfonlayer}{bg}    % select the background layer
-		\node[draw,thick,dotted,red,minimum width=7em, fill=blue!10, fit=(task1) (task2) (task3)] (Virtualized) {};
+		\node[draw,thick,dotted,red,minimum width=7em, fill=blue!10, fit=(task1) (task2) (t3) ] (Virtualized) {};
 		\draw[<->] ([xshift=1.5cm]Virtualized.north) --node[sloped, anchor=south] {Virtualization Window} ([xshift=1.5cm]Virtualized.south);

### 2017-11-12 paper/coalescing.tex
First two paragraphs of coalescing section rewritten

@@
-\sys dynamically {\em coalesces} consecutively executing tasks at run time.
-When two tasks are coalesced, the first of the two tasks does not commit its updates to protected memory locations when it reaches a \transition statement.
-Instead, execution immediately continues from the beginning of the second of the two tasks.
-At the end of the second of the two coalesced tasks, \sys commits the pages that were updated by either of the coalescing tasks.
-The purpose of coalescing is to better amortize the run time overhead of committing updated pages when a task ends.
+Often, energy available at the transiently-powered device is sufficient to run multiple consecutive tasks without page commit at each task boundary.
+In this case task-based execution model can better amortize the run time overhead of committing updated pages when a task ends.
+Hence many tasks can be \emph{merged} into a single virtual task---a process we call \emph{coalescing}.
+
+\sys, as the first task-based execution model, dynamically coalesces consecutively executing tasks at run time.
+When $X$ tasks are coalesced, the $X-1$ tasks do not commit their updates to protected memory locations when reaching a \transition statement.
+Instead, execution immediately continues from the beginning of the $X$th task.
+At the end of the $X$th coalesced task, \sys commits the pages that were updated by any of the $X$ coalescing tasks.
 If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed once.

### 2017-11-13 paper/intro.tex
fixing a fig
Pass over Alexei revision of intro

@@
 \begin{figure}
-	%\includegraphics[width=.5\columnwidth]{presentation/images/alpaca-war.pdf}
-    \caption{\label{fig:coalesce}Task decomposition is required for
-intermittent execution but task transitions introduce overhead.}
+	%\includegraphics[width=\columnwidth]{figures/}
+    \caption{\label{fig:coalesce}Task decomposition is required for intermittent execution but task transitions introduce overhead.}
 \end{figure}
@@
 have executed {\em atomically}, despite arbitrarily-timed power failures.
-The programmer explicitly expresses task-to-task control flow.
-%, which may be conditionally input-dependent.
-%
-Figure~\ref{fig:decomp} illustrates how the decomposition into tasks affects
+The programmer explicitly expresses task-to-task control flow. Figure~\ref{fig:coalesce} illustrates how the decomposition into tasks affects
 the time the application takes to complete.
-%
-Each task transition introduces the overhead of tracking and atomically
-committing the modifications to non-volatile memory by the task, to maintain
-consistency of program state.
-%
-Naively minimizing the number of task transitions, (i.e. the task count),
-\emph{at compile time} risks creating a task that requires more energy than the
-device can buffer in its storage capacitor.
-%
+Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory by the task, to maintain consistency of program state.
+Naively minimizing the number of task transitions, (i.e.
+the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its storage capacitor.
 Such a large task would fail to complete unless the environment provides
 sufficient incoming energy to supplement the energy in the capacitor.
-%
-To eliminate this risk, the programmer (or the compiler) must decompose the
-program conservatively into many small tasks, at the cost of higher task
-transition overhead.
+To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.
 
-One approach to decrease the task count, without assumming a minimum level of
+One approach to decrease the task count, without assuming a minimum level of
 incoming power always available, is to predict task energy consumption and
 choose a decomposition specific to the device capacitor size.
-%
-Howerver, statically predicting energy of arbitrary input-dependent code with
-peripheral access is a problem without a general solution.
-%
-Furthermore, a static decomposition approach cannot produce decompositions
-portable across devices with different storage capacitorss.
-%
-In this paper, we propose a \emph{dynamic} task decomposition approach, that
-accepts any static decomposition and opportunistically \emph{coalesces} tasks
-at runtime when there is sufficient energy.
+However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution.
+Furthermore, a static decomposition approach cannot produce decompositions portable across devices with different storage capacitors.
 
@@
 
-\textbf{Task Coalescing.} To add support for dynamic task
-dvision to a task-based programming model, we propose a
-\emph{task coalescing} mechanism: merging of atomic tasks
-into a single (larger) task while preserving atomicity.
-Merging two atomic tasks might produce a non-atomic task,
-that may leave memory inconsistent after a partial
-execution. A merge breaks atomicity when the accesses to
-non-volatile memory from the previously atomic merged tasks
-form a \emph{write-after-read (WAR)} dependency.
+\textbf{Task Coalescing.} To address this challenge we propose a \emph{dynamic} task decomposition approach, that accepts any static decomposition and opportunistically \emph{coalesces} tasks at runtime when there is sufficient energy.
+The \emph{task coalescing} is a mechanism where atomic tasks are merged into a single (larger) task while preserving atomicity.
+
+Unfortunately, merging two atomic tasks might produce a non-atomic task, that may leave memory inconsistent after a partial execution.
+A merge breaks atomicity when the accesses to non-volatile memory from the previously atomic merged tasks form a \emph{write-after-read (WAR)} dependency.
 %
@@
 %
-Although WAR dependencies introduced by the \emph{static}
-task division can be eliminated at compile time~\cite{alpaca}, those introduced \emph{dynamically} at
-run-time cannot.
-%
-To maintain memory consistency in merged tasks, we propose a
-mechanism for \emph{memory virtualization} that buffers the
-updates to non-volatile variables in volatile memory before
-committing them at the dynamic task boundary.
+Although WAR dependencies introduced by the \emph{static} task division can be eliminated at compile time~\cite{alpaca}, those introduced \emph{dynamically} at
+run-time cannot. To maintain memory consistency in merged tasks, we propose a
+mechanism for \emph{memory virtualization} that buffers the updates to non-volatile variables in volatile memory before committing them at the dynamic task boundary.
 %
@@
 %consistency.
-%
 
-Task coallescing removes the optimization burden of choosing
-the best task decomposition from the programmer, because it
-accepts any decomposition and improves it dynamically.
-%
-To further reduce the programmer effort, we propose a
-compiler pass for \emph{automatic decomposition} of programs
-into (small) \emph{atomic} tasks.
-%
-The compiler identifies non-volatile variables shared across
-tasks as tasks are created, and instruments reads and writes
-of those variables, using memory virtualization, to keep
-the data consistent in the presence of power loss.
-%
-Despite being limited to a subset of the C language, the
-automatic task decomposition allowed us to port several
-applications to an intermittent platform with a moderate
-effort.
+Task coalescing removes the optimization burden of choosing the best task decomposition from the programmer, because it accepts any decomposition and improves it dynamically.
+To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks.
+The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss.
+Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
+applications to an intermittent platform with a moderate effort.
 
-We develop \sys%
-\footnote{To be released at \url{http://anonymized.link}}
-%
-: a task coallescing system for task-based
-programming models on intermittently-powered devices. The
-capabilities required for effective coallescing mark our
-contributions:
+\textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: a task coalescing system for task-based programming models on intermittently-powered devices.
+The capabilities required for effective coalescing mark our contributions:
 
 \begin{itemize}
-\item Two adaptive coallescing strategies for selecting boundaries to coallesce based on past execution behavior
-\item Software memory virtualization for transiently-powered devices that preserves atomicity of coallesced tasks
-\item Compiler pass for automatic decomposition into small atomic tasks for adaptive coallescing at runtime
+	\item Two adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
+	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
+	\item Compiler pass for automatic decomposition into small atomic tasks for adaptive coalescing at runtime.
 \end{itemize}
@@
 
-We evaluated \sys end-to-end on a real energy-harvesting
-device~\cite{wisp} running existing benchmarks ~\cite{chain}
-and new programs. A comparison to a state-of-the-art
-task-based system~\cite{chain} showed that \sys recovers
-performance on conservative task decompositions missed by
-systems with static decomposition.
+We evaluated \sys end-to-end on a real energy-harvesting device~\cite{wisp} running existing benchmarks ~\cite{chain} and new programs.
+A comparison to a state-of-the-art task-based system~\cite{chain} showed that \sys recovers
+performance on conservative task decompositions missed by systems with static decomposition.
 %

### 2017-11-13 paper/background.tex

@@
 
-Energy harvesting devices operate using energy extracted from sources such as RF and solar energy.
+Energy harvesting devices operate using energy extracted from sources such as radio frequency transmissions and solar energy.
 These devices elide tethered power or a battery, instead collecting energy into a capacitor, operating when sufficient energy accumulates, and upon depleting the energy, turning off and recharging.
@@
 
-\subsection{Energy Harvesting and Intermittent Computation}
+\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
+We assume a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM.
+Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
+We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
+\sys supports I/O behaviour similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.
+
+\subsection{Intermittent Execution}
 \label{sec:background_consistency}
@@
 %
-Software running on an energy-harvesting device executes {\em intermittently} because buffered energy is only sometimes available.
+Software running on an energy-harvesting device executes {\em intermittently} because buffered energy is only available \emph{sometimes}.
 An intermittent execution is composed of operating periods interspersed with power failures~\cite{dino,chain,alpaca,ratchet}.
@@
 \begin{figure}
-	\centering
-	%\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
-	%\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
-	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
-	\label{fig:code_demo_incosistency}
+	\begin{subfigure}[t]{\linewidth}
+		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
+		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
+	\end{subfigure}
+	\begin{subfigure}[t]{\linewidth}
+		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
+		\caption{Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
+    \end{subfigure}
+	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}\label{fig:code_demo_incosistency}
 \end{figure}
+
+%\begin{figure}
+%	\centering
+%	\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
+%	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
+%	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
+%	\label{fig:code_demo_incosistency}
+%\end{figure}
 
@@
 Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value.
-The situation occurs when code includes a write after read (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}.
+The situation occurs when code includes a WAR dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}.
 
-Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench~\cite{hicks_mibench2_2016}.
+Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
 The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
@@
 After restarting, line 9 reads the updated value of $\texttt{Rmnd}$, producing an incorrect $\texttt{data}$ value.
-%
 Checkpoint-based systems risk violating memory consistency~\cite{dino}.
@@
 
-\subsection{Task-based Intermittent Computation}
+\subsection{Task-based Intermittent Programming}
 \label{section:background_task_computing}
@@
 
-\noindent \textbf{Memory Consistency Enforcement Overhead:} Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
+\textbf{Memory Consistency Enforcement Overhead.} Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
 Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf.
@@
 \begin{figure}
-	\centering
+\begin{subfigure}[t]{.49\columnwidth}
+	\centering \includegraphics[width=\columnwidth]{figures/fram_write.eps}
+	\caption{Cost of \emph{write} operation.}
+\end{subfigure}%
+\begin{subfigure}[t]{.49\columnwidth}
+	\centering \includegraphics[width=\columnwidth]{figures/fram_read.eps}
+	\caption{Cost of \emph{read} operation.}
+\end{subfigure}
+	\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write and read operation.\todo{pico jole$\rightarrow$Picojoule}{Amjad}}\label{fig:framEnergy}
+\end{figure}
+
+%\begin{figure}
+%\centering
 	%\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
 	%\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
-	\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
-	\label{fig:framEnergy}
-\end{figure}
+	%\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
+	%\label{fig:framEnergy}
+%\end{figure}
 
-To understand these non-volatile memory overheads, we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations for the MSP430FR5969~\cite{msp430datasheet} MCU.
+To understand these non-volatile memory overheads, for the MSP430FR5969~\cite{msp430datasheet} MCU we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations.
 We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device across 1600 read and write accesses to each memory type, randomizing to avoid caching.
@@
 
-\noindent \textbf{Fixed-size Tasks are Inflexible and Inefficient:} If a
-programmer writes a task that consumes {\em more} energy than the device can
-buffer, the task will never complete using buffered energy. {\em preventing
-forward progress} by repeatedly re-executing. If a programmer writes tasks that
-consume far {\em less} energy than a device can buffer, the system may operate
-{\em inefficiently}. When a task and its successor both complete, the two tasks
-were interleaved by a task boundary that preserves progress and state (i.e.,
-checkpoint or commit). However, absent a power failure, the work to preserve
-state was {\em unnecessary}, incurring overhead on the tasks' executions.
-Avoiding excessively costly, non-terminating tasks and short, high-overhead
-tasks is a programming challenge, given fixed hardware with a fixed energy
-buffering capacity. Buffer sizes from prior work vary widely from $20\mu F$~\cite{rodriguez_tbcs_2015}
-to $0.1F$~\cite{moo}.
-Sizing tasks complicates 
-{\em porting} code from one device to another. An excessively costly task 
-on a device with a small energy buffer may be a relatively short,
-task on a device with a larger buffer. The
-key problem is that using existing systems, the programmer statically sizes tasks for a fixed energy buffer. \sys's {\em task coalescing}, described
+\textbf{Fixed-size Tasks are Inflexible and Inefficient.} If a programmer writes a task that consumes {\em more} energy than the device can buffer, the task will never complete using buffered energy.
+{\em preventing forward progress} by repeatedly re-executing.
+If a programmer writes tasks that consume far {\em less} energy than a device can buffer, the system may operate {\em inefficiently}.
+When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).
+However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions.
+Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity.
+Buffer sizes from prior work vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}.
+Sizing tasks complicates {\em porting} code from one device to another.
+An excessively costly task on a device with a small energy buffer may be a relatively short, task on a device with a larger buffer.
+The key problem is that using existing systems, the programmer statically sizes tasks for a fixed energy buffer.
+\sys's {\em task coalescing}, described
 in Section~\ref{sec:task_coalescing}, is motivated by these observations.
-
-\subsection{Hardware Assumptions}
-\label{sec:background_hardware}
-
-\sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
-We assume a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform~\cite{wisp} is equipped with a mixture of SRAM and FRAM.
-Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
-We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
-\sys supports I/O behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.

### 2017-11-13 paper/intro.tex
intro: draft of fig1

@@
 \begin{figure}
-	%\includegraphics[width=\columnwidth]{figures/}
-    \caption{\label{fig:coalesce}Task decomposition is required for intermittent execution but task transitions introduce overhead.}
+	\includegraphics[width=0.9\columnwidth]{figures/coalescing-horiz}
+    \caption{\label{fig:coalesce}Task coalescing at runtime reduces time and
+energy overhead in task-based intermittent programming models by performing
+fewer commits and buffering updates to non-volatile variables in volatile
+memory.}
 \end{figure}

### 2017-11-13 paper/coalescing.tex
Work on task coaleascing (not ready yet)

@@
 
-\sys, as the first task-based execution model, dynamically coalesces consecutively executing tasks at run time.
+\textbf{Process of Task Coalescing.} \sys, as the first task-based execution model, dynamically coalesces consecutively executing tasks at run time.
 When $X$ tasks are coalesced, the $X-1$ tasks do not commit their updates to protected memory locations when reaching a \transition statement.
@@
 If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed once.
+\sys repeatedly executes $X-1$ coalesced tasks, followed by a committing task.
 
-\textbf{Task Coalescing Problem:} \sys can coalesce an arbitrary number of consecutive tasks.
+\textbf{Coalescing Speed/Wasted Effort Trade-off.} \sys can coalesce an arbitrary number of consecutive tasks.
 As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance.
-As more tasks coalesce, the amount of potentially {\em wasted work} also increases.
-If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of the tasks in the long sequence.
-\sys moderates the risk of coalescing and capitalizes on its benefit by dynamically, \emph{adaptively} determining a coalescing factor that combines many tasks without suffering an excess of wasted work.
-
+However, as more tasks coalesce, the amount of potentially {\em wasted work} also increases.
+If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of the tasks in this long sequence.
 The challenge to coalescing tasks is determining how many tasks to coalesce before committing.
 During execution, an executing task is either a {\em coalesced} task, which does not commit its state before transitioning or a {\em committing} task, which does commit its state before transitioning.
-\sys monitors a program's execution to adaptively determine $C$, the number of tasks to coalesce before committing; \sys repeatedly executes $C-1$ coalesced tasks,
-followed by a committing task.
 
-\textbf{Task Coalescing Strategies:} There are infinite possible strategies for coalescing (infinite amount of $(C_U,C_D)$ pairs). We will consider the three fundamental ones:
+\textbf{Task Coalescing Strategies.} The simplest strategy will coalesce/de-coalesce fixed number of task at every successful/unsuccessful commit operation, where task coalescing factor increases by $C_U$ at each successful virtual task execution and decreases by $C_D$, otherwise.
+There are naturally infinite possible strategies for coalescing, i.e.
+infinite amount of $(C_U,C_D)$ pairs.
+The caveat of predefined $(C_U,C_D)$ pair is inability to adapt to changing energy availability conditions.
+A \emph{good} coalescing strategy must moderates the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work.
+Therefore we propose to consider two fundamental coalescing strategies for \sys:
 
 \begin{enumerate}
-	\item \textbf{$N$ up/$N$ down:} where task coalescing factor increases by $N$ at each successful virtual task execution and decreases by $N$, otherwise;
-	\item \textbf{History-driven:} where coalescing will consider history of power failures into deciding on $(C_U,C_D)$ values on the fly;
-	\item \textbf{Program profiling-driven:} where coalescing strategy takes into account \emph{expected size} of each task in the application to pick the best coalescing factor.
+	\item \textbf{Power Interrupt History-driven Coalescing:} ; and 
+	\item \textbf{Power Interrupt History-driven Coalescing with Weights:} where coalescing strategy takes into account \emph{expected size} of each task in the application to pick the best coalescing factor.
 \end{enumerate}
 
-\subsection{Operation of $N$ up/$N$ Down Coalescing:} \todo{To be written}{Przemek}
-
-\subsection{Operation of History-driven Coalescing}: \sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the current operational period.
+\subsection{Operation of Power Interrupt History-driven Coalescing.} \sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the current operational period.
 \sys begins execution in a coalesced task with $C$ initialized to half of the number of tasks executed during the previous operating period and $S$ initialized to $0$.
@@
 If more tasks complete in an operating period, \sys attempts to coalesce more tasks.
-If fewer tasks
-complete, \sys attempts to coalesce fewer tasks. If no committing task
-completes due to a power failure, \sys updates $C$ on reboot to half of $S$
-(the number of tasks completed before the power failure) and continues.
+If fewer tasks complete, \sys attempts to coalesce fewer tasks.
+If no committing task completes due to a power failure, \sys updates $C$ on reboot to half of $S$ (the number of tasks completed before the power failure) and continues.
 
-\noindent\textbf{Task Coalescing Illustration:}
-Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks
-executing in \sys. At left, there is no coalescing and tasks incur the cost of
-commit at each transition. The center shows the same tasks executing as
-coalesced groups of two tasks $C=2$, avoiding one commit.  At right, \sys
-coalesces three tasks ($C=3$) avoiding two commits.
-Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value
+\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys.
+At left, there is no coalescing and tasks incur the cost of commit at each transition.
+The center shows the same tasks executing as coalesced groups of two tasks $C=2$, avoiding one commit.
+At right, \sys
+coalesces three tasks ($C=3$) avoiding two commits. Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value
 of $C$ based on $S$. \sys executes tasks in the first operating period, with
@@
 	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}
-	\caption{Result of benchmark applications (Section~ref{}) task size measurement, measured in the number of instruction cycles.
+	\caption{Result of benchmark applications (Section~\ref{}) task size measurement, measured in the number of instruction cycles.
 We observe very large variations of task sizes per each benchmark.

### 2017-11-13 paper/paper.tex
Replace \bibliographystyle with ACM one
Title for paper header created
subfigure fixed
text for virtualization
moving order of main sections: coalescing, paging, compiler
Colons in sections added; computation->computing
minor changes

@@
 %\usepackage{booktabs}   %% For formal tables: http://ctan.org/pkg/booktabs
-%\usepackage{subcaption} %% For complex figures with subfigures/subcaptions http://ctan.org/pkg/subcaption
-
+\usepackage{subcaption} %% For complex figures with subfigures/subcaptions http://ctan.org/pkg/subcaption
 %\usepackage[normalem]{ulem}
+%\usepackage{float}
+\usepackage{soul}
 \usepackage{amssymb}
@@
 
-%\usepackage{float}
 %\usepackage[keeplastbox]{flushend} %Brandon: this broke the build for me
@@
 %\RequirePackage[font={normalsize,sf,bf}]{caption}
-\RequirePackage[position=bottom]{subfig}
+%\RequirePackage[position=bottom]{subfig}
 %\captionsetup[table]{aboveskip=0.5em, belowskip=0.5em}
@@
 \newcommand{\todo}[2]{\textcolor{red}{$\bullet$ \textbf{Task:} #1; \textbf{Responsible:} #2}}
+\newcommand{\add}[1]{\textcolor{blue}{#1}}
+\newcommand{\remove}[1]{\textcolor{red}{\st{#1}}}
+\newcommand{\replace}[2]{\remove{#1} \add{#2}}
 
+\newcommand{\transition}{{\tt next\_task}\xspace}
 \newcommand{\sys}{Coala\xspace}
-\newcommand{\transition}{{\tt next\_task}\xspace}
-
+\newcommand{\sysmed}{\sys: Intermittent Computing $\ldots$}
+\newcommand{\sysfull}{\sys: Adaptive Task Coalescing for Intermittent Computing on Energy-harvesting Devices}
 %newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization} %Old title
-\newcommand{\sysfull}{\sys: Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization}
 
@@
 
-\title{\sysfull}
+\title[\sysmed]{\sysfull}
 
@@
 
-\section{Intermittent Computation: Background}
+\section{Intermittent Computing: Background}
 \label{sec:background}
@@
 
-\section{\sys Automatic Task Generation}
-\label{sec:compiler}
+\section{\sys: Task Coalescing}
+\label{sec:task_coalescing}
 
-\input{compiler}
+\input{coalescing}
 
-\section{\sys Memory Virtualization}
+\section{\sys: Memory Virtualization}
 \label{sec:memory_virtulaization}
@@
 
-\section{\sys Task Coalescing}
-\label{sec:task_coalescing}
+\section{\sys: Automatic Task Generation}
+\label{sec:compiler}
 
-\input{coalescing}
+\input{compiler}
 
@@
 
-\bibliographystyle{plain}
+\bibliographystyle{ACM-Reference-Format}
+%\bibliographystyle{plain}
 \bibliography{bib}

### 2017-11-14 paper/intro.tex
intro: redrawn coalescing figure and prose expanded

--- 2017-11-13 paper/intro.tex
@@
 \begin{figure}
-	%\includegraphics[width=\columnwidth]{figures/}
-    \caption{\label{fig:coalesce}Task decomposition is required for intermittent execution but task transitions introduce overhead.}
+	\includegraphics[width=0.85\columnwidth]{figures/coalescing-is}
+    \caption{\label{fig:coalesce}\textbf{An execution of three
+statically-defined tasks without (top) and with (bottom) dynamic
+coalescing.} Task coalescing at runtime reduces time and energy overhead in
+task-based intermittent programming models by performing fewer commits and
+buffering updates to non-volatile (NV) variables in volatile (V) memory.}
 \end{figure}
@@
 New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
-However, they have their own limitations that need to be addressed---the core one being \emph{inability to adapt} to changing energy conditions.
+However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.
 
@@
 have executed {\em atomically}, despite arbitrarily-timed power failures.
-The programmer explicitly expresses task-to-task control flow. Figure~\ref{fig:coalesce} illustrates how the decomposition into tasks affects
-the time the application takes to complete.
-Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory by the task, to maintain consistency of program state.
+The programmer explicitly expresses task-to-task control flow. The execution
+in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
+tasks affects the time the application takes to complete.
+Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory by the task, to maintain consistency of program state~\cite{chain,alpaca}.
 Naively minimizing the number of task transitions, (i.e.
@@
 
-\textbf{Task Coalescing.} To address this challenge we propose a \emph{dynamic} task decomposition approach, that accepts any static decomposition and opportunistically \emph{coalesces} tasks at runtime when there is sufficient energy.
-The \emph{task coalescing} is a mechanism where atomic tasks are merged into a single (larger) task while preserving atomicity.
+\textbf{Task Coalescing.} To address this challenge we propose a
+\emph{dynamic} task decomposition approach, that accepts any static
+decomposition and opportunistically coalesces tasks at runtime when
+there is sufficient energy. \emph{Task coalescing} is a mechanism where
+atomic tasks are merged into a single (larger) task while preserving
+atomicity of the merged task. If power fails while a colesced task is
+running, execution will restart at the first task in the group.
+Figure~\ref{fig:coalescing} (bottom) illustrates how dynamic coalescing
+changes the earlier execution and highlights the change in memory model
+necessary to not break atomicity.
 
-Unfortunately, merging two atomic tasks might produce a non-atomic task, that may leave memory inconsistent after a partial execution.
-A merge breaks atomicity when the accesses to non-volatile memory from the previously atomic merged tasks form a \emph{write-after-read (WAR)} dependency.
+Unfortunately, naively merging two atomic tasks might produce a non-atomic
+task, that may leave memory inconsistent after a partial execution. A merge
+breaks atomicity when the accesses to non-volatile memory from the
+previously atomic merged tasks form a \emph{write-after-read (WAR)}
+dependency.
 %

### 2017-11-14 paper/coalescing.tex

@@
 A \emph{good} coalescing strategy must moderates the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work.
-Therefore we propose to consider two fundamental coalescing strategies for \sys:
+Therefore we propose to consider two fundamental coalescing strategies for \sys, which we will later assess and compare experimentally:
 
 \begin{enumerate}
-	\item \textbf{Power Interrupt History-driven Coalescing:} ; and 
-	\item \textbf{Power Interrupt History-driven Coalescing with Weights:} where coalescing strategy takes into account \emph{expected size} of each task in the application to pick the best coalescing factor.
+	\item \textbf{Power Interrupt History-based Coalescing (HDC):} when more tasks are coalesced without power interrupt $C_U$ increases gradually, and $C_D$ increases, as more task are interrupted before committing; and
+	\item \textbf{Weighted Power Interrupt History-based Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
 \end{enumerate}
 
-\subsection{Operation of Power Interrupt History-driven Coalescing.} \sys determines $C$ for a given operation based on the number of tasks completed, $S$, during the current operational period.
-\sys begins execution in a coalesced task with $C$ initialized to half of the number of tasks executed during the previous operating period and $S$ initialized to $0$.
-The system executes $C-1$ coalesced tasks followed by one committing task.
-After each commit, \sys assigns $C$ to half of $S$.
-If more tasks complete in an operating period, \sys attempts to coalesce more tasks.
-If fewer tasks complete, \sys attempts to coalesce fewer tasks.
-If no committing task completes due to a power failure, \sys updates $C$ on reboot to half of $S$ (the number of tasks completed before the power failure) and continues.
+\subsection{Power Interrupt History-driven Coalescing}
 
-\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys.
+With this strategy \sys calculates only one parameter $C=C_U=C_D$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period.
+\sys begins execution in a coalesced task with $S_t=0$.
+When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows.
+After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$.
+This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks.
+However, tasks coalescing are increased by a half the speed of the successful task completion.
+This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt.
+In the power failure case, if fewer tasks complete, \sys attempts to coalesce fewer tasks.
+
+%\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys.
 At left, there is no coalescing and tasks incur the cost of commit at each transition.
 The center shows the same tasks executing as coalesced groups of two tasks $C=2$, avoiding one commit.
-At right, \sys
-coalesces three tasks ($C=3$) avoiding two commits. Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value
-of $C$ based on $S$. \sys executes tasks in the first operating period, with
-$C$ equal to half the number of tasks completed in the prior execution period.
-In the example, $C$ starts at eight. After completing eight tasks, \sys commits
-and reduces $C$ by half to four.  \sys completes two of four tasks and power
-fails.  During the operating period, \sys executed a total of ten tasks (i.e.,
-$S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$).
-\sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$
-to half of $S$, which is three (after rounding up). \sys coalesces three tasks
-and again commits.  At this point, $S=8$; \sys assigns $C=4$ and execution
-continues.
+At right, \sys coalesces three tasks ($C=3$) avoiding two commits.
+Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$.
+\sys executes tasks in the first operating period, with $C$ equal to half the number of tasks completed in the prior execution period.
+In the example, $C$ starts at eight.
+After completing eight tasks, \sys commits and reduces $C$ by half to four.
+\sys completes two of four tasks and power fails.
+During the operating period, \sys executed a total of ten tasks (i.e., $S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$).
+\sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$ to half of $S$, which is three (after rounding up).
+\sys coalesces three tasks and again commits.
+At this point, $S=8$; \sys assigns $C=4$ and execution continues.
 
-\subsection{Operation of Program Profiling-driven Coalescing:} \todo{To be written}{Przemek}
+\subsection{Weighted Power Interrupt History-driven Coalescing}
+
+The previous coalescing strategy is oblivious to the task size.
+Although many tasks can be coalesced, their non-equal size will determine that some coalesced tasks will be more susceptible to power interrupts than others.
+To understand that effect we provide the following illustration.
+
+\textbf{Task Size Profiling.} To measure how large the distribution of tasks is, we have used a set of program benchmarks (refer to Section~\ref{sec:results_software} for details and to the beginnin of Section~\ref{sec:methodology_evaluation} for system setup).
+Each benchmark was run for predefined period of time (X\,seconds \todo{Provide exact number}{Amjad}), on non-intermittent power, and at each task evocation, its execution time was measured.
+If a task was evoked more than once, longer execution time replaced the short one---this way we obtained conservative values of task sizes.
+At the end of task measurement, normalized values of task execution are assigned for each task, the shortest one being one task, the longest one X tasks \todo{Provide exact number}{Amjad}.
+Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}.
+We observe huge variations of task sizes---both within a benchmark, as well per benchmark, which calls for the use of this information in coalescing.
 
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}
-	\caption{Result of benchmark applications (Section~\ref{}) task size measurement, measured in the number of instruction cycles.
+	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}%coalescing_timeline
+	\caption{Result of benchmark applications (Section~\ref{sec:results_software}) task size measurement, measured in the number of instruction cycles.
 We observe very large variations of task sizes per each benchmark.
@@
 
-Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}.
-Each benchmark was run for predefined period of time, on non-intermittent power, and at each task evocation, its execution time was measured.
-If a task was evoked more than once, longer execution time replaces the short one---this way we obtain conservative values of task sizes.
-At the end of task measurement, normalized values of task execution are assigned for each task, the shortest one being one task, the longest one X tasks.
+In the second coalescing strategy an application needs to be profiled, as above, before it is loaded to the intermittently-powered device.
 
-\begin{figure}
-	\centering
+%\begin{figure}
+%	\centering
 	%\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
 	%\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
-	\caption{\sys coalesces tasks to avoid overheads.}
-	\label{fig:coalescing_example}
-\end{figure}
+	%\caption{\sys coalesces tasks to avoid overheads.}
+	%\label{fig:coalescing_example}
+%\end{figure}
 

### 2017-11-14 paper/compiler.tex
partial pass over compiler.  I got interrupted, but I will come back to this latter today

@@
-\sys provides compiler support for \emph{automatic task generation} that simplifies writing task-based code by automatically translating C code into a task-based executable.
-Doing so requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task, or are made explicit in inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
-The \sys compiler, based on LLVM, first analyzes the control-flow graph (CFG) and, using a rule-based heuristic, determines at which points to place a task boundary.
-After identifying where to put task boundaries, the \sys compiler transforms the imperative code into tasks, instrumenting protected memory accesses.
+We studied \sys's effectiveness for systems with automatically generated task
+boundaries by developing a simple compiler prototype that transforms an
+arbitrary C into a program with tasks. The compiler is representative of,
+although not a direct implementation of, a system that assists in writing
+task-based programs by automatically inserting tasks (e.g., inspired by
+Ratchet~\cite{ratchet}). 
+%
+Translating C code to use tasks requires identifying program points to use as
+task boundaries, defining a set of task functions that begin and end at those
+boundaries, ensuring that all existing control-flow paths are either contained
+within a task or are made explicit with inter-task control-flow, and identifying
+and instrumenting accesses to task-shared variables that need to be protected
+by \sys at runtime. Our simple prototype, which is implemented using LLVM, analyzes the
+control-flow graph (CFG) and uses a rule-based heuristic to determine at which
+points to place a task boundary. After identifying where to put task
+boundaries, the \sys compiler transforms the imperative code into tasks,
+instrumenting protected memory accesses.
 
@@
 
-\sys analyzes a program to determine where to put task boundaries to divide the program into tasks.
-A task is a single-entry, multiple-exit region of the CFG and in a correct task decomposition, no task should ever consume more energy than the device's energy buffer can store.
-There are several factors that determine whether tasks will execute efficiently.
-A task decomposition should minimize the number of variables that are accessed by multiple tasks and, consequently, must be protected by \sys's paging mechanism.
-Tasks should be roughly the same size, simplifying \sys's dynamic reasoning about coalescing by eliminating the need to consider variable task cost at run time.
-To minimize overhead introduced at task boundaries (i.e., commit), a task decomposition should maximize task size, which minimizes the frequency of task transitions.
+\sys analyzes a program to determine where to put task boundaries. 
+%
+A task is single-entry, multiple-exit part of the CFG that our compiler excises
+into a new function.  
+%
+There are several factors that determine whether tasks will execute
+efficiently. A task decomposition should minimize the number of variables that
+are accessed by multiple tasks and, consequently, must be protected by \sys's
+paging mechanism. 
+%
+Different tasks should be approximately the same size to make reasoning about
+coalescing simple by eliminating the need to consider variable run time costs. 
+%
+To minimize task boundary overheads (e.g., commit), a task decomposition should
+maximize task size, which minimizes the frequency of task transitions.
 
@@
 
-The input to the task boundary identification analysis is a standard C program (as shown in Figure~\ref{fig:compiler_overview}) and the output of \sys's analysis is the CFG, with some vertices marked as {\em task boundaries}.
-The \sys compiler builds the program's CFG (as shown in
-Figure~\ref{fig:compiler_overview}).
-To simplify inter-procedural analysis, our prototype inlines all function calls into the main function in the CFG.
-Recursive function call cannot be handled, so the programmer has to rewrite it into a non-recursive call if there is one.
-The approach of always inlining function calls will introduce code duplication.
-If the code size is problematic, alternative approach is to make each function call a task boundary.
-We leave implementation of such alternative strategy as a future work.
+The task boundary identification analysis processes a standard C program (as
+shown in Figure~\ref{fig:compiler_overview}) and produces a CFG with some
+vertices marked as {\em task boundaries}. The \sys compiler builds the
+program's CFG as shown in Figure~\ref{fig:compiler_overview}. In our simple reference prototype, 
+we implemented inter-procedural analysis by inlining all functions called.   Recursive function call cannot be handled, so the
+programmer has to rewrite it into a non-recursive call if there is one. The
+approach of always inlining function calls will introduce code duplication. If
+the code size is problematic, alternative approach is to make each function
+call a task boundary. We leave implementation of such alternative strategy as a
+future work.
 %
@@
 %
-We first describe how \sys measures each of these costs and then describe \sys's procedure for placing boundaries in the CFG.
+We first describe how \sys measures each of these costs and then describe
+\sys's procedure for placing boundaries in the CFG.
 

### 2017-11-14 paper/systemdescription.tex
tweaks to system and evaluation
System description revised: small tweakts to the text - looks good

@@
+\sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
+\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable}, {\em efficient} and {\em flexible}.
+\sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.
+
 \begin{figure}
@@
 
-\sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
-\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em efficient}, {\em flexible}, and {\em programmable}.
-\sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.
-
-\noindent \textbf{\sys Programming Model:} To use \sys, a programmer first writes plain, imperative C code.
+\textbf{\sys Programming Model.} To use \sys, a programmer first writes plain, imperative C code.
 The programmer then has an option to {\em manually} or {\em automatically} translate the program into code for \sys's task-based programming model.
@@
 \sys consists of its compiler and the run time software support for its execution model and includes several lightweight programming model extensions similar to those in prior work~\cite{chain,alpaca}.
-Table~\ref{tab:viper_syntax} provides an overview of these extensions, which can be used manually by a programmer or added to a program automatically, using \sys's compiler.
+Table~\ref{tab:viper_syntax} summarizes these extensions, which can be used manually by a programmer or added to a program automatically, using \sys's compiler.
 As in prior task-based intermittent programming models~\cite{chain,alpaca}, a \sys program is decomposed into tasks, which are functions that are not allowed to have any direct callers.
 The programmer identifies a task by adding the {\tt task} keyword to the function's declaration.
-A programmer invokes one task, {\tt t} from another task {\tt u} using {\tt next\_task( t )}, which atomically commits the state of {\tt u} and transfers control to the entry point of {\tt t}.
+A programmer invokes one task, {\tt t} from another task {\tt u} using {\tt next\_task(\,t\,)}, which atomically commits the state of {\tt u} and transfers control to the entry point of {\tt t}.
 A program has a single {\em origin} task that executes first during the program's first execution; after a power failure, in subsequent operating periods, \sys restarts from the task that was executing just before the failure.
@@
 		\texttt{next\_task()} & Transition to a new task\\
-		\texttt{RP} & Read-protected variable\\
-		\texttt{WP} & Write-protected variable\\
+		\texttt{RVAR} & Read-protected variable\\
+		\texttt{WVAR} & Write-protected variable\\
 		\hline
 	\end{tabular}
-	\caption{Summary of \sys's programming model extensions.}
+	\caption{\sys's programming model extensions.}
 	\label{tab:viper_syntax}
@@
 
-\noindent \textbf{\sys Memory Manager:} When the binary executes, \sys's runtime support affects its execution through a {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}) and a {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}).
+\textbf{\sys Memory Manager.} When the binary executes, \sys's runtime support affects its execution through a {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}) and {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}).
 The \sys binary executes its tasks correctly, with consistent memory state, due to \sys's virtual memory mechanism.
@@
 
-\noindent \textbf{\sys Task Scheduler:} \sys runs a lightweight task scheduler that uses {\em task coalescing} to make \sys's statically defined tasks efficiently use available energy.
+\textbf{\sys Task Scheduler.} \sys runs a lightweight task scheduler that uses {\em task coalescing} to make \sys's statically defined tasks efficiently use available energy.
 By default, tasks run in sequence and each task commits as it completes, potentially incurring unnecessary overhead for commits between consecutive, non-failing tasks.

### 2017-11-14 paper/relatedwork.tex
Related work section revised: small tweakts to the text - looks good

@@
-{\noindent \textbf{Intermittently-Powered Devices:}} There is a large body of research on energy harvesting and wirelessly powered embedded devices.
+\textbf{Intermittently-Powered Devices.} There is a large body of research on energy harvesting and wirelessly powered embedded devices.
 Several recent reviews provide a broad in-depth overview~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
@@
 intermittent computing hardware platforms has led to the development of tooling
-and instrumentation for such systems
-building~\cite{hester_sensys_2014,hester_sensys_2015,edb,stork,wisent}.
+and instrumentation for such systems building~\cite{hester_sensys_2014,hester_sensys_2015,edb,stork,wisent}.
 
-{\noindent \textbf{Intermittent Execution Environments:}} Early work on Energy-harvesting runtime systems assumed simple, small computations not exceeding a predictable burst of energy~\cite{dewdrop} or long power-cycles/priority-based scheduling~\cite{sorber_sensys_2007}.
+\textbf{Intermittent Execution Environments.} Early work on energy-harvesting runtime systems assumed simple, small computations not exceeding a predictable burst of energy~\cite{dewdrop} or long power-cycles/priority-based scheduling~\cite{sorber_sensys_2007}.
 These efforts did not consider data consistency with intermittence.
-Works supporting intermittent computing using dynamic, on-demand checkpoints for consistency include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall}, Hibernus++~\cite{hibernusplusplus} and HarvOS~\cite{mottola2017harvos}.
+Works supporting intermittent computing using dynamic, on-demand checkpoints include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall}, Hibernus++~\cite{hibernusplusplus} and HarvOS~\cite{mottola2017harvos}.
 Dynamic checkpointing systems may checkpoint
@@
 requirements.
-Moreover, some systems require hardware to measure stored energy, incurring energy cost.
+Moreover, some require hardware to measure stored energy, incurring energy cost.
 Moreover, these systems do not checkpoint non-volatile state, compromising memory consistency.
@@
 
-Recent task-based systems are Chain~\cite{chain} and
-Alpaca~\cite{alpaca}.
+Recent task-based systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}.
 Using static tasks, they eliminate checkpointing volatile state.
@@
 
-{{\noindent \bf Memory Virtualization:}} Prior work on embedded systems has studied a variety of memory virtualization strategies that relate to \sys.
+\textbf{Memory Virtualization.} Prior work on embedded systems has studied a variety of memory virtualization strategies relating to \sys.
 TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory management.
@@
 The goal of \sys is to provide consistent, intermittent execution, leveraging the benefits of efficient bulk copying.
-In contrast, these prior efforts were focused more on programmability and runtime reliability properties provided by virtual memory.
+In contrast, prior efforts focused more on programmability and runtime reliability properties provided by virtual memory.
 
 Another \sys-related prior effort is on unbounded, page-based transactional memory and deterministic parallel runtime systems~\cite{pagebasedtm,grace}.
-These works have a different mechanism and purpose than \sys---ensuring that data are consistent and
-deterministically updated during concurrent executions. What makes these
+These works have a different mechanism and purpose than \sys---ensuring that data are consistent and deterministically updated during concurrent executions.
+What makes these
 efforts related is that they manage state to ensure consistency at the

### 2017-11-14 paper/compiler.tex
pass over all but generation

@@
 vertices marked as {\em task boundaries}. The \sys compiler builds the
-program's CFG as shown in Figure~\ref{fig:compiler_overview}. In our simple reference prototype, 
-we implemented inter-procedural analysis by inlining all functions called.   Recursive function call cannot be handled, so the
-programmer has to rewrite it into a non-recursive call if there is one. The
-approach of always inlining function calls will introduce code duplication. If
-the code size is problematic, alternative approach is to make each function
-call a task boundary. We leave implementation of such alternative strategy as a
-future work.
+program's CFG as shown in Figure~\ref{fig:compiler_overview}. In our simple
+reference prototype, we implemented inter-procedural analysis by inlining all
+functions called.  We opted not to support recursion, which is uncommon in
+embedded code. We leave to future work better approaches to interprocedural
+analysis and task transformation.
+
 %
@@
 %
-We first describe how \sys measures each of these costs and then describe
-\sys's procedure for placing boundaries in the CFG.
 
@@
 
-\sys's task cost analysis considers estimated task energy cost, variable liveness, and the impact of splitting loops across tasks.
-The result of this analysis determines where to place boundaries in the CFG.
+\sys's task cost analysis considers estimated run time, the liveness of
+variables and the cost of splitting loops across tasks. The result of the task
+cost analysis determines where to place boundaries in the CFG.
 
-{\noindent \bf Estimated energy cost.} \sys assigns each basic block in the CFG a {\em cost}, which is an estimate of the energy consumption of the block.
-Precisely modelling the energy cost of arbitrary C code is a difficult problem and beyond the scope of this work; we simply use the number of intermediate representation (IR) instructions in a block as a proxy for energy because more instructions mean more energy consumption.
-Moreover, we assume as an input to the compiler the maximum number of operations that the device can execute without exhausting its buffered energy.
-This parameter is highly device-dependent, and we assume that a developer or device manufacturer determines the number empirically, as we did in this work through direct measurement.
-\sys's analysis tries to ensure that the estimated energy cost of different tasks is roughly the same and does not exceed given energy budget (which is part of the boundary placement process described later).
+{\noindent \bf Estimated run time.} \sys assigns each basic block in the CFG a
+{\em cost}, which is an estimate of the run time of the block.  Precisely
+modeling the run time and energy cost of arbitrary C code is difficult and
+beyond the scope of our simple compiler prototype.  Instead, our prototype counts
+LLVM intermediate representation (IR) instructions and assigns a higher cost to
+a block with more operations, assuming that the block runs for a longer period
+of time and consumes more energy.  
+%
+To make the cost estimate useful, our compiler prototype estimates the maximum
+cost that the device's energy buffering capacitor supports.  \sys estimates
+this quantity by measuring the number of arithmetic operations that can execute
+without a power interruption, starting from a full capacitor. This maximum number
+of operations is a limit on the cost of a task -- any longer task risks not
+executing to completion.
 
-{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of variable that {\em may} be {\em alive} across the edge (i.e., variables whose value may be consumed later on).
-Variables that live across an edge must be protected by \sys's memory virtualization mechanism when boundary is placed at the edge and the resulting inter-task variable increases the run time overhead.
-\sys's tries to minimize the overhead of protected variables by placing boundaries to minimize the number of inter-task live variables.
+{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of
+variables that {\em may} be {\em alive} across the edge (i.e., variables
+assigned before the edge and may be read after the edge). Variables that live
+across an edge must be protected by \sys's memory virtualization mechanism when
+\sys places a task boundary on the edge.  \sys heuristically minimizes variable
+virtualization overhead by placing boundaries to minimize the number of
+live variables that span task boundaries.
 
-{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis must account for the cost of multiple iterations of the loop body. \sys assumes that most loops are
-annotated with a static loop bound. For loops that do not have a static bound, \sys conservatively assumes the loop will be too large to fit within a single task.
+{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis
+must account for the cost of multiple iterations of the loop body. \sys assumes
+that most loops are annotated with a static loop bound. For loops that do not
+have a static bound, \sys conservatively assumes the loop will be too large to
+fit within a single task. Assuming loop bounds is a reasonable simplification:
+while not always the case, long loops in embedded code are often statically
+bounded, especially when code aims to meet strict resource requirements.
 
@@
 
-The analysis tries to minimize the dynamic overhead of traversing task boundaries by avoiding placing a task boundary on a loop with a large bound, when possible.
-At the same time, the analysis must avoid placing a loop in a single task if the total energy cost of all of its iterations is likely to exceed the device's energy buffer.
-\sys first multiplies the estimated energy cost of the blocks in the loop body by the loop bound to determine estimate the cost of the loop's execution.
-If the loop's cost exceeds the device's energy buffer, \sys can place a task boundary on the loop's back-edge, converting each iteration into its own task.
+The analysis tries to minimize the dynamic overhead of traversing task
+boundaries by avoiding placing a task boundary on a loop with a large bound,
+when possible. At the same time, the analysis must avoid placing a loop in a
+single task if the total energy cost of all of its iterations is likely to
+exceed the device's energy buffer. \sys first multiplies the estimated energy
+cost of the blocks in the loop body by the loop bound to determine estimate the
+cost of the loop's execution. If the loop's cost exceeds the device's energy
+buffer, \sys can place a task boundary on the loop's back-edge, converting each
+iteration into its own task.
 

### 2017-11-14 paper/conclusions.tex

@@
 This paper presented \sys: a new task-based runtime for intermittently-powered devices that does not require any hardware support.
-\sys core new features are: (i) automatic task-generation---removing programmer intervention in the program re-composition to the intermittent domain \emph{completely}, (ii) memory virtualization---improving the execution speeds of broad set of benchmarks executed on popular energy-harvesting embedded platform, and (iii) (adaptive) task coalescing---enabling fast execution despite changing energy arrivals and enabling re-use of the same code on intermittent systems with varying energy reservoirs.
+\sys core new features are: (i)  adaptive task coalescing---enabling fast execution despite changing energy arrivals and enabling re-use of the same code on intermittent systems with varying energy reservoirs, (ii) memory virtualization---improving the execution speeds of broad set of benchmarks executed on popular energy-harvesting embedded platform, and (iii) automatic task-generation---removing programmer from the program re-composition to the intermittent domain \emph{almost completely}.
 \sys, with its task creation/execution automation, responses to the need that all existing predecessors of \sys advocated.

### 2017-11-15 paper/paging.tex
paging

@@
 \sys paging scheme partitions the non-volatile memory address space into pages of {\em protected data}.
-As they are accessed by a task, pages of protected data are swapped into a {\em page buffer} that is stored in volatile memory and holds a single page.
-When a modified page is swapped out, it is buffered in a dedicated commit buffer for later.
+As they are accessed by a task, pages of protected data are swapped into a {\em page buffer} \texttt{pagesBuf} that is stored in volatile memory and holds a pre-determined number of pages.
+When any modified page in the page buffer is swapped out, it is buffered in a dedicated {\em commit buffer} \texttt{pagesTemp} for later.
 \sys uses two-phase commit to write buffered, modified pages back to their original memory locations atomically, ensuring data consistency even with power failures.
@@
 The interface includes two operations: \texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and \texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory location {\tt var}.
-{\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory and translates it into a \emph{virtual address} in the page buffer \texttt{pageBuf} in volatile memory.
+{\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory and translates it into a \emph{virtual address} in the page buffer in volatile memory.
 In \sys, a virtual address is composed of a \emph{page tag} and a \emph{page offset}.
@@
 		\State $\texttt{tag}\leftarrow \texttt{getTag(var)}$ 
-		\If { \texttt{tag} != \texttt{CrntPagTag} }	\Comment{Check if the page is in the page buffer}
+		\If { \texttt{tag} is not found in \texttt{pagesBuf} }	\Comment{Check if the page is in the page buffer}
 		\State	\texttt{PageFault(tag)} \Comment{Page is not in the page buffer, bring it}
+		\Else 
+		\State \texttt{idx} $\leftarrow$ \texttt{getPageIndex()} \Comment{Get the page index in the page buffer}
 		\EndIf
 				\State $\texttt{offset}\leftarrow \texttt{getOffset(var)}$ 		
-		\State \texttt{return pageBuf[offset]}  \Comment{Return directly from page buffer}
+		\State \texttt{return pagesBuf[idx][offset]}  \Comment{Return directly from page buffer}
 	\end{algorithmic}
@@
 After address translation, \sys attempts to access the protected variable's location in the volatile paging buffer.
-\sys keeps track of the page tag for the page currently resident in the paging buffer in {\tt CrntPagTag}, which is stored in non-volatile memory.
-When a task accesses a variable, it compares the variable's page tag to {\tt CrntPagTag}.
-If an accessed variable's page is equal to {\tt CrntPagTag} then the accessed variable's page is resident in the page buffer.
+\sys keeps track of the page tags for the pages currently resident in the paging buffer.
+When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt pagesBuf}.
+If the accessed variable's page is resident in the page buffer, then the tag of one of the pages in this buffer is equal to the accessed variable's page tag.
 The executing memory operation indexes the page buffer using the variable's page offset and manipulates that byte (i.e., either reading or writing).
-If the accessed variable's page tag is not equal to {\tt CrntPagTag} then the variable's page is not resident in the page buffer; the operation incurs a {\em page fault}.
-At a page fault, \sys must swap out the page that is resident in the page buffer and swap in the the variable's page.
+If the accessed variable's page tag is not found then the variable's page is not resident in the page buffer; the operation incurs a {\em page fault}.
+At a page fault, \sys must swap out one of the pages that is resident in the page buffer and swap in the variable's page.
 
 Algorithm~\ref{algo:rwar} shows pseudo-code of \texttt{RVAR} and we omit a listing for {\tt WVAR} as it is very similar except it writes the memory location and returns nothing.
-If the page tag of the variable \texttt{var} and \texttt{CrntPagTag} are not equal (e.g.
+If the page tag of the variable \texttt{var} is not found in \texttt{pagesBuf} (e.g.
 Lines 2--3), there is a page fault that requires a new page to be swapped in to the page buffer.
@@
 	\begin{algorithmic}[1]
-		\If {\texttt{isDirty(CrntPagTag)} }	\Comment{Check if the active page is modified}
-		\State \texttt{pageCommit(pagesTemp)} \Comment{Phase I commit the active page}
+		\State \texttt{VictimTag} $\leftarrow$ \texttt{selectVictimPage()} \Comment{Select a page to swap out}
+		\If {\texttt{isDirty(VictimTag)} }	\Comment{Check if the victim page is modified}
+		\State \texttt{pageCommit(VictimTag,pagesTemp)} \Comment{Phase I commit the victim page}
 		\EndIf
@@
 		\EndIf 
-		\State \texttt{CrntPagTag} = \texttt{tag} \Comment{Update tag}
 	\end{algorithmic}
@@
 
-A page fault on a memory access requires \sys to swap out whatever page of data is in the page buffer, preserving updates made to that page, and to copy the accessed page of protected data into the page buffer.
+A page fault on a memory access requires \sys to swap out one of the pages in the page buffer, i.e.
+\emph{victim page}, preserving updates made to that page, and to copy the accessed page of protected data into the page buffer.
 Algorithm~\ref{algo:pagefault} shows the how \sys handles a page fault.
+In our current implementation, we used a \emph{first-in-first-out} page replacement algorithm to select the victim page in \texttt{pagesBuf} (Line 1).
 If both the page being swapped out and the page being swapped in are not {\em dirty}---which means that they do not contain bytes modified by the executing task---then handling a page fault simply requires copying the accessed page into the page buffer.
-If either the {\em resident} page being swapped out or the {\em accessed} page being swapped in is dirty, \sys must take extra steps when handling a page fault.
+If either the {\em victim} page being swapped out or the {\em accessed} page being swapped in is dirty, \sys must take extra steps when handling a page fault.
 
-\paragraph{Dirty resident page.} If a task modified any byte in the resident page (i.e.,  using \texttt{WVAR}), then the page is dirty.
+\paragraph{Dirty victim page.} If a task modified any byte in the victim page (i.e.,  using \texttt{WVAR}), then the page is dirty.
 Before swapping in the accessed page, \sys must first preserve the updated values in the dirty page so that they can eventually be committed back to their locations in memory when the task ends.
-To preserve dirty values, \sys copies the resident page to its dedicated memory region called \texttt{pagesTemp}, which is acts as a double buffer for each page of protected memory.
-Dirty pages copied into {\tt pagesTemp} during a task remain there until the task completes, at which point \sys commits the modified pages from {\tt pagesTemp} back to memory.
+To preserve dirty values, \sys copies the victim page to its dedicated memory region called \texttt{pagesTemp}, which is acts as a double buffer for each page of protected memory.
+Dirty pages copied into {\tt pagesTemp} during a task remain there until the task completes, at which point \sys commits the modified pages from {\tt pagesTemp} back to memory (Lines 2--3).
 
 \paragraph{Dirty accessed page.} If an accessed page is modified, \sys should swap it into the page buffer from its location in {\tt pagesTemp}, rather than swapping it in from its original location in memory.
-As Algorithm~\ref{algo:pagefault} shows in Line 3, \sys checks if the requested page was previously modified by the executing task.
+As Algorithm~\ref{algo:pagefault} shows in Line 4, \sys checks if the requested page was previously modified by the executing task.
 To track whether a page is modified, \sys stores a dirty bit with the page that is set when a task modifies the page.
-If an accessed page's dirty bit is set, the page must be swapped in from \texttt{pagesTemp} rather than its original memory location (Line 4), furnishing the task with the page's modified values.
-If the accessed page is not dirty, \sys swaps it in from its original location in memory (Line 6).
-After swapping the page in, \sys sets \texttt{CrntPagTag} to the tag of the accessed page (Line 7).
+If an accessed page's dirty bit is set, the page must be swapped in from \texttt{pagesTemp} rather than its original memory location (Line 5), furnishing the task with the page's modified values.
+If the accessed page is not dirty, \sys swaps it in from its original location in memory (Line 7).
 \sys uses DMA block copies for all page swapping operations, leveraging the efficiency of hardware support.
@@
 Since variable \texttt{x} is kept in a different page in \texttt{pagesOrg}, accessing it creates a page fault.
-At this point, \sys swaps the active page in \texttt{pageBuf} by committing it into the non-volatile {\tt pagesTemp} region and copying the corresponding page from \texttt{pagesOrg} to \texttt{pageBuf}.
+At this point, \sys swaps the active page in \texttt{pagesBuf} by committing it into the non-volatile {\tt pagesTemp} region and copying the corresponding page from \texttt{pagesOrg} to \texttt{pageBuf}.
 When the task completes, \sys sets the commit bit, copies the dirty page from \texttt{pagesTmp} to its original location in \texttt{pagesOrg} and unsets the commit bit.

### 2017-11-15 paper/compiler.tex

@@
 
-After completing program cost analysis, \sys inserts task boundaries into the CFG in two steps.
-The first step is to insert task boundaries on the back-edges of all loops that have an energy cost greater than the device's energy buffer capacity, producing an initial boundary placement.
-The second step is to insert task boundaries at additional points in the code, sub-dividing tasks that have an estimated cost that exceeds the energy buffer's capacity.
+After completing program cost analysis, \sys inserts task boundaries into the
+CFG in two steps. The first step is to insert task boundaries on the back-edges
+of all loops that have an energy cost greater than the device's energy buffer
+capacity, producing an initial boundary placement. The second step is to insert
+task boundaries at additional points in the code, sub-dividing tasks that have
+an estimated cost that exceeds the energy buffer's capacity.
 
-To sub-divide these tasks, \sys does depth-first traversals of all paths in the CFG, ignoring back-edges.
-Traversals start from the beginning of the main function, or from an existing task boundary.
-\sys accumulates the cost of basic blocks during the traversal until reaching another task boundary or an exit node in the CFG.
-At the end of a traversal, if the accumulated cost exceeds to device's energy capacity \sys must split the traversed path with a task boundary.
-Each edge on the path is a candidate location for a boundary.
+To sub-divide these tasks, the compiler prototype does depth-first traversals
+of all paths in the CFG, ignoring back-edges. Traversals start from the
+beginning of the main function, or from an existing task boundary. The
+prototype accumulates the cost of basic blocks during the traversal until
+reaching another task boundary or an exit node in the CFG. At the end of a
+traversal, if the accumulated cost exceeds to estimated device energy capacity the
+compiler prototype must split the traversed path with a task boundary. Each
+edge on the path is a candidate location for a boundary.
 
-\sys uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score.
-The heuristic tries to avoid placing boundaries on loops, to minimize the number of live variables that must be protected across the new boundary, and to maximize the number of operations along the path.
-A boundary candidate's score is computed as
+\sys uses a heuristic to compute a score for each candidate and places a
+boundary at the candidate with the highest score. The heuristic tries to avoid
+placing boundaries on loops, tries to minimize the number of live variables 
+protected across the new boundary, and tries to maximize the number of operations
+along the path. A boundary candidate's score is computed as
 %
@@
 %
-where $N_{\text{IR}}$ is the length of the path being traversed up to the candidate, $N_{\text{PV}}$ is the number of protected variables that will be newly introduced by placing boundary at the candidate, and $N_{\text{loop}}$ is the product of the loop bounds of all loops (i.e., nested loops) containing the boundary candidate.
-\sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the values among boundary candidates.
-The weights, $w_{0}, w_{1}$ are between zero and one, and are hand-tuned parameters that define the relative importance of $N_{\text{IR}}$,
-$N_{\text{PV}}$, and $N_{\text{loop}}$.
-We empirically determined these weight values by measuring the runtime for different weight values across applications.
-We found that varying weights varies runtime by only around 5\% in most cases and in our experimental evaluation, we used $w_{0} = w_{1} = 0.5$.
+where $N_{\text{IR}}$ is the length of the path being traversed up to the
+candidate, $N_{\text{PV}}$ is the number of protected variables that will be
+newly introduced by placing boundary at the candidate, and $N_{\text{loop}}$ is
+the product of the loop bounds of all loops (to handle nested loops) containing the
+boundary candidate. \sys normalizes $N_{\text{IR}}$ and $N_{\text{PV}}$ to the
+values among boundary candidates. The weights, $w_{0}, w_{1}$ are between zero
+and one, and are hand-tuned parameters that define the relative importance of
+$N_{\text{IR}}$,
+$N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight
+values by measuring the runtime for different weight values across
+applications. We found that varying weights varies runtime by only around 5\%
+in most cases and in our experimental evaluation, we used $w_{0} = w_{1} =
+0.5$.
 
-After placing task boundaries, the compiler must ensure that all tasks are single-entry.
-To do so, \sys computes the dominator tree of the CFG and checks that every vertex within a task is dominated by the entry vertex of the task.
-If this dominance condition is unmet for a CFG node, the compiler places an additional task boundary at that node, meeting the condition.
-Figure~\ref{fig:compiler_overview} illustrates boundary placement: the first three vertices make up one task and the remainder become a second task.
+After placing task boundaries, the compiler must ensure that all tasks are
+single-entry. To do so, \sys computes the dominator tree of the CFG and checks
+that every vertex within a task is dominated by the entry vertex of the task.
+If this dominance condition is unmet for a CFG node, the compiler places an
+additional task boundary at that node and re-evaluates the condition, repeatedly
+inserting a boundary at each node for which the dominance condition is unmet.
+Figure~\ref{fig:compiler_overview} illustrates boundary placement: the first
+three vertices make up one task and the remainder become a second task.
 
@@
 
-After placing task boundaries during analysis, \sys's transformation analysis restructures the code into tasks defined by the placed boundaries.
-The analysis traverses the CFG starting from an arbitrary task boundary.
-On reaching a subsequent boundary, the transformation analysis identifies all paths between the two task boundaries and copies those paths into a new task function.
-The analysis continues until all paths are contained within a task.
+After placing task boundaries during analysis, \sys's transformation analysis
+restructures the each task into its own function. The analysis
+traverses the CFG starting from an arbitrary task boundary. On reaching a
+subsequent boundary, the transformation analysis identifies all paths between
+the two task boundaries and copies those paths into a new task function. The
+analysis continues until all paths are contained within a task.
 
-After forming tasks, variables that are live beyond the end of the task become protected, which requires renaming those variables and updating references to renamed variable names.
-For example, Figure~\ref{fig:compiler_overview} shows a boundary placed within the live range of {\tt b}; {\tt b} becomes protected and accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in Figure~\ref{fig:compiler_overview}).
-{\tt a} is task-local and need not be protected, but only declared in the task.
-Constants and read-only variable are never protected.
-When writes in multiple, newly formed tasks may update the same location (i.e., at a PHI node in the CFG), \sys re-writes the accesses to refer to the same global name.
-Subsequent reads simply read the protected variable.
+After forming tasks, variables that are live beyond the end of the task become
+protected, which requires renaming those variables and updating references to
+renamed variable names. For example, Figure~\ref{fig:compiler_overview} shows a
+boundary placed within the live range of {\tt b}; {\tt b} becomes protected and
+accesses to {\tt b} should be redirected to {\tt p\_b} (as shown in
+Figure~\ref{fig:compiler_overview}). {\tt a} is task-local and need not be
+protected, but only declared in the task. Constants and read-only variable are
+never protected. When writes in multiple, newly formed tasks may update the
+same location (i.e., at a PHI node in the CFG), \sys re-writes the accesses to
+refer to the same protected, global name. Subsequent reads simply read the
+protected variable.
 
-After forming tasks, some branches in a task may refer to targets in another task.
-\sys patches the CFG to re-write those branches to instead be \transition statements that transition to the entry point of the other task.
-Such a ``broken'' branch will always refer to a task's entry node because \sys ensures that all tasks have only one entry.
+After forming tasks, some branches in a task may refer to targets in another
+task. \sys patches the CFG to re-write those branches to instead be \transition
+statements that transition to the entry point of the other task.  Cross-task
+branches will always refer to a task's entry node because each task boundary
+was placed and guaranteed to be single-entry using the compiler's dominance
+analysis.
 
-The compiler re-writes accesses to protected variables to use \sys's memory virtualization API.
-Pointers pose a challenge to re-writing protected accesses because unprotected pointers may point to protected data.
-Leveraging LLVM's {\tt basic} pointer alias analysis the compiler conservatively re-writes all accesses that {\em may} read or write protected data.
+The compiler re-writes accesses to protected variables to use \sys's memory
+virtualization API. Pointers pose a challenge to re-writing protected accesses
+because unprotected pointers may point to protected data. Leveraging LLVM's
+{\tt basic} pointer alias analysis the compiler conservatively re-writes all
+accesses that {\em may} read or write protected data.
 

### 2017-11-15 paper/compiler.tex
review of compiler section

@@
 Ratchet~\cite{ratchet}).
-%
-Translating C code to use tasks requires identifying program points to use as
-task boundaries, defining a set of task functions that begin and end at those
-boundaries, ensuring that all existing control-flow paths are either contained
-within a task or are made explicit with inter-task control-flow, and identifying
-and instrumenting accesses to task-shared variables that need to be protected
-by \sys at runtime. Our simple prototype, which is implemented using LLVM, analyzes the
-control-flow graph (CFG) and uses a rule-based heuristic to determine at which
-points to place a task boundary. After identifying where to put task
+Translating C code to use tasks requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task or are made explicit with inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
+Our simple prototype, implemented using LLVM, analyzes the control-flow graph (CFG) and uses a rule-based heuristic to determine at which points to place a task boundary.
+After identifying where to put task
 boundaries, the \sys compiler transforms the imperative code into tasks,
@@
 %
-A task is single-entry, multiple-exit part of the CFG that our compiler excises
+A task is single-entry, multiple-exit part of the CFG that \sys compiler excises
 into a new function.  
 %
-There are several factors that determine whether tasks will execute
-efficiently. A task decomposition should minimize the number of variables that
-are accessed by multiple tasks and, consequently, must be protected by \sys's
-paging mechanism. 
+There are several factors determining efficient task execution.
+A task decomposition should minimize the number of variables that are accessed by multiple tasks and, consequently, must be protected by \sys's paging mechanism.
 %
@@
 	\centering
-	\includegraphics[width=.85\columnwidth]{figures/compiler.pdf}
+	\includegraphics[width=\columnwidth]{figures/compiler.pdf}
 	\caption{\sys compiles code to use tasks.}
@@
 shown in Figure~\ref{fig:compiler_overview}) and produces a CFG with some
-vertices marked as {\em task boundaries}. The \sys compiler builds the
-program's CFG as shown in Figure~\ref{fig:compiler_overview}. In our simple
+vertices marked as {\em task boundaries}. In our simple
 reference prototype, we implemented inter-procedural analysis by inlining all
 functions called.  We opted not to support recursion, which is uncommon in
-embedded code. We leave to future work better approaches to interprocedural
+embedded code. We leave to future work better approaches to inter-procedural
 analysis and task transformation.
@@
 
-{\noindent \bf Estimated run time.} \sys assigns each basic block in the CFG a
-{\em cost}, which is an estimate of the run time of the block.  Precisely
-modeling the run time and energy cost of arbitrary C code is difficult and
+\textbf{Estimated Run Time.} \sys assigns each basic block in the CFG a
+{\em cost}, i.e. an estimate of the run time of the block. Precisely
+modelling the run time and energy cost of arbitrary C code is difficult and
 beyond the scope of our simple compiler prototype. Instead, our prototype counts
@@
 of time and consumes more energy.
-%
-To make the cost estimate useful, our compiler prototype estimates the maximum
-cost that the device's energy buffering capacitor supports.  \sys estimates
-this quantity by measuring the number of arithmetic operations that can execute
-without a power interruption, starting from a full capacitor. This maximum number
-of operations is a limit on the cost of a task -- any longer task risks not
-executing to completion.
+To make the cost estimate useful, our compiler prototype estimates the maximum cost that the device's energy buffering capacitor supports.
+\sys does this by measuring the number of arithmetic operations that can execute without a power interruption, starting from a full capacitor.
+This maximum number of operations is a limit on the cost of a task---any longer task risks not executing to completion.
 
-{\noindent \bf Live variable cost.} Each CFG edge is associated with a set of
-variables that {\em may} be {\em alive} across the edge (i.e., variables
-assigned before the edge and may be read after the edge). Variables that live
+\textbf{Live Variable Cost.} Each CFG edge is associated with a set of
+variables that {\em may} be {\em alive} across the edge, i.e., variables
+assigned before the edge and may be read after the edge. Variables that live
 across an edge must be protected by \sys's memory virtualization mechanism when
@@
 
-{\noindent \bf Loop cost.} Loops are a challenge for \sys because the analysis
+\textbf{Loop Cost.} Loops are a challenge for \sys because the analysis
 must account for the cost of multiple iterations of the loop body. \sys assumes
@@
 and one, and are hand-tuned parameters that define the relative importance of
-$N_{\text{IR}}$,
-$N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight
-values by measuring the runtime for different weight values across
+$N_{\text{IR}}$, $N_{\text{PV}}$, and $N_{\text{loop}}$. We empirically determined these weight values by measuring the runtime for different weight values across
 applications. We found that varying weights varies runtime by only around 5\%
@@
 that every vertex within a task is dominated by the entry vertex of the task.
-If this dominance condition is unmet for a CFG node, the compiler places an
+If this condition is unmet for a CFG node, the compiler places an
 additional task boundary at that node and re-evaluates the condition, repeatedly
@@
 After placing task boundaries during analysis, \sys's transformation analysis
-restructures the each task into its own function. The analysis
-traverses the CFG starting from an arbitrary task boundary. On reaching a
-subsequent boundary, the transformation analysis identifies all paths between
-the two task boundaries and copies those paths into a new task function. The
-analysis continues until all paths are contained within a task.
+restructures the each task into its own function.
+The analysis traverses the CFG starting from an arbitrary task boundary.
+On reaching a subsequent boundary, the transformation analysis identifies all paths between the two task boundaries and copies those paths into a new task function.
+The analysis continues until all paths are contained within a task.
 
@@
 never protected. When writes in multiple, newly formed tasks may update the
-same location (i.e., at a PHI node in the CFG), \sys re-writes the accesses to
+same location, i.e., at a PHI node in the CFG, \sys re-writes the accesses to
 refer to the same protected, global name. Subsequent reads simply read the

### 2017-11-15 paper/paging.tex
review of paging section

@@
 
-\paragraph{Dirty victim page.} If a task modified any byte in the victim page (i.e.,  using \texttt{WVAR}), then the page is dirty.
+\textbf{Dirty Victim Page.} If a task modified any byte in the victim page (i.e.,  using \texttt{WVAR}), then the page is dirty.
 Before swapping in the accessed page, \sys must first preserve the updated values in the dirty page so that they can eventually be committed back to their locations in memory when the task ends.
@@
 
-\paragraph{Dirty accessed page.} If an accessed page is modified, \sys should swap it into the page buffer from its location in {\tt pagesTemp}, rather than swapping it in from its original location in memory.
+\textbf{Dirty Accessed Page.} If an accessed page is modified, \sys should swap it into the page buffer from its location in {\tt pagesTemp}, rather than swapping it in from its original location in memory.
 As Algorithm~\ref{algo:pagefault} shows in Line 4, \sys checks if the requested page was previously modified by the executing task.

### 2017-11-15 paper/paging.tex
review of paging section (second pass)

@@
+\sys's paging and buffering scheme is a key enabler of task coalescing, which we described in Section~\ref{sec:task_coalescing}.
+Paging simplifies coalescing because deferring a commit of paged data at a task boundary entails leaving paged data buffered only, not requiring additional state management logic at task boundaries.
+
 \sys introduces a dynamic memory virtualization mechanism that allows tasks to operate primarily through interaction with efficient, volatile SRAM, rather than non-volatile FRAM which, as we showed in Figure~\ref{fig:framEnergy}, is more costly to access than SRAM.
@@
 \sys uses two-phase commit to write buffered, modified pages back to their original memory locations atomically, ensuring data consistency even with power failures.
-\sys's page-based memory model allows moving large blocks of data between non-volatile and volatile memory, which \sys does efficiently using common,
-hardware DMA block-copy operations, unlike prior work~\cite{chain,alpaca} that moves one variable at a time.
-Moreover, \sys's paging and buffering scheme is a key enabler of task coalescing, which we describe in Section~\ref{sec:task_coalescing}.
-Paging simplifies coalescing because deferring a commit of paged data at a task boundary entails leaving paged data buffered only, not requiring additional state management logic at task boundaries.
+\sys's page-based memory model allows moving large blocks of data between non-volatile and volatile memory, which \sys does efficiently using common, hardware DMA block-copy operations, unlike prior work~\cite{chain,alpaca} that moves one variable at a time.
 
@@
 \sys virtualizes the protected, non-volatile data that a task can access through a modified memory interface.
-The interface includes two operations: \texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and \texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory location {\tt var}.
+The interface includes two operations (Table~\ref{tab:viper_syntax}): \texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and \texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory location {\tt var}.
 {\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory and translates it into a \emph{virtual address} in the page buffer in volatile memory.
@@
 If power fails during commit, the commit bit remains set and \sys resumes committing before moving on to execute the next task.
-To commit pages from {\tt pagesTemp}, \sys iterates over the pages' dirty bits and for each page with its dirty bit set, \sys DMA block-copies the page back to its
-original memory location, then clears its dirty bit.
+To commit pages from {\tt pagesTemp}, \sys iterates over the pages' dirty bits and for each page with its dirty bit set, \sys DMA block-copies the page back to its original memory location, then clears its dirty bit.
 
@@
 A power failure erases the contents of the page buffer, preventing a re-executing task from observing updates in the page buffer from a previous execution attempt.
-On a reboot after a power failure,
-\sys clears all dirty bits in {\tt pagesTemp} {\em if the commit bit is unset}.
+On a reboot after a power failure, \sys clears all dirty bits in {\tt pagesTemp} {\em if the commit bit is unset}.
 Clearing these dirty bits ensures that all accesses to protected variables correctly access their original memory locations.

### 2017-11-15 paper/compiler.tex
review of compiler section (second pass)

@@
 arbitrary C into a program with tasks. The compiler is representative of,
-although not a direct implementation of, a system that assists in writing
+although \emph{not a direct implementation} of, a system that assists in writing
 task-based programs by automatically inserting tasks (e.g., inspired by
 Ratchet~\cite{ratchet}).
-Translating C code to use tasks requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task or are made explicit with inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
-Our simple prototype, implemented using LLVM, analyzes the control-flow graph (CFG) and uses a rule-based heuristic to determine at which points to place a task boundary.
-After identifying where to put task
-boundaries, the \sys compiler transforms the imperative code into tasks,
-instrumenting protected memory accesses.
+
+\textbf{Automatic Task Generation Flow.} Transforming C code to use tasks requires identifying program points to use as task boundaries, defining a set of task functions that begin and end at those boundaries, ensuring that all existing control-flow paths are either contained within a task or are made explicit with inter-task control-flow, and identifying and instrumenting accesses to task-shared variables that need to be protected by \sys at runtime.
+Our simple prototype, implemented using LLVM, analyzes the control-flow graph (CFG) and uses a rule-based heuristic to determine where to place a task boundary.
+After identifying task boundaries placement, the \sys compiler transforms the imperative code into tasks, instrumenting protected memory accesses.
 
@@
 \sys analyzes a program to determine where to put task boundaries.
-%
-A task is single-entry, multiple-exit part of the CFG that \sys compiler excises
-into a new function.  
-%
+A task is single-entry, multiple-exit part of the CFG that \sys compiler excises into a new function.
 There are several factors determining efficient task execution.
 A task decomposition should minimize the number of variables that are accessed by multiple tasks and, consequently, must be protected by \sys's paging mechanism.
-%
-Different tasks should be approximately the same size to make reasoning about
-coalescing simple by eliminating the need to consider variable run time costs. 
-%
-To minimize task boundary overheads (e.g., commit), a task decomposition should
-maximize task size, which minimizes the frequency of task transitions.
+Different tasks should be approximately the same size to make reasoning about coalescing simple by eliminating the need to consider variable run time costs.
+To minimize task boundary overheads (e.g., commit), a task decomposition should maximize task size, which minimizes the frequency of task transitions.
 
@@
 The task boundary identification analysis processes a standard C program (as
-shown in Figure~\ref{fig:compiler_overview}) and produces a CFG with some
-vertices marked as {\em task boundaries}. In our simple
-reference prototype, we implemented inter-procedural analysis by inlining all
-functions called.  We opted not to support recursion, which is uncommon in
-embedded code. We leave to future work better approaches to inter-procedural
-analysis and task transformation.
+shown in Figure~\ref{fig:compiler_overview}) and produces a CFG with some vertices marked as {\em task boundaries}.
+In our simple reference prototype, we implemented inter-procedural analysis by inlining all functions called.
+We opted not to support recursion, which is uncommon in embedded code.
+We leave to future work better approaches to inter-procedural analysis and task transformation.
 
-%
 %\sys uses a cost-based task boundary placement analysis that considers per-block energy consumption estimates, inter-task variable liveness, and the cost of partitioning loops.
-%
 
@@
 
-\textbf{Live Variable Cost.} Each CFG edge is associated with a set of
-variables that {\em may} be {\em alive} across the edge, i.e., variables
+\textbf{Live Variable Cost.} Each CFG edge associates with a set of
+variables that {\em may} be {\em alive} across the edge, i.e. variables
 assigned before the edge and may be read after the edge. Variables that live
-across an edge must be protected by \sys's memory virtualization mechanism when
+across an edge must be protected by \sys's memory virtualization when
 \sys places a task boundary on the edge. \sys heuristically minimizes variable

### 2017-11-15 paper/coalescing.tex
coalescing section revised

@@
 Often, energy available at the transiently-powered device is sufficient to run multiple consecutive tasks without page commit at each task boundary.
-In this case task-based execution model can better amortize the run time overhead of committing updated pages when a task ends.
+This way task-based execution model can better amortize the overhead of committing updated pages when a task ends.
 Hence many tasks can be \emph{merged} into a single virtual task---a process we call \emph{coalescing}.
@@
 The caveat of predefined $(C_U,C_D)$ pair is inability to adapt to changing energy availability conditions.
-A \emph{good} coalescing strategy must moderates the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work.
+A \emph{good} coalescing strategy must moderate the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work.
 Therefore we propose to consider two fundamental coalescing strategies for \sys, which we will later assess and compare experimentally:
@@
 \begin{enumerate}
-	\item \textbf{Power Interrupt History-based Coalescing (HDC):} when more tasks are coalesced without power interrupt $C_U$ increases gradually, and $C_D$ increases, as more task are interrupted before committing; and
-	\item \textbf{Weighted Power Interrupt History-based Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
+	\item \textbf{Interrupt History Coalescing (HDC):} when more tasks are coalesced without power interrupt, $C_U$ increases, and $C_D$ increases, as more task are interrupted before committing; and 
+	\item \textbf{Weighted Interrupt History Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
 \end{enumerate}
 
-\subsection{Power Interrupt History-driven Coalescing}
+\subsection{Power Interrupt History Coalescing}
 
@@
 However, tasks coalescing are increased by a half the speed of the successful task completion.
-This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt.
-In the power failure case, if fewer tasks complete, \sys attempts to coalesce fewer tasks.
+This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt \sys does not immediately reduce coalescing factor to one, and with successful page commit of coalescing task, \sys gradually increases coalescing factor.
+In both cases \sys reduces the effect of outliers, i.e.
+short bursts of energy or short bursts of frequent power interrupts.
 
@@
 
-\subsection{Weighted Power Interrupt History-driven Coalescing}
+\subsection{Weighted Power Interrupt History Coalescing}
 
@@
 
-\textbf{Task Size Profiling.} To measure how large the distribution of tasks is, we have used a set of program benchmarks (refer to Section~\ref{sec:results_software} for details and to the beginnin of Section~\ref{sec:methodology_evaluation} for system setup).
-Each benchmark was run for predefined period of time (X\,seconds \todo{Provide exact number}{Amjad}), on non-intermittent power, and at each task evocation, its execution time was measured.
+\textbf{Task Size Profiling.} We have used a set of program benchmarks (refer to Section~\ref{sec:results_software} for details and to the beginning of Section~\ref{sec:methodology_evaluation} for system setup).
+Each benchmark was run for predefined period of time (in case of Figure~\ref{fig:task_profiling} 10\,seconds), on non-intermittent power, and at each task evocation, its execution time was measured.
 If a task was evoked more than once, longer execution time replaced the short one---this way we obtained conservative values of task sizes.
-At the end of task measurement, normalized values of task execution are assigned for each task, the shortest one being one task, the longest one X tasks \todo{Provide exact number}{Amjad}.
+At the end of task measurement, normalized integer values of task execution for task $x$, $W_x$, are collected---the shortest one being one task, the longest one $W_{x,\max}$ (in case of Figure~\ref{fig:task_profiling} $W_{x,\max}=5$).
 Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}.
@@
 
-In the second coalescing strategy an application needs to be profiled, as above, before it is loaded to the intermittently-powered device.
+With the task size information, in the second (weighted) coalescing strategy an application needs to be profiled, as above, before it is loaded to the intermittently-powered device.
+Then the process of coalescing is the same as for the (non-weighted) power interrupt history coalescing, with the exception that $S_t=W_x \times S_t$, i.e.
+number of traversed tasks before commit are a function of their size.
 

### 2017-11-16 paper/intro.tex
intro revised

@@
 
+% <<<<<<< HEAD
+% \begin{figure}
+% 	\includegraphics[width=.5\columnwidth]{figures/alpaca-war.pdf}
+% 	\caption{\label{fig:virtualization}Adaptive task division at run-time might lead to write-after-read dependencies (variable \ttfamily{x}) that do not exist with static task division.\todo{Box with Task 1,2,3 must be the same size as merged task; Merged task must be named as Task 4; x must be marked in text as having WAR; "Task X" name should be next to the task box; "Window" - small letter}{Sinan}}
+% \end{figure}
+% =======
+
 \textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
@@
 tasks affects the time the application takes to complete.
-Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory by the task, to maintain consistency of program state~\cite{chain,alpaca}.
+Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
 Naively minimizing the number of task transitions, (i.e.
-the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its storage capacitor.
-Such a large task would fail to complete unless the environment provides
-sufficient incoming energy to supplement the energy in the capacitor.
+the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its energy storage.
+Such a large task would fail to complete unless the environment provides sufficient incoming energy to supplement the energy in the capacitor.
 To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.
@@
 However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution.
-Furthermore, a static decomposition approach cannot produce decompositions portable across devices with different storage capacitors.
+Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.
 
@@
 
-\textbf{Task Coalescing.} To address this challenge we propose a
-\emph{dynamic} task decomposition approach, that accepts any static
-decomposition and opportunistically coalesces tasks at runtime when
-there is sufficient energy. \emph{Task coalescing} is a mechanism where
-atomic tasks are merged into a single (larger) task while preserving
-atomicity of the merged task. If power fails while a colesced task is
-running, execution will restart at the first task in the group.
-Figure~\ref{fig:coalescing} (bottom) illustrates how dynamic coalescing
-changes the earlier execution and highlights the change in memory model
-necessary to not break atomicity.
+\textbf{Task Coalescing.} To address this challenge we propose a \emph{dynamic} task decomposition, that accepts any static decomposition and opportunistically coalesces tasks at runtime when there is sufficient energy.
+With \emph{task coalescing} atomic tasks are merged into a single (larger) task while preserving atomicity of the merged task.
+If power fails while a coalesced task is running, execution will restart at the first task in the group.
+Figure~\ref{fig:coalescing} (bottom) illustrates how dynamic coalescing changes the earlier execution and highlights the change in memory model necessary to not break atomicity.
 
-Unfortunately, naively merging two atomic tasks might produce a non-atomic
-task, that may leave memory inconsistent after a partial execution. A merge
-breaks atomicity when the accesses to non-volatile memory from the
-previously atomic merged tasks form a \emph{write-after-read (WAR)}
-dependency.
+Unfortunately, naively merging two atomic tasks might produce a non-atomic task, that may leave memory inconsistent after a partial execution.
+A merge breaks atomicity when the accesses to non-volatile memory from the previously atomic merged tasks form a \emph{write-after-read (WAR)} dependency.
 %
@@
 
-Task coalescing removes the optimization burden of choosing the best task decomposition from the programmer, because it accepts any decomposition and improves it dynamically.
+Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically.
 To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks.
@@
 
+% <<<<<<< HEAD
+% We develop \sys%
+% \footnote{To be released at \url{http://anonymized.link}}
+% %
+% : a task coallescing system for task-based
+% programming models on intermittently-powered devices. The
+% capabilities required for effective coallescing mark our
+% contributions:
+% >>>>>>> 96e5d9f9b4e97ce8700c375bd8e11708870f5aba
+% =======
 \textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: a task coalescing system for task-based programming models on intermittently-powered devices.
 The capabilities required for effective coalescing mark our contributions:
+% >>>>>>> 84d7711b6c6f6424cdad97efc8aad96e52f316ff
 

### 2017-11-16 paper/systemdescription.tex
overview: programs accessible

@@
 \sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
-\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable}, {\em efficient} and {\em flexible}.
+\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em accessible}, {\em efficient} and {\em flexible}.
 \sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.

### 2017-11-16 paper/systemdescription.tex
overview: shortenning

@@
 
-\sys consists of its compiler and the run time software support for its execution model and includes several lightweight programming model extensions similar to those in prior work~\cite{chain,alpaca}.
-Table~\ref{tab:viper_syntax} summarizes these extensions, which can be used manually by a programmer or added to a program automatically, using \sys's compiler.
-As in prior task-based intermittent programming models~\cite{chain,alpaca}, a \sys program is decomposed into tasks, which are functions that are not allowed to have any direct callers.
-The programmer identifies a task by adding the {\tt task} keyword to the function's declaration.
-A programmer invokes one task, {\tt t} from another task {\tt u} using {\tt next\_task(\,t\,)}, which atomically commits the state of {\tt u} and transfers control to the entry point of {\tt t}.
-A program has a single {\em origin} task that executes first during the program's first execution; after a power failure, in subsequent operating periods, \sys restarts from the task that was executing just before the failure.
-
-A \sys program manipulates {\em protected} state---accessed by more than one task---using the {\tt RP} and {\tt WP} operations, which read and write (respectively) a protected variable.
-These operations convey to \sys's runtime system which data must be protected through paging by the memory manager.
+\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists of
+constructs for defining tasks~\cite{chain,alpaca} and accessors for
+protected variables shared between tasks. These anotations are added either by
+the programmer (in the manual translation mode) or the compiler (in automatic
+translation mode). As in prior task-based intermittent programming
+models~\cite{chain,alpaca}, tasks are functions that are not allowed to have
+any direct callers but receive control by explicit {\tt next\_task} statements.
+In \sys, a task transition atomically commits the memory operations of the
+current task.  The programmer identifies each task by adding the {\tt task}
+keyword to the function's declaration and designates one task as the {\em
+origin} task at which execution begins. After a power failure,
+\sys restarts from the last (partially) executed task.
+%
+A \sys program manipulates {\em protected} state---accessed by more than one
+task---using the \emph{read variable} ({\tt RVAR}) and \emph{write variable}
+({\tt WVAR}) operations. These operations convey to \sys's memory manager which
+data must be buffered in volatile memory and committed.
 
@@
 		\hline
-		Keyword & Description\\
+		\textbf{Keyword} & \textbf{Description} \\
 		\hline\hline
-		\texttt{task foo(){...}} & Task declaration\\
-		\texttt{origin task foo(){...}} & Origin task declaration\\
-		\texttt{next\_task()} & Transition to a new task\\
-		\texttt{RVAR} & Read-protected variable\\
-		\texttt{WVAR} & Write-protected variable\\
+		\texttt{task foo() \{...\}} & Task definition\\
+		\texttt{origin task foo() \{...\}} & Origin task definition\\
+		\texttt{next\_task(t)} & Transition to task \texttt{t}\\
+		\texttt{u = RVAR(v)} & Read protected variable \texttt{v} into \texttt{u}\\
+		\texttt{WVAR(v, u)} & Write \texttt{u} to protected variable \texttt{v}\\
 		\hline
 	\end{tabular}
-	\caption{\sys's programming model extensions.}
+	\caption{\sys's programming interface.}
 	\label{tab:viper_syntax}
@@
 
-\textbf{\sys Memory Manager.} When the binary executes, \sys's runtime support affects its execution through a {\em (coalescing) task scheduler} (described in Section~\ref{sec:task_coalescing}) and {\em (virtualizing) memory manager} (described in Section~\ref{sec:memory_virtulaization}).
-The \sys binary executes its tasks correctly, with consistent memory state, due to \sys's virtual memory mechanism.
-Relying on programmer or compiler annotation of accesses to memory locations that might be shared between tasks, \sys's memory virtualization {\em pages} these data to SRAM from FRAM as they are needed.
-Tasks operate, accessing SRAM only, demand-paging data to accommodate a SRAM smaller than FRAM.
-When a task completes, \sys commits modified, paged data back to its original FRAM location atomically ensuring subsequent tasks page in correct memory state.
+\textbf{\sys Memory Manager.} At runtime \sys's \emph{virtualizing memory
+manager} (Section~\ref{sec:memory_virtulaization}) protects changes to
+non-volatile data against becoming inconsistent after a power failure.  Relying
+on annotations of accesses to protected variables, \sys's memory manager {\em
+pages} these data into volatile memory from non-volatile memory as they are
+needed.  Tasks cannot modify protected variables directly in non-volatile
+memory.  If the working set exceeds the size of volatile memory, pages are
+swapped out to non-volatile memory without committing the changes, leaving
+original locations of protected variables unchanged.  When a coalesced task
+completes, \sys atomically commits modified pages to their original locations
+in non-volatile memory.
 
-\textbf{\sys Task Scheduler.} \sys runs a lightweight task scheduler that uses {\em task coalescing} to make \sys's statically defined tasks efficiently use available energy.
-By default, tasks run in sequence and each task commits as it completes, potentially incurring unnecessary overhead for commits between consecutive, non-failing tasks.
-\sys dynamically coalesces consecutive tasks by deferring the commit of an earlier task until the completion of a later task.
-\sys's memory management mechanism facilitates coalescing because \sys can defer an earlier task's commit by simply buffering modified, paged data as the later task executes.
-After coalescing a sequence of tasks, \sys commits all tasks' updates to FRAM, together, combining tasks updates to the same pages and minimizing the commit overhead.
+\textbf{\sys Task Scheduler.} The scheduler {\em coalesces}
+statically\hyp{}defined tasks into dynamic tasks sized to match the available
+energy. By default, tasks run in sequence and each task commits as it
+completes, potentially incurring unnecessary overhead for commits between
+consecutive, non-failing tasks. \sys dynamically coalesces consecutive tasks by
+deferring the commit of an earlier task until the completion of a later task.
+\sys's memory management mechanism is essential for coalescing, because commits
+are deferred by keeping dirty pages in volatile memory across coalesced
+boundaries.  After a coalesced sequence of tasks completes, \sys commits pages
+updated by all of the tasks to non-volatile memory, reducing the commit
+overhead through batching.

### 2017-11-16 paper/evaluation.tex
removed 'except sort'
removed problematic header
evaluation section revised up to section 7.2

@@
 
-\subsection{Methodology}
-\label{sec:results_hardware}
-
 We evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp} energy-harvesting
-platform build around a MSP430FR5969~\cite{wolverine} MCU with 64kB of FRAM.
+platform build around a MSP430FR5969~\cite{wolverine} MCU with 64\,kB of FRAM.
 We used a TI Flash Emulation Tool (FET)~\cite{fet} as a power supply during
@@
 3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ
-8\,dBic~\cite{atlas2015}. We ccontrolled the RFID reader from a workstation
+8\,dBic~\cite{atlas2015}. We controlled the RFID reader from a workstation
 running Ubuntu 10.4 using \emph{sllurp}~\cite{sllrp_github}. We set the
@@
 For distance-controlled experiments, we affixed the WISP at varying lengths to
-hollow paper tubes placed flat on a lying antenna. We did timing measurements
-using the Salea~\cite{saleae} logic analyzer.
+hollow paper tubes placed flat on a lying antenna. Timing measurements
+were done using the Salea~\cite{saleae} logic analyzer.
 
 For some experiments (noted below) we used an alternative setup where the
-MSP-TS430RGZ48C launchpad was used for continuous power experiments. We used
+MSP-TS430RGZ48C launchpad was used for continuous power experiments and used
 internal timer for measuring run time. For harvested energy, we used
@@
 reader set at 30\,dBm transmission power at 915\,MHz. We used EDB~\cite{edb}
-to measure run time. Applications were cross-compiled using GCC and Clang,
-allowing us to use LLVM for the \sys compiler. We used version 3.8.0 of both
-Clang and LLVM.
+to measure run time. Applications were cross-compiled using GCC and Clang (both version 3.8.0), allowing us to use LLVM for the \sys compiler.
 
@@
 which encrypts a fixed string with a stored key; (v) \textbf{ar}: which
-performs nearest neighbour classification of randomly generated data modeling a
+performs nearest neighbour classification of randomly generated data modelling a
 three-axis accelerometer. This set of benchmarks overlaps with those used
@@
 We evaluated \sys to show that it is easy to use and efficient across our
-benchmarks.  We evaluate \sys's ease of use by characterizing \sys's automatic
-task compiler.  We evaluate \sys efficiency by comparing its runtime
-performance to Chain.  We characterize \sys by studying how its performance
-varies due to coalescing and when using automatic task compilation versus
-manual task compilation.
+benchmarks.
+%We evaluate \sys's ease of use by characterizing \sys's automatic task compiler.
+We evaluate \sys efficiency by comparing its runtime performance to Chain.
+We characterize \sys by studying how its performance varies due to coalescing and when using automatic task compilation versus manual task compilation.
 
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coalescing}
+	\includegraphics[width=\columnwidth]{figures/coalStrategies}
 	\caption{\sys coalescing performance measured per application on fixed power (left-most bar per application) and at 20 cm and 40 cm from the RFID reader (center and right-most two bars per application) compared to \sys without coalescing.}
@@
 observe that the page size is not the same for each application, although 64
-byte pages perform well for all applications except sort. The increased rate of
+byte pages perform well for all applications. The increased rate of
 page faults is responsible for the overhead with small page sizes; smaller
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/pagSizeOverhead}
+	\includegraphics[width=\columnwidth]{figures/ramPagsSizes}
 	\includegraphics[width=\columnwidth]{figures/pagefault}

### 2017-11-16 paper/coalescing.tex
coalescing: edits to first half

@@
-Often, energy available at the transiently-powered device is sufficient to run multiple consecutive tasks without page commit at each task boundary.
-This way task-based execution model can better amortize the overhead of committing updated pages when a task ends.
-Hence many tasks can be \emph{merged} into a single virtual task---a process we call \emph{coalescing}.
+When incoming energy is sufficient to run multiple consecutive tasks,
+committing state after each task is wasteful. \sys mitigates this
+waste by deferring the expensive commits and performing them in a
+batch. In the resulting execution, several tasks are \emph{merged} into a
+single virtual task---a process we call \emph{coalescing}. 
+%
+When $N$ tasks are coalesced, $N-1$ boundaries are crossed and some number of
+pages are dirtied. The first $N-2$ executions of \transition statement skip
+committing dirty pages to non-volatile memory, leaving the work of committing
+pages modified by any of the $N$ tasks to the statement $N-1$.
+Coalescing is likely to reduce the total number of non-volatile memory
+accesses, because when coalesced tasks access the same pages of data, only the
+last version of the page needs to be committed.
 
-\textbf{Process of Task Coalescing.} \sys, as the first task-based execution model, dynamically coalesces consecutively executing tasks at run time.
-When $X$ tasks are coalesced, the $X-1$ tasks do not commit their updates to protected memory locations when reaching a \transition statement.
-Instead, execution immediately continues from the beginning of the $X$th task.
-At the end of the $X$th coalesced task, \sys commits the pages that were updated by any of the $X$ coalescing tasks.
-If consecutive tasks access the same pages of data, the pages accessed by both tasks only need to be committed once.
-\sys repeatedly executes $X-1$ coalesced tasks, followed by a committing task.
+\textbf{Trade-off between Speed and Wasted Effort.} \sys can coalesce an
+arbitrary number of consecutive tasks. However, as more tasks coalesce, their
+collective commit overhead amortizes better, but the risk of wasting work also
+increases. If power fails during a long sequence of coalesced tasks, execution
+will restart from the last commit, i.e. the first task in the sequence, losing
+the progress made by any of the coalesced tasks.  The challenge to coalescing
+tasks is determining how many tasks to coalesce before committing.
+%
+%During execution, an executing task is either a {\em coalesced} task, which
+%does not commit its state before transitioning or a {\em committing} task,
+%which does commit its state before transitioning.
 
-\textbf{Coalescing Speed/Wasted Effort Trade-off.} \sys can coalesce an arbitrary number of consecutive tasks.
-As more tasks coalesce, their collective commit overhead amortizes better, potentially improving performance.
-However, as more tasks coalesce, the amount of potentially {\em wasted work} also increases.
-If power fails during a long sequence of coalesced tasks the lack of an intervening commit requires execution to restart from the {\em first} task in the sequence, failing to preserve the progress of any of the tasks in this long sequence.
-The challenge to coalescing tasks is determining how many tasks to coalesce before committing.
-During execution, an executing task is either a {\em coalesced} task, which does not commit its state before transitioning or a {\em committing} task, which does commit its state before transitioning.
-
-\textbf{Task Coalescing Strategies.} The simplest strategy will coalesce/de-coalesce fixed number of task at every successful/unsuccessful commit operation, where task coalescing factor increases by $C_U$ at each successful virtual task execution and decreases by $C_D$, otherwise.
-There are naturally infinite possible strategies for coalescing, i.e.
-infinite amount of $(C_U,C_D)$ pairs.
-The caveat of predefined $(C_U,C_D)$ pair is inability to adapt to changing energy availability conditions.
-A \emph{good} coalescing strategy must moderate the risk of coalescing and capitalize on its benefit by dynamically, \emph{adaptively} determining a coalescing factors $(C_U,C_D)$ that combines many tasks without suffering an excess of wasted work.
-Therefore we propose to consider two fundamental coalescing strategies for \sys, which we will later assess and compare experimentally:
+\textbf{Task Coalescing Strategies.} 
+A good coalescing strategy must moderate the risk of wasted work and capitalize
+on benefits of deferred commits by adapting the target size of the dynamic task
+to the energy available at runtime.
+%
+The general structure of our strategies is to increase the target task size by
+some positive delta $C_\text{up}$ after the dynamic coalesced task completes,
+and decrease it by some negative delta $C_\text{down}$ upon a reboot after a
+power failure.
+%
+The choice of $C_\text{up}$ and $C_\text{down}$, including their units, defines
+the strategy and its performance.
+%
+For example, a naive strategy that fixes $C_\text{up}$ and $C_\text{down}$ to
+``one static task``, will be very slow to react to changes in incoming energy,
+since its dynamic task size can only change by the size of one (small) static
+task per charge-discharge cycle.
+%
+Working from these observations, we propose and evaluate two coalescing
+strategies for \sys:
 

### 2017-11-16 paper/coalescing.tex
coalescing: partial edits to strategies

@@
 \textbf{Task Coalescing Strategies.}
+
+The likelihood that a task group will complete depends on the total amount of
+energy consumed by all tasks in the group, i.e. the size of the dynamic task.
+%
 A good coalescing strategy must moderate the risk of wasted work and capitalize
@@
 Working from these observations, we propose and evaluate two coalescing
-strategies for \sys:
+strategies for \sys: \textbf{History-aware Coalescing (HC)} and
+\textbf{Task-size- and History-aware Coalescing (TSHC)}.
 
-\begin{enumerate}
-	\item \textbf{Interrupt History Coalescing (HDC):} when more tasks are coalesced without power interrupt, $C_U$ increases, and $C_D$ increases, as more task are interrupted before committing; and 
-	\item \textbf{Weighted Interrupt History Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
-\end{enumerate}
+%\begin{enumerate}
+%	\item \textbf{Interrupt History Coalescing (HDC):} when more tasks are coalesced without power interrupt, $C_U$ increases, and $C_D$ increases, as more task are interrupted before committing; and 
+%	\item \textbf{Weighted Interrupt History Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
+%\end{enumerate}
 
-\subsection{Power Interrupt History Coalescing}
+\subsection{HC: History-aware Coalescing}
 
+\todo{clarify}{AC}
 With this strategy \sys calculates only one parameter $C=C_U=C_D$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period.
@@
 
-\subsection{Weighted Power Interrupt History Coalescing}
+\subsection{TSHC: Task-size- and History-aware Coalescing}
 
-The previous coalescing strategy is oblivious to the task size.
-Although many tasks can be coalesced, their non-equal size will determine that some coalesced tasks will be more susceptible to power interrupts than others.
-To understand that effect we provide the following illustration.
+The history-aware coalescing policy used the number of tasks in the group as a
+rough proxy for the energy cost of the task group.
+%
+With some up-front investment of developer effort and modest time and space
+overhead, a more precise proxy for task size can be obtained by profiling the
+tasks at design time.
+%
+The TSHC policy approximates task energy cost from a profile of task execution
+times.
+%
+Profiling execution time instead of direct energy consumption is
+advantageous, because the former does not require specialized
+current\hyp{}measurement hardware and can be readily done on any
+off-the-shelf microcontroller.
+%
+To use the TSHC policy, the application is first profiled, as will be described
+in the experiment below, and the task execution time for each task is written
+into a read-only table in non-volatile memory.
+%
+At runtime, the decision to coalesce or commit at each task boundary proceeds
+similarly with the HC policy, but the task completion counter, $S_t$ has units
+of ``task execution time''.\todo{be precise}{AC}
 
-\textbf{Task Size Profiling.} We have used a set of program benchmarks (refer to Section~\ref{sec:results_software} for details and to the beginning of Section~\ref{sec:methodology_evaluation} for system setup).
-Each benchmark was run for predefined period of time (in case of Figure~\ref{fig:task_profiling} 10\,seconds), on non-intermittent power, and at each task evocation, its execution time was measured.
-If a task was evoked more than once, longer execution time replaced the short one---this way we obtained conservative values of task sizes.
-At the end of task measurement, normalized integer values of task execution for task $x$, $W_x$, are collected---the shortest one being one task, the longest one $W_{x,\max}$ (in case of Figure~\ref{fig:task_profiling} $W_{x,\max}=5$).
-Result of the task size measurement is provided in Figure~\ref{fig:task_profiling}.
-We observe huge variations of task sizes---both within a benchmark, as well per benchmark, which calls for the use of this information in coalescing.
+To quantify the importance of task size in the coalescing decision, we measured
+the distribution of task sizes in our benchmarks (described in
+Sections~\ref{sec:results_software}).
+%
+Each benchmark was run for 10 seconds on non-intermittent power, and a timestamp
+in clock cycles was taken at each task boundary. For each execution of
+a task, its size was estimated as the maximum difference in timestamps at its
+boundaries.
+%
+The CDF for the distribution of task sizes in Figure~\ref{fig:task_profiling}
+shows the wide range of task sizes across and within benchmarks, and explains
+the advantage of the TSHC policy.
+
+%At the end of
+%task measurement, normalized integer values of task execution for task $x$, $W_x$, are
+%collected---the shortest one being one task, the longest one $W_{x,\max}$ (in
+%case of Figure~\ref{fig:task_profiling} $W_{x,\max}=5$). Result of the task
+%size measurement is provided in Figure~\ref{fig:task_profiling}.
 
@@
 	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}%coalescing_timeline
-	\caption{Result of benchmark applications (Section~\ref{sec:results_software}) task size measurement, measured in the number of instruction cycles.
-We observe very large variations of task sizes per each benchmark.
+	\caption{\textbf{Distribution of task sizes, estimated by task execution times.} The
+wide variation in task size motivates accounting for task size in the coalescing strategy.
 \todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
@@
 \end{figure}
-
-With the task size information, in the second (weighted) coalescing strategy an application needs to be profiled, as above, before it is loaded to the intermittently-powered device.
-Then the process of coalescing is the same as for the (non-weighted) power interrupt history coalescing, with the exception that $S_t=W_x \times S_t$, i.e.
-number of traversed tasks before commit are a function of their size.
 

### 2017-11-16 paper/evaluation.tex
evaluation section revised (up to 7.2.1)

@@
 	\includegraphics[width=\columnwidth]{figures/coala_chain_clang}
-	\caption{Performance of \sys applications (divided manually into tasks) at fixed power (left-most bar per application) and at multiple WISP to RFID reader antenna distances, \{20, 40\}\,cm (center and right-most bar per application) compared against Chain (results are normalized).}
+	\caption{\textbf{Performance of \sys applications (divided manually into tasks) at multiple WISP to RFID reader antenna distances}: fixed power (0\,m to the antenna) and \{0.2, 0.4\}\,m (left, center and right-most bar per application, respectively), compared against Chain (results are normalized).}
 	\label{fig:runtime}
@@
 
-We evaluate \sys's run time performance and compare the results to the run time
-performance of Chain.  We show the results in Figure~\ref{fig:runtime}. For
-each application we measured its execution time on continuous power (the first
-two bars of each group) and on intermittent power at 20cm and 40 cm from the
-RFID antenna (the second two and third two bars of each group, respectively).
+\textbf{\sys versus Task-based Model.} We evaluate \sys's run time performance and compare \sys to the run time performance of Chain.
+Results are given in Figure~\ref{fig:runtime}.
+For each application we measured its execution time on continuous power, i.e.
+0\,m RFID antenna/WISP distance (the first two bars of each group) and on intermittent power at 20\,cm and 40\,cm from the RFID antenna (the second two and third two bars of each group, respectively).
 The data show the average execution time of each application run repeatedly for 20 seconds, with runtimes normalized to \sys's average.
-The results show that \sys provides 
-a performance benefit compared to Chain for most applications.  The performance benefit
-is greatest for applications that have small tasks (bc,
-sort, cuckoo) at all distances on intermittent power and with fixed power. In applications that  
-have larger tasks (e.g. dft) or access more different pages of memory, \sys incurs
-overhead from memory virtualization that cause its performance to be comparable to
-(or in the single case of cem, worse than) Chain.
+The results show that \sys provides a performance benefit compared to Chain for most applications.
+The performance benefit is greatest for applications that have small tasks (bc, sort, cuckoo) at all distances on intermittent power and with fixed power.
+In applications that have larger tasks (e.g.
+dft) or access more different memory pages, \sys incurs overhead from memory virtualization that cause its performance to be comparable to (or in the \emph{single} cem case, worse than) Chain.
 
-We evaluated the impact that \sys's dynamic task coalescing had on run time
-performance.  Figure~\ref{fig:coalescing} shows the run time performance of
-\sys's coalescing, compared to \sys running without coalescing enabled with
-data normalized to the run time performance of \sys with coalescing.  We show
-data from experiments with continuous power (the first two bars in each bar
-group), and with intermittent power, with the WISP at 20 cm and 40cm from the
+\textbf{Coalescing Strategies.} We evaluated the impact that \sys's dynamic task coalescing had on run time performance.
+Figure~\ref{fig:coalescing} shows the run time performance of \sys's coalescing, compared to \sys running without coalescing enabled with data normalized to the run time performance of \sys with coalescing.
+We show data from experiments with continuous power (the first two bars in each bar
+group), and with intermittent power, with the WISP at 20\,cm and 40\,cm from the
 RFID reader (the second and third groups of two bars in each group).
-%
-The data illustrate that coalescing improves \sys's performance by eliminating
-the overhead of committing after each task.  The performance benefit varies,
-providing up to 4.5x performance improvement (for cuckoo filtering, at 40cm).
-The cuckoo application (like other top-performers, bc and sort) has small tasks
-that \sys can easily coalesce to eliminate many commits. In contrast, for some
-applications, e.g. cem, coalescing does not improve performance. The reason for
-cem's poor performance is that cem uses a large LZW dictionary structure that
-spans multiple pages, but has very little {\em locality}.  The absence of
-locality means that coalesced tasks do not access the same pages, and as a
-result, do not amortize commit overhead. The lack of locality causes
-performance to regress to be similar to the non-coalescing case.
+The data illustrate that coalescing improves \sys's performance by eliminating the overhead of committing after each task.
+The performance benefit varies, providing up to 4.5x performance improvement (for cuckoo filtering, at 40cm).
+The cuckoo application (like other top-performers, bc and sort) has small tasks that \sys can easily coalesce to eliminate many commits.
+In contrast, for some applications, e.g.
+cem, coalescing does not improve performance.
+The reason for cem's poor performance is that cem uses a large LZW dictionary structure that spans multiple pages, but has very little {\em locality}.
+The absence of locality means that coalesced tasks do not access the same pages, and as a result, do not amortize commit overhead.
+The lack of locality causes performance to regress to be similar to the non-coalescing case.
+
+%Uniformly distributed random pattern of power interrupts with an average on and off time of 15\,ms and 10\,ms respectively.
 
@@
 	\includegraphics[width=\columnwidth]{figures/coalStrategies}
-	\caption{\sys coalescing performance measured per application on fixed power (left-most bar per application) and at 20 cm and 40 cm from the RFID reader (center and right-most two bars per application) compared to \sys without coalescing.}
+	\caption{\textbf{\sys coalescing strategies (`WHD' and `HD') performance per application on intermittent power compared to \sys without coalescing (`NO').} Coalescing provides benefit for all applications, while we observe little difference between `WHD' and `HD'.}
 	\label{fig:coalescing}
@@
 
-\subsubsection{Evaluating \sys's Compiler}
+\subsubsection{\sys's Compiler}
 \label{sec:result_compiler_time}
-\label{sec:results_compiler}
 
-We characterized the behavior of, and evaluated the efficiency of the \sys
-compiler.  Table~\ref{table:compiler_result} shows, for each application, for
-manually written benchmarks (Man.) and automatically compiled benchmarks
-(Comp.) the difference in the number of bytes of protected data (columns 2 and
-3), the number of tasks (columns 4 and 5), the number of \sys protected data
-API access calls (columns 6 and 7), the lines of code in each application for
-\sys and Chain (columns 8 and 9), and compilation time using \sys's compiler
+We characterized the behavior and the efficiency of the \sys compiler.
+Table~\ref{table:compiler_result} shows, for each application, for manually written (Man.) and automatically compiled (Comp.) benchmarks the difference in the number of bytes of protected data (columns 2 and 3), the number of tasks (columns 4 and 5), the number of \sys protected data API access calls (columns 6 and 7), the lines of code in each application for \sys and Chain (columns 8 and 9), and compilation time using \sys's compiler
 (column 10).

### 2017-11-16 paper/evaluation.tex
evaluation section revised (compiler section)

@@
 
-\subsubsection{\sys's Compiler}
-\label{sec:result_compiler_time}
-
-We characterized the behavior and the efficiency of the \sys compiler.
-Table~\ref{table:compiler_result} shows, for each application, for manually written (Man.) and automatically compiled (Comp.) benchmarks the difference in the number of bytes of protected data (columns 2 and 3), the number of tasks (columns 4 and 5), the number of \sys protected data API access calls (columns 6 and 7), the lines of code in each application for \sys and Chain (columns 8 and 9), and compilation time using \sys's compiler
-(column 10).
-
-\begin{table}[t]
-	\centering
-	\footnotesize
-        \renewcommand{\tabcolsep}{1pt}
-	\begin{tabular}{|l|cc|cc|cc|cc|c|}
-		\hline
-		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
-		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
-		\hline\hline
-		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
-		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
-		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
-		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
-		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
-		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
-		%dd &  &  &  &  &  &  &  & 287 &  \\
-		\hline
-	\end{tabular}
-	\caption{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns).
-$^\dagger$dft does not compile with Clang}
-\label{table:compiler_result}
-\end{table}
-
-On average, compiler-generated code introduces about 0.3\% more bytes of
-protected data, suggesting that liveness analysis captures natural boundaries
-in the use of variables also caught by a manual implementation.
-Compiler-generated code generates about 16\% fewer tasks for applications, on
-average.  The results indicate that the \sys compiler selects task boundaries
-differently than a programmer would manually, favoring fewer, larger tasks.
-The \sys compiler inserts protected data access operations conservatively when
-a variable {\em may} alias a protected variables, due to the limitations of
-pointer aliasing. Consequently, the \sys compiler introduces 48\% more
-protected data accesses than programmer-written code. The run time results in
-Section~\ref{sec:result_compiler_time} show that the increase in protected data
-accesses sometimes degrades performance. Overall, the data show that the
-compiler automatically generates tasks similarly to manually written tasks.
-Doing so, \sys's compiler reaps the static task benefits of \sys, but
-simplifies programming by sparing the programmer the job of task decomposition.
-
-We measured compile time for \sys's compiler's automatic task generation for
-each application\footnote{We could not compile the dft application since LLVM
-inserts mathematical built-in functions (i.e. \texttt{\_\_muldf3},
-\texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not
-provided by the LLVM MSP430 back-end.}. We compiled the applications on a
-4-core virtual machine with 1\,GB of memory running Arch Linux on an Intel Core
-i5-6200U at 2.30\,GHz. We collected results using the Linux system time
-collected in seconds. Most of the compile time was spent computing pointer
-alias analysis, while calculating the possible live ranges of variables.
-Applications such as RSA that have a larger code size and that manipulate more
-variables have a longer compilation time, due to a larger search space for
-pointer aliasing. Nonetheless, the compile time is orders of magnitude shorter
-than manually defining tasks, which requires hours of programmer time. The
-characterization of \sys's compiler suggests that the \sys compiler enables
-simpler task-based intermittent programming for \sys.
-
-In addition to the characterization of the \sys compiler in Table~\ref{table:compiler_result},  
-we experimentally assessed the performance of the \sys compiler, compared to the performance
-of the manually written benchmarks, as written in prior work~\cite{alpaca,chain}. 
-%
-Figure~\ref{fig:comp_user} compares the performance of manually written
-benchmark code to the performance of code automatically compiled using \sys.
-The figure shows run time for both cases, normalized to the run time of the
-manually written program, running on continuous power.
-%
-The data show that in all cases except for ar and rsa, the run time performance
-of compiled code generated by the \sys compiler has comparable performance to
-manually written code.  With comparable performance in most cases, the data
-suggest that the \sys compiler not only makes programming simpler, but does so,
-in most cases, without a high run time overhead.  In the cases of both ar and
-rsa, the code manipulates a large amount of data that is stored in an array and
-passed into functions by reference.   We expect that the data being manipulated
-by reference is leading to a large amount of may-alias references in the
-compiler's analysis, which leads to a larger number of accesses to protected
-data in the automatically compiled code, compared to the manually written code.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/comp_user}
-	\caption{\sys compiler code overhead measured in terms of execution time per application (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
-	\label{fig:comp_user}
-\end{figure}
-
 %\begin{table}
@@
 
+\subsubsection{\sys's Compiler}
+\label{sec:result_compiler_time}
+
+\begin{table}[t]
+	\centering
+	\footnotesize
+	\renewcommand{\tabcolsep}{1pt}
+	\begin{tabular}{|l|cc|cc|cc|cc|c|}
+		\hline
+		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
+		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
+		\hline\hline
+		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
+		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
+		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
+		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
+		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
+		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
+		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
+		%dd &  &  &  &  &  &  &  & 287 &  \\
+		\hline
+	\end{tabular}
+	\caption{\textbf{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns)}; $^\dagger$dft does not compile with Clang}\label{table:compiler_result}
+\end{table}
+
+\textbf{\sys Compiler Generated Code Characterization.} The behaviour and the efficiency metrics of the \sys compiler are given in Table~\ref{table:compiler_result}. 
+%
+%for each application, for manually written (Man.) and automatically compiled (Comp.) benchmarks the difference in the number of bytes of protected data (columns 2 and 3), the number of tasks (columns 4 and 5), the number of \sys protected data API access calls (columns 6 and 7), the lines of code in each application for \sys and Chain (columns 8 and 9), and compilation time using \sys's compiler (column 10).
+%
+On average, compiler-generated code introduces about 0.3\% more bytes of protected data, suggesting that liveness analysis captures natural boundaries in the use of variables also caught by a manual implementation.
+Compiler-generated code generates about 16\% fewer tasks for applications, on average.
+The results indicate that the \sys compiler selects task boundaries differently than a programmer would manually, favoring fewer, larger tasks.
+The \sys compiler inserts protected data access operations conservatively when a variable {\em may} alias a protected variables, due to the limitations of pointer aliasing.
+Consequently, the \sys compiler introduces 48\% more protected data accesses than programmer-written code.
+The run time results in Section~\ref{sec:result_compiler_time} show that the increase in protected data
+accesses sometimes degrades performance.
+Overall, the data show that the compiler automatically generates tasks similarly to manually written tasks.
+Doing so, \sys's compiler reaps the static task benefits of \sys, but simplifies programming by sparing the programmer the job of task decomposition.
+
+\textbf{Compile Time.} We measured compile time for \sys's compiler's automatic task generation for each application\footnote{We could not compile the dft application since LLVM inserts mathematical built-in functions (i.e.
+\texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}.
+We compiled the applications on a 4-core virtual machine with 1\,GB of memory running Arch Linux on an Intel Core i5-6200U at 2.30\,GHz.
+We collected results using the Linux system time collected in seconds.
+Most of the compile time was spent computing pointer alias analysis, while calculating the possible live ranges of variables.
+Applications such as rsa that have a larger code size and manipulate more variables have a longer compilation time, due to a larger search space for pointer aliasing.
+Nonetheless, the compile time is orders of magnitude shorter than manually defining tasks, which requires hours of programmer time.
+%The characterization of \sys's compiler suggests that the \sys compiler enables simpler task-based intermittent programming for \sys.
+
+\textbf{Run Time of \sys Compiler Code.} In addition to the characterization of the \sys compiler in Table~\ref{table:compiler_result}, we experimentally assessed the performance of the \sys compiler, compared to the performance of the manually written benchmarks, as written in prior work~\cite{alpaca,chain}.
+Figure~\ref{fig:comp_user} compares the performance of manually written benchmark code to the performance of code automatically compiled using \sys.
+The figure shows run time for both cases, normalized to the run time of the manually written program, running on continuous power.
+The data show that in all cases except for ar and rsa, the run time performance of compiled code generated by the \sys compiler has comparable performance to manually written code.
+With comparable performance in most cases, the data suggest that the \sys compiler not only makes programming simpler, but does so, in most cases, without a high run time overhead.
+In the cases of both ar and rsa, the code manipulates a large amount of data that is stored in an array and passed into functions by reference.
+We expect that the data being manipulated by reference is leading to a large amount of may-alias references in the compiler's analysis, which leads to a larger number of accesses to protected data in the automatically compiled code, compared to the manually written code.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/comp_user}
+	\caption{\textbf{\sys compiler code overhead measured in terms of execution time per application} (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
+	\label{fig:comp_user}
+\end{figure}
+
 %\subsubsection{Case Study}

### 2017-11-16 paper/evaluation.tex
evaluation section revised (paging section)

@@
 
-We evaluated the effect that paging with different page sizes has on \sys's
-performance and we show data in Figure~\ref{fig:page_size}. We show two plots:
-the top plot shows the run time performance normalized to the best
-per-application performance, using pages of different sizes in \sys, indicated
-with the labels on top of each cem bar; the bottom plot shows the number of
-page faults for the same set of executions normalized to the maximum of each
-set.
-%
-We measured performance as MCU clock cycles using values read out by TI's CCS
-IDE. 
-%
-The results show a clear ``bathtub curve'' in the performance of each
-application, as the number of page faults varies. The data show that across
-application there is a page size that minimizes \emph{both} execution time.  We
-observe that the page size is not the same for each application, although 64
-byte pages perform well for all applications. The increased rate of
-page faults is responsible for the overhead with small page sizes; smaller
-pages require accessing more different pages, leading to higher paging costs.
+We evaluated the effect that paging with different page sizes has on \sys's performance and we show data in Figure~\ref{fig:page_size}; The top plot shows the run time performance normalized to the best per-application performance, using pages of different sizes in \sys, indicated with the labels at each bar; the bottom plot shows the number of page faults for the same set of executions normalized to the maximum of each set.
+We measured performance as MCU clock cycles using values read out by TI's CCS IDE.
+The results show a clear ``bathtub curve'' in the performance of each application, as the number of page faults varies.
+The data show that across application there is a page size that minimizes \emph{both} execution time.
+We observe that the page size is not the same for each application, although 64 byte pages perform well for all applications.
+The increased rate of page faults is responsible for the overhead with small page sizes; smaller pages require accessing more different pages, leading to higher paging costs.
 With larger page sizes, the overhead is higher than with moderate pages, too.
-The explanation for these higher overheads is that large pages have a higher
-commit cost. Even if an application accesses few memory locations, \sys pages
-data at full page size, degrading performance by sometimes paging in and out
-unnecessary data.  The data reveal a reasonable default page size of 64 bytes
-and when developing an application, a developer should consider varying the
-page size to moderate poor performance.
+The explanation for these higher overheads is that large pages have a higher commit cost.
+Even if an application accesses few memory locations, \sys pages data at full page size, degrading performance by sometimes paging in and out unnecessary data.
+The data reveal a reasonable default page size of 64 bytes and when developing an application, a developer should consider varying the page size to moderate poor performance.
 
@@
 	\includegraphics[width=\columnwidth]{figures/pagefault}
-	\caption{\sys's page size affects overhead and page fault count.}
-	\label{fig:IPOSPerformance}
+	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X: no page faults.}
 	\label{fig:page_size}

### 2017-11-17 paper/evaluation.tex
evaluation section revised completely (questions need to be answered by Amjad)

@@
 
+\textbf{Coalescing Strategies.} Figure~\ref{fig:coalescing} shows the run time performance of both \sys's coalescing, compared to \sys running without coalescing enabled with data normalized to the run time performance of \sys with coalescing\todo{which one: WHD, or HD?}{Amjad}.
+We show data from experiments with intermittent power, where intermittency was generated in software \todo{Say briefly how we generate full reset with MSP430}{Amjad} by creating uniformly distributed random pattern of power interrupts with an average power availability and unavailability time of 15\,ms and 10\,ms respectively.
+Each application ran with the same generated pattern for the fairness of comparison.
+
+While we see the benefit from coalescing for all applications (except for dft \todo{why coalescing for dft does not work but in Fig.
+9 does?}{Amjad}), the runtime difference between WHD and HD strategies is actually small \todo{say why this difference is small}{Amjad}.
+For this reason we shall proceed with HD coalescing strategy in subsequent evaluations.
+
 \begin{figure}
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coala_chain_clang}
-	\caption{\textbf{Performance of \sys applications (divided manually into tasks) at multiple WISP to RFID reader antenna distances}: fixed power (0\,m to the antenna) and \{0.2, 0.4\}\,m (left, center and right-most bar per application, respectively), compared against Chain (results are normalized).}
-	\label{fig:runtime}
+	\includegraphics[width=\columnwidth]{figures/coalStrategies}
+	\caption{\textbf{\sys coalescing strategies (`WHD' and `HD') performance per application on intermittent power compared to \sys without coalescing (`NO').} Coalescing provides benefit for all applications, while we observe little difference between `WHD' and `HD'.}
+	\label{fig:coalescing}
 \end{figure}
@@
 Results are given in Figure~\ref{fig:runtime}.
-For each application we measured its execution time on continuous power, i.e.
+For each application we measured its execution time using real wireless power provision with RFID reader for continuous power, i.e.
 0\,m RFID antenna/WISP distance (the first two bars of each group) and on intermittent power at 20\,cm and 40\,cm from the RFID antenna (the second two and third two bars of each group, respectively).
@@
 
-\textbf{Coalescing Strategies.} We evaluated the impact that \sys's dynamic task coalescing had on run time performance.
-Figure~\ref{fig:coalescing} shows the run time performance of \sys's coalescing, compared to \sys running without coalescing enabled with data normalized to the run time performance of \sys with coalescing.
-We show data from experiments with continuous power (the first two bars in each bar
-group), and with intermittent power, with the WISP at 20\,cm and 40\,cm from the
-RFID reader (the second and third groups of two bars in each group).
-The data illustrate that coalescing improves \sys's performance by eliminating the overhead of committing after each task.
-The performance benefit varies, providing up to 4.5x performance improvement (for cuckoo filtering, at 40cm).
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coala_chain_clang}
+	\caption{\textbf{Performance of \sys applications (divided manually into tasks) at multiple WISP to RFID reader antenna distances}: fixed power (0\,m to the antenna) and \{0.2, 0.4\}\,m (left, center and right-most bar per application, respectively), compared against Chain (results normalized).}
+	\label{fig:runtime}
+\end{figure}
+
+The data also illustrate that coalescing improves \sys's performance by eliminating the overhead of committing after each task.
+The performance benefit varies, providing up to 4.5x performance improvement.
 The cuckoo application (like other top-performers, bc and sort) has small tasks that \sys can easily coalesce to eliminate many commits.
@@
 The lack of locality causes performance to regress to be similar to the non-coalescing case.
-
-%Uniformly distributed random pattern of power interrupts with an average on and off time of 15\,ms and 10\,ms respectively.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/coalStrategies}
-	\caption{\textbf{\sys coalescing strategies (`WHD' and `HD') performance per application on intermittent power compared to \sys without coalescing (`NO').} Coalescing provides benefit for all applications, while we observe little difference between `WHD' and `HD'.}
-	\label{fig:coalescing}
-\end{figure}
 

### 2017-11-17 paper/intro.tex
revisions abs and intro

@@
 If power fails while a coalesced task is running, execution will restart at the first task in the group.
-Figure~\ref{fig:coalescing} (bottom) illustrates how dynamic coalescing changes the earlier execution and highlights the change in memory model necessary to not break atomicity.
+Figure~\ref{fig:coalesce} (bottom) illustrates how dynamic coalescing changes the earlier execution and highlights the change in memory model necessary to not break atomicity.
 
@@
 %
-\todo{Provide concrete numeric results supporting this claim}{Amjad/Przemek}
+%\todo{Provide concrete numeric results supporting this claim}{Amjad/Przemek}
 

### 2017-11-17 paper/coalescing.tex
coalescing: whitespace fix
coalescing and evaluation revision

@@
 \textbf{Task Coalescing Strategies.} 
-
+%
 The likelihood that a task group will complete depends on the total amount of
@@
 \todo{clarify}{AC}
-With this strategy \sys calculates only one parameter $C=C_U=C_D$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period.
+With this strategy \sys calculates only one parameter $C=C_\text{up}=C_\text{down}$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period.
 \sys begins execution in a coalesced task with $S_t=0$.
@@
 	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}%coalescing_timeline
-	\caption{\textbf{Distribution of task sizes, estimated by task execution times.} The
-wide variation in task size motivates accounting for task size in the coalescing strategy.
-\todo{Describe X axis, no capitals in titles - except first word; markers in addition to colors per app; remove 4.7}{Amjad}}
+	\caption{\textbf{Distribution of task sizes, estimated by task execution times.} The wide variation in task size motivates accounting for task size in the coalescing strategy.}
 	\label{fig:task_profiling}

### 2017-11-17 paper/paging.tex
paging: more details and specifics, incl. algo for commit

--- 2017-11-15 paper/paging.tex
@@
-\sys introduces a dynamic memory virtualization mechanism that allows tasks to operate primarily through interaction with efficient, volatile SRAM, rather than non-volatile FRAM which, as we showed in Figure~\ref{fig:framEnergy}, is more costly to access than SRAM.
-\sys supports memory virtualization with a mechanism that moves pages of the address space from their location in a large, non-volatile memory to a small, volatile memory that tasks use during their execution.
+\sys exposes non-volatile memory for \emph{protected} variables to
+the application as a virtual address space partitioned into pages.
+Paged virtual memory is a key enabler of task coalescing, which
+we described in Section~\ref{sec:task_coalescing}. Virtualization simplifies
+coalescing, because the system gets the freedom to move application's data
+between volatile and non-volatile memory transparently.
+%
+\sys manages memory such that application tasks operate primarily from
+efficient volatile SRAM rather than non-volatile FRAM, which is more costly to access than SRAM,
+as we showed in Figure~\ref{fig:framEnergy}. \sys runtime implements memory
+virtualization by moving pages of the address space from the large,
+non-volatile memory to a small, volatile memory on demand and writing all dirty
+pages back atomically. Since memory moves are at page granularity, \sys
+exploits hardware\hyp{}accelerated block copies.
 
-\subsection{Paged Non-Volatile Virtual Memory}
+%\subsection{Paged Non-Volatile Virtual Memory}
 
-\sys paging scheme partitions the non-volatile memory address space into pages of {\em protected data}.
-As they are accessed by a task, pages of protected data are swapped into a {\em page buffer} \texttt{pagesBuf} that is stored in volatile memory and holds a pre-determined number of pages.
-When any modified page in the page buffer is swapped out, it is buffered in a dedicated {\em commit buffer} \texttt{pagesTemp} for later.
-\sys uses two-phase commit to write buffered, modified pages back to their original memory locations atomically, ensuring data consistency even with power failures.
-\sys's page-based memory model allows moving large blocks of data between non-volatile and volatile memory, which \sys does efficiently using common,
-hardware DMA block-copy operations, unlike prior work~\cite{chain,alpaca} that moves one variable at a time.
-Moreover, \sys's paging and buffering scheme is a key enabler of task coalescing, which we describe in Section~\ref{sec:task_coalescing}.
-Paging simplifies coalescing because deferring a commit of paged data at a task boundary entails leaving paged data buffered only, not requiring additional state management logic at task boundaries.
+As they are
+accessed by a task, pages are swapped from their home location
+in the direct\hyp{}mapped non-volatile memory, called
+\texttt{\underline{store}} (underline denotes allocation in non-volatile
+memory) into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The
+\texttt{working} buffer is allocated to take up the volatile memory that
+remains after space for stack is allocated.
+%
+\sys uses two-phase commit to write dirty pages from the \texttt{working} buffer
+back to their home location in \texttt{\underline{store}} atomically,
+ensuring data consistency even with power failures.
+%
+When any dirty page is swapped out from the \texttt{working} buffer due to
+eviction or task transition from last coalesced task, the first phase writes
+the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
+buffer, and the second phase writes it from
+\texttt{\underline{shadow}} to \texttt{\underline{store}}.
+%
+Unline prior work~\cite{chain,alpaca} that moves scattered variables one at a
+time between volatile and non-volatile memory, \sys's page-based
+design moves contiguous blocks of data.
+%
+\sys takes full advantage of hardware\hyp{}accelerated block copies using
+Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x
+fewer cycles on MSP430FRxxxx.
 
-\subsubsection{Address Translation and Variable Access}
+\subsection{Address Translation and Variable Access}
 
-\sys virtualizes the protected, non-volatile data that a task can access through a modified memory interface.
-The interface includes two operations: \texttt{RVAR(var)}, which reads the value of the memory location {\tt var} and \texttt{WVAR(var,val)}, which writes the value {\tt val} into the memory location {\tt var}.
-{\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory and translates it into a \emph{virtual address} in the page buffer in volatile memory.
-In \sys, a virtual address is composed of a \emph{page tag} and a \emph{page offset}.
-The page tag identifies in which page of non-volatile memory the memory location resides and the page offset identifies a particular byte of that page.
+A \sys task must access protected non-volatile variables through a restricted memory
+interface. The interface includes \texttt{RVAR(var)}, which reads the value of
+variable {\tt var} and \texttt{WVAR} \texttt{(var,val)}, which assigns
+value {\tt val} to variable {\tt var}. The implementation of {\tt
+RVAR} is shown in Algorithm~\ref{algo:rwar} and {\tt WVAR} implementation is
+similar except for an assignment instead of the return.
+%
+{\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory into a \emph{virtual address} in the page buffer in volatile memory.
+In \sys, a virtual address is composed of a \emph{page tag} that identifies the page and a \emph{page offset} that identifies a byte.
 
@@
 
-\begin{algorithm}[t]
-	\caption{\texttt{RVAR(var)} pseudo-code}
+\begin{algorithm}
+	\caption{\texttt{RVAR}(variable $v$)}
 	\label{algo:rwar}
@@
 	\begin{algorithmic}[1]
-		\State $\texttt{tag}\leftarrow \texttt{getTag(var)}$ 
-		\If { \texttt{tag} is not found in \texttt{pagesBuf} }	\Comment{Check if the page is in the page buffer}
-		\State	\texttt{PageFault(tag)} \Comment{Page is not in the page buffer, bring it}
-		\Else 
-		\State \texttt{idx} $\leftarrow$ \texttt{getPageIndex()} \Comment{Get the page index in the page buffer}
+		\State $t \leftarrow \Call{GetTag}{v}$ 
+        \State $i \gets \{ j\ |\ \Call{GetTag}{\texttt{working}[j]} = t \}$ \Comment{Look in the page buffer}
+		\If { $i = \emptyset$ } \Comment{Is the variable in a resident page?}
+		\State	$i \gets \Call{PageFault}{t}$ \Comment{Page in from non-volatile memory}
 		\EndIf
-				\State $\texttt{offset}\leftarrow \texttt{getOffset(var)}$ 		
-		\State \texttt{return pagesBuf[idx][offset]}  \Comment{Return directly from page buffer}
+		\State $o \leftarrow \Call{GetOffset}{v}$ 		
+		\State \Return \texttt{working}[$i$][$o$]  \Comment{Return directly from page buffer}
 	\end{algorithmic}
@@
 
-After address translation, \sys attempts to access the protected variable's location in the volatile paging buffer.
-\sys keeps track of the page tags for the pages currently resident in the paging buffer.
-When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt pagesBuf}.
-If the accessed variable's page is resident in the page buffer, then the tag of one of the pages in this buffer is equal to the accessed variable's page tag.
-The executing memory operation indexes the page buffer using the variable's page offset and manipulates that byte (i.e., either reading or writing).
-If the accessed variable's page tag is not found then the variable's page is not resident in the page buffer; the operation incurs a {\em page fault}.
-At a page fault, \sys must swap out one of the pages that is resident in the page buffer and swap in the variable's page.
+After address translation, \sys attempts to access the protected variable's
+location in the volatile paging buffer. \sys keeps track of the page tags for
+the pages currently resident in the paging buffer. When a task accesses a
+variable, it compares the variable's page tag to tags of the pages in {\tt
+working} (Line 2).
+%
+%If the accessed variable's page is resident in the page buffer,
+%then the tag of one of the pages in this buffer is equal to the accessed
+%variable's page tag.
+%
+If the accessed variable's page tag is not found in the page buffer, the
+operation incurs a {\em page fault} (Line 4). At a page fault, \sys must swap
+out one of the pages that is resident in the page buffer (to the non-volatile
+commit buffer, \texttt{\underline{shadow}}) and swap in the variable's page.
+%
+The byte is accessed in the page buffer at the index of the resident page and
+the variable's page offset (Lines 5-6).
 
-Algorithm~\ref{algo:rwar} shows pseudo-code of \texttt{RVAR} and we omit a listing for {\tt WVAR} as it is very similar except it writes the memory location and returns nothing.
-If the page tag of the variable \texttt{var} is not found in \texttt{pagesBuf} (e.g.
+%Algorithm~\ref{algo:rwar} shows pseudo-code of \texttt{RVAR} and we omit a listing for {\tt WVAR} as it is very similar except it writes the memory location and returns nothing.
+If the page tag of the variable \texttt{var} is not found in \texttt{working} (e.g.
 Lines 2--3), there is a page fault that requires a new page to be swapped in to the page buffer.
 
-\subsubsection{Page Faults and Page Swapping}
+\subsection{Page Faults and Page Swapping}
 
-\begin{algorithm}[t]
-	\caption{\texttt{PageFault(tag)} pseudo-code}
+\begin{algorithm}
+	\caption{\texttt{PageFault}(tag $t$)}
 	\label{algo:pagefault}
@@
 	\begin{algorithmic}[1]
-		\State \texttt{VictimTag} $\leftarrow$ \texttt{selectVictimPage()} \Comment{Select a page to swap out}
-		\If {\texttt{isDirty(VictimTag)} }	\Comment{Check if the victim page is modified}
-		\State \texttt{pageCommit(VictimTag,pagesTemp)} \Comment{Phase I commit the victim page}
+        \State \LeftComment{$i_\text{victim}$: volatile index to next free slot or page to be evicted, set to 0 at boot}
+        \State \LeftComment{\texttt{\underline{numShadowed}}: non-volatile, reset to 0 on boot unless in phase 2 commit}
+        \State $t_\text{victim} \gets \Call{Tag}{\texttt{working}[i_\text{victim}]}$
+		\If {\Call{IsDirty}{working[$i_\text{victim}$]} }	\Comment{Are we evicting a dirty page?}
+		\State $\texttt{\underline{shadow}}[t_\text{victim}] \xleftarrow{\text{DMA}} \texttt{working}[i_\text{victim}]$
+            \Comment{Swap out the page}
+        \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track as list for quick traversal}
+            \State $\texttt{\underline{shadowList}}[\texttt{numUncommited++}] \gets t_\text{victim}$
         \EndIf
-		\If {\texttt{isDirty(tag)}} \Comment{Check if the requested page has been modified}
-		\State {\texttt{getPage(tag,pagesTemp)}} \Comment{Get page from the temporary buffer}
+		\EndIf
+		\If {\Call{IsValid}{\underline{shadow}[$t$]}} \Comment{Had the page been swapped out earlier?}
+		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
+            \Comment{Get page from commit buffer}
 		\Else
-		\State \texttt{getPage(tag,pagesOrg)} \Comment{Get page from the original location}
+		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{store}}[t]$
+            \Comment{Get page from backing store}
 		\EndIf 
+        \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBuf}|$ \Comment{FIFO eviction policy}
 	\end{algorithmic}
@@
 
-A page fault on a memory access requires \sys to swap out one of the pages in the page buffer, i.e.
-\emph{victim page}, preserving updates made to that page, and to copy the accessed page of protected data into the page buffer.
-Algorithm~\ref{algo:pagefault} shows the how \sys handles a page fault.
-In our current implementation, we used a \emph{first-in-first-out} page replacement algorithm to select the victim page in \texttt{pagesBuf} (Line 1).
-If both the page being swapped out and the page being swapped in are not {\em dirty}---which means that they do not contain bytes modified by the executing task---then handling a page fault simply requires copying the accessed page into the page buffer.
-If either the {\em victim} page being swapped out or the {\em accessed} page being swapped in is dirty, \sys must take extra steps when handling a page fault.
+A page fault on a memory access requires \sys to swap out one of the pages in
+the page buffer, i.e. \emph{victim page}, preserving updates made to that page,
+and to copy the accessed page of protected data into the page buffer.
+Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault. A victim
+page in the \texttt{working} buffer is selected for eviction according to
+\emph{first-in-first-out} replacement policy (Lines 2 and 12).  If the page
+being swapped out is not dirty and the page being accessed had not been dirtied
+and swapped out to the \texttt{shadow} buffer since last boot, then
+handling a page fault consists of copying the accessed page from
+\texttt{\underline{store}} into the \texttt{working} buffer (Line 11).
+Otherwise, the page fault handler must take the following extra steps.
 
-\textbf{Dirty Victim Page.} If a task modified any byte in the victim page (i.e.,  using \texttt{WVAR}), then the page is dirty.
-Before swapping in the accessed page, \sys must first preserve the updated values in the dirty page so that they can eventually be committed back to their locations in memory when the task ends.
-To preserve dirty values, \sys copies the victim page to its dedicated memory region called \texttt{pagesTemp}, which is acts as a double buffer for each page of protected memory.
-Dirty pages copied into {\tt pagesTemp} during a task remain there until the task completes, at which point \sys commits the modified pages from {\tt pagesTemp} back to memory (Lines 2--3).
+If a task modified any byte in the victim page
+(i.e.,  using \texttt{WVAR}), then the page is dirty. Before swapping in the
+accessed page, \sys must first preserve the updated values in the dirty page so
+that they can be committed back to their locations in memory
+when the dynamic task ends (but not earlier). To preserve dirty values, \sys
+copies the victim page to the non-volatile \texttt{\underline{shadow}}
+buffer (Line 5) and records the page's tag in a non-volatile list (Lines 6-7)
+to support iteration over shadow pages having traversing all slots in the
+buffer, necessary in the second phase of the commit.
+%
+%Dirty pages remain in {\tt \underline{shadow}} buffer until the
+%dynamic task completes, at which point \sys commits the modified
+%pages from {\tt \underline{shadow}} back to \texttt{\underline{store}}.
+%
+If the accessed page was previously modified and
+swapped out since the last boot (Line 8), the most recent contents of the page is in {\tt
+\underline{shadow}}, not \texttt{\underline{store}} (Line 9).
 
-\textbf{Dirty Accessed Page.} If an accessed page is modified, \sys should swap it into the page buffer from its location in {\tt pagesTemp}, rather than swapping it in from its original location in memory.
-As Algorithm~\ref{algo:pagefault} shows in Line 4, \sys checks if the requested page was previously modified by the executing task.
-To track whether a page is modified, \sys stores a dirty bit with the page that is set when a task modifies the page.
-If an accessed page's dirty bit is set, the page must be swapped in from \texttt{pagesTemp} rather than its original memory location (Line 5), furnishing the task with the page's modified values.
-If the accessed page is not dirty, \sys swaps it in from its original location in memory (Line 7).
-\sys uses DMA block copies for all page swapping operations, leveraging the efficiency of hardware support.
+%As
+%Algorithm~\ref{algo:pagefault} shows in Line 4, \sys checks if the requested
+%page was previously modified by the executing task. To track whether a page is
+%modified, \sys stores a dirty bit with the page and sets it when the page is
+%written to. If an accessed page's dirty bit is set, the page must be swapped in
+%from \texttt{\underline{shadow}} rather than its original memory location (Line 5),
+%furnishing the task with the page's modified values. If the accessed page is
+%not dirty, \sys swaps it in from its original location in memory (Line 7).
 
-\subsubsection{Committing a Page}
+%
+%\sys
+%uses DMA block copies for all page swapping operations, leveraging the
+%efficiency of hardware support.
 
-When a task completes, it must {\em commit} all (dirty) pages stored in \texttt{pagesTemp} back to their original memory locations.
-\sys maintains a global, non-volatile {\em commit bit} that indicates that a task is committing.
-The commit bit is set before the first page is committed and the bit is unset after the last page is committed.
-If power fails during commit, the commit bit remains set and \sys resumes committing before moving on to execute the next task.
-To commit pages from {\tt pagesTemp}, \sys iterates over the pages' dirty bits and for each page with its dirty bit set, \sys DMA block-copies the page back to its
-original memory location, then clears its dirty bit.
+\subsection{Atomic Two-Phase Commit of Dirty Pages}
 
-\subsection{Paging Example}
+When the last task in a coalesced group completes, \sys must commit \emph{all}
+dirty pages in the \texttt{working} buffer and the swapped out pages in
+\texttt{\underline{shadow}} buffer back to their original locations in
+\texttt{\underline{store}}. If a task is allowed to (re-)execute after only
+\emph{some} pages have been copied to \texttt{\underline{store}}, that task
+might see protected variables in an inconsistent state.
 
-Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task using \sys's paging mechanism.
+To make the commit atomic, \sys copies the pages in two phases, implemented by
+Algorithm~\ref{algo:commit}.
+%
+The first phase copies dirty pages in \texttt{working}
+buffer to the non-volatile \texttt{\underline{shadow}} buffer (Line 4). 
+The second phase, copies pages from \texttt{\underline{shadow}} to \texttt{store} (Line 12).
+If power fails during the first phase, the whole
+commit is aborted, and the dynamic coalesced task restarts from its first
+static task. If power fails in the second phase, the commit will be safely
+resumed on next boot.
+The non-volatile metadata for the second phase is the \texttt{\underline{committing}}
+bit, which is set before the first page is copied (Line 9) and cleared after the last
+page is copied (Line 14), and \texttt{\underline{numShadowed}} counter, which
+indexes into the \texttt{\underline{shadowList}} (Line 11), is
+decremented after each page is copied (Line 13), and cleared once the phase
+completes (Line 18).
+
+\sys non-volatile \texttt{\underline{shadowList/numShadowed}} implementation is
+a more efficient alternative to a simpler implementation that keeps a valid bit
+in each entry of \texttt{\underline{shadow}} and, on every boot into second
+phase, traverses all entries, copying the pages with a set bit, and
+clearing the bit.
+%
+\sys assumes that the write to a non-volatile word in memory is atomic.
+
+\begin{algorithm}[t]
+	\caption{Two-phase commit}
+	\label{algo:commit}
+	\scriptsize
+	%\small
+	\begin{algorithmic}[1]
+        \Procedure{CommitPhase1}{} \Comment{Invoked from \texttt{next\_task} on last coalesced task}
+            \For {$i \in 0..|\texttt{working}|-1$}
+                \State $t \gets \Call{Tag}{\texttt{working}[i]}$
+                \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
+                \State $\texttt{\underline{shadowList}}[\texttt{\underline{numShadowed}}] \gets t$
+                \State $\texttt{\underline{numUncommited}} \gets \texttt{\underline{numUncommited}} + 1$
+            \EndFor
+            \State \Call{CommitPhase2}{}
+        \EndProcedure
+        \Procedure{CommitPhase2}{}
+            \State $\texttt{\underline{committing}} \gets \textbf{true}$
+            \While{\texttt{\underline{numUncommited}} > 0}
+                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{numUncommited}} - 1]$
+                \State $\texttt{\underline{store}}[t] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
+                \State $\texttt{\underline{shadow}}[t] \gets \textbf{nil}$ \Comment{Clear entry for page fault handler}
+                \State $\texttt{\underline{numUncommited}} \gets \texttt{\underline{numUncommited}} - 1 $
+            \EndWhile
+            \State $\texttt{\underline{committing}} \gets \textbf{false}$
+        \EndProcedure
+        \Procedure{OnBoot}{} \Comment{Invoked on every boot, before transitioning to a task}
+            \If { \texttt{\underline{committing}} } \Call{CommitPhase2}{}
+            \EndIf 
+            \State $\texttt{\underline{numShadowed}} \gets 0$, $i_\text{victim} \gets 0$
+        \EndProcedure
+	\end{algorithmic}
+\end{algorithm}
+
+%\subsection{Paging Example}
+%
+%Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task using \sys's paging mechanism.
 The task manipulates the protected variables {\tt x}, {\tt y}, and {\tt z}.
@@
 Since variable \texttt{x} is kept in a different page in \texttt{pagesOrg}, accessing it creates a page fault.
-At this point, \sys swaps the active page in \texttt{pagesBuf} by committing it into the non-volatile {\tt pagesTemp} region and copying the corresponding page from \texttt{pagesOrg} to \texttt{pageBuf}.
+At this point, \sys swaps the active page in \texttt{working} by committing it into the non-volatile {\tt \underline{shadow}} region and copying the corresponding page from \texttt{pagesOrg} to \texttt{pageBuf}.
 When the task completes, \sys sets the commit bit, copies the dirty page from \texttt{pagesTmp} to its original location in \texttt{pagesOrg} and unsets the commit bit.
 
-\begin{figure}[t]
-	\centering
-	\includegraphics[width=0.75\columnwidth]{figures/sram-buffer}
-	\caption{The interaction between the task, volatile buffer and the non-volatile memory (FRAM).
+%\begin{figure}[t]
+%	\centering
+%	\includegraphics[width=0.75\columnwidth]{figures/sram-buffer}
+%	\caption{The interaction between the task, volatile buffer and the non-volatile memory (FRAM).
 Initially, the persistent variable \texttt{x} is not in the volatile buffer but \texttt{y} and \texttt{z} are.}
-	\label{fig:volatile-buffer}
-\end{figure}
+%	\label{fig:volatile-buffer}
+%\end{figure}
 
-\subsection{Memory Consistency}
+%\subsection{Memory Consistency}
 
-\sys's paging mechanism ensures that a task only ever executes using consistent protected data.
-During task execution, modifications to protected data do not affect their original memory locations because a task reads and writes the volatile page buffer only, and modified pages are swapped out to \texttt{pagesTemp} until commit.
-A power failure erases the contents of the page buffer, preventing a re-executing task from observing updates in the page buffer from a previous execution attempt.
-On a reboot after a power failure,
-\sys clears all dirty bits in {\tt pagesTemp} {\em if the commit bit is unset}.
-Clearing these dirty bits ensures that all accesses to protected variables correctly access their original memory locations.
-The non-volatile commit bit persists across power failures and \sys repeatedly attempts to commit until it succeeds, despite arbitrary power failures.
+\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent
+protected data. During task execution, modifications to protected data do not
+affect their original memory locations, because a task reads and writes the
+volatile \texttt{working} buffer only, and modified pages are kept in
+\texttt{\underline{shadow}} until commit. A power failure erases the contents
+of the page buffer, preventing a re-executing task from observing updates in
+the page buffer from a previous execution attempt. Clearing \texttt{\underline{shadow}}
+entries as part of the second phase commit (Line 13), ensures that all accesses to
+protected variables in subsequent tasks correctly access their original memory
+locations in \texttt{\underline{store}}.
+%
+%The non-volatile commit bit persists
+%across power failures and \sys repeatedly attempts to commit until it succeeds,
+%despite arbitrary power failures.
 

### 2017-11-17 paper/abstract.tex

@@
 Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing.
-If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve execution times.
-\todo{Provide concrete numbers for improvement}{Amjad} \todo{Provide concrete numbers for checkpointing improvement}{Amjad}
+If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve execution times up to 4.5 times compared to static task code.

### 2017-11-17 paper/background.tex
background: add a paragraph break
rename the fram access figures

@@
 		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
-		\caption{Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
+		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
     \end{subfigure}
-	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}\label{fig:code_demo_incosistency}
+	\caption{\textbf{Code example demonstrating effect of write after read on volatile memory checkpointing.}}\label{fig:code_demo_incosistency}
 \end{figure}
@@
 	\end{tabular}
-	\caption{Memory consistency enforcement overheads of intermittent execution models.
-\emph{Reg.} includes the entire register file, \emph{PC} is the program counter, \emph{Channel data} are variables explicitly task-shared by the programmer, and \emph{WAR variables} are variables involved in WAR dependences (\emph{NV}: non-volatile).}
+	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
 	\label{table:chechpoint_comparison}
@@
 
-Task-based models incur costs that motivate \sys.
+%Task-based models incur costs that motivate \sys.
 
@@
 \begin{subfigure}[t]{.49\columnwidth}
-	\centering \includegraphics[width=\columnwidth]{figures/fram_write.eps}
-	\caption{Cost of \emph{write} operation.}
+	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
+	\caption{Cost of \emph{write} operation}
 \end{subfigure}%
 \begin{subfigure}[t]{.49\columnwidth}
-	\centering \includegraphics[width=\columnwidth]{figures/fram_read.eps}
-	\caption{Cost of \emph{read} operation.}
+	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
+	\caption{Cost of \emph{read} operation}
 \end{subfigure}
-	\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write and read operation.\todo{pico jole$\rightarrow$Picojoule}{Amjad}}\label{fig:framEnergy}
+	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.}}\label{fig:framEnergy}
 \end{figure}
@@
 
-\textbf{Fixed-size Tasks are Inflexible and Inefficient.} If a programmer writes a task that consumes {\em more} energy than the device can buffer, the task will never complete using buffered energy.
-{\em preventing forward progress} by repeatedly re-executing.
+\textbf{Fixed-size Tasks are Inflexible and Inefficient.} If a programmer writes a task that consumes {\em more} energy than the device can buffer, the task will never complete using buffered energy {\em preventing forward progress} by repeatedly re-executing.
 If a programmer writes tasks that consume far {\em less} energy than a device can buffer, the system may operate {\em inefficiently}.
@@
 However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions.
+
 Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity.

### 2017-11-17 paper/evaluation.tex

@@
 	\includegraphics[width=\columnwidth]{figures/coalStrategies}
-	\caption{\textbf{\sys coalescing strategies (`WHD' and `HD') performance per application on intermittent power compared to \sys without coalescing (`NO').} Coalescing provides benefit for all applications, while we observe little difference between `WHD' and `HD'.}
+	\caption{\textbf{\sys coalescing strategies (`WHD' and `HD') performance per application on intermittent power compared to \sys without coalescing (`NO').} Coalescing provides benefit for all applications, while we observe little difference between `WHD' and `HD'.
+\todo{Change WHD to TSHC and HD to HC}{Amjad}}
 	\label{fig:coalescing}

### 2017-11-17 paper/coalescing.tex
coalescing: more general strategy definitions

@@
 
-\todo{clarify}{AC}
-With this strategy \sys calculates only one parameter $C=C_\text{up}=C_\text{down}$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period.
+Assuming tasks of similar energy costs, an accessible predictor of a
+suitable dynamic task size is the number of successfully completed static tasks
+in the recent past.
+%
+The HC policy keeps track of the number of static task boundaries crossed since
+the last boot to set a \emph{target} dynamic task size.
+%
+This target is expressed in units of ``one static task'' and persisted in
+non-volatile memory.
+%
+When target number of boundaries is crossed, the policy stops coalescing,
+thereby setting the size of the dynamic task at the target.
+
+The target is conservatively initialized to zero before the origin task runs
+for the first time, i.e. $T_0 \gets 0$.
+%
+After each successful end of a dynamic task, the policy adjusts the target
+upwards, to explore the possibility for larger dynamic tasks.
+%
+The optimistic upward adjustment of the target is counteracted by the downward
+adjustment that takes place after a dynamic task fails to complete after energy
+is exhausted, i.e. on each reboot.
+%
+The adjustments are geometric by a parametrizable factor $\alpha$, i.e.
+$T_{i+1} \gets \lceil \alpha T_i \rceil$, where $\alpha  >
+1$ for upward and $\alpha < 1$ for downward adjustment.
+%
+In our experiments, we empirically tuned the controller parameters to $\alpha_\text{up} =
+1.5$ and $\alpha_\text{down} = 0.5$.
+%
+The absolute values of the parameters provide a reaction time that is large
+enough to filter out outliers, due to short peaks or lulls in energy supply or
+due to disparate task energy cost.
+%
+The asymmetry reflects conservatism: fast reaction to a decreased energy
+availability -- which prioritizes reducing wasted executions over attempts to
+extract maximum gain -- and slower reaction to increased energy -- which
+cautiously pushes for a larger gain.
+
+%With this strategy \sys calculates only one parameter $C=C_\text{up}=C_\text{down}$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period.
 \sys begins execution in a coalesced task with $S_t=0$.
@@
 short bursts of energy or short bursts of frequent power interrupts.
-
+%
 %\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys.
@@
 
-The history-aware coalescing policy used the number of tasks in the group as a
+The HC policy used the number of tasks in the group as a
 rough proxy for the energy cost of the task group.
@@
 %
-Profiling execution time instead of direct energy consumption is
-advantageous, because the former does not require specialized
-current\hyp{}measurement hardware and can be readily done on any
-off-the-shelf microcontroller.
+Profiling execution time instead of direct energy consumption is advantageous,
+because the former does not require any external current\hyp{}measurement
+hardware and can be readily done on any off-the-shelf microcontroller.
 %
@@
 %
-At runtime, the decision to coalesce or commit at each task boundary proceeds
-similarly with the HC policy, but the task completion counter, $S_t$ has units
-of ``task execution time''.\todo{be precise}{AC}
+At runtime, the decision for whether to continue coalescing at each task
+boundary is determined based on the adaptive target size of the dynamic task as
+in the HC policy, but the target and the crossed boundary counter have units of
+``task execution time.''
 
-To quantify the importance of task size in the coalescing decision, we measured
+To motivate accounting for task size in coalescing decisions, we measured
 the distribution of task sizes in our benchmarks (described in

### 2017-11-17 paper/evaluation.tex
eval: todo about 0 instead of X

@@
 	\includegraphics[width=\columnwidth]{figures/pagefault}
-	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X: no page faults.}
+	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X: no page faults.
+\todo{change X to 0, remove the definition of X from caption}{Amjad}}
 	\label{fig:page_size}

### 2017-11-17 paper/evaluation.tex
some minor eval edits for a submittable safety draft

--- 2017-11-17 paper/evaluation.tex
@@
 
-\textbf{Coalescing Strategies.} Figure~\ref{fig:coalescing} shows the run time performance of both \sys's coalescing, compared to \sys running without coalescing enabled with data normalized to the run time performance of \sys with coalescing\todo{which one: WHD, or HD?}{Amjad}.
-We show data from experiments with intermittent power, where intermittency was generated in software \todo{Say briefly how we generate full reset with MSP430}{Amjad} by creating uniformly distributed random pattern of power interrupts with an average power availability and unavailability time of 15\,ms and 10\,ms respectively.
+\textbf{Coalescing Strategies.} Figure~\ref{fig:coalescing} shows the run time
+of \sys's two coalescing strategies normalized to the run time without
+coalescing. We show data from experiments with intermittent power, where intermittency was generated in software
+%\todo{Say briefly how we generate full reset with MSP430}{Amjad}
+by triggering a reset from software at intervals that were uniformly distributed with an average power availability and unavailability interval of 15\,ms and 10\,ms respectively.
 Each application ran with the same generated pattern for the fairness of comparison.
-
-While we see the benefit from coalescing for all applications (except for dft \todo{why coalescing for dft does not work but in Fig.
-9 does?}{Amjad}), the runtime difference between WHD and HD strategies is actually small \todo{say why this difference is small}{Amjad}.
-For this reason we shall proceed with HD coalescing strategy in subsequent evaluations.
+%
+The results show the benefit from coalescing for most applications, with the
+exception of dft, where wasted work dominated the gains from coalescing.
+ %\todo{why coalescing for dft does not work but in Fig. 9 does?}{Amjad}
+The difference in performance between HC and TSHC strategies is highly
+application-dependent, with some applications benefiting from size awareness
+while others suffering, due to decisions based on inaccurate prediction of the
+actual task energy consumption.
+The TSHC gains are particularly small (or negative) for applications with
+narrower range of task sizes in Figure~\ref{fig:task_profiling}.
+%\todo{say why this difference is small}{Amjad}.
+%For this reason we shall proceed with HD coalescing strategy in subsequent evaluations.
 
@@
 	\includegraphics[width=\columnwidth]{figures/coalStrategies}
-	\caption{\textbf{\sys coalescing strategies (`WHD' and `HD') performance per application on intermittent power compared to \sys without coalescing (`NO').} Coalescing provides benefit for all applications, while we observe little difference between `WHD' and `HD'.}
+	\caption{\textbf{\sys coalescing strategies performance per application on intermittent power compared to \sys without coalescing (`NO').}}
+%\todo{Change WHD to TSHC and HD to HC}{Amjad}}
 	\label{fig:coalescing}
@@
 	\includegraphics[width=\columnwidth]{figures/pagefault}
-	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X: no page faults.}
+	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X = zero page faults.}
+%\todo{change X to 0, remove the definition of X from caption}{Amjad}}
 	\label{fig:page_size}

### 2017-11-17 paper/intro.tex
intro: enlarge figure
intro

@@
 \begin{figure}
-	\includegraphics[width=0.85\columnwidth]{figures/coalescing-is}
+	\includegraphics[width=0.95\columnwidth]{figures/coalescing-is}
     \caption{\label{fig:coalesce}\textbf{An execution of three
@@
 
-Unfortunately, naively merging two atomic tasks might produce a non-atomic task, that may leave memory inconsistent after a partial execution.
-A merge breaks atomicity when the accesses to non-volatile memory from the previously atomic merged tasks form a \emph{write-after-read (WAR)} dependency.
+Unfortunately, naively merging two atomic tasks might produce a non-atomic task, that may leave non-volatile memory inconsistent after a partial execution.
+A merge breaks atomicity when atomic merged tasks form a \emph{write-after-read (WAR)} dependency---for instance if two tasks \texttt{\{x++\}; \{vector[x]=v;\}} are merged and if the power failure occurs after \texttt{x++}, the value \texttt{x} will be increased twice when the merged task is restarted, that leads to an inconsistency.
 %
@@
 % =======
-\textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: a task coalescing system for task-based programming models on intermittently-powered devices.
+\textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: the first task coalescing system for task-based programming models on intermittently-powered devices.
 The capabilities required for effective coalescing mark our contributions:

### 2017-11-17 paper/systemdescription.tex
overview: whitespace
system description background
sec 4

@@
 constructs for defining tasks~\cite{chain,alpaca} and accessors for
-protected variables shared between tasks. These anotations are added either by
+protected variables shared between tasks. These annotations are added either by
 the programmer (in the manual translation mode) or the compiler (in automatic
@@
 any direct callers but receive control by explicit {\tt next\_task} statements.
-In \sys, a task transition atomically commits the memory operations of the
-current task.  The programmer identifies each task by adding the {\tt task}
+In \sys, a task transition does not atomically commit the memory operations of the current task---commit is deferred until the boundary of the coalesced task is reached.
+The programmer identifies each task by adding the {\tt task}
 keyword to the function's declaration and designates one task as the {\em
 origin} task at which execution begins. After a power failure,
-\sys restarts from the last (partially) executed task.
+\sys restarts from the last (partially) executed coalesced task.
 %
@@
 
-\textbf{\sys Task Scheduler.} The scheduler {\em coalesces}
-statically\hyp{}defined tasks into dynamic tasks sized to match the available
-energy. By default, tasks run in sequence and each task commits as it
-completes, potentially incurring unnecessary overhead for commits between
+\textbf{\sys Task Scheduler.} The scheduler
+{\em coalesces} statically\hyp{}defined tasks into dynamic tasks sized to match
+the available energy. By default, tasks run in sequence and each task commits
+as it completes, potentially incurring unnecessary overhead for commits between
 consecutive, non-failing tasks. \sys dynamically coalesces consecutive tasks by

### 2017-11-17 paper/paging.tex
paging: numUncomitted -> numShadowed rename bug
sec 5 - typos

@@
 %
-Unline prior work~\cite{chain,alpaca} that moves scattered variables one at a
+Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a
 time between volatile and non-volatile memory, \sys's page-based
@@
         \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track as list for quick traversal}
-            \State $\texttt{\underline{shadowList}}[\texttt{numUncommited++}] \gets t_\text{victim}$
+            \State $\texttt{\underline{shadowList}}[\texttt{numShadowed++}] \gets t_\text{victim}$
         \EndIf
@@
                 \State $\texttt{\underline{shadowList}}[\texttt{\underline{numShadowed}}] \gets t$
-                \State $\texttt{\underline{numUncommited}} \gets \texttt{\underline{numUncommited}} + 1$
+                \State $\texttt{\underline{numShadowed}} \gets \texttt{\underline{numShadowed}} + 1$
             \EndFor
@@
             \State $\texttt{\underline{committing}} \gets \textbf{true}$
-            \While{\texttt{\underline{numUncommited}} > 0}
-                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{numUncommited}} - 1]$
+            \While{\texttt{\underline{numShadowed}} > 0}
+                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{numShadowed}} - 1]$
                 \State $\texttt{\underline{store}}[t] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
                 \State $\texttt{\underline{shadow}}[t] \gets \textbf{nil}$ \Comment{Clear entry for page fault handler}
-                \State $\texttt{\underline{numUncommited}} \gets \texttt{\underline{numUncommited}} - 1 $
+                \State $\texttt{\underline{numShadowed}} \gets \texttt{\underline{numShadowed}} - 1 $
             \EndWhile

### 2017-11-17 paper/evaluation.tex
eval: match the legend: NONE
explain the blue section of a page fault bar

@@
 	\includegraphics[width=\columnwidth]{figures/coalStrategies}
-	\caption{\textbf{\sys coalescing strategies performance per application on intermittent power compared to \sys without coalescing (`NO').}}
+	\caption{\textbf{\sys coalescing strategies performance per application on intermittent power compared to \sys without coalescing (`NONE').}}
 %\todo{Change WHD to TSHC and HD to HC}{Amjad}}
@@
 	\includegraphics[width=\columnwidth]{figures/pagefault}
-	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X = zero page faults.}
+	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X = zero page faults; the blue section of a bar represents a full page swap (replacing pages in both FRAM and SRAM).}
 %\todo{change X to 0, remove the definition of X from caption}{Amjad}}

### 2017-11-17 paper/evaluation.tex
color bars in fig 9 explained

@@
 	\includegraphics[width=\columnwidth]{figures/pagefault}
-	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X = zero page faults; the blue section of a bar represents a full page swap (replacing pages in both FRAM and SRAM).}
+	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X = zero page faults; red bar: half-page swap, blue bar: full page swap (replacing pages in both FRAM and SRAM).}\vspace{-0.5cm}
 %\todo{change X to 0, remove the definition of X from caption}{Amjad}}
@@
 	\end{tabular}
-	\caption{\textbf{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns)}; $^\dagger$dft does not compile with Clang}\label{table:compiler_result}
+	\caption{\textbf{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns)}; $^\dagger$dft does not compile with Clang}\label{table:compiler_result}\vspace{-0.5cm}
 \end{table}

### 2017-11-30 IPE_Poster/example.tex
major update (most important update is the paging)

% Copyright (c) 2013 Joost van Zwieten
%
% Permission is hereby granted, free of charge, to any person obtaining a copy
% of this software and associated documentation files (the "Software"), to deal
% in the Software without restriction, including without limitation the rights
% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
% copies of the Software, and to permit persons to whom the Software is
% furnished to do so, subject to the following conditions:
%
% The above copyright notice and this permission notice shall be included in
% all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
% FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
% THE SOFTWARE.
%
\documentclass{tudelftposter}

% optional, makes QR code clickable
\usepackage[hidelinks,implicit=false,bookmarks=false]{hyperref}

\title{The Title}

\addauthornote{mail}[1]{\ttfamily j.f.nash@tudelft.nl}
\addauthornote{diam}[2]{Delft Institute of Applied Mathematics, TU Delft}
\addauthornote{tp}[3]{Department of Theoretical Physics, TU Delft}

\addauthor[mail,diam]{J.F. Nash}
\addauthor[diam,tp]{C.F. Gau{\ss}}
\addauthor[diam]{J. von Neumann}
\addauthor[tp]{R.P. Feynman}

\addfootimage(c:right column.center)[Delft Institute of Applied Mathematics]{tudelft}
\ifqrcodessupported
  % NOTE: compile this document with lualatex instead of pdflatex
  \addfootqrcode(l:left column.left)[tudelft-poster repository]{http://github.com/joostvanzwieten/tudelft-poster}
\else
  \addfootobject*(l:left column.left)[research web page]{%
    \begin{tikzpicture}
      \node[red,draw,inner sep=1ex] {%
        \begin{minipage}{15cm}
          \footnotesize%
          Please compile this document with lualatex to see a QR code here:\\
          \null\texttt{\quad lualatex \jobname}
        \end{minipage}};
    \end{tikzpicture}}
\fi

\begin{document}

\section{Introduction}

Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit,
vestibulum ut, placerat ac, adipiscing vitae, felis. Curabitur dictum gravida
mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna.
Donec vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et
netus et malesuada fames ac turpis egestas. Mauris ut leo. Cras viverra metus
rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices.  Phasellus eu
tellus sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium
quis, viverra ac, nunc. Praesent eget sem vel leo ultrices bibendum. Aenean
faucibus.  Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla.
Curabitur auctor semper nulla. Donec varius orci eget risus. Duis nibh mi,
congue eu, accumsan eleifend, sagittis quis, diam. Duis eget orci sit amet orci
dignissim rutrum.  See Figure \ref{fig:test}.

\begin{figure}
  \centering
  \framebox{\rule{0pt}{5cm}\rule{5cm}{0pt}}
  \caption{test}
  \label{fig:test}
\end{figure}

Nam dui ligula, fringilla a, euismod sodales, sollicitudin vel, wisi. Morbi
auctor lorem non justo. Nam lacus libero, pretium at, lobortis vitae, ultricies
et, tellus. Donec aliquet, tortor sed accumsan bibendum, erat ligula aliquet
magna, vitae ornare odio metus a mi nascetur ridiculus mus.
\begin{equation}
  \int_\Omega v_i \left(
    u_{i,t} + u_j u_{i,j} + p_{,i} - R^{-1} u_{i,jj}
  \right) =
  \int_\Omega v_i f_i.
\end{equation}
Morbi ac orci et nisl hendrerit mollis.  Suspendisse ut massa. Cras nec ante.
Pellentesque a nulla.  Cum sociis natoque penatibus et magnis dis parturient
montes, nascetur ridiculus mus. Aliquam tincidunt urna. Nulla ullamcorper
vestibulum turpis. Pellentesque cursus luctus mauris.

\section{Spam}

Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at,
tincidunt tristique, libero. Vivamus viverra fermentum felis. Donec nonummy
pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa ac
quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas
lacinia. Nam ipsum ligula, eleifend at, accumsan nec, suscipit a, ipsum.  Morbi
blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia nulla
vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent
euismod nunc eu purus. Donec bibendum quam in tellus. Nullam cursus pulvinar
lectus. Donec et mi.  Nam vulputate metus eu enim. Vestibulum pellentesque felis
eu massa.

\section{Eggs}

Quisque ullamcorper placerat ipsum. Cras nibh.  Morbi vel justo vitae lacus
tincidunt ultrices. Lorem ipsum dolor sit amet, consectetuer adipiscing elit. In
hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida
sollicitudin, felis odio placerat quam, ac pulvinar elit purus eget enim. Nunc
vitae tortor. Proin tempus nibh sit amet nisl.  Vivamus quis tortor vitae risus
porta vehicula.

\experimentalblockright{%
  SPAM:
  Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at,
  tincidunt tristique, libero.}

Fusce mauris. Vestibulum luctus nibh at lectus.  Sed bibendum, nulla a faucibus
semper, leo velit ultricies tellus, ac venenatis arcu wisi vel nisl. Vestibulum
diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis
porttitor. Vestibulum porttitor. Nulla facilisi. Sed a turpis eu lacus commodo.
\begin{equation}
  u_{i,i} = 0.
\end{equation}
Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis
dui, et vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis
cursus enim ut augue. Cras ac magna. Cras nulla. Nulla egestas. Curabitur a leo.
Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur
consectetuer.

\section{Conclusion}

Suspendisse vel felis. Ut lorem lorem, interdum eu, tincidunt sit amet, laoreet
vitae, arcu. Aenean faucibus pede eu ante.
\begin{itemize}
  \item Quisque vehicula, urna sed ultricies auctor, pede lorem egestas dui, et
  convallis elit erat sed nulla.
  \item Donec luctus.
  \item Curabitur et nunc.
\end{itemize}
Integer arcu est, nonummy in, fermentum faucibus, egestas vel, odio.

\end{document}
% vim: tw=80:ts=2:sts=2:sw=2:et:fdm=marker:fmr=[[[,]]]

### 2017-11-30 IPE_Poster/example_tikz.tex
major update (most important update is the paging)

--- 2017-11-30 IPE_Poster/example.tex
@@
 \usepackage[hidelinks,implicit=false,bookmarks=false]{hyperref}
+
+\usepackage{amsmath}
 
@@
 
+\newlength{\textboxwidth}
+\newlength{\textboxcolumnsep}\setlength{\textboxcolumnsep}{1em}
+
+\setlength\arrayrulewidth{2pt}
+\def\centeredhrulefill{\leaders\hbox{\rule[.5ex]{5pt}{\arrayrulewidth}}\hfill}
+
+% source: http://tex.stackexchange.com/questions/23100/looking-for-an-ignorespacesandpars
+\makeatletter
+\def\ignorespacesandpars{%
+  \begingroup
+  \catcode13=10
+  \@ifnextchar\par
+    {\endgroup\expandafter\ignorespacesandpars\@gobble}%
+    {\endgroup}%
+}
+\makeatother
+
+\def\textbox(#1) at (#2) [#3] <#4,#5> title #6{%
+  \def\textboxname{#1}%
+  \def\textboxpos{#2}%
+  \def\textboxalign{#3}%
+  \def\textboxncolumns{#4}%
+  \setlength{\textboxwidth}{#5}%
+  \def\textboxtitle{#6}%
+%
+  \ifdefstring{\textboxncolumns}{1}{%
+    % single column
+    \def\atbeginofminipage{%
+      \vspace{.5\baselineskip}%
+      \par\noindent\ignorespacesandpars}%
+    \def\atendofminipage{}%
+  }{%
+    % multiple columns
+    \setlength{\textboxwidth}{\textboxncolumns\textboxwidth}%
+    \addtolength{\textboxwidth}{\textboxncolumns\textboxcolumnsep}%
+    \addtolength{\textboxwidth}{-\textboxcolumnsep}%
+    \def\atbeginofminipage{%
+      \setlength{\columnsep}{1em}%
+      \begin{multicols}{\textboxncolumns}%
+        \par\noindent\ignorespacesandpars}%
+    \def\atendofminipage{%
+      \end{multicols}}%
+  }%
+%
+  \node[inner sep=0pt,outer sep=0cm] (#1) at (#2) [#3] \bgroup%
+    \begin{minipage}{\textboxwidth}%
+      \setlength{\parskip}{0pt}%
+      \setlength{\parindent}{1em}%
+      \par\noindent%
+      {% title
+        \color{tudcyan}\scshape%
+        \centeredhrulefill\hspace{1em}%
+        \textboxtitle%
+        \hspace{1em}\centeredhrulefill}%
+      \atbeginofminipage}%
+\def\endtextbox{%
+      \atendofminipage%
+    \end{minipage}%
+  \egroup;%
+}
+
 \begin{document}
+\begin{tikzpicture}[remember picture,overlay,line width=\arrayrulewidth]
 
-\section{Introduction}
+\begin{textbox}(introduction)
+at (body.north west) [below right] <2,24cm>
+title {Introduction}
 
@@
   mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna.
-Donec vehicula augue eu neque. Pellentesque habitant morbi tristique senectus et
-netus et malesuada fames ac turpis egestas. Mauris ut leo. Cras viverra metus
-rhoncus sem. Nulla et lectus vestibulum urna fringilla ultrices.  Phasellus eu
-tellus sit amet tortor gravida placerat. Integer sapien est, iaculis in, pretium
-quis, viverra ac, nunc. Praesent eget sem vel leo ultrices bibendum. Aenean
-faucibus.  Morbi dolor nulla, malesuada eu, pulvinar at, mollis ac, nulla.
-Curabitur auctor semper nulla. Donec varius orci eget risus. Duis nibh mi,
-congue eu, accumsan eleifend, sagittis quis, diam. Duis eget orci sit amet orci
-dignissim rutrum.  See Figure \ref{fig:test}.
+  Donec vehicula augue eu neque.
 
-\begin{figure}
-  \centering
-  \framebox{\rule{0pt}{5cm}\rule{5cm}{0pt}}
-  \caption{test}
-  \label{fig:test}
-\end{figure}
+  Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac
+  turpis egestas. Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus
+  vestibulum urna fringilla ultrices.  Phasellus eu tellus sit amet tortor
+  gravida placerat.
+
+  Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget
+  sem vel leo ultrices bibendum. Aenean faucibus.  Morbi dolor nulla, malesuada
+  eu, pulvinar at, mollis ac, nulla.  Curabitur auctor semper nulla. Donec
+  varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis
+  quis, diam. Duis eget orci sit amet orci dignissim rutrum.
+
+\end{textbox}
+
+\begin{textbox}(spam)
+at (body.north east) [below left] <1,24cm>
+title {Spam}
+
+  \begin{itemize}
+    \item Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat
+    at, tincidunt tristique, libero. Vivamus viverra fermentum felis. Donec
+    nonummy pellentesque ante.
+    \item Phasellus adipiscing semper elit. Proin fermentum massa ac quam. Sed
+    diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas
+    lacinia.
+    \item Nam ipsum ligula, eleifend at, accumsan nec, suscipit a, ipsum.  Morbi
+    blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia
+    nulla vitae enim.
+  \end{itemize}
+
+\end{textbox}
+
+\begin{textbox}(eggs)
+at ([shift={(0pt,-30cm)}]body.north west) [below right] <1,40cm>
+title {Eggs}
+
+  Quisque ullamcorper placerat ipsum. Cras nibh.  Morbi vel justo vitae lacus
+  tincidunt ultrices. Lorem ipsum dolor sit amet, consectetuer adipiscing elit. In
+  hac habitasse platea dictumst.
+
+  Integer tempus convallis augue. Etiam facilisis.  Nunc elementum fermentum
+  wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio
+  placerat quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin
+  tempus nibh sit amet nisl.  Vivamus quis tortor vitae risus porta vehicula.
+
+\end{textbox}
+
+\begin{textbox}(foobar)
+at (body.south west) [above right] <1,35cm>
+title {Foo \& Bar}%
 
@@
 
-\section{Spam}
+\end{textbox}
 
-Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at,
-tincidunt tristique, libero. Vivamus viverra fermentum felis. Donec nonummy
-pellentesque ante. Phasellus adipiscing semper elit. Proin fermentum massa ac
-quam. Sed diam turpis, molestie vitae, placerat a, molestie nec, leo. Maecenas
-lacinia. Nam ipsum ligula, eleifend at, accumsan nec, suscipit a, ipsum.  Morbi
-blandit ligula feugiat magna. Nunc eleifend consequat lorem. Sed lacinia nulla
-vitae enim. Pellentesque tincidunt purus vel magna. Integer non enim. Praesent
-euismod nunc eu purus. Donec bibendum quam in tellus. Nullam cursus pulvinar
-lectus. Donec et mi.  Nam vulputate metus eu enim. Vestibulum pellentesque felis
-eu massa.
-
-\section{Eggs}
-
-Quisque ullamcorper placerat ipsum. Cras nibh.  Morbi vel justo vitae lacus
-tincidunt ultrices. Lorem ipsum dolor sit amet, consectetuer adipiscing elit. In
-hac habitasse platea dictumst. Integer tempus convallis augue. Etiam facilisis.
-Nunc elementum fermentum wisi. Aenean placerat. Ut imperdiet, enim sed gravida
-sollicitudin, felis odio placerat quam, ac pulvinar elit purus eget enim. Nunc
-vitae tortor. Proin tempus nibh sit amet nisl.  Vivamus quis tortor vitae risus
-porta vehicula.
-
-\experimentalblockright{%
-  SPAM:
-  Nulla malesuada porttitor diam. Donec felis erat, congue non, volutpat at,
-  tincidunt tristique, libero.}
-
-Fusce mauris. Vestibulum luctus nibh at lectus.  Sed bibendum, nulla a faucibus
-semper, leo velit ultricies tellus, ac venenatis arcu wisi vel nisl. Vestibulum
-diam. Aliquam pellentesque, augue quis sagittis posuere, turpis lacus congue
-quam, in hendrerit risus eros eget felis. Maecenas eget erat in sapien mattis
-porttitor. Vestibulum porttitor. Nulla facilisi. Sed a turpis eu lacus commodo.
-\begin{equation}
-  u_{i,i} = 0.
-\end{equation}
-Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis
-dui, et vehicula libero dui cursus dui. Mauris tempor ligula sed lacus. Duis
-cursus enim ut augue. Cras ac magna. Cras nulla. Nulla egestas. Curabitur a leo.
-Quisque egestas wisi eget nunc. Nam feugiat lacus vel est. Curabitur
-consectetuer.
-
-\section{Conclusion}
+\begin{textbox}(conclusion)
+at (body.south east) [above left] <1,35cm>
+title {Conclusion}%
 
@@
 
+\end{textbox}
+
+\node (fig1)
+at ([shift={(0pt,-37cm)}]body.north east) [below left] {%
+  \begin{tikzpicture}[x=5cm,y=5cm,line width=\arrayrulewidth]
+    \draw (0,0) -- (4,0) node[anchor=north] {$x$};
+    \draw (0,0) -- (0,3) node[anchor=east] {$f(x)$};
+    \draw
+      (0,0) node[anchor=north] {0}
+      (1,0) node[anchor=north] {1}
+      (2,0) node[anchor=north] {2}
+      (3,0) node[anchor=north] {3};
+    \draw [tudcyan] plot [smooth, tension=1]
+      coordinates {(0,0) (1,2) (2,1) (3,3)};
+  \end{tikzpicture}};
+\node at ([shift={(-2ex,0)}]fig1.south west) [above left] {%
+  \begin{minipage}{15cm}
+    \itshape\raggedleft%
+    {\scshape Figure 1}\\
+    \small%
+    Morbi fringilla, wisi in dignissim interdum, justo lectus sagittis dui, et
+    vehicula libero dui cursus dui
+  \end{minipage}};
+\draw[tudcyan] ([shift={(-1ex,0)}]fig1.south west)--([shift={(-1ex,0)}]fig1.north west);
+
+\end{tikzpicture}
 \end{document}

### 2017-11-30 IPE_Poster/ipe_poster.tex
major update (most important update is the paging)

--- 2017-11-30 IPE_Poster/example_tikz.tex
@@
 % optional, makes QR code clickable
-\usepackage[hidelinks,implicit=false,bookmarks=false]{hyperref}
+% \usepackage[hidelinks,implicit=false,bookmarks=false]{hyperref}
 
@@
 
-\title{The Title}
+\newcommand{\sys}{IPE}
+\newcommand{\sysfull}{Intermittently-Powered Executor}
 
-\addauthornote{mail}[1]{\ttfamily j.f.nash@tudelft.nl}
-\addauthornote{diam}[2]{Delft Institute of Applied Mathematics, TU Delft}
-\addauthornote{tp}[3]{Department of Theoretical Physics, TU Delft}
+\title{\sys{}: \sysfull{}}
 
-\addauthor[mail,diam]{J.F. Nash}
-\addauthor[diam,tp]{C.F. Gau{\ss}}
-\addauthor[diam]{J. von Neumann}
-\addauthor[tp]{R.P. Feynman}
+\addauthor{\large Amjad Y. Majid}
 
-\addfootimage(c:right column.center)[Delft Institute of Applied Mathematics]{tudelft}
-\ifqrcodessupported
-  % NOTE: compile this document with lualatex instead of pdflatex
-  \addfootqrcode(l:left column.left)[tudelft-poster repository]{http://github.com/joostvanzwieten/tudelft-poster}
-\else
-  \addfootobject*(l:left column.left)[research web page]{%
-    \begin{tikzpicture}
-      \node[red,draw,inner sep=1ex] {%
-        \begin{minipage}{15cm}
-          \footnotesize%
-          Please compile this document with lualatex to see a QR code here:\\
-          \null\texttt{\quad lualatex \jobname}
-        \end{minipage}};
-    \end{tikzpicture}}
-\fi
+\addauthornote{mail}{a.y.majid@tudelft.nl}
+\addauthornote{uni}{Delft University of Technology, Delft, NL}
+
+% \addfootimage(c:right column.center)[Delft Institute of Applied Mathematics]{tudelft}
+% \addfootqrcode(l:left column.left)[latex-poster-class repository]{https://github.com/tudelft-diam-na/latex-poster-class}
 
@@
       {% title
-        \color{tudcyan}\scshape%
+        \color{tudcyan}\rm\scshape%
         \centeredhrulefill\hspace{1em}%
@@
 
-  Lorem ipsum dolor sit amet, consectetuer adipiscing elit. Ut purus elit,
-  vestibulum ut, placerat ac, adipiscing vitae, felis. Curabitur dictum gravida
-  mauris. Nam arcu libero, nonummy eget, consectetuer id, vulputate a, magna.
-  Donec vehicula augue eu neque.
+Enabling battery-free devices is a mandatory step towards an environment friendly Internet of Things (IoT).
+However, removing the batteries requires IoT to operate on an harvested power supply.
+This make sustaining long computation very challenging.
+Correspondingly, there are two approaches to enable long-running computations on intermittently-powered IoT: (i) checkpointing, where the volatile state of a program is frequently saved to the non-volatile memory, and (ii) task-based approach, where a programmer splits the code into small idempotent modules.
+Results show that task-based approach performs better than checkpointing.
+However, the task-based model is static approach that does not take advantage of variation in the amount of the harvested energy or adapts to different harvester configuration.
+Therefore, we introduce \sysfull{} which pushes the boundaries of the task-based model by enabling an application to enlarge its task size, on the fly, when there is redundant energy.
 
-  Pellentesque habitant morbi tristique senectus et netus et malesuada fames ac
-  turpis egestas. Mauris ut leo. Cras viverra metus rhoncus sem. Nulla et lectus
-  vestibulum urna fringilla ultrices.  Phasellus eu tellus sit amet tortor
-  gravida placerat.
-
-  Integer sapien est, iaculis in, pretium quis, viverra ac, nunc. Praesent eget
-  sem vel leo ultrices bibendum. Aenean faucibus.  Morbi dolor nulla, malesuada
-  eu, pulvinar at, mollis ac, nulla.  Curabitur auctor semper nulla. Donec
-  varius orci eget risus. Duis nibh mi, congue eu, accumsan eleifend, sagittis
-  quis, diam. Duis eget orci sit amet orci dignissim rutrum.
+% \sysfull{} (\sys{}) is runtime library that facilitates tasks navigation and preserves data/memory consistency of IPDs.
+\sys{} aims at reducing energy consumption and applications execution time.
+It optimizes the commit rate (saving tasks contexts into the non-volatile memory), subject to the number of power interrupts and completed tasks.
 
@@
 
-\begin{textbox}(eggs)
-at ([shift={(0pt,-30cm)}]body.north west) [below right] <1,40cm>
-title {Eggs}
+\begin{textbox}(initresults)
+at ([shift={(0pt,-29cm)}]body.north west) [below right] <1,40cm>
+title {preliminary results}
 
-  Quisque ullamcorper placerat ipsum. Cras nibh.  Morbi vel justo vitae lacus
-  tincidunt ultrices. Lorem ipsum dolor sit amet, consectetuer adipiscing elit. In
-  hac habitasse platea dictumst.
+  \centering
+    \adjincludegraphics[width=.47\columnwidth]{./fig/envAwareness}
 
-  Integer tempus convallis augue. Etiam facilisis.  Nunc elementum fermentum
-  wisi. Aenean placerat. Ut imperdiet, enim sed gravida sollicitudin, felis odio
-  placerat quam, ac pulvinar elit purus eget enim. Nunc vitae tortor. Proin
-  tempus nibh sit amet nisl.  Vivamus quis tortor vitae risus porta vehicula.
+  \vspace{-1cm}
+
+  \justify
+  \noindent\textbf{App execution time, on a varying distance from the reader}
+
+  \centering
+      \adjincludegraphics[width=.47\columnwidth]{./fig/fram_write}
+
+    \vspace{-1cm}
+   \textbf{Energy cost of FRAM and RAM write access}
 
@@
     \itshape\raggedleft%
-    {\scshape Figure 1}\\
+    {\rm\scshape Figure 1}\\
     \small%

### 2017-11-30 IPE_Poster/poster.tex
major update (most important update is the paging)

% Copyright (c) 2013 Joost van Zwieten
%
% Permission is hereby granted, free of charge, to any person obtaining a copy
% of this software and associated documentation files (the "Software"), to deal
% in the Software without restriction, including without limitation the rights
% to use, copy, modify, merge, publish, distribute, sublicense, and/or sell
% copies of the Software, and to permit persons to whom the Software is
% furnished to do so, subject to the following conditions:
%
% The above copyright notice and this permission notice shall be included in
% all copies or substantial portions of the Software.
%
% THE SOFTWARE IS PROVIDED "AS IS", WITHOUT WARRANTY OF ANY KIND, EXPRESS OR
% IMPLIED, INCLUDING BUT NOT LIMITED TO THE WARRANTIES OF MERCHANTABILITY,
% FITNESS FOR A PARTICULAR PURPOSE AND NONINFRINGEMENT. IN NO EVENT SHALL THE
% AUTHORS OR COPYRIGHT HOLDERS BE LIABLE FOR ANY CLAIM, DAMAGES OR OTHER
% LIABILITY, WHETHER IN AN ACTION OF CONTRACT, TORT OR OTHERWISE, ARISING FROM,
% OUT OF OR IN CONNECTION WITH THE SOFTWARE OR THE USE OR OTHER DEALINGS IN
% THE SOFTWARE.
%
\documentclass[onecolumn]{tudelftposter}

% optional, makes QR code clickable
\usepackage[hidelinks,implicit=false,bookmarks=false]{hyperref}

\usepackage{graphicx}				% graphics support (i.e. \includegraphics)
\usepackage[tight]{subfigure}			% DEPRECATED - superseded by subfig and subcaption
\usepackage{amssymb}				% extended symbol collection
\usepackage{amsmath}				% extended math typesetting enhancements
\usepackage{hyperref}				% references and citations as pdf hyperlink
%\usepackage{url}
\usepackage{cite}					% enhanced citations (grouping and spacing)
\usepackage{xcolor}					% color support for graphics
%\usepackage[ancient,keeplastbox]{flushend}	% flush columns to equal length (ancient to maintain column order, keeplastbox to fix last reference having wrong indentation)
\usepackage{todonotes}				% show todo notes
\usepackage{xspace} 			% keep space at end of macros (use \xspace)
\usepackage[mode=text]{siunitx} 	% nice SI units and numerals
\usepackage[capitalize]{cleveref}		% automatic reference naming (e.g. Fig. n, Table n)
\usepackage[inline]{enumitem}			% inline lists
\usepackage{acronym} 			% automatic acronyms, long version on first usage
\usepackage{microtype}				% minor small typographic improvements
\usepackage{tikz}					% drawing library
\usepackage{pgfplots}				% graph drawing library
\usepackage{pgfplotstable}			% table plugin for pgfplots
\usepackage{tikzscale}				% scale tikz pictures as if they were 'boring' normal pictures
\usepackage{helvet}					% sans serif font (for figures)
\usepackage[eulergreek]{sansmath}	% sans serif math (for figures)
\usepackage{algorithm}				% algorithm float
\usepackage[noend]{algpseudocode}	% pseudocode algorithms
\usepackage{xpatch}
\usepackage{tabularx}				% full column width tables
\usepackage{booktabs}				% visually nicer tables
\usepackage{amsthm}
\usepackage[export]{adjustbox}
\usepackage{ragged2e}

\usetikzlibrary{%
	 arrows%
	,automata%
	,calc%
	,fit%
	,patterns%
	,positioning%
	,shadows%
	,shapes.misc%
}

\tikzset{cross/.style={cross out, draw=black, minimum size=2*(#1-\pgflinewidth), inner sep=0pt, outer sep=0pt},
%default radius will be 1pt.
cross/.default={3pt}}

% math typesetting options
\DeclareMathOperator\erfc{erfc}

\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{theorem}{Theorem}
\newtheorem{problem}{Problem}
\newtheorem{corollary}{Corollary}
\newtheorem{design}{Design}
\newtheorem{observation}{Observation}

% some definitions like project code name, easy to change
%\acrodef{<acronym>}[<short name>]{<full name>}
%\acrodef{twi}[I\textsuperscript{2}C]{I\textsuperscript{2}
\acrodef{ack}[{\texttt{ACK}}]{Acknowledgement}
\acrodef{adc}[ADC]{Analog to Digital Converter}
\acrodef{api}[API]{Application Programming Interface}
\acrodef{asm}[ASM]{Assembly}
\acrodef{ble}[BLE]{Bluetooth Low Energy}
\acrodef{blisp}[BLISP]{platform combined of \acs{ble} and \acs{wisp}}
\acrodef{cots}[COTS]{Commercial off-the-Shelf}
\acrodef{crfid}[CRFID]{Computational RFID}
\acrodef{dut}[DUT]{Device Under Test}
\acrodef{epcc1g2}[EPC~C1G2]{EPCglobal Class\,1 Generation\,2}
\acrodef{epc}[{\texttt{EPC}}]{Electronic Product Code}
\acrodef{fram}[FRAM]{Ferroelectric Random Access Memory}
\acrodef{gpio}[GPIO]{General Purpose Input/Output}
\acrodef{gsm}[GSM]{Global System for Mobile communication}
\acrodef{ide}[IDE]{Integrated Developement Environment}
\acrodef{iot}[IoT]{Internet of Things}
\acrodef{llrp}[LLRP]{Low-Level Reader Protocol}
\acrodef{lsb}[LSB]{Least Significant Bit}
\acrodef{nfc}[NFC]{Near Field Communication}
\acrodef{ocr}[OCR]{Optical Character Recognition}
\acrodef{pcb}[PCB]{Printed Circuit Board}
\acrodef{rfid}[RFID]{Radio Frequency Identification}
\acrodef{rf}[RF]{Radio Frequency}
\acrodef{rn16}[{\texttt{RN16}}]{16bit Random Number}
\acrodef{rssi}[RSSI]{Received Signal Strength Indication}
\acrodef{sdk}[SDK]{Software Development Kit}
\acrodef{soc}[SoC]{System on Chip}
\acrodef{spi}[SPI]{Serial Peripheral Interface}
\acrodef{twi}[TWI]{Two Wire Interface}
\acrodef{uart}[UART]{Universal Asynchronous Receiver/Transmitter}
\acrodef{usb}[USB]{Universal Serial Bus}
\acrodef{vlc}[VLC]{Visual Light Communication}
\acrodef{wisp}[WISP]{Wireless Identification and Sensing Platform}

\newcommand{\acresetwithnoexpand}{%
	\acresetall
	\acused{gsm}
	\acused{rfid}
	\acused{rf}
	\acused{rn16}
}

\newlength{\textboxwidth}
\newlength{\textboxcolumnsep}\setlength{\textboxcolumnsep}{1em}

\setlength\arrayrulewidth{2pt}
\def\centeredhrulefill{\leaders\hbox{\rule[.5ex]{5pt}{\arrayrulewidth}}\hfill}

% source: http://tex.stackexchange.com/questions/23100/looking-for-an-ignorespacesandpars
\makeatletter
\def\ignorespacesandpars{%
  \begingroup
  \catcode13=10
  \@ifnextchar\par
    {\endgroup\expandafter\ignorespacesandpars\@gobble}%
    {\endgroup}%
}
\makeatother

\def\textbox(#1) at (#2) [#3] <#4,#5> title #6{%
  \def\textboxname{#1}%
  \def\textboxpos{#2}%
  \def\textboxalign{#3}%
  \def\textboxncolumns{#4}%
  \setlength{\textboxwidth}{#5}%
  \def\textboxtitle{#6}%
%
  \ifdefstring{\textboxncolumns}{1}{%
    % single column
    \def\atbeginofminipage{%
      \vspace{.5\baselineskip}%
      \par\noindent\ignorespacesandpars}%
    \def\atendofminipage{}%
  }{%
    % multiple columns
    \setlength{\textboxwidth}{\textboxncolumns\textboxwidth}%
    \addtolength{\textboxwidth}{\textboxncolumns\textboxcolumnsep}%
    \addtolength{\textboxwidth}{-\textboxcolumnsep}%
    \def\atbeginofminipage{%
      \setlength{\columnsep}{1em}%
      \begin{multicols}{\textboxncolumns}%
        \par\noindent\ignorespacesandpars}%
    \def\atendofminipage{%
      \end{multicols}}%
  }%
%
  \node[inner sep=0pt,outer sep=0cm] (#1) at (#2) [#3] \bgroup%
    \begin{minipage}{\textboxwidth}%
      \setlength{\parskip}{0pt}%
      \setlength{\parindent}{1em}%
      \par\noindent%
      {% title
        \color{tudcyan}\scshape%
        \centeredhrulefill\hspace{1em}%
        \textboxtitle%
        \hspace{1em}\centeredhrulefill}%
      \atbeginofminipage}%
\def\endtextbox{%
      \atendofminipage%
    \end{minipage}%
  \egroup;%
}

%\title{BLISP: Enhancing Backscatter Radio with Active Radio for Computational RFIDs}

\begin{document}
\begin{tikzpicture}[remember picture, overlay, line width=\arrayrulewidth]

%-------------------------- start a block

\begin{textbox}(summary)
at (body.north west) [below right] <1,.47\textwidth>
title {Summary}

IPE abstracts the task based execution model and adapts it during the run time to optimize the execution time.

\end{textbox}
%------------------|

%-------------------------- start a block

\begin{textbox}(Intermittent Execution Models)
at (summary.north west) [below right] <1,.47\textwidth>
title {Intermittent Execution Models}

\begin{itemize}

\item Checkpointing

  \begin{enumerate}

    \item \emph{VM Checkpointing}: 
    \begin{enumerate}
    \item It does not scale with RAM size.
    \item \emph{Roll back chkpt}: FRAM interaction problem.
    \item \emph{Push forward chkpt}:  peripherals reinitialization problem (i.e DMA).
    \end{enumerate}

    \item \emph{NVM Checkpointing}: 
    \begin{enumerate}
    \item FRAM consume more energy than RAM 
    \item FRAM is slower 
    \end{enumerate}

  \end{enumerate}

  \centering
      \adjincludegraphics[width=\columnwidth]{./fig/fram_write}

    \vspace{-1cm}
   \textbf{Energy cost of FRAM and RAM write access}

\end{itemize}  
\begin{itemize}

\item Task based

  \begin{enumerate}
  \item static division 
  \item high memory consumption (multi-versioning)
  \item static dependencies  
  \item code modification 
  \end{enumerate}

    \vspace{1cm}
    \centering
      \adjincludegraphics[width=0.80\columnwidth]{./fig/dynamic_war}

   \textbf{Selective saving results in dynamic war dependencies. }

\end{itemize}

\end{textbox}
%------------------|

%-------------------------- start a block

\begin{textbox}(Motivation)
at (body.north east) [below left] <1,.47\textwidth>
title {Motivation}

  \centering
    \adjincludegraphics[width=\columnwidth]{./fig/envAwareness}

  \vspace{-1cm}

  \justify
  \noindent\textbf{App execution time, on a varying distance from the reader}

  \noindent 
\vspace{2cm}

\end{textbox}
%------------------|

%-------------------------- start a block

\begin{textbox}(Vision)
at (Motivation.south east) [below left] <1,.47\textwidth>
title {Vision}

\noindent I'm envision a system that switches mode (heavy vs. light state saving) based on predicting the amount of the harvested energy.

  \centering
    \adjincludegraphics[width=\columnwidth]{./fig/virtualTaskSize}

  \vspace{-1cm}

  \justify
  \noindent\textbf{diminishing returns}

\end{textbox}
%------------------|

\end{tikzpicture}

\end{document}
% vim: tw=80:ts=2:sts=2:sw=2:et:fdm=marker:fmr=[[[,]]]

### 2017-11-30 DMA/ipos-gcc-dma/include/contFlags.h

@@ -1,0 +1,46 @@
+/*
+ * contFalgs.h
+ *
+ *  Created on: 3 Nov 2017
+ *      Author: amjad
+ */
+
+#ifndef CONTFLAGS_H_
+#define CONTFLAGS_H_
+
+
+/****************************************
+        Paging control flags
+****************************************/
+#define PAG_FAULT_CNTR 1
+
+
+/****************************************
+        Coalescing control flags
+****************************************/
+
+// PLEASE SELECT AN ALGORITHIN
+#define NO_COALESCING 0
+#define SLO_CHNG_ALGO 0
+#define FST_CHNG_ALGO 0
+#define FST_CHNG_TSK_AWAR_ALGO 1
+
+// algorithm selection flags control
+#if     FST_CHNG_ALGO && SLO_CHNG_ALGO || \
+        FST_CHNG_ALGO && FST_CHNG_TSK_AWAR_ALGO || \
+        FST_CHNG_ALGO && NO_COALESCING || \
+        SLO_CHNG_ALGO && FST_CHNG_TSK_AWAR_ALGO || \
+        SLO_CHNG_ALGO && NO_COALESCING || \
+        FST_CHNG_TSK_AWAR_ALGO && NO_COALESCING
+#error "More than one algorithm is enabled"
+#endif
+
+#if FST_CHNG_ALGO ==0  && SLO_CHNG_ALGO ==0 && FST_CHNG_TSK_AWAR_ALGO ==0  && NO_COALESCING ==0
+#error "No coalescing algorithm is selected"
+#endif
+
+// Logging
+#define COAL_TSK_SIZ_SEND 1
+#define NUM_PWR_INTR__MEM_DUMP 0
+
+#endif /* CONTFLAGS_H_ */

### 2017-11-30 DMA/ipos-gcc-dma/include/dataProtec.h

@@ -21,49 +21,25 @@
  */
 
 
-uint8_t* __return_addr(uint8_t* var) ;
-uint8_t* __return_addr_wr(uint8_t* var) ;
-uint8_t* __return_addr_no_check(unsigned int* var) ;
-uint8_t* __return_addr_wr_no_check(unsigned int* var) ;
-void __sendPagTemp(unsigned int pagTag);
-void __bringPagTemp(unsigned int pagTag);
-void __bringPagROM(unsigned int pagTag);
-void __sendPagROM(unsigned int pagTag);
-unsigned int __pageSwap(unsigned int * varAddr);
-void __pagesSwap(unsigned int * CurnPag, unsigned int * varAddr);
-void __bringPersisCrntPag(unsigned int curntPag);
+
+uint8_t* __return_addr_no_check(uint8_t*var) ;
+uint8_t* __return_addr_wr_no_check(uint8_t* var) ;
+void __sendPagTemp(unsigned int from, unsigned int to);
+void __movPag(unsigned int from, unsigned int to);
+void __pageSwap(uint8_t * varAddr, uint8_t * dirtyPag, unsigned int *curtPagHdr, unsigned int ramPagsBuf  );
+
+
 void __pagsCommit();
-void __bringCrntPagROM();
 
 // Memory access interface
 
-#define DIRTY_PAGE  0      // make it 0 to commit only the dirty pages [This macro does not work with P() ]
+#define CLR_DIRTY_PAGE  0
 
-#define __VAR_ADDR(var)                 ((unsigned int) (&(var)) )
+//#define __VAR_ADDR(var)                 ((unsigned int) (&(var)) )
 
-#define __IS_VAR_IN_CRNT_PAG(var)       ( ( __VAR_ADDR(var)  >= CrntPagHeader ) && \
-                                        ( __VAR_ADDR(var)  <  (CrntPagHeader+PAG_SIZE) ))
+#define __IS_VAR_IN_CRNT_PAG(var, ramPagsBuf)       ( ( ((unsigned int)var)  >= ramPagsBuf.crntPagHdr ) && ( ((unsigned int)var)   <  (ramPagsBuf.crntPagHdr+PAG_SIZE) ))
 
-#define __VAR_PT_IN_RAM(var)            (  (__typeof__(var)*) (  (__VAR_ADDR(var) - CrntPagHeader) + RAM_PAG )  )
-
-
-//#define WVAR(var, val)   if( __IS_VAR_IN_CRNT_PAG(var) )\
-//                                { \
[…]

### 2017-11-30 DMA/ipos-gcc-dma/include/global.h

@@ -13,37 +13,43 @@
 // Special memory locaitons
 
 #define END_ROM         0xFF70   //  using the address 0xFF7F disables DMA transfer
-#define END_RAM         0x2200
+#define BGN_RAM         0x1F00
 #define APP_MEM         (8 * 1024)  // 12 KB [Linker script might need to be adjusted]
-#define PAG_SIZE        128   // 64 128 256 512 1024
-#define NUM_PAG         (APP_MEM/PAG_SIZE)  // 1KB
-#define RAM_PAG         (END_RAM - PAG_SIZE)
-//#define TOT_PAG_SIZE    (PAG_SIZE * NUM_PAG)
+#define RAM_BUF_SIZE    0x400  // 1024 bytes
+#define PAG_SIZE        256   // 64 128 256 512 1024
+#define RAM_BUF_LEN     (RAM_BUF_SIZE/PAG_SIZE)
+#define NUM_PRS_PAGS    (APP_MEM/PAG_SIZE)  // 1KB
 #define PAG_SIZE_W      (PAG_SIZE/2)  //1KB
-#define BIGEN_ROM       ( (END_ROM - APP_MEM) - APP_MEM  ) // 0xBF70
-
-//#define TASKS_STRUC     0x400  // 1KB
-//#define LIST_HEAD       (BIGEN_ROM - TASKS_STRUC)           //2 bytes
+#define BGN_ROM         ( (END_ROM - APP_MEM) - APP_MEM  ) // 0xBF70
+#define LIST_HEAD       (BIGEN_ROM - TASKS_STRUC)           //2 bytes
 
 // function pointer
 typedef void (* funcPt)(void);
-
-extern unsigned int __coalTskAddr;
-extern volatile unsigned int  __totNumTask;
-extern unsigned int __persis_CrntPagHeader;
-extern unsigned int __pagsInTemp[];
-extern unsigned int __persis_pagsInTemp[];
-extern volatile uint8_t dirtyPag ;
-extern uint32_t __temp_temp;
-extern unsigned int CrntPagHeader;  // Holds the address of the first byte of a page
-extern unsigned int __coalTaskCntr;
-
 //This is a task interface. It is shared between the user and IPOS
 typedef struct _taskId{
   funcPt * func;
   uint8_t id;
   uint8_t riskLevel;
 }taskId;
+
+typedef struct {
+    unsigned int ramPagAddr;
+    uint8_t dirtyPag;
+    unsigned int crntPagHdr;
[…]

### 2017-11-30 DMA/ipos-gcc-dma/ipos/dataProtec.c

@@ -1,51 +1,34 @@
-
-
 #include "dataProtec.h"
-/*####################################
-              Paging
-#####################################*/
+#include "contFlags.h"
+//#include <mspDebugger.h>
 
-     volatile uint8_t dirtyPag = DIRTY_PAGE;
      uint32_t __temp_temp=0;
      // TODO THIS ARRAY CONSUME TOO MUCH OF THE SRAM
-     unsigned int __pagsInTemp[NUM_PAG] = {0};  // This can be pretty big and consume large section of the SRAM, therefore I shift it to FRAM
-__nv unsigned int __persis_pagsInTemp[NUM_PAG ] = {0};
+     unsigned int __pagsInTemp[NUM_PRS_PAGS] = {0};  // This can be pretty big and consume large section of the SRAM, therefore I shift it to FRAM
+__nv unsigned int __persis_pagsInTemp[NUM_PRS_PAGS] = {0};
 
-     unsigned int CrntPagHeader = BIGEN_ROM;
-__nv unsigned int __persis_CrntPagHeader = BIGEN_ROM;
+
+    volatile unsigned int swapIndex = 0;
+    ramPagMeta ramPagsBuf[RAM_BUF_LEN] = {{BGN_RAM , CLR_DIRTY_PAGE,  BGN_ROM }, 0};
+//__nv unsigned int __prs_crntPagHdr = BGN_ROM;
 
 __nv unsigned int __pageFaultCounter = 0;
+__nv unsigned int __fullpageFaultCounter = 0;
+__nv unsigned int __cntr=0;   // to enable incremental commit
 
-/*
- * Bring a page to its ROM buffer
- */
-// void __bringCrntPagROM()
-// {
-//     // Configure DMA channel 1
-//     __data16_write_addr((unsigned short) &DMA0SA,(unsigned long) ( __persis_CrntPagHeader) );
-//                                             // Source block address
-//     __data16_write_addr((unsigned short) &DMA0DA,(unsigned long) RAM_PAG );
-//                                             // Destination single address
-//     DMA0SZ = PAG_SIZE_W;                            // Block size
-//     DMA0CTL = DMADT_5 | DMASRCINCR_3 | DMADSTINCR_3; // Rpt, inc
-//     DMA0CTL |= DMAEN;                         // Enable DMA0
-
-//     DMA0CTL |= DMAREQ;                      // Trigger block transfer
-
[…]

### 2017-11-30 DMA/ipos-gcc-dma/ipos/pTCB.c

@@ -45,7 +45,7 @@ void os_addTasks(unsigned char numTasks, taskId tasks[]){
     if( funcBlocker != 0xAD)
     {
         // each task needs 6 byte of meta data.
-        unsigned int __temp_head =  ((BIGEN_ROM - numTasks * 6) - 2);
+        unsigned int __temp_head =  ((BGN_ROM - numTasks * 6) - 2);
         __head =  (unsigned int*)  __temp_head;
         unsigned char i = 0;
         unsigned int cnt=0;

### 2017-11-30 DMA/ipos-gcc-dma/ipos/psched.c

@@ -3,8 +3,8 @@
 #include <mspDebugger.h>
 
 // PLEASE SELECT AN ALGORITHIN
-#define NO_COALESCING 0
-#define SLO_CHNG_ALGO 1
+#define NO_COALESCING 1
+#define SLO_CHNG_ALGO 0
 #define FST_CHNG_ALGO 0
 #define FST_CHNG_TSK_AWAR_ALGO 0
 
@@ -99,16 +99,24 @@ void os_scheduler()
 
     if (__commit_flag == COMMITTING)
     {
-        // firm transition
+        __pagsCommit(); // Commit the dirty pages
         __coalTskAddr = (unsigned int) __temp_coalTskAddr;
-
-        // Commit the dirty pages
-        __pagsCommit();
         __commit_flag = COMMIT_FINISH;
     }
 
+    // initialize ram pages buffer
+
+    unsigned int __rb_cnt;
+    unsigned int __ramAddr = BGN_RAM;
+    for(__rb_cnt=1; __rb_cnt < RAM_BUF_LEN; __rb_cnt++)
+    {
+        __ramAddr += PAG_SIZE;
+        ramPagsBuf[__rb_cnt].ramPagAddr = __ramAddr;
+    }
+
     // Recover the virtual state
-    __bringPagROM(__persis_CrntPagHeader);
+    __movPag( BGN_ROM , BGN_RAM);
+
     __realTask = (unsigned int *) __coalTskAddr;
 
 
@@ -122,24 +130,36 @@ void os_scheduler()
 
         // transition stage (from virtual -> persistent)
         __temp_coalTskAddr = __realTask;
-        __sendPagTemp(CrntPagHeader);
 
-        unsigned int p_cnt;
-        for (p_cnt=0; __pagsInTemp[p_cnt]; p_cnt++)
+        // moving the dirty pages to PagTempBuffer 
+        unsigned int _wp_cntr;
+
+        for(_wp_cntr = 0; (_wp_cntr < RAM_BUF_LEN)  && ramPagsBuf[_wp_cntr].crntPagHdr ; _wp_cntr++)
         {
-            //move the dirty pages addresses to a persistent buffer
-            __persis_pagsInTemp[p_cnt] = __pagsInTemp[p_cnt];
+            if(ramPagsBuf[_wp_cntr].dirtyPag)
+            {
[…]

### 2017-11-30 IPE_Poster/data/energy_distance/add_50cm_WISP.m

@@ -1,0 +1,74 @@
+#!/usr/bin/env octave
+% Shift measured WISP values by loss level corresponding to 50cm free-air
+
+% Reference: [1] TODO.
+%
+% Code development: Ivar in 't Veen (I.J.G.intVeen@student.tudelft.nl)
+%
+% Last update: November 1, 2015
+
+
+% Copyright (c) 2015, Embedded Software Group,
+% Delft University of Technology, The Netherlands. All rights reserved.
+%
+% Redistribution and use in source and binary forms, with or without
+% modification, are permitted provided that the following conditions
+% are met:
+%
+% 1. Redistributions of source code must retain the above copyright notice,
+% this list of conditions and the following disclaimer.
+%
+% 2. Redistributions in binary form must reproduce the above copyright
+% notice, this list of conditions and the following disclaimer in the
+% documentation and/or other materials provided with the distribution.
+%
+% 3. Neither the name of the copyright holder nor the names of its
+% contributors may be used to endorse or promote products derived from this
+% software without specific prior written permission.
+%
+% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
+% "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
+% LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A
+% PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
+% HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
+% SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED
+% TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR
+% PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
+% LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING
[…]

### 2017-11-30 IPE_Poster/data/energy_distance/generate_theoretical.m

@@ -1,0 +1,81 @@
+#!/usr/bin/env octave
+% Generate theoretical energy/byte values
+
+% Reference: [1] TODO.
+%
+% Code development: Ivar in 't Veen (I.J.G.intVeen@student.tudelft.nl)
+%
+% Last update: November 1, 2015
+
+
+% Copyright (c) 2015, Embedded Software Group,
+% Delft University of Technology, The Netherlands. All rights reserved.
+%
+% Redistribution and use in source and binary forms, with or without
+% modification, are permitted provided that the following conditions
+% are met:
+%
+% 1. Redistributions of source code must retain the above copyright notice,
+% this list of conditions and the following disclaimer.
+%
+% 2. Redistributions in binary form must reproduce the above copyright
+% notice, this list of conditions and the following disclaimer in the
+% documentation and/or other materials provided with the distribution.
+%
+% 3. Neither the name of the copyright holder nor the names of its
+% contributors may be used to endorse or promote products derived from this
+% software without specific prior written permission.
+%
+% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS
+% "AS IS" AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT
+% LIMITED TO, THE IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A
+% PARTICULAR PURPOSE ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT
+% HOLDER OR CONTRIBUTORS BE LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL,
+% SPECIAL, EXEMPLARY, OR CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED
+% TO, PROCUREMENT OF SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR
+% PROFITS; OR BUSINESS INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF
+% LIABILITY, WHETHER IN CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING
[…]

### 2017-11-30 apps/DFT/main.c

@@ -252,7 +252,7 @@ void init()
     P3OUT &= ~BIT5;
     P3DIR |=BIT5;
 
-#if 0
+#if 1
     CSCTL0_H = CSKEY >> 8;                // Unlock CS registers
 //    CSCTL1 = DCOFSEL_4 |  DCORSEL;                   // Set DCO to 16MHz
     CSCTL1 = DCOFSEL_6;                   // Set DCO to 8MHz
@@ -270,7 +270,7 @@ void init()
 #endif
 
 #ifdef AUTO_RST
-    mr_auto_rand_reseter(13000); // every 12 msec the MCU will be reseted
+    mr_auto_rand_reseter(50000); // every 12 msec the MCU will be reseted
 #endif
 
 }
@@ -303,12 +303,12 @@ int main(void) {
     // uart_sendText("Start\n\r", 7);
 
 
-       taskId tasks[] = {  {discTimeSign,1},
-                           {dft_outer_loop,2},
-                           {dft_real,3 },
-                           {dft_im, 4},
-                           {dft_power, 5},
-                           {dft_end,6}
+       taskId tasks[] = {  {discTimeSign,1, 4},
+                           {dft_outer_loop,2, 1},
+                           {dft_real,3, 5 },
+                           {dft_im, 4, 5},
+                           {dft_power, 5, 1},
+                           {dft_end,6, 2}
        };
        //This function should be called only once
        os_addTasks(6, tasks );

### 2017-11-30 apps/ar/main.c

@@ -1,0 +1,602 @@
+#include <msp430.h>
+#include <stdint.h>
+#include <stdbool.h>
+#include <stdlib.h>
+
+#include <mspReseter.h>
+#include "mspProfiler.h"
+#include "mspDebugger.h"
+
+#include <ipos.h>
+#include <accel.h>
+
+#include <msp-math.h>
+
+#define TSK_SIZ
+//#define AUTO_RST
+//#define LOG_INFO
+//#define RAISE_PIN 1
+
+// Number of samples to discard before recording training set
+#define NUM_WARMUP_SAMPLES 3
+
+#define ACCEL_WINDOW_SIZE 3
+#define MODEL_SIZE 16
+#define SAMPLE_NOISE_FLOOR 10 // TODO: made up value
+
+// Number of classifications to complete in one experiment
+#define SAMPLES_TO_COLLECT 128
+
+#define TASK_NUM 11
+#define OFFSET(src, dest) src <= dest ? dest - src : TASK_NUM + dest - src
+
+unsigned volatile *timer = &TBCTL;
+typedef threeAxis_t_8 accelReading;
+typedef accelReading accelWindow[ACCEL_WINDOW_SIZE];
+
+typedef struct {
+    unsigned meanmag;
+    unsigned stddevmag;
+} features_t;
+
+typedef enum {
+    CLASS_STATIONARY,
+    CLASS_MOVING,
+} class_t;
+
+
+typedef enum {
+    MODE_IDLE = 3,
+    MODE_TRAIN_STATIONARY = 2,
+    MODE_TRAIN_MOVING = 1,
+    MODE_RECOGNIZE = 0, // default
+} run_mode_t;
+
+enum task_index {
+    t_init,
+    t_selectMode,
+    t_resetStats,
+    t_sample,
+    t_transform,
+    t_featurize,
+    t_classify,
+    t_stats,
+    t_warmup,
+    t_train,
+    t_idle
+};
+
+void task_init();
+void task_selectMode();
+void task_resetStats();
+void task_sample();
+void task_transform();
+void task_featurize();
+void task_classify();
+void task_stats();
+void task_warmup();
+void task_train();
+void task_idle();
+
+__p uint16_t _v_pinState;
+__p unsigned _v_discardedSamplesCount;
+__p run_mode_t _v_class;
+__p unsigned _v_totalCount;
+__p unsigned _v_movingCount;
+__p unsigned _v_stationaryCount;
[…]

### 2017-11-30 apps/bc/main.c

@@ -6,10 +6,10 @@
 #include <ipos.h>
 
 
-//#define TSK_SIZ
+#define TSK_SIZ
 //#define AUTO_RST
 //#define LOG_INFO
-
+//#define RAISE_PIN 1
 
 #define SEED 4L
 #define ITER 100
@@ -311,10 +311,10 @@ void task_AR_btbl_bitcount() {
     unsigned char * Ptr = (unsigned char *) &RP(_v_seed) ;
     int Accu ;
 
-    Accu  = bits[ RP(*Ptr++) ];
-    Accu += bits[ RP(*Ptr++) ];
-    Accu += bits[ RP(*Ptr++) ];
-    Accu += bits[ RP(*Ptr) ];
+    Accu  = bits[ *Ptr++ ];
+    Accu += bits[ *Ptr++ ];
+    Accu += bits[ *Ptr++ ];
+    Accu += bits[ *Ptr ];
     WP(_v_n_5)+= Accu;
     uint32_t tmp_seed = RP(_v_seed);
     WP(_v_seed) = tmp_seed + 13;
@@ -373,10 +373,13 @@ void task_end() {
         P3OUT &=~BIT5;
     }
     pinCont=0;
+    PAGCMT();
 
 #ifdef TSK_SIZ
      cp_sendRes("task_end \0");
 #endif
+
+     while(1);
 }
 
 void init() {
@@ -384,7 +387,10 @@ void init() {
     WDTCTL = WDTPW | WDTHOLD;   // Stop watchdog timer
     PM5CTL0 &= ~LOCKLPM5;       // Lock LPM5.
 
-    P3DIR = BIT5;
+#if RAISE_PIN
+  P3OUT &=~BIT5;
+  P3DIR |=BIT5;
+#endif
 
 #if 0
     CSCTL0_H = CSKEY >> 8;                // Unlock CS registers
@@ -396,6 +402,7 @@ void init() {
 #endif
 
 #ifdef TSK_SIZ
+    uart_init();
     cp_init();
 #endif
 
@@ -404,22 +411,22 @@ void init() {
 #endif
 
 #ifdef AUTO_RST
-    mr_auto_rand_reseter(13000); // every 12 msec the MCU will be reseted
+    mr_auto_rand_reseter(25000); // every 12 msec the MCU will be reseted
 #endif
 
 }
 
 int main(void) {
     init();
-    taskId tasks[] = {{task_init,        1, 1},
-        {task_select_func,               2, 1},
[…]

### 2017-11-30 apps/cem_delft/main.c

@@ -5,7 +5,7 @@
 #include <ipos.h>
 
 //#define TSK_SIZ
-//#define AUTO_RST
+#define AUTO_RST
 //#define LOG_INFO
 
 
@@ -49,8 +49,6 @@ void task_print();
 void task_done();
 
 
-
-__p node_t _v_dict[DICT_SIZE];
 __p letter_t _v_letter;
 __p unsigned _v_letter_idx;
 __p sample_t _v_prev_sample;
@@ -66,6 +64,8 @@ __p node_t _v_parent_node;
 __p node_t _v_sibling_node;
 __p index_t _v_symbol;
 __p node_t _v_compressed_data[BLOCK_SIZE];
+__p node_t _v_dict[DICT_SIZE];
+
 
 static sample_t acquire_sample(letter_t prev_sample)
 {
@@ -101,19 +101,23 @@ void task_init_dict()
     cp_reset();
 #endif
 
-    int i = RP(_v_letter);
+    uint16_t i = RP(_v_letter);
+
+    // uart_sendStr("_v_letter \0");
+    //     uart_sendHex16(_v_letter);
+    // uart_sendStr("\n\r \0");
+
     WP(_v_dict[i].letter) = i ;
     WP(_v_dict[i].sibling) =  NIL;
     WP(_v_dict[i].child) = NIL;
-    i++;
-    WP(_v_letter) = i;
+    WP(_v_letter)++;
     if (i < NUM_LETTERS) {
         os_jump(0);
     } else {
         WP(_v_node_count) =  NUM_LETTERS;
 //        os_jump(1);
-        ;
     }
+
 #ifdef TSK_SIZ
     cp_sendRes("task_init_dict \0");
 #endif
@@ -125,17 +129,21 @@ void task_sample()
     cp_reset();
 #endif
 
+    // uart_sendStr("task_sample \0");
+    // uart_sendStr("\n\r \0");
+
     unsigned next_letter_idx = RP(_v_letter_idx) + 1;
     if (next_letter_idx == NUM_LETTERS_IN_SAMPLE)
         next_letter_idx = 0;
 
-    WP(_v_letter_idx) = next_letter_idx;
     if (RP(_v_letter_idx) == 0) {
-//        os_jump(1);
-        ;
+        WP(_v_letter_idx) = next_letter_idx;
+        // os_jump(1);
     } else {
+        WP(_v_letter_idx) = next_letter_idx;
         os_jump(2);
     }
+
 #ifdef TSK_SIZ
[…]

### 2017-11-30 apps/cuckoo_delft/main.c

@@ -11,7 +11,7 @@
 #include <ipos.h>
 
 
-//#define TSK_SIZ
+#define TSK_SIZ
 //#define AUTO_RST
 //#define LOG_INFO
 
@@ -124,6 +124,10 @@ unsigned i;
 
 void task_init()
 {
+
+        P3OUT |= BIT5;
+        P3OUT &= ~BIT5;
+
 #ifdef TSK_SIZ
        cp_reset();
 #endif
@@ -450,9 +454,11 @@ void task_lookup_done()
 
 void task_print_stats()
 {
-//    __no_operation();
-    pinRaised=1;
 
+        P3OUT |= BIT5;
+        P3OUT &= ~BIT5;
+
+        PAGCMT();
 }
 
 void task_done()
@@ -461,15 +467,13 @@ void task_done()
        cp_reset();
 #endif
 
-    if(pinRaised){
-        P3OUT |= BIT5;
-        P3OUT &= ~BIT5;
-    }
-    pinRaised=0;
 
 #ifdef TSK_SIZ
      cp_sendRes("task_done \0");
 #endif
+
+     while(1);
+
 }
 
 
@@ -481,8 +485,6 @@ void init()
     P3OUT &= ~BIT5;
     P3DIR |=BIT5;
 
-    srand(0);
-
 #if 0
     CSCTL0_H = CSKEY >> 8;                // Unlock CS registers
 //    CSCTL1 = DCOFSEL_4 |  DCORSEL;      // Set DCO to 16MHz
@@ -493,6 +495,7 @@ void init()
 #endif
 
 #ifdef TSK_SIZ
+    uart_init();
     cp_init();
 #endif
 
@@ -504,11 +507,14 @@ void init()
     mr_auto_rand_reseter(13000); // every 12 msec the MCU will be reseted
 #endif
 
+    srand(0);
+
 }
+
 int main(void) {
     init();
 
-    taskId tasks[] = {{task_init,       1,1 },
+    taskId tasks[] = {{task_init,       1,5 },
         {task_init_array,               2,1} ,
         {task_generate_key,             3,1 },
         {task_calc_indexes,             4,1 },

### 2017-11-30 apps/rsa/main.c

@@ -80,7 +80,7 @@ void init()
     P3OUT &=~BIT5;
     P3DIR |=BIT5;
 
-#if 0
+#if 1
     CSCTL0_H = CSKEY >> 8;                // Unlock CS registers
     //    CSCTL1 = DCOFSEL_4 |  DCORSEL;                   // Set DCO to 16MHz
     CSCTL1 = DCOFSEL_6;                   // Set DCO to 8MHz
@@ -98,7 +98,7 @@ void init()
 #endif
 
 #ifdef AUTO_RST
-    mr_auto_rand_reseter(13000); // every 12 msec the MCU will be reseted
+    mr_auto_rand_reseter(25000); // every 12 msec the MCU will be reseted
 #endif
 
 }
@@ -414,6 +414,7 @@ void decrypt_print()
         P3OUT &=~BIT5;
     }
     protect=0;
+    PAGCMT();
 
     __enable_interrupt();
 

### 2017-11-30 apps/sortAlgo/main.c

@@ -48,6 +48,7 @@ __nv unsigned protect = 0;
 unsigned int in_i, in_j, arr_i, arr_j;
 void task_inner_loop()
 {
+
 #if TSK_SIZ
     cp_reset();
 #endif
@@ -126,6 +127,10 @@ void task_finish()
 #if TSK_SIZ
     cp_sendRes("\ntask_finish \0");
 #endif
+
+//cp_sendRes("\sort_100 \0");
+
+while(1);
 }
 
 void init()
@@ -139,9 +144,20 @@ void init()
   P3DIR |=BIT5;
 #endif
 
+#if 1
+    CSCTL0_H = CSKEY >> 8;                // Unlock CS registers
+    //    CSCTL1 = DCOFSEL_4 |  DCORSEL;                   // Set DCO to 16MHz
+    CSCTL1 = DCOFSEL_6;                   // Set DCO to 8MHz
+    CSCTL2 =  SELM__DCOCLK;               // MCLK = DCO
+    CSCTL3 = DIVM__1;                     // divide the DCO frequency by 1
+    CSCTL0_H = 0;
+#endif
+
 #ifdef TSK_SIZ
     cp_init();
 #endif
+
+    uart_init();
 
 #ifdef LOG_INFO
     uart_init();

### 2017-11-30 develpment_workspace/paging/dataProtec.c

@@ -16,8 +16,8 @@
 unsigned int __pagsInTemp[NUM_PAG] = {0};
 __nv unsigned int __persis_pagsInTemp[NUM_PAG] = {0};
 
-unsigned int CrntPagHeader = BIGEN_ROM;
-__nv unsigned int __persis_CrntPagHeader = BIGEN_ROM;
+unsigned int CrntPagHeader = BGN_ROM;
+__nv unsigned int __persis_CrntPagHeader = BGN_ROM;
 __nv uint8_t pageCommit = 0;
 
 uint16_t dirtyPag=0;
@@ -61,7 +61,7 @@ void __sendPagTemp(unsigned int pagHeader)
 
 
     // Find the page index
-    unsigned int pageGuess = pagHeader - BIGEN_ROM ; // possible values are 0, 0x400, 0x800, 0xc00 ...
+    unsigned int pageGuess = pagHeader - BGN_ROM ; // possible values are 0, 0x400, 0x800, 0xc00 ...
     uint8_t idx=0;
     unsigned int __pageSizes = PAG_SIZE;
     while( pageGuess >= __pageSizes )
@@ -139,7 +139,7 @@ unsigned int __pageSwap(unsigned int * varAddr)
     //2// Find the requested page 
     unsigned int ReqPagTag;
     unsigned int ReqPagTag_dirty = (unsigned int) varAddr;
-    unsigned int __temp_pagSize = BIGEN_ROM+PAG_SIZE ; // the upper limit of the first page
+    unsigned int __temp_pagSize = BGN_ROM+PAG_SIZE ; // the upper limit of the first page
     // TODO we are not checking if the var is not in any page !
     uint8_t idx=0;
     while( ! (ReqPagTag_dirty <= __temp_pagSize) )  // if the var is not with the page
@@ -209,7 +209,7 @@ unsigned int __pageSwap_w(unsigned int * varAddr)
     //2// Find the requested page
     unsigned int ReqPagTag;
     unsigned int ReqPagTag_dirty = (unsigned int) varAddr;
-    unsigned int __temp_pagSize = BIGEN_ROM+PAG_SIZE ; // the upper limit of the first page
+    unsigned int __temp_pagSize = BGN_ROM+PAG_SIZE ; // the upper limit of the first page
     // TODO we are not checking if the var is not in any page !
     unsigned char idx=0;
 

### 2017-11-30 develpment_workspace/paging/dataProtec.h

@@ -25,7 +25,7 @@
  */
 
 
-// The location of any page is given by (BIGEN_ROM + pag_tag - PAG_SIZE)
+// The location of any page is given by (BGN_ROM + pag_tag - PAG_SIZE)
 // The TEMP location of any page is given by ( END_ROM - TOT_PAG_SIZE + pag_tag - PAG_SIZE )
 #define END_ROM         0xFF70    //  using the address 0xFF7F disables DMA transfer
 #define END_RAM         0x2100
@@ -42,7 +42,7 @@
 #define RAM_PAG         (END_RAM - PAG_SIZE)
 #define TOT_PAG_SIZE    (PAG_SIZE * NUM_PAG)
 #define PAG_SIZE_W      (PAG_SIZE/2)  //1KB
-#define BIGEN_ROM       ( (END_ROM - TOT_PAG_SIZE) - TOT_PAG_SIZE  ) // 0xBF70
+#define BGN_ROM       ( (END_ROM - TOT_PAG_SIZE) - TOT_PAG_SIZE  ) // 0xBF70
 
 
 void __sendPagTemp(unsigned int pagTag);

### 2017-11-30 experiments/Paper_data/CoalaChain_Clang/CoalaChain.py

@@ -35,7 +35,7 @@ def figureSetting():
 
     #set size of markers, e.g., circles representing points
     #set numpoints for legend
-    plt.rcParams['legend.numpoints'] = 1
+    # plt.rcParams['legend.numpoints'] = 1
 
 
 chain_cont   = [0.355809,    0.091599,    1.651908,   5.771600664062499  , 0.309126,    0.278821,    0.600697]
@@ -60,7 +60,7 @@ data = [ np.transpose([chain_cont, coal_cont]),
 # print(data)
 # exit()         
 
-f = plt.figure(figsize=(8,2.6))
+f = plt.figure(figsize=(8,2.5))
 figureSetting()    
 
 s=0
@@ -72,11 +72,21 @@ for i in range(1,8):
     # print(np.arange(6) + s)
     # exit()
 
-    [us, co, _,__,___,____]= plt.bar( np.arange(6) + s, data , 
+    [us, co, _,__,___,____]= plt.bar( np.arange(6)+0.5 + s, data , 
             width=1, 
             color=['#addd8e','#31a354', '#addd8e','#31a354', '#addd8e','#31a354'], 
              linewidth=0.5)
     s+=6.5
+
+# add text at the bottom of the bars 
+
+pos=np.array([1,3,5])-0.3
+lbl = ['0m', '.2m','.4m' ]
+d=0
+for h in range(7):
+    for i in range( 3 ):
+        plt.text( pos[i]+d , 0.1 , lbl[i] , fontsize=9, color='#404040', rotation=90, verticalalignment='bottom')
+    d+=6.5
 
 plt.legend([us, co], ["Coala", "Chain"], loc='upper left')
 
@@ -86,7 +96,7 @@ ax.xaxis.set_major_formatter(ticker.FixedFormatter( apps ))
 
 
 # exit()
-plt.yticks([1,2,3])
+plt.yticks([1,2,3,4])
 plt.xlim(0,45)
 plt.ylabel("Norm. runtime")
 plt.tight_layout()

### 2017-11-30 experiments/Paper_data/compiler_user/CompUser.py

@@ -1,123 +1,0 @@
-"""
-    IPOS: performance 
-    Metric : time
-    Task   : execute certain apps
-
-
-@ Author  : Amjad Yousef Majid
-@ Gmail ID: amjad.y.majid 
-@ Date    : 30/June/2017 
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.ticker as ticker
-
-def executionTimeCuckoo(fh):
-    """
-    Input: file handler 
-    return: the average execution time of a program
-    """
-    lim=0
-    numbers = []
-    for num in fh:
-        if num[-2:-1] == '1' and lim < 10:
-            lim+=1;
-            #print(num)
-            numbers.append( float(num.split(',')[0] ) )
-    aDiff = 0
-    for i in range(0, len(numbers)-1, 2):
-        aDiff += (numbers[i+1] - numbers[i])
-#     aDiff *=1000  # make the diff in milliseconds
-
-    time = (aDiff / int(len(numbers)/2) )
-    fh.close()
-    return time
-
-def executionTime(fh):
-    """
-    Input: file handler 
-    return: the average execution time of a program
-    """
-    numbers = []
-    for num in fh:
-        if num[-2:-1] == '1':
-            #print(num)
-            numbers.append( float(num.split(',')[0] ) )
-    aDiff = 0
-    for i in range(0, len(numbers)-1, 2):
-        aDiff += (numbers[i+1] - numbers[i])
-#     aDiff *=1000  # make the diff in milliseconds
-
-    time = (aDiff / int(len(numbers)/2) )
-    fh.close()
-    return time
-
-def figureSetting():
-    """
-    Set a figure params
-    """
-    #set figure linewidth
-    plt.rcParams['axes.linewidth']=2
-
-    #set font size for labels on axes
-    plt.rcParams['axes.labelsize'] = 16
-
-    #set size of numbers on x-axis
-    plt.rcParams['xtick.labelsize'] = 16
-    #set size of numbers on y-axis
-    plt.rcParams['ytick.labelsize'] = 16
-
-    #set size of ticks on x-axis
-    plt.rcParams['xtick.major.size'] = 7
-    #set size of ticks on y-axis
-    plt.rcParams['ytick.major.size'] = 7
-
[…]

### 2017-11-30 experiments/newExperiments/scripts/CoalaChain.py

@@ -1,0 +1,95 @@
+"""
+    IPOS: performance 
+    Metric : time
+    Task   : execute certain apps
+
+
+@ Author  : Amjad Yousef Majid
+@ Gmail ID: amjad.y.majid 
+@ Date    : 30/June/2017 
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+from csf import *
+
+figureSetting()
+
+
+
+chain_cont   = [0.355809,    0.091599,    1.651908,   5.771600664062499  , 0.309126,    0.278821,    0.600697]
+coal_cont    = [0.56662,     0.049746,    1.213208,    3.5161449166666663 , 0.235423,    0.089789,   0.137456]
+              
+chain_20   = [0.499513513513513,   0.129282608695652,  2.3233125,  5.772032160714283 , 0.43731914893617,    0.394844444444445,   0.839388888888889]
+coal_20    = [0.805625,            0.0685652173913044,  1.70852173913043,    3.517230833333335  , 0.345225,   0.121641025641026,   0.193191489361702]
+                        
+chain_40   = [1.37189473684211,    0.289142857142857,   4.58725 ,  8.892858390624998 ,             0.976735294117647 ,  0.914209302325582 ,  1.55934146341463]
+coal_40    = [1.68121621621622,   0.172515873015873,   2.46094117647059 , 7.607710203125,   0.785947368421053   , 0.31355       ,      0.42868085106383]
+
+    
+distances = ["cont.csv",'20cm.csv', '40cm.csv']
+
+apps    =   ['cem', 'cuckoo', 'ar',  'dft', 'rsa', 'bc', 'sort']
+
+
+data = [ np.transpose([chain_cont, coal_cont]), 
+         np.transpose([chain_20, coal_20    ]), 
+         np.transpose([chain_40, coal_40 ]) ]
+
+# print(data)
+# exit()         
+
+f = plt.figure(figsize=(8,2.5))
+figureSetting()    
+
+s=0
+for i in range(1,8):
+
+    data = [ coal_cont[i-1]/coal_cont[i-1], chain_cont[i-1]/coal_cont[i-1], 
+            coal_20[i-1]/coal_20[i-1], chain_20[i-1]/coal_20[i-1], 
[…]

### 2017-11-30 experiments/newExperiments/scripts/CompUser.py

@@ -1,0 +1,123 @@
+"""
+    IPOS: performance 
+    Metric : time
+    Task   : execute certain apps
+
+
+@ Author  : Amjad Yousef Majid
+@ Gmail ID: amjad.y.majid 
+@ Date    : 30/June/2017 
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.ticker as ticker
+
+def executionTimeCuckoo(fh):
+    """
+    Input: file handler 
+    return: the average execution time of a program
+    """
+    lim=0
+    numbers = []
+    for num in fh:
+        if num[-2:-1] == '1' and lim < 10:
+            lim+=1;
+            #print(num)
+            numbers.append( float(num.split(',')[0] ) )
+    aDiff = 0
+    for i in range(0, len(numbers)-1, 2):
+        aDiff += (numbers[i+1] - numbers[i])
+#     aDiff *=1000  # make the diff in milliseconds
+
+    time = (aDiff / int(len(numbers)/2) )
+    fh.close()
+    return time
+
+def executionTime(fh):
+    """
+    Input: file handler 
+    return: the average execution time of a program
+    """
+    numbers = []
+    for num in fh:
+        if num[-2:-1] == '1':
+            #print(num)
+            numbers.append( float(num.split(',')[0] ) )
+    aDiff = 0
+    for i in range(0, len(numbers)-1, 2):
+        aDiff += (numbers[i+1] - numbers[i])
+#     aDiff *=1000  # make the diff in milliseconds
+
+    time = (aDiff / int(len(numbers)/2) )
+    fh.close()
+    return time
+
+def figureSetting():
+    """
+    Set a figure params
+    """
+    #set figure linewidth
+    plt.rcParams['axes.linewidth']=2
+
+    #set font size for labels on axes
+    plt.rcParams['axes.labelsize'] = 16
+
+    #set size of numbers on x-axis
+    plt.rcParams['xtick.labelsize'] = 16
+    #set size of numbers on y-axis
+    plt.rcParams['ytick.labelsize'] = 16
+
+    #set size of ticks on x-axis
+    plt.rcParams['xtick.major.size'] = 7
+    #set size of ticks on y-axis
+    plt.rcParams['ytick.major.size'] = 7
+
+    #set size of markers, e.g., circles representing points
[…]

### 2017-11-30 experiments/newExperiments/scripts/TimeToCompletion_dif_algos_blined-measurement.py

@@ -21,6 +21,9 @@ filesContainer = filesFinder(apps, 'simulatedPwrInter/TimeToCompletion/*/*.csv')
 # print(apps)
 # print('---------------')
 # print(filesContainer)
+
+# exit()
+
 names=[]
 times=[]
 for appFiles in filesContainer:
@@ -43,23 +46,25 @@ f = plt.figure(figsize=(8,2.5))
 s=0
 for i, row in enumerate(times):
 
-        [fst, slw, smt] = plt.bar( np.array([0,0.5,1]) +i*1.5+s, 
-            np.array( [row[2], row[0], row[1]] )/float(row[0]), 
-            width=0.5, 
+        [slw,fst, smt] = plt.bar( np.array([0,0.3,0.6]) +s, 
+            np.array( [row[1], row[0], row[2]] )/float(row[1]), 
+            width=0.3, 
             color=['#addd8e','#31a354','#61a654'], 
              linewidth=0.5)
-        s+=0.3
+        s+=1
 
 ax = plt.axes()
-ax.xaxis.set_major_locator(ticker.FixedLocator(  [0.75, 2.5 , 3.1, 4.4, 5.7, 7 ] ))
+ax.xaxis.set_major_locator(ticker.FixedLocator( np.arange(s)+0.5 ))
 ax.xaxis.set_major_formatter(ticker.FixedFormatter( names ))
 
 # plt.xlim(0,s-0.1)
-plt.ylabel("Norm. runtime")
+plt.ylabel("Norm. Runtime")
+plt.yticks([0.0, 0.4, 0.8, 1.2])
 plt.tight_layout()
-plt.legend([fst, slw, smt], ["Smart", "Fast", "Slow"], loc='best')
+plt.legend([smt, fst, slw], ["TSHC", "HC", "NONE"], loc='lower left',     fontsize= 12)
 # plt.ylim(0,3.2) 
-f.savefig("../figures/fast_slow_algos.pdf",format="pdf", dpi=1200)
+f.savefig("../figures/coalStrategies.pdf",format="pdf", dpi=1200)
+f.savefig("../figures/coalStrategies.eps",format="eps", dpi=1200)
 plt.show()
 
 

### 2017-11-30 experiments/newExperiments/scripts/difCap.py

@@ -1,0 +1,108 @@
+#! /usr/bin/python
+
+
+"""
+    IPOS: performance 
+    Metric : time
+    Task   : execute apps
+
+
+@ Author  : Amjad Yousef Majid
+@ Gmail ID: amjad.y.majid 
+@ Date    : 17/October/2017 
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+from csf import *
+
+apps =  appsSelector()
+filesContainer = filesFinder(apps, 'simulatedPwrInter/difCap/*/*.csv')
+
+print(apps)
+print('---------------')
+print(filesContainer)
+
+# exit()
+
+
+names=[]
+times=[]
+for appFiles in filesContainer:
+    appTimes=[]
+    for file in appFiles:
+        # l = labelMaker(file)
+        # name = nameMaker(file)
+        time = executionTime( open( '../data/'+file) )
+        appTimes.append(time)
+    # names.append(name)
+    times.append(appTimes)
+
+# print(times)
+
+# exit()
+
+coal = [0.8210784916666664, 0.8251448288043478,  0.8292796825,  0.8634430216346155]
+noCoal = [1.2018005568181822,  1.204097878472222, 1.2085183641304347,  1.2322385520833326]
+
+
+data = np.transpose([coal, noCoal])
+
+
+# cem     = cem / float( cem[0])
+# cuckoo  = cuckoo / float( cuckoo[0])
+# ar     = ar /float( ar[0])
+# dft     = dft/ float( dft[0])
+# rsa     = rsa / float( rsa[0])
+# bc      = bc / float( bc[0])
+# sort    = sort / float( sort[0])
+
+
+
+
+# apps = [ cuckoo, ar, dft, rsa, bc, sort]
+# appsName = [ "cuckoo", "ar", "dft", "rsa", "bc", "sort"]
+# numOfApps = len(apps)
+
+# basePag = 32
+# numPags =  ( basePag , 2 * basePag, 4 * basePag, 8 * basePag, 16 * basePag, 32 * basePag  )   
+
+
+figureSetting()
+f = plt.figure(figsize=(8,2.5))
+
+# xlabelsPos=[]
+# s = 0
+# barsGroupsGap = 0.04
+# barWidth = (1.0/len(apps[0]))
+# barsPos = ( np.arange( len(apps[0]) )/float(len(apps[0])) ) # bars position for a single set of bars
[…]

### 2017-11-30 experiments/newExperiments/scripts/pagFalts.py

@@ -1,0 +1,98 @@
+#!/usr/bin/python
+
+
+"""
+    IPOS: performance 
+    Metric : time
+    Task   : execute apps
+
+
+@ Author  : Amjad Yousef Majid
+@ Gmail ID: amjad.y.majid 
+@ Date    : 17/October/2017 
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+from csf import *
+
+
+
+
+
+
+
+
+
+cem      = np.array( [ 77,39,21,16-1,197-6,5286-1992 ]) 
+cuckoo   = np.array( [ 9, 5, 3, 2, 1, 0 ]) 
+ar   	 = np.array( [ 4,2,1,0,0,0 ]) 
+dft      = np.array( [ 8, 4, 2, 1, 0, 0]) 
+rsa      = np.array( [ 8, 4, 2, 1, 0, 0])   
+bc       = np.array( [ 0,0,0,0,0,0 ]) 
+sort     = np.array( [ 6, 3, 1, 0, 0, 0 ]) 
+
+# cem      = np.array( [ 0,0,0,1,6,1992  ]) 
+
+
+
+# cem     = cem / float( cem[0])
+# cuckoo  = cuckoo / float( cuckoo[0])
+# ar     = ar /float( ar[0])
+# dft     = dft/ float( dft[0])
+# rsa     = rsa / float( rsa[0])
+# bc      = bc / float( bc[0])
+# sort    = sort / float( sort[0])
+
+
+
+
+apps = [ cem, cuckoo, ar, dft, rsa, bc, sort]
+appsName = [ "cem", "cuckoo", "ar", "dft", "rsa", "bc", "sort"]
+numOfApps = len(apps)
+
+basePag = 32
+numPags =  ( basePag , 2 * basePag, 4 * basePag, 8 * basePag, 16 * basePag, 32 * basePag  )   
+
+# zeroPagFalt = ['0', '0', ]
+
+figureSetting()
+f = plt.figure(figsize=(8,2.5))
+
+xlabelsPos=[]
+s = 0
+barsGroupsGap = 0.04
+barWidth = (1.0/len(apps[0]))
+barsPos = ( np.arange( len(apps[0]) )/float(len(apps[0])) ) # bars position for a single set of bars
+zeros = []
+for i in range(numOfApps):
+    [p32, p64, p128, p256, p512, p1024] = plt.bar(  barsPos + s ,apps[i], width= barWidth, color='r', bottom=0.01 )
+
[…]

### 2017-11-30 experiments/newExperiments/scripts/ramPage_timeToCompletion.py

@@ -1,80 +1,0 @@
-#! /usr/bin/python
-
-
-"""
-    IPOS: performance 
-    Metric : time
-    Task   : execute apps
-
-
-@ Author  : Amjad Yousef Majid
-@ Gmail ID: amjad.y.majid 
-@ Date    : 17/October/2017 
-"""
-import numpy as np
-import matplotlib.pyplot as plt
-from csf import *
-
-numPags =  ("single",  "double",  "triple" )   
-
-############### page size of 128 bytes ##############
-
-cem      = np.array( [ 3691384,       2081138,    2658959 ]) 
-cuckoo   = np.array( [ 443089,        513159,     622667  ]) 
-crc      = np.array( [ 82567,         84605,      86641   ]) 
-dft      = np.array( [ 26267755,      26010302,   26254522]) 
-rsa      = np.array( [ 2196414,       1771543,    2260314 ]) 
-bc       = np.array( [ 831246,        849276,     868791  ]) 
-sort     = np.array( [ 12242871,      7765337,    7920915 ]) 
-
-############### page size of 64 bytes ##############
-
-# cem      = np.array( [ 3249640,       1869253,    2205335 ]) 
-# cuckoo   = np.array( [ 382723,        440148,     552546  ]) 
-# crc      = np.array( [ 70087,         72061,      74033   ]) 
-# dft      = np.array( [ 26372167,      26306501,   25931295]) 
-# rsa      = np.array( [ 2277448,       2070869,    1815520 ]) 
-# bc       = np.array( [ 713362,        726396,     745847  ]) 
-# sort     = np.array( [ 13915418,      9599409,    8834494 ]) 
-
-
-cem     = cem / float( cem[0])
-cuckoo  = cuckoo / float( cuckoo[0])
-crc     = crc /float( crc[0])
-dft     = dft/ float( dft[0])
-rsa     = rsa / float( rsa[0])
-bc      = bc / float( bc[0])
-sort    = sort / float( sort[0])
-
-
-
-
-apps = [cem, cuckoo, crc, dft, rsa, bc, sort]
-appsName = ["cem", "cuckoo", "crc", "dft", "rsa", "bc", "sort"]
-
-figureSetting()
-f = plt.figure(figsize=(8,2.5))
-
-xlabelsPos=[]
-s = 0
-for i in range(7):
[…]

### 2017-11-30 experiments/newExperiments/scripts/ramPages_timeToCompletion.py

@@ -15,40 +15,21 @@ import numpy as np
 import matplotlib.pyplot as plt
 from csf import *
 
-############### page size of 128 bytes ##############
-
-# cem      = np.array( [ 3691384,       2081138,    2658959 ]) 
-# cuckoo   = np.array( [ 443089,        513159,     622667  ]) 
-# crc      = np.array( [ 82567,         84605,      86641   ]) 
-# dft      = np.array( [ 26267755,      26010302,   26254522]) 
-# rsa      = np.array( [ 2196414,       1771543,    2260314 ]) 
-# bc       = np.array( [ 831246,        849276,     868791  ]) 
-# sort     = np.array( [ 12242871,      7765337,    7920915 ]) 
-
-############### page size of 64 bytes ##############
-
-# cem      = np.array( [ 3249640,       1869253,    2205335 ]) 
-# cuckoo   = np.array( [ 382723,        440148,     552546  ]) 
-# crc      = np.array( [ 70087,         72061,      74033   ]) 
-# dft      = np.array( [ 26372167,      26306501,   25931295]) 
-# rsa      = np.array( [ 2277448,       2070869,    1815520 ]) 
-# bc       = np.array( [ 713362,        726396,     745847  ]) 
-# sort     = np.array( [ 13915418,      9599409,    8834494 ]) 
 
 
-
-# cem      = np.array( [ 3691384,       2081138,    2658959 ]) 
+cem      = np.array( [ 13251011,5115064,4283901,4220972,4248560,13867514 ]) 
 cuckoo   = np.array( [ 454057, 427650, 457850, 552558, 768143, 1222542 ]) 
+ar   	 = np.array( [ 483818,452052,469499,421585,471081,542725 ]) 
 dft      = np.array( [ 26058854, 26625854, 25964064, 25928743, 26197667, 26763421]) 
-rsa      = np.array( [ 1929844, 1829844, 1849674, 1987705, 2248648, 2910148])    ## the fist result of RSA must be re-produced !! 
+rsa      = np.array( [ 2614059, 1829844, 1849674, 1987705, 2248648, 2910148])    ## the fist result of RSA must be re-produced !! 
[…]

### 2017-11-30 experiments/newExperiments/scripts/realTaskSizeDist.py

@@ -56,7 +56,8 @@ def main():
 	f = plt.figure("Apps distributions", figsize=(8,3))
 	ax = plt.axes()
 	ax.set_xscale('log')
-	for app in apps_names :
+	# linSty = ['b--', 'r:', 'g.', 'b:', "m-", ',', '>']
+	for i, app in enumerate(apps_names) :
 		c = generateRandomColor()
 		# if not normPlotting( apps_dict[app], c  ): 
 		# 	print(apps)

### 2017-11-30 experiments/newExperiments/scripts/serialConnection.py

@@ -6,16 +6,34 @@ import time
 from csf import *
 # get app name 
 appName=unified_input("enter app's name: ")
-fileName=unified_input("file's name (fast, slow, smart): ")
+fileName=unified_input("file's name ( (f)ast, (s)low, smar(t) ): ")
+
+if fileName == 'f':
+	fileName = 'fast'
+elif fileName == 's':
+	fileName = 'slow'
+elif fileName == 't':
+	fileName = 'smart'
+
+
 fileName +='Change.txt'
-pwrPtn = unified_input("Select from (camel, toward, fixed)")
+pwrPtn = unified_input("Select from ( (c)amel, (t)oward, (f)ixed)")
+
+if pwrPtn == 'c':
+	pwrPtn = 'camel'
+elif pwrPtn == 't':
+	pwrPtn = 'toward'
+elif pwrPtn == 'f':
+	pwrPtn = 'fixed'
 
 Togo=int( unified_input("Number of lines to be received: ") )
 lineLen=int( unified_input("line's length: ") )
 
 # check if the required directory exist
 basePath= '../data/'
-tailPath = 'simulatedPwrInter/CoalescedTaskSize/'+pwrPtn+'PowerPattern/'
+# tailPath = 'simulatedPwrInter/CoalescedTaskSize/'+pwrPtn+'PowerPattern/'
+tailPath = 'realTaskSize/'
+
 
 dirPath = os.path.join(os.path.join(basePath,appName), tailPath)
 
@@ -31,11 +49,17 @@ cp = os.getcwd()
 
 fname =  open( os.path.join(dirPath,fileName) , 'w')
 
-portName = "/dev/cu.usbmodem1413"
+portsName = ["/dev/cu.usbmodem1413","/dev/cu.usbmodem1423","/dev/cu.usbmodem1433" ]
 baudRate = 115200
-
-ser = serial.Serial(portName, baudRate, timeout=5)
-
+# port=None
+# e=None
+# ser=None
+# while port in portsName:
+# 	try: 
+ser = serial.Serial(portsName[1], baudRate, timeout=5)
+	# except e:
+	# 		print("Port error: ", str(e) )
+	# 		continue
 
 while Togo:
 		try:

### 2017-11-30 experiments/otherData/chain_viper/viper_chain.py

@@ -100,7 +100,7 @@ plt.ylabel('Seconds')
 num_bars = len(data)
 # hatches = ['+', '/']
 colors = ['#edf8b1', '#7fcdbb']
-labels = ["VIPER", "Chain"]
+labels = ["IPE", "Chain"]
 
 gap = 0.1                         # space each bars group 0.1 from the next group
 bar_width = (1- gap) / num_bars   # divide the remaining distance equally between the bars          
@@ -142,7 +142,7 @@ for i, row in enumerate(data):    # enumerate returns a data unit (a row) and it
 
 plt.legend(loc='upper left')
 # plt.ylim(0,3.2) 
-f.savefig("../figures/viperChain.eps",format="eps", dpi=1200)
+f.savefig("../figures/ipeChain.eps",format="eps", dpi=1200)
 plt.show()
 
 

### 2017-11-30 experiments/otherData/experiment_kiwan/main_delft_cem.c

@@ -101,7 +101,7 @@ void task_init_dict()
 {
 	LOG("init dict: letter %u\r\n", P(_v_letter));
 
-	int i = P(_v_letter);	
+	uint16_t i = P(_v_letter);	
 	PW(_v_dict[i].letter) = P(_v_letter); 
 	PW(_v_dict[i].sibling) = NIL; 
 	PW(_v_dict[i].child) = NIL; 

### 2018-03-22 paper/background.tex
sec 2
some TODOs for oopsla18

@@
 Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
-We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
-\sys supports I/O behaviour similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.
+We do not require a particular non-volatile memory technology, nor do we require architectural additions to commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
+\sys supports I/O behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.
 
@@
 
+\xxx{reference work by Intel}
+
 Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into tasks.
@@
 		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
+        \cite{intel} & TODO \\ 
 		\hline
@@
 \textbf{Memory Consistency Enforcement Overhead.} Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
+\xxx{Intel: undo and redo logging cost} \xxx{todo for table}.
 Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf.
@@
 Sizing tasks complicates {\em porting} code from one device to another.
-An excessively costly task on a device with a small energy buffer may be a relatively short, task on a device with a larger buffer.
-The key problem is that using existing systems, the programmer statically sizes tasks for a fixed energy buffer.
-\sys's {\em task coalescing}, described
+An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer.
+The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer
+either risks non-termination or poor performance due to high overhead. \sys's {\em task coalescing}, described
 in Section~\ref{sec:task_coalescing}, is motivated by these observations.

### 2018-03-22 paper/intro.tex
some TODOs for oopsla18

@@
 
-There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
-New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
+There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl} \xxx{tictpl is not HW, is it?
+also it doesn't deal with inconsistency; this paragraph doesn't include Hibernus, so it should not include CPTL either.} New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
 However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.
@@
 static decomposition of a program (by a programmer) into a collection of tasks.
+\xxx{add \cite{intel} work; separate by-programmer and by-compiler; both cases
+hit the energy estimation problem; also intel work proposes a method for decomposing
+into one of many decompositions, but does not give a good solution for choosing one}
 Tasks can include arbitrary computation and, upon completion, are guaranteed to
@@
 incoming power always available, is to predict task energy consumption and
-choose a decomposition specific to the device capacitor size.
-However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution.
-Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.
+choose a decomposition specific to the device capacitor size~\cite{cleancut}.
+However, statically predicting energy of arbitrary input-dependent code with
+peripheral access is a problem without a general solution. Furthermore, a
+static decomposition approach prevents portability across devices with
+different storage capacitors.
 
@@
 The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss.
-Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
-applications to an intermittent platform with a moderate effort.
+\xxx{Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
+applications to an intermittent platform with a moderate effort.}
 
@@
 \begin{itemize}
-	\item Two adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
+	\item \xxx{Two} adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
 	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;

### 2018-03-22 paper/systemdescription.tex
some TODOs for oopsla18

@@
-\sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
-\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em accessible}, {\em efficient} and {\em flexible}.
-\sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.
+\sys is a new programming and execution model for intermittent computing on
+energy-harvesting devices. \sys addresses the challenges outlined in
+Section~\ref{sec:background} to make task-based intermittent programs {\em
+accessible}, {\em efficient} and {\em flexible}. \sys accomplishes this goal
+with a new program analysis and run time system for intermittent applications,
+refer to Figure~\ref{fig:system_overview}.
 

### 2018-04-08 paper/newCoalescing.tex
start updating the coalescing text

\begin{algorithm}
	\caption{Coalescing}
	\label{algo:genCoalescing}
	\scriptsize
	%\small
	\begin{algorithmic}[1]
        \State $H_\text{p} \gets H_\text{c}$
        \State $H_\text{p} \gets 0$ 
        \State $H_\text{B} \leftarrow $ \Call{$f_\text{reboot}$}{B, $H_\text{p}$}

        \While{$\texttt{true}$}
	        \State $ B_\text{c} \gets B$
	        \While{$ B_\text{c} > 0$}
		        \State \Call{execute\_task}{$T_\texttt{i}$}
		        % \State $W$ \gets \Call{f_\text{weight}}{$T_\texttt{i}$}
		        \State $B_\text{c} \gets B_\text{c} - W$
				\State $H_\text{c} \gets H_\text{c} + W$
	        \EndWhile
	        % \State $B$ \gets \Call{f\_compl}{$\texttt{B},\texttt{H_\text{c}},\texttt{H_\text{p}}$}
	        \State \Call{commit\_to\_fram}{}
        \EndWhile
	\end{algorithmic}
\end{algorithm}

%% Coalescing motivation
Finding an optimal task size, given random energy conditions is an open question. Obviously, a static approach (i.e., pure compiler-based) does not have the potential to answer it. Therefore, \sys champions runtime-based methods to approximate the ideal task size under random energy shots arrival. Furthermore, \sys advocates making intermittently powered software energy-aware is the key for \emph{efficient code execution and probability}.

% %% address the challenge and define the optimal task
\sys advances its execution in a state-less (or virtual) manner, and then it frequently saves its forward progress. The longer \sys virtually progresses, the less committing overhead it introduces. However, long state-less execution results in a considerable re-execution penalty; All the tasks that have been virtually executed, must be re-executed after a power interrupt. As such, an optimal task is a task that occupies with a single commit an entire power cycle, which is therefore necessarily of a varying length.

% %% How does Coala approximates an ideal task
\sys adapts its virtual execution by mean of \emph{coalescing} static tasks, when energy permits, to optimally execute an intermittently powered program. It devises a number of coalescing algorithms, and states that a good coalescing strategy must moderate the risk of wasted work (re-execution penalty) and capitalize on benefits of deferred commits. To describe the proposed strategies we will use the following parameters: \\
(i) \textit{target budget B}: total number of units to coalesce; (ii) \textit{current budget $B_\text{c}$}: remaining number of tasks to be coalesced before committing to non-volatile memory; (iii) \textit{history H}: total number of tasks executed in the current discharge cycle, initialized to the number of executed tasks in the last power cycle.

The general structure of the proposed coalescing strategies is summaries in algorithm~\ref{algo:genCoalescing}, where where $H_\text{p}$, $H_\text{c}$, $B$ and $B_\text{c}$ represent past history, current history, target budget and current budget, respectively. $T_\text{i}$ is the i-th task being executed and $W$ is its weight.
A coalescing strategy and its performance are defined by the functions: \\
(i) \emph{$f_\text{reboot}$}: which updates the target budget after a reboot; (ii) \emph{$f_\text{weight}$}: which returns the weight of the task just executed and its unit; (iii) \emph{$f_\text{compl}$}: which updates the target budget after a successful completion of a dynamic coalesced task. The following sections detail the different strategies \sys proposes and the needs that originated from one strategy leading to the conception of another one.

\subsection{Slow Coalescing}
\label{subsec:slowCoalescing}

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{figures/slowCoal}%coalescing_timeline
	\caption{The Slow coalescing algorithm \emph{linearly} updates its coalescing target}
	\label{fig:slowCoal}
\end{figure}

The need for \emph{adaptiveness}, that is coalescing static tasks, can be fulfilled by a naive strategy, having the target budget $B$ to increase by $x$ static tasks after a successful completion of a coalesced task, and decrease it by $x$ static task upon a reboot. In this case: $f_\text{reboot} = B - x$; $f_\text{weight} =  1$; $f_\text{compl} = B + x$; \\
Due to the linear behavior of the target adaptation, this algorithm is slow in reacting to the changes in energy conditions; the energy required by a coalesced task or the available energy. In particular, this algorithm can experience a large number of \emph{repeated power} failures without forward progress.

\subsection{History-aware Coalescing (HC)}
\label{subsec:hCCoalescing}

### 2018-04-08 apps/sort/main.c

@@ -4,7 +4,7 @@
 #include "mspProfiler.h"
 #include "mspDebugger.h"
 
-//#define TSK_SIZ 0
+#define TSK_SIZ 1
 //#define RAISE_PIN 1
 //#define PWR_INT_SIM 1
 
@@ -154,10 +154,11 @@ void init()
 #endif
 
 #ifdef TSK_SIZ
+    uart_init();
     cp_init();
 #endif
 
-    uart_init();
+//    uart_init();
 
 #ifdef LOG_INFO
     uart_init();

### 2018-04-08 coala/src/psched.c

@@ -1,6 +1,6 @@
 #include <psched.h>
 #include <msp430fr5969.h>
-#include <mspDebugger.h>
+//#include <mspDebugger.h>
 
 // PLEASE SELECT AN ALGORITHIN
 #define NO_COALESCING 1

### 2018-04-08 experiments/newExperiments/scripts/csf.py

@@ -13,6 +13,7 @@ import colors
 from collections import defaultdict
 import sys
 
+plt.style.use('seaborn-darkgrid')
 
 colorCnts=0
 
@@ -141,6 +142,7 @@ def figureSetting():
     """
     Set a figure params
     """
+    plt.rcParams['font.family'] = ['Tahoma', 'DejaVu Sans','Lucida Grande', 'Verdana']
     #set figure linewidth
     plt.rcParams['axes.linewidth']=2
 

### 2018-04-08 experiments/newExperiments/scripts/slowCoalescingBehavior.py

@@ -1,0 +1,76 @@
+#! /usr/bin/python
+
+
+"""
+    IPOS: performance 
+    Metric : time
+    Task   : execute apps
+
+
+@ Author  : Amjad Yousef Majid
+@ Gmail ID: amjad.y.majid 
+@ Date    : 8/April/2017 
+"""
+import numpy as np
+import matplotlib.pyplot as plt
+from csf import *
+
+coalTaskSize = [1,2,3,4,5, 6,5,4,3, 2,3,4]
+
+figureSetting()
+f = plt.figure(figsize=(8,2.5))
+
+
+
+plt.plot(coalTaskSize)
+
+data_len = len(coalTaskSize)
+for i in range(data_len-1):
+    if coalTaskSize[i+1] > coalTaskSize[i]:
+        plt.plot(i, coalTaskSize[i], 'o', color='g')
+    else:
+        plt.plot(i, coalTaskSize[i], 'X', color='r')
+
+
+
+plt.annotate('Repeated failures',
+             fontsize=14,
+             xytext = (5,1.2),  # text location
+             xy=(5,5.8), 
+                    # arrows points to 
+             ha = 'center',    # horizontal alignment 
+             va = 'bottom',    # vertical alignment
+             arrowprops={'facecolor': 'k', 'width': 2, "headwidth":7, 'shrink' : 0.01}, 
+            )
+
+plt.annotate("", 
+             fontsize=14,
+             xytext = (5,2),  # text location
+             xy=(5.9,4.8), 
+                    # arrows points to 
+             ha = 'center',    # horizontal alignment 
+             va = 'bottom',    # vertical alignment
+             arrowprops={'facecolor': 'k', 'width': 2, "headwidth":7, 'shrink' : 0.01}, 
+            )
+
+plt.annotate("", xytext = (5,2),  # text location
+             fontsize=14,
+             xy=(6.9,3.8), 
+                    # arrows points to 
+             ha = 'center',    # horizontal alignment 
+             va = 'bottom',    # vertical alignment
+             arrowprops={'facecolor': 'k', 'width': 2, "headwidth":7, 'shrink' : 0.01}, 
+            )
+
+plt.text(0.7,5 , 'successful commit', fontsize=14, color='k', rotation=26) 
+
+plt.ylabel("Coal. Task Size")
+plt.xlabel("Time")
+plt.tight_layout()
[…]

### 2018-04-09 paper/newCoalescing.tex
put back the old coalecing section

@@
 Finding an optimal task size, given random energy conditions is an open question.
-Obviously, a static approach (i.e., pure compiler-based) does not have the potential to answer it.
+Obviously, a static approach does not have the potential to answer it.
 Therefore, \sys champions runtime-based methods to approximate the ideal task size under random energy shots arrival.
 Furthermore, \sys advocates making intermittently powered software energy-aware is the key for \emph{efficient code execution and probability}.
+%However, given the extremely limited recourses any optimization technique must adhere to the principle of simplicity, otherwise it introduces a non-tolerable overhead.
 
@@
 \sys advances its execution in a state-less (or virtual) manner, and then it frequently saves its forward progress.
-The longer \sys virtually progresses, the less committing overhead it introduces.
-However, long state-less execution results in a considerable re-execution penalty; All the tasks that have been virtually executed, must be re-executed after a power interrupt.
+The longer \sys virtually progresses, the less committing (saving data to non-volatile memory) overhead it introduces.
+However, long state-less execution results in a considerable re-execution penalty---all the tasks that have been virtually executed must be re-executed after a power interrupt.
 As such, an optimal task is a task that occupies with a single commit an entire power cycle, which is therefore necessarily of a varying length.
@@
 
-\subsection{Slow Coalescing}
+\subsection{SC: Slow Coalescing}
 \label{subsec:slowCoalescing}
@@
 
-The need for \emph{adaptiveness}, that is coalescing static tasks, can be fulfilled by a naive strategy, having the target budget $B$ to increase by $x$ static tasks after a successful completion of a coalesced task, and decrease it by $x$ static task upon a reboot.
+The need for \emph{adaptiveness}, that is coalescing static tasks, can be fulfilled by a naive strategy, having the target budget $B$ to increase by $x$ static tasks after a successful completion of a coalesced task, and to decrease by the same number of tasks upon a reboot.
 In this case: $f_\text{reboot} = B - x$; $f_\text{weight} =  1$; $f_\text{compl} = B + x$; \\
@@
 
-\subsection{History-aware Coalescing (HC)}
+\subsection{HC: History-aware Coalescing}
 \label{subsec:hCCoalescing}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/HCCoal}%coalescing_timeline
+	\caption{The History-aware coalescing algorithm  updates its coalescing target}
+	\label{fig:HCCoal}
+\end{figure}
+
+The History-aware algorithm uses its recent execution history as metric to estimate the optimal coalesce task size.
+Relying on the history of execution enables this algorithm to react very quickly to discharge rates variation.

### 2018-04-09 paper/abstract.tex
rewrite the abstract

@@
-Computation consistency on intermittently-powered devices can be ensured by: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) via tasks, where a programmer splits the code into small idempotent sections.
+% Connecting the work to a real-world problem
+Battery-free computers promise a sustainable Internet of things by forgo the batteries and scarf the energy from the environment. 
+% The main problem that we are tackling
+However, relying on the surrounding energy when it is available is at odds with current execution model of long-running computation. 
+% The already proposed solution to the main problem
+Scientists have proposed a discrete-like execution model that alternates between progressing the execution and protecting the execution progress, against power interrupts.
+% The problem of the proposed solutions 
+Generally, the proposed techniques can be classified as (i) checkpointing systems, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based systems, where a programmer splits the code into small idempotent sections.
+Despite the superiority of the task-based techniques, they suffer from a fixed overhead and they are completely blind to the energy conditions.
+% Our main claim 
+In this paper we introduce \sys an \emph{efficient} and \emph{adaptive} task-based execution model.
+% show off
+\sys tunes its frequency of switching between progressing and protecting based on the available energy, but 
+% One of the main functions of an OS is to abstract the hardware and to be less dependent on it. Coala does that too!
+tt does not require hardware support to refine its estimation. 
+% main benefits of Coala
+\sys defaults to progress on a multi-tasks scale, when energy permits (\emph{efficiency}), and on a sub-task scale when needed {\emph{guarantee}}. 
+% I really hope it will support our claims!! haha
+Results show the \sys is more efficient than the current state-of-art execution model by X-times and more reliable.
+
+% old abstract 
+%Computation consistency on intermittently-powered devices can be ensured by: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) via tasks, where a programmer splits the code into small idempotent sections.
 Results so far suggest that a task-based approach performs better than checkpointing.

### 2018-04-10 paper/intro.tex
add task splitting, correct the provided example

--- 2017-11-17 paper/intro.tex
@@
 Unfortunately, naively merging two atomic tasks might produce a non-atomic task, that may leave non-volatile memory inconsistent after a partial execution.
-A merge breaks atomicity when atomic merged tasks form a \emph{write-after-read (WAR)} dependency---for instance if two tasks \texttt{\{x++\}; \{vector[x]=v;\}} are merged and if the power failure occurs after \texttt{x++}, the value \texttt{x} will be increased twice when the merged task is restarted, that leads to an inconsistency.
+A merge breaks atomicity when atomic merged tasks form a \emph{write-after-read (WAR)} dependency---for instance if two tasks \texttt{\{x=y+1\}; \{y=x;\}} are merged and if the power failure occurs after \texttt{x=y}, the value \texttt{x} will be increased twice when the merged task is restarted, that leads to an inconsistency.
 %
@@
 
+\textbf{Task Splitting}: The static task execution model assumes that a single task execution is guaranteed by the hardware.
+If this assumption is violated, due to hardware inconsistency or inaccurate task devision startegy, the system enters a non-termination loop of failed attempts to reach the next task boundery.
+logically, better than relying on a rondom process, energy shots arrival, to break out of this loop, the system should execute an "emergency-like" routine to terminate it and preserve forward progress.
+
 Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically.
@@
 	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
-	\item Compiler pass for automatic decomposition into small atomic tasks for adaptive coalescing at runtime.
+	\item Task splitting startegy that preserve forward progress and memory consistency;
 \end{itemize}

### 2018-04-10 paper/paper.tex
med langth title corrected
fix in latex in main file
reducing section spacing globally
overview: shortenning
remove repetative coala from section titles
paging: more details and specifics, incl. algo for commit
space enlargements to fill the last page
start updating the coalescing text
put back the old coalecing section
add task splitting, correct the provided example
git push test

@@
 
-\acmConference[Submitted to PLDI'18]{ACM SIGPLAN Conference on Programming Language Design and Implementation}{June 18--22, 2018}{Philadelphia, PA, USA}
+\acmConference[Submitted to OOPSLA'18]{ACM SIGPLAN Conference on Programming Language Design and Implementation}{June 18--22, 2018}{Philadelphia, PA, USA}
 
@@
 \usepackage{listings}
+\usepackage{hyphenat}
+
+\usepackage{titlesec}
+
+%\titlespacing{\section}{0pt}{0.5\baselineskip}{\parskip}
+%\titlespacing{\subsection}{0pt}{0.5\baselineskip}{\parskip}
+%\titlespacing{\subsubsection}{0pt}{0.5\baselineskip}{\parskip}
+\titlespacing{\section}{2pt}{0.8\baselineskip}{\parskip}
+\titlespacing{\subsection}{2pt}{0.6\baselineskip}{\parskip}
+\titlespacing{\subsubsection}{2pt}{0.6\baselineskip}{\parskip}
+
+\setlength{\textfloatsep}{5pt}% Remove \textfloatsep
 
@@
 \hypersetup{colorlinks=false,linkcolor=blue,urlcolor=blue}
-\urlstyle{rm}
-\linespread{0.98}
 
@@
 
+\algnewcommand{\LeftComment}[2][.5\linewidth]{%
+  \leavevmode\makebox[#1][l]{$\triangleright$~#2}\hfill}
+
 % Tighter than defaults:
-\setlength{\sectionbelowskip}{1pt}
+%\setlength{\sectionbelowskip}{1pt}
+\setlength{\sectionbelowskip}{4pt}
 \setlength{\sectionaboveskip}{5pt plus 1.5pt minus 1.5pt}
@@
 \newcommand{\sys}{Coala\xspace}
-\newcommand{\sysmed}{\sys: Intermittent Computing $\ldots$}
+\newcommand{\sysmed}{\sys: Adaptive Task Coalescing $\ldots$}
 \newcommand{\sysfull}{\sys: Adaptive Task Coalescing for Intermittent Computing on Energy-harvesting Devices}
@@
 
-\section{\sys: Task Coalescing}
+\section{Task Coalescing}
 \label{sec:task_coalescing}
@@
 \input{coalescing}
+% \input{newCoalescing}
 
-\section{\sys: Memory Virtualization}
+\section{Memory Virtualization}
 \label{sec:memory_virtulaization}
@@
 
-\section{\sys: Automatic Task Generation}
+\section{Automatic Task Generation}
 \label{sec:compiler}

### 2018-04-10 paper/paper.tex
made paper OOPSLA 2018 style complaint

@@
-\documentclass[sigplan,10pt,review,anonymous]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
 
-\acmConference[Submitted to OOPSLA'18]{ACM SIGPLAN Conference on Programming Language Design and Implementation}{June 18--22, 2018}{Philadelphia, PA, USA}
-
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
+\acmArticle{1}
 \acmYear{2018}
-\acmISBN{} % \acmISBN{978-x-xxxx-xxxx-x/YY/MM}
+\acmMonth{1}
 \acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
 \startPage{1}
+
 \setcopyright{none}
 
-%% Some recommended packages.
-%\usepackage{booktabs}   %% For formal tables: http://ctan.org/pkg/booktabs
-\usepackage{subcaption} %% For complex figures with subfigures/subcaptions http://ctan.org/pkg/subcaption
-%\usepackage[normalem]{ulem}
-%\usepackage{float}
+\usepackage{booktabs}   % For formal tables: http://ctan.org/pkg/booktabs
+\usepackage{subcaption} % For complex figures with subfigures/subcaptions http://ctan.org/pkg/subcaption
 \usepackage{soul}
@@
 
-\usepackage{titlesec}
+\hypersetup{colorlinks=false,linkcolor=blue,urlcolor=blue}
+
+%\usepackage[normalem]{ulem}
+%\usepackage{float}
+
+%\usepackage{titlesec} %P{rzemek: ommented, has conflict https://tex.stackexchange.com/questions/321999/conflict-between-makeuppercase-in-title-formatting-and-usepackagexcolor
 
@@
 %\titlespacing{\subsubsection}{0pt}{0.5\baselineskip}{\parskip}
-\titlespacing{\section}{2pt}{0.8\baselineskip}{\parskip}
-\titlespacing{\subsection}{2pt}{0.6\baselineskip}{\parskip}
-\titlespacing{\subsubsection}{2pt}{0.6\baselineskip}{\parskip}
+%\titlespacing{\section}{2pt}{0.8\baselineskip}{\parskip}
+%\titlespacing{\subsection}{2pt}{0.6\baselineskip}{\parskip}
+%\titlespacing{\subsubsection}{2pt}{0.6\baselineskip}{\parskip}
 
-\setlength{\textfloatsep}{5pt}% Remove \textfloatsep
+%\setlength{\textfloatsep}{5pt}% Remove \textfloatsep
 
@@
 %\captionsetup[figure]{aboveskip=0.5em, belowskip=0em}
-\captionsetup[subfloat]{font={small,sf}}
-
-\hypersetup{colorlinks=false,linkcolor=blue,urlcolor=blue}
+%\captionsetup[subfloat]{font={small,sf}}
 
@@
 \newcommand{\sys}{Coala\xspace}
-\newcommand{\sysmed}{\sys: Adaptive Task Coalescing $\ldots$}
+\newcommand{\sysmed}{\sys: Adaptive Task Coalescing for Intermittent Computing $\ldots$}
 \newcommand{\sysfull}{\sys: Adaptive Task Coalescing for Intermittent Computing on Energy-harvesting Devices}
@@
 
-\title[\sysmed]{\sysfull}
+%% Title information
+\title[\sysmed]{\sysfull}         %% [Short Title] is optional;
+%% when present, will be used in
+%% header instead of Full Title.
+%\titlenote{}             %% \titlenote is optional;
+%% can be repeated if necessary;
+%% contents suppressed with 'anonymous'
+%\subtitle{}                     %% \subtitle is optional
+%\subtitlenote{}       %% \subtitlenote is optional;
+%% can be repeated if necessary;
+%% contents suppressed with 'anonymous'
 
@@
 
-%\author{Kiwan Maeng}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Carnegie Mellon University}
-%	\streetaddress{4720 Forbes Avenue}
-%   \city{Pittsburgh} 
-%	\state{PA} 
-%	\postcode{15213}
-%}
-%\email{kmaeng@andrew.cmu.edu}
+\author{Kiwan Maeng}
+\authornote{}
+\orcid{}
+\affiliation{%
+	\institution{Carnegie Mellon University}
+	\streetaddress{4720 Forbes Avenue}
+   \city{Pittsburgh} 
+	\state{PA} 
+	\postcode{15213}
+}
+\email{kmaeng@andrew.cmu.edu}
 
-%\author{Alexei Colin}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Carnegie Mellon University}
-%	\streetaddress{4720 Forbes Avenue}
-%	\city{Pittsburgh} 
-%	\state{PA} 
-%	\postcode{15213}
-%}
-%\email{acolin@andrew.cmu.edu}
+\author{Alexei Colin}
+\authornote{}
+\orcid{}
+\affiliation{%
+	\institution{Carnegie Mellon University}
+	\streetaddress{4720 Forbes Avenue}
+	\city{Pittsburgh} 
+	\state{PA} 
+	\postcode{15213}
+}
+\email{acolin@andrew.cmu.edu}
 
-%\author{Brandon Lucia}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Carnegie Mellon University}
-%	\streetaddress{4720 Forbes Avenue}
-%	\city{Pittsburgh} 
-%	\state{PA} 
-%	\postcode{15213}
-%}
-%\email{blucia@andrew.cmu.edu}
+\author{Brandon Lucia}
+\authornote{}
+\orcid{}
+\affiliation{%
+	\institution{Carnegie Mellon University}
+	\streetaddress{4720 Forbes Avenue}
+	\city{Pittsburgh} 
+	\state{PA} 
+	\postcode{15213}
+}
+\email{blucia@andrew.cmu.edu}
 
-%\author{Amjad Yousef Majid}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Delft University of Technology}
-%	\streetaddress{Mekelweg 4}
-%	\city{Delft, The Netherlands} 
-%	\state{Zuid Holland} 
-%	\postcode{2628\,CD}
-%}
-%\email{a.y.majid@tudelft.nl}
+\author{Carlo Delle Donne}
+\authornote{}
+\orcid{}
+\affiliation{%
+	\institution{Delft University of Technology}
+	\streetaddress{Van Mourik Broekmanweg 6}
+	\city{Delft, The Netherlands} 
+	\state{Zuid Holland} 
+	\postcode{2628\,XE}
+}
+\email{c.delledonne@student.tudelft.nl}
 
-%\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Delft University of Technology}
-%	\streetaddress{Mekelweg 4}
-%	\city{Delft, The Netherlands} 
-%	\state{Zuid Holland} 
-%	\postcode{2628\,CD}
-%}
-%\email{k.s.yildirim@tudelft.nl}
+\author{Amjad Yousef Majid}
+\authornote{}
+\orcid{}
+\affiliation{%
+	\institution{Delft University of Technology}
+	\streetaddress{Van Mourik Broekmanweg 6}
+	\city{Delft, The Netherlands} 
+	\state{Zuid Holland} 
+	\postcode{2628\,XE}
+}
+\email{a.y.majid@tudelft.nl}
 
-%\author{Przemys{\l}aw Pawe{\l}czak}
-%\authornote{}
-%\orcid{}
-%\affiliation{%
-%	\institution{Delft University of Technology}
-%	\streetaddress{Mekelweg 4}
-%	\city{Delft, The Netherlands} 
-%	\state{Zuid Holland} 
-%	\postcode{2628\,CD}
-%}
-%\email{p.pawelczak@tudelft.nl}
+\author{Kas{\i}m Sinan Y{\i}ld{\i}r{\i}m}
+\authornote{}
+\orcid{}
+\affiliation{%
+	\institution{Delft University of Technology}
+	\streetaddress{Van Mourik Broekmanweg 6}
+	\city{Delft, The Netherlands} 
+	\state{Zuid Holland} 
+	\postcode{2628\,XE}
+}
+\email{k.s.yildirim@tudelft.nl}
 
-%\renewcommand{\shortauthors}{A. Y. Majid et al.}
+\author{Przemys{\l}aw Pawe{\l}czak}
+\authornote{}
+\orcid{}
+\affiliation{%
+	\institution{Delft University of Technology}
+	\streetaddress{Van Mourik Broekmanweg 6}
+	\city{Delft, The Netherlands} 
+	\state{Zuid Holland} 
+	\postcode{2628\,XE}
+}
+\email{p.pawelczak@tudelft.nl}
+
+\renewcommand{\shortauthors}{A. Y. Majid et al.}
 
@@
 
-%\begin{CCSXML}
-%	<ccs2012>
-%	<concept>
-%	<concept_id>10010520.10010553.10010562.10010564</concept_id>
-%	<concept_desc>Computer systems organization~Embedded 
-%	software</concept_desc>
-%	<concept_significance>500</concept_significance>
-%	</concept>
-%	<concept>
-%	<concept_id>10010520.10010575.10010578</concept_id>
-%	<concept_desc>Computer systems organization~Availability</concept_desc>
-%	<concept_significance>300</concept_significance>
-%	</concept>
-%	<concept>
-%	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
-%	<concept_desc>Software and its engineering~Scheduling</concept_desc>
-%	<concept_significance>300</concept_significance>
-%	</concept>
-%	<concept>
-%	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
-%	<concept_desc>Software and its engineering~Communications 
-%	management</concept_desc>
-%	<concept_significance>300</concept_significance>
-%	</concept>
-%	<concept>
-%	<concept_id>10011007.10010940.10010971.10010564</concept_id>
-%	<concept_desc>Software and its engineering~Embedded 
-%	software</concept_desc>
-%	<concept_significance>300</concept_significance>
-%	</concept>
-%	</ccs2012>
-%\end{CCSXML}
+\begin{CCSXML}
+	<ccs2012>
+	<concept>
+	<concept_id>10010520.10010553.10010562.10010564</concept_id>
+	<concept_desc>Computer systems organization~Embedded 
+	software</concept_desc>
+	<concept_significance>500</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10010520.10010575.10010578</concept_id>
+	<concept_desc>Computer systems organization~Availability</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10011007.10010940.10010941.10010949.10010957.10010688</concept_id>
+	<concept_desc>Software and its engineering~Scheduling</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10011007.10010940.10010941.10010949.10010965</concept_id>
+	<concept_desc>Software and its engineering~Communications 
+	management</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	<concept>
+	<concept_id>10011007.10010940.10010971.10010564</concept_id>
+	<concept_desc>Software and its engineering~Embedded 
+	software</concept_desc>
+	<concept_significance>300</concept_significance>
+	</concept>
+	</ccs2012>
+\end{CCSXML}
 
-%\ccsdesc[500]{Computer systems organization~Embedded software}
-%\ccsdesc[300]{Computer systems organization~Availability}
-%\ccsdesc[300]{Software and its engineering~Scheduling}
-%\ccsdesc[300]{Software and its engineering~Communications management}
-%\ccsdesc[300]{Software and its engineering~Embedded software}
+\ccsdesc[500]{Computer systems organization~Embedded software}
+\ccsdesc[300]{Computer systems organization~Availability}
+\ccsdesc[300]{Software and its engineering~Scheduling}
+\ccsdesc[300]{Software and its engineering~Communications management}
+\ccsdesc[300]{Software and its engineering~Embedded software}
 
-%\keywords{Energy harvesting, transient operation, operating system}
+\keywords{Energy harvesting, transient operation, operating system}
 
@@
 
-\input{coalescing}
-% \input{newCoalescing}
+%\input{coalescing}
+\input{newCoalescing}
 
@@
 
-%\section{Discussion}
-%\label{sec:discussion}
+\section{Discussion}
+\label{sec:discussion}
 
-%\input{discussion}
+\input{discussion}
 
@@
 \bibliography{bib}
+\citestyle{acmauthoryear}
 

### 2018-04-10 paper/adaptiveModel.tex
random

--- 2018-04-09 paper/newCoalescing.tex
@@
-\begin{algorithm}
-	\caption{Coalescing}
-	\label{algo:genCoalescing}
-	\scriptsize
-	%\small
-	\begin{algorithmic}[1]
-        \State $H_\text{p} \gets H_\text{c}$
-        \State $H_\text{p} \gets 0$ 
-        \State $H_\text{B} \leftarrow $ \Call{$f_\text{reboot}$}{B, $H_\text{p}$}
-
-        \While{$\texttt{true}$}
-	        \State $ B_\text{c} \gets B$
-	        \While{$ B_\text{c} > 0$}
-		        \State \Call{execute\_task}{$T_\texttt{i}$}
-		        % \State $W$ \gets \Call{f_\text{weight}}{$T_\texttt{i}$}
-		        \State $B_\text{c} \gets B_\text{c} - W$
-				\State $H_\text{c} \gets H_\text{c} + W$
-	        \EndWhile
-	        % \State $B$ \gets \Call{f\_compl}{$\texttt{B},\texttt{H_\text{c}},\texttt{H_\text{p}}$}
-	        \State \Call{commit\_to\_fram}{}
-        \EndWhile
-	\end{algorithmic}
-\end{algorithm}
-
 %% Coalescing motivation
@@
 As such, an optimal task is a task that occupies with a single commit an entire power cycle, which is therefore necessarily of a varying length.
-
-% %% How does Coala approximates an ideal task
-\sys adapts its virtual execution by mean of \emph{coalescing} static tasks, when energy permits, to optimally execute an intermittently powered program.
-It devises a number of coalescing algorithms, and states that a good coalescing strategy must moderate the risk of wasted work (re-execution penalty) and capitalize on benefits of deferred commits.
-To describe the proposed strategies we will use the following parameters: \\
-(i) \textit{target budget B}: total number of units to coalesce; (ii) \textit{current budget $B_\text{c}$}: remaining number of tasks to be coalesced before committing to non-volatile memory; (iii) \textit{history H}: total number of tasks executed in the current discharge cycle, initialized to the number of executed tasks in the last power cycle.
-
-The general structure of the proposed coalescing strategies is summaries in algorithm~\ref{algo:genCoalescing}, where where $H_\text{p}$, $H_\text{c}$, $B$ and $B_\text{c}$ represent past history, current history, target budget and current budget, respectively.
-$T_\text{i}$ is the i-th task being executed and $W$ is its weight.
-A coalescing strategy and its performance are defined by the functions: \\
-(i) \emph{$f_\text{reboot}$}: which updates the target budget after a reboot; (ii) \emph{$f_\text{weight}$}: which returns the weight of the task just executed and its unit; (iii) \emph{$f_\text{compl}$}: which updates the target budget after a successful completion of a dynamic coalesced task.
-The following sections detail the different strategies \sys proposes and the needs that originated from one strategy leading to the conception of another one.
-
-\subsection{SC: Slow Coalescing}
-\label{subsec:slowCoalescing}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/slowCoal}%coalescing_timeline
-	\caption{The Slow coalescing algorithm \emph{linearly} updates its coalescing target}
-	\label{fig:slowCoal}
-\end{figure}
-
-The need for \emph{adaptiveness}, that is coalescing static tasks, can be fulfilled by a naive strategy, having the target budget $B$ to increase by $x$ static tasks after a successful completion of a coalesced task, and to decrease by the same number of tasks upon a reboot.
-In this case: $f_\text{reboot} = B - x$; $f_\text{weight} =  1$; $f_\text{compl} = B + x$; \\
-Due to the linear behavior of the target adaptation, this algorithm is slow in reacting to the changes in energy conditions; the energy required by a coalesced task or the available energy.
-In particular, this algorithm can experience a large number of \emph{repeated power} failures without forward progress.
-
-\subsection{HC: History-aware Coalescing}
-\label{subsec:hCCoalescing}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/HCCoal}%coalescing_timeline
-	\caption{The History-aware coalescing algorithm  updates its coalescing target}
-	\label{fig:HCCoal}
-\end{figure}
-
-The History-aware algorithm uses its recent execution history as metric to estimate the optimal coalesce task size.
-Relying on the history of execution enables this algorithm to react very quickly to discharge rates variation.

### 2018-04-10 paper/intro.tex
random

@@
 We evaluated \sys end-to-end on a real energy-harvesting device~\cite{wisp} running existing benchmarks ~\cite{chain} and new programs.
-A comparison to a state-of-the-art task-based system~\cite{chain} showed that \sys recovers
+A comparison to a state-of-the-art task-based system~\cite{aplaca} showed that \sys recovers
 performance on conservative task decompositions missed by systems with static decomposition.

### 2018-04-10 paper/newCoalescing.tex
random

@@
 \end{algorithm}
-
-%% Coalescing motivation
-Finding an optimal task size, given random energy conditions is an open question.
-Obviously, a static approach does not have the potential to answer it.
-Therefore, \sys champions runtime-based methods to approximate the ideal task size under random energy shots arrival.
-Furthermore, \sys advocates making intermittently powered software energy-aware is the key for \emph{efficient code execution and probability}.
-%However, given the extremely limited recourses any optimization technique must adhere to the principle of simplicity, otherwise it introduces a non-tolerable overhead.
-
-% %% address the challenge and define the optimal task
-\sys advances its execution in a state-less (or virtual) manner, and then it frequently saves its forward progress.
-The longer \sys virtually progresses, the less committing (saving data to non-volatile memory) overhead it introduces.
-However, long state-less execution results in a considerable re-execution penalty---all the tasks that have been virtually executed must be re-executed after a power interrupt.
-As such, an optimal task is a task that occupies with a single commit an entire power cycle, which is therefore necessarily of a varying length.
 

### 2018-04-10 paper/paper.tex
random

--- 2018-04-10 paper/paper.tex
@@
 
-\acmConference[Submitted to OOPSLA'18]{ACM SIGPLAN Conference on Programming Language Design and Implementation}{June 18--22, 2018}{Philadelphia, PA, USA}
+\acmConference[Submitted to PLDI'18]{ACM SIGPLAN Conference on Programming Language Design and Implementation}{June 18--22, 2018}{Philadelphia, PA, USA}
 
@@
 \newcommand{\transition}{{\tt next\_task}\xspace}
-\newcommand{\sys}{Coala\xspace}
-\newcommand{\sysmed}{\sys: Adaptive Task Coalescing $\ldots$}
-\newcommand{\sysfull}{\sys: Adaptive Task Coalescing for Intermittent Computing on Energy-harvesting Devices}
+\newcommand{\sys}{ROO\xspace}
+%\newcommand{\sys}{Coala\xspace}
+\newcommand{\sysmed}{\sys: Adaptive Intermittent Execution Model $\ldots$}
+%\newcommand{\sysmed}{\sys: Adaptive Task Coalescing $\ldots$}
+%\newcommand{\sysfull}{\sys: Adaptive Task Coalescing for Intermittent Computing on Energy-harvesting Devices}
+\newcommand{\sysfull}{\sys: Adaptive Intermittent Execution Model for Energy-harvesting Devices}
 %newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization} %Old title
@@
 \input{background}
+\section{\sys: Adaptive Intermittent Execution Model}
+\label{sec:adaptiveExecModel}
+\input{adaptiveModel.tex}
 
-\section{\sys: System Overview}
-\label{sec:overall_system}
+\subsection{\sys: Coalescing}
+\label{sec:implementation}
+\input{newCoalescing}
+\input{coalescing}
 
+\subsection{\sys: Memory Virtualization}
+\label{sec:memory_virtulaization}
+\input{paging}
+
+\section{\sys Implementation}
+\label{sec:implementation}
 \input{systemdescription}
-
-\section{Task Coalescing}
-\label{sec:task_coalescing}
-
-\input{coalescing}
-% \input{newCoalescing}
-
-\section{Memory Virtualization}
-\label{sec:memory_virtulaization}
-
-\input{paging}
 

### 2018-04-10 paper/intro.tex
change of cite format

--- 2018-04-10 paper/intro.tex
@@
-Advances in processor efficiency along with the development of energy-harvesting systems has created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
-These devices operate using ambient energy, such as radio frequency transmissions~\cite{rf_powered_computing_gollakota_2014}, light~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
-Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo}, such devices are a promising technology for use in the Internet of Things~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat}.
+Advances in processor efficiency along with the development of energy-harvesting systems has created a new category of devices that require neither a battery nor a tethered power supply~\citep{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
+These devices operate using ambient energy, such as radio frequency transmissions~\citep{rf_powered_computing_gollakota_2014}, light~\citep{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\citep{gorlatova_sigmetrics_2014}.
+Incorporating compute, storage, sensing, and communication hardware~\citep{wisp5,moo}, such devices are a promising technology for use in the Internet of Things~\citep{ku_cst_2016}, in-body~\citep{nadeau_naturebio_2017} and on-body~\citep{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\citep{kicksat}.
 
-Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}. An
+Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\citep{hicks_isca_2017,lucia_snapl_2017}. An
 energy-harvesting device buffers energy in a small storage, e.g.
-a capacitor~\cite{gorlatova_tmc_2013},~\cite{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating.
+a capacitor~\citep{gorlatova_tmc_2013},~\citep{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating.
 Harvestable energy sources are low-power compared to a platform's operating level.
@@
 3c]{nadeau_naturebio_2017}.
-Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\cite{tan_infocom_2016},~\cite{mementos},~\cite{nvp}.
+Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\citep{tan_infocom_2016},~\citep{mementos},~\citep{nvp}.
 
@@
 %		\centering\large Figure
-%		\caption{\centering WISP~\cite{wisp} against an RFID antenna}\label{fig:1a}
+%		\caption{\centering WISP~\citep{wisp} against an RFID antenna}\label{fig:1a}
 %	\end{subfigure}%
@@
 %		\centering\large Figure
-%		\caption{\centering Program X~\cite{hicks_mibench2_2016} (split into X and Y tasks~\cite{chain}) at two distances}\label{fig:1b}
+%		\caption{\centering Program X~\citep{hicks_mibench2_2016} (split into X and Y tasks~\citep{chain}) at two distances}\label{fig:1b}
 %    \end{subfigure}
@@
 
-\textbf{Data Consistency in Intermittent Computing.} Software in an energy-harvesting system operates in the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, with execution interrupted by failure periods.
+\textbf{Data Consistency in Intermittent Computing.} Software in an energy-harvesting system operates in the {\em intermittent execution model}~\citep{dino,lucia_snapl_2017}, with execution interrupted by failure periods.
 When power fails, a device loses volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
-While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\cite{dino,edb}.
+While capturing periodic checkpoints~\citep{mementos,quickrecall} and sleep scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\citep{dino,edb}.
 
-There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\cite{hicks_isca_2017,idetic,nvp,tictpl}.
-New architectures require hardware changes and are inapplicable to today's systems~\cite{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
+There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\citep{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\citep{hicks_isca_2017,idetic,nvp,tictpl}.
+New architectures require hardware changes and are inapplicable to today's systems~\citep{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
 However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.
@@
 \textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
-task-based} programming and execution models~\cite{alpaca,chain} advocate for
+task-based} programming and execution models~\citep{alpaca,chain} advocate for
 static decomposition of a program (by a programmer) into a collection of tasks.
@@
 tasks affects the time the application takes to complete.
-Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
+Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\citep{chain,alpaca}.
 Naively minimizing the number of task transitions, (i.e.
@@
 %applications (the same application X, but divided into X and Y tasks as
-%in~\cite{chain}) running on RFID antenna-powered Computational
-%RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna
+%in~\citep{chain}) running on RFID antenna-powered Computational
+%RFID~\citep{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna
 %distances (near---stable energy supply/far---high energy intermittency). At
@@
 %
-Although WAR dependencies introduced by the \emph{static} task division can be eliminated at compile time~\cite{alpaca}, those introduced \emph{dynamically} at
+Although WAR dependencies introduced by the \emph{static} task division can be eliminated at compile time~\citep{alpaca}, those introduced \emph{dynamically} at
 run-time cannot. To maintain memory consistency in merged tasks, we propose a
@@
 
-%Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\cite{dino,chain} or may preclude the use of volatile memory~\cite{ratchet}.
-Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\cite{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\cite{wolverine}).
+%Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\citep{dino,chain} or may preclude the use of volatile memory~\citep{ratchet}.
+Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\citep{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\citep{wolverine}).
 How to connect this trade-off with task virtualization remains an open question.
@@
 
-We evaluated \sys end-to-end on a real energy-harvesting device~\cite{wisp} running existing benchmarks ~\cite{chain} and new programs.
-A comparison to a state-of-the-art task-based system~\cite{chain} showed that \sys recovers
+We evaluated \sys end-to-end on a real energy-harvesting device~\citep{wisp} running existing benchmarks ~\citep{chain} and new programs.
+A comparison to a state-of-the-art task-based system~\citep{chain} showed that \sys recovers
 performance on conservative task decompositions missed by systems with static decomposition.
@@
 While \sys's complier is still \emph{limited in scope}, e.g.
-is unable to handle instruction jumps, normalized loops and recursions, it already eliminates the labour-intensive process of writing tasks~\cite{chain,alpaca}, i.e., hours of programmer time.
+is unable to handle instruction jumps, normalized loops and recursions, it already eliminates the labour-intensive process of writing tasks~\citep{chain,alpaca}, i.e., hours of programmer time.
 

### 2018-04-10 paper/adaptiveModel.tex

@@
 % %% address the challenge and define the optimal task
-\sys advances its execution in a state-less (or virtual) manner, and then it frequently saves its forward progress.
+\textbf{Trade-off between Speed and Wasted Effort.} \sys advances its execution in a state-less (or virtual) manner, and then it frequently saves its forward progress.
 The longer \sys virtually progresses, the less committing (saving data to non-volatile memory) overhead it introduces.
@@
 As such, an optimal task is a task that occupies with a single commit an entire power cycle, which is therefore necessarily of a varying length.
+
+% \textbf{Trade-off between Speed and Wasted Effort.} \sys can coalesce an
+% arbitrary number of consecutive tasks. However, as more tasks coalesce, their
+% collective commit overhead amortizes better, but the risk of wasting work also
+% increases. If power fails during a long sequence of coalesced tasks, execution
+% will restart from the last commit, i.e. the first task in the sequence, losing
+% the progress made by any of the coalesced tasks.  The challenge to coalescing
+% tasks is determining how many tasks to coalesce before committing.
+
+\textbf{Task Coalescing.} 
+The need for \emph{adaptiveness}, that is coalescing static tasks, can be fulfilled by a naive strategy that adapts the target (coalesced task size) by a constant value.
+However, due to the linear behavior of the target adaptation, this approach is slow in reacting to the changes in energy conditions.
+Actually, even if an coalescing algorithm alters its target exponentially, We still can consider it as a slow algorithm.
+The reason for describing these algorithms as slow ones is that their targets enlarging and shrinking processes requires the same number of steps to to change from one target to another.
+For example, if a coalescing algorithm starts with a target size one and then double its target size after each successful commit, it requires five commits to reach the target 16.
+However, if this algorithm needs to shrink its target to one to preserve the forward progress it requires the system to fail five times!
+
+To This end, we can concluded that a good coalescing algorithm has to separate its adaptation processes.

### 2018-04-10 paper/intro.tex
last comments from OOPSLA 2018 branch by Alexei ported manually to main branch

@@
 3c]{nadeau_naturebio_2017}.
-Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\citep{tan_infocom_2016},~\citep{mementos},~\citep{nvp}.
+Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.
 
@@
 There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\citep{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\citep{hicks_isca_2017,idetic,nvp,tictpl}.
-New architectures require hardware changes and are inapplicable to today's systems~\citep{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
+\textcolor{red}{tictpl is not HW, is it?
+also it doesn't deal with inconsistency; this paragraph doesn't include Hibernus, so it should not include CPTL either.} New architectures require hardware changes and are inapplicable to today's systems~\citep{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
 However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.
@@
 static decomposition of a program (by a programmer) into a collection of tasks.
+\textcolor{red}{add \cite{intel} work; separate by-programmer and by-compiler; both cases hit the energy estimation problem; also intel work proposes a method for decomposing into one of many decompositions, but does not give a good solution for choosing one}
 Tasks can include arbitrary computation and, upon completion, are guaranteed to
@@
 incoming power always available, is to predict task energy consumption and
-choose a decomposition specific to the device capacitor size.
+choose a decomposition specific to the device capacitor size~\cite{cleancut}.
 However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution.
@@
 The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss.
-Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
-applications to an intermittent platform with a moderate effort.
+\textcolor{red}{Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several	applications to an intermittent platform with a moderate effort.}
 
@@
 \begin{itemize}
-	\item Two adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
+	\item \textcolor{red}{XXX} adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
 	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;

### 2018-04-10 paper/background.tex

--- 2017-11-17 paper/background.tex
@@
 We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
-\sys supports I/O behaviour similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.
+\sys supports I/O behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.
 
@@
 \label{sec:background_consistency}
+
+\textcolor{red}{reference work by Intel}
 
@@
 		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
+		~\cite{intel} & XXX \\
 		\hline
@@
 An excessively costly task on a device with a small energy buffer may be a relatively short, task on a device with a larger buffer.
-The key problem is that using existing systems, the programmer statically sizes tasks for a fixed energy buffer.
+The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead.
 \sys's {\em task coalescing}, described

### 2018-04-10 paper/abstract.tex
small modificaiton

@@
 % Connecting the work to a real-world problem
-Battery-free computers promise a sustainable Internet of things by forgo the batteries and scarf the energy from the environment. 
+%Intermittently powered computers promise a sustainable Internet of things by forgo the batteries and scarf the energy from the environment. 
+The Internet of Things (IoT) must be energy neutral. This can only be achieved when the batteries are removed and the energy to operate IoT comes from the ambient energy sources.
 % The main problem that we are tackling
-However, relying on the surrounding energy when it is available is at odds with current execution model of long-running computation. 
+However, relying on the highly varying, unpredictable and intermittent surrounding energy is at odds with the current programming model of a long-running computation: long computations require long periods of stable uninterrupted energy supply.
 % The already proposed solution to the main problem
-Scientists have proposed a discrete-like execution model that alternates between progressing the execution and protecting the execution progress, against power interrupts.
+The core programming model solution to the intermittently-powered computation a discrete-like execution that alternates between progressing and protecting the execution progress against power interrupts.
 % The problem of the proposed solutions 
-Generally, the proposed techniques can be classified as (i) checkpointing systems, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based systems, where a programmer splits the code into small idempotent sections.
-Despite the superiority of the task-based techniques, they suffer from a fixed overhead and they are completely blind to the energy conditions.
+Generally, the intermittent computation systems can be classified into (i) checkpointing systems, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based systems, where a programmer splits the code into small idempotent sections.
+Despite the superiority of the task-based systems over checkpointing they suffer from a fixed overhead and they are completely oblivious to the energy conditions.
 % Our main claim 
-In this paper we introduce \sys an \emph{efficient} and \emph{adaptive} task-based execution model.
+In this paper we introduce \sys: an \emph{efficient} and \emph{adaptive} task-based execution model.
 % show off
-\sys tunes its frequency of switching between progressing and protecting based on the available energy, but 
+\sys programs tune its frequency of switching between progressing and protecting based on the available energy, taking the burden of finding the right balance from the programer's shoulders.
 % One of the main functions of an OS is to abstract the hardware and to be less dependent on it. Coala does that too!
-tt does not require hardware support to refine its estimation. 
+\sys programs do not require hardware support to refine its estimation. 
 % main benefits of Coala
@@
 
-% old abstract 
+% old PLDI abstract 
 %Computation consistency on intermittently-powered devices can be ensured by: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) via tasks, where a programmer splits the code into small idempotent sections.

### 2018-04-10 paper/adaptiveModel.tex
brain dump of the conceptual ideas behind task coalescing
minor modifications

@@
 However, due to the linear behavior of the target adaptation, this approach is slow in reacting to the changes in energy conditions.
-Actually, even if an coalescing algorithm alters its target exponentially, We still can consider it as a slow algorithm.
-The reason for describing these algorithms as slow ones is that their targets enlarging and shrinking processes requires the same number of steps to to change from one target to another.
+Actually, even if a coalescing algorithm alters its target exponentially, We still can consider it as a slow algorithm.
+The reason for describing these algorithms as slow ones is that their targets enlarging and shrinking processes require the same number of steps to change from one target to another.
 For example, if a coalescing algorithm starts with a target size one and then double its target size after each successful commit, it requires five commits to reach the target 16.
@@
 
-To This end, we can concluded that a good coalescing algorithm has to separate its adaptation processes.
+To This end, we can conclude that a good coalescing algorithm has to separate its adaptation processes.
+In particular, a coalescing strategy should prevent repeated power failures that result from inadequate target estimation.
+In order for a coalescing strategy to do an educated guess, it has to be energy-aware.
+For that, it can rely on hardware support or on the history of execution.
+% We need to justify our choice of using the history of execution in the Coala sections. (generality of the software approach, and avoiding additional circuitry ...)
+
+\textcolor{red}{explain the idea behind the need for task splitting and how it must be done such that it is not limited by the commit size itself}

### 2018-04-10 paper/abstract.tex
abstract revised

--- 2018-04-09 paper/abstract.tex
@@
 % Connecting the work to a real-world problem
-Battery-free computers promise a sustainable Internet of things by forgo the batteries and scarf the energy from the environment. 
+%Intermittently powered computers promise a sustainable Internet of things by forgo the batteries and scarf the energy from the environment. 
+Massively deployed Internet of Things (IoT) must be energy neutral. This can only be achieved when the IoT is batteryless and the energy to operate the IoT comes from the ambient energy sources.
 % The main problem that we are tackling
-However, relying on the surrounding energy when it is available is at odds with current execution model of long-running computation. 
+However, relying on the highly varying, unpredictable and intermittent harvested energy is at odds with the principal programming model: program executions require stable and uninterrupted energy supply.
 % The already proposed solution to the main problem
-Scientists have proposed a discrete-like execution model that alternates between progressing the execution and protecting the execution progress, against power interrupts.
+The core programming model for the intermittently-powered computation is a discrete-like one that alternates between progressing and protecting the execution progress against power interrupts.
 % The problem of the proposed solutions 
-Generally, the proposed techniques can be classified as (i) checkpointing systems, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based systems, where a programmer splits the code into small idempotent sections.
-Despite the superiority of the task-based techniques, they suffer from a fixed overhead and they are completely blind to the energy conditions.
+Generally, the intermittent computation models can be classified into (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based programming, where a programmer splits the code into small idempotent sections.
+Despite the superiority of the task-based systems over checkpointing, expressed in lower memory overhead, task-based programming is completely oblivious to the energy conditions.
+Simply, it cannot increase or decrease the size of the task on the fly (the moment of commit to the non-volatile memory) depending on how long the energy burst is.
 % Our main claim 
-In this paper we introduce \sys an \emph{efficient} and \emph{adaptive} task-based execution model.
+To address this challenge, this paper introduces \sys: \emph{adaptive} task-based execution model that does not require any dedicated hardware to support task adaptability, which at the same time is \emph{efficient}.
 % show off
-\sys tunes its frequency of switching between progressing and protecting based on the available energy, but 
-% One of the main functions of an OS is to abstract the hardware and to be less dependent on it. Coala does that too!
-tt does not require hardware support to refine its estimation. 
+\sys programs tune frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period from the programer's shoulders.
 % main benefits of Coala
@@
 % I really hope it will support our claims!! haha
-Results show the \sys is more efficient than the current state-of-art execution model by X-times and more reliable.
+Results show the \sys is more efficient than the current state-of-art execution model by \textcolor{red}{X times}.
 
-% old abstract 
+% old PLDI abstract 
 %Computation consistency on intermittently-powered devices can be ensured by: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) via tasks, where a programmer splits the code into small idempotent sections.

### 2018-04-11 paper/intro.tex
pass over the intro.  it is too long, but the content is there

--- 2018-04-10 paper/intro.tex
@@
-Advances in processor efficiency along with the development of energy-harvesting systems has created a new category of devices that require neither a battery nor a tethered power supply~\citep{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
-These devices operate using ambient energy, such as radio frequency transmissions~\citep{rf_powered_computing_gollakota_2014}, light~\citep{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\citep{gorlatova_sigmetrics_2014}.
-Incorporating compute, storage, sensing, and communication hardware~\citep{wisp5,moo}, such devices are a promising technology for use in the Internet of Things~\citep{ku_cst_2016}, in-body~\citep{nadeau_naturebio_2017} and on-body~\citep{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\citep{kicksat}.
+Advances in processor efficiency along with the development of
+energy-harvesting systems has created a new category of devices that require
+neither a battery nor a tethered power
+supply~\citep{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These
+devices operate using ambient energy, such as radio frequency
+transmissions~\citep{rf_powered_computing_gollakota_2014},
+light~\citep{margolies_infocom_2016,margolies_tosn_2016}, and
+vibration~\citep{gorlatova_sigmetrics_2014}. Incorporating compute, storage,
+sensing, and communication hardware~\citep{wisp5,moo,capybara}, such devices are a
+promising technology for use in the Internet of Things~\citep{ku_cst_2016},
+in-body~\citep{nadeau_naturebio_2017} and
+on-body~\citep{bandodkar_electroanalysis_2015} medical systems, and
+energy-harvesting nano-satellites~\citep{kicksat,capybara}.
 
-Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\citep{hicks_isca_2017,lucia_snapl_2017}. An
-energy-harvesting device buffers energy in a small storage, e.g.
-a capacitor~\citep{gorlatova_tmc_2013},~\citep{gunduz_commag_2014} and when a threshold amount of energy accumulates, begins operating.
-Harvestable energy sources are low-power compared to a platform's operating level.
-A device operates briefly and when buffered energy is depleted, shuts down and recharges to operate again later.
-As an example, recharge time may be tens of seconds in radio frequency powered medical device~\cite[Fig.
+Energy-harvesting devices create unique challenges because they operate {\em
+intermittently} when energy is
+available~\citep{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
+buffers energy in a small storage capacitor~\citep{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
+threshold amount of energy has accumulated. 
+%
+Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
+power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
+down and recharges to operate again later. As an example, recharge times may be
+tens of seconds in radio frequency powered medical device~\cite[Fig.
 3c]{nadeau_naturebio_2017}.
-Moreover, charge and discharge times vary by device (due to different capacitor sizes) and some may fail $\approx$10 to $\approx$100 times per second~\citep{tan_infocom_2016},~\citep{mementos},~\citep{nvp}.
+The recharge and discharge time --- which corresponds to the device's inactive and active time --- varies with the energy buffering capacitor's~\cite{capybara} and some devices fail and restart operating $\approx$10 to
+$\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.
 
@@
 
-\textbf{Data Consistency in Intermittent Computing.} Software in an energy-harvesting system operates in the {\em intermittent execution model}~\citep{dino,lucia_snapl_2017}, with execution interrupted by failure periods.
-When power fails, a device loses volatile state, i.e., registers, stack, SRAM, and retains its non-volatile state, i.e., FRAM, Flash.
-While capturing periodic checkpoints~\citep{mementos,quickrecall} and sleep scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave volatile and non-volatile state inconsistent, leading to unrecoverable failures~\citep{dino,edb}.
+\textbf{Data Consistency in Intermittent Computing.}  Software in an
+energy-harvesting system operates according to the {\em intermittent execution
+model}~\citep{dino,lucia_snapl_2017}, which corresponds to the
+operation-failure-restart cycles of an energy-harvesting device. In an
+intermittent software execution, an {\em operating period} proceeds for an
+arbitrary duration (dictated by energy availability) before being interrupted
+by a {\em failure period}. In a failure period, a device loses the volatile
+state in its registers, stack, SRAM, and retains the state of any non-volatile
+memory, such as FRAM or Flash. While capturing periodic
+checkpoints~\citep{mementos,quickrecall} and sleep
+scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution
+progress, failures can leave non-volatile state incorrectly, partially updated,
+and may leave checkpointed volatile state inconsistent with non-volatile state.
+These inconsistencies cause intermittent executio to deviate from
+continuously-powered behavior, often leading to an unrecoverable
+application failure~\citep{dino,edb}.
 
-There are two main approaches for dealing with data inconsistency for intermittently-powered devices: (i) \emph{programming and execution models}~\citep{dino,ratchet,chain,alpaca} and (ii) \emph{architectures}~\citep{hicks_isca_2017,idetic,nvp,tictpl}.
-New architectures require hardware changes and are inapplicable to today's systems~\citep{hicks_isca_2017,nvp}, therefore new programming models and compilers are more suitable (as of now).
-However, they have their own limitations that need to be addressed---the core one being the rigidity of the \emph{static} decomposition into tasks.
+Prior work developed two main approaches to dealing with data inconsistency for
+intermittently-powered devices: (i) \emph{software-based programming and
+execution models}~\citep{dino,ratchet,chain,alpaca} and (ii)
+\emph{hardware-based computer architecture
+support}~\citep{hicks_isca_2017,idetic,nvp}.  Complex architectural changes are
+expensive to design, verify, and manufacture.  New architectures are also
+inapplicable to existing systems~\citep{hicks_isca_2017,nvp}. Software
+approaches are simpler and applicable to existing devices today, but prior
+software approaches have a number of limitations that are the focus of the work
+in this paper.  A key limitation of prior software approaches that we address in this paper is the {\em inflexibility}
+of a program statically decomposed into tasks.
+
+%\textcolor{red}{tictpl is not HW, is it? also it doesn't deal with inconsistency; this paragraph doesn't include Hibernus, so it should not include CPTL either.}
 
@@
 
-\textbf{Task Decomposition of Intermittent Programs.} Recently proposed {\em
-task-based} programming and execution models~\citep{alpaca,chain} advocate for
-static decomposition of a program (by a programmer) into a collection of tasks.
-Tasks can include arbitrary computation and, upon completion, are guaranteed to
-have executed {\em atomically}, despite arbitrarily-timed power failures.
-The programmer explicitly expresses task-to-task control flow. The execution
-in Figure~\ref{fig:coalesce} (top) illustrates how the decomposition into
-tasks affects the time the application takes to complete.
-Each task transition introduces the overhead of tracking and atomically committing the modifications to non-volatile memory, to maintain consistency of program state~\citep{chain,alpaca}.
-Naively minimizing the number of task transitions, (i.e.
-the task count), \emph{at compile time} risks creating a task that requires more energy than the device can buffer in its energy storage.
-Such a large task would fail to complete unless the environment provides sufficient incoming energy to supplement the energy in the capacitor.
-To eliminate this risk, the programmer (or the compiler) must decompose the program conservatively into many small tasks, at the cost of \emph{higher task transition overhead}.
+%Moved intel note to related work
+\textbf{Task Decomposition of Intermittent Programs.} {\em
+Task-based} programming and execution models require a
+programmer~\citep{alpaca,chain} or a compiler~\cite{intelehcompiler} to
+statically decompose a program into a collection of {\em tasks}.  A task can
+include arbitrary computation and these existing systems guarantee that each
+tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
+%
+The programmer explicitly expresses task-to-task control flow, or the compiler
+converts existing program control-flow into task-to-task control flow to sequence 
+the tasks that it defines.
+%
+Figure~\ref{fig:coalesce} (top) illustrates how a program's tasks execute and
+shows how tasks can impose a run-time overhead. 
+%
+At each task transition, the system incurs an overhead to track and atomically
+commit modifications to the non-volatile memory, to maintain consistency of
+program state~\citep{chain,alpaca}.  
+%
+The more task transitions a program executes at run time, the more overhead.
 
-One approach to decrease the task count, without assuming a minimum level of
-incoming power always available, is to predict task energy consumption and
-choose a decomposition specific to the device capacitor size.
-However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution.
-Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.
+A savvy programmer may thus create very large tasks in an attempt to
+minimize overhead by minimizing the number of task transitions.  However, a
+very large task may require more energy to complete than a device's fixed
+hardware energy buffer can hold.  Such a large task fails to completely execute
+assuming only buffered energy is available (i.e., assuming the worst case).
+While some environments may have sufficient harvestable energy replenish the
+capacitor during operation, enabling the task to complete, assuming cooperation
+from the physical environment to avoid non-termination is a risky programming
+proposition.  To eliminate this risk, existing systems require the programmer
+or compiler to decompose a  program into small tasks, all of which complete
+using only buffered energy.  These constraints on task sizing leave an
+unsatisfying dilemma: large, efficient tasks that risk non-termination, or
+small tasks that are guaranteed to complete, but incur a high task transition
+and commit overhead.  In this work, we propose a third way, using a novel
+technique called {\em dynamic task coalescing} that efficiently executes small
+tasks, avoiding unnecessary overheads and avoiding the risk of non-termination.
+
+%Brandon: I moved Cleancut to the related work section.  THere's too much in the intro already
 
@@
 
-\textbf{Task Coalescing.} To address this challenge we propose a \emph{dynamic} task decomposition, that accepts any static decomposition and opportunistically coalesces tasks at runtime when there is sufficient energy.
-With \emph{task coalescing} atomic tasks are merged into a single (larger) task while preserving atomicity of the merged task.
-If power fails while a coalesced task is running, execution will restart at the first task in the group.
-Figure~\ref{fig:coalesce} (bottom) illustrates how dynamic coalescing changes the earlier execution and highlights the change in memory model necessary to not break atomicity.
+\textbf{Task Coalescing.} In this work, we introduce a new system called \sys, which executes a task-based
+intermittent program by dynamically coalescing its tasks.
+\sys accepts any static decomposition (i.e., from the programmer or from a compiler) 
+and coalesces its tasks at runtime. Two consecutive, coalesced tasks execute  
+with no commit or task transition overhead between them, instead performing
+task-end commit actions at the end second task only.
+%
+When there is sufficient energy to execute both tasks, two distinct, small
+tasks are effectively executed as a single large task, preserving the atomicity
+of both.  If power fails during a coalesced task, execution restarts from the
+{\em first} of the coalesced tasks.  Figure~\ref{fig:coalesce} (bottom)
+illustrates how dynamic coalescing changes the earlier execution and highlights
+the change in memory model necessary to not break atomicity.
 
-Unfortunately, naively merging two atomic tasks might produce a non-atomic task, that may leave non-volatile memory inconsistent after a partial execution.
-A merge breaks atomicity when atomic merged tasks form a \emph{write-after-read (WAR)} dependency---for instance if two tasks \texttt{\{x=y+1\}; \{y=x;\}} are merged and if the power failure occurs after \texttt{x=y}, the value \texttt{x} will be increased twice when the merged task is restarted, that leads to an inconsistency.
+%Brandon: This is too detailed for the intro:
+%Unfortunately, naively merging two atomic tasks might produce a non-atomic
+%task, that may leave non-volatile memory inconsistent after a partial
+%execution. A merge breaks atomicity when atomic merged tasks form a
+%\emph{write-after-read (WAR)} dependency---for instance if two tasks
+%\texttt{\{x=y+1\}; \{y=x;\}} are merged and if the power failure occurs after
+%\texttt{x=y}, the value \texttt{x} will be increased twice when the merged task
+%is restarted, that leads to an inconsistency.
+
 %
@@
 %
-Although WAR dependencies introduced by the \emph{static} task division can be eliminated at compile time~\citep{alpaca}, those introduced \emph{dynamically} at
-run-time cannot. To maintain memory consistency in merged tasks, we propose a
-mechanism for \emph{memory virtualization} that buffers the updates to non-volatile variables in volatile memory before committing them at the dynamic task boundary.
+%Brandon: Sentence too long.
+While a compiler can use static data privatization and commit instrumentation
+(i.e., redo-logging) to eliminate statically identifiable, inter-task data
+dependences~\citep{alpaca}, a \emph{dynamic} dependence between two coalesced
+tasks requires dynamic privatization and commit actions. 
+%
+To ensure consistency and respect inter-task dependences, \sys uses a novel
+approach to \emph{memory virtualization} that buffers non-volatile variable
+updates in volatile memory during coalesced task execution, before committing
+them to non-volatile memory at the dynamic task boundary.
 %
@@
 
-\textbf{Task Splitting}: The static task execution model assumes that a single task execution is guaranteed by the hardware.
-If this assumption is violated, due to hardware inconsistency or inaccurate task devision startegy, the system enters a non-termination loop of failed attempts to reach the next task boundery.
-logically, better than relying on a rondom process, energy shots arrival, to break out of this loop, the system should execute an "emergency-like" routine to terminate it and preserve forward progress.
+\textbf{Task Splitting}: A static task decomposition model assumes that each
+single task can execute to completion.  If the hardware energy buffer provides
+inadequate energy to execute each single task to completion, a program will not
+terminate~\cite{cleancut}. To avoid non-termination under adversarial 
+energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
+Partial commit avoids non-termination by committing the intermediate state of a long-running
+task that has repeatedly failed and restarted.  Partial commit violates
+task atomicity, but preserves forward progress; if a programmer knows that
+task atomicity is crucial to correctness, they can disable partial commit
+instead risking non-termination.
 
-Task coalescing removes the burden from the programmer, because it accepts any task decomposition and improves it dynamically.
-To further reduce the programmer effort, we propose a compiler pass for \emph{automatic decomposition} of programs into (small) \emph{atomic} tasks.
-The compiler identifies non-volatile variables shared across tasks as tasks are created, and instruments reads and writes of those variables, using memory virtualization, to keep the data consistent in the presence of power loss.
-Despite being limited to a subset of the C language, the automatic task decomposition allowed us to port several
-applications to an intermittent platform with a moderate effort.
+%Task coalescing removes the burden from the programmer, because it accepts any
+%task decomposition and improves it dynamically. To further reduce the
+%programmer effort, we propose a compiler pass for \emph{automatic
+%decomposition} of programs into (small) \emph{atomic} tasks. The compiler
+%identifies non-volatile variables shared across tasks as tasks are created, and
+%instruments reads and writes of those variables, using memory virtualization,
+%to keep the data consistent in the presence of power loss.
+%
+%\textcolor{red}{Despite being limited to a subset of the C language, the
+%automatic task decomposition allowed us to port several	applications to an
+%intermittent platform with a moderate effort.}
 
-% <<<<<<< HEAD
-% We develop \sys%
-% \footnote{To be released at \url{http://anonymized.link}}
-% %
-% : a task coallescing system for task-based
-% programming models on intermittently-powered devices. The
-% capabilities required for effective coallescing mark our
-% contributions:
-% >>>>>>> 96e5d9f9b4e97ce8700c375bd8e11708870f5aba
-% =======
-\textbf{Contributions.} We develop \sys\footnote{To be released at \url{http://anonymized.link}}: the first task coalescing system for task-based programming models on intermittently-powered devices.
-The capabilities required for effective coalescing mark our contributions:
-% >>>>>>> 84d7711b6c6f6424cdad97efc8aad96e52f316ff
+\textbf{Contributions.} We develop \sys\footnote{To be released at
+\url{http://anonymized.link}}: the first task coalescing system for task-based
+programming models on intermittently-powered devices. The capabilities required
+for effective coalescing mark our contributions:
 
 \begin{itemize}
-	\item Two adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
+	\item \textcolor{red}{XXX} adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
 	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
@@
 
-We evaluated \sys end-to-end on a real energy-harvesting device~\citep{wisp} running existing benchmarks ~\citep{chain} and new programs.
-A comparison to a state-of-the-art task-based system~\citep{chain} showed that \sys recovers
-performance on conservative task decompositions missed by systems with static decomposition.
+We evaluated \sys end-to-end on a real energy-harvesting device~\citep{wisp}
+running existing benchmarks~\citep{chain,alpaca} and new programs. A comparison
+to a state-of-the-art task-based system~\citep{chain,alpaca} showed that \sys
+recovers performance on conservative task decompositions missed by systems with
+static decomposition.
 %

### 2018-04-11 paper/intro.tex
wrapfig for reasonable fig size

@@
 
-\begin{figure}
-	\includegraphics[width=0.95\columnwidth]{figures/coalescing-is}
-    \caption{\label{fig:coalesce}\textbf{An execution of three
+\begin{wrapfigure}{t!}{0.5\textwidth}
+    \centering
+    \includegraphics[width=0.5\textwidth]{figures/coalescing-is}
+    \caption{\textbf{An execution of three
 statically-defined tasks without (top) and with (bottom) dynamic
@@
 buffering updates to non-volatile (NV) variables in volatile (V) memory.}
-\end{figure}
+    \label{fig:coalesce}
+\end{wrapfigure}
 

### 2018-04-11 paper/intro.tex
cleancut and intel paper added to bib (Brandon: intelehcompiler ->baghsorkhi_cgo_2018)

@@
 Task-based} programming and execution models require a
-programmer~\citep{alpaca,chain} or a compiler~\cite{intelehcompiler} to
+programmer~\citep{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
 statically decompose a program into a collection of {\em tasks}.  A task can

### 2018-04-11 paper/implementation.tex
placeholder/braindump for coala implementation

\textcolor{red}{THIS IS UNEDITED AND SHOULD NOT BE READ YET - Placeholder}

Coala's API adds only 5 syntactic constructs to C-based languages: \texttt{COALA\_TASK}, used to declare a new task; \texttt{COALA\_PV}, used to declare a new protected variable; \texttt{COALA\_NEXT\_TASK} used to mark the next task to be scheduled after the currently executing one completes; \texttt{RP}, used to mark a protected variable read; \texttt{WP}, used to mark a protected variable write.

Moreover, the programmer is required to explicitly call two API functions inside the main body: \texttt{COALA\_INIT}, passing the global initialization task to run upon the first boot as argument; \texttt{COALA\_RUN}, which hands the control to Coala's scheduler.

Finally, when the programmes intends to typedef a struct, to use it as a protected variable, they need to declare each structure's member by using the following macro: \texttt{COALA\_SM}, used to ensure proper memory alignment required by Coala's page handler.

The following table summarizes Coala's API methods and their arguments. $T$ denotes the set of all tasks, while $V$ denotes the set of all protected variables. These methods trigger specific kernel functions, whose detailed implementation is given in the next paragraph.

\begin{table}
\centering
\begin{tabular}{| c | c |}
	\hline
	\textbf{Method} & \textbf{Arguments} \\
	\hline\hline
	\texttt{COALA\_TASK}($t$, $w_t$) & $t \in T$: new task name, $w_t$: weight of task $t$ \\
	\hline
	\texttt{COALA\_PV}(type, $v$ [, size]) & type: var. type, $v \in V$: new protected var. name, size (optional)  array size\\
	\hline
	\texttt{COALA\_NEXT\_TASK}($t$) & $t \in T$ : next task to schedule\\
	\hline
	$u$ := RP($v$) & $v \in V$: protected variable to be read, $u$: value \\
	\hline	
	WP($v$) := $u$ &  $v \in V$: protected variable to be written, $u$: value \\
	\hline
	\texttt{COALA\_INIT}(t) & $t \in T$: task to be scheduled on first boot \\
	\hline
	\texttt{COALA\_RUN}() & {} \\
	\hline
	\texttt{COALA\_SM}(type, $m$ [, size]) &  type: struct member's type, $m$: struct member's name, size (optional): array size \\
	\hline
\end{tabular}
\caption{Coala API}
\label{table:coala_api}
\end{table}

An example usage of Coala’s API is given in the following code snippet. (TBW)

\subsection{Kernel functions implementation}

New tasks. The API method \texttt{COALA\_TASK} statically allocates a non-volatile constant variable holding the task weight, and declares a void function with the name passed as an argument.

New protected variables. The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by RP and WP to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in non-volatile memory.

Task transitions. By calling \texttt{COALA\_NEXT\_TASK}, the user tells Coala’s scheduler what task to execute next. Note that the currently executing is not interrupted, and only once it completes the next task is run.

Protected reads and writes. When RP and WP are used, Coala searches the volatile memory for the page the protected variable belongs to. If such page is not found in RAM, then a page fault occurs, and the page is fetched from the non-volatile memory. The two functions return a handle to the variable’s volatile memory location. Unlike RP, WP marks the page being written as dirty, so that it can be committed to its non-volatile location.

Initialization and origin task. The behavior of the method \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the addition that all the preliminary kernel initializations are performed inside this function. Moreover, the origin task is scheduled only upon the first boot.

Execution. The application control is passed to Coala’s scheduler by calling \texttt{COALA\_RUN} at the end of the main function. A high-level state machine of the scheduler is depicted in Figure \textcolor{red}{[draw and ref figure]}. Depending on the coalescing algorithm, state transitions follow a different pattern.

\subsection{Paging}

Memory virtualization. Working (volatile) buffer, store and shadow (non-volatile) buffers. No major changes since last version.

Page allocation and page tags. When the application accesses a protected variable with RP or WP, Coala has to search for the variable in RAM first, and in non-volatile memory if it not found in RAM. Actually, the kernel searches for pages, not single variables. A page search can be very fast if each page is identified by a unique tag, and if every variable belonging to a page bears the tag in its address. Thus, retrieving a page tag from a protected variable’s address boils down to extracting the upper bits of the address itself. The remaining lower bits denote the variable’s offset in its page.

\begin{figure}
	\caption{Illustration of page tags \textcolor{red}{to be drawn}}
\end{figure}

In order to obtain unique page tags, some alignment requirements have to be met. First, the page size $S$ has to be a power of two. Moreover, pages in non-volatile memory have to be aligned by $S$. In the example above, $S$ is 256, and every page’s first address is a multiple of 256. Note that, by aligning pages by $2n$, the alignment is also valid for all sizes $2^x$ with $x \leq n$.

Page fault and page swap (with DMA). When a page is not found in RAM, it is fetched from FRAM. No major changes since last version.

\subsection{Scheduling and committing}

Target budget initialization. Upon a reboot, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy [...].

Coalescing state recall. The last (coalesced) task has to be recalled. More specifically, the function pointer of the last coalesced task has to be retrieved from NV memory, where is had been saved during the last successful commit. [...].

Partial execution restoration. In case there is a valid partial execution checkpoint, the program has to resume from there [...].

Coalesced scheduling. Static tasks are coalesced and run [...].

Intermediate commit and ultimate commit. When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to non-volatile memory. Since last version, no major changes were made to the intermediate commit (SRAM to shadow buffer), but an optimization was applied to the ultimate commit (shadow buffer to store buffer). Store buffer and shadow buffer used to be physically separated in memory. Now, the role of a block of memory (i.e. a page) is interchangeable, meaning that, during the ultimate commit, the shadow buffer pages that were written during the previous commit phase become store buffer pages, and their relative store buffer pages become shadow buffer pages. A page’s role is retrieved using an indirection table, i.e. an array that holds a zero for all the pages whose store buffer resides in the lower memory block, and a one for all the pages whose store buffer is located in the upper memory block. This allows for a lighter commit process, since fewer page moves are performed.

\begin{figure}
	\caption{Intermediate commit and ultimate commit Illustration \textcolor{red}{to be drawn}}
\end{figure}

Target budget update. Before starting a new coalesced task, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy [...].

### 2018-04-12 paper/systemdescription.tex
re-wrote the overview part

--- 2017-11-16 paper/systemdescription.tex
@@
-\sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
-\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em accessible}, {\em efficient} and {\em flexible}.
-\sys accomplishes this goal with a constellation of a new programming model, compiler, and run time software system support, refer to Figure~\ref{fig:system_overview}.
+\sys is a new programming and execution model for intermittent computing on
+energy-harvesting devices. \sys addresses the challenges outlined in
+Section~\ref{sec:background} to make task-based intermittent programs {\em
+programmable} and {\em efficient}. \sys accomplishes this goal with a
+constellation of a new programming model, compiler, and run time software
+system support, that supports dynamically adaptive task-based execution.
+Figure~\ref{fig:system_overview} shows an overview of \sys.
 
@@
 
-\textbf{\sys Programming Model.} To use \sys, a programmer first writes plain, imperative C code.
-The programmer then has an option to {\em manually} or {\em automatically} translate the program into code for \sys's task-based programming model.
-To manually translate, the programmer decomposes the program into tasks, and annotates memory accesses that manipulate data shared by multiple tasks.
-To automatically translate, the programmer simply uses \sys's compiler (described in Section~\ref{sec:compiler}), which decomposes a program into tasks and annotates accesses to task-shared data.
-After translating to task-based code and compiling, the programmer has an executable \sys binary.
+\textbf{\sys Programming and Execution Model.} 
+To use \sys, a programmer first writes plain, imperative C code. The programmer
+then must decompose the program into tasks. To manually decompose a program
+into tasks, the programmer designates a set of functions as tasks, sequences
+control-flow between these tasks, and annotates memory accesses that manipulate
+data shared by multiple tasks.
 
-\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists of
-constructs for defining tasks~\cite{chain,alpaca} and accessors for
-protected variables shared between tasks. These anotations are added either by
-the programmer (in the manual translation mode) or the compiler (in automatic
-translation mode). As in prior task-based intermittent programming
-models~\cite{chain,alpaca}, tasks are functions that are not allowed to have
-any direct callers but receive control by explicit {\tt next\_task} statements.
-In \sys, a task transition atomically commits the memory operations of the
-current task.  The programmer identifies each task by adding the {\tt task}
-keyword to the function's declaration and designates one task as the {\em
-origin} task at which execution begins. After a power failure,
-\sys restarts from the last (partially) executed task.
+\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists
+of primitives for defining tasks~\cite{chain,alpaca} and accessors for
+protected variables shared between tasks. As in prior task-based intermittent
+programming models~\cite{chain,alpaca}, tasks are functions that are not
+allowed to have any direct callers but receive control via explicit {\tt
+next\_task} statements only.  The programmer identifies each task
+by adding the {\tt task} keyword to the function's declaration and designates
+one task as the {\em origin} task at which execution begins. After a power
+failure, \sys restarts the last partially executed task.
 %
@@
 task---using the \emph{read variable} ({\tt RVAR}) and \emph{write variable}
-({\tt WVAR}) operations. These operations convey to \sys's memory manager which
-data must be buffered in volatile memory and committed.
+({\tt WVAR}) operations. These operations convey to \sys that
+the referenced data are task-shared, meaning that they must be buffered in
+volatile memory and committed when the executing task ends.
+%
+Commit is a key overhead associated with task-based operation and \sys's
+coalescing strategy (described in Section~\ref{sec:coalescing} avoids commit
+when a task ends by continuing to buffer task-shared variables for multiple
+coalesced tasks, and committing only when all tasks complete.
 
@@
 
-\textbf{\sys Memory Manager.} At runtime \sys's \emph{virtualizing memory
-manager} (Section~\ref{sec:memory_virtulaization}) protects changes to
-non-volatile data against becoming inconsistent after a power failure.  Relying
-on annotations of accesses to protected variables, \sys's memory manager {\em
-pages} these data into volatile memory from non-volatile memory as they are
-needed.  Tasks cannot modify protected variables directly in non-volatile
-memory.  If the working set exceeds the size of volatile memory, pages are
-swapped out to non-volatile memory without committing the changes, leaving
-original locations of protected variables unchanged.  When a coalesced task
-completes, \sys atomically commits modified pages to their original locations
-in non-volatile memory.
+The programmer can decompose a program into tasks either {\em manually} or {\em
+automatically}. To do so manually,  requires reasoning similar to prior
+task-based systems~\cite{chain,alpaca}.  
+%
+A programmer may also opt to use compiler support to automatically decompose a
+program into tasks, leveraging recent work~\cite{cleancut,baghsorkhi_cgo_2018}.
+%
+Without loss of generality, we assume throughout this work that the programmer
+manually decomposed the program into tasks; \sys's behavior with automatically
+decomposed code would be identical.
+%
 
-\textbf{\sys Task Scheduler.} The scheduler {\em coalesces}
-statically\hyp{}defined tasks into dynamic tasks sized to match the available
-energy. By default, tasks run in sequence and each task commits as it
-completes, potentially incurring unnecessary overhead for commits between
-consecutive, non-failing tasks. \sys dynamically coalesces consecutive tasks by
-deferring the commit of an earlier task until the completion of a later task.
-\sys's memory management mechanism is essential for coalescing, because commits
-are deferred by keeping dirty pages in volatile memory across coalesced
-boundaries.  After a coalesced sequence of tasks completes, \sys commits pages
-updated by all of the tasks to non-volatile memory, reducing the commit
-overhead through batching.
+Regardless of whether automatically or manually decomposed into tasks, the
+programmer compiles their task-based code, and links to the \sys runtime,
+producing a \sys-enabled binary.
+%
+The \sys runtime implements \sys's task-based programming and execution model.
+%
+The runtime system also includes \sys's novel {\em virtualizing memory manager}
+and its {\em task coalescing manager}, both of which are essential to \sys's
+adaptive task-based execution.
+
+\textbf{\sys Task Coalescing Manager.} \sys relies on its task coalescing
+manager to dynamically {\em coalesce} statically\hyp{}defined tasks to avoid
+inessential overheads associated with completing tasks.  
+%
+By default, tasks run in a sequence and each task commits its task-shared state
+as it completes.  The time and energy cost of a task's commit is unnecessary if
+the task and its successor both complete without a power failure.  The key
+insight that \sys leverages is that the first task could have deferred its
+commit to the second task. {\em coalescing} tasks by deferring the first commit
+avoids the fixed cost of the first commit (task transitioning overheads).
+Coalescing also amortizes per-variable commit cost, committing each  location
+accessed by both tasks only after the second task, rather than once per task.
+%
+
+\textbf{\sys Virtual Memory Manager.} 
+\sys is able to efficiently coalesce tasks because of its efficient virtual
+memory manager,
+%
+which is described in Section~\ref{sec:memory_virtualization}. 
+%
+\sys's memory manager paginates memory and ensures that data in a page remain consistent
+despite power interruptions. 
+%
+\sys allows a task to manipulate data in a volatile copy of a page only.
+%
+Pages swap between volatile and non-volatile memory, depending on the capacity
+of the volatile memory and the program's access pattern. 
+%
+\sys tracks a task's memory accesses efficiently at page granularity (rather
+than using, e.g., costly word-granular tracking).
+%
+When a task ends, each page it accessed commits from volatile memory (or from a
+non-volatile swap region for dirty pages) back to the non-volatile main memory.
+%
+Pages efficiently, atomically commit using a two-phase commit procedure
+accelerated using hardware support for direct memory access (DMA). 
+%

### 2018-04-12 paper/intro.tex
edits in intro and todos in relatedwork

@@
 and may leave checkpointed volatile state inconsistent with non-volatile state.
-These inconsistencies cause intermittent executio to deviate from
+These inconsistencies cause intermittent execution to deviate from
 continuously-powered behavior, often leading to an unrecoverable
@@
 
-A savvy programmer may thus create very large tasks in an attempt to
-minimize overhead by minimizing the number of task transitions.  However, a
-very large task may require more energy to complete than a device's fixed
-hardware energy buffer can hold.  Such a large task fails to completely execute
-assuming only buffered energy is available (i.e., assuming the worst case).
-While some environments may have sufficient harvestable energy replenish the
-capacitor during operation, enabling the task to complete, assuming cooperation
-from the physical environment to avoid non-termination is a risky programming
-proposition.  To eliminate this risk, existing systems require the programmer
-or compiler to decompose a  program into small tasks, all of which complete
-using only buffered energy.  These constraints on task sizing leave an
-unsatisfying dilemma: large, efficient tasks that risk non-termination, or
-small tasks that are guaranteed to complete, but incur a high task transition
-and commit overhead.  In this work, we propose a third way, using a novel
-technique called {\em dynamic task coalescing} that efficiently executes small
-tasks, avoiding unnecessary overheads and avoiding the risk of non-termination.
+A savvy programmer may thus create very large tasks in an attempt to minimize
+overhead by minimizing the number of task transitions.  However, a very large
+task may require more energy to complete than a device's fixed hardware energy
+buffer can hold.  Such a large task fails to completely execute assuming only
+buffered energy is available (i.e., assuming the worst case).  While some
+environments may have sufficient harvestable energy to replenish a device's
+capacitor during operation, prolonging the execution period and enabling the
+task to complete, assuming cooperation from the physical environment to avoid
+non-termination is a risky programming proposition.  To eliminate this risk,
+existing systems require the programmer or compiler to decompose a  program
+into small tasks, all of which complete using only buffered energy.  These
+constraints on task sizing leave an unsatisfying dilemma: large, efficient
+tasks that risk non-termination, or small tasks that are guaranteed to
+complete, but incur a high task transition and commit overhead.  In this work,
+we propose a third way, using a novel technique called {\em dynamic task
+coalescing} that efficiently executes small tasks, avoiding unnecessary
+overheads while avoiding the risk of non-termination.
 
@@
 
-\textbf{Task Coalescing.} In this work, we introduce a new system called \sys, which executes a task-based
+In this work, we introduce a new system called \sys, which executes a task-based
 intermittent program by dynamically coalescing its tasks.
@@
 
-\textbf{Task Splitting}: A static task decomposition model assumes that each
+A static task decomposition model assumes that each
 single task can execute to completion.  If the hardware energy buffer provides
@@
 
-\textbf{Contributions.} We develop \sys\footnote{To be released at
-\url{http://anonymized.link}}: the first task coalescing system for task-based
-programming models on intermittently-powered devices. The capabilities required
-for effective coalescing mark our contributions:
+\textbf{Contributions.} To summarize, \sys's main contributions are:
 
 \begin{itemize}
-	\item \textcolor{red}{XXX} adaptive coalescing strategies for selecting boundaries to coalesce based on past execution behaviour;
-	\item Software memory virtualisation for transiently-powered devices that preserves atomicity of coalesced tasks;
-	\item Task splitting startegy that preserve forward progress and memory consistency;
+\item The first {\em adaptive} task-based intermittent execution model that dynamically coalesces tasks to avoid unnecessary commit overhead. 
+\item The first software memory virtualization mechanism for an intermittent system that \sys uses to preserve the atomicity of coalesced tasks.
+\item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
+\item A fully-realized prototype of \sys that runs on real energy-harvesting devices~\cite{wisp,capy}, to be released as open source after publication. 
+\item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work~\cite{alpaca}, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to varied energy conditions.
 \end{itemize}
 
-%Existing programming and execution models may require more accesses to non-volatile memory (for multi-versioning)~\citep{dino,chain} or may preclude the use of volatile memory~\citep{ratchet}.
-Effective use of both volatile and non-volatile memory is important for \emph{efficiency}, because non-volatile memory has higher access energy and latency than SRAM~\citep{nvp}, and \emph{generality}, because devices often have much more FRAM than SRAM (e.g., 64 times more in~\citep{wolverine}).
-How to connect this trade-off with task virtualization remains an open question.
-
-%\textbf{Research Question and Contributions:} Specifically, we ask (i) how to use software support to \emph{dynamically adapt} the effective size of a task, while respecting programmer-specified task atomicity, and (ii) how to minimize run time and energy consumption while \emph{automatically maintaining memory consistency} during execution?
-
-%To address this question, we develop {\bf \sys}: a new programming and execution model efficiently executing tasks which avoids non-termination across a range of energy buffer sizes, without burdening the programmer\footnote{Code will be available via \href{http://anonymized.link}{http://anonymized.link}.}.
-To accomplish this, \sys introduces two new capabilities:
-
-%\begin{itemize}
-%	\item {\bf Dynamic task coalescing:} \sys's task coalescing mechanism dynamically executes multiple tasks as a single task.
-A programmer or compiler can specify small tasks that will execute on a device with a small energy buffer.
-Coalescing tasks allows running the same tasks on a device with a larger energy buffer, avoiding run time overheads associated with ending one task and beginning another.
-%	%
-%	\item {\bf Volatile memory virtualization with transiently-powered device:} Colascing through task virtualization requires memory virtualization.
-\sys's uses SRAM as working memory, which \sys dynamically populates with pages of data from FRAM on demand, during a task's execution.
-When a task ends, \sys commits dirty pages to FRAM using DMA block copies, ensuring task atomicity and memory consistency.
-To the best of our knowledge this is the first time when memory virtualization has been demonstrated on a transient power budget.
-%\end{itemize}
-
-We evaluated \sys end-to-end on a real energy-harvesting device~\citep{wisp}
-running existing benchmarks~\citep{chain,alpaca} and new programs. A comparison
-to a state-of-the-art task-based system~\citep{chain,alpaca} showed that \sys
-recovers performance on conservative task decompositions missed by systems with
-static decomposition.
-%
-%and can flexibly target a variety of platforms without rewriting code.
-%
-%\todo{Provide concrete numeric results supporting this claim}{Amjad/Przemek}
-
-%In this work we also address the problem of task construction burden imposed on the programmer.
-That is, we introduce \sys's compiler that transforms arbitrary C code into a graph of tasks that reflects the original program's control- and data-flow constraints.
-The compiler identifies data shared by multiple of the newly created tasks and instruments reads and writes of those data so that \sys's memory virtualization mechanism keeps those data consistent.
-While \sys's complier is still \emph{limited in scope}, e.g.
-is unable to handle instruction jumps, normalized loops and recursions, it already eliminates the labour-intensive process of writing tasks~\citep{chain,alpaca}, i.e., hours of programmer time.
-
-%\sys's core contribution is the introduction of dynamic task coalescing mechanism.
-It allows the programmer or a compiler to intuitively decompose the program into small tasks that amortize fixed per-task overheads, yet present no risk of exceeding device energy capacity.
-As such a decomposition executes, \sys~{\em coalesces dynamically} consecutive tasks.
-Coalescing elides the commits of coalesced tasks by buffering multiple tasks' updates in an FRAM commit buffer.
-Periodically, as the span of the coalesced task grows, \sys ends coalescing and commits the state of the coalesced task.
-If a power failure interrupts a sequence of coalesced tasks, \sys adaptively reduces the number of tasks in that sequence that it will coalesce, committing sooner in future executions.
-Consequently, \sys's coalescing mechanism allows a program to execute efficiently across a range of energy buffer sizes, avoiding transition overheads in larger buffers, and ensuring progress in smaller buffers.
-
-%We propose and explore two data swapping strategies for intermittently-powered systems based on FRAM/SRAM architecture.
-The first strategy is {\em demand paging}, which swaps the SRAM page with a new page from FRAM, buffering the swapped-out page until commit.
-The second strategy is {\em buffered direct access}, which directly reads and writes FRAM relying on dynamic double-buffering to ensure memory is consistent at commit.
+Section~\ref{sec:background} provides background on intermittent computing.
+Section~\ref{sec:overview} provides an overview of \sys, while
+Sections~\ref{sec:coalescing} and~\ref{sec:memory} describe \sys's coalescing
+and virtualization mechanisms.Section~\ref{sec:discussion} discusses \sys
+design issues.  Sections~\ref{sec:methodology} and~\ref{sec:eval} describe
+\sys's methodology and evaluation.  Section~\ref{sec:related} puts \sys in the
+context of related work and Section~\ref{sec:conc} concludes and discusses
+future work.

### 2018-04-12 paper/relatedwork.tex
edits in intro and todos in relatedwork

@@
 
+One approach to decrease the task count, without assuming a minimum level of
+incoming power always available, is to predict task energy consumption and
+choose a decomposition specific to the device capacitor size~\cite{cleancut}.
+However, statically predicting energy of arbitrary input-dependent code with
+peripheral access is a problem without a general solution. Furthermore, a
+static decomposition approach prevents portability across devices with
+different storage capacitors.
+
+\textcolor{red}{TODO: make this one paragraph with the CleanCut stuff.  add \cite{intel} work; separate by-programmer and by-compiler;
+both cases hit the energy estimation problem; also intel work proposes a method
+for decomposing into one of many decompositions, but does not give a good
+solution for choosing one}
+
 \textbf{Memory Virtualization.} Prior work on embedded systems has studied a variety of memory virtualization strategies relating to \sys.

### 2018-04-12 paper/partialcommit.tex
section for partial commit added

\subsection{Task Downsizing} %Partial Commit/Partial Execution

\begin{figure}
	\caption{\textcolor{red}{Figure to describe partial commit is needed}}
\end{figure}

\subsubsection{When Task Downsizing is Triggered}

If power fails consecutively twice before the system can complete one single static task, it means that the static task requires more energy than currently available in a discharge cycle, so it is necessary to have a mechanism to execute a chunk of the task, save the partial progress through the task, and continue try to run the rest of the task.  If a power failure occurs after saving partial task progress, the system makes forward progress, albeit without task atomicity. Coala may save partial task progress multiple times before completing a very long task.  To avoid the fail-stop case of non-termination, Coala's partial commit mechanism will split tasks that have not been marked as atomic by the programer.

A partial commit saves the stack, registers and the volatile pages to the non-volatile memory.

Coala starts with a big random guess and down-scales it using a binary search until it succeeds in preserving the partial execution progress. The obtained value is the initial seed for the next partial commit procedure.

\textcolor{red}{More variables describing the partial commit algorithm: we do not explore as many options for partial commit as we do for coalescing}

\subsection{Why partial commit is not Coala's default behavior}

\emph{Static deployment of partial commit stubs} This approach reduces programmer effort - they still need to apply Coala's API and distribute trigger points for partial commits - but it introduces an additional overhead by saving the stack and the register file each time it commits the pages.

\emph{Dynamic deployment of partial commit stubs:} This approach does not rely on any information from the programmer to trigger the commit process. However, it utilizes a trial and error approach to optimize the commit rate over the entire program execution time. Therefore, it suffers from a large cumulative re-execution penalty.

### 2018-04-12 paper/intro.tex
capybara reference added

@@
 buffers energy in a small storage capacitor~\citep{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
-threshold amount of energy has accumulated. 
-%
-Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
+threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
 power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
@@
 3c]{nadeau_naturebio_2017}.
-The recharge and discharge time --- which corresponds to the device's inactive and active time --- varies with the energy buffering capacitor's~\cite{capybara} and some devices fail and restart operating $\approx$10 to
+The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor's~\cite{capybara} and some devices fail and restart operating $\approx$10 to
 $\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.
-
-%\begin{figure}
-%	\begin{subfigure}[t]{.35\linewidth}
-%		\centering\large Figure
-%		\caption{\centering WISP~\citep{wisp} against an RFID antenna}\label{fig:1a}
-%	\end{subfigure}%
-%	\begin{subfigure}[t]{.65\linewidth}
-%		\centering\large Figure
-%		\caption{\centering Program X~\citep{hicks_mibench2_2016} (split into X and Y tasks~\citep{chain}) at two distances}\label{fig:1b}
-%    \end{subfigure}
-%	\caption{Impact of an inadequate task sizing on the computation speed of a moving transiently-powered device.}\label{fig:1}
-%\end{figure}
 
@@
 software approaches have a number of limitations that are the focus of the work
-in this paper.  A key limitation of prior software approaches that we address in this paper is the {\em inflexibility}
-of a program statically decomposed into tasks.
+in this paper.  A key limitation of prior software approaches that we address in this paper is the {\em inflexibility} of a program statically decomposed into tasks.
 
-%\textcolor{red}{tictpl is not HW, is it? also it doesn't deal with inconsistency; this paragraph doesn't include Hibernus, so it should not include CPTL either.}
-
-% <<<<<<< HEAD
-% \begin{figure}
-% 	\includegraphics[width=.5\columnwidth]{figures/alpaca-war.pdf}
-% 	\caption{\label{fig:virtualization}Adaptive task division at run-time might lead to write-after-read dependencies (variable \ttfamily{x}) that do not exist with static task division.\todo{Box with Task 1,2,3 must be the same size as merged task; Merged task must be named as Task 4; x must be marked in text as having WAR; "Task X" name should be next to the task box; "Window" - small letter}{Sinan}}
-% \end{figure}
-% =======
-
-%Moved intel note to related work
 \textbf{Task Decomposition of Intermittent Programs.} {\em
@@
 overheads while avoiding the risk of non-termination.
-
-%Brandon: I moved Cleancut to the related work section.  THere's too much in the intro already
 
@@
 
-%
 %variables in non-volatile memory at \emph{run-time} that can
@@
 %results.
-%
-%Brandon: Sentence too long.
+
 While a compiler can use static data privatization and commit instrumentation

### 2018-04-12 paper/intro.tex
fix reference in the intro

@@
 inadequate energy to execute each single task to completion, a program will not
-terminate~\cite{cleancut}. To avoid non-termination under adversarial 
+terminate~\cite{cleancut_2018}. To avoid non-termination under adversarial 
 energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
@@
 \item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
-\item A fully-realized prototype of \sys that runs on real energy-harvesting devices~\cite{wisp,capy}, to be released as open source after publication. 
+\item A fully-realized prototype of \sys that runs on real energy-harvesting devices~\cite{wisp,capybara}, to be released as open source after publication. 
 \item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work~\cite{alpaca}, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to varied energy conditions.
@@
 Sections~\ref{sec:coalescing} and~\ref{sec:memory} describe \sys's coalescing
-and virtualization mechanisms.Section~\ref{sec:discussion} discusses \sys
+and virtualization mechanisms. Section~\ref{sec:discussion} discusses \sys
 design issues. Sections~\ref{sec:methodology} and~\ref{sec:eval} describe

### 2018-04-12 paper/intro.tex
added a link to Coala repository website in the intro

@@
 \item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
-\item A fully-realized prototype of \sys that runs on real energy-harvesting devices~\cite{wisp,capybara}, to be released as open source after publication. 
+\item A fully-realized prototype of \sys that runs on real energy-harvesting devices~\cite{wisp,capybara}, to be released as open source after publication\footnote{An anonymized version of the Coala repository is already accessible via~\cite{coala_website} for inspection.}.
 \item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work~\cite{alpaca}, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to varied energy conditions.

### 2018-04-12 paper/background.tex
added a todo and changed the title
editorial revisions to background

@@
 We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys addresses.
+
+%\begin{figure}
+\begin{wrapfigure}{t!}{0.5\textwidth}
+	\begin{subfigure}[t]{\linewidth}
+		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
+		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
+	\end{subfigure}
+	\begin{subfigure}[t]{\linewidth}
+		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
+		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
+	\end{subfigure}
+	\caption{\textbf{Code example demonstrating effect of write after read on volatile memory checkpointing.}}\label{fig:code_demo_incosistency}
+\end{wrapfigure}
+%\end{figure}
 
@@
 These devices elide tethered power or a battery, instead collecting energy into a capacitor, operating when sufficient energy accumulates, and upon depleting the energy, turning off and recharging.
+\TODO{We need a figure that shows what intermittent execution looks like.}
 %
@@
 
-\textcolor{red}{reference work by Intel}
+\textcolor{red}{reference work by Intel~\cite{baghsorkhi_cgo_2018}}
 
@@
 Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
-
-\begin{figure}
-	\begin{subfigure}[t]{\linewidth}
-		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
-		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
-	\end{subfigure}
-	\begin{subfigure}[t]{\linewidth}
-		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
-		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
-    \end{subfigure}
-	\caption{\textbf{Code example demonstrating effect of write after read on volatile memory checkpointing.}}\label{fig:code_demo_incosistency}
-\end{figure}
 
@@
 		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
-		~\cite{intel} & XXX \\
+		Region Formation~\cite{baghsorkhi_cgo_2018} & \textcolor{red}{XXX} \\
 		\hline
@@
 \end{subfigure}
-	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.}}\label{fig:framEnergy}
+	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
 \end{figure}

### 2018-04-12 paper/intro.tex
Coala into \sys in the intro

--- 2018-04-12 paper/intro.tex
@@
 \item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
-\item A fully-realized prototype of \sys that runs on real energy-harvesting devices~\cite{wisp,capybara}, to be released as open source after publication. 
+\item A fully-realized prototype of \sys that runs on real energy-harvesting devices~\cite{wisp,capybara}, to be released as open source after publication\footnote{An anonymized version of the \sys repository is already accessible via~\cite{coala_website} for inspection.}.
 \item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work~\cite{alpaca}, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to varied energy conditions.

### 2018-04-12 paper/systemdescription.tex
system description updated: added downscaling and removed reference to complier

@@
-\sys is a new programming and execution model for intermittent computing on
-energy-harvesting devices. \sys addresses the challenges outlined in
-Section~\ref{sec:background} to make task-based intermittent programs {\em
-programmable} and {\em efficient}. \sys accomplishes this goal with a
-constellation of a new programming model, compiler, and run time software
-system support, that supports dynamically adaptive task-based execution.
+\sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
+\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable} and {\em efficient}.
+\sys accomplishes this goal with a constellation of a new programming model and run time software system support, that supports dynamically adaptive task-based execution.
 Figure~\ref{fig:system_overview} shows an overview of \sys.
@@
 	\includegraphics[width=\columnwidth]{figures/viper_block_diagram.pdf}
-	\caption{\sys top-level view.}
+	\caption{\sys top-level view. \textcolor{red}{redraw the figure: remove compiler, add task downscaling}}
 	\label{fig:system_overview}
@@
 
-\textbf{\sys Programming and Execution Model.} 
-To use \sys, a programmer first writes plain, imperative C code. The programmer
-then must decompose the program into tasks. To manually decompose a program
-into tasks, the programmer designates a set of functions as tasks, sequences
-control-flow between these tasks, and annotates memory accesses that manipulate
-data shared by multiple tasks.
+\textbf{\sys Programming and Execution Model.}  To use \sys, a programmer first writes plain, imperative C code. The programmer
+then must decompose the program into tasks.
+To manually decompose a program into tasks, the programmer designates a set of functions as tasks, sequences control-flow between these tasks, and annotates memory accesses that manipulate data shared by multiple tasks.
 
-\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists
-of primitives for defining tasks~\cite{chain,alpaca} and accessors for
-protected variables shared between tasks. As in prior task-based intermittent
-programming models~\cite{chain,alpaca}, tasks are functions that are not
-allowed to have any direct callers but receive control via explicit {\tt
-next\_task} statements only.  The programmer identifies each task
-by adding the {\tt task} keyword to the function's declaration and designates
-one task as the {\em origin} task at which execution begins. After a power
-failure, \sys restarts the last partially executed task.
+%\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists of primitives for defining tasks~\cite{chain,alpaca} and accessors for protected variables shared between tasks.
 %
-A \sys program manipulates {\em protected} state---accessed by more than one
-task---using the \emph{read variable} ({\tt RVAR}) and \emph{write variable}
-({\tt WVAR}) operations. These operations convey to \sys that
-the referenced data are task-shared, meaning that they must be buffered in
-volatile memory and committed when the executing task ends.
+%As in prior task-based intermittent programming models~\cite{chain,alpaca}, tasks are functions that are not allowed to have any direct callers but receive control via explicit {\tt next\_task} statements only.
+The programmer identifies each task by adding the {\tt task} keyword to the function's declaration and designates one task as the {\em origin} task at which execution begins.
+After a power failure, \sys restarts the last partially executed task.
 %
-Commit is a key overhead associated with task-based operation and \sys's
-coalescing strategy (described in Section~\ref{sec:coalescing} avoids commit
-when a task ends by continuing to buffer task-shared variables for multiple
-coalesced tasks, and committing only when all tasks complete.
+%A \sys program manipulates {\em protected} state---accessed by more than one task---using the \emph{read variable} ({\tt RVAR}) and \emph{write variable} ({\tt WVAR}) operations.
+These operations convey to \sys that the referenced data are task-shared, meaning that they must be buffered in volatile memory and committed when the executing task ends.
+%
+%Commit is a key overhead associated with task-based operation and \sys's coalescing strategy (described in Section~\ref{sec:coalescing} avoids commit when a task ends by continuing to buffer task-shared variables for multiple coalesced tasks, and committing only when all tasks complete.
 
-\begin{table}
-	\centering
-	\footnotesize
-	\begin{tabular}{|c|c|}
-		\hline
-		\textbf{Keyword} & \textbf{Description} \\
-		\hline\hline
-		\texttt{task foo() \{...\}} & Task definition\\
-		\texttt{origin task foo() \{...\}} & Origin task definition\\
-		\texttt{next\_task(t)} & Transition to task \texttt{t}\\
-		\texttt{u = RVAR(v)} & Read protected variable \texttt{v} into \texttt{u}\\
-		\texttt{WVAR(v, u)} & Write \texttt{u} to protected variable \texttt{v}\\
-		\hline
-	\end{tabular}
-	\caption{\sys's programming interface.}
-	\label{tab:viper_syntax}
-\end{table}
+%\begin{table}
+%	\centering
+%	\footnotesize
+%	\begin{tabular}{|c|c|}
+%		\hline
+%		\textbf{Keyword} & \textbf{Description} \\
+%		\hline\hline
+%		\texttt{task foo() \{...\}} & Task definition\\
+%		\texttt{origin task foo() \{...\}} & Origin task definition\\
+%		\texttt{next\_task(t)} & Transition to task \texttt{t}\\
+%		\texttt{u = RVAR(v)} & Read protected variable \texttt{v} into \texttt{u}\\
+%		\texttt{WVAR(v, u)} & Write \texttt{u} to protected variable \texttt{v}\\
+%		\hline
+%	\end{tabular}
+%	\caption{\sys's programming interface.}
+%	\label{tab:viper_syntax}
+%\end{table}
 
-The programmer can decompose a program into tasks either {\em manually} or {\em
-automatically}. To do so manually,  requires reasoning similar to prior
+The programmer can decompose a program into tasks either {\em manually}. To do so manually,  requires reasoning similar to prior
 task-based systems~\cite{chain,alpaca}.  
 %
-A programmer may also opt to use compiler support to automatically decompose a
-program into tasks, leveraging recent work~\cite{cleancut,baghsorkhi_cgo_2018}.
+%A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut,baghsorkhi_cgo_2018}.
 %
-Without loss of generality, we assume throughout this work that the programmer
-manually decomposed the program into tasks; \sys's behavior with automatically
-decomposed code would be identical.
+%Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with automatically decomposed code would be identical.
+
+%Regardless of whether automatically or manually decomposed into tasks, the
 %
+The programmer compiles their task-based code, and links to the \sys runtime, producing a \sys-enabled binary.
+The \sys runtime implements \sys's task-based programming and execution model.
+The runtime system also includes \sys's novel {\em virtualizing memory manager} and its {\em task coalescing manager}, and {\em task downscaling manager}, both of which are essential to \sys's adaptive task-based execution.
 
-Regardless of whether automatically or manually decomposed into tasks, the
-programmer compiles their task-based code, and links to the \sys runtime,
-producing a \sys-enabled binary.
-%
-The \sys runtime implements \sys's task-based programming and execution model.
-%
-The runtime system also includes \sys's novel {\em virtualizing memory manager}
-and its {\em task coalescing manager}, both of which are essential to \sys's
-adaptive task-based execution.
+\textbf{\sys Task Coalescing Manager.} \sys relies on its task coalescing manager to dynamically {\em coalesce} statically\hyp{}defined tasks to avoid inessential overheads associated with completing tasks.
+By default, tasks run in a sequence and each task commits its task-shared state as it completes.
+The time and energy cost of a task's commit is unnecessary if the task and its successor both complete without a power failure.
+The key insight that \sys leverages is that the first task could have deferred its commit to the second task.
+{\em coalescing} tasks by deferring the first commit avoids the fixed cost of the first commit (task transitioning overheads).
+Coalescing also amortizes per-variable commit cost, committing each  location accessed by both tasks only after the second task, rather than once per task.
 
-\textbf{\sys Task Coalescing Manager.} \sys relies on its task coalescing
-manager to dynamically {\em coalesce} statically\hyp{}defined tasks to avoid
-inessential overheads associated with completing tasks.  
-%
-By default, tasks run in a sequence and each task commits its task-shared state
-as it completes.  The time and energy cost of a task's commit is unnecessary if
-the task and its successor both complete without a power failure.  The key
-insight that \sys leverages is that the first task could have deferred its
-commit to the second task. {\em coalescing} tasks by deferring the first commit
-avoids the fixed cost of the first commit (task transitioning overheads).
-Coalescing also amortizes per-variable commit cost, committing each  location
-accessed by both tasks only after the second task, rather than once per task.
-%
+\textbf{\sys Task Downscaling Manager.} \textcolor{red}{to be written}
 
-\textbf{\sys Virtual Memory Manager.} 
-\sys is able to efficiently coalesce tasks because of its efficient virtual
-memory manager,
-%
-which is described in Section~\ref{sec:memory_virtualization}. 
-%
-\sys's memory manager paginates memory and ensures that data in a page remain consistent
-despite power interruptions. 
-%
+\textbf{\sys Virtual Memory Manager.}  \sys is able to efficiently coalesce tasks because of its efficient virtual memory manager, which is described in Section~\ref{sec:memory_virtualization}.
+\sys's memory manager paginates memory and ensures that data in a page remain consistent despite power interruptions.
 \sys allows a task to manipulate data in a volatile copy of a page only.
-%
-Pages swap between volatile and non-volatile memory, depending on the capacity
-of the volatile memory and the program's access pattern. 
-%
-\sys tracks a task's memory accesses efficiently at page granularity (rather
-than using, e.g., costly word-granular tracking).
-%
-When a task ends, each page it accessed commits from volatile memory (or from a
-non-volatile swap region for dirty pages) back to the non-volatile main memory.
-%
-Pages efficiently, atomically commit using a two-phase commit procedure
-accelerated using hardware support for direct memory access (DMA). 
-%
+Pages swap between volatile and non-volatile memory, depending on the capacity of the volatile memory and the program's access pattern.
+\sys tracks a task's memory accesses efficiently at page granularity (rather than using, e.g., costly word-granular tracking).
+When a task ends, each page it accessed commits from volatile memory (or from a non-volatile swap region for dirty pages) back to the non-volatile main memory.
+Pages efficiently, atomically commit using a two-phase commit procedure accelerated using hardware support for direct memory access (DMA).

### 2018-04-12 paper/adaptiveModel.tex
add brain dump about Coala the execution model

@@
+\textcolor{red}{THIS BRAIN DUMP ABOUT THE EXECUTION MODEL - Placeholder}
+
+\sys execution model operates on task-based programming model, but It progresses on a dynamic, virtual task granularity.
+A virtual task can be either multiple static tasks \emph{coalesced} together, when the energy permits, or a sub-task, when it is necessary to preserve forward progress.
+
+\sys execution model decouples the process of coalescing task from the process to downsizing the virtual task size.
+This separation is necessary to prevent the system from suffering from repeated power failures when energy becomes scares after it was abandoned or when the energy required to finish a static task increases significantly as compared to the previous static tasks.
+
+\textcolor{red}{Not very clear.
+Here with downsizing you are referring to the target, but, based on the previous sentence, the reader might mistake it with the task splitting feature.
+The reader might not know yet the re-execution penalty problem, so it might be hard to get what you mean with decoupling the two processes.}
+
+For example, if \sys uses a linear adaptation strategy and it was progressing on 20 static task per a virtual task, then the system has to default back to a single static task per a virtual task it may need to reboot 20 times before preserving the progress of a single static task!
+
+\textcolor{red}{
+Here you are already suggesting (and disproving) a coalescing strategy, but the reader should be exposed to the challenges before doing this, in my opinion.
+Here instead I would mention why coalescing up is needed (to reduce the number of commits), and why coalescing down is necessary (to reduce the re-execution penalty), using something like this figure https://goo.gl/GiJgEz}
+
+This adaptive execution model estimates the energy - using a pure-software technique - to tune the size of the virtual task.
+Therefore, its virtual task upscaling and downscaling processes are independent and, in case it is needed due to scarce energy, it only takes two reboots to shrink the virtual task size to a single static task.
+\emph{The adaptive execution model of \sys is efficient because it is energy aware.}
+
+Coala's execution model is neither limited by a static task size to progress, nor by the commit size, i.e.
+the amount of data that needs to be saved to the non-volatile memory to preserve execution progress.
+Coala can partition compile-time tasks, if it fails on a single static task more than once, or merge them, if saving the state to non-volatile memory is redundant because a power failure is not expected soon.
+This unique technique makes the commit size proportional to the number of state-less instructions executed.
+In other words, if energy allows executing only a small chunk of code, the amount of data to be committed shrinks accordingly.
+
 %% Coalescing motivation

### 2018-04-13 paper/evaluation.tex
update the paging related figures; NO FULL PAGE SWAP OBSERVED

@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/ramPagsSizes}
-	\includegraphics[width=\columnwidth]{figures/pagefault}
-	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64, $\ldots$,1024\}\,bytes, per application (top) and respective page faults (bottom)}; X = zero page faults; red bar: half-page swap, blue bar: full page swap (replacing pages in both FRAM and SRAM).}\vspace{-0.5cm}
+	\includegraphics[width=\columnwidth]{figures/page_exec-time}
+	\includegraphics[width=\columnwidth]{figures/pagePulls}
+	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,bytes, per application (top) and respective page pulls (bottom)}).}\vspace{-0.5cm}
 %\todo{change X to 0, remove the definition of X from caption}{Amjad}}

### 2018-04-13 paper/intro.tex
new intro figure
change the intro figure

@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/coalescing-is}
-    \caption{\textbf{An execution of three
-statically-defined tasks without (top) and with (bottom) dynamic
-coalescing.} Task coalescing at runtime reduces time and energy overhead in
+    \includegraphics[width=0.5\textwidth]{figures/Coala_Intro_Figure.pdf}
+    \caption{ %\textbf{An execution of three statically-defined tasks without (top) and with (bottom) dynamiccoalescing.}
+        Task coalescing at runtime reduces time and energy overhead in
 task-based intermittent programming models by performing fewer commits and

### 2018-04-13 paper/evaluation.tex

@@
 
-We evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp} energy-harvesting
-platform build around a MSP430FR5969~\cite{wolverine} MCU with 64\,kB of FRAM.
-We used a TI Flash Emulation Tool (FET)~\cite{fet} as a power supply during
-experiments with continuous power. For energy measurements, we used
-EDB~\cite{edb}. In experiments with wireless power, we powered the WISP using
-an RFID reader (Impinj R1000 with firmware version
-3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ
-8\,dBic~\cite{atlas2015}. We controlled the RFID reader from a workstation
-running Ubuntu 10.4 using \emph{sllurp}~\cite{sllrp_github}. We set the
-transmitted power of the RFID reader to 30\,dBm at 915\,MHz center frequency.
-For distance-controlled experiments, we affixed the WISP at varying lengths to
-hollow paper tubes placed flat on a lying antenna. Timing measurements
-were done using the Salea~\cite{saleae} logic analyzer.
+We evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp} energy-harvesting platform build around a MSP430FR5969~\cite{wolverine} MCU with 64\,kB of FRAM.
+We used a TI Flash Emulation Tool (FET)~\cite{fet} as a power supply during experiments with continuous power.
+For energy measurements, we used EDB~\cite{edb}.
+In experiments with wireless power, we powered the WISP using an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
+We controlled the RFID reader from a workstation running Ubuntu 10.4 using \emph{sllurp}~\cite{sllrp_github}.
+We set the transmitted power of the RFID reader to 30\,dBm at 915\,MHz center frequency.
+For distance-controlled experiments, we affixed the WISP at varying lengths to hollow paper tubes placed flat on a lying antenna.
+Timing measurements were done using the Salea~\cite{saleae} logic analyzer.
 
-For some experiments (noted below) we used an alternative setup where the
-MSP-TS430RGZ48C launchpad was used for continuous power experiments and used
-internal timer for measuring run time. For harvested energy, we used
-WISP\,5.0~\cite{wisp5,wisp} hardware and ThingMagic Astra-EX integrated RFID
-reader set at 30\,dBm transmission power at 915\,MHz. We used EDB~\cite{edb}
-to measure run time. Applications were cross-compiled using GCC and Clang (both version 3.8.0), allowing us to use LLVM for the \sys compiler.
+For some experiments (noted below) we used an alternative setup where the MSP-TS430RGZ48C launchpad was used for continuous power experiments and used internal timer for measuring run time.
+For harvested energy, we used WISP\,5.0~\cite{wisp5,wisp} hardware and ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power at 915\,MHz.
+We used EDB~\cite{edb} to measure run time.
+Applications were cross-compiled using GCC and Clang (both version 3.8.0), allowing us to use LLVM for the \sys compiler.
 
@@
 
-We compare \sys against state-of-the-art task-based runtime Chain~\cite{chain}.
-We do not compare \sys to other, earlier systems, such as DINO~\cite{dino} or
-Mementos~\cite{mementos} (cf. Section~\ref{sec:background_consistency}), as
-Chain is the only task-based execution model available to which we can compare
-\sys.
+%We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
+We do not compare \sys against Chain~\cite{chain}, another task-based system, which is inferior to Alpaca in its performance.
+Also, we do not compare \sys to checkpointing-based systems, such as DINO~\cite{dino} or Mementos~\cite{mementos} (cf.
+Section~\ref{sec:background_consistency}), as both of them are build on top LLVM compiler, which performance is inferior to GCC compiler used in the evaluation of \sys \todo{Please check if the LLVM compiler argument makes sense}.
 
-We evaluated \sys using benchmarks from the literature~\cite{chain,alpaca} and
-several new programs: (i) \textbf{bc}: counts bits in byte pseudo-random string
-using several algorithms, and compares results for correctness; (ii)
-\textbf{cem}, which emulates an embedded temperature sensing engine and in
-which measurements are randomly generated and compressed using LZW with a fixed
-entry dictionary and fixed byte block; (iii) \textbf{cuckoo}: which implements
-a Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random
-numbers, then queries the filter to recover the sequence; (iv) \textbf{rsa}:
-which encrypts a fixed string with a stored key; (v) \textbf{ar}: which
-performs nearest neighbour classification of randomly generated data modelling a
-three-axis accelerometer. This set of benchmarks overlaps with those used
-in~\cite{chain,alpaca}. We implemented several new applications: (vi)
-\textbf{sort}: selection sort algorithm on a fixed sequence of numbers, and
-(vii) \textbf{dft}: discrete Fourier transform. All applications are compared
-in columns eight and nine of Table~\ref{table:compiler_result} showing the
-number of source lines of code compared to Chain.
-
-%\begin{table}[t]
-%	\centering
-%	\footnotesize
-%	\begin{tabular}{|c|c|c|}
-%		\hline
-%		Application & SLOC (\sys) & SLOC (Chain~\cite{chain})\\
-%		\hline\hline
-%		{bc} & 351 & 588 \\ %59\%
-%		{cem} & 388 & 721 \\ %53\%
-%		{cuckoo} & 483 & 762 \\ %63\%
-%		{rsa} & 887 & 1233 \\ %71\%
-%		{ar} & 483 & 762 \\ %63\%
-%		{dft} & {} & {} \\ %Two resolutions: 4 and 8 Bytes
-%		{dd} & {} & {} \\ %Data decompression size: 100 bytes
-%		{sort} & {} & {} \\
-%		\hline
-%	\end{tabular}
-%\caption{List of benchmark programs used in \sys evaluation; SLOC: source lines of code.}
-%\label{table:benchmark_list}
-%\end{table}
+%We evaluated \sys using benchmarks that originated from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.
+Specifically, we have implemented three benchmarks used by Chain~\cite{chain} and Alpaca~\cite{alpaca} (source code: \cite{coala_website}), namely: \textbf{ar}: nearest neighbor classification of randomly generated data modeling a three-axis accelerometer; \textbf{bc}: counts bits in byte pseudo-random string using several algorithms, and compares results for correctness; \textbf{cuckoo}: Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence.
+Additionally, we have ported three new benchmarks from~\cite{mibench}, namely: \textbf{fft}: Fast Fourier Transform algorithm \todo{specify details of the implementation}; and \textbf{sort}: selection sort algorithm on a fixed sequence of numbers.
+\todo{justify the selection of these benchmarks} All applications are summarized in Table~\ref{table:compiler_result} in terms of memory footprint, source lines of code and number of tasks.
+\todo{update table with new data}
 
@@
 	\end{tabular}
-	\caption{\textbf{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns)}; $^\dagger$dft does not compile with Clang}\label{table:compiler_result}\vspace{-0.5cm}
+	\caption{\textbf{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns)}; $^\dagger$dft does not compile with Clang }\label{table:compiler_result}\vspace{-0.5cm}
 \end{table}
 
-\textbf{\sys Compiler Generated Code Characterization.} The behaviour and the efficiency metrics of the \sys compiler are given in Table~\ref{table:compiler_result}. 
+%
+
+\textbf{\sys Compiler Generated Code Characterization.} The behaviour and the efficiency metrics of the \sys compiler are given in Table~\ref{table:compiler_result}.  \todo{update table with new data}
 %

### 2018-04-13 paper/evaluation.tex
complier removed from evaluation results

@@
-%We present the experimental results of \sys's performance.
-We start with describing the hardware environment (Section~\ref{sec:results_hardware}) followed by the software environment (Section~\ref{sec:results_software}) and we concluded by the evaluation results (Section~\ref{sec:results_evaluation}).
-
-We evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp} energy-harvesting platform build around a MSP430FR5969~\cite{wolverine} MCU with 64\,kB of FRAM.
-We used a TI Flash Emulation Tool (FET)~\cite{fet} as a power supply during experiments with continuous power.
-For energy measurements, we used EDB~\cite{edb}.
-In experiments with wireless power, we powered the WISP using an RFID reader (Impinj R1000 with firmware version 3.2.4.240~\cite{r1000_data_sheet}) connected to the Liard RFMAX S9028PCRJ 8\,dBic~\cite{atlas2015}.
-We controlled the RFID reader from a workstation running Ubuntu 10.4 using \emph{sllurp}~\cite{sllrp_github}.
-We set the transmitted power of the RFID reader to 30\,dBm at 915\,MHz center frequency.
-For distance-controlled experiments, we affixed the WISP at varying lengths to hollow paper tubes placed flat on a lying antenna.
-Timing measurements were done using the Salea~\cite{saleae} logic analyzer.
-
-For some experiments (noted below) we used an alternative setup where the MSP-TS430RGZ48C launchpad was used for continuous power experiments and used internal timer for measuring run time.
-For harvested energy, we used WISP\,5.0~\cite{wisp5,wisp} hardware and ThingMagic Astra-EX integrated RFID reader set at 30\,dBm transmission power at 915\,MHz.
-We used EDB~\cite{edb} to measure run time.
-Applications were cross-compiled using GCC and Clang (both version 3.8.0), allowing us to use LLVM for the \sys compiler.
-
-\subsection{Benchmarks and Baseline}
-\label{sec:results_software}
-
-%We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
-We do not compare \sys against Chain~\cite{chain}, another task-based system, which is inferior to Alpaca in its performance.
-Also, we do not compare \sys to checkpointing-based systems, such as DINO~\cite{dino} or Mementos~\cite{mementos} (cf.
-Section~\ref{sec:background_consistency}), as both of them are build on top LLVM compiler, which performance is inferior to GCC compiler used in the evaluation of \sys \todo{Please check if the LLVM compiler argument makes sense}.
-
-%We evaluated \sys using benchmarks that originated from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.
-Specifically, we have implemented three benchmarks used by Chain~\cite{chain} and Alpaca~\cite{alpaca} (source code: \cite{coala_website}), namely: \textbf{ar}: nearest neighbor classification of randomly generated data modeling a three-axis accelerometer; \textbf{bc}: counts bits in byte pseudo-random string using several algorithms, and compares results for correctness; \textbf{cuckoo}: Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence.
-Additionally, we have ported three new benchmarks from~\cite{mibench}, namely: \textbf{fft}: Fast Fourier Transform algorithm \todo{specify details of the implementation}; and \textbf{sort}: selection sort algorithm on a fixed sequence of numbers.
-\todo{justify the selection of these benchmarks} All applications are summarized in Table~\ref{table:compiler_result} in terms of memory footprint, source lines of code and number of tasks.
-\todo{update table with new data}
-
 \subsection{\sys Evaluation Results}
@@
 
-\subsubsection{\sys's Compiler}
-\label{sec:result_compiler_time}
+%\subsubsection{\sys's Compiler}
+%\label{sec:result_compiler_time}
 
@@
 
-\textbf{\sys Compiler Generated Code Characterization.} The behaviour and the efficiency metrics of the \sys compiler are given in Table~\ref{table:compiler_result}.  \todo{update table with new data}
+%\textbf{\sys Compiler Generated Code Characterization.} The behaviour and the efficiency metrics of the \sys compiler are given in Table~\ref{table:compiler_result}.
+\todo{update table with new data}
 %
@@
 %
-On average, compiler-generated code introduces about 0.3\% more bytes of protected data, suggesting that liveness analysis captures natural boundaries in the use of variables also caught by a manual implementation.
+%On average, compiler-generated code introduces about 0.3\% more bytes of protected data, suggesting that liveness analysis captures natural boundaries in the use of variables also caught by a manual implementation.
 Compiler-generated code generates about 16\% fewer tasks for applications, on average.
@@
 Consequently, the \sys compiler introduces 48\% more protected data accesses than programmer-written code.
-The run time results in Section~\ref{sec:result_compiler_time} show that the increase in protected data
-accesses sometimes degrades performance.
+The run time results in Section~\ref{sec:result_compiler_time} show that the increase in protected data accesses sometimes degrades performance.
 Overall, the data show that the compiler automatically generates tasks similarly to manually written tasks.
@@
 
-\textbf{Compile Time.} We measured compile time for \sys's compiler's automatic task generation for each application\footnote{We could not compile the dft application since LLVM inserts mathematical built-in functions (i.e.
+%\textbf{Compile Time.} We measured compile time for \sys's compiler's automatic task generation for each application\footnote{We could not compile the dft application since LLVM inserts mathematical built-in functions (i.e.
 \texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}.
@@
 
-\textbf{Run Time of \sys Compiler Code.} In addition to the characterization of the \sys compiler in Table~\ref{table:compiler_result}, we experimentally assessed the performance of the \sys compiler, compared to the performance of the manually written benchmarks, as written in prior work~\cite{alpaca,chain}.
+%\textbf{Run Time of \sys Compiler Code.} In addition to the characterization of the \sys compiler in Table~\ref{table:compiler_result}, we experimentally assessed the performance of the \sys compiler, compared to the performance of the manually written benchmarks, as written in prior work~\cite{alpaca,chain}.
 Figure~\ref{fig:comp_user} compares the performance of manually written benchmark code to the performance of code automatically compiled using \sys.
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/comp_user}
-	\caption{\textbf{\sys compiler code overhead measured in terms of execution time per application} (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
-	\label{fig:comp_user}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\columnwidth]{figures/comp_user}
+%	\caption{\textbf{\sys compiler code overhead measured in terms of execution time per application} (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
+%	\label{fig:comp_user}
+%\end{figure}
 

### 2018-04-13 paper/hwswsetup.tex

Our hardware and software suite used to evaluate \sys novel intermittent execution model represents the  real implementation of low-power intermittent system. As our implementation makes no system assumptions (we execute \sys model on a \emph{real} hardware powered by \emph{real} intermittent source), therefore promises to generate the ultimate set of \sys performance metrics.

\subsection{Intermittent Platform and Hardware Toolchain}
\label{sec:results_hardware_software}

We evaluated \sys in two power supply environments: (i)  \emph{intermittent} and (ii) \emph{fixed} (to benchmark the intermittent power supply results). In the experiments with intermittent power supply, we evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp}---energy-harvesting platform build around a MSP430FR5969~\cite{wolverine} MCU with 64\,kB of FRAM. We powered the WISP using RF signal generator~\cite{} emitting 20\,dBm sinusoidal wave at 915\,MHz. \TODO{check if true} Signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic antenna~\cite{atlas2015}. For distance-controlled experiments, we affixed WISP at two paper protrusions around 10\,cm from the surface of the table (parallel to, and at the edge of, table surface). Antenna of the signal generator was facing WISP directly and located at three distances, $d=\{15, 30, 50\}$\,cm, from WISP. For task coalescing experiments, an additional distance of $d=60$\,cm was considered. Both antennas were parallel to each other and no physical objects were located between the WISP and signal generator antenna. For continuous power experiments we used an alternative setup with MSP-TS430RGZ48C launchpad~\cite{}, which had the same MCU as WISP. \TODO{Is this the right launchpad model?} In all the experiments the timing measurements were performed using the Saleae~\cite{saleae} logic analyzer. \TODO{How about paging measurements?}

\subsection{Software Benchmarks}
\label{sec:software_benchmarks}

We evaluated \sys using benchmarks that originated from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}. Specifically, we have implemented three benchmarks used by Chain~\cite{chain} and Alpaca~\cite{alpaca} (source code is accessible~\cite{coala_website}), namely: \textbf{ar}: nearest neighbor classification of randomly generated data modeling a three-axis accelerometer; \textbf{bc}: counts bits in byte pseudo-random string using several algorithms, and compares results for correctness; \textbf{cuckoo}: Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence. Additionally, we have ported three new benchmarks from~\cite{mibench}, namely: \textbf{fft}: continuously repeating Fast Fourier Transform on three pre-generated 128 sample vector; and \textbf{sort}: selection sort algorithm on a fixed sequence of numbers. \todo{justify the selection of these benchmarks} All applications are summarized in Table~\ref{table:compiler_result} in terms of memory footprint, source lines of code and number of tasks. \TODO{update table with new data}

All applications were compiled using GCC version 3.8.0. \TODO{Check if this still holds} We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}. We do not compare \sys against Chain~\cite{chain}, another task-based system, which is inferior to Alpaca in its performance. Also, we do not compare \sys to checkpointing-based systems, such as DINO~\cite{dino} or Mementos~\cite{mementos} (cf. Section~\ref{sec:background_consistency}), as both of them are build on top LLVM compiler, which performance is inferior to GCC compiler used in the evaluation of \sys \todo{Please check if the LLVM compiler argument makes sense}.

### 2018-04-13 paper/paging.tex
alg 1 moved to top
added todos section
added some things to paging

@@
 non-volatile memory to a small, volatile memory on demand and writing all dirty
-pages back atomically. Since memory moves are at page granularity, \sys
-exploits hardware\hyp{}accelerated block copies.
+pages back to non-volatile memory atomically. Since memory moves at page granularity, \sys
+exploits hardware\hyp{}accelerated block copies using DMA.
 
@@
 
-As they are
-accessed by a task, pages are swapped from their home location
+Paging moves data between non-volatile and volatile memory. A page has a home
+location in non-volatile memory which we refer to as the page's {\tt store}.
+As a task accesses pages, they are swapped from their home location
 in the direct\hyp{}mapped non-volatile memory, called
-\texttt{\underline{store}} (underline denotes allocation in non-volatile
-memory) into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The
+\texttt{\underline{store}} (underlined locations are non-volatile) 
+into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The
 \texttt{working} buffer is allocated to take up the volatile memory that
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
 	\caption{\texttt{RVAR}(variable $v$)}
@@
 %despite arbitrary power failures.
+\subsection{Dynamic vs. Static Privatization}
+\TODO{Put paragraph about Alpaca's static privatization algo vs. Coala's dynamic privatization algo}
+\TODO{pass by ref might not work the right way}
+\begin{verbatim}
 
+GLOBAL_(x)
+
+foo(*y){ 
+  *y++; 
+}
+
+TASK( ){ 
+  ... ;  
+  foo( &GV(x) ) ; 
+}
+
+\end{verbatim}
 %\subsection{Managing Paging Overhead}

### 2018-04-13 paper/hwswsetup.tex
apps into table

@@
 We powered the WISP using RF signal generator~\cite{} emitting 20\,dBm sinusoidal wave at 915\,MHz.
-\TODO{check if true} Signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic antenna~\cite{atlas2015}.
+Signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic antenna~\cite{atlas2015}.
 For distance-controlled experiments, we affixed WISP at two paper protrusions around 10\,cm from the surface of the table (parallel to, and at the edge of, table surface).
@@
 Both antennas were parallel to each other and no physical objects were located between the WISP and signal generator antenna.
-For continuous power experiments we used an alternative setup with MSP-TS430RGZ48C launchpad~\cite{}, which had the same MCU as WISP.
-\TODO{Is this the right launchpad model?} In all the experiments the timing measurements were performed using the Saleae~\cite{saleae} logic analyzer.
-\TODO{How about paging measurements?}
+For continuous power experiments we used an alternative setup with MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}, which had the same MCU as WISP.
+In all the experiments the timing measurements were performed using either the Saleae~\cite{saleae} logic analyzer (for intermittent power experiments) or TI Code Composer Studio (for fixed power experiments).
+Both type of experiments are explained in the following section.
 
@@
 \label{sec:software_benchmarks}
+
+\begin{table}
+	\centering
+	\footnotesize
+	\begin{tabular}{| c | c |}
+		\hline
+		Application & Description \\
+		\hline\hline
+		\textbf{ar} & nearest neighbor classification of randomly generated data modeling a three-axis accelerometer\\
+		\textbf{bc} & counts bits in byte pseudo-random string using several algorithms, and compares results for correctness\\
+		\textbf{cuckoo} & Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence\\
+		\textbf{fft} & continuously repeating Fast Fourier Transform on three pre-generated 128 sample vector\\
+		\textbf{sort} & selection sort algorithm on a fixed sequence of numbers\\
+		\hline
+	\end{tabular}
+\caption{Suite of benchmark applications used in evaluating \sys performance. All applications have been chosen from MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.}
+\label{table:benchmark_table}
+\end{table}
 

### 2018-04-13 paper/hwswsetup.tex
methodology section revised (almost complete)

@@
 In the experiments with intermittent power supply, we evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp}---energy-harvesting platform build around a MSP430FR5969~\cite{wolverine} MCU with 64\,kB of FRAM.
-We powered the WISP using RF signal generator~\cite{} emitting 20\,dBm sinusoidal wave at 915\,MHz.
+We powered the WISP using RF signal generator~\cite{} \todo{Add reference to signal generator}{Przemek} emitting 20\,dBm sinusoidal wave at 915\,MHz.
 Signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic antenna~\cite{atlas2015}.
@@
 	\footnotesize
-	\begin{tabular}{| c | c |}
+	\begin{tabular}{| r | p{0.85\columnwidth} |}
 		\hline
@@
 		\textbf{ar} & nearest neighbor classification of randomly generated data modeling a three-axis accelerometer\\
+		\hline
 		\textbf{bc} & counts bits in byte pseudo-random string using several algorithms, and compares results for correctness\\
-		\textbf{cuckoo} & Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence\\
+		\hline
+		\textbf{cuckoo} & Bloom-filter-like Cuckoo filter: first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence\\
+		\hline
+		\textbf{dijkstra} &  Dijkstra algorithm executed continously on a predefined adjecency matrix\\
+		\hline
 		\textbf{fft} & continuously repeating Fast Fourier Transform on three pre-generated 128 sample vector\\
+		\hline
 		\textbf{sort} & selection sort algorithm on a fixed sequence of numbers\\
@@
 	\end{tabular}
-\caption{Suite of benchmark applications used in evaluating \sys performance. All applications have been chosen from MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.}
+\caption{Suite of benchmark applications used in evaluating \sys performance.
+All applications have been chosen from MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.\todo{Consider adding more information here}{Brandon}}
 \label{table:benchmark_table}
@@
 
-We evaluated \sys using benchmarks that originated from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.
-Specifically, we have implemented three benchmarks used by Chain~\cite{chain} and Alpaca~\cite{alpaca} (source code is accessible~\cite{coala_website}), namely: \textbf{ar}: nearest neighbor classification of randomly generated data modeling a three-axis accelerometer; \textbf{bc}: counts bits in byte pseudo-random string using several algorithms, and compares results for correctness; \textbf{cuckoo}: Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence.
-Additionally, we have ported three new benchmarks from~\cite{mibench}, namely: \textbf{fft}: continuously repeating Fast Fourier Transform on three pre-generated 128 sample vector; and \textbf{sort}: selection sort algorithm on a fixed sequence of numbers.
-\todo{justify the selection of these benchmarks} All applications are summarized in Table~\ref{table:compiler_result} in terms of memory footprint, source lines of code and number of tasks.
-\TODO{update table with new data}
+We evaluated \sys using benchmarks that originate from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.
+Set of implemented benchmarks is presented in Table~\ref{table:benchmark_table} and their implementation is available via~\cite{coala_website}.
+Three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) were implemented by Chain~\cite{chain} and Alpaca~\cite{alpaca}.
+Three remaining ones (\textbf{dijkstra}, \textbf{fft}, \textbf{sort}) are added as new.
+\todo{justify the selection of these benchmarks}{Przemek, Carlo, Amjad, Brandon}
+
+%Specifically, we have implemented three benchmarks used by Chain~\cite{chain} and Alpaca~\cite{alpaca} (source code is accessible~\cite{coala_website}), namely: \textbf{ar}: nearest neighbor classification of randomly generated data modeling a three-axis accelerometer; \textbf{bc}: counts bits in byte pseudo-random string using several algorithms, and compares results for correctness; \textbf{cuckoo}: Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence.
+Additionally, we have ported three new benchmarks from~\cite{mibench}, namely: \textbf{dijkstra}: Dijkstra algorithm executed continously on a predefined adjecency matrix; \textbf{fft}: continuously repeating Fast Fourier Transform on three pre-generated 128 sample vector; and \textbf{sort}: selection sort algorithm on a fixed sequence of numbers.
+\todo{justify the selection of these benchmarks}
 
 All applications were compiled using GCC version 3.8.0.
-\TODO{Check if this still holds} We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
+\todo{Check if this still holds}{Przemek} We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
 We do not compare \sys against Chain~\cite{chain}, another task-based system, which is inferior to Alpaca in its performance.
 Also, we do not compare \sys to checkpointing-based systems, such as DINO~\cite{dino} or Mementos~\cite{mementos} (cf.
-Section~\ref{sec:background_consistency}), as both of them are build on top LLVM compiler, which performance is inferior to GCC compiler used in the evaluation of \sys \todo{Please check if the LLVM compiler argument makes sense}.
+Section~\ref{sec:background_consistency}), as both of them are build on top LLVM compiler, which performance is inferior to GCC compiler used in the evaluation of \sys \todo{Please check if the LLVM compiler argument makes sense}{Brandon}.

### 2018-04-13 paper/hwswsetup.tex
methodology section revised (almost complete)

@@
 We evaluated \sys in two power supply environments: (i)  \emph{intermittent} and (ii) \emph{fixed} (to benchmark the intermittent power supply results).
-In the experiments with intermittent power supply, we evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp}---energy-harvesting platform build around a MSP430FR5969~\cite{wolverine} MCU with 64\,kB of FRAM.
-We powered the WISP using RF signal generator~\cite{} \todo{Add reference to signal generator}{Przemek} emitting 20\,dBm sinusoidal wave at 915\,MHz.
+In the experiments with intermittent power supply, we evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp}---energy-harvesting platform build around a MSP430FR5969~\cite{wolverine} MCU with with 64\,kB of non-volatile (FRAM) memory and and 2\,kB of volatile memory (SRAM).
+MCU was running on X\,MHz clock cycle \todo{Check clock frequency}{Przemek} We powered the WISP using RF signal generator~\cite{} \todo{Add reference to signal generator}{Przemek} emitting 20\,dBm sinusoidal wave at 915\,MHz.
 Signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic antenna~\cite{atlas2015}.

### 2018-04-13 paper/evaluation.tex
update the eval results
cleanup of evaluation

@@
-\subsection{\sys Evaluation Results}
+Our evaluation illustrates key findings about \sys. We show that: \todo{Make a list of core findings here once results are solid}{Przemek}
+
+\subsection{\sys Runtime Performance}
 \label{sec:results_evaluation}
 
-We evaluated \sys to show that it is easy to use and efficient across our
-benchmarks.
-%We evaluate \sys's ease of use by characterizing \sys's automatic task compiler.
-We evaluate \sys efficiency by comparing its runtime performance to Chain.
-We characterize \sys by studying how its performance varies due to coalescing and when using automatic task compilation versus manual task compilation.
+We evaluated \sys to show that it is easy to use and efficient across our benchmarks.
+We evaluate \sys efficiency by comparing its runtime performance to Alpaca.
+We characterize \sys by studying how its performance varies due to coalescing.
 
-%\begin{table}[t]
-%	\centering
-%	\footnotesize
-%	\begin{tabular}{|c|c|c|c|c|c|}
-%		\hline
-%		bc & cem & cuckoo & rsa & ar & sort \\
-%		\hline\hline
-%		3 & 2 & 6 & 86 & 34 & < 1 \\
-%		\hline
-%	\end{tabular}
-%	\caption{\sys compile time per application (seconds).}
-%\label{table:compile_time}
-%\end{table}
-%
-%\subsubsection{Task Coalescing Performance}
-%\label{sec:results_coalescing}
-
-%List of figures: 
-%\begin{itemize}
-%	\item Fig. 1: Viper only/per app/fixed power/(hand/complier)/no coalescing/clang [Resuls are ready by Kiwan]
-%	\item Fig. 2: Viper only/per app/per distance (fixed/20/40/60)/(hand/complier)/coalescing [Resuls are ready by Kiwan]
-%	\item Fig. 3a: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/clang [Resuls are ready by Kiwan]
-%	\item Fig. 3b: (Viper Coalescing/Chain)/per app/per distance (fixed/20/40/60)/complier/gcc [Resuls are not ready: Kiwan needs to make them]
-%	\item Fig. 4: exactly as Fig 12 in the paper (4 apps only)
-%	\item Fig. 5: average virtual task size/gcc/fixed distance/per app 
-%\end{itemize}
-
-\subsubsection{\sys Runtime Performance}
+\subsubsection{\sys Task Adaptation}
 \label{sec:result_coalescing}
 
-\textbf{Coalescing Strategies.} Figure~\ref{fig:coalescing} shows the run time
-of \sys's two coalescing strategies normalized to the run time without
+We first need to assess the performance of all coalescing strategies intriduced by us in Section~\ref{}.
+
+\textbf{Coalescing Strategies.} Figure~\ref{fig:coalescing} shows the run time of \sys's two coalescing strategies normalized to the run time without
 coalescing. We show data from experiments with intermittent power, where intermittency was generated in software
@@
 Each application ran with the same generated pattern for the fairness of comparison.
-%
 The results show the benefit from coalescing for most applications, with the
@@
 while others suffering, due to decisions based on inaccurate prediction of the
-actual task energy consumption.
-The TSHC gains are particularly small (or negative) for applications with
-narrower range of task sizes in Figure~\ref{fig:task_profiling}.
+actual task energy consumption. The TSHC gains are particularly small (or negative) for applications with narrower range of task sizes in Figure~\ref{fig:task_profiling}.
 %\todo{say why this difference is small}{Amjad}.
@@
 	\includegraphics[width=\columnwidth]{figures/coalStrategies}
-	\caption{\textbf{\sys coalescing strategies performance per application on intermittent power compared to \sys without coalescing (`NONE').}}
-%\todo{Change WHD to TSHC and HD to HC}{Amjad}}
+	\caption{\sys coalescing strategies performance per application on intermittent power compared to \sys without coalescing. \todo{Add new figure; comment on the result}{Przemek}}
 	\label{fig:coalescing}
@@
 
-\textbf{\sys versus Task-based Model.} We evaluate \sys's run time performance and compare \sys to the run time performance of Chain.
+\textbf{\sys versus Task-based Model.} We evaluate \sys's run time performance and compare \sys to the run time performance of Alpaca.
 Results are given in Figure~\ref{fig:runtime}.
-For each application we measured its execution time using real wireless power provision with RFID reader for continuous power, i.e.
-0\,m RFID antenna/WISP distance (the first two bars of each group) and on intermittent power at 20\,cm and 40\,cm from the RFID antenna (the second two and third two bars of each group, respectively).
+For each application we measured its execution time using real wireless power provision from the RF signal generator (refer to Section~\ref{}).
+
 The data show the average execution time of each application run repeatedly for 20 seconds, with runtimes normalized to \sys's average.
@@
 dft) or access more different memory pages, \sys incurs overhead from memory virtualization that cause its performance to be comparable to (or in the \emph{single} cem case, worse than) Chain.
+\todo{Rewrite conclusions}{Przemek}
 
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coala_chain_clang}
+	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
 	\caption{\textbf{Performance of \sys applications (divided manually into tasks) at multiple WISP to RFID reader antenna distances}: fixed power (0\,m to the antenna) and \{0.2, 0.4\}\,m (left, center and right-most bar per application, respectively), compared against Chain (results normalized).}
@@
 
-%\begin{table}
-%	\centering
-%	\footnotesize
-%	\begin{tabular}{|r|c|c|c|}
-%		\hline
-%		{} & bc & dft & cem \\
-%		\hline\hline
-%		Avg. no. coalesced tasks & 56 & 3 & 26 \\
-%		Avg. task size (MCU cycles) & 657 & 9680 & 722 \\
-%		\hline
-%	\end{tabular}
-%\caption{Average \sys virtual task size (at 60\,cm WISP to RFID antenna distance) per manually task split application.}
-%\label{tab:aveVirtuTaskSize}
-%\end{table}
+\textbf{Task Downscaling.} \todo{Provide text; add results}{Przemek}
 
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\columnwidth]{figures/averageVirtualTaskSize}
-%	\caption{Average \sys virtual task size at fixed power per application. Top (green): number of coalesced tasks; bottom (red): average size of real task.}
-%	\label{fig:aveVirtuTaskSize}
-%\end{figure}
+\subsection{Characterization of \sys Overhead}
+\label{sec:coala_overhead}
 
@@
 
-%\subsubsection{\sys's Compiler}
-%\label{sec:result_compiler_time}
+\subsection{\sys Programs Characterization}
+\label{sec:results_program_characterization}
+
+\sys program characterization is provide in Table~\ref{table:compiler_result}. \todo{Finalize this section}{Przemek}
 
@@
 	\end{tabular}
-	\caption{\textbf{Comparison between \sys and Chain source code line number (SLOC column) and comparison between compiler-generated \sys code (Comp.) and hand-written \sys code (Man.) (other columns)}; $^\dagger$dft does not compile with Clang }\label{table:compiler_result}\vspace{-0.5cm}
+		\caption{Comparison between \sys and Alpaca benchmarks.}
+		\label{table:compiler_result}\vspace{-0.5cm}
 \end{table}
-
-%
-
-%\textbf{\sys Compiler Generated Code Characterization.} The behaviour and the efficiency metrics of the \sys compiler are given in Table~\ref{table:compiler_result}.
-\todo{update table with new data}
-%
-%for each application, for manually written (Man.) and automatically compiled (Comp.) benchmarks the difference in the number of bytes of protected data (columns 2 and 3), the number of tasks (columns 4 and 5), the number of \sys protected data API access calls (columns 6 and 7), the lines of code in each application for \sys and Chain (columns 8 and 9), and compilation time using \sys's compiler (column 10).
-%
-%On average, compiler-generated code introduces about 0.3\% more bytes of protected data, suggesting that liveness analysis captures natural boundaries in the use of variables also caught by a manual implementation.
-Compiler-generated code generates about 16\% fewer tasks for applications, on average.
-The results indicate that the \sys compiler selects task boundaries differently than a programmer would manually, favoring fewer, larger tasks.
-The \sys compiler inserts protected data access operations conservatively when a variable {\em may} alias a protected variables, due to the limitations of pointer aliasing.
-Consequently, the \sys compiler introduces 48\% more protected data accesses than programmer-written code.
-The run time results in Section~\ref{sec:result_compiler_time} show that the increase in protected data accesses sometimes degrades performance.
-Overall, the data show that the compiler automatically generates tasks similarly to manually written tasks.
-Doing so, \sys's compiler reaps the static task benefits of \sys, but simplifies programming by sparing the programmer the job of task decomposition.
-
-%\textbf{Compile Time.} We measured compile time for \sys's compiler's automatic task generation for each application\footnote{We could not compile the dft application since LLVM inserts mathematical built-in functions (i.e.
-\texttt{\_\_muldf3}, \texttt{\_\_floatsidf}, \texttt{\_\_adddf3}) whose implementation were not provided by the LLVM MSP430 back-end.}.
-We compiled the applications on a 4-core virtual machine with 1\,GB of memory running Arch Linux on an Intel Core i5-6200U at 2.30\,GHz.
-We collected results using the Linux system time collected in seconds.
-Most of the compile time was spent computing pointer alias analysis, while calculating the possible live ranges of variables.
-Applications such as rsa that have a larger code size and manipulate more variables have a longer compilation time, due to a larger search space for pointer aliasing.
-Nonetheless, the compile time is orders of magnitude shorter than manually defining tasks, which requires hours of programmer time.
-%The characterization of \sys's compiler suggests that the \sys compiler enables simpler task-based intermittent programming for \sys.
-
-%\textbf{Run Time of \sys Compiler Code.} In addition to the characterization of the \sys compiler in Table~\ref{table:compiler_result}, we experimentally assessed the performance of the \sys compiler, compared to the performance of the manually written benchmarks, as written in prior work~\cite{alpaca,chain}.
-Figure~\ref{fig:comp_user} compares the performance of manually written benchmark code to the performance of code automatically compiled using \sys.
-The figure shows run time for both cases, normalized to the run time of the manually written program, running on continuous power.
-The data show that in all cases except for ar and rsa, the run time performance of compiled code generated by the \sys compiler has comparable performance to manually written code.
-With comparable performance in most cases, the data suggest that the \sys compiler not only makes programming simpler, but does so, in most cases, without a high run time overhead.
-In the cases of both ar and rsa, the code manipulates a large amount of data that is stored in an array and passed into functions by reference.
-We expect that the data being manipulated by reference is leading to a large amount of may-alias references in the compiler's analysis, which leads to a larger number of accesses to protected data in the automatically compiled code, compared to the manually written code.
-
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\columnwidth]{figures/comp_user}
-%	\caption{\textbf{\sys compiler code overhead measured in terms of execution time per application} (dft excluded as it was not compiled by \sys compiler, see Table~\ref{table:compiler_result}); WISP was powered continuously.}
-%	\label{fig:comp_user}
-%\end{figure}
-
-%\subsubsection{Case Study}
-%\label{sec:case_study}
-
-%The final experiment assessing usefulness of \sys's use in real application.
-For this we have a implemented an battery-less wirelessly-powered sound detector as an example.
-To the best of our knowledge, this is the first case study that demonstrates a intermittently-powered system supplied by harvested energy at runtime, which continuously interacts with real-life signals.
-
-%The application aims at detecting a specific tone in the audio signal (in the implementation: 1\,kHz) based on on-the-fly audio measurement.
-For this we have connected an analog MEMS microphone~\cite{microphone} output to AUX3 port of WISP\,5.1 (while the microphone was powered directly by WISP, which was then wirelessly power-supplied by the RFID reader).
-Audio signal is sampled by the microcontroller's ADC and passed to the FFT function (internal function of the TI's Digital Signal Processing library~\cite{ti_dsp}) to find a tone above a predefined threshold.
-The output of the detection is signalled by the WISP on-board LED.
-The video of demonstrating the working system is available via \url{https://link_to_a_video}.
-
-%Expected results:
-%\begin{itemize}
-%	\item Compare execution time: \sys against Chain for X\,s long audio signal detection at two antenna distances
-%	\item Make a video of the demonstration 
-%\end{itemize}
-
-%\todo{Check if we really need this application}{Brandon}
-%\todo{Provide results and text to this section---if time allows (this result is not critical)}{Amjad/Przemek}

### 2018-04-13 paper/evaluation.tex
table holder for apps evaluation

@@
 
-\begin{table}[t]
-	\centering
-	\footnotesize
-	\renewcommand{\tabcolsep}{1pt}
-	\begin{tabular}{|l|cc|cc|cc|cc|c|}
+\begin{table}
+	\begin{tabular}{| c | c | c | c | c |}
 		\hline
-		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
-		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
+		Application & \multicolumn{2}{| c |}{Memory footprint} & No. tasks & SLOC \\
+		{} & \sys & Alpaca & {} & {} \\
 		\hline\hline
-		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
-		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
-		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
-		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
-		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
-		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
-		%dd &  &  &  &  &  &  &  & 287 &  \\
+		\textbf{ar} & --- & --- & --- & ---\\
 		\hline
@@
 \end{table}
+
+%\begin{table}[t]
+%	\centering
+%	\renewcommand{\tabcolsep}{1pt}
+%	\begin{tabular}{|l|cc|cc|cc|cc|c|}
+%		\hline
+%		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
+%		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
+%		\hline\hline
+%		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
+%		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
+%		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
+%		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
+%		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
+%		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
+%		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
+%		%dd &  &  &  &  &  &  &  & 287 &  \\
+%		\hline
+%	\end{tabular}
+%		\caption{Comparison between \sys and Alpaca benchmarks.}
+%		\label{table:compiler_result}\vspace{-0.5cm}
+%\end{table}

### 2018-04-14 paper/background.tex
intellectually modtivating coala design

@@
 in Section~\ref{sec:task_coalescing}, is motivated by these observations.
+
+\textbf{Intellectually Motivating the Design of \sys's Execution Model.}  
+The challenge of \em{efficiently} executing a program under frequent power failures, lies in the fact the the program needs to be energy aware.
+This challenge becomes even bigger when \em{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
+\em{flexibility} is key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software  technology.
+\em{\sys champions an efficient, general, and flexible intermittent execution.}

### 2018-04-14 paper/background.tex
parsing latest commit by Amjad to background section

--- 2018-04-12 paper/background.tex
@@
 The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead.
-\sys's {\em task coalescing}, described
-in Section~\ref{sec:task_coalescing}, is motivated by these observations.
+\sys's {\em task coalescing}, described in Section~\ref{sec:task_coalescing}, is motivated by these observations.
+
+\textbf{Motivating the Design of \sys's Execution Model.} The challenge of \em{efficiently} executing a program under frequent power failures, lies in the fact the the program needs to be energy aware.
+This challenge becomes even bigger when \em{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
+\em{Fllexibility} is key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software  technology.
+\em{\sys champions an efficient, general, and flexible intermittent execution.}

### 2018-04-14 paper/background.tex
intellectually modtivating coala design

--- 2018-04-12 paper/background.tex
@@
 The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead.
-\sys's {\em task coalescing}, described
-in Section~\ref{sec:task_coalescing}, is motivated by these observations.
+\sys's {\em task coalescing}, described in Section~\ref{sec:task_coalescing}, is motivated by these observations.
+
+\textbf{Motivating the Design of \sys's Execution Model.} The challenge of \emph{efficiently} executing a program under frequent power failures, lies in the fact that the program needs to be energy aware.
+This challenge becomes even bigger when \emph{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
+\emph{Flexibility} is a key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software technology.
+\emph{\sys champions an efficient, general, and flexible intermittent execution.}

### 2018-04-14 paper/background.tex
intellectually modtivating coala design

@@
 \textbf{Motivating the Design of \sys's Execution Model.} The challenge of \emph{efficiently} executing a program under frequent power failures, lies in the fact that the program needs to be energy aware.
-This challenge becomes even bigger when \emph{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
+This challenge becomes even bigger when \emph{generality} is a design requirement, and consequently, the system must not require a specific hardware circuitry to offer its efficiency.
 \emph{Flexibility} is a key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software technology.

### 2018-04-14 paper/background.tex
first pass over motivating coala execution model design

--- 2018-04-14 paper/background.tex
@@
 
-\textbf{Motivating the Design of \sys's Execution Model.} The challenge of \em{efficiently} executing a program under frequent power failures, lies in the fact the the program needs to be energy aware.
-This challenge becomes even bigger when \em{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
-\em{Fllexibility} is key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software  technology.
-\em{\sys champions an efficient, general, and flexible intermittent execution.}
+\subsection {Motivating the Design of \sys's Execution Model.} 
+%% Coala’s model motivation
+The challenge of \emph{efficiently} executing a program under frequent power failures, lies in the fact that the program needs to be energy aware.  This challenge becomes even bigger when 
+Hardware support is not allowed to advocate the approach \emph{generality} and to eliminate hardware energy consumption overhead.
+Furthermore, \emph{flexibility} is a key for wide adaptation, therefore, \sys's execution model should not limit a programmer to a subset of the functionalities of the underlying software technology---\emph{\sys champions an efficient, general, and flexible intermittent execution.}
+
+%% How does Coala’s execution model progress?
+\sys execution model operates on task-based programming model, but It progresses on a dynamic, virtual task granularity.
+A virtual task can be either multiple static tasks \emph{coalesced} together, when the energy permits, or a sub-task, when it is necessary to preserve forward progress.
+However, as more tasks coalesce, their  collective commit overhead amortizes better, but the risk of wasting work also increases.
+If power fails during a long sequence of coalesced tasks, execution will restart from the last commit, i.e.
+the first task in the sequence, losing  the progress made by any of the coalesced tasks.
+The challenge to coalescing tasks is determining how many tasks to coalesce before committing.
+
+%% How does Coala’s execution model decide on its step size when there is abandoned energy?
+\sys uses the recent execution history---pure software technique---as a metric to estimate the amount of the available energy and to set the adequate virtual task size accordingly.
+This fundamental feature prevents \sys from blind virtual task adaptation which can result in a repeated power failures problem.
+For example, If \sys enlarges the virtual task size by $x$ static tasks on a virtual task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, 10 times to reduce the virtual task size from 10 to one when $x=1$.
+
+%% How does Coala’s execution model decide on its step size when there is scarce energy?
+
+Coala's execution model is neither limited by a static task size to progress, nor by the commit size, i.e.
+the amount of data that needs to be saved to the non-volatile memory to preserve execution progress.
+Coala can partition compile-time tasks, if it fails on a single static task more than once, or merge them, if saving the state to non-volatile memory is redundant because a power failure is not expected soon.
+This unique technique makes the commit size proportional to the number of state-less instructions executed.
+In other words, if energy allows executing only a small chunk of code, the amount of data to be committed shrinks accordingly.

### 2018-04-14 paper/newCoalescing.tex
added placeholder for partial commit
remove the history our algo and its fig

--- 2018-04-09 paper/newCoalescing.tex
@@
+\subsection{Task Coalescing}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coal_Intro_Figure.pdf}%coalescing_timeline
+	\caption{ When $N$ tasks coalesced $N-1$ commit operations is skipped.}
+	\label{fig:coalIntro}
+\end{figure}
+
+When incoming energy is sufficient to run multiple consecutive tasks,
+committing state after each task is wasteful. \sys mitigates this
+waste by deferring the expensive commits and performing them in a
+batch. In the resulting execution, several tasks are \emph{merged} into a
+single virtual task---a process we call \emph{coalescing}. 
+%
+When $N$ tasks are coalesced, $N-1$ boundaries are crossed and some number of
+pages are dirtied. The first $N-2$ executions of \transition statement skip
+committing dirty pages to non-volatile memory, leaving the work of committing
+pages modified by any of the $N$ tasks to the statement $N-1$.
+Coalescing is likely to reduce the total number of non-volatile memory
+accesses, because when coalesced tasks access the same pages of data, only the
+last version of the page needs to be committed.  \textcolor{red}{maybe is possible to demonstrate with data? count number of executed non-volatile vs volatile accesses}
+
 \begin{algorithm}
@@
 
-\subsection{SC: Slow Coalescing}
+\subsubsection{SC: Slow Coalescing}
 \label{subsec:slowCoalescing}
@@
 
-\subsection{HC: History-aware Coalescing}
-\label{subsec:hCCoalescing}
+% \subsubsection{HC: History-aware Coalescing}
+% \label{subsec:hCCoalescing}
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/HCCoal}%coalescing_timeline
-	\caption{The History-aware coalescing algorithm  updates its coalescing target}
-	\label{fig:HCCoal}
-\end{figure}
+% \begin{figure}
+% 	\centering
+% 	\includegraphics[width=\columnwidth]{figures/HCCoal}%coalescing_timeline
+% 	\caption{The History-aware coalescing algorithm  updates its coalescing target}
+% 	\label{fig:HCCoal}
+% \end{figure}
 
-The History-aware algorithm uses its recent execution history as metric to estimate the optimal coalesce task size.
+% The History-aware algorithm uses its recent execution history as metric to estimate the optimal coalesce task size.
 Relying on the history of execution enables this algorithm to react very quickly to discharge rates variation.

### 2018-04-14 paper/newCoalescing.tex
remove some redundant text

@@
 %% Coalescing motivation
-Finding an optimal task size, given random energy conditions is an open question.
+
+% Finding an optimal task size, given random energy conditions is an open question.
 Obviously, a static approach does not have the potential to answer it.
@@
 Furthermore, \sys advocates making intermittently powered software energy-aware is the key for \emph{efficient code execution and probability}.
-%However, given the extremely limited recourses any optimization technique must adhere to the principle of simplicity, otherwise it introduces a non-tolerable overhead.
 
 % %% address the challenge and define the optimal task
-\sys advances its execution in a state-less (or virtual) manner, and then it frequently saves its forward progress.
+
+% \sys advances its execution in a state-less (or virtual) manner, and then it frequently saves its forward progress.
 The longer \sys virtually progresses, the less committing (saving data to non-volatile memory) overhead it introduces.
@@
 
-% \subsubsection{HC: History-aware Coalescing}
-% \label{subsec:hCCoalescing}
+\subsubsection{EC: Energy-aware Coalescing}
+\label{subsec:ECoalescing}
 

### 2018-04-14 paper/hwswsetup.tex
methodologu section revised

@@
-Our hardware and software suite used to evaluate \sys novel intermittent execution model represents the  real implementation of low-power intermittent system.
-As our implementation makes no system assumptions (we execute \sys model on a \emph{real} hardware powered by \emph{real} intermittent source), therefore promises to generate the ultimate set of \sys performance metrics.
+Our implementation of \sys makes no system assumptions: we execute \sys on a \emph{real} low-power embedded computing platform powered by a \emph{real} intermittent energy harvesting source.
+This approach shall lead to undisputed set of \sys performance metrics.
 
@@
 In the experiments with intermittent power supply, we evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp}---energy-harvesting platform build around a MSP430FR5969~\cite{wolverine} MCU with with 64\,kB of non-volatile (FRAM) memory and and 2\,kB of volatile memory (SRAM).
-MCU was running on X\,MHz clock cycle \todo{Check clock frequency}{Przemek} We powered the WISP using RF signal generator~\cite{} \todo{Add reference to signal generator}{Przemek} emitting 20\,dBm sinusoidal wave at 915\,MHz.
+MCU was operating at 1\,MHz clock cycle.
+We powered the WISP using RF signal generator~\cite{} \todo{Add reference to signal generator}{Przemek} emitting 20\,dBm sinusoidal wave at 915\,MHz.
 Signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic antenna~\cite{atlas2015}.
-For distance-controlled experiments, we affixed WISP at two paper protrusions around 10\,cm from the surface of the table (parallel to, and at the edge of, table surface).
+For distance-controlled experiments, we affixed WISP at two paper protrusions around 10\,cm from the surface of the table (parallel to, and at the edge of, the table surface).
 Antenna of the signal generator was facing WISP directly and located at three distances, $d=\{15, 30, 50\}$\,cm, from WISP.
-For task coalescing experiments, an additional distance of $d=60$\,cm was considered.
-Both antennas were parallel to each other and no physical objects were located between the WISP and signal generator antenna.
+For the task downscaling experiments, an additional distance of $d=60$\,cm was considered.
+Both antennas were parallel to each other and no physical objects were located between the WISP and the signal generator antenna.
 For continuous power experiments we used an alternative setup with MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}, which had the same MCU as WISP.
 In all the experiments the timing measurements were performed using either the Saleae~\cite{saleae} logic analyzer (for intermittent power experiments) or TI Code Composer Studio (for fixed power experiments).
-Both type of experiments are explained in the following section.
+\todo{provide version and more details on TI CCS}{Przemek} Both type of experiments are explained in the following section.
 
@@
 \caption{Suite of benchmark applications used in evaluating \sys performance.
-All applications have been chosen from MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.\todo{Consider adding more information here}{Brandon}}
+All applications have been chosen from MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.\todo{Consider adding more information here}{Przemek, Carlo}}
 \label{table:benchmark_table}
@@
 
-We evaluated \sys using benchmarks that originate from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.
-Set of implemented benchmarks is presented in Table~\ref{table:benchmark_table} and their implementation is available via~\cite{coala_website}.
-Three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) were implemented by Chain~\cite{chain} and Alpaca~\cite{alpaca}.
-Three remaining ones (\textbf{dijkstra}, \textbf{fft}, \textbf{sort}) are added as new.
+We evaluated \sys using six benchmarks summarized in Table~\ref{table:benchmark_table}, with their implementation is available via~\cite{coala_website}.
+Five originate from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}: three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) were implemented by Chain~\cite{chain} and Alpaca~\cite{alpaca}, while two (\textbf{dijkstra}, \textbf{fft}) are added as new.
+The final one (\textbf{sort}) has been written from scratch.
 \todo{justify the selection of these benchmarks}{Przemek, Carlo, Amjad, Brandon}
@@
 All applications were compiled using GCC version 3.8.0.
-\todo{Check if this still holds}{Przemek} We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
+\todo{Check if this still holds.
+provide optimization flag}{Przemek} We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
+For correct comparison, all benchmarks of \sys were divided into tasks in the same way as Alpaca.
+Moreover, since original Alpaca runtime was written in Clang, we transformed every benchmark manually into GCC-complaint code.
+\todo{Provide more information about porting to GCC from LLVM}{Brandon, Carlo}
+
 We do not compare \sys against Chain~\cite{chain}, another task-based system, which is inferior to Alpaca in its performance.

### 2018-04-14 paper/background.tex
last part of background revised - commented out the last paragraph added by Amjad: repeats what earlier paragraph said

--- 2018-04-14 paper/background.tex
@@
 
-Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity.
-Buffer sizes from prior work vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}.
+Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity.
+Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}.
 Sizing tasks complicates {\em porting} code from one device to another.
-An excessively costly task on a device with a small energy buffer may be a relatively short, task on a device with a larger buffer.
+An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer.
 The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead.
-\sys's {\em task coalescing}, described in Section~\ref{sec:task_coalescing}, is motivated by these observations.
+\sys's execution model based on {\em task coalescing}---merging or dissecting multiple tasks {\em during} execution based on the program execution behavior, and {\em task downscaling}---shortening task boundaries during execution when even the smallest task does not fit the energy buffer, both models described in Section~\ref{sec:task_adaptation}, is motivated by these observations.
 
-\textbf{Motivating the Design of \sys's Execution Model.} The challenge of \emph{efficiently} executing a program under frequent power failures, lies in the fact that the program needs to be energy aware.
-This challenge becomes even bigger when \emph{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
-\emph{Flexibility} is a key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software technology.
-\emph{\sys champions an efficient, general, and flexible intermittent execution.}
+%\textbf{Motivating the Design of \sys's Execution Model.} The challenge of \em{efficiently} executing a program under frequent power failures, lies in the fact the the program needs to be energy aware.
+This challenge becomes even bigger when \em{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
+\em{Fllexibility} is key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software technology.
+\em{\sys champions an efficient, general, and flexible intermittent execution.}

### 2018-04-14 paper/background.tex
editorial changes to background: table 1 and intro passage to sec 2.3.1

@@
 		\hline\hline
-		Mementos~\cite{mementos}                             & Reg. + Stack     \\
-		DINO~\cite{dino}                                     & Reg. + WAR variables \\
-		Chain~\cite{chain}                                   & PC   + Channel data\\
+		Mementos~\cite{mementos}	& Registers + Stack     \\
+		DINO~\cite{dino}	& Registers + WAR variables \\
+		Chain~\cite{chain}	& PC + channel data\\
 		Alpaca~\cite{alpaca}	& PC + WAR variables \\
-		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
+		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
 		Region Formation~\cite{baghsorkhi_cgo_2018} & \textcolor{red}{XXX} \\
@@
 	\end{tabular}
-	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
+	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{PC}: program counter, \emph{channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
 	\label{table:chechpoint_comparison}
@@
 \subsubsection{Costs of Task-based Models}
+\label{sec:cost_task-based}
 
-%Task-based models incur costs that motivate \sys.
+Despite its advantages, the task-based programming model for intermittently-powered computing devices has its limitations.
+Two major ones being: {\em memory overhead} and {\em inflexibility to changing energy conditions}.
 
@@
 
-%\begin{figure}
-%\centering
-	%\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
-	%\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
-	%\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
-	%\label{fig:framEnergy}
-%\end{figure}
-
 To understand these non-volatile memory overheads, for the MSP430FR5969~\cite{msp430datasheet} MCU we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations.
@@
 Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device.
-In an intermittent device, energy cost corresponds to decreased run time between power failures.
+This observation is also consistent with the TI specification~\cite{}.
+\todo{add citation to TI specification on FRAM energy cost}{Przemek} In an intermittent device, energy cost corresponds to decreased run time between power failures.
 The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.

### 2018-04-14 paper/intro.tex
sinan figure re-added to intro
minor changes
remove the history our algo and its fig
intro figure text cleaned up

--- 2018-04-12 paper/intro.tex
@@
+%%%%%%%%%%%%%%%%%%%%%% IMPORTANT INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% SRAM read/write latency is in the order of 8ns 		 [ Intermittent computation without hardware support of programmer intervention]
+% FRAM read/write latency is in the orfer of 60ns/100ns  [Texas Instrument MSP430FR573x DataSheet]
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+
 Advances in processor efficiency along with the development of
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/coalescing-is}
-    \caption{\textbf{An execution of three
-statically-defined tasks without (top) and with (bottom) dynamic
-coalescing.} Task coalescing at runtime reduces time and energy overhead in
-task-based intermittent programming models by performing fewer commits and
-buffering updates to non-volatile (NV) variables in volatile (V) memory.}
+    \includegraphics[width=0.5\textwidth]{figures/graffle/intro-figure.pdf} By Sinan
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits and buffering updates to non-volatile (NV) variables in volatile (V) memory.}
     \label{fig:coalesce}

### 2018-04-14 paper/newCoalescing.tex
slow colaescing figure wrapped
intro colaescing figure wrapped

@@
 
-\begin{figure}
+\begin{wrapfigure}{t!}{0.5\textwidth}
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coal_Intro_Figure.pdf}%coalescing_timeline
-	\caption{ When $N$ tasks coalesced $N-1$ commit operations is skipped.}
+	\includegraphics[width=0.5\columnwidth]{figures/coal_Intro_Figure.pdf}
+	\caption{When $N$ tasks coalesced $N-1$ commit operations are skipped.}
 	\label{fig:coalIntro}
-\end{figure}
+\end{wrapfigure}
 
@@
 
-\subsubsection{SC: Slow Coalescing}
+\subsubsection{Slow Coalescing}
 \label{subsec:slowCoalescing}
 
-\begin{figure}
+\begin{wrapfigure}{t!}{0.5\textwidth}
 	\centering
-	\includegraphics[width=\columnwidth]{figures/slowCoal}%coalescing_timeline
-	\caption{The Slow coalescing algorithm \emph{linearly} updates its coalescing target}
+	\includegraphics[width=0.5\columnwidth]{figures/slowCoal}
+	\caption{The {\em slow coalescing} algorithm \emph{linearly} updates its coalescing target.}
 	\label{fig:slowCoal}
-\end{figure}
+\end{wrapfigure}
 
@@
 
-\subsubsection{EC: Energy-aware Coalescing}
+\subsubsection{Energy-aware Coalescing}
 \label{subsec:ECoalescing}

### 2018-04-14 paper/background.tex
text to be considered by Brandon, Przemek

@@
+\textcolor{red}{Text to be considered by: Brandon Przemek}
+
+\textcolor{red}{The challenge of \emph{efficiently} executing a program under frequent power failures, lies in the fact that the \emph{program needs to be energy aware}.
+\sys uses the recent execution history---pure software technique---as a metric to estimate the amount of the available energy and to set the adequate virtual task size accordingly.
+This fundamental feature prevents \sys from blind virtual task adaptation which can result in a repeated power failures problem.
+For example, If \sys enlarges the virtual task size by $x$ static tasks on a virtual task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, 10 times to reduce a virtual task size from 10 to one, when $x=1$}
+
+\textcolor{red}{Why hardware based checkpoint systems are static? and why \sys's partial commit is dynamic?}
+
+\textcolor{red}{(i) Stopping at arbitrary locations in the code may result in incorrect computation or even crash the program.
+If a system checkpoints between, for example, the DMA initialization and the DMA call the program will crash.
+One might think that a simple solution is to disable interrupts at these critical regions.
+Although this is the correct solution, the checkpoint-based system will "theoretically" stuck (non-termination problem) because it has to checkpoint at this exact location since it is a threshold-based system, so it is a static system.
+However, \sys's partial commit is not threshold based therefore it allows the programmer to disable interrupts at certain locations and it will partially commit around it.}
+
+\textcolor{red}{(ii) A Checkpoint system normally checkpoints all volatile data.
+Therefore, it suffers from moving large amount of data on each checkpoint.
+However, \sys's partial commit technique relies on a paging mechanism, which makes the size of the commit proportional to the number of state-less instructions executed.
+In other words, if energy allows executing only a small chunk of code, the amount of data to be committed shrinks accordingly.
+As such, \sys's execution model is neither limited by a static task size to progress, nor by the commit size,
+}
+
+\textcolor{red}{
+\emph{flexibility} is a key for wide adaptation, therefore, \sys's execution model should not limit a programmer to a subset of the functionalities of the underlying software technology}
+
 %Here, we provide background on energy harvesting systems and the intermittent software execution model.

### 2018-04-14 paper/partialcommit.tex
small fixes cleanup

@@
-\subsection{Task Downsizing} %Partial Commit/Partial Execution
+\subsection{Task Downsizing}
+\label{sec:task_downsizing}
 
@@
 If a power failure occurs after saving partial task progress, the system makes forward progress, albeit without task atomicity.
-Coala may save partial task progress multiple times before completing a very long task.
-To avoid the fail-stop case of non-termination, Coala's partial commit mechanism will split tasks that have not been marked as atomic by the programer.
+\sys may save partial task progress multiple times before completing a very long task.
+To avoid the fail-stop case of non-termination, \sys's partial commit mechanism will split tasks that have not been marked as atomic by the programmer.
 
@@
 
-Coala starts with a big random guess and down-scales it using a binary search until it succeeds in preserving the partial execution progress.
+\sys starts with a big random guess and down-scales it using a binary search until it succeeds in preserving the partial execution progress.
 The obtained value is the initial seed for the next partial commit procedure.
@@
 
-\subsection{Why partial commit is not Coala's default behavior}
+\subsection{Why partial commit is not \sys's default behavior}
 
-\emph{Static deployment of partial commit stubs} This approach reduces programmer effort - they still need to apply Coala's API and distribute trigger points for partial commits - but it introduces an additional overhead by saving the stack and the register file each time it commits the pages.
+\emph{Static deployment of partial commit stubs} This approach reduces programmer effort---they still need to apply \sys's API and distribute trigger points for partial commits---but it introduces an additional overhead by saving the stack and the register file each time it commits the pages.
 

### 2018-04-14 paper/hwswsetup.tex

@@
 We evaluated \sys using six benchmarks summarized in Table~\ref{table:benchmark_table}, with their implementation is available via~\cite{coala_website}.
-Five originate from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}: three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) were implemented by Chain~\cite{chain} and Alpaca~\cite{alpaca}, while two (\textbf{dijkstra}, \textbf{fft}) are added as new.
+Five benchmarks originate from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}: three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) were already implemented by Chain~\cite{chain} and Alpaca~\cite{alpaca}, while two (\textbf{dijkstra}, \textbf{fft}) are added as new.
 The final one (\textbf{sort}) has been written from scratch.

### 2018-04-14 paper/systemdescription.tex

@@
 
-\begin{figure}
+\begin{wrapfigure}{t!}{0.5\textwidth}
 	\centering
-	\includegraphics[width=\columnwidth]{figures/viper_block_diagram.pdf}
-	\caption{\sys top-level view. \textcolor{red}{redraw the figure: remove compiler, add task downscaling}}
+	\includegraphics[width=0.5\columnwidth]{figures/graffle/overview.pdf}
+	\caption{\sys top-level view. \textcolor{red}{redraw the figure: change font type}}
 	\label{fig:system_overview}
-\end{figure}
+\end{wrapfigure}
 
-\textbf{\sys Programming and Execution Model.}  To use \sys, a programmer first writes plain, imperative C code. The programmer
-then must decompose the program into tasks.
+\textbf{\sys Programming and Execution Model.}  To use \sys, a programmer first writes plain, imperative C code.
+The programmer then must decompose the program into tasks.
 To manually decompose a program into tasks, the programmer designates a set of functions as tasks, sequences control-flow between these tasks, and annotates memory accesses that manipulate data shared by multiple tasks.
+In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.
 
-%\sys's programming interface, listed in Table~\ref{tab:viper_syntax}, consists of primitives for defining tasks~\cite{chain,alpaca} and accessors for protected variables shared between tasks.
-%
-%As in prior task-based intermittent programming models~\cite{chain,alpaca}, tasks are functions that are not allowed to have any direct callers but receive control via explicit {\tt next\_task} statements only.
-The programmer identifies each task by adding the {\tt task} keyword to the function's declaration and designates one task as the {\em origin} task at which execution begins.
-After a power failure, \sys restarts the last partially executed task.
-%
-%A \sys program manipulates {\em protected} state---accessed by more than one task---using the \emph{read variable} ({\tt RVAR}) and \emph{write variable} ({\tt WVAR}) operations.
-These operations convey to \sys that the referenced data are task-shared, meaning that they must be buffered in volatile memory and committed when the executing task ends.
-%
-%Commit is a key overhead associated with task-based operation and \sys's coalescing strategy (described in Section~\ref{sec:coalescing} avoids commit when a task ends by continuing to buffer task-shared variables for multiple coalesced tasks, and committing only when all tasks complete.
+A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
+Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with automatically decomposed code would be identical.
 
-%\begin{table}
-%	\centering
-%	\footnotesize
-%	\begin{tabular}{|c|c|}
-%		\hline
-%		\textbf{Keyword} & \textbf{Description} \\
-%		\hline\hline
-%		\texttt{task foo() \{...\}} & Task definition\\
-%		\texttt{origin task foo() \{...\}} & Origin task definition\\
-%		\texttt{next\_task(t)} & Transition to task \texttt{t}\\
-%		\texttt{u = RVAR(v)} & Read protected variable \texttt{v} into \texttt{u}\\
-%		\texttt{WVAR(v, u)} & Write \texttt{u} to protected variable \texttt{v}\\
-%		\hline
-%	\end{tabular}
-%	\caption{\sys's programming interface.}
-%	\label{tab:viper_syntax}
-%\end{table}
-
-The programmer can decompose a program into tasks either {\em manually}. To do so manually,  requires reasoning similar to prior
-task-based systems~\cite{chain,alpaca}.  
-%
-%A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut,baghsorkhi_cgo_2018}.
-%
-%Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with automatically decomposed code would be identical.
-
-%Regardless of whether automatically or manually decomposed into tasks, the
-%
 The programmer compiles their task-based code, and links to the \sys runtime, producing a \sys-enabled binary.
@@
 Coalescing also amortizes per-variable commit cost, committing each location accessed by both tasks only after the second task, rather than once per task.
+Details of coalescing are presented in Section~\ref{sec:task_coalescing}.
 
-\textbf{\sys Task Downscaling Manager.} \textcolor{red}{to be written}
+\textbf{\sys Task Downscaling Manager.} Details of task downscaling are presented in Section~\ref{sec:task_downsizing}. \textcolor{red}{to be written}
 
@@
 Pages efficiently, atomically commit using a two-phase commit procedure accelerated using hardware support for direct memory access (DMA).
+Details of Virtual Memory Manager are presented in Section~\ref{sec:memory_virtulaization}.

### 2018-04-15 paper/implementation.tex
added placeholder for partial commit
paging figure
Table 2 revised (ready) and API complete

@@
-\textcolor{red}{THIS IS UNEDITED AND SHOULD NOT BE READ YET - Placeholder}
+We proceed with the explanation of the implementation of \sys programming model. \sys implementation is light and adds very little on top of a plain C language.
 
-Coala's API adds only 5 syntactic constructs to C-based languages: \texttt{COALA\_TASK}, used to declare a new task; \texttt{COALA\_PV}, used to declare a new protected variable; \texttt{COALA\_NEXT\_TASK} used to mark the next task to be scheduled after the currently executing one completes; \texttt{RP}, used to mark a protected variable read; \texttt{WP}, used to mark a protected variable write.
+\subsection{\sys Application Programming Interface}
+\label{sec:coala_api}
 
-Moreover, the programmer is required to explicitly call two API functions inside the main body: \texttt{COALA\_INIT}, passing the global initialization task to run upon the first boot as argument; \texttt{COALA\_RUN}, which hands the control to Coala's scheduler.
+Coala's API adds only five syntactic constructs to a C-based language: (i) \texttt{COALA\_TASK}: declaration of a new task, (ii) \texttt{COALA\_NEXT\_TASK}: declaration of the next task to be scheduled after the currently executing one completes, (iii) \texttt{COALA\_PV}: declaration of a new protected variable, (iv) \texttt{RP}: declaration of a protected variable read, and (v) \texttt{WP}: declaration of a protected variable write.
+Moreover, the programmer is required to explicitly call two \sys API functions inside the main program body: (i) \texttt{COALA\_INIT}: to pass as an argument the global initialization task to run upon the first boot , and (ii) \texttt{COALA\_RUN}: which hands the control to \sys's scheduler.
+Finally, when the programmer intends to \texttt{typedef} a \texttt{struct} to use it as a protected variable, she/he need to declare each structure's member by using the \texttt{COALA\_SM} macro to ensure proper memory alignment required by \sys's page handler.
 
-Finally, when the programmes intends to typedef a struct, to use it as a protected variable, they need to declare each structure's member by using the following macro: \texttt{COALA\_SM}, used to ensure proper memory alignment required by Coala's page handler.
-
-The following table summarizes Coala's API methods and their arguments.
-$T$ denotes the set of all tasks, while $V$ denotes the set of all protected variables.
-These methods trigger specific kernel functions, whose detailed implementation is given in the next paragraph.
+Table~\ref{table:coala_api} summarizes \sys's API methods and their arguments.
+\todo{Add example of a Coala program}{Carlo, Przemek} These methods trigger specific \sys kernel functions, whose explanation is given in the following sections.
 
@@
 \centering
-\begin{tabular}{| c | c |}
+\begin{tabular}{| r | p{0.7\columnwidth} |}
 	\hline
-	\textbf{Method} & \textbf{Arguments} \\
+	{Method} & {Arguments} \\
 	\hline\hline
-	\texttt{COALA\_TASK}($t$, $w_t$) & $t \in T$: new task name, $w_t$: weight of task $t$ \\
+	\texttt{COALA\_INIT}(t) & $t \in \mathbf{T}$: task to be scheduled on a first boot \\
 	\hline
-	\texttt{COALA\_PV}(type, $v$ [, size]) & type: var. type, $v \in V$: new protected var. name, size (optional)  array size\\
+	\texttt{COALA\_RUN}() & --- \\
 	\hline
-	\texttt{COALA\_NEXT\_TASK}($t$) & $t \in T$ : next task to schedule\\
+	\texttt{COALA\_TASK}($t$, $w_t$) & $t \in \mathbf{T}$: new task name, $w_t$: weight of task $t$ \\
 	\hline
-	$u$ := RP($v$) & $v \in V$: protected variable to be read, $u$: value \\
+	\texttt{COALA\_NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : next task to be scheduled\\
 	\hline
-	WP($v$) := $u$ &  $v \in V$: protected variable to be written, $u$: value \\
+	\texttt{COALA\_PV}($p$, $v$ [, $s$]) & $p$: variable type, $v \in \mathbf{V}$: new protected variable name, $s$: array size\\
 	\hline
-	\texttt{COALA\_INIT}(t) & $t \in T$: task to be scheduled on first boot \\
+	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to be read, $u$: value \\
 	\hline	
-	\texttt{COALA\_RUN}() & {} \\
+	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to be written, $u$: value \\
 	\hline
-	\texttt{COALA\_SM}(type, $m$ [, size]) &  type: struct member's type, $m$: struct member's name, size (optional): array size \\
+	\texttt{COALA\_SM}($p$, $m$ [, $s$]) & $p$: \texttt{struct} member's type, $m$: \texttt{struct} member's name, $s$: array size \\
 	\hline
 \end{tabular}
-\caption{Coala API}
+\caption{Summary of \sys API; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
 \label{table:coala_api}
@@
 
-An example usage of Coala’s API is given in the following code snippet. (TBW)
+\subsection{\sys Kernel Functions Implementation}
 
-\subsection{Kernel functions implementation}
+\textcolor{red}{THIS IS UNEDITED AND SHOULD NOT BE READ YET - Placeholder}
 
-New tasks. The API method \texttt{COALA\_TASK} statically allocates a non-volatile constant variable holding the task weight, and declares a void function with the name passed as an argument.
+\textbf{New tasks.} The API method \texttt{COALA\_TASK} statically allocates a non-volatile constant variable holding the task weight, and declares a void function with the name passed as an argument.
 
@@
 
-\subsection{Paging}
+\subsection{\sys Paging}
 
@@
 \begin{figure}
+	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
 	\caption{Illustration of page tags \textcolor{red}{to be drawn}}
@@
 
-\subsection{Scheduling and committing}
+\subsection{\sys Scheduling and Committing}
 
@@
 Target budget update. Before starting a new coalesced task, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy [...].
+
+\textcolor{red}{API of the partial commit needs to be explained as well}

### 2018-04-15 paper/implementation.tex

@@
 
-\textcolor{red}{THIS IS UNEDITED AND SHOULD NOT BE READ YET - Placeholder}
+\textbf{New Tasks.} The API method \texttt{COALA\_TASK} statically allocates a non-volatile constant variable holding the task weight, and declares a void function with the name passed as an argument.
 
-\textbf{New tasks.} The API method \texttt{COALA\_TASK} statically allocates a non-volatile constant variable holding the task weight, and declares a void function with the name passed as an argument.
+\textbf{New Protected Variables.} The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by \texttt{RP} and \texttt{WP} API methods to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in a non-volatile memory.
 
-New protected variables.
-The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by RP and WP to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in non-volatile memory.
+\textbf{Task Transitions.} By calling \texttt{COALA\_NEXT\_TASK} API method, the programmer tells \sys's scheduler what task to execute next.
+The \sys kernel implementation is built on the assumption that the currently executing task is not interrupted, and only once it completes the next task is run.
+\todo{Provide information on the implementation - this paragraph repeats simply what previous section did}{Carlo}
 
-Task transitions.
-By calling \texttt{COALA\_NEXT\_TASK}, the user tells Coala’s scheduler what task to execute next.
-Note that the currently executing is not interrupted, and only once it completes the next task is run.
+\textbf{Protected Reads and Writes.} When \texttt{RP} and \texttt{WP} API mentods are used, \sys searches the volatile memory for the page the protected variable belongs to.
+If such page is not found in RAM, then a page fault occurs, and the page is fetched from the non-volatile memory.
+The two functions return a handle to the variable's volatile memory location.
+Unlike \texttt{RP}, \texttt{WP} marks the page being written as dirty, so that it can be committed to its non-volatile location.
 
-Protected reads and writes.
-When RP and WP are used, Coala searches the volatile memory for the page the protected variable belongs to.
-If such page is not found in RAM, then a page fault occurs, and the page is fetched from the non-volatile memory.
-The two functions return a handle to the variable’s volatile memory location.
-Unlike RP, WP marks the page being written as dirty, so that it can be committed to its non-volatile location.
-
-Initialization and origin task.
-The behavior of the method \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the addition that all the preliminary kernel initializations are performed inside this function.
+\textbf{Initialization and Origin Task.} The behavior of the API method \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the addition that all the preliminary kernel initializations are performed inside this function.
 Moreover, the origin task is scheduled only upon the first boot.
 
-Execution.
-The application control is passed to Coala’s scheduler by calling \texttt{COALA\_RUN} at the end of the main function.
-A high-level state machine of the scheduler is depicted in Figure \textcolor{red}{[draw and ref figure]}.
-Depending on the coalescing algorithm, state transitions follow a different pattern.
+\begin{wrapfigure}{t!}{0.5\textwidth}
+	\centering
+	%\includegraphics[width=0.5\columnwidth]{figures/graffle/overview.pdf}
+	\caption{\sys state machine implementation \todo{Draw a figure}{Sinan, Carlo}}
+	\label{fig:coala_state_machine}
+\end{wrapfigure}
 
-\subsection{\sys Paging}
+\textbf{Execution.} The application control is passed to \sys's scheduler by calling \texttt{COALA\_RUN} at the end of the main function.
+A high-level state machine of the scheduler is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure; provide more text on the execution}{Carlo}.
+
+\subsection{\sys Paging Implementation}
+
+\textcolor{red}{THIS IS UNEDITED AND SHOULD NOT BE READ YET - Placeholder}
 
@@
 \begin{figure}
+	\centering
 	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
-	\caption{Illustration of page tags \textcolor{red}{to be drawn}}
+	\caption{Illustration of \sys page tags.}
 \end{figure}

### 2018-04-15 paper/implementation.tex
Section 6.3 revised

@@
 
-\textbf{New Protected Variables.} The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by \texttt{RP} and \texttt{WP} API methods to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in a non-volatile memory.
+\noindent \textbf{New Protected Variables.} The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by \texttt{RP} and \texttt{WP} API methods to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in a non-volatile memory.
 
-\textbf{Task Transitions.} By calling \texttt{COALA\_NEXT\_TASK} API method, the programmer tells \sys's scheduler what task to execute next.
+\noindent \textbf{Task Transitions.} By calling \texttt{COALA\_NEXT\_TASK} API method, the programmer tells \sys's scheduler what task to execute next.
 The \sys kernel implementation is built on the assumption that the currently executing task is not interrupted, and only once it completes the next task is run.
@@
 
-\textbf{Protected Reads and Writes.} When \texttt{RP} and \texttt{WP} API mentods are used, \sys searches the volatile memory for the page the protected variable belongs to.
+\noindent \textbf{Protected Reads and Writes.} When \texttt{RP} and \texttt{WP} API mentods are used, \sys searches the volatile memory for the page the protected variable belongs to.
 If such page is not found in RAM, then a page fault occurs, and the page is fetched from the non-volatile memory.
@@
 
-\textbf{Initialization and Origin Task.} The behavior of the API method \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the addition that all the preliminary kernel initializations are performed inside this function.
+\noindent \textbf{Initialization and Origin Task.} The behavior of the API method \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the addition that all the preliminary kernel initializations are performed inside this function.
 Moreover, the origin task is scheduled only upon the first boot.
@@
 
-\textbf{Execution.} The application control is passed to \sys's scheduler by calling \texttt{COALA\_RUN} at the end of the main function.
+\noindent \textbf{Execution.} The application control is passed to \sys's scheduler by calling \texttt{COALA\_RUN} at the end of the main function.
 A high-level state machine of the scheduler is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure; provide more text on the execution}{Carlo}.
@@
 
-\textcolor{red}{THIS IS UNEDITED AND SHOULD NOT BE READ YET - Placeholder}
+\noindent \textbf{Memory Virtualization.} The memory virtualization is implemented using three buffers: (i) working (volatile) buffer, (ii) store (non-volatile) buffer and (iii) shadow (non-volatile) buffers.
+\todo{More explanation required here}{Carlo}
 
-Memory virtualization. Working (volatile) buffer, store and shadow (non-volatile) buffers. No major changes since last version.
-
-Page allocation and page tags.
-When the application accesses a protected variable with RP or WP, Coala has to search for the variable in RAM first, and in non-volatile memory if it not found in RAM.
-Actually, the kernel searches for pages, not single variables.
+\noindent \textbf{Page Allocation and Page Tags.} When the \sys application accesses a protected variable with either \texttt{RP} or \texttt{WP}, \sys has to search for the variable in RAM first, and then in a non-volatile memory if it not found in RAM.
+In reality, the \sys kernel searches for pages, not single variables.
 A page search can be very fast if each page is identified by a unique tag, and if every variable belonging to a page bears the tag in its address.
-Thus, retrieving a page tag from a protected variable’s address boils down to extracting the upper bits of the address itself.
-The remaining lower bits denote the variable’s offset in its page.
+Thus, retrieving a page tag from a protected variable's address boils down to extracting the upper bits of the address itself.
+The remaining lower bits denote the variable's offset in its page.
+The complete process is depicted in Figure~\ref{figure:coala_page_tags}.
 
@@
 	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
-	\caption{Illustration of \sys page tags.}
+	\caption{Illustration of \sys page tags mechanism.}
+	\label{figure:coala_page_tags}
 \end{figure}
 
-In order to obtain unique page tags, some alignment requirements have to be met.
-First, the page size $S$ has to be a power of two.
+We note that in order to obtain unique page tags, some alignment requirements have to be met.
+First, the page size $S$ has to be the power of two.
 Moreover, pages in non-volatile memory have to be aligned by $S$.
-In the example above, $S$ is 256, and every page’s first address is a multiple of 256.
-Note that, by aligning pages by $2n$, the alignment is also valid for all sizes $2^x$ with $x \leq n$.
+Note that, by aligning pages by $2^n$, the alignment is also valid for all sizes $2^x$, where $x \leq n$.
 
-Page fault and page swap (with DMA). When a page is not found in RAM, it is fetched from FRAM. No major changes since last version.
+\noindent \textbf{DMA Page Fault and Page Swap.} When a page is not found in RAM, it is fetched from FRAM. \todo{More explanation required here}{Carlo}
 
 \subsection{\sys Scheduling and Committing}
+
+\textcolor{red}{THIS IS UNEDITED AND SHOULD NOT BE READ YET - Placeholder}
 

### 2018-04-15 paper/evaluation.tex

@@
+\textcolor{red}{ highlight the difference between Alpaca GCC and Alpaca Clang.
+Alpaca GCC is about 30\%-40\% faster than Alpaca Clang.
+Writing Alpaca program by hand is a very trick thing and it is very error prone.
+Reason: a programmer needs to identify the write-after-read dependencies, change variables' names, and modify instructions accordingly, in general writing Alpaca code by hand is not a recommended, for that a compiler is needed}
+
 Our evaluation illustrates key findings about \sys. We show that: \todo{Make a list of core findings here once results are solid}{Przemek}
@@
 
-We first need to assess the performance of all coalescing strategies intriduced by us in Section~\ref{}.
+We first need to assess the performance of all coalescing strategies introduced by us in Section~\ref{}.
 

### 2018-04-15 paper/implementation.tex
Section 6.4 revised

@@
 	\centering
-	%\includegraphics[width=0.5\columnwidth]{figures/graffle/overview.pdf}
+	%\includegraphics[width=0.5\columnwidth]{figures/graffle/state-machine.pdf}
 	\caption{\sys state machine implementation \todo{Draw a figure}{Sinan, Carlo}}
@@
 
-\textcolor{red}{THIS IS UNEDITED AND SHOULD NOT BE READ YET - Placeholder}
+\noindent \textbf{Target Budget Initialization.} Upon a reboot, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy. \todo{expand text}{Carlo}
 
-Target budget initialization. Upon a reboot, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy [...].
+\noindent \textbf{Coalescing State Recall.} The last (coalesced) task has to be recalled.
+More specifically, the function pointer of the last coalesced task has to be retrieved from a non-volatile memory, where is had been saved during the last successful commit.
+\todo{expand text}{Carlo}
 
-Coalescing state recall.
-The last (coalesced) task has to be recalled.
-More specifically, the function pointer of the last coalesced task has to be retrieved from NV memory, where is had been saved during the last successful commit.
-[...].
+\noindent \textbf{Partial Execution Restoration.} In case there is a valid partial execution checkpoint, the program has to resume from there \todo{expand text}{Carlo}
 
-Partial execution restoration. In case there is a valid partial execution checkpoint, the program has to resume from there [...].
+\noindent \textbf{Coalesced Scheduling.} \todo{expand text}{Carlo}
 
-Coalesced scheduling. Static tasks are coalesced and run [...].
-
-Intermediate commit and ultimate commit.
-When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to non-volatile memory.
+\noindent \textbf{Intermediate Commit and Ultimate Commit.} When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to a non-volatile memory.
 Since last version, no major changes were made to the intermediate commit (SRAM to shadow buffer), but an optimization was applied to the ultimate commit (shadow buffer to store buffer).
@@
 
-\begin{figure}
-	\caption{Intermediate commit and ultimate commit Illustration \textcolor{red}{to be drawn}}
-\end{figure}
+\begin{wrapfigure}{t!}{0.5\textwidth}
+	\centering
+	%\includegraphics[width=0.5\columnwidth]{figures/graffle/commits-illustration.pdf}
+	\caption{Intermediate commit and ultimate commit Illustration \todo{to be drawn}{sinan}}
+	\label{fig:intermediate_ultimate-commit}
+\end{wrapfigure}
 
-Target budget update. Before starting a new coalesced task, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy [...].
+\noindent \textbf{Target Budget Update.} Before starting a new coalesced task, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy.
+\todo{expand text}{Carlo}
 
-\textcolor{red}{API of the partial commit needs to be explained as well}
+\todo{API of the partial commit needs to be explained as well}{Carlo}

### 2018-04-15 paper/implementation.tex
Section 6.2 revised and enriched

@@
 
-Coala's API adds only five syntactic constructs to a C-based language: (i) \texttt{COALA\_TASK}: declaration of a new task, (ii) \texttt{COALA\_NEXT\_TASK}: declaration of the next task to be scheduled after the currently executing one completes, (iii) \texttt{COALA\_PV}: declaration of a new protected variable, (iv) \texttt{RP}: declaration of a protected variable read, and (v) \texttt{WP}: declaration of a protected variable write.
-Moreover, the programmer is required to explicitly call two \sys API functions inside the main program body: (i) \texttt{COALA\_INIT}: to pass as an argument the global initialization task to run upon the first boot , and (ii) \texttt{COALA\_RUN}: which hands the control to \sys's scheduler.
-Finally, when the programmer intends to \texttt{typedef} a \texttt{struct} to use it as a protected variable, she/he need to declare each structure's member by using the \texttt{COALA\_SM} macro to ensure proper memory alignment required by \sys's page handler.
+Coala's API adds only five syntactic constructs to a C-based language: (i) \texttt{COALA\_TASK}, to declare a new task, (ii) \texttt{COALA\_NEXT\_TASK}, to mark the next task to be run after the currently executing one completes, (iii) \texttt{COALA\_PV}, to declare a new protected variable, (iv) \texttt{RP} and \texttt{WP}, to mark protected variables' reads and writes, and (v) \texttt{COALA\_SM}, to declare \texttt{typedef}'d \texttt{struct} members, when used as protected variables, to ensure proper memory alignment required by \sys's page handler.
+Moreover, the programmer is required to explicitly call two API functions inside the main program body: (i) \texttt{COALA\_INIT}, to pass as an argument the global initialization task to run upon the first boot, and (ii) \texttt{COALA\_RUN}, which hands the control to \sys's task manager.
 
@@
 	\hline\hline
-	\texttt{COALA\_INIT}(t) & $t \in \mathbf{T}$: task to be scheduled on a first boot \\
+	\texttt{COALA\_INIT}($t$) & $t \in \mathbf{T}$: task to be scheduled on the first boot \\
 	\hline
@@
 	\hline
-	\texttt{COALA\_NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : next task to be scheduled\\
+	\texttt{COALA\_NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
 	\hline
-	\texttt{COALA\_PV}($p$, $v$ [, $s$]) & $p$: variable type, $v \in \mathbf{V}$: new protected variable name, $s$: array size\\
+	\texttt{COALA\_PV}($p$, $v$ [, $s$]) & $p$: variable type, $v \in \mathbf{V}$: new protected variable name, $s$: array size \\
 	\hline
-	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to be read, $u$: value \\
+	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to be read, $u$: destination operand \\
 	\hline	
-	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to be written, $u$: value \\
+	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to be written, $u$: source operand \\
 	\hline
@@
 \end{tabular}
-\caption{Summary of \sys API; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
+\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
 \label{table:coala_api}
@@
 
-\noindent \textbf{New Protected Variables.} The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by \texttt{RP} and \texttt{WP} API methods to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in a non-volatile memory.
+\noindent \textbf{New Protected Variables.} The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by \texttt{RP} and \texttt{WP} API methods to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in non-volatile memory.
 
-\noindent \textbf{Task Transitions.} By calling \texttt{COALA\_NEXT\_TASK} API method, the programmer tells \sys's scheduler what task to execute next.
-The \sys kernel implementation is built on the assumption that the currently executing task is not interrupted, and only once it completes the next task is run.
-\todo{Provide information on the implementation - this paragraph repeats simply what previous section did}{Carlo}
+\noindent \textbf{Task Transitions.} By calling \texttt{COALA\_NEXT\_TASK}, the programmer tells \sys's task manager which task to execute next.
+The implementation of \sys is built on the assumption that the currently executing task is not interrupted, and only once it completes the next task is run.
 
-\noindent \textbf{Protected Reads and Writes.} When \texttt{RP} and \texttt{WP} API mentods are used, \sys searches the volatile memory for the page the protected variable belongs to.
+\noindent \textbf{Protected Reads and Writes.} When \texttt{RP} and \texttt{WP} API methods are used, \sys searches the volatile memory for the page the protected variable belongs to.
 If such page is not found in RAM, then a page fault occurs, and the page is fetched from the non-volatile memory.
 The two functions return a handle to the variable's volatile memory location.
-Unlike \texttt{RP}, \texttt{WP} marks the page being written as dirty, so that it can be committed to its non-volatile location.
+Unlike \texttt{RP}, \texttt{WP} marks the page being written as dirty, so that it can be copied to its non-volatile location during the commit phase.
 
@@
 
-\noindent \textbf{Execution.} The application control is passed to \sys's scheduler by calling \texttt{COALA\_RUN} at the end of the main function.
-A high-level state machine of the scheduler is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure; provide more text on the execution}{Carlo}.
+\noindent \textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
+A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}.
+Upon reboot, the coalescing target is initialized according to the coalescing strategy.
+Then, the second phase of the commit is resumed, in case the last power failure occurred during this operation.
+The shadow buffer list is cleared, and the metadata of the volatile buffer is initialized.
+The pointer to the last coalescing task is recalled from non-volatile memory.
+Before running, Coala checks whether there is any partial execution to resume, in which case the volatile state, including the PC, is restored.
+Otherwise, coalescing starts: the current budget is replenished; tasks are run, the history is increased and the current budget is decreased; when the latter is zero, coalescing stops, the pointer to the next task is saved and the commit phase starts.
+Coalescing resumes after commit phase two ends and the target is updated based on the coalescing strategy.
+\todo{Go over keywords (coalescing target, current budget, commit phases etc.) and make them consistent with the rest of the paper}{Amjad, Carlo}
 

### 2018-04-15 paper/implementation.tex
little corrections in the implementation
Section 6 enriched

@@
-We proceed with the explanation of the implementation of \sys programming model. \sys implementation is light and adds very little on top of a plain C language.
+We proceed with the explanation of the implementation of \sys programming model, which is light and whose API adds very little on top of a plain C language.
 
-\subsection{\sys Application Programming Interface}
+\subsection{Application Programming Interface}
 \label{sec:coala_api}
@@
 
-\subsection{\sys Kernel Functions Implementation}
+\subsection{Kernel Functions Implementation}
 
@@
 
-\noindent \textbf{New Protected Variables.} The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by \texttt{RP} and \texttt{WP} API methods to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in non-volatile memory.
+\textbf{New Protected Variables.} The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by \texttt{RP} and \texttt{WP} API methods to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in non-volatile memory.
 
-\noindent \textbf{Task Transitions.} By calling \texttt{COALA\_NEXT\_TASK}, the programmer tells \sys's task manager which task to execute next.
+\textbf{Task Transitions.} By calling \texttt{COALA\_NEXT\_TASK}, the programmer tells \sys's task manager which task to execute next.
 The implementation of \sys is built on the assumption that the currently executing task is not interrupted, and only once it completes the next task is run.
 
-\noindent \textbf{Protected Reads and Writes.} When \texttt{RP} and \texttt{WP} API methods are used, \sys searches the volatile memory for the page the protected variable belongs to.
+\textbf{Protected Reads and Writes.} When \texttt{RP} and \texttt{WP} API methods are used, \sys searches the volatile memory for the page the protected variable belongs to.
 If such page is not found in RAM, then a page fault occurs, and the page is fetched from the non-volatile memory.
@@
 
-\noindent \textbf{Initialization and Origin Task.} The behavior of the API method \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the addition that all the preliminary kernel initializations are performed inside this function.
+\textbf{Initialization and Origin Task.} The behavior of the API method \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the addition that all the preliminary kernel initializations are performed inside this function.
 Moreover, the origin task is scheduled only upon the first boot.
@@
 
-\noindent \textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
+\textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
 A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}.
 Upon reboot, the coalescing target is initialized according to the coalescing strategy.
-Then, the second phase of the commit is resumed, in case the last power failure occurred during this operation.
+Then, the ultimate commit is resumed, in case the last power failure occurred during this operation.
 The shadow buffer list is cleared, and the metadata of the volatile buffer is initialized.
@@
 Before running, Coala checks whether there is any partial execution to resume, in which case the volatile state, including the PC, is restored.
-Otherwise, coalescing starts: the current budget is replenished; tasks are run, the history is increased and the current budget is decreased; when the latter is zero, coalescing stops, the pointer to the next task is saved and the commit phase starts.
-Coalescing resumes after commit phase two ends and the target is updated based on the coalescing strategy.
+Otherwise, coalescing starts: the current budget is replenished; tasks are run, the history is increased and the current budget is decreased; when the latter is zero, coalescing stops, the pointer to the next task is saved and the (intermediate) commit starts.
+Coalescing resumes after the ultimate commit is complete and the target is updated based on the coalescing strategy.
 \todo{Go over keywords (coalescing target, current budget, commit phases etc.) and make them consistent with the rest of the paper}{Amjad, Carlo}
 
-\subsection{\sys Paging Implementation}
+\subsection{Paging Implementation}
 
-\noindent \textbf{Memory Virtualization.} The memory virtualization is implemented using three buffers: (i) working (volatile) buffer, (ii) store (non-volatile) buffer and (iii) shadow (non-volatile) buffers.
-\todo{More explanation required here}{Carlo}
+\sys's memory virtualization is implemented using three buffers: (i) working (volatile) buffer, (ii) store (non-volatile) buffer and (iii) shadow (non-volatile) buffer.
+The store buffer holds the latest safe copy of all the pages, and it is read-only up to, and including, the ultimate commit.
+Upon reboot, the working buffer is empty, and it gets filled with pages from non-volatile memory on demand.
+All reads and writes to protected variables happen in volatile memory.
+The shadow buffer serves as an intermediate non-volatile buffer where dirty, volatile pages are copied during the intermediate commit.
+Moreover, it contains pages evicted from SRAM when a \emph{full page fault} occurs.
 
-\noindent \textbf{Page Allocation and Page Tags.} When the \sys application accesses a protected variable with either \texttt{RP} or \texttt{WP}, \sys has to search for the variable in RAM first, and then in a non-volatile memory if it not found in RAM.
-In reality, the \sys kernel searches for pages, not single variables.
+\textbf{Page Allocation and Page Tags.} When a \sys application accesses a protected variable with either \texttt{RP} or \texttt{WP}, \sys has to search for the variable in RAM first, and then, if not found, in non-volatile memory.
+In reality, the page manager searches for pages, not single variables.
 A page search can be very fast if each page is identified by a unique tag, and if every variable belonging to a page bears the tag in its address.
@@
 The remaining lower bits denote the variable's offset in its page.
-The complete process is depicted in Figure~\ref{figure:coala_page_tags}.
+The complete process is depicted in Figure~\ref{figure:coala_page_tags} \todo{this figure doesn't really show page tags and offsets, but paging in general}.
+We note that, in order to obtain unique page tags, some alignment requirements have to be met.
+First, the page size $S$ has to be the power of two.
+Moreover, pages in non-volatile memory have to be aligned by $S$.
+Note that, by aligning pages by $2^n$, the alignment is also valid for all sizes $2^x$, where $x \leq n$.
 
@@
 
-We note that in order to obtain unique page tags, some alignment requirements have to be met.
-First, the page size $S$ has to be the power of two.
-Moreover, pages in non-volatile memory have to be aligned by $S$.
-Note that, by aligning pages by $2^n$, the alignment is also valid for all sizes $2^x$, where $x \leq n$.
+\textbf{Page Fault and Page Swap.} When a page is not found in RAM, a \emph{page fault} it has to be fetched from FRAM.
+In case that page was in the working buffer and it had been evicted, its most recent non-volatile copy resides in the shadow buffer.
+For this reason, \sys searches this buffer first.
+If the requested page is not found in the shadow buffer, then it will be pulled from its store location.
+To make room for the requested page in SRAM, a volatile page might have to be evicted.
+If the victim page is dirty, we say that a \emph{full page fault} has occurred, and the victim page has to be copied to the shadow buffer.
+Page moves are accelerated by the MCU's direct memory access (DMA) module.
 
-\noindent \textbf{DMA Page Fault and Page Swap.} When a page is not found in RAM, it is fetched from FRAM. \todo{More explanation required here}{Carlo}
+% It is important to mention that store and shadow buffers are virtual.
+By default, \sys allocates two memory banks for pages, $B0$ and $B1$, and all the pages have their store copy in $B0$.
+A dirty page that during the intermediate commit, or due to eviction, is copied to its shadow location in $B1$, will end up having its store location in $B1$ after the ultimate commit.
+Thus, during the ultimate commit
 
-\subsection{\sys Scheduling and Committing}
+\subsection{Coalescing Loop and Commit Phase}
 
-\noindent \textbf{Target Budget Initialization.} Upon a reboot, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy. \todo{expand text}{Carlo}
+As described above, the main coalescing loop alternates between executing tasks and saving progresses.
+Static tasks are coalesced as long as the current budget is not null, after which the two commit phases are triggered, the coalescing target is updated and the current budget is replenished.
 
-\noindent \textbf{Coalescing State Recall.} The last (coalesced) task has to be recalled.
-More specifically, the function pointer of the last coalesced task has to be retrieved from a non-volatile memory, where is had been saved during the last successful commit.
-\todo{expand text}{Carlo}
+\textbf{Intermediate Commit and Ultimate Commit.} The reason why the commit process is split in two phases is two-fold: ensure consistency by using atomic operations, and preserving performance by enabling an incremental commit mechanism.
+When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to the non-volatile shadow buffer.
+This is what we call \emph{intermediate commit}: it has to be completed in one shot, and if power fails during this phase the application will have to resume from the beginning of the last coalesced task.
+When the intermediate commit is done, a flag can be atomically raised to signify that a valid, persistent copy of dirty pages is available in non-volatile memory.
+From this point on, the commit process can be resumed even if power fails.
+To understand the last phase, referred to as \emph{ultimate commit}, it has to be mentioned that store and shadow buffers are virtual.
+By default, \sys allocates two memory banks for non-volatile pages, $M_0$ and $M_1$, and all the pages have their initial store copy in $M_0$.
+A dirty page that during the intermediate commit, or due to eviction, is copied to its shadow location in $M_x$, will end up having its store location in $M_x$ after the ultimate commit.
+In other words, during the ultimate commit dirty pages are not physically moved, instead their store and shadow locations are swapped.
+This trick makes \sys save some energy during this phase, avoiding redundant page moves.
+As for page swaps, the intermediate commit makes use of the on-board DMA.
 
-\noindent \textbf{Partial Execution Restoration.} In case there is a valid partial execution checkpoint, the program has to resume from there \todo{expand text}{Carlo}
-
-\noindent \textbf{Coalesced Scheduling.} \todo{expand text}{Carlo}
-
-\noindent \textbf{Intermediate Commit and Ultimate Commit.} When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to a non-volatile memory.
-Since last version, no major changes were made to the intermediate commit (SRAM to shadow buffer), but an optimization was applied to the ultimate commit (shadow buffer to store buffer).
-Store buffer and shadow buffer used to be physically separated in memory.
-Now, the role of a block of memory (i.e.
-a page) is interchangeable, meaning that, during the ultimate commit, the shadow buffer pages that were written during the previous commit phase become store buffer pages, and their relative store buffer pages become shadow buffer pages.
-A page’s role is retrieved using an indirection table, i.e.
-an array that holds a zero for all the pages whose store buffer resides in the lower memory block, and a one for all the pages whose store buffer is located in the upper memory block.
-This allows for a lighter commit process, since fewer page moves are performed.
+\subsection{Partial execution}
 
@@
 \end{wrapfigure}
-
-\noindent \textbf{Target Budget Update.} Before starting a new coalesced task, the target budget (or target coalesced task size) has to be updated depending on the coalescing strategy.
-\todo{expand text}{Carlo}
-
-\todo{API of the partial commit needs to be explained as well}{Carlo}

### 2018-04-15 paper/hwswsetup.tex

--- 2018-04-14 paper/hwswsetup.tex
@@
 We evaluated \sys using six benchmarks summarized in Table~\ref{table:benchmark_table}, with their implementation is available via~\cite{coala_website}.
-Five originate from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}: three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) were implemented by Chain~\cite{chain} and Alpaca~\cite{alpaca}, while two (\textbf{dijkstra}, \textbf{fft}) are added as new.
+Five benchmarks originate from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}: three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) were already implemented by Chain~\cite{chain} and Alpaca~\cite{alpaca}, while two (\textbf{dijkstra}, \textbf{fft}\footnote{Specifically, we re-implemented FFT based on the TI's accelerated FFT library~\cite{}.\todo{add reference}{Przemek}}) are added as new.
 The final one (\textbf{sort}) has been written from scratch.

### 2018-04-15 paper/implementation.tex

@@
 
-\textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
+\noindent \textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
 A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}.
 Upon reboot, the coalescing target is initialized according to the coalescing strategy.
-Then, the ultimate commit is resumed, in case the last power failure occurred during this operation.
+Then, the second phase of the commit is resumed, in case the last power failure occurred during this operation.
 The shadow buffer list is cleared, and the metadata of the volatile buffer is initialized.
 The pointer to the last coalescing task is recalled from non-volatile memory.
-Before running, Coala checks whether there is any partial execution to resume, in which case the volatile state, including the PC, is restored.
-Otherwise, coalescing starts: the current budget is replenished; tasks are run, the history is increased and the current budget is decreased; when the latter is zero, coalescing stops, the pointer to the next task is saved and the (intermediate) commit starts.
-Coalescing resumes after the ultimate commit is complete and the target is updated based on the coalescing strategy.
+Before running, \sys checks whether there is any partial execution to resume, in which case the volatile state, including the program counter, is restored.
+Otherwise, coalescing starts: the current budget is replenished; tasks are run, the history is increased and the current budget is decreased; when the latter is zero, coalescing stops, the pointer to the next task is saved and the commit phase starts.
+Coalescing resumes after commit phase two ends and the target is updated based on the coalescing strategy.
 \todo{Go over keywords (coalescing target, current budget, commit phases etc.) and make them consistent with the rest of the paper}{Amjad, Carlo}

### 2018-04-15 paper/evaluation.tex
Table 4 revised and ready to be filled in with data

@@
-\textcolor{red}{ highlight the difference between Alpaca GCC and Alpaca Clang.
+\textcolor{red}{highlight the difference between Alpaca GCC and Alpaca Clang.
 Alpaca GCC is about 30\%-40\% faster than Alpaca Clang.
@@
 
+%\sys is efficient and easy to use across our benchmarks.
+
 \subsection{\sys Runtime Performance}
@@
 
-We evaluated \sys to show that it is easy to use and efficient across our benchmarks.
-We evaluate \sys efficiency by comparing its runtime performance to Alpaca.
-We characterize \sys by studying how its performance varies due to coalescing.
+Our aim is to assess \sys as detailed as possible.
+For this we have characterized \sys runtime performance (execution time) considering task adaptation: task coalescing and downscaling.
+We also characterize \sys overhead showing precisely what is the source of \sys superior performance.
+Additionally, we characterize \sys code.
 
@@
 
-We first need to assess the performance of all coalescing strategies introduced by us in Section~\ref{}.
+We characterize \sys by studying how its performance varies due to task adaptation. We start with evaluating the task coalescing mechanism.
 
-\textbf{Coalescing Strategies.} Figure~\ref{fig:coalescing} shows the run time of \sys's two coalescing strategies normalized to the run time without
-coalescing. We show data from experiments with intermittent power, where intermittency was generated in software
+\textbf{Coalescing Strategies.} We first need to assess the performance of all coalescing strategies introduced by us in Section~\ref{sec:task_coalescing}.
+Figure~\ref{fig:coalescing} shows the run time of \sys's two coalescing strategies normalized to the run time without coalescing.
+We show data from experiments with intermittent power, where intermittency was generated in software
 %\todo{Say briefly how we generate full reset with MSP430}{Amjad}
@@
 		\hline
-		Application & \multicolumn{2}{| c |}{Memory footprint} & No. tasks & SLOC \\
+		\multirow{2}{*}{Application} & \multicolumn{2}{ c |}{Memory footprint} & \multirow{2}{*}{No. tasks} & \multirow{2}{*}{SLOC} \\
+		\cline{2-3}
 		{} & \sys & Alpaca & {} & {} \\
@@
 				\textbf{ar} & --- & --- & --- & ---\\
+		\hline
+				\textbf{bc} & --- & --- & --- & ---\\
+		\hline
+				\textbf{cuckoo} & --- & --- & --- & ---\\
+		\hline
+				\textbf{dijkstra} & --- & --- & --- & ---\\
+		\hline
+				\textbf{fft} & --- & --- & --- & ---\\
+		\hline
+				\textbf{sort} & --- & --- & --- & ---\\
 		\hline

### 2018-04-15 paper/implementation.tex
Section 6.5 (partial execution) added

@@
 We note that, in order to obtain unique page tags, some alignment requirements have to be met.
-First, the page size $S$ has to be the power of two.
-Moreover, pages in non-volatile memory have to be aligned by $S$.
+First, the page size $S$ has to be a power of two.
+Moreover, pages have to be aligned by $S$.
 Note that, by aligning pages by $2^n$, the alignment is also valid for all sizes $2^x$, where $x \leq n$.
@@
 
-\textbf{Page Fault and Page Swap.} When a page is not found in RAM, a \emph{page fault} it has to be fetched from FRAM.
+\textbf{Page Fault and Page Swap.} When a page is not found in RAM, a \emph{page fault} occurs, and it has to be fetched from FRAM.
 In case that page was in the working buffer and it had been evicted, its most recent non-volatile copy resides in the shadow buffer.
@@
 To make room for the requested page in SRAM, a volatile page might have to be evicted.
-If the victim page is dirty, we say that a \emph{full page fault} has occurred, and the victim page has to be copied to the shadow buffer.
+If the victim page is dirty, we say that a \emph{full page fault} occurs, and the victim page has to be copied to the shadow buffer.
 Page moves are accelerated by the MCU's direct memory access (DMA) module.
@@
 
-As described above, the main coalescing loop alternates between executing tasks and saving progresses.
+As described above, the main coalescing loop alternates between executing tasks and saving progress.
 Static tasks are coalesced as long as the current budget is not null, after which the two commit phases are triggered, the coalescing target is updated and the current budget is replenished.
 
-\textbf{Intermediate Commit and Ultimate Commit.} The reason why the commit process is split in two phases is two-fold: ensure consistency by using atomic operations, and preserving performance by enabling an incremental commit mechanism.
+\textbf{Intermediate Commit and Ultimate Commit.} The reason why the commit process is split in two phases is two-fold: ensure consistency by using atomic operations, and preserve performance by enabling an incremental commit mechanism.
 When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to the non-volatile shadow buffer.
@@
 
+Among \sys's novelties, the task-splitting feature is the most effective when it comes to preservation of forward progress.
+Its implementation adds hardly any computational overhead, and is supported by a hardware timer running "in background", whose energy consumption is extremely low (around $3 \mu A/MHz$~\cite{msp430datasheet}).
+The goal is to interrupt a static task if it cannot be run to completion with a full capacitor's energy, commit its partial results, save volatile state and program counter and continue.
+Coala speculates and assumes that a static task is too large and has to be split when power fails two times consecutively in the same static task, and the coalescing target is one, because, after the first power failure, the large task is run with a fresh energy budget.
+To "split" a static task, Coala schedules a timer interrupt.
+If this interrupt occurs before the capacitor is depleted, Coala commits all the dirty volatile pages (using the two-phase commit explained above), saves stack and volatile registers to non-volatile memory and then continues with the task.
+In case a power failure occurs after this operation, but before the static task completes, the execution at the next reboot starts exactly where Coala had saved the partial execution.
+A task can be split in multiple sub-tasks, as many as needed to progress forward.
+Regarding the timer interrupt, it is initially scheduled to fire after a large time interval.
+If power fails before the interrupt fires, Coala reduces the interval.
+If the partial execution is committed successfully, Coala enlarges the interval.
+These adjustments follow a binary search technique, and make the interval converge very fast to a value optimal for preserving forward progress and avoiding redundant commits.
+
+For some applications, specific pieces of a task cannot be split and have to be run atomically.
+For instance, when sampling an analog signal, continuity between samples has to be ensured, otherwise the digital array would represent a wrong, non-continuous input signal.
+In such cases, the programmer can use two dedicated API guards before and after the atomic piece of code, namely \texttt{COALA\_DISABLE\_PE} and \texttt{COALA\_ENABLE\_PE}, to respectively halt and resume the partial execution timer.
+
 \begin{wrapfigure}{t!}{0.5\textwidth}

### 2018-04-15 paper/evaluation.tex
added intro paragraph about manual code transform by alpaca

@@
-\textcolor{red}{highlight the difference between Alpaca GCC and Alpaca Clang.
+%\textcolor{red}{highlight the difference between Alpaca GCC and Alpaca Clang.
 Alpaca GCC is about 30\%-40\% faster than Alpaca Clang.
@@
 Our aim is to assess \sys as detailed as possible.
-For this we have characterized \sys runtime performance (execution time) considering task adaptation: task coalescing and downscaling.
+For this we have characterized \sys runtime performance (execution time), based on a set of benchmarks, considering task adaptation: task coalescing and downscaling.
 We also characterize \sys overhead showing precisely what is the source of \sys superior performance.
 Additionally, we characterize \sys code.
+
+\sys is compared to Alpaca runtime.
+We remind the reader that the benchmarks for Alpaca were hand-transformed in a time-consuming manual process from Clang (LLVM-generated code) to GCC.
+The manual code transformation was required as \sys was written with GCC in mind, and Alpaca cannot support any other complier than LLVM.
 

### 2018-04-15 paper/intro.tex
by Sinan - removed
intro some modifications

@@
 power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
-down and recharges to operate again later. As an example, recharge times may be
+down and recharges to operate again later---corresponds to {\em intermittent execution model}~\citep{dino,lucia_snapl_2017} composed of
+operation-power failure-restart cycles. As an example, recharge times may be
 tens of seconds in radio frequency powered medical device~\cite[Fig.
@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/graffle/intro-figure.pdf} By Sinan
-    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits and buffering updates to non-volatile (NV) variables in volatile (V) memory.}
+    \includegraphics[width=0.5\textwidth]{figures/graffle/intro-figure.pdf}
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits.
+On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
     \label{fig:coalesce}
@@
 
-\textbf{Data Consistency in Intermittent Computing.}  Software in an
-energy-harvesting system operates according to the {\em intermittent execution
-model}~\citep{dino,lucia_snapl_2017}, which corresponds to the
-operation-failure-restart cycles of an energy-harvesting device. In an
-intermittent software execution, an {\em operating period} proceeds for an
-arbitrary duration (dictated by energy availability) before being interrupted
-by a {\em failure period}. In a failure period, a device loses the volatile
+\textbf{Data Consistency in Intermittent Computing.}  
+%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\citep{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device.
+In an intermittent software execution, an {\em operating period} proceeds for anarbitrary duration (dictated by energy availability) before being interruptedby a {\em failure period}.
+In a failure period,
+Upon power failures, a device loses the volatile
 state in its registers, stack, SRAM, and retains the state of any non-volatile
@@
 continuously-powered behavior, often leading to an unrecoverable
-application failure~\citep{dino,edb}.
-
-Prior work developed two main approaches to dealing with data inconsistency for
+application failure~\citep{dino,edb}. Prior work developed two main approaches to dealing with data inconsistency for
 intermittently-powered devices: (i) \emph{software-based programming and
@@
 statically decompose a program into a collection of {\em tasks}.  A task can
-include arbitrary computation and these existing systems guarantee that each
-tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
+include arbitrary computation that should be executed despite arbitrarily-timed power failures.
+%and these existing systems guarantee that each
+%tasks executes {\em atomically}, despite arbitrarily-timed power failures.  
 %
@@
 At each task transition, the system incurs an overhead to track and atomically
-commit modifications to the non-volatile memory, to maintain consistency of
+\emph{commit} modifications to the non-volatile memory, to maintain consistency of
 program state~\citep{chain,alpaca}.  
 %
-The more task transitions a program executes at run time, the more overhead.
+The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.
 
@@
 complete, but incur a high task transition and commit overhead.  In this work,
-we propose a third way, using a novel technique called {\em dynamic task
-coalescing} that efficiently executes small tasks, avoiding unnecessary
-overheads while avoiding the risk of non-termination.
+we propose a third way, using a novel technique called {\em adaptive task-execution by dynamic task
+coalescing/splitting} that efficiently executes small tasks, avoiding unnecessary overheads while avoiding the risk of non-termination.
 

### 2018-04-15 paper/evaluation.tex
add figures to eval section

--- 2018-04-15 paper/evaluation.tex
@@
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/overallOverhead}
+	\caption{ overall overhead breakdown: nc: No Coalecing, sc: Slow Coalescing, ea: Energy Aware coalescing \todo{}{Przemek}}
+	\label{fig:overallOverheadBreakdown}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
+	\caption{ Coalescing algorithm efficiency:  nc: No Coalecing, sc: Slow Coalescing, ea: Energy Aware coalescing, eat: Energy \& Task Aware coalescing \todo{}{Przemek}}
+	\label{fig:coalEfficiency}
+\end{figure}
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/memAccess}
+	\caption{ Memory access overhead breakdown:  nc: No Coalecing, sc: Slow Coalescing, ea: Energy Aware coalescing \todo{}{Przemek}}
+	\label{fig:coalEfficiency}
+\end{figure}
+
 \textcolor{red}{highlight the difference between Alpaca GCC and Alpaca Clang.

### 2018-04-15 paper/intro.tex

@@
 complete, but incur a high task transition and commit overhead.  In this work,
-we propose a third way, using a novel technique called {\em adaptive task-execution by dynamic task
+we propose a third way, using a novel technique called {\em adaptive execution by dynamic task
 coalescing/splitting} that efficiently executes small tasks, avoiding unnecessary overheads while avoiding the risk of non-termination.

### 2018-04-15 paper/systemdescription.tex

@@
 Coalescing also amortizes per-variable commit cost, committing each location accessed by both tasks only after the second task, rather than once per task.
+\textcolor{red}{\sys's coalescing manager is an \emph{energy-aware manager}.
+It uses the recent execution history---pure software technique---as a metric to estimate the amount of the available energy and to set the adequate coalesced task size accordingly.
+This fundamental feature prevents \sys from blind coalesced task adaptation which can result in a repeated power failures problem.
+For example, If \sys enlarges the coalesced task size by $x$ static tasks on a coalesced task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, 10 times to reduce a coalesced task size from 10 to one, when $x=1$}
+
 Details of coalescing are presented in Section~\ref{sec:task_coalescing}.
 
-\textbf{\sys Task Downscaling Manager.} Details of task downscaling are presented in Section~\ref{sec:task_downsizing}. \textcolor{red}{to be written}
+\textbf{\sys Task Downscaling Manager.} \sys's downscaling manager is responsible for preserving the forward progress when \sys fails repeatedly on a single static task.
+The downscaling manager utilizes a timer-based technique to split a big task and ensure a partial progress.
+
+ Details of task downscaling are presented in Section~\ref{sec:task_downsizing}. \textcolor{red}{to be written}
+
+ \textcolor{red}{Why hardware based checkpoint systems are static? and why \sys's partial commit is dynamic?}
+
+\textcolor{red}{(i) Stopping at arbitrary locations in the code may result in incorrect computation or even crash the program.
+If a system checkpoints between, for example, the DMA initialization and the DMA call the program will crash.
+One might think that a simple solution is to disable interrupts at these critical regions.
+Although this is the correct solution, the checkpoint-based system will "theoretically" stuck (non-termination problem) because it has to checkpoint at this exact location since it is a threshold-based system, so it is a static system.
+However, \sys's partial commit is not threshold based therefore it allows the programmer to disable interrupts at certain locations and it will partially commit around it.}
+
+\textcolor{red}{(ii) A Checkpoint system normally checkpoints all volatile data.
+Therefore, it suffers from moving large amount of data on each checkpoint.
+However, \sys's partial commit technique relies on a paging mechanism, which makes the size of the commit proportional to the number of state-less instructions executed.
+In other words, if energy allows executing only a small chunk of code, the amount of data to be committed shrinks accordingly.
+As such, \sys's execution model is neither limited by a static task size to progress, nor by the commit size,
+}
 

### 2018-04-15 paper/abstract.tex
Abstract polished

@@
 %Intermittently powered computers promise a sustainable Internet of things by forgo the batteries and scarf the energy from the environment. 
-Massively deployed Internet of Things (IoT) must be energy neutral. This can only be achieved when the IoT is batteryless and the energy to operate the IoT comes from the ambient energy sources.
+Massively deployed Internet of Things (IoT) must be energy neutral. This can only be achieved when the IoT is batteryless and the energy to operate the IoT comes from ambient energy sources.
 % The main problem that we are tackling
@@
 % The problem of the proposed solutions 
-Generally, the intermittent computation models can be classified into (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based programming, where a programmer splits the code into small idempotent sections.
-Despite the superiority of the task-based systems over checkpointing, expressed in lower memory overhead, task-based programming is completely oblivious to the energy conditions.
+Generally, intermittent computation models can be classified into (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based programming, where a programmer splits the code into small idempotent sections.
+Despite the superiority of task-based systems over checkpointing, expressed in lower memory overhead, task-based programming is completely oblivious to the energy conditions.
 Simply, it cannot increase or decrease the size of the task on the fly (the moment of commit to the non-volatile memory) depending on how long the energy burst is.
 % Our main claim 
-To address this challenge, this paper introduces \sys: \emph{adaptive} task-based execution model that does not require any dedicated hardware to support task adaptability, which at the same time is \emph{efficient}.
+To address this challenge, this paper introduces \sys, an \emph{adaptive} task-based execution model that does not require any dedicated hardware to support task adaptability, which at the same time is \emph{efficient}.
 % show off
-\sys programs tune frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period from the programer's shoulders.
+\sys programs tune frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period off the programmer's shoulders.
 % main benefits of Coala
-\sys defaults to progress on a multi-tasks scale, when energy permits (\emph{efficiency}), and on a sub-task scale when needed {\emph{guarantee}}. 
+\sys defaults to progress on a multi-tasks scale, when energy permits (\emph{efficiency}), and on a sub-task scale when needed (\emph{guarantee}).
 % I really hope it will support our claims!! haha
-Results show the \sys is more efficient than the current state-of-art execution model by \textcolor{red}{X times}.
+Results show that \sys is more efficient than the current state-of-art execution model by \textcolor{red}{X times}.
 

### 2018-04-15 paper/hwswsetup.tex
revision of text about manual code transform by alpaca

@@
 The final one (\textbf{sort}) has been written from scratch.
-\todo{justify the selection of these benchmarks}{Przemek, Carlo, Amjad, Brandon}
+\todo{justify the selection of these benchmarks}{Przemek, Carlo, Amjad, Brandon} All applications were compiled using GCC version 3.8.0.
+\todo{Check if this still holds.
+provide optimization flag}{Przemek}
 
@@
 
-All applications were compiled using GCC version 3.8.0.
-\todo{Check if this still holds.
-provide optimization flag}{Przemek} We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
+\textbf{Comparison with Alpaca and Alpaca Compiler Limitations.} We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
 For correct comparison, all benchmarks of \sys were divided into tasks in the same way as Alpaca.
-Moreover, since original Alpaca runtime was written in Clang, we transformed every benchmark manually into GCC-complaint code.
-\todo{Provide more information about porting to GCC from LLVM}{Brandon, Carlo}
+Since original Alpaca runtime was written in Clang, we had to transform every benchmark manually into GCC-complaint code.
+\todo{Provide more information about porting to GCC from LLVM}{Brandon, Carlo} This was a time-consuming process since a programmer needs to identify WAR dependencies \todo{check if we use WAR or write-after-read or anything else}{Przemek}, then change variables' names and modify instructions accordingly.
+Nota bene, our experiments based on the benchmarks we used here show that Alpaca compiled with GCC is about 30\%-40\% faster than Alpaca compiled with Clang.
+\todo{Provide concrete numbers for all applications}{Amjad} The speed improvement we got for Alpaca with GCC is however sacrificed by the above-mentioned error-prone manual code transformation process for which no compiler support exists thus far.
+\sys \emph{does not require any of such transformations}.
 
-We do not compare \sys against Chain~\cite{chain}, another task-based system, which is inferior to Alpaca in its performance.
+\textbf{Comparison of \sys against other Systems.} We do not compare \sys against Chain~\cite{chain}, another task-based system, which is inferior to Alpaca in its performance.
 Also, we do not compare \sys to checkpointing-based systems, such as DINO~\cite{dino} or Mementos~\cite{mementos} (cf.

### 2018-04-15 paper/intro.tex

@@
 
-In this work, we introduce a new system called \sys, which executes a task-based
-intermittent program by dynamically coalescing its tasks.
+This paper introduces a new system called \sys, which executes a task-based
+intermittent program by dynamically coalescing/splitting its tasks.
 \sys accepts any static decomposition (i.e., from the programmer or from a compiler) 
-and coalesces its tasks at runtime. Two consecutive, coalesced tasks execute  
+and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute  
 with no commit or task transition overhead between them, instead performing
@@
 of both.  If power fails during a coalesced task, execution restarts from the
-{\em first} of the coalesced tasks.  Figure~\ref{fig:coalesce} (bottom)
+{\em first} of the coalesced tasks.  Figure~\ref{fig:coalesce} (top)
 illustrates how dynamic coalescing changes the earlier execution and highlights
 the change in memory model necessary to not break atomicity.
+On the other hand, a large task that requires more energy to complete than a device's fixed energy buffer is split by \sys to enable task termination, as depicted in Figure~\ref{fig:coalesce} (bottom).
 
@@
 %results.
-
-While a compiler can use static data privatization and commit instrumentation
+\textbf{Challenges.} While a compiler can use static data privatization and commit instrumentation
 (i.e., redo-logging) to eliminate statically identifiable, inter-task data
@@
 %consistency.
-
 A static task decomposition model assumes that each

### 2018-04-15 paper/evaluation.tex
revision of text about manual code transform by alpaca
Resolved merge conflict on evalutaion

@@
 
-\textcolor{red}{highlight the difference between Alpaca GCC and Alpaca Clang.
-Alpaca GCC is about 30\%-40\% faster than Alpaca Clang.
-Writing Alpaca program by hand is a very trick thing and it is very error prone.
-Reason: a programmer needs to identify the write-after-read dependencies, change variables' names, and modify instructions accordingly, in general writing Alpaca code by hand is not a recommended, for that a compiler is needed}
-
 Our evaluation illustrates key findings about \sys. We show that: \todo{Make a list of core findings here once results are solid}{Przemek}
-
-%\sys is efficient and easy to use across our benchmarks.
 

### 2018-04-15 paper/abstract.tex
programming model used too often - removed one instance

@@
 % The main problem that we are tackling
-However, relying on the highly varying, unpredictable and intermittent harvested energy is at odds with the principal programming model: program executions require stable and uninterrupted energy supply.
+However, relying on the highly varying, unpredictable and intermittent harvested energy is at odds with the core assumption that program executions require stable and uninterrupted energy supply. 
 % The already proposed solution to the main problem

### 2018-04-15 paper/systemdescription.tex
emphasize the benefits of Coala in the system overview

--- 2018-04-14 paper/systemdescription.tex
@@
 Coalescing also amortizes per-variable commit cost, committing each location accessed by both tasks only after the second task, rather than once per task.
+\textcolor{red}{\sys's coalescing manager is an \emph{energy-aware manager}.
+It uses the recent execution history---pure software technique---as a metric to estimate the amount of the available energy and to set the adequate coalesced task size accordingly.
+This fundamental feature prevents \sys from blind coalesced task adaptation which can result in a repeated power failures problem.
+For example, If \sys enlarges the coalesced task size by $x$ static tasks on a coalesced task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, 10 times to reduce a coalesced task size from 10 to one, when $x=1$}
+
 Details of coalescing are presented in Section~\ref{sec:task_coalescing}.
 
-\textbf{\sys Task Downscaling Manager.} Details of task downscaling are presented in Section~\ref{sec:task_downsizing}. \textcolor{red}{to be written}
+\textbf{\sys Task Downscaling Manager.} \sys's downscaling manager is responsible for preserving the forward progress when \sys repeatedly fails on a single static task.
+The downscaling manager utilizes a timer-based technique and relies on the virtual memory manager to gain a full \emph{flexibility} on where to split a task.
+These features distinguish \sys task splitting technique from the traditional checkpointing.
+A checkpointing system that uses voltage threshold to place a checkpoint might cause a program to crash if the checkpoint happened to be placed between a peripheral initialization and its call.
+If a programmer disables the interrupts around this section of the code, the program may fail indefinitely to progress since the checkpoint may only be deferred.
+On a repeated failure the \sys task downscaling manager places the \emph{partial commit} after a shorter time than the one before to guarantee forward progress.
+Details of task downscaling are presented in Section~\ref{sec:task_downsizing}.}
 

### 2018-04-15 paper/systemdescription.tex
enrich the system overview

@@
 It uses the recent execution history---pure software technique---as a metric to estimate the amount of the available energy and to set the adequate coalesced task size accordingly.
-This fundamental feature prevents \sys from blind coalesced task adaptation which can result in a repeated power failures problem.
-For example, If \sys enlarges the coalesced task size by $x$ static tasks on a coalesced task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, 10 times to reduce a coalesced task size from 10 to one, when $x=1$}
-
+This fundamental feature prevents \sys from blind coalesced task adaptation which may result in a repeated power failure problem.
+For example, If \sys enlarges the coalesced task size by $x$ static tasks on a coalesced task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, 10 times to reduce a coalesced task size from 10 to one, when $x=1$}.
 Details of coalescing are presented in Section~\ref{sec:task_coalescing}.
 
-\textbf{\sys Task Downscaling Manager.} \sys's downscaling manager is responsible for preserving the forward progress when \sys repeatedly fails on a single static task.
+\textbf{\sys Task Downscaling Manager.} \textcolor{red}{\sys's downscaling manager is responsible for preserving the forward progress when \sys repeatedly fails on a single static task.
 The downscaling manager utilizes a timer-based technique and relies on the virtual memory manager to gain a full \emph{flexibility} on where to split a task.

### 2018-04-15 paper/systemdescription.tex
enrich the system overview

@@
 A checkpointing system that uses voltage threshold to place a checkpoint might cause a program to crash if the checkpoint happened to be placed between a peripheral initialization and its call.
-If a programmer disables the interrupts around this section of the code, the program may fail indefinitely to progress since the checkpoint may only be deferred.
-On a repeated failure the \sys task downscaling manager places the \emph{partial commit} after a shorter time than the one before to guarantee forward progress.
+If a programmer disables interrupts between the peripheral initialization and the call, the program may fail indefinitely to progress since the checkpoint may only be deferred.
+The \sys task downscaling manager, on the other hand,  triggers  a \emph{partial commit} faster than the previous one on a repeated failure to guarantee forward progress.
 Details of task downscaling are presented in Section~\ref{sec:task_downsizing}.}

### 2018-04-15 paper/intro.tex
References to sections updated, coala temporary online link changed, refs to alpaca and capybara removed from summary of contributions

@@
 \item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
-\item A fully-realized prototype of \sys that runs on real energy-harvesting devices~\cite{wisp,capybara}, to be released as open source after publication\footnote{An anonymized version of the \sys repository is already accessible via~\cite{coala_website} for inspection.}.
-\item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work~\cite{alpaca}, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to varied energy conditions.
+\item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication\footnote{The \sys repository will be accessible via~\cite{coala_website} immediately after publication.}.
+\item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to varied energy conditions.
 \end{itemize}
@@
 Section~\ref{sec:background} provides background on intermittent computing.
-Section~\ref{sec:overview} provides an overview of \sys, while
-Sections~\ref{sec:coalescing} and~\ref{sec:memory} describe \sys's coalescing
-and virtualization mechanisms. Section~\ref{sec:discussion} discusses \sys
-design issues. Sections~\ref{sec:methodology} and~\ref{sec:eval} describe
-\sys's methodology and evaluation. Section~\ref{sec:related} puts \sys in the
-context of related work and Section~\ref{sec:conc} concludes and discusses
+Section~\ref{sec:systemdescription} provides an overview of \sys, while
+Sections~\ref{sec:task_adaptation} describe \sys's task adaptation mechanism (coalescing and downscaling).
+Memory virtualization mechanisms is provided in Section~\ref{sec:memory_virtulaization}.
+Implementation details of \sys are give in Section~\ref{sec:implementation}.
+Section~\ref{sec:discussion} discusses \sys
+design issues. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe
+\sys's methodology and evaluation, respectively. Section~\ref{sec:related_work} puts \sys in the
+context of related work and Section~\ref{sec:conclusions} concludes and discusses
 future work.

### 2018-04-15 paper/hwswsetup.tex
update paragraph about checkpointing

@@
 Also, we do not compare \sys to checkpointing-based systems, such as DINO~\cite{dino} or Mementos~\cite{mementos} (cf.
-Section~\ref{sec:background_consistency}), as both of them are build on top LLVM compiler, which performance is inferior to GCC compiler used in the evaluation of \sys \todo{Please check if the LLVM compiler argument makes sense}{Brandon}.
+Section~\ref{sec:background_consistency}).
+\todo{Provide reasons why we do not compare against checkpointing}{Brandon}.

### 2018-04-15 paper/evaluation.tex

@@
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/overallOverhead}
-	\caption{ overall overhead breakdown: nc: No Coalecing, sc: Slow Coalescing, ea: Energy Aware coalescing \todo{}{Przemek}}
-	\label{fig:overallOverheadBreakdown}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
-	\caption{ Coalescing algorithm efficiency:  nc: No Coalecing, sc: Slow Coalescing, ea: Energy Aware coalescing, eat: Energy \& Task Aware coalescing \todo{}{Przemek}}
-	\label{fig:coalEfficiency}
-\end{figure}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/memAccess}
-	\caption{ Memory access overhead breakdown:  nc: No Coalecing, sc: Slow Coalescing, ea: Energy Aware coalescing \todo{}{Przemek}}
-	\label{fig:coalEfficiency}
-\end{figure}
-
 Our evaluation illustrates key findings about \sys. We show that: \todo{Make a list of core findings here once results are solid}{Przemek}
@@
 
+\indent \textbf{Overall \sys Overhead Breakdown.} The result is presented in Figure~\ref{fig:overallOverheadBreakdown}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/overallOverhead}
+	\caption{Overall \sys overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.}
+	\label{fig:overallOverheadBreakdown}
+\end{figure}
+
+\textbf{\sys Coalescing Algorithm Efficiency.} The result is presented in Figure~\ref{fig:coalEfficiency}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
+	\caption{\sys coalescing algorithm efficiency; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing, eta: energy and task-aware coalescing.}
+	\label{fig:coalEfficiency}
+\end{figure}
+
+\textbf{\sys Memory Access Overhead Breakdown.} The result is presented in Figure~\ref{fig:coalmemory}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/memAccess}
+	\caption{\sys memory access overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.}
+	\label{fig:coalmemory}
+\end{figure}
+
 \subsubsection{\sys Paging Performance}

### 2018-04-15 paper/hwswsetup.tex

@@
 Both antennas were parallel to each other and no physical objects were located between the WISP and the signal generator antenna.
-For continuous power experiments we used an alternative setup with MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}, which had the same MCU as WISP.
+Every benchmark we used (refer to next section) was run repeatedly for X\,s and the result was averaged for the duration of the experiment.
+\todo{check if this is correct}{Carlo} For continuous power experiments we used an alternative setup with MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}, which had the same MCU as WISP.
 In all the experiments the timing measurements were performed using either the Saleae~\cite{saleae} logic analyzer (for intermittent power experiments) or TI Code Composer Studio (for fixed power experiments).

### 2018-04-15 paper/evaluation.tex
new text to evaluation

@@
 \sys is compared to Alpaca runtime.
-We remind the reader that the benchmarks for Alpaca were hand-transformed in a time-consuming manual process from Clang (LLVM-generated code) to GCC.
+We remind the reader that the benchmarks for Alpaca were hand-transformed in a time-consuming manual process from Clang (LLVM-generated code) to GCC (see Section~\ref{sec:implementation}).
 The manual code transformation was required as \sys was written with GCC in mind, and Alpaca cannot support any other complier than LLVM.
@@
 
-\textbf{Coalescing Strategies.} We first need to assess the performance of all coalescing strategies introduced by us in Section~\ref{sec:task_coalescing}.
-Figure~\ref{fig:coalescing} shows the run time of \sys's two coalescing strategies normalized to the run time without coalescing.
-We show data from experiments with intermittent power, where intermittency was generated in software
-%\todo{Say briefly how we generate full reset with MSP430}{Amjad}
-by triggering a reset from software at intervals that were uniformly distributed with an average power availability and unavailability interval of 15\,ms and 10\,ms respectively.
-Each application ran with the same generated pattern for the fairness of comparison.
-The results show the benefit from coalescing for most applications, with the
-exception of dft, where wasted work dominated the gains from coalescing.
- %\todo{why coalescing for dft does not work but in Fig. 9 does?}{Amjad}
-The difference in performance between HC and TSHC strategies is highly
-application-dependent, with some applications benefiting from size awareness
-while others suffering, due to decisions based on inaccurate prediction of the
-actual task energy consumption. The TSHC gains are particularly small (or negative) for applications with narrower range of task sizes in Figure~\ref{fig:task_profiling}.
-%\todo{say why this difference is small}{Amjad}.
-%For this reason we shall proceed with HD coalescing strategy in subsequent evaluations.
+\textbf{Coalescing Strategies Performance.} We first need to assess the performance of all coalescing strategies introduced by us in Section~\ref{sec:task_coalescing}.
+Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies normalized to the \sys runtime without coalescing.
+We show data from experiments measured at 15\,cm WISP/RF generator antenna distance.
 
+The results show the benefit from coalescing for all applications.
+The difference in performance between all strategies is highly application-dependent.
+The gain from energy and task-aware strategy (smart) compared to task-aware only (fast) are particularly small (or negative) for applications with large task sizes, i.e.
+\textbf{fft}.
+
+%\begin{wrapfigure}{t!}{0.5\textwidth}
 \begin{figure}
@@
 \end{figure}
+%\end{wrapfigure}
 
-\textbf{\sys versus Task-based Model.} We evaluate \sys's run time performance and compare \sys to the run time performance of Alpaca.
+\textbf{\sys versus Task-based Model.} We now evaluate \sys's run time performance and compare \sys to the run time performance of Alpaca.
 Results are given in Figure~\ref{fig:runtime}.
-For each application we measured its execution time using real wireless power provision from the RF signal generator (refer to Section~\ref{}).
+For each application we measured its execution time using real wireless power provisioned from the RF signal generator at three distances (refer again to Section~\ref{sec:implementation}).
 
-The data show the average execution time of each application run repeatedly for 20 seconds, with runtimes normalized to \sys's average.
-The results show that \sys provides a performance benefit compared to Chain for most applications.
-The performance benefit is greatest for applications that have small tasks (bc, sort, cuckoo) at all distances on intermittent power and with fixed power.
-In applications that have larger tasks (e.g.
-dft) or access more different memory pages, \sys incurs overhead from memory virtualization that cause its performance to be comparable to (or in the \emph{single} cem case, worse than) Chain.
-\todo{Rewrite conclusions}{Przemek}
+The data show the average execution time of each application with runtimes normalized to \sys's average.
+The results show that \sys provides a performance benefit compared to Alpaca for most applications (all except for \textbf{ar} and \textbf{cuckoo}).
+The performance benefit is greatest for applications that have small tasks (\textbf{bc}, \textbf{sort}, \textbf{dijkstra}) at all distances on intermittent power.
+In applications that have larger tasks or access more different memory pages, \sys incurs overhead from memory virtualization that cause its performance to be comparable to (or worse than) Alpaca.
 
@@
 	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
-	\caption{\textbf{Performance of \sys applications (divided manually into tasks) at multiple WISP to RFID reader antenna distances}: fixed power (0\,m to the antenna) and \{0.2, 0.4\}\,m (left, center and right-most bar per application, respectively), compared against Chain (results normalized).}
+	\caption{Performance of \sys applications at multiple WISP to RF generator antenna distances ()\{0.15, 0.3, 0.5\}\,m (left, center and right-most bar per application, respectively), compared against Alpaca (results normalized).}
 	\label{fig:runtime}
@@
 
-\subsection{Characterization of \sys Overhead}
+\subsubsection{Characterization of \sys Overhead}
 \label{sec:coala_overhead}

### 2018-04-15 paper/adaptiveModel.tex
add brain dump about Coala the execution model

@@
-\textcolor{red}{THIS BRAIN DUMP ABOUT THE EXECUTION MODEL - Placeholder}
+\subsection {Motivating the Design of \sys's Execution Model.} The challenge of \emph{efficiently} executing a program under frequent power failures, lies in the fact that the program needs to be energy aware.
+This challenge becomes even bigger when \emph{generality} is a design requirement, and consequently, the system must not require a specific hardware circuitry to offer its efficiency.
+\emph{Flexibility} is a key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software technology.
+\emph{\sys champions an efficient, general, and flexible intermittent execution.}
 

### 2018-04-15 paper/background.tex
limitations of usability by Alpaca added

--- 2018-04-14 paper/background.tex
@@
 	\end{subfigure}
-	\caption{\textbf{Code example demonstrating effect of write after read on volatile memory checkpointing.}}\label{fig:code_demo_incosistency}
+	\caption{\textbf{Code example demonstrating checkpointing gives rise to data inconsistency.}}\label{fig:code_demo_incosistency}
 \end{wrapfigure}
@@
 These devices elide tethered power or a battery, instead collecting energy into a capacitor, operating when sufficient energy accumulates, and upon depleting the energy, turning off and recharging.
-\TODO{We need a figure that shows what intermittent execution looks like.}
+%\TODO{We need a figure that shows what intermittent execution looks like.}
 %
@@
 %
-Many platforms enable intermittent, battery-less energy harvesting-based computation.
+There are several energy harvesting battery-less hardware platforms.
 For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
-Other intermittently-powered platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
+Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
 
@@
 
-\textcolor{red}{reference work by Intel~\cite{baghsorkhi_cgo_2018}}
+%\textcolor{red}{reference work by Intel~\cite{baghsorkhi_cgo_2018}}
 
@@
 %
-Software running on an energy-harvesting device executes {\em intermittently} because buffered energy is only available \emph{sometimes}.
+Software running on an energy-harvesting device executes {\em intermittently} because energy sources are not always available to harvest and buffer sufficient operational energy.
 An intermittent execution is composed of operating periods interspersed with power failures~\cite{dino,chain,alpaca,ratchet}.
 The frequency of failures depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods.
-Energy-harvesters provide input power orders of magnitude less than operating power, making recharging negligible during operation.
-
-Intermittent execution is different from continuous execution.
+%Energy-harvesters provide input power orders of magnitude less than operating power, making recharging negligible during operation.
+%Intermittent execution is different from continuous execution. 
 A power failure clears volatile  state (registers, stack, and globals) and non-volatile memory (e.g., FRAM) persists.
-At a failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
+Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
 Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
@@
 Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value.
-The situation occurs when code includes a WAR dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}.
-
-Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
+The situation occurs when code includes a WAR dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
+%Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
 The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
-
-Despite checkpointing at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
+%Despite checkpointing at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
 Figure~\ref{fig:crc_example_war} shows an intermittent execution.
@@
 Task flow happens at programmer demarcated transition points that may be conditional on program values.
-Task-based programming abstractions guarantee that tasks execute {\em atomically}, regardless of power failures.
+%Task-based programming abstractions guarantee that tasks execute {\em atomically}, regardless of power failures. 
 Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent.
@@
 
-Despite its advantages, the task-based programming model for intermittently-powered computing devices has its limitations.
-Two major ones being: {\em memory overhead} and {\em inflexibility to changing energy conditions}.
+Despite their advantages, the task-based programming models have their limitations. Two major ones being: {\em memory overhead} and {\em inflexibility to dynamic energy harvesting conditions}.
 
@@
 
+%Przemek placeholder
+%\textbf{Task-based Models have Syntactic Limitations.} State of the art task-based programming models, Alpaca~\cite{alpaca}, has substantial limitations that prohibit writing generic programs.
+First, for Alpaca (and Chain~\cite{chain}), accesses to protected variables must be applied \emph{only} by using the variable's name, i.e.
+reference by variable address is not possible.
+This limitation implies that existing third-party libraries, such as MCU-tailored FFT library~\cite{}, cannot be implemented to support Alpaca or Chain (not to mention the impossibility of the development of an intermittent-safe \texttt{memcpy}).
+Furthermore, if a protected variable needs to be accessed in a code path that is not predictable by the compiler, for instance an Interrupt Service Routine, Alpaca cannot protect that access.
+
 \textbf{Fixed-size Tasks are Inflexible and Inefficient.} If a programmer writes a task that consumes {\em more} energy than the device can buffer, the task will never complete using buffered energy {\em preventing forward progress} by repeatedly re-executing.

### 2018-04-15 paper/background_conflict-20180414-165358.tex

--- 2018-04-14 paper/background.tex
@@
 		\hline\hline
-		Mementos~\cite{mementos}	& Registers + Stack     \\
-		DINO~\cite{dino}	& Registers + WAR variables \\
-		Chain~\cite{chain}	& PC + channel data\\
+		Mementos~\cite{mementos}                             & Reg. + Stack     \\
+		DINO~\cite{dino}                                     & Reg. + WAR variables \\
+		Chain~\cite{chain}                                   & PC   + Channel data\\
 		Alpaca~\cite{alpaca}                                 & PC   + WAR variables \\
-		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
+		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Reg. (requires NV main memory) \\
 		Region Formation~\cite{baghsorkhi_cgo_2018} & \textcolor{red}{XXX} \\
@@
 	\end{tabular}
-	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{PC}: program counter, \emph{channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
+	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{Reg.}: the entire register file, \emph{PC}: program counter, \emph{Channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
 	\label{table:chechpoint_comparison}
@@
 \subsubsection{Costs of Task-based Models}
-\label{sec:cost_task-based}
 
-Despite its advantages, the task-based programming model for intermittently-powered computing devices has its limitations.
-Two major ones being: {\em memory overhead} and {\em inflexibility to changing energy conditions}.
+%Task-based models incur costs that motivate \sys.
 
@@
 
+%\begin{figure}
+%\centering
+	%\subfloat[Energy cost of \emph{write} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_write.eps}}
+	%\subfloat[Energy cost of \emph{read} operation]{\includegraphics[width=0.49\columnwidth]{figures/fram_read.eps}}
+	%\caption{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write (left) and read (right) operation.}
+	%\label{fig:framEnergy}
+%\end{figure}
+
 To understand these non-volatile memory overheads, for the MSP430FR5969~\cite{msp430datasheet} MCU we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations.
@@
 Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device.
-This observation is also consistent with the TI specification~\cite{}.
-\todo{add citation to TI specification on FRAM energy cost}{Przemek} In an intermittent device, energy cost corresponds to decreased run time between power failures.
+In an intermittent device, energy cost corresponds to decreased run time between power failures.
 The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.
@@
 
-Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity.
-Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}.
+Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware with a fixed energy buffering capacity.
+Buffer sizes from prior work vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}.
 Sizing tasks complicates {\em porting} code from one device to another.
-An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer.
+An excessively costly task on a device with a small energy buffer may be a relatively short, task on a device with a larger buffer.
 The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead.
-\sys's execution model based on {\em task coalescing}---merging or dissecting multiple tasks {\em during} execution based on the program execution behavior, and {\em task downscaling}---shortening task boundaries during execution when even the smallest task does not fit the energy buffer, both models described in Section~\ref{sec:task_adaptation}, is motivated by these observations.
+\sys's {\em task coalescing}, described in Section~\ref{sec:task_coalescing}, is motivated by these observations.
 
-%\textbf{Motivating the Design of \sys's Execution Model.} The challenge of \em{efficiently} executing a program under frequent power failures, lies in the fact the the program needs to be energy aware.
-This challenge becomes even bigger when \em{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
-\em{Fllexibility} is key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software technology.
-\em{\sys champions an efficient, general, and flexible intermittent execution.}
+\subsection {Motivating the Design of \sys's Execution Model.} 
+%% Coala’s model motivation
+The challenge of \emph{efficiently} executing a program under frequent power failures, lies in the fact that the program needs to be energy aware.  This challenge becomes even bigger when 
+Hardware support is not allowed to advocate the approach \emph{generality} and to eliminate hardware energy consumption overhead.
+Furthermore, \emph{flexibility} is a key for wide adaptation, therefore, \sys's execution model should not limit a programmer to a subset of the functionalities of the underlying software technology---\emph{\sys champions efficient, general, and flexible intermittent execution.}
+
+%% How does Coala’s execution model progress?
+\sys execution model operates on task-based programming model, but It progresses on a dynamic, virtual task granularity.
+A virtual task can be either multiple static tasks \emph{coalesced} together, when the energy permits, or a sub-task, when it is necessary to preserve forward progress.
+However, as more tasks coalesce, their  collective commit overhead amortizes better, but the risk of wasting work also increases.
+If power fails during a long sequence of coalesced tasks, execution will restart from the last commit, i.e.
+the first task in the sequence, losing  the progress made by any of the coalesced tasks.
+The challenge to coalescing tasks is determining how many tasks to coalesce before committing.
+
+%% How does Coala’s execution model decide on its step size when there is abandoned energy?
+\sys uses the recent execution history---pure software technique---as a metric to estimate the amount of the available energy and to set the adequate virtual task size accordingly.
+This fundamental feature prevents \sys from blind virtual task adaptation which can result in a repeated power failures problem.
+For example, If \sys enlarges the virtual task size by $x$ static tasks on a virtual task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, 10 times to reduce the virtual task size from 10 to one when $x=1$.
+
+%% How does Coala’s execution model decide on its step size when there is scarce energy?
+
+Coala's execution model is neither limited by a static task size to progress, nor by the commit size, i.e.
+the amount of data that needs to be saved to the non-volatile memory to preserve execution progress.
+Coala can partition compile-time tasks, if it fails on a single static task more than once, or merge them, if saving the state to non-volatile memory is redundant because a power failure is not expected soon.
+This unique technique makes the commit size proportional to the number of state-less instructions executed.
+In other words, if energy allows executing only a small chunk of code, the amount of data to be committed shrinks accordingly.

### 2018-04-15 paper/intro.tex

@@
 Advances in processor efficiency along with the development of
-energy-harvesting systems has created a new category of devices that require
+energy-harvesting systems have created a new category of devices that require
 neither a battery nor a tethered power
@@
 3c]{nadeau_naturebio_2017}.
-The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor's~\cite{capybara} and some devices fail and restart operating $\approx$10 to
+The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
 $\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.
@@
 continuously-powered behavior, often leading to an unrecoverable
-application failure~\citep{dino,edb}. Prior work developed two main approaches to dealing with data inconsistency for
+application failure~\citep{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for
 intermittently-powered devices: (i) \emph{software-based programming and
@@
 
-A savvy programmer may thus create very large tasks in an attempt to minimize
+A savvy programmer may thus create very large tasks in an attempt to reduce
 overhead by minimizing the number of task transitions. However, a very large
@@
 we propose a third way, using a novel technique called {\em adaptive execution by dynamic task
-coalescing/splitting} that efficiently executes small tasks, avoiding unnecessary overheads while avoiding the risk of non-termination.
+coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads while avoiding the risk of non-termination.
 
@@
 with no commit or task transition overhead between them, instead performing
-task-end commit actions at the end second task only.
+task-end commit actions at the end of the second task only.
 %
@@
 approach to \emph{memory virtualization} that buffers non-volatile variable
-updates in volatile memory during coalesced task execution, before committing
+updates in volatile memory during a coalesced task execution, before committing
 them to non-volatile memory at the dynamic task boundary.
@@
 energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
-Partial commit avoids non-termination by committing the intermediate state of a long-running
-task that has repeatedly failed and restarted.  Partial commit violates
+Partial execution avoids non-termination by committing the intermediate state of a long-running
+task that has repeatedly failed and restarted. Partial execution violates
 task atomicity, but preserves forward progress; if a programmer knows that
-task atomicity is crucial to correctness, they can disable partial commit
-instead risking non-termination.
+task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
 
@@
 \item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication\footnote{The \sys repository will be accessible via~\cite{coala_website} immediately after publication.}.
-\item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to varied energy conditions.
+\item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to diverse energy conditions.
 \end{itemize}

### 2018-04-15 paper/partialcommit.tex

@@
 
+We proceed with the second task adaptation feature of \sys, namely reduction of the task size boundary to enable execution of a program on any energy buffer size.
+
 \begin{figure}
-	\caption{\textcolor{red}{Figure to describe partial commit is needed}}
+	\caption{\todo{Figure to describe partial commit is needed}{Sinan, Carlo}}
 \end{figure}
@@
 A partial commit saves the stack, registers and the volatile pages to the non-volatile memory.
-
 \sys starts with a big random guess and down-scales it using a binary search until it succeeds in preserving the partial execution progress.
@@
 
-\subsection{Why partial commit is not \sys's default behavior}
+\todo{Proof of the maximum granularity of task downscaling - show it is limited by the clock}{Carlo}
 
-\emph{Static deployment of partial commit stubs} This approach reduces programmer effort---they still need to apply \sys's API and distribute trigger points for partial commits---but it introduces an additional overhead by saving the stack and the register file each time it commits the pages.
+\subsubsection{Why Task Downsizing is not \sys's Default Behavior}
 
-\emph{Dynamic deployment of partial commit stubs:} This approach does not rely on any information from the programmer to trigger the commit process.
-However, it utilizes a trial and error approach to optimize the commit rate over the entire program execution time.
+Task downsizing is the core novelty of \sys when it comes to code portability and re-usability. However, it has limitations, which force  it to be a non-default \sys behavior.
+
+\textbf{Static deployment of partial commit stubs.} This approach reduces programmer effort, although the programmer still needs to apply \sys's API and distribute trigger points for partial commits.
+On the other hand, \sys' downsizing introduces an additional overhead by saving the stack and the register file each time it commits the pages.
+
+\textbf{Dynamic deployment of partial commit stubs.} \sys downsizing does not rely on any information from the programmer to trigger the commit process.
+However, \sys utilizes a trial and error approach to optimize the commit rate over the entire program execution time.
 Therefore, it suffers from a large cumulative re-execution penalty.

### 2018-04-15 paper/newCoalescing.tex
some changes to the coalescing, still a lot of work needed

@@
 \subsection{Task Coalescing}
+\label{sec:task_coalescing}
 
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/coal_Intro_Figure.pdf}
-	\caption{When $N$ tasks coalesced $N-1$ commit operations are skipped.}
-	\label{fig:coalIntro}
-\end{wrapfigure}
+% \begin{wrapfigure}{t!}{0.5\textwidth}
+% 	\centering
+% 	\includegraphics[width=0.5\columnwidth]{figures/coal_Intro_Figure.pdf}
+% 	\caption{When $N$ tasks coalesced $N-1$ commit operations are skipped.}
+% 	\label{fig:coalIntro}
+% \end{wrapfigure}
 
@@
 pages modified by any of the $N$ tasks to the statement $N-1$.
-Coalescing is likely to reduce the total number of non-volatile memory
+Coalescing reduces the total number of non-volatile memory
 accesses, because when coalesced tasks access the same pages of data, only the
-last version of the page needs to be committed.  \textcolor{red}{maybe is possible to demonstrate with data? count number of executed non-volatile vs volatile accesses}
+last version of the page needs to be committed.  
+\textcolor{red}{maybe is possible to demonstrate with data? count number of executed non-volatile vs volatile accesses}
 
-\begin{algorithm}
+\textbf{Trade-off between Speed and Wasted Effort.} \sys can coalesce an
+arbitrary number of consecutive tasks. However, as more tasks coalesce, their
+collective commit overhead amortizes better, but the risk of wasting work also
+increases. If power fails during a long sequence of coalesced tasks, execution
+will restart from the last commit, i.e. the first task in the sequence, losing
+the progress made by any of the coalesced tasks.  The challenge to coalescing
+tasks is determining how many tasks to coalesce before committing.
+%
+%During execution, an executing task is either a {\em coalesced} task, which
+%does not commit its state before transitioning or a {\em committing} task,
+%which does commit its state before transitioning.
+
+\textbf{Task Coalescing Strategies.} 
+%
+\begin{algorithm}[t]
 	\caption{Coalescing}
@@
 \end{algorithm}
+%
+%
+The likelihood that a task group will complete depends on the total amount of
+energy consumed by all tasks in the group, i.e. the size of the dynamic task.
+%
+A good coalescing strategy must moderate the risk of wasted work and capitalize
+on benefits of deferred commits by adapting the target size of the dynamic task
+to the energy available at runtime.
+
+To describe the proposed strategies we will use the following parameters: \\
+(i) \textit{target budget B}: total number of static tasks to be coalesced; (ii) \textit{current budget $B_\text{c}$}: remaining number of tasks to be coalesced before committing to non-volatile memory; (iii) \textit{history H}: up on reboot, it is the number of executed static tasks between the last two power interrupts.
+While executing, it is the number of executed static tasks since the last power interrupt.
 
@@
 % %% How does Coala approximates an ideal task
-\sys adapts its virtual execution by mean of \emph{coalescing} static tasks, when energy permits, to optimally execute an intermittently powered program.
-It devises a number of coalescing algorithms, and states that a good coalescing strategy must moderate the risk of wasted work (re-execution penalty) and capitalize on benefits of deferred commits.
-To describe the proposed strategies we will use the following parameters: \\
-(i) \textit{target budget B}: total number of units to coalesce; (ii) \textit{current budget $B_\text{c}$}: remaining number of tasks to be coalesced before committing to non-volatile memory; (iii) \textit{history H}: total number of tasks executed in the current discharge cycle, initialized to the number of executed tasks in the last power cycle.
 
@@
 
-\subsubsection{Slow Coalescing}
-\label{subsec:slowCoalescing}
+\subsubsection{Energy-blind Coalescing}
+\label{subsec:energyBlind}
 
@@
 	\includegraphics[width=0.5\columnwidth]{figures/slowCoal}
-	\caption{The {\em slow coalescing} algorithm \emph{linearly} updates its coalescing target.}
+	\caption{The {\em energy blind coalescing} algorithm \emph{linearly} updates its coalescing target.}
 	\label{fig:slowCoal}
@@
 \label{subsec:ECoalescing}
-
-% \begin{figure}
-% 	\centering
-% 	\includegraphics[width=\columnwidth]{figures/HCCoal}%coalescing_timeline
-% 	\caption{The History-aware coalescing algorithm  updates its coalescing target}
-% 	\label{fig:HCCoal}
-% \end{figure}
-
-% The History-aware algorithm uses its recent execution history as metric to estimate the optimal coalesce task size.
-Relying on the history of execution enables this algorithm to react very quickly to discharge rates variation.

### 2018-04-15 paper/newCoalescing.tex
correct algo 1

@@
 \textcolor{red}{maybe is possible to demonstrate with data? count number of executed non-volatile vs volatile accesses}
-
+%
 \textbf{Trade-off between Speed and Wasted Effort.} \sys can coalesce an
@@
 	        % \State $B$ \gets \Call{f\_compl}{$\texttt{B},\texttt{H_\text{c}},\texttt{H_\text{p}}$}
+	         \State $B \leftarrow $ \Call{$f_\text{reboot}$}{B, $H_\text{c}$,$H_\text{p}$}
 	        \State \Call{commit\_to\_fram}{}
@@
 to the energy available at runtime. 
-
-To describe the proposed strategies we will use the following parameters: \\
-(i) \textit{target budget B}: total number of static tasks to be coalesced; (ii) \textit{current budget $B_\text{c}$}: remaining number of tasks to be coalesced before committing to non-volatile memory; (iii) \textit{history H}: up on reboot, it is the number of executed static tasks between the last two power interrupts.
+%
+To describe the proposed strategies we will use the following parameters:
+\begin{itemize}
+\item target budget $B$: total number of static tasks to be coalesced.
+\item current budget $B_\text{c}$: remaining number of tasks to be coalesced before committing to non-volatile memory; 
+\item history $H$: up on reboot, it is the number of executed static tasks between the last two power interrupts.
 While executing, it is the number of executed static tasks since the last power interrupt.
+\end{itemize}
 

### 2018-04-15 paper/newCoalescing.tex
correct algo 1
corrct the function calls for the energy blind algo

@@
 	        \State $ B_\text{c} \gets B$
+	        %
 	        \While{$ B_\text{c} > 0$}
 		        \State \Call{execute\_task}{$T_\texttt{i}$}
-		        % \State $W$ \gets \Call{f_\text{weight}}{$T_\texttt{i}$}
+		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
 		        \State $B_\text{c} \gets B_\text{c} - W$
@@
 	        \EndWhile
-	        % \State $B$ \gets \Call{f\_compl}{$\texttt{B},\texttt{H_\text{c}},\texttt{H_\text{p}}$}
-	         \State $B \leftarrow $ \Call{$f_\text{reboot}$}{B, $H_\text{c}$,$H_\text{p}$}
-	        \State \Call{commit\_to\_fram}{}
+	        %
+	         \State $B \leftarrow $ \Call{$f_\text{compl}$}{B, $H_\text{c}$,$H_\text{p}$}
+	        \State \Call{commit\_to\_fram}{\null}
         \EndWhile
@@
 $T_\text{i}$ is the i-th task being executed and $W$ is its weight.
-A coalescing strategy and its performance are defined by the functions: \\
-(i) \emph{$f_\text{reboot}$}: which updates the target budget after a reboot; (ii) \emph{$f_\text{weight}$}: which returns the weight of the task just executed and its unit; (iii) \emph{$f_\text{compl}$}: which updates the target budget after a successful completion of a dynamic coalesced task.
+A coalescing strategy and its performance are defined by the functions:
+\begin{itemize}
+\item $f_\text{reboot}$: which updates the target budget after a reboot.
+\item $f_\text{weight}$: which returns the weight of the task just executed and its unit. 
+\item $f_\text{compl}$ : which updates the target budget after a successful completion of a dynamic coalesced task.
+\end{itemize}
 The following sections detail the different strategies \sys proposes and the needs that originated from one strategy leading to the conception of another one.
@@
 The need for \emph{adaptiveness}, that is coalescing static tasks, can be fulfilled by a naive strategy, having the target budget $B$ to increase by $x$ static tasks after a successful completion of a coalesced task, and to decrease by the same number of tasks upon a reboot.
-In this case: $f_\text{reboot} = B - x$; $f_\text{weight} =  1$; $f_\text{compl} = B + x$; \\
+In this case:
+\begin{itemize}
+\item $f_\text{reboot}(B, H_\text{p}) = B - x $;
+\item $f_\text{weight}(T_\text{i}) =  1$; 
+\item $f_\text{compl}(B,H_\text{c},H_\text{p}) = B + x$; 
+\end{itemize}
 Due to the linear behavior of the target adaptation, this algorithm is slow in reacting to the changes in energy conditions; the energy required by a coalesced task or the available energy.

### 2018-04-15 paper/newCoalescing.tex
describe the energy-aware coalecing algo

@@
 \label{subsec:ECoalescing}
+This algorithm tunes its target budget based on the recent history of execution.
+It takes advantage of the guarantees that the energy buffer offers on a fresh start by coalescing many static tasks.
+Then it approaches the expected power failure carefully by reduces the task target size by x\%, amongst 25\%, 50\% and 75\% the 50\% shows the best performance and therefore it is the default of \sys.
+If the algorithm concludes that it wrongly estimated the time of the power failure, it will force target re-tuning based on the new history of execution.
+
+\begin{wrapfigure}{}{0.5\textwidth}
+	\centering
+	\includegraphics[width=0.5\columnwidth]{figures/energy_aware_algo}
+	\caption{paceholder}
+	\label{fig:energyAware}
+\end{wrapfigure}

### 2018-04-16 paper/hwswsetup.tex
methodology: edits
* merged App Characterization into table 3

--- 2018-04-15 paper/hwswsetup.tex
@@
-Our implementation of \sys makes no system assumptions: we execute \sys on a \emph{real} low-power embedded computing platform powered by a \emph{real} intermittent energy harvesting source.
-This approach shall lead to undisputed set of \sys performance metrics.
+We prototype \sys and use its API to implement a set of benchmark
+applications representative of the embedded domain. We build the applications
+and deploy the binaries onto a real energy-harvesting device. In this section
+we describe the setup for the experiments presented in
+Section~\ref{sec:evaluation} and characterize the benchmark applications.
 
-\subsection{Intermittent Platform and Hardware Toolchain}
+\subsection{Intermittently-powered Platform and Experimental Setup}
 \label{sec:results_hardware_software}
 
-We evaluated \sys in two power supply environments: (i)  \emph{intermittent} and (ii) \emph{fixed} (to benchmark the intermittent power supply results).
-In the experiments with intermittent power supply, we evaluated \sys using a WISP\,5.1~\cite{wisp5,wisp}---energy-harvesting platform build around a MSP430FR5969~\cite{wolverine} MCU with with 64\,kB of non-volatile (FRAM) memory and and 2\,kB of volatile memory (SRAM).
-MCU was operating at 1\,MHz clock cycle.
-We powered the WISP using RF signal generator~\cite{} \todo{Add reference to signal generator}{Przemek} emitting 20\,dBm sinusoidal wave at 915\,MHz.
-Signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic antenna~\cite{atlas2015}.
-For distance-controlled experiments, we affixed WISP at two paper protrusions around 10\,cm from the surface of the table (parallel to, and at the edge of, the table surface).
-Antenna of the signal generator was facing WISP directly and located at three distances, $d=\{15, 30, 50\}$\,cm, from WISP.
-For the task downscaling experiments, an additional distance of $d=60$\,cm was considered.
-Both antennas were parallel to each other and no physical objects were located between the WISP and the signal generator antenna.
-For continuous power experiments we used an alternative setup with MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}, which had the same MCU as WISP.
-In all the experiments the timing measurements were performed using either the Saleae~\cite{saleae} logic analyzer (for intermittent power experiments) or TI Code Composer Studio (for fixed power experiments).
-\todo{provide version and more details on TI CCS}{Przemek} Both type of experiments are explained in the following section.
+We used two platforms to evaluate \sys: (1) an \emph{intermittently-powered}
+energy-harvesting device, and (2) a \emph{continuously-powered} evaluation
+board.
+%
+Every benchmark listed in the next section was run repeatedly for X\,s on
+each platform and the result was averaged for the duration of the experiment.
+\todo{check if this is correct}{Carlo}
+
+Experiments on the intermittently-powered platform used the
+WISP\,5.1~\cite{wisp5,wisp} RF-powered energy-harvesting device.
+%
+The WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
+non-volatile (FRAM) memory and 2\,kB of volatile (SRAM) memory and was
+configured to 1\,MHz clock speed.
+%
+We powered the WISP using an RF signal generator~\cite{} \todo{Add reference
+to signal generator}{Przemek} emitting 20\,dBm sinusoidal wave at 915\,MHz.
+The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
+antenna~\cite{atlas2015}.
+%
+The antenna was oriented towards and in parallel with the WISP antenna, and
+no objects obstructed the path.
+%
+We affixed the WISP with a paper harness at the edge of a table, so that the
+WISP was parallel to the table surface at a height of 10~cm.
+%
+For distance-controlled experiments we positioned the signal generator
+antenna at a fixed distance from the WISP. The distances were $d=\{15, 30,
+50\}$\,cm for the main experiments and 60~cm for the task downscaling
+experiments.
+%
+To obtain execution time, the software toggled GPIO pins at the beginning and
+end of the application, and the Saleae~\cite{saleae} logic analyzer measured
+the interval between edges in the signal.
+
+The experiments which are not sensitive to the behavior of an
+energy-harvesting power supply were conducted on an MSP-EXP430FR5969
+launchpad~\cite{MSP-EXP430FR5969_launchpad} board powered continuously with a
+power supply.
+%
+This platform has the same MCU as the WISP.
+%
+Execution time was measured using the instrumentation features in TI Code
+Composer Studio IDE.
+%
+\todo{provide version and more details on TI CCS}{Przemek}
 
@@
 	\footnotesize
-	\begin{tabular}{| r | p{0.85\columnwidth} |}
+	\begin{tabular}{| r|r|r | p{0.65\columnwidth} |}
 		\hline
-		Application & Description \\
+		Application & No. tasks & SLOC & Description \\
 		\hline\hline
-		\textbf{ar} & nearest neighbor classification of randomly generated data modeling a three-axis accelerometer\\
+        \textbf{ar} & & & Activity recognition using a nearest neighbor
+classifier with a deterministic input from a randomly-generated sequence of
+accelerometer data\\
 		\hline
-		\textbf{bc} & counts bits in byte pseudo-random string using several algorithms, and compares results for correctness\\
+        \textbf{bc} & & & Counts set bits in a pseudo-random sequence of bytes
+using several algorithms and compares results for correctness\\
 		\hline
-		\textbf{cuckoo} & Bloom-filter-like Cuckoo filter: first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence\\
+        \textbf{cuckoo} & & & Inserts and queries a sequence of pseudo-random
+values in a Cuckoo filter, a set membership data structure with deletion\\
 		\hline
-		\textbf{dijkstra} &  Dijkstra algorithm executed continously on a predefined adjecency matrix\\
+        \textbf{dijkstra} &  & & Dijkstra shortest path algorithm executed on a
+predefined adjacency matrix\\
 		\hline
-		\textbf{fft} & continuously repeating Fast Fourier Transform on three pre-generated 128 sample vector\\
+        \textbf{fft} & & & Fast Fourier Transform on three pre-generated 128
+sample vectors\\
 		\hline
-		\textbf{sort} & selection sort algorithm on a fixed sequence of numbers\\
+		\textbf{sort} & & & Selection sort of a fixed sequence of numbers\\
 		\hline
 	\end{tabular}
-\caption{Suite of benchmark applications used in evaluating \sys performance.
-All applications have been chosen from MiBench testing suite~\cite{mibench,hicks_mibench2_2016}.\todo{Consider adding more information here}{Przemek, Carlo}}
+\caption{Description and characteristics of benchmark applications used to
+evaluate \sys.}
 \label{table:benchmark_table}
@@
 
-We evaluated \sys using six benchmarks summarized in Table~\ref{table:benchmark_table}, with their implementation is available via~\cite{coala_website}.
-Five benchmarks originate from the MiBench testing suite~\cite{mibench,hicks_mibench2_2016}: three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) were already implemented by Chain~\cite{chain} and Alpaca~\cite{alpaca}, while two (\textbf{dijkstra}, \textbf{fft}\footnote{Specifically, we re-implemented FFT based on the TI's accelerated FFT library~\cite{}.\todo{add reference}{Przemek}}) are added as new.
-The final one (\textbf{sort}) has been written from scratch.
-\todo{justify the selection of these benchmarks}{Przemek, Carlo, Amjad, Brandon} All applications were compiled using GCC version 3.8.0.
-\todo{Check if this still holds.
-provide optimization flag}{Przemek}
+We evaluated \sys using six benchmarks summarized in
+Table~\ref{table:benchmark_table}. %
+Three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) are from prior
+work~\cite{chain,alpaca} and three benchmarks (\textbf{dijkstra},
+\textbf{fft}, \textbf{sort}) are new implementations of applications
+from the MiBench suite~\cite{mibench,hicks_mibench2_2016}. The
+\textbf{fft} implementation is based on the Accelerated FFT library from
+TI~\cite{ti-fft}.
+%
+\todo{add reference}{Przemek}
+%
+All applications were compiled using TI GCC~\cite{ti-gcc} version 3.8.0 with
+\texttt{-O1}.
+%
+\todo{Check if this still holds. provide optimization flag}{Przemek} 
+%
+The source code for all benchmarks is released at \cite{coala_website}.
 
@@
 
-\textbf{Comparison with Alpaca and Alpaca Compiler Limitations.} We compare \sys against state-of-the-art task-based runtime Alpaca~\cite{alpaca}.
-For correct comparison, all benchmarks of \sys were divided into tasks in the same way as Alpaca.
-Since original Alpaca runtime was written in Clang, we had to transform every benchmark manually into GCC-complaint code.
-\todo{Provide more information about porting to GCC from LLVM}{Brandon, Carlo} This was a time-consuming process since a programmer needs to identify WAR dependencies \todo{check if we use WAR or write-after-read or anything else}{Przemek}, then change variables' names and modify instructions accordingly.
-Nota bene, our experiments based on the benchmarks we used here show that Alpaca compiled with GCC is about 30\%-40\% faster than Alpaca compiled with Clang.
-\todo{Provide concrete numbers for all applications}{Amjad} The speed improvement we got for Alpaca with GCC is however sacrificed by the above-mentioned error-prone manual code transformation process for which no compiler support exists thus far.
-\sys \emph{does not require any of such transformations}.
+\textbf{Comparison with Alpaca using the GCC compiler.} We compare
+\sys against Alpaca~\cite{alpaca}, a state-of-the-art task-based system for
+intermittent computing. For a fair comparison the task decomposition is the
+same in both \sys and Alpaca implementations of the benchmarks. We compare
+both systems using GCC, the compiler with the best performance for the
+MSP430 architecture.
 
-\textbf{Comparison of \sys against other Systems.} We do not compare \sys against Chain~\cite{chain}, another task-based system, which is inferior to Alpaca in its performance.
-Also, we do not compare \sys to checkpointing-based systems, such as DINO~\cite{dino} or Mementos~\cite{mementos} (cf.
-Section~\ref{sec:background_consistency}).
-\todo{Provide reasons why we do not compare against checkpointing}{Brandon}.
+Since Alpaca's compiler pass is implemented only for LLVM, we could not use it
+to instrument the benchmarks and compile them with GCC. Instead, we
+\emph{manually} performed the instrumentation done by the compiler pass.
+We performed the instrumentation at the source level, adding C
+statements to the program, instead of at the level of binary
+intermediate-representation (LLVM IR) as the compiler pass would. The result is
+equivalent from the perspective of overheads. The
+instrumentation consists of identifying write-after-read dependencies and
+adding code and memory allocations to make a private copy of the
+affected variables. For array variables, instrumentation also includes adding
+code and memory allocations to track the first write to the array.
+
+Manual instrumentation of Alpaca code is not a general approach, but only
+served as a practical alternative to porting the compiler pass from LLVM to
+GCC.
+%
+Our experiments show that manually-instrumented Alpaca code compiled with GCC
+is about 30\%-40\% faster than Alpaca code instrumented automatically by the
+Alpaca LLVM compiler pass and compiled with LLVM.
+%
+\todo{Provide concrete numbers for all applications}{Amjad}
+%
+This performance difference is due to the early experimental state of the
+MSP430 backend for LLVM.
+%
+By compiling both systems with the same compiler (GCC), we ensure that our
+comparison in Section~\ref{sec:evaluation} is fair.
+
+% Unnecessary
+%\textbf{Comparison of \sys against other Systems.} We do not compare \sys
+%against Chain~\cite{chain}, another task-based system, which is inferior to
+%Alpaca in its performance. Also, we do not compare \sys to
+%checkpointing-based systems, such as DINO~\cite{dino} or
+%Mementos~\cite{mementos} (cf. Section~\ref{sec:background_consistency}).
+%\todo{Provide reasons why we do not compare against checkpointing}{Brandon}.

### 2018-04-16 paper/evaluation.tex
methodology: edits
* merged App Characterization into table 3
text on Fig 11

@@
 Our aim is to assess \sys as detailed as possible.
-For this we have characterized \sys runtime performance (execution time), based on a set of benchmarks, considering task adaptation: task coalescing and downscaling.
+For this we have characterized \sys runtime performance (execution time), based on a set of benchmarks, considering core task adaptation mechanisms of \sys: \emph{task coalescing} and \emph{task downscaling}.
 We also characterize \sys overhead showing precisely what is the source of \sys superior performance.
@@
 
-\textbf{Coalescing Strategies Performance.} We first need to assess the performance of all coalescing strategies introduced by us in Section~\ref{sec:task_coalescing}.
+\textbf{Coalescing Strategies Performance.} We first need to assess the performance of all coalescing strategies introduced in Section~\ref{sec:task_coalescing}.
 Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies normalized to the \sys runtime without coalescing.
-We show data from experiments measured at 15\,cm WISP/RF generator antenna distance.
-
+\todo{Check if this normalization is true}{Przemek} We show data from experiments measured at 15\,cm WISP to RF generator antenna distance.
 The results show the benefit from coalescing for all applications.
 The difference in performance between all strategies is highly application-dependent.
-The gain from energy and task-aware strategy (smart) compared to task-aware only (fast) are particularly small (or negative) for applications with large task sizes, i.e.
+The gain from energy and task-aware strategy (EA, smart) compared to task-aware only (TA, fast) are particularly small (or negative) for applications with large task sizes, i.e.
 \textbf{fft}.
+\todo{Provide more justification}{Przemek}
 
@@
 	\includegraphics[width=\columnwidth]{figures/coalStrategies}
-	\caption{\sys coalescing strategies performance per application on intermittent power compared to \sys without coalescing. \todo{Add new figure; comment on the result}{Przemek}}
+	\caption{\sys coalescing strategies performance per application on intermittent power (RF generator at 15\,cm from the WISP) compared to \sys without coalescing.
+\sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.
+\todo{no coalescing missing; change names of algos; reduce Y width by 40\%; names of apps bold}{Przemek}}
 	\label{fig:coalescing}
@@
 
-\subsection{\sys Programs Characterization}
-\label{sec:results_program_characterization}
-
-\sys program characterization is provide in Table~\ref{table:compiler_result}. \todo{Finalize this section}{Przemek}
-
-\begin{table}
-	\begin{tabular}{| c | c | c | c | c |}
-		\hline
-		\multirow{2}{*}{Application} & \multicolumn{2}{ c |}{Memory footprint} & \multirow{2}{*}{No. tasks} & \multirow{2}{*}{SLOC} \\
-		\cline{2-3}
-		{} & \sys & Alpaca & {} & {} \\
-		\hline\hline
-				\textbf{ar} & --- & --- & --- & ---\\
-		\hline
-				\textbf{bc} & --- & --- & --- & ---\\
-		\hline
-				\textbf{cuckoo} & --- & --- & --- & ---\\
-		\hline
-				\textbf{dijkstra} & --- & --- & --- & ---\\
-		\hline
-				\textbf{fft} & --- & --- & --- & ---\\
-		\hline
-				\textbf{sort} & --- & --- & --- & ---\\
-		\hline
-	\end{tabular}
-		\caption{Comparison between \sys and Alpaca benchmarks.}
-		\label{table:compiler_result}\vspace{-0.5cm}
-\end{table}
+%\subsection{\sys Memory Footprint?}
+%\label{sec:results_program_characterization}
+%
+%\sys program characterization is provide in Table~\ref{table:compiler_result}. \todo{Finalize this section}{Przemek}
+%
+%\begin{table}
+%	\begin{tabular}{| c | c | c | c | c |}
+%		\hline
+%		\multirow{2}{*}{Application} & \multicolumn{2}{ c |}{Memory footprint} & \multirow{2}{*}{No. tasks} & \multirow{2}{*}{SLOC} \\
+%		\cline{2-3}
+%		{} & \sys & Alpaca & {} & {} \\
+%		\hline\hline
+%				\textbf{ar} & --- & --- & --- & ---\\
+%		\hline
+%				\textbf{bc} & --- & --- & --- & ---\\
+%		\hline
+%				\textbf{cuckoo} & --- & --- & --- & ---\\
+%		\hline
+%				\textbf{dijkstra} & --- & --- & --- & ---\\
+%		\hline
+%				\textbf{fft} & --- & --- & --- & ---\\
+%		\hline
+%				\textbf{sort} & --- & --- & --- & ---\\
+%		\hline
+%	\end{tabular}
+%		\caption{Comparison between \sys and Alpaca benchmarks.}
+%		\label{table:compiler_result}\vspace{-0.5cm}
+%\end{table}
 

### 2018-04-16 paper/evaluation.tex

@@
 
+\todo{what is the page size for all apps selected?}{Przemek}
+
 \textbf{Coalescing Strategies Performance.} We first need to assess the performance of all coalescing strategies introduced in Section~\ref{sec:task_coalescing}.
@@
 \todo{Check if this normalization is true}{Przemek} We show data from experiments measured at 15\,cm WISP to RF generator antenna distance.
+
 The results show the benefit from coalescing for all applications.
+Coalescing improves \sys's performance by eliminating the overhead of committing after each task.
 The difference in performance between all strategies is highly application-dependent.
@@
 \todo{Provide more justification}{Przemek}
+
+The \textbf{cuckoo} application (like other top-performers, \textbf{bc} and \textbf{sort}) has small tasks that \sys can easily coalesce to eliminate many commits.
+In contrast, for some applications, e.g.
+\textbf{ar}, coalescing does not improve performance.
+
+%The reason for cem's poor performance is that cem uses a large LZW dictionary structure that spans multiple pages, but has very little {\em locality}.
+The absence of locality means that coalesced tasks do not access the same pages, and as a result, do not amortize commit overhead.
+The lack of locality causes performance to regress to be similar to the non-coalescing case.
 
@@
 \sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.
-\todo{no coalescing missing; change names of algos; reduce Y width by 40\%; names of apps bold}{Przemek}}
+\todo{no coalescing missing; change names of algos; reduce Y width by 40\%; names of apps bold and capitalized}{Przemek}}
 	\label{fig:coalescing}
@@
 
-\textbf{\sys versus Task-based Model.} We now evaluate \sys's run time performance and compare \sys to the run time performance of Alpaca.
-Results are given in Figure~\ref{fig:runtime}.
+\textbf{\sys versus non-adaptive Task-based Model.} We now evaluate \sys's run time performance and compare \sys to the performance of Alpaca runtime, i.e.
+a non-adaptive task-based system.
+Results are presented in Figure~\ref{fig:runtime}.
 For each application we measured its execution time using real wireless power provisioned from the RF signal generator at three distances (refer again to Section~\ref{sec:implementation}).
@@
 The data show the average execution time of each application with runtimes normalized to \sys's average.
-The results show that \sys provides a performance benefit compared to Alpaca for most applications (all except for \textbf{ar} and \textbf{cuckoo}).
-The performance benefit is greatest for applications that have small tasks (\textbf{bc}, \textbf{sort}, \textbf{dijkstra}) at all distances on intermittent power.
-In applications that have larger tasks or access more different memory pages, \sys incurs overhead from memory virtualization that cause its performance to be comparable to (or worse than) Alpaca.
+\todo{verify this statement}{Przemek} The results show that \sys provides a performance benefit compared to Alpaca for most applications (all except for \textbf{ar} and \textbf{cuckoo}).
+The performance benefit is the greatest for applications that perform repeating read and write operations, particularly if involving arrays (\textbf{bc}, \textbf{sort}, \textbf{dijkstra}).
+This effect is seen at all distances on intermittent power.
+In applications with protected variables (or array elements) accessed next to each other but residing in different SRAM or FRAM pages \sys incurs overhead from memory virtualization that cause its performance to be comparable to (or worse than) Alpaca (\textbf{ar}, \textbf{dijkstra}).
+This effect will be investigated in detail in the next section when we will look deep into \sys overhead.
+
+Another critical observation is that \textbf{fft} benchmark did not complete at any distance larger than 0.15\,cm.
+As we will show now, \sys tast downscaling mechanism will be able to solve this issue.
 
@@
 	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
-	\caption{Performance of \sys applications at multiple WISP to RF generator antenna distances ()\{0.15, 0.3, 0.5\}\,m (left, center and right-most bar per application, respectively), compared against Alpaca (results normalized).}
+	\caption{Performance of \sys applications at multiple WISP to RF generator antenna distances \{0.15, 0.3, 0.5\}\,m (left, center and right-most bar per application, respectively), compared against Alpaca (results normalized).
+Results show that \sys performs on average better than Alpaca for majority of the benchmarks.
+For specific benchmarks \sys performs worse than Alpaca \todo{state why it is sometimes worse}{Przemek} \todo{Add non-coalescing result; names of apps bold; reduce Y axis by 40\%; tiny spaces in-between bars for the same app}{Amjad}}
 	\label{fig:runtime}
@@
 
-The data also illustrate that coalescing improves \sys's performance by eliminating the overhead of committing after each task.
-The performance benefit varies, providing up to 4.5x performance improvement.
-The cuckoo application (like other top-performers, bc and sort) has small tasks that \sys can easily coalesce to eliminate many commits.
-In contrast, for some applications, e.g.
-cem, coalescing does not improve performance.
-The reason for cem's poor performance is that cem uses a large LZW dictionary structure that spans multiple pages, but has very little {\em locality}.
-The absence of locality means that coalesced tasks do not access the same pages, and as a result, do not amortize commit overhead.
-The lack of locality causes performance to regress to be similar to the non-coalescing case.
+\textbf{Task Downscaling.} The result is presented in Figure~\ref{fig:task_downscaling} \todo{Provide text; add results}{Przemek}
 
-\textbf{Task Downscaling.} \todo{Provide text; add results}{Przemek}
+\begin{figure}
+	\centering
+	\caption{Task downscaling performance. \todo{Generate taks downscaling figure}{Carlo, Amajda}}
+	\label{fig:task_downscaling}
+\end{figure}
 
@@
 
+\todo{replace slow coalescing with steady coalescing}{Przemek}
+
 \begin{figure}
@@
 	\includegraphics[width=\columnwidth]{figures/overallOverhead}
-	\caption{Overall \sys overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.}
+	\caption{Overall \sys overhead breakdown; NC: no coalescing, SC: slow coalescing, EA: energy-aware coalescing.
+\todo{names of coalescing strategies in capitals; names of apps bold; reduce figure size by 40\%; app -> application; \& -> and; overall overhead breakdown -> overhead breakdown}{Amjad}}
 	\label{fig:overallOverheadBreakdown}
@@
 
+Fixed power or intermittent?
+
 \begin{figure}
@@
 	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
-	\caption{\sys coalescing algorithm efficiency; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing, eta: energy and task-aware coalescing.}
+	\caption{\sys coalescing algorithm efficiency; NC: no coalescing, SC: slow coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.
+\todo{names of coalescing strategies in capitals; names of apps bold; reduce figure size by 40\%; app -> application; \& -> and;  colaescing algorithm efficiency -> overhead breakdown}{Amjad}}
 	\label{fig:coalEfficiency}
@@
 	\includegraphics[width=\columnwidth]{figures/memAccess}
-	\caption{\sys memory access overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.}
+	\caption{\sys memory access overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.
+\todo{names of coalescing strategies in capitals; names of apps bold; reduce figure size by 40\%; Memory access over -> overhead breakdown}{Amjad}}
 	\label{fig:coalmemory}
@@
 	\includegraphics[width=\columnwidth]{figures/pagePulls}
-	\caption{\textbf{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,bytes, per application (top) and respective page pulls (bottom)}).}\vspace{-0.5cm}
-%\todo{change X to 0, remove the definition of X from caption}{Amjad}}
+	\caption{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,bytes, per application (top) and respective page pulls (bottom)).\todo{add page numbers below for each bar for page faults; \# -> No.
+page faults; app names in bold; reduce Y size by 40\%; small space after 128 and 256 in top graph; white background}{Amjad}}
 	\label{fig:page_size}

### 2018-04-16 paper/evaluation.tex
minor changes

@@
 
-%\begin{wrapfigure}{t!}{0.5\textwidth}
-\begin{figure}
+\begin{wrapfigure}{t!}{0.5\textwidth}
+%\begin{figure}
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coalStrategies}
+	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}
 	\caption{\sys coalescing strategies performance per application on intermittent power (RF generator at 15\,cm from the WISP) compared to \sys without coalescing.
-\sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.
-\todo{no coalescing missing; change names of algos; reduce Y width by 40\%; names of apps bold and capitalized}{Przemek}}
+\sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.}
 	\label{fig:coalescing}
-\end{figure}
-%\end{wrapfigure}
+%\end{figure}
+\end{wrapfigure}
 
@@
 Results show that \sys performs on average better than Alpaca for majority of the benchmarks.
-For specific benchmarks \sys performs worse than Alpaca \todo{state why it is sometimes worse}{Przemek} \todo{Add non-coalescing result; names of apps bold; reduce Y axis by 40\%; tiny spaces in-between bars for the same app}{Amjad}}
+For specific benchmarks \sys performs worse than Alpaca \todo{state why it is sometimes worse}{Przemek}}
 	\label{fig:runtime}
@@
 	\includegraphics[width=\columnwidth]{figures/overallOverhead}
-	\caption{Overall \sys overhead breakdown; NC: no coalescing, SC: slow coalescing, EA: energy-aware coalescing.
-\todo{names of coalescing strategies in capitals; names of apps bold; reduce figure size by 40\%; app -> application; \& -> and; overall overhead breakdown -> overhead breakdown}{Amjad}}
+	\caption{Overall \sys overhead breakdown; NC: no coalescing, SC: slow coalescing, EA: energy-aware coalescing.}
 	\label{fig:overallOverheadBreakdown}
@@
 	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
-	\caption{\sys coalescing algorithm efficiency; NC: no coalescing, SC: slow coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.
-\todo{names of coalescing strategies in capitals; names of apps bold; reduce figure size by 40\%; app -> application; \& -> and;  colaescing algorithm efficiency -> overhead breakdown}{Amjad}}
+	\caption{\sys coalescing algorithm efficiency; NC: no coalescing, SC: slow coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.}
 	\label{fig:coalEfficiency}
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/page_exec-time}
-	\includegraphics[width=\columnwidth]{figures/pagePulls}
+	\includegraphics[width=0.49\columnwidth]{figures/page_exec-time}
+	\includegraphics[width=0.49\columnwidth]{figures/pagePulls}
 	\caption{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,bytes, per application (top) and respective page pulls (bottom)).\todo{add page numbers below for each bar for page faults; \# -> No.

### 2018-04-16 paper/evaluation.tex
removed todo from figures

@@
 	\includegraphics[width=\columnwidth]{figures/memAccess}
-	\caption{\sys memory access overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.
-\todo{names of coalescing strategies in capitals; names of apps bold; reduce figure size by 40\%; Memory access over -> overhead breakdown}{Amjad}}
+	\caption{\sys memory access overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.}
 	\label{fig:coalmemory}
@@
 	\includegraphics[width=0.49\columnwidth]{figures/pagePulls}
-	\caption{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,bytes, per application (top) and respective page pulls (bottom)).\todo{add page numbers below for each bar for page faults; \# -> No.
-page faults; app names in bold; reduce Y size by 40\%; small space after 128 and 256 in top graph; white background}{Amjad}}
+	\caption{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,bytes, per application (top) and respective page pulls (bottom).}
 	\label{fig:page_size}

### 2018-04-16 paper/evaluation.tex
Working on evaluation text

@@
-Our evaluation illustrates key findings about \sys. We show that: \todo{Make a list of core findings here once results are solid}{Przemek}
+\begin{wrapfigure}{t!}{0.5\textwidth}
+	%\begin{figure}
+	\centering
+	\includegraphics[width=0.5\columnwidth]{figures/coalEfficiency}
+	\caption{Normalized \sys runtime for all coalescing strategies compared to non-coalesced \sys with overall \sys overhead breakdown; NC: no coalescing, EB: energy-blind coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.
+All \sys coalescing strategies significantly reduce the overhead of (task) commits.
+EA and ETA perform much better than EB a they reduce re-execution further.}
+	\label{fig:overallOverheadBreakdown}
+	%\end{figure}
+\end{wrapfigure}
 
-\subsection{\sys Runtime Performance}
-\label{sec:results_evaluation}
+Our results illustrate key findings about \sys.
+We show through extensive qualitative and quantitative evaluation that (i) \sys \emph{provides computation progress} at the partial task level, guaranteeing execution at the worst intermittent energy conditions where state-of-the-art runtimes fail; (ii) usually \emph{improve execution speed} of applications compared to the state-of-the-art runtimes; (iii) \sys through coalescing mechanism \emph{reduces the memory protection overhead} significantly; and (iv) that \sys is \emph{more generic to use} and allows to interact with code written for non-intermittent devices much easier.
 
-Our aim is to assess \sys as detailed as possible.
+\textbf{Key Performance Indicators.} Our aim is to assess \sys as detailed as possible.
 For this we have characterized \sys runtime performance (execution time), based on a set of benchmarks, considering core task adaptation mechanisms of \sys: \emph{task coalescing} and \emph{task downscaling}.
-We also characterize \sys overhead showing precisely what is the source of \sys superior performance.
-Additionally, we characterize \sys code.
+We also characterize \sys overhead showing precisely what is the source of \sys superior performance investigating task re-execution and commit penalty and memory page access patterns.
 
-\sys is compared to Alpaca runtime.
-We remind the reader that the benchmarks for Alpaca were hand-transformed in a time-consuming manual process from Clang (LLVM-generated code) to GCC (see Section~\ref{sec:implementation}).
+We remind the reader that \sys is compared to Alpaca runtime and the benchmarks for Alpaca were hand-transformed in a time-consuming manual process from Clang (LLVM-generated code) to GCC (see Section~\ref{sec:implementation}).
 The manual code transformation was required as \sys was written with GCC in mind, and Alpaca cannot support any other complier than LLVM.
 
-\subsubsection{\sys Task Adaptation}
+%Then, we need to know how many tasks \sys can coalesce within a given execution scenario.
+We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and the amount of tasks coalesced.
+Result is given in Table~\ref{tab:aveVirtuTaskSize}.
+We see clearly that \sys manages to coalesce more tasks as individual tasks are small.
+As the size of individual task increases, e.g., as in the case of dft application, the number of coalesced tasks is also small.
+This result clearly shows that for the benefit of coalescing and for code portability, initial code should be split by \emph{as small tasks as possible}.
+This will help \sys to find the best possible virtual task size to minimise its runtime.
+
+\subsection{Characterization of \sys Overhead}
+\label{sec:coala_overhead}
+
+We start with the assessment of \sys Overhead for all benchmarks. This will lead us to the selection of the best set of coalescing strategies.
+
+\indent \textbf{\sys Coalescing Algorithm Efficiency.} We have measured the time spent on executing (i) user code, (ii) task re-execution and (iii) task commits for all coalescing strategies (EB, EA, and ETA) and all benchmarks.
+We compared our results to non-coalescing \sys (NC).
+The experiment was performed at 15\,cm WISP to RF generator antenna distance and normalized \todo{to what}{Przemek}.
+The result is presented in Figure~\ref{fig:overallOverheadBreakdown}.
+We clearly see that any \sys coalescing strategy reduces the cost of task commit for \emph{all} benchmarks.
+EA and ETA strategies reduce the re-execution cost of tasks further.
+For this reason we proceed with the investigation of \sys without considering EB coalescing.
+
+\textbf{Overall \sys Overhead Breakdown.} The result is presented in Figure~\ref{fig:coalEfficiency}.
+
+\begin{figure}
+	\centering
+	\includegraphics[width=0.49\columnwidth]{figures/overallOverhead}
+	\includegraphics[width=0.49\columnwidth]{figures/memAccess}
+	\caption{\sys coalescing algorithm efficiency; NC: no coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing. Left: normalized runtime, Right: memory access breakdown.}
+	\label{fig:coalEfficiency}
+\end{figure}
+
+\textbf{\sys Memory Access Overhead Breakdown.} The result is presented in Figure~\ref{fig:coalmemory}.
+
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\columnwidth]{figures/memAccess}
+%	\caption{\sys memory access overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.}
+%	\label{fig:coalmemory}
+%\end{figure}
+
+\subsection{\sys Task Adaptation}
 \label{sec:result_coalescing}
@@
 
-\subsubsection{Characterization of \sys Overhead}
-\label{sec:coala_overhead}
-
-%Then, we need to know how many tasks \sys can coalesce within a given execution scenario.
-We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and the amount of tasks coalesced.
-Result is given in Table~\ref{tab:aveVirtuTaskSize}.
-We see clearly that \sys manages to coalesce more tasks as individual tasks are small.
-As the size of individual task increases, e.g., as in the case of dft application, the number of coalesced tasks is also small.
-This result clearly shows that for the benefit of coalescing and for code portability, initial code should be split by \emph{as small tasks as possible}.
-This will help \sys to find the best possible virtual task size to minimise its runtime.
-
-\indent \textbf{Overall \sys Overhead Breakdown.} The result is presented in Figure~\ref{fig:overallOverheadBreakdown}.
-
-\todo{replace slow coalescing with steady coalescing}{Przemek}
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/overallOverhead}
-	\caption{Overall \sys overhead breakdown; NC: no coalescing, SC: slow coalescing, EA: energy-aware coalescing.}
-	\label{fig:overallOverheadBreakdown}
-\end{figure}
-
-\textbf{\sys Coalescing Algorithm Efficiency.} The result is presented in Figure~\ref{fig:coalEfficiency}.
-
-Fixed power or intermittent?
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
-	\caption{\sys coalescing algorithm efficiency; NC: no coalescing, SC: slow coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.}
-	\label{fig:coalEfficiency}
-\end{figure}
-
-\textbf{\sys Memory Access Overhead Breakdown.} The result is presented in Figure~\ref{fig:coalmemory}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/memAccess}
-	\caption{\sys memory access overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.}
-	\label{fig:coalmemory}
-\end{figure}
-
-\subsubsection{\sys Paging Performance}
+\subsection{\sys Paging Performance}
 \label{sec:results_memory_management}

### 2018-04-16 paper/hwswsetup.tex
Methodology details added in section 9

@@
 %
-Every benchmark listed in the next section was run repeatedly for X\,s on
-each platform and the result was averaged for the duration of the experiment.
-\todo{check if this is correct}{Carlo}
+Every benchmark listed in the next section was run repeatedly on
+each platform for a more or less long time window, depending on the benchmark,
+to capture a diverse power trace and produce a high number of complete runs of all applications.
+The results were averaged across the duration of the experiment.
 
@@
 We powered the WISP using an RF signal generator~\cite{} \todo{Add reference
-to signal generator}{Przemek} emitting 20\,dBm sinusoidal wave at 915\,MHz.
+to signal generator}{Przemek} emitting  a 20-\,dBm sinusoidal wave at 915\,MHz.
 The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
@@
 %
-To obtain execution time, the software toggled GPIO pins at the beginning and
-end of the application, and the Saleae~\cite{saleae} logic analyzer measured
+To obtain execution time, the software toggled GPIO pins at sections of the code
+under profile, and the Saleae~\cite{saleae} logic analyzer measured
 the interval between edges in the signal.
@@
 %
-All applications were compiled using TI GCC~\cite{ti-gcc} version 3.8.0 with
-\texttt{-O1}.
+All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 with
+\texttt{-O1} as optimization flag.
 %
-\todo{Check if this still holds. provide optimization flag}{Przemek} 
-%
-The source code for all benchmarks is released at \cite{coala_website}.
+The source code for all benchmarks will be released at \cite{coala_website}.
 

### 2018-04-16 paper/evaluation.tex
Working on evaluation text

@@
 	\includegraphics[width=0.5\columnwidth]{figures/coalEfficiency}
-	\caption{Normalized \sys runtime for all coalescing strategies compared to non-coalesced \sys with overall \sys overhead breakdown; NC: no coalescing, EB: energy-blind coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.
+	\caption{Normalized \sys time usage for all coalescing strategies compared to non-coalesced \sys with overall \sys overhead breakdown; NC: no coalescing, EB: energy-blind coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.
 All \sys coalescing strategies significantly reduce the overhead of (task) commits.
@@
 
-We start with the assessment of \sys Overhead for all benchmarks. This will lead us to the selection of the best set of coalescing strategies.
+We start with the assessment of \sys Overhead for all benchmarks. This will lead us to the selection of the best set of coalescing strategies to proceed with.
 
@@
 We compared our results to non-coalescing \sys (NC).
-The experiment was performed at 15\,cm WISP to RF generator antenna distance and normalized \todo{to what}{Przemek}.
+The experiment was performed at 15\,cm WISP to RF generator antenna distance and normalized per benchmark.
+All benchmarks were using a 256\,B memory page size except for \textbf{bc} which was using 16\,B page.
+Those page sizes were selected for a large set of potential page sizes and were found to maximize the performance of the respective benchmark.
+Detailed investigation of \sys paging performance will be given in Section~\ref{sec:results_memory_management}.
+
 The result is presented in Figure~\ref{fig:overallOverheadBreakdown}.
@@
 
-\textbf{Overall \sys Overhead Breakdown.} The result is presented in Figure~\ref{fig:coalEfficiency}.
+\textbf{Overall \sys Overhead Breakdown.} We now proceed with investigating how \sys manipulates task commits: task transitions and memory accesses (including breakdown per SRAM and FRAM access).
+The result for all benchmarks are measured in the same way as for the results presented earlier (i.e.
+given in Figure~\ref{fig:overallOverheadBreakdown}) and using the same page sizes as in the earlier experiment.
+The result is presented in Figure~\ref{fig:coalEfficiency}.
 
@@
 	\includegraphics[width=0.49\columnwidth]{figures/memAccess}
-	\caption{\sys coalescing algorithm efficiency; NC: no coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing. Left: normalized runtime, Right: memory access breakdown.}
+	\caption{\sys coalescing algorithm efficiency.
+Left: load split per task manipulation step; right: detailed breakdown of memory access presented in the left figure); NC: no coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.}
 	\label{fig:coalEfficiency}
@@
 
-\textbf{\sys Memory Access Overhead Breakdown.} The result is presented in Figure~\ref{fig:coalmemory}.
+Looking into Figure~\ref{fig:coalEfficiency} (left) we see that no coalescing (NC) \sys has a huge cost of task transitions and commits (which recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown}).
+For both coalescing strategies (EA and ETA) memory access constitutes about 30\% of the whole \sys runtime across all benchmarks.
 
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\columnwidth]{figures/memAccess}
-%	\caption{\sys memory access overhead breakdown; nc: no coalescing, sc: slow coalescing, ea: energy-aware coalescing.}
-%	\label{fig:coalmemory}
-%\end{figure}
+\textbf{\sys Memory Access Overhead Breakdown.} Let us now look into the results on the \sys memory operation process, i.e. we will show specifically which parts of memory from Figure~
+\ref{fig:coalEfficiency} (left) are accessed.
+The result is presented in Figure~\ref{fig:coalEfficiency} (right).
+We immediately see that majority of operations are targeting the last SRAM page and very little operations on the FRAM pages.
+There are very few operations involving other SRAM pages than the last one (only visible for \textbf{cuckoo} and \textbf{dijkstra}).
 
@@
 
-We characterize \sys by studying how its performance varies due to task adaptation. We start with evaluating the task coalescing mechanism.
+We characterize \sys by studying how its execution time varies due to task adaptation. We look at \sys in isolation and in comparison to Alpaca.]
 
-\todo{what is the page size for all apps selected?}{Przemek}
-
-\textbf{Coalescing Strategies Performance.} We first need to assess the performance of all coalescing strategies introduced in Section~\ref{sec:task_coalescing}.
-Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies normalized to the \sys runtime without coalescing.
+\textbf{Coalescing Strategies Performance.}  Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies normalized to the \sys runtime without coalescing.
 \todo{Check if this normalization is true}{Przemek} We show data from experiments measured at 15\,cm WISP to RF generator antenna distance.
@@
 \sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.}
-	\label{fig:coalescing}
+	\label{fig:coalescing} norm per no coalescinhg	
 %\end{figure} 

### 2018-04-16 paper/evaluation.tex
Working on evaluation text

@@
 All \sys coalescing strategies significantly reduce the overhead of (task) commits.
-EA and ETA perform much better than EB a they reduce re-execution further.}
+EA and ETA perform much better than EB a they reduce re-execution further.
+\todo{replace user code with task code; increase font by two more points everywhere}{Amjad}}
 	\label{fig:overallOverheadBreakdown}
@@
 	\caption{\sys coalescing algorithm efficiency.
-Left: load split per task manipulation step; right: detailed breakdown of memory access presented in the left figure); NC: no coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.}
+Left: load split per task manipulation step; right: detailed breakdown of memory access presented in the left figure); NC: no coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.
+\todo{Check if Y axis names are correct; replace task code with user code}{Amjad, Carlo}}
 	\label{fig:coalEfficiency}
@@
 
-We characterize \sys by studying how its execution time varies due to task adaptation. We look at \sys in isolation and in comparison to Alpaca.]
+We characterize \sys by studying how its execution time varies due to task adaptation. We look at \sys in isolation and in comparison to Alpaca.
 
-\textbf{Coalescing Strategies Performance.}  Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies normalized to the \sys runtime without coalescing.
-\todo{Check if this normalization is true}{Przemek} We show data from experiments measured at 15\,cm WISP to RF generator antenna distance.
+\textbf{\sys Coalescing Execution Time.}  Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies for all benchmarks normalized to the \sys runtime without coalescing.
+We show data from experiments measured in the same was as for the results in Section~\ref{sec:coala_overhead}.
 
-The results show the benefit from coalescing for all applications.
+The result in Figure~\ref{fig:coalescing} recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown} showing the benefit in \emph {normalized runtime} from coalescing (for EA and ETA case) for all applications compared to non-coalescing (NC) case.
 Coalescing improves \sys's performance by eliminating the overhead of committing after each task.
-The difference in performance between all strategies is highly application-dependent.
-The gain from energy and task-aware strategy (EA, smart) compared to task-aware only (TA, fast) are particularly small (or negative) for applications with large task sizes, i.e.
-\textbf{fft}.
-\todo{Provide more justification}{Przemek}
-
-The \textbf{cuckoo} application (like other top-performers, \textbf{bc} and \textbf{sort}) has small tasks that \sys can easily coalesce to eliminate many commits.
-In contrast, for some applications, e.g.
-\textbf{ar}, coalescing does not improve performance.
-
-%The reason for cem's poor performance is that cem uses a large LZW dictionary structure that spans multiple pages, but has very little {\em locality}.
-The absence of locality means that coalesced tasks do not access the same pages, and as a result, do not amortize commit overhead.
-The lack of locality causes performance to regress to be similar to the non-coalescing case.
+The difference in performance between coalescing and no coalescing is, however, highly application-dependent.
+Then, the gain from energy and task-aware strategy (ETA) compared to energy-aware only (EA) are almost non-existent for applications, except for \textbf{fft}, i.e.
+the one with with large tasks.
+Furthermore, the \textbf{cuckoo} application (like other top-performers, \textbf{bc} and \textbf{sort}) has small tasks that \sys can easily coalesce to eliminate many commits.
 
@@
 	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}
-	\caption{\sys coalescing strategies performance per application on intermittent power (RF generator at 15\,cm from the WISP) compared to \sys without coalescing.
-\sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.}
-	\label{fig:coalescing} norm per no coalescinhg	
+	\caption{\sys coalescing strategies (EA and ETA) performance per application on intermittent power (RF generator at 15\,cm from the WISP) compared to \sys without coalescing (NC).
+\sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.
+\todo{NC, EA, ETA labels below bars}{Amjad}}
+	\label{fig:coalescing}
 %\end{figure} 

### 2018-04-16 paper/background.tex

@@
 
+\subsubsection{Costs of Task-based Models}
+\label{sec:cost_task-based}
+
+Despite their advantages, the task-based programming models have their limitations.
+Two major ones being: {\em memory consistency enforcement overhead} and {\em inflexibility to dynamic energy harvesting conditions}.
+
 \begin{table}
@@
 
-\subsubsection{Costs of Task-based Models}
-\label{sec:cost_task-based}
-
-Despite their advantages, the task-based programming models have their limitations. Two major ones being: {\em memory overhead} and {\em inflexibility to dynamic energy harvesting conditions}.
-
 \textbf{Memory Consistency Enforcement Overhead.} Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
@@
 Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.
-
-\begin{figure}
-\begin{subfigure}[t]{.49\columnwidth}
-	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
-	\caption{Cost of \emph{write} operation}
-\end{subfigure}%
-\begin{subfigure}[t]{.49\columnwidth}
-	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
-	\caption{Cost of \emph{read} operation}
-\end{subfigure}
-	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
-\end{figure}
-
-To understand these non-volatile memory overheads, for the MSP430FR5969~\cite{msp430datasheet} MCU we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations.
+% We do not need to emphasize the following figure! (Sinan)
+%\begin{figure}
+%\begin{subfigure}[t]{.49\columnwidth}
+%	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
+%	\caption{Cost of \emph{write} operation}
+%\end{subfigure}%
+%\begin{subfigure}[t]{.49\columnwidth}
+%	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
+%	\caption{Cost of \emph{read} operation}
+%\end{subfigure}
+%	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
+%\end{figure}
+However, a non-volatile memory access consumes around 1.5 times the energy of a volatile access~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
+%To understand these non-volatile memory overheads, for the MSP430FR5969~\cite{msp430datasheet} MCU we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations.
 We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device across 1600 read and write accesses to each memory type, randomizing to avoid caching.
@@
 This observation is also consistent with the TI specification~\cite{}.
-\todo{add citation to TI specification on FRAM energy cost}{Przemek} In an intermittent device, energy cost corresponds to decreased run time between power failures.
-The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.
+\todo{add citation to TI specification on FRAM energy cost}{Przemek}
+One of the primary motivation of \sys is to promote accessing SRAM more frequently rather than FRAM in order to reduce memory access energy overhead and  to increase the time between power failures.
+
+%The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.
 

### 2018-04-16 paper/background.tex
section 2 done

@@
 Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value.
-The situation occurs when code includes a WAR dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
+The situation occurs when code includes a \emph{write-after-read} (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
 %Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
@@
 However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions.
-
 Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity.
@@
 \sys's execution model based on {\em task coalescing}---merging or dissecting multiple tasks {\em during} execution based on the program execution behavior, and {\em task downscaling}---shortening task boundaries during execution when even the smallest task does not fit the energy buffer, both models described in Section~\ref{sec:task_adaptation}, is motivated by these observations.
+The main enabler for task coalescing is \emph{memory virtualization} that resolves dynamic WAR dependencies, which is described in Section~\ref{sec:memory_virtulaization}.
 

### 2018-04-16 paper/evaluation.tex
Working on evaluation text

@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
+	\caption{Performance of \sys applications at multiple WISP to RF generator antenna distances \{0.15, 0.3, 0.5\}\,m (left, center and right-most bar per application, respectively), compared against Alpaca (results normalized) for ETA coalescing strategy.
+Results show that \sys performs on average better than Alpaca for majority of the benchmarks.
+For specific benchmarks \sys performs worse than Alpaca \todo{state why it is sometimes worse}{Carlo}}
+	\label{fig:runtime}
+\end{figure}
+
 The data show the average execution time of each application with runtimes normalized to \sys's average.
@@
 Another critical observation is that \textbf{fft} benchmark did not complete at any distance larger than 0.15\,cm.
-As we will show now, \sys tast downscaling mechanism will be able to solve this issue.
+As we will show now, \sys task downscaling mechanism will be able to solve this issue.
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
-	\caption{Performance of \sys applications at multiple WISP to RF generator antenna distances \{0.15, 0.3, 0.5\}\,m (left, center and right-most bar per application, respectively), compared against Alpaca (results normalized).
-Results show that \sys performs on average better than Alpaca for majority of the benchmarks.
-For specific benchmarks \sys performs worse than Alpaca \todo{state why it is sometimes worse}{Przemek}}
-	\label{fig:runtime}
-\end{figure}
-
-\textbf{Task Downscaling.} The result is presented in Figure~\ref{fig:task_downscaling} \todo{Provide text; add results}{Przemek}
-
-\begin{figure}
-	\centering
-	\caption{Task downscaling performance. \todo{Generate taks downscaling figure}{Carlo, Amajda}}
-	\label{fig:task_downscaling}
-\end{figure}
+\textbf{Task Downscaling Execution Time.} We have enabled task downscaling and measured the execution time for \textbf{fft} benchmark at all distances that this benchmark could not execute (refer again to Figure~\ref{fig:coalescing} and $\infty$ marks).
+\sys could execute the benchmark successfully with the execution time of X, Y and Z for distances $d=\{0.3, 0.5.
+0.6\}$\,cm, respectively.
 

### 2018-04-16 paper/background.tex
Section 2

@@
 The situation occurs when code includes a \emph{write-after-read} (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
-%Figure~\ref{fig:code_demo_incosistency} illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
+%Figure~\ref{fig:code_demo_incosistency} illgit statuustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
 The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
@@
 %\end{figure}
-However, a non-volatile memory access consumes around 1.5 times the energy of a volatile access~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
-%To understand these non-volatile memory overheads, for the MSP430FR5969~\cite{msp430datasheet} MCU we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations.
+To understand the non-volatile memory access overheads, we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations for the MSP430FR5969~\cite{msp430datasheet} MCU.
 We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device across 1600 read and write accesses to each memory type, randomizing to avoid caching.
-Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device.
+Our experiment showed that a non-volatile memory access consumes around 1.5 times the energy of a volatile access.
+%~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
+% Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device.
 This observation is also consistent with the TI specification~\cite{}.
 \todo{add citation to TI specification on FRAM energy cost}{Przemek}
-One of the primary motivation of \sys is to promote accessing SRAM more frequently rather than FRAM in order to reduce memory access energy overhead and  to increase the time between power failures.
+Based on this observation, one of the primary motivation of \sys is to promote accessing SRAM more frequently rather than FRAM in order to reduce memory access energy overhead and to increase the time between power failures.
 

### 2018-04-16 paper/intro.tex
little edits to intro

@@
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/graffle/intro-figure.pdf}
-    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits.
+    \includegraphics[width=0.5\textwidth]{figures/intro-figure.pdf}
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \textt{cmt} in the figure).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
@@
 %
-Figure~\ref{fig:coalesce} (top) illustrates how a program's tasks execute and
+Figure~\ref{fig:coalesce} illustrates how a program's tasks execute and
 shows how tasks can impose a run-time overhead.
-%
-At each task transition, the system incurs an overhead to track and atomically
-\emph{commit} modifications to the non-volatile memory, to maintain consistency of
-program state~\citep{chain,alpaca}.  
-%
+At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\citep{chain,alpaca}.
 The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.
@@
 \item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
-\item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication\footnote{The \sys repository will be accessible via~\cite{coala_website} immediately after publication.}.
+\item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication via~\cite{coala_website}. 
 \item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to diverse energy conditions.

### 2018-04-16 paper/intro.tex
little edits to intro and abstarct

@@
     \includegraphics[width=0.5\textwidth]{figures/intro-figure.pdf}
-    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \textt{cmt} in the figure).
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}

### 2018-04-16 paper/implementation.tex
Sections 5 and 6 revised

@@
 
-\subsection{Kernel Functions Implementation}
+\subsection{Kernel Functions}
 
@@
 
-\noindent \textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
+\textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
 A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}.
@@
 
-\subsection{Paging Implementation}
+\subsection{Paging}
 
-\sys's memory virtualization is implemented using three buffers: (i) working (volatile) buffer, (ii) store (non-volatile) buffer and (iii) shadow (non-volatile) buffer.
-The store buffer holds the latest safe copy of all the pages, and it is read-only up to, and including, the ultimate commit.
-Upon reboot, the working buffer is empty, and it gets filled with pages from non-volatile memory on demand.
+As described in Section~\ref{sec:memory_virtulaization}, \sys's memory virtualization is implemented using three buffers: \texttt{working} (volatile) buffer, \texttt{\underline{store}} (non-volatile) buffer and \texttt{\underline{shadow}} (non-volatile) buffer.
+The \texttt{\underline{store}} buffer holds the latest safe copy of all the pages, and it is read-only up to, and including, the ultimate commit.
+Upon reboot, the \texttt{working} buffer is empty, and it gets filled with pages from non-volatile memory on demand.
 All reads and writes to protected variables happen in volatile memory.
-The shadow buffer serves as an intermediate non-volatile buffer where dirty, volatile pages are copied during the intermediate commit.
+The \texttt{\underline{shadow}} buffer serves as an intermediate non-volatile buffer where dirty, volatile pages are copied during the intermediate commit.
 Moreover, it contains pages evicted from SRAM when a \emph{full page fault} occurs.
 
-\textbf{Page Allocation and Page Tags.} When a \sys application accesses a protected variable with either \texttt{RP} or \texttt{WP}, \sys has to search for the variable in RAM first, and then, if not found, in non-volatile memory.
+When a \sys application accesses a protected variable with either \texttt{RP} or \texttt{WP}, \sys has to search for the variable in RAM first, and then, if not found, in non-volatile memory.
 In reality, the page manager searches for pages, not single variables.
@@
 Note that, by aligning pages by $2^n$, the alignment is also valid for all sizes $2^x$, where $x \leq n$.
+For our experiments, we used 32-, 64-, 128- and 256-byte pages.
+16\,KB were allocated in non-volatile memory for \texttt{\underline{shadow}} and \texttt{\underline{store}} (8\,KB each), and the volatile page buffer was 1\,KB, half the total SRAM size.
+The main advantage of having large pages is searches through page buffers are shorter.
+On the other hand, page swaps are more bulky.
 
@@
 \end{figure}
-
-\textbf{Page Fault and Page Swap.} When a page is not found in RAM, a \emph{page fault} occurs, and it has to be fetched from FRAM.
-In case that page was in the working buffer and it had been evicted, its most recent non-volatile copy resides in the shadow buffer.
-For this reason, \sys searches this buffer first.
-If the requested page is not found in the shadow buffer, then it will be pulled from its store location.
-To make room for the requested page in SRAM, a volatile page might have to be evicted.
-If the victim page is dirty, we say that a \emph{full page fault} occurs, and the victim page has to be copied to the shadow buffer.
-Page moves are accelerated by the MCU's direct memory access (DMA) module.
 

### 2018-04-16 paper/newCoalescing.tex
describe the energy-aware coalecing algo
update the ea algor

--- 2018-04-15 paper/newCoalescing.tex
@@
+Energy-aware task coalescing strategies gain \sys efficiency, and the flexible time-based task splitting strategy enables \sys to progress where its counterparts fail.
+
 \subsection{Task Coalescing}
@@
 
-When incoming energy is sufficient to run multiple consecutive tasks,
+\noindent When incoming energy is sufficient to run multiple consecutive tasks,
 committing state after each task is wasteful. \sys mitigates this
@@
 single virtual task---a process we call \emph{coalescing}.
-%
-When $N$ tasks are coalesced, $N-1$ boundaries are crossed and some number of
+
+\noindent When $N$ tasks are coalesced, $N-1$ boundaries are crossed and some number of
 pages are dirtied. The first $N-2$ executions of \transition statement skip
 committing dirty pages to non-volatile memory, leaving the work of committing
-pages modified by any of the $N$ tasks to the statement $N-1$.
-Coalescing reduces the total number of non-volatile memory
-accesses, because when coalesced tasks access the same pages of data, only the
-last version of the page needs to be committed.  
+pages modified by any of the $N$ tasks to the statement $N-1$.Therefore, coalescing static tasks is a twofold benefit: (i) it reduces the total number of non-volatile memory
+accesses when variables are accessed many times within a coalesced task and committed only conly at the coalesced task boundery; (ii) it fasten applicaiton exeution be skipping unnecessary data commits operations when there is abandond energy.
+
 \textcolor{red}{maybe is possible to demonstrate with data? count number of executed non-volatile vs volatile accesses}
 %
-\textbf{Trade-off between Speed and Wasted Effort.} \sys can coalesce an
+
+\noindent \textbf{Trade-off between Speed and Wasted Effort.} \sys can coalesce an
 arbitrary number of consecutive tasks. However, as more tasks coalesce, their
@@
 will restart from the last commit, i.e. the first task in the sequence, losing
-the progress made by any of the coalesced tasks.  The challenge to coalescing
-tasks is determining how many tasks to coalesce before committing.
+the progress made by any of the coalesced tasks.
+Therefore, a good coalescing strategy must moderate the risk of wasted work and capitalize on benefits of deferred commits by adapting the target size of the dynamic task
+to the energy available at runtime. 
 %
-%During execution, an executing task is either a {\em coalesced} task, which
-%does not commit its state before transitioning or a {\em committing} task,
-%which does commit its state before transitioning.
 
-\textbf{Task Coalescing Strategies.} 
+\noindent \textbf{Task Coalescing Strategies.} 
+% % Carlo pseudo code 
+% \begin{algorithm}[t]
+% 	\caption{Coalescing}
+% 	\label{algo:genCoalescing}
+% 	\scriptsize
+% 	%\small
+% 	\begin{algorithmic}[1]
+%         \State $H_\text{p} \gets H_\text{c}$
+%         \State $H_\text{p} \gets 0$ 
+%         \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$B, H_\text{p}$}
+
+%         \While{$\texttt{true}$}
+% 	        \State $ B_\text{c} \gets B$
+% 	        %
+% 	        \While{$ B_\text{c} > 0$}
+% 		        \State \Call{execute\_task}{$T_\texttt{i}$}
+% 		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
+% 		        \State $B_\text{c} \gets B_\text{c} - W$
+% 				\State $H_\text{c} \gets H_\text{c} + W$
+% 	        \EndWhile
+% 	        %
+% 	         \State $B \leftarrow $ \Call{$f_\text{compl}$}{B, $H_\text{c}$,$H_\text{p}$}
+% 	        \State \Call{commit\_to\_fram}{\null}
+%         \EndWhile
+% 	\end{algorithmic}
+% \end{algorithm}
+% %
+% To describe the proposed strategies we will use the following parameters:
+% \begin{itemize}
+% \item target budget $B$: total number of static tasks to be coalesced.
+% \item current budget $B_\text{c}$: remaining number of tasks to be coalesced before committing to non-volatile memory; 
+% \item history $H$: up on reboot, it is the number of executed static tasks between the last two power interrupts.
+While executing, it is the number of executed static tasks since the last power interrupt.
+% \end{itemize}
 %
+%
+% Modifiying Carlo's pseudo code a little 
 \begin{algorithm}[t]
@@
 	\begin{algorithmic}[1]
-        \State $H_\text{p} \gets H_\text{c}$
-        \State $H_\text{p} \gets 0$ 
-        \State $H_\text{B} \leftarrow $ \Call{$f_\text{reboot}$}{B, $H_\text{p}$}
-
+        \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$H$}
+        \State $H \gets 0$ 
         \While{$\texttt{true}$}       
-	        \State $ B_\text{c} \gets B$
-	        \While{$ B_\text{c} > 0$}
+	        \State $ i \gets 0$
+	        %
+	        \While{$ i <= B$}
 		        \State \Call{execute\_task}{$T_\texttt{i}$}
-		        % \State $W$ \gets \Call{f_\text{weight}}{$T_\texttt{i}$}
-		        \State $B_\text{c} \gets B_\text{c} - W$
-				\State $H_\text{c} \gets H_\text{c} + W$
+		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
+		        \State $i \gets i + W$
+				\State $H \gets H + W$
 	        \EndWhile
-	        % \State $B$ \gets \Call{f\_compl}{$\texttt{B},\texttt{H_\text{c}},\texttt{H_\text{p}}$}
-	         \State $B \leftarrow $ \Call{$f_\text{reboot}$}{B, $H_\text{c}$,$H_\text{p}$}
-	        \State \Call{commit\_to\_fram}{}
+	        %
+	        \State \Call{commit\_to\_fram}{\null}
+	        \State $B \leftarrow $ \Call{$f_\text{coalesc\_task\_maker}$}{$B$}
         \EndWhile
@@
 %
-The likelihood that a task group will complete depends on the total amount of
-energy consumed by all tasks in the group, i.e. the size of the dynamic task.
-%
-A good coalescing strategy must moderate the risk of wasted work and capitalize
-on benefits of deferred commits by adapting the target size of the dynamic task
-to the energy available at runtime. 
-%
-To describe the proposed strategies we will use the following parameters:
-\begin{itemize}
-\item target budget $B$: total number of static tasks to be coalesced.
-\item current budget $B_\text{c}$: remaining number of tasks to be coalesced before committing to non-volatile memory; 
-\item history $H$: up on reboot, it is the number of executed static tasks between the last two power interrupts.
-While executing, it is the number of executed static tasks since the last power interrupt.
-\end{itemize}
-
-%% Coalescing motivation
-
-% Finding an optimal task size, given random energy conditions is an open question.
-Obviously, a static approach does not have the potential to answer it.
-Therefore, \sys champions runtime-based methods to approximate the ideal task size under random energy shots arrival.
-Furthermore, \sys advocates making intermittently powered software energy-aware is the key for \emph{efficient code execution and probability}.
-
-% %% address the challenge and define the optimal task
-
-% \sys advances its execution in a state-less (or virtual) manner, and then it frequently saves its forward progress.
-The longer \sys virtually progresses, the less committing (saving data to non-volatile memory) overhead it introduces.
-However, long state-less execution results in a considerable re-execution penalty---all the tasks that have been virtually executed must be re-executed after a power interrupt.
-As such, an optimal task is a task that occupies with a single commit an entire power cycle, which is therefore necessarily of a varying length.
-
-% %% How does Coala approximates an ideal task
-
-The general structure of the proposed coalescing strategies is summaries in algorithm~\ref{algo:genCoalescing}, where where $H_\text{p}$, $H_\text{c}$, $B$ and $B_\text{c}$ represent past history, current history, target budget and current budget, respectively.
-$T_\text{i}$ is the i-th task being executed and $W$ is its weight.
-A coalescing strategy and its performance are defined by the functions: \\
-(i) \emph{$f_\text{reboot}$}: which updates the target budget after a reboot; (ii) \emph{$f_\text{weight}$}: which returns the weight of the task just executed and its unit; (iii) \emph{$f_\text{compl}$}: which updates the target budget after a successful completion of a dynamic coalesced task.
+The general structure of the proposed coalescing strategies is summaries in algorithm~\ref{algo:genCoalescing}, where, $B$ (target budget) is total number of static tasks to be coalesced; $T_\text{i}$ is the i-th task being executed; $W_\text{i}$ is weight of $T_\text{i}$; and $H$ (history) is the number of executed static tasks between the last two power interrupts, on a reboot, and it is the number of executed static tasks since the last power interrupt, while executing.
+Furthermore, we use the following function templates to characterize the proposed coalescing algorithms: $f_\text{reboot}$ which updates the target budget after a reboot.
+$f_\text{weight}$ which returns the weight of the task and its unit.
+$f_\text{coalesc\_task\_maker}$ which specifies the size of the coalesced task size.
 The following sections detail the different strategies \sys proposes and the needs that originated from one strategy leading to the conception of another one.
 
-\subsubsection{Energy-blind Coalescing}
+\subsubsection{Energy-blind Coalescing (EB)}
 \label{subsec:energyBlind}
@@
 	\caption{The {\em energy blind coalescing} algorithm \emph{linearly} updates its coalescing target.}
-	\label{fig:slowCoal}
+	\label{fig:energyBlind}
 \end{wrapfigure}
@@
 The need for \emph{adaptiveness}, that is coalescing static tasks, can be fulfilled by a naive strategy, having the target budget $B$ to increase by $x$ static tasks after a successful completion of a coalesced task, and to decrease by the same number of tasks upon a reboot.
-In this case: $f_\text{reboot} = B - x$; $f_\text{weight} =  1$; $f_\text{compl} = B + x$; \\
-Due to the linear behavior of the target adaptation, this algorithm is slow in reacting to the changes in energy conditions; the energy required by a coalesced task or the available energy.
-In particular, this algorithm can experience a large number of \emph{repeated power} failures without forward progress.
+In this case:
+\begin{itemize}
+\item $f_\text{reboot}(B) = B - x $;
+\item $f_\text{weight}(T_\text{i}) =  1$; 
+\item $f_\text{coalesc\_task\_maker}(B) = B + x$; 
+\end{itemize}
+Since this algorithm relies on an \em{energy blind counting process} that enlarges the coalesced task size by $x$ static tasks on a successful completion, and reduces its size by $x$ on a reboot.
+this algorithm is slow in reacting to the changes in energy conditions; the energy required by a coalesced task or the available energy.
+In particular, this algorithm can experience a large number of \emph{repeated power} failures without forward progress, (see figure~\ref{fig:energyBlind}).
+In other words, if a coalesced task size is enlarged ten times, due to ample amount of energy or coalescing small static tasks, and then the algorithm can only progress by a single static task, due to a change in energy condition, the devise has to reboot for ten times to have an adequate coalesced task size.
+Moreover, This algorithm tend to experience a large re-execution penalty due to its unawareness of the energy conditions and it optimistic behavior.
 
-\subsubsection{Energy-aware Coalescing}
+\subsubsection{Energy-aware Coalescing (EA)}
 \label{subsec:ECoalescing}
+The Energy-aware coalescing algorithm overcomes the limitations of the previous algorithm, with out losing generality or requiring hardware support, by using the recent execution history as a metric---pure software technique---to estimate the available energy and to set the size of the coalescing task accordingly.
+Moreover, it takes advantage of the guarantees that the energy buffer offers on a fresh start by coalescing many static tasks.
+Then it approaches the expected power failure carefully by reduces the task target size by x\%, amongst 25\%, 50\% and 75\% the 50\% shows the best performance and therefore it is the default of \sys.
+if twice the number of the estimated tasks has been executed in the same power discharging cycle the algorithm concludes a wrong power failure time estimation has happened and enforce target re-tuning based on its new history.
+The EA algorithm is characterized by the following functions:
+
+\begin{itemize}
+\item $f_\text{reboot}(B) = floor(H / 2);$
+\item $f_\text{weight}(T_\text{i}) =  W_\text{i}$; 
+\item $f_\text{coalesc\_task\_maker}(B) = floor(H / 2);$ 
+\end{itemize}
+
+\begin{wrapfigure}{}{0.5\textwidth}
+	\centering
+	\includegraphics[width=0.5\columnwidth]{figures/energy-aware-coal.pdf}
+	\caption{placeholder}
+	\label{fig:energyAware}
+\end{wrapfigure}

### 2018-04-16 paper/background.tex

@@
 	\end{subfigure}
-	\caption{\textbf{Code example demonstrating checkpointing gives rise to data inconsistency.}}\label{fig:code_demo_incosistency}
+	\caption{Code example demonstrating checkpointing gives rise to data inconsistency.}\label{fig:code_demo_incosistency}
 \end{wrapfigure}
@@
 	\end{tabular}
-	\caption{\textbf{Memory consistency enforcement overheads of intermittent execution models;} \emph{PC}: program counter, \emph{channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
+	\caption{Memory consistency enforcement overheads of intermittent execution models; \emph{PC}: program counter, \emph{channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
 	\label{table:chechpoint_comparison}

### 2018-04-16 paper/evaluation.tex

@@
 	\caption{\sys coalescing strategies (EA and ETA) performance per application on intermittent power (RF generator at 15\,cm from the WISP) compared to \sys without coalescing (NC).
-\sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.
-\todo{NC, EA, ETA labels below bars}{Amjad}}
+\sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.}
 	\label{fig:coalescing}
@@
 
-\begin{figure}
+\begin{wrapfigure}{t!}{0.5\textwidth}
+%\begin{figure}
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
+	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
 	\caption{Performance of \sys applications at multiple WISP to RF generator antenna distances \{0.15, 0.3, 0.5\}\,m (left, center and right-most bar per application, respectively), compared against Alpaca (results normalized) for ETA coalescing strategy.
@@
 	\label{fig:runtime}
-\end{figure}
+%\end{figure}
+\end{wrapfigure}
 
@@
 In applications with protected variables (or array elements) accessed next to each other but residing in different SRAM or FRAM pages \sys incurs overhead from memory virtualization that cause its performance to be comparable to (or worse than) Alpaca (\textbf{ar}, \textbf{dijkstra}).
-This effect will be investigated in detail in the next section when we will look deep into \sys overhead.
 
@@
 \textbf{Task Downscaling Execution Time.} We have enabled task downscaling and measured the execution time for \textbf{fft} benchmark at all distances that this benchmark could not execute (refer again to Figure~\ref{fig:coalescing} and $\infty$ marks).
-\sys could execute the benchmark successfully with the execution time of X, Y and Z for distances $d=\{0.3, 0.5.
-0.6\}$\,cm, respectively.
+\sys could execute the benchmark successfully with the execution time of X\,s at a distances of $d= 0.6$\,m, i.e.
+the one that exceeds the furthest distance ($d=0.5$\,m) presented in Figure~\ref{fig:runtime}, respectively.
 
@@
 
-We evaluated the effect that paging with different page sizes has on \sys's performance and we show data in Figure~\ref{fig:page_size}; The top plot shows the run time performance normalized to the best per-application performance, using pages of different sizes in \sys, indicated with the labels at each bar; the bottom plot shows the number of page faults for the same set of executions normalized to the maximum of each set.
-We measured performance as MCU clock cycles using values read out by TI's CCS IDE.
-The results show a clear ``bathtub curve'' in the performance of each application, as the number of page faults varies.
+We have also evaluated the effect that paging with different page sizes has on \sys's performance.
+We show data in Figure~\ref{fig:page_size}; The left plot shows the run time performance normalized to the best per-application performance, using pages of different sizes in \sys; the right plot shows the number of page faults for the same set of executions normalized to the maximum of each set.
+We measured performance as MCU clock cycles using values read out by TI's Code Composer Studio IDE.
+
+\textbf{Effect of Page Size on \sys Runtime.} The results (Figure~\ref{fig:page_size}, left) show a ``bathtub curve'' in the performance of each application (the effect would be more profound with a larger set of pages than we considered), as the number of page faults varies.
 The data show that across application there is a page size that minimizes \emph{both} execution time.
-We observe that the page size is not the same for each application, although 64 byte pages perform well for all applications.
-The increased rate of page faults is responsible for the overhead with small page sizes; smaller pages require accessing more different pages, leading to higher paging costs.
+We observe that the page size is not the same for each application, although 128\,B pages perform well for all applications.
+
+\textbf{Page Size versus Page Faults in \sys Runtime.} The increased rate of page faults (Figure~\ref{fig:page_size}, right) is responsible for the overhead with small page sizes; smaller pages require accessing more different pages, leading to higher paging costs.
 With larger page sizes, the overhead is higher than with moderate pages, too.
@@
 Even if an application accesses few memory locations, \sys pages data at full page size, degrading performance by sometimes paging in and out unnecessary data.
-The data reveal a reasonable default page size of 64 bytes and when developing an application, a developer should consider varying the page size to moderate poor performance.
+The data reveal a reasonable default page size of 128\,B and when developing an application, a programmer should consider varying the page size to moderate poor performance.
 
@@
 	\includegraphics[width=0.49\columnwidth]{figures/pagePulls}
-	\caption{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,bytes, per application (top) and respective page pulls (bottom).}
+	\caption{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,B (left) and respective page pulls (bottom) for all benchmarks.}
 	\label{fig:page_size}
@@
 %\label{sec:results_program_characterization}
-%
-%\sys program characterization is provide in Table~\ref{table:compiler_result}. \todo{Finalize this section}{Przemek}
 %

### 2018-04-16 paper/paging.tex

@@
 
-Paging moves data between non-volatile and volatile memory. A page has a home
-location in non-volatile memory which we refer to as the page's {\tt store}.
+Paging moves data between non-volatile and volatile memory.
+A page has a home location in non-volatile memory which we refer to as the page's \texttt{\underline{store}} (underlined locations are non-volatile).
 As a task accesses pages, they are swapped from their home location
-in the direct\hyp{}mapped non-volatile memory, called
-\texttt{\underline{store}} (underlined locations are non-volatile) 
-into the fully\hyp{}associative \texttt{working} page buffer in volatile memory. The
-\texttt{working} buffer is allocated to take up the volatile memory that
-remains after space for stack is allocated.
+into the fully\hyp{}associative \texttt{working} page buffer in volatile memory.
+The \texttt{working} buffer is allocated to take up the volatile memory that remains after space for stack is allocated.
 %
-\sys uses two-phase commit to write dirty pages from the \texttt{working} buffer
-back to their home location in \texttt{\underline{store}} atomically,
-ensuring data consistency even with power failures.
+\sys uses a two-phase commit to write dirty pages from the \texttt{working} buffer back to their home location in \texttt{\underline{store}} atomically, ensuring data consistency even with power failures.
 %
-When any dirty page is swapped out from the \texttt{working} buffer due to
-eviction or task transition from last coalesced task, the first phase writes
-the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
-buffer, and the second phase writes it from
-\texttt{\underline{shadow}} to \texttt{\underline{store}}.
+When any dirty page is swapped out from the \texttt{working} buffer due to eviction or task transition from last coalesced task, the first phase writes the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer, and the second phase writes it from \texttt{\underline{shadow}} to \texttt{\underline{store}}.
 %
-Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a
-time between volatile and non-volatile memory, \sys's page-based
-design moves contiguous blocks of data.
+Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a time between volatile and non-volatile memory, \sys's page-based design moves contiguous blocks of data.
 %
-\sys takes full advantage of hardware\hyp{}accelerated block copies using
-Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x
-fewer cycles on MSP430FRxxxx.
+\sys takes full advantage of hardware\hyp{}accelerated block copies using Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x fewer cycles on MSP430FRxxxx.
 
@@
 
-A \sys task must access protected non-volatile variables through a restricted memory
-interface. The interface includes \texttt{RVAR(var)}, which reads the value of
-variable {\tt var} and \texttt{WVAR} \texttt{(var,val)}, which assigns
-value {\tt val} to variable {\tt var}. The implementation of {\tt
-RVAR} is shown in Algorithm~\ref{algo:rwar} and {\tt WVAR} implementation is
-similar except for an assignment instead of the return.
+A \sys task must access protected non-volatile variables through a restricted memory interface, as will be described in Section~\ref{sec:coala_api}.
+The interface includes \texttt{RP(var)}, to read the value of variable {\tt var}, and \texttt{WP(var)} to assign a value to variable {\tt var}.
+The implementation of {\tt RP} is shown in Algorithm~\ref{algo:rwar}, and {\tt WP}'s implementation is similar except for marking the accessed page as dirty.
 %
-{\tt RVAR} and {\tt WVAR} operations translate a variable's physical address in non-volatile memory into a \emph{virtual address} in the page buffer in volatile memory.
+{\tt RP} and {\tt WP} operations translate a variable's physical address in non-volatile memory into a \emph{virtual address} in the page buffer in volatile memory.
 In \sys, a virtual address is composed of a \emph{page tag} that identifies the page and a \emph{page offset} that identifies a byte.
@@
 \begin{algorithm}[t]
-	\caption{\texttt{RVAR}(variable $v$)}
+	\caption{\texttt{RP}(variable $v$)}
 	\label{algo:rwar}
@@
 
-After address translation, \sys attempts to access the protected variable's
-location in the volatile paging buffer. \sys keeps track of the page tags for
-the pages currently resident in the paging buffer. When a task accesses a
-variable, it compares the variable's page tag to tags of the pages in {\tt
-working} (Line 2).
+After address translation, \sys attempts to access the protected variable's location in the volatile paging buffer.
+\sys keeps track of the page tags for the pages currently resident in the paging buffer.
+When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt working} (Line 2).
 %
@@
 %
-If the accessed variable's page tag is not found in the page buffer, the
-operation incurs a {\em page fault} (Line 4). At a page fault, \sys must swap
-out one of the pages that is resident in the page buffer (to the non-volatile
-commit buffer, \texttt{\underline{shadow}}) and swap in the variable's page.
+If the accessed variable's page tag is not found in the page buffer, the operation incurs a {\em page fault} (Line 4).
+At a page fault, \sys must swap out one of the pages that is resident in the page buffer (to the non-volatile \texttt{\underline{shadow}}) if \texttt{working} is full, and swap in the variable's page.
 %
-The byte is accessed in the page buffer at the index of the resident page and
-the variable's page offset (Lines 5-6).
+The byte is accessed in the page buffer at the index of the resident page and the variable's page offset (Lines 5-6).
 
@@
         \State \LeftComment{$i_\text{victim}$: volatile index to next free slot or page to be evicted, set to 0 at boot}
-        \State \LeftComment{\texttt{\underline{numShadowed}}: non-volatile, reset to 0 on boot unless in phase 2 commit}
+        \State \LeftComment{\texttt{\underline{shadowCount}}: non-volatile, reset to 0 on boot unless in phase 2 commit}
         \State $t_\text{victim} \gets \Call{Tag}{\texttt{working}[i_\text{victim}]}$
-		\If {\Call{IsDirty}{working[$i_\text{victim}$]} }	\Comment{Are we evicting a dirty page?}
+		\If {\Call{isDirty}{working[$i_\text{victim}$]} }	\Comment{Are we evicting a dirty page?}
 		\State $\texttt{\underline{shadow}}[t_\text{victim}] \xleftarrow{\text{DMA}} \texttt{working}[i_\text{victim}]$
@@
         \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track as list for quick traversal}
-            \State $\texttt{\underline{shadowList}}[\texttt{numShadowed++}] \gets t_\text{victim}$
+            \State $\texttt{\underline{shadowList}}[\texttt{shadowCount++}] \gets t_\text{victim}$
         \EndIf
 		\EndIf
-		\If {\Call{IsValid}{\underline{shadow}[$t$]}} \Comment{Had the page been swapped out earlier?}
+		\If {\Call{exists}{\underline{shadow}[$t$]}} \Comment{Had the page been swapped out earlier?}
 		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
@@
 		\EndIf 
-        \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBuf}|$ \Comment{FIFO eviction policy}
+        \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBufLength}|$ \Comment{FIFO eviction policy}
 	\end{algorithmic}
@@
 
-A page fault on a memory access requires \sys to swap out one of the pages in
-the page buffer, i.e. \emph{victim page}, preserving updates made to that page,
-and to copy the accessed page of protected data into the page buffer.
-Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault. A victim
-page in the \texttt{working} buffer is selected for eviction according to
-\emph{first-in-first-out} replacement policy (Lines 2 and 12).  If the page
-being swapped out is not dirty and the page being accessed had not been dirtied
-and swapped out to the \texttt{shadow} buffer since last boot, then
-handling a page fault consists of copying the accessed page from
-\texttt{\underline{store}} into the \texttt{working} buffer (Line 11).
+If the \texttt{working} buffer is full, a page fault on a memory access requires \sys to swap out one of the pages in the volatile buffer, the \emph{victim page}, preserving updates made to that page, and to copy the accessed page of protected data into the page buffer.
+Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault.
+A victim page in the \texttt{working} buffer is selected for eviction according to \emph{first-in-first-out} replacement policy (Lines 2 and 12).
+If the page being swapped out is not dirty and the page being accessed had not been dirtied and swapped out to the \texttt{shadow} buffer since last boot, then handling a page fault consists of copying the accessed page from \texttt{\underline{store}} into the \texttt{working} buffer (Line 11).
 Otherwise, the page fault handler must take the following extra steps.
-
-If a task modified any byte in the victim page
-(i.e.,  using \texttt{WVAR}), then the page is dirty. Before swapping in the
-accessed page, \sys must first preserve the updated values in the dirty page so
-that they can be committed back to their locations in memory
-when the dynamic task ends (but not earlier). To preserve dirty values, \sys
-copies the victim page to the non-volatile \texttt{\underline{shadow}}
-buffer (Line 5) and records the page's tag in a non-volatile list (Lines 6-7)
-to support iteration over shadow pages having traversing all slots in the
-buffer, necessary in the second phase of the commit.
+%
+If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty.
+Before swapping in the accessed page, \sys must first preserve the updated values in the dirty page so that they can be committed back to their locations in memory when the dynamic task ends (but not earlier).
+To preserve dirty values, \sys copies the victim page to the non-volatile \texttt{\underline{shadow}} buffer (Line 5) and records the page's tag in a non-volatile list (Lines 6-7) to support iteration over shadow pages having traversing all slots in the buffer, necessary in the second phase of the commit.
 %
@@
 %
-If the accessed page was previously modified and
-swapped out since the last boot (Line 8), the most recent contents of the page is in {\tt
+If the accessed page was previously modified and swapped out since the last boot (Line 8), the most recent version of the page is in {\tt
 \underline{shadow}}, not \texttt{\underline{store}} (Line 9).
@@
 
-When the last task in a coalesced group completes, \sys must commit \emph{all}
-dirty pages in the \texttt{working} buffer and the swapped out pages in
-\texttt{\underline{shadow}} buffer back to their original locations in
-\texttt{\underline{store}}. If a task is allowed to (re-)execute after only
-\emph{some} pages have been copied to \texttt{\underline{store}}, that task
-might see protected variables in an inconsistent state.
+When the last task in a coalesced group completes, \sys must commit \emph{all} dirty pages in the \texttt{working} buffer and the swapped out pages in \texttt{\underline{shadow}} buffer back to their original locations in \texttt{\underline{store}}.
+If a task is allowed to (re-)execute after only \emph{some} pages have been committed to \texttt{\underline{store}}, that task might see protected variables in an inconsistent state.
 
-To make the commit atomic, \sys copies the pages in two phases, implemented by
-Algorithm~\ref{algo:commit}.
+To make the commit atomic, \sys commits the pages in two phases, implemented by Algorithm~\ref{algo:commit}.
 %
-The first phase copies dirty pages in \texttt{working}
-buffer to the non-volatile \texttt{\underline{shadow}} buffer (Line 4). 
-The second phase, copies pages from \texttt{\underline{shadow}} to \texttt{store} (Line 12).
-If power fails during the first phase, the whole
-commit is aborted, and the dynamic coalesced task restarts from its first
-static task. If power fails in the second phase, the commit will be safely
-resumed on next boot.
-The non-volatile metadata for the second phase is the \texttt{\underline{committing}}
-bit, which is set before the first page is copied (Line 9) and cleared after the last
-page is copied (Line 14), and \texttt{\underline{numShadowed}} counter, which
-indexes into the \texttt{\underline{shadowList}} (Line 11), is
-decremented after each page is copied (Line 13), and cleared once the phase
-completes (Line 18).
+The first phase copies dirty pages from the \texttt{working} buffer to the non-volatile \texttt{\underline{shadow}} buffer (Line 4).
+The second phase commits pages from \texttt{\underline{shadow}} to \texttt{\underline{store}} (Line 12).
+If power fails during the first phase, the whole commit is aborted, and the dynamic coalesced task restarts from its first static task.
+If power fails in the second phase, the commit will be safely resumed on next boot.
+The non-volatile metadata for the second phase are: the \texttt{\underline{committing}} bit, which is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16); \texttt{\underline{shadowCount}} counter, which is cleared once the phase completes (Line 14); \texttt{\underline{commitIndex}}, which indexes the next page to be committed (Line 11) and is cleared at the end of the phase (Line 15).
 
-\sys non-volatile \texttt{\underline{shadowList/numShadowed}} implementation is
-a more efficient alternative to a simpler implementation that keeps a valid bit
-in each entry of \texttt{\underline{shadow}} and, on every boot into second
-phase, traverses all entries, copying the pages with a set bit, and
-clearing the bit.
-%
-\sys assumes that the write to a non-volatile word in memory is atomic.
+\sys non-volatile \texttt{\underline{shadowList/shadowCount/commitIndex}} implementation is a more efficient alternative to a simpler implementation that keeps a valid bit in each entry of \texttt{\underline{shadow}} and, on every boot into second phase, traverses all entries, committing the pages with a set bit, and clearing the bit.
+During the second commit phase, pages are not physically copied from \texttt{\underline{shadow}} to \texttt{\underline{store}}, but a table pointing at the right buffer for each page is updated.
+This is explained in more details in Section~\ref{sec:implementation}.
 
@@
                 \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
-                \State $\texttt{\underline{shadowList}}[\texttt{\underline{numShadowed}}] \gets t$
-                \State $\texttt{\underline{numShadowed}} \gets \texttt{\underline{numShadowed}} + 1$
+                \State $\texttt{\underline{shadowList}}[\texttt{\underline{shadowCount}}] \gets t$
+                \State $\texttt{\underline{shadowCount}} \gets \texttt{\underline{shadowCount}} + 1$
             \EndFor
@@
             \State $\texttt{\underline{committing}} \gets \textbf{true}$
-            \While{\texttt{\underline{numShadowed}} > 0}
-                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{numShadowed}} - 1]$
-                \State $\texttt{\underline{store}}[t] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
-                \State $\texttt{\underline{shadow}}[t] \gets \textbf{nil}$ \Comment{Clear entry for page fault handler}
-                \State $\texttt{\underline{numShadowed}} \gets \texttt{\underline{numShadowed}} - 1 $
+            \While{\texttt{\underline{commitIndex}} < \texttt{\underline{shadowCount}}} \Comment{\texttt{\underline{commitIndex}} set to 0 at first boot}
+                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{commitIndex}}]$
+                \State \Call{commitToStore}{\texttt{\underline{shadow}}[t]}
+                \State $\texttt{\underline{commitIndex}} \gets \texttt{\underline{commitIndex}} + 1$
             \EndWhile
+            \State $\texttt{\underline{shadowCount}} \gets 0$
+            \State $\texttt{\underline{commitIndex}} \gets 0$
             \State $\texttt{\underline{committing}} \gets \textbf{false}$
@@
             \EndIf 
-            \State $\texttt{\underline{numShadowed}} \gets 0$, $i_\text{victim} \gets 0$
+            \State $\texttt{\underline{shadowCount}} \gets 0$, $i_\text{victim} \gets 0$
         \EndProcedure
@@
 
-\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent
-protected data. During task execution, modifications to protected data do not
-affect their original memory locations, because a task reads and writes the
-volatile \texttt{working} buffer only, and modified pages are kept in
-\texttt{\underline{shadow}} until commit. A power failure erases the contents
-of the page buffer, preventing a re-executing task from observing updates in
-the page buffer from a previous execution attempt. Clearing \texttt{\underline{shadow}}
-entries as part of the second phase commit (Line 13), ensures that all accesses to
-protected variables in subsequent tasks correctly access their original memory
-locations in \texttt{\underline{store}}.
+\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent protected data.
+During task execution, modifications to protected data do not affect their original memory locations, because a task reads and writes the volatile \texttt{working} buffer only, and modified pages are kept in \texttt{\underline{shadow}} until commit.
+A power failure erases the contents of the page buffer, preventing a re-executing task from observing updates in the page buffer from a previous execution attempt.
+Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14), ensures that all accesses to protected variables in subsequent tasks correctly access their original memory locations in \texttt{\underline{store}}.
+\sys assumes that the write to a non-volatile word in memory is atomic.
 %

### 2018-04-16 paper/newCoalescing.tex
update the ea algo

@@
 \end{itemize}
-Since this algorithm relies on an \em{energy blind counting process} that enlarges the coalesced task size by $x$ static tasks on a successful completion, and reduces its size by $x$ on a reboot.
+Since this algorithm relies on an \emph{energy blind counting process} that enlarges the coalesced task size by $x$ static tasks on a successful completion, and reduces its size by $x$ on a reboot.
 this algorithm is slow in reacting to the changes in energy conditions; the energy required by a coalesced task or the available energy.

### 2018-04-16 paper/systemdescription.tex
system descripton updates

@@
 
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/graffle/overview.pdf}
-	\caption{\sys top-level view. \textcolor{red}{redraw the figure: change font type}}
-	\label{fig:system_overview}
-\end{wrapfigure}
-
 \textbf{\sys Programming and Execution Model.}  To use \sys, a programmer first writes plain, imperative C code.
@@
 In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.
+
+\begin{wrapfigure}{t!}{0.5\textwidth}
+	\centering
+	\includegraphics[width=0.5\columnwidth]{figures/graffle/overview.pdf}
+	\caption{\sys top-level view. \sys is composed of two core components: \emph{Task Scheduler} and \emph{Memory Manager}.\todo{Redraw figure to map new terms}{Carlo}}
+	\label{fig:system_overview}
+\end{wrapfigure}
 
@@
 The key insight that \sys leverages is that the first task could have deferred its commit to the second task.
-{\em coalescing} tasks by deferring the first commit avoids the fixed cost of the first commit (task transitioning overheads).
+{\em Coalescing} tasks by deferring the first commit avoids the fixed cost of the first commit (task transitioning overheads).
 Coalescing also amortizes per-variable commit cost, committing each location accessed by both tasks only after the second task, rather than once per task.
-\textcolor{red}{\sys's coalescing manager is an \emph{energy-aware manager}.
-It uses the recent execution history---pure software technique---as a metric to estimate the amount of the available energy and to set the adequate coalesced task size accordingly.
+
+\sys's coalescing manager is an \emph{energy-aware} and \emph{software-based only} manager in a sense it uses the recent execution history as a metric to estimate the amount of the available energy to set the adequate coalesced task size accordingly.
 This fundamental feature prevents \sys from blind coalesced task adaptation which may result in a repeated power failure problem.
-For example, If \sys enlarges the coalesced task size by $x$ static tasks on a coalesced task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, 10 times to reduce a coalesced task size from 10 to one, when $x=1$}.
-Details of coalescing are presented in Section~\ref{sec:task_coalescing}.
+For example, If \sys enlarges the coalesced task size by $x$ static tasks on a coalesced task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, ten times to reduce a coalesced task size from ten to one, when $x=1$.
+Details of coalescing will be given in Section~\ref{sec:task_coalescing}.
 
-\textbf{\sys Task Downscaling Manager.} \textcolor{red}{\sys's downscaling manager is responsible for preserving the forward progress when \sys repeatedly fails on a single static task.
+\textbf{\sys Task Downscaling Manager.} \sys's task downscaling manager is responsible for preserving the forward progress when \sys repeatedly fails on a single static task.
 The downscaling manager utilizes a timer-based technique and relies on the virtual memory manager to gain a full \emph{flexibility} on where to split a task.
@@
 The \sys task downscaling manager, on the other hand, triggers a \emph{partial commit} faster than the previous one on a repeated failure to guarantee forward progress.
-Details of task downscaling are presented in Section~\ref{sec:task_downsizing}.}
+Details of task downscaling will be given in Section~\ref{sec:task_downsizing}.
 
-\textbf{\sys Virtual Memory Manager.} \sys is able to efficiently coalesce tasks because of its efficient virtual memory manager, which is described in Section~\ref{sec:memory_virtualization}.
+\textbf{\sys Virtual Memory Manager.} \sys is able to efficiently coalesce tasks because of its efficient virtual memory manager.
 \sys's memory manager paginates memory and ensures that data in a page remain consistent despite power interruptions.
@@
 Pages efficiently, atomically commit using a two-phase commit procedure accelerated using hardware support for direct memory access (DMA).
-Details of Virtual Memory Manager are presented in Section~\ref{sec:memory_virtulaization}.
+Details of Virtual Memory Manager are will be given in Section~\ref{sec:memory_virtulaization}.

### 2018-04-16 paper/relatedwork.tex
related

--- 2017-11-14 paper/relatedwork.tex
@@
 
-\textbf{Intermittent Execution Environments.} Early work on energy-harvesting runtime systems assumed simple, small computations not exceeding a predictable burst of energy~\cite{dewdrop} or long power-cycles/priority-based scheduling~\cite{sorber_sensys_2007}.
+\textbf{Checkpointing-Based Execution Environments.} Early work on energy-harvesting runtime systems assumed simple, small computations not exceeding a predictable burst of energy~\cite{dewdrop} or long power-cycles/priority-based scheduling~\cite{sorber_sensys_2007}.
 These efforts did not consider data consistency with intermittence.
-Works supporting intermittent computing using dynamic, on-demand checkpoints include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall}, Hibernus++~\cite{hibernusplusplus} and HarvOS~\cite{mottola2017harvos}.
-Dynamic checkpointing systems may checkpoint
-at any point, making it difficult to implement application-level atomicity
-requirements.
-Moreover, some require hardware to measure stored energy, incurring energy cost.
-Moreover, these systems do not checkpoint non-volatile state, compromising memory consistency.
-DINO~\cite{dino} selectively versions non-volatile memory and checkpoints volatile state, ensuring progress and memory consistency.
+Works supporting intermittent computing using dynamic, on-demand checkpoints include Mementos~\cite{mementos}, DINO~\cite{dino}, Quickrecall~\cite{quickrecall}, Hibernus++~\cite{hibernusplusplus}, Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} and HarvOS~\cite{mottola2017harvos}.
+Dynamic checkpointing systems may checkpoint at any point, making it difficult to implement application-level atomicity requirements.
+Among aforementioned prior work, Mementos does not checkpoint non-volatile state---compromising memory consistency.
+To overcome this, DINO selectively versions non-volatile memory and checkpoints the volatile state, ensuring both progress and memory consistency.
+However, checkpointing, in general is not scalable since its time and energy cost grows with the size of the volatile memory---giving possibility to exceed device's energy budget.
+QuickRecall, Clank and Ratchet propose a relatively lightweight checkpointing since they can only be
+applied to devices with  non-volatile memory only and they require leveraging special hardware.
+
+% Moreover, some require hardware to measure stored energy, incurring energy cost.
+The major drawback of these systems is that they do not checkpoint non-volatile state, compromising memory consistency.
+In order to overcome this, DINO~\cite{dino} selectively versions non-volatile memory and checkpoints volatile state, ensuring progress and memory consistency.
 Rachet~\cite{ratchet} assumes all memory is non-volatile and checkpoints volatile state.
-Clang~\cite{hicks_isca_2017} takes a Ratchet-like approach leveraging special hardware, eliminating checkpointing and versioning overhead.
+Clank~\cite{hicks_isca_2017} takes a Ratchet-like approach leveraging special hardware, reducing checkpointing and versioning overhead.
 
-Recent task-based systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}.
+\textbf{Task-Based Execution Environments.} Recent task-based systems are Chain~\cite{chain}, Alpaca~\cite{alpaca} and Mayfly~\cite{hester_sensys_2017}.
 Using static tasks, they eliminate checkpointing volatile state.
-Using channel-based memory models~\cite{chain} or automatic privatization and redo-logging~\cite{alpaca} they avoid checkpointing overheads.
+Using channel-based memory models~\cite{chain,hester_sensys_2017} or automatic privatization and redo-logging~\cite{alpaca} they avoid checkpointing overheads.
 Moreover, task-based models facilitate specifying application-level atomicity properties.
-
 \sys relates to the above efforts but is distinct in approach and mechanism.
 \sys relates closely to task-based models because \sys also relies on statically defined tasks to avoid checkpointing volatile state.
-\sys's automatic compilation spares the programmer's work of writing task-based code that prior systems demand.
+%\sys's automatic compilation spares the programmer's work of writing task-based code that prior systems demand. 
 \sys's memory virtualization provides an alternative for ensuring memory consistency, optimized for bulk accesses to task-shared data with high locality.
 \sys's task coalescing optimization ameliorates the key overheads associated with prior systems.
+
+\textbf{Optimal Task Division.}One approach to optimize the task division, without assuming a minimum level of
+incoming power always available, is to predict task energy consumption and
+choose a static decomposition specific to the device capacitor size~\cite{cleancut_2018}.
+Another approach is to enumerate different versions of a program with different task sizes ranging from conservatively small to confidently large by using compiler support and selecting the best among them ~\cite{baghsorkhi_cgo_2018}.
+However, statically predicting energy of arbitrary input-dependent code with peripheral access is a problem without a general solution.
+Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.
+
+%\textcolor{red}{TODO: make this one paragraph with the CleanCut stuff.  add \cite{intel} work; separate by-programmer and by-compiler;
+%both cases hit the energy estimation problem; also intel work proposes a method
+%for decomposing into one of many decompositions, but does not give a good
+%solution for choosing one}
 

### 2018-04-16 paper/newCoalescing.tex

@@
 In particular, this algorithm can experience a large number of \emph{repeated power} failures without forward progress, (see figure~\ref{fig:energyBlind}).
-In other words, if a coalesced task size is enlarged ten times, due to ample amount of energy or coalescing small static tasks, and then the algorithm can only progress by a single static task, due to a change in energy condition, the devise has to reboot for ten times to have an adequate coalesced task size.
-Moreover, This algorithm tend to experience a large re-execution penalty due to its unawareness of the energy conditions and it optimistic behavior.
+In other words, if a coalesced task size is enlarged ten times, due to ample amount of energy or coalescing small static tasks, and then the algorithm can only progress by a single static task, due to a change in energy condition, the device has to reboot for ten times to have an adequate coalesced task size.
+Moreover, This algorithm tends to experience a large re-execution penalty due to its unawareness of the energy conditions and it optimistic behavior.
 
 \subsubsection{Energy-aware Coalescing (EA)}
-\label{subsec:ECoalescing}
+\label{subsec:energyAware}
 The Energy-aware coalescing algorithm overcomes the limitations of the previous algorithm, with out losing generality or requiring hardware support, by using the recent execution history as a metric---pure software technique---to estimate the available energy and to set the size of the coalescing task accordingly.
@@
 \end{wrapfigure}
+
+\subsubsection{Energy- and Task-aware Coalescing (EA)}
+\label{subsec:energyTaskAware}

### 2018-04-16 paper/paging.tex
paging

@@
 the application as a virtual address space partitioned into pages.
-Paged virtual memory is a key enabler of task coalescing, which
-we described in Section~\ref{sec:task_coalescing}. Virtualization simplifies
-coalescing, because the system gets the freedom to move application's data
-between volatile and non-volatile memory transparently.
+\emph{Paged virtual memory} is a key enabler of task coalescing since it keeps the non-volatile memory consistent despite dynamic WAR dependencies incurred by coalesced code regions.
+Thanks to memory virtualization, \sys have the freedom to move application's data between volatile and non-volatile memory transparently.
 %
-\sys manages memory such that application tasks operate primarily from
-efficient volatile SRAM rather than non-volatile FRAM, which is more costly to access than SRAM,
-as we showed in Figure~\ref{fig:framEnergy}. \sys runtime implements memory
+%,. Virtualization simplifies
+%coalescing, because the system gets the freedom to move application's data
+%between volatile and non-volatile memory transparently.
+%
+To this end, \sys isolates FRAM from the tasks of the application so that they interface only SRAM---which is cheaper to read/write as we indicated in Section~\ref{sec:background}.
+%
+%manages memory such that application tasks operate primarily from
+%efficient volatile SRAM rather than non-volatile FRAM, which is more costly to access than SRAM,
+%as we showed in Figure~\ref{fig:framEnergy}. 
+\sys runtime implements memory
 virtualization by moving pages of the address space from the large,
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
+	\caption{Illustration of task-virtual memory interaction in \sys.
+Tasks only interact with the working buffer in SRAM (indicated as page buffer).
+The shadow buffer is used to hold the dirty bages.
+The indirection table is used to speed-up page two-phase commit operation: it eliminates the copy operation in the second phase.}
+	\label{figure:coala_page_tags}
+\end{figure}
+
+\textbf{System Overview.} The overall picture of the paging operations is illustrated in Figure~\ref{figure:coala_page_tags}.
 Paging moves data between non-volatile and volatile memory.

### 2018-04-16 paper/newCoalescing.tex

@@
 \item $f_\text{reboot}(B) = floor(H / 2);$
-\item $f_\text{weight}(T_\text{i}) =  W_\text{i}$; 
+\item $f_\text{weight}(T_\text{i}) =  1$; 
 \item $f_\text{coalesc\_task\_maker}(B) = floor(H / 2);$ 
@@
 
-\subsubsection{Energy- and Task-aware Coalescing (EA)}
+\subsubsection{Energy- and Task-aware Coalescing (ETA)}
 \label{subsec:energyTaskAware}
+There are two factors that affect the accuracy of predicting a power interrupt.
+The stability of the incoming energy and the static tasks sizes distribution.
+If the task sizes are not uniformly distributed then the algorithm estimation might be off by a relatively large margin, for example, if there are two tasks of size $1s$ and $10s$ then EA will count them as two and might estimate that the target should be of size two static tasks.
+However, if the following two tasks, on one hand, are of size $1s$ then the algorithm will waste energy on unnecessary commits.
+On the other hand, if the following two tasks were of size $10s$ then the algorithm might fail, on the second task, and lose progress of $+10$ tasks.
+To overcome this limitation the counting process of ETA algorithm takes tasks' weight into consideration to approximate a uniform task size distribution, from the counting process perspective.
+As such ETA algorithm adapts the characterization of the EA algorithm by changing the weight function to be, $f_\text{weight}(T_\text{i}) =  W_\text{i}$;

### 2018-04-16 paper/paging.tex
paging
motivating paging

@@
+\textcolor{red}{[additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain.
+And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
+
 \sys exposes non-volatile memory for \emph{protected} variables to

### 2018-04-16 paper/abstract.tex
little edits to intro and abstarct
add some additional info to abstract

@@
 %Intermittently powered computers promise a sustainable Internet of things by forgo the batteries and scarf the energy from the environment. 
-Massively deployed Internet of Things (IoT) must be energy neutral. This can only be achieved when the IoT is batteryless and the energy to operate the IoT comes from ambient energy sources.
+Massively deployed Internet of Things (IoT) must be energy neutral. This can only be achieved when the IoT is battery-free and the energy to operate the IoT comes from ambient energy sources.
 % The main problem that we are tackling
@@
 Generally, intermittent computation models can be classified into (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based programming, where a programmer splits the code into small idempotent sections.
-Despite the superiority of task-based systems over checkpointing, expressed in lower memory overhead, task-based programming is completely oblivious to the energy conditions.
+Despite the superiority of task-based systems over checkpointing, task-based programming is completely oblivious to the energy conditions.
 Simply, it cannot increase or decrease the size of the task on the fly (the moment of commit to the non-volatile memory) depending on how long the energy burst is.
@@
 \sys defaults to progress on a multi-tasks scale, when energy permits (\emph{efficiency}), and on a sub-task scale when needed (\emph{guarantee}).
+Furthermore, \sys's \emph{memory virtualization} gains \sys full data protection against runtime triggered and compile-time defined memory accesses.
 % I really hope it will support our claims!! haha
-Results show that \sys is more efficient than the current state-of-art execution model by \textcolor{red}{X times}.
+Results show that \sys is more efficient than the current state-of-art execution model by up to a factor of two, and able to progress where its counterparts fail.
 

### 2018-04-16 paper/abstract.tex
add some additional info to abstract

@@
 % show off
-\sys programs tune frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period off the programmer's shoulders.
+\sys programs tune the frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period off the programmer's shoulders.
 % main benefits of Coala
@@
 % I really hope it will support our claims!! haha
-Results show that \sys is more efficient than the current state-of-art execution model by up to a factor of two, and able to progress where its counterparts fail.
-
-% old PLDI abstract 
-%Computation consistency on intermittently-powered devices can be ensured by: (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) via tasks, where a programmer splits the code into small idempotent sections.
-Results so far suggest that a task-based approach performs better than checkpointing.
-However, existing tasks-based approaches do not adapt to changing energy levels and storage.
-This makes the code inefficient (when executed on a larger energy buffer than intended) or can potentially completely disable the program (in the opposite case).
-Therefore, we present \sys: a task-based execution runtime, enabling code portability by means of a new concept of task coalescing.
-If a code is written for a very small energy buffer, \sys is able to dynamically coalesce these tasks and constructs a bigger (virtual) task to improve execution times up to 4.5 times compared to static task code.
+Results show that \sys is more efficient than the current state-of-art execution model by up to a factor of two and able to progress where its counterparts fail.

### 2018-04-16 paper/intro.tex
intro is revised

@@
 and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute  
-with no commit or task transition overhead between them, instead performing
+with no commit overhead between them, instead performing
 task-end commit actions at the end of the second task only.
@@
 {\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
-illustrates how dynamic coalescing changes the earlier execution and highlights
-the change in memory model necessary to not break atomicity.
-On the other hand, a large task that requires more energy to complete than a device's fixed energy buffer is split by \sys to enable task termination, as depicted in Figure~\ref{fig:coalesce} (bottom).
+illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution. 
+On the other hand, a large task that requires more energy to complete than a device buffered energy 
+is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).
 

### 2018-04-16 paper/evaluation.tex
Eval revised and enhanced

@@
 	\includegraphics[width=0.5\columnwidth]{figures/coalEfficiency}
-	\caption{Normalized \sys time usage for all coalescing strategies compared to non-coalesced \sys with overall \sys overhead breakdown; NC: no coalescing, EB: energy-blind coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.
+	\caption{Percent time usage for all coalescing strategies (including no-coalescing) comparing useful execution to intermittency overhead.
+NC: no coalescing, EB: energy-blind coalescing, EA: energy-aware coalescing, ETA: energy- and task-aware coalescing.
 All \sys coalescing strategies significantly reduce the overhead of (task) commits.
-EA and ETA perform much better than EB a they reduce re-execution further.
-\todo{replace user code with task code; increase font by two more points everywhere}{Amjad}}
+EA and ETA perform much better than EB since they reduce re-execution penalty.
+\todo{rename legend entry 'user code' with 'useful task code'; rename y-axis to 'Time usage (\%)'; increase font by two more points everywhere}{Amjad}}
 	\label{fig:overallOverheadBreakdown}
@@
 Our results illustrate key findings about \sys.
-We show through extensive qualitative and quantitative evaluation that (i) \sys \emph{provides computation progress} at the partial task level, guaranteeing execution at the worst intermittent energy conditions where state-of-the-art runtimes fail; (ii) usually \emph{improve execution speed} of applications compared to the state-of-the-art runtimes; (iii) \sys through coalescing mechanism \emph{reduces the memory protection overhead} significantly; and (iv) that \sys is \emph{more generic to use} and allows to interact with code written for non-intermittent devices much easier.
+We show through extensive qualitative and quantitative evaluation that (i) \sys \emph{provides computation progress} at the partial task level, guaranteeing execution at the worst intermittent energy conditions where state-of-the-art runtimes fail; (ii) usually \emph{improves execution speed} of applications compared to state-of-the-art runtimes; (iii) \sys, through coalescing mechanism, \emph{reduces the memory protection overhead} significantly; and (iv) that \sys is \emph{more generic to use} and allows to interact with code written for non-intermittent devices more easily.
 
 \textbf{Key Performance Indicators.} Our aim is to assess \sys as detailed as possible.
-For this we have characterized \sys runtime performance (execution time), based on a set of benchmarks, considering core task adaptation mechanisms of \sys: \emph{task coalescing} and \emph{task downscaling}.
-We also characterize \sys overhead showing precisely what is the source of \sys superior performance investigating task re-execution and commit penalty and memory page access patterns.
+For this we have characterized \sys's runtime performance (execution time), based on a set of benchmarks, considering core task adaptation mechanisms of the system: \emph{task coalescing} and \emph{task downscaling}.
+We also characterize \sys's overhead showing precisely what the source of \sys superior performance is, investigating task re-execution and commit penalty and memory access patterns.
 
@@
 
-\subsection{Characterization of \sys Overhead}
+\subsection{Characterization of \sys's Overhead}
 \label{sec:coala_overhead}
 
-We start with the assessment of \sys Overhead for all benchmarks. This will lead us to the selection of the best set of coalescing strategies to proceed with.
+We start with the assessment of \sys's overhead for all benchmarks. This will lead us to the selection of the best coalescing strategy to proceed with.
 
-\indent \textbf{\sys Coalescing Algorithm Efficiency.} We have measured the time spent on executing (i) user code, (ii) task re-execution and (iii) task commits for all coalescing strategies (EB, EA, and ETA) and all benchmarks.
+\indent \textbf{Coalescing Algorithms Efficiency.} We have measured the time spent on executing (i) useful task code, (ii) wasted task code (that have to be re-executed due to uncommitted progress), and (iii) commits for all coalescing strategies (EB, EA and ETA) and all benchmarks.
 We compared our results to non-coalescing \sys (NC).
-The experiment was performed at 15\,cm WISP to RF generator antenna distance and normalized per benchmark.
-All benchmarks were using a 256\,B memory page size except for \textbf{bc} which was using 16\,B page.
-Those page sizes were selected for a large set of potential page sizes and were found to maximize the performance of the respective benchmark.
-Detailed investigation of \sys paging performance will be given in Section~\ref{sec:results_memory_management}.
-
+The experiment was performed at 15\,cm WISP-to-exciter distance and normalized per benchmark.
+All benchmarks were using 256-B pages, except for \textbf{bc} which was using 32-B pages.
+Those page sizes are the ones yielding the best performance, selected among the set mentioned in Section~\ref{sec:implementation}.
+Detailed investigation of \sys's paging performance will be given in Section~\ref{sec:results_memory_management}.
+%
 The result is presented in Figure~\ref{fig:overallOverheadBreakdown}.
-We clearly see that any \sys coalescing strategy reduces the cost of task commit for \emph{all} benchmarks.
-EA and ETA strategies reduce the re-execution cost of tasks further.
-For this reason we proceed with the investigation of \sys without considering EB coalescing.
-
-\textbf{Overall \sys Overhead Breakdown.} We now proceed with investigating how \sys manipulates task commits: task transitions and memory accesses (including breakdown per SRAM and FRAM access).
-The result for all benchmarks are measured in the same way as for the results presented earlier (i.e.
-given in Figure~\ref{fig:overallOverheadBreakdown}) and using the same page sizes as in the earlier experiment.
-The result is presented in Figure~\ref{fig:coalEfficiency}.
+We clearly see that all coalescing strategies make a better use of the resources for \emph{all} benchmarks (except for EB in \textbf{bc}).
+With a more thorough look, one can notice that the re-execution penalty, i.e.
+the time wasted running tasks whose progress could not be committed due to a power failure, is minimized by NC.
+This is obvious, since this strategy makes the system commit after each static task.
+On the other hand, the burden on committing so frequently reflects in a degraded performance with respect to coalescing, motivating one of the needs raised by this work.
+EA and ETA strategies provide the best balance between re-execution penalty and commit cost. For this reason we proceed with the investigation of \sys without considering EB coalescing.
 
@@
 	\includegraphics[width=0.49\columnwidth]{figures/memAccess}
-	\caption{\sys coalescing algorithm efficiency.
-Left: load split per task manipulation step; right: detailed breakdown of memory access presented in the left figure); NC: no coalescing, EA: energy-aware coalescing, ETA: energy and task-aware coalescing.
-\todo{Check if Y axis names are correct; replace task code with user code}{Amjad, Carlo}}
+	\caption{Kernel overhead.
+Left: high-level load distribution; right: detailed breakdown of memory accesses overhead from the left figure).
+NC: no coalescing, EA: energy-aware coalescing, ETA: energy- and task-aware coalescing.
+\todo{rename legend entry 'task code' to 'user code'; rename y-axis to 'Load (\%)'}{Amjad}}
 	\label{fig:coalEfficiency}
@@
 
-Looking into Figure~\ref{fig:coalEfficiency} (left) we see that no coalescing (NC) \sys has a huge cost of task transitions and commits (which recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown}).
-For both coalescing strategies (EA and ETA) memory access constitutes about 30\% of the whole \sys runtime across all benchmarks.
+\textbf{Kernel Overhead Breakdown.} We now proceed with investigating how much utilization \sys's kernel steals to user code, quantifying task transitions and memory accesses overhead.
+For a broader picture, we profile kernel load for NC as well.
+The result for all benchmarks are measured in the same way as for the results presented above (i.e.
+given in Figure~\ref{fig:overallOverheadBreakdown}) and using the same page sizes as in the previous experiment.
+The outcome is presented in Figure~\ref{fig:coalEfficiency} (left side).
+%
+As expected, non-coalescing (NC) \sys introduces a huge cost of task transitions and commits (which recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown}).
+For both coalescing strategies (EA and ETA) protected variables' accesses constitutes about 30\% of the whole load across all benchmarks.
+In fact, dynamic address translation is the most critical bottleneck for \sys.
 
-\textbf{\sys Memory Access Overhead Breakdown.} Let us now look into the results on the \sys memory operation process, i.e. we will show specifically which parts of memory from Figure~
-\ref{fig:coalEfficiency} (left) are accessed.
+\textbf{Protected Memory Accesses Breakdown.} Figure~\ref{fig:coalEfficiency} (right side) zooms into the overhead introduced by \sys's protected memory accesses.
+% Let us now look into the results on the \sys memory operation process, i.e.
+we will show specifically which parts of memory from Figure~\ref{fig:coalEfficiency} (left) are accessed.
 The result is presented in Figure~\ref{fig:coalEfficiency} (right).
-We immediately see that majority of operations are targeting the last SRAM page and very little operations on the FRAM pages.
-There are very few operations involving other SRAM pages than the last one (only visible for \textbf{cuckoo} and \textbf{dijkstra}).
+We immediately see that the majority of operations are targeting the most recently accessed SRAM page, and very few operations include a page swap from FRAM.
+There are very few cases involving SRAM pages other than the last one (only visible for \textbf{cuckoo} and \textbf{dijkstra}).
+This distinction is important because, when searching the \texttt{\underline{working}} buffer, the most recently accessed page is the first one being checked by \sys.
+This choice is highly motivated by this result.
+It is important to mention that memory access patterns are fully shaped by the application, and those ones which use a lot of protected memory can generate a higher rate of page swaps.
 
-\subsection{\sys Task Adaptation}
+\subsection{Benchmark execution time}
 \label{sec:result_coalescing}
-
-We characterize \sys by studying how its execution time varies due to task adaptation. We look at \sys in isolation and in comparison to Alpaca.
-
-\textbf{\sys Coalescing Execution Time.}  Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies for all benchmarks normalized to the \sys runtime without coalescing.
-We show data from experiments measured in the same was as for the results in Section~\ref{sec:coala_overhead}.
-
-The result in Figure~\ref{fig:coalescing} recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown} showing the benefit in \emph {normalized runtime} from coalescing (for EA and ETA case) for all applications compared to non-coalescing (NC) case.
-Coalescing improves \sys's performance by eliminating the overhead of committing after each task.
-The difference in performance between coalescing and no coalescing is, however, highly application-dependent.
-Then, the gain from energy and task-aware strategy (ETA) compared to energy-aware only (EA) are almost non-existent for applications, except for \textbf{fft}, i.e.
-the one with with large tasks.
-Furthermore, the \textbf{cuckoo} application (like other top-performers, \textbf{bc} and \textbf{sort}) has small tasks that \sys can easily coalesce to eliminate many commits.
 
@@
 	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}
-	\caption{\sys coalescing strategies (EA and ETA) performance per application on intermittent power (RF generator at 15\,cm from the WISP) compared to \sys without coalescing (NC).
-\sys task coalescing strategy provides significant improvements (from 0.25, \textbf{ar}, up to 0.70, \textbf{sort}) compared to a non-coalesced system.}
+	\caption{Coalescing strategies (EA and ETA) run time per application on intermittent power (RF generator at 15\,cm from the WISP), compared to \sys without coalescing (NC).
+Coalescing provides significant improvements: from 0.25 (\textbf{ar}) up to 0.70 (\textbf{sort}) compared to a non-coalescing system.}
 	\label{fig:coalescing}
@@
 
-\textbf{\sys versus non-adaptive Task-based Model.} We now evaluate \sys's run time performance and compare \sys to the performance of Alpaca runtime, i.e.
-a non-adaptive task-based system.
+We characterize \sys by studying how its execution time varies due to task adaptation. We look at \sys in isolation and in comparison to Alpaca.
+
+\textbf{\sys versus \sys.} Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies for all benchmarks normalized to NC's performances.
+We show data from experiments measured in the same way as for the results in Section~\ref{sec:coala_overhead}.
+
+The trend in Figure~\ref{fig:coalescing} recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown} showing the benefit, in \emph {normalized runtime}, of coalescing (considering EA and ETA), for all applications.
+Coalescing improves performances by eliminating the overhead of committing after each task.
+The difference in performance between coalescing and no coalescing is, however, highly application-dependent.
+Then, the gain from the energy- and task-aware strategy (ETA) compared to energy-aware-only (EA) are almost non-existent for applications, except for \textbf{fft}.
+Actually, in some cases energy awareness solely introduces overhead.
+This behavior reflects an uniform task decomposition.
+When tasks have similar run time, it is pleonastic to count their load.
+As for \textbf{fft}, tasks are not so uniform, and this makes energy awareness necessary (for this benchmark, NC is better than EA).
+Furthermore, benchmarks like \textbf{bc}, \textbf{cuckoo}, \textbf{dijkstra} and \textbf{sort} have small tasks that \sys can easily coalesce to eliminate many unnecessary commits.
+
+\textbf{\sys versus non-adaptive Task-based Model.} We now evaluate \sys's run time and compare it to Alpaca, a non-adaptive task-based system.
 Results are presented in Figure~\ref{fig:runtime}.
-For each application we measured its execution time using real wireless power provisioned from the RF signal generator at three distances (refer again to Section~\ref{sec:implementation}).
+For each application we measured its execution time using real wireless power provisioned by the RF signal generator at three distances (refer again to Section~\ref{sec:implementation}).
+%
+The plot presents the average execution time of each application with, for Coala and Alpaca, normalized to the latter.
+The results show that \sys provides a performance benefit compared to Alpaca for most applications (all except for \textbf{ar} and \textbf{cuckoo} at 30 and 50 cm).
+The improvement is the greatest for applications that perform repeated write after read operations, particularly if involving arrays (\textbf{dijkstra}, \textbf{fft} and \textbf{sort}).
+This effect is seen at all distances on intermittent power.
+In applications with protected variables (or array elements) accessed next to each other, but residing in different SRAM or FRAM pages, \sys incurs overhead from memory virtualization that causes its performance to be comparable to (or worse than) Alpaca (\textbf{ar}, \textbf{cuckoo}).
+%
+Another critical observation is that \textbf{fft} benchmark did not complete at any distance larger than 0.15\,cm.
+As we will show now, \sys's task downscaling mechanism, which was intendedly disabled for this experiment, will be able to solve this issue.
 
@@
 
-The data show the average execution time of each application with runtimes normalized to \sys's average.
-\todo{verify this statement}{Przemek} The results show that \sys provides a performance benefit compared to Alpaca for most applications (all except for \textbf{ar} and \textbf{cuckoo}).
-The performance benefit is the greatest for applications that perform repeating read and write operations, particularly if involving arrays (\textbf{bc}, \textbf{sort}, \textbf{dijkstra}).
-This effect is seen at all distances on intermittent power.
-In applications with protected variables (or array elements) accessed next to each other but residing in different SRAM or FRAM pages \sys incurs overhead from memory virtualization that cause its performance to be comparable to (or worse than) Alpaca (\textbf{ar}, \textbf{dijkstra}).
+\textbf{Task Downscaling Execution Time.} We have enabled task downscaling and measured the execution time for \textbf{fft} benchmark at 60\,cm from the exciter, even farther than the farthest distance at which, without this feature, this benchmark could not run to completion (refer again to Figure~\ref{fig:coalescing} and $\infty$ marks).
+\sys could execute the benchmark successfully with an average on-time of 201\,ms for NC and of 180\,ms for ETA, respectively 2.16x and 2.31x slower than at 15\,cm, which is not bad compared to the infinite run time of a static system (Alpaca in this case).
 
-Another critical observation is that \textbf{fft} benchmark did not complete at any distance larger than 0.15\,cm.
-As we will show now, \sys task downscaling mechanism will be able to solve this issue.
-
-\textbf{Task Downscaling Execution Time.} We have enabled task downscaling and measured the execution time for \textbf{fft} benchmark at all distances that this benchmark could not execute (refer again to Figure~\ref{fig:coalescing} and $\infty$ marks).
-\sys could execute the benchmark successfully with the execution time of X\,s at a distances of $d= 0.6$\,m, i.e.
-the one that exceeds the furthest distance ($d=0.5$\,m) presented in Figure~\ref{fig:runtime}, respectively.
-
-\subsection{\sys Paging Performance}
+\subsection{Paging Performance}
 \label{sec:results_memory_management}
 
-We have also evaluated the effect that paging with different page sizes has on \sys's performance.
-We show data in Figure~\ref{fig:page_size}; The left plot shows the run time performance normalized to the best per-application performance, using pages of different sizes in \sys; the right plot shows the number of page faults for the same set of executions normalized to the maximum of each set.
+We have also evaluated the effect that paging with different page sizes has on \sys's performance, shown in Figure~\ref{fig:page_size}.
+The left-hand side plot shows the run time performance normalized to the best per-application performance, using pages of different sizes; the right-hand side plot shows the number of page faults for the same set of executions normalized to the minimum of each set.
 We measured performance as MCU clock cycles using values read out by TI's Code Composer Studio IDE.
 
-\textbf{Effect of Page Size on \sys Runtime.} The results (Figure~\ref{fig:page_size}, left) show a ``bathtub curve'' in the performance of each application (the effect would be more profound with a larger set of pages than we considered), as the number of page faults varies.
-The data show that across application there is a page size that minimizes \emph{both} execution time.
+\textbf{Effect of Page Size on Run Time.} The results (Figure~\ref{fig:page_size}, left) show a ``bathtub curve'' in the performance of each application (the effect would be more profound with a larger set of pages than we considered), as the number of page faults varies.
+The data suggest that across applications there is a page size that minimizes execution time.
 We observe that the page size is not the same for each application, although 128\,B pages perform well for all applications.
@@
 Even if an application accesses few memory locations, \sys pages data at full page size, degrading performance by sometimes paging in and out unnecessary data.
-The data reveal a reasonable default page size of 128\,B and when developing an application, a programmer should consider varying the page size to moderate poor performance.
+The data reveal a reasonable default page size of 128\,B and, when developing an application, a programmer should consider varying the page size to moderate poor performance.
 

### 2018-04-16 paper/evaluation.tex
Eval revised and enhanced

@@
 	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
-	\caption{Performance of \sys applications at multiple WISP to RF generator antenna distances \{0.15, 0.3, 0.5\}\,m (left, center and right-most bar per application, respectively), compared against Alpaca (results normalized) for ETA coalescing strategy.
+	\caption{Performance of \sys applications at multiple WISP-to-exciter distances \{0.15, 0.3, 0.5\}\,m, compared against Alpaca (results normalized to Alpaca's performances) for ETA coalescing strategy.
 Results show that \sys performs on average better than Alpaca for majority of the benchmarks.
-For specific benchmarks \sys performs worse than Alpaca \todo{state why it is sometimes worse}{Carlo}}
+For specific benchmarks \sys is comparable with, or worse than, Alpaca.
 	\label{fig:runtime}

### 2018-04-16 paper/evaluation.tex
revision of evaluation up to section 9.2

--- 2018-04-16 paper/evaluation.tex
@@
 Our results illustrate key findings about \sys.
-We show through extensive qualitative and quantitative evaluation that (i) \sys \emph{provides computation progress} at the partial task level, guaranteeing execution at the worst intermittent energy conditions where state-of-the-art runtimes fail; (ii) usually \emph{improves execution speed} of applications compared to state-of-the-art runtimes; (iii) \sys, through coalescing mechanism, \emph{reduces the memory protection overhead} significantly; and (iv) that \sys is \emph{more generic to use} and allows to interact with code written for non-intermittent devices more easily.
+We show through extensive qualitative and quantitative evaluation that (i) \sys \emph{provides computation progress} at the partial task level, guaranteeing execution at the worst intermittent energy conditions where state-of-the-art runtimes fail; (ii) \sys usually \emph{improves execution speed} of applications compared to state-of-the-art runtimes; (iii) \sys, through coalescing mechanism, \emph{reduces the memory protection overhead} significantly; and  that (iv) \sys is \emph{more generic to use} and allows to interact with code written for non-intermittent devices easily.
 
@@
 \indent \textbf{Coalescing Algorithms Efficiency.} We have measured the time spent on executing (i) useful task code, (ii) wasted task code (that have to be re-executed due to uncommitted progress), and (iii) commits for all coalescing strategies (EB, EA and ETA) and all benchmarks.
-We compared our results to non-coalescing \sys (NC).
+We compared our results to non-coalescing (NC).
 The experiment was performed at 15\,cm WISP-to-exciter distance and normalized per benchmark.
-All benchmarks were using 256-B pages, except for \textbf{bc} which was using 32-B pages.
-Those page sizes are the ones yielding the best performance, selected among the set mentioned in Section~\ref{sec:implementation}.
-Detailed investigation of \sys's paging performance will be given in Section~\ref{sec:results_memory_management}.
-%
+All benchmarks were using 256\,B pages, except for \textbf{bc} which was using 32\,B pages.
+Those page sizes are the ones yielding acceptable performance\footnote{Detailed investigation of \sys's paging performance will be given in Section~\ref{sec:results_memory_management}.}, selected among the set mentioned in Section~\ref{sec:implementation}.
+
 The result is presented in Figure~\ref{fig:overallOverheadBreakdown}.
@@
 	\includegraphics[width=0.49\columnwidth]{figures/memAccess}
-	\caption{Kernel overhead.
-Left: high-level load distribution; right: detailed breakdown of memory accesses overhead from the left figure).
+	\caption{\sys kernel overhead.
+Left: high-level load distribution; right: detailed breakdown of memory accesses overhead (from the left figure).
 NC: no coalescing, EA: energy-aware coalescing, ETA: energy- and task-aware coalescing.
@@
 
-\textbf{Kernel Overhead Breakdown.} We now proceed with investigating how much utilization \sys's kernel steals to user code, quantifying task transitions and memory accesses overhead.
+\textbf{\sys Kernel Overhead Breakdown.} We now proceed with investigating how much utilization \sys's kernel steals to user code, quantifying task transitions and memory accesses overhead.
 For a broader picture, we profile kernel load for NC as well.
@@
 
-We characterize \sys by studying how its execution time varies due to task adaptation. We look at \sys in isolation and in comparison to Alpaca.
+We proceed with characterizing \sys by studying how its execution time varies due to task adaptation. We look at \sys in isolation and in comparison to Alpaca.
 
-\textbf{\sys versus \sys.} Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies for all benchmarks normalized to NC's performances.
+\textbf{\sys Cross-Evaluation.} Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies for all benchmarks normalized to NC's performances.
 We show data from experiments measured in the same way as for the results in Section~\ref{sec:coala_overhead}.
@@
 
-\textbf{\sys versus non-adaptive Task-based Model.} We now evaluate \sys's run time and compare it to Alpaca, a non-adaptive task-based system.
+\textbf{\sys versus non-adaptive Task-based Model.} We now evaluate \sys's run time and compare it to Alpaca---a non-adaptive task-based system.
 Results are presented in Figure~\ref{fig:runtime}.
 For each application we measured its execution time using real wireless power provisioned by the RF signal generator at three distances (refer again to Section~\ref{sec:implementation}).
-%
-The plot presents the average execution time of each application with, for Coala and Alpaca, normalized to the latter.
-The results show that \sys provides a performance benefit compared to Alpaca for most applications (all except for \textbf{ar} and \textbf{cuckoo} at 30 and 50 cm).
+The plot presents the average execution time of each application for Coala and Alpaca, normalized to the latter.
+The results show that \sys provides a performance benefit compared to Alpaca for most applications (all except for \textbf{ar} and \textbf{cuckoo} at 30 and 50\,cm).
 The improvement is the greatest for applications that perform repeated write after read operations, particularly if involving arrays (\textbf{dijkstra}, \textbf{fft} and \textbf{sort}).
@@
 
-\textbf{Task Downscaling Execution Time.} We have enabled task downscaling and measured the execution time for \textbf{fft} benchmark at 60\,cm from the exciter, even farther than the farthest distance at which, without this feature, this benchmark could not run to completion (refer again to Figure~\ref{fig:coalescing} and $\infty$ marks).
-\sys could execute the benchmark successfully with an average on-time of 201\,ms for NC and of 180\,ms for ETA, respectively 2.16x and 2.31x slower than at 15\,cm, which is not bad compared to the infinite run time of a static system (Alpaca in this case).
+\textbf{Task Downscaling Execution Time.} We have enabled task downscaling and measured the execution time for \textbf{fft} benchmark at 60\,cm from the exciter, that is, even \emph{farther} than the farthest distance at which (without this feature) this benchmark could not run to completion (refer again to Figure~\ref{fig:coalescing} and $\infty$ marks).
+\sys could execute the benchmark successfully with an average `on' time (the time spent for useful computation excluding energy breaks) of 201\,ms for NC and of 180\,ms for ETA, respectively 2.16$\times$ and 2.31$\times$ slower than at 15\,cm, which is impressive compared to the infinite run time of Alpaca, being a static system.
 

### 2018-04-16 paper/intro.tex

@@
 To ensure consistency and respect inter-task dependences, \sys uses a novel
-approach to \emph{memory virtualization} that buffers non-volatile variable
+approach \emph{power failure immune memory virtualization} that buffers non-volatile variable
 updates in volatile memory during a coalesced task execution, before committing
@@
 task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
-
+%
+% !!!!
+% It should be emphasized that unlike a voltage-threshold-based protection approaches that may suffer from a non-termination problem if the checkpoint has to be placed within the interrupt-disabled region, a timer-based approach adapts its partial commit location on repeated failures.
+%
 %Task coalescing removes the burden from the programmer, because it accepts any
@@
 %intermittent platform with a moderate effort.}
-
+%
 \textbf{Contributions.} To summarize, \sys's main contributions are: 
-
+%
 \begin{itemize}
 \item The first {\em adaptive} task-based intermittent execution model that dynamically coalesces tasks to avoid unnecessary commit overhead. 
-\item The first software memory virtualization mechanism for an intermittent system that \sys uses to preserve the atomicity of coalesced tasks.
+\item The first power interrupts immune memory virtualization mechanism for an intermittent system that \sys uses to preserve the atomicity of coalesced tasks.
 \item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
@@
 \end{itemize}
-
+%
 Section~\ref{sec:background} provides background on intermittent computing.
@@
 Sections~\ref{sec:task_adaptation} describe \sys's task adaptation mechanism (coalescing and downscaling).
-Memory virtualization mechanisms is provided in Section~\ref{sec:memory_virtulaization}.
+Memory virtualization mechanism is provided in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are give in Section~\ref{sec:implementation}.

### 2018-04-16 paper/relatedwork.tex
edits to related work

--- 2017-11-14 paper/relatedwork.tex
@@
-\textbf{Intermittently-Powered Devices.} There is a large body of research on energy harvesting and wirelessly powered embedded devices.
-Several recent reviews provide a broad in-depth overview~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
-As computing power consumption continues to decrease, a new wave of {\em ambiently} powered systems is emerging, powered by e.g.
-radio waves~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014}.
-Radio-powered devices~\cite{wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}
-are related to \sys---they are the technological foundation of an
-important category of intermittent computing devices. The emergence of
-intermittent computing hardware platforms has led to the development of tooling
-and instrumentation for such systems building~\cite{hester_sensys_2014,hester_sensys_2015,edb,stork,wisent}.
+\textbf{Intermittently-powered Devices.} There is a large body of research on
+energy-harvesting and wirelessly-powered embedded devices covered in recent
+surveys~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
+As computing power consumption continues to decrease, a new wave of
+systems that are powered by radio waves (RF) is
+emerging~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014,wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
+RF-powered devices are an important category of intermittent computing devices
+that \sys targets.  The emergence of intermittent computing hardware platforms
+has led to the development of instrumentation, debugging, and prototyping tools
+for such systems
+~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,wisent,flicker}.
 
-\textbf{Intermittent Execution Environments.} Early work on energy-harvesting runtime systems assumed simple, small computations not exceeding a predictable burst of energy~\cite{dewdrop} or long power-cycles/priority-based scheduling~\cite{sorber_sensys_2007}.
-These efforts did not consider data consistency with intermittence.
-Works supporting intermittent computing using dynamic, on-demand checkpoints include Mementos~\cite{mementos}, Quickrecall~\cite{quickrecall}, Hibernus++~\cite{hibernusplusplus} and HarvOS~\cite{mottola2017harvos}.
-Dynamic checkpointing systems may checkpoint
-at any point, making it difficult to implement application-level atomicity
-requirements.
-Moreover, some require hardware to measure stored energy, incurring energy cost.
-Moreover, these systems do not checkpoint non-volatile state, compromising memory consistency.
-DINO~\cite{dino} selectively versions non-volatile memory and checkpoints volatile state, ensuring progress and memory consistency.
-Rachet~\cite{ratchet} assumes all memory is non-volatile and checkpoints volatile state.
-Clang~\cite{hicks_isca_2017} takes a Ratchet-like approach leveraging special hardware, eliminating checkpointing and versioning overhead.
+\textbf{Checkpointing-based Systems.} Early work on energy-harvesting runtime
+systems, such as Dewdrop~\cite{dewdrop}, assumed simple computations
+that complete on a predictable burst of energy.
+Eon~\cite{sorber_sensys_2007} has scheduled such computation based on priority.
+Support for computation that spans power failures was first achieved with
+statically-placed conditional checkpoints in Mementos~\cite{mementos}.
 
-Recent task-based systems are Chain~\cite{chain} and Alpaca~\cite{alpaca}.
-Using static tasks, they eliminate checkpointing volatile state.
-Using channel-based memory models~\cite{chain} or automatic privatization and redo-logging~\cite{alpaca} they avoid checkpointing overheads.
-Moreover, task-based models facilitate specifying application-level atomicity properties.
+Mementos left the possibility of non-volatile state becoming inconsistent with
+volatile state saved in the checkpoint, rendering in-place writes to data
+structures in non-volatile memory unsafe~\cite{mspcdino}.
+%
+DINO~\cite{dino} addressed the consistency problem by selectively versioning
+non-volatile state within the checkpoints. Ratchet~\cite{ratchet} ensured
+consistency by placing a checkpoint at the beginning of each idempotent region
+in the code.  Clank~\cite{hicks_isca_2017} ensured consistency with custom
+hardware that dynamically tracks read-after-write dependencies in memory
+accesses and checkpoints on demand. The consistency problem was also
+approached with a combination of undo and redo logging in
+software~\cite{baghsorkhi_cgo_2018}.
+%
+Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and
+Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all
+volatile state immediately before power failure and halting the execution.
+Unlike \sys, such systems rely on introspection hardware to monitor supply
+voltage and on accurate worst-case bounds on checkpoint cost.
 
-\sys relates to the above efforts but is distinct in approach and mechanism.
-\sys relates closely to task-based models because \sys also relies on statically defined tasks to avoid checkpointing volatile state.
-\sys's automatic compilation spares the programmer's work of writing task-based code that prior systems demand.
-\sys's memory virtualization provides an alternative for ensuring memory consistency, optimized for bulk accesses to task-shared data with high locality.
-\sys's task coalescing optimization ameliorates the key overheads associated with prior systems.
+In all of the above systems, with the exception of DINO~\cite{dino},
+checkpoints are dynamic, i.e. the programmer does not have explicit control
+over the point at which the code may be resumed after power failure.
+%
+Dynamic checkpointing systems make it difficult for the programmer to respect
+application-level atomicity constraints, such as correlating sensor readings.
+%
+Checkpointing systems that copy most volatile state scale poorly as the
+size of volatile memory increases.  QuickRecall, Clank and Ratchet reduce
+the copying overhead by allocating the stack in non-volatile memory, which
+requires more time and energy to access than volatile memory (cf.
+Section~\ref{sec:cost_task-based}) and is not a viable option for off-chip non-volatile
+memory. In addition, unlike \sys, QuickRecall and Clank require custom hardware.
 
-\textbf{Memory Virtualization.} Prior work on embedded systems has studied a variety of memory virtualization strategies relating to \sys.
-TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory management.
-Later work extended the memory manager to support memory virtualization backed by flash memory~\cite{sensornetvm} and to ensure memory and type safety~\cite{tinyosmemorysafety}.
-SOS~\cite{sos}, Contiki~\cite{contiki}, and
-t-kernel~\cite{tkernel} also developed memory management abstractions that
-virtualize memory size and provide safe and indirect access.
-Mat\'e~\cite{mate} developed full virtual machines support for sensor nodes, virtualizing not just memory resources, but other states and peripherals.
-The goal of \sys is to provide consistent, intermittent execution, leveraging the benefits of efficient bulk copying.
-In contrast, prior efforts focused more on programmability and runtime reliability properties provided by virtual memory.
+\textbf{Task-based Systems.} An alternative to checkpointing are recent systems
+based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and
+Mayfly~\cite{hester_sensys_2017}. Using static tasks, they eliminate
+the need to checkpoint volatile state. Using channel-based memory
+models~\cite{chain,hester_sensys_2017} or automatic privatization and
+redo-logging~\cite{alpaca} they avoid checkpointing overheads. Moreover,
+task-based models facilitate respecting application-level atomicity
+constraints.
+%
+\sys relates closely to task-based models because \sys also relies on
+statically defined tasks to avoid checkpointing volatile state. 
+%
+However, unlike prior systems, \sys coalesces its statically-defined tasks at
+runtime into more efficient dynamic tasks that adapt to changing energy
+conditions.
+%
+\sys's mechanism for ensuring memory consistency also differs from the
+channel-based~\cite{chain} and privatization-based~\cite{alpaca} mechanisms in
+prior systems. \sys keeps memory consistent through memory virtualization
+optimized for bulk accesses to task-shared data with high locality.
 
-Another \sys-related prior effort is on unbounded, page-based transactional memory and deterministic parallel runtime systems~\cite{pagebasedtm,grace}.
-These works have a different mechanism and purpose than \sys---ensuring that data are consistent and deterministically updated during concurrent executions.
-What makes these
-efforts related is that they manage state to ensure consistency at the
-granularity of pages to amortize checking and tracking costs. \sys also tracks
-accesses at the granularity of a page and uses its paging mechanism to ensure
-consistency. Moreover, \sys's paging implementation, which keeps a shadow page
-for each page to use during commit, is similar to the shadow paging scheme
-use for transactional commit~\cite{pagebasedtm}.
+\textbf{Task Decomposition.} In contrast to \sys's dynamic construction of
+coallesced tasks at runtime, prior work has proposed to optimize task size
+at compile time. CleanCut~\cite{cleancut_2018} program analysis statically
+estimated energy consumption and split the program into tasks until all tasks
+consumed less energy than the device can store.
+%
+An alternative program analysis~\cite{baghsorkhi_cgo_2018} generated different
+versions of a program with different task sizes and empirically selected the
+best among them~\cite{baghsorkhi_cgo_2018}.
+%
+HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program
+analysis to place a minimal number of conditional checkpoints that test the
+energy level at runtime before copying state, like Mementos~\cite{mementos}.
+
+Unlike \sys, such compiler-based approaches, face the challenge of statically
+predictiong energy consumption of arbitrary input-dependent code with
+peripheral access, which is an problem without a general solution. Furthermore,
+a static decomposition approach prevents portability across devices with
+different storage capacitors. In contrast, \sys avoids forcing any assumptions
+at compile time and adapts to energy storage capacity and incoming energy
+conditions at runtime.
+
+\textbf{Memory Virtualization.} Prior work on embedded systems has studied a
+variety of memory virtualization strategies relating to \sys.
+TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory
+management. Later work extended the memory manager to support memory
+virtualization backed by flash memory~\cite{sensornetvm} and to ensure memory
+and type safety~\cite{tinyosmemorysafety}. SOS~\cite{sos},
+Contiki~\cite{contiki}, and t-kernel~\cite{tkernel} also developed memory
+management abstractions that virtualize memory size and provide safe and
+indirect access. Mat\'e~\cite{mate} developed full virtual machines support for
+sensor nodes, virtualizing not just memory resources, but other states and
+peripherals. The goal of \sys is to provide consistent, intermittent execution,
+leveraging the benefits of efficient bulk copying. In contrast, prior efforts
+focused more on programmability and runtime reliability properties provided by
+virtual memory.
+
+A related domain is unbounded, page-based transactional memory and
+deterministic parallel runtime systems~\cite{pagebasedtm,grace}.  These works
+have a different mechanism and purpose than \sys---ensuring that data are
+consistent and deterministically updated during concurrent executions.  Their
+similarity with \sys is in managing state to ensure consistency at
+the granularity of pages to amortize checking and tracking costs.
+\sys's paging implementation, which keeps a shadow page for each page to use
+during commit, is similar to the shadow paging scheme use for transactional
+commit~\cite{pagebasedtm}.

### 2018-04-16 paper/evaluation.tex

@@
 The manual code transformation was required as \sys was written with GCC in mind, and Alpaca cannot support any other complier than LLVM.
-
-%Then, we need to know how many tasks \sys can coalesce within a given execution scenario.
-We ran a subset of applications (split by tasks manually) with continuous power supply and measured the size of tasks for each application and the amount of tasks coalesced.
-Result is given in Table~\ref{tab:aveVirtuTaskSize}.
-We see clearly that \sys manages to coalesce more tasks as individual tasks are small.
-As the size of individual task increases, e.g., as in the case of dft application, the number of coalesced tasks is also small.
-This result clearly shows that for the benefit of coalescing and for code portability, initial code should be split by \emph{as small tasks as possible}.
-This will help \sys to find the best possible virtual task size to minimise its runtime.
 
@@
 The outcome is presented in Figure~\ref{fig:coalEfficiency} (left side).
-%
 As expected, non-coalescing (NC) \sys introduces a huge cost of task transitions and commits (which recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown}).
@@
 \textbf{Protected Memory Accesses Breakdown.} Figure~\ref{fig:coalEfficiency} (right side) zooms into the overhead introduced by \sys's protected memory accesses.
-% Let us now look into the results on the \sys memory operation process, i.e.
-we will show specifically which parts of memory from Figure~\ref{fig:coalEfficiency} (left) are accessed.
-The result is presented in Figure~\ref{fig:coalEfficiency} (right).
 We immediately see that the majority of operations are targeting the most recently accessed SRAM page, and very few operations include a page swap from FRAM.
@@
 In applications with protected variables (or array elements) accessed next to each other, but residing in different SRAM or FRAM pages, \sys incurs overhead from memory virtualization that causes its performance to be comparable to (or worse than) Alpaca (\textbf{ar}, \textbf{cuckoo}).
-%
 Another critical observation is that \textbf{fft} benchmark did not complete at any distance larger than 0.15\,cm.
@@
 Results show that \sys performs on average better than Alpaca for majority of the benchmarks.
-For specific benchmarks \sys is comparable with, or worse than, Alpaca.
+For specific benchmarks \sys is comparable with, or worse than, Alpaca.}
 	\label{fig:runtime}
@@
 
-\textbf{Effect of Page Size on Run Time.} The results (Figure~\ref{fig:page_size}, left) show a ``bathtub curve'' in the performance of each application (the effect would be more profound with a larger set of pages than we considered), as the number of page faults varies.
+\textbf{Effect of Page Size on \sys Runtime.} The results (Figure~\ref{fig:page_size}, left side) show a ``bathtub curve'' in the performance of each application (the effect would be more profound with a larger set of pages than we considered), as the number of page faults varies.
 The data suggest that across applications there is a page size that minimizes execution time.
@@
 
-\textbf{Page Size versus Page Faults in \sys Runtime.} The increased rate of page faults (Figure~\ref{fig:page_size}, right) is responsible for the overhead with small page sizes; smaller pages require accessing more different pages, leading to higher paging costs.
+\textbf{Page Size versus Page Faults in \sys Runtime.} The increased rate of page faults (Figure~\ref{fig:page_size}, right side) is responsible for the overhead with small page sizes; smaller pages require accessing more different pages, leading to higher paging costs.
 With larger page sizes, the overhead is higher than with moderate pages, too.

### 2018-04-16 paper/hwswsetup.tex
methodology: toolchain version
todos removed from methodology

@@
 
-We used two platforms to evaluate \sys: (1) an \emph{intermittently-powered}
-energy-harvesting device, and (2) a \emph{continuously-powered} evaluation
+We used two platforms to evaluate \sys: (i) an \emph{intermittently-powered}
+energy-harvesting device, and (i) a \emph{continuously-powered} evaluation
 board.
@@
 %
-We powered the WISP using an RF signal generator~\cite{} \todo{Add reference
-to signal generator}{Przemek} emitting  a 20-\,dBm sinusoidal wave at 915\,MHz.
+We powered the WISP using an RF signal generator emitting  a 20-\,dBm sinusoidal wave at 915\,MHz.
 The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
@@
 We affixed the WISP with a paper harness at the edge of a table, so that the
-WISP was parallel to the table surface at a height of 10~cm.
+WISP was parallel to the table surface at a height of 10\,cm.
 %
@@
 antenna at a fixed distance from the WISP. The distances were $d=\{15, 30,
-50\}$\,cm for the main experiments and 60~cm for the task downscaling
+50\}$\,cm for the main experiments and 60\,cm for the task downscaling
 experiments.
@@
 %
-All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 with
-\texttt{-O1} as optimization flag.
+All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0
+from toolchain version 5.01.01.00 with \texttt{-O1} as optimization flag.
 %
@@
 is about 30\%-40\% faster than Alpaca code instrumented automatically by the
-Alpaca LLVM compiler pass and compiled with LLVM.
-%
-\todo{Provide concrete numbers for all applications}{Amjad}
-%
-This performance difference is due to the early experimental state of the
-MSP430 backend for LLVM.
+Alpaca LLVM compiler pass and compiled with LLVM. This performance difference is due to the early experimental state of the MSP430 backend for LLVM.
 %

### 2018-04-16 paper/implementation.tex
paging
revisions to implementation

@@
 
-Table~\ref{table:coala_api} summarizes \sys's API methods and their arguments.
-\todo{Add example of a Coala program}{Carlo, Przemek} These methods trigger specific \sys kernel functions, whose explanation is given in the following sections.
+Table~\ref{table:coala_api} summarizes \sys's API methods and their arguments. These methods trigger specific \sys kernel functions, whose explanation is given in the following sections.
 
@@
 
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	\centering
-	%\includegraphics[width=0.5\columnwidth]{figures/graffle/state-machine.pdf}
-	\caption{\sys state machine implementation \todo{Draw a figure}{Sinan, Carlo}}
-	\label{fig:coala_state_machine}
-\end{wrapfigure}
+%\begin{wrapfigure}{t!}{0.5\textwidth}
+%	\centering
+%	%\includegraphics[width=0.5\columnwidth]{figures/graffle/state-machine.pdf}
+%	\caption{\sys state machine implementation \todo{Draw a figure}{Sinan, Carlo}}
+%	\label{fig:coala_state_machine}
+%\end{wrapfigure}
+
+%A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}.
 
 \textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
-A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}.
 Upon reboot, the coalescing target is initialized according to the coalescing strategy.
@@
 The remaining lower bits denote the variable's offset in its page.
-The complete process is depicted in Figure~\ref{figure:coala_page_tags} \todo{this figure doesn't really show page tags and offsets, but paging in general}.
+%The complete process is depicted in Figure~\ref{figure:coala_page_tags} \todo{this figure doesn't really show page tags and offsets, but paging in general}.
 We note that, in order to obtain unique page tags, some alignment requirements have to be met.
@@
 
-\begin{figure}
-	\centering
-	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
-	\caption{Illustration of \sys page tags mechanism.}
-	\label{figure:coala_page_tags}
-\end{figure}
+%\begin{figure}
+%	\centering
+%	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
+%	\caption{Illustration of \sys page tags mechanism.}
+%	\label{figure:coala_page_tags}
+%\end{figure}
 
@@
 
-\subsection{Partial execution}
+\subsection{Partial Execution}
 
 Among \sys's novelties, the task-splitting feature is the most effective when it comes to preservation of forward progress.
-Its implementation adds hardly any computational overhead, and is supported by a hardware timer running "in background", whose energy consumption is extremely low (around $3 \mu A/MHz$~\cite{msp430datasheet}).
-The goal is to interrupt a static task if it cannot be run to completion with a full capacitor's energy, commit its partial results, save volatile state and program counter and continue.
-Coala speculates and assumes that a static task is too large and has to be split when power fails two times consecutively in the same static task, and the coalescing target is one, because, after the first power failure, the large task is run with a fresh energy budget.
-To "split" a static task, Coala schedules a timer interrupt.
-If this interrupt occurs before the capacitor is depleted, Coala commits all the dirty volatile pages (using the two-phase commit explained above), saves stack and volatile registers to non-volatile memory and then continues with the task.
-In case a power failure occurs after this operation, but before the static task completes, the execution at the next reboot starts exactly where Coala had saved the partial execution.
+Its implementation adds hardly any computational overhead, and is supported by a hardware timer running ``in the background'', whose energy consumption is extremely low, i,e.
+$\approx$3\,$\mu$ A/MHz~\cite{msp430datasheet}.
+The goal is to interrupt a static task if it cannot be run to a completion with a full capacitor's energy, commit its partial results, save volatile state and program counter and continue.
+\sys speculates and assumes that a static task is too large and has to be split when power fails two times consecutively in the same static task, and the coalescing target is one, because, after the first power failure, the large task is run with a fresh energy budget.
+To ``split'' a static task, \sys schedules a timer interrupt.
+If this interrupt occurs before the capacitor is depleted, \sys commits all the dirty volatile pages (using the two-phase commit explained above), saves stack and volatile registers to non-volatile memory and then continues with the task.
+In case a power failure occurs after this operation, but before the static task completes, the execution at the next reboot starts exactly where \sys had saved the partial execution.
 A task can be split in multiple sub-tasks, as many as needed to progress forward.
 Regarding the timer interrupt, it is initially scheduled to fire after a large time interval.
-If power fails before the interrupt fires, Coala reduces the interval.
-If the partial execution is committed successfully, Coala enlarges the interval.
+If power fails before the interrupt fires, \sys reduces the interval.
+If the partial execution is committed successfully, \sys enlarges the interval.
 These adjustments follow a binary search technique, and make the interval converge very fast to a value optimal for preserving forward progress and avoiding redundant commits.
@@
 
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	\centering
-	%\includegraphics[width=0.5\columnwidth]{figures/graffle/commits-illustration.pdf}
-	\caption{Intermediate commit and ultimate commit Illustration \todo{to be drawn}{sinan}}
-	\label{fig:intermediate_ultimate-commit}
-\end{wrapfigure}
+%\begin{wrapfigure}{t!}{0.5\textwidth}
+%	\centering
+%	%\includegraphics[width=0.5\columnwidth]{figures/graffle/commits-illustration.pdf}
+%	\caption{Intermediate commit and ultimate commit Illustration \todo{to be drawn}{sinan}}
+%	\label{fig:intermediate_ultimate-commit}
+%\end{wrapfigure}

### 2018-04-16 paper/implementation.tex
partial commit exec revision

@@
 Among \sys's novelties, the task-splitting feature is the most effective when it comes to preservation of forward progress.
-Its implementation adds hardly any computational overhead, and is supported by a hardware timer running ``in the background'', whose energy consumption is extremely low, i,e.
-$\approx$3\,$\mu$ A/MHz~\cite{msp430datasheet}.
+Its implementation adds very little computational overhead, and is supported by a hardware timer running ``in the background'', whose energy consumption is extremely low, i,e.
+$\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}.
 The goal is to interrupt a static task if it cannot be run to a completion with a full capacitor's energy, commit its partial results, save volatile state and program counter and continue.
-\sys speculates and assumes that a static task is too large and has to be split when power fails two times consecutively in the same static task, and the coalescing target is one, because, after the first power failure, the large task is run with a fresh energy budget.
+
+\textbf{Partial Commit Algorithm.} \sys speculates and assumes that a static task is too large and has to be split when power fails two times consecutively in the same static task, and the coalescing target is one, because, after the first power failure, the large task is run with a fresh energy budget.
 To ``split'' a static task, \sys schedules a timer interrupt.
@@
 A task can be split in multiple sub-tasks, as many as needed to progress forward.
-Regarding the timer interrupt, it is initially scheduled to fire after a large time interval.
+The timer interrupt is initially scheduled to fire after a large time interval.
 If power fails before the interrupt fires, \sys reduces the interval.
@@
 
-For some applications, specific pieces of a task cannot be split and have to be run atomically.
+\textbf{Discussion.} For some applications, specific pieces of a task cannot be split and have to be run atomically.
 For instance, when sampling an analog signal, continuity between samples has to be ensured, otherwise the digital array would represent a wrong, non-continuous input signal.

### 2018-04-16 paper/paging.tex
edits (small) to paging

@@
 The shadow buffer is used to hold the dirty bages.
-The indirection table is used to speed-up page two-phase commit operation: it eliminates the copy operation in the second phase.}
+The indirection table is used to speed-up two-phase commit operation: it eliminates the copy operation in the second phase.}
 	\label{figure:coala_page_tags}
@@
 %
-\sys takes full advantage of hardware\hyp{}accelerated block copies using Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5x fewer cycles on MSP430FRxxxx.
+\sys takes full advantage of hardware\hyp{}accelerated block copies using Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5$\times$ fewer cycles on MSP430FRxxxx.
 
@@
 %
-The byte is accessed in the page buffer at the index of the resident page and the variable's page offset (Lines 5-6).
+The byte is accessed in the page buffer at the index of the resident page and the variable's page offset (Lines 5--6).
 

### 2018-04-16 paper/newCoalescing.tex
ETA algo text improvedf

@@
-Energy-aware task coalescing strategies gain \sys efficiency, and the flexible time-based task splitting strategy enables \sys to progress where its counterparts fail.
+Energy-aware task coalescing strategies bring \sys efficiency, and the flexible time-based task splitting strategy enables \sys to progress where its counterparts fail.
 
@@
 
-\noindent When incoming energy is sufficient to run multiple consecutive tasks,
-committing state after each task is wasteful. \sys mitigates this
-waste by deferring the expensive commits and performing them in a
-batch. In the resulting execution, several tasks are \emph{merged} into a
-single virtual task---a process we call \emph{coalescing}.
+When incoming energy is sufficient to run multiple consecutive tasks, committing state after each task is wasteful.
+\sys mitigates this waste by deferring the expensive commits and performing them in a batch.
+In the resulting execution, several tasks are \emph{merged} into a single virtual task---a process we call \emph{coalescing}.
 
@@
 \label{subsec:energyTaskAware}
+
 There are two factors that affect the accuracy of predicting a power interrupt.
 The stability of the incoming energy and the static tasks sizes distribution.
-If the task sizes are not uniformly distributed then the algorithm estimation might be off by a relatively large margin, for example, if there are two tasks of size $1s$ and $10s$ then EA will count them as two and might estimate that the target should be of size two static tasks.
+If the task sizes are not uniformly distributed then the algorithm estimation might be off by a relatively large margin.
+For example, if there are two tasks of size $1s$ and $10s$ then EA will count them as two and might estimate that the target should be of size two static tasks.
 However, if the following two tasks, on one hand, are of size $1s$ then the algorithm will waste energy on unnecessary commits.
@@
 To overcome this limitation the counting process of ETA algorithm takes tasks' weight into consideration to approximate a uniform task size distribution, from the counting process perspective.
-As such ETA algorithm adapts the characterization of the EA algorithm by changing the weight function to be, $f_\text{weight}(T_\text{i}) =  W_\text{i}$;
+As such ETA algorithm adapts the characterization of the EA algorithm by changing the weight function to be $f_\text{weight}(T_\text{i}) = W_\text{i}$.

### 2018-04-16 paper/hwswsetup.tex
add cite to DSP lib from TI (actually was already there)
methodology: added note on short distance

--- 2018-04-16 paper/hwswsetup.tex
@@
 experiments.
+%
+We chose relatively shorter distances than in a deployment scenario to reduce
+charge time and run experiments faster, after having confirmed ourselves a
+result also reported in prior work~\cite{alpaca} that the amount of energy
+available in one charge cycle --- the sum of stored energy and incoming energy
+during execution -- is the same at distances beyond 50~cm where incoming energy
+becomes negligible.
 %

### 2018-04-16 paper/hwswsetup.tex

--- 2018-04-16 paper/hwswsetup.tex
@@
 \textbf{fft} implementation is based on the Accelerated FFT library from
-TI~\cite{ti-fft}.
-%
-\todo{add reference}{Przemek}
-%
-All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0
+TI~\cite{ti_dsp}. All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0
 from toolchain version 5.01.01.00 with \texttt{-O1} as optimization flag.

### 2018-04-16 paper/implementation.tex

@@
 
-\textbf{Intermediate Commit and Ultimate Commit.} The reason why the commit process is split in two phases is two-fold: ensure consistency by using atomic operations, and preserve performance by enabling an incremental commit mechanism.
+\textbf{Intermediate Commit and Ultimate Commit.} The reason why the commit process is split in two phases is two fold: \emph{ensure consistency} by using atomic operations, and \emph{preserve performance} by enabling an incremental commit mechanism.
 When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to the non-volatile shadow buffer.

### 2018-04-16 paper/background.tex
moved quickrecall ref to HW based solution

@@
 Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
-Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos,quickrecall}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic}.
+Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.
 

### 2018-04-16 paper/paging.tex
added section 5.4

@@
 %despite arbitrary power failures.
-\subsection{Dynamic vs. Static Privatization}
-\TODO{Put paragraph about Alpaca's static privatization algo vs. Coala's dynamic privatization algo}
-\TODO{pass by ref might not work the right way}
-\begin{verbatim}
+\subsection{Dynamic Paging of \sys}
 
-GLOBAL_(x)
+One design choice \sys made was to insert dynamic check for paging on every
+possible problematic memory operation using {\tt RP} and {\tt WP} macros.
+Since dynamic checks incur overhead on every read and write, it has a higher
+overhead for accessing memory than a static approach,
+e.g., Alpaca, where a system statically analyze the memory operation and try to insert only optimal
+number of memory protection code (usually at the beginning of each task). 
+%
+However, \sys's dynamic paging has several benefits over static approach.
+First, static approach has limitations, prohibiting certain types of programming pattern. For example,
+in the presence of arbitrary pointer operation, function calls using function pointer, or an interrupt
+within a task, the system cannot statically analyze the memory behavior. Thus, static
+approaches either prohibits the programmer from using certain programming pattern~\cite{alpaca}, 
+or will have to conservatively insert memory protection code that can be unnecessary.
+Second, static approach cannot handle task coalescing, because with coalescing the scope to be
+searched statically is not determined at compile time. 
+By choosing to check every memory access dynamically, \sys supports programming pattern that
+Alpaca does not support, such as arbitrary pointer operation, while also being able to
+support task coalescing, which yields extra performance benefits.
+%
+%Moreover, if the programmer wants to optimize for high performance, it is possible for the
+%programmer to place {\tt RP} and {\tt WP} only minimally by carefully examining the code.
+%We argue that the
 
-foo(*y){ 
-  *y++; 
-}
-
-TASK( ){ 
-  ... ;  
-  foo( &GV(x) ) ; 
-}
-
-\end{verbatim}
+%\TODO{Put paragraph about Alpaca's static privatization algo vs. Coala's dynamic privatization algo}
+%\TODO{pass by ref might not work the right way}
+%\begin{verbatim}
+%
+%GLOBAL_(x)
+%
+%foo(*y){ 
+%  *y++; 
+%}
+%
+%TASK( ){ 
+%  ... ;  
+%  foo( &GV(x) ) ; 
+%}
+%
+%\end{verbatim}
 %\subsection{Managing Paging Overhead}

### 2018-04-16 paper/background.tex
fine-tuning of background

@@
 The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead.
-\sys's execution model based on {\em task coalescing}---merging or dissecting multiple tasks {\em during} execution based on the program execution behavior, and {\em task downscaling}---shortening task boundaries during execution when even the smallest task does not fit the energy buffer, both models described in Section~\ref{sec:task_adaptation}, is motivated by these observations.
-The main enabler for task coalescing is \emph{memory virtualization} that resolves dynamic WAR dependencies, which is described in Section~\ref{sec:memory_virtulaization}.
+\sys's execution model based on {\em task coalescing} and {\em task downscaling}.
+Coalescing dynamically merges multiple tasks based on execution behavior to amortize the cost of committing each task separately.
+Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too large to complete using the device's buffered energy.
+Section~\ref{sec:task_adaptation}describes \sys's adaptive coalescing and downscaling mechanisms in detail.
+The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, which is described in Section~\ref{sec:memory_virtulaization}.
 

### 2018-04-17 paper/background.tex
background, methodology: TODOs in tables

--- 2018-04-16 paper/background.tex
@@
 	\end{tabular}
-	\caption{Memory consistency enforcement overheads of intermittent execution models; \emph{PC}: program counter, \emph{channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
+	\caption{Memory consistency enforcement overheads of intermittent execution models; \emph{PC}: program counter, \emph{channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).
+\todo{region formation}{Alexei}}
 	\label{table:chechpoint_comparison}

### 2018-04-17 paper/hwswsetup.tex
methodology: cite Intel paper on early state of LLVM MSP430 backend
background, methodology: TODOs in tables

@@
 \caption{Description and characteristics of benchmark applications used to
-evaluate \sys.}
+evaluate \sys. \todo{SLOC and Number of Tasks}{Amjad}}
 \label{table:benchmark_table}
@@
 is about 30\%-40\% faster than Alpaca code instrumented automatically by the
-Alpaca LLVM compiler pass and compiled with LLVM. This performance difference is due to the early experimental state of the MSP430 backend for LLVM.
+Alpaca LLVM compiler pass and compiled with LLVM. This performance difference
+is due to the early experimental state of the MSP430 backend for
+LLVM~\cite{baghsorkhi_cgo_2018}.
 %

### 2018-04-17 paper/hwswsetup.tex
methodology: minor edits

@@
 intermittent computing. For a fair comparison the task decomposition is the
-same in both \sys and Alpaca implementations of the benchmarks. We compare
+same in \sys and Alpaca implementations of the benchmarks. We compare
 both systems using GCC, the compiler with the best performance for the
@@
 
-Since Alpaca's compiler pass is implemented only for LLVM, we could not use it
-to instrument the benchmarks and compile them with GCC. Instead, we
-\emph{manually} performed the instrumentation done by the compiler pass.
-We performed the instrumentation at the source level, adding C
+Since Alpaca's compiler pass is implemented only for LLVM, we could not use
+that implementation to instrument the benchmarks and compile them with GCC.
+Instead, we \emph{manually} performed the instrumentation done by the compiler
+pass.  We performed the instrumentation at the source level, adding C
 statements to the program, instead of at the level of binary
 intermediate-representation (LLVM IR) as the compiler pass would. The result is
-equivalent from the perspective of overheads. The
-instrumentation consists of identifying write-after-read dependencies and
-adding code and memory allocations to make a private copy of the
-affected variables. For array variables, instrumentation also includes adding
-code and memory allocations to track the first write to the array.
+equivalent from the perspective of overheads. The instrumentation consists of
+identifying write-after-read dependencies and adding code and memory
+allocations to make a private copy of the affected variables. For array
+variables, instrumentation also includes adding code and memory allocations to
+track the first write to the array.
 

### 2018-04-17 paper/evaluation.tex
eval: my pass

@@
 	\includegraphics[width=0.5\columnwidth]{figures/coalEfficiency}
-	\caption{Percent time usage for all coalescing strategies (including no-coalescing) comparing useful execution to intermittency overhead.
-NC: no coalescing, EB: energy-blind coalescing, EA: energy-aware coalescing, ETA: energy- and task-aware coalescing.
-All \sys coalescing strategies significantly reduce the overhead of (task) commits.
-EA and ETA perform much better than EB since they reduce re-execution penalty.
-\todo{rename legend entry 'user code' with 'useful task code'; rename y-axis to 'Time usage (\%)'; increase font by two more points everywhere}{Amjad}}
+    \caption{Breakdown of overhead for the proposed coalescing strategies
+relative to the baseline without coalescing.
+NC: no coalescing, EB: energy-blind coalescing, EA: energy-aware coalescing, ETA:
+energy- and task-aware coalescing. All coalescing strategies reduce total
+overhead and maximize useful work. EA and ETA perform better than EB.
+\todo{rename legend entry 'user code' with 'useful task code'; rename y-axis to
+'Time usage (\%)'; increase font by
+two more points everywhere}{Amjad}}
 	\label{fig:overallOverheadBreakdown}
@@
 
-Our results illustrate key findings about \sys.
-We show through extensive qualitative and quantitative evaluation that (i) \sys \emph{provides computation progress} at the partial task level, guaranteeing execution at the worst intermittent energy conditions where state-of-the-art runtimes fail; (ii) \sys usually \emph{improves execution speed} of applications compared to state-of-the-art runtimes; (iii) \sys, through coalescing mechanism, \emph{reduces the memory protection overhead} significantly; and  that (iv) \sys is \emph{more generic to use} and allows to interact with code written for non-intermittent devices easily.
+Our evaluation quantitavely demonstrates that \sys (i) \emph{reduces the memory
+protection overhead} by coalescing, (ii) \emph{improves execution speed} for
+conservative static task decompostions compared to state-of-the-art runtimes; %
+and (iii) \emph{provides computation progress} at the sub-task granularity,
+preventing non-termination at adverse energy conditions where state-of-the-art
+runtimes fail. We measured the execution time of the benchmarks implemented
+using \sys with each variant of the two proposed adaptation mechanisms:
+\emph{task coalescing} and \emph{task downscaling}. We compared \sys's best
+configurations against a state-of-the art system, Alpaca~\cite{alpaca}. We have
+also broken down \sys's overhead to explain the source of \sys's improved
+performance.
 
-\textbf{Key Performance Indicators.} Our aim is to assess \sys as detailed as possible.
-For this we have characterized \sys's runtime performance (execution time), based on a set of benchmarks, considering core task adaptation mechanisms of the system: \emph{task coalescing} and \emph{task downscaling}.
-We also characterize \sys's overhead showing precisely what the source of \sys superior performance is, investigating task re-execution and commit penalty and memory access patterns.
-
-We remind the reader that \sys is compared to Alpaca runtime and the benchmarks for Alpaca were hand-transformed in a time-consuming manual process from Clang (LLVM-generated code) to GCC (see Section~\ref{sec:implementation}).
-The manual code transformation was required as \sys was written with GCC in mind, and Alpaca cannot support any other complier than LLVM.
-
-\subsection{Characterization of \sys's Overhead}
+\subsection{Characterization of Overhead}
 \label{sec:coala_overhead}
 
-We start with the assessment of \sys's overhead for all benchmarks. This will lead us to the selection of the best coalescing strategy to proceed with.
+%\indent \textbf{Overhead Reduced by Coalescing.}
 
-\indent \textbf{Coalescing Algorithms Efficiency.} We have measured the time spent on executing (i) useful task code, (ii) wasted task code (that have to be re-executed due to uncommitted progress), and (iii) commits for all coalescing strategies (EB, EA and ETA) and all benchmarks.
-We compared our results to non-coalescing (NC).
-The experiment was performed at 15\,cm WISP-to-exciter distance and normalized per benchmark.
-All benchmarks were using 256\,B pages, except for \textbf{bc} which was using 32\,B pages.
-Those page sizes are the ones yielding acceptable performance\footnote{Detailed investigation of \sys's paging performance will be given in Section~\ref{sec:results_memory_management}.}, selected among the set mentioned in Section~\ref{sec:implementation}.
+\subsubsection{Overhead Reduced by Coalescing}
+\label{sec:overhead-coalescing}
 
-The result is presented in Figure~\ref{fig:overallOverheadBreakdown}.
-We clearly see that all coalescing strategies make a better use of the resources for \emph{all} benchmarks (except for EB in \textbf{bc}).
-With a more thorough look, one can notice that the re-execution penalty, i.e.
-the time wasted running tasks whose progress could not be committed due to a power failure, is minimized by NC.
-This is obvious, since this strategy makes the system commit after each static task.
-On the other hand, the burden on committing so frequently reflects in a degraded performance with respect to coalescing, motivating one of the needs raised by this work.
-EA and ETA strategies provide the best balance between re-execution penalty and commit cost. For this reason we proceed with the investigation of \sys without considering EB coalescing.
+For each coalescing strategy from Section~\ref{sec:task_adaptation} (EB, EA,
+and ETA) and for a baseline without coalescing (NC), we have measured the time
+spent on (i) executing useful task code, including \sys's overhead on access to
+protected variables, (ii) re-executing task code after a power failure, and
+(iii) committing state to non-volatile memory at the end of each (coalesced)
+task.  In this experiment the WISP was 15\,cm away from the signal generator
+antenna.
+%
+The page size was set to the best performing value (32~B for \textbf{bc} and
+256~B pages for the rest) identified in detailed evaluation of paging in
+Section~\ref{sec:results_memory_management}.
+
+The overhead incurred by each coalescing strategy is broken down in
+Figure~\ref{fig:overallOverheadBreakdown}. Without coalescing enabled (NC), the
+re-execution penalty is smallest, because the amount of work that can happen
+between commits and may have to be re-executed if interrupted is smallest when
+work from multiple static tasks is not combined.
+%
+However, any gain from a reduced re-execution penalty is cancelled out by
+the increased commit overhead that is incurred at the end of each static
+task.
+%
+Across all benchmarks, all coalescing strategies reduce more commit overhead
+than the re-execution overhead they add.
+%
+This net overhead reduction is greatest in EA and ETA strategies compared
+to the EB strategy. We attribute this discrepancy to EB's slow adjustment
+of the target task size without regard to execution history.
+%
+In the subsequent experiments, we focus on the better-performing EA and ETA
+strategies.
 
@@
 	\includegraphics[width=0.49\columnwidth]{figures/memAccess}
-	\caption{\sys kernel overhead.
-Left: high-level load distribution; right: detailed breakdown of memory accesses overhead (from the left figure).
-NC: no coalescing, EA: energy-aware coalescing, ETA: energy- and task-aware coalescing.
-\todo{rename legend entry 'task code' to 'user code'; rename y-axis to 'Load (\%)'}{Amjad}}
+	\caption{\sys kernel overhead: a breakdown of time spent in
+the application and in the runtime (left) and a breakdown of protected memory
+accesses into three types with different costs. NC: no coalescing, EA:
+energy-aware coalescing, ETA: energy- and task-aware coalescing. \todo{rename
+legend entry 'task code' to 'user code'; rename y-axis to 'Load (\%)'}{Amjad}}
 	\label{fig:coalEfficiency}
@@
 
-\textbf{\sys Kernel Overhead Breakdown.} We now proceed with investigating how much utilization \sys's kernel steals to user code, quantifying task transitions and memory accesses overhead.
-For a broader picture, we profile kernel load for NC as well.
-The result for all benchmarks are measured in the same way as for the results presented above (i.e.
-given in Figure~\ref{fig:overallOverheadBreakdown}) and using the same page sizes as in the previous experiment.
-The outcome is presented in Figure~\ref{fig:coalEfficiency} (left side).
-As expected, non-coalescing (NC) \sys introduces a huge cost of task transitions and commits (which recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown}).
-For both coalescing strategies (EA and ETA) protected variables' accesses constitutes about 30\% of the whole load across all benchmarks.
-In fact, dynamic address translation is the most critical bottleneck for \sys.
+%\textbf{\sys Kernel Overhead Breakdown.}
 
-\textbf{Protected Memory Accesses Breakdown.} Figure~\ref{fig:coalEfficiency} (right side) zooms into the overhead introduced by \sys's protected memory accesses.
-We immediately see that the majority of operations are targeting the most recently accessed SRAM page, and very few operations include a page swap from FRAM.
-There are very few cases involving SRAM pages other than the last one (only visible for \textbf{cuckoo} and \textbf{dijkstra}).
-This distinction is important because, when searching the \texttt{\underline{working}} buffer, the most recently accessed page is the first one being checked by \sys.
-This choice is highly motivated by this result.
-It is important to mention that memory access patterns are fully shaped by the application, and those ones which use a lot of protected memory can generate a higher rate of page swaps.
+\subsubsection{\sys Kernel Overhead Breakdown}
 
-\subsection{Benchmark execution time}
+Figure~\ref{fig:coalEfficiency} (left) breaks down the time spent on executing
+useful task code and committing further into time spent by the application and
+time spent by \sys's runtime.
+%
+This dataset is generated in a separate experiment with the same setup as the
+previous experiment in Section~\ref{sec:overhead-coalescing}.
+%
+Consistent with the previous results in
+Figure~\ref{fig:overallOverheadBreakdown}, commit overhead is highest when
+coalescing is not used to reduce the number of commits.
+%
+However, coalescing increases the cost of each access to a protected variable,
+since work in the address translation in the paging system increases as the
+task size grows.
+%
+For both coalescing strategies (EA and ETA) accesses to protected variables
+constitute about 30\% of the runtime overhead. Dynamic address translation
+necessary on each protected access is the most critical bottleneck for \sys.
+
+%\textbf{Protected Memory Accesses Breakdown.}
+
+\subsubsection{Protected Memory Accesses Breakdown}
+
+Figure~\ref{fig:coalEfficiency} (right) breaks down protected memory accesses
+into three categories.
+%
+Each type of protected access incurs a different amount of overhead: access to
+the same page in SRAM as the previous access is cheapest, access to a different
+SRAM page has a larger cost, and access to a page that needs to be swapped in
+from FRAM into SRAM is most expensive.
+%
+The results in the figure show that the overwhelming majority of accesses are
+of the cheapest kind, which motivated us to optimize this access type in our
+implementation (cf. Section~\ref{sec:impl:paging}).
+%
+The search for the accessed page in the \texttt{\underline{working}} buffer,
+will return immediately when the page is the most recently accessed page.
+%
+Only \textbf{cuckoo}, \textbf{dijkstra}, and \textbf{fft} have non-negligible
+number of accesses to a different SRAM page, which is due to the larger working
+set and a less regular access pattern in these applications.
+%
+In general, memory access patterns are shaped by the application, and the more
+program state is protected, the higher the rate of page swaps.
+
+\subsection{Execution time}
 \label{sec:result_coalescing}
 
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	%\begin{figure}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}
-	\caption{Coalescing strategies (EA and ETA) run time per application on intermittent power (RF generator at 15\,cm from the WISP), compared to \sys without coalescing (NC).
+Having shown in Section~\ref{sec:coala_overhead} that coalescing reduces
+overhead, we now investigate the outcome of this reduction on the total
+execution time. We first investigate difference variants of \sys and
+then compare the best variant to Alpaca.
+
+%\textbf{\sys Speedup with Coalescing.}
+
+\subsubsection{\sys Speedup with Coalescing}
+
+Figure~\ref{fig:coalescing} shows the \sys run time with two coalescing
+strategies (EA, ETA) normalized to the run time without coalescing (NC).
+%
+The experimental setup is the same as that in Section~\ref{sec:coala_overhead}.
+%
+The results in Figure~\ref{fig:coalescing} (left) show that all benchmarks
+complete faster with coalescing than without coalescing: from 25\% faster
+(\textbf{ar}) up to 70\% faster (\textbf{sort}).
+%
+This speedup is a consequence of the reduced overhead demonstrated in
+Section~\ref{sec:coala_overhead}.
+%
+However, the magnitude of the speedup is (1) highly application-dependent and
+(2) largely similar across the two coalescing strategies, with the exception of
+\textbf{fft}.
+%
+In some cases (\textbf{bc}, \textbf{cuckoo}, \textbf{sort}) the energy-aware
+feature in the coalescing algorithm is counter-productive.
+%
+This occurs in task decompositions with a uniform task energy, where counting
+tasks disregarding their energy provides an equal amount of information with a
+smaller effort.
+%
+In \textbf{fft}, tasks are not uniform, and accounting for task energies is
+beneficial. In fact, the lack of task energy awareness is detrimental: with EA
+\textbf{fft} runs slower than without any coalescing (NC).
+%
+The speedup is highest for \textbf{bc}, \textbf{cuckoo}, \textbf{dijkstra} and
+\textbf{sort}, because their tasks are small and are easily coalesced,
+eliminating many unnecessary commits.
+
+\begin{figure*}
+	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}%
+	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
+    \caption{\sys's coalescing performance. Left: Application execution time
+with coalescing strategies (EA and ETA) normalized to the execution time
+without coalescing. Right: Execution time with the ETA variant of \sys,
+normalized to the execution time with Alpaca, as a function of distance to the
+energy source.}
+	\label{fig:coalescing}
+	\label{fig:runtime}
+\end{figure*}
+
+%\begin{wrapfigure}{t!}{0.5\textwidth}
+%	%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}
+%	\caption{Coalescing strategies (EA and ETA) run time per application on intermittent power (RF generator at 15\,cm from the WISP), compared to \sys without coalescing (NC).
 Coalescing provides significant improvements: from 0.25 (\textbf{ar}) up to 0.70 (\textbf{sort}) compared to a non-coalescing system.}
-	\label{fig:coalescing}
-	%\end{figure} 
-\end{wrapfigure}
+%	\label{fig:coalescing}
+%	%\end{figure} 
+%\end{wrapfigure}
+%
+%\begin{wrapfigure}{t!}{0.5\textwidth}
+%	%\begin{figure}
+%	\centering
+%	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
+%    \caption{Application execution time with the ETA variant of \sys,
+%normalized to the execution time with Alpaca, as a function of distance to the
+%signal generator antenna. \sys outperforms Alpaca on the majority but not all
+%of the benchmarks.}
+%	\label{fig:runtime}
+%%\end{figure}
+%\end{wrapfigure}
 
-We proceed with characterizing \sys by studying how its execution time varies due to task adaptation. We look at \sys in isolation and in comparison to Alpaca.
+%\textbf{Benefits of Adaptive Tasks.}
 
-\textbf{\sys Cross-Evaluation.} Figure~\ref{fig:coalescing} shows the run time of \sys's three coalescing strategies for all benchmarks normalized to NC's performances.
-We show data from experiments measured in the same way as for the results in Section~\ref{sec:coala_overhead}.
+\subsubsection{Benefits of Adaptive Tasks}
 
-The trend in Figure~\ref{fig:coalescing} recapitulates the result presented in Figure~\ref{fig:overallOverheadBreakdown} showing the benefit, in \emph {normalized runtime}, of coalescing (considering EA and ETA), for all applications.
-Coalescing improves performances by eliminating the overhead of committing after each task.
-The difference in performance between coalescing and no coalescing is, however, highly application-dependent.
-Then, the gain from the energy- and task-aware strategy (ETA) compared to energy-aware-only (EA) are almost non-existent for applications, except for \textbf{fft}.
-Actually, in some cases energy awareness solely introduces overhead.
-This behavior reflects an uniform task decomposition.
-When tasks have similar run time, it is pleonastic to count their load.
-As for \textbf{fft}, tasks are not so uniform, and this makes energy awareness necessary (for this benchmark, NC is better than EA).
-Furthermore, benchmarks like \textbf{bc}, \textbf{cuckoo}, \textbf{dijkstra} and \textbf{sort} have small tasks that \sys can easily coalesce to eliminate many unnecessary commits.
+We now compare \sys's performance to Alpaca~\cite{alpaca}---a
+\emph{non-adaptive} task-based system with tasks fixed at compile time.
+Figure~\ref{fig:runtime} (right) shows the average execution time of each
+application for Coala and Alpaca, normalized to the latter. The results show
+that \sys provides a performance benefit compared to Alpaca for most
+applications, with the exception of \textbf{ar} and \textbf{cuckoo} at some
+distances.  The speedup is greatest for applications with repeated write after
+read dependencies throughout their code, particularly involving arrays
+(\textbf{dijkstra}, \textbf{fft} and \textbf{sort}). \sys's virtualized memory
+successfully amortizes the overhead of protecting memory that is accessed in
+such patterns.  In applications without locality among accesses to protected
+variables, including array elements, \sys incurs overhead from memory
+virtualization that causes its performance to be comparable to (or worse than)
+Alpaca (\textbf{ar}, \textbf{cuckoo}).
 
-\textbf{\sys versus non-adaptive Task-based Model.} We now evaluate \sys's run time and compare it to Alpaca---a non-adaptive task-based system.
-Results are presented in Figure~\ref{fig:runtime}.
-For each application we measured its execution time using real wireless power provisioned by the RF signal generator at three distances (refer again to Section~\ref{sec:implementation}).
-The plot presents the average execution time of each application for Coala and Alpaca, normalized to the latter.
-The results show that \sys provides a performance benefit compared to Alpaca for most applications (all except for \textbf{ar} and \textbf{cuckoo} at 30 and 50\,cm).
-The improvement is the greatest for applications that perform repeated write after read operations, particularly if involving arrays (\textbf{dijkstra}, \textbf{fft} and \textbf{sort}).
-This effect is seen at all distances on intermittent power.
-In applications with protected variables (or array elements) accessed next to each other, but residing in different SRAM or FRAM pages, \sys incurs overhead from memory virtualization that causes its performance to be comparable to (or worse than) Alpaca (\textbf{ar}, \textbf{cuckoo}).
-Another critical observation is that \textbf{fft} benchmark did not complete at any distance larger than 0.15\,cm.
-As we will show now, \sys's task downscaling mechanism, which was intendedly disabled for this experiment, will be able to solve this issue.
+The \textbf{fft} benchmark did not complete at distances larger than
+15\,cm~\footnote{At distances greater than 15~cm, energy incoming during
+execution is negligible and stored energy is insufficient to complete some of
+the static tasks.} with neither Alpaca, nor Coala \emph{without the downscaling
+mechanism}. This is marked with $\infty$ signs in Figure~\ref{fig:coalescing}
+(left).  This non-terminating condition cannot be avoided in Alpaca, due to the
+fixed, non-adaptive nature of its tasks.  In contrast, our next experiment
+shows how \sys's task downscaling adaptation mechanism eliminates this
+non-termination, without changing the static tasks and without recompiling the
+application.
 
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	%\begin{figure}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
-	\caption{Performance of \sys applications at multiple WISP-to-exciter distances \{0.15, 0.3, 0.5\}\,m, compared against Alpaca (results normalized to Alpaca's performances) for ETA coalescing strategy.
-Results show that \sys performs on average better than Alpaca for majority of the benchmarks.
-For specific benchmarks \sys is comparable with, or worse than, Alpaca.}
-	\label{fig:runtime}
-%\end{figure}
-\end{wrapfigure}
+In a separate experiment, we have enabled task downscaling in \sys and measured
+the execution time for \textbf{fft} benchmark at 60\,cm from the signal
+generator antenna --- even \emph{farther} than the distance at which this
+benchmark failed to terminate in the previous experiment. \sys executed the
+benchmark successfully without (NC) and with coalescing (ETA). The application
+took longer to complete than at a shorter distance of 15~cm in the previous
+experiment, because of smaller (negligible) incoming energy at the larger
+distance. The variant without coalescing (NC) slowed down by 2.16$\times$ and the
+variant with ETA coalescing slowed down by 2.31$\times$ slower, but at both
+distances, ETA completed faster than NC.
+%
+For comparison, the run time with Alpaca is effectively infinite at this
+distance, which is a consequence of its non-adaptive static tasks.
 
-\textbf{Task Downscaling Execution Time.} We have enabled task downscaling and measured the execution time for \textbf{fft} benchmark at 60\,cm from the exciter, that is, even \emph{farther} than the farthest distance at which (without this feature) this benchmark could not run to completion (refer again to Figure~\ref{fig:coalescing} and $\infty$ marks).
-\sys could execute the benchmark successfully with an average `on' time (the time spent for useful computation excluding energy breaks) of 201\,ms for NC and of 180\,ms for ETA, respectively 2.16$\times$ and 2.31$\times$ slower than at 15\,cm, which is impressive compared to the infinite run time of Alpaca, being a static system.
-
-\subsection{Paging Performance}
+\subsection{Virtual Memory Performance}
 \label{sec:results_memory_management}
 
-We have also evaluated the effect that paging with different page sizes has on \sys's performance, shown in Figure~\ref{fig:page_size}.
-The left-hand side plot shows the run time performance normalized to the best per-application performance, using pages of different sizes; the right-hand side plot shows the number of page faults for the same set of executions normalized to the minimum of each set.
-We measured performance as MCU clock cycles using values read out by TI's Code Composer Studio IDE.
+We characterize the performance of \sys's virtual memory sub-system in an
+experiment on a continuously-powered evaluation board described in
+Section~\ref{sec:methodology}.  Figure~\ref{fig:page_size} quantifies the
+effects of page size.
 
-\textbf{Effect of Page Size on \sys Runtime.} The results (Figure~\ref{fig:page_size}, left side) show a ``bathtub curve'' in the performance of each application (the effect would be more profound with a larger set of pages than we considered), as the number of page faults varies.
-The data suggest that across applications there is a page size that minimizes execution time.
-We observe that the page size is not the same for each application, although 128\,B pages perform well for all applications.
+%\textbf{Effect of Page Size on Runtime.}
 
-\textbf{Page Size versus Page Faults in \sys Runtime.} The increased rate of page faults (Figure~\ref{fig:page_size}, right side) is responsible for the overhead with small page sizes; smaller pages require accessing more different pages, leading to higher paging costs.
-With larger page sizes, the overhead is higher than with moderate pages, too.
-The explanation for these higher overheads is that large pages have a higher commit cost.
-Even if an application accesses few memory locations, \sys pages data at full page size, degrading performance by sometimes paging in and out unnecessary data.
-The data reveal a reasonable default page size of 128\,B and, when developing an application, a programmer should consider varying the page size to moderate poor performance.
+\subsubsection{Effect of Page Size on Runtime}
+
+Figure~\ref{fig:page_size} (left) shows the execution time as a function
+of page size, normalized to the lowest per-application performance among the
+set of page sizes.
+%
+With some exceptions, a ``bathtub curve'' can be observed in the data.  For
+some applications one of the two edges of the bathtub shape is missing, because
+we did not evaluate a sufficiently wide range of page sizes.
+%
+The right edge of the bathtub is a consequence of the time spent copying
+unnecessary data along with data of interest, because copies are at page
+granularity.
+%
+The left edge of the bathtub is due to the overhead of swapping pages in
+upon access.
+%
+The swap-in overhead is proportional to the number of page faults.
+%
+Page faults are more likely with smaller pages, as visible in the plot on the
+right in Figure~\ref{fig:page_size}.
+%
+The data suggest that there is a page size, distinct from the largest or
+smallest, that minimizes execution time. The best page size is not the same for
+each application. Nevertheless, if a choice must be made for all applications,
+128\,B pages perform well for most of them.
+
+%\textbf{Effect of Page Size on Page Faults.}
+
+\subsubsection{Effect of Page Size on Page Faults}
+
+Figure~\ref{fig:page_size} (right) shows the number of page faults as a
+function of page size.
+%
+The smaller the page the more likely that a memory access will land
+outside that page and that a new page will need to be swapped in.
+%
+This trend is visible for all applications, exception \textbf{bc}. The total
+amount of data accessed by \textbf{bc}, as well as its working set, is small.
+Even with the smallest page, all accesses are contained within that page, and
+no page faults occur. Without any page faults to begin with, increasing the
+page size has no effect.
 
@@
 	\includegraphics[width=0.49\columnwidth]{figures/pagePulls}
-	\caption{\sys normalized runtime for various page sizes, \{32, 64,128,256\}\,B (left) and respective page pulls (bottom) for all benchmarks.}
+    \caption{Effect of page size on execution time (left) and on the number of
+page faults (right). The execution time is normalized to the lowest in each set
+of page sizes.}
 	\label{fig:page_size}

### 2018-04-17 paper/systemdescription.tex

@@
 
-The programmer compiles their task-based code, and links to the \sys runtime, producing a \sys-enabled binary.
-The \sys runtime implements \sys's task-based programming and execution model.
-The runtime system also includes \sys's novel {\em virtualizing memory manager} and its {\em task coalescing manager}, and {\em task downscaling manager}, both of which are essential to \sys's adaptive task-based execution.
+The programmer compiles their task-based code, and links to the \sys runtime,
+producing a \sys-enabled binary. The \sys runtime implements \sys's task-based
+programming and execution model. The runtime system also includes \sys's novel
+{\em virtualizing memory manager} and its {\em adaptive task scheduler}, both
+of which are essential to \sys's adaptive task-based execution.
 
-\textbf{\sys Task Coalescing Manager.} \sys relies on its task coalescing manager to dynamically {\em coalesce} statically\hyp{}defined tasks to avoid inessential overheads associated with completing tasks.
-By default, tasks run in a sequence and each task commits its task-shared state as it completes.
-The time and energy cost of a task's commit is unnecessary if the task and its successor both complete without a power failure.
-The key insight that \sys leverages is that the first task could have deferred its commit to the second task.
-{\em Coalescing} tasks by deferring the first commit avoids the fixed cost of the first commit (task transitioning overheads).
-Coalescing also amortizes per-variable commit cost, committing each location accessed by both tasks only after the second task, rather than once per task.
+\textbf{\sys Adaptive Task Scheduler.} 
+\sys's adaptive task scheduler decides when to coalesce consecutive tasks
+together to amortize commit and transition costs, and when to downscale a task,
+breaking it into multiple tasks, to avoid non-termination.
 
-\sys's coalescing manager is an \emph{energy-aware} and \emph{software-based only} manager in a sense it uses the recent execution history as a metric to estimate the amount of the available energy to set the adequate coalesced task size accordingly.
-This fundamental feature prevents \sys from blind coalesced task adaptation which may result in a repeated power failure problem.
-For example, If \sys enlarges the coalesced task size by $x$ static tasks on a coalesced task completion and reduces it by the same number of static tasks on a power reboot, then \sys may need to reboot, for instance, ten times to reduce a coalesced task size from ten to one, when $x=1$.
-Details of coalescing will be given in Section~\ref{sec:task_coalescing}.
+\sys's scheduler dynamically {\em coalesces} statically\hyp{}defined tasks to
+avoid inessential overheads associated with completing tasks. By default, tasks
+run in a sequence and each task commits its task-shared state as it completes.
+The time and energy cost of a task's commit is unnecessary if the task and its
+successor both complete without a power failure. The key insight that \sys
+leverages is that the first task could have deferred its commit to the second
+task. {\em Coalescing} tasks by deferring the first commit avoids the fixed
+cost of the first commit (task transitioning overheads). Coalescing also
+amortizes per-variable commit cost, committing each location accessed by both
+tasks only after the second task, rather than once per task.  \sys makes
+\emph{energy-aware} scheduling decisions, using a short history of recently
+executed tasks to estimate available energy and guide future decisions. \sys
+relies on {\em software only}, avoiding the need for additional circuitry to,
+e.g., measure available energy.
 
-\textbf{\sys Task Downscaling Manager.} \sys's task downscaling manager is responsible for preserving the forward progress when \sys repeatedly fails on a single static task.
-The downscaling manager utilizes a timer-based technique and relies on the virtual memory manager to gain a full \emph{flexibility} on where to split a task.
-These features distinguish \sys task splitting technique from the traditional checkpointing.
-A checkpointing system that uses voltage threshold to place a checkpoint might cause a program to crash if the checkpoint happened to be placed between a peripheral initialization and its call.
-If a programmer disables interrupts between the peripheral initialization and the call, the program may fail indefinitely to progress since the checkpoint may only be deferred.
-The \sys task downscaling manager, on the other hand, triggers a \emph{partial commit} faster than the previous one on a repeated failure to guarantee forward progress.
-Details of task downscaling will be given in Section~\ref{sec:task_downsizing}.
+%to set the adequate coalesced task
+%size accordingly. This fundamental feature prevents \sys from blind coalesced
+%task adaptation which may result in a repeated power failure problem. For
+%example, If \sys enlarges the coalesced task size by $x$ static tasks on a
+%coalesced task completion and reduces it by the same number of static tasks on
+%a power reboot, then \sys may need to reboot, for instance, ten times to reduce
+%a coalesced task size from ten to one, when $x=1$. Details of coalescing will
+%be given in Section~\ref{sec:task_coalescing}.
 
-\textbf{\sys Virtual Memory Manager.} \sys is able to efficiently coalesce tasks because of its efficient virtual memory manager.
-\sys's memory manager paginates memory and ensures that data in a page remain consistent despite power interruptions.
-\sys allows a task to manipulate data in a volatile copy of a page only.
-Pages swap between volatile and non-volatile memory, depending on the capacity of the volatile memory and the program's access pattern.
-\sys tracks a task's memory accesses efficiently at page granularity (rather than using, e.g., costly word-granular tracking).
-When a task ends, each page it accessed commits from volatile memory (or from a non-volatile swap region for dirty pages) back to the non-volatile main memory.
-Pages efficiently, atomically commit using a two-phase commit procedure accelerated using hardware support for direct memory access (DMA).
-Details of Virtual Memory Manager are will be given in Section~\ref{sec:memory_virtulaization}.
+\sys's scheduler {\em downscales} a task to preserve forward progress when a
+single task repeatedly fails to execute to completion.  For a repeatedly-failing
+task, the scheduler uses a timer that fires at a decreasing interval to
+determine where to execute a {\em partial commit} of the task. Partial
+commit saves progress through a task that runs for too long to complete
+using the device's buffered energy. Eventually, after a series of partial
+commits, the too-long task completes --- partial commit favors progress over task atomicity.  
+When a task repeatedly fails, the scheduler uses a timer that counts an operating interval
+that decreases in length with each failure and \sys partially commits at the end of the interval. 
+Section~\ref{sec:task_coalescing} provides details of \sys's task scheduler.
+%and relies on the virtual memory manager to gain a full \emph{flexibility} on
+%where to split a task. These features distinguish \sys task splitting technique
+%from the traditional checkpointing. A checkpointing system that uses voltage
+%threshold to place a checkpoint might cause a program to crash if the
+%checkpoint happened to be placed between a peripheral initialization and its
+%call. If a programmer disables interrupts between the peripheral initialization
+%and the call, the program may fail indefinitely to progress since the
+%checkpoint may only be deferred. The \sys task downscaling manager, on the
+%other hand, triggers a \emph{partial commit} faster than the previous one on a
+%repeated failure to guarantee forward progress. Details of task downscaling
+%will be given in Section~\ref{sec:task_downsizing}.
+
+\textbf{\sys Virtual Memory Manager.} \sys is able to efficiently coalesce
+tasks because of its efficient virtual memory manager. \sys's memory manager
+paginates memory and ensures that data in a page remain consistent despite
+power interruptions. \sys allows a task to manipulate data in a volatile copy
+of a page only. Pages swap between volatile and non-volatile memory, depending
+on the capacity of the volatile memory and the program's access pattern. \sys
+tracks a task's memory accesses efficiently at page granularity (rather than
+using, e.g., costly word-granular tracking). When a task ends, each page it
+accessed commits from volatile memory (or from a non-volatile swap region for
+dirty pages) back to the non-volatile main memory. Pages efficiently,
+atomically commit using a two-phase commit procedure accelerated using hardware
+support for direct memory access (DMA). Section~\ref{sec:memory_virtualization}
+provides details of \sys's memory manager.

### 2018-04-17 paper/evaluation.tex
eval: downscaling -> downsizing

@@
 using \sys with each variant of the two proposed adaptation mechanisms:
-\emph{task coalescing} and \emph{task downscaling}. We compared \sys's best
+\emph{task coalescing} and \emph{task downsizing}. We compared \sys's best
 configurations against a state-of-the art system, Alpaca~\cite{alpaca}. We have
@@
 Figure~\ref{fig:runtime} (right) shows the average execution time of each
-application for Coala and Alpaca, normalized to the latter. The results show
+application for \sys and Alpaca, normalized to the latter. The results show
 that \sys provides a performance benefit compared to Alpaca for most
@@
 execution is negligible and stored energy is insufficient to complete some of
-the static tasks.} with neither Alpaca, nor Coala \emph{without the downscaling
+the static tasks.} with neither Alpaca, nor \sys \emph{without the downsizing
 mechanism}. This is marked with $\infty$ signs in Figure~\ref{fig:coalescing}
@@
 fixed, non-adaptive nature of its tasks.  In contrast, our next experiment
-shows how \sys's task downscaling adaptation mechanism eliminates this
+shows how \sys's task downsizing adaptation mechanism eliminates this
 non-termination, without changing the static tasks and without recompiling the
@@
 
-In a separate experiment, we have enabled task downscaling in \sys and measured
+In a separate experiment, we have enabled task downsizing in \sys and measured
 the execution time for \textbf{fft} benchmark at 60\,cm from the signal

### 2018-04-17 paper/coalescing.tex
some TODOs for oopsla18
last comments from OOPSLA 2018 branch by Alexei ported manually to main branch
heavy pass over coalescing

--- 2017-11-17 paper/coalescing.tex
@@
-When incoming energy is sufficient to run multiple consecutive tasks,
-committing state after each task is wasteful. \sys mitigates this
-waste by deferring the expensive commits and performing them in a
-batch. In the resulting execution, several tasks are \emph{merged} into a
-single virtual task---a process we call \emph{coalescing}. 
+\sys's design includes a novel task scheduler that uses efficient,
+energy-aware task coalescing to amortize task overheads, and uses a timer-based
+task-splitting mechanism to avoid non-termination of tasks too long
+for a device's energy buffer.
+
+\subsection{Task Coalescing}
+\label{sec:task_coalescing}
+
+% \begin{wrapfigure}{t!}{0.5\textwidth}
+% 	\centering
+% 	\includegraphics[width=0.5\columnwidth]{figures/coal_Intro_Figure.pdf}
+% 	\caption{When $N$ tasks coalesced $N-1$ commit operations are skipped.}
+% 	\label{fig:coalIntro}
+% \end{wrapfigure}
+
+When a device's buffered energy is sufficient to run multiple consecutive tasks
+without a power failure, committing state after each task is an unnecessary
+overhead.   \sys eliminates this overhead by {\em coalescing} a sequence of
+consecutive tasks. \sys coalesces a sequence of tasks by deferring the commit
+operations for all tasks to the end of the sequence, performing their commits
+as a batch.  In general, committing involves moving state manipulated by a task
+into its permanent location in non-volatile main memory.  In particular, \sys's
+commit procedure copies dirty pages of memory that a task updated between fast,
+volatile working memory and slower, non-volatile main memory; we defer the
+details of paging and commit to Section~\ref{sec:memory_virtualization}.
+
+When \sys coalesces a sequence of $N$ tasks, \sys avoids the cost of commit at
+the first $N-1$ boundaries.   The first $N-1$ inter-task transitions leave
+dirty pages in working memory, deferring the work of committing dirty modified
+until the commit of the last task in the sequence.  Coalescing tasks brings a
+double benefit. First, coalescing reduces the total number of relatively
+high-latency, high-energy non-volatile memory accesses because tasks primarily
+access variables in relatively low-latency, low-energy volatile memory.
+Second, coalescing eliminates the unnecessary instructions that implement task
+commit, which our measurements in Section~\ref{sec:evaluation} show constitute
+a substantial run-time overhead.
+
+%it fasten applicaiton exeution be skipping unnecessary data commits operations
+%when there is abandond energy.
+
+The effectiveness of coalescing tasks hinges on having an effective {\em
+coalescing strategy}, which determines how many consecutive tasks to coalesce.
+An effective strategy must be {\em aggressive} enough, attempting to coalesce a
+large number of consecutive tasks to amortize a large amount of commit overhead.   
 %
-When $N$ tasks are coalesced, $N-1$ boundaries are crossed and some number of
-pages are dirtied. The first $N-2$ executions of \transition statement skip
-committing dirty pages to non-volatile memory, leaving the work of committing
-pages modified by any of the $N$ tasks to the statement $N-1$.
-Coalescing is likely to reduce the total number of non-volatile memory
-accesses, because when coalesced tasks access the same pages of data, only the
-last version of the page needs to be committed.
+If power fails during a long sequence of coalesced tasks, execution will
+restart from the last commit, i.e. the first task in the sequence, losing the
+progress made by any of the coalesced tasks.
+%
+An effective strategy must also be {\em
+conservative} enough, attempting to coalesce only as many tasks as will execute
+to completion using the device's available energy, but no more.
+%
+An effective coalescing strategy moderates the risk of a power failure during a
+long coalesced task and capitalizes on the benefits of deferred commits by
+adapting the target size of the dynamic task to the energy available at
+runtime.
 
-\textbf{Trade-off between Speed and Wasted Effort.} \sys can coalesce an
-arbitrary number of consecutive tasks. However, as more tasks coalesce, their
-collective commit overhead amortizes better, but the risk of wasting work also
-increases. If power fails during a long sequence of coalesced tasks, execution
-will restart from the last commit, i.e. the first task in the sequence, losing
-the progress made by any of the coalesced tasks.  The challenge to coalescing
-tasks is determining how many tasks to coalesce before committing.
+%\textcolor{red}{maybe is possible to demonstrate with data? count number of executed non-volatile vs volatile accesses}
 %
-%During execution, an executing task is either a {\em coalesced} task, which
-%does not commit its state before transitioning or a {\em committing} task,
-%which does commit its state before transitioning.
 
-\textbf{Task Coalescing Strategies.} 
-%
-The likelihood that a task group will complete depends on the total amount of
-energy consumed by all tasks in the group, i.e. the size of the dynamic task.
-%
-A good coalescing strategy must moderate the risk of wasted work and capitalize
-on benefits of deferred commits by adapting the target size of the dynamic task
-to the energy available at runtime.
-%
-The general structure of our strategies is to increase the target task size by
-some positive delta $C_\text{up}$ after the dynamic coalesced task completes,
-and decrease it by some negative delta $C_\text{down}$ upon a reboot after a
-power failure.
-%
-The choice of $C_\text{up}$ and $C_\text{down}$, including their units, defines
-the strategy and its performance.
-%
-For example, a naive strategy that fixes $C_\text{up}$ and $C_\text{down}$ to
-``one static task``, will be very slow to react to changes in incoming energy,
-since its dynamic task size can only change by the size of one (small) static
-task per charge-discharge cycle.
-%
-Working from these observations, we propose and evaluate two coalescing
-strategies for \sys: \textbf{History-aware Coalescing (HC)} and
-\textbf{Task-size- and History-aware Coalescing (TSHC)}.
-
-%\begin{enumerate}
-%	\item \textbf{Interrupt History Coalescing (HDC):} when more tasks are coalesced without power interrupt, $C_U$ increases, and $C_D$ increases, as more task are interrupted before committing; and 
-%	\item \textbf{Weighted Interrupt History Coalescing (WHDC):} where coalescing strategy follows the one of HDC, while also taking into account the \emph{expected execution time} of each task in the application.
-%\end{enumerate}
-
-\subsection{HC: History-aware Coalescing}
-
-\todo{clarify}{AC}
-With this strategy \sys calculates only one parameter $C=C_\text{up}=C_\text{down}$ based on the number of tasks completed, until $t$th task commit, $S_t$, during the current operational period.
-\sys begins execution in a coalesced task with $S_t=0$.
-When $S_t=C_{t-1}-1$ coalesced tasks have been executed without power interrupt (where $C_{t-1}$ defines $C$ until $t-1$ task commit) a task commit follows.
-After each commit, \sys assigns new coalescing factor as $C_{t}=\lceil \sum_{k=1}^{t}S_{k-1}/2\rceil$.
-This implies that if more tasks complete in an operating period, \sys attempts to coalesce more tasks.
-However, tasks coalescing are increased by a half the speed of the successful task completion.
-This way we adapt to the coalescing speed/wasted effort trade-off: at the potential power interrupt \sys does not immediately reduce coalescing factor to one, and with successful page commit of coalescing task, \sys gradually increases coalescing factor.
-In both cases \sys reduces the effect of outliers, i.e.
-short bursts of energy or short bursts of frequent power interrupts.
-
-%\textbf{Task Coalescing Illustration.} Figure~\ref{fig:coalescing_example_states} shows a sequence of three tasks executing in \sys.
-At left, there is no coalescing and tasks incur the cost of commit at each transition.
-The center shows the same tasks executing as coalesced groups of two tasks $C=2$, avoiding one commit.
-At right, \sys coalesces three tasks ($C=3$) avoiding two commits.
-Figure~\ref{fig:coalescing_example_flowchart} shows how \sys adapts the value of $C$ based on $S$.
-\sys executes tasks in the first operating period, with $C$ equal to half the number of tasks completed in the prior execution period.
-In the example, $C$ starts at eight.
-After completing eight tasks, \sys commits and reduces $C$ by half to four.
-\sys completes two of four tasks and power fails.
-During the operating period, \sys executed a total of ten tasks (i.e., $S=10$) and when \sys restarts, $C$ is assigned to five (i.e., half of $S$).
-\sys completes the five coalesced tasks $S=5$ and after committing, assigns $C$ to half of $S$, which is three (after rounding up).
-\sys coalesces three tasks and again commits.
-At this point, $S=8$; \sys assigns $C=4$ and execution continues.
-
-\subsection{TSHC: Task-size- and History-aware Coalescing}
-
-The history-aware coalescing policy used the number of tasks in the group as a
-rough proxy for the energy cost of the task group.
-%
-With some up-front investment of developer effort and modest time and space
-overhead, a more precise proxy for task size can be obtained by profiling the
-tasks at design time.
-%
-The TSHC policy approximates task energy cost from a profile of task execution
-times.
-%
-Profiling execution time instead of direct energy consumption is
-advantageous, because the former does not require specialized
-current\hyp{}measurement hardware and can be readily done on any
-off-the-shelf microcontroller.
-%
-To use the TSHC policy, the application is first profiled, as will be described
-in the experiment below, and the task execution time for each task is written
-into a read-only table in non-volatile memory.
-%
-At runtime, the decision to coalesce or commit at each task boundary proceeds
-similarly with the HC policy, but the task completion counter, $S_t$ has units
-of ``task execution time''.\todo{be precise}{AC}
-
-To quantify the importance of task size in the coalescing decision, we measured
-the distribution of task sizes in our benchmarks (described in
-Sections~\ref{sec:results_software}).
-%
-Each benchmark was run for 10 seconds on non-intermittent power, and a timestamp
-in clock cycles was taken at each task boundary. For each execution of
-a task, its size was estimated as the maximum difference in timestamps at its
-boundaries.
-%
-The CDF for the distribution of task sizes in Figure~\ref{fig:task_profiling}
-shows the wide range of task sizes across and within benchmarks, and explains
-the advantage of the TSHC policy.
-
-%At the end of
-%task measurement, normalized integer values of task execution for task $x$, $W_x$, are
-%collected---the shortest one being one task, the longest one $W_{x,\max}$ (in
-%case of Figure~\ref{fig:task_profiling} $W_{x,\max}=5$). Result of the task
-%size measurement is provided in Figure~\ref{fig:task_profiling}.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/TskDist.pdf}%coalescing_timeline
-	\caption{\textbf{Distribution of task sizes, estimated by task execution times.} The wide variation in task size motivates accounting for task size in the coalescing strategy.}
-	\label{fig:task_profiling}
-\end{figure}
-
-%\begin{figure}
-%	\centering
-	%\subfloat[\sys coalesces consecutive tasks.]{\includegraphics[width=\columnwidth]{figures/coalescing_states}\label{fig:coalescing_example_states}}\\
-	%\subfloat[\sys dynamically adapts the amount of coalescing.]{\includegraphics[width=\columnwidth]{figures/coalescing_timeline}\label{fig:coalescing_example_flowchart}}
-	%\caption{\sys coalesces tasks to avoid overheads.}
-	%\label{fig:coalescing_example}
-%\end{figure}
-
-%However, it takes a conservative approach by making the virtual task size equal
-%to the half of the number of successfully executed real tasks before the last
-%power interrupt (see lines~\ref{algo:coalescing:executionHistory1},
-%\ref{algo:coalescing:executionHistory2} of Algorithm~\ref{algo:coalescing}). 
-%
-%We
-%advocate for the algorithm conservative approach by the following two reasons:
-%(i) the power interrupts are not uniformly distributed over the tasks; and (ii)
-%the re-execution penalty, in case of a power interrupt, grows linearly with the
-%size of a virtual task. 
-%
-%Furthermore, the algorithm considers the charging rate
-%by allowing the real tasks counter (RTC), which represents the execution
-%history between the last two power interrupts, to increase from one up to
-%multiple virtual tasks (see line~\ref{algo:coalescing:realTaskCounter} of
-%Algorithm~\ref{algo:coalescing}). 
-%
-%As a consequence, the virtual task size
-%changes rapidly (since it is half of the execution history) in response to the
-%change in the charging rate. 
-%
-%However, the algorithm sets a limit to the size of
-%the virtual task for these two reasons: (i) having an infinity virtual task
-%size, results in a certain computation progress loss; and (ii) the benefit of
-%task merging is governed by the diminishing return principle (which will be
-%demonstrated experimentally in Section~\ref{sec:results_coalescing}).
-%
-%The algorithm protects itself from power interrupts by going into three stages:
-%(i) virtual progressing, where the operations are done on the volatile memory
-%(see
-%lines~\ref{algo:coalescing:virtualProgressing1}--\ref{algo:coalescing:virtualProgressing2}
-%of Algorithm~\ref{algo:coalescing} ); 
-%(ii) transition stage, where the volatile
-%state is moved to a temporary persistent buffer. 
-%
-%Until the end of the second
-%stage if a power is interrupted, then all the virtual progress will be
-%cancelled and the virtual task will re-executed and re-initialized from a
-%consistent input (see line~\ref{algo:coalescing:virtualProgressing1} of
-%Algorithm~\ref{algo:coalescing}); 
-%
-%and (iii) persistent commit stage, upon
-%entering this stage the algorithm will make a firm transition (see
-%lines~\ref{algo:coalescing:firmTransition1},
-%\ref{algo:coalescing:firmTransition2} of Algorithm~\ref{algo:coalescing}) and
-%it will not go back to the other stages unless all the data is committed from
-%the persistent buffer to non-volatile memory. 
-%
-%Since the data is being moved in one direction and from a persistent location,
-%power interrupt is tolerable at this stage.
-
-%\begin{algorithm}[t]
-%	\caption{\sys task coalescing mechanism}
-%	\label{algo:coalescing}
-%	\scriptsize
-%	\begin{algorithmic}[1]
-%		\State $\text{p}$  \Comment{Indicates persistent variables, initialized \textbf{only} during device programming}
-%		\State $\text{RT} \in \{\sys~\text{static tasks}\}$  \Comment{Real task}
-%		\State $\text{VT} \subset \{~\text{RT's}\}$  \Comment{Virtual task}
-%		\State $\text{RTC}$  \Comment{RT counter}
-%		\State $\text{VTC}$  \Comment{VT counter}
-%		% \State $|\text{VT}|$ \Comment{VT size measured in RT}
-%		\State $\text{TVT}$ \Comment{Temporary VT}
-%		\State $\text{VT}_{\max}$ \Comment{Maximum VT size measured in RT (coalescing upper bound)}
-%		\vspace{0.1cm}
-%		
-%		\State $\text{p RTC} = x $ 
-%		\State $\text{p VT}_{\max} = y$
-%		\State $\text{p VT} = \text{RT}$ 
-%		\State $\text{p TVT} = \text{VT} $ 
-%		\State $\text{VTC} = 0 $ 
-%		\vspace{0.1cm}
-%
-%		\Function{scheduler}{\null}
-%
-%			\State $\text{VTC} = \text{RTC/2} $  \label{algo:coalescing:executionHistory1}
-%			\State $\text{RTC} = 0 $
-%			\State $\text{RT} = \text{VT}$ \Comment{Recover virtual state} \label{algo:coalescing:virtualProgressing1}
-%			\If{commiting}
-%				\State $\text{RT}=\text{TVT}$
-%				\State $\texttt{goto} \ \ \ref{commitStage}$ \label{algo:coalescing:firmTransition1}
-%			\EndIf
-%			\vspace{0.1cm}
-%
-%			\While {True}
-%				\While{$\text{VTC}--$}
-%					\Function{execute}{$\text{RT}$}
-%					\EndFunction
-%					\If{$\text{RTC} < \text{VT}_{\max}$}
-%						\State $\text{RTC}++$  \label{algo:coalescing:realTaskCounter}
-%					\EndIf
-%					\State $\text{RT} \leftarrow \text{RT}_{next}$ \Comment{Virtual progressing}
-%				\EndWhile        \label{algo:coalescing:virtualProgressing2}
-%
-%				\State \Comment{Virtual task is finished}
-%				\State $\text{Data} \rightarrow \text{pTempBuf}$ \Comment{pTempBuf: persistent temporary buffer}
-%				\State $\text{TVT} = \text{RT}$ 
-%				\State $\text{commiting} = \text{True}$ \label{algo:coalescing:firmTransition2}
-%				\State 
-%				\State $\text{VT} = \text{TVT}$ \label{commitStage}
-%				\State $\text{pTempBuf} \rightarrow \text{FRAM}$ 
-%				\State $\text{VTC} = \text{RTC/2}$ 		\Comment{Set VT size} \label{algo:coalescing:executionHistory2}
-%				\State $\text{commiting} = \text{False}$ 
-%
-%			\EndWhile
-%
-%		\EndFunction
-%			
-%	\end{algorithmic}
-%\end{algorithm}
-
-%\subsection{Power Interrupt Immune Scheduler}
-%
-%\input{relativeJumpAlgo.tex} %Task jumping algorithm
-%
-%It utilizes a persistent circular buffer (i.e.
-persistent linked list) to keep the state of a program across power failures.
-\sys provides an API to enable a programmer to have a full control over the execution flow of the program, i.e.
-(un)blocking a task or re-execute the same task which is particularly important in the intermittent execution to emulate a persistent loop.
-\todo{Expand this section}{Amjad}
-%
-%\begin{algorithm}
-%	\caption{Opportunistic virtual Task size}
-%	\label{algo:fixVirtTask}
+\noindent \textbf{Task Coalescing Strategies.} 
+% % Carlo pseudo code 
+% \begin{algorithm}[t]
+% 	\caption{Coalescing}
+% 	\label{algo:genCoalescing}
 % 	\scriptsize
@@
 % 	\begin{algorithmic}[1]
-%		\State $VT \subset \text{\{\sys Tasks\}} $  \Comment{$VT:$ Virtual Task}
-%		\State VTS : VT size
-%		\vspace{0.1cm}
-%		
-%		\While {$True$}
-%		\State $VT \leftarrow VT_{next}$
-%		\vspace{0.1cm}
-%		\While {execute $VT$} 
-%		\If { $\text{power failed twice}$ }				
-%		\State $VTS--$  
-%		\EndIf
+%         \State $H_\text{p} \gets H_\text{c}$
+%         \State $H_\text{p} \gets 0$ 
+%         \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$B, H_\text{p}$}
+
+%         \While{$\texttt{true}$}
+% 	        \State $ B_\text{c} \gets B$
+% 	        %
+% 	        \While{$ B_\text{c} > 0$}
+% 		        \State \Call{execute\_task}{$T_\texttt{i}$}
+% 		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
+% 		        \State $B_\text{c} \gets B_\text{c} - W$
+% 				\State $H_\text{c} \gets H_\text{c} + W$
 % 	        \EndWhile
-%		
-%		\vspace{0.1cm}
-%		\If {$ \text{All tasks executed}$}
-%		\State $VTS++$
-%		\EndIf
+% 	        %
+% 	         \State $B \leftarrow $ \Call{$f_\text{compl}$}{B, $H_\text{c}$,$H_\text{p}$}
+% 	        \State \Call{commit\_to\_fram}{\null}
 %         \EndWhile
 % 	\end{algorithmic}
-%\end{algorithm}
+% \end{algorithm}
+% %
+% To describe the proposed strategies we will use the following parameters:
+% \begin{itemize}
+% \item target budget $B$: total number of static tasks to be coalesced.
+% \item current budget $B_\text{c}$: remaining number of tasks to be coalesced before committing to non-volatile memory; 
+% \item history $H$: up on reboot, it is the number of executed static tasks between the last two power interrupts.
+While executing, it is the number of executed static tasks since the last power interrupt.
+% \end{itemize}
+%
+%
+% Modifiying Carlo's pseudo code a little 
+\begin{algorithm}[t]
+	\caption{Coalescing}
+	\label{algo:genCoalescing}
+	\scriptsize
+	%\small
+	\begin{algorithmic}[1]
+        \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$H$}
+        \State $H \gets 0$ 
+        \While{$\texttt{true}$}       
+	        \State $ i \gets 0$
+	        %
+	        \While{$ i <= B$}
+		        \State \Call{execute\_task}{$T_\texttt{i}$}
+		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
+		        \State $i \gets i + W$
+				\State $H \gets H + W$
+	        \EndWhile
+	        %
+	        \State \Call{commit\_to\_fram}{\null}
+	        \State $B \leftarrow $ \Call{$f_\text{commit}$}{$B$}
+        \EndWhile
+	\end{algorithmic}
+\end{algorithm}
+%
+%
+Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing
+strategy. In the algorithm, $B$ (target budget) is total number of static tasks
+that \sys will next attempt to coalesce. $T_\text{i}$ is the $i^{th}$ task
+executed since the last power failure. $H$ (history) is the number tasks
+(coalesced or not) executed between the two most recent power failures.
+$W_\text{i}$ is the {\em weight} of a task $T_\text{i}$.  A task's weight is an
+arbitrary quantity associated with the task that represents its cost in time or
+energy to execute; different coalescing strategies may apply different weights
+to a task.  For example, a uniform weighting scheme sets the weight of each
+task to one, making them all effectively equal in cost to the coalescing
+algorithm.  A different policy might use a weighting scheme that sets a task's
+weight equal to its run time relative to the task that executes for the
+shortest duration, which has unity weight.
+
+Different coalescing strategies adhere mainly to the template in
+Algorithm~\ref{algo:genCoalescing}, varying in only a few {\em characteristic
+operations} that the algorithm leaves deliberately abstract.  A strategy's
+reboot update function, $f_\text{reboot}$, updates $B$, the target budget,
+after a reboot. A strategy's weight lookup function, $f_\text{weight}$, returns
+the a task's weight.  A strategy's commit update function, $f_\text{commit}$,
+updates $B$ after a successful commit. The following sections detail several
+possible coalescing strategies that we developed for \sys.
+
+\subsubsection{Energy-oblivious Coalescing (EB)}
+\label{subsec:energyBlind}
+
+\begin{wrapfigure}{t!}{0.5\textwidth}
+	\centering
+	\includegraphics[width=0.5\columnwidth]{figures/slowCoal}
+	\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
+	\label{fig:energyBlind}
+\end{wrapfigure}
+
+A very simple {\em energy-oblivious} coalescing strategy treats all tasks as
+having unity weight and varies the size of a coalesced task linearly. In such a
+scheme, the number of tasks to coalesce, $B$, increases by a constant $x$ when
+a coalesced sequence of tasks commits, and decreases by the same constant $x$
+when a coalesced sequence of tasks fails to complete (i.e., after a power
+failure).  The characteristic operations of the energy-oblivious strategy are:
+
+\begin{itemize}
+\item $f_\text{reboot}(B) = B - x $;
+\item $f_\text{weight}(T_\text{i}) =  1$; 
+\item $f_\text{commit}(B) = B + x$; 
+\end{itemize}
+
+The energy-oblivious strategy reacts slowly to the variation in the energy
+required to execute different tasks and to the in the effective quantum of
+energy available to the device (i.e., variation in energy buffer size). 
+%
+Figure~\ref{fig:energyBlind} shows how this strategy may experience a 
+large number of repeated power failures without forward progress.
+%
+With the successful commit of each coalesced task sequence, the coalescing
+target increases.  Eventually, the target may be too high and a coalesced task
+composed of only fewer tasks will commit without interruption by a power
+failure.
+%
+The strategy then slowly decreases the coalescing target, eventually reaching a
+coalescing target that allows completion.
+%
+
+A key limitation of this algorithm lies in the equality of the target decrease
+in $f_\text{reboot}(B)$ and the increase in $f_\text{reboot}(B)$.  If, after
+some number $k$ successful commits the target must decrease back to its
+original value before the $k$ commits, the algorithm must then decrease $B$ by
+$k \times x$.  To accumulate this decrease, this strategy requires $k$ power
+failures, after each of which the strategy decreases $B$ by $x$, eventually
+decreasing by $k \times x$, but only after a long time.
+
+\subsubsection{History-Guided Coalescing (HG)}
+\label{subsec:energyAware}
+The History-Guided (HG) coalescing strategy adapts the coalescing target more
+quickly, addressing a key limitation of the EB coalescing strategy.  
+%
+EA gives each task unity weight, as with EB.  
+%
+HG differs from EB in how it updates $B$ on commit, and after a power failure.
+%
+HG assumes that as the duration of execution since a power failure increases,
+the amount of energy remaining decreases, and the risk of failing during a long
+sequence of coalesced tasks increases.  
+%
+Consequently, HG decreases the coalescing target by half with each successful commit.
+%
+HG is {\em history-guided} because it initializes the coalescing target to be
+equal to $H$, the total number of tasks executed between the most recent pair
+of power failures.
+%
+While a history of $H$ suggests that there is sufficient energy to coalesce and
+successfully commit $H$ tasks (i.e., to set $B = H$), HG conservatively sets $B$ equal to half
+of $H$ moderating the risk of a very long coalesced task. 
+%
+The fraction of the history by which HG adjusts $B$ is arbitrary.
+We experimented with a range of fraction values, settling on one half as the
+default for HG, because one half had the highest performance.
+
+The EA algorithm is characterized by the following functions:
+
+\begin{itemize}
+\item $f_\text{reboot}(B) = \lfloor(H / 2)\rfloor;$
+\item $f_\text{weight}(T_\text{i}) =  1$; 
+\item $f_\text{commit}(B) = \lfloor(B / 2)\rfloor;$ 
+\end{itemize}
+
+Figure~\ref{fig:energyAware} illustrates the operation of the history-guided
+coalescing strategy.  Each white box is a task and each red box is a commit.
+Time goes left to right and consecutive white boxes with no intervening commit
+are coalesced tasks.  A gray box is a task interrupted by a power failure.  At
+the top, the figure shows the $i^{th}$ execution period, which has a history
+value of $H^{(i)} = 6$.  $B$ is initially three and decreases to one.  Once the
+value of $B$ reaches one, HG is, in effect, expecting a power failure,
+justifying the decrease in the coalescing target.  After the power failure, HG
+first updates the history $H = H^{i+1} = 8$.  HG then uses the value of $H$ to
+update $B$, setting $B$ equal to four and continuing execution.
+
+\begin{wrapfigure}{}{0.5\textwidth}
+	\centering
+	\includegraphics[width=0.5\columnwidth]{figures/energy-aware-coal.pdf}
+	\caption{placeholder}
+	\label{fig:energyAware}
+\end{wrapfigure}
+
+%, without losing generality or requiring hardware support,
+%Moreover, it takes advantage of the guarantees that the energy buffer offers on a fresh start by coalescing many static tasks.
+
+\subsubsection{Weighted History-Guided Coalescing (WHG)}
+\label{subsec:energyTaskAware}
+
+The Weighted History-Guided (WHG) coalescing strategy accounts for the
+different energy and time cost of a program's tasks when setting the coalescing
+target.
+%
+Each different task in the program consumes a different amount of energy to
+execute to completion.  
+%
+The EB and HG coalescing strategies assume each task has uniform energy cost:
+for these strategies, $B$ simply corresponds to a target {\em count} of tasks
+coalesced, regardless of the individual cost of each task.
+%
+However, if one task executes for ten seconds and another executes for one
+second, counting tasks misjudges the amount of {\em work} in the coalesced
+tasks.
+
+Instead, WHG associates a non-unity weight with each task in the program. 
+%
+WHG tracks the sum of the weights of tasks in a coalesced sequence of tasks.
+%
+When the sum of weights reaches $B$, the target, WHG commits the coalesced
+task.
+%
+WHG differs extends EB and HG to explicitly account for the variation in energy
+cost of a program's tasks, eliminating the assumption that all tasks do the
+same work, which is a key limitation.
+%
+WHG has the same characteristic operations as HG, except that
+$f_\text{weight}(T_\text{i}) = W_\text{i}$.
+
+The effectiveness of the WHG algorithm hinges on correctly identifying the
+weight of each task, which WHG assumes is {\em statically} available.
+%
+Profiling the time and energy cost of tasks in a program is a difficult,
+orthogonal problem~\cite{cleancut,baghsorkhi_cgo_2018}.
+%
+WHG could use the result of an arbitrarily sophisticated profiling procedure.
+%
+To produce a concrete result in this paper, we give WHG access to a simple
+profile of task run time collected offline using a single fixed input.
+%
+WHG stores the profile in a lookup table that maps a task's identifier to its
+weight, making the information available to \sys's scheduler at run time.

### 2018-04-17 paper/partialcommit.tex
pass over the partial commit biz

--- 2018-04-14 paper/partialcommit.tex
@@
-\subsection{Task Downsizing}
+\subsection{Task Downscaling}
 \label{sec:task_downsizing}
 
-\begin{figure}
-	\caption{\textcolor{red}{Figure to describe partial commit is needed}}
-\end{figure}
+\sys uses {\em task downscaling} to make progress through a task that is too
+large to complete using the energy that the device can buffer.  Task
+downscaling executes part of the long task and interrupts its execution with a
+{\em partial commit}, after which the long task continues.  If power fails, the
+partial commit preserves the progress through the task, and execution resumes
+from the point of the partial commit, rather than from the start of the task.
+Eventually, after some number of partial commits and power failures, the task
+will complete and execution will move on to a task with an energy demand
+accommodated by the device's energy buffer. Task downscaling violates task
+atomicity in a non-terminating execution, favoring continued progress over
+atomicity.  If a program requires a task's atomicity over its progress, 
+the programmer can disable task downscaling for that task.
 
-\subsubsection{When Task Downsizing is Triggered}
+The key design issue for task downscaling is deciding when to partially commit.
+\sys only applies task downscaling to a task that is likely to be
+non-terminating.  A task is likely to be non-terminating if it fails to execute
+completely twice consecutively. The second incomplete execution executes after
+a power failure, when the device will have fully recharged its energy buffer.
+If a task cannot complete when executing with a full energy buffer, the task is
+non-terminating.
 
-If power fails consecutively twice before the system can complete one single static task, it means that the static task requires more energy than currently available in a discharge cycle, so it is necessary to have a mechanism to execute a chunk of the task, save the partial progress through the task, and continue try to run the rest of the task.
-If a power failure occurs after saving partial task progress, the system makes forward progress, albeit without task atomicity.
-\sys may save partial task progress multiple times before completing a very long task.
-To avoid the fail-stop case of non-termination, \sys's partial commit mechanism will split tasks that have not been marked as atomic by the programmer.
+After identifying a task as likely non-terminating, \sys must decide when
+during the task's execution to partially commit the task's state.
+%
+\sys sets a timer at the start of the likely non-terminating task, initializing
+it to a very large value (e.g., an estimate of the maximum execution time using
+the device's energy buffer).
+%
+If the timer expires before power fails, \sys partially commits, retaining 
+the timer value to use for future partial commits. 
+%
+If the timer does not expire before power fails, \sys halves the timer and
+tries again.
+%
+\sys's exponential decrease of the timer value converges rapidly to a useable
+timer value.
 
-A partial commit saves the stack, registers and the volatile pages to the non-volatile memory.
+%Putting these in comments to make space look right
+%\textcolor{red}{More variables describing the partial commit algorithm: we do not explore as many options for partial commit as we do for coalescing}
 
-\sys starts with a big random guess and down-scales it using a binary search until it succeeds in preserving the partial execution progress.
-The obtained value is the initial seed for the next partial commit procedure.
+%\todo{Proof of the maximum granularity of task downscaling - show it is limited by the clock}{Carlo}
 
-\textcolor{red}{More variables describing the partial commit algorithm: we do not explore as many options for partial commit as we do for coalescing}
-
-\subsection{Why partial commit is not \sys's default behavior}
-
-\emph{Static deployment of partial commit stubs} This approach reduces programmer effort---they still need to apply \sys's API and distribute trigger points for partial commits---but it introduces an additional overhead by saving the stack and the register file each time it commits the pages.
-
-\emph{Dynamic deployment of partial commit stubs:} This approach does not rely on any information from the programmer to trigger the commit process.
-However, it utilizes a trial and error approach to optimize the commit rate over the entire program execution time.
+%\subsubsection{Why Task Downsizing is not \sys's Default Behavior}
+%I do not understand what this section is about so I am removing it.  If someone wants to take another crack at re-writing it feel free.
+%Task downsizing is the core novelty of \sys when it comes to code portability and re-usability. However, it has limitations, which force  it to be a non-default \sys behavior. 
+%
+%\textbf{Static deployment of partial commit stubs.} This approach reduces programmer effort, although the programmer still needs to apply \sys's API and distribute trigger points for partial commits.
+On the other hand, \sys' downsizing introduces an additional overhead by saving the stack and the register file each time it commits the pages.
+%
+%\textbf{Dynamic deployment of partial commit stubs.} \sys downsizing does not rely on any information from the programmer to trigger the commit process.
+However, \sys utilizes a trial and error approach to optimize the commit rate over the entire program execution time.
 Therefore, it suffers from a large cumulative re-execution penalty.

### 2018-04-17 paper/paging.tex
fixed 5.4
partial commit on the paging section

@@
 
-\sys exposes non-volatile memory for \emph{protected} variables to
-the application as a virtual address space partitioned into pages.
-\emph{Paged virtual memory} is a key enabler of task coalescing since it keeps the non-volatile memory consistent despite dynamic WAR dependencies incurred by coalesced code regions.
-Thanks to memory virtualization, \sys have the freedom to move application's data between volatile and non-volatile memory transparently.
+\sys's {\em virtual memory manager} exposes non-volatile memory for
+\emph{protected} variables to the application as a virtual address space
+partitioned into pages.  
+%
+A program's tasks read and write these protected variables, using them like any
+other data in the program. 
+%
+The virtual memory manager automatically finds a variable in its page and  
+allows a task to directly manipulate its value.
+%
+When a task completes, the virtual memory manager ensures that updated pages 
+are committed to main memory and visible to subesequently executed tasks.
+%
+If a power failure interrupts a task, the virtual memory manager ensures that
+the task sees consistent memory state when it restarts its execution. 
+%
+The virtual memory manager ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
+%
+%
 %
@@
 %
-To this end, \sys isolates FRAM from the tasks of the application so that they interface only SRAM---which is cheaper to read/write as we indicated in Section~\ref{sec:background}.
+
+\sys's virtual memory manager \sys abstracts non-volatile storage (e.g.,
+variables in FRAM) from tasks, which simply manipulate protected variables.
+%
+Protected variables remain consistent and are durable across power failures.
+%
+While ensuring variables remain durable, \sys avoids accessing non-volatile
+memory when possible, by copying pages of data into a {\em working memory} in the device's SRAM, which
+has a lower latency and energy cost to access than the device's FRAM (as we
+showed in Section~\ref{sec:background}).
+%
+%
+On a commit, \sys atomically copies all updated (i.e., ``dirty'') pages from their
+space in the working memory back into their location in the non-volatile
+memory.
+%
+SRAM has a limited size and there is an extended working memory containing {\em
+shadow pages} in the larger FRAM.  A shadow page stores the data from an
+updated pages that need to be committed back to non-volatile memory, but that
+exceeded the size of the volatile working memory. 
 %
@@
 %as we showed in Figure~\ref{fig:framEnergy}. 
-\sys runtime implements memory
-virtualization by moving pages of the address space from the large,
-non-volatile memory to a small, volatile memory on demand and writing all dirty
-pages back to non-volatile memory atomically. Since memory moves at page granularity, \sys
-exploits hardware\hyp{}accelerated block copies using DMA.
-
-%\subsection{Paged Non-Volatile Virtual Memory}
+\sys moves pages of data efficiently using hardware accelerated Direct Memory
+Access (DMA) support, which copies a block of memory from one place to another
+as a single operation.
 
@@
 
-\textbf{System Overview.} The overall picture of the paging operations is illustrated in Figure~\ref{figure:coala_page_tags}.
-Paging moves data between non-volatile and volatile memory.
-A page has a home location in non-volatile memory which we refer to as the page's \texttt{\underline{store}} (underlined locations are non-volatile).
-As a task accesses pages, they are swapped from their home location
-into the fully\hyp{}associative \texttt{working} page buffer in volatile memory.
-The \texttt{working} buffer is allocated to take up the volatile memory that remains after space for stack is allocated.
+\textbf{System Overview.} Figure~\ref{figure:coala_page_tags} provides an
+overall view of \sys's virtual memory management system.  A page has a home
+location in non-volatile memory which we refer to as the page's
+\texttt{\underline{store}} (underlined locations are non-volatile).  When a
+task accesses a page, the memory manager swaps the page from its home location
+into the fully\hyp{}associative \texttt{working} buffer that is stored in
+volatile memory.  The \texttt{working} buffer is allocated to take up the
+volatile memory that remains after space for stack is allocated.
 %
-\sys uses a two-phase commit to write dirty pages from the \texttt{working} buffer back to their home location in \texttt{\underline{store}} atomically, ensuring data consistency even with power failures.
+\sys uses a two-phase commit to write dirty pages from the \texttt{working}
+buffer back to their home location in \texttt{\underline{store}} atomically,
+ensuring data consistency even in the presence of a power failure during commit.
 %
-When any dirty page is swapped out from the \texttt{working} buffer due to eviction or task transition from last coalesced task, the first phase writes the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer, and the second phase writes it from \texttt{\underline{shadow}} to \texttt{\underline{store}}.
+The memory manager is forced to swap out a dirty page from the \texttt{working} buffer 
+when a task accesses a page that is not in the working buffer, and there is no
+free page in the working buffer (i.e., a page fault). On a page fault, \sys
+moves the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
+page buffer. Later, to commit the page, \sys writes it from its
+\texttt{\underline{shadow}} location back to its home location in
+\texttt{\underline{store}}.
 %
-Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a time between volatile and non-volatile memory, \sys's page-based design moves contiguous blocks of data.
+%BML: this point is true, but without caches, it is not a performance factor, so I removed it.
+%Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a
+%time between volatile and non-volatile memory, \sys's page-based design moves
+%contiguous blocks of data.
 %
-\sys takes full advantage of hardware\hyp{}accelerated block copies using Direct Memory Access (DMA) hardware in common microcontrollers --- e.g. 5$\times$ fewer cycles on MSP430FRxxxx.
+%BML: this seems gratuitous because DMA is very common and we've already introduced it.
+%\sys takes full advantage of hardware\hyp{}accelerated block copies using
+%Direct Memory Access (DMA) hardware in common microcontrollers --- e.g.
+%5$\times$ fewer cycles on MSP430FRxxxx.
 
@@
 
-A \sys task must access protected non-volatile variables through a restricted memory interface, as will be described in Section~\ref{sec:coala_api}.
-The interface includes \texttt{RP(var)}, to read the value of variable {\tt var}, and \texttt{WP(var)} to assign a value to variable {\tt var}.
-The implementation of {\tt RP} is shown in Algorithm~\ref{algo:rwar}, and {\tt WP}'s implementation is similar except for marking the accessed page as dirty.
+A \sys task must access protected non-volatile variables through \sys's
+restricted memory interface, which Section~\ref{sec:coala_api} describes in
+detail.  The interface includes \texttt{RP(var)}, to read the value of variable
+{\tt var}, and \texttt{WP(var)} to assign a value to variable {\tt var}. The
+implementation of {\tt RP} is shown in Algorithm~\ref{algo:rwar}, and {\tt
+WP}'s implementation is similar except that {\tt WP} marks the accessed page as
+dirty.
 %
-{\tt RP} and {\tt WP} operations translate a variable's physical address in non-volatile memory into a \emph{virtual address} in the page buffer in volatile memory.
-In \sys, a virtual address is composed of a \emph{page tag} that identifies the page and a \emph{page offset} that identifies a byte.
+{\tt RP} and {\tt WP} operations translate a variable's physical address in
+non-volatile memory into a \emph{virtual address} in the page buffer in
+volatile memory. In \sys, a virtual address is composed of a \emph{page tag}
+that identifies the page and a \emph{page offset} that identifies a byte.
 
@@
 
-After address translation, \sys attempts to access the protected variable's location in the volatile paging buffer.
-\sys keeps track of the page tags for the pages currently resident in the paging buffer.
-When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt working} (Line 2).
+After address translation, \sys attempts to access the protected variable's
+location in the volatile paging buffer. \sys keeps track of the page tags for
+the pages currently resident in the paging buffer. When a task accesses a
+variable, it compares the variable's page tag to tags of the pages in {\tt
+working} (Line 2).
 %
@@
 %
-If the accessed variable's page tag is not found in the page buffer, the operation incurs a {\em page fault} (Line 4).
-At a page fault, \sys must swap out one of the pages that is resident in the page buffer (to the non-volatile \texttt{\underline{shadow}}) if \texttt{working} is full, and swap in the variable's page.
+If the accessed variable's page tag is not found in the page buffer, the
+operation incurs a {\em page fault} (Line 4). At a page fault, \sys must swap
+out one of the pages that is resident in the page buffer (to the non-volatile
+\texttt{\underline{shadow}}) if \texttt{working} is full, and swap in the
+variable's page.
 %
-The byte is accessed in the page buffer at the index of the resident page and the variable's page offset (Lines 5--6).
+The byte is accessed in the page buffer at the index of the resident page and
+the variable's page offset (Lines 5--6).
 
@@
 
-If the \texttt{working} buffer is full, a page fault on a memory access requires \sys to swap out one of the pages in the volatile buffer, the \emph{victim page}, preserving updates made to that page, and to copy the accessed page of protected data into the page buffer.
-Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault.
-A victim page in the \texttt{working} buffer is selected for eviction according to \emph{first-in-first-out} replacement policy (Lines 2 and 12).
-If the page being swapped out is not dirty and the page being accessed had not been dirtied and swapped out to the \texttt{shadow} buffer since last boot, then handling a page fault consists of copying the accessed page from \texttt{\underline{store}} into the \texttt{working} buffer (Line 11).
+If the \texttt{working} buffer is full, a page fault on a memory access
+requires \sys to swap out one of the pages in the volatile buffer (a 
+\emph{victim page}) preserving updates made to that page, and to copy the
+accessed page of protected data into the working buffer.
+Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault. A victim
+page in the \texttt{working} buffer is selected for eviction according to
+a \emph{first-in-first-out} replacement policy (Lines 2 and 12). If the page
+being swapped out is not dirty and the page being accessed had not been dirtied
+and swapped out to the \texttt{shadow} buffer since last boot, then handling a
+page fault consists of copying the accessed page from
+\texttt{\underline{store}} into the \texttt{working} buffer (Line 11).
 Otherwise, the page fault handler must take the following extra steps.
 %
-If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty.
-Before swapping in the accessed page, \sys must first preserve the updated values in the dirty page so that they can be committed back to their locations in memory when the dynamic task ends (but not earlier).
-To preserve dirty values, \sys copies the victim page to the non-volatile \texttt{\underline{shadow}} buffer (Line 5) and records the page's tag in a non-volatile list (Lines 6-7) to support iteration over shadow pages having traversing all slots in the buffer, necessary in the second phase of the commit.
+If a task modified any byte in the victim page (i.e., using \texttt{WP}), then
+the page is dirty. Before swapping in the accessed page, \sys must first
+preserve the updated values in the dirty page so that they can be committed
+back to their locations in memory when the dynamic task ends (but not earlier).
+To preserve dirty values, \sys copies the victim page to the non-volatile
+\texttt{\underline{shadow}} buffer (Line 5) and records the page's tag in a
+non-volatile list (Lines 6-7) to support iteration over shadow pages having
+traversing all slots in the buffer, necessary in the second phase of the
+commit.
 %
@@
 number of memory protection code (usually at the beginning of each task).
-%
+
 However, \sys's dynamic paging has several benefits over static approach.

### 2018-04-17 paper/background.tex
pass over background

--- 2018-04-16 paper/background.tex
@@
 These devices elide tethered power or a battery, instead collecting energy into a capacitor, operating when sufficient energy accumulates, and upon depleting the energy, turning off and recharging.
-%\TODO{We need a figure that shows what intermittent execution looks like.}
 %
@@
 Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
-We do not require a particular non-volatile memory technology, nor do we require architectural additions commodity processors~\cite{su_date_2017,ratchet,quickrecall,nvp}.
+\sys do not require architectural additions to commodity processors as in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.
 \sys supports I/O behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.
@@
 %Intermittent execution is different from continuous execution. 
-A power failure clears volatile  state (registers, stack, and globals) and non-volatile memory (e.g., FRAM) persists.
+A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists.
 Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
-Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mementos,mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.
+Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.
 
@@
 
-Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}.
+Checkpointing volatile state alone do not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}.
+Especially, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependence between operations that manipulate 
+non-volatile memory, because it allows {\em writes} from a failed attempt to persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}
+---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution 
+using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
+Additional method to maintain data consistent in the presence of WAR dependency is necessary for correct execution.
+%Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}.
 Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value.
@@
 Checkpoint-based systems risk violating memory consistency~\cite{dino}.
-To maintain consistency, prior systems~\cite{dino,ratchet} {\em version} a subset of non-volatile data with the checkpoint.
+%To maintain consistency, prior systems~\cite{dino} {\em version} a subset of non-volatile data with the checkpoint.
 Task-based systems~\cite{chain,alpaca}, which we focus on, ensure consistency using programmer, compiler, and runtime support.
@@
 
-Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into tasks.
-A task is a function with no caller containing arbitrary computation, sensing, and communication.
-A programmer describes task control-flow as a {\em task graph}.
-Task flow happens at programmer demarcated transition points that may be conditional on program values.
+Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into tasks,
+which is a region of code that can contain arbitrary computation, sensing, and communication. 
+Task-based models keep progress at the granularity of tasks by repeatedly trying to re-execute
+each task on power failure until it successfully finishes, only then moving on to the next task.
+Since these models do not rely on taking an expensive checkpoint, it is usually faster than the
+checkpoint-based solution~\cite{chain, alpaca}.
+\sys also follows the paradigm of the task-based programming model, where the programmer
+explicitly expresses the application as a sequence of tasks and the transition between them.
+%Task-based runtime systems ensure that repeated task executions are idempotent. 
+%The key to idempotence is ensuring that non-volatile updates made by an interrupted task are never visible to a future task execution.  
+%A programmer describes task control-flow as a {\em task graph}. Task flow happens at programmer demarcated transition points that may be conditional on program values. 
 %Task-based programming abstractions guarantee that tasks execute {\em atomically}, regardless of power failures.
-Task-based runtime systems ensure task-atomic semantics by ensuring that repeated task executions are idempotent.
-The key to idempotence is ensuring that non-volatile updates made by an interrupted task are never visible to a future task execution.
 
-There are several run time strategies to ensuring task idempotence.
+Task-based models also suffer from data inconsistency when there is a WAR dependence.
+Prior systems tried to tackle the problem by a compiler-automated redo-logging for the variables
+that is part of the dependence~\cite{alpaca}, or by statically creating a multiple copies of the problematic variable
+to ensure that no task reads and writes the same copy~\cite{chain}.
+%There are several run time strategies to ensuring task idempotence.
 One approach~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{ratchet} did for checkpoints), execute the task using private copies of those data, and commit the private copies on task completion.
@@
 
-\subsubsection{Costs of Task-based Models}
+\subsection{Costs of Previous Models}
 \label{sec:cost_task-based}
 
-Despite their advantages, the task-based programming models have their limitations.
-Two major ones being: {\em memory consistency enforcement overhead} and {\em inflexibility to dynamic energy harvesting conditions}.
+Prior systems, including the task-based models, suffer from two major overheads:
+{\em frequent non-volatile memory access} and {\em incorrectly size tasks}.
 
@@
 		Mementos~\cite{mementos}	& Registers + Stack     \\
-		DINO~\cite{dino}	& Registers + WAR variables \\
-		Chain~\cite{chain}	& PC + channel data\\
-		Alpaca~\cite{alpaca}	& PC + WAR variables \\
+		DINO~\cite{dino}	& Registers + Stack + WAR NV variables used in task\\
+		Chain~\cite{chain}	& PC + NV variables used in task\\
+		Alpaca~\cite{alpaca}	& PC + WAR NV variables used in task\\
 		Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
-		Region Formation~\cite{baghsorkhi_cgo_2018} & \textcolor{red}{XXX} \\
+		Region Formation~\cite{baghsorkhi_cgo_2018} & Registers + Updated variables in task \\
 		\hline
 	\end{tabular}
-	\caption{Memory consistency enforcement overheads of intermittent execution models; \emph{PC}: program counter, \emph{channel data}: variables explicitly task-shared by the programmer, \emph{WAR variables}: variables involved in WAR dependences (\emph{NV}: non-volatile).}
+	\caption{Non-volatile memory access for data consistency; \emph{PC}: program counter, \emph{WAR variables}: variables involved in WAR dependences, \emph{NV}: non-volatile.}
 	\label{table:chechpoint_comparison}
@@
 
-\textbf{Memory Consistency Enforcement Overhead.} Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
+\textbf{Frequent Non-volatile Memory Access.} 
+Previous systems frequently access the non-volatile memory directly, which we claim is a major source of inefficiency.
+Checkpoint-based models copy the entire volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile
+memory on every checkpoint.
+Task-based models, which do not take checkpoints, still have to copy a subset of the non-volatile data to maintain memory consistent in WAR dependence (e.g., redo-logging).
+Table~\ref{table:chechpoint_comparison} summarizes the data that gets copies to and from the non-volatile memory for memory consistency and forward progress.
+Ratchet~\cite{ratchet} and Clank~\cite{hicks_isca_2017}, which uses a whole non-volatile memory, obviously also has a high number of non-volatile memory access
+for normal operations in addition to taking checkpoints.
+
+%Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
 Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf.
@@
 %\end{figure}
-To understand the non-volatile memory access overheads, we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory operations for the MSP430FR5969~\cite{msp430datasheet} MCU.
-We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device across 1600 read and write accesses to each memory type, randomizing to avoid caching.
-Our experiment showed that a non-volatile memory access consumes around 1.5 times the energy of a volatile access.
+To understand the expense of non-volatile memory access, we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory 
+operations for the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device 
+across 1600 read and write accesses to each memory type, randomizing to avoid caching. Our experiment showed that a non-volatile memory access 
+consumes around 1.5 times the energy of a volatile access. The result shows that non-volatile memory access is more expensive than the volatile
+memory access, and prior works with frequent non-volatile memory accesses is inefficient.
 %~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
@@
 \todo{add citation to TI specification on FRAM energy cost}{Przemek}
-Based on this observation, one of the primary motivation of \sys is to promote accessing SRAM more frequently rather than FRAM in order to reduce memory access energy overhead and to increase the time between power failures.
+Based on this observation, one of the primary motivation of \sys is to promote accessing SRAM more frequently rather than FRAM in order to 
+reduce memory access energy overhead and to increase the time between power failures.
 
@@
 
-\textbf{Fixed-size Tasks are Inflexible and Inefficient.} If a programmer writes a task that consumes {\em more} energy than the device can buffer, the task will never complete using buffered energy {\em preventing forward progress} by repeatedly re-executing.
-If a programmer writes tasks that consume far {\em less} energy than a device can buffer, the system may operate {\em inefficiently}.
-When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).
+\textbf{Incorrectly Sized Tasks.} Compiler-inserted checkpoints or programmer-defined tasks can both be incorrect or inefficient.
+If the code between checkpoints or the task (for brevity, we will refer to both as ''tasks'' in this section) consumes {\em more} 
+energy than the device can buffer, the task will never complete 
+using buffered energy {\em preventing forward progress}, making the system incorrect. 
+%
+If the task consumes far {\em less} energy than a device can buffer, the system may operate {\em inefficiently},
+checkpointing or transitioning between tasks too often than needed. 
+%
+Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is challenging both to the
+programmer and the compiler, since estimating the exact energy use of an arbitrary code is complicated.
+Consequently, existing systems either risks non-termination to achieve high performance, or 
+use conservatively small tasks that result in poor performance. 
+When {\em porting} apps from one device to another, the problem gets much worse.
+An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer, and vice versa. 
+Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
+exemplifying the problem.
+%
+As a solution, \sys's execution model is based on {\em task coalescing} and {\em task downscaling}.  
+Coalescing dynamically merges multiple tasks based on execution behavior to amortize the overhead between tasks if tasks are small.
+Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too 
+large to complete using the device's buffered energy.  Section~\ref{sec:task_adaptation} describes \sys's adaptive coalescing and 
+downscaling mechanisms in detail. The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, 
+which is described in Section~\ref{sec:memory_virtulaization}.
+%When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).
 However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions.

### 2018-04-17 paper/background.tex
shortened and changed bg text

@@
 programmer and the compiler, since estimating the exact energy use of an arbitrary code is complicated.
-Consequently, existing systems either risks non-termination to achieve high performance, or 
-use conservatively small tasks that result in poor performance. 
-When {\em porting} apps from one device to another, the problem gets much worse.
-An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer, and vice versa. 
-Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
-exemplifying the problem.
+When {\em porting} apps from one device to another, the problem gets much worse because
+an excessively costly task on one device may be a relatively short task on a device with a larger buffer, and vice versa. 
+Buffer sizes of different platforms vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
+complicating the problem.
 %
-As a solution, \sys's execution model is based on {\em task coalescing} and {\em task downscaling}.  
-Coalescing dynamically merges multiple tasks based on execution behavior to amortize the overhead between tasks if tasks are small.
-Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too 
-large to complete using the device's buffered energy.  Section~\ref{sec:task_adaptation} describes \sys's adaptive coalescing and 
-downscaling mechanisms in detail. The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, 
-which is described in Section~\ref{sec:memory_virtulaization}.
+As a solution, \sys's execution model presents {\em task coalescing}, which merges multiple tasks to amortize the overhead
+when tasks are too small, and {\em task downscaling}, which breaks up a task when the task is too large. 
 %When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).

### 2018-04-17 paper/paging.tex
done with paging

@@
 page in the \texttt{working} buffer is selected for eviction according to
-a \emph{first-in-first-out} replacement policy (Lines 2 and 12). If the page
-being swapped out is not dirty and the page being accessed had not been dirtied
-and swapped out to the \texttt{shadow} buffer since last boot, then handling a
-page fault consists of copying the accessed page from
-\texttt{\underline{store}} into the \texttt{working} buffer (Line 11).
-Otherwise, the page fault handler must take the following extra steps.
+a \emph{first-in-first-out} (FIFO) replacement policy (Lines 2 and 12). 
+%
+While we opted for the simplicity of FIFO, any other replacement policy, such
+as Least Recently Used (LRU) could work in its place.
 %
 If a task modified any byte in the victim page (i.e., using \texttt{WP}), then
-the page is dirty. Before swapping in the accessed page, \sys must first
-preserve the updated values in the dirty page so that they can be committed
-back to their locations in memory when the dynamic task ends (but not earlier).
-To preserve dirty values, \sys copies the victim page to the non-volatile
+the page is dirty.  If the page being swapped out is not dirty and the page
+being accessed had not been dirtied and swapped out to the \texttt{shadow}
+buffer since the last reboot, then handling a page fault requires copying the
+accessed page from its home in the  \texttt{\underline{store}} into the
+\texttt{working} buffer (Line 11).  Otherwise, the page fault handler must take
+the following extra steps.
+%
+Before swapping in the accessed page, \sys must first preserve the updated
+values in the dirty page so that they can be committed back to their locations
+in memory when the task commits (but not earlier).  To preserve the dirty
+values, \sys copies the victim page to the non-volatile
 \texttt{\underline{shadow}} buffer (Line 5) and records the page's tag in a
-non-volatile list (Lines 6-7) to support iteration over shadow pages having
-traversing all slots in the buffer, necessary in the second phase of the
-commit.
+non-volatile commit list (Lines 6-7). \sys uses the commit list to iterate over shadow pages 
+at commit time, to copy the contents of shadow pages back to the main memory {\tt store}. 
 %
@@
 %
-If the accessed page was previously modified and swapped out since the last boot (Line 8), the most recent version of the page is in {\tt
+If the accessed page was previously modified and swapped out since the last
+power failure (Line 8), the most recent version of the page is in {\tt
 \underline{shadow}}, not \texttt{\underline{store}} (Line 9).
@@
 
-When the last task in a coalesced group completes, \sys must commit \emph{all} dirty pages in the \texttt{working} buffer and the swapped out pages in \texttt{\underline{shadow}} buffer back to their original locations in \texttt{\underline{store}}.
-If a task is allowed to (re-)execute after only \emph{some} pages have been committed to \texttt{\underline{store}}, that task might see protected variables in an inconsistent state.
+When the last task in a sequence of coalesced tasks completes, \sys must commit
+\emph{all} dirty pages in the \texttt{working} buffer and in the
+\texttt{\underline{shadow}} buffer, copying them back to their original
+locations in the main memory \texttt{\underline{store}}. If a task were allowed
+to (re-)execute after only some pages have been committed to
+\texttt{\underline{store}}, the task might see protected variables in an
+inconsistent state; \sys's two-phase commit prevents this inconsistency.
 
-To make the commit atomic, \sys commits the pages in two phases, implemented by Algorithm~\ref{algo:commit}.
+To make the commit atomic, \sys commits the pages in two phases, implemented by
+Algorithm~\ref{algo:commit}.
 %
-The first phase copies dirty pages from the \texttt{working} buffer to the non-volatile \texttt{\underline{shadow}} buffer (Line 4).
-The second phase commits pages from \texttt{\underline{shadow}} to \texttt{\underline{store}} (Line 12).
-If power fails during the first phase, the whole commit is aborted, and the dynamic coalesced task restarts from its first static task.
-If power fails in the second phase, the commit will be safely resumed on next boot.
-The non-volatile metadata for the second phase are: the \texttt{\underline{committing}} bit, which is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16); \texttt{\underline{shadowCount}} counter, which is cleared once the phase completes (Line 14); \texttt{\underline{commitIndex}}, which indexes the next page to be committed (Line 11) and is cleared at the end of the phase (Line 15).
-
-\sys non-volatile \texttt{\underline{shadowList/shadowCount/commitIndex}} implementation is a more efficient alternative to a simpler implementation that keeps a valid bit in each entry of \texttt{\underline{shadow}} and, on every boot into second phase, traverses all entries, committing the pages with a set bit, and clearing the bit.
-During the second commit phase, pages are not physically copied from \texttt{\underline{shadow}} to \texttt{\underline{store}}, but a table pointing at the right buffer for each page is updated.
-This is explained in more details in Section~\ref{sec:implementation}.
+The first phase copies dirty pages from the \texttt{working} buffer to the
+non-volatile \texttt{\underline{shadow}} buffer (Line 4). The second phase
+commits pages from \texttt{\underline{shadow}} to \texttt{\underline{store}}
+(Line 12).  If power fails during the first phase, the whole commit is aborted,
+and the execution restarts from the most recently committed point (i.e., from
+the first task in a coalesced sequence of tasks). If power fails in the second
+phase, the commit process safely resumes on reboot.  The second phase of commit
+depends on some run-time metadata. The \texttt{\underline{committing}} bit
+indicates that a commit is in progress and is set before the first page is
+committed (Line 9) and cleared after the last page is committed (Line 16).  The
+\texttt{\underline{shadowCount}} records the number of dirty shadow pages to
+commit and \sys clears the counter when commit completes (Line 14).  The
+\texttt{\underline{commitIndex}} indexes the next page to be committed (Line
+11) and \sys clears the index at the end of the phase (Line 15).
+%
+\sys's commit is efficient because it maintains an index dirty pages instead of
+iterating through all potentially dirty pages to check a dirty bit during
+commit. 
+%
+\sys's implementation is also efficient because the memory manager does not
+copy from \texttt{\underline{shadow}} to \texttt{\underline{store}}, instead
+swapping pointers in an indirection table that maintains the pages as a double
+buffer.  Section~\ref{sec:implementation} describes double buffering in more
+detail.
 
@@
 
-\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent protected data.
-During task execution, modifications to protected data do not affect their original memory locations, because a task reads and writes the volatile \texttt{working} buffer only, and modified pages are kept in \texttt{\underline{shadow}} until commit.
-A power failure erases the contents of the page buffer, preventing a re-executing task from observing updates in the page buffer from a previous execution attempt.
-Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14), ensures that all accesses to protected variables in subsequent tasks correctly access their original memory locations in \texttt{\underline{store}}.
+\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only
+ever executes using consistent protected data. During task execution,
+modifications to protected data do not affect their original memory locations,
+because a task reads and writes the volatile \texttt{working} buffer only, and
+modified pages are kept in \texttt{\underline{shadow}} until commit.  A power
+failure erases the contents of the page buffer, preventing a re-executing task
+from observing updates in the page buffer from a previous execution attempt.
+Clearing \texttt{\underline{shadowCount}} as part of the second phase commit
+(Line 14), ensures that all accesses to protected variables in subsequent tasks
+correctly access their original memory locations in \texttt{\underline{store}}.
 \sys assumes that the write to a non-volatile word in memory is atomic.
@@
 
-One design choice \sys made was to insert dynamic check for paging on every
-possible problematic memory operation using {\tt RP} and {\tt WP} macros.
-Since dynamic checks incur overhead on every read and write, it has a higher
-overhead for accessing memory than a static approach,
-e.g., Alpaca, where a system statically analyze the memory operation and try to insert only optimal
-number of memory protection code (usually at the beginning of each task).
+\sys asks the programmer to use its {\tt RP} and {\tt WP} paging API methods on
+every access to a protected variable.  These API invocations present a risk of
+high overhead because there is a dynamic check on every read and write. An
+alternative approach explored by prior work~\cite{alpaca} relied on a compiler
+to analyze memory operations and try to insert memory
+protection code mostly at the beginning of a task, leaving most memory accesses uninstrumented.
 
-However, \sys's dynamic paging has several benefits over static approach.
-First, static approach has limitations, prohibiting certain types of programming pattern. For example,
-in the presence of arbitrary pointer operation, function calls using function pointer, or an interrupt
-within a task, the system cannot statically analyze the memory behavior. Thus, static
-approaches either prohibits the programmer from using certain programming pattern~\cite{alpaca}, 
-or will have to conservatively insert memory protection code that can be unnecessary.
-Second, static approach cannot handle task coalescing, because with coalescing the scope to be
-searched statically is not determined at compile time. 
-By choosing to check every memory access dynamically, \sys supports programming pattern that
-Alpaca does not support, such as arbitrary pointer operation, while also being able to
-support task coalescing, which yields extra performance benefits.
+Despite the risk of per-access overhead, \sys's dynamic memory protection
+scheme brings several benefits over a static approach.  First, the limitations
+of static analysis preclude some uses of pointers due to potential pointer
+aliasing. For example, in the presence of arbitrary pointer operation, a
+function call using a function pointer, or an interrupt within a task, the
+system cannot statically analyze the memory behavior.   Second, a static approach
+cannot handle task coalescing because a protected variable's lifetime (i.e., from first use to commit) is unknown 
+at compile time. \sys's dynamic, per-access instrumentation supports arbitrary use of pointers 
+and enables task coalescing.
 %

### 2018-04-17 paper/hwswsetup.tex
slight fix

@@
 same in \sys and Alpaca implementations of the benchmarks. We compare
-both systems using GCC, the compiler with the best performance for the
-MSP430 architecture.
+both systems using GCC.
 
@@
 allocations to make a private copy of the affected variables. For array
-variables, instrumentation also includes adding code and memory allocations to
-track the first write to the array.
+variables, instrumentation also includes adding code for 
+dynamically tracking the array access.
 

### 2018-04-17 paper/paging.tex
reduced algorithms

@@
 
-\begin{algorithm}[t]
-	\caption{\texttt{RP}(variable $v$)}
+%\begin{algorithm}[t]
+\begin{wrapfigure}{t}{0.4\textwidth}
+\fbox{\parbox{\linewidth}{
+	\captionof{algorithm}{\texttt{RP}(variable $v$)}
 	\label{algo:rwar}
@@
 		\State $t \leftarrow \Call{GetTag}{v}$ 
-        \State $i \gets \{ j\ |\ \Call{GetTag}{\texttt{working}[j]} = t \}$ \Comment{Look in the page buffer}
-		\If { $i = \emptyset$ } \Comment{Is the variable in a resident page?}
-		\State	$i \gets \Call{PageFault}{t}$ \Comment{Page in from non-volatile memory}
+        \State $i \gets \{ j\ |\ \Call{GetTag}{\texttt{working}[j]} = t \}$ \Comment{Search page}
+		\If { $i = \emptyset$ } \Comment{Var in a resident page?}
+		\State	$i \gets \Call{PageFault}{t}$ \Comment{Swap page in}
 		\EndIf
 		\State $o \leftarrow \Call{GetOffset}{v}$ 		
-		\State \Return \texttt{working}[$i$][$o$]  \Comment{Return directly from page buffer}
-	\end{algorithmic}
-\end{algorithm}
+		\State \Return \texttt{working}[$i$][$o$]  \Comment{Return from page}
+%	\end{algorithmic}
+%\end{algorithm}
+	\end{algorithmic}}}
+\end{wrapfigure}
 
@@
 
-\begin{algorithm}
-	\caption{\texttt{PageFault}(tag $t$)}
+%\begin{algorithm}
+\begin{wrapfigure}{t}{0.5\textwidth}
+\fbox{\parbox{\linewidth}{
+	\captionof{algorithm}{\texttt{PageFault}(tag $t$)}
 	\label{algo:pagefault}
@@
 	\begin{algorithmic}[1]
-        \State \LeftComment{$i_\text{victim}$: volatile index to next free slot or page to be evicted, set to 0 at boot}
-        \State \LeftComment{\texttt{\underline{shadowCount}}: non-volatile, reset to 0 on boot unless in phase 2 commit}
+        \State \LeftComment{$i_\text{victim}$: index to free slot or page to be evicted, initially 0}
+        \State \LeftComment{\texttt{\underline{shadowCount}}: non-volatile, 0 on boot unless in phase 2 commit}
         \State $t_\text{victim} \gets \Call{Tag}{\texttt{working}[i_\text{victim}]}$
-		\If {\Call{isDirty}{working[$i_\text{victim}$]} }	\Comment{Are we evicting a dirty page?}
+		\If {\Call{isDirty}{working[$i_\text{victim}$]} }	\Comment{Evicting a dirty page?}
 		\State $\texttt{\underline{shadow}}[t_\text{victim}] \xleftarrow{\text{DMA}} \texttt{working}[i_\text{victim}]$
             \Comment{Swap out the page}
-        \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track as list for quick traversal}
+        \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track for quick traversal}
             \State $\texttt{\underline{shadowList}}[\texttt{shadowCount++}] \gets t_\text{victim}$
@@
 		\EndIf
-		\If {\Call{exists}{\underline{shadow}[$t$]}} \Comment{Had the page been swapped out earlier?}
+		\If {\Call{exists}{\underline{shadow}[$t$]}} \Comment{Page swapped out earlier?}
 		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
-            \Comment{Get page from commit buffer}
+            \Comment{Get from commit buffer}
 		\Else
 		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{store}}[t]$
-            \Comment{Get page from backing store}
+            \Comment{Get from backing store}
 		\EndIf 
-        \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBufLength}|$ \Comment{FIFO eviction policy}
-	\end{algorithmic}
-\end{algorithm}
+        \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBufLength}|$ \Comment{FIFO evict}
+%	\end{algorithmic}
+%\end{algorithm}
+	\end{algorithmic}}}
+\end{wrapfigure}
 
@@
 
-\begin{algorithm}[t]
+%\begin{algorithm}[t]
+\begin{wrapfigure}{t}{0.5\textwidth}
+\fbox{\parbox{\linewidth}{
+	\captionof{algorithm}{\texttt{PageFault}(tag $t$)}
 	\caption{Two-phase commit}
@@
 	\begin{algorithmic}[1]
-        \Procedure{CommitPhase1}{} \Comment{Invoked from \texttt{next\_task} on last coalesced task}
+        \Procedure{CommitPhase1}{} \Comment{On last coalesced task}
             \For {$i \in 0..|\texttt{working}|-1$}
@@
         \EndProcedure
-        \Procedure{CommitPhase2}{}
+        \Procedure{CommitPhase2}{} \Comment{\texttt{\underline{commitIndex}} 0 on first boot}
             \State $\texttt{\underline{committing}} \gets \textbf{true}$
-            \While{\texttt{\underline{commitIndex}} < \texttt{\underline{shadowCount}}} \Comment{\texttt{\underline{commitIndex}} set to 0 at first boot}
+            \While{\texttt{\underline{commitIndex}} < \texttt{\underline{shadowCount}}}
                 \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{commitIndex}}]$
@@
         \EndProcedure
-        \Procedure{OnBoot}{} \Comment{Invoked on every boot, before transitioning to a task}
+        \Procedure{OnBoot}{} \Comment{Invoked on every boot}
             \If { \texttt{\underline{committing}} } \Call{CommitPhase2}{}
@@
         \EndProcedure
-	\end{algorithmic}
-\end{algorithm}
+%	\end{algorithmic}
+%\end{algorithm}
+	\end{algorithmic}}}
+\end{wrapfigure}
 

### 2018-04-17 paper/implementation.tex

--- 2018-04-16 paper/implementation.tex
@@
-We proceed with the explanation of the implementation of \sys programming model, which is light and whose API adds very little on top of a plain C language.
+We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittence-safe using \sys.
 
@@
 
-Coala's API adds only five syntactic constructs to a C-based language: (i) \texttt{COALA\_TASK}, to declare a new task, (ii) \texttt{COALA\_NEXT\_TASK}, to mark the next task to be run after the currently executing one completes, (iii) \texttt{COALA\_PV}, to declare a new protected variable, (iv) \texttt{RP} and \texttt{WP}, to mark protected variables' reads and writes, and (v) \texttt{COALA\_SM}, to declare \texttt{typedef}'d \texttt{struct} members, when used as protected variables, to ensure proper memory alignment required by \sys's page handler.
-Moreover, the programmer is required to explicitly call two API functions inside the main program body: (i) \texttt{COALA\_INIT}, to pass as an argument the global initialization task to run upon the first boot, and (ii) \texttt{COALA\_RUN}, which hands the control to \sys's task manager.
+Coala's API adds only five syntactic constructs to a C-based language.
+\texttt{COALA\_TASK} is a function annotation that declares a function to be a
+task. \texttt{COALA\_NEXT\_TASK}, indicates the task to run next, after an
+executing task completes.  \texttt{COALA\_PV} declare a new protected variable.
+The programmer uses \texttt{RP} and \texttt{WP}, to read and write protected
+respectively.  \texttt{COALA\_SM} identifies protected \texttt{struct} fields
+and tells \sys to ensure that the struct fields are page-aligned in memory.
+The programmer is required to explicitly call two API functions at the start of
+{\tt main()}: \texttt{COALA\_INIT}, which does global initialization upon first
+boot and \texttt{COALA\_RUN} which hands the control to \sys's task manager to
+resume task execution.
 
-Table~\ref{table:coala_api} summarizes \sys's API methods and their arguments. These methods trigger specific \sys kernel functions, whose explanation is given in the following sections.
+Table~\ref{table:coala_api} summarizes \sys's API methods and their arguments.
 
@@
 
-\subsection{Kernel Functions}
+\subsection{API Details}
 
-\textbf{New Tasks.} The API method \texttt{COALA\_TASK} statically allocates a non-volatile constant variable holding the task weight, and declares a void function with the name passed as an argument.
+\textbf{New Tasks.} The \texttt{COALA\_TASK} annotation on a function
+declaration statically allocates a non-volatile constant variable holding a
+task's weight and declares that the function is a task.
 
-\textbf{New Protected Variables.} The API method \texttt{COALA\_PV} statically allocates (and aligns in memory) a non-volatile variable, whose address will be used by \texttt{RP} and \texttt{WP} API methods to locate the most recent location of the variable itself, which could be either the location of the variable in volatile memory or its location in non-volatile memory.
+\textbf{Protected Variables.} The \texttt{COALA\_PV} annotation on a
+variable statically allocates and aligns a protected non-volatile variable. The
+variable must then be accessed \texttt{RP} and \texttt{WP} API methods at
+runtime to ensure correct operation.
 
-\textbf{Task Transitions.} By calling \texttt{COALA\_NEXT\_TASK}, the programmer tells \sys's task manager which task to execute next.
-The implementation of \sys is built on the assumption that the currently executing task is not interrupted, and only once it completes the next task is run.
+\textbf{Task Transitions.} A call to \texttt{COALA\_NEXT\_TASK} immediately
+transfers control to a specified next task and this API can be invoked along
+any arbitrary control path (i.e., to dynamically determine the next task).
+Like prior work~\cite{chain,dino,ratchet}, \sys's task control-flow model
+assumes that an executing task is never interrupted.
 
-\textbf{Protected Reads and Writes.} When \texttt{RP} and \texttt{WP} API methods are used, \sys searches the volatile memory for the page the protected variable belongs to.
-If such page is not found in RAM, then a page fault occurs, and the page is fetched from the non-volatile memory.
-The two functions return a handle to the variable's volatile memory location.
-Unlike \texttt{RP}, \texttt{WP} marks the page being written as dirty, so that it can be copied to its non-volatile location during the commit phase.
+\textbf{Protected Reads and Writes.} The programmer uses the \texttt{RP} and \texttt{WP} API
+methods to access protected variables.  These methods cause \sys to find a variable's page.
+If the page is not resident in working memory, the method leads to a page fault,
+which fetches the page from non-volatile memory. Unlike \texttt{RP},
+\texttt{WP} marks the variable's page as dirty, allowing for efficient 
+commit back to non-volatile memory.
 
-\textbf{Initialization and Origin Task.} The behavior of the API method \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the addition that all the preliminary kernel initializations are performed inside this function.
-Moreover, the origin task is scheduled only upon the first boot.
+\textbf{Initialization.} The behavior of the API method
+\texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the
+addition that all the preliminary kernel initializations are performed inside
+this function.
 
@@
 
-\textbf{Execution.} The application control is passed to \sys's execution manager by calling \texttt{COALA\_RUN} at the end of the main function.
-Upon reboot, the coalescing target is initialized according to the coalescing strategy.
-Then, the second phase of the commit is resumed, in case the last power failure occurred during this operation.
-The shadow buffer list is cleared, and the metadata of the volatile buffer is initialized.
-The pointer to the last coalescing task is recalled from non-volatile memory.
-Before running, \sys checks whether there is any partial execution to resume, in which case the volatile state, including the program counter, is restored.
-Otherwise, coalescing starts: the current budget is replenished; tasks are run, the history is increased and the current budget is decreased; when the latter is zero, coalescing stops, the pointer to the next task is saved and the commit phase starts.
-Coalescing resumes after commit phase two ends and the target is updated based on the coalescing strategy.
-\todo{Go over keywords (coalescing target, current budget, commit phases etc.) and make them consistent with the rest of the paper}{Amjad, Carlo}
+\textbf{Execution.} The programmer passes control to \sys's task scheduler 
+by calling \texttt{COALA\_RUN} after device initialization. 
+%
+First the scheduler updates the coalescing target using the coalescing
+strategy.  Second, the scheduler resumes a commit, if one is in progress. Third,
+the scheduler clears the list of dirty shadow pages. The scheduler then 
+sets the program counter to the next task to run, which \sys tracks in  
+non-volatile memory.  
+%
+Before executing the task, \sys checks whether there is a partially committed
+task to resume, which requires \sys to restore volatile state, including the program
+counter.  
+%
+If there is no in-progress, partially committed task, \sys starts executing and
+coalescing tasks. 
+%\todo{Go over keywords (coalescing target, current budget, commit phases etc.) and make them consistent with the rest of the paper}{Amjad, Carlo}
 
 \subsection{Paging}
+\label{sec:impl:paging}
 
-As described in Section~\ref{sec:memory_virtulaization}, \sys's memory virtualization is implemented using three buffers: \texttt{working} (volatile) buffer, \texttt{\underline{store}} (non-volatile) buffer and \texttt{\underline{shadow}} (non-volatile) buffer.
-The \texttt{\underline{store}} buffer holds the latest safe copy of all the pages, and it is read-only up to, and including, the ultimate commit.
-Upon reboot, the \texttt{working} buffer is empty, and it gets filled with pages from non-volatile memory on demand.
-All reads and writes to protected variables happen in volatile memory.
-The \texttt{\underline{shadow}} buffer serves as an intermediate non-volatile buffer where dirty, volatile pages are copied during the intermediate commit.
-Moreover, it contains pages evicted from SRAM when a \emph{full page fault} occurs.
+As described in Section~\ref{sec:memory_virtulaization}, \sys's memory
+virtualization scheme uses three buffers: a volatile \texttt{working} buffer, a
+non-volatile main memory \texttt{\underline{store}}, and a non-volatile
+\texttt{\underline{shadow}} page buffer. 
+%
+The main memory \texttt{\underline{store}} holds the last committed copy of all 
+pages. 
+%
+After reboot, the \texttt{working} buffer is empty. Execution fills 
+the working buffer with pages as the program accesses them. 
+%
+The program primarily reads and writes data paged into the volatile working memory. 
+%
+The non-volatile \texttt{\underline{shadow}} buffer serves as an intermediate,
+non-volatile backing store for dirty pages evicted from the space-constrained
+working buffer.
 
-When a \sys application accesses a protected variable with either \texttt{RP} or \texttt{WP}, \sys has to search for the variable in RAM first, and then, if not found, in non-volatile memory.
-In reality, the page manager searches for pages, not single variables.
-A page search can be very fast if each page is identified by a unique tag, and if every variable belonging to a page bears the tag in its address.
-Thus, retrieving a page tag from a protected variable's address boils down to extracting the upper bits of the address itself.
-The remaining lower bits denote the variable's offset in its page.
-%The complete process is depicted in Figure~\ref{figure:coala_page_tags} \todo{this figure doesn't really show page tags and offsets, but paging in general}.
-We note that, in order to obtain unique page tags, some alignment requirements have to be met.
-First, the page size $S$ has to be a power of two.
-Moreover, pages have to be aligned by $S$.
-Note that, by aligning pages by $2^n$, the alignment is also valid for all sizes $2^x$, where $x \leq n$.
-For our experiments, we used 32-, 64-, 128- and 256-byte pages.
-16\,KB were allocated in non-volatile memory for \texttt{\underline{shadow}} and \texttt{\underline{store}} (8\,KB each), and the volatile page buffer was 1\,KB, half the total SRAM size.
-The main advantage of having large pages is searches through page buffers are shorter.
-On the other hand, page swaps are more bulky.
+To access a protected variable with either \texttt{RP}
+or \texttt{WP}, \sys searches for the variable in its working buffer first. If
+found, the program accesses the data in the working buffer.  If the page is not 
+in the working memory, the memory manager fetches it from non-volatile memory. 
+%
+\sys uses address-based page tagging to make finding a variable efficient.  The
+upper bits of a variable's memory address identify its page, and the lower bits
+denote the variable's offset in its page. The total number of pages in memory,
+$p$, determines the number of tag bits, which is $log(P)$.
 
+Page tagging imposes a data alignment requirement.
+Page size $S$ must be a power of two. 
+%
+Pages must be aligned to an $S$-byte boundary. 
+%
+We experimented with 32-, 64-, 128- and 256-byte pages, an 8kB non-volatile \texttt{\underline{shadow}} buffer and
+an 8kB non-volatile \texttt{\underline{store}}, and a volatile working buffer
+of 1kB, which is half of the total SRAM size. 
 %\begin{figure}
@@
 
-\subsection{Coalescing Loop and Commit Phase}
+\subsection{Coalescing and Commit}
 
-As described above, the main coalescing loop alternates between executing tasks and saving progress.
-Static tasks are coalesced as long as the current budget is not null, after which the two commit phases are triggered, the coalescing target is updated and the current budget is replenished.
+\sys uses a two-phase commit process to atomically commit pages despite
+possible power interruptions during commit.  When a coalesced task completes,
+\sys backs up dirty pages in volatile working memory to the non-volatile
+shadow buffer using DMA. 
+%
+If power fails while backing up these pages, commit fails and the execution
+resumes from the last committed point.  After backing up pages from working
+memory to the shadow buffer, \sys sets a flag that indicates an on-going atomic
+commit.  At this point, commit is irrevocable and will continue until it
+completes despite power failures.   
+%
+Each page is double buffered, with one copy always in the shadow page buffer
+and one copy always in the main memory store.  \sys uses double buffering to
+commit dirty pages efficiently by swapping which buffer is which, without
+copying any data.
+%
 
-\textbf{Intermediate Commit and Ultimate Commit.} The reason why the commit process is split in two phases is two-fold: ensure consistency by using atomic operations, and preserve performance by enabling an incremental commit mechanism.
-When a coalesced task is successfully run to completion, all the dirty pages in SRAM have to be persistently saved to the non-volatile shadow buffer.
-This is what we call \emph{intermediate commit}: it has to be completed in one shot, and if power fails during this phase the application will have to resume from the beginning of the last coalesced task.
-When the intermediate commit is done, a flag can be atomically raised to signify that a valid, persistent copy of dirty pages is available in non-volatile memory.
-From this point on, the commit process can be resumed even if power fails.
-To understand the last phase, referred to as \emph{ultimate commit}, it has to be mentioned that store and shadow buffers are virtual.
-By default, \sys allocates two memory banks for non-volatile pages, $M_0$ and $M_1$, and all the pages have their initial store copy in $M_0$.
-A dirty page that during the intermediate commit, or due to eviction, is copied to its shadow location in $M_x$, will end up having its store location in $M_x$ after the ultimate commit.
-In other words, during the ultimate commit dirty pages are not physically moved, instead their store and shadow locations are swapped.
-This trick makes \sys save some energy during this phase, avoiding redundant page moves.
-As for page swaps, the intermediate commit makes use of the on-board DMA.
+\subsection{Task Downscaling}
 
-\subsection{Partial Execution}
+\sys's task downscaling preserves forward progress by partially committing an
+ongoing, non-terminating task. 
+%
+Partial commit relies on the existence of a hardware timer, which is a very
+common feature in a typical MCU and consumes very little energy (i.e.,
+$\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).
 
-Among \sys's novelties, the task-splitting feature is the most effective when it comes to preservation of forward progress.
-Its implementation adds hardly any computational overhead, and is supported by a hardware timer running ``in the background'', whose energy consumption is extremely low, i,e.
-$\approx$3\,$\mu$ A/MHz~\cite{msp430datasheet}.
-The goal is to interrupt a static task if it cannot be run to a completion with a full capacitor's energy, commit its partial results, save volatile state and program counter and continue.
-\sys speculates and assumes that a static task is too large and has to be split when power fails two times consecutively in the same static task, and the coalescing target is one, because, after the first power failure, the large task is run with a fresh energy budget.
-To ``split'' a static task, \sys schedules a timer interrupt.
-If this interrupt occurs before the capacitor is depleted, \sys commits all the dirty volatile pages (using the two-phase commit explained above), saves stack and volatile registers to non-volatile memory and then continues with the task.
-In case a power failure occurs after this operation, but before the static task completes, the execution at the next reboot starts exactly where \sys had saved the partial execution.
-A task can be split in multiple sub-tasks, as many as needed to progress forward.
-Regarding the timer interrupt, it is initially scheduled to fire after a large time interval.
-If power fails before the interrupt fires, \sys reduces the interval.
-If the partial execution is committed successfully, \sys enlarges the interval.
-These adjustments follow a binary search technique, and make the interval converge very fast to a value optimal for preserving forward progress and avoiding redundant commits.
-
-For some applications, specific pieces of a task cannot be split and have to be run atomically.
-For instance, when sampling an analog signal, continuity between samples has to be ensured, otherwise the digital array would represent a wrong, non-continuous input signal.
-In such cases, the programmer can use two dedicated API guards before and after the atomic piece of code, namely \texttt{COALA\_DISABLE\_PE} and \texttt{COALA\_ENABLE\_PE}, to respectively halt and resume the partial execution timer.
+ \textbf{Partial Commit and Task Atomicity.} Some applications prevent
+downscaling a task because of a need for task atomicity.  For example, sampling
+an analog signal requires consecutive samples at a known interval, or the
+digitally sampled signal is meaningless.  In such a case, the programmer can
+disable partial commit for a task or a span of code, marking  the code with a
+pair of \texttt{COALA\_DISABLE\_PE} and \texttt{COALA\_ENABLE\_PE} annotations.
+These annotations respectively halt and resume the partial execution timer.
 

### 2018-04-17 paper/paper.tex
ROO title
add the adaptive model section
change the title
yet another title
wrapfig for reasonable fig size
Compiler section commented out
placeholder/braindump for coala implementation
added a todo and changed the title
added placeholder for partial commit
change title
change section name
add brain dump about Coala the execution model
methodology secion almost done
added some things to paging
push the system overview to the top
New section titles made consistent
month of PACMPL corrected
Table 4 revised and ready to be filled in with data
Sections 5 and 6 revised
solve conflict
discussion sec

--- 2018-04-10 paper/paper.tex
@@
 \acmJournal{PACMPL}
-\acmVolume{1}
+\acmVolume{X}
 \acmNumber{OOPSLA} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
+\acmArticle{X}
 \acmYear{2018}
-\acmMonth{1}
+\acmMonth{10}
 \acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
@@
 \setcopyright{none}
+
+\bibliographystyle{ACM-Reference-Format}
+%\bibliographystyle{plain}
+\citestyle{acmauthoryear}
 
@@
 \usepackage{hyphenat}
+\usepackage{wrapfig}
+\usepackage{multirow}
+\usepackage{capt-of}
 
@@
 
-\newcommand{\transition}{{\tt next\_task}\xspace}
+\newcommand{\transition}{{\tt coala\_next\_task}\xspace}
 \newcommand{\sys}{Coala\xspace}
-\newcommand{\sysmed}{\sys: Adaptive Task Coalescing for Intermittent Computing $\ldots$}
-\newcommand{\sysfull}{\sys: Adaptive Task Coalescing for Intermittent Computing on Energy-harvesting Devices}
+\newcommand{\sysmed}{\sys: Adaptive Task-Based Intermittent Execution}
+\newcommand{\sysfull}{\sys: Adaptive Task-Based Intermittent Execution}
+%\newcommand{\sysmed}{\sys: Adaptive Intermittent Execution Model$\ldots$}
+%\newcommand{\sysfull}{\sys: Adaptive Intermittent Execution Model for Energy-harvesting Devices}
+%\newcommand{\sysmed}{\sys: Adaptive Task Coalescing for Intermittent Computing $\ldots$}
+%\newcommand{\sysfull}{\sys: Adaptive Task Coalescing for Intermittent Computing on Energy-harvesting Devices}
 %newcommand{\sysfull}{\sys: Efficient Task-based Intermittent Computing with Adaptive Task Coalescing and Memory Virtualization} %Old title
@@
 
-\author{Brandon Lucia}
-\authornote{}
-\orcid{}
-\affiliation{%
-	\institution{Carnegie Mellon University}
-	\streetaddress{4720 Forbes Avenue}
-	\city{Pittsburgh} 
-	\state{PA} 
-	\postcode{15213}
-}
-\email{blucia@andrew.cmu.edu}
-
 \author{Carlo Delle Donne}
@@
 \email{k.s.yildirim@tudelft.nl}
+
+\author{Brandon Lucia}
+\authornote{}
+\orcid{}
+\affiliation{%
+	\institution{Carnegie Mellon University}
+	\streetaddress{4720 Forbes Avenue}
+	\city{Pittsburgh} 
+	\state{PA} 
+	\postcode{15213}
+}
+\email{blucia@andrew.cmu.edu}
 
@@
 
-\ccsdesc[500]{Computer systems organization~Embedded software}
-\ccsdesc[300]{Computer systems organization~Availability}
-\ccsdesc[300]{Software and its engineering~Scheduling}
-\ccsdesc[300]{Software and its engineering~Communications management}
-\ccsdesc[300]{Software and its engineering~Embedded software}
-
-\keywords{Energy harvesting, transient operation, operating system}
+%\ccsdesc[500]{Computer systems organization~Embedded software}
+%\ccsdesc[300]{Computer systems organization~Availability}
+%\ccsdesc[300]{Software and its engineering~Scheduling}
+%\ccsdesc[300]{Software and its engineering~Communications management}
+%\ccsdesc[300]{Software and its engineering~Embedded software}
+%
+%\keywords{Energy harvesting, transient operation, operating system}
 
@@
 
-\section{\sys: System Overview}
-\label{sec:overall_system}
+\section{\sys Programming Model and \sys System Overview}
+\label{sec:systemdescription}
 
@@
 
-\section{Task Coalescing}
-\label{sec:task_coalescing}
+\section{\sys Task Adaptation}
+\label{sec:task_adaptation}
 
-%\input{coalescing}
-\input{newCoalescing}
+\input{coalescing}
 
-\section{Memory Virtualization}
+\input{partialcommit}
+
+\section{\sys Memory Management}
+% \section{\sys Memory Virtualization}
 \label{sec:memory_virtulaization}
@@
 
-\section{Automatic Task Generation}
-\label{sec:compiler}
+\section{\sys Implementation}
+\label{sec:implementation}
 
-\input{compiler}
+\input{implementation}
 
-\section{Discussion}
-\label{sec:discussion}
+%\section{Discussion}
+%\label{sec:discussion}
 
-\input{discussion}
+%\input{discussion}
 
-\section{Methodology and Evaluation}
-\label{sec:methodology_evaluation}
+\section{Methodology}
+\label{sec:methodology}
+
+\input{hwswsetup}
+
+\section{\sys Evaluation}
+\label{sec:evaluation}
 
@@
 
-\bibliographystyle{ACM-Reference-Format}
-%\bibliographystyle{plain}
 \bibliography{bib}
-\citestyle{acmauthoryear}
 

### 2018-04-17 paper/background.tex

@@
 
-Energy harvesting devices operate using energy extracted from sources such as radio frequency transmissions and solar energy.
-These devices elide tethered power or a battery, instead collecting energy into a capacitor, operating when sufficient energy accumulates, and upon depleting the energy, turning off and recharging.
+Energy harvesting devices operate using energy extracted from sources such as
+radio frequency transmissions and solar energy. These devices elide tethered
+power or a battery, instead collecting energy into a capacitor, operating when
+sufficient energy accumulates, and upon depleting the energy, turning off and
+recharging.
 %
@@
 %
-There are several energy harvesting battery-less hardware platforms.
-For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
-Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
+There are several energy harvesting battery-less hardware platforms. For
+instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine}
+WISP~\cite{wisp5} (with its variants such as
+WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
+NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
+such as~\cite{medusa_farsens_2017}. Other platforms include ambient backscatter
+tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less
+phone~\cite{talla_imwut_2017}.
 
@@
 
-\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
-We assume a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM.
-Our implementation leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
-\sys do not require architectural additions to commodity processors as in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.
-\sys supports I/O behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and unsafe, asynchronous I/O.
+\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and
+future energy-harvesting platforms based around general purpose, commodity
+computing components~\cite{wisp,msp430datasheet}. We assume a device with a
+memory system that has fast, byte-addressable volatile and non-volatile memory;
+in particular, our target platform, WISP~\cite{wisp}, is equipped with a
+mixture of SRAM and FRAM. Our implementation leverages hardware support for
+fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys do not
+require architectural additions to commodity processors as
+in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}. \sys supports I/O
+behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and
+unsafe, asynchronous I/O.
 
@@
 %
-Software running on an energy-harvesting device executes {\em intermittently} because energy sources are not always available to harvest and buffer sufficient operational energy.
-An intermittent execution is composed of operating periods interspersed with power failures~\cite{dino,chain,alpaca,ratchet}.
-The frequency of failures depends on the size of the device's energy storage buffer: a larger buffer allows longer operating periods.
+Software running on an energy-harvesting device executes {\em intermittently}
+because energy sources are not always available to harvest and buffer
+sufficient operational energy. An intermittent execution is composed of
+operating periods interspersed with power
+failures~\cite{dino,chain,alpaca,ratchet}. The frequency of failures depends on
+the size of the device's energy storage buffer: a larger buffer allows longer
+operating periods. 
 %Energy-harvesters provide input power orders of magnitude less than operating power, making recharging negligible during operation.
 %Intermittent execution is different from continuous execution. 
-A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists.
-Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
-Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.
+A power failure clears volatile state (e.g., registers and SRAM) while
+non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
+to a prior point in the execution: by default, to the beginning of {\tt
+main()}. Early intermittent systems preserved progress by periodically
+checkpointing volatile execution context to non-volatile
+memory~\cite{mementos}, sometimes using hardware
+support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.
 
@@
 
-Checkpointing volatile state alone do not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}.
-Especially, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependence between operations that manipulate 
-non-volatile memory, because it allows {\em writes} from a failed attempt to persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}
----see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution 
-using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
-Additional method to maintain data consistent in the presence of WAR dependency is necessary for correct execution.
+Checkpointing volatile state alone do not ensure data consistency when the
+system can directly manipulate non-volatile memory~\cite{mspcdino}.
+Especially, data can get inconsistent when code includes a
+\emph{write-after-read} (WAR) dependence between operations that manipulate
+non-volatile memory, because it allows {\em writes} from a failed attempt to
+persist across failure and be read by the {\em
+reads}~\cite{ratchet,dino,alpaca} ---see
+Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
+inconsistent in an intermittent execution using the cyclic redundancy check
+(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
+maintain data consistent in the presence of WAR dependency is necessary for
+correct execution.
 %Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}.
@@
 
-Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into tasks,
-which is a region of code that can contain arbitrary computation, sensing, and communication. 
-Task-based models keep progress at the granularity of tasks by repeatedly trying to re-execute
-each task on power failure until it successfully finishes, only then moving on to the next task.
-Since these models do not rely on taking an expensive checkpoint, it is usually faster than the
-checkpoint-based solution~\cite{chain, alpaca}.
-\sys also follows the paradigm of the task-based programming model, where the programmer
-explicitly expresses the application as a sequence of tasks and the transition between them.
+Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to
+decompose their program into tasks, which is a region of code that can contain
+arbitrary computation, sensing, and communication.  Task-based models keep
+progress at the granularity of tasks by repeatedly trying to re-execute each
+task on power failure until it successfully finishes, only then moving on to
+the next task.  Since these models do not rely on taking an expensive
+checkpoint, it is usually faster than the checkpoint-based
+solution~\cite{chain, alpaca}.  \sys also follows the paradigm of the
+task-based programming model, where the programmer explicitly expresses the
+application as a sequence of tasks and the transition between them.
 %Task-based runtime systems ensure that repeated task executions are idempotent. 
@@
 
-Task-based models also suffer from data inconsistency when there is a WAR dependence.
-Prior systems tried to tackle the problem by a compiler-automated redo-logging for the variables
-that is part of the dependence~\cite{alpaca}, or by statically creating a multiple copies of the problematic variable
-to ensure that no task reads and writes the same copy~\cite{chain}.
+Task-based models also suffer from data inconsistency when there is a WAR
+dependence.  Prior systems tried to tackle the problem by a compiler-automated
+redo-logging for the variables that is part of the dependence~\cite{alpaca}, or
+by statically creating a multiple copies of the problematic variable to ensure
+that no task reads and writes the same copy~\cite{chain}.
 %There are several run time strategies to ensuring task idempotence.
@@
 Prior systems, including the task-based models, suffer from two major overheads:
-{\em frequent non-volatile memory access} and {\em incorrectly size tasks}.
+{\em frequent non-volatile memory access} and {\em incorrectly sized, inefficient or non-terminating tasks}.
 
@@
 \textbf{Frequent Non-volatile Memory Access.} 
-Previous systems frequently access the non-volatile memory directly, which we claim is a major source of inefficiency.
-Checkpoint-based models copy the entire volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile
-memory on every checkpoint.
-Task-based models, which do not take checkpoints, still have to copy a subset of the non-volatile data to maintain memory consistent in WAR dependence (e.g., redo-logging).
-Table~\ref{table:chechpoint_comparison} summarizes the data that gets copies to and from the non-volatile memory for memory consistency and forward progress.
-Ratchet~\cite{ratchet} and Clank~\cite{hicks_isca_2017}, which uses a whole non-volatile memory, obviously also has a high number of non-volatile memory access
-for normal operations in addition to taking checkpoints.
+Previous systems frequently access the non-volatile memory directly, which 
+is a potential source of inefficiency.  Checkpointing systems copy all volatile
+state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at
+every checkpoint.  Task-based models, which do not take checkpoints, back up a
+subset of non-volatile data to maintain memory consistency using, e.g.,
+redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that
+gets copies to and from the non-volatile memory in service of memory
+consistency and progress preservation.  Ratchet~\cite{ratchet} and
+Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile
+perform high frequency non-volatile memory accesses, in to checkpointing
+frequently.
 
@@
 %\end{figure}
-To understand the expense of non-volatile memory access, we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory 
-operations for the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device 
-across 1600 read and write accesses to each memory type, randomizing to avoid caching. Our experiment showed that a non-volatile memory access 
-consumes around 1.5 times the energy of a volatile access. The result shows that non-volatile memory access is more expensive than the volatile
-memory access, and prior works with frequent non-volatile memory accesses is inefficient.
+We measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
+the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor
+voltage drop on a capacitor powering the device across 1600 read and write
+accesses to each memory, randomizing accesses to avoid caching effects. Our
+data showed that a non-volatile memory access consumes around 1.5 times the
+energy of a volatile access. The data confirm that, because non-volatile memory
+accesses are more expensive than volatile memory accesses, prior systems that
+frequently access non-volatile memory may be artificially inefficient.
 %~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
@@
 \todo{add citation to TI specification on FRAM energy cost}{Przemek}
-Based on this observation, one of the primary motivation of \sys is to promote accessing SRAM more frequently rather than FRAM in order to 
-reduce memory access energy overhead and to increase the time between power failures.
+\sys is motivated by this observation: \sys 
+moves data pages from FRAM to SRAM to maximize SRAM accesses and minimize FRAM
+accesses, to decrease energy consumption and increase the time between power
+failures.
 
@@
 
-\textbf{Incorrectly Sized Tasks.} Compiler-inserted checkpoints or programmer-defined tasks can both be incorrect or inefficient.
-If the code between checkpoints or the task (for brevity, we will refer to both as ''tasks'' in this section) consumes {\em more} 
-energy than the device can buffer, the task will never complete 
-using buffered energy {\em preventing forward progress}, making the system incorrect. 
+\textbf{Incorrectly Sized Tasks.} Compiler-inserted checkpoints or
+programmer-defined tasks can both both incorrect and non-terminating or
+inefficient.  If code between consecutive checkpoints or in a task (we refer to
+both scenarios as ''code in a task'' in this section) consumes {\em more}
+energy than the fixed, maximum energy that the device can buffer, then the task
+will never be able to complete using buffered energy.  Such a task is
+non-terminating, {\em prevents forward progress}, and makes the program
+incorrect. 
 %
-If the task consumes far {\em less} energy than a device can buffer, the system may operate {\em inefficiently},
-checkpointing or transitioning between tasks too often than needed. 
+If the task consumes far {\em less} energy than a device can buffer, the system
+may operate {\em inefficiently}, checkpointing or transitioning between tasks
+more often than needed. 
 %
-Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is challenging both to the
-programmer and the compiler, since estimating the exact energy use of an arbitrary code is complicated.
-When {\em porting} apps from one device to another, the problem gets much worse because
-an excessively costly task on one device may be a relatively short task on a device with a larger buffer, and vice versa. 
-Buffer sizes of different platforms vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
+Avoiding excessively costly, non-terminating tasks and short, high-overhead
+tasks is challenging both to the programmer and the compiler, because
+estimating the exact energy use of an arbitrary code is complicated.  When {\em
+porting} a program from one platform to another, the problem becomes worse
+because an excessively costly task on one device may be a relatively short task
+on a device with a larger buffer.  Buffer sizes on different platforms vary
+widely from, e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
 complicating the problem.
 %
-As a solution, \sys's execution model presents {\em task coalescing}, which merges multiple tasks to amortize the overhead
-when tasks are too small, and {\em task downscaling}, which breaks up a task when the task is too large. 
+As a solution, \sys's execution model presents {\em task coalescing}, which
+merges multiple tasks to amortize the overhead when tasks are too small, and
+{\em task downscaling}, which breaks up a task when the task is too large to complete. 
 %When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).

### 2018-04-17 paper/implementation.tex
impl: shorten
impl: typesetting: log not the product of l o and g

@@
 
-Coala's API adds only five syntactic constructs to a C-based language.
+Coala's API adds only five syntactic constructs to a C-based language, summarized
+in Table~\ref{table:coala_api}.
 \texttt{COALA\_TASK} is a function annotation that declares a function to be a
@@
 resume task execution.
-
-Table~\ref{table:coala_api} summarizes \sys's API methods and their arguments.
 
@@
 denote the variable's offset in its page. The total number of pages in memory,
-$p$, determines the number of tag bits, which is $log(P)$.
+$p$, determines the number of tag bits, which is $\log P$.
 

### 2018-04-17 paper/implementation.tex
impl: specify mcu

@@
 an 8kB non-volatile \texttt{\underline{store}}, and a volatile working buffer
-of 1kB, which is half of the total SRAM size. 
+of 1kB, which is half of the total SRAM size in the MSP430FR6989 microcontroller.
 %\begin{figure}

### 2018-04-17 paper/implementation.tex
impl: missing preposition
impl: interruptible by async events

@@
 
+\textbf{Task Transitions.} A call to \texttt{COALA\_NEXT\_TASK} immediately
+transfers control to a specified next task and can be invoked along
+any control path to dynamically determine the next task.
+Like prior work~\cite{chain,dino,ratchet}, \sys's control-flow model
+assumes that an executing task is not interruptible by asynchronous events.
+
 \textbf{Protected Variables.} The \texttt{COALA\_PV} annotation on a
 variable statically allocates and aligns a protected non-volatile variable. The
-variable must then be accessed \texttt{RP} and \texttt{WP} API methods at
+variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
 runtime to ensure correct operation.
-
-\textbf{Task Transitions.} A call to \texttt{COALA\_NEXT\_TASK} immediately
-transfers control to a specified next task and this API can be invoked along
-any arbitrary control path (i.e., to dynamically determine the next task).
-Like prior work~\cite{chain,dino,ratchet}, \sys's task control-flow model
-assumes that an executing task is never interrupted.
 

### 2018-04-17 paper/implementation.tex
impl: shorten

@@
 \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the
-addition that all the preliminary kernel initializations are performed inside
-this function.
+addition of performing preliminary kernel initializations, including hardware setup.
 

### 2018-04-17 paper/implementation.tex
impl: remove redundant text

@@
-We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittence-safe using \sys.
+We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittence-safe.
 
@@
 in Table~\ref{table:coala_api}.
-\texttt{COALA\_TASK} is a function annotation that declares a function to be a
-task. \texttt{COALA\_NEXT\_TASK}, indicates the task to run next, after an
-executing task completes.  \texttt{COALA\_PV} declare a new protected variable.
-The programmer uses \texttt{RP} and \texttt{WP}, to read and write protected
-respectively.  \texttt{COALA\_SM} identifies protected \texttt{struct} fields
-and tells \sys to ensure that the struct fields are page-aligned in memory.
-The programmer is required to explicitly call two API functions at the start of
-{\tt main()}: \texttt{COALA\_INIT}, which does global initialization upon first
-boot and \texttt{COALA\_RUN} which hands the control to \sys's task manager to
-resume task execution.
+
+%\texttt{COALA\_TASK} is a function annotation that declares a function to be a
+%task. \texttt{COALA\_NEXT\_TASK}, indicates the task to run next, after an
+%executing task completes.  \texttt{COALA\_PV} declare a new protected variable.
+%The programmer uses \texttt{RP} and \texttt{WP}, to read and write protected
+%respectively.  \texttt{COALA\_SM} identifies protected \texttt{struct} fields
+%and tells \sys to ensure that the struct fields are page-aligned in memory.
+%The programmer is required to explicitly call two API functions at the start of
+%{\tt main()}: \texttt{COALA\_INIT}, which does global initialization upon first
+%boot and \texttt{COALA\_RUN} which hands the control to \sys's task manager to
+%resume task execution.
 
@@
 
-\subsection{API Details}
+%\subsection{API Details}
 

### 2018-04-17 paper/paging.tex
paging: remove stray alg caption
paging: fwd ref to API

@@
 \fbox{\parbox{\linewidth}{
-	\captionof{algorithm}{\texttt{PageFault}(tag $t$)}
-	\caption{Two-phase commit}
+	\captionof{algorithm}{Two-phase commit}
+	%\caption{Two-phase commit}
 	\label{algo:commit}
@@
 \sys asks the programmer to use its {\tt RP} and {\tt WP} paging API methods on
-every access to a protected variable.  These API invocations present a risk of
-high overhead because there is a dynamic check on every read and write. An
-alternative approach explored by prior work~\cite{alpaca} relied on a compiler
-to analyze memory operations and try to insert memory
-protection code mostly at the beginning of a task, leaving most memory accesses uninstrumented.
+every access to a protected variable (cf. Section~\ref{sec:coala_api}).  These
+API invocations present a risk of high overhead because there is a dynamic
+check on every read and write. An alternative approach explored by prior
+work~\cite{alpaca} relied on a compiler to analyze memory operations and try to
+insert memory protection code mostly at the beginning of a task, leaving most
+memory accesses uninstrumented.
 
@@
 system cannot statically analyze the memory behavior.   Second, a static approach
-cannot handle task coalescing because a protected variable's lifetime (i.e., from first use to commit) is unknown 
-at compile time. \sys's dynamic, per-access instrumentation supports arbitrary use of pointers 
-and enables task coalescing.
+cannot handle task coalescing, because a protected variable's lifetime (i.e.,
+from first use to commit) is unknown at compile time. \sys's dynamic,
+per-access instrumentation supports arbitrary use of pointers and enables task
+coalescing.
 %

### 2018-04-17 paper/intro.tex
fig 1 to 0.4 column
intro: tighten the contribs

@@
 
-\begin{wrapfigure}{t!}{0.5\textwidth}
+\begin{wrapfigure}{t!}{0.4\textwidth}
     \centering
-    \includegraphics[width=0.5\textwidth]{figures/intro-figure.pdf}
+    \includegraphics[width=0.4\textwidth]{figures/intro-figure.pdf}
     \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
@@
 \begin{itemize}
-\item The first {\em adaptive} task-based intermittent execution model that dynamically coalesces tasks to avoid unnecessary commit overhead. 
-\item The first power interrupts immune memory virtualization mechanism for an intermittent system that \sys uses to preserve the atomicity of coalesced tasks.
-\item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
+\item The first {\em adaptive} task-based intermittent execution model that
+dynamically \emph{coalesces} tasks to reduce commit overhead and
+\emph{downscales} tasks to ensure forward progress.
+\item The first memory virtualization mechanism that preserves the atomicity of coalesced tasks under power failures.
+%\item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
 \item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication via~\cite{coala_website}. 
-\item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to diverse energy conditions.
+\item An evaluation directly comparing \sys to a state of the art task-based
+intermittent programming and execution model from prior work, showing that on a
+suite of benchmarks from the literature and several new workloads, \sys often
+has higher performance and is more flexible to diverse energy conditions. 
 \end{itemize}
@@
 Implementation details of \sys are give in Section~\ref{sec:implementation}.
-Section~\ref{sec:discussion} discusses \sys
-design issues. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe
-\sys's methodology and evaluation, respectively. Section~\ref{sec:related_work} puts \sys in the
+Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's methodology and evaluation, respectively.
+Section~\ref{sec:related_work} puts \sys in the
 context of related work and Section~\ref{sec:conclusions} concludes and discusses

### 2018-04-17 paper/intro.tex
intro

--- 2018-04-16 paper/intro.tex
@@
 %intermittent platform with a moderate effort.}
+%
 
 \textbf{Contributions.} To summarize, \sys's main contributions are: 
-
+%
 \begin{itemize}
-\item The first {\em adaptive} task-based intermittent execution model that dynamically coalesces tasks to avoid unnecessary commit overhead. 
-\item The first software memory virtualization mechanism for an intermittent system that \sys uses to preserve the atomicity of coalesced tasks.
-\item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
+\item The first {\em adaptive} task-based intermittent execution model that
+dynamically \emph{coalesces} tasks to reduce commit overhead and
+\emph{downscales} tasks to ensure forward progress.
+\item The first memory virtualization mechanism that preserves the atomicity of coalesced tasks under power failures.
+%\item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
 \item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication via~\cite{coala_website}. 
-\item An evaluation directly comparing \sys to a state of the art task-based intermittent programming and execution model from prior work, showing that on a suite of benchmarks from the literature and several new workloads, \sys often has higher performance and is more flexible to diverse energy conditions.
+\item An evaluation directly comparing \sys to a state of the art task-based
+intermittent programming and execution model from prior work, showing that on a
+suite of benchmarks from the literature and several new workloads, \sys often
+has higher performance and is more flexible to diverse energy conditions. 
 \end{itemize}
-
+%
 Section~\ref{sec:background} provides background on intermittent computing.
@@
 Sections~\ref{sec:task_adaptation} describe \sys's task adaptation mechanism (coalescing and downscaling).
-Memory virtualization mechanisms is provided in Section~\ref{sec:memory_virtulaization}.
+Memory virtualization mechanism is provided in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are give in Section~\ref{sec:implementation}.
-Section~\ref{sec:discussion} discusses \sys
-design issues. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe
-\sys's methodology and evaluation, respectively. Section~\ref{sec:related_work} puts \sys in the
+Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's methodology and evaluation, respectively.
+Section~\ref{sec:related_work} puts \sys in the
 context of related work and Section~\ref{sec:conclusions} concludes and discusses

### 2018-04-17 paper/abstract.tex

@@
 % Connecting the work to a real-world problem
-%Intermittently powered computers promise a sustainable Internet of things by forgo the batteries and scarf the energy from the environment. 
-Massively deployed Internet of Things (IoT) must be energy neutral. This can only be achieved when the IoT is battery-free and the energy to operate the IoT comes from ambient energy sources.
+Embedded devices in the Internet of Things (IoT) gain new deployment
+opportunities when their hardware is free of batteries and powered by ambient
+energy.
 % The main problem that we are tackling
-However, relying on the highly varying, unpredictable and intermittent harvested energy is at odds with the core assumption that program executions require stable and uninterrupted energy supply. 
+Energy harvested from the environment, however, is unpredictable
+and can only power a device intermittently.
 % The already proposed solution to the main problem
-The core programming model for the intermittently-powered computation is a discrete-like one that alternates between progressing and protecting the execution progress against power interrupts.
+In the intermittent execution model, computation is frequently interrupted by
+power failures, and to make forward progress, the system must save intermediate
+program state into non-volatile memory.
+
 % The problem of the proposed solutions 
-Generally, intermittent computation models can be classified into (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based programming, where a programmer splits the code into small idempotent sections.
-Despite the superiority of task-based systems over checkpointing, task-based programming is completely oblivious to the energy conditions.
-Simply, it cannot increase or decrease the size of the task on the fly (the moment of commit to the non-volatile memory) depending on how long the energy burst is.
+Task-based programming models that execute at the granularity of
+statically-defined tasks, offer an efficient alternative to checkpointing.
+%
+Existing task-based systems, however, fix their task size at compile time and
+operate without any regard to the changing energy conditions at runtime.
+%
+When tasks are sized without accounting for available energy, state may be
+persisted either more frequently than unnecessary or not frequently enough for
+the program to make progress and terminate.
+%
 % Our main claim 
-To address this challenge, this paper introduces \sys, an \emph{adaptive} task-based execution model that does not require any dedicated hardware to support task adaptability, which at the same time is \emph{efficient}.
+To address this challenge, we propose Coala, an \emph{adaptive} task-based
+execution model that colesces or downscales static tasks at runtime in response
+to changing energy availability.
 % show off
-\sys programs tune the frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period off the programmer's shoulders.
 % main benefits of Coala
-\sys defaults to progress on a multi-tasks scale, when energy permits (\emph{efficiency}), and on a sub-task scale when needed (\emph{guarantee}).
-Furthermore, \sys's \emph{memory virtualization} gains \sys full data protection against runtime triggered and compile-time defined memory accesses.
-% I really hope it will support our claims!! haha
-Results show that \sys is more efficient than the current state-of-art execution model by up to a factor of two and able to progress where its counterparts fail.
+Coala ensures that power failures cannot leave the program state in
+non-volatile memory inconsistent using a specialized memory virtualization
+mechanism.
+%
+Our evaluation on a real energy-harvesting device demonstrates the benefits of
+adaptive task coalescing and downscaling, and highlights Coala's performance
+improvements over a state-of-art task-based system.

### 2018-04-17 paper/hwswsetup.tex
pass over hw

@@
 		\hline\hline
-        \textbf{ar} & & & Activity recognition using a nearest neighbor
+        \textbf{ar} &10 &500 & Activity recognition using a nearest neighbor
 classifier with a deterministic input from a randomly-generated sequence of
@@
 		\hline
-        \textbf{bc} & & & Counts set bits in a pseudo-random sequence of bytes
+        \textbf{bc} &10 &419 & Counts set bits in a pseudo-random sequence of bytes
 using several algorithms and compares results for correctness\\
 		\hline
-        \textbf{cuckoo} & & & Inserts and queries a sequence of pseudo-random
+        \textbf{cuckoo} &14 &499 & Inserts and queries a sequence of pseudo-random
 values in a Cuckoo filter, a set membership data structure with deletion\\
 		\hline
-        \textbf{dijkstra} &  & & Dijkstra shortest path algorithm executed on a
+        \textbf{dijkstra} &5 &227 & Dijkstra shortest path algorithm executed on a
 predefined adjacency matrix\\
 		\hline
-        \textbf{fft} & & & Fast Fourier Transform on three pre-generated 128
+        \textbf{fft} &8 &516 & Fast Fourier Transform on three pre-generated 128
 sample vectors\\
 		\hline
-		\textbf{sort} & & & Selection sort of a fixed sequence of numbers\\
+		\textbf{sort} &181 &4 & Selection sort of a fixed sequence of numbers\\
 		\hline
@@
 \caption{Description and characteristics of benchmark applications used to
-evaluate \sys. \todo{SLOC and Number of Tasks}{Amjad}}
+evaluate \sys.}
 \label{table:benchmark_table}

### 2018-04-17 paper/evaluation.tex

@@
 energy- and task-aware coalescing. All coalescing strategies reduce total
-overhead and maximize useful work. EA and ETA perform better than EB.
-\todo{rename legend entry 'user code' with 'useful task code'; rename y-axis to
-'Time usage (\%)'; increase font by
-two more points everywhere}{Amjad}}
+overhead and maximize useful work. EA and ETA perform better than EB.}
 	\label{fig:overallOverheadBreakdown}
@@
 accesses into three types with different costs. NC: no coalescing, EA:
-energy-aware coalescing, ETA: energy- and task-aware coalescing. \todo{rename
-legend entry 'task code' to 'user code'; rename y-axis to 'Load (\%)'}{Amjad}}
+energy-aware coalescing, ETA: energy- and task-aware coalescing.}
 	\label{fig:coalEfficiency}
@@
 
+\begin{figure}
+	\centering
+	\includegraphics[width=0.49\columnwidth]{figures/page_exec-time}
+	\includegraphics[width=0.49\columnwidth]{figures/pagePulls}
+    \caption{Effect of page size on execution time (left) and on the number of
+page faults (right). The execution time is normalized to the lowest in each set
+of page sizes.}
+	\label{fig:page_size}
+\end{figure}
+
 \subsubsection{Effect of Page Size on Runtime}
@@
 page size has no effect.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.49\columnwidth]{figures/page_exec-time}
-	\includegraphics[width=0.49\columnwidth]{figures/pagePulls}
-    \caption{Effect of page size on execution time (left) and on the number of
-page faults (right). The execution time is normalized to the lowest in each set
-of page sizes.}
-	\label{fig:page_size}
-\end{figure}
 

### 2018-04-17 paper/paging.tex
abstract, and had to move figures to fix layout
expand algorithm 3 and 4 to make them more readable, sill within the pages limit
remove the red text on top

@@
-\textcolor{red}{[additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain.
+% \textcolor{red}{[additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain.
 And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
@@
 
-%\begin{algorithm}
-\begin{wrapfigure}{t}{0.5\textwidth}
-\fbox{\parbox{\linewidth}{
-	\captionof{algorithm}{\texttt{PageFault}(tag $t$)}
+If the \texttt{working} buffer is full, a page fault on a memory access
+requires \sys to swap out one of the pages in the volatile buffer (a 
+\emph{victim page}) preserving updates made to that page, and to copy the
+accessed page of protected data into the working buffer.
+Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault. A victim
+page in the \texttt{working} buffer is selected for eviction according to
+a \emph{first-in-first-out} (FIFO) replacement policy (Lines 2 and 12). 
+%
+While we opted for the simplicity of FIFO, any other replacement policy, such
+as Least Recently Used (LRU) could work in its place.
+
+% \begin{algorithm}
+% \begin{wrapfigure}{t}{0.5\textwidth}
+% \fbox{\parbox{}{
+% 	\captionof{algorithm}{\texttt{PageFault}(tag $t$)}
+% 	\label{algo:pagefault}
+% 	\scriptsize
+\begin{algorithm}
+	\caption{\texttt{PageFault}(tag $t$)}
 	\label{algo:pagefault}
@@
         \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBufLength}|$ \Comment{FIFO evict}
+	\end{algorithmic}
+\end{algorithm}
+
 % 	\end{algorithmic}
-%\end{algorithm}
-	\end{algorithmic}}}
-\end{wrapfigure}
+% \end{algorithm}
+	% \end{algorithmic}}}
+% \end{wrapfigure}
 
-If the \texttt{working} buffer is full, a page fault on a memory access
-requires \sys to swap out one of the pages in the volatile buffer (a 
-\emph{victim page}) preserving updates made to that page, and to copy the
-accessed page of protected data into the working buffer.
-Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault. A victim
-page in the \texttt{working} buffer is selected for eviction according to
-a \emph{first-in-first-out} (FIFO) replacement policy (Lines 2 and 12). 
-%
-While we opted for the simplicity of FIFO, any other replacement policy, such
-as Least Recently Used (LRU) could work in its place.
-%
 If a task modified any byte in the victim page (i.e., using \texttt{WP}), then
@@
 
-%\begin{algorithm}[t]
-\begin{wrapfigure}{t}{0.5\textwidth}
-\fbox{\parbox{\linewidth}{
-	\captionof{algorithm}{Two-phase commit}
-	%\caption{Two-phase commit}
-	\label{algo:commit}
-	\scriptsize
-	%\small
-	\begin{algorithmic}[1]
-        \Procedure{CommitPhase1}{} \Comment{On last coalesced task}
-            \For {$i \in 0..|\texttt{working}|-1$}
-                \State $t \gets \Call{Tag}{\texttt{working}[i]}$
-                \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
-                \State $\texttt{\underline{shadowList}}[\texttt{\underline{shadowCount}}] \gets t$
-                \State $\texttt{\underline{shadowCount}} \gets \texttt{\underline{shadowCount}} + 1$
-            \EndFor
-            \State \Call{CommitPhase2}{}
-        \EndProcedure
-        \Procedure{CommitPhase2}{} \Comment{\texttt{\underline{commitIndex}} 0 on first boot}
-            \State $\texttt{\underline{committing}} \gets \textbf{true}$
-            \While{\texttt{\underline{commitIndex}} < \texttt{\underline{shadowCount}}}
-                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{commitIndex}}]$
-                \State \Call{commitToStore}{\texttt{\underline{shadow}}[t]}
-                \State $\texttt{\underline{commitIndex}} \gets \texttt{\underline{commitIndex}} + 1$
-            \EndWhile
-            \State $\texttt{\underline{shadowCount}} \gets 0$
-            \State $\texttt{\underline{commitIndex}} \gets 0$
-            \State $\texttt{\underline{committing}} \gets \textbf{false}$
-        \EndProcedure
-        \Procedure{OnBoot}{} \Comment{Invoked on every boot}
-            \If { \texttt{\underline{committing}} } \Call{CommitPhase2}{}
-            \EndIf 
-            \State $\texttt{\underline{shadowCount}} \gets 0$, $i_\text{victim} \gets 0$
-        \EndProcedure
-%	\end{algorithmic}
-%\end{algorithm}
-	\end{algorithmic}}}
-\end{wrapfigure}
-
 %\subsection{Paging Example}
@@
 %despite arbitrary power failures.
+
+%\begin{algorithm}[t]
+% \begin{wrapfigure}{t}{0.5\textwidth}
+% \fbox{\parbox{\linewidth}{
+% 	\captionof{algorithm}{Two-phase commit}
+% 	%\caption{Two-phase commit}
+% 	\label{algo:commit}
+% 	\scriptsize
+\begin{algorithm}[t]
+	\caption{Two-phase commit}
+	\label{algo:commit}
+	\scriptsize
+	%\small
+	\begin{algorithmic}[1]
+        \Procedure{CommitPhase1}{} \Comment{On last coalesced task}
+            \For {$i \in 0..|\texttt{working}|-1$}
+                \State $t \gets \Call{Tag}{\texttt{working}[i]}$
+                \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
+                \State $\texttt{\underline{shadowList}}[\texttt{\underline{shadowCount}}] \gets t$
+                \State $\texttt{\underline{shadowCount}} \gets \texttt{\underline{shadowCount}} + 1$
+            \EndFor
+            \State \Call{CommitPhase2}{}
+        \EndProcedure
+        \Procedure{CommitPhase2}{} \Comment{\texttt{\underline{commitIndex}} 0 on first boot}
+            \State $\texttt{\underline{committing}} \gets \textbf{true}$
+            \While{\texttt{\underline{commitIndex}} < \texttt{\underline{shadowCount}}}
+                \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{commitIndex}}]$
+                \State \Call{commitToStore}{\texttt{\underline{shadow}}[t]}
+                \State $\texttt{\underline{commitIndex}} \gets \texttt{\underline{commitIndex}} + 1$
+            \EndWhile
+            \State $\texttt{\underline{shadowCount}} \gets 0$
+            \State $\texttt{\underline{commitIndex}} \gets 0$
+            \State $\texttt{\underline{committing}} \gets \textbf{false}$
+        \EndProcedure
+        \Procedure{OnBoot}{} \Comment{Invoked on every boot}
+            \If { \texttt{\underline{committing}} } \Call{CommitPhase2}{}
+            \EndIf 
+            \State $\texttt{\underline{shadowCount}} \gets 0$, $i_\text{victim} \gets 0$
+        \EndProcedure
+	\end{algorithmic}
+\end{algorithm}
+% 	\end{algorithmic}}}
+% \end{wrapfigure}
+
 \subsection{Dynamic Paging of \sys}

### 2018-04-17 paper/coalescing.tex
Fig 5 updated
minor modifications

@@
 % Modifiying Carlo's pseudo code a little 
-\begin{algorithm}[t]
-	\caption{Coalescing}
+%\begin{algorithm}[t]
+%	\caption{Coalescing}
+%	\label{algo:genCoalescing}
+%	\scriptsize
+%	%\small
+%	\begin{algorithmic}[1]
+%        \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$H$}
+%        \State $H \gets 0$ 
+%        \While{$\texttt{true}$}       
+%	        \State $ i \gets 0$
+%	        %
+%	        \While{$ i <= B$}
+%		        \State \Call{execute\_task}{$T_\texttt{i}$}
+%		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
+%		        \State $i \gets i + W$
+%				\State $H \gets H + W$
+%	        \EndWhile
+%	        %
+%	        \State \Call{commit\_to\_fram}{\null}
+%	        \State $B \leftarrow $ \Call{$f_\text{commit}$}{$B$}
+%        \EndWhile
+%	\end{algorithmic}
+%\end{algorithm}
+
+\begin{wrapfigure}{t}{0.3\textwidth}
+\fbox{\parbox{\linewidth}{
+	\captionof{algorithm}{Coalescing}
 	\label{algo:genCoalescing}
@@
         \EndWhile
-	\end{algorithmic}
-\end{algorithm}
+	\end{algorithmic}}}
+\end{wrapfigure}
+%	\begin{subfigure}[t]{\linewidth}
+%	\end{subfigure}
 %
@@
 
-\subsubsection{Energy-oblivious Coalescing (EB)}
+\subsubsection{Energy-oblivious Coalescing (EO)}
 \label{subsec:energyBlind}
-
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/slowCoal}
-	\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
-	\label{fig:energyBlind}
-\end{wrapfigure}
 
@@
 \end{itemize}
+
+\begin{wrapfigure}{t!}{0.5\textwidth}
+	\centering
+	\includegraphics[width=0.5\columnwidth]{figures/slowCoal}
+	\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
+	\label{fig:energyBlind}
+\end{wrapfigure}
 
@@
 The History-Guided (HG) coalescing strategy adapts the coalescing target more
-quickly, addressing a key limitation of the EB coalescing strategy.  
+quickly, addressing a key limitation of the EO coalescing strategy.  
 %
-EA gives each task unity weight, as with EB.  
+EA gives each task unity weight, as with EO.  
 %
-HG differs from EB in how it updates $B$ on commit, and after a power failure.
+HG differs from EO in how it updates $B$ on commit, and after a power failure.
 %
@@
 the top, the figure shows the $i^{th}$ execution period, which has a history
-value of $H^{(i)} = 6$.  $B$ is initially three and decreases to one.  Once the
+value of $H^{(i)} = 4$.  $B$ is initially two and decreases to one.  Once the
 value of $B$ reaches one, HG is, in effect, expecting a power failure,
@@
 	\includegraphics[width=0.5\columnwidth]{figures/energy-aware-coal.pdf}
-	\caption{placeholder}
+	\caption{The {\em history-guided coalescing} algorithm uses its recent history of execution to estimate the available amount of energy and set the coalescing target accordingly.}
 	\label{fig:energyAware}
@@
 %
-The EB and HG coalescing strategies assume each task has uniform energy cost:
+The EO and HG coalescing strategies assume each task has uniform energy cost:
 for these strategies, $B$ simply corresponds to a target {\em count} of tasks
@@
 %
-WHG differs extends EB and HG to explicitly account for the variation in energy
+WHG differs extends EO and HG to explicitly account for the variation in energy
 cost of a program's tasks, eliminating the assumption that all tasks do the

### 2018-04-17 paper/hwswsetup.tex
minor modifications

@@
 		\hline\hline
-        \textbf{ar} &10 &500 & Activity recognition using a nearest neighbor
+        \textbf{ar} &10 &428 & Activity recognition using a nearest neighbor
 classifier with a deterministic input from a randomly-generated sequence of
@@
 		\hline
-        \textbf{bc} &10 &419 & Counts set bits in a pseudo-random sequence of bytes
+        \textbf{bc} &10 &371 & Counts set bits in a pseudo-random sequence of bytes
 using several algorithms and compares results for correctness\\
 		\hline
-        \textbf{cuckoo} &14 &499 & Inserts and queries a sequence of pseudo-random
+        \textbf{cuckoo} &14 &426 & Inserts and queries a sequence of pseudo-random
 values in a Cuckoo filter, a set membership data structure with deletion\\
 		\hline
-        \textbf{dijkstra} &5 &227 & Dijkstra shortest path algorithm executed on a
+        \textbf{dijkstra} &5 &198 & Dijkstra shortest path algorithm executed on a
 predefined adjacency matrix\\
 		\hline
-        \textbf{fft} &8 &516 & Fast Fourier Transform on three pre-generated 128
+        \textbf{fft} &8 &449 & Fast Fourier Transform on three pre-generated 128
 sample vectors\\
 		\hline
-		\textbf{sort} &181 &4 & Selection sort of a fixed sequence of numbers\\
+		\textbf{sort} &167 &4 & Selection sort of a fixed sequence of numbers\\
 		\hline

### 2018-04-17 paper/intro.tex
abstract, and had to move figures to fix layout
introduction typo

@@
 
-\begin{wrapfigure}{t!}{0.4\textwidth}
-    \centering
-    \includegraphics[width=0.4\textwidth]{figures/intro-figure.pdf}
-    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
-On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
-    \label{fig:coalesce}
-\end{wrapfigure}
-
 \textbf{Data Consistency in Intermittent Computing.}  
@@
 in this paper. A key limitation of prior software approaches that we address in this paper is the {\em inflexibility} of a program statically decomposed into tasks.
+
+\begin{wrapfigure}{t!}{0.4\textwidth}
+    \centering
+    \includegraphics[width=0.4\textwidth]{figures/intro-figure.pdf}
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
+On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
+    \label{fig:coalesce}
+\end{wrapfigure}
 
@@
 To ensure consistency and respect inter-task dependences, \sys uses a novel
-approach \emph{power failure immune memory virtualization} that buffers non-volatile variable
+approach \emph{power failure-immune memory virtualization} that buffers non-volatile variable
 updates in volatile memory during a coalesced task execution, before committing

### 2018-04-17 paper/abstract.tex
abstract with comments removed (to speedup pasting to hotcrp)
shorter abstract from alexei

@@
-% Connecting the work to a real-world problem
-Embedded devices in the Internet of Things (IoT) gain new deployment
-opportunities when their hardware is free of batteries and powered by ambient
-energy.
-% The main problem that we are tackling
-Energy harvested from the environment, however, is unpredictable
-and can only power a device intermittently.
-% The already proposed solution to the main problem
-In the intermittent execution model, computation is frequently interrupted by
-power failures, and to make forward progress, the system must save intermediate
-program state into non-volatile memory.
+%Old version of abstarct
 
-% The problem of the proposed solutions 
-Task-based programming models that execute at the granularity of
-statically-defined tasks, offer an efficient alternative to checkpointing.
-%
-Existing task-based systems, however, fix their task size at compile time and
-operate without any regard to the changing energy conditions at runtime.
-%
-When tasks are sized without accounting for available energy, state may be
-persisted either more frequently than unnecessary or not frequently enough for
-the program to make progress and terminate.
-%
-% Our main claim 
-To address this challenge, we propose Coala, an \emph{adaptive} task-based
-execution model that colesces or downscales static tasks at runtime in response
-to changing energy availability.
-% show off
-% main benefits of Coala
-Coala ensures that power failures cannot leave the program state in
-non-volatile memory inconsistent using a specialized memory virtualization
-mechanism.
-%
-Our evaluation on a real energy-harvesting device demonstrates the benefits of
-adaptive task coalescing and downscaling, and highlights Coala's performance
-improvements over a state-of-art task-based system.
+%Massively deployed Internet of Things (IoT) must be energy neutral.
+This can only be achieved when the IoT is battery-free and the energy to operate the IoT comes from ambient energy sources.
+However, relying on the highly varying, unpredictable and intermittent harvested energy is at odds with the core assumption that program executions require stable and uninterrupted energy supply.
+The core programming model for the intermittently-powered computation is a discrete-like one that alternates between progressing and protecting the execution progress against power interrupts.
+Generally, intermittent computation models can be classified into (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based programming, where a programmer splits the code into small idempotent sections.
+Despite the superiority of task-based systems over checkpointing, task-based programming is completely oblivious to the energy conditions.
+Simply, it cannot increase or decrease the size of the task on the fly (the moment of commit to the non-volatile memory) depending on how long the energy burst is.
+To address this challenge, this paper introduces \sys, an \emph{adaptive} task-based execution model that does not require any dedicated hardware to support task adaptability, which at the same time is \emph{efficient}.
+\sys programs tune the frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period off the programmer's shoulders.
+\sys defaults to progress on a multi-tasks scale, when energy permits (\emph{efficiency}), and on a sub-task scale when needed (\emph{guarantee}).
+Furthermore, \sys's \emph{memory virtualization} gains \sys full data protection against runtime triggered and compile-time defined memory accesses.
+Results show that \sys is more efficient than the current state-of-art execution model by up to a factor of two and able to progress where its counterparts fail.
+
+%Alexei's version of abstract
+
+Embedded devices in the Internet of Things (IoT) gain new deployment opportunities when their hardware is free of batteries and powered by ambient energy.
+Energy harvested from the environment, however, is unpredictable and can only power a device intermittently.
+In the intermittent execution model, computation is frequently interrupted by power failures, and to make forward progress, the system must save intermediate program state into non-volatile memory.
+Task-based programming models that execute at the granularity of statically-defined tasks, offer an efficient alternative to checkpointing.
+Existing task-based systems, however, fix their task size at compile time and operate without any regard to the changing energy conditions at runtime.
+When tasks are sized without accounting for available energy, state may be persisted either more frequently than unnecessary or not frequently enough for the program to make progress and terminate.
+To address this challenge, we propose \sys, an \emph{adaptive} task-based execution model that colesces or downscales static tasks at runtime in response to changing energy availability.
+\sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.
+Our evaluation on a real energy-harvesting device demonstrates the benefits of adaptive task coalescing and downscaling, and highlights \sys's performance improvements over a state-of-art task-based system.

### 2018-04-17 paper/background.tex
edits to background

@@
-%Here, we provide background on energy harvesting systems and the intermittent software execution model.
-We then discuss the key limitations of existing task-based execution models for intermittent computing that \sys addresses.
-
-%\begin{figure}
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	\begin{subfigure}[t]{\linewidth}
-		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
-		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
-	\end{subfigure}
-	\begin{subfigure}[t]{\linewidth}
-		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
-		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
-	\end{subfigure}
-	\caption{Code example demonstrating checkpointing gives rise to data inconsistency.}\label{fig:code_demo_incosistency}
-\end{wrapfigure}
-%\end{figure}
-
 \subsection{Energy Harvesting Systems}
@@
 
-%\textcolor{red}{reference work by Intel~\cite{baghsorkhi_cgo_2018}}
+\begin{wrapfigure}{t!}{0.5\textwidth}
+	\begin{subfigure}[t]{\linewidth}
+		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
+		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
+	\end{subfigure}
+	\begin{subfigure}[t]{\linewidth}
+		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
+		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
+	\end{subfigure}
+	\caption{Code example demonstrating checkpointing gives rise to data inconsistency.}\label{fig:code_demo_incosistency}
+\end{wrapfigure}
+%\end{figure}
 
-%Energy-harvesting devices execute software according to the {\em intermittent execution model}.
-Physically, a device charges until a threshold, operates briefly until its energy is depleted, shuts down recharges, and repeats the cycle.
-%
 Software running on an energy-harvesting device executes {\em intermittently}
@@
 	\end{tabular}
-	\caption{Non-volatile memory access for data consistency; \emph{PC}: program counter, \emph{WAR variables}: variables involved in WAR dependences, \emph{NV}: non-volatile.}
+	\caption{Non-volatile memory access for data consistency; \emph{PC}: program counter, \emph{WAR variables}: variables involved in Write after Read (WAR) dependences, \emph{NV}: non-volatile.}
 	\label{table:chechpoint_comparison}
@@
 %\end{figure}
-We measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
+As an illustration, we measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
 the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor

### 2018-04-17 paper/coalescing.tex
coalescing edits

@@
 volatile working memory and slower, non-volatile main memory; we defer the
-details of paging and commit to Section~\ref{sec:memory_virtualization}.
+details of paging and commit to Section~\ref{sec:memory_virtulaization}.
 
@@
 
-\noindent \textbf{Task Coalescing Strategies.} 
 % % Carlo pseudo code 
@@
 %
-Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing
-strategy. In the algorithm, $B$ (target budget) is total number of static tasks
-that \sys will next attempt to coalesce. $T_\text{i}$ is the $i^{th}$ task
+\noindent \textbf{Task Coalescing Strategies.} Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing strategy.
+In the algorithm, $B$ (target budget) is total number of static tasks
+that \sys will next attempt to coalesce. $T_\text{i}$ is the $i$th task
 executed since the last power failure. $H$ (history) is the number tasks
@@
 a coalesced sequence of tasks commits, and decreases by the same constant $x$
-when a coalesced sequence of tasks fails to complete (i.e., after a power
-failure).  The characteristic operations of the energy-oblivious strategy are:
+when a coalesced sequence of tasks fails to complete, i.e., after a power
+failure. The characteristic operations of the energy-oblivious strategy are:
+
+\begin{wrapfigure}{t!}{0.3\textwidth}
+	\centering
+	\includegraphics[width=0.3\columnwidth]{figures/slowCoal}
+	\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
+	\label{fig:energyBlind}
+\end{wrapfigure}
 
@@
 \end{itemize}
-
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/slowCoal}
-	\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
-	\label{fig:energyBlind}
-\end{wrapfigure}
 
@@
 \label{subsec:energyAware}
+
 The History-Guided (HG) coalescing strategy adapts the coalescing target more
@@
 
+\begin{wrapfigure}{}{0.5\textwidth}
+	\centering
+	\includegraphics[width=0.5\columnwidth]{figures/energy-aware-coal.pdf}
+	\caption{The {\em history-guided coalescing} algorithm uses its recent history of execution to estimate the available amount of energy and set the coalescing target accordingly.}
+	\label{fig:energyAware}
+\end{wrapfigure}
+
 The EA algorithm is characterized by the following functions:
@@
 are coalesced tasks.  A gray box is a task interrupted by a power failure.  At
-the top, the figure shows the $i^{th}$ execution period, which has a history
+the top, the figure shows the $i$th execution period, which has a history
 value of $H^{(i)} = 4$.  $B$ is initially two and decreases to one.  Once the
@@
 update $B$, setting $B$ equal to four and continuing execution.
-
-\begin{wrapfigure}{}{0.5\textwidth}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/energy-aware-coal.pdf}
-	\caption{The {\em history-guided coalescing} algorithm uses its recent history of execution to estimate the available amount of energy and set the coalescing target accordingly.}
-	\label{fig:energyAware}
-\end{wrapfigure}
 

### 2018-04-17 paper/coalescing.tex
coalescing edits
pass over coalescing

--- 2018-04-17 paper/coalescing.tex
@@
 \sys's design includes a novel task scheduler that uses efficient,
-energy-aware task coalescing to amortize task overheads, and uses a timer-based
+energy-aware task coalescing to amortize static task overheads, and uses a timer-based
 task-splitting mechanism to avoid non-termination of tasks too long
@@
 Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing
-strategy. In the algorithm, $B$ (target budget) is total number of static tasks
+strategy. In the algorithm, $B$ (target budget) is the total number of static tasks
 that \sys will next attempt to coalesce. $T_\text{i}$ is the $i^{th}$ task
@@
 The energy-oblivious strategy reacts slowly to the variation in the energy
-required to execute different tasks and to the in the effective quantum of
+required to execute different tasks and to the variation in the effective quantum of
 energy available to the device (i.e., variation in energy buffer size). 

### 2018-04-17 paper/coalescing.tex

--- 2018-04-17 paper/coalescing.tex
@@
 %
+
 \noindent \textbf{Task Coalescing Strategies.} Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing strategy.
@@
 Profiling the time and energy cost of tasks in a program is a difficult,
-orthogonal problem~\cite{cleancut,baghsorkhi_cgo_2018}.
+orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
 %

### 2018-04-17 paper/partialcommit.tex

@@
 Eventually, after some number of partial commits and power failures, the task
-will complete and execution will move on to a task with an energy demand
+will complete and execution will proceed to a task with an energy demand
 accommodated by the device's energy buffer. Task downscaling violates task

### 2018-04-17 paper/systemdescription.tex
move fig 3 up the page

@@
+\begin{wrapfigure}{t!}{0.45\textwidth}
+	\centering
+	\includegraphics[width=0.45\columnwidth]{figures/graffle/overview.pdf}
+	\caption{\sys top-level view. \sys is composed of two core components: \emph{Task Scheduler} and \emph{Memory Manager}.}
+	\label{fig:system_overview}
+\end{wrapfigure}
+
 \sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
@@
 In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.
-
-\begin{wrapfigure}{t!}{0.5\textwidth}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/graffle/overview.pdf}
-	\caption{\sys top-level view. \sys is composed of two core components: \emph{Task Scheduler} and \emph{Memory Manager}.\todo{Redraw figure to map new terms}{Carlo}}
-	\label{fig:system_overview}
-\end{wrapfigure}
 
@@
 using the device's buffered energy. Eventually, after a series of partial
-commits, the too-long task completes --- partial commit favors progress over task atomicity.  
+commits, the too-long task completes---partial commit favors progress over task atomicity.  
 When a task repeatedly fails, the scheduler uses a timer that counts an operating interval
@@
 atomically commit using a two-phase commit procedure accelerated using hardware
-support for direct memory access (DMA). Section~\ref{sec:memory_virtualization}
+support for direct memory access (DMA). Section~\ref{sec:memory_virtulaization}
 provides details of \sys's memory manager.

### 2018-04-17 paper/hwswsetup.tex
correct the table and provide ccs version number

@@
 %
-Execution time was measured using the instrumentation features in TI Code
-Composer Studio IDE.
+Execution time was measured using the clock features in TI Code
+Composer Studio IDE version 7.1.
 %
-\todo{provide version and more details on TI CCS}{Przemek}
 
@@
 		\hline
-		\textbf{sort} &167 &4 & Selection sort of a fixed sequence of numbers\\
+		\textbf{sort} &4 &167 & Selection sort of a fixed sequence of numbers\\
 		\hline

### 2018-04-17 paper/paging.tex
paging edits

@@
 % 	\scriptsize
+
 \begin{algorithm}[t]
@@
 API invocations present a risk of high overhead because there is a dynamic
-check on every read and write. An alternative approach explored by prior
-work~\cite{alpaca} relied on a compiler to analyze memory operations and try to
-insert memory protection code mostly at the beginning of a task, leaving most
+check on every read and write.
+An alternative approach~\cite{alpaca} relied on a compiler to analyze memory operations and try to insert memory protection code mostly at the beginning of a task, leaving most
 memory accesses uninstrumented.

### 2018-04-17 paper/abstract.tex

@@
 Existing task-based systems, however, fix their task size at compile time and operate without any regard to the changing energy conditions at runtime.
-When tasks are sized without accounting for available energy, state may be persisted either more frequently than unnecessary or not frequently enough for the program to make progress and terminate.
-To address this challenge, we propose \sys, an \emph{adaptive} task-based execution model that colesces or downscales static tasks at runtime in response to changing energy availability.
+When tasks are sized without accounting for available energy, state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
+To address these challenges, we propose \sys, an \emph{adaptive} task-based execution model that  progresses on a multi-tasks scale, when energy permits (\emph{efficiency}), and preseves the computation progress on a sub-task scale when needed.
 \sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.

### 2018-04-17 paper/implementation.tex
typos in implementation

@@
 %
-We experimented with 32-, 64-, 128- and 256-byte pages, an 8kB non-volatile \texttt{\underline{shadow}} buffer and
-an 8kB non-volatile \texttt{\underline{store}}, and a volatile working buffer
-of 1kB, which is half of the total SRAM size in the MSP430FR6989 microcontroller.
+We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
+an 8\,KB non-volatile \texttt{\underline{store}}, and a volatile working buffer
+of 1\,KB, which is half of the total SRAM size in the MSP430FR6989 microcontroller.
 %\begin{figure}

### 2018-04-17 paper/hwswsetup.tex
methodology typos

@@
 result also reported in prior work~\cite{alpaca} that the amount of energy
-available in one charge cycle --- the sum of stored energy and incoming energy
-during execution -- is the same at distances beyond 50~cm where incoming energy
+available in one charge cycle---the sum of stored energy and incoming energy
+during execution---is the same at distances beyond 50\,cm where incoming energy
 becomes negligible.

### 2018-04-17 paper/intro.tex
Final pass on intro and background

@@
 
-\begin{wrapfigure}{t!}{0.4\textwidth}
+\begin{wrapfigure}{t!}{0.45\textwidth}
     \centering
-    \includegraphics[width=0.4\textwidth]{figures/intro-figure.pdf}
+    \includegraphics[width=0.45\textwidth]{figures/intro-figure.pdf}
     \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
@@
 Section~\ref{sec:systemdescription} provides an overview of \sys, while
-Sections~\ref{sec:task_adaptation} describe \sys's task adaptation mechanism (coalescing and downscaling).
+Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism (coalescing and downscaling).
 Memory virtualization mechanism is provided in Section~\ref{sec:memory_virtulaization}.
-Implementation details of \sys are give in Section~\ref{sec:implementation}.
+Implementation details of \sys are given in Section~\ref{sec:implementation}.
 Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's methodology and evaluation, respectively.

### 2018-04-17 paper/evaluation.tex
correct the algo names

@@
 relative to the baseline without coalescing.
-NC: no coalescing, EB: energy-blind coalescing, EA: energy-aware coalescing, ETA:
+NC: no coalescing, EB: energy-blind coalescing, HG: history-guided coalescing, WHG:
 energy- and task-aware coalescing. All coalescing strategies reduce total
-overhead and maximize useful work. EA and ETA perform better than EB.}
+overhead and maximize useful work. HG and WHG perform better than EB.}
 	\label{fig:overallOverheadBreakdown}
@@
 
-For each coalescing strategy from Section~\ref{sec:task_adaptation} (EB, EA,
-and ETA) and for a baseline without coalescing (NC), we have measured the time
+For each coalescing strategy from Section~\ref{sec:task_adaptation} (EB, HG,
+and WHG) and for a baseline without coalescing (NC), we have measured the time
 spent on (i) executing useful task code, including \sys's overhead on access to
@@
 %
-This net overhead reduction is greatest in EA and ETA strategies compared
+This net overhead reduction is greatest in HG and WHG strategies compared
 to the EB strategy. We attribute this discrepancy to EB's slow adjustment
@@
 %
-In the subsequent experiments, we focus on the better-performing EA and ETA
+In the subsequent experiments, we focus on the better-performing HG and WHG
 strategies.
@@
 the application and in the runtime (left) and a breakdown of protected memory
-accesses into three types with different costs. NC: no coalescing, EA:
-energy-aware coalescing, ETA: energy- and task-aware coalescing.}
+accesses into three types with different costs. NC: no coalescing, HG:
+history-guided coalescing, WHG: weighted history-guided coalescing.}
 	\label{fig:coalEfficiency}
@@
 %
-For both coalescing strategies (EA and ETA) accesses to protected variables
+For both coalescing strategies (HG and WHG) accesses to protected variables
 constitute about 30\% of the runtime overhead. Dynamic address translation
@@
 Figure~\ref{fig:coalescing} shows the \sys run time with two coalescing
-strategies (EA, ETA) normalized to the run time without coalescing (NC).
+strategies (HG, WHG) normalized to the run time without coalescing (NC).
 %
@@
 %
-In some cases (\textbf{bc}, \textbf{cuckoo}, \textbf{sort}) the energy-aware
+In some cases (\textbf{bc}, \textbf{cuckoo}, \textbf{sort}) the history-guided
 feature in the coalescing algorithm is counter-productive.
@@
 In \textbf{fft}, tasks are not uniform, and accounting for task energies is
-beneficial. In fact, the lack of task energy awareness is detrimental: with EA
+beneficial. In fact, the lack of task energy awareness is detrimental: with HG
 \textbf{fft} runs slower than without any coalescing (NC).
@@
     \caption{\sys's coalescing performance. Left: Application execution time
-with coalescing strategies (EA and ETA) normalized to the execution time
-without coalescing. Right: Execution time with the ETA variant of \sys,
+with coalescing strategies (HG and WHG) normalized to the execution time
+without coalescing. Right: Execution time with the WHG variant of \sys,
 normalized to the execution time with Alpaca, as a function of distance to the
@@
 %	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}
-%	\caption{Coalescing strategies (EA and ETA) run time per application on intermittent power (RF generator at 15\,cm from the WISP), compared to \sys without coalescing (NC).
+%	\caption{Coalescing strategies (HG and WHG) run time per application on intermittent power (RF generator at 15\,cm from the WISP), compared to \sys without coalescing (NC).
 Coalescing provides significant improvements: from 0.25 (\textbf{ar}) up to 0.70 (\textbf{sort}) compared to a non-coalescing system.}
@@
 %	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
-%    \caption{Application execution time with the ETA variant of \sys,
+%    \caption{Application execution time with the WHG variant of \sys,
 %normalized to the execution time with Alpaca, as a function of distance to the
@@
 benchmark failed to terminate in the previous experiment. \sys executed the
-benchmark successfully without (NC) and with coalescing (ETA). The application
+benchmark successfully without (NC) and with coalescing (WHG). The application
 took longer to complete than at a shorter distance of 15~cm in the previous
@@
 distance. The variant without coalescing (NC) slowed down by 2.16$\times$ and the
-variant with ETA coalescing slowed down by 2.31$\times$ slower, but at both
-distances, ETA completed faster than NC.
+variant with WHG coalescing slowed down by 2.31$\times$ slower, but at both
+distances, WHG completed faster than NC.
 %

### 2018-04-17 paper/evaluation.tex
correct the algo names

@@
 relative to the baseline without coalescing.
-NC: no coalescing, EB: energy-blind coalescing, HG: history-guided coalescing, WHG:
+NC: no coalescing, EO: energy-blind coalescing, HG: history-guided coalescing, WHG:
 energy- and task-aware coalescing. All coalescing strategies reduce total
-overhead and maximize useful work. HG and WHG perform better than EB.}
+overhead and maximize useful work. HG and WHG perform better than EO.}
 	\label{fig:overallOverheadBreakdown}
@@
 
-For each coalescing strategy from Section~\ref{sec:task_adaptation} (EB, HG,
+For each coalescing strategy from Section~\ref{sec:task_adaptation} (EO, HG,
 and WHG) and for a baseline without coalescing (NC), we have measured the time
@@
 This net overhead reduction is greatest in HG and WHG strategies compared
-to the EB strategy. We attribute this discrepancy to EB's slow adjustment
+to the EO strategy. We attribute this discrepancy to EO's slow adjustment
 of the target task size without regard to execution history.

### 2018-04-17 paper/abstract.tex
abstract typo

@@
 When tasks are sized without accounting for available energy, state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
-To address these challenges, we propose \sys, an \emph{adaptive} task-based execution model that  progresses on a multi-tasks scale, when energy permits (\emph{efficiency}), and preseves the computation progress on a sub-task scale when needed.
+To address these challenges, we propose \sys, an \emph{adaptive} task-based execution model that progresses on a multi-tasks scale, when energy permits (\emph{efficiency}), and preserves the computation progress on a sub-task scale when needed.
 \sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.

### 2018-04-17 paper/abstract.tex
abstract typo

@@
 
-Embedded devices in the Internet of Things (IoT) gain new deployment opportunities when their hardware is free of batteries and powered by ambient energy.
+Embedded devices in the Internet of Things gain new deployment opportunities when their hardware is free of batteries and powered by ambient energy.
 Energy harvested from the environment, however, is unpredictable and can only power a device intermittently.

### 2018-04-17 paper/coalescing.tex
Final pass on coalescing

@@
 A key limitation of this algorithm lies in the equality of the target decrease
-in $f_\text{reboot}(B)$ and the increase in $f_\text{reboot}(B)$.  If, after
+in $f_\text{reboot}(B)$ and the increase in $f_\text{commit}(B)$.  If, after
 some number $k$ successful commits the target must decrease back to its
@@
 %
-EA gives each task unity weight, as with EO.  
+HG gives each task unity weight, as with EO.  
 %
@@
 
-The EA algorithm is characterized by the following functions:
+The HG algorithm is characterized by the following functions:
 

### 2018-04-17 paper/implementation.tex
indent

@@
 
-\textbf{New Tasks.} The \texttt{COALA\_TASK} annotation on a function
+\noindent\textbf{New Tasks.} The \texttt{COALA\_TASK} annotation on a function
 declaration statically allocates a non-volatile constant variable holding a
@@
 
-\textbf{Task Transitions.} A call to \texttt{COALA\_NEXT\_TASK} immediately
+\noindent\textbf{Task Transitions.} A call to \texttt{COALA\_NEXT\_TASK} immediately
 transfers control to a specified next task and can be invoked along
@@
 
-\textbf{Protected Variables.} The \texttt{COALA\_PV} annotation on a
+\noindent\textbf{Protected Variables.} The \texttt{COALA\_PV} annotation on a
 variable statically allocates and aligns a protected non-volatile variable. The
@@
 
-\textbf{Protected Reads and Writes.} The programmer uses the \texttt{RP} and \texttt{WP} API
+\noindent\textbf{Protected Reads and Writes.} The programmer uses the \texttt{RP} and \texttt{WP} API
 methods to access protected variables.  These methods cause \sys to find a variable's page.
@@
 
-\textbf{Initialization.} The behavior of the API method
+\noindent\textbf{Initialization.} The behavior of the API method
 \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the
@@
 
-\textbf{Execution.} The programmer passes control to \sys's task scheduler 
+\noindent\textbf{Execution.} The programmer passes control to \sys's task scheduler 
 by calling \texttt{COALA\_RUN} after device initialization. 

### 2018-04-17 paper/systemdescription.tex
Overview figure (3) changed

@@
 	\centering
-	\includegraphics[width=0.45\columnwidth]{figures/graffle/overview.pdf}
-	\caption{\sys top-level view. \sys is composed of two core components: \emph{Task Scheduler} and \emph{Memory Manager}.}
+	\includegraphics[width=0.45\columnwidth]{figures/overview.pdf}
+	\caption{System's top-level view. \sys is composed of two core components: \emph{Adaptive Task Scheduler} and \emph{Virtual Memory Manager}.}
 	\label{fig:system_overview}
@@
 
-\textbf{\sys Programming and Execution Model.}  To use \sys, a programmer first writes plain, imperative C code.
+\textbf{Programming and Execution Model.}  To use \sys, a programmer first writes plain, imperative C code.
 The programmer then must decompose the program into tasks.
@@
 
-The programmer compiles their task-based code, and links to the \sys runtime,
-producing a \sys-enabled binary. The \sys runtime implements \sys's task-based
+The programmer compiles their task-based code, and links to \sys runtime,
+producing a \sys-enabled binary. The runtime implements \sys's task-based
 programming and execution model. The runtime system also includes \sys's novel
-{\em virtualizing memory manager} and its {\em adaptive task scheduler}, both
+{\em virtual memory manager} and its {\em adaptive task scheduler}, both
 of which are essential to \sys's adaptive task-based execution.
 
-\textbf{\sys Adaptive Task Scheduler.} 
+\textbf{Adaptive Task Scheduler.} 
 \sys's adaptive task scheduler decides when to coalesce consecutive tasks
@@
 
-\textbf{\sys Virtual Memory Manager.} \sys is able to efficiently coalesce
+\textbf{Virtual Memory Manager.} \sys is able to efficiently coalesce
 tasks because of its efficient virtual memory manager. \sys's memory manager

### 2018-04-17 paper/evaluation.tex
correct the algo names
evaluation edits

@@
 relative to the baseline without coalescing.
-NC: no coalescing, EO: energy-blind coalescing, HG: history-guided coalescing, WHG:
+NC: no coalescing, EO: energy-Oblivious coalescing, HG: history-guided coalescing, WHG:
 energy- and task-aware coalescing. All coalescing strategies reduce total
@@
 The \textbf{fft} benchmark did not complete at distances larger than
-15\,cm~\footnote{At distances greater than 15~cm, energy incoming during
+15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
 execution is negligible and stored energy is insufficient to complete some of
@@
 the execution time for \textbf{fft} benchmark at 60\,cm from the signal
-generator antenna --- even \emph{farther} than the distance at which this
+generator antenna---even \emph{farther} than the distance at which this
 benchmark failed to terminate in the previous experiment. \sys executed the
 benchmark successfully without (NC) and with coalescing (WHG). The application
-took longer to complete than at a shorter distance of 15~cm in the previous
+took longer to complete than at a shorter distance of 15\,cm in the previous
 experiment, because of smaller (negligible) incoming energy at the larger

### 2018-08-02 paper/asplos19/paper/evaluation.tex
preparing for asplos19

--- 2018-04-17 paper/evaluation.tex
@@
-\begin{wrapfigure}{t!}{0.5\textwidth}
+% \begin{wrapfigure}{t!}{0.5\textwidth}
+\begin{figure}
 	%\begin{figure}
@@
 relative to the baseline without coalescing.
-NC: no coalescing, EO: energy-Oblivious coalescing, HG: history-guided coalescing, WHG:
+NC: no coalescing, EO: energy-oblivious coalescing, HG: history-guided coalescing, WHG:
 energy- and task-aware coalescing. All coalescing strategies reduce total
@@
 	\label{fig:overallOverheadBreakdown}
-	%\end{figure}
-\end{wrapfigure}
+\end{figure}
+% \end{wrapfigure}
 
@@
 %
-The search for the accessed page in the \texttt{\underline{working}} buffer,
+The search for the accessed page in the \texttt{working} buffer,
 will return immediately when the page is the most recently accessed page.
@@
 
-\begin{figure*}
+\begin{figure}
 	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}%
@@
 	\label{fig:runtime}
-\end{figure*}
+\end{figure}
 

### 2018-08-02 paper/asplos19/paper/intro.tex
preparing for asplos19

--- 2018-04-17 paper/intro.tex
@@
 neither a battery nor a tethered power
-supply~\citep{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These
+supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These
 devices operate using ambient energy, such as radio frequency
-transmissions~\citep{rf_powered_computing_gollakota_2014},
-light~\citep{margolies_infocom_2016,margolies_tosn_2016}, and
-vibration~\citep{gorlatova_sigmetrics_2014}. Incorporating compute, storage,
-sensing, and communication hardware~\citep{wisp5,moo,capybara}, such devices are a
-promising technology for use in the Internet of Things~\citep{ku_cst_2016},
-in-body~\citep{nadeau_naturebio_2017} and
-on-body~\citep{bandodkar_electroanalysis_2015} medical systems, and
-energy-harvesting nano-satellites~\citep{kicksat,capybara}.
+transmissions~\cite{rf_powered_computing_gollakota_2014},
+light~\cite{margolies_infocom_2016,margolies_tosn_2016}, and
+vibration~\cite{gorlatova_sigmetrics_2014}. Incorporating compute, storage,
+sensing, and communication hardware~\cite{wisp5,moo,capybara}, such devices are a
+promising technology for use in the Internet of Things~\cite{ku_cst_2016},
+in-body~\cite{nadeau_naturebio_2017} and
+on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and
+energy-harvesting nano-satellites~\cite{kicksat,capybara}.
 
@@
 intermittently} when energy is
-available~\citep{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
-buffers energy in a small storage capacitor~\citep{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
+available~\cite{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
+buffers energy in a small storage capacitor~\cite{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
 threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
 power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
-down and recharges to operate again later---corresponds to {\em intermittent execution model}~\citep{dino,lucia_snapl_2017} composed of
+down and recharges to operate again later---corresponds to {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of
 operation-power failure-restart cycles. As an example, recharge times may be
@@
 The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
-$\approx$100 times per second~\citep{tan_infocom_2016,mementos,nvp}.
+$\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.
 
 \textbf{Data Consistency in Intermittent Computing.}  
-%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\citep{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device.
+%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device.
 In an intermittent software execution, an {\em operating period} proceeds for anarbitrary duration (dictated by energy availability) before being interruptedby a {\em failure period}.
@@
 memory, such as FRAM or Flash. While capturing periodic
-checkpoints~\citep{mementos,quickrecall} and sleep
-scheduling~\citep{dewdrop,hibernus,hibernusplusplus} help preserve execution
+checkpoints~\cite{mementos,quickrecall} and sleep
+scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution
 progress, failures can leave non-volatile state incorrectly, partially updated,
@@
 continuously-powered behavior, often leading to an unrecoverable
-application failure~\citep{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for
+application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for
 intermittently-powered devices: (i) \emph{software-based programming and
-execution models}~\citep{dino,ratchet,chain,alpaca} and (ii)
+execution models}~\cite{dino,ratchet,chain,alpaca} and (ii)
 \emph{hardware-based computer architecture
-support}~\citep{hicks_isca_2017,idetic,nvp}. Complex architectural changes are
+support}~\cite{hicks_isca_2017,idetic,nvp}. Complex architectural changes are
 expensive to design, verify, and manufacture. New architectures are also
-inapplicable to existing systems~\citep{hicks_isca_2017,nvp}. Software
+inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software
 approaches are simpler and applicable to existing devices today, but prior
@@
 
-\begin{wrapfigure}{t!}{0.45\textwidth}
+\begin{figure} %{t!}{0.45\textwidth}
     \centering
-    \includegraphics[width=0.45\textwidth]{figures/intro-figure.pdf}
+    \includegraphics[width=\columnwidth]{figures/intro-figure.pdf}
     \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
@@
     \label{fig:coalesce}
-\end{wrapfigure}
+\end{figure}
 
@@
 Task-based} programming and execution models require a
-programmer~\citep{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
+programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
 statically decompose a program into a collection of {\em tasks}.  A task can
@@
 shows how tasks can impose a run-time overhead.
-At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\citep{chain,alpaca}.
+At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
 The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.
@@
 %applications (the same application X, but divided into X and Y tasks as
-%in~\citep{chain}) running on RFID antenna-powered Computational
-%RFID~\citep{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna
+%in~\cite{chain}) running on RFID antenna-powered Computational
+%RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna
 %distances (near---stable energy supply/far---high energy intermittency). At
@@
 (i.e., redo-logging) to eliminate statically identifiable, inter-task data
-dependences~\citep{alpaca}, a \emph{dynamic} dependence between two coalesced
+dependences~\cite{alpaca}, a \emph{dynamic} dependence between two coalesced
 tasks requires dynamic privatization and commit actions. 

### 2018-08-02 paper/asplos19/paper/texFigCode/dma.tex
preparing for asplos19

--- 2017-08-04 paper/figures/sram-buffer.tex
@@
 
-	\node[punkt, below = -3cm of sram,fill=red!30] (x-sram) {x};
+	\node[punkt, below = -3cm of sram,fill=blue!30] (x-sram) {x};
 	\node[punkt, below = 0.3cm of x-sram,fill=blue!30] (y-sram) {y};
@@
 	\path[every node/.style={font=\sffamily\small}]
-	([yshift=0.3cm] task1.east) edge [->, dotted, thick, bend left] node {(1) get x} (x-sram.west)
-	(x-sram.east) edge [->, dotted, thick, bend left] node[xshift=0.5cm] {(2) read x} (x1-fram.west)
+	 (fram1.west) edge [->, dotted, thick, bend right] node[xshift=-1cm] {(1) flush in (DMA)}([yshift=0.3cm] sram.east)
+
+	([yshift=0.2cm]x-sram.west) edge [->, dotted, thick, bend right] node[xshift=-0.5cm] {(2) read x} ([yshift=0.6cm]task1.east)
 
@@
 
-	(x-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(6) commit-I x} (x2-fram.west)
-	(z-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(7) commit-I z} (z2-fram.west)
-	(y-sram.east) edge [->, dotted, thick, bend right] node[xshift=-0.4cm] {(8) commit-I y} (y2-fram.west)
+	(sram.east) edge [->, dotted, thick, bend right] node[align=center,xshift=-0.5cm] {(6)  flush out \\ (DMA)} (fram2.west)
 
-	(x2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(9) commit-II x} (x1-fram.east)
-	(y2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(10) commit-II y} (y1-fram.east)
-	(z2-fram.east) edge [->, dotted, thick, bend right] node[xshift=2.5cm] {(11) commit-II z} (z1-fram.east)
+	(fram2.east) edge [->, dotted, thick, bend right] node[align=center,xshift=1.5cm] {(7) flush out \\ (DMA)} (fram1.east)
 	;

### 2018-08-02 paper/asplos19/paper/texFigCode/dynamic-chain-2.tex
preparing for asplos19

--- 2017-06-27 paper/presentation/images/chain-virtualization-II.tex
@@
 	\node[punkt, right =0.9cm of task2] (task5) {Task (5)};
-	\node[punkt, right =2cm of task4] (task6) {Task (6)};
+	\node[punkt, right =1.6cm of task4] (task6) {Task (6)};
 
@@
 		\node[draw,thick,dotted,red,minimum width=7em, fill=blue!10, fit=(task1) (task2) (task3)] (Virtualized) {};
-		\draw[<->] ([xshift=-1.7cm]Virtualized.north) --node[sloped, anchor=south] {Virtualization Window} ([xshift=-1.7cm]Virtualized.south);
+		\draw[<->] ([xshift=-1.4cm]Virtualized.north) --node[sloped, anchor=south] {} ([xshift=-1.4cm]Virtualized.south);
+		% \draw[] ([xshift=-1.5cm]Virtualized.north) --node[sloped, anchor=south] {} ([xshift=-1.5cm]Virtualized.south);
 	\end{pgfonlayer}

### 2018-08-02 paper/asplos19/background.tex
restructure the paper directory

--- 2018-04-17 paper/background.tex
@@
 
-\begin{wrapfigure}{t!}{0.5\textwidth}
+% \begin{wrapfigure}{t!}{0.5\textwidth}
+\begin{figure}
 	\begin{subfigure}[t]{\linewidth}
@@
 	\caption{Code example demonstrating checkpointing gives rise to data inconsistency.}\label{fig:code_demo_incosistency}
-\end{wrapfigure}
-%\end{figure}
+% \end{wrapfigure}
+\end{figure}
 
@@
 		Mementos~\cite{mementos}	& Registers + Stack     \\
-		DINO~\cite{dino}	& Registers + Stack + WAR NV variables used in task\\
+		DINO~\cite{dino}	& Registers + Stack + WAR NV variables \\%used in task\\
 		Chain~\cite{chain}	& PC + NV variables used in task\\
@@
 redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that
-gets copies to and from the non-volatile memory in service of memory
+gets copied to and from the non-volatile memory in service of memory
 consistency and progress preservation.  Ratchet~\cite{ratchet} and
 Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile
-perform high frequency non-volatile memory accesses, in to checkpointing
+perform high frequency non-volatile memory accesses, into checkpointing
 frequently.
@@
 inefficient.  If code between consecutive checkpoints or in a task (we refer to
-both scenarios as ''code in a task'' in this section) consumes {\em more}
-energy than the fixed, maximum energy that the device can buffer, then the task
+both scenarios as ''code in a task'' in this section) consumes {\em more} than the fixed,
+maximum energy that the device can buffer, then the task
 will never be able to complete using buffered energy.  Such a task is

### 2018-08-02 paper/asplos19/hwswsetup.tex
restructure the paper directory

--- 2018-04-17 paper/hwswsetup.tex
@@
 	\footnotesize
-	\begin{tabular}{| r|r|r | p{0.65\columnwidth} |}
+	\begin{tabular}{| r|r|r | p{0.54\columnwidth} |}
 		\hline
-		Application & No. tasks & SLOC & Description \\
+		App.&tasks&SLOC&Description\\
 		\hline\hline
         \textbf{ar} &10 &428 & Activity recognition using a nearest neighbor
-classifier with a deterministic input from a randomly-generated sequence of
-accelerometer data\\
+classifier\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
 		\hline
@@
 		\hline
-        \textbf{dijkstra} &5 &198 & Dijkstra shortest path algorithm executed on a
-predefined adjacency matrix\\
+        \textbf{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
 		\hline
-        \textbf{fft} &8 &449 & Fast Fourier Transform on three pre-generated 128
-sample vectors\\
+        \textbf{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
 		\hline
-		\textbf{sort} &4 &167 & Selection sort of a fixed sequence of numbers\\
+		\textbf{sort} &4 &167 & Selection sort algorithm\\
 		\hline
@@
 By compiling both systems with the same compiler (GCC), we ensure that our
-comparison in Section~\ref{sec:evaluation} is fair.
+comparison in Section~\ref{sec:evaluation} is fair.\vspace{2cm}
 

### 2018-08-02 paper/asplos19/implementation.tex
restructure the paper directory

--- 2018-04-17 paper/implementation.tex
@@
 \centering
-\begin{tabular}{| r | p{0.7\columnwidth} |}
+\begin{tabular}{| r | p{0.67\columnwidth} |}
 	\hline
@@
 	\hline\hline
-	\texttt{COALA\_INIT}($t$) & $t \in \mathbf{T}$: task to be scheduled on the first boot \\
+	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
 	\hline
-	\texttt{COALA\_RUN}() & --- \\
+	\texttt{RUN}() & --- \\
 	\hline
-	\texttt{COALA\_TASK}($t$, $w_t$) & $t \in \mathbf{T}$: new task name, $w_t$: weight of task $t$ \\
+	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
 	\hline
-	\texttt{COALA\_NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
+	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
 	\hline
-	\texttt{COALA\_PV}($p$, $v$ [, $s$]) & $p$: variable type, $v \in \mathbf{V}$: new protected variable name, $s$: array size \\
+	\texttt{PV}($p$, $v$ [, $s$]) & $p$: variable type, $v \in \mathbf{V}$: new protected variable name, $s$: array size \\
 	\hline
@@
 	\hline
-	\texttt{COALA\_SM}($p$, $m$ [, $s$]) & $p$: \texttt{struct} member's type, $m$: \texttt{struct} member's name, $s$: array size \\
+	\texttt{SM}($p$, $m$ [, $s$]) & $p$: \texttt{struct} member's type, $m$: \texttt{struct} member's name, $s$: array size \\
 	\hline

### 2018-08-02 paper/asplos19/texFigCode/sram-buffer.tex
restructure the paper directory

--- 2017-08-11 paper/texFigCode/sram-buffer.tex
@@
 	\node[right = 0.5cm of sram] (ref) {};
-	\node[draw,very thick,minimum width=8em, minimum height=30em, right = of ref] (fram1) {};
+	\node[draw,very thick,minimum width=8em, minimum height=21em, right = of ref] (fram1) {};
 
@@
 
-	\node[draw,below = -2.7cm of fram1,minimum width=6em, minimum height=5em] (pagesTmp) {};
-	\node[punkt,draw=none, below = -2.5cm of fram1,fill=black!5] (y2-fram) {y};
+	\node[draw,dotted,below = -2.5cm of fram1,minimum width=6em, minimum height=5em] (pagesTmp) {};
+	\node[punkt,draw=none, below = -2.2cm of fram1,fill=black!5] (y2-fram) {y};
 	\node[punkt,draw=none, below = 0.1cm of y2-fram,fill=black!5] (z2-fram) {z};
@@
 
-	(pagesTmp.east) edge [->, bend right,dotted,thick] node{(6) \texttt{commit}} (pagesOrg2.east);
+	(pagesTmp.east) edge [->, bend right,dotted,thick] node[yshift=-0.2cm]{(6) \texttt{commit}} (pagesOrg2.east);
 

### 2018-08-02 paper/oopsla18/background.tex
restructure the paper directory

--- 2018-08-02 paper/asplos19/background.tex
@@
 
-% \begin{wrapfigure}{t!}{0.5\textwidth}
-\begin{figure}
+\begin{wrapfigure}{t!}{0.5\textwidth}
 	\begin{subfigure}[t]{\linewidth}
@@
 	\caption{Code example demonstrating checkpointing gives rise to data inconsistency.}\label{fig:code_demo_incosistency}
-% \end{wrapfigure}
-\end{figure}
+\end{wrapfigure}
+%\end{figure}
 
@@
 		Mementos~\cite{mementos}	& Registers + Stack     \\
-		DINO~\cite{dino}	& Registers + Stack + WAR NV variables \\%used in task\\
+		DINO~\cite{dino}	& Registers + Stack + WAR NV variables used in task\\
 		Chain~\cite{chain}	& PC + NV variables used in task\\

### 2018-08-02 paper/oopsla18/conclusions.tex
restructure the paper directory

Software for intermittently-powered energy-harvesting devices requires a
runtime system for performing computations across power failures.  \sys is a
new task-based runtime system that fulfills this need.  \sys's distinguishing
feature is its adaptability to changing energy conditions at runtime.  When
more energy is available, \sys makes faster progress through the computation by
coalescing statically-defined tasks together according to one of three proposed
policies. When less energy is available, progress is latched at sub-task
granularity. \sys's virtual memory system ensures that program state in
non-volatile memory remains consistent and amortizes the cost of state transfer
between volatile and non-volatile memory.  Our evaluation across embedded
applications on a real energy-harvesting device demonstrates the utility of
adaptive coalescing and downsizing, as well as the practicality of the proposed
implementation.

### 2018-08-02 paper/asplos19/abstract.tex
changes

--- 2018-04-17 paper/abstract.tex
@@
 Results show that \sys is more efficient than the current state-of-art execution model by up to a factor of two and able to progress where its counterparts fail.
-
+%
 %Alexei's version of abstract
-
-Embedded devices in the Internet of Things gain new deployment opportunities when their hardware is free of batteries and powered by ambient energy.
+%
+% Embedded devices in the Internet of Things gain new deployment opportunities when their hardware is free of batteries and powered by ambient energy.
 Energy harvested from the environment, however, is unpredictable and can only power a device intermittently.
@@
 Our evaluation on a real energy-harvesting device demonstrates the benefits of adaptive task coalescing and downscaling, and highlights \sys's performance improvements over a state-of-art task-based system.
+%
+Freeing embedded devices from batteries is a key factor in developing the energy neutral Internet of Things. 
+Energy harvesting, however, is unpredictable and can only power a device intermittently.
+Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save intermediate programs states into non-volatile memory frequently, to preserve a program forward progress.
+Despite the superiority of the task-based approach, however, it fixes their task size at compile time and operates obliviously to the changing energy conditions.
+When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
+To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model.
+\sys progresses on a multi-tasks scale, when energy permits and preserves the computation progress on a sub-task scale if necessary.
+\sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.
+Our evaluation shows that \sys is $2X$ faster than the state-of-the-art system in most cases and is able to progress where static systems fail.

### 2018-08-02 paper/asplos19/intro.tex
New Makefile and more tiny intro figure

--- 2018-08-02 paper/asplos19/paper/intro.tex
@@
 down and recharges to operate again later---corresponds to {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of
-operation-power failure-restart cycles. As an example, recharge times may be
-tens of seconds in radio frequency powered medical device~\cite[Fig.
-3c]{nadeau_naturebio_2017}.
+operation-power failure-restart cycles. 
+%As an example, recharge times may be tens of seconds in radio frequency powered medical device~\cite[Fig.3c]{nadeau_naturebio_2017}. 
 The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
@@
 state in its registers, stack, SRAM, and retains the state of any non-volatile
-memory, such as FRAM or Flash. While capturing periodic
+memory, such as FRAM. While capturing periodic
 checkpoints~\cite{mementos,quickrecall} and sleep
@@
 inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software
-approaches are simpler and applicable to existing devices today, but prior
-software approaches have a number of limitations that are the focus of the work
-in this paper. A key limitation of prior software approaches that we address in this paper is the {\em inflexibility} of a program statically decomposed into tasks.
+approaches are simpler and applicable to existing devices today.
+Therefore, this work focuses on a key limitation of task-based software approaches, namely, the {\em inflexibility} of statically decomposing a program into tasks.
 
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/intro-figure.pdf}
+    \includegraphics[width=\columnwidth]{figures/intro-figure-tiny.pdf}
     \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
@@
 %
-The programmer explicitly expresses task-to-task control flow, or the compiler
-converts existing program control-flow into task-to-task control flow to sequence 
-the tasks that it defines.
+The programmer (or a compiler) explicitly expresses task-to-task control flow.
 %
@@
 
-A savvy programmer may thus create very large tasks in an attempt to reduce
+A programmer may thus create very large tasks in an attempt to reduce
 overhead by minimizing the number of task transitions. However, a very large
 task may require more energy to complete than a device's fixed hardware energy
-buffer can hold. Such a large task fails to completely execute assuming only
-buffered energy is available (i.e., assuming the worst case). While some
-environments may have sufficient harvestable energy to replenish a device's
-capacitor during operation, prolonging the execution period and enabling the
-task to complete, assuming cooperation from the physical environment to avoid
-non-termination is a risky programming proposition. To eliminate this risk,
-existing systems require the programmer or compiler to decompose a program
+buffer can hold which may lead to a program non-termination problem. To eliminate this risk,
+existing systems require the programmer to decompose a program
 into small tasks, all of which complete using only buffered energy. These
@@
 we propose a third way, using a novel technique called {\em adaptive execution by dynamic task
-coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads while avoiding the risk of non-termination.
+coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.
 
@@
 intermittent program by dynamically coalescing/splitting its tasks.
-\sys accepts any static decomposition (i.e., from the programmer or from a compiler) 
-and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute  
+\sys accepts any static decomposition and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute
 with no commit overhead between them, instead performing

### 2018-08-03 paper/asplos19/abstract.tex

@@
 Energy harvesting, however, is unpredictable and can only power a device intermittently.
-Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save intermediate programs states into non-volatile memory frequently, to preserve a program forward progress.
-Despite the superiority of the task-based approach, however, it fixes their task size at compile time and operates obliviously to the changing energy conditions.
-When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
+Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve execution progress of a program.
+Despite the superiority of the task-based approach, however, it fixes the task size at compile time and operates obliviously to changing energy conditions.
+When tasks are sized without accounting for available energy, state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.% When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
 To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model.

### 2018-08-03 paper/asplos19/intro.tex
first iteration on challenges and contributions

@@
 %results.
-\textbf{Challenges.} While a compiler can use static data privatization and commit instrumentation
-(i.e., redo-logging) to eliminate statically identifiable, inter-task data
-dependences~\cite{alpaca}, a \emph{dynamic} dependence between two coalesced
-tasks requires dynamic privatization and commit actions.
+\textbf{Challenges and Contributions} 
+\begin{itemize}
+\item\textit{challenge} 1) Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
+\textit{contribution---1} \sys addresses this challenge by using its \textit{recent execution history} as a metric to estimate the environment energy conditions.
+Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.
 %
-To ensure consistency and respect inter-task dependences, \sys uses a novel
-approach \emph{power failure-immune memory virtualization} that buffers non-volatile variable
-updates in volatile memory during a coalesced task execution, before committing
-them to non-volatile memory at the dynamic task boundary.
+\item\textit{challenge} 2) Given dynamic variables dependencies, how to ensure efficient data protection against power interrupts?
+\textit{contribution---2} \sys ensures the consistency of the data by using a novel
+approach called \emph{dynamic group privatization}.
+Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e.
+protecting all global variables) in each dynamic task transition.
+\sys, therefore, utilizes a real-time data tracking to enable partial data protection on a task transition.
+Individual variables tacking, however, slows down a system dramatically.
+Therefore, \sys  relies on a coarse locality principle and protects the global variables in a batch.
+Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
+
+\item\textit{challenge} 3) A static task decomposition model assumes that each
+single task can execute to completion. If the hardware energy buffer provides
+inadequate energy to execute each single task to completion, a program will not
+terminate~\cite{cleancut_2018}. under such conditions how to enable the dynamic execution model to progress on a sub-task level? \textit{contribution---3} To avoid non-termination under adversarial
+energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
+Partial execution avoids non-termination by committing the intermediate state of a long-running
+task that has repeatedly failed and restarted. Partial execution violates
+task atomicity, but preserves forward progress; if a programmer knows that
+task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
+
+% \textbf{dynamic privatization}: while a compiler can use static data privatization and commit instrumentation
+% (i.e., redo-logging) to eliminate statically identifiable, inter-task data
+% dependences~\cite{alpaca}, a \emph{dynamic} dependence between two coalesced
+% tasks requires dynamic privatization and commit actions.
+% %
+
 %
@@
 %consistency.
-A static task decomposition model assumes that each
-single task can execute to completion. If the hardware energy buffer provides
-inadequate energy to execute each single task to completion, a program will not
-terminate~\cite{cleancut_2018}. To avoid non-termination under adversarial
-energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
-Partial execution avoids non-termination by committing the intermediate state of a long-running
-task that has repeatedly failed and restarted. Partial execution violates
-task atomicity, but preserves forward progress; if a programmer knows that
-task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
+% \textbf{task non-termination}:
+
 %
@@
 %
+\end{itemize}
 
-\textbf{Contributions.} To summarize, \sys's main contributions are:
-%
-\begin{itemize}
-\item The first {\em adaptive} task-based intermittent execution model that
-dynamically \emph{coalesces} tasks to reduce commit overhead and
-\emph{downscales} tasks to ensure forward progress.
-\item The first memory virtualization mechanism that preserves the atomicity of coalesced tasks under power failures.
-%\item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
-\item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication via~\cite{coala_website}.
-\item An evaluation directly comparing \sys to a state of the art task-based
-intermittent programming and execution model from prior work, showing that on a
-suite of benchmarks from the literature and several new workloads, \sys often
-has higher performance and is more flexible to diverse energy conditions.
-\end{itemize}
+% \textbf{Contributions.} To summarize, \sys's main contributions are:
+% %
+% \begin{itemize}
+% \item The first {\em adaptive} task-based intermittent execution model that
+% dynamically \emph{coalesces} tasks to reduce commit overhead and
+% \emph{downscales} tasks to ensure forward progress.
+% \item The first memory virtualization mechanism that preserves the atomicity of coalesced tasks under power failures.
+% %\item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
+% \item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication via~\cite{coala_website}.
+% \item An evaluation directly comparing \sys to a state of the art task-based
+% intermittent programming and execution model from prior work, showing that on a
+% suite of benchmarks from the literature and several new workloads, \sys often
+% has higher performance and is more flexible to diverse energy conditions.
+% \end{itemize}
 %

### 2018-08-03 paper/asplos19/intro.tex
change the paragraph about paper structure

@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/intro-figure-tiny.pdf}
+    \includegraphics[width=\columnwidth]{figures/intro-figure-vert.pdf}
     \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
@@
 Section~\ref{sec:systemdescription} provides an overview of \sys, while
-Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism (coalescing and downscaling).
-Memory virtualization mechanism is provided in Section~\ref{sec:memory_virtulaization}.
+Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
+Dynamic group privatization mechanism is provided in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are given in Section~\ref{sec:implementation}.

### 2018-08-03 paper/asplos19/background.tex
revice the background

@@
 require architectural additions to commodity processors as
-in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}. \sys supports I/O
-behavior similar to~\cite{alpaca,chain}, allowing safe, synchronous I/O and
-unsafe, asynchronous I/O.
+in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.
 
@@
 persist across failure and be read by the {\em
-reads}~\cite{ratchet,dino,alpaca} ---see
+reads}~\cite{ratchet,dino,alpaca}\TODO{figure 2 dependent text ---see
 Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
@@
 maintain data consistent in the presence of WAR dependency is necessary for
-correct execution.
+correct execution.}
 %Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}.
@@
 Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to
-decompose their program into tasks, which is a region of code that can contain
+decompose their program into \textbf{tasks}, which is a region of code that can contain
 arbitrary computation, sensing, and communication.  Task-based models keep
@@
 %\end{figure}
-As an illustration, we measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
-the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor
-voltage drop on a capacitor powering the device across 1600 read and write
-accesses to each memory, randomizing accesses to avoid caching effects. Our
-data showed that a non-volatile memory access consumes around 1.5 times the
-energy of a volatile access. The data confirm that, because non-volatile memory
-accesses are more expensive than volatile memory accesses, prior systems that
-frequently access non-volatile memory may be artificially inefficient.
+
+% As an illustration, we measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
+% the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor
+% voltage drop on a capacitor powering the device across 1600 read and write
+% accesses to each memory, randomizing accesses to avoid caching effects. Our
+% data showed that a non-volatile memory access consumes around 1.5 times the
+% energy of a volatile access. The data confirm that, because non-volatile memory
+% accesses are more expensive than volatile memory accesses, prior systems that
+% frequently access non-volatile memory may be artificially inefficient.
+
 %~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
@@
 \todo{add citation to TI specification on FRAM energy cost}{Przemek}
-\sys is motivated by this observation: \sys 
-moves data pages from FRAM to SRAM to maximize SRAM accesses and minimize FRAM
-accesses, to decrease energy consumption and increase the time between power
-failures.
+
+% \sys is motivated by this observation: \sys 
+% moves data pages from FRAM to SRAM to maximize SRAM accesses and minimize FRAM
+% accesses, to decrease energy consumption and increase the time between power
+% failures.
 
@@
 
-\textbf{Incorrectly Sized Tasks.} Compiler-inserted checkpoints or
-programmer-defined tasks can both both incorrect and non-terminating or
-inefficient.  If code between consecutive checkpoints or in a task (we refer to
-both scenarios as ''code in a task'' in this section) consumes {\em more} than the fixed,
+\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or
+programmer-defined tasks can be both non-terminating and/or
+inefficient.  If a task (or code between two checkpoints) consumes more than the fixed,
 maximum energy that the device can buffer, then the task
@@
 %
-If the task consumes far {\em less} energy than a device can buffer, the system
-may operate {\em inefficiently}, checkpointing or transitioning between tasks
-more often than needed. 
+If the task consumes far less energy than a device can buffer, the system
+may operate {\em inefficiently}, saving the program state more often than needed.
 %
-Avoiding excessively costly, non-terminating tasks and short, high-overhead
-tasks is challenging both to the programmer and the compiler, because
-estimating the exact energy use of an arbitrary code is complicated.  When {\em
-porting} a program from one platform to another, the problem becomes worse
-because an excessively costly task on one device may be a relatively short task
-on a device with a larger buffer.  Buffer sizes on different platforms vary
-widely from, e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
-complicating the problem.
+Avoiding excessively costly, non-terminating, tasks and short, high-overhead,
+tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.  
+Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}) are considered the problem of porting a program become much worse
+because a large task on one device may be a relatively short task
+on a device with a larger buffer.
 %
-As a solution, \sys's execution model presents {\em task coalescing}, which
+\textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which
 merges multiple tasks to amortize the overhead when tasks are too small, and

### 2018-08-03 paper/asplos19/background.tex
partial commit

@@
 mixture of SRAM and FRAM. Our implementation leverages hardware support for
-fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys do not
+fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys does not
 require architectural additions to commodity processors as
@@
 task-based programming model, where the programmer explicitly expresses the
-application as a sequence of tasks and the transition between them.
+application as a sequence of tasks and the transitions between them.
 %Task-based runtime systems ensure that repeated task executions are idempotent. 

### 2018-08-03 paper/asplos19/systemdescription.tex
partial commit

--- 2018-04-17 paper/systemdescription.tex
@@
-\begin{wrapfigure}{t!}{0.45\textwidth}
+% \begin{wrapfigure}{t!}{0.45\textwidth}
+\begin{figure}
 	\centering
-	\includegraphics[width=0.45\columnwidth]{figures/overview.pdf}
+	\includegraphics[width=\columnwidth]{figures/overview.pdf}
 	\caption{System's top-level view. \sys is composed of two core components: \emph{Adaptive Task Scheduler} and \emph{Virtual Memory Manager}.}
 	\label{fig:system_overview}
-\end{wrapfigure}
+% \end{wrapfigure}
+\end{figure}
 
@@
 
-\textbf{Programming and Execution Model.}  To use \sys, a programmer first writes plain, imperative C code.
-The programmer then must decompose the program into tasks.
-To manually decompose a program into tasks, the programmer designates a set of functions as tasks, sequences control-flow between these tasks, and annotates memory accesses that manipulate data shared by multiple tasks.
-In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.
+\textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by  designates a top-level set of functions as tasks, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
+% In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.
 
-A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
+% A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
 Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with automatically decomposed code would be identical.
@@
 The programmer compiles their task-based code, and links to \sys runtime,
-producing a \sys-enabled binary. The runtime implements \sys's task-based
-programming and execution model. The runtime system also includes \sys's novel
-{\em virtual memory manager} and its {\em adaptive task scheduler}, both
-of which are essential to \sys's adaptive task-based execution.
+producing a \sys-enabled binary. The runtime library relies on \sys's novel
+{\em virtual memory manager} and its {\em adaptive task scheduler} to execute the binary 
+efficiently.
 

### 2018-08-03 paper/asplos19/systemdescription.tex
partial commit

@@
 
-\textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by  designates a top-level set of functions as tasks, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
+\textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by  encapsulating the conde in set of top-level set of functions (tasks), sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
 % In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.

### 2018-08-03 paper/asplos19/abstract.tex
some cleaning up and partical commit

@@
 Energy harvesting, however, is unpredictable and can only power a device intermittently.
-Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve execution progress of a program.
+Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve the execution progress of a program.
 Despite the superiority of the task-based approach, however, it fixes the task size at compile time and operates obliviously to changing energy conditions.
-When tasks are sized without accounting for available energy, state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.% When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
+When tasks are sized without accounting for available energy, the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.% When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
 To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model.
-\sys progresses on a multi-tasks scale, when energy permits and preserves the computation progress on a sub-task scale if necessary.
+\sys progresses on a multi-tasks scale when energy permits and preserves the computation progress on a sub-task scale if necessary.
 \sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.

### 2018-08-03 paper/asplos19/background.tex
some cleaning up and partical commit

@@
 decompose their program into \textbf{tasks}, which is a region of code that can contain
-arbitrary computation, sensing, and communication.  Task-based models keep
-progress at the granularity of tasks by repeatedly trying to re-execute each
-task on power failure until it successfully finishes, only then moving on to
+arbitrary computation, sensing, and communication.
+Task-based models progress at the granularity of tasks.
+They re-execute each task interrupted by a power failure until it successfully finishes, only then moving on to
 the next task. Since these models do not rely on taking an expensive
-checkpoint, it is usually faster than the checkpoint-based
+checkpoint, they are usually faster than the checkpoint-based
 solution~\cite{chain, alpaca}.  \sys also follows the paradigm of the
@@
 application as a sequence of tasks and the transitions between them.
-%Task-based runtime systems ensure that repeated task executions are idempotent. 
-%The key to idempotence is ensuring that non-volatile updates made by an interrupted task are never visible to a future task execution.  
-%A programmer describes task control-flow as a {\em task graph}. Task flow happens at programmer demarcated transition points that may be conditional on program values. 
-%Task-based programming abstractions guarantee that tasks execute {\em atomically}, regardless of power failures.
 
@@
 redo-logging for the variables that is part of the dependence~\cite{alpaca}, or
-by statically creating a multiple copies of the problematic variable to ensure
+by statically creating multiple copies of the problematic variable to ensure
 that no task reads and writes the same copy~\cite{chain}.
@@
 \textbf{Frequent Non-volatile Memory Access.} 
-Previous systems frequently access the non-volatile memory directly, which 
-is a potential source of inefficiency.  Checkpointing systems copy all volatile
+Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.  Checkpointing systems copy all volatile
 state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at
@@
 Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile
-perform high frequency non-volatile memory accesses, into checkpointing
+perform high-frequency non-volatile memory accesses, into checkpointing
 frequently.
-
-%Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}.
-Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf.
-\cite[Sec.
-2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory.
-Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.
-% We do not need to emphasize the following figure! (Sinan)
-%\begin{figure}
-%\begin{subfigure}[t]{.49\columnwidth}
-%	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
-%	\caption{Cost of \emph{write} operation}
-%\end{subfigure}%
-%\begin{subfigure}[t]{.49\columnwidth}
-%	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
-%	\caption{Cost of \emph{read} operation}
-%\end{subfigure}
-%	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
-%\end{figure}
-
-% As an illustration, we measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
-% the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor
-% voltage drop on a capacitor powering the device across 1600 read and write
-% accesses to each memory, randomizing accesses to avoid caching effects. Our
-% data showed that a non-volatile memory access consumes around 1.5 times the
-% energy of a volatile access. The data confirm that, because non-volatile memory
-% accesses are more expensive than volatile memory accesses, prior systems that
-% frequently access non-volatile memory may be artificially inefficient.
-
-%~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
-% Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device.
-This observation is also consistent with the TI specification~\cite{}.
-\todo{add citation to TI specification on FRAM energy cost}{Przemek}
-
-% \sys is motivated by this observation: \sys 
-% moves data pages from FRAM to SRAM to maximize SRAM accesses and minimize FRAM
-% accesses, to decrease energy consumption and increase the time between power
-% failures.
-
-%The cost of non-volatile memory motivates \sys, which virtualizes memory to primarily use SRAM.
-
-%The above result reassures the intuitive observation that a decrease of checkpoint frequency (CP) decreases the energy cost associated with each state preservation method and reduces operation cost which in most simplistic terms is defined as $\mathcal{O}(\text{CP})+\mathcal{O}(\text{PL})$, where PL is the cost of program execution without checkpointing.
-On the other hand, this increases task/checkpoint re-execution time.
-The ideal solution is a runtime that \emph{adaptively} changes task size during runtime, which however exposes a second problem.
-
-%\begin{table}
-%	\centering
-%	\footnotesize
-%	\begin{tabular}{|c|c|}
-%		\hline
-%		Platform name & Storage capacitor size \\
-%		\hline \hline
-%		%Moo~\cite{moo} & 0.1\,F \\
-%		WISPCam~\cite{naderiparizi_rfid_2015} & 6.08\,mF \\ %tested [11.24, 17.45, 21.98]\,mF
-%		%NFC-WISP~\cite{zhao_rfid_2015} & 300\,$\mu$F \\
-%		NeuralWISP~\cite{holleman_biocas_2008} & 100\,$\mu$F \\
-%		WISP~\cite{wisp5} & 47\,$\mu$F \\
-%		{\em BioImpedance} sensor~\cite{rodriguez_tbcs_2015} & 20\,$\mu$F \\
-%		{\em Ingestible} sensor~\cite{nadeau_naturebio_2017} & 220\,$\mu$F\\
-%		\hline
-%	\end{tabular} 
-%	\caption{{\em Default} energy storage sizes for example battery-less platforms. Observe a huge variation in storage capacity.}
-%	%We note that values for other representative platforms~\cite{medusa_farsens_2017,talla_imwut_2017,liu_sigcomm_2013,parks_sigcomm_2014} were not reported in their respective papers.
-%	\label{table:capacitor}
-%\end{table}
-
-%Przemek placeholder
-%\textbf{Task-based Models have Syntactic Limitations.} State of the art task-based programming models, Alpaca~\cite{alpaca}, has substantial limitations that prohibit writing generic programs.
-First, for Alpaca (and Chain~\cite{chain}), accesses to protected variables must be applied \emph{only} by using the variable's name, i.e.
-reference by variable address is not possible.
-This limitation implies that existing third-party libraries, such as MCU-tailored FFT library~\cite{}, cannot be implemented to support Alpaca or Chain (not to mention the impossibility of the development of an intermittent-safe \texttt{memcpy}).
-Furthermore, if a protected variable needs to be accessed in a code path that is not predictable by the compiler, for instance an Interrupt Service Routine, Alpaca cannot protect that access.
 
@@
 {\em task downscaling}, which breaks up a task when the task is too large to complete.
-%When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit).
-However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions.
-Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity.
-Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}.
-Sizing tasks complicates {\em porting} code from one device to another.
-An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer.
-The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead.
-\sys's execution model based on {\em task coalescing} and {\em task downscaling}.
-Coalescing dynamically merges multiple tasks based on execution behavior to amortize the cost of committing each task separately.
-Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too large to complete using the device's buffered energy.
-Section~\ref{sec:task_adaptation}describes \sys's adaptive coalescing and downscaling mechanisms in detail.
-The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, which is described in Section~\ref{sec:memory_virtulaization}.
-
-%\textbf{Motivating the Design of \sys's Execution Model.} The challenge of \em{efficiently} executing a program under frequent power failures, lies in the fact the the program needs to be energy aware.
-This challenge becomes even bigger when \em{generality} is a design requirement and such the system must not require a specific hardware circuitry to offer its efficiency.
-\em{Fllexibility} is key for wide adaptation, therefore, \sys's execution model should not limit the programmer to a subset of the functionality of the underlying software technology.
-\em{\sys champions an efficient, general, and flexible intermittent execution.}

### 2018-08-03 paper/asplos19/systemdescription.tex
some cleaning up and partical commit

--- 2018-08-03 paper/asplos19/systemdescription.tex
@@
 \sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable} and {\em efficient}.
-\sys accomplishes this goal with a constellation of a new programming model and run time software system support, that supports dynamically adaptive task-based execution.
+\sys accomplishes this goal with a constellation of a new programming model and runtime software system support, that supports dynamically adaptive task-based execution.
 Figure~\ref{fig:system_overview} shows an overview of \sys.
 
-\textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by  designates a top-level set of functions as tasks, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
+\textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by encapsulating the code in a top-level set of functions, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
 % In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.
@@
 % A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
-Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with automatically decomposed code would be identical.
+Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with the automatically decomposed code would be identical.
 
@@
 producing a \sys-enabled binary. The runtime library relies on \sys's novel
-{\em virtual memory manager} and its {\em adaptive task scheduler} to execute the binary 
-efficiently.
+{\em virtual memory manager} and its {\em adaptive task scheduler} to adapt its execution with the energy conditions.
 
 \textbf{Adaptive Task Scheduler.} 
-\sys's adaptive task scheduler decides when to coalesce consecutive tasks
-together to amortize commit and transition costs, and when to downscale a task,
-breaking it into multiple tasks, to avoid non-termination.
-
-\sys's scheduler dynamically {\em coalesces} statically\hyp{}defined tasks to
-avoid inessential overheads associated with completing tasks. By default, tasks
-run in a sequence and each task commits its task-shared state as it completes.
-The time and energy cost of a task's commit is unnecessary if the task and its
-successor both complete without a power failure. The key insight that \sys
-leverages is that the first task could have deferred its commit to the second
-task. {\em Coalescing} tasks by deferring the first commit avoids the fixed
-cost of the first commit (task transitioning overheads). Coalescing also
-amortizes per-variable commit cost, committing each location accessed by both
-tasks only after the second task, rather than once per task.  \sys makes
-\emph{energy-aware} scheduling decisions, using a short history of recently
-executed tasks to estimate available energy and guide future decisions. \sys
-relies on {\em software only}, avoiding the need for additional circuitry to,
-e.g., measure available energy.
-
-%to set the adequate coalesced task
-%size accordingly. This fundamental feature prevents \sys from blind coalesced
-%task adaptation which may result in a repeated power failure problem. For
-%example, If \sys enlarges the coalesced task size by $x$ static tasks on a
-%coalesced task completion and reduces it by the same number of static tasks on
-%a power reboot, then \sys may need to reboot, for instance, ten times to reduce
-%a coalesced task size from ten to one, when $x=1$. Details of coalescing will
-%be given in Section~\ref{sec:task_coalescing}.
+\sys's adaptive task scheduler makes \emph{energy-aware} scheduling decisions to group tasks together to amortize commit and transition costs, or to downscale a task, breaking it into multiple tasks, to avoid non-termination.
+\sys's scheduler uses its recent execution history---no hardware dependency---as a metric to estimate energy availability.
 

### 2018-08-03 paper/asplos19/background.tex
New WAR figure (and explanation)

@@
 
-% \begin{wrapfigure}{t!}{0.5\textwidth}
 \begin{figure}
-	\begin{subfigure}[t]{\linewidth}
-		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
-		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
+    \begin{subfigure}{0.49\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
+        \caption{WAR-affected code}
+        \label{fig:war-example}
     \end{subfigure}
-	\begin{subfigure}[t]{\linewidth}
-		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
-		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
+    \begin{subfigure}{0.49\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
+        \caption{Incorrect execution}
+        \label{fig:war-execution}
     \end{subfigure}
-	\caption{Code example demonstrating checkpointing gives rise to data inconsistency.}\label{fig:code_demo_incosistency}
-% \end{wrapfigure}
+    \caption{WAR dependence example. \texttt{NV} marks non-volatile variables,
+    while \texttt{checkp()} is a checkpoint of volatile state.}
+    \label{fig:war}
 \end{figure}
@@
 
-Checkpointing volatile state alone do not ensure data consistency when the
+Checkpointing volatile state alone does not ensure data consistency when the
 system can directly manipulate non-volatile memory~\cite{mspcdino}.
-Especially, data can get inconsistent when code includes a
+Precisely, data can get inconsistent when code includes a
 \emph{write-after-read} (WAR) dependence between operations that manipulate
@@
 persist across failure and be read by the {\em
-reads}~\cite{ratchet,dino,alpaca}\TODO{figure 2 dependent text ---see
-Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
-inconsistent in an intermittent execution using the cyclic redundancy check
-(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
-maintain data consistent in the presence of WAR dependency is necessary for
-correct execution.}
-%Checkpoints of volatile state preserve intermittent progress, but do not ensure data consistency~\cite{dino,chain,ratchet}.
-Data may become inconsistent if an attempt to execute some computation {\em writes} to a non-volatile variable, then power fails, then a second attempt to re-execute the same computation incorrectly {\em reads} the value written in the first attempt, rather than the variable's original value.
-The situation occurs when code includes a \emph{write-after-read} (WAR) dependence between operations that manipulate non-volatile variables~\cite{ratchet,dino,alpaca}---see Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
-%Figure~\ref{fig:code_demo_incosistency} illgit statuustrates how state can become inconsistent in an intermittent execution using the cyclic redundancy check (CRC) code from MIBench2~\cite{hicks_mibench2_2016}.
-The code computes the CRC for an $\texttt{nBytes}$ byte message $\texttt{msg}$ with remainder $\texttt{Rmnd}$.
-%Despite checkpointing at line 3 in Fig.~\ref{fig:crc_example}, the code may compute  $\texttt{data}$ incorrectly because of the read of $\texttt{Rmnd}$ on line 4 and the write of $\texttt{Rmnd}$ on line 5.
-Figure~\ref{fig:crc_example_war} shows an intermittent execution.
-The second iteration writes $\texttt{Rmnd}$, but power fails before the checkpoint.
-After restarting, line 9 reads the updated value of $\texttt{Rmnd}$, producing an incorrect $\texttt{data}$ value.
-Checkpoint-based systems risk violating memory consistency~\cite{dino}.
-%To maintain consistency, prior systems~\cite{dino} {\em version} a subset of non-volatile data with the checkpoint.
-Task-based systems~\cite{chain,alpaca}, which we focus on, ensure consistency using programmer, compiler, and runtime support.
+reads}~\cite{ratchet,dino,alpaca}.
+Figure~\ref{fig:war} illustrates how state can become
+inconsistent in an intermittent execution using a simple example of an average
+operation over an array of integers.
+The variable \texttt{sum} introduces the WAR, being read and written sequentially
+by the increment operation, thus being erroneously incremented when a power
+failure occurs before updating the index \texttt{i}.
 

### 2018-08-03 paper/asplos19/background.tex
New WAR figure

--- 2018-08-03 paper/asplos19/background.tex
@@
 
-% \begin{wrapfigure}{t!}{0.5\textwidth}
 \begin{figure}
-	\begin{subfigure}[t]{\linewidth}
-		\centering \includegraphics[width=\columnwidth]{figures/crc_example}
-		\caption{Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration.}\label{fig:crc_example}
+    \begin{subfigure}{0.49\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
+        \caption{Subcaption}
+        \label{fig:war-example}
     \end{subfigure}
-	\begin{subfigure}[t]{\linewidth}
-		\centering \includegraphics[width=\columnwidth]{figures/crc_example_war}
-		\caption{Execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)}\label{fig:crc_example_war}
+    \begin{subfigure}{0.49\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
+        \caption{Subcaption}
+        \label{fig:war-execution}
     \end{subfigure}
-	\caption{Code example demonstrating checkpointing gives rise to data inconsistency.}\label{fig:code_demo_incosistency}
-% \end{wrapfigure}
+    \caption{Main caption.}
+    \label{fig:war}
 \end{figure}

### 2018-08-03 paper/asplos19/intro.tex
partial commit

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
 Advances in processor efficiency along with the development of
@@
 energy-harvesting nano-satellites~\cite{kicksat,capybara}.
-
+%
 Energy-harvesting devices create unique challenges because they operate {\em
@@
 operation-power failure-restart cycles. 
-%As an example, recharge times may be tens of seconds in radio frequency powered medical device~\cite[Fig.3c]{nadeau_naturebio_2017}. 
 The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
 $\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.
-
+%
 \textbf{Data Consistency in Intermittent Computing.}
-%Software in an energy-harvesting system operates according to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017}, which corresponds to the operation-failure-restart cycles of an energy-harvesting device.
-In an intermittent software execution, an {\em operating period} proceeds for anarbitrary duration (dictated by energy availability) before being interruptedby a {\em failure period}.
-In a failure period,
 Upon power failures, a device loses the volatile
@@
 Therefore, this work focuses on a key limitation of task-based software approaches, namely, the {\em inflexibility} of statically decomposing a program into tasks.
-
-\begin{figure} %{t!}{0.45\textwidth}
+%
+\begin{figure}
     \centering
     \includegraphics[width=\columnwidth]{figures/intro-figure-vert.pdf}
-    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{cmt} in the figure).
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C} in the figure).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
@@
 \end{figure}
-
+%
 \textbf{Task Decomposition of Intermittent Programs.} {\em
@@
 include arbitrary computation that should be executed despite arbitrarily-timed power failures.
-%and these existing systems guarantee that each
-%tasks executes {\em atomically}, despite arbitrarily-timed power failures.
-%
 The programmer (or a compiler) explicitly expresses task-to-task control flow.
-%
 Figure~\ref{fig:coalesce} illustrates how a program's tasks execute and
@@
 The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.
-
+%
 A programmer may thus create very large tasks in an attempt to reduce
@@
 coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.
-
-%A key challenge is that the length of a software
-%task's execution is \emph{limited by the fixed total amount of energy} that a
-%device can buffer in hardware. A task's code is static, but the duration of its
-%execution may be input-dependent and is difficult to predict. To illustrate
-%this, we refer to Figure~\ref{fig:1} showing the execution time of two
-%applications (the same application X, but divided into X and Y tasks as
-%in~\cite{chain}) running on RFID antenna-powered Computational
-%RFID~\cite{wisp,rf_powered_computing_gollakota_2014} at two device-to-antenna
-%distances (near---stable energy supply/far---high energy intermittency). At
-%short distance: execution will takes too long caused by runtime cost of
-%marshalling excessively small tasks. At far distance: program \emph{might never
-%execute} if task execution consumes more energy than the system can buffer.
-%This calls for at-runtime adaptive task division of any transiently-powered
-%application---namely \emph{task virtualization}.
-
+%
 This paper introduces a new system called \sys, which executes a task-based
@@
 is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).
-
-%Brandon: This is too detailed for the intro:
-%Unfortunately, naively merging two atomic tasks might produce a non-atomic
-%task, that may leave non-volatile memory inconsistent after a partial
-%execution. A merge breaks atomicity when atomic merged tasks form a
-%\emph{write-after-read (WAR)} dependency---for instance if two tasks
-%\texttt{\{x=y+1\}; \{y=x;\}} are merged and if the power failure occurs after
-%\texttt{x=y}, the value \texttt{x} will be increased twice when the merged task
-%is restarted, that leads to an inconsistency.
-
-%variables in non-volatile memory at \emph{run-time} that can
-%break the atomicity of the \emph{virtual task}. Consider the
-%example depicted in Fig.~\ref{fig:virtualization}: Tasks 1,
-%2 and 3 being executed consecutively. All tasks in this
-%example are atomic since they do not have WAR dependency on
-%the persistent variables they are accessing, e.g. \emph{x}
-%is only read and \emph{y} is only written within Task 1.
-%Now, suppose three tasks have been virtualized into a single
-%one at run-time, namely Task 4: since \emph{x} is now first
-%\emph{read} and then \emph{written}, a WAR dependency on
-%\emph{x} is introduced dynamically at
-%run-time---unfortunately Task 4 is \emph{no more atomic}
-%since its re-execution will not always produce the same
-%results.
+%
 \textbf{Challenges and Contributions} 
@@
 Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
-
+%
 \item\textit{challenge} 3) A static task decomposition model assumes that each
@@
 task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
-
-% \textbf{dynamic privatization}: while a compiler can use static data privatization and commit instrumentation
-% (i.e., redo-logging) to eliminate statically identifiable, inter-task data
-% dependences~\cite{alpaca}, a \emph{dynamic} dependence between two coalesced
-% tasks requires dynamic privatization and commit actions.
-% %
-
-%
-%Therefore, we require a \emph{new execution
-%model} that keeps each virtual task atomic by committing the
-%modified persistent variables at the boundary defined
-%dynamically at run-time---keeping the non-volatile memory
-%unmodified upon a power-interrupt and preserving its
-%consistency.
-% \textbf{task non-termination}:
-
-%
-% !!!!
-% It should be emphasized that unlike a voltage-threshold-based protection approaches that may suffer from a non-termination problem if the checkpoint has to be placed within the interrupt-disabled region, a timer-based approach adapts its partial commit location on repeated failures.
-%
-%Task coalescing removes the burden from the programmer, because it accepts any
-%task decomposition and improves it dynamically. To further reduce the
-%programmer effort, we propose a compiler pass for \emph{automatic
-%decomposition} of programs into (small) \emph{atomic} tasks. The compiler
-%identifies non-volatile variables shared across tasks as tasks are created, and
-%instruments reads and writes of those variables, using memory virtualization,
-%to keep the data consistent in the presence of power loss.
-%
-%\textcolor{red}{Despite being limited to a subset of the C language, the
-%automatic task decomposition allowed us to port several	applications to an
-%intermittent platform with a moderate effort.}
-%
 \end{itemize}
-
-% \textbf{Contributions.} To summarize, \sys's main contributions are:
-% %
-% \begin{itemize}
-% \item The first {\em adaptive} task-based intermittent execution model that
-% dynamically \emph{coalesces} tasks to reduce commit overhead and
-% \emph{downscales} tasks to ensure forward progress.
-% \item The first memory virtualization mechanism that preserves the atomicity of coalesced tasks under power failures.
-% %\item A fall-back dynamic task splitting mechanism that preserves forward progress despite executing tasks that are too large for a device's energy supply.
-% \item A fully-realized prototype of \sys that runs on real energy-harvesting devices, to be released as open source after publication via~\cite{coala_website}.
-% \item An evaluation directly comparing \sys to a state of the art task-based
-% intermittent programming and execution model from prior work, showing that on a
-% suite of benchmarks from the literature and several new workloads, \sys often
-% has higher performance and is more flexible to diverse energy conditions.
-% \end{itemize}
 %

### 2018-08-03 paper/asplos19/systemdescription.tex
partial commit

@@
-% \begin{wrapfigure}{t!}{0.45\textwidth}
 \begin{figure}
@@
 	\label{fig:system_overview}
-% \end{wrapfigure}
 \end{figure}
-
+%
 \sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
@@
 \textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by encapsulating the code in a top-level set of functions, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
-% In other words, it requires reasoning similar to prior task-based systems~\cite{chain,alpaca}.
-
-% A programmer may also opt to use compiler support to automatically decompose a program into tasks, leveraging recent work~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
-Without loss of generality, we assume throughout this work that the programmer manually decomposed the program into tasks; \sys's behavior with the automatically decomposed code would be identical.
-
 The programmer compiles their task-based code, and links to \sys runtime,
@@
 \textbf{Adaptive Task Scheduler.} 
-\sys's adaptive task scheduler makes \emph{energy-aware} scheduling decisions to group tasks together to amortize commit and transition costs, or to downscale a task, breaking it into multiple tasks, to avoid non-termination.
-\sys's scheduler uses its recent execution history---no hardware dependency---as a metric to estimate energy availability.
+\sys's adaptive task scheduler makes \emph{energy-aware} scheduling decisions to group tasks together or split a task.
+By coalescing tasks \sys amortizes commit and transition costs, and by a task splitting, after repeated failures on a single task, it avoids the task non-termination problem.
+The scheduler uses its recent execution history---no hardware dependency---as a metric to estimate energy availability and eventually to decide on the dynamic task size.
 
-\sys's scheduler {\em downscales} a task to preserve forward progress when a
-single task repeatedly fails to execute to completion.  For a repeatedly-failing
-task, the scheduler uses a timer that fires at a decreasing interval to
-determine where to execute a {\em partial commit} of the task. Partial
-commit saves progress through a task that runs for too long to complete
-using the device's buffered energy. Eventually, after a series of partial
-commits, the too-long task completes---partial commit favors progress over task atomicity.  
-When a task repeatedly fails, the scheduler uses a timer that counts an operating interval
-that decreases in length with each failure and \sys partially commits at the end of the interval. 
-Section~\ref{sec:task_coalescing} provides details of \sys's task scheduler.
-%and relies on the virtual memory manager to gain a full \emph{flexibility} on
-%where to split a task. These features distinguish \sys task splitting technique
-%from the traditional checkpointing. A checkpointing system that uses voltage
-%threshold to place a checkpoint might cause a program to crash if the
-%checkpoint happened to be placed between a peripheral initialization and its
-%call. If a programmer disables interrupts between the peripheral initialization
-%and the call, the program may fail indefinitely to progress since the
-%checkpoint may only be deferred. The \sys task downscaling manager, on the
-%other hand, triggers a \emph{partial commit} faster than the previous one on a
-%repeated failure to guarantee forward progress. Details of task downscaling
-%will be given in Section~\ref{sec:task_downsizing}.
-
+% HERE -----------------------------
 \textbf{Virtual Memory Manager.} \sys is able to efficiently coalesce

### 2018-08-03 paper/asplos19/background.tex
partial commit

--- 2018-08-03 paper/asplos19/background.tex
@@
 recharging.
-%
-%Batteryless operation has a number of important advantages, making intermittent computing an important research domain.
-Supplying power to billions~\cite{gartner_iot} of embedded computers using batteries is not sustainable.
-The European Commission estimates that more than 160 kilotons of consumer batteries enter the European Union annually~\cite{eu_batteries_2016}.
-Batteries are an environmental risk, are fragile, are limited in their number of charge/discharge cycles, and may require costly physical maintenance that is difficult or impossible deployed (e.g.
-in space~\cite{kicksat}).
-By contrast, super-capacitors are durable, promising millions of charge/discharge cycles~\cite[Sec.
-I]{ongaro_pwre_2012}.
-The main limitation in moving to capacitor-based energy storage is that capacitor energy density---and consequently operating discharge time---is orders of magnitude less than a battery.
-%
-%Given current technology development, battery-less systems are best suited for very long-term sensing and monitoring where access to recharge is either prohibitive or impossible.
-These include battery-less image capture and processing~\cite{naderiparizi_rfid_2015}, animal monitoring~\cite{thomas_jbcs_2012} or implantable~\cite{rodriguez_tbcs_2015} and digestible~\cite{nadeau_naturebio_2017} sensors.
-%
 There are several energy harvesting battery-less hardware platforms. For
@@
 phone~\cite{talla_imwut_2017}.
-
-%In all of the above, the main source of energy harvested is the electromagnetic radiation in the radio frequency range (ambient transmitters such as high power TV transmitters~\cite{liu_sigcomm_2013} or dedicated RFID antenna~\cite{wisp5,moo,talla_imwut_2017,medusa_farsens_2017,holleman_biocas_2008,naderiparizi_rfid_2015}).
-Naturally, other forms of energy harvesting sources exist, including temperature gradient, (micro-)motions, light/sun radiation, vibrations, and body fluid flow (blood, gastric acid).
-Several recent surveys discussing energy harvesting, low-power, embedded systems and intermittent computing at a high level~\cite{paradiso_pvc_2005,soyata_csm_2016,prasad_comst_2014,ku_cst_2016,lucia_snapl_2017}.
 
@@
         \includegraphics[width=\columnwidth]{figures/war-example.pdf}
-        \caption{WAR-affected code}
+        \caption{Subcaption}
         \label{fig:war-example}
@@
         \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
-        \caption{Incorrect execution}
+        \caption{Subcaption}
         \label{fig:war-execution}
     \end{subfigure}
-    \caption{WAR dependence example. \texttt{NV} marks non-volatile variables,
-    while \texttt{checkp()} is a checkpoint of volatile state.}
+    \caption{Main caption.}
     \label{fig:war}
@@
 operating periods.
-%Energy-harvesters provide input power orders of magnitude less than operating power, making recharging negligible during operation.
-%Intermittent execution is different from continuous execution. 
+
 A power failure clears volatile state (e.g., registers and SRAM) while
@@
 
-%\begin{figure}
-%	\centering
-%	\subfloat[Simplified C code snippet of a CRC calculation from~\cite{hicks_mibench2_2016}: per-byte message division by a polynomial; \texttt{NV} denotes non-volatile variable declaration]{\includegraphics[width=\columnwidth]{figures/crc_example}\label{fig:crc_example}}\\
-%	\subfloat[Consecutive execution steps of the loop body in the snippet above: non-volatile checkpointing did not guarantee data consistency as data has been manipulated (line 9) with stale reminder (line 3)]{\includegraphics[width=\columnwidth]{figures/crc_example_war}\label{fig:crc_example_war}}
-%	\caption{Code example demonstrating effect of write after read on volatile memory checkpointing.}
-%	\label{fig:code_demo_incosistency}
-%\end{figure}
-
-Checkpointing volatile state alone does not ensure data consistency when the
+Checkpointing volatile state alone do not ensure data consistency when the
 system can directly manipulate non-volatile memory~\cite{mspcdino}.
-Precisely, data can get inconsistent when code includes a
+Especially, data can get inconsistent when code includes a
 \emph{write-after-read} (WAR) dependence between operations that manipulate
@@
 persist across failure and be read by the {\em
-reads}~\cite{ratchet,dino,alpaca}.
-Figure~\ref{fig:war} illustrates how state can become
-inconsistent in an intermittent execution using a simple example of an average
-operation over an array of integers.
-The variable \texttt{sum} introduces the WAR, being read and written sequentially
-by the increment operation, thus being erroneously incremented when a power
-failure occurs before updating the index \texttt{i}.
+reads}~\cite{ratchet,dino,alpaca}\TODO{figure 2 dependent text ---see
+Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
+inconsistent in an intermittent execution using the cyclic redundancy check
+(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
+maintain data consistent in the presence of WAR dependency is necessary for
+correct execution.}
 
@@
 that no task reads and writes the same copy~\cite{chain}.
-%There are several run time strategies to ensuring task idempotence.
-One approach~\cite{alpaca}, is to identify non-volatile data involved in WAR dependences in a task (like DINO~\cite{dino} and Ratchet~\cite{ratchet} did for checkpoints), execute the task using private copies of those data, and commit the private copies on task completion.
-Another way is to statically create multiple versions of non-volatile data shared by tasks and ensure that no task reads and writes the same version~\cite{chain}.
-Regardless of the strategy, task-based systems execute statically-defined tasks atomically, completing in one or more attempt.
 

### 2018-08-03 paper/asplos19/coalescing.tex
restructure the paper directory
New HG figure inserted (and explained)

--- 2018-04-17 paper/coalescing.tex
@@
 
-\begin{wrapfigure}{t}{0.3\textwidth}
+% \begin{wrapfigure}{t}{0.3\textwidth}
+\begin{figure}{t}
 \fbox{\parbox{\linewidth}{
@@
 	%\small
+	%
 	\begin{algorithmic}[1]
-        \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$H$}
+        \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$z$}
         \State $H \gets 0$ 
@@
 	\end{algorithmic}}}
-\end{wrapfigure}
+% \end{wrapfigure}
+\end{figure}
+
 %	\begin{subfigure}[t]{\linewidth}
@@
 
-\begin{wrapfigure}{t!}{0.3\textwidth}
+\begin{figure}
     \centering
-	\includegraphics[width=0.3\columnwidth]{figures/slowCoal}
+    \begin{subfigure}[b]{0.49\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/slowCoal}
 		\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
 		\label{fig:energyBlind}
-\end{wrapfigure}
+    \end{subfigure}
+    \begin{subfigure}[b]{0.49\columnwidth}
+       	\includegraphics[width=\columnwidth]{figures/energy-aware-coal.pdf}
+		\caption{The {\em history-guided coalescing} algorithm uses its recent history of execution to estimate the available amount of energy and set the coalescing target accordingly.}
+		\label{fig:energyAware}
+    \end{subfigure}
+    % \caption{Pictures of animals}\label{fig:animals}
+\end{figure}
+
+% % \begin{wrapfigure}{t!}{0.3\textwidth}
+% \begin{figure}
+% 	\centering
+% 	\includegraphics[width=0.48\columnwidth]{figures/slowCoal}
+% 	\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
+% 	\label{fig:energyBlind}
+% % \end{wrapfigure}
+% \end{figure}
+
+% % \begin{wrapfigure}{}{0.5\textwidth}
+% \begin{figure}
+% 	\centering
+% 	\includegraphics[width=0.48\columnwidth]{figures/energy-aware-coal.pdf}
+% 	\caption{The {\em history-guided coalescing} algorithm uses its recent history of execution to estimate the available amount of energy and set the coalescing target accordingly.}
+% 	\label{fig:energyAware}
+% % \end{wrapfigure}
+% \end{figure}
 
@@
 
+\begin{figure*}
+    \includegraphics[width=\linewidth]{figures/hg-coal-horiz.pdf}
+    \caption{HG coalescing sample execution across two power cycles.}
+    \label{fig:hg-coal}
+\end{figure*}
+
 The History-Guided (HG) coalescing strategy adapts the coalescing target more
@@
 
-\begin{wrapfigure}{}{0.5\textwidth}
-	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/energy-aware-coal.pdf}
-	\caption{The {\em history-guided coalescing} algorithm uses its recent history of execution to estimate the available amount of energy and set the coalescing target accordingly.}
-	\label{fig:energyAware}
-\end{wrapfigure}
-
 The HG algorithm is characterized by the following functions:
@@
 \begin{itemize}
-\item $f_\text{reboot}(B) = \lfloor(H / 2)\rfloor;$
+\item $f_\text{reboot}(H) = \lfloor(H / 2)\rfloor;$
 \item $f_\text{weight}(T_\text{i}) =  1$; 
@@
 
-Figure~\ref{fig:energyAware} illustrates the operation of the history-guided
-coalescing strategy.  Each white box is a task and each red box is a commit.
-Time goes left to right and consecutive white boxes with no intervening commit
-are coalesced tasks.  A gray box is a task interrupted by a power failure.  At
-the top, the figure shows the $i$th execution period, which has a history
-value of $H^{(i)} = 4$.  $B$ is initially two and decreases to one.  Once the
+Figure~\ref{fig:hg-coal} illustrates the operation of the history-guided
+coalescing strategy.
+The figure shows two power cycles following the $i$th, the latter having an
+execution history $H^{(i)} = 4$.
+Based on that, $B$ is initially set to two and then decreases to one. Once the
 value of $B$ reaches one, HG is, in effect, expecting a power failure,
-justifying the decrease in the coalescing target.  After the power failure, HG
-first updates the history $H = H^{i+1} = 8$. HG then uses the value of $H$ to
-update $B$, setting $B$ equal to four and continuing execution.
+justifying the conservative approach. After the power failure, HG
+uses the most recent history $H^{(i+1)} = 8$ to set $B = 4$
+and continue execution.
 

### 2018-08-04 paper/asplos19/paging.tex
restructure the paper directory
Figure 6 updated and explained through section 5

--- 2018-04-17 paper/paging.tex
@@
     \centering
-	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
-	\caption{Illustration of task-virtual memory interaction in \sys.
+    \includegraphics[width=\columnwidth]{figures/mem-man.pdf}
+    \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP},
+    and order of occurrence. Steps 2, 3, 4 and 5 are conditional. Tasks only have direct access to the
+    \texttt{working} buffer.}
+    \label{figure:mem-man}
+\end{figure}
+
+% \begin{figure}
+	% \centering
+	% \includegraphics[width=\columnwidth]{figures/paging.pdf}
+	% \caption{Illustration of task-virtual memory interaction in \sys.
 Tasks only interact with the working buffer in SRAM (indicated as page buffer).
@@
 The indirection table is used to speed-up two-phase commit operation: it eliminates the copy operation in the second phase.}
-	\label{figure:coala_page_tags}
-\end{figure}
+	% \label{figure:coala_page_tags}
+% \end{figure}
 
-\textbf{System Overview.} Figure~\ref{figure:coala_page_tags} provides an
-overall view of \sys's virtual memory management system.  A page has a home
-location in non-volatile memory which we refer to as the page's
+\textbf{System Overview.}
+Figure~\ref{figure:mem-man} gives an overview of \sys's virtual memory management system,
+whose details are provided throughout this section.
+A page has a home location in non-volatile memory which we refer to as the page's
 \texttt{\underline{store}} (underlined locations are non-volatile).  When a
@@
 
-A \sys task must access protected non-volatile variables through \sys's
-restricted memory interface, which Section~\ref{sec:coala_api} describes in
-detail.  The interface includes \texttt{RP(var)}, to read the value of variable
-{\tt var}, and \texttt{WP(var)} to assign a value to variable {\tt var}. The
+A task must access protected non-volatile variables through \sys's
+restricted memory interface. The interface includes \texttt{RP}$(v)$, to read the value of variable
+$v$, and \texttt{WP}$(v)$ to assign a value to variable $v$. The
 implementation of {\tt RP} is shown in Algorithm~\ref{algo:rwar}, and {\tt
@@
 %\begin{algorithm}[t]
-\begin{wrapfigure}{t}{0.4\textwidth}
+\begin{figure} %{t}{0.4\textwidth}
 \fbox{\parbox{\linewidth}{
@@
 	\end{algorithmic}}}
-\end{wrapfigure}
+\end{figure}
 
-After address translation, \sys attempts to access the protected variable's
-location in the volatile paging buffer. \sys keeps track of the page tags for
+After address translation (Figure~\ref{figure:mem-man}, Step 6), a task
+accesses the protected variable's location in the volatile paging buffer
+(Step 7). \sys keeps track of the page tags for
 the pages currently resident in the paging buffer. When a task accesses a
 variable, it compares the variable's page tag to tags of the pages in {\tt
-working} (Line 2).
+working} (Algorithm~\ref{algo:rwar}, Line 2).
 %
@@
 If the accessed variable's page tag is not found in the page buffer, the
-operation incurs a {\em page fault} (Line 4). At a page fault, \sys must swap
-out one of the pages that is resident in the page buffer (to the non-volatile
-\texttt{\underline{shadow}}) if \texttt{working} is full, and swap in the
-variable's page.
+operation incurs a {\em page fault} (Line 4).
+% At a page fault, \sys must swap
+% out one of the pages that is resident in the page buffer (to the non-volatile
+% \texttt{\underline{shadow}}) if \texttt{working} is full, and swap in the
+% variable's page.
 %
@@
 
+When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory
+manager first searched the variable's page in the \texttt{working} buffer
+(Figure~\ref{figure:mem-man}, Step 1).
+If the page is not found there, a page fault is incurred and a new page needs to
+be swapped in.
 If the \texttt{working} buffer is full, a page fault on a memory access
 requires \sys to swap out one of the pages in the volatile buffer (a
-\emph{victim page}) preserving updates made to that page, and to copy the
-accessed page of protected data into the working buffer.
-Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault. A victim
-page in the \texttt{working} buffer is selected for eviction according to
-a \emph{first-in-first-out} (FIFO) replacement policy (Lines 2 and 12). 
+\emph{victim page}) preserving updates made to that page.
+% and to copy the accessed page of protected data into the working buffer.
+% Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault. A victim
+% page in the \texttt{working} buffer is selected for eviction according to
+% a \emph{first-in-first-out} (FIFO) replacement policy (Lines 2 and 12).
 %
@@
 % 	\scriptsize
-\begin{algorithm}
-	\caption{\texttt{PageFault}(tag $t$)}
-	\label{algo:pagefault}
-	\scriptsize
-	%\small
-	\begin{algorithmic}[1]
-        \State \LeftComment{$i_\text{victim}$: index to free slot or page to be evicted, initially 0}
-        \State \LeftComment{\texttt{\underline{shadowCount}}: non-volatile, 0 on boot unless in phase 2 commit}
-        \State $t_\text{victim} \gets \Call{Tag}{\texttt{working}[i_\text{victim}]}$
-		\If {\Call{isDirty}{working[$i_\text{victim}$]} }	\Comment{Evicting a dirty page?}
-		\State $\texttt{\underline{shadow}}[t_\text{victim}] \xleftarrow{\text{DMA}} \texttt{working}[i_\text{victim}]$
-            \Comment{Swap out the page}
-        \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track for quick traversal}
-            \State $\texttt{\underline{shadowList}}[\texttt{shadowCount++}] \gets t_\text{victim}$
-        \EndIf
-		\EndIf
-		\If {\Call{exists}{\underline{shadow}[$t$]}} \Comment{Page swapped out earlier?}
-		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
-            \Comment{Get from commit buffer}
-		\Else
-		\State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{store}}[t]$
-            \Comment{Get from backing store}
-		\EndIf 
-        \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBufLength}|$ \Comment{FIFO evict}
-	\end{algorithmic}
-\end{algorithm}
+% \begin{algorithm}
+	% \caption{\texttt{PageFault}(tag $t$)}
+	% \label{algo:pagefault}
+	% \scriptsize
+	% %\small
+	% \begin{algorithmic}[1]
+        % \State \LeftComment{$i_\text{victim}$: index to free slot or page to be evicted, initially 0}
+        % \State \LeftComment{\texttt{\underline{shadowCount}}: non-volatile, 0 on boot unless in phase 2 commit}
+        % \State $t_\text{victim} \gets \Call{Tag}{\texttt{working}[i_\text{victim}]}$
+		% \If {\Call{isDirty}{working[$i_\text{victim}$]} }	\Comment{Evicting a dirty page?}
+		% \State $\texttt{\underline{shadow}}[t_\text{victim}] \xleftarrow{\text{DMA}} \texttt{working}[i_\text{victim}]$
+            % \Comment{Swap out the page}
+        % \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track for quick traversal}
+            % \State $\texttt{\underline{shadowList}}[\texttt{shadowCount++}] \gets t_\text{victim}$
+        % \EndIf
+		% \EndIf
+		% \If {\Call{exists}{\underline{shadow}[$t$]}} \Comment{Page swapped out earlier?}
+		% \State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
+            % \Comment{Get from commit buffer}
+		% \Else
+		% \State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{store}}[t]$
+            % \Comment{Get from backing store}
+		% \EndIf
+        % \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBufLength}|$ \Comment{FIFO evict}
+	% \end{algorithmic}
+% \end{algorithm}
 
@@
 If a task modified any byte in the victim page (i.e., using \texttt{WP}), then
-the page is dirty.  If the page being swapped out is not dirty and the page
-being accessed had not been dirtied and swapped out to the \texttt{shadow}
-buffer since the last reboot, then handling a page fault requires copying the
-accessed page from its home in the  \texttt{\underline{store}} into the
-\texttt{working} buffer (Line 11).  Otherwise, the page fault handler must take
-the following extra steps.
-%
-Before swapping in the accessed page, \sys must first preserve the updated
-values in the dirty page so that they can be committed back to their locations
-in memory when the task commits (but not earlier).  To preserve the dirty
-values, \sys copies the victim page to the non-volatile
-\texttt{\underline{shadow}} buffer (Line 5) and records the page's tag in a
-non-volatile commit list (Lines 6-7). \sys uses the commit list to iterate over shadow pages 
-at commit time, to copy the contents of shadow pages back to the main memory {\tt store}. 
+the page is dirty and its changes has to be persisted to non-volatile memory
+(Figure~\ref{figure:mem-man}, Step 2).
+If the accessed page was previously modified and swapped out since the last
+power failure, the most recent version of the page is in {\tt
+\underline{shadow}} (Step 3), otherwise it has to be retrieved from
+{\tt \underline{store}} (Step 4).
+Finally, the page is copied to the volatile buffer (Step 5).
 %
@@
 %
-If the accessed page was previously modified and swapped out since the last
-power failure (Line 8), the most recent version of the page is in {\tt
-\underline{shadow}}, not \texttt{\underline{store}} (Line 9).
 

### 2018-08-04 paper/asplos19/abstract.tex
partial commit

@@
-%Old version of abstarct
-
-%Massively deployed Internet of Things (IoT) must be energy neutral.
-This can only be achieved when the IoT is battery-free and the energy to operate the IoT comes from ambient energy sources.
-However, relying on the highly varying, unpredictable and intermittent harvested energy is at odds with the core assumption that program executions require stable and uninterrupted energy supply.
-The core programming model for the intermittently-powered computation is a discrete-like one that alternates between progressing and protecting the execution progress against power interrupts.
-Generally, intermittent computation models can be classified into (i) checkpointing, where the volatile state of a program is frequently saved to non-volatile memory, or (ii) task-based programming, where a programmer splits the code into small idempotent sections.
-Despite the superiority of task-based systems over checkpointing, task-based programming is completely oblivious to the energy conditions.
-Simply, it cannot increase or decrease the size of the task on the fly (the moment of commit to the non-volatile memory) depending on how long the energy burst is.
-To address this challenge, this paper introduces \sys, an \emph{adaptive} task-based execution model that does not require any dedicated hardware to support task adaptability, which at the same time is \emph{efficient}.
-\sys programs tune the frequency of switching between computation progressing and program state protection based on the available energy, taking the burden of finding the right computation/protection period off the programmer's shoulders.
-\sys defaults to progress on a multi-tasks scale, when energy permits (\emph{efficiency}), and on a sub-task scale when needed (\emph{guarantee}).
-Furthermore, \sys's \emph{memory virtualization} gains \sys full data protection against runtime triggered and compile-time defined memory accesses.
-Results show that \sys is more efficient than the current state-of-art execution model by up to a factor of two and able to progress where its counterparts fail.
-%
-%Alexei's version of abstract
-%
-% Embedded devices in the Internet of Things gain new deployment opportunities when their hardware is free of batteries and powered by ambient energy.
-Energy harvested from the environment, however, is unpredictable and can only power a device intermittently.
-In the intermittent execution model, computation is frequently interrupted by power failures, and to make forward progress, the system must save intermediate program state into non-volatile memory.
-Task-based programming models that execute at the granularity of statically-defined tasks, offer an efficient alternative to checkpointing.
-Existing task-based systems, however, fix their task size at compile time and operate without any regard to the changing energy conditions at runtime.
-When tasks are sized without accounting for available energy, state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
-To address these challenges, we propose \sys, an \emph{adaptive} task-based execution model that progresses on a multi-tasks scale, when energy permits (\emph{efficiency}), and preserves the computation progress on a sub-task scale when needed.
-\sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.
-Our evaluation on a real energy-harvesting device demonstrates the benefits of adaptive task coalescing and downscaling, and highlights \sys's performance improvements over a state-of-art task-based system.
-%
 Freeing embedded devices from batteries is a key factor in developing the energy neutral Internet of Things. 
@@
 Despite the superiority of the task-based approach, however, it fixes the task size at compile time and operates obliviously to changing energy conditions.
-When tasks are sized without accounting for available energy, the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.% When a program divided into tasks for heterogeneous energy buffers it will be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
+When tasks are sized without accounting for available energy, the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
 To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model.

### 2018-08-04 paper/asplos19/background.tex
partial commit

@@
         \includegraphics[width=\columnwidth]{figures/war-example.pdf}
-        \caption{Subcaption}
+        \caption{WAR-affected code}
         \label{fig:war-example}
@@
         \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
-        \caption{Subcaption}
+        \caption{Incorrect execution}
         \label{fig:war-execution}
     \end{subfigure}
-    \caption{Main caption.}
+    \caption{WAR dependence example. \texttt{NV} marks non-volatile variables,
+    while \texttt{checkp()} is a checkpoint of volatile state.}
     \label{fig:war}
@@
 support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. 
-
-Checkpointing volatile state alone do not ensure data consistency when the
+%
+Checkpointing volatile state alone does not ensure data consistency when the
 system can directly manipulate non-volatile memory~\cite{mspcdino}.
-Especially, data can get inconsistent when code includes a
+Precisely, data can get inconsistent when code includes a
 \emph{write-after-read} (WAR) dependence between operations that manipulate
 non-volatile memory, because it allows {\em writes} from a failed attempt to
-persist across failure and be read by the {\em
-reads}~\cite{ratchet,dino,alpaca}\TODO{figure 2 dependent text ---see
-Figure~\ref{fig:code_demo_incosistency} that illustrates how state can become
-inconsistent in an intermittent execution using the cyclic redundancy check
-(CRC) code from MIBench2~\cite{hicks_mibench2_2016}.  Additional method to
-maintain data consistent in the presence of WAR dependency is necessary for
-correct execution.}
+persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
+Figure~\ref{fig:war} illustrates how state can become
+inconsistent in an intermittent execution using a simple example of an average
+operation over an array of integers.
+The variable \texttt{sum} introduces the WAR, being read and written sequentially
+by the increment operation, thus being erroneously incremented when a power
+failure occurs before updating the index \texttt{i}.
 

### 2018-08-04 paper/asplos19/evaluation.tex
partial commit

--- 2018-08-02 paper/asplos19/paper/evaluation.tex
@@
 	\centering
-	\includegraphics[width=0.5\columnwidth]{figures/coalEfficiency}
+	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
     \caption{Breakdown of overhead for the proposed coalescing strategies
@@
 	\centering
-	\includegraphics[width=0.49\columnwidth]{figures/overallOverhead}
-	\includegraphics[width=0.49\columnwidth]{figures/memAccess}
+	\includegraphics[width=\columnwidth]{figures/overallOverhead}
+	\includegraphics[width=\columnwidth]{figures/memAccess}
 	\caption{\sys kernel overhead: a breakdown of time spent in
@@
 \begin{figure}
-	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}%
-	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
+	\includegraphics[width=\columnwidth]{figures/coalStrategies}%
     \caption{\sys's coalescing performance. Left: Application execution time
 with coalescing strategies (HG and WHG) normalized to the execution time
-without coalescing. Right: Execution time with the WHG variant of \sys,
-normalized to the execution time with Alpaca, as a function of distance to the
-energy source.}
+without coalescing.}
 	\label{fig:coalescing}
+\end{figure}
+
+\begin{figure}
+	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
+    \caption{\sys's execution time, normalized to the execution time 
+    with Alpaca, as a function of distance to the energy source 
+    measured in meter.}
 	\label{fig:runtime}
 \end{figure}
-
-%\begin{wrapfigure}{t!}{0.5\textwidth}
-%	%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.5\columnwidth]{figures/coalStrategies}
-%	\caption{Coalescing strategies (HG and WHG) run time per application on intermittent power (RF generator at 15\,cm from the WISP), compared to \sys without coalescing (NC).
-Coalescing provides significant improvements: from 0.25 (\textbf{ar}) up to 0.70 (\textbf{sort}) compared to a non-coalescing system.}
-%	\label{fig:coalescing}
-%	%\end{figure} 
-%\end{wrapfigure}
-%
-%\begin{wrapfigure}{t!}{0.5\textwidth}
-%	%\begin{figure}
-%	\centering
-%	\includegraphics[width=0.5\columnwidth]{figures/coala_alpaca_gcc}
-%    \caption{Application execution time with the WHG variant of \sys,
-%normalized to the execution time with Alpaca, as a function of distance to the
-%signal generator antenna. \sys outperforms Alpaca on the majority but not all
-%of the benchmarks.}
-%	\label{fig:runtime}
-%%\end{figure}
-%\end{wrapfigure}
-
-%\textbf{Benefits of Adaptive Tasks.}
 

### 2018-08-04 paper/asplos19/background.tex
WAR figure updated

@@
 operation over an array of integers.
-The variable \texttt{sum} introduces the WAR, being read and written sequentially
-by the increment operation, thus being erroneously incremented when a power
-failure occurs before updating the index \texttt{i}.
+The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially
+by the increment operation.
+If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).
 

### 2018-08-04 paper/asplos19/systemdescription.tex
revising system description

@@
 
-% HERE -----------------------------
 \textbf{Virtual Memory Manager.} \sys is able to efficiently coalesce
-tasks because of its efficient virtual memory manager. \sys's memory manager
-paginates memory and ensures that data in a page remain consistent despite
-power interruptions. \sys allows a task to manipulate data in a volatile copy
-of a page only. Pages swap between volatile and non-volatile memory, depending
-on the capacity of the volatile memory and the program's access pattern. \sys
-tracks a task's memory accesses efficiently at page granularity (rather than
-using, e.g., costly word-granular tracking). When a task ends, each page it
-accessed commits from volatile memory (or from a non-volatile swap region for
-dirty pages) back to the non-volatile main memory. Pages efficiently,
-atomically commit using a two-phase commit procedure accelerated using hardware
-support for direct memory access (DMA). Section~\ref{sec:memory_virtulaization}
-provides details of \sys's memory manager.
+tasks because of its efficient virtual memory manager (VMM).
+the VMM privatizes a group of memory pages demanded by a coalesced task, and limits an application access only to the fast volatile memory.
+It persists all pages modifications in non-volatile memory on a coalesced task transition.
+Section~\ref{sec:memory_virtulaization} provides details of VMM.

### 2018-08-04 paper/asplos19/background.tex
revise intro

@@
 persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
-Figure~\ref{fig:war} illustrates how state can become
+Fig.~\ref{fig:war} illustrates how state can become
 inconsistent in an intermittent execution using a simple example of an average

### 2018-08-04 paper/asplos19/intro.tex
revise intro

@@
 %
-\textbf{Task Decomposition of Intermittent Programs.} {\em
-Task-based} programming and execution models require a
+\textbf{Task Decomposition of Intermittent Programs.} Task-based programming and execution models require a
 programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
-statically decompose a program into a collection of {\em tasks}.  A task can
+statically decompose a program into a collection of tasks.  A \emph{task}, a top level function, can
 include arbitrary computation that should be executed despite arbitrarily-timed power failures.
 The programmer (or a compiler) explicitly expresses task-to-task control flow.
-Figure~\ref{fig:coalesce} illustrates how a program's tasks execute and
-shows how tasks can impose a run-time overhead.
-At each task transition, the system incurs an overhead to track and atomically \emph{commit} modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
+Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and
+shows how tasks transitions can impose a run-time overhead.
+At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
 The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.
 %
-A programmer may thus create very large tasks in an attempt to reduce
-overhead by minimizing the number of task transitions. However, a very large
-task may require more energy to complete than a device's fixed hardware energy
-buffer can hold which may lead to a program non-termination problem. To eliminate this risk,
-existing systems require the programmer to decompose a program
-into small tasks, all of which complete using only buffered energy. These
-constraints on task sizing leave an unsatisfying dilemma: large, efficient
-tasks that risk non-termination, or small tasks that are guaranteed to
-complete, but incur a high task transition and commit overhead. In this work,
-we propose a third way, using a novel technique called {\em adaptive execution by dynamic task
-coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.
+A programmer may thus create very large tasks in an attempt to reduce task transitions
+overhead. However, a very large task may require more energy to complete than a device's fixed hardware energy
+buffer can hold which may lead to a task non-termination problem. To eliminate this risk,
+existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress.
+These constraints on task sizing leave an unsatisfying dilemma: large, efficient tasks that risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
+In this work,
+we propose a third way, using a novel technique called \emph{adaptive task size execution by task coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.
 %
-This paper introduces a new system called \sys, which executes a task-based
+This paper introduces a new system called \emph{\sys}, which executes a task-based
 intermittent program by dynamically coalescing/splitting its tasks.
-\sys accepts any static decomposition and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute
-with no commit overhead between them, instead performing
-task-end commit actions at the end of the second task only.
-%
-When there is sufficient energy to execute both tasks, two distinct, small
-tasks are effectively executed as a single large task, preserving the atomicity
-of both. If power fails during a coalesced task, execution restarts from the
-{\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
-illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution.
-On the other hand, a large task that requires more energy to complete than a device buffered energy
-is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).
-%
+\sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce} (top)) or splits them (Fig.~\ref{fig:coalesce} (bottom)) based on the estimated energy availability. 
+\sys does not requires hardware support for energy estimation to make its scheduling decisions.
+
+% \sys accepts any static decomposition and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute
+% with no commit overhead between them, instead performing task-end commit actions at the end of the second task only.
+% %
+% When there is sufficient energy to execute both tasks, two distinct, small
+% tasks are effectively executed as a single large task, preserving the atomicity
+% of both. If power fails during a coalesced task, execution restarts from the
+% {\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
+% illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution.
+% On the other hand, a large task that requires more energy to complete than a device buffered energy
+% is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).
+
 \textbf{Challenges and Contributions} 

### 2018-08-04 paper/asplos19/intro.tex
revise the intro

@@
 
-% \sys accepts any static decomposition and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute
-% with no commit overhead between them, instead performing task-end commit actions at the end of the second task only.
-% %
-% When there is sufficient energy to execute both tasks, two distinct, small
-% tasks are effectively executed as a single large task, preserving the atomicity
-% of both. If power fails during a coalesced task, execution restarts from the
-% {\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
-% illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution.
-% On the other hand, a large task that requires more energy to complete than a device buffered energy
-% is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).
-
 \textbf{Challenges and Contributions} 
 \begin{itemize}
-\item\textit{challenge} 1) Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
+\item\textit{challenge} 1) Given unpredictable energy shots arrival, how to specify the size of a coalesced task on the fly?
 \textit{contribution---1} \sys addresses this challenge by using its \textit{recent execution history} as a metric to estimate the environment energy conditions.
-Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.
+Furthermore, It mitigates the risk of coalesced task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.
 %
 \item\textit{challenge} 2) Given dynamic variables dependencies, how to ensure efficient data protection against power interrupts?
-\textit{contribution---2} \sys ensures the consistency of the data by using a novel
-approach called \emph{dynamic group privatization}.
+\textit{contribution---2} The \sys's Virtual Memory Manager (VMM) ensures the consistency of the data by using a novel
+approach called \emph{dynamic page privatization}.
 Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e.
-protecting all global variables) in each dynamic task transition.
+protecting all global variables) in each colesced task transition.
 \sys, therefore, utilizes a real-time data tracking to enable partial data protection on a task transition.
 Individual variables tacking, however, slows down a system dramatically.
-Therefore, \sys  relies on a coarse locality principle and protects the global variables in a batch.
+Therefore, \sys  relies on a coarse locality principle and privatize memory pages.
 Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
@@
 energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
-Partial execution avoids non-termination by committing the intermediate state of a long-running
-task that has repeatedly failed and restarted. Partial execution violates
+Partial commit avoids non-termination by committing the intermediate state of a long-running
+task that has repeatedly failed and restarted. Partial commit violates
 task atomicity, but preserves forward progress; if a programmer knows that
-task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
+task atomicity is crucial to correctness, they can disable partial execution around a piece of code.\textcolor{red}{how this is different from Rachet timer}
 \end{itemize}

### 2018-08-04 paper/asplos19/coalescing.tex
change the title
revise upto coalescing strategies

@@
 for a device's energy buffer. 
-
+% 
 \subsection{Task Coalescing}
 \label{sec:task_coalescing}
-
-% \begin{wrapfigure}{t!}{0.5\textwidth}
-% 	\centering
-% 	\includegraphics[width=0.5\columnwidth]{figures/coal_Intro_Figure.pdf}
-% 	\caption{When $N$ tasks coalesced $N-1$ commit operations are skipped.}
-% 	\label{fig:coalIntro}
-% \end{wrapfigure}
-
+% 
 When a device's buffered energy is sufficient to run multiple consecutive tasks
 without a power failure, committing state after each task is an unnecessary
-overhead. \sys eliminates this overhead by {\em coalescing} a sequence of
-consecutive tasks. \sys coalesces a sequence of tasks by deferring the commit
-operations for all tasks to the end of the sequence, performing their commits
-as a batch.  In general, committing involves moving state manipulated by a task
-into its permanent location in non-volatile main memory.  In particular, \sys's
+overhead. \sys eliminates this overhead by coalescing a sequence of tasks and deferring the commit
+operations for all tasks to the end of the sequence. 
+In general, committing involves moving state manipulated by a task
+into its permanent location in non-volatile memory.  In particular, \sys's
 commit procedure copies dirty pages of memory that a task updated between fast,
 volatile working memory and slower, non-volatile main memory; we defer the
-details of paging and commit to Section~\ref{sec:memory_virtulaization}.
+details of paging privatization and commit to Section~\ref{sec:memory_virtulaization}.
 
-When \sys coalesces a sequence of $N$ tasks, \sys avoids the cost of commit at
-the first $N-1$ boundaries. The first $N-1$ inter-task transitions leave
-dirty pages in working memory, deferring the work of committing dirty modified
-until the commit of the last task in the sequence.  Coalescing tasks brings a
-double benefit. First, coalescing reduces the total number of relatively
+Coalescing tasks brings a double benefit. First, coalescing reduces the total number of relatively
 high-latency, high-energy non-volatile memory accesses because tasks primarily
@@
 
-%it fasten applicaiton exeution be skipping unnecessary data commits operations
-%when there is abandond energy.
+An effective colescing strategy must be {\em aggressive} enough, attempting to coalesce a
+large number of tasks to amortize a large amount of commit overhead. However, it must be 
+{\em conservative} enough, attempting to coalesce only as many tasks as will execute
+to completion given a certain energy conditions, reducing the risk or coalesced task re-execution penalty.
 
-The effectiveness of coalescing tasks hinges on having an effective {\em
-coalescing strategy}, which determines how many consecutive tasks to coalesce.
-An effective strategy must be {\em aggressive} enough, attempting to coalesce a
-large number of consecutive tasks to amortize a large amount of commit overhead.   
+% An effective coalescing strategy moderates the risk of a power failure during a
+% long coalesced task and capitalizes on the benefits of deferred commits by
+% adapting the target size of the dynamic task to the energy available at
+% runtime. 
 %
-If power fails during a long sequence of coalesced tasks, execution will
-restart from the last commit, i.e. the first task in the sequence, losing the
-progress made by any of the coalesced tasks.
-%
-An effective strategy must also be {\em
-conservative} enough, attempting to coalesce only as many tasks as will execute
-to completion using the device's available energy, but no more.
-%
-An effective coalescing strategy moderates the risk of a power failure during a
-long coalesced task and capitalizes on the benefits of deferred commits by
-adapting the target size of the dynamic task to the energy available at
-runtime.
-
-%\textcolor{red}{maybe is possible to demonstrate with data? count number of executed non-volatile vs volatile accesses}
-%
-
-% % Carlo pseudo code 
-% \begin{algorithm}[t]
-% 	\caption{Coalescing}
-% 	\label{algo:genCoalescing}
-% 	\scriptsize
-% 	%\small
-% 	\begin{algorithmic}[1]
-%         \State $H_\text{p} \gets H_\text{c}$
-%         \State $H_\text{p} \gets 0$ 
-%         \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$B, H_\text{p}$}
-
-%         \While{$\texttt{true}$}
-% 	        \State $ B_\text{c} \gets B$
-% 	        %
-% 	        \While{$ B_\text{c} > 0$}
-% 		        \State \Call{execute\_task}{$T_\texttt{i}$}
-% 		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
-% 		        \State $B_\text{c} \gets B_\text{c} - W$
-% 				\State $H_\text{c} \gets H_\text{c} + W$
-% 	        \EndWhile
-% 	        %
-% 	         \State $B \leftarrow $ \Call{$f_\text{compl}$}{B, $H_\text{c}$,$H_\text{p}$}
-% 	        \State \Call{commit\_to\_fram}{\null}
-%         \EndWhile
-% 	\end{algorithmic}
-% \end{algorithm}
-% %
-% To describe the proposed strategies we will use the following parameters:
-% \begin{itemize}
-% \item target budget $B$: total number of static tasks to be coalesced.
-% \item current budget $B_\text{c}$: remaining number of tasks to be coalesced before committing to non-volatile memory; 
-% \item history $H$: up on reboot, it is the number of executed static tasks between the last two power interrupts.
-While executing, it is the number of executed static tasks since the last power interrupt.
-% \end{itemize}
-%
-%
-% Modifiying Carlo's pseudo code a little 
-%\begin{algorithm}[t]
-%	\caption{Coalescing}
-%	\label{algo:genCoalescing}
-%	\scriptsize
-%	%\small
-%	\begin{algorithmic}[1]
-%        \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$H$}
-%        \State $H \gets 0$ 
-%        \While{$\texttt{true}$}       
-%	        \State $ i \gets 0$
-%	        %
-%	        \While{$ i <= B$}
-%		        \State \Call{execute\_task}{$T_\texttt{i}$}
-%		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
-%		        \State $i \gets i + W$
-%				\State $H \gets H + W$
-%	        \EndWhile
-%	        %
-%	        \State \Call{commit\_to\_fram}{\null}
-%	        \State $B \leftarrow $ \Call{$f_\text{commit}$}{$B$}
-%        \EndWhile
-%	\end{algorithmic}
-%\end{algorithm}
-
-% \begin{wrapfigure}{t}{0.3\textwidth}
 \begin{figure}{t}
@@
 \end{figure}
-
-%	\begin{subfigure}[t]{\linewidth}
-%	\end{subfigure}
 % 
-%
-
-\noindent \textbf{Task Coalescing Strategies.} Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing strategy.
-In the algorithm, $B$ (target budget) is total number of static tasks
+\subsection{Task Coalescing Strategies} 
+Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing strategy. In the algorithm, $B$ (target budget) is total number of static tasks
 that \sys will next attempt to coalesce. $T_\text{i}$ is the $i$th task
@@
 possible coalescing strategies that we developed for \sys. 
-
+% 
 \subsubsection{Energy-oblivious Coalescing (EO)}
 \label{subsec:energyBlind}
-
+% 
 A very simple {\em energy-oblivious} coalescing strategy treats all tasks as
@@
 failure. The characteristic operations of the energy-oblivious strategy are: 
-
-\begin{figure}
-    \centering
-    \begin{subfigure}[b]{0.49\columnwidth}
-        \includegraphics[width=\columnwidth]{figures/slowCoal}
-		\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
-		\label{fig:energyBlind}
-    \end{subfigure}
-    \begin{subfigure}[b]{0.49\columnwidth}
-       	\includegraphics[width=\columnwidth]{figures/energy-aware-coal.pdf}
-		\caption{The {\em history-guided coalescing} algorithm uses its recent history of execution to estimate the available amount of energy and set the coalescing target accordingly.}
-		\label{fig:energyAware}
-    \end{subfigure}
-    % \caption{Pictures of animals}\label{fig:animals}
-\end{figure}
-
-% % \begin{wrapfigure}{t!}{0.3\textwidth}
-% \begin{figure}
-% 	\centering
-% 	\includegraphics[width=0.48\columnwidth]{figures/slowCoal}
-% 	\caption{The {\em energy-oblivious coalescing} algorithm \emph{linearly} updates its coalescing target.}
-% 	\label{fig:energyBlind}
-% % \end{wrapfigure}
-% \end{figure}
-
-% % \begin{wrapfigure}{}{0.5\textwidth}
-% \begin{figure}
-% 	\centering
-% 	\includegraphics[width=0.48\columnwidth]{figures/energy-aware-coal.pdf}
-% 	\caption{The {\em history-guided coalescing} algorithm uses its recent history of execution to estimate the available amount of energy and set the coalescing target accordingly.}
-% 	\label{fig:energyAware}
-% % \end{wrapfigure}
-% \end{figure}
-
+%
 \begin{itemize}
@@
 \end{itemize}
-
+%
 The energy-oblivious strategy reacts slowly to the variation in the energy
@@
 %
-Figure~\ref{fig:energyBlind} shows how this strategy may experience a 
+Fig.~\ref{fig:energyBlind} shows how this strategy may experience a 
 large number of repeated power failures without forward progress.
@@
 %
-
 A key limitation of this algorithm lies in the equality of the target decrease
@@
 decreasing by $k \times x$, but only after a long time. 
-
+%
 \subsubsection{History-Guided Coalescing (HG)}
 \label{subsec:energyAware}
-
+%
 \begin{figure*}
     \includegraphics[width=\linewidth]{figures/hg-coal-horiz.pdf}
-    \caption{HG coalescing sample execution across two power cycles.}
+    \caption{HG coalescing sample execution across two power cycles. \textcolor{red}{explain the meaning of the letters and the logical connection between the left and the right (or cycle 1 and 2)}}
     \label{fig:hg-coal}
 \end{figure*}
-
+%
 The History-Guided (HG) coalescing strategy adapts the coalescing target more
 quickly, addressing a key limitation of the EO coalescing strategy.  
-%
 HG gives each task unity weight, as with EO.  
-%
 HG differs from EO in how it updates $B$ on commit, and after a power failure.
-%
 HG assumes that as the duration of execution since a power failure increases,
@@
 sequence of coalesced tasks increases.  
-%
 Consequently, HG decreases the coalescing target by half with each successful commit.
-%
 HG is {\em history-guided} because it initializes the coalescing target to be
@@
 of power failures.
-%
 While a history of $H$ suggests that there is sufficient energy to coalesce and
@@
 of $H$ moderating the risk of a very long coalesced task. 
-%
 The fraction of the history by which HG adjusts $B$ is arbitrary.
@@
 The HG algorithm is characterized by the following functions: 
-
+%
 \begin{itemize}
@@
 \end{itemize}
-
-Figure~\ref{fig:hg-coal} illustrates the operation of the history-guided
+%
+Fig.~\ref{fig:hg-coal} illustrates the operation of the history-guided
 coalescing strategy.
@@
 and continue execution.
-
-%, without losing generality or requiring hardware support,
-%Moreover, it takes advantage of the guarantees that the energy buffer offers on a fresh start by coalescing many static tasks.
-
+%
 \subsubsection{Weighted History-Guided Coalescing (WHG)}
 \label{subsec:energyTaskAware}
-
+%
 The Weighted History-Guided (WHG) coalescing strategy accounts for the

### 2018-08-04 paper/asplos19/coalescing.tex
change the env for the coalescing equations

@@
 An effective colescing strategy must be {\em aggressive} enough, attempting to coalesce a
-large number of tasks to amortize a large amount of commit overhead. However, it must be 
+large number of tasks to amortize a large amount of commit overhead. However, it must also be 
 {\em conservative} enough, attempting to coalesce only as many tasks as will execute
-to completion given a certain energy conditions, reducing the risk or coalesced task re-execution penalty.
+to completion given a certain energy conditions, reducing the risk or long coalesced task re-execution penalty.
 
@@
 that \sys will next attempt to coalesce. $T_\text{i}$ is the $i$th task
-executed since the last power failure. $H$ (history) is the number tasks
-(coalesced or not) executed between the two most recent power failures.
-$W_\text{i}$ is the {\em weight} of a task $T_\text{i}$.  A task's weight is an
+executed since the last power failure. $H$ (history) is the total number of
+tasks between the last executed task and the power failure that preceding it.
+$W_\text{i}$ is the {\em weight} of a task $T_\text{i}$.  A task's weight $W$ is an
 arbitrary quantity associated with the task that represents its cost in time or
@@
 updates $B$ after a successful commit. The following sections detail several
-possible coalescing strategies that we developed for \sys. 
+possible coalescing strategies that we developed for \sys, albeit not the only possible.
+In fact, our system allows developers to design their own strategy to implement and link to Coala’s task scheduler.
 % 
-\subsubsection{Energy-oblivious Coalescing (EO)}
+\subsubsection{Energy-Oblivious Coalescing (EO)}
 \label{subsec:energyBlind}
 % 
-A very simple {\em energy-oblivious} coalescing strategy treats all tasks as
+A very simple {\em energy-Oblivious} coalescing strategy treats all tasks as
 having unity weight and varies the size of a coalesced task linearly. In such a
@@
 %
-\begin{itemize}
-\item $f_\text{reboot}(B) = B - x $;
-\item $f_\text{weight}(T_\text{i}) =  1$; 
-\item $f_\text{commit}(B) = B + x$; 
-\end{itemize}
+\begin{equation}
+	\begin{split}
+		f_\text{reboot}(B)  		&	 = B - x \\
+		f_\text{weight}(T_\text{i}) & 	 =  1 \\
+		f_\text{commit}(B) 			& 	 = B + x
+	\end{split}
+\end{equation}
 %
@@
 %
-The History-Guided (HG) coalescing strategy adapts the coalescing target more
+The \emph{History-Guided} (HG) coalescing strategy adapts the coalescing target more
 quickly, addressing a key limitation of the EO coalescing strategy.  
@@
 %
-\begin{itemize}
-\item $f_\text{reboot}(H) = \lfloor(H / 2)\rfloor;$
-\item $f_\text{weight}(T_\text{i}) =  1$; 
-\item $f_\text{commit}(B) = \lfloor(B / 2)\rfloor;$ 
-\end{itemize}
+\begin{equation}
+	\begin{split}
+		 f_\text{reboot}(H) 			& = \lfloor(H / 2)\rfloor \\
+		 f_\text{weight}(T_\text{i}) 	& =  1 \\
+		 f_\text{commit}(B) 			& = \lfloor(B / 2)\rfloor
+	\end{split}
+\end{equation}
 %

### 2018-08-04 paper/asplos19/intro.tex
intro modified

--- 2018-08-04 paper/asplos19/intro.tex
@@
 %
-\textbf{Data Consistency in Intermittent Computing.}
+%\textbf{Data Consistency in Intermittent Computing.}
+
 Upon power failures, a device loses the volatile
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/intro-figure-vert.pdf}
+    \includegraphics[width=0.7\columnwidth]{figures/intro-figure-vert.pdf}
     \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C} in the figure).
@@
 %
-\textbf{Task Decomposition of Intermittent Programs.} Task-based programming and execution models require a
+%\textbf{Task Decomposition of Intermittent Programs.}
+
+\noindent\textbf{Drawback of Static Task Decomposition.} 
+Task-based programming and execution models require a
 programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
@@
 overhead. However, a very large task may require more energy to complete than a device's fixed hardware energy
-buffer can hold which may lead to a task non-termination problem. To eliminate this risk,
+buffer can hold which may lead to a task \emph{non-termination} problem. To eliminate this risk,
 existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress.
@@
 %
+
+\noindent\textbf{Challenges and Contributions.}
 This paper introduces a new system called \emph{\sys}, which executes a task-based
-intermittent program by dynamically coalescing/splitting its tasks.
-\sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce} (top)) or splits them (Fig.~\ref{fig:coalesce} (bottom)) based on the estimated energy availability. 
-\sys does not requires hardware support for energy estimation to make its scheduling decisions.
+intermittent program by \emph{dynamically} coalescing/splitting its tasks.
+\sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce} (top)) or splits them (Fig.~\ref{fig:coalesce} (bottom)) based on the estimated energy availability without demanding any hardware support.
+The unique challenges and contributions revealed by this work are listed as follows (C1--C3):
+
+% \sys does not requires hardware support for energy estimation to make its scheduling decisions.
 
@@
 
-\textbf{Challenges and Contributions} 
-\begin{itemize}
-\item\textit{challenge} 1) Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
+%\textbf{Challenges and Contributions}
+
+\noindent\emph{C1) Dynamic Task Sizing:} Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
+\sys addresses this challenge by using a novel approach: \textit{recent execution history} as a metric to estimate the environment energy conditions and to specify task sizes.
+%Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.
+
+\noindent\emph{C2) Dynamic Memory Consistency Handling:} Specifying task sizes on the fly gives rise to the need for dynamic memory consistency handling: how to detect dynamic variables dependencies and ensure data protection efficiently against power interrupts?
+\sys ensures the consistency of the data by using a novel
+approach called \emph{dynamic group privatization}--- %Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e.
+protecting all global variables) in each dynamic task transition.
+utilizing a real-time data tracking to enable partial data protection on a task transition.
+Individual variables tacking, however, slows down a system dramatically.
+Therefore, \sys  relies on a coarse locality principle and protects the global variables in a batch.
+Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
+
+\noindent\emph{C3) Ensuring Task Termination:} A static task decomposition model assumes that each
+single task can execute to completion. If the hardware energy buffer provides
+inadequate energy to execute each single task to completion, a program will not
+terminate~\cite{cleancut_2018}. Under such conditions how to enable the dynamic execution model to progress on a sub-task level? To avoid non-termination under adversarial
+energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism.
+Partial execution avoids non-termination by committing the intermediate state of a long-running
+task that has repeatedly failed and restarted. 
+%Partial execution violates task atomicity, but preserves forward progress; if a programmer knows that task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
+
+%\begin{itemize}
+%\item\textit{challenge} 1) Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
 \textit{contribution---1} \sys addresses this challenge by using its \textit{recent execution history} as a metric to estimate the environment energy conditions.
 Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.
-%
-\item\textit{challenge} 2) Given dynamic variables dependencies, how to ensure efficient data protection against power interrupts?
+%%
+%\item\textit{challenge} 2) Given dynamic variables dependencies, how to ensure efficient data protection against power interrupts?
 \textit{contribution---2} \sys ensures the consistency of the data by using a novel
-approach called \emph{dynamic group privatization}.
+%approach called \emph{dynamic group privatization}.
 Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e.
@@
 Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
+%%
+%\item\textit{challenge} 3) A static task decomposition model assumes that each
+%single task can execute to completion. If the hardware energy buffer provides
+%inadequate energy to execute each single task to completion, a program will not
+%terminate~\cite{cleancut_2018}. under such conditions how to enable the dynamic execution model to progress on a sub-task level? \textit{contribution---3} To avoid non-termination under adversarial
+%energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
+%Partial execution avoids non-termination by committing the intermediate state of a long-running
+%task that has repeatedly failed and restarted. Partial execution violates
+%task atomicity, but preserves forward progress; if a programmer knows that
+%task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
+%\end{itemize}
 %
-\item\textit{challenge} 3) A static task decomposition model assumes that each
-single task can execute to completion. If the hardware energy buffer provides
-inadequate energy to execute each single task to completion, a program will not
-terminate~\cite{cleancut_2018}. under such conditions how to enable the dynamic execution model to progress on a sub-task level? \textit{contribution---3} To avoid non-termination under adversarial
-energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
-Partial execution avoids non-termination by committing the intermediate state of a long-running
-task that has repeatedly failed and restarted. Partial execution violates
-task atomicity, but preserves forward progress; if a programmer knows that
-task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
-\end{itemize}
-%
-Section~\ref{sec:background} provides background on intermittent computing.
+The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
 Section~\ref{sec:systemdescription} provides an overview of \sys, while

### 2018-08-04 paper/asplos19/systemdescription.tex
revise intro

@@
 \sys accomplishes this goal with a constellation of a new programming model and runtime software system support, that supports dynamically adaptive task-based execution.
-Figure~\ref{fig:system_overview} shows an overview of \sys.
+Fig.~\ref{fig:system_overview} shows an overview of \sys.
 
@@
 
-\textbf{Adaptive Task Scheduler.} 
-\sys's adaptive task scheduler makes \emph{energy-aware} scheduling decisions to group tasks together or split a task.
+\textbf{Adaptive Task Scheduler(ATS).} 
+\sys's adaptive task scheduler (ATS) makes \emph{energy-aware} scheduling decisions to group tasks together or split a task.
 By coalescing tasks \sys amortizes commit and transition costs, and by a task splitting, after repeated failures on a single task, it avoids the task non-termination problem.
@@
 
-\textbf{Virtual Memory Manager.} \sys is able to efficiently coalesce
+\textbf{Virtual Memory Manager (VMM).} \sys is able to efficiently coalesce
 tasks because of its efficient virtual memory manager (VMM).

### 2018-08-04 paper/asplos19/intro.tex
intro modified

@@
 
-\noindent\textbf{Drawback of Static Task Decomposition.} 
+\noindent\textbf{Crucial Drawback of Static Task Decomposition.} 
 Task-based programming and execution models require a
@@
 %
+
 The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.

### 2018-08-04 paper/asplos19/intro.tex
intro modified

@@
 
+\todo{Add a summary of our extensive evaluations.}{Amjad, Carlo} To asses the benefits of \sys over existing task-based systems, we .... 
+Our evaluation shows that \sys is $2X$ more energy-efficient than the state-of-the-art system in most cases and is able to progress where static systems fail---\emph{the only task-based system that guarantees task termination in any arbitrary task division}.
+
 The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.

### 2018-08-04 paper/asplos19/coalescing.tex
Energy guided

@@
 % 
-A very simple {\em energy-Oblivious} coalescing strategy treats all tasks as
+A very simple {\em energy-oblivious} coalescing strategy treats all tasks as
 having unity weight and varies the size of a coalesced task linearly. In such a
@@
 required to execute different tasks and to the variation in the effective quantum of
-energy available to the device (i.e., variation in energy buffer size). 
-%
-Fig.~\ref{fig:energyBlind} shows how this strategy may experience a 
-large number of repeated power failures without forward progress.
-%
-With the successful commit of each coalesced task sequence, the coalescing
-target increases.  Eventually, the target may be too high and a coalesced task
+energy available to the device. With the successful commit of each coalesced task sequence, the coalescing target increases.  Eventually, the target may be too high and a coalesced task
 composed of only fewer tasks will commit without interruption by a power
@@
 %
-The strategy then slowly decreases the coalescing target, eventually reaching a
+The strategy then linearly decreases the coalescing target, eventually reaching a
 coalescing target that allows completion.
@@
 A key limitation of this algorithm lies in the equality of the target decrease
-in $f_\text{reboot}(B)$ and the increase in $f_\text{commit}(B)$.  If, after
-some number $k$ successful commits the target must decrease back to its
-original value before the $k$ commits, the algorithm must then decrease $B$ by
-$k \times x$.  To accumulate this decrease, this strategy requires $k$ power
-failures, after each of which the strategy decreases $B$ by $x$, eventually
-decreasing by $k \times x$, but only after a long time. 
+in $f_\text{reboot}$ and the increase in $f_\text{commit}$.
+Let us assume $x=1$.
+If, after $k$ successful commits the target must decrease to its original value $B-k$ due to an energy drop, this strategy requires $k$ successive power failures.
 %
-\subsubsection{History-Guided Coalescing (HG)}
+\subsubsection{Energy-Guided Coalescing (EG)}
 \label{subsec:energyAware}
@@
 %
-The \emph{History-Guided} (HG) coalescing strategy adapts the coalescing target more
-quickly, addressing a key limitation of the EO coalescing strategy.  
-HG gives each task unity weight, as with EO.  
-HG differs from EO in how it updates $B$ on commit, and after a power failure.
-HG assumes that as the duration of execution since a power failure increases,
-the amount of energy remaining decreases, and the risk of failing during a long
-sequence of coalesced tasks increases.  
-Consequently, HG decreases the coalescing target by half with each successful commit.
-HG is {\em history-guided} because it initializes the coalescing target to be
-equal to $H$, the total number of tasks executed between the most recent pair
-of power failures.
+The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions.
+It uses its recent execution history as
+mean to measure energy availability and alters its target accordingly.
+By relying on the history of execution EG eliminates the problem of frequent power failures a single coalesced task.
+In fact EG can reduce its target to one from any coalesced target in at most two successive power failures.
+For instance, if the execution is interrupted after finishing a single task the history will be one ($H = 1$) and the target will be set to one also.
 While a history of $H$ suggests that there is sufficient energy to coalesce and
-successfully commit $H$ tasks (i.e., to set $B = H$), HG conservatively sets $B$ equal to half
-of $H$ moderating the risk of a very long coalesced task. 
-The fraction of the history by which HG adjusts $B$ is arbitrary.
-We experimented with a range of fraction values, settling on one half as the
-default for HG, because one half had the highest performance.
-
-The HG algorithm is characterized by the following functions: 
+successfully commit $H$ tasks (i.e., to set $B = H$).
+EG conservatively sets $B$ equal to half of $H$ moderating the risk of a very long coalesced task.
+We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
+The EG algorithm is characterized by the following functions:
 %
@@
 	\begin{split}
-		 f_\text{reboot}(H) 			& = \lfloor(H / 2)\rfloor \\
+		 f_\text{reboot}(H) 			& = \lceil(H / 2)\rceil \\
 		 f_\text{weight}(T_\text{i}) 	& =  1 \\
-		 f_\text{commit}(B) 			& = \lfloor(B / 2)\rfloor
+		 f_\text{commit}(B) 			& = \lceil(B / 2)\rceil
 	\end{split}
@@
 %
-Fig.~\ref{fig:hg-coal} illustrates the operation of the history-guided
-coalescing strategy.
+Fig.~\ref{fig:hg-coal} illustrates the operation of the EG.
 The figure shows two power cycles following the $i$th, the latter having an
@@
 Based on that, $B$ is initially set to two and then decreases to one. Once the
-value of $B$ reaches one, HG is, in effect, expecting a power failure,
-justifying the conservative approach. After the power failure, HG
+value of $B$ reaches one, EG is, in effect, expecting a power failure,
+justifying the conservative approach. After the power failure, EG
 uses the most recent history $H^{(i+1)} = 8$ to set $B = 4$

### 2018-08-04 paper/asplos19/coalescing.tex
revise WEG algo

@@
 %
-\subsubsection{Weighted History-Guided Coalescing (WHG)}
+\subsubsection{Weighted Energy-Guided Coalescing (WEG)}
 \label{subsec:energyTaskAware}
 %
-The Weighted History-Guided (WHG) coalescing strategy accounts for the
+The \emph{Weighted Energy-Guided} (WEG) coalescing strategy accounts for the
 different energy and time cost of a program's tasks when setting the coalescing
@@
 %
-The EO and HG coalescing strategies assume each task has uniform energy cost:
+The EO and EG coalescing strategies assume each task has uniform energy cost:
 for these strategies, $B$ simply corresponds to a target {\em count} of tasks
@@
 
-Instead, WHG associates a non-unity weight with each task in the program. 
+Instead, WEG associates a non-unity weight with each task in the program. 
 %
-WHG tracks the sum of the weights of tasks in a coalesced sequence of tasks.
+WEG tracks the sum of the weights of tasks in a coalesced sequence of tasks.
 %
-When the sum of weights reaches $B$, the target, WHG commits the coalesced
+When the sum of weights reaches $B$, the target, WEG commits the coalesced
 task.
 %
-WHG differs extends EO and HG to explicitly account for the variation in energy
+WEG differs from EO and HG by explicitly accounting for the variation in energy
 cost of a program's tasks, eliminating the assumption that all tasks do the
@@
 %
-WHG has the same characteristic operations as HG, except that
+WEG has the same characteristic operations as HG, except that
 $f_\text{weight}(T_\text{i}) = W_\text{i}$.
 
-The effectiveness of the WHG algorithm hinges on correctly identifying the
-weight of each task, which WHG assumes is {\em statically} available.
+The effectiveness of the WEG algorithm hinges on correctly identifying the
+weight of each task, which WEG assumes is {\em statically} available.
 %
@@
 %
-WHG could use the result of an arbitrarily sophisticated profiling procedure.
+WEG could use the result of an arbitrarily sophisticated profiling procedure.
 %
-To produce a concrete result in this paper, we give WHG access to a simple
+To produce a concrete result in this paper, we give WEG access to a simple
 profile of task run time collected offline using a single fixed input.
 %
-WHG stores the profile in a lookup table that maps a task's identifier to its
+WEG stores the profile in a lookup table that maps a task's identifier to its
 weight, making the information available to \sys's scheduler at run time.

### 2018-08-04 paper/asplos19/relatedwork.tex
add Rachet task splitting differences

--- 2018-04-16 paper/relatedwork.tex
@@
 consistency by placing a checkpoint at the beginning of each idempotent region
-in the code.  Clank~\cite{hicks_isca_2017} ensured consistency with custom
+in the code.
+Ratchet has a similar technique to task splitting with some essential differences: (i) It does not need to handle non-volatile memory accesses, since it assumes a unified non-volatile memory, whereas \sys has to save the volatile pages in non-volatile memory; (ii) The checkpoint is of a fixed size, while \sys privatize variant number of memory pages.
+Clank~\cite{hicks_isca_2017} ensured consistency with custom
 hardware that dynamically tracks read-after-write dependencies in memory

### 2018-08-04 paper/asplos19/abstract.tex

@@
 Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve the execution progress of a program.
-Despite the superiority of the task-based approach, however, it fixes the task size at compile time and operates obliviously to changing energy conditions.
-When tasks are sized without accounting for available energy, the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
+Despite the superiority of the task-based approach, tasks' size is fixed at compile-time, and agnostic to energy conditions.
+% it fixes the task size at compile time and operates obliviously to changing energy conditions.
+% When tasks are sized without accounting for available energy,
+Thus,
+the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
 To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model.
-\sys progresses on a multi-tasks scale when energy permits and preserves the computation progress on a sub-task scale if necessary.
-\sys ensures that power failures cannot leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.
-Our evaluation shows that \sys is $2X$ faster than the state-of-the-art system in most cases and is able to progress where static systems fail.
+\sys progresses on a multi-task scale when energy permits and preserves the computation progress on a sub-task scale if necessary.
+\sys ensures that power failures do not leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.
+Our evaluation on a real energy-harvesting platform shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, and is able to progress where static systems fail.
+% Our evaluation shows that \sys is $2X$ faster than the state-of-the-art system in most cases and is able to progress where static systems fail.

### 2018-08-04 paper/asplos19/abstract.tex
abstract revised

@@
-Freeing embedded devices from batteries is a key factor in developing the energy neutral Internet of Things.
-Energy harvesting, however, is unpredictable and can only power a device intermittently.
+Freeing embedded devices from batteries and powering them from ambient energy is a key factor in developing the energy-neutral Internet of Things.
+Ambient energy is, however, unpredictable and can only power a device intermittently.
 Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve the execution progress of a program.
@@
 % When tasks are sized without accounting for available energy,
-Thus,
-the state may be persisted either more frequently than necessary or not frequently enough for the program to make progress and terminate.
+Thus, the state may be saved either more frequently than necessary or not frequently enough for the program to progress and terminate.
 To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model.

### 2018-08-04 paper/asplos19/hwswsetup.tex

@@
 We used two platforms to evaluate \sys: (i) an \emph{intermittently-powered}
-energy-harvesting device, and (i) a \emph{continuously-powered} evaluation
-board.
+energy-harvesting device (WISP\,5.1~\cite{wisp5,wisp}), and (ii) a \emph{continuously-powered} evaluation
+board (MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}).
 %
 Every benchmark listed in the next section was run repeatedly on
-each platform for a more or less long time window, depending on the benchmark,
-to capture a diverse power trace and produce a high number of complete runs of all applications.
+each platform for a few minutes to ensure capturing a diverse power 
+trace and produce a high number of complete runs of all applications.
 The results were averaged across the duration of the experiment.
 
-Experiments on the intermittently-powered platform used the
-WISP\,5.1~\cite{wisp5,wisp} RF-powered energy-harvesting device.
+% Experiments on the intermittently-powered platform used the
+% WISP\,5.1~\cite{wisp5,wisp} RF-powered energy-harvesting device.
 %
@@
 %
-We powered the WISP using an RF signal generator emitting  a 20-\,dBm sinusoidal wave at 915\,MHz.
+We powered the WISP using an RF signal generator emitting  a 20\,dBm sinusoidal wave at 915\,MHz.
 The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
@@
 %
-We affixed the WISP with a paper harness at the edge of a table, so that the
-WISP was parallel to the table surface at a height of 10\,cm.
+We affixed the WISP with a paper harness at the edge of a table at a height of 10\,cm.
 %
-For distance-controlled experiments we positioned the signal generator
-antenna at a fixed distance from the WISP. The distances were $d=\{15, 30,
-50\}$\,cm for the main experiments and 60\,cm for the task downscaling
+For distance-controlled experiments we positioned WISP at $d=\{15, 30,
+50\}$\,cm from the exciter antenna and 60\,cm was used for the task downscaling
 experiments.
 %
-We chose relatively shorter distances than in a deployment scenario to reduce
-charge time and run experiments faster, after having confirmed ourselves a
-result also reported in prior work~\cite{alpaca} that the amount of energy
-available in one charge cycle---the sum of stored energy and incoming energy
-during execution---is the same at distances beyond 50\,cm where incoming energy
-becomes negligible.
+% We chose relatively shorter distances than in a deployment scenario to reduce
+% charge time and run experiments faster, after having confirmed ourselves a
+% result also reported in prior work~\cite{alpaca} that the amount of energy
+% available in one charge cycle---the sum of stored energy and incoming energy
+% during execution---is the same at distances beyond 50\,cm where incoming energy
+% becomes negligible.
 %
@@
 under profile, and the Saleae~\cite{saleae} logic analyzer measured
-the interval between edges in the signal.
-
-The experiments which are not sensitive to the behavior of an
-energy-harvesting power supply were conducted on an MSP-EXP430FR5969
-launchpad~\cite{MSP-EXP430FR5969_launchpad} board powered continuously with a
-power supply.
-%
-This platform has the same MCU as the WISP.
-%
-Execution time was measured using the clock features in TI Code
-Composer Studio IDE version 7.1.
+the interval between edges in the signal. When the board is continously powered the 
+execution time was measured using the clock features in TI Code Composer Studio IDE version 7.1.
 %

### 2018-08-04 paper/asplos19/systemdescription.tex

@@
 \sys accomplishes this goal with a constellation of a new programming model and runtime software system support, that supports dynamically adaptive task-based execution.
-Fig.~\ref{fig:system_overview} shows an overview of \sys.
-
+Fig.~\ref{fig:system_overview} shows an overview of \sys.\\
+%
 \textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by encapsulating the code in a top-level set of functions, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
@@
 producing a \sys-enabled binary. The runtime library relies on \sys's novel
-{\em virtual memory manager} and its {\em adaptive task scheduler} to adapt its execution with the energy conditions.
-
+{\em virtual memory manager} and its {\em adaptive task scheduler} to adapt its execution with the energy conditions.\\
+%
 \textbf{Adaptive Task Scheduler(ATS).} 
@@
 The scheduler uses its recent execution history---no hardware dependency---as a metric to estimate energy availability and eventually to decide on the dynamic task size.
-
+\\
+%
 \textbf{Virtual Memory Manager (VMM).} \sys is able to efficiently coalesce

### 2018-08-04 paper/asplos19/intro.tex
Intro revised (Carlo)

--- 2018-08-04 paper/asplos19/intro.tex
@@
 buffers energy in a small storage capacitor~\cite{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
-threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., nW to $\mu$W) compared to a platform's operating
-power level (hundreds of $\mu$W to mW). A device operates briefly until it depletes its buffered energy, after which, the device shuts
-down and recharges to operate again later---corresponds to {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of
+threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., \nano\watt\ to \micro\watt) compared to a platform's operating
+power level (hundreds of \micro\watt\ to \milli\watt). A device operates briefly until it depletes its buffered energy, after which, it shuts
+down and recharges to operate again later---corresponding to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of
 operation-power failure-restart cycles. 
-The recharge and discharge time---which corresponds to the device's inactive and active time---varies with the energy buffering capacitor~\cite{capybara} and some devices fail and restart operating $\approx$10 to
+The recharge and discharge intervals---which correspond to the device's inactive and active time---vary with the energy buffering capacitor~\cite{capybara} \note{not only, though} and some devices fail and restart operating $\approx$10 to
 $\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.
@@
 scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution
-progress, failures can leave non-volatile state incorrectly, partially updated,
+progress, failures can leave non-volatile state incorrect, partially updated,
 and may leave checkpointed volatile state inconsistent with non-volatile state.
@@
 approaches are simpler and applicable to existing devices today.
-Therefore, this work focuses on a key limitation of task-based software approaches, namely, the {\em inflexibility} of statically decomposing a program into tasks.
+Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.
+
 %
@@
     \includegraphics[width=0.7\columnwidth]{figures/intro-figure-vert.pdf}
-    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C} in the figure).
+    \caption{\note{All this white padding is ugly} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C} in the figure).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
@@
 
-\noindent\textbf{Drawback of Static Task Decomposition.} 
+\noindent\textbf{Crucial Drawback of Static Task Decomposition.} 
+\note{Why no indent?}
 Task-based programming and execution models require a
@@
 Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and
-shows how tasks transitions can impose a run-time overhead.
+shows how task transitions can impose a run-time overhead.
 At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
@@
 \noindent\textbf{Challenges and Contributions.}
+\note{Why no indent?}
 This paper introduces a new system called \emph{\sys}, which executes a task-based
 intermittent program by \emph{dynamically} coalescing/splitting its tasks.
-\sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce} (top)) or splits them (Fig.~\ref{fig:coalesce} (bottom)) based on the estimated energy availability without demanding any hardware support.
-The unique challenges and contributions revealed by this work are listed as follows (C1--C3):
+\sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce}, top) or splits them (Fig.~\ref{fig:coalesce},bottom) based on the estimated energy availability without demanding any hardware support.
+The unique challenges and contributions revealed by this work are listed below:
+\note{it takes more space, but a bullet list makes C1-C3 stand out more properly}
 
@@
 %\textbf{Challenges and Contributions} 
-
-\noindent\emph{C1) Dynamic Task Sizing:} Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
+\noindent\note{Carlo's rephrasing} \emph{C1) Task Transition Overhead:} given unpredictable incoming energy, how to save progress as few times as possible?
+\sys tries to minimize task transition overhead by estimating energy conditions at run-time using \emph{recent execution history} as a metric.
+% \noindent\emph{C1) Dynamic Task Sizing:} Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
 \sys addresses this challenge by using a novel approach: \textit{recent execution history} as a metric to estimate the environment energy conditions and to specify task sizes.
@@
 
-\noindent\emph{C2) Dynamic Memory Consistency Handling:} Specifying task sizes on the fly gives rise to the need for dynamic memory consistency handling: how to detect dynamic variables dependencies and ensure data protection efficiently against power interrupts?
-\sys ensures the consistency of the data by using a novel
+\noindent\emph{C2) Dynamic Memory Consistency Handling:} merging static tasks on the fly gives rise to the need for dynamic memory consistency handling.
+How to dynamically detect inter-task data dependencies and ensure efficient protection against power interrupts?
+\sys does so by using a novel
 approach called \emph{dynamic group privatization}--- %Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e.
 protecting all global variables) in each dynamic task transition.
-utilizing a real-time data tracking to enable partial data protection on a task transition.
-Individual variables tacking, however, slows down a system dramatically.
-Therefore, \sys  relies on a coarse locality principle and protects the global variables in a batch.
-Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
+performing real-time dependency tracking to enable protection on a task transition.
+Individual variables tracking, however, slows down a system dramatically.
+Therefore, \sys protects global variables in a batch.
+Moreover, it optimizes data transfers by using Direct Memory Access (DMA).
 
-\noindent\emph{C3) Ensuring Task Termination:} A static task decomposition model assumes that each
+\noindent\emph{C3) Ensuring Task Termination:} a static task decomposition model assumes that each
 single task can execute to completion. If the hardware energy buffer provides
 inadequate energy to execute each single task to completion, a program will not
-terminate~\cite{cleancut_2018}. Under such conditions how to enable the dynamic execution model to progress on a sub-task level? To avoid non-termination under adversarial
-energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism.
+terminate~\cite{cleancut_2018}. Under such conditions, how to enable the dynamic execution model to progress on a sub-task level? To avoid non-termination under adverse
+energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism \note{should we instead cite Ratchet?}.
 Partial execution avoids non-termination by committing the intermediate state of a long-running
@@
 %
+
+To asses the benefits of \sys over existing task-based systems, we implemented and tested six
+benchmarks on a real energy-harvesting platform.
+Our evaluation shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, and is able to progress where static systems fail.
+
 The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
 Section~\ref{sec:systemdescription} provides an overview of \sys, while
-Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
-Dynamic group privatization mechanism is provided in Section~\ref{sec:memory_virtulaization}.
+Section~\ref{sec:task_adaptation} describes \sys's energy-adaptive mechanism.
+The dynamic group privatization is explained in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are given in Section~\ref{sec:implementation}.
-Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's methodology and evaluation, respectively.
-Section~\ref{sec:related_work} puts \sys in the
+Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results.
+Section~\ref{sec:related_work} positions \sys in the
 context of related work and Section~\ref{sec:conclusions} concludes and discusses

### 2018-08-04 paper/asplos19/intro.tex
changes to the intro (not ready yet)

--- 2018-08-04 paper/asplos19/intro.tex
@@
     \includegraphics[width=0.7\columnwidth]{figures/intro-figure-vert.pdf}
-    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C} in the figure).
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{T(x)}).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
@@
 Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and
-shows how tasks transitions can impose a run-time overhead.
+shows how tasks transitions can impose a runtime overhead.
 At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
-The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.
+The more task transitions a program requires, the more commit overhead is incurred by the system at runtime.
 %
@@
 existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress.
-These constraints on task sizing leave an unsatisfying dilemma: large, efficient tasks that risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
+These constraints on task sizing lead to the following dilemma: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
 In this work,
 we propose a third way, using a novel technique called \emph{adaptive task size execution by task coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.
-%
 
-\noindent\textbf{Challenges and Contributions.}
+\noindent\textbf{Our Contributions.}
 This paper introduces a new system called \emph{\sys}, which executes a task-based
@@
 \sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce} (top)) or splits them (Fig.~\ref{fig:coalesce} (bottom)) based on the estimated energy availability without demanding any hardware support.
-The unique challenges and contributions revealed by this work are listed as follows (C1--C3):
+The unique challenges and contributions revealed by this work are listed as follows:
 

### 2018-08-04 paper/asplos19/hwswsetup.tex
section 7 revised

@@
 
-\subsection{Intermittently-powered Platform and Experimental Setup}
+\subsection{Experimental Setup}
 \label{sec:results_hardware_software}
@@
 %
-Every benchmark listed in the next section was run repeatedly on
+Every benchmark listed in Section~\ref{sec:evaluation} was run repeatedly on
 each platform for a few minutes to ensure capturing a diverse power 
@@
 
-% Experiments on the intermittently-powered platform used the
-% WISP\,5.1~\cite{wisp5,wisp} RF-powered energy-harvesting device.
-%
 The WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
@@
 %
-We affixed the WISP with a paper harness at the edge of a table at a height of 10\,cm.
+We affixed the WISP with a paper harness at the edge of a table at a height, from the table suface, of 10\,cm.
 %
@@
 %
-
 \subsection{Software Benchmarks}
@@
 
-We evaluated \sys using six benchmarks summarized in
-Table~\ref{table:benchmark_table}. %
-Three benchmarks (\textbf{ar}, \textbf{bc}, \textbf{cuckoo}) are from prior
-work~\cite{chain,alpaca} and three benchmarks (\textbf{dijkstra},
-\textbf{fft}, \textbf{sort}) are new implementations of applications
-from the MiBench suite~\cite{mibench,hicks_mibench2_2016}. The
-\textbf{fft} implementation is based on the Accelerated FFT library from
-TI~\cite{ti_dsp}. All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0
-from toolchain version 5.01.01.00 with \texttt{-O1} as optimization flag.
-%
+We evaluated \sys using six benchmarks that are oftenly used in embedded systems (summarized in
+Table~\ref{table:benchmark_table}).
+All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 toolchain with \texttt{-O1} as optimization flag.
 The source code for all benchmarks will be released at \cite{coala_website}.
 
-%Specifically, we have implemented three benchmarks used by Chain~\cite{chain} and Alpaca~\cite{alpaca} (source code is accessible~\cite{coala_website}), namely: \textbf{ar}: nearest neighbor classification of randomly generated data modeling a three-axis accelerometer; \textbf{bc}: counts bits in byte pseudo-random string using several algorithms, and compares results for correctness; \textbf{cuckoo}: Bloom-filter-like Cuckoo filter that first hashes a sequence of pseudo-random numbers, then queries the filter to recover the sequence.
-Additionally, we have ported three new benchmarks from~\cite{mibench}, namely: \textbf{dijkstra}: Dijkstra algorithm executed continously on a predefined adjecency matrix; \textbf{fft}: continuously repeating Fast Fourier Transform on three pre-generated 128 sample vector; and \textbf{sort}: selection sort algorithm on a fixed sequence of numbers.
-\todo{justify the selection of these benchmarks}
-
 \textbf{Comparison with Alpaca using the GCC compiler.} We compare
-\sys against Alpaca~\cite{alpaca}, a state-of-the-art task-based system for
-intermittent computing. For a fair comparison the task decomposition is the
-same in \sys and Alpaca implementations of the benchmarks. We compare
-both systems using GCC.
+\sys against Alpaca~\cite{alpaca}, the state-of-the-art task-based system for
+intermittent computing. For a fair comparison the task decomposition of the benchmarks are ensured to be the
+same for both systems.
 
@@
 Instead, we \emph{manually} performed the instrumentation done by the compiler
-pass.  We performed the instrumentation at the source level, adding C
-statements to the program, instead of at the level of binary
-intermediate-representation (LLVM IR) as the compiler pass would. The result is
-equivalent from the perspective of overheads. The instrumentation consists of
-identifying write-after-read dependencies and adding code and memory
-allocations to make a private copy of the affected variables. For array
-variables, instrumentation also includes adding code for 
-dynamically tracking the array access.
-
-Manual instrumentation of Alpaca code is not a general approach, but only
-served as a practical alternative to porting the compiler pass from LLVM to
-GCC.
+pass. The instrumentation consists of identifying write-after-read dependencies 
+and adding code and memory allocations to make a private copy of the affected variables. 
+For array variables, instrumentation also includes adding code for 
+dynamically tracking the array access. Manual instrumentation of Alpaca code is not a general approach, 
+but only served as a practical alternative to porting the compiler pass from LLVM to GCC.
 %
@@
 By compiling both systems with the same compiler (GCC), we ensure that our
-comparison in Section~\ref{sec:evaluation} is fair.\vspace{2cm}
+comparison in Section~\ref{sec:evaluation} is fair.
 

### 2018-08-04 paper/asplos19/intro.tex
changes to the intro (Przemek) conflict from Carlo resolved

--- 2018-08-04 paper/asplos19/intro.tex
@@
     \includegraphics[width=0.7\columnwidth]{figures/intro-figure-vert.pdf}
-    \caption{\note{All this white padding is ugly} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C} in the figure).
+    \caption{\note{All this white padding is ugly} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{T(x)}).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
@@
 The programmer (or a compiler) explicitly expresses task-to-task control flow.
-Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and
-shows how task transitions can impose a run-time overhead.
+Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and shows how tasks transitions can impose a runtime overhead.
 At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
-The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.
+The more task transitions a program requires, the more commit overhead is incurred by the system at runtime.
 %
@@
 existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress.
-These constraints on task sizing leave an unsatisfying dilemma: large, efficient tasks that risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
+These constraints on task sizing lead to the following dilemma: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
 In this work,
 we propose a third way, using a novel technique called \emph{adaptive task size execution by task coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.
-%
 

### 2018-08-04 paper/asplos19/background.tex
Background revised (Carlo)

@@
 Energy harvesting devices operate using energy extracted from sources such as
-radio frequency transmissions and solar energy. These devices elide tethered
-power or a battery, instead collecting energy into a capacitor, operating when
-sufficient energy accumulates, and upon depleting the energy, turning off and
-recharging.
+radio frequency transmissions and light. These devices elide tethered
+power and batteries, instead collecting energy into a capacitor, operating when
+sufficient charge accumulates, and turning off for recharge upon depletion of the buffer.
 There are several energy harvesting battery-less hardware platforms. For
@@
     \end{subfigure}
-    \caption{WAR dependence example. \texttt{NV} marks non-volatile variables,
+    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables,
     while \texttt{checkp()} is a checkpoint of volatile state.}
@@
 Software running on an energy-harvesting device executes {\em intermittently}
-because energy sources are not always available to harvest and buffer
+because power sources are not always available to harvest and buffer
 sufficient operational energy. An intermittent execution is composed of
 operating periods interspersed with power
-failures~\cite{dino,chain,alpaca,ratchet}. The frequency of failures depends on
-the size of the device's energy storage buffer: a larger buffer allows longer
-operating periods.
+failures~\cite{dino,chain,alpaca,ratchet}.
+The frequency of failures depends on the size of the device's energy storage buffer
+(a larger buffer allows longer operating periods), current consumption and
+incoming energy.
 
@@
 Precisely, data can get inconsistent when code includes a
-\emph{write-after-read} (WAR) dependence between operations that manipulate
-non-volatile memory, because it allows {\em writes} from a failed attempt to
-persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
+\emph{write-after-read} (WAR) dependency between operations that manipulate
+non-volatile memory.
+% because it allows {\em writes} from a failed attempt to
+% persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
 Fig.~\ref{fig:war} illustrates how state can become
@@
 Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to
-decompose their program into \textbf{tasks}, which is a region of code that can contain
+decompose their program into \textbf{tasks}, which are regions of code that can contain
 arbitrary computation, sensing, and communication.
@@
 They re-execute each task interrupted by a power failure until it successfully finishes, only then moving on to
-the next task. Since these models do not rely on taking an expensive
+the next task. Since these models do not rely on capturing an expensive
 checkpoint, they are usually faster than the checkpoint-based
@@
 
-Task-based models also suffer from data inconsistency when there is a WAR
-dependence.  Prior systems tried to tackle the problem by a compiler-automated
-redo-logging for the variables that is part of the dependence~\cite{alpaca}, or
+Task-based models also suffer from data inconsistencies when WAR dependencies are involved.
+Prior systems tried to tackle this problem by a compiler-automated
+redo-logging for the variables that are part of the dependency~\cite{alpaca}, or
 by statically creating multiple copies of the problematic variable to ensure
@@
 
-Prior systems, including the task-based models, suffer from two major overheads:
-{\em frequent non-volatile memory access} and {\em incorrectly sized, inefficient or non-terminating tasks}.
+Prior systems, including task-based models, have two major drawbacks:
+{\em frequent non-volatile memory accesses} and {\em incorrectly sized tasks}
+(too small or too large).
 
@@
 \textbf{Frequent Non-volatile Memory Access.} 
+\note{I want to emphasize again that we do not evaluate NV memory accesses}
 Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.  Checkpointing systems copy all volatile
@@
 consistency and progress preservation.  Ratchet~\cite{ratchet} and
-Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile
+Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile,
 perform high-frequency non-volatile memory accesses, into checkpointing
-frequently.
+frequently \note{is "into" a misspell?}.
 
@@
 maximum energy that the device can buffer, then the task
-will never be able to complete using buffered energy.  Such a task is
+will never be able to complete using buffered energy only.  Such a task is
 non-terminating, {\em prevents forward progress}, and makes the program
-incorrect. 
+deadlock. 
 %
@@
 tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.  
-Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}) are considered the problem of porting a program become much worse
+Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \micro\farad~\cite{rodriguez_tbcs_2015} to 0.1 \farad~\cite{moo}) are considered, the problem of porting a program becomes much worse
 because a large task on one device may be a relatively short task
@@
 merges multiple tasks to amortize the overhead when tasks are too small, and
-{\em task downscaling}, which breaks up a task when the task is too large to complete.
+{\em task downscaling}, which breaks up a task when it is too large to complete.

### 2018-08-04 paper/asplos19/intro.tex
Background revised (Carlo)

@@
     \includegraphics[width=0.7\columnwidth]{figures/intro-figure-vert.pdf}
-    \caption{\note{All this white padding is ugly} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{T(x)}).
+    \caption{\note{All this white padding is ugly} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}

### 2018-08-04 paper/asplos19/implementation.tex
implementation: partial commit

@@
 
-Coala's API adds only five syntactic constructs to a C-based language, summarized
+Coala's API adds only eight syntactic constructs to a C-based language, summarized
 in Table~\ref{table:coala_api}.
@@
 declaration statically allocates a non-volatile constant variable holding a
-task's weight and declares that the function is a task.
-
-\noindent\textbf{Task Transitions.} A call to \texttt{COALA\_NEXT\_TASK} immediately
-transfers control to a specified next task and can be invoked along
-any control path to dynamically determine the next task.
-Like prior work~\cite{chain,dino,ratchet}, \sys's control-flow model
-assumes that an executing task is not interruptible by asynchronous events.
-
+task's weight and declares that the function is a task.\\
+%
+\noindent\textbf{Task Transitions.} A call to \texttt{COALA\_NEXT\_TASK} marks the task to be executed at the 
+end of the current one task and it can be invoked along
+any control path to dynamically determine the next task.\\
+%
 \noindent\textbf{Protected Variables.} The \texttt{COALA\_PV} annotation on a
-variable statically allocates and aligns a protected non-volatile variable. The
+variable statically allocates a protected non-volatile variable. The
 variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
-runtime to ensure correct operation.
-
-\noindent\textbf{Protected Reads and Writes.} The programmer uses the \texttt{RP} and \texttt{WP} API
-methods to access protected variables.  These methods cause \sys to find a variable's page.
-If the page is not resident in working memory, the method leads to a page fault,
-which fetches the page from non-volatile memory. Unlike \texttt{RP},
-\texttt{WP} marks the variable's page as dirty, allowing for efficient 
-commit back to non-volatile memory.
-
+runtime to ensure correct operation.\\
+%
 \noindent\textbf{Initialization.} The behavior of the API method
 \texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the
-addition of performing preliminary kernel initializations, including hardware setup.
-
-%\begin{wrapfigure}{t!}{0.5\textwidth}
-%	\centering
-%	%\includegraphics[width=0.5\columnwidth]{figures/graffle/state-machine.pdf}
-%	\caption{\sys state machine implementation \todo{Draw a figure}{Sinan, Carlo}}
-%	\label{fig:coala_state_machine}
-%\end{wrapfigure}
-
-%A high-level state machine of the execution manager is depicted in Figure~\ref{fig:coala_state_machine} \todo{Draw a figure reflecting the following description}{Sinan, Carlo}.
-
+addition of performing preliminary kernel initializations, including hardware setup.\\
+%
 \noindent\textbf{Execution.} The programmer passes control to \sys's task scheduler 
@@
 coalescing tasks.
-%\todo{Go over keywords (coalescing target, current budget, commit phases etc.) and make them consistent with the rest of the paper}{Amjad, Carlo}
 
-\subsection{Paging}
+\subsection{\textcolor{red}{to be revised after Virtual Mem.}Paging}
 \label{sec:impl:paging}
@@
 of 1\,KB, which is half of the total SRAM size in the MSP430FR6989 microcontroller.
-%\begin{figure}
-%	\centering
-%	\includegraphics[width=\textwidth]{figures/graffle/paging.pdf}
-%	\caption{Illustration of \sys page tags mechanism.}
-%	\label{figure:coala_page_tags}
-%\end{figure}
-
-% It is important to mention that store and shadow buffers are virtual.
-By default, \sys allocates two memory banks for pages, $B0$ and $B1$, and all the pages have their store copy in $B0$.
-A dirty page that during the intermediate commit, or due to eviction, is copied to its shadow location in $B1$, will end up having its store location in $B1$ after the ultimate commit.
-Thus, during the ultimate commit
-
+%
 \subsection{Coalescing and Commit}
-
+%
 \sys uses a two-phase commit process to atomically commit pages despite
@@
 %
-
 \subsection{Task Downscaling}
-
+%
 \sys's task downscaling preserves forward progress by partially committing an
@@
 These annotations respectively halt and resume the partial execution timer.
-
-%\begin{wrapfigure}{t!}{0.5\textwidth}
-%	\centering
-%	%\includegraphics[width=0.5\columnwidth]{figures/graffle/commits-illustration.pdf}
-%	\caption{Intermediate commit and ultimate commit Illustration \todo{to be drawn}{sinan}}
-%	\label{fig:intermediate_ultimate-commit}
-%\end{wrapfigure}

### 2018-08-04 paper/asplos19/hwswsetup.tex
minor modifications

@@
 		\hline
-        \textbf{bc} &10 &371 & Counts set bits in a pseudo-random sequence of bytes
-using several algorithms and compares results for correctness\\
+        \textbf{bc} &10 &371 & Counts bits in bytes sequences
+using several algorithms and check for correctness\\
 		\hline
         \textbf{cuckoo} &14 &426 & Inserts and queries a sequence of pseudo-random
-values in a Cuckoo filter, a set membership data structure with deletion\\
+values in a Cuckoo filter.\\
 		\hline

### 2018-08-04 paper/asplos19/relatedwork.tex
minor modifications

@@
 for such systems
-~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,wisent,flicker}.
-
+~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,wisent,flicker}. \\
+%
 \textbf{Checkpointing-based Systems.} Early work on energy-harvesting runtime
@@
 Support for computation that spans power failures was first achieved with
-statically-placed conditional checkpoints in Mementos~\cite{mementos}.
-
+statically-placed conditional checkpoints in Mementos~\cite{mementos}. \\
+%
 Mementos left the possibility of non-volatile state becoming inconsistent with
@@
 Unlike \sys, such systems rely on introspection hardware to monitor supply
-voltage and on accurate worst-case bounds on checkpoint cost.
-
+voltage and on accurate worst-case bounds on checkpoint cost. \\
+%
 In all of the above systems, with the exception of DINO~\cite{dino},
@@
 Section~\ref{sec:cost_task-based}) and is not a viable option for off-chip non-volatile
-memory. In addition, unlike \sys, QuickRecall and Clank require custom hardware.
-
+memory. In addition, unlike \sys, QuickRecall and Clank require custom hardware.  \\
+%
 \textbf{Task-based Systems.} An alternative to checkpointing are recent systems
@@
 prior systems. \sys keeps memory consistent through memory virtualization
-optimized for bulk accesses to task-shared data with high locality.
-
+optimized for bulk accesses to task-shared data with high locality. \\
+%
 \textbf{Task Decomposition.} In contrast to \sys's dynamic construction of
@@
 analysis to place a minimal number of conditional checkpoints that test the
-energy level at runtime before copying state, like Mementos~\cite{mementos}.
-
+energy level at runtime before copying state, like Mementos~\cite{mementos}. \\
+%
 Unlike \sys, such compiler-based approaches, face the challenge of statically
@@
 at compile time and adapts to energy storage capacity and incoming energy
-conditions at runtime.
-
+conditions at runtime. \\
+%
 \textbf{Memory Virtualization.} Prior work on embedded systems has studied a
@@
 focused more on programmability and runtime reliability properties provided by
-virtual memory.
-
+virtual memory. \\
+%
 A related domain is unbounded, page-based transactional memory and

### 2018-08-04 paper/asplos19/hwswsetup.tex
minor modifications

@@
 All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 toolchain with \texttt{-O1} as optimization flag.
-The source code for all benchmarks will be released at \cite{coala_website}.
-
+The source code for all benchmarks will be released at \cite{coala_website}.\\
+%
 \textbf{Comparison with Alpaca using the GCC compiler.} We compare
@@
 intermittent computing. For a fair comparison the task decomposition of the benchmarks are ensured to be the
-same for both systems.
-
+same for both systems.\\
+%
 Since Alpaca's compiler pass is implemented only for LLVM, we could not use
@@
 comparison in Section~\ref{sec:evaluation} is fair.
-
-% Unnecessary
-%\textbf{Comparison of \sys against other Systems.} We do not compare \sys
-%against Chain~\cite{chain}, another task-based system, which is inferior to
-%Alpaca in its performance. Also, we do not compare \sys to
-%checkpointing-based systems, such as DINO~\cite{dino} or
-%Mementos~\cite{mementos} (cf. Section~\ref{sec:background_consistency}).
-%\todo{Provide reasons why we do not compare against checkpointing}{Brandon}.

### 2018-08-05 paper/asplos19/intro.tex
comment on the intro fig

@@
     \includegraphics[width=0.7\columnwidth]{figures/intro-figure-vert.pdf}
-    \caption{\note{All this white padding is ugly} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
+    \caption{\note{small static task subplot and large static task subplot can be re-arranged to fit next to each other, so we end up with two layers plot} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}

### 2018-08-05 paper/asplos19/paging.tex
comment on the intro fig

@@
 And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
-
+%
 \sys's {\em virtual memory manager} exposes non-volatile memory for
 \emph{protected} variables to the application as a virtual address space
-partitioned into pages.
+partitioned into privatized pages, a technique we call \emph{dynamic pages privatization}.
 %
@@
 %
-The virtual memory manager automatically finds a variable in its page and
-allows a task to directly manipulate its value.
+The virtual memory manager automatically privatize the pages associated with the accessed protected variables.
 %
@@
 %
-If a power failure interrupts a task, the virtual memory manager ensures that
+If the power interrupts a task, the virtual memory manager ensures that
 the task sees consistent memory state when it restarts its execution.
@@
 The virtual memory manager ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
-%
-%
-%
-%,. Virtualization simplifies
-%coalescing, because the system gets the freedom to move application's data
-%between volatile and non-volatile memory transparently.
-%
 
 \sys's virtual memory manager \sys abstracts non-volatile storage (e.g.,
-variables in FRAM) from tasks, which simply manipulate protected variables.
+variables in FRAM) from tasks.
 %
@@
 showed in Section~\ref{sec:background}).
-%
 %
@@
 %
-%manages memory such that application tasks operate primarily from
-%efficient volatile SRAM rather than non-volatile FRAM, which is more costly to access than SRAM,
-%as we showed in Figure~\ref{fig:framEnergy}.
 \sys moves pages of data efficiently using hardware accelerated Direct Memory
 Access (DMA) support, which copies a block of memory from one place to another
-as a single operation.
-
+as a single operation.\\
+%
 \begin{figure}
@@
 \end{figure}
-
-% \begin{figure}
-	% \centering
-	% \includegraphics[width=\columnwidth]{figures/paging.pdf}
-	% \caption{Illustration of task-virtual memory interaction in \sys.
-Tasks only interact with the working buffer in SRAM (indicated as page buffer).
-The shadow buffer is used to hold the dirty bages.
-The indirection table is used to speed-up two-phase commit operation: it eliminates the copy operation in the second phase.}
-	% \label{figure:coala_page_tags}
-% \end{figure}
-
+%
 \textbf{System Overview.}
@@
 %
-%BML: this point is true, but without caches, it is not a performance factor, so I removed it.
-%Unlike prior work~\cite{chain,alpaca} that moves scattered variables one at a
-%time between volatile and non-volatile memory, \sys's page-based design moves
-%contiguous blocks of data.
+\subsection{Address Translation and Variable Access}
 %
-%BML: this seems gratuitous because DMA is very common and we've already introduced it.
-%\sys takes full advantage of hardware\hyp{}accelerated block copies using
-%Direct Memory Access (DMA) hardware in common microcontrollers --- e.g.
-%5$\times$ fewer cycles on MSP430FRxxxx.
-
-\subsection{Address Translation and Variable Access}
-
 A task must access protected non-volatile variables through \sys's
@@
 that identifies the page and a \emph{page offset} that identifies a byte.
-
-%How an $a$-bit memory address decomposes into a tag and an offset depends on the number (and size) of each page of protected memory.
-To address {\tt p} pages of size {\tt s}, a tag comprises $log(p)$ bits, the offset comprises $log(s)$ bits and $log(p) + log(s) = a$.
-For example, in a 16-bit address space (i.e., 16\,kB of protected memory) with 512 pages of 128 bytes each, a page's tag is the high-order nine bits of the address and a byte's page offset is the remaining 7 bits.
-
-%\begin{algorithm}[t]
-\begin{figure} %{t}{0.4\textwidth}
+%
+\begin{figure}
 \fbox{\parbox{\linewidth}{
@@
 	\scriptsize
-	%\small
 	\begin{algorithmic}[1]
@@
 		\State \Return \texttt{working}[$i$][$o$]  \Comment{Return from page}
-%	\end{algorithmic}
-%\end{algorithm}
 	\end{algorithmic}}}
 \end{figure}
-
+%
 After address translation (Figure~\ref{figure:mem-man}, Step 6), a task
@@
 %
-%If the accessed variable's page is resident in the page buffer,
-%then the tag of one of the pages in this buffer is equal to the accessed
-%variable's page tag.
-%
 If the accessed variable's page tag is not found in the page buffer, the
 operation incurs a {\em page fault} (Line 4).
-% At a page fault, \sys must swap
-% out one of the pages that is resident in the page buffer (to the non-volatile
-% \texttt{\underline{shadow}}) if \texttt{working} is full, and swap in the
-% variable's page.
 %
@@
 the variable's page offset (Lines 5--6).
-
-%Algorithm~\ref{algo:rwar} shows pseudo-code of \texttt{RVAR} and we omit a listing for {\tt WVAR} as it is very similar except it writes the memory location and returns nothing.
-If the page tag of the variable \texttt{var} is not found in \texttt{working} (e.g.
-Lines 2--3), there is a page fault that requires a new page to be swapped in to the page buffer.
-
+%
 \subsection{Page Faults and Page Swapping}
-
+%
 When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory
@@
 \emph{victim page}) preserving updates made to that page.
-% and to copy the accessed page of protected data into the working buffer.
-% Algorithm~\ref{algo:pagefault} shows how \sys handles a page fault. A victim
-% page in the \texttt{working} buffer is selected for eviction according to
-% a \emph{first-in-first-out} (FIFO) replacement policy (Lines 2 and 12).
 %
 While we opted for the simplicity of FIFO, any other replacement policy, such
-as Least Recently Used (LRU) could work in its place.
-
-% \begin{algorithm}
-% \begin{wrapfigure}{t}{0.5\textwidth}
-% \fbox{\parbox{}{
-% 	\captionof{algorithm}{\texttt{PageFault}(tag $t$)}
-% 	\label{algo:pagefault}
-% 	\scriptsize
-% \begin{algorithm}
-	% \caption{\texttt{PageFault}(tag $t$)}
-	% \label{algo:pagefault}
-	% \scriptsize
-	% %\small
-	% \begin{algorithmic}[1]
-        % \State \LeftComment{$i_\text{victim}$: index to free slot or page to be evicted, initially 0}
-        % \State \LeftComment{\texttt{\underline{shadowCount}}: non-volatile, 0 on boot unless in phase 2 commit}
-        % \State $t_\text{victim} \gets \Call{Tag}{\texttt{working}[i_\text{victim}]}$
-		% \If {\Call{isDirty}{working[$i_\text{victim}$]} }	\Comment{Evicting a dirty page?}
-		% \State $\texttt{\underline{shadow}}[t_\text{victim}] \xleftarrow{\text{DMA}} \texttt{working}[i_\text{victim}]$
-            % \Comment{Swap out the page}
-        % \If {$t_\text{victim} \not\in \texttt{\underline{shadowList}}$} \Comment{Keep track for quick traversal}
-            % \State $\texttt{\underline{shadowList}}[\texttt{shadowCount++}] \gets t_\text{victim}$
-        % \EndIf
-		% \EndIf
-		% \If {\Call{exists}{\underline{shadow}[$t$]}} \Comment{Page swapped out earlier?}
-		% \State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{shadow}}[t]$
-            % \Comment{Get from commit buffer}
-		% \Else
-		% \State $\texttt{working}[i_\text{victim}] \xleftarrow{\text{DMA}} \texttt{\underline{store}}[t]$
-            % \Comment{Get from backing store}
-		% \EndIf
-        % \State $i_\text{victim} \gets (i_\text{victim} + 1) \mod |\texttt{pageBufLength}|$ \Comment{FIFO evict}
-	% \end{algorithmic}
-% \end{algorithm}
-
-% 	\end{algorithmic}
-% \end{algorithm}
-	% \end{algorithmic}}}
-% \end{wrapfigure}
-
+as Least Recently Used (LRU) could work in its place. \\
+%
 If a task modified any byte in the victim page (i.e., using \texttt{WP}), then
@@
 {\tt \underline{store}} (Step 4).
-Finally, the page is copied to the volatile buffer (Step 5).
+Finally, the page is copied to the volatile buffer (Step 5). \\
 %
-%Dirty pages remain in {\tt \underline{shadow}} buffer until the
-%dynamic task completes, at which point \sys commits the modified
-%pages from {\tt \underline{shadow}} back to \texttt{\underline{store}}.
+\subsection{Atomic Two-Phase Commit of Dirty Pages}
 %
-
-%As
-%Algorithm~\ref{algo:pagefault} shows in Line 4, \sys checks if the requested
-%page was previously modified by the executing task. To track whether a page is
-%modified, \sys stores a dirty bit with the page and sets it when the page is
-%written to. If an accessed page's dirty bit is set, the page must be swapped in
-%from \texttt{\underline{shadow}} rather than its original memory location (Line 5),
-%furnishing the task with the page's modified values. If the accessed page is
-%not dirty, \sys swaps it in from its original location in memory (Line 7).
-
-%
-%\sys
-%uses DMA block copies for all page swapping operations, leveraging the
-%efficiency of hardware support.
-
-\subsection{Atomic Two-Phase Commit of Dirty Pages}
-
 When the last task in a sequence of coalesced tasks completes, \sys must commit
@@
 \texttt{\underline{store}}, the task might see protected variables in an
-inconsistent state; \sys's two-phase commit prevents this inconsistency.
-
+inconsistent state; \sys's two-phase commit prevents this inconsistency. \\
+%
 To make the commit atomic, \sys commits the pages in two phases, implemented by
@@
 buffer.  Section~\ref{sec:implementation} describes double buffering in more
-detail.
-
-%\subsection{Paging Example}
+detail. \\
 %
-%Figure~\ref{fig:volatile-buffer} illustrates the execution of a simple task using \sys's paging mechanism.
-The task manipulates the protected variables {\tt x}, {\tt y}, and {\tt z}.
-The variables {\tt y} and {\tt z} are in the same page and they are currently in the volatile page buffer---on accessing {\tt y} and {\tt z}, the task needs only to access \texttt{pageBuf}.
-Since variable \texttt{x} is kept in a different page in \texttt{pagesOrg}, accessing it creates a page fault.
-At this point, \sys swaps the active page in \texttt{working} by committing it into the non-volatile {\tt \underline{shadow}} region and copying the corresponding page from \texttt{pagesOrg} to \texttt{pageBuf}.
-When the task completes, \sys sets the commit bit, copies the dirty page from \texttt{pagesTmp} to its original location in \texttt{pagesOrg} and unsets the commit bit.
-
-%\begin{figure}[t]
-%	\centering
-%	\includegraphics[width=0.75\columnwidth]{figures/sram-buffer}
-%	\caption{The interaction between the task, volatile buffer and the non-volatile memory (FRAM).
-Initially, the persistent variable \texttt{x} is not in the volatile buffer but \texttt{y} and \texttt{z} are.}
-%	\label{fig:volatile-buffer}
-%\end{figure}
-
-%\subsection{Memory Consistency}
-
 \textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only
@@
 correctly access their original memory locations in \texttt{\underline{store}}.
-\sys assumes that the write to a non-volatile word in memory is atomic.
+\sys assumes that the write to a non-volatile word in memory is atomic. \\
 %
-%The non-volatile commit bit persists
-%across power failures and \sys repeatedly attempts to commit until it succeeds,
-%despite arbitrary power failures.
-
-%\begin{algorithm}[t]
-% \begin{wrapfigure}{t}{0.5\textwidth}
-% \fbox{\parbox{\linewidth}{
-% 	\captionof{algorithm}{Two-phase commit}
-% 	%\caption{Two-phase commit}
-% 	\label{algo:commit}
-% 	\scriptsize
-
 \begin{algorithm}[t]
@@
 \end{algorithm}
-% 	\end{algorithmic}}}
-% \end{wrapfigure}
-
+%
 \subsection{Dynamic Paging of \sys}
-
+%
 \sys asks the programmer to use its {\tt RP} and {\tt WP} paging API methods on
@@
 An alternative approach~\cite{alpaca} relied on a compiler to analyze memory operations and try to insert memory protection code mostly at the beginning of a task, leaving most
-memory accesses uninstrumented.
-
+memory accesses uninstrumented.\\
+%
 Despite the risk of per-access overhead, \sys's dynamic memory protection
@@
 coalescing.
-%
-%Moreover, if the programmer wants to optimize for high performance, it is possible for the
-%programmer to place {\tt RP} and {\tt WP} only minimally by carefully examining the code.
-%We argue that the
-
-%\TODO{Put paragraph about Alpaca's static privatization algo vs. Coala's dynamic privatization algo}
-%\TODO{pass by ref might not work the right way}
-%\begin{verbatim}
-%
-%GLOBAL_(x)
-%
-%foo(*y){
-%  *y++;
-%}
-%
-%TASK( ){
-%  ... ;
-%  foo( &GV(x) ) ;
-%}
-%
-%\end{verbatim}
-%\subsection{Managing Paging Overhead}
-%
-%The efficiency of \sys's paging scheme relies on the fact that data in pages are contiguous, which makes them amenable to being manipulated by hardware assisted DMA bulk copy operations.
-To justify this design choice we experimentally evaluated the efficiency of using DMA block copies, compared to explicit, software copy loops.
-We copied blocks of data of different size between memory regions on a WISP~\cite{wisp} using a hand-written loop and using DMA (refer to Section~\ref{sec:results_hardware} for hardware setup details).
-Figure~\ref{fig:dmaTimeEnergy} shows that for moving large blocks of data like a page, hardware DMA is considerably better than software in both run time and energy efficiency.
-\sys's design moves data in pages, not individual words, leveraging the increased efficiency of DMA data movement.
-%
-%\begin{figure}[t]
-%	\centering
-%	\subfloat[Time needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/dmaSize_time.eps} }
-%	\subfloat[Energy needed to transfer a block of data]{\includegraphics[width=0.49\columnwidth]{figures/energyConsumptionDMA_SW.eps}}
-%	\caption{Time and energy consumption of moving a block of data from SRAM to FRAM: CPU intervention versus Direct Memory Access (DMA).\todo{Explain HW/SW setup; make fonts larger, unify legends and axes}{Amjad}}
-%	\label{fig:dmaTimeEnergy}
-%\end{figure}

### 2018-08-05 paper/asplos19/background.tex
SIunits package errors fixed

@@
 tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.  
-Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \micro\farad~\cite{rodriguez_tbcs_2015} to 0.1 \farad~\cite{moo}) are considered, the problem of porting a program becomes much worse
+Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse
 because a large task on one device may be a relatively short task

### 2018-08-05 paper/asplos19/intro.tex
SIunits package errors fixed

@@
 buffers energy in a small storage capacitor~\cite{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
-threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., \nano\watt\ to \micro\watt) compared to a platform's operating
-power level (hundreds of \micro\watt\ to \milli\watt). A device operates briefly until it depletes its buffered energy, after which, it shuts
+threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., \si{\nano\watt}\ to \si{\micro\watt}) compared to a platform's operating
+power level (hundreds of \si{\micro\watt}\ to \si{\milli\watt}). A device operates briefly until it depletes its buffered energy, after which, it shuts
 down and recharges to operate again later---corresponding to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of

### 2018-08-05 paper/asplos19/paging.tex
partial commit

@@
 %
-\sys's {\em virtual memory manager} exposes non-volatile memory for
-\emph{protected} variables to the application as a virtual address space
-partitioned into privatized pages, a technique we call \emph{dynamic pages privatization}.
+\sys's {\em virtual memory manager} (VMM) abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \emph{private} and \emph{shadow} buffers.
+Each buffer is divided into pages.
+The VMM prohibits applications from directly access these buffers.
+Instead, it redirects any data request to the \emph{working} buffer which is located in the volatile memory (SRAM) has a lower latency and energy cost to access than the device's FRAM.
+It populates working buffer with the privatized pages from non-volatile memory.
+When a coalesced task completes, the VMM ensures that updated pages are committed to main memory and visible to subsequently executed tasks.
+If the power interrupts a task, the VMM ensures that the task sees consistent memory state when it restarts its execution.
+The VMM ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
 %
-A program's tasks read and write these protected variables, using them like any
-other data in the program.
-%
-The virtual memory manager automatically privatize the pages associated with the accessed protected variables.
-%
-When a task completes, the virtual memory manager ensures that updated pages
-are committed to main memory and visible to subesequently executed tasks.
-%
-If the power interrupts a task, the virtual memory manager ensures that
-the task sees consistent memory state when it restarts its execution.
-%
-The virtual memory manager ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
-
-\sys's virtual memory manager \sys abstracts non-volatile storage (e.g.,
-variables in FRAM) from tasks.
-%
-Protected variables remain consistent and are durable across power failures.
-%
-While ensuring variables remain durable, \sys avoids accessing non-volatile
-memory when possible, by copying pages of data into a {\em working memory} in the device's SRAM, which
-has a lower latency and energy cost to access than the device's FRAM (as we
-showed in Section~\ref{sec:background}).
-%
-On a commit, \sys atomically copies all updated (i.e., ``dirty'') pages from their
-space in the working memory back into their location in the non-volatile
-memory.
+On a commit, the VMM atomically copies all ``dirty'' pages back into their location in the non-volatile memory.
 %
@@
 %
-\sys moves pages of data efficiently using hardware accelerated Direct Memory
+the VMM moves pages of data efficiently using hardware accelerated Direct Memory
 Access (DMA) support, which copies a block of memory from one place to another
@@
 A page has a home location in non-volatile memory which we refer to as the page's
-\texttt{\underline{store}} (underlined locations are non-volatile).  When a
-task accesses a page, the memory manager swaps the page from its home location
+\texttt{\underline{private}} (underlined locations are non-volatile).  When a
+task accesses a page, the VMM swaps the page from its home location
 into the fully\hyp{}associative \texttt{working} buffer that is stored in
@@
 %
-\sys uses a two-phase commit to write dirty pages from the \texttt{working}
-buffer back to their home location in \texttt{\underline{store}} atomically,
-ensuring data consistency even in the presence of a power failure during commit.
+The VMM utilizes a two-phase commit to write dirty pages from the \texttt{working}
+buffer back to their home location in \texttt{\underline{private}}. It optimizes this process by using an indirection table to abstract the location of the \texttt{\underline{private}} buffer and
+to eliminate the need for the second page copy.
 %
@@
 moves the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
-page buffer. Later, to commit the page, \sys writes it from its
-\texttt{\underline{shadow}} location back to its home location in
-\texttt{\underline{store}}.
+page buffer. \\
 %

### 2018-08-05 paper/asplos19/paging.tex
partial commit paging

@@
 %
-\sys's {\em virtual memory manager} (VMM) abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \emph{private} and \emph{shadow} buffers.
+\sys's {\em virtual memory manager} (VMM) abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers.
 Each buffer is divided into pages.
 The VMM prohibits applications from directly access these buffers.
-Instead, it redirects any data request to the \emph{working} buffer which is located in the volatile memory (SRAM) has a lower latency and energy cost to access than the device's FRAM.
-It populates working buffer with the privatized pages from non-volatile memory.
+Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
+It populates \texttt{working} buffer with the privatized pages from non-volatile memory.
 When a coalesced task completes, the VMM ensures that updated pages are committed to main memory and visible to subsequently executed tasks.
-If the power interrupts a task, the VMM ensures that the task sees consistent memory state when it restarts its execution.
+If the power interrupts a task, the VMM ensures the consistency of the data by relying on its \texttt{\underline{private}} buffer.
 The VMM ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
@@
 %
-The memory manager is forced to swap out a dirty page from the \texttt{working} buffer
-when a task accesses a page that is not in the working buffer, and there is no
-free page in the working buffer (i.e., a page fault). On a page fault, \sys
-moves the page to a direct\hyp{}mapped non-volatile \texttt{\underline{shadow}}
-page buffer. \\
+The VMM is forced to swap out a dirty page from the \texttt{working} buffer
+when a task requests a page that is not in the \texttt{working} buffer, and there is no
+free page in the working buffer (i.e., a page fault). On a page fault, the VMM
+moves the page to the direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer. \\
 %

### 2018-08-05 paper/asplos19/paging.tex
Page Faults and Page Swapping revised

@@
 {\tt RP} and {\tt WP} operations translate a variable's physical address in
-non-volatile memory into a \emph{virtual address} in the page buffer in
+non-volatile memory into a \emph{virtual address} in \texttt{working} buffer in
 volatile memory. In \sys, a virtual address is composed of a \emph{page tag}
@@
 After address translation (Figure~\ref{figure:mem-man}, Step 6), a task
-accesses the protected variable's location in the volatile paging buffer
+accesses the protected variable's location in the volatile \texttt{working} buffer
 (Step 7). \sys keeps track of the page tags for
-the pages currently resident in the paging buffer. When a task accesses a
+the pages currently resident in the \texttt{working} buffer. When a task accesses a
 variable, it compares the variable's page tag to tags of the pages in {\tt
@@
 If the \texttt{working} buffer is full, a page fault on a memory access
-requires \sys to swap out one of the pages in the volatile buffer (a
+requires VMM to swap out one of the pages in the \texttt{working} buffer (a
 \emph{victim page}) preserving updates made to that page.
@@
 \texttt{\underline{shadow}} buffer, copying them back to their original
-locations in the main memory \texttt{\underline{store}}. If a task were allowed
+locations in the main memory \texttt{\underline{private}}. If a task were allowed
 to (re-)execute after only some pages have been committed to

### 2018-08-05 paper/asplos19/intro.tex
intro

@@
 approaches are simpler and applicable to existing devices today.
-Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.
+\note{Talk about CP-based systems and talk why we focused task-based instead...}Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.
 
@@
     \centering
-    \includegraphics[width=0.7\columnwidth]{figures/intro-figure-vert.pdf}
+    \includegraphics[width=\columnwidth]{figures/intro-figure-v3.pdf}
     \caption{\note{small static task subplot and large static task subplot can be re-arranged to fit next to each other, so we end up with two layers plot} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
@@
 These constraints on task sizing lead to the following dilemma: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
-In this work,
-we propose a third way, using a novel technique called \emph{adaptive task size execution by task coalescing/splitting} that efficiently executes small tasks, trimming unnecessary overheads, while avoiding the risk of non-termination.
 
@@
 \note{Why no indent?}
-This paper introduces a new system called \emph{\sys}, which executes a task-based
-intermittent program by \emph{dynamically} coalescing/splitting its tasks.
+This paper introduces \sys, a new task-based system that employs \emph{adaptive task size execution by task coalescing/splitting}.
+By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overheads \emph{dynamically}, meanwhile avoiding the risk of non-termination.
 \sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce}, top) or splits them (Fig.~\ref{fig:coalesce},bottom) based on the estimated energy availability without demanding any hardware support.
+\note{\textbf{To the best of our knowledge, \sys is the only task-based system that guarantees task termination and ....(something related with Coalescing...)  }}
+
 The unique challenges and contributions revealed by this work are listed below:

### 2018-08-05 paper/asplos19/paging.tex
first pass over paging

@@
 accesses the protected variable's location in the volatile \texttt{working} buffer
-(Step 7). \sys keeps track of the page tags for
+(Step 7). The VMM keeps track of the page tags for
 the pages currently resident in the \texttt{working} buffer. When a task accesses a
@@
 \underline{shadow}} (Step 3), otherwise it has to be retrieved from
-{\tt \underline{store}} (Step 4).
+{\tt \underline{private}} (Step 4).
 Finally, the page is copied to the volatile buffer (Step 5). \\
@@
 \texttt{\underline{shadow}} buffer, copying them back to their original
-locations in the main memory \texttt{\underline{private}}. If a task were allowed
-to (re-)execute after only some pages have been committed to
-\texttt{\underline{store}}, the task might see protected variables in an
-inconsistent state; \sys's two-phase commit prevents this inconsistency. \\
-%
+locations in the main memory \texttt{\underline{private}}.\\
 To make the commit atomic, \sys commits the pages in two phases, implemented by
@@
 non-volatile \texttt{\underline{shadow}} buffer (Line 4). The second phase
-commits pages from \texttt{\underline{shadow}} to \texttt{\underline{store}}
+commits pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}
 (Line 12).  If power fails during the first phase, the whole commit is aborted,
-and the execution restarts from the most recently committed point (i.e., from
-the first task in a coalesced sequence of tasks). If power fails in the second
+and the execution restarts from the most recently committed point (i.e., from the beginning of a coalesced task). If power fails in the second
 phase, the commit process safely resumes on reboot.  The second phase of commit
@@
 \texttt{\underline{shadowCount}} records the number of dirty shadow pages to
-commit and \sys clears the counter when commit completes (Line 14).  The
+commit and the VMM clears the counter when commit completes (Line 14).  The
 \texttt{\underline{commitIndex}} indexes the next page to be committed (Line
-11) and \sys clears the index at the end of the phase (Line 15).
+11) and the VMM clears the index at the end of the phase (Line 15).
 %
@@
 %
-\sys's implementation is also efficient because the memory manager does not
-copy from \texttt{\underline{shadow}} to \texttt{\underline{store}}, instead
+\sys's implementation is also efficient because the VMM does not
+copy from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead
 swapping pointers in an indirection table that maintains the pages as a double
@@
 ever executes using consistent protected data. During task execution,
-modifications to protected data do not affect their original memory locations,
+modifications to protected data do not affect the \texttt{\underline{private}} buffer,
 because a task reads and writes the volatile \texttt{working} buffer only, and
 modified pages are kept in \texttt{\underline{shadow}} until commit.  A power
-failure erases the contents of the page buffer, preventing a re-executing task
-from observing updates in the page buffer from a previous execution attempt.
+failure erases the contents of the \texttt{working} buffer, preventing a re-executing task
+from observing updates from a previous execution attempt.
 Clearing \texttt{\underline{shadowCount}} as part of the second phase commit
 (Line 14), ensures that all accesses to protected variables in subsequent tasks
-correctly access their original memory locations in \texttt{\underline{store}}.
-\sys assumes that the write to a non-volatile word in memory is atomic. \\
+correctly access their original memory locations in \texttt{\underline{private}}.\\
 %
@@
 check on every read and write. 
-An alternative approach~\cite{alpaca} relied on a compiler to analyze memory operations and try to insert memory protection code mostly at the beginning of a task, leaving most
-memory accesses uninstrumented.\\
 %
 Despite the risk of per-access overhead, \sys's dynamic memory protection
-scheme brings several benefits over a static approach.  First, the limitations
-of static analysis preclude some uses of pointers due to potential pointer
+scheme brings several benefits over a static approach (i.e.~\cite{alpaca}).  
+First, the limitations of static analysis preclude some uses of pointers due to potential pointer
 aliasing. For example, in the presence of arbitrary pointer operation, a

### 2018-08-05 paper/asplos19/abstract.tex

@@
-Freeing embedded devices from batteries and powering them from ambient energy is a key factor in developing the energy-neutral Internet of Things.
+Energy-neutral Internet of Things requires freeing embedded devices from batteries and powering them from ambient energy.
 Ambient energy is, however, unpredictable and can only power a device intermittently.
-Therefore, the two paradigms of intermittent execution (checkpointing and task-based) save the intermediate program state into non-volatile memory frequently, to preserve the execution progress of a program.
-Despite the superiority of the task-based approach, tasks' size is fixed at compile-time, and agnostic to energy conditions.
-% it fixes the task size at compile time and operates obliviously to changing energy conditions.
-% When tasks are sized without accounting for available energy,
+Therefore, the two paradigms of intermittent execution, program checkpointing and task-based programming, save the intermediate program state into non-volatile memory frequently to preserve the execution progress.
+Despite the superiority of the task-based approach over checkpointing, tasks' size is fixed at compile-time and agnostic to energy conditions.
 Thus, the state may be saved either more frequently than necessary or not frequently enough for the program to progress and terminate.
-To address these challenges, we propose \sys, an \emph{adaptive} and \emph{efficient} task-based execution model.
+To address these challenges, we propose \sys, an adaptive and efficient task-based execution model.
 \sys progresses on a multi-task scale when energy permits and preserves the computation progress on a sub-task scale if necessary.
-\sys ensures that power failures do not leave the program state in non-volatile memory inconsistent using a specialized memory virtualization mechanism.
+\sys' specialized memory virtualization mechanism ensures that power failures do not leave the program state in non-volatile memory inconsistent.
 Our evaluation on a real energy-harvesting platform shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, and is able to progress where static systems fail.
-% Our evaluation shows that \sys is $2X$ faster than the state-of-the-art system in most cases and is able to progress where static systems fail.

### 2018-08-05 paper/asplos19/intro.tex
revisions to intro (not complete)

@@
-%%%%%%%%%%%%%%%%%%%%%% IMPORTANT INFO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% SRAM read/write latency is in the order of 8ns 		 [ Intermittent computation without hardware support of programmer intervention]
-% FRAM read/write latency is in the orfer of 60ns/100ns  [Texas Instrument MSP430FR573x DataSheet]
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-Advances in processor efficiency along with the development of
-energy-harvesting systems have created a new category of devices that require
-neither a battery nor a tethered power
-supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}. These
-devices operate using ambient energy, such as radio frequency
-transmissions~\cite{rf_powered_computing_gollakota_2014},
-light~\cite{margolies_infocom_2016,margolies_tosn_2016}, and
-vibration~\cite{gorlatova_sigmetrics_2014}. Incorporating compute, storage,
-sensing, and communication hardware~\cite{wisp5,moo,capybara}, such devices are a
-promising technology for use in the Internet of Things~\cite{ku_cst_2016},
-in-body~\cite{nadeau_naturebio_2017} and
-on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and
-energy-harvesting nano-satellites~\cite{kicksat,capybara}.
-%
-Energy-harvesting devices create unique challenges because they operate {\em
-intermittently} when energy is
-available~\cite{hicks_isca_2017,lucia_snapl_2017}. An energy-harvesting device
-buffers energy in a small storage capacitor~\cite{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a
-threshold amount of energy has accumulated. Harvestable energy sources are low-power (e.g., \si{\nano\watt}\ to \si{\micro\watt}) compared to a platform's operating
-power level (hundreds of \si{\micro\watt}\ to \si{\milli\watt}). A device operates briefly until it depletes its buffered energy, after which, it shuts
-down and recharges to operate again later---corresponding to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of
-operation-power failure-restart cycles. 
-The recharge and discharge intervals---which correspond to the device's inactive and active time---vary with the energy buffering capacitor~\cite{capybara} \note{not only, though} and some devices fail and restart operating $\approx$10 to
-$\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.
-%
-%\textbf{Data Consistency in Intermittent Computing.}
+Advances in processor efficiency along with the development of energy-harvesting systems have created a new category of devices that require neither a battery nor a tethered power supply~\cite{prasad_comst_2014,lucia_snapl_2017,soyata_csm_2016}.
+These devices operate using ambient energy, such as radio frequency transmissions~\cite{rf_powered_computing_gollakota_2014},
+light~\cite{margolies_infocom_2016,margolies_tosn_2016}, and vibration~\cite{gorlatova_sigmetrics_2014}.
+Incorporating compute, storage, sensing, and communication hardware~\cite{wisp5,moo,capybara}, such devices are a promising technology for use in the Internet of Things~\cite{ku_cst_2016}, in-body~\cite{nadeau_naturebio_2017} and on-body~\cite{bandodkar_electroanalysis_2015} medical systems, and energy-harvesting nano-satellites~\cite{kicksat,capybara}.
+Energy-harvesting devices create unique challenges because they operate {\em intermittently} when energy is available~\cite{hicks_isca_2017,lucia_snapl_2017}.
+An energy-harvesting device buffers energy in a small storage capacitor~\cite{gorlatova_tmc_2013,gunduz_commag_2014} and operates when a threshold amount of energy has accumulated.
+Harvestable energy sources are low-power (e.g., \si{\nano\watt}\ to \si{\micro\watt}) compared to a platform's operating power level (hundreds of \si{\micro\watt}\ to \si{\milli\watt}).
+A device operates briefly until it depletes its buffered energy, after which, it shuts down and recharges to operate again later---corresponding to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of operation-power failure-restart cycles.
+The recharge and discharge intervals---which correspond to the device's inactive and active time---vary depending on the underlying hardware, such as the size of the energy buffering capacitor~\cite{capybara}, and some devices discharge and restart $\approx$10 to $\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.
 
-Upon power failures, a device loses the volatile
-state in its registers, stack, SRAM, and retains the state of any non-volatile
-memory, such as FRAM. While capturing periodic
-checkpoints~\cite{mementos,quickrecall} and sleep
-scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution
-progress, failures can leave non-volatile state incorrect, partially updated,
-and may leave checkpointed volatile state inconsistent with non-volatile state.
-These inconsistencies cause intermittent execution to deviate from
-continuously-powered behavior, often leading to an unrecoverable
-application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for
-intermittently-powered devices: (i) \emph{software-based programming and
-execution models}~\cite{dino,ratchet,chain,alpaca} and (ii)
-\emph{hardware-based computer architecture
-support}~\cite{hicks_isca_2017,idetic,nvp}. Complex architectural changes are
-expensive to design, verify, and manufacture. New architectures are also
-inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software
-approaches are simpler and applicable to existing devices today.
+Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM.
+While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated, and may leave checkpointed volatile state inconsistent with non-volatile state.
+These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, often leading to an unrecoverable application failure~\cite{dino,edb}.
+Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based computer architecture support}~\cite{hicks_isca_2017,idetic,nvp}.
+Complex architectural changes are expensive to design, verify, and manufacture.
+New architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}.
+Software approaches are simpler and applicable to existing devices today.
 \note{Talk about CP-based systems and talk why we focused task-based instead...}Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.
 
-%
 \begin{figure}
@@
     \includegraphics[width=\columnwidth]{figures/intro-figure-v3.pdf}
-    \caption{\note{small static task subplot and large static task subplot can be re-arranged to fit next to each other, so we end up with two layers plot} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
+    \caption{\note{small static task subplot and large static task subplot can be re-arranged to fit next to each other, so we end up with two layers plot} \note{Add distinction: top---legacy task-based system; bottom---Coala} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
@@
 
-\noindent\textbf{Crucial Drawback of Static Task Decomposition.} 
-\note{Why no indent?}
-Task-based programming and execution models require a
-programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to
-statically decompose a program into a collection of tasks. A \emph{task}, a top level function, can
-include arbitrary computation that should be executed despite arbitrarily-timed power failures.
+\textbf{Crucial Drawback of Static Task Decomposition.} Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks.
+A \emph{task}, a top level function, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.
 The programmer (or a compiler) explicitly expresses task-to-task control flow.
@@
 The more task transitions a program requires, the more commit overhead is incurred by the system at runtime.
-%
-A programmer may thus create very large tasks in an attempt to reduce task transitions
-overhead. However, a very large task may require more energy to complete than a device's fixed hardware energy
-buffer can hold which may lead to a task \emph{non-termination} problem. To eliminate this risk,
-existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress.
+A programmer may thus create very large tasks in an attempt to reduce task transitions overhead.
+However, a very large task may require more energy to complete than a device's fixed hardware energy buffer can hold which may lead to a task \emph{non-termination} problem.
+To eliminate this risk, existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress.
 These constraints on task sizing lead to the following dilemma: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
 
-\noindent\textbf{Challenges and Contributions.}
-\note{Why no indent?}
-This paper introduces \sys, a new task-based system that employs \emph{adaptive task size execution by task coalescing/splitting}.
+\textbf{Challenges and Contributions.} This paper introduces \sys, a new task-based system that employs \emph{adaptive task size execution by task coalescing/splitting}.
 By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overheads \emph{dynamically}, meanwhile avoiding the risk of non-termination.

### 2018-08-05 paper/asplos19/intro.tex
update intro fig

--- 2018-08-05 paper/asplos19/intro.tex
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/intro-figure-v3.pdf}
-    \caption{\note{small static task subplot and large static task subplot can be re-arranged to fit next to each other, so we end up with two layers plot} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
+    \includegraphics[width=\columnwidth]{figures/intro-figure-horiz.pdf}
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}

### 2018-08-05 paper/asplos19/intro.tex

--- 2018-08-05 paper/asplos19/intro.tex
@@
 \note{Talk about CP-based systems and talk why we focused task-based instead...}Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.
-
+%
 \begin{figure}
     \centering
-    \includegraphics[width=\columnwidth]{figures/intro-figure-v3.pdf}
-    \caption{\note{small static task subplot and large static task subplot can be re-arranged to fit next to each other, so we end up with two layers plot} \note{Add distinction: top---legacy task-based system; bottom---Coala} Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
+    \includegraphics[width=\columnwidth]{figures/intro-figure-horiz.pdf}
+    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
 On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
@@
 %\textbf{Task Decomposition of Intermittent Programs.} 
-
+%
 \textbf{Crucial Drawback of Static Task Decomposition.} Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks.

### 2018-08-05 paper/asplos19/intro.tex

@@
 Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based computer architecture support}~\cite{hicks_isca_2017,idetic,nvp}.
-Complex architectural changes are expensive to design, verify, and manufacture.
-New architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}.
+Complex hardware architectural changes are expensive to design, verify, and manufacture.
+New hardware architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}.
 Software approaches are simpler and applicable to existing devices today.
-\note{Talk about CP-based systems and talk why we focused task-based instead...}Therefore, this work focuses on a key limitation of task-based software approaches, that is, the {\em inflexibility} of statically decomposing a program into tasks.
+Therefore, this work focuses on software approaches.
+In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} of statically decomposing a program into tasks.
 %
@@
 \end{figure}
-%
+
 %\textbf{Task Decomposition of Intermittent Programs.}
-%
+
 \textbf{Crucial Drawback of Static Task Decomposition.} Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks.
@@
 To eliminate this risk, existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress.
-These constraints on task sizing lead to the following dilemma: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete, but incur a high task transition and commit overhead.
+These constraints on task sizing lead to the following \emph{dilemma}: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete but incur a high task transition and commit overhead?
 
@@
 %\textbf{Challenges and Contributions}
-\noindent\note{Carlo's rephrasing} \emph{C1) Task Transition Overhead:} given unpredictable incoming energy, how to save progress as few times as possible?
+
+\noindent \emph{C1) Task Transition Overhead:} given unpredictable incoming energy, how to save progress as few times as possible?
 \sys tries to minimize task transition overhead by estimating energy conditions at run-time using \emph{recent execution history} as a metric.
@@
 
-To asses the benefits of \sys over existing task-based systems, we implemented and tested six
-benchmarks on a real energy-harvesting platform.
+To asses the benefits of \sys over existing task-based systems, we implemented and tested six benchmarks on a real energy-harvesting platform.
 Our evaluation shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, and is able to progress where static systems fail.
@@
 The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
-Section~\ref{sec:systemdescription} provides an overview of \sys, while
-Section~\ref{sec:task_adaptation} describes \sys's energy-adaptive mechanism.
+Section~\ref{sec:systemdescription} provides an overview of \sys, while Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
 The dynamic group privatization is explained in Section~\ref{sec:memory_virtulaization}.
@@
 Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results.
-Section~\ref{sec:related_work} positions \sys in the
-context of related work and Section~\ref{sec:conclusions} concludes and discusses
-future work.
+Section~\ref{sec:related_work} positions \sys in the context of related work and Section~\ref{sec:conclusions} concludes and discusses future work.

### 2018-08-05 paper/asplos19/implementation.tex
remove COALA

@@
 
-%\texttt{COALA\_TASK} is a function annotation that declares a function to be a
-%task. \texttt{COALA\_NEXT\_TASK}, indicates the task to run next, after an
-%executing task completes.  \texttt{COALA\_PV} declare a new protected variable.
+%\texttt{TASK} is a function annotation that declares a function to be a
+%task. \texttt{NEXT\_TASK}, indicates the task to run next, after an
+%executing task completes.  \texttt{PV} declare a new protected variable.
 %The programmer uses \texttt{RP} and \texttt{WP}, to read and write protected
-%respectively.  \texttt{COALA\_SM} identifies protected \texttt{struct} fields
+%respectively.  \texttt{SM} identifies protected \texttt{struct} fields
 %and tells \sys to ensure that the struct fields are page-aligned in memory.
 %The programmer is required to explicitly call two API functions at the start of
-%{\tt main()}: \texttt{COALA\_INIT}, which does global initialization upon first
-%boot and \texttt{COALA\_RUN} which hands the control to \sys's task manager to
+%{\tt main()}: \texttt{INIT}, which does global initialization upon first
+%boot and \texttt{RUN} which hands the control to \sys's task manager to
 %resume task execution.
@@
 
-\noindent\textbf{New Tasks.} The \texttt{COALA\_TASK} annotation on a function
+\noindent\textbf{New Tasks.} The \texttt{TASK} annotation on a function
 declaration statically allocates a non-volatile constant variable holding a
@@
 %
-\noindent\textbf{Task Transitions.} A call to \texttt{COALA\_NEXT\_TASK} marks the task to be executed at the 
+\noindent\textbf{Task Transitions.} A call to \texttt{NEXT\_TASK} marks the task to be executed at the 
 end of the current one task and it can be invoked along
@@
 %
-\noindent\textbf{Protected Variables.} The \texttt{COALA\_PV} annotation on a
+\noindent\textbf{Protected Variables.} The \texttt{PV} annotation on a
 variable statically allocates a protected non-volatile variable. The
@@
 \noindent\textbf{Initialization.} The behavior of the API method
-\texttt{COALA\_INIT} is very similar to \texttt{COALA\_NEXT\_TASK}, with the
+\texttt{INIT} is very similar to \texttt{NEXT\_TASK}, with the
 addition of performing preliminary kernel initializations, including hardware setup.\\
@@
 \noindent\textbf{Execution.} The programmer passes control to \sys's task scheduler 
-by calling \texttt{COALA\_RUN} after device initialization. 
+by calling \texttt{RUN} after device initialization. 
 %
@@
 disable partial commit for a task or a span of code, marking  the code with a
-pair of \texttt{COALA\_DISABLE\_PE} and \texttt{COALA\_ENABLE\_PE} annotations.
+pair of \texttt{DISABLE\_PE} and \texttt{ENABLE\_PE} annotations.
 These annotations respectively halt and resume the partial execution timer.

### 2018-08-05 paper/asplos19/implementation.tex
implemenation: paging

@@
 
-\subsection{\textcolor{red}{to be revised after Virtual Mem.}Paging}
+\subsection{Paging}
 \label{sec:impl:paging}
@@
 virtualization scheme uses three buffers: a volatile \texttt{working} buffer, a
-non-volatile main memory \texttt{\underline{store}}, and a non-volatile
-\texttt{\underline{shadow}} page buffer. 
+non-volatile \texttt{\underline{private}}, and a non-volatile
+\texttt{\underline{shadow}} buffer. 
 %
-The main memory \texttt{\underline{store}} holds the last committed copy of all 
+The main memory \texttt{\underline{private}} holds the last committed copy of all 
 pages. 
@@
 After reboot, the \texttt{working} buffer is empty. Execution fills 
-the working buffer with pages as the program accesses them. 
+the \texttt{working} buffer with privatized pages as the program accesses them. 
 %
-The program primarily reads and writes data paged into the volatile working memory. 
-%
-The non-volatile \texttt{\underline{shadow}} buffer serves as an intermediate,
-non-volatile backing store for dirty pages evicted from the space-constrained
-working buffer.
+The \texttt{\underline{shadow}} buffer serves as an intermediate,
+non-volatile backing \texttt{\underline{private}} for dirty pages evicted from the space-constrained
+\texttt{working} buffer.
 
 To access a protected variable with either \texttt{RP}
-or \texttt{WP}, \sys searches for the variable in its working buffer first. If
-found, the program accesses the data in the working buffer.  If the page is not 
-in the working memory, the memory manager fetches it from non-volatile memory. 
+or \texttt{WP}, \sys searches for the variable in its \texttt{working} buffer first.
+ If the page is not, the VMM fetches it from non-volatile memory. 
 %
@@
 %
-Pages must be aligned to an $S$-byte boundary. 
+Pages must be aligned to an $S$-byte boundary for efficient memory access. 
 %
 We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
-an 8\,KB non-volatile \texttt{\underline{store}}, and a volatile working buffer
-of 1\,KB, which is half of the total SRAM size in the MSP430FR6989 microcontroller.
+an 8\,KB non-volatile \texttt{\underline{private}}, and a \texttt{working} buffer
+of 1\,KB.
 %

### 2018-08-05 paper/asplos19/intro.tex

@@
 
-\textbf{Challenges and Contributions.} This paper introduces \sys, a new task-based system that employs \emph{adaptive task size execution by task coalescing/splitting}.
+\textbf{Challenges and Contributions.} This paper introduces \sys: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
 By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overheads \emph{dynamically}, meanwhile avoiding the risk of non-termination.
-\sys accepts any static decomposition and it coalesces (groups) tasks (Fig.~\ref{fig:coalesce}, top) or splits them (Fig.~\ref{fig:coalesce},bottom) based on the estimated energy availability without demanding any hardware support.
-\note{\textbf{To the best of our knowledge, \sys is the only task-based system that guarantees task termination and ....(something related with Coalescing...)  }}
+\sys accepts any static decomposition and it coalesces (groups) tasks or splits them (see again Fig.~\ref{fig:coalesce}) based on the estimated energy availability without demanding any hardware support.
+To the best of our knowledge, \sys is the only task-based system that guarantees forward progress for any statically-defined tasks or energy storage sizes.
 
-The unique challenges and contributions revealed by this work are listed below:
-\note{it takes more space, but a bullet list makes C1-C3 stand out more properly}
+The unique contributions of \sys in relation to challenges of task-based systems revealed by this work are as follows:
 
-% \sys does not requires hardware support for energy estimation to make its scheduling decisions.
+\begin{itemize}
 
-% \sys accepts any static decomposition and coalesces/splits its tasks at runtime. Two consecutive, coalesced tasks execute
-% with no commit overhead between them, instead performing task-end commit actions at the end of the second task only.
-% %
-% When there is sufficient energy to execute both tasks, two distinct, small
-% tasks are effectively executed as a single large task, preserving the atomicity
-% of both. If power fails during a coalesced task, execution restarts from the
-% {\em first} of the coalesced tasks. Figure~\ref{fig:coalesce} (top)
-% illustrates how dynamic coalescing removes unnecessary overhead and shortens intermittent program execution.
-% On the other hand, a large task that requires more energy to complete than a device buffered energy
-% is split by \sys to enable task termination and preserve execution progress, as depicted in Figure~\ref{fig:coalesce} (bottom).
+\item[C1] \emph{Overcoming Task Transition Overhead:} given unpredictable incoming energy, how to save computation progress as rare as possible?
+\sys tries to minimize task transition overhead by estimating energy conditions at run-time using \emph{recent execution history} as a metric.
 
-%\textbf{Challenges and Contributions}
-
-\noindent \emph{C1) Task Transition Overhead:} given unpredictable incoming energy, how to save progress as few times as possible?
-\sys tries to minimize task transition overhead by estimating energy conditions at run-time using \emph{recent execution history} as a metric.
-% \noindent\emph{C1) Dynamic Task Sizing:} Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
-\sys addresses this challenge by using a novel approach: \textit{recent execution history} as a metric to estimate the environment energy conditions and to specify task sizes.
-%Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.
-
-\noindent\emph{C2) Dynamic Memory Consistency Handling:} merging static tasks on the fly gives rise to the need for dynamic memory consistency handling.
+\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly gives rise to the need for dynamic memory consistency handling.
 How to dynamically detect inter-task data dependencies and ensure efficient protection against power interrupts?
-\sys does so by using a novel
-approach called \emph{dynamic group privatization}--- %Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e.
-protecting all global variables) in each dynamic task transition.
-performing real-time dependency tracking to enable protection on a task transition.
+\sys does so by using a novel approach called \emph{dynamic group privatization}---performing real-time dependency tracking to enable protection on a task transition.
 Individual variables tracking, however, slows down a system dramatically.
@@
 
-\noindent\emph{C3) Ensuring Task Termination:} a static task decomposition model assumes that each
-single task can execute to completion. If the hardware energy buffer provides
-inadequate energy to execute each single task to completion, a program will not
-terminate~\cite{cleancut_2018}. Under such conditions, how to enable the dynamic execution model to progress on a sub-task level? To avoid non-termination under adverse
-energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism \note{should we instead cite Ratchet?}.
-Partial execution avoids non-termination by committing the intermediate state of a long-running
-task that has repeatedly failed and restarted.
-%Partial execution violates task atomicity, but preserves forward progress; if a programmer knows that task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
+\item[C3] \emph{Ensuring Task Termination:} a static task decomposition model assumes that each single task can execute to completion.
+If the hardware energy buffer provides inadequate energy to execute each single task to completion, a program will not terminate~\cite{cleancut_2018}.
+Under such conditions, how to enable the dynamic execution model to progress on a sub-task level?
+To avoid non-termination under adverse energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism \note{should we instead cite Ratchet?}.
+Partial execution avoids non-termination by committing the intermediate state of a long-running task that has repeatedly failed and restarted.
 
-%\begin{itemize}
-%\item\textit{challenge} 1) Given unpredictable energy shots arrival, how to specify the size of a dynamic task on the fly?
-\textit{contribution---1} \sys addresses this challenge by using its \textit{recent execution history} as a metric to estimate the environment energy conditions.
-Furthermore, It mitigates the risk of dynamic task re-execution by taking advantage of the certainty that the energy buffer offers and approaches the expected power interrupt carefully.
-%%
-%\item\textit{challenge} 2) Given dynamic variables dependencies, how to ensure efficient data protection against power interrupts?
-\textit{contribution---2} \sys ensures the consistency of the data by using a novel
-%approach called \emph{dynamic group privatization}.
-Varying variables dependencies forces a static approach (i.e.~\cite{alpaca}) to operate by considering a worst case scenario (i.e.
-protecting all global variables) in each dynamic task transition.
-\sys, therefore, utilizes a real-time data tracking to enable partial data protection on a task transition.
-Individual variables tacking, however, slows down a system dramatically.
-Therefore, \sys  relies on a coarse locality principle and protects the global variables in a batch.
-Moreover, it optimizes the data transfer by using Direct Memory Access (DMA).
-%%
-%\item\textit{challenge} 3) A static task decomposition model assumes that each
-%single task can execute to completion. If the hardware energy buffer provides
-%inadequate energy to execute each single task to completion, a program will not
-%terminate~\cite{cleancut_2018}. under such conditions how to enable the dynamic execution model to progress on a sub-task level? \textit{contribution---3} To avoid non-termination under adversarial
-%energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
-%Partial execution avoids non-termination by committing the intermediate state of a long-running
-%task that has repeatedly failed and restarted. Partial execution violates
-%task atomicity, but preserves forward progress; if a programmer knows that
-%task atomicity is crucial to correctness, they can disable partial execution around a piece of code.
-%\end{itemize}
-%
+\end{itemize}
 
 To asses the benefits of \sys over existing task-based systems, we implemented and tested six benchmarks on a real energy-harvesting platform.
-Our evaluation shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, and is able to progress where static systems fail.
+Our evaluation shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, and is able to progress where static task-based systems fail.
 

### 2018-08-05 paper/asplos19/implementation.tex
revise implemenation

@@
 We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittence-safe.
-
+%
 \subsection{Application Programming Interface}
 \label{sec:coala_api}
-
+%
 Coala's API adds only eight syntactic constructs to a C-based language, summarized
 in Table~\ref{table:coala_api}.
-
-%\texttt{TASK} is a function annotation that declares a function to be a
-%task. \texttt{NEXT\_TASK}, indicates the task to run next, after an
-%executing task completes.  \texttt{PV} declare a new protected variable.
-%The programmer uses \texttt{RP} and \texttt{WP}, to read and write protected
-%respectively.  \texttt{SM} identifies protected \texttt{struct} fields
-%and tells \sys to ensure that the struct fields are page-aligned in memory.
-%The programmer is required to explicitly call two API functions at the start of
-%{\tt main()}: \texttt{INIT}, which does global initialization upon first
-%boot and \texttt{RUN} which hands the control to \sys's task manager to
-%resume task execution.
-
+%
 \begin{table}
@@
 \end{table}
-
-%\subsection{API Details}
-
+%
 \noindent\textbf{New Tasks.} The \texttt{TASK} annotation on a function
@@
 %
-\sys uses a two-phase commit process to atomically commit pages despite
-possible power interruptions during commit.  When a coalesced task completes,
-\sys backs up dirty pages in volatile working memory to the non-volatile
-shadow buffer using DMA. 
+The coalescing algorithm starts with a random guess about the size of the coalescing task, in our implementation the target is initialized to four.
+It then refines its guess based on the number of the static-tasks being successfully executed.
 %
-If power fails while backing up these pages, commit fails and the execution
-resumes from the last committed point.  After backing up pages from working
-memory to the shadow buffer, \sys sets a flag that indicates an on-going atomic
-commit.  At this point, commit is irrevocable and will continue until it
-completes despite power failures.   
-%
-Each page is double buffered, with one copy always in the shadow page buffer
-and one copy always in the main memory store.  \sys uses double buffering to
-commit dirty pages efficiently by swapping which buffer is which, without
-copying any data.
+When a coalesced task completes, \sys save all the modifications in its \texttt{\underline{private}} using DMA.
+In order to ensure the atomicity of the commit operation all the modified pages are copied to the \texttt{\underline{shadow}} buffer first.
+The \sys set a flag to indicates an on-going \emph{atomic commit} to update the private pages.
+For efficiency reasons \sys uses an indirection table to point to the private consistent pages and eliminate the need for moving pages for the second time.
 %

### 2018-08-05 paper/asplos19/intro.tex

@@
 A programmer may thus create very large tasks in an attempt to reduce task transitions overhead.
-However, a very large task may require more energy to complete than a device's fixed hardware energy buffer can hold which may lead to a task \emph{non-termination} problem.
-To eliminate this risk, existing systems require the programmer to decompose a program into relatively small tasks to preserve execution progress.
+However, a large task may require more energy to complete than a device's fixed hardware energy buffer can hold, which may lead to a task \emph{non-termination} problem.
+To eliminate this risk, existing systems require the programmer to decompose a program into small tasks to preserve execution progress.
 These constraints on task sizing lead to the following \emph{dilemma}: should large tasks be used, that are efficient but risk non-termination, or small tasks that are guaranteed to complete but incur a high task transition and commit overhead?
 
-\textbf{Challenges and Contributions.} This paper introduces \sys: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
+\textbf{Challenges and Contributions.} We introduce \sys: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
 By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overheads \emph{dynamically}, meanwhile avoiding the risk of non-termination.
 \sys accepts any static decomposition and it coalesces (groups) tasks or splits them (see again Fig.~\ref{fig:coalesce}) based on the estimated energy availability without demanding any hardware support.
-To the best of our knowledge, \sys is the only task-based system that guarantees forward progress for any statically-defined tasks or energy storage sizes.
+To the best of our knowledge, \sys is the only system that guarantees forward progress for any statically-defined tasks or energy storage sizes.
 
@@
 
-\item[C1] \emph{Overcoming Task Transition Overhead:} given unpredictable incoming energy, how to save computation progress as rare as possible?
+\item[C1] \emph{Overcoming Task Transition Overhead:} given unpredictable incoming energy, how to save computation progress, i.e.
+saving state at task transitions, as rare as possible?
 \sys tries to minimize task transition overhead by estimating energy conditions at run-time using \emph{recent execution history} as a metric.
 
-\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly gives rise to the need for dynamic memory consistency handling.
-How to dynamically detect inter-task data dependencies and ensure efficient protection against power interrupts?
-\sys does so by using a novel approach called \emph{dynamic group privatization}---performing real-time dependency tracking to enable protection on a task transition.
+\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly gives rises  the need for dynamic memory consistency handling.
+This leads to the second challenge: how to dynamically detect inter-task data dependencies and ensure efficient protection against power interrupts?
+\sys addresses this challenge by using a novel approach called \emph{dynamic group privatization}---performing real-time dependency tracking to enable protection on a task transition.
 Individual variables tracking, however, slows down a system dramatically.
@@
 
-\item[C3] \emph{Ensuring Task Termination:} a static task decomposition model assumes that each single task can execute to completion.
-If the hardware energy buffer provides inadequate energy to execute each single task to completion, a program will not terminate~\cite{cleancut_2018}.
-Under such conditions, how to enable the dynamic execution model to progress on a sub-task level?
-To avoid non-termination under adverse energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism \note{should we instead cite Ratchet?}.
+\item[C3] \emph{Ensuring Task Termination:} a static task decomposition model assumes that each task can execute to completion.
+If the hardware energy buffer provides inadequate energy to execute each task to completion, a program will not terminate~\cite{cleancut_2018}.
+This leads to a third challenge: how to enable the dynamic execution model to progress on a sub-task level?
+To avoid non-termination under adverse energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism, inspired by~\cite{ratchet}.
 Partial execution avoids non-termination by committing the intermediate state of a long-running task that has repeatedly failed and restarted.

### 2018-08-05 paper/asplos19/background.tex
revisions to background (incomplete)

@@
 
-Energy harvesting devices operate using energy extracted from sources such as
-radio frequency transmissions and light. These devices elide tethered
-power and batteries, instead collecting energy into a capacitor, operating when
-sufficient charge accumulates, and turning off for recharge upon depletion of the buffer.
-There are several energy harvesting battery-less hardware platforms. For
-instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine}
-WISP~\cite{wisp5} (with its variants such as
-WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or
-NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones
-such as~\cite{medusa_farsens_2017}. Other platforms include ambient backscatter
-tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less
-phone~\cite{talla_imwut_2017}.
+Energy harvesting devices operate using energy extracted from ambient sources, e.g.
+radio frequency transmissions, light.
+These devices elide tethered power and batteries, instead collecting energy into a capacitor, operating when sufficient charge accumulates, and turning off for recharge upon depletion of the buffer.
+There are several energy harvesting battery-less hardware platforms.
+For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
+Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
 
-\textbf{Hardware Assumptions.} \sys is designed for the demands of existing and
-future energy-harvesting platforms based around general purpose, commodity
-computing components~\cite{wisp,msp430datasheet}. We assume a device with a
-memory system that has fast, byte-addressable volatile and non-volatile memory;
-in particular, our target platform, WISP~\cite{wisp}, is equipped with a
-mixture of SRAM and FRAM. Our implementation leverages hardware support for
-fast, bulk-copying between memories via DMA~\cite{msp430datasheet}. \sys does not
-require architectural additions to commodity processors as
-in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.
+\textbf{What Hardware does \sys Target.} \sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
+\sys targets a device with a
+memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM.
+\sys leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
+\sys does not require architectural additions to commodity processors as in~\cite{su_date_2017,hicks_isca_2017,quickrecall,nvp}.
 
@@
     \end{subfigure}
-    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables,
-    while \texttt{checkp()} is a checkpoint of volatile state.}
+    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state. \task{replace checkp(); with checkpoint();}}
     \label{fig:war}
@@
 
-Software running on an energy-harvesting device executes {\em intermittently}
-because power sources are not always available to harvest and buffer
-sufficient operational energy. An intermittent execution is composed of
-operating periods interspersed with power
-failures~\cite{dino,chain,alpaca,ratchet}.
-The frequency of failures depends on the size of the device's energy storage buffer
-(a larger buffer allows longer operating periods), current consumption and
-incoming energy.
+Software running on an energy-harvesting device executes {\em intermittently} because power sources are not always available to harvest and buffer sufficient operational energy.
+An intermittent execution is composed of operating periods interspersed with power failures~\cite{dino,chain,alpaca,ratchet}.
+The frequency of failures depends on the size of the device's energy storage buffer (a larger buffer allows longer operating periods), current consumption and incoming energy.
 
-A power failure clears volatile state (e.g., registers and SRAM) while
-non-volatile memory (e.g., FRAM) persists. Upon a power failure, control flows
-to a prior point in the execution: by default, to the beginning of {\tt
-main()}. Early intermittent systems preserved progress by periodically
-checkpointing volatile execution context to non-volatile
-memory~\cite{mementos}, sometimes using hardware
-support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}. 
-%
-Checkpointing volatile state alone does not ensure data consistency when the
-system can directly manipulate non-volatile memory~\cite{mspcdino}.
-Precisely, data can get inconsistent when code includes a
-\emph{write-after-read} (WAR) dependency between operations that manipulate
-non-volatile memory.
-% because it allows {\em writes} from a failed attempt to
-% persist across failure and be read by the {\em reads}~\cite{ratchet,dino,alpaca}.
-Fig.~\ref{fig:war} illustrates how state can become
-inconsistent in an intermittent execution using a simple example of an average
-operation over an array of integers.
-The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially
-by the increment operation.
+A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists.
+Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
+Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.
+Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}.
+Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory.
+Fig.~\ref{fig:war} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
+The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation.
 If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).
@@
 
-Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to
-decompose their program into \textbf{tasks}, which are regions of code that can contain
-arbitrary computation, sensing, and communication.
+Task-based execution models~\cite{dino,chain,alpaca} ask the programmer to decompose their program into \textbf{tasks}, which are regions of code that can contain arbitrary computation, sensing, and communication.
 Task-based models progress at the granularity of tasks.
-They re-execute each task interrupted by a power failure until it successfully finishes, only then moving on to
-the next task. Since these models do not rely on capturing an expensive
-checkpoint, they are usually faster than the checkpoint-based
-solution~\cite{chain, alpaca}.  \sys also follows the paradigm of the
-task-based programming model, where the programmer explicitly expresses the
-application as a sequence of tasks and the transitions between them.
+They re-execute each task interrupted by a power failure until it successfully finishes, only then moving on to the next task.
+Since these models do not rely on capturing an expensive checkpoint, they are usually faster than the checkpoint-based solution~\cite{chain, alpaca}.
+\sys also follows the paradigm of the task-based programming model, where the programmer explicitly expresses the application as a sequence of tasks and the transitions between them.
 
 Task-based models also suffer from data inconsistencies when WAR dependencies are involved.
-Prior systems tried to tackle this problem by a compiler-automated
-redo-logging for the variables that are part of the dependency~\cite{alpaca}, or
-by statically creating multiple copies of the problematic variable to ensure
-that no task reads and writes the same copy~\cite{chain}.
+Prior systems tried to tackle this problem by a compiler-automated redo-logging for the variables that are part of the dependency~\cite{alpaca}, or by statically creating multiple copies of the problematic variable to ensure that no task reads and writes the same copy~\cite{chain}.
 
@@
 
-Prior systems, including task-based models, have two major drawbacks:
-{\em frequent non-volatile memory accesses} and {\em incorrectly sized tasks}
-(too small or too large).
+Prior systems, including task-based models, have two major drawbacks: {\em frequent non-volatile memory accesses} and {\em incorrectly sized tasks} (too small or too large).
 

### 2018-08-05 paper/asplos19/background.tex
revisions to background (incomplete)
revisions to background (complete)

@@
 
-\textbf{What Hardware does \sys Target.} \sys is designed for the demands of existing and future energy-harvesting platforms based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
-\sys targets a device with a
-memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM.
+\textbf{What Hardware does \sys Target.} \sys is designed for the demands of existing and future \emph{tiny embedded energy-harvesting platforms} based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
+\sys targets a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM.
 \sys leverages hardware support for fast, bulk-copying between memories via DMA~\cite{msp430datasheet}.
@@
         \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
-        \caption{Incorrect execution}
+        \caption{WAR-affected code execution}
         \label{fig:war-execution}
     \end{subfigure}
-    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state. \task{replace checkp(); with checkpoint();}}
+    \caption{WAR dependency example.
+\texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state.
+\note{replace checkp(); with checkpoint();} \note{reduce font size by one point}}
     \label{fig:war}
@@
 If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).
+\note{in the above example it should \ref{fig:war-example}, Line 6, first and then \ref{fig:war-execution}, Lines 5 and 8}
 
@@
 
-\textbf{Frequent Non-volatile Memory Access.} 
-\note{I want to emphasize again that we do not evaluate NV memory accesses}
-Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.  Checkpointing systems copy all volatile
-state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at
-every checkpoint.  Task-based models, which do not take checkpoints, back up a
-subset of non-volatile data to maintain memory consistency using, e.g.,
-redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that
-gets copied to and from the non-volatile memory in service of memory
-consistency and progress preservation.  Ratchet~\cite{ratchet} and
-Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile,
-perform high-frequency non-volatile memory accesses, into checkpointing
-frequently \note{is "into" a misspell?}.
+\textbf{Frequent Non-volatile Memory Access.}  \note{I want to emphasize again that we do not evaluate NV memory accesses} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.
+Checkpointing systems copy all volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at every checkpoint.
+Task-based models, which do not take checkpoints, back up a subset of non-volatile data to maintain memory consistency using, e.g., redo-logging.
+Table~\ref{table:chechpoint_comparison} summarizes the data that gets copied to and from the non-volatile memory in service of memory consistency and progress preservation.
+Ratchet~\cite{ratchet} and Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile, perform high-frequency non-volatile memory accesses, checkpointing frequently.
 
-\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or
-programmer-defined tasks can be both non-terminating and/or
-inefficient.  If a task (or code between two checkpoints) consumes more than the fixed,
-maximum energy that the device can buffer, then the task
-will never be able to complete using buffered energy only.  Such a task is
-non-terminating, {\em prevents forward progress}, and makes the program
-deadlock. 
-%
-If the task consumes far less energy than a device can buffer, the system
-may operate {\em inefficiently}, saving the program state more often than needed.
-%
-Avoiding excessively costly, non-terminating, tasks and short, high-overhead,
-tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.  
-Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse
-because a large task on one device may be a relatively short task
-on a device with a larger buffer.
-%
-\textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which
-merges multiple tasks to amortize the overhead when tasks are too small, and
-{\em task downscaling}, which breaks up a task when it is too large to complete.
+\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or programmer-defined tasks can be both non-terminating and/or inefficient.
+If a task (or code between two checkpoints) consumes more than the fixed, maximum energy that the device can buffer, then the task will never be able to complete using buffered energy only.
+Such a task is non-terminating, {\em prevents forward progress}, and makes the program deadlock.
+If the task consumes far less energy than a device can buffer, the system may operate {\em inefficiently}, saving the program state more often than needed.
+Avoiding excessively costly, non-terminating, tasks and short, high-overhead, tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.
+Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse because a large task on one device may be a relatively short task on a device with a larger buffer.
+\textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which merges multiple tasks to amortize the overhead when tasks are too small, and {\em task downscaling}, which breaks up a task when it is too large to complete.

### 2018-08-05 paper/asplos19/evaluation.tex
partial commit

@@
-% \begin{wrapfigure}{t!}{0.5\textwidth}
 \begin{figure}
-	%\begin{figure}
 	\centering
@@
 % \end{wrapfigure}
-
-Our evaluation quantitavely demonstrates that \sys (i) \emph{reduces the memory
-protection overhead} by coalescing, (ii) \emph{improves execution speed} for
-conservative static task decompostions compared to state-of-the-art runtimes; %
-and (iii) \emph{provides computation progress} at the sub-task granularity,
-preventing non-termination at adverse energy conditions where state-of-the-art
-runtimes fail. We measured the execution time of the benchmarks implemented
-using \sys with each variant of the two proposed adaptation mechanisms:
-\emph{task coalescing} and \emph{task downsizing}. We compared \sys's best
-configurations against a state-of-the art system, Alpaca~\cite{alpaca}. We have
-also broken down \sys's overhead to explain the source of \sys's improved
-performance.
-
+%
+Building on top of the existing work, we have selected the most efficient run-time library~\cite{alpaca} and compared the performance of \sys against it.
+Our evaluation quantitatively demonstrates that \sys (i) \emph{reduces the memory
+protection overhead} by coalescing, (ii) \emph{improves execution speed}, in most cases; %
+and (iii) by mean of task-downscaling \sys is able to progress where static system stuck in task non-termination loop. 
+We have also broken down \sys's overhead to explain the source of \sys's improved performance.
+%
 \subsection{Characterization of Overhead}
 \label{sec:coala_overhead}
-
-%\indent \textbf{Overhead Reduced by Coalescing.}
 
@@
 
-For each coalescing strategy from Section~\ref{sec:task_adaptation} (EO, HG,
-and WHG) and for a baseline without coalescing (NC), we have measured the time
+For each coalescing strategy from Section~\ref{sec:task_adaptation} (EO, EG,
+and WEG) and for a baseline without coalescing (NC), we have measured the time
 spent on (i) executing useful task code, including \sys's overhead on access to
@@
 %
-However, any gain from a reduced re-execution penalty is cancelled out by
+However, any gain from a reduced re-execution penalty is canceled out by
 the increased commit overhead that is incurred at the end of each static
@@
 %
-This net overhead reduction is greatest in HG and WHG strategies compared
+This net overhead reduction is greatest in EG and WEG strategies compared
 to the EO strategy. We attribute this discrepancy to EO's slow adjustment
-of the target task size without regard to execution history.
+of the target task size without regard to the energy conditions.
 %
-In the subsequent experiments, we focus on the better-performing HG and WHG
+In the subsequent experiments, we focus on the better-performing EG and WEG
 strategies.
-
+%
 \begin{figure}
@@
 	\caption{\sys kernel overhead: a breakdown of time spent in
-the application and in the runtime (left) and a breakdown of protected memory
+the application and in the runtime (top) and a breakdown of protected memory
 accesses into three types with different costs. NC: no coalescing, HG:
@@
 \end{figure}
-
-%\textbf{\sys Kernel Overhead Breakdown.}
-
+%
 \subsubsection{\sys Kernel Overhead Breakdown}
-
-Figure~\ref{fig:coalEfficiency} (left) breaks down the time spent on executing
+%
+Figure~\ref{fig:coalEfficiency} (top) breaks down the time spent on executing
 useful task code and committing further into time spent by the application and
@@
 %
-For both coalescing strategies (HG and WHG) accesses to protected variables
+For both coalescing strategies (EG and WEG) accesses to protected variables
 constitute about 30\% of the runtime overhead. Dynamic address translation
 necessary on each protected access is the most critical bottleneck for \sys.
-
-%\textbf{Protected Memory Accesses Breakdown.}
-
+%
 \subsubsection{Protected Memory Accesses Breakdown}
-
-Figure~\ref{fig:coalEfficiency} (right) breaks down protected memory accesses
+%
+Figure~\ref{fig:coalEfficiency} (bottom) breaks down protected memory accesses
 into three categories.
@@
 %
-The search for the accessed page in the \texttt{working} buffer,
-will return immediately when the page is the most recently accessed page.
-%
 Only \textbf{cuckoo}, \textbf{dijkstra}, and \textbf{fft} have non-negligible
@@
 program state is protected, the higher the rate of page swaps.
-
+%
 \subsection{Execution time}
 \label{sec:result_coalescing}
-
+%
 Having shown in Section~\ref{sec:coala_overhead} that coalescing reduces
@@
 execution time. We first investigate difference variants of \sys and
-then compare the best variant to Alpaca.
-
-%\textbf{\sys Speedup with Coalescing.}
-
+then compare the best variant to Alpaca~\cite{alpaca}.
+%
 \subsubsection{\sys Speedup with Coalescing}
@@
 Figure~\ref{fig:coalescing} shows the \sys run time with two coalescing
-strategies (HG, WHG) normalized to the run time without coalescing (NC).
+strategies (EG, WEG) normalized to the run time without coalescing (NC).
 %
@@
 %
-The results in Figure~\ref{fig:coalescing} (left) show that all benchmarks
+The results in Figure~\ref{fig:coalescing} show that all benchmarks
 complete faster with coalescing than without coalescing: from 25\% faster
@@
 %
-In some cases (\textbf{bc}, \textbf{cuckoo}, \textbf{sort}) the history-guided
+In some cases (\textbf{bc}, \textbf{cuckoo}, \textbf{sort}) the Weight Energy-guided
 feature in the coalescing algorithm is counter-productive.
@@
 In \textbf{fft}, tasks are not uniform, and accounting for task energies is
-beneficial. In fact, the lack of task energy awareness is detrimental: with HG
+beneficial. In fact, the lack of task energy awareness is detrimental: with EG
 \textbf{fft} runs slower than without any coalescing (NC).
@@
 eliminating many unnecessary commits.
-
+%
 \begin{figure}
 	\includegraphics[width=\columnwidth]{figures/coalStrategies}%
-    \caption{\sys's coalescing performance. Left: Application execution time
-with coalescing strategies (HG and WHG) normalized to the execution time
+    \caption{\sys's coalescing performance.  Application execution time
+with coalescing strategies (EG and WEG) normalized to the execution time
 without coalescing.}
@@
 \subsubsection{Benefits of Adaptive Tasks}
-
+%
 We now compare \sys's performance to Alpaca~\cite{alpaca}---a
@@
 	\centering
-	\includegraphics[width=0.49\columnwidth]{figures/page_exec-time}
-	\includegraphics[width=0.49\columnwidth]{figures/pagePulls}
+	\includegraphics[width=\columnwidth]{figures/page_exec-time}
+	\includegraphics[width=\columnwidth]{figures/pagePulls}
     \caption{Effect of page size on execution time (left) and on the number of

### 2018-08-05 paper/asplos19/systemdescription.tex
revisions to system description (complete)

--- 2018-08-04 paper/asplos19/systemdescription.tex
@@
 	\includegraphics[width=\columnwidth]{figures/overview.pdf}
-	\caption{System's top-level view. \sys is composed of two core components: \emph{Adaptive Task Scheduler} and \emph{Virtual Memory Manager}.}
+	\caption{\sys top-level view.
+\sys core innovation is \emph{Adaptive Task Scheduler}.
+The scheduler to work requires \emph{Virtual Memory Manager}, another \sys's innovation.
+\note{Replace the figure with a better block diagram}}
 	\label{fig:system_overview}
@@
 \sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
-\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable} and {\em efficient}.
+\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable}, {\em efficient} and {\em portable}.
 \sys accomplishes this goal with a constellation of a new programming model and runtime software system support, that supports dynamically adaptive task-based execution.
@@
 
-\textbf{Programming and Execution Model.}  To use \sys, a programmer must convert a plain C code into tasks by encapsulating the code in a top-level set of functions, sequences the control-flow between these tasks, and annotates memory accesses that manipulate global data.
-The programmer compiles their task-based code, and links to \sys runtime,
-producing a \sys-enabled binary. The runtime library relies on \sys's novel
-{\em virtual memory manager} and its {\em adaptive task scheduler} to adapt its execution with the energy conditions.
+\textbf{Programming and Execution Model.}  To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate global data.
+The programmer compiles their task-based code, and links to \sys runtime, producing a \sys-enabled binary.
+The runtime library relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
+The scheduler itself required specialized memory management, which we denote as {\em virtual memory manager}.
 
-\textbf{Adaptive Task Scheduler(ATS).} 
-\sys's adaptive task scheduler (ATS) makes \emph{energy-aware} scheduling decisions to group tasks together or split a task.
+\textbf{Adaptive Task Scheduler.} \sys's adaptive task scheduler (ATS) makes \emph{energy-aware} scheduling decisions to group tasks together or split a task.
 By coalescing tasks \sys amortizes commit and transition costs, and by a task splitting, after repeated failures on a single task, it avoids the task non-termination problem.
-The scheduler uses its recent execution history---no hardware dependency---as a metric to estimate energy availability and eventually to decide on the dynamic task size.
+The scheduler uses only its recent execution history---i.e.
+it is hardware independent---as a metric to estimate energy availability and eventually to decide on the dynamic task size.
+Section~\ref{sec:task_adaptation} provides details on the ATS.
 
-\textbf{Virtual Memory Manager (VMM).} \sys is able to efficiently coalesce
-tasks because of its efficient virtual memory manager (VMM).
+\textbf{Virtual Memory Manager.} \sys is able to coalesce tasks \emph{only because of} its virtual memory manager (VMM).
 the VMM privatizes a group of memory pages demanded by a coalesced task, and limits an application access only to the fast volatile memory.
 It persists all pages modifications in non-volatile memory on a coalesced task transition.
-Section~\ref{sec:memory_virtulaization} provides details of VMM.
+Section~\ref{sec:memory_virtulaization} provides details on the VMM.

### 2018-08-05 paper/asplos19/background.tex
section 2

@@
 
-\textbf{What Hardware does \sys Target.} \sys is designed for the demands of existing and future \emph{tiny embedded energy-harvesting platforms} based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
+\textbf{Target Hardware of \sys.} \sys is designed for the demands of existing and future \emph{tiny embedded energy-harvesting platforms} based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
 \sys targets a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM.
@@
 
-\textbf{Frequent Non-volatile Memory Access.}  \note{I want to emphasize again that we do not evaluate NV memory accesses} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.
+\textbf{Direct Non-volatile Memory Access.}  \note{I want to emphasize again that we do not evaluate NV memory accesses} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.
 Checkpointing systems copy all volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at every checkpoint.
@@
 Ratchet~\cite{ratchet} and Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile, perform high-frequency non-volatile memory accesses, checkpointing frequently.
+On the other hand, \sys accesses non-volatile memory by using only DMA and without CPU intervention: \sys moves a batch of variables from non-volatile memory into volatile memory and programs are only allowed to access individual variables in volatile memory directly.
 

### 2018-08-05 paper/asplos19/evaluation.tex
revise evaluation

@@
 The overhead incurred by each coalescing strategy is broken down in
-Figure~\ref{fig:overallOverheadBreakdown}. Without coalescing enabled (NC), the
+Fig.~\ref{fig:overallOverheadBreakdown}. Without coalescing enabled (NC), the
 re-execution penalty is smallest, because the amount of work that can happen
@@
 %
-Figure~\ref{fig:coalEfficiency} (top) breaks down the time spent on executing
+Fig.~\ref{fig:coalEfficiency} (top) breaks down the time spent on executing
 useful task code and committing further into time spent by the application and
@@
 Consistent with the previous results in
-Figure~\ref{fig:overallOverheadBreakdown}, commit overhead is highest when
+Fig.~\ref{fig:overallOverheadBreakdown}, commit overhead is highest when
 coalescing is not used to reduce the number of commits.
@@
 %
-Figure~\ref{fig:coalEfficiency} (bottom) breaks down protected memory accesses
+Fig.~\ref{fig:coalEfficiency} (bottom) breaks down protected memory accesses
 into three categories.
@@
 
-Figure~\ref{fig:coalescing} shows the \sys run time with two coalescing
+Fig.~\ref{fig:coalescing} shows the \sys run time with two coalescing
 strategies (EG, WEG) normalized to the run time without coalescing (NC).
@@
 %
-The results in Figure~\ref{fig:coalescing} show that all benchmarks
+The results in Fig.~\ref{fig:coalescing} show that all benchmarks
 complete faster with coalescing than without coalescing: from 25\% faster
@@
 \emph{non-adaptive} task-based system with tasks fixed at compile time.
-Figure~\ref{fig:runtime} (right) shows the average execution time of each
+Fig.~\ref{fig:runtime} shows the average execution time of each
 application for \sys and Alpaca, normalized to the latter. The results show
 that \sys provides a performance benefit compared to Alpaca for most
-applications, with the exception of \textbf{ar} and \textbf{cuckoo} at some
-distances.  The speedup is greatest for applications with repeated write after
+applications. The speedup is greatest for applications with repeated write after
 read dependencies throughout their code, particularly involving arrays
-(\textbf{dijkstra}, \textbf{fft} and \textbf{sort}). \sys's virtualized memory
+(\textit{dijkstra}, \textit{fft} and \textit{sort}). \sys's VMM
 successfully amortizes the overhead of protecting memory that is accessed in
@@
 virtualization that causes its performance to be comparable to (or worse than)
-Alpaca (\textbf{ar}, \textbf{cuckoo}).
+Alpaca (\textit{ar}, \textit{cuckoo}), on average \sys 26\% faster than Alpaca.
 
-The \textbf{fft} benchmark did not complete at distances larger than
+The \textit{fft} benchmark did not complete at distances larger than
 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
@@
 the static tasks.} with neither Alpaca, nor \sys \emph{without the downsizing
-mechanism}. This is marked with $\infty$ signs in Figure~\ref{fig:coalescing}
-(left).  This non-terminating condition cannot be avoided in Alpaca, due to the
+mechanism}. This is marked with $\infty$ signs in Fig.~\ref{fig:coalescing}
+. This non-terminating condition cannot be avoided in Alpaca, due to the
 fixed, non-adaptive nature of its tasks.  In contrast, our next experiment
@@
 In a separate experiment, we have enabled task downsizing in \sys and measured
-the execution time for \textbf{fft} benchmark at 60\,cm from the signal
-generator antenna---even \emph{farther} than the distance at which this
-benchmark failed to terminate in the previous experiment. \sys executed the
-benchmark successfully without (NC) and with coalescing (WHG). The application
+the execution time for \textit{fft} benchmark at 60\,cm from the signal
+generator antenna. \sys executed the
+benchmark successfully without (NC) and with coalescing (WEG). The application
 took longer to complete than at a shorter distance of 15\,cm in the previous
@@
 distance. The variant without coalescing (NC) slowed down by 2.16$\times$ and the
-variant with WHG coalescing slowed down by 2.31$\times$ slower, but at both
-distances, WHG completed faster than NC.
+variant with WEG coalescing slowed down by 2.31$\times$ slower, but at both
+distances, WEG completed faster than NC.
 %
@@
 experiment on a continuously-powered evaluation board described in
-Section~\ref{sec:methodology}.  Figure~\ref{fig:page_size} quantifies the
+Section~\ref{sec:methodology}.  Fig.~\ref{fig:page_size} quantifies the
 effects of page size.
-
-%\textbf{Effect of Page Size on Runtime.}
-
+%
 \begin{figure}
@@
 	\includegraphics[width=\columnwidth]{figures/pagePulls}
-    \caption{Effect of page size on execution time (left) and on the number of
-page faults (right). The execution time is normalized to the lowest in each set
+    \caption{Effect of page size on execution time (top) and on the number of
+page faults (bottom). The execution time is normalized to the lowest in each set
 of page sizes.}
@@
 \end{figure}
-
+%
 \subsubsection{Effect of Page Size on Runtime}
-
-Figure~\ref{fig:page_size} (left) shows the execution time as a function
+%
+Fig.~\ref{fig:page_size} (top) shows the execution time as a function
 of page size, normalized to the lowest per-application performance among the
 set of page sizes.
-%
-With some exceptions, a ``bathtub curve'' can be observed in the data.  For
-some applications one of the two edges of the bathtub shape is missing, because
-we did not evaluate a sufficiently wide range of page sizes.
-%
-The right edge of the bathtub is a consequence of the time spent copying
-unnecessary data along with data of interest, because copies are at page
-granularity.
-%
-The left edge of the bathtub is due to the overhead of swapping pages in
-upon access.
-%
-The swap-in overhead is proportional to the number of page faults.
-%
-Page faults are more likely with smaller pages, as visible in the plot on the
-right in Figure~\ref{fig:page_size}.
-%
 The data suggest that there is a page size, distinct from the largest or
@@
 128\,B pages perform well for most of them.
-
-%\textbf{Effect of Page Size on Page Faults.}
-
+%
 \subsubsection{Effect of Page Size on Page Faults}
-
-Figure~\ref{fig:page_size} (right) shows the number of page faults as a
+%
+Fig.~\ref{fig:page_size} (bottom) shows the number of page faults as a
 function of page size.
@@
 no page faults occur. Without any page faults to begin with, increasing the
-page size has no effect.
-
-%\subsection{\sys Memory Footprint?}
-%\label{sec:results_program_characterization}
-%
-%\begin{table}
-%	\begin{tabular}{| c | c | c | c | c |}
-%		\hline
-%		\multirow{2}{*}{Application} & \multicolumn{2}{ c |}{Memory footprint} & \multirow{2}{*}{No. tasks} & \multirow{2}{*}{SLOC} \\
-%		\cline{2-3}
-%		{} & \sys & Alpaca & {} & {} \\
-%		\hline\hline
-%				\textbf{ar} & --- & --- & --- & ---\\
-%		\hline
-%				\textbf{bc} & --- & --- & --- & ---\\
-%		\hline
-%				\textbf{cuckoo} & --- & --- & --- & ---\\
-%		\hline
-%				\textbf{dijkstra} & --- & --- & --- & ---\\
-%		\hline
-%				\textbf{fft} & --- & --- & --- & ---\\
-%		\hline
-%				\textbf{sort} & --- & --- & --- & ---\\
-%		\hline
-%	\end{tabular}
-%		\caption{Comparison between \sys and Alpaca benchmarks.}
-%		\label{table:compiler_result}\vspace{-0.5cm}
-%\end{table}
-
-%\begin{table}[t]
-%	\centering
-%	\renewcommand{\tabcolsep}{1pt}
-%	\begin{tabular}{|l|cc|cc|cc|cc|c|}
-%		\hline
-%		{} & \multicolumn{2}{c|}{{\bf Prot. bytes}} & \multicolumn{2}{c|}{{\bf \# Tasks}} & \multicolumn{2}{c|}{{\bf \# Prot. acc.}} & \multicolumn{2}{c|}{\bf SLOC} & {\bf Comp.} \\
-%		App & Man. & Comp. & Man. & Comp. & Man. & Comp. & \multicolumn{1}{l}{\sys} & \multicolumn{1}{r|}{Chain~\cite{chain}} & {\bf time} \\
-%		\hline\hline
-%		bc & 22 & 22 & 10 & 15 & 81 & 93 & 351 &588 & 3\\
-%		cem & 3492 & 3242 & 12 & 9 & 92 & 123 & 388 &721 & 2\\
-%		cuckoo & 282 & 288 & 15 & 6 & 90 & 76 & 483 &762 & 6\\
-%		rsa & 332 & 250 & 20 & 27 & 130 & 296 & 887 &1233 & 86\\
-%		ar & 166 & 218 & 11 & 6 & 112 & 333 & 483 &762 & 34\\
-%		sort & 104 & 104 & 4 & 2 & 70 & 23 & 180 & 287 & $<$1\\
-%		dft$^\dagger$ & --- & --- & --- & --- & --- & --- & 222 & 293 & ---\\
-%		%dd &  &  &  &  &  &  &  & 287 &  \\
-%		\hline
-%	\end{tabular}
-%		\caption{Comparison between \sys and Alpaca benchmarks.}
-%		\label{table:compiler_result}\vspace{-0.5cm}
-%\end{table}
+page size result only in overhead.

### 2018-08-05 paper/asplos19/evaluation.tex
minor modification

@@
 
-In a separate experiment, we have enabled task downsizing in \sys and measured
+\textcolor{red}{In a separate experiment, we have enabled task downsizing in \sys and measured
 the execution time for \textit{fft} benchmark at 60\,cm from the signal
@@
 variant with WEG coalescing slowed down by 2.31$\times$ slower, but at both
-distances, WEG completed faster than NC.
+distances, WEG completed faster than NC.}
 %

### 2018-08-05 paper/asplos19/systemdescription.tex
update the system overview fig

@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/overview.pdf}
+	\includegraphics[width=\columnwidth]{figures/system_overview.pdf}
 	\caption{\sys top-level view.
-\sys core innovation is \emph{Adaptive Task Scheduler}.
-The scheduler to work requires \emph{Virtual Memory Manager}, another \sys's innovation.
-\note{Replace the figure with a better block diagram}}
+\sys core innovation are \emph{Adaptive Task Scheduler (TA)} and \emph{Virtual Memory Manager (VMM)}.
+(PM) stands for Phyical Memory, (App) for applicaiton, and (Tx) for Task.
 	\label{fig:system_overview}

### 2018-08-06 paper/asplos19/systemdescription.tex
revisions to system description (after Amjad's new figure) - text compression

@@
 	\caption{\sys top-level view.
-\sys core innovation are \emph{Adaptive Task Scheduler (TA)} and \emph{Virtual Memory Manager (VMM)}.
-(PM) stands for Phyical Memory, (App) for applicaiton, and (Tx) for Task.
+\sys core innovation are \emph{Adaptive Task Scheduler (TA)} and \emph{Virtual Memory Manager (VMM)}; \emph{PM}: phyical memory, \emph{App}: applicaiton, \emph{Tx}: task.}
 	\label{fig:system_overview}
 \end{figure}
-%
+
 \sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
-\sys addresses the challenges outlined in Section~\ref{sec:background} to make task-based intermittent programs {\em programmable}, {\em efficient} and {\em portable}.
-\sys accomplishes this goal with a constellation of a new programming model and runtime software system support, that supports dynamically adaptive task-based execution.
+\sys addresses the challenges given in Section~\ref{sec:background} making task-based intermittent programs {\em programmable}, {\em efficient} and {\em portable}.
+\sys accomplishes this goal with a new programming model and runtime software system, that supports dynamically adaptive task-based execution.
 Fig.~\ref{fig:system_overview} shows an overview of \sys.
@@
 \textbf{Programming and Execution Model.}  To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate global data.
-The programmer compiles their task-based code, and links to \sys runtime, producing a \sys-enabled binary.
-The runtime library relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
-The scheduler itself required specialized memory management, which we denote as {\em virtual memory manager}.
+The programmer compiles the task-based code, and links to \sys runtime, producing a \sys-enabled binary.
+The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
+The scheduler itself requires custom memory management, denoted as {\em virtual memory manager}.
 
 \textbf{Adaptive Task Scheduler.} \sys's adaptive task scheduler (ATS) makes \emph{energy-aware} scheduling decisions to group tasks together or split a task.
-By coalescing tasks \sys amortizes commit and transition costs, and by a task splitting, after repeated failures on a single task, it avoids the task non-termination problem.
+By coalescing tasks \sys \emph{amortizes commit and transition costs}, and by a task splitting, after repeated failures on a single task, it \emph{avoids the task non-termination problem}.
 The scheduler uses only its recent execution history---i.e.
 it is hardware independent---as a metric to estimate energy availability and eventually to decide on the dynamic task size.
-Section~\ref{sec:task_adaptation} provides details on the ATS.
+Section~\ref{sec:task_adaptation} outlines ATS.
 
-\textbf{Virtual Memory Manager.} \sys is able to coalesce tasks \emph{only because of} its virtual memory manager (VMM).
-the VMM privatizes a group of memory pages demanded by a coalesced task, and limits an application access only to the fast volatile memory.
-It persists all pages modifications in non-volatile memory on a coalesced task transition.
-Section~\ref{sec:memory_virtulaization} provides details on the VMM.
+\textbf{Virtual Memory Manager.} \sys can coalesce tasks \emph{only} due to its virtual memory manager (VMM).
+VMM privatizes a group of memory pages demanded by a coalesced task and limits an application access only to the fast volatile memory.
+VMM keeps all pages modifications in non-volatile memory on a coalesced task transition.
+Section~\ref{sec:memory_virtulaization} outlines VMM.

### 2018-08-06 paper/asplos19/coalescing.tex
change the style of algo 1
revisions to coalescing (not complete)

@@
-\sys's design includes a novel task scheduler that uses efficient,
-energy-aware task coalescing to amortize static task overheads, and uses a timer-based
-task-splitting mechanism to avoid non-termination of tasks too long
-for a device's energy buffer. 
-% 
+\sys's design includes a novel task scheduler.
+It uses efficient, energy-aware task coalescing to amortize static task overheads, and a timer-based task-splitting mechanism to avoid non-termination of tasks too long for a device's energy buffer.
+
 \subsection{Task Coalescing}
 \label{sec:task_coalescing}
-% 
-When a device's buffered energy is sufficient to run multiple consecutive tasks
-without a power failure, committing state after each task is an unnecessary
-overhead. \sys eliminates this overhead by coalescing a sequence of tasks and deferring the commit
+
+When a device's buffered energy is sufficient to run multiple consecutive tasks without a power failure, committing state after each task is an unnecessary overhead.
+\sys eliminates this overhead by coalescing a sequence of tasks and deferring the commit
 operations for all tasks to the end of the sequence.
-In general, committing involves moving state manipulated by a task
-into its permanent location in non-volatile memory.  In particular, \sys's
-commit procedure copies dirty pages of memory that a task updated between fast,
-volatile working memory and slower, non-volatile main memory; we defer the
-details of paging privatization and commit to Section~\ref{sec:memory_virtulaization}.
+In general, committing involves moving state manipulated by a task into its permanent location in non-volatile memory.
+In particular, \sys's commit procedure copies dirty pages of memory that a task updated between fast, volatile working memory and slower, non-volatile main memory; we defer the details of paging privatization and commit to Section~\ref{sec:memory_virtulaization}.
 
-Coalescing tasks brings a double benefit. First, coalescing reduces the total number of relatively
-high-latency, high-energy non-volatile memory accesses because tasks primarily
-access variables in relatively low-latency, low-energy volatile memory.
-Second, coalescing eliminates the unnecessary instructions that implement task
-commit, which our measurements in Section~\ref{sec:evaluation} show constitute
-a substantial run-time overhead.
+Coalescing has a double benefit.
+First, coalescing reduces the total number of relatively high-latency, high-energy non-volatile memory accesses because tasks primarily access variables in relatively low-latency, low-energy volatile memory.
+Second, coalescing eliminates the unnecessary instructions that implement task commit, which constitute a substantial run-time overhead, as our measurements in Section~\ref{sec:evaluation} show.
 
-An effective colescing strategy must be {\em aggressive} enough, attempting to coalesce a
-large number of tasks to amortize a large amount of commit overhead. However, it must also be 
-{\em conservative} enough, attempting to coalesce only as many tasks as will execute
-to completion given a certain energy conditions, reducing the risk or long coalesced task re-execution penalty.
+An effective coalescing strategy must be {\em aggressive} enough, attempting to coalesce a large number of tasks to amortize commit overhead.
+However, it must also be  {\em conservative} enough, attempting to coalesce only as many tasks as will execute to completion given a certain energy conditions, reducing the risk or long coalesced task re-execution penalty.
 
-% An effective coalescing strategy moderates the risk of a power failure during a
-% long coalesced task and capitalizes on the benefits of deferred commits by
-% adapting the target size of the dynamic task to the energy available at
-% runtime. 
-%
-\begin{figure}{t}
-\fbox{\parbox{\linewidth}{
-	\captionof{algorithm}{Coalescing}
+\begin{algorithm}[t]
+	\caption{Coalescing}
 	\label{algo:genCoalescing}
 	\scriptsize
-	%\small
-	%
 	\begin{algorithmic}[1]
         \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$z$}
-        \State $H \gets 0$ 
+        \State $H \gets 0$ \Comment{\note{add comments for some lines to get rid of white spaces}}
         \While{$\texttt{true}$}       
@@
         \EndWhile
-	\end{algorithmic}}}
-% \end{wrapfigure}
-\end{figure}
+	\end{algorithmic}
+\end{algorithm}
 % 
 \subsection{Task Coalescing Strategies}
-Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing strategy. In the algorithm, $B$ (target budget) is total number of static tasks
-that \sys will next attempt to coalesce. $T_\text{i}$ is the $i$th task
+
+Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing strategy.
+In the algorithm, $B$ (target budget) is total number of static tasks that \sys will next attempt to coalesce.
+$T_\text{i}$ is the $i$th task
 executed since the last power failure. $H$ (history) is the total number of

### 2018-08-06 paper/asplos19/coalescing.tex
revisions to coalescing (not complete)

@@
 \end{algorithm}
-% 
+
 \subsection{Task Coalescing Strategies}
+\label{subsec:coalescingStrategies}
 
@@
 In the algorithm, $B$ (target budget) is total number of static tasks that \sys will next attempt to coalesce.
-$T_\text{i}$ is the $i$th task
-executed since the last power failure. $H$ (history) is the total number of
-tasks between the last executed task and the power failure that preceding it.
-$W_\text{i}$ is the {\em weight} of a task $T_\text{i}$.  A task's weight $W$ is an
-arbitrary quantity associated with the task that represents its cost in time or
-energy to execute; different coalescing strategies may apply different weights
-to a task.  For example, a uniform weighting scheme sets the weight of each
-task to one, making them all effectively equal in cost to the coalescing
-algorithm.  A different policy might use a weighting scheme that sets a task's
-weight equal to its run time relative to the task that executes for the
-shortest duration, which has unity weight.
+$T_\text{i}$ is the $i$th task executed since the last power failure.
+History, $H$, is the total number of tasks between the last executed task and the power failure that preceding it.
+$W_\text{i}$ is the {\em weight} of a task $T_\text{i}$.
+A task's weight $W$ is an arbitrary quantity associated with the task that represents its cost in time or energy to execute; different coalescing strategies may apply different weights to a task.
+For example, a uniform weighting scheme sets the weight of each task to one, making them all effectively equal in cost to the coalescing algorithm.
+A different policy might use a weighting scheme that sets a task's weight equal to its run time relative to the task that executes for the shortest duration, which has unity weight.
 
-Different coalescing strategies adhere mainly to the template in
-Algorithm~\ref{algo:genCoalescing}, varying in only a few {\em characteristic
-operations} that the algorithm leaves deliberately abstract.  A strategy's
-reboot update function, $f_\text{reboot}$, updates $B$, the target budget,
-after a reboot. A strategy's weight lookup function, $f_\text{weight}$, returns
-the a task's weight.  A strategy's commit update function, $f_\text{commit}$,
-updates $B$ after a successful commit. The following sections detail several
-possible coalescing strategies that we developed for \sys, albeit not the only possible.
-In fact, our system allows developers to design their own strategy to implement and link to Coala’s task scheduler.
-% 
-\subsubsection{Energy-Oblivious Coalescing (EO)}
+Different coalescing strategies adhere mainly to the template in Algorithm~\ref{algo:genCoalescing}, varying in only a few {\em characteristic operations} that the algorithm leaves deliberately abstract.
+A strategy's reboot update function, $f_\text{reboot}$, updates $B$, the target budget, after a reboot.
+A strategy's weight lookup function, $f_\text{weight}$, returns the a task's weight.
+A strategy's commit update function, $f_\text{commit}$, updates $B$ after a successful commit.
+The following sections detail several possible coalescing strategies that we developed for \sys, albeit not the only possible.
+In fact, \sys \emph{allows programmers to design their own strategy} to implement and link to \sys’s task scheduler.
+
+\subsubsection{Energy-Oblivious Coalescing}
 \label{subsec:energyBlind}
-% 
-A very simple {\em energy-oblivious} coalescing strategy treats all tasks as
-having unity weight and varies the size of a coalesced task linearly. In such a
-scheme, the number of tasks to coalesce, $B$, increases by a constant $x$ when
-a coalesced sequence of tasks commits, and decreases by the same constant $x$
-when a coalesced sequence of tasks fails to complete, i.e., after a power
-failure. The characteristic operations of the energy-oblivious strategy are: 
+
+A very simple {\em energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of a coalesced task linearly.
+In such a scheme, the number of tasks to coalesce, $B$, increases by a constant $x$ when a coalesced sequence of tasks commits, and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure.
+The characteristic operations of the energy-oblivious strategy are
 %
@@
 	\begin{split}
-		f_\text{reboot}(B)  		&	 = B - x \\
-		f_\text{weight}(T_\text{i}) & 	 =  1 \\
-		f_\text{commit}(B) 			& 	 = B + x
+		f_\text{reboot}(B) & = B - x, \\
+		f_\text{weight}(T_\text{i}) & =  1, \\
+		f_\text{commit}(B) & = B + x.
 	\end{split}
@@
 %
-The energy-oblivious strategy reacts slowly to the variation in the energy
-required to execute different tasks and to the variation in the effective quantum of
-energy available to the device. With the successful commit of each coalesced task sequence, the coalescing target increases.  Eventually, the target may be too high and a coalesced task
-composed of only fewer tasks will commit without interruption by a power
-failure.
-%
-The strategy then linearly decreases the coalescing target, eventually reaching a
-coalescing target that allows completion.
-%
-A key limitation of this algorithm lies in the equality of the target decrease
-in $f_\text{reboot}$ and the increase in $f_\text{commit}$.
+The energy-oblivious strategy reacts slowly to the variation in the energy required to execute different tasks and to the variation in the effective quantum of energy available to the device.
+With the successful commit of each coalesced task sequence, the coalescing target increases.
+Eventually, the target may be too high and a coalesced task composed of only fewer tasks will commit without interruption by a power failure.
+The strategy then linearly decreases the coalescing target, eventually reaching a coalescing target that allows completion.
+A key limitation of this algorithm lies in the equality of the target decrease in $f_\text{reboot}$ and the increase in $f_\text{commit}$.
 Let us assume $x=1$.
 If, after $k$ successful commits the target must decrease to its original value $B-k$ due to an energy drop, this strategy requires $k$ successive power failures.
-%
-\subsubsection{Energy-Guided Coalescing (EG)}
+
+\subsubsection{Energy-Guided Coalescing}
 \label{subsec:energyAware}
-%
+
 \begin{figure*}
@@
 \end{figure*}
-%
+
 The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions.
-It uses its recent execution history as
-mean to measure energy availability and alters its target accordingly.
+It uses its recent execution history as  mean to measure energy availability and alters its target accordingly.
 By relying on the history of execution EG eliminates the problem of frequent power failures a single coalesced task.
@@
 For instance, if the execution is interrupted after finishing a single task the history will be one ($H = 1$) and the target will be set to one also.
-While a history of $H$ suggests that there is sufficient energy to coalesce and
-successfully commit $H$ tasks (i.e., to set $B = H$).
+While a history of $H$ suggests that there is sufficient energy to coalesce and successfully commit $H$ tasks (i.e., to set $B = H$).
 EG conservatively sets $B$ equal to half of $H$ moderating the risk of a very long coalesced task.
 We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
-The EG algorithm is characterized by the following functions:
+The EG algorithm is characterized by the following functions
 %
@@
 	\begin{split}
-		 f_\text{reboot}(H) 			& = \lceil(H / 2)\rceil \\
-		 f_\text{weight}(T_\text{i}) 	& =  1 \\
-		 f_\text{commit}(B) 			& = \lceil(B / 2)\rceil
+		 f_\text{reboot}(H) & = \lceil(H / 2)\rceil,\\
+		 f_\text{weight}(T_\text{i}) & = 1, \\
+		 f_\text{commit}(B) & = \lceil(B / 2)\rceil.
 	\end{split}
@@
 %
-Fig.~\ref{fig:hg-coal} illustrates the operation of the EG.
-The figure shows two power cycles following the $i$th, the latter having an
+Fig.~\ref{fig:hg-coal} illustrates the operation of the EG. The figure shows two power cycles following the $i$th, the latter having an
 execution history $H^{(i)} = 4$.
-Based on that, $B$ is initially set to two and then decreases to one. Once the
-value of $B$ reaches one, EG is, in effect, expecting a power failure,
-justifying the conservative approach. After the power failure, EG
-uses the most recent history $H^{(i+1)} = 8$ to set $B = 4$
-and continue execution.
-%
-\subsubsection{Weighted Energy-Guided Coalescing (WEG)}
+Based on that, $B$ is initially set to two and then decreases to one.
+Once the value of $B$ reaches one, EG is, in effect, expecting a power failure, justifying the conservative approach.
+After the power failure, EG uses the most recent history $H^{(i+1)} = 8$ to set $B = 4$ and continue execution.
+
+\subsubsection{Weighted Energy-Guided Coalescing}
 \label{subsec:energyTaskAware}
-%
-The \emph{Weighted Energy-Guided} (WEG) coalescing strategy accounts for the
-different energy and time cost of a program's tasks when setting the coalescing
-target.
-%
-Each different task in the program consumes a different amount of energy to
-execute to completion.  
-%
-The EO and EG coalescing strategies assume each task has uniform energy cost:
-for these strategies, $B$ simply corresponds to a target {\em count} of tasks
-coalesced, regardless of the individual cost of each task.
-%
-However, if one task executes for ten seconds and another executes for one
-second, counting tasks misjudges the amount of {\em work} in the coalesced
-tasks.
+
+The \emph{Weighted Energy-Guided} (WEG) coalescing strategy accounts for the different energy and time cost of a program's tasks when setting the coalescing target.
+Each different task in the program consumes a different amount of energy to execute to completion.
+The EO and EG coalescing strategies assume each task has uniform energy cost: for these strategies, $B$ simply corresponds to a target {\em count} of tasks coalesced, regardless of the individual cost of each task.
+However, if one task executes for ten seconds and another executes for one second, counting tasks misjudges the amount of {\em work} in the coalesced tasks.
 
 Instead, WEG associates a non-unity weight with each task in the program.
-%
 WEG tracks the sum of the weights of tasks in a coalesced sequence of tasks.
-%
-When the sum of weights reaches $B$, the target, WEG commits the coalesced
-task.
-%
-WEG differs from EO and HG by explicitly accounting for the variation in energy
-cost of a program's tasks, eliminating the assumption that all tasks do the
-same work, which is a key limitation.
-%
-WEG has the same characteristic operations as HG, except that
-$f_\text{weight}(T_\text{i}) = W_\text{i}$.
+When the sum of weights reaches $B$, the target, WEG commits the coalesced task.
+WEG differs from EO and HG by explicitly accounting for the variation in energy cost of a program's tasks, eliminating the assumption that all tasks do the same work, which is a key limitation.
+WEG has the same characteristic operations as HG, except that $f_\text{weight}(T_\text{i}) = W_\text{i}$.
 
-The effectiveness of the WEG algorithm hinges on correctly identifying the
-weight of each task, which WEG assumes is {\em statically} available.
-%
-Profiling the time and energy cost of tasks in a program is a difficult,
-orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
-%
+The effectiveness of the WEG algorithm hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available.
+Profiling the time and energy cost of tasks in a program is a difficult, orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
 WEG could use the result of an arbitrarily sophisticated profiling procedure.
-%
-To produce a concrete result in this paper, we give WEG access to a simple
-profile of task run time collected offline using a single fixed input.
-%
-WEG stores the profile in a lookup table that maps a task's identifier to its
-weight, making the information available to \sys's scheduler at run time.
+To produce a concrete result in this paper, we give WEG access to a simple profile of task run time collected offline using a single fixed input.
+WEG stores the profile in a lookup table that maps a task's identifier to its weight, making the information available to \sys's scheduler at run time.

### 2018-08-06 paper/asplos19/coalescing.tex
Coalescing section adapted

@@
 The following sections detail several possible coalescing strategies that we developed for \sys, albeit not the only possible.
-In fact, \sys \emph{allows programmers to design their own strategy} to implement and link to \sys’s task scheduler.
+In fact, \sys \emph{allows programmers to design their own strategy} to implement and link to \sys's task scheduler.
 
@@
 	\end{split}
+	\label{eq:eo}
 \end{equation}
@@
 	\end{split}
+	\label{eq:eg}
 \end{equation}
@@
 WEG differs from EO and HG by explicitly accounting for the variation in energy cost of a program's tasks, eliminating the assumption that all tasks do the same work, which is a key limitation.
-WEG has the same characteristic operations as HG, except that $f_\text{weight}(T_\text{i}) = W_\text{i}$.
+WEG is characterized the same way as HG (equation (\ref{eq:eg})), except that $f_\text{weight}(T_\text{i}) = W_\text{i}$.
 
-The effectiveness of the WEG algorithm hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available.
+The effectiveness of the WEG hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available.
 Profiling the time and energy cost of tasks in a program is a difficult, orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}.

### 2018-08-06 paper/asplos19/paging.tex
typo
update algo 2
Paging revised (not complete)

@@
-% \textcolor{red}{[additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain.
+%\note{additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain.
 And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
-%
-\sys's {\em virtual memory manager} (VMM) abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers.
+
+\sys's {\em virtual memory manager} is a core enabler of \sys task adaptation mechanism.
+Its overview is given in Figure~\ref{figure:mem-man}.
+VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers.
 Each buffer is divided into pages.
@@
 The VMM ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
-%
-On a commit, the VMM atomically copies all ``dirty'' pages back into their location in the non-volatile memory.
-%
-SRAM has a limited size and there is an extended working memory containing {\em
-shadow pages} in the larger FRAM.  A shadow page stores the data from an
-updated pages that need to be committed back to non-volatile memory, but that
-exceeded the size of the volatile working memory.
-%
-the VMM moves pages of data efficiently using hardware accelerated Direct Memory
-Access (DMA) support, which copies a block of memory from one place to another
-as a single operation.\\
-%
+On a commit, the VMM atomically copies, through two-phase commit, all ``dirty'' pages back into their location in the non-volatile memory.
+SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM.
+A shadow page stores the data from an updated pages that need to be committed back to non-volatile memory, but that exceeded the size of the volatile working memory.
+The VMM moves pages of data efficiently using hardware accelerated DMA support.
+
 \begin{figure}
@@
     \includegraphics[width=\columnwidth]{figures/mem-man.pdf}
-    \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP},
-    and order of occurrence. Steps 2, 3, 4 and 5 are conditional. Tasks only have direct access to the
-    \texttt{working} buffer.}
+    \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP}, and its order of occurrence; steps 2, 3, 4 and 5 are conditional.
+Tasks only have direct access to the \texttt{working} buffer.}
     \label{figure:mem-man}
 \end{figure}
-%
-\textbf{System Overview.}
-Figure~\ref{figure:mem-man} gives an overview of \sys's virtual memory management system,
-whose details are provided throughout this section.
-A page has a home location in non-volatile memory which we refer to as the page's
-\texttt{\underline{private}} (underlined locations are non-volatile).  When a
-task accesses a page, the VMM swaps the page from its home location
-into the fully\hyp{}associative \texttt{working} buffer that is stored in
-volatile memory.  The \texttt{working} buffer is allocated to take up the
-volatile memory that remains after space for stack is allocated.
-%
-The VMM utilizes a two-phase commit to write dirty pages from the \texttt{working}
-buffer back to their home location in \texttt{\underline{private}}. It optimizes this process by using an indirection table to abstract the location of the \texttt{\underline{private}} buffer and
-to eliminate the need for the second page copy.
-%
-The VMM is forced to swap out a dirty page from the \texttt{working} buffer
-when a task requests a page that is not in the \texttt{working} buffer, and there is no
-free page in the working buffer (i.e., a page fault). On a page fault, the VMM
-moves the page to the direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer. \\
-%
+
+%\textbf{System Overview.} Figure~\ref{figure:mem-man} gives an overview of \sys's virtual memory management system.
+A page has a home location in non-volatile memory which we refer to as the page's \texttt{\underline{private}}.
+When a task accesses a page, the VMM swaps the page from its home location into the fully\hyp{}associative \texttt{working} buffer that is stored in volatile memory.
+The \texttt{working} buffer is allocated to take up the volatile memory that remains after space for stack is allocated.
+
+%The VMM utilizes a two-phase commit to write dirty pages from the \texttt{working} buffer back to their home location in \texttt{\underline{private}}.
+It optimizes this process by using an indirection table to abstract the location of the \texttt{\underline{private}} buffer and to eliminate the need for the second page copy.
+The VMM is forced to swap out a dirty page from the \texttt{working} buffer when a task requests a page that is not in the \texttt{working} buffer, and there is no free page in the working buffer (i.e., a page fault).
+On a page fault, the VMM moves the page to the direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer.
+
 \subsection{Address Translation and Variable Access}
-%
-A task must access protected non-volatile variables through \sys's
-restricted memory interface. The interface includes \texttt{RP}$(v)$, to read the value of variable
-$v$, and \texttt{WP}$(v)$ to assign a value to variable $v$. The
-implementation of {\tt RP} is shown in Algorithm~\ref{algo:rwar}, and {\tt
-WP}'s implementation is similar except that {\tt WP} marks the accessed page as
-dirty.
-%
-{\tt RP} and {\tt WP} operations translate a variable's physical address in
-non-volatile memory into a \emph{virtual address} in \texttt{working} buffer in
-volatile memory. In \sys, a virtual address is composed of a \emph{page tag}
-that identifies the page and a \emph{page offset} that identifies a byte.
-%
-\begin{figure}
-\fbox{\parbox{\linewidth}{
+
+A task must access protected non-volatile variables through \sys's restricted memory interface.
+The interface includes \texttt{RP}$(v)$, to read the value of variable $v$, and \texttt{WP}$(v)$ to assign a value to variable $v$.
+The implementation of {\tt RP} is shown in Algorithm~\ref{algo:rwar}, and {\tt WP}'s implementation is similar except that {\tt WP} marks the accessed page as dirty.
+{\tt RP} and {\tt WP} operations translate a variable's physical address in non-volatile memory into a \emph{virtual address} in \texttt{working} buffer in volatile memory.
+In \sys, a virtual address is composed of a \emph{page tag} that identifies the page and a \emph{page offset} that identifies a byte.
+
+\begin{algorithm}[t]
     \captionof{algorithm}{\texttt{RP}(variable $v$)}
@@
 		\State \Return \texttt{working}[$i$][$o$]  \Comment{Return from page}
-	\end{algorithmic}}}
-\end{figure}
-%
-After address translation (Figure~\ref{figure:mem-man}, Step 6), a task
-accesses the protected variable's location in the volatile \texttt{working} buffer
-(Step 7). The VMM keeps track of the page tags for
-the pages currently resident in the \texttt{working} buffer. When a task accesses a
-variable, it compares the variable's page tag to tags of the pages in {\tt
-working} (Algorithm~\ref{algo:rwar}, Line 2).
-%
-If the accessed variable's page tag is not found in the page buffer, the
-operation incurs a {\em page fault} (Line 4).
-%
-The byte is accessed in the page buffer at the index of the resident page and
-the variable's page offset (Lines 5--6).
-%
+	\end{algorithmic}
+\end{algorithm}
+
+After address translation (Figure~\ref{figure:mem-man}, Step 6), a task accesses the protected variable's location in the volatile \texttt{working} buffer (Step 7).
+The VMM keeps track of the page tags for the pages currently resident in the \texttt{working} buffer.
+When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt working} (Algorithm~\ref{algo:rwar}, Line 2).
+If the accessed variable's page tag is not found in the page buffer, the operation incurs a {\em page fault} (Line 4).
+The byte is accessed in the page buffer at the index of the resident page and the variable's page offset (Lines 5--6).
+
 \subsection{Page Faults and Page Swapping}
-%
-When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory
-manager first searched the variable's page in the \texttt{working} buffer
-(Figure~\ref{figure:mem-man}, Step 1).
+
+When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory manager first searched the variable's page in the \texttt{working} buffer (Figure~\ref{figure:mem-man}, Step 1).
 If the page is not found there, a page fault is incurred and a new page needs to
 be swapped in.
-If the \texttt{working} buffer is full, a page fault on a memory access
-requires VMM to swap out one of the pages in the \texttt{working} buffer (a
-\emph{victim page}) preserving updates made to that page.
-%
-While we opted for the simplicity of FIFO, any other replacement policy, such
-as Least Recently Used (LRU) could work in its place. \\
-%
-If a task modified any byte in the victim page (i.e., using \texttt{WP}), then
-the page is dirty and its changes has to be persisted to non-volatile memory
-(Figure~\ref{figure:mem-man}, Step 2).
-If the accessed page was previously modified and swapped out since the last
-power failure, the most recent version of the page is in {\tt
-\underline{shadow}} (Step 3), otherwise it has to be retrieved from
-{\tt \underline{private}} (Step 4).
-Finally, the page is copied to the volatile buffer (Step 5). \\
-%
+If the \texttt{working} buffer is full, a page fault on a memory access requires VMM to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
+While we opted for the simplicity of FIFO, any other replacement policy, such as Least Recently Used (LRU) could work in its place.
+If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes has to be persisted to non-volatile memory (Figure~\ref{figure:mem-man}, Step 2).
+If the accessed page was previously modified and swapped out since the last power failure, the most recent version of the page is in {\tt \underline{shadow}} (Step 3), otherwise it has to be retrieved from {\tt \underline{private}} (Step 4).
+Finally, the page is copied to the volatile buffer (Step 5).
+
 \subsection{Atomic Two-Phase Commit of Dirty Pages}
-%
-When the last task in a sequence of coalesced tasks completes, \sys must commit
-\emph{all} dirty pages in the \texttt{working} buffer and in the
-\texttt{\underline{shadow}} buffer, copying them back to their original
-locations in the main memory \texttt{\underline{private}}.\\
-To make the commit atomic, \sys commits the pages in two phases, implemented by
-Algorithm~\ref{algo:commit}.
-%
-The first phase copies dirty pages from the \texttt{working} buffer to the
-non-volatile \texttt{\underline{shadow}} buffer (Line 4). The second phase
-commits pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}
-(Line 12).  If power fails during the first phase, the whole commit is aborted,
-and the execution restarts from the most recently committed point (i.e., from the beginning of a coalesced task). If power fails in the second
-phase, the commit process safely resumes on reboot.  The second phase of commit
-depends on some run-time metadata. The \texttt{\underline{committing}} bit
-indicates that a commit is in progress and is set before the first page is
-committed (Line 9) and cleared after the last page is committed (Line 16).  The
-\texttt{\underline{shadowCount}} records the number of dirty shadow pages to
-commit and the VMM clears the counter when commit completes (Line 14).  The
-\texttt{\underline{commitIndex}} indexes the next page to be committed (Line
-11) and the VMM clears the index at the end of the phase (Line 15).
-%
-\sys's commit is efficient because it maintains an index dirty pages instead of
-iterating through all potentially dirty pages to check a dirty bit during
-commit.
-%
-\sys's implementation is also efficient because the VMM does not
-copy from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead
-swapping pointers in an indirection table that maintains the pages as a double
-buffer.  Section~\ref{sec:implementation} describes double buffering in more
-detail. \\
-%
-\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only
-ever executes using consistent protected data. During task execution,
-modifications to protected data do not affect the \texttt{\underline{private}} buffer,
-because a task reads and writes the volatile \texttt{working} buffer only, and
-modified pages are kept in \texttt{\underline{shadow}} until commit.  A power
-failure erases the contents of the \texttt{working} buffer, preventing a re-executing task
-from observing updates from a previous execution attempt.
-Clearing \texttt{\underline{shadowCount}} as part of the second phase commit
-(Line 14), ensures that all accesses to protected variables in subsequent tasks
-correctly access their original memory locations in \texttt{\underline{private}}.\\
-%
+
+When the last task in a sequence of coalesced tasks completes, \sys must commit \emph{all} dirty pages in the \texttt{working} buffer and in the \texttt{\underline{shadow}} buffer, copying them back to their original locations in the main memory \texttt{\underline{private}}.
+
+To make the commit atomic, \sys commits the pages in two phases, implemented by Algorithm~\ref{algo:commit}.
+The first phase copies dirty pages from the \texttt{working} buffer to the non-volatile \texttt{\underline{shadow}} buffer (Line 4).
+The second phase commits pages from \texttt{\underline{shadow}} to \texttt{\underline{private}} (Line 12).
+If power fails during the first phase, the whole commit is aborted, and the execution restarts from the most recently committed point (i.e., from the beginning of a coalesced task).
+If power fails in the second
+phase, the commit process safely resumes on reboot.
+The second phase of commit depends on some run-time metadata.
+The \texttt{\underline{committing}} bit indicates that a commit is in progress and is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16).
+The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to commit and the VMM clears the counter when commit completes (Line 14).
+The \texttt{\underline{commitIndex}} indexes the next page to be committed (Line 11) and the VMM clears the index at the end of the phase (Line 15).
+
+\sys's commit is efficient because it maintains an index dirty pages instead of iterating through all potentially dirty pages to check a dirty bit during commit.
+\sys's implementation is also efficient because the VMM does not copy from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead swapping pointers in an indirection table that maintains the pages as a double buffer.
+Section~\ref{sec:implementation} describes double buffering in more detail.
+
+\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent protected data.
+During task execution, modifications to protected data do not affect the \texttt{\underline{private}} buffer, because a task reads and writes the volatile \texttt{working} buffer only, and modified pages are kept in \texttt{\underline{shadow}} until commit.
+A power failure erases the contents of the \texttt{working} buffer, preventing a re-executing task from observing updates from a previous execution attempt.
+Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14), ensures that all accesses to protected variables in subsequent tasks correctly access their original memory locations in \texttt{\underline{private}}.
+
 \begin{algorithm}[t]
@@
                 \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{commitIndex}}]$
-                \State \Call{commitToStore}{\texttt{\underline{shadow}}[t]}
+                \State \Call{commitToPrivate}{\texttt{\underline{shadow}}[t]}
                 \State $\texttt{\underline{commitIndex}} \gets \texttt{\underline{commitIndex}} + 1$
@@
 \end{algorithm}
-%
+
 \subsection{Dynamic Paging of \sys}
-%
-\sys asks the programmer to use its {\tt RP} and {\tt WP} paging API methods on
-every access to a protected variable (cf. Section~\ref{sec:coala_api}).  These
-API invocations present a risk of high overhead because there is a dynamic
-check on every read and write. 
-%
-Despite the risk of per-access overhead, \sys's dynamic memory protection
-scheme brings several benefits over a static approach (i.e.~\cite{alpaca}).  
-First, the limitations of static analysis preclude some uses of pointers due to potential pointer
-aliasing. For example, in the presence of arbitrary pointer operation, a
-function call using a function pointer, or an interrupt within a task, the
-system cannot statically analyze the memory behavior.   Second, a static approach
-cannot handle task coalescing, because a protected variable's lifetime (i.e.,
-from first use to commit) is unknown at compile time. \sys's dynamic,
-per-access instrumentation supports arbitrary use of pointers and enables task
-coalescing.
+
+\sys asks the programmer to use its {\tt RP} and {\tt WP} paging API methods on every access to a protected variable (cf.
+Section~\ref{sec:coala_api}).
+These API invocations present a risk of high overhead because there is a dynamic check on every read and write.
+Despite the risk of per-access overhead, \sys's dynamic memory protection scheme brings several benefits over a static approach (i.e.~\cite{alpaca}).
+First, the limitations of static analysis preclude some uses of pointers due to potential pointer aliasing.
+For example, in the presence of arbitrary pointer operation, a function call using a function pointer, or an interrupt within a task, the system cannot statically analyze the memory behavior.
+Second, a static approach cannot handle task coalescing, because a protected variable's lifetime, i.e.,from first use to commit, is unknown at compile time.
+\sys's dynamic, per-access instrumentation supports arbitrary use of pointers and enables task coalescing.

### 2018-08-06 paper/asplos19/relatedwork.tex
related work

@@
 surveys~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
-As computing power consumption continues to decrease, a new wave of
-systems that are powered by radio waves (RF) is
+As power requirements continue to decrease, a new wave of
+systems powered by radio waves (RF) is
 emerging~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014,wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
-RF-powered devices are an important category of intermittent computing devices
-that \sys targets.  The emergence of intermittent computing hardware platforms
+%RF-powered devices are an important category of intermittent computing devices
+%that \sys targets.  
+The emergence of such hardware platforms
 has led to the development of instrumentation, debugging, and prototyping tools
@@
 size of volatile memory increases.  QuickRecall, Clank and Ratchet reduce
-the copying overhead by allocating the stack in non-volatile memory, which
-requires more time and energy to access than volatile memory (cf.
-Section~\ref{sec:cost_task-based}) and is not a viable option for off-chip non-volatile
-memory. In addition, unlike \sys, QuickRecall and Clank require custom hardware.  \\
+the copying overhead by allocating the stack in non-volatile memory---requires more time and energy to access than volatile memory (cf.
+Section~\ref{sec:cost_task-based}) and not a viable option for off-chip non-volatile
+memory. In addition, QuickRecall and Clank require custom hardware.  \\
 %
@@
 %
-\sys relates closely to task-based models because \sys also relies on
+\sys also relies on
 statically defined tasks to avoid checkpointing volatile state. 
@@
 %
-\textbf{Task Decomposition.} In contrast to \sys's dynamic construction of
+\textbf{Task Decomposition.} In contrast to \sys's construction of
 coallesced tasks at runtime, prior work has proposed to optimize task size

### 2018-08-06 paper/asplos19/intro.tex
intro

@@
 
-\item[C1] \emph{Overcoming Task Transition Overhead:} given unpredictable incoming energy, how to save computation progress, i.e.
-saving state at task transitions, as rare as possible?
+\item[C1] \emph{Overcoming Task Transition Overhead:} given unpredictable incoming energy, how to save computation state at task transitions as rare as possible?
 \sys tries to minimize task transition overhead by estimating energy conditions at run-time using \emph{recent execution history} as a metric.
@@
 Individual variables tracking, however, slows down a system dramatically.
-Therefore, \sys protects global variables in a batch.
-Moreover, it optimizes data transfers by using Direct Memory Access (DMA).
+Therefore, \sys keeps memory consistent through \emph{memory virtualization} optimized for bulk accesses to task-shared data with high locality.
+
+%Therefore, \sys protects global variables in a batch. Moreover, it optimizes data transfers by using Direct Memory Access (DMA).
 

### 2018-08-06 paper/asplos19/systemdescription.tex
Section 3 updated (with new figure as well)

@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/system_overview.pdf}
-	\caption{\sys top-level view.
-\sys core innovation are \emph{Adaptive Task Scheduler (TA)} and \emph{Virtual Memory Manager (VMM)}; \emph{PM}: phyical memory, \emph{App}: applicaiton, \emph{Tx}: task.}
+	\includegraphics[width=\columnwidth]{figures/system-overview.pdf}
+	\caption{\sys top-level view. ATS: adaptive task scheduler, VMM: virtual memory manager, PM: physical memory.}
 	\label{fig:system_overview}
@@
 
-\sys is a new programming and execution model for intermittent computing on energy-harvesting devices.
-\sys addresses the challenges given in Section~\ref{sec:background} making task-based intermittent programs {\em programmable}, {\em efficient} and {\em portable}.
-\sys accomplishes this goal with a new programming model and runtime software system, that supports dynamically adaptive task-based execution.
-Fig.~\ref{fig:system_overview} shows an overview of \sys.
+\sys is a new programming and execution model for intermittent operation on energy-harvesting devices.
+It addresses the challenges given in Section~\ref{sec:background} making task-based intermittent applications {\em programmable}, {\em efficient} and {\em portable}, introducing a new programming model and runtime software system that supports adaptive task-based execution.
+Figure~\ref{fig:system_overview} shows an overview of \sys.
 
-\textbf{Programming and Execution Model.}  To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate global data.
-The programmer compiles the task-based code, and links to \sys runtime, producing a \sys-enabled binary.
+\textbf{Programming and Execution Model.}  To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate task-shared data.
+Then, they compile the task-based code, and link to \sys's runtime, producing a \sys-enabled binary.
 The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
-The scheduler itself requires custom memory management, denoted as {\em virtual memory manager}.
+The scheduler itself \textbf{requires} custom memory management, denoted as {\em virtual memory manager}.
 
@@
 By coalescing tasks \sys \emph{amortizes commit and transition costs}, and by a task splitting, after repeated failures on a single task, it \emph{avoids the task non-termination problem}.
-The scheduler uses only its recent execution history---i.e.
-it is hardware independent---as a metric to estimate energy availability and eventually to decide on the dynamic task size.
-Section~\ref{sec:task_adaptation} outlines ATS.
+The scheduler uses only its recent execution history---i.e., it is hardware independent---as a metric to estimate energy availability and eventually to decide on the dynamic task size.
+Section~\ref{sec:task_adaptation} describes ATS.
 
-\textbf{Virtual Memory Manager.} \sys can coalesce tasks \emph{only} due to its virtual memory manager (VMM).
-VMM privatizes a group of memory pages demanded by a coalesced task and limits an application access only to the fast volatile memory.
-VMM keeps all pages modifications in non-volatile memory on a coalesced task transition.
+\textbf{Virtual Memory Manager.} \sys can coalesce tasks \emph{only} thanks to its virtual memory manager (VMM).
+VMM privatizes a group of memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.
+VMM keeps all page modifications in non-volatile memory on a coalesced task transition.
 Section~\ref{sec:memory_virtulaization} outlines VMM.

### 2018-08-06 paper/asplos19/relatedwork.tex
chopping down related work

--- 2018-08-04 paper/asplos19/relatedwork.tex
@@
-\textbf{Intermittently-powered Devices.} There is a large body of research on
-energy-harvesting and wirelessly-powered embedded devices covered in recent
-surveys~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
-As computing power consumption continues to decrease, a new wave of
-systems that are powered by radio waves (RF) is
-emerging~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014,wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
-RF-powered devices are an important category of intermittent computing devices
-that \sys targets.  The emergence of intermittent computing hardware platforms
-has led to the development of instrumentation, debugging, and prototyping tools
-for such systems
+\textbf{Intermittently-powered Devices.} There is a large body of research on energy-harvesting and ambient-powered embedded devices summarized by~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
+For instance, a new wave of embedded systems powered by radio waves is emerging~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014,wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
+The emergence of such hardware platforms
+has led to the development of instrumentation, debugging, and prototyping tools for such systems
 ~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,wisent,flicker}.
 
-\textbf{Checkpointing-based Systems.} Early work on energy-harvesting runtime
-systems, such as Dewdrop~\cite{dewdrop}, assumed simple computations
-that complete on a predictable burst of energy.
+\textbf{Checkpointing-based Systems.} Early work on energy-harvesting runtimes, such as Dewdrop~\cite{dewdrop}, assumed simple computations that complete on a predictable burst of energy.
 Eon~\cite{sorber_sensys_2007} has scheduled such computation based on priority.
-Support for computation that spans power failures was first achieved with
-statically-placed conditional checkpoints in Mementos~\cite{mementos}.
+Support for computation that spans power failures was first achieved with statically-placed conditional checkpoints in Mementos~\cite{mementos}.
+Mementos left the possibility of non-volatile state becoming inconsistent with volatile state saved in the checkpoint, rendering in-place writes to data structures in non-volatile memory unsafe~\cite{mspcdino}.
+DINO~\cite{dino} addressed the consistency problem by selectively versioning non-volatile state within the checkpoints.
+Ratchet~\cite{ratchet} ensured consistency by placing a checkpoint at the beginning of each idempotent region in the code.
+Ratchet has a similar technique to task splitting with some essential differences: (i) It does not need to handle non-volatile memory accesses, since it assumes a unified non-volatile memory, whereas \sys has to save the volatile pages in non-volatile memory; (ii) The checkpoint is of a fixed size, while \sys privatize variant number of memory pages.
+Clank~\cite{hicks_isca_2017} ensured consistency with custom hardware that dynamically tracks read-after-write dependencies in memory accesses and checkpoints on demand.
+The consistency problem was also approached with a combination of undo and redo logging in software~\cite{baghsorkhi_cgo_2018}.
+Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all volatile state immediately before power failure and halting the execution.
+Unlike \sys, such systems rely on introspection hardware to monitor supply voltage and on accurate worst-case bounds on checkpoint cost.
 
-Mementos left the possibility of non-volatile state becoming inconsistent with
-volatile state saved in the checkpoint, rendering in-place writes to data
-structures in non-volatile memory unsafe~\cite{mspcdino}.
-%
-DINO~\cite{dino} addressed the consistency problem by selectively versioning
-non-volatile state within the checkpoints. Ratchet~\cite{ratchet} ensured
-consistency by placing a checkpoint at the beginning of each idempotent region
-in the code.
-Ratchet has a similar technique to task splitting with some essential differences: (i) It does not need to handle non-volatile memory accesses, since it assumes a unified non-volatile memory, whereas \sys has to save the volatile pages in non-volatile memory; (ii) The checkpoint is of a fixed size, while \sys privatize variant number of memory pages.
-Clank~\cite{hicks_isca_2017} ensured consistency with custom
-hardware that dynamically tracks read-after-write dependencies in memory
-accesses and checkpoints on demand. The consistency problem was also
-approached with a combination of undo and redo logging in
-software~\cite{baghsorkhi_cgo_2018}.
-%
-Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and
-Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all
-volatile state immediately before power failure and halting the execution.
-Unlike \sys, such systems rely on introspection hardware to monitor supply
-voltage and on accurate worst-case bounds on checkpoint cost.
+In all of the above systems, with the exception of DINO~\cite{dino}, checkpoints are dynamic, i.e.
+the programmer does not have explicit control over the point at which the code may be resumed after power failure.
+Dynamic checkpointing systems make it difficult for the programmer to respect application-level atomicity constraints, such as correlating sensor readings.
+Checkpointing systems that copy most volatile state scale poorly as the size of volatile memory increases.
+QuickRecall, Clank and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory---requires more time and energy to access than volatile memory (cf.
+Section~\ref{sec:cost_task-based}) and not a viable option for off-chip non-volatile memory.
+In addition, QuickRecall and Clank require custom hardware.
 
-In all of the above systems, with the exception of DINO~\cite{dino},
-checkpoints are dynamic, i.e. the programmer does not have explicit control
-over the point at which the code may be resumed after power failure.
-%
-Dynamic checkpointing systems make it difficult for the programmer to respect
-application-level atomicity constraints, such as correlating sensor readings.
-%
-Checkpointing systems that copy most volatile state scale poorly as the
-size of volatile memory increases.  QuickRecall, Clank and Ratchet reduce
-the copying overhead by allocating the stack in non-volatile memory, which
-requires more time and energy to access than volatile memory (cf.
-Section~\ref{sec:cost_task-based}) and is not a viable option for off-chip non-volatile
-memory. In addition, unlike \sys, QuickRecall and Clank require custom hardware.
+\textbf{Task-based Systems.} An alternative to checkpointing are recent systems based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and Mayfly~\cite{hester_sensys_2017}.
+Using static tasks, they eliminate the need to checkpoint volatile state.
+Using channel-based memory models~\cite{chain,hester_sensys_2017} or automatic privatization and redo-logging~\cite{alpaca} they avoid checkpointing overheads.
+Moreover, task-based models facilitate respecting application-level atomicity constraints.
+\sys also relies on
+statically defined tasks to avoid checkpointing volatile state.
+However, unlike prior systems, \sys coalesces its statically-defined tasks at runtime into more efficient dynamic tasks that adapt to changing energy conditions.
+\sys's mechanism for ensuring memory consistency also differs from the channel-based~\cite{chain} and privatization-based~\cite{alpaca} mechanisms in prior systems.
+\sys keeps memory consistent through memory virtualization optimized for bulk accesses to task-shared data with high locality.
 
-\textbf{Task-based Systems.} An alternative to checkpointing are recent systems
-based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and
-Mayfly~\cite{hester_sensys_2017}. Using static tasks, they eliminate
-the need to checkpoint volatile state. Using channel-based memory
-models~\cite{chain,hester_sensys_2017} or automatic privatization and
-redo-logging~\cite{alpaca} they avoid checkpointing overheads. Moreover,
-task-based models facilitate respecting application-level atomicity
-constraints.
-%
-\sys relates closely to task-based models because \sys also relies on
-statically defined tasks to avoid checkpointing volatile state. 
-%
-However, unlike prior systems, \sys coalesces its statically-defined tasks at
-runtime into more efficient dynamic tasks that adapt to changing energy
-conditions.
-%
-\sys's mechanism for ensuring memory consistency also differs from the
-channel-based~\cite{chain} and privatization-based~\cite{alpaca} mechanisms in
-prior systems. \sys keeps memory consistent through memory virtualization
-optimized for bulk accesses to task-shared data with high locality.
+\textbf{Task Decomposition.} In contrast to \sys's construction of coalesced tasks at runtime, prior work has proposed to optimize task size at compile time.
+CleanCut~\cite{cleancut_2018} program analysis statically estimated energy consumption and split the program into tasks until all tasks consumed less energy than the device can store.
+An alternative program analysis~\cite{baghsorkhi_cgo_2018} generated different versions of a program with different task sizes and empirically selected the best among them~\cite{baghsorkhi_cgo_2018}.
 
-\textbf{Task Decomposition.} In contrast to \sys's dynamic construction of
-coallesced tasks at runtime, prior work has proposed to optimize task size
-at compile time. CleanCut~\cite{cleancut_2018} program analysis statically
-estimated energy consumption and split the program into tasks until all tasks
-consumed less energy than the device can store.
-%
-An alternative program analysis~\cite{baghsorkhi_cgo_2018} generated different
-versions of a program with different task sizes and empirically selected the
-best among them~\cite{baghsorkhi_cgo_2018}.
-%
-HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program
-analysis to place a minimal number of conditional checkpoints that test the
-energy level at runtime before copying state, like Mementos~\cite{mementos}.
+HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program analysis to place a minimal number of conditional checkpoints that test the energy level at runtime before copying state, like Mementos~\cite{mementos}.
 
-Unlike \sys, such compiler-based approaches, face the challenge of statically
-predictiong energy consumption of arbitrary input-dependent code with
-peripheral access, which is an problem without a general solution. Furthermore,
-a static decomposition approach prevents portability across devices with
-different storage capacitors. In contrast, \sys avoids forcing any assumptions
-at compile time and adapts to energy storage capacity and incoming energy
-conditions at runtime.
+Unlike \sys, such compiler-based approaches, face the challenge of statically predictiong energy consumption of arbitrary input-dependent code with peripheral access, which is an problem without a general solution.
+Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.
+In contrast, \sys avoids forcing any assumptions at compile time and adapts to energy storage capacity and incoming energy conditions at runtime.
 
-\textbf{Memory Virtualization.} Prior work on embedded systems has studied a
-variety of memory virtualization strategies relating to \sys.
-TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory
-management. Later work extended the memory manager to support memory
-virtualization backed by flash memory~\cite{sensornetvm} and to ensure memory
-and type safety~\cite{tinyosmemorysafety}. SOS~\cite{sos},
-Contiki~\cite{contiki}, and t-kernel~\cite{tkernel} also developed memory
-management abstractions that virtualize memory size and provide safe and
-indirect access. Mat\'e~\cite{mate} developed full virtual machines support for
-sensor nodes, virtualizing not just memory resources, but other states and
-peripherals. The goal of \sys is to provide consistent, intermittent execution,
-leveraging the benefits of efficient bulk copying. In contrast, prior efforts
-focused more on programmability and runtime reliability properties provided by
-virtual memory.
+\textbf{Memory Virtualization.} Prior work on embedded systems has studied a variety of memory virtualization strategies relating to \sys.
+TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory management.
+Later work extended the memory manager to support memory virtualization backed by flash memory~\cite{sensornetvm} and to ensure memory and type safety~\cite{tinyosmemorysafety}.
+SOS~\cite{sos}, Contiki~\cite{contiki}, and t-kernel~\cite{tkernel} also developed memory management abstractions that virtualize memory size and provide safe and indirect access.
+Mat\'e~\cite{mate} developed full virtual machines support for sensor nodes, virtualizing not just memory resources, but other states and peripherals.
+The goal of \sys is to provide consistent, intermittent execution, leveraging the benefits of efficient bulk copying.
+In contrast, prior efforts focused more on programmability and runtime reliability properties provided by virtual memory.
 
-A related domain is unbounded, page-based transactional memory and
-deterministic parallel runtime systems~\cite{pagebasedtm,grace}.  These works
-have a different mechanism and purpose than \sys---ensuring that data are
-consistent and deterministically updated during concurrent executions.  Their
-similarity with \sys is in managing state to ensure consistency at
-the granularity of pages to amortize checking and tracking costs.
-\sys's paging implementation, which keeps a shadow page for each page to use
-during commit, is similar to the shadow paging scheme use for transactional
-commit~\cite{pagebasedtm}.
+A related domain is unbounded, page-based transactional memory and deterministic parallel runtime systems~\cite{pagebasedtm,grace}.
+These works have a different mechanism and purpose than \sys---ensuring that data are consistent and deterministically updated during concurrent executions.
+Their similarity with \sys is in managing state to ensure consistency at the granularity of pages to amortize checking and tracking costs.
+\sys's paging implementation, which keeps a shadow page for each page to use during commit, is similar to the shadow paging scheme use for transactional commit~\cite{pagebasedtm}.

### 2018-08-06 paper/asplos19/relatedwork.tex
chopping down related work (left for now but we might still cut some)

@@
 \textbf{Checkpointing-based Systems.} Early work on energy-harvesting runtimes, such as Dewdrop~\cite{dewdrop}, assumed simple computations that complete on a predictable burst of energy.
-Eon~\cite{sorber_sensys_2007} has scheduled such computation based on priority.
+%
+%Eon~\cite{sorber_sensys_2007} has scheduled such computation based on priority.
+%
 Support for computation that spans power failures was first achieved with statically-placed conditional checkpoints in Mementos~\cite{mementos}. 
-Mementos left the possibility of non-volatile state becoming inconsistent with volatile state saved in the checkpoint, rendering in-place writes to data structures in non-volatile memory unsafe~\cite{mspcdino}.
+%Mementos left the possibility of non-volatile state becoming inconsistent with volatile state saved in the checkpoint, rendering in-place writes to data structures in non-volatile memory unsafe~\cite{mspcdino}.
 DINO~\cite{dino} addressed the consistency problem by selectively versioning non-volatile state within the checkpoints.
 Ratchet~\cite{ratchet} ensured consistency by placing a checkpoint at the beginning of each idempotent region in the code.
-Ratchet has a similar technique to task splitting with some essential differences: (i) It does not need to handle non-volatile memory accesses, since it assumes a unified non-volatile memory, whereas \sys has to save the volatile pages in non-volatile memory; (ii) The checkpoint is of a fixed size, while \sys privatize variant number of memory pages.
+Ratchet has a similar technique to task splitting with the core difference: the checkpoint is of a fixed size, while \sys privatize variant number of memory pages.
+%
+%Ratchet has a similar technique to task splitting with some essential differences: (i) It does not need to handle non-volatile memory accesses, since it assumes a unified non-volatile memory, whereas \sys has to save the volatile pages in non-volatile memory; (ii) The checkpoint is of a fixed size, while \sys privatize variant number of memory pages.
+%
 Clank~\cite{hicks_isca_2017} ensured consistency with custom hardware that dynamically tracks read-after-write dependencies in memory accesses and checkpoints on demand.
@@
 An alternative program analysis~\cite{baghsorkhi_cgo_2018} generated different versions of a program with different task sizes and empirically selected the best among them~\cite{baghsorkhi_cgo_2018}.
-
 HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program analysis to place a minimal number of conditional checkpoints that test the energy level at runtime before copying state, like Mementos~\cite{mementos}.
-
-Unlike \sys, such compiler-based approaches, face the challenge of statically predictiong energy consumption of arbitrary input-dependent code with peripheral access, which is an problem without a general solution.
+Unlike \sys, such compiler-based approaches, face the challenge of statically predicting energy consumption of arbitrary input-dependent code with peripheral access, which is an problem without a general solution.
 Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.

### 2018-08-06 paper/asplos19/conclusions.tex
chopping down conclusions

Software for intermittently-powered energy-harvesting devices requires a dedicated runtime system. \sys is a new task-based system whose distinguishing feature is its adaptability to changing energy conditions at runtime. When more energy is available, \sys makes faster progress through the computation by coalescing statically-defined tasks. When less energy is available, progress is latched at sub-task granularity. \sys's virtual memory system ensures that program state in non-volatile memory remains consistent and amortizes the cost of state transfer between volatile and non-volatile memory. Our evaluation across applications on a real embedded energy-harvesting device demonstrates the utility of \sys, as well as the practicality of the proposed implementation.

### 2018-08-06 paper/asplos19/background.tex
backgroud pragraph cut on FRAM

@@
     \end{subfigure}
-    \caption{WAR dependency example.
-\texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state.
-\note{replace checkp(); with checkpoint();} \note{reduce font size by one point}}
+    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state.}
     \label{fig:war}
@@
 The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation.
-If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 7), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 6 and 9).
-\note{in the above example it should \ref{fig:war-example}, Line 6, first and then \ref{fig:war-execution}, Lines 5 and 8}
+If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 6), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 5 and 8).
 
@@
 
-\textbf{Direct Non-volatile Memory Access.}  \note{I want to emphasize again that we do not evaluate NV memory accesses} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.
+%\textbf{Direct Non-volatile Memory Access.} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.
 Checkpointing systems copy all volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at every checkpoint.

### 2018-08-06 paper/asplos19/coalescing.tex
Task coalescing recectioned
Get rid of redundancies in sections 4, 5 and 6

@@
 In general, committing involves moving state manipulated by a task into its permanent location in non-volatile memory.
-In particular, \sys's commit procedure copies dirty pages of memory that a task updated between fast, volatile working memory and slower, non-volatile main memory; we defer the details of paging privatization and commit to Section~\ref{sec:memory_virtulaization}.
+In particular, \sys's commit procedure copies dirty pages of memory that a task updated between fast, volatile working memory and slower, non-volatile main memory.
+We defer the details of paging privatization and commit to Section~\ref{sec:memory_virtulaization}.
+An effective coalescing strategy must be {\em aggressive} enough, attempting to coalesce a large number of tasks to amortize commit overhead.
+However, it must also be  {\em conservative} enough, attempting to coalesce only as many tasks as will execute to completion given certain energy conditions, reducing the risk or long coalesced task re-execution penalty.
 
-Coalescing has a double benefit.
+\subsubsection{Generic Design of Task Coalescing Strategies}
+\label{subsec:coalescingGeneral}
+
+% Coalescing has a double benefit.
 First, coalescing reduces the total number of relatively high-latency, high-energy non-volatile memory accesses because tasks primarily access variables in relatively low-latency, low-energy volatile memory.
 Second, coalescing eliminates the unnecessary instructions that implement task commit, which constitute a substantial run-time overhead, as our measurements in Section~\ref{sec:evaluation} show.
-
-An effective coalescing strategy must be {\em aggressive} enough, attempting to coalesce a large number of tasks to amortize commit overhead.
-However, it must also be  {\em conservative} enough, attempting to coalesce only as many tasks as will execute to completion given a certain energy conditions, reducing the risk or long coalesced task re-execution penalty.
 
@@
 	\begin{algorithmic}[1]
-        \State $B \leftarrow $ \Call{$f_\text{reboot}$}{$z$}
+        \State $C \leftarrow $ \Call{$f_\text{reboot}$}{$z$}
         \State $H \gets 0$ \Comment{\note{add comments for some lines to get rid of white spaces}}
@@
 	        %
-	        \While{$ i <= B$}
+	        \While{$ i <= C$}
 		        \State \Call{execute\_task}{$T_\texttt{i}$}
@@
 	        \State \Call{commit\_to\_fram}{\null}
-	        \State $B \leftarrow $ \Call{$f_\text{commit}$}{$B$}
+	        \State $C \leftarrow $ \Call{$f_\text{commit}$}{$C$}
         \EndWhile
@@
 
-\subsection{Task Coalescing Strategies}
+Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing strategy.
+In the algorithm, $C$ (coalescing target) is total number of static tasks that \sys will next attempt to coalesce.
+$T_\text{i}$ is the $i$th task executed since the last power failure.
+$H$ (history) is the total number of tasks executed since the last power failure.
+$W_\text{i}$ is the {\em weight} of a task.
+A task's weight is an arbitrary quantity associated with the task that represents its cost in time or energy to execute.
+Different coalescing strategies may apply different weights to a task,
+e.g., $W_\text{i} = 1\ \forall i$, or $W_\text{i} = \alpha E(T_\text{i})$ where $E(T_\text{i})$ is the average execution time of $T_\text{i}$ and $\alpha$ is a constant.
+% For example, a uniform weighting scheme sets the weight of each task to one, making them all effectively equal in cost to the coalescing algorithm.
+A different policy might use a weighting scheme that sets a task's weight equal to its run time relative to the task that executes for the shortest duration, which has unity weight.
+
+\subsubsection{Example Task Coalescing Strategies}
 \label{subsec:coalescingStrategies}
 
-Algorithm~\ref{algo:genCoalescing} shows the general structure of a coalescing strategy.
-In the algorithm, $B$ (target budget) is total number of static tasks that \sys will next attempt to coalesce.
-$T_\text{i}$ is the $i$th task executed since the last power failure.
-History, $H$, is the total number of tasks between the last executed task and the power failure that preceding it.
-$W_\text{i}$ is the {\em weight} of a task $T_\text{i}$.
-A task's weight $W$ is an arbitrary quantity associated with the task that represents its cost in time or energy to execute; different coalescing strategies may apply different weights to a task.
-For example, a uniform weighting scheme sets the weight of each task to one, making them all effectively equal in cost to the coalescing algorithm.
-A different policy might use a weighting scheme that sets a task's weight equal to its run time relative to the task that executes for the shortest duration, which has unity weight.
-
 Different coalescing strategies adhere mainly to the template in Algorithm~\ref{algo:genCoalescing}, varying in only a few {\em characteristic operations} that the algorithm leaves deliberately abstract.
-A strategy's reboot update function, $f_\text{reboot}$, updates $B$, the target budget, after a reboot.
-A strategy's weight lookup function, $f_\text{weight}$, returns the a task's weight.
-A strategy's commit update function, $f_\text{commit}$, updates $B$ after a successful commit.
-The following sections detail several possible coalescing strategies that we developed for \sys, albeit not the only possible.
+A strategy's reboot update function, $f_\text{reboot}$, updates $C$, the coalescing target, after a reboot.
+A strategy's weight lookup function, $f_\text{weight}$, returns a task's weight.
+A strategy's commit update function, $f_\text{commit}$, updates $C$ after a successful commit.
+The following paragraphs detail several possible coalescing strategies that we developed for \sys, albeit not the only possible.
 In fact, \sys \emph{allows programmers to design their own strategy} to implement and link to \sys's task scheduler.
 
-\subsubsection{Energy-Oblivious Coalescing}
+\paragraph{Energy-Oblivious Coalescing.}
 \label{subsec:energyBlind}
@@
 A very simple {\em energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of a coalesced task linearly.
-In such a scheme, the number of tasks to coalesce, $B$, increases by a constant $x$ when a coalesced sequence of tasks commits, and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure.
-The characteristic operations of the energy-oblivious strategy are
+In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits, and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure.
+The characteristic operations of EO are
 %
@@
 	\begin{split}
-		f_\text{reboot}(B) & = B - x, \\
+		f_\text{reboot}(C) & = C - x, \\
 		f_\text{weight}(T_\text{i}) & =  1, \\
-		f_\text{commit}(B) & = B + x.
+		f_\text{commit}(C) & = C + x.
 	\end{split}
@@
 With the successful commit of each coalesced task sequence, the coalescing target increases.
-Eventually, the target may be too high and a coalesced task composed of only fewer tasks will commit without interruption by a power failure.
-The strategy then linearly decreases the coalescing target, eventually reaching a coalescing target that allows completion.
+Eventually, the target may be too high and only a coalesced task composed of a few units will commit without interruption by a power failure.
+The strategy then linearly decreases $C$, eventually reaching a value that allows completion.
 A key limitation of this algorithm lies in the equality of the target decrease in $f_\text{reboot}$ and the increase in $f_\text{commit}$.
 Let us assume $x=1$.
-If, after $k$ successful commits the target must decrease to its original value $B-k$ due to an energy drop, this strategy requires $k$ successive power failures.
+If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive power failures.
 
-\subsubsection{Energy-Guided Coalescing}
+\paragraph{Energy-Guided Coalescing.}
 \label{subsec:energyAware}
@@
     \includegraphics[width=\linewidth]{figures/hg-coal-horiz.pdf}
-    \caption{HG coalescing sample execution across two power cycles. \textcolor{red}{explain the meaning of the letters and the logical connection between the left and the right (or cycle 1 and 2)}}
+    \caption{EG coalescing sample execution across two power cycles. C: coalescing target, H: history.}
     \label{fig:hg-coal}
@@
 The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions.
-It uses its recent execution history as  mean to measure energy availability and alters its target accordingly.
-By relying on the history of execution EG eliminates the problem of frequent power failures a single coalesced task.
-In fact EG can reduce its target to one from any coalesced target in at most two successive power failures.
-For instance, if the execution is interrupted after finishing a single task the history will be one ($H = 1$) and the target will be set to one also.
-While a history of $H$ suggests that there is sufficient energy to coalesce and successfully commit $H$ tasks (i.e., to set $B = H$).
-EG conservatively sets $B$ equal to half of $H$ moderating the risk of a very long coalesced task.
-We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
+It uses its recent execution history H to estimate energy availability and alters its target accordingly.
 The EG algorithm is characterized by the following functions
@@
 	\begin{split}
-		 f_\text{reboot}(H) & = \lceil(H / 2)\rceil,\\
+		 f_\text{reboot}(H) & = \lceil \rho H \rceil,\\
 		 f_\text{weight}(T_\text{i}) & = 1, \\
-		 f_\text{commit}(B) & = \lceil(B / 2)\rceil.
+		 f_\text{commit}(C) & = \lceil \gamma C \rceil,
 	\end{split}
@@
 %
-Fig.~\ref{fig:hg-coal} illustrates the operation of the EG. The figure shows two power cycles following the $i$th, the latter having an
+where $\rho$ and $\gamma$ belong to $[0, 1]$.
+By relying on the history of execution EG eliminates the problem of frequent power failures on a single coalesced task.
+In fact, at any reboot, EG will {\em conservatively} try to coalesce only a fraction ($\rho$) of tasks that had successfully completed during the last power cycle.
+% In fact EG can reduce its target to one from any coalesced target in at most two successive power failures.
+For instance, if the execution is interrupted after finishing a single task the history will be one ($H = 1$) and the target will be set to one also.
+While a history of $H$ suggests that there is sufficient energy to coalesce and successfully commit $H$ tasks (i.e., to set $C = H$).
+EG conservatively sets $C$ equal to half of $H$ moderating the risk of a very long coalesced task.
+We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
+Figure~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$. The figure shows two power cycles following the $i$th, the latter having an
 execution history $H^{(i)} = 4$.
-Based on that, $B$ is initially set to two and then decreases to one.
-Once the value of $B$ reaches one, EG is, in effect, expecting a power failure, justifying the conservative approach.
-After the power failure, EG uses the most recent history $H^{(i+1)} = 8$ to set $B = 4$ and continue execution.
+Based on that, $C$ is initially set to two and then decreases to one.
+Once the value of $C$ reaches one, EG is, in effect, expecting a power failure, justifying the conservative approach.
+After the power failure, EG uses the most recent history $H^{(i+1)} = 8$ to set $C = 4$ and continue execution.
 
-\subsubsection{Weighted Energy-Guided Coalescing}
+\paragraph{Weighted Energy-Guided Coalescing.}
 \label{subsec:energyTaskAware}
 
-The \emph{Weighted Energy-Guided} (WEG) coalescing strategy accounts for the different energy and time cost of a program's tasks when setting the coalescing target.
+The \emph{Weighted Energy-Guided} (WEG) coalescing strategy accounts for non-uniform energy and time costs of a program's tasks when setting $C$.
 Each different task in the program consumes a different amount of energy to execute to completion.
-The EO and EG coalescing strategies assume each task has uniform energy cost: for these strategies, $B$ simply corresponds to a target {\em count} of tasks coalesced, regardless of the individual cost of each task.
-However, if one task executes for ten seconds and another executes for one second, counting tasks misjudges the amount of {\em work} in the coalesced tasks.
-
-Instead, WEG associates a non-unity weight with each task in the program.
-WEG tracks the sum of the weights of tasks in a coalesced sequence of tasks.
-When the sum of weights reaches $B$, the target, WEG commits the coalesced task.
-WEG differs from EO and HG by explicitly accounting for the variation in energy cost of a program's tasks, eliminating the assumption that all tasks do the same work, which is a key limitation.
-WEG is characterized the same way as HG (equation (\ref{eq:eg})), except that $f_\text{weight}(T_\text{i}) = W_\text{i}$.
-
-The effectiveness of the WEG hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available.
+The EO and EG coalescing strategies assume that each task has the same cost: for these strategies, $C$ simply corresponds to a target {\em count} of tasks to coalesce, regardless of the individual cost of each task.
+However, if one task executes for ten seconds and another executes for one second, counting tasks misjudges the amount of {\em work} in the coalesced tasks (and the history).
+%
+Instead, WEG associates a non-unity weight with each task in the program, and tracks the sum of the weights of tasks in a coalesced sequence.
+When the sum of weights reaches $C$, the target, WEG commits the coalesced task.
+% WEG differs from EO and HG by explicitly accounting for the variation in energy cost of a program's tasks, eliminating the assumption that all tasks do the same work, which is a key limitation.
+WEG is characterized the same way as EG (equation (\ref{eq:eg})), except that $f_\text{weight}(T_\text{i}) = W_\text{i}$.
+%
+% The effectiveness of the WEG hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available.
 Profiling the time and energy cost of tasks in a program is a difficult, orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}.

### 2018-08-06 paper/asplos19/paging.tex
Get rid of redundancies in sections 4, 5 and 6

@@
 And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
-
-\sys's {\em virtual memory manager} is a core enabler of \sys task adaptation mechanism.
+%
+\sys's {\em virtual memory manager} (VMM) is a core enabler of \sys task adaptation mechanism.
 Its overview is given in Figure~\ref{figure:mem-man}.
-VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers.
-Each buffer is divided into pages.
-The VMM prohibits applications from directly access these buffers.
+VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
+\texttt{\underline{private}} holds the consistent version of pages after each commit (and on reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
+The VMM prohibits applications from directly accessing these buffers.
 Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
-It populates \texttt{working} buffer with the privatized pages from non-volatile memory.
-When a coalesced task completes, the VMM ensures that updated pages are committed to main memory and visible to subsequently executed tasks.
-If the power interrupts a task, the VMM ensures the consistency of the data by relying on its \texttt{\underline{private}} buffer.
+It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
+When a coalesced task completes, the VMM ensures that {\em dirty} (i.e., modified) pages are committed to FRAM and visible to subsequent tasks.
+If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
+% If the power interrupts a task, the VMM ensures the consistency of the data by relying on its \texttt{\underline{private}} buffer.
 The VMM ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
-On a commit, the VMM atomically copies, through two-phase commit, all ``dirty'' pages back into their location in the non-volatile memory.
-SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM.
-A shadow page stores the data from an updated pages that need to be committed back to non-volatile memory, but that exceeded the size of the volatile working memory.
-The VMM moves pages of data efficiently using hardware accelerated DMA support.
+On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
+% SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM.
+A shadow page stores the data from an updated page that needs to be committed back to non-volatile memory, but that exceeded the size of the volatile working memory.
 
@@
     \includegraphics[width=\columnwidth]{figures/mem-man.pdf}
-    \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP}, and its order of occurrence; steps 2, 3, 4 and 5 are conditional.
-Tasks only have direct access to the \texttt{working} buffer.}
+    \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP}, and order of occurrence. Steps 2, 3, 4 and 5 are conditional.}
     \label{figure:mem-man}
@@
 The interface includes \texttt{RP}$(v)$, to read the value of variable $v$, and \texttt{WP}$(v)$ to assign a value to variable $v$.
-The implementation of {\tt RP} is shown in Algorithm~\ref{algo:rwar}, and {\tt WP}'s implementation is similar except that {\tt WP} marks the accessed page as dirty.
-{\tt RP} and {\tt WP} operations translate a variable's physical address in non-volatile memory into a \emph{virtual address} in \texttt{working} buffer in volatile memory.
+The implementation of {\tt RP} is shown in Algorithm~\ref{algo:rwar}, and {\tt WP}'s implementation is similar except that accessed page are marked as dirty.
+{\tt RP} and {\tt WP} operations translate a variable's physical address in non-volatile memory into a \emph{virtual address} in \texttt{working} in volatile memory.
 In \sys, a virtual address is composed of a \emph{page tag} that identifies the page and a \emph{page offset} that identifies a byte.
@@
 
-When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory manager first searched the variable's page in the \texttt{working} buffer (Figure~\ref{figure:mem-man}, Step 1).
-If the page is not found there, a page fault is incurred and a new page needs to
-be swapped in.
-If the \texttt{working} buffer is full, a page fault on a memory access requires VMM to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
-While we opted for the simplicity of FIFO, any other replacement policy, such as Least Recently Used (LRU) could work in its place.
-If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes has to be persisted to non-volatile memory (Figure~\ref{figure:mem-man}, Step 2).
+When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory manager first searches the variable's page in the \texttt{working} buffer (Figure~\ref{figure:mem-man}, Step 1).
+If the page is not found there, a page fault is incurred and a new page needs to be swapped in.
+If the \texttt{working} buffer is full, a page fault on a memory access requires the VMM to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
+If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes have to be persisted to non-volatile memory (Figure~\ref{figure:mem-man}, Step 2).
 If the accessed page was previously modified and swapped out since the last power failure, the most recent version of the page is in {\tt \underline{shadow}} (Step 3), otherwise it has to be retrieved from {\tt \underline{private}} (Step 4).
@@
 
-When the last task in a sequence of coalesced tasks completes, \sys must commit \emph{all} dirty pages in the \texttt{working} buffer and in the \texttt{\underline{shadow}} buffer, copying them back to their original locations in the main memory \texttt{\underline{private}}.
+When the last task in a sequence of coalesced tasks completes, \sys must commit \emph{all} dirty pages in \texttt{working} and \texttt{\underline{shadow}}, copying them back to their original locations in \texttt{\underline{private}}.
 
-To make the commit atomic, \sys commits the pages in two phases, implemented by Algorithm~\ref{algo:commit}.
-The first phase copies dirty pages from the \texttt{working} buffer to the non-volatile \texttt{\underline{shadow}} buffer (Line 4).
+To make the commit atomic, \sys commits dirty pages in two phases, as shown in Algorithm~\ref{algo:commit}.
+The first phase copies dirty pages from \texttt{working} to the non-volatile \texttt{\underline{shadow}} (Line 4).
 The second phase commits pages from \texttt{\underline{shadow}} to \texttt{\underline{private}} (Line 12).
@@
 phase, the commit process safely resumes on reboot.
-The second phase of commit depends on some run-time metadata.
+The second phase depends on some run-time metadata.
 The \texttt{\underline{committing}} bit indicates that a commit is in progress and is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16).
@@
 
-\sys's commit is efficient because it maintains an index dirty pages instead of iterating through all potentially dirty pages to check a dirty bit during commit.
-\sys's implementation is also efficient because the VMM does not copy from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead swapping pointers in an indirection table that maintains the pages as a double buffer.
-Section~\ref{sec:implementation} describes double buffering in more detail.
+\sys's commit is efficient because it maintains an index of dirty pages instead of iterating through all potentially dirty pages to check their state.
+Another source of efficiency lies in the second phase of the commit:
+% \sys's implementation is also efficient because
+the VMM does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.
 
-\textbf{Memory Consistency.} \sys's paging mechanism ensures that a task only ever executes using consistent protected data.
+\paragraph{Memory Consistency.}
+\sys's paging mechanism ensures that a task only ever executes using consistent protected data.
 During task execution, modifications to protected data do not affect the \texttt{\underline{private}} buffer, because a task reads and writes the volatile \texttt{working} buffer only, and modified pages are kept in \texttt{\underline{shadow}} until commit.
 A power failure erases the contents of the \texttt{working} buffer, preventing a re-executing task from observing updates from a previous execution attempt.
-Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14), ensures that all accesses to protected variables in subsequent tasks correctly access their original memory locations in \texttt{\underline{private}}.
+Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14) ensures that all accesses to protected variables in subsequent tasks correctly access their consistent memory locations in \texttt{\underline{private}}.
 
@@
 
-\sys asks the programmer to use its {\tt RP} and {\tt WP} paging API methods on every access to a protected variable (cf.
+\sys asks the programmer to use its {\tt RP} and {\tt WP} API methods on every access to a protected variable (cf.
 Section~\ref{sec:coala_api}).
 These API invocations present a risk of high overhead because there is a dynamic check on every read and write.
-Despite the risk of per-access overhead, \sys's dynamic memory protection scheme brings several benefits over a static approach (i.e.~\cite{alpaca}).
+Despite the risk of per-access overhead, \sys's dynamic memory protection scheme brings several benefits over a static approach (i.e.,~\cite{alpaca}).
 First, the limitations of static analysis preclude some uses of pointers due to potential pointer aliasing.
-For example, in the presence of arbitrary pointer operation, a function call using a function pointer, or an interrupt within a task, the system cannot statically analyze the memory behavior.
-Second, a static approach cannot handle task coalescing, because a protected variable's lifetime, i.e.,from first use to commit, is unknown at compile time.
+For example, in the presence of arbitrary pointer operations, a function call using a function pointer, or an interrupt within a task, the system cannot statically analyze the memory behavior.
+Second, a static approach \textit{cannot handle task coalescing}, because a protected variable's lifetime, i.e., from first use to commit, is unknown at compile time.
 \sys's dynamic, per-access instrumentation supports arbitrary use of pointers and enables task coalescing.

### 2018-08-06 paper/asplos19/hwswsetup.tex
Minor changes to section 7

@@
 		\hline\hline
-        \textbf{ar} &10 &428 & Activity recognition using a nearest neighbor
-classifier\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
+        \textbf{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
 		\hline
-        \textbf{bc} &10 &371 & Counts bits in bytes sequences
-using several algorithms and check for correctness\\
+        \textbf{bc} &10 &371 & several bitcount algorithms\\
+        % \textbf{bc} &10 &371 & Counts bits in bytes sequences
+% using several algorithms and check for correctness\\
 		\hline
-        \textbf{cuckoo} &14 &426 & Inserts and queries a sequence of pseudo-random
-values in a Cuckoo filter.\\
+        \textbf{cuckoo} &14 &426 & inserts and queries a sequence of pseudo-random
+values in a Cuckoo filter\\
 		\hline
@@
 		\hline
-		\textbf{sort} &4 &167 & Selection sort algorithm\\
+		\textbf{sort} &4 &167 & selection sort algorithm\\
 		\hline
 	\end{tabular}
-\caption{Description and characteristics of benchmark applications used to
-evaluate \sys.}
+\caption{Characteristics of benchmarks used for evaluation.}
 \label{table:benchmark_table}
@@
 All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 toolchain with \texttt{-O1} as optimization flag.
-The source code for all benchmarks will be released at \cite{coala_website}.\\
+The source code for all benchmarks will be released at \cite{coala_website}.
+
+\paragraph{Comparison with Alpaca using the GCC compiler.}
+
+We compare \sys against Alpaca~\cite{alpaca}, the state-of-the-art task-based system for
+intermittent computing. For a fair comparison the task decomposition of the benchmarks are ensured to be the
+same for both systems.
 %
-\textbf{Comparison with Alpaca using the GCC compiler.} We compare
-\sys against Alpaca~\cite{alpaca}, the state-of-the-art task-based system for
-intermittent computing. For a fair comparison the task decomposition of the benchmarks are ensured to be the
-same for both systems.\\
+\note{try to reduce what follows}
 %

### 2018-08-06 paper/asplos19/evaluation.tex
revise evaluation

--- 2018-08-05 paper/asplos19/evaluation.tex
@@
 	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
-    \caption{Breakdown of overhead for the proposed coalescing strategies
-relative to the baseline without coalescing.
-NC: no coalescing, EO: energy-oblivious coalescing, HG: history-guided coalescing, WHG:
+    \caption{Breakdown of overhead for the proposed coalescing strategies.
+NC: no coalescing, EO: energy-oblivious coalescing, EG: Energy-guided coalescing, WEG:
 energy- and task-aware coalescing. All coalescing strategies reduce total
-overhead and maximize useful work. HG and WHG perform better than EO.}
+overhead and maximize useful work. EG and WEG perform better than EO.}
 	\label{fig:overallOverheadBreakdown}
@@
 %
-Building on top of the existing work, we have selected the most efficient run-time library~\cite{alpaca} and compared the performance of \sys against it.
+Building on top of the existing work, we have selected the most efficient task-based run-time library~\cite{alpaca} and compared \sys against it.
 Our evaluation quantitatively demonstrates that \sys (i) \emph{reduces the memory
 protection overhead} by coalescing, (ii) \emph{improves execution speed}, in most cases; %
-and (iii) by mean of task-downscaling \sys is able to progress where static system stuck in task non-termination loop. 
+and (iii) \sys is able to progress where static systems suffer from a task non-termination loop. 
 We have also broken down \sys's overhead to explain the source of \sys's improved performance.
@@
 and WEG) and for a baseline without coalescing (NC), we have measured the time
-spent on (i) executing useful task code, including \sys's overhead on access to
-protected variables, (ii) re-executing task code after a power failure, and
+spent on (i) executing task code (ii) re-executing task code after a power failure, and
 (iii) committing state to non-volatile memory at the end of each (coalesced)
@@
 antenna.
-%
-The page size was set to the best performing value (32~B for \textbf{bc} and
-256~B pages for the rest) identified in detailed evaluation of paging in
-Section~\ref{sec:results_memory_management}.
 
@@
 the application and in the runtime (top) and a breakdown of protected memory
-accesses into three types with different costs. NC: no coalescing, HG:
-history-guided coalescing, WHG: weighted history-guided coalescing.}
+accesses into three types with different costs. NC: no coalescing, EG:
+Energy-guided coalescing, WEG: weighted Energy-guided coalescing.}
 	\label{fig:coalEfficiency}
@@
 Fig.~\ref{fig:coalEfficiency} (top) breaks down the time spent on executing 
-useful task code and committing further into time spent by the application and
-time spent by \sys's runtime.
-%
-This dataset is generated in a separate experiment with the same setup as the
-previous experiment in Section~\ref{sec:overhead-coalescing}.
-%
-Consistent with the previous results in
-Fig.~\ref{fig:overallOverheadBreakdown}, commit overhead is highest when
-coalescing is not used to reduce the number of commits.
+task code and committing further into time spent by the application and
+time spent by \sys's runtime. When coalescing is not enabled the commit overhead is 
+highest.
 %
 However, coalescing increases the cost of each access to a protected variable,
-since work in the address translation in the paging system increases as the
-task size grows.
+since VMM has to spend more on average time in searching for the requested page.
 %
@@
 of the cheapest kind, which motivated us to optimize this access type in our
-implementation (cf. Section~\ref{sec:impl:paging}).
+implementation (see Section~\ref{sec:impl:paging}).
 %
-Only \textbf{cuckoo}, \textbf{dijkstra}, and \textbf{fft} have non-negligible
+Only \textit{cuckoo}, \textit{dijkstra}, and \textit{fft} have non-negligible
 number of accesses to a different SRAM page, which is due to the larger working
@@
 %
-The experimental setup is the same as that in Section~\ref{sec:coala_overhead}.
-%
 The results in Fig.~\ref{fig:coalescing} show that all benchmarks
 complete faster with coalescing than without coalescing: from 25\% faster
-(\textbf{ar}) up to 70\% faster (\textbf{sort}).
+(\textit{ar}) up to 70\% faster (\textit{sort}).
 %
@@
 (2) largely similar across the two coalescing strategies, with the exception of
-\textbf{fft}.
+\textit{fft}.
 %
-In some cases (\textbf{bc}, \textbf{cuckoo}, \textbf{sort}) the Weight Energy-guided
+In some cases (\textit{bc}, \textit{cuckoo}, \textit{sort}) the Weight Energy-guided
 feature in the coalescing algorithm is counter-productive.
@@
 %
-In \textbf{fft}, tasks are not uniform, and accounting for task energies is
+In \textit{fft}, tasks are not uniform, and accounting for task energies is
 beneficial. In fact, the lack of task energy awareness is detrimental: with EG
-\textbf{fft} runs slower than without any coalescing (NC).
+\textit{fft} runs slower than without any coalescing (NC).
 %
-The speedup is highest for \textbf{bc}, \textbf{cuckoo}, \textbf{dijkstra} and
-\textbf{sort}, because their tasks are small and are easily coalesced,
+The speedup is highest for \textit{bc}, \textit{cuckoo}, \textit{dijkstra} and
+\textit{sort}, because their tasks are relatively small and are easily coalesced,
 eliminating many unnecessary commits.
@@
 	\includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
-    \caption{\sys's execution time, normalized to the execution time 
+    \caption{\sys's execution time, normalized, where possible, to the execution time 
     with Alpaca, as a function of distance to the energy source 
@@
 Fig.~\ref{fig:runtime} shows the average execution time of each
-application for \sys and Alpaca, normalized to the latter. The results show
+application for \sys and Alpaca, normalized, where possible, to the latter. The results show
 that \sys provides a performance benefit compared to Alpaca for most
@@
 virtualization that causes its performance to be comparable to (or worse than)
-Alpaca (\textit{ar}, \textit{cuckoo}), on average \sys 26\% faster than Alpaca.
+Alpaca (\textit{ar}, \textit{cuckoo}).
 
-The \textit{fft} benchmark did not complete at distances larger than
-15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
+Due to its static progressing behavior Alpaca was unable to finish the \textit{fft} benchmark on a distance larger than 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
 execution is negligible and stored energy is insufficient to complete some of
-the static tasks.} with neither Alpaca, nor \sys \emph{without the downsizing
-mechanism}. This is marked with $\infty$ signs in Fig.~\ref{fig:coalescing}
-. This non-terminating condition cannot be avoided in Alpaca, due to the
-fixed, non-adaptive nature of its tasks.  In contrast, our next experiment
-shows how \sys's task downsizing adaptation mechanism eliminates this
-non-termination, without changing the static tasks and without recompiling the
-application.
+the static tasks.}.
+This is marked with $\infty$ signs in Fig.~\ref{fig:coalescing}.
+\sys, however, manged to complete \textit{fft} application by enabling its task downscaling on the distances 30\,cm and 50\,cm.
 
-In a separate experiment, we have enabled task downsizing in \sys and measured
-the execution time for \textit{fft} benchmark at 60\,cm from the signal
-generator antenna. \sys executed the
-benchmark successfully without (NC) and with coalescing (WEG). The application
-took longer to complete than at a shorter distance of 15\,cm in the previous
-experiment, because of smaller (negligible) incoming energy at the larger
-distance. The variant without coalescing (NC) slowed down by 2.16$\times$ and the
-variant with WEG coalescing slowed down by 2.31$\times$ slower, but at both
-distances, WEG completed faster than NC.
-%
-For comparison, the run time with Alpaca is effectively infinite at this
-distance, which is a consequence of its non-adaptive static tasks.
+Overall, \sys shows better performance than its counterpart, and it is able to overcome the \textit{big task} (i.e. \textit{fft} tasks) problem that the static task-based systems suffer from.
 
@@
 %
-This trend is visible for all applications, exception \textbf{bc}. The total
-amount of data accessed by \textbf{bc}, as well as its working set, is small.
+This trend is visible for all applications, exception \textit{bc}. The total
+amount of data accessed by \textit{bc}, as well as its working set, is small.
 Even with the smallest page, all accesses are contained within that page, and

### 2018-08-06 paper/asplos19/implementation.tex

@@
-We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittence-safe.
+We implemented \sys's programming and execution model as a runtime library and API that a programmer can use to make a plain C program intermittency-safe.
 %
@@
 %
-Coala's API adds only eight syntactic constructs to a C-based language, summarized
+Coala's API adds only a few syntactic constructs to a C-based language, summarized
 in Table~\ref{table:coala_api}.
@@
 \begin{table}
+\footnotesize
 \centering
-\begin{tabular}{| r | p{0.67\columnwidth} |}
+\begin{tabular}{| r | p{0.7\columnwidth} |}
 	\hline
@@
 	\hline
-	\texttt{PV}($p$, $v$ [, $s$]) & $p$: variable type, $v \in \mathbf{V}$: new protected variable name, $s$: array size \\
+	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
 	\hline
-	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to be read, $u$: destination operand \\
+	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
 	\hline	
-	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to be written, $u$: source operand \\
+	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
 	\hline
-	\texttt{SM}($p$, $m$ [, $s$]) & $p$: \texttt{struct} member's type, $m$: \texttt{struct} member's name, $s$: array size \\
+	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
+	\hline
+	\texttt{DISABLE\_PC}() & --- \\
+	\hline
+	\texttt{ENABLE\_PC}() & --- \\
 	\hline
@@
 \end{table}
-%
-\noindent\textbf{New Tasks.} The \texttt{TASK} annotation on a function
+
+\paragraph{New Tasks.} The \texttt{TASK} annotation on a function
 declaration statically allocates a non-volatile constant variable holding a
-task's weight and declares that the function is a task.\\
-%
-\noindent\textbf{Task Transitions.} A call to \texttt{NEXT\_TASK} marks the task to be executed at the 
+task's weight and declares that the function is a task.
+
+\paragraph{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed at the 
 end of the current one task and it can be invoked along
-any control path to dynamically determine the next task.\\
-%
-\noindent\textbf{Protected Variables.} The \texttt{PV} annotation on a
+any control path to dynamically determine the next task.
+
+\paragraph{Protected Variables.} The \texttt{PV} annotation on a
 variable statically allocates a protected non-volatile variable. The
 variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
-runtime to ensure correct operation.\\
-%
-\noindent\textbf{Initialization.} The behavior of the API method
+run-time to ensure correct operation.
+
+\paragraph{Initialization.} The behavior of the API method
 \texttt{INIT} is very similar to \texttt{NEXT\_TASK}, with the
-addition of performing preliminary kernel initializations, including hardware setup.\\
-%
-\noindent\textbf{Execution.} The programmer passes control to \sys's task scheduler 
+addition of performing preliminary kernel initializations, including hardware setup.
+
+\paragraph{Execution.} The programmer passes control to \sys's task scheduler 
 by calling \texttt{RUN} after device initialization. 
@@
 
-\subsection{Paging}
-\label{sec:impl:paging}
-
-As described in Section~\ref{sec:memory_virtulaization}, \sys's memory
-virtualization scheme uses three buffers: a volatile \texttt{working} buffer, a
-non-volatile \texttt{\underline{private}}, and a non-volatile
-\texttt{\underline{shadow}} buffer. 
 %
-The main memory \texttt{\underline{private}} holds the last committed copy of all 
-pages. 
+\subsection{Task Coalescing}
 %
-After reboot, the \texttt{working} buffer is empty. Execution fills 
-the \texttt{working} buffer with privatized pages as the program accesses them. 
-%
-The \texttt{\underline{shadow}} buffer serves as an intermediate,
-non-volatile backing \texttt{\underline{private}} for dirty pages evicted from the space-constrained
-\texttt{working} buffer.
-
-To access a protected variable with either \texttt{RP}
-or \texttt{WP}, \sys searches for the variable in its \texttt{working} buffer first.
- If the page is not, the VMM fetches it from non-volatile memory. 
-%
-\sys uses address-based page tagging to make finding a variable efficient.  The
-upper bits of a variable's memory address identify its page, and the lower bits
-denote the variable's offset in its page. The total number of pages in memory,
-$p$, determines the number of tag bits, which is $\log P$.
-
-Page tagging imposes a data alignment requirement.
-Page size $S$ must be a power of two. 
-%
-Pages must be aligned to an $S$-byte boundary for efficient memory access. 
-%
-We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
-an 8\,KB non-volatile \texttt{\underline{private}}, and a \texttt{working} buffer
-of 1\,KB.
-%
-\subsection{Coalescing and Commit}
-%
-The coalescing algorithm starts with a random guess about the size of the coalescing task, in our implementation the target is initialized to four.
+% The coalescing algorithm starts with a random guess about the size of the coalescing task, in our implementation the target is initialized to four.
 It then refines its guess based on the number of the static-tasks being successfully executed.
 %
-When a coalesced task completes, \sys save all the modifications in its \texttt{\underline{private}} using DMA.
-In order to ensure the atomicity of the commit operation all the modified pages are copied to the \texttt{\underline{shadow}} buffer first.
-The \sys set a flag to indicates an on-going \emph{atomic commit} to update the private pages.
-For efficiency reasons \sys uses an indirection table to point to the private consistent pages and eliminate the need for moving pages for the second time.
+\paragraph{Parameters.}
+Equations (\ref{eq:eo}) and (\ref{eq:eg}) parametrize the behavior of the coalescing strategies in terms of $x$, $\rho$ and $\gamma$.
+We experimented with a range of values, and then used the ones that yielded the best performance: $x = 1$ (for EO) and $\rho = \gamma = 1$ (for EG/WEG).
+
+\paragraph{Weights for WEG.}
+The effectiveness of the WEG hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available.
+Profiling the time and energy cost of tasks in a program is a difficult, orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
+WEG could use the result of an arbitrarily sophisticated profiling procedure.
+To produce a concrete result in this paper, we give WEG access to a simple profile of task run time collected offline using a single fixed input.
+WEG stores the profile in a lookup table that maps a task's identifier to its weight, making the information available to \sys's scheduler at run time.
+
 %
@@
 %
-\sys's task downscaling preserves forward progress by partially committing an
-ongoing, non-terminating task. 
+% \sys's task downscaling preserves forward progress by partially committing an
+% ongoing, non-terminating task. 
+%
+\paragraph{Timer Strategy.}
+After identifying a task as likely non-terminating, \sys must decide when during the task's execution to partially commit the task's state.
+\sys sets a timer at the start of the likely non-terminating task, initializing it to a very large value (e.g., an estimate of the maximum execution time using the device's energy buffer).
+If the timer expires before power fails, \sys partially commits, retaining  the timer value to use for future partial commits.
+If the timer does not expire before power fails, \sys halves the timer and tries again.
+The exponential decrease of the timer value converges rapidly to a usable one.
 %
@@
 
- \textbf{Partial Commit and Task Atomicity.} Some applications prevent
+ \paragraph{Partial Commit and Task Atomicity.}
+ Some applications may prevent
 downscaling a task because of a need for task atomicity.  For example, sampling
@@
 disable partial commit for a task or a span of code, marking  the code with a
-pair of \texttt{DISABLE\_PE} and \texttt{ENABLE\_PE} annotations.
-These annotations respectively halt and resume the partial execution timer.
+pair of \texttt{DISABLE\_PC} and \texttt{ENABLE\_PC} annotations.
+These annotations respectively halt and resume the partial commit timer.
+
+\subsection{Paging}
+\label{sec:impl:paging}
+
+% As described in Section~\ref{sec:memory_virtulaization}, \sys's memory
+% virtualization scheme uses three buffers: a volatile \texttt{working} buffer, a
+% non-volatile \texttt{\underline{private}}, and a non-volatile
+% \texttt{\underline{shadow}} buffer. 
+%
+% The main memory \texttt{\underline{private}} holds the last committed copy of all 
+% pages. 
+%
+% After reboot, the \texttt{working} buffer is empty. Execution fills 
+% the \texttt{working} buffer with privatized pages as the program accesses them. 
+%
+% The \texttt{\underline{shadow}} buffer serves as an intermediate,
+% non-volatile backing \texttt{\underline{private}} for dirty pages evicted from the space-constrained
+% \texttt{working} buffer.
+
+% To access a protected variable with either \texttt{RP}
+% or \texttt{WP}, \sys searches for the variable in its \texttt{working} buffer first.
+ % If the page is not, the VMM fetches it from non-volatile memory. 
+%
+\paragraph{Efficiency.}
+\sys uses address-based page tagging to make finding a variable efficient.  The
+upper bits of a variable's memory address identify its page, and the lower bits
+denote the variable's offset in its page. The total number of pages in memory,
+$P$, determines the number of tag bits, which is $\log_2 P$.
+%
+Furthermore, the VMM moves pages of data efficiently using hardware accelerated DMA support.
+
+\paragraph{Alignment.}
+Page tagging imposes a data alignment requirement.
+The page size $S$ must be a power of two. 
+%
+Pages must be aligned to an $S$-byte boundary for efficient memory access. 
+%
+To preserve alignment, when typedef'ing a C structure for protected variables, the programmer has to use the API method \texttt{SM} on all members of the structure (only when defining the structure).
+%
+We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
+an 8\,KB non-volatile \texttt{\underline{private}} buffer, and a \texttt{working} buffer
+of 1\,KB.
+
+\paragraph{Page Eviction.}
+When a page in \texttt{working} has to be swapped out to make room for a newly requested one, an eviction policy has to be chosen.
+While we opted for the simplicity of FIFO, any other replacement policy, such as Least Recently Used (LRU), could work in its place.

### 2018-08-06 paper/asplos19/partialcommit.tex
revise partial commit
Downscaling section revised

--- 2018-04-17 paper/partialcommit.tex
@@
 
-\sys uses {\em task downscaling} to make progress through a task that is too
-large to complete using the energy that the device can buffer.  Task
-downscaling executes part of the long task and interrupts its execution with a
-{\em partial commit}, after which the long task continues. If power fails, the
-partial commit preserves the progress through the task, and execution resumes
-from the point of the partial commit, rather than from the start of the task.
-Eventually, after some number of partial commits and power failures, the task
-will complete and execution will proceed to a task with an energy demand
-accommodated by the device's energy buffer. Task downscaling violates task
-atomicity in a non-terminating execution, favoring continued progress over
-atomicity. If a program requires a task's atomicity over its progress, 
-the programmer can disable task downscaling for that task.
+\sys uses {\em task downscaling} to make progress through a task that is too large to complete using the energy that the device can buffer.
+Task downscaling executes part of the long task and interrupts its execution with a {\em partial commit}, after which the long task continues.
+If power fails, the partial commit preserves the progress through the task, and execution resumes from the point of the partial commit, rather than from the start of the task.
+Eventually, after some number of partial commits and power failures, the task will complete and execution will proceed.
+Task downscaling violates task atomicity in a non-terminating execution, favoring continued progress over atomicity.
+If a program requires a task's atomicity to be preserved, they can disable task downscaling for that task.
 
 The key design issue for task downscaling is deciding when to partially commit.
-\sys only applies task downscaling to a task that is likely to be
-non-terminating.  A task is likely to be non-terminating if it fails to execute
-completely twice consecutively. The second incomplete execution executes after
-a power failure, when the device will have fully recharged its energy buffer.
-If a task cannot complete when executing with a full energy buffer, the task is
-non-terminating.
+\sys only applies task downscaling to a task that is likely to be non-terminating.
+A task is likely to be non-terminating if it fails to execute completely twice consecutively.
+The second incomplete execution executes after a power failure, when the device will have fully recharged its energy buffer.
+If a task cannot complete when executing with a full energy buffer, the task is marked as non-terminating.
 
-After identifying a task as likely non-terminating, \sys must decide when
-during the task's execution to partially commit the task's state.
-%
-\sys sets a timer at the start of the likely non-terminating task, initializing
-it to a very large value (e.g., an estimate of the maximum execution time using
-the device's energy buffer).
-%
-If the timer expires before power fails, \sys partially commits, retaining 
-the timer value to use for future partial commits. 
-%
-If the timer does not expire before power fails, \sys halves the timer and
-tries again.
-%
-\sys's exponential decrease of the timer value converges rapidly to a useable
-timer value.
-
-%Putting these in comments to make space look right
-%\textcolor{red}{More variables describing the partial commit algorithm: we do not explore as many options for partial commit as we do for coalescing}
-
-%\todo{Proof of the maximum granularity of task downscaling - show it is limited by the clock}{Carlo}
-
-%\subsubsection{Why Task Downsizing is not \sys's Default Behavior}
-%I do not understand what this section is about so I am removing it.  If someone wants to take another crack at re-writing it feel free.
-%Task downsizing is the core novelty of \sys when it comes to code portability and re-usability. However, it has limitations, which force  it to be a non-default \sys behavior. 
-%
-%\textbf{Static deployment of partial commit stubs.} This approach reduces programmer effort, although the programmer still needs to apply \sys's API and distribute trigger points for partial commits.
-On the other hand, \sys' downsizing introduces an additional overhead by saving the stack and the register file each time it commits the pages.
-%
-%\textbf{Dynamic deployment of partial commit stubs.} \sys downsizing does not rely on any information from the programmer to trigger the commit process.
-However, \sys utilizes a trial and error approach to optimize the commit rate over the entire program execution time.
-Therefore, it suffers from a large cumulative re-execution penalty.
+% After identifying a task as likely non-terminating, \sys must decide when during the task's execution to partially commit the task's state.
+\sys sets a timer at the start of the likely non-terminating task, initializing it to a very large value (e.g., an estimate of the maximum execution time using the device's energy buffer).
+If the timer expires before power fails, \sys partially commits, retaining  the timer value to use for future partial commits.
+If the timer does not expire before power fails, \sys halves the timer and tries again.
+The exponential decrease of the timer value converges rapidly to a usable one.

### 2018-08-06 paper/asplos19/coalescing.tex

@@
 	\begin{algorithmic}[1]
-        \State $C \leftarrow $ \Call{$f_\text{reboot}$}{$z$}
-        \State $H \gets 0$ \Comment{\note{add comments for some lines to get rid of white spaces}}
+        \State $C \leftarrow $ \Call{$f_\text{reboot}$}{$z$} \Comment{Update coalescing target by reboot update function}
+        \State $H \gets 0$ \Comment{Initialize history: total number of tasks executed since last power failure}
         \While{$\texttt{true}$}       
@@
 		        \State \Call{execute\_task}{$T_\texttt{i}$}
-		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$}
+		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$} \Comment{Assign new task-dependent weight}
 		        \State $i \gets i + W$
@@
 A task's weight is an arbitrary quantity associated with the task that represents its cost in time or energy to execute.
-Different coalescing strategies may apply different weights to a task,
-e.g., $W_\text{i} = 1\ \forall i$, or $W_\text{i} = \alpha E(T_\text{i})$ where $E(T_\text{i})$ is the average execution time of $T_\text{i}$ and $\alpha$ is a constant.
+Different coalescing strategies may apply different weights to a task, e.g., $W_\text{i} = 1\ \forall i$, or $W_\text{i} = \alpha E(T_\text{i})$ where $E(T_\text{i})$ is the average execution time of $T_\text{i}$ and $\alpha$ is a constant.
 % For example, a uniform weighting scheme sets the weight of each task to one, making them all effectively equal in cost to the coalescing algorithm.
@@
     \includegraphics[width=\linewidth]{figures/hg-coal-horiz.pdf}
-    \caption{EG coalescing sample execution across two power cycles. C: coalescing target, H: history.}
+    \caption{EG coalescing sample execution across two power cycles. \emph{C}: coalescing target, \emph{H}: history.}
     \label{fig:hg-coal}
@@
 The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions.
-It uses its recent execution history H to estimate energy availability and alters its target accordingly.
+It uses its recent execution history, H, to estimate energy availability and alters its target accordingly.
 The EG algorithm is characterized by the following functions
@@
 %
-where $\rho$ and $\gamma$ belong to $[0, 1]$.
+where $\rho, \gamma \in [0, 1]$.
 By relying on the history of execution EG eliminates the problem of frequent power failures on a single coalesced task.
@@
 We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
-Figure~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$. The figure shows two power cycles following the $i$th, the latter having an
-execution history $H^{(i)} = 4$.
+Figure~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$.
+The figure shows two power cycles following the $i$th, the latter having an execution history $H^{(i)} = 4$.
 Based on that, $C$ is initially set to two and then decreases to one.

### 2018-08-06 paper/asplos19/paging.tex

@@
 And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
-%
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalesced-war.pdf}
+	\caption{induction of WAR dependency by coalescing: without extra protection task coalescing induces WAR dependency (at variable $x$) even though individual tasks are WAR protected.}
+	\label{figure:coalescedWar}
+\end{figure}
+
 \sys's {\em virtual memory manager} (VMM) is a core enabler of \sys task adaptation mechanism.
-Its overview is given in Figure~\ref{figure:mem-man}.
+Its necessity is illustrated by example given in Figure~\ref{figure:coalescedWar}: even though each task might be WAR-protected, coalescing breaks this insurance.
+Therefore we need a new mechanism that will keep WAR-protection despite coalescing.
+
+VMM overview is given in Figure~\ref{figure:mem-man}.
 VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.

### 2018-08-06 paper/asplos19/paging.tex
paging revised

--- 2018-08-06 paper/asplos19/paging.tex
@@
 And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
-%
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalesced-war.pdf}
+	\caption{induction of WAR dependency by coalescing: without extra protection, which static runtime cannot reason about, task coalescing induces WAR dependency (at variable $x$) even though individual tasks are WAR protected.}
+	\label{figure:coalescedWar}
+\end{figure}
+
 \sys's {\em virtual memory manager} (VMM) is a core enabler of \sys task adaptation mechanism.
-Its overview is given in Figure~\ref{figure:mem-man}.
+Its necessity is illustrated by example given in Figure~\ref{figure:coalescedWar}: even though each task might be WAR-protected, coalescing breaks the global WAR insurance.
+Therefore we need a new memory management mechanism that will keep WAR-protection despite coalescing, which in case of \sys is exemplified by VMM.
+
+VMM overview is given in Figure~\ref{figure:mem-man}.
 VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.

### 2018-08-06 paper/asplos19/hwswsetup.tex
Methodology revised

@@
-We prototype \sys and use its API to implement a set of benchmark
-applications representative of the embedded domain. We build the applications
-and deploy the binaries onto a real energy-harvesting device. In this section
-we describe the setup for the experiments presented in
-Section~\ref{sec:evaluation} and characterize the benchmark applications.
+We prototype \sys and use its API to implement a set of benchmark applications representative of the embedded domain.
+We build the applications and deploy the binaries onto a real energy-harvesting device.
+
+%In this section we describe the setup for the experiments presented in Section~\ref{sec:evaluation} and characterize the benchmark applications.
 
@@
 
-We evaluated \sys using six benchmarks that are oftenly used in embedded systems (summarized in
-Table~\ref{table:benchmark_table}).
+We evaluated \sys using six benchmarks that are often used in embedded systems (summarized in Table~\ref{table:benchmark_table}).
 All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 toolchain with \texttt{-O1} as optimization flag.
@@
 
-We compare \sys against Alpaca~\cite{alpaca}, the state-of-the-art task-based system for
-intermittent computing. For a fair comparison the task decomposition of the benchmarks are ensured to be the
-same for both systems.
+We compare \sys against Alpaca~\cite{alpaca}, the state-of-the-art task-based system for intermittent computing.
+For a fair comparison the task decomposition of the benchmarks are ensured to be the same for both systems.
+
+Since Alpaca's compiler pass is implemented only for LLVM, we could not use that implementation to instrument the benchmarks and compile them with GCC.
+Instead, we \emph{manually} performed the instrumentation done by the compiler pass.
+The instrumentation consists of identifying WAR dependencies  and adding code and memory allocations to make a private copy of the affected variables.
+For array variables, instrumentation also includes adding code for dynamically tracking the array access.
+
+%Manual instrumentation of Alpaca code is not a general approach, but only served as a practical alternative to porting the compiler pass from LLVM to GCC.
 %
-\note{try to reduce what follows}
+%Our experiments show that manually-instrumented Alpaca code compiled with GCC is about 30\%-40\% faster than Alpaca code instrumented automatically by the Alpaca LLVM compiler pass and compiled with LLVM.
+This performance difference is due to the early experimental state of the MSP430 backend for LLVM~\cite{baghsorkhi_cgo_2018}.
 %
-Since Alpaca's compiler pass is implemented only for LLVM, we could not use
-that implementation to instrument the benchmarks and compile them with GCC.
-Instead, we \emph{manually} performed the instrumentation done by the compiler
-pass. The instrumentation consists of identifying write-after-read dependencies 
-and adding code and memory allocations to make a private copy of the affected variables. 
-For array variables, instrumentation also includes adding code for 
-dynamically tracking the array access. Manual instrumentation of Alpaca code is not a general approach, 
-but only served as a practical alternative to porting the compiler pass from LLVM to GCC.
-%
-Our experiments show that manually-instrumented Alpaca code compiled with GCC
-is about 30\%-40\% faster than Alpaca code instrumented automatically by the
-Alpaca LLVM compiler pass and compiled with LLVM. This performance difference
-is due to the early experimental state of the MSP430 backend for
-LLVM~\cite{baghsorkhi_cgo_2018}.
-%
-By compiling both systems with the same compiler (GCC), we ensure that our
-comparison in Section~\ref{sec:evaluation} is fair.
+By compiling both systems with the same compiler (GCC), we ensure that our comparison in Section~\ref{sec:evaluation} is fair.

### 2018-08-06 paper/asplos19/intro.tex
revise intro

@@
 Harvestable energy sources are low-power (e.g., \si{\nano\watt}\ to \si{\micro\watt}) compared to a platform's operating power level (hundreds of \si{\micro\watt}\ to \si{\milli\watt}).
-A device operates briefly until it depletes its buffered energy, after which, it shuts down and recharges to operate again later---corresponding to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of operation-power failure-restart cycles.
-The recharge and discharge intervals---which correspond to the device's inactive and active time---vary depending on the underlying hardware, such as the size of the energy buffering capacitor~\cite{capybara}, and some devices discharge and restart $\approx$10 to $\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.
+A device operates briefly until it depletes its buffered energy, after which, it shuts down and recharges to operate again later---corresponding to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of operation---power failure---restart cycles.
+The recharge and discharge intervals---which correspond to the device's inactive and active time---vary depending on the underlying hardware, such as the size of the energy buffering~\cite{capybara}, and energy conditions.
+For example, some devices discharge and restart $\approx$10 to $\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.
 
 Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM.
-While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated, and may leave checkpointed volatile state inconsistent with non-volatile state.
+While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated.
 These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, often leading to an unrecoverable application failure~\cite{dino,edb}.
@@
 Therefore, this work focuses on software approaches.
-In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} of statically decomposing a program into tasks.
+In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} and \emph{energy-unawareness} of statically decomposing a program into tasks.
 %
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/intro-figure-horiz.pdf}
+    \includegraphics[width=\columnwidth]{figures/intro-figure.pdf}
     \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
@@
 The programmer (or a compiler) explicitly expresses task-to-task control flow.
-Fig.~\ref{fig:coalesce} illustrates how a program's tasks execute and shows how tasks transitions can impose a runtime overhead.
+Fig.~\ref{fig:coalesce} (left) illustrates how a program's tasks execute and shows how tasks transitions can impose a runtime overhead.
 At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
@@
 A programmer may thus create very large tasks in an attempt to reduce task transitions overhead.
-However, a large task may require more energy to complete than a device's fixed hardware energy buffer can hold, which may lead to a task \emph{non-termination} problem.
+However, a large task may require more energy to complete than a device's fixed hardware energy buffer can hold, which may lead to a task \emph{non-termination} problem (see Fig.~\ref{fig:coalesce}(right)).
 To eliminate this risk, existing systems require the programmer to decompose a program into small tasks to preserve execution progress.
@@
 
-\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly gives rises  the need for dynamic memory consistency handling.
-This leads to the second challenge: how to dynamically detect inter-task data dependencies and ensure efficient protection against power interrupts?
+\item[C2] \emph{Dynamic Memory Consistency Handling:} merging static tasks on the fly rises  the need for dynamic memory consistency handling.
+This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts?
 \sys addresses this challenge by using a novel approach called \emph{dynamic group privatization}---performing real-time dependency tracking to enable protection on a task transition.
@@
 To asses the benefits of \sys over existing task-based systems, we implemented and tested six benchmarks on a real energy-harvesting platform.
-Our evaluation shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, and is able to progress where static task-based systems fail.
+Our evaluation not only shows that \sys reduces run time by up to 54\%, and by 26\% on average, but also is able to continue the execution where static task-based systems fail.
 

### 2018-08-06 paper/asplos19/intro.tex
update intro fig caption

@@
     \includegraphics[width=\columnwidth]{figures/intro-figure.pdf}
-    \caption{Task coalescing at runtime reduces time and energy overhead in task-based intermittent programming models by performing fewer commits (denoted as \texttt{C}) of individual tasks (denoted as \texttt{Tx}).
-On the other hand, task splitting reduces wasted computation and enables termination for bigger tasks.}
+    \caption{Task coalescing reduces time and energy overhead by performing fewer commits (denoted as \texttt{C}) than the individual tasks (denoted as \texttt{Tx}) commits (left).
+On the other hand, task splitting reduces wasted computation and enables termination of big tasks, requiring more energy than the energy buffer can hold (right).}
     \label{fig:coalesce}

### 2018-08-06 paper/asplos19/abstract.tex
abstract alignment for hotCRP
update the abstract

@@
 \sys progresses on a multi-task scale when energy permits and preserves the computation progress on a sub-task scale if necessary.
-\sys' specialized memory virtualization mechanism ensures that power failures do not leave the program state in non-volatile memory inconsistent.
-Our evaluation on a real energy-harvesting platform shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, and is able to progress where static systems fail.
+\sys's specialized memory virtualization mechanism ensures that power failures do not leave the program state in non-volatile memory inconsistent.
+Our evaluation on a real energy-harvesting platform not only shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, but also it is able to \emph{progress} where static systems \emph{fail}.

### 2018-08-06 paper/asplos19/paging.tex
Fig. 5

--- 2018-08-06 paper/asplos19/paging.tex
@@
 And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
-%
+
+\begin{figure}
+	\centering
+	\includegraphics[width=\columnwidth]{figures/coalesced-war.pdf}
+	\caption{A WAR dependency (at variable $x$) induced by task coalescing at runtime. Such dynamic dependencies cannot be resolved at compile time and they break memory consistency.
+%		induction of WAR dependency by coalescing: without extra protection, which static runtime cannot reason about, task coalescing induces WAR dependency (at variable $x$) even though individual tasks are WAR protected.
+}
+	\label{figure:coalescedWar}
+\end{figure}
+
 \sys's {\em virtual memory manager} (VMM) is a core enabler of \sys task adaptation mechanism.
-Its overview is given in Figure~\ref{figure:mem-man}.
+Its necessity is illustrated by example given in Figure~\ref{figure:coalescedWar}: resolving WAR dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent with task coalescing.
+Since task coalescing induces dynamic WAR dependencies,  a new memory consistency handling technique is mandatory so that these dependencies are resolved and handled at runtime.
+
+VMM overview is given in Figure~\ref{figure:mem-man}.
 VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.

### 2018-08-07 paper/asplos19/background.tex
revise background

@@
     \end{subfigure}
-    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkp()} is a checkpoint of volatile state.}
+    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkpoint()} is a checkpoint of volatile state.}
     \label{fig:war}
 \end{figure}
+
+\begin{table}
+    \centering
+    \footnotesize
+    \begin{tabular}{|c|c|}
+        \hline
+        Model & Data Copied to/from NVRAM \\
+        \hline\hline
+        Mementos~\cite{mementos}    & Registers + Stack     \\
+        DINO~\cite{dino}    & Registers + Stack + WAR NV variables \\%used in task\\
+        Chain~\cite{chain}  & PC + NV variables used in task\\
+        Alpaca~\cite{alpaca}    & PC + WAR NV variables used in task\\
+        Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
+        Region Formation~\cite{baghsorkhi_cgo_2018} & Registers + Updated variables in task \\
+        \hline
+    \end{tabular}
+    \caption{Non-volatile memory access for data consistency; \emph{PC}: program counter, \emph{WAR variables}: variables involved in Write after Read (WAR) dependences, \emph{NV}: non-volatile.}
+    \label{table:chechpoint_comparison}
+\end{table}
 
@@
 Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
-Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes using hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}.
+Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes requiring hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}, Table~\ref{table:chechpoint_comparison} summarizes the data that gets copied to and from the non-volatile memory in service of memory consistency and progress preservation.
 Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}.
@@
 
-\begin{table}
-    \centering
-    \footnotesize
-    \begin{tabular}{|c|c|}
-        \hline
-        Model & Data Copied to/from NVRAM \\
-        \hline\hline
-        Mementos~\cite{mementos}	& Registers + Stack     \\
-        DINO~\cite{dino}	& Registers + Stack + WAR NV variables \\%used in task\\
-        Chain~\cite{chain}	& PC + NV variables used in task\\
-        Alpaca~\cite{alpaca}	& PC + WAR NV variables used in task\\
-        Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
-        Region Formation~\cite{baghsorkhi_cgo_2018} & Registers + Updated variables in task \\
-        \hline
-    \end{tabular}
-    \caption{Non-volatile memory access for data consistency; \emph{PC}: program counter, \emph{WAR variables}: variables involved in Write after Read (WAR) dependences, \emph{NV}: non-volatile.}
-    \label{table:chechpoint_comparison}
-\end{table}
-
 %\textbf{Direct Non-volatile Memory Access.} Previous systems frequently access the non-volatile memory directly, which is a potential source of inefficiency.

### 2018-08-07 paper/asplos19/intro.tex
revised and ready

@@
 This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts?
-\sys addresses this challenge by using a novel approach called \emph{dynamic group privatization}---performing real-time dependency tracking to enable protection on a task transition.
+\sys addresses this challenge by using a novel approach called \emph{dynamic page privatization}---performing real-time dependency tracking to enable protection on a task transition.
 Individual variables tracking, however, slows down a system dramatically.
@@
 Section~\ref{sec:systemdescription} provides an overview of \sys, while Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
-The dynamic group privatization is explained in Section~\ref{sec:memory_virtulaization}.
+The dynamic page privatization is explained in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are given in Section~\ref{sec:implementation}.

### 2018-08-07 paper/asplos19/abstract.tex
Abstract revised

@@
+%Therefore, the two paradigms of intermittent execution, program checkpointing and task-based programming, save the intermediate program state into non-volatile memory frequently to preserve the execution progress.
+Despite the superiority of the task-based approach over checkpointing, tasks' size is fixed at compile-time and agnostic to energy conditions.
+
 Energy-neutral Internet of Things requires freeing embedded devices from batteries and powering them from ambient energy.
 Ambient energy is, however, unpredictable and can only power a device intermittently.
-Therefore, the two paradigms of intermittent execution, program checkpointing and task-based programming, save the intermediate program state into non-volatile memory frequently to preserve the execution progress.
-Despite the superiority of the task-based approach over checkpointing, tasks' size is fixed at compile-time and agnostic to energy conditions.
-Thus, the state may be saved either more frequently than necessary or not frequently enough for the program to progress and terminate.
-To address these challenges, we propose \sys, an adaptive and efficient task-based execution model.
+Therefore, the paradigm of intermittent execution is to save the program state into non-volatile memory frequently to preserve the execution progress.
+In task-based intermittent programming the state is saved at task transition.
+Tasks are fixed at compile-time and agnostic to energy conditions.
+Thus, the state may be saved either more often than necessary or not often enough for the program to progress and terminate.
+To address this challenges, we propose \sys, an adaptive and efficient task-based execution model.
 \sys progresses on a multi-task scale when energy permits and preserves the computation progress on a sub-task scale if necessary.

### 2018-08-07 paper/asplos19/systemdescription.tex

@@
 It addresses the challenges given in Section~\ref{sec:background} making task-based intermittent applications {\em programmable}, {\em efficient} and {\em portable}, introducing a new programming model and runtime software system that supports adaptive task-based execution.
-Figure~\ref{fig:system_overview} shows an overview of \sys.
+Fig.~\ref{fig:system_overview} shows an overview of \sys.
 
 \textbf{Programming and Execution Model.}  To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate task-shared data.
-Then, they compile the task-based code, and link to \sys's runtime, producing a \sys-enabled binary.
+Then, he compile the task-based code, and link to \sys's runtime, producing a \sys-enabled binary.
 The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
-The scheduler itself \textbf{requires} custom memory management, denoted as {\em virtual memory manager}.
+Facilitating efficient task adaptation requires dynamic memory protection which is \sys's \emph{virtual memory manager} through a process we call \emph{dynamic page privatization}.
 
@@
 By coalescing tasks \sys \emph{amortizes commit and transition costs}, and by a task splitting, after repeated failures on a single task, it \emph{avoids the task non-termination problem}.
-The scheduler uses only its recent execution history---i.e., it is hardware independent---as a metric to estimate energy availability and eventually to decide on the dynamic task size.
+The scheduler uses its recent execution history---i.e., it is hardware independent---as a metric to estimate energy availability and eventually to decide on the coalesced task size.
 Section~\ref{sec:task_adaptation} describes ATS.
 
-\textbf{Virtual Memory Manager.} \sys can coalesce tasks \emph{only} thanks to its virtual memory manager (VMM).
+\textbf{Virtual Memory Manager.} \sys can coalesce tasks thanks to its virtual memory manager (VMM).
 VMM privatizes a group of memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.

### 2018-08-07 paper/asplos19/intro.tex
Final pass on intro (Carlo)

@@
 Harvestable energy sources are low-power (e.g., \si{\nano\watt}\ to \si{\micro\watt}) compared to a platform's operating power level (hundreds of \si{\micro\watt}\ to \si{\milli\watt}).
-A device operates briefly until it depletes its buffered energy, after which, it shuts down and recharges to operate again later---corresponding to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of operation---power failure---restart cycles.
-The recharge and discharge intervals---which correspond to the device's inactive and active time---vary depending on the underlying hardware, such as the size of the energy buffering~\cite{capybara}, and energy conditions.
+A device operates briefly until it depletes its buffered energy, after which, it shuts down and recharges to operate again later---corresponding to the {\em intermittent execution model}~\cite{dino,lucia_snapl_2017} composed of operation-power failure-restart cycles.
+The recharge and discharge intervals---which correspond to the device's inactive and active time---vary depending on the underlying hardware, such as the size of the energy buffer~\cite{capybara}, and energy conditions.
 For example, some devices discharge and restart $\approx$10 to $\approx$100 times per second~\cite{tan_infocom_2016,mementos,nvp}.
@@
 While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated.
-These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, often leading to an unrecoverable application failure~\cite{dino,edb}.
-Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based computer architecture support}~\cite{hicks_isca_2017,idetic,nvp}.
+These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, leading to an unrecoverable application failure~\cite{dino,edb}.
+Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based architectural support}~\cite{hicks_isca_2017,idetic,nvp}.
 Complex hardware architectural changes are expensive to design, verify, and manufacture.
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/intro-figure.pdf}
-    \caption{Task coalescing reduces time and energy overhead by performing fewer commits (denoted as \texttt{C}) than the individual tasks (denoted as \texttt{Tx}) commits (left).
+    \includegraphics[width=\columnwidth]{figures/intro-figure-v5.pdf}
+    \caption{Task coalescing reduces commit overhead (left), while task splitting enables termination of large tasks (right).
+    S: static legacy task-based runtime, C: \sys.}
+    % \caption{Task coalescing reduces time and energy overhead by performing fewer commits (denoted as \texttt{C}) than the individual tasks (denoted as \texttt{Tx}) commits (left).
 On the other hand, task splitting reduces wasted computation and enables termination of big tasks, requiring more energy than the energy buffer can hold (right).}
@@
 
-\textbf{Crucial Drawback of Static Task Decomposition.} Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks.
-A \emph{task}, a top level function, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.
+\paragraph{Drawbacks of Static Task Decomposition.}
+Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks.
+\emph{Tasks}, top level functions, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.
 The programmer (or a compiler) explicitly expresses task-to-task control flow.
-Fig.~\ref{fig:coalesce} (left) illustrates how a program's tasks execute and shows how tasks transitions can impose a runtime overhead.
+Fig.~\ref{fig:coalesce} (left) illustrates how a program's tasks execute and shows how task transitions can affect run time.
 At each transition, the system incurs an overhead to track and atomically commit modifications to the non-volatile memory, to maintain consistency of program state~\cite{chain,alpaca}.
-The more task transitions a program requires, the more commit overhead is incurred by the system at runtime.
+The more task transitions a program requires, the more commit overhead is incurred by the system at run-time.
 A programmer may thus create very large tasks in an attempt to reduce task transitions overhead.
-However, a large task may require more energy to complete than a device's fixed hardware energy buffer can hold, which may lead to a task \emph{non-termination} problem (see Fig.~\ref{fig:coalesce}(right)).
+However, a large task may require more energy to complete than a device's fixed hardware energy buffer can hold, which may lead to a task \emph{non-termination} problem (Fig.~\ref{fig:coalesce}, right).
 To eliminate this risk, existing systems require the programmer to decompose a program into small tasks to preserve execution progress.
@@
 
-\textbf{Challenges and Contributions.} We introduce \sys: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
-By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overheads \emph{dynamically}, meanwhile avoiding the risk of non-termination.
+\paragraph{Challenges and Contributions.}
+We introduce \sys: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
+By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overhead \emph{dynamically}, meanwhile avoiding the risk of non-termination.
 \sys accepts any static decomposition and it coalesces (groups) tasks or splits them (see again Fig.~\ref{fig:coalesce}) based on the estimated energy availability without demanding any hardware support.
-To the best of our knowledge, \sys is the only system that guarantees forward progress for any statically-defined tasks or energy storage sizes.
+To the best of our knowledge, \sys is the only system that guarantees forward progress for any statically-defined task  set or energy storage.
 
-The unique contributions of \sys in relation to challenges of task-based systems revealed by this work are as follows:
+The unique contributions of \sys in relation to challenges of task-based systems revealed by this work are listed below.
 
@@
 The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
-Section~\ref{sec:systemdescription} provides an overview of \sys, while Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
+Section~\ref{sec:systemdescription} illustrates an overview of \sys.
+Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
 The dynamic page privatization is explained in Section~\ref{sec:memory_virtulaization}.

### 2018-08-07 paper/asplos19/coalescing.tex
coalescing revised

@@
 
-When a device's buffered energy is sufficient to run multiple consecutive tasks without a power failure, committing state after each task is an unnecessary overhead.
+When a device's buffered energy is sufficient to run multiple tasks without a power failure, committing state after each task is an unnecessary overhead.
 \sys eliminates this overhead by coalescing a sequence of tasks and deferring the commit
@@
 Let us assume $x=1$.
-If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive power failures.
+If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive power failures, before being able to progress.
 

### 2018-08-07 paper/asplos19/background.tex
Final pass on sections 2 and 3 (Carlo)

@@
 
-Energy harvesting devices operate using energy extracted from ambient sources, e.g.
-radio frequency transmissions, light.
+Energy harvesting devices operate using energy extracted from ambient sources, e.g., radio frequency transmissions, light, etc.
 These devices elide tethered power and batteries, instead collecting energy into a capacitor, operating when sufficient charge accumulates, and turning off for recharge upon depletion of the buffer.
-There are several energy harvesting battery-less hardware platforms.
+There are several energy harvesting battery-less platforms.
 For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
 Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
-
-\textbf{Target Hardware of \sys.} \sys is designed for the demands of existing and future \emph{tiny embedded energy-harvesting platforms} based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
+%
+% \textbf{Target Hardware of \sys.}
+\sys is designed for the demands of existing and future \emph{tiny embedded energy-harvesting platforms} based around general purpose, commodity computing components~\cite{wisp,msp430datasheet}.
 \sys targets a device with a memory system that has fast, byte-addressable volatile and non-volatile memory; in particular, our target platform, WISP~\cite{wisp}, is equipped with a mixture of SRAM and FRAM.
@@
 \label{sec:background_consistency}
-
-\begin{figure}
-    \begin{subfigure}{0.49\columnwidth}
-        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
-        \caption{WAR-affected code}
-        \label{fig:war-example}
-    \end{subfigure}
-    \begin{subfigure}{0.49\columnwidth}
-        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
-        \caption{WAR-affected code execution}
-        \label{fig:war-execution}
-    \end{subfigure}
-    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, while \texttt{checkpoint()} is a checkpoint of volatile state.}
-    \label{fig:war}
-\end{figure}
 
@@
 
+\begin{figure}
+    \begin{subfigure}{0.49\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/war-example.pdf}
+        \caption{WAR-affected code}
+        \label{fig:war-example}
+    \end{subfigure}
+    \begin{subfigure}{0.49\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/war-execution.pdf}
+        \caption{WAR-affected execution}
+        \label{fig:war-execution}
+    \end{subfigure}
+    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, \texttt{checkpoint()} is a checkpoint of volatile state.}
+    \label{fig:war}
+\end{figure}
+
 \subsection{Task-based Intermittent Programming}
@@
 
-\textbf{Incorrectly Sized Tasks Problem.} Compiler-inserted checkpoints or programmer-defined tasks can be both non-terminating and/or inefficient.
+\paragraph{Incorrectly Sized Tasks.}
+Compiler-inserted checkpoints or programmer-defined tasks can be both non-terminating and/or inefficient.
 If a task (or code between two checkpoints) consumes more than the fixed, maximum energy that the device can buffer, then the task will never be able to complete using buffered energy only.
@@
 If the task consumes far less energy than a device can buffer, the system may operate {\em inefficiently}, saving the program state more often than needed.
-Avoiding excessively costly, non-terminating, tasks and short, high-overhead, tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.
+Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is challenging, because estimating the exact energy use of an arbitrary code is complicated.
 Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse because a large task on one device may be a relatively short task on a device with a larger buffer.

### 2018-08-07 paper/asplos19/systemdescription.tex
Final pass on sections 2 and 3 (Carlo)

@@
 
-\textbf{Programming and Execution Model.}  To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate task-shared data.
-Then, he compile the task-based code, and link to \sys's runtime, producing a \sys-enabled binary.
+\paragraph{Programming and Execution Model.}
+To use \sys, a programmer must (i) convert a plain C code into tasks by encapsulating the code in a top-level set of functions, (ii) sequence the control-flow between these tasks, and (iii) annotate memory accesses that manipulate task-shared data.
+Then, they compile and link the code against \sys's runtime, producing a \sys-enabled binary.
 The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
-Facilitating efficient task adaptation requires dynamic memory protection which is \sys's \emph{virtual memory manager} through a process we call \emph{dynamic page privatization}.
+Facilitating efficient task adaptation requires dynamic memory protection which \sys's \emph{virtual memory manager} handles through a process we call \emph{dynamic page privatization}.
 
-\textbf{Adaptive Task Scheduler.} \sys's adaptive task scheduler (ATS) makes \emph{energy-aware} scheduling decisions to group tasks together or split a task.
-By coalescing tasks \sys \emph{amortizes commit and transition costs}, and by a task splitting, after repeated failures on a single task, it \emph{avoids the task non-termination problem}.
+\paragraph{Adaptive Task Scheduler.}
+\sys's adaptive task scheduler (ATS) makes \emph{energy-aware} scheduling decisions to group tasks together or split a task.
+By coalescing tasks \sys \emph{amortizes commit and transition costs}, and by splitting a task, after it repeatedly failed to complete, it \emph{avoids the task non-termination problem}.
 The scheduler uses its recent execution history---i.e., it is hardware independent---as a metric to estimate energy availability and eventually to decide on the coalesced task size.
@@
 
-\textbf{Virtual Memory Manager.} \sys can coalesce tasks thanks to its virtual memory manager (VMM).
-VMM privatizes a group of memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.
+\paragraph{Virtual Memory Manager.}
+\sys can coalesce tasks thanks to its virtual memory manager (VMM).
+VMM privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.
 VMM keeps all page modifications in non-volatile memory on a coalesced task transition.

### 2018-08-07 paper/asplos19/hwswsetup.tex
Last pass on section 7 (Carlo)

@@
 %
-Every benchmark listed in Section~\ref{sec:evaluation} was run repeatedly on
+Every benchmark used in Section~\ref{sec:evaluation} was run repeatedly on
 each platform for a few minutes to ensure capturing a diverse power 
@@
 
-The WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
+WISP contains the MSP430FR5969~\cite{wolverine} MCU with 64\,kB of
 non-volatile (FRAM) memory and 2\,kB of volatile (SRAM) memory and was
@@
 %
-We powered the WISP using an RF signal generator emitting a 20\,dBm sinusoidal wave at 915\,MHz.
+We powered WISP using an RF signal generator emitting a 20\,dBm sinusoidal wave at 915\,MHz.
 The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
@@
 %
-The antenna was oriented towards and in parallel with the WISP antenna, and
+The antenna was oriented towards and in parallel with WISP's antenna, and
 no objects obstructed the path.
 %
-We affixed the WISP with a paper harness at the edge of a table at a height, from the table suface, of 10\,cm.
+We affixed WISP with a paper harness at the edge of a table at a height, from the table suface, of 10\,cm.
 %
 For distance-controlled experiments we positioned WISP at $d=\{15, 30,
-50\}$\,cm from the exciter antenna and 60\,cm was used for the task downscaling
-experiments.
+50\}$\,cm from the exciter antenna.
 %
@@
 under profile, and the Saleae~\cite{saleae} logic analyzer measured
-the interval between edges in the signal. When the board is continously powered the 
+intervals between edges in the signal. For continous power experiments, the 
 execution time was measured using the clock features in TI Code Composer Studio IDE version 7.1.
@@
 		\hline
-        \textbf{cuckoo} &14 &426 & inserts and queries a sequence of pseudo-random
-values in a Cuckoo filter\\
+                \textbf{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
 		\hline
@@
 We evaluated \sys using six benchmarks that are often used in embedded systems (summarized in Table~\ref{table:benchmark_table}).
-All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 toolchain with \texttt{-O1} as optimization flag.
+All applications were compiled using MSP430 GCC~\cite{ti-gcc} version 6.4.0 with \texttt{-O1} as optimization flag.
 The source code for all benchmarks will be released at \cite{coala_website}.
@@
 For a fair comparison the task decomposition of the benchmarks are ensured to be the same for both systems.
-
+%
 Since Alpaca's compiler pass is implemented only for LLVM, we could not use that implementation to instrument the benchmarks and compile them with GCC.
@@
 The instrumentation consists of identifying WAR dependencies  and adding code and memory allocations to make a private copy of the affected variables.
-For array variables, instrumentation also includes adding code for dynamically tracking the array access.
-
+% For array variables, instrumentation also includes adding code for dynamically tracking the array access. 
+%
 %Manual instrumentation of Alpaca code is not a general approach, but only served as a practical alternative to porting the compiler pass from LLVM to GCC.

### 2018-08-07 paper/asplos19/intro.tex
Final pass on section 4 (Carlo) and new figure 5
change memory manager to page privatizer

@@
 This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts?
-\sys addresses this challenge by using a novel approach called \emph{dynamic page privatization}---performing real-time dependency tracking to enable protection on a task transition.
+\sys addresses this challenge by using a novel approach called \emph{page privatization}---performing real-time dependency tracking to enable protection on a task transition.
 Individual variables tracking, however, slows down a system dramatically.
@@
 Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
-The dynamic page privatization is explained in Section~\ref{sec:memory_virtulaization}.
+The page privatization is explained in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are given in Section~\ref{sec:implementation}.

### 2018-08-07 paper/asplos19/paging.tex
change memory manager to page privatizer

@@
-%\note{additional motivation for coala memory virtualization] we should emphasize that Coala programming model removes the burden of identifying the write-after-read dependencies from the programmer  as compared to Chain.
-And it does not require a specific compiler that cannot consider a runtime dependent behavior like Interrupts service handling to guarantee memory protection, as in Alpaca.}
-
 \begin{figure}
@@
 
-\sys's {\em virtual memory manager} (VMM) is a core enabler of \sys task adaptation mechanism.
-Its necessity is illustrated by example given in Figure~\ref{figure:coalescedWar}: resolving WAR dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent with task coalescing.
-Since task coalescing induces dynamic WAR dependencies,  a new memory consistency handling technique is mandatory so that these dependencies are resolved and handled at runtime.
+Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when task are coalesced (see Fig.~\ref{figure:coalescedWar}).
+Therefore, \sys introduces its {\em page privatizer} (PP) which enables efficient task coalescing.
 
-VMM overview is given in Figure~\ref{figure:mem-man}.
-VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
+PP overview is given in Fig.~\ref{figure:mem-man}.
+PP abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
 \texttt{\underline{private}} holds the consistent version of pages after each commit (and on reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
-The VMM prohibits applications from directly accessing these buffers.
+The PP prohibits applications from directly accessing these buffers.
 Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
 It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
-When a coalesced task completes, the VMM ensures that {\em dirty} (i.e., modified) pages are committed to FRAM and visible to subsequent tasks.
+When a coalesced task completes, the PP ensures that {\em dirty} (i.e., modified) pages are committed to FRAM and visible to subsequent tasks.
 If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
-% If the power interrupts a task, the VMM ensures the consistency of the data by relying on its \texttt{\underline{private}} buffer.
-The VMM ensures consistency even if a task or a sequence of tasks include WAR dependences that could compromise consistency~\cite{ratchet,dino}.
-On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
+On a commit, the PP atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
 % SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM.
@@
     \includegraphics[width=\columnwidth]{figures/mem-man.pdf}
-    \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP}, and order of occurrence. Steps 2, 3, 4 and 5 are conditional.}
+    \caption{Interaction between task, page privatizer and memory buffers upon \texttt{RP} and \texttt{WP}, and order of occurrence. Steps 2, 3, 4 and 5 are conditional.}
     \label{figure:mem-man}
 \end{figure}
-
-%\textbf{System Overview.} Figure~\ref{figure:mem-man} gives an overview of \sys's virtual memory management system.
-A page has a home location in non-volatile memory which we refer to as the page's \texttt{\underline{private}}.
-When a task accesses a page, the VMM swaps the page from its home location into the fully\hyp{}associative \texttt{working} buffer that is stored in volatile memory.
-The \texttt{working} buffer is allocated to take up the volatile memory that remains after space for stack is allocated.
-
-%The VMM utilizes a two-phase commit to write dirty pages from the \texttt{working} buffer back to their home location in \texttt{\underline{private}}.
-It optimizes this process by using an indirection table to abstract the location of the \texttt{\underline{private}} buffer and to eliminate the need for the second page copy.
-The VMM is forced to swap out a dirty page from the \texttt{working} buffer when a task requests a page that is not in the \texttt{working} buffer, and there is no free page in the working buffer (i.e., a page fault).
-On a page fault, the VMM moves the page to the direct\hyp{}mapped non-volatile \texttt{\underline{shadow}} buffer.
 
@@
 
-After address translation (Figure~\ref{figure:mem-man}, Step 6), a task accesses the protected variable's location in the volatile \texttt{working} buffer (Step 7).
-The VMM keeps track of the page tags for the pages currently resident in the \texttt{working} buffer.
+After address translation (Fig.~\ref{figure:mem-man}, Step 6), a task accesses the protected variable's location in the volatile \texttt{working} buffer (Step 7).
+The PP keeps track of the page tags for the pages currently resident in the \texttt{working} buffer.
 When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt working} (Algorithm~\ref{algo:rwar}, Line 2).
@@
 
-When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory manager first searches the variable's page in the \texttt{working} buffer (Figure~\ref{figure:mem-man}, Step 1).
+When accessing a protected variable with \texttt{RP} or \texttt{WP}, the PP first searches the variable's page in the \texttt{working} buffer (Fig.~\ref{figure:mem-man}, Step 1).
 If the page is not found there, a page fault is incurred and a new page needs to be swapped in.
-If the \texttt{working} buffer is full, a page fault on a memory access requires the VMM to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
-If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes have to be persisted to non-volatile memory (Figure~\ref{figure:mem-man}, Step 2).
+If the \texttt{working} buffer is full, a page fault on a memory access requires the PP to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
+If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes have to be persisted to non-volatile memory (Fig.~\ref{figure:mem-man}, Step 2).
 If the accessed page was previously modified and swapped out since the last power failure, the most recent version of the page is in {\tt \underline{shadow}} (Step 3), otherwise it has to be retrieved from {\tt \underline{private}} (Step 4).
@@
 The \texttt{\underline{committing}} bit indicates that a commit is in progress and is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16).
-The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to commit and the VMM clears the counter when commit completes (Line 14).
-The \texttt{\underline{commitIndex}} indexes the next page to be committed (Line 11) and the VMM clears the index at the end of the phase (Line 15).
+The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to commit and the PP clears the counter when commit completes (Line 14).
+The \texttt{\underline{commitIndex}} indexes the next page to be committed (Line 11) and the PP clears the index at the end of the phase (Line 15).
 
@@
 % \sys's implementation is also efficient because
-the VMM does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.
+the PP does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.
 

### 2018-08-07 paper/asplos19/systemdescription.tex
change memory manager to page privatizer

@@
 	\includegraphics[width=\columnwidth]{figures/system-overview.pdf}
-	\caption{\sys top-level view. ATS: adaptive task scheduler, VMM: virtual memory manager, PM: physical memory.}
+	\caption{\sys top-level view. ATS: adaptive task scheduler, PP: Page Privatizer, PM: physical memory.}
 	\label{fig:system_overview}
@@
 The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
-Facilitating efficient task adaptation requires dynamic memory protection which \sys's \emph{virtual memory manager} handles through a process we call \emph{dynamic page privatization}.
+Facilitating efficient task adaptation requires dynamic memory protection which is what \sys's \emph{page privatizer} does.
 
@@
 
-\paragraph{Virtual Memory Manager.}
-\sys can coalesce tasks thanks to its virtual memory manager (VMM).
-VMM privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.
-VMM keeps all page modifications in non-volatile memory on a coalesced task transition.
-Section~\ref{sec:memory_virtulaization} outlines VMM.
+\paragraph{Page Privatizer.}
+\sys can coalesce tasks thanks to its pager privatization (PP) technique.
+PP privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.
+PP keeps all page modifications in non-volatile memory on a coalesced task transition.
+Section~\ref{sec:memory_virtulaization} outlines the PP mechanism.

### 2018-08-07 paper/asplos19/coalescing.tex
coalescing revised

@@
 
-When a device's buffered energy is sufficient to run multiple tasks without a power failure, committing state after each task is an unnecessary overhead.
-\sys eliminates this overhead by coalescing a sequence of tasks and deferring the commit
+When a device's buffered energy is sufficient to run multiple tasks without a power failure, committing state after each task is unnecessary overhead.
+\sys reduces this overhead by coalescing a sequence of tasks and deferring commit
 operations for all tasks to the end of the sequence.
 In general, committing involves moving state manipulated by a task into its permanent location in non-volatile memory.
-In particular, \sys's commit procedure copies dirty pages of memory that a task updated between fast, volatile working memory and slower, non-volatile main memory.
+In particular, \sys's commit procedure copies dirty pages of memory that a task updated from fast, volatile working memory to slower, non-volatile main memory.
 We defer the details of paging privatization and commit to Section~\ref{sec:memory_virtulaization}.
 An effective coalescing strategy must be {\em aggressive} enough, attempting to coalesce a large number of tasks to amortize commit overhead.
-However, it must also be  {\em conservative} enough, attempting to coalesce only as many tasks as will execute to completion given certain energy conditions, reducing the risk or long coalesced task re-execution penalty.
+However, it must also be  {\em conservative} enough, coalescing only as many tasks as will execute to completion given certain energy conditions, reducing the risk of re-execution penalty for long coalesced tasks.
 
@@
         \State $C \leftarrow $ \Call{$f_\text{reboot}$}{$z$} \Comment{Update coalescing target by reboot update function}
-        \State $H \gets 0$ \Comment{Initialize history: total number of tasks executed since last power failure}
+        \State $H \gets 0$ \Comment{Initialize history}
         \While{$\texttt{true}$}       
@@
 	        \State \Call{commit\_to\_fram}{\null}
-	        \State $C \leftarrow $ \Call{$f_\text{commit}$}{$C$}
+                \State $C \leftarrow $ \Call{$f_\text{commit}$}{$C$} \Comment{Update coalescing target by commit update function}
         \EndWhile
@@
 Different coalescing strategies adhere mainly to the template in Algorithm~\ref{algo:genCoalescing}, varying in only a few {\em characteristic operations} that the algorithm leaves deliberately abstract.
-A strategy's reboot update function, $f_\text{reboot}$, updates $C$, the coalescing target, after a reboot.
-A strategy's weight lookup function, $f_\text{weight}$, returns a task's weight.
-A strategy's commit update function, $f_\text{commit}$, updates $C$ after a successful commit.
-The following paragraphs detail several possible coalescing strategies that we developed for \sys, albeit not the only possible.
-In fact, \sys \emph{allows programmers to design their own strategy} to implement and link to \sys's task scheduler.
+The reboot update function, $f_\text{reboot}$, updates $C$, the coalescing target, after a reboot.
+The weight lookup function $f_\text{weight}$ returns a task's weight.
+The commit update function $f_\text{commit}$ updates $C$ after a successful commit.
+The following paragraphs detail three coalescing strategies that we developed for \sys, albeit not the only possible.
+In fact, \sys \emph{allows programmers to design their own strategy} to implement and link against \sys's task scheduler.
 
@@
 
-A very simple {\em energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of a coalesced task linearly.
+A very simple {\em energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly.
 In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits, and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure.
@@
 %
-The energy-oblivious strategy reacts slowly to the variation in the energy required to execute different tasks and to the variation in the effective quantum of energy available to the device.
-With the successful commit of each coalesced task sequence, the coalescing target increases.
+EO reacts slowly to the variation in the energy required to execute different tasks and to the variation in the effective quantum of energy available to the device.
+With the successful commit of each coalesced task sequence, $C$ increases.
 Eventually, the target may be too high and only a coalesced task composed of a few units will commit without interruption by a power failure.
@@
 By relying on the history of execution EG eliminates the problem of frequent power failures on a single coalesced task.
-In fact, at any reboot, EG will {\em conservatively} try to coalesce only a fraction ($\rho$) of tasks that had successfully completed during the last power cycle.
+In fact, at each reboot, EG will {\em conservatively} try to coalesce only a fraction ($\rho$) of tasks that had successfully completed during the last power cycle.
 % In fact EG can reduce its target to one from any coalesced target in at most two successive power failures.
@@
 We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
-Figure~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$.
+Fig.~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$.
 The figure shows two power cycles following the $i$th, the latter having an execution history $H^{(i)} = 4$.
 Based on that, $C$ is initially set to two and then decreases to one.
-Once the value of $C$ reaches one, EG is, in effect, expecting a power failure, justifying the conservative approach.
-After the power failure, EG uses the most recent history $H^{(i+1)} = 8$ to set $C = 4$ and continue execution.
+Once the value of $C$ reaches one, EG is expecting a power failure, justifying the conservative approach.
+After the reboot, EG uses the most recent history $H^{(i+1)} = 8$ to set $C = 4$ and continue execution.
 
@@
 The \emph{Weighted Energy-Guided} (WEG) coalescing strategy accounts for non-uniform energy and time costs of a program's tasks when setting $C$.
-Each different task in the program consumes a different amount of energy to execute to completion.
+Each different task in the program consumes a different amount of energy to run to completion.
 The EO and EG coalescing strategies assume that each task has the same cost: for these strategies, $C$ simply corresponds to a target {\em count} of tasks to coalesce, regardless of the individual cost of each task.

### 2018-08-07 paper/asplos19/partialcommit.tex
coalescing revised

@@
+% After identifying a task as likely non-terminating, \sys must decide when during the task's execution to partially commit the task's state.
+\sys sets a timer at the start of the likely non-terminating task, initializing it to a very large value (e.g., an estimate of the maximum execution time using the device's energy buffer).
+If the timer expires before power fails, \sys partially commits, retaining  the timer value to use for future partial commits.
+If the timer does not expire before power fails, \sys halves the timer and tries again.
+The exponential decrease of the timer value converges rapidly to a usable one.
+\label{sec:task_downsizing}
 \subsection{Task Downscaling}
-\label{sec:task_downsizing}
-
 \sys uses {\em task downscaling} to make progress through a task that is too large to complete using the energy that the device can buffer.
@@
 Task downscaling violates task atomicity in a non-terminating execution, favoring continued progress over atomicity.
-If a program requires a task's atomicity to be preserved, they can disable task downscaling for that task.
-
+If a programmer requires a task's atomicity to be preserved, they can disable task downscaling for that task.
 The key design issue for task downscaling is deciding when to partially commit.
 \sys only applies task downscaling to a task that is likely to be non-terminating.
-A task is likely to be non-terminating if it fails to execute completely twice consecutively.
-The second incomplete execution executes after a power failure, when the device will have fully recharged its energy buffer.
+A task is likely to be non-terminating if it fails to run to completion twice consecutively.
+The second incomplete run executes after a power failure, when the device will have fully recharged its energy buffer.
 If a task cannot complete when executing with a full energy buffer, the task is marked as non-terminating.
-
-% After identifying a task as likely non-terminating, \sys must decide when during the task's execution to partially commit the task's state.
-\sys sets a timer at the start of the likely non-terminating task, initializing it to a very large value (e.g., an estimate of the maximum execution time using the device's energy buffer).
-If the timer expires before power fails, \sys partially commits, retaining  the timer value to use for future partial commits.
-If the timer does not expire before power fails, \sys halves the timer and tries again.
-The exponential decrease of the timer value converges rapidly to a usable one.

### 2018-08-07 paper/asplos19/conclusions.tex
update conclusions

@@
 When less energy is available, progress is latched at sub-task granularity.
-\sys's virtual memory system ensures that program state in non-volatile memory remains consistent and amortizes the cost of state transfer between volatile and non-volatile memory.
+\sys's page privatization system ensures that program state in non-volatile memory remains consistent and amortizes the cost of state transfer between volatile and non-volatile memory.
 Our evaluation across applications on a real embedded energy-harvesting device demonstrates the utility of \sys, as well as the practicality of the proposed implementation.

### 2018-08-07 paper/asplos19/abstract.tex
revisions of abstract and intro (Sinan's comments applied)

@@
 \sys's specialized memory virtualization mechanism ensures that power failures do not leave the program state in non-volatile memory inconsistent.
-Our evaluation on a real energy-harvesting platform not only shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, but also it is able to \emph{progress} where static systems \emph{fail}.
+Our evaluation on a real energy-harvesting platform not only shows that \sys reduces run time by up to 54\% as compared to a state-of-the-art system, but also it is able to \emph{progress} where static systems fail.

### 2018-08-07 paper/asplos19/intro.tex
revisions of abstract and intro (Sinan's comments applied)

--- 2018-08-07 paper/asplos19/intro.tex
@@
     \caption{Task coalescing reduces commit overhead (left), while task splitting enables termination of large tasks (right).
-    S: static legacy task-based runtime, C: \sys.}
+    S: static legacy task-based runtime, C: \sys (this work).}
     % \caption{Task coalescing reduces time and energy overhead by performing fewer commits (denoted as \texttt{C}) than the individual tasks (denoted as \texttt{Tx}) commits (left).
@@
 \sys accepts any static decomposition and it coalesces (groups) tasks or splits them (see again Fig.~\ref{fig:coalesce}) based on the estimated energy availability without demanding any hardware support.
-To the best of our knowledge, \sys is the only system that guarantees forward progress for any statically-defined task  set or energy storage.
+To the best of our knowledge, \sys is the only system that eliminates restructuring and re-compilation of applications considering device's energy buffer---enabling energy-storage independency for the applications, while keeping execution efficient.
 
@@
 This leads to a third challenge: how to enable the dynamic execution model to progress on a sub-task level?
-To avoid non-termination under adverse energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism, inspired by~\cite{ratchet}.
+To avoid non-termination under adverse energy conditions, \sys uses a timer-based {\em partial task commit} mechanism.
 Partial execution avoids non-termination by committing the intermediate state of a long-running task that has repeatedly failed and restarted.
@@
 To asses the benefits of \sys over existing task-based systems, we implemented and tested six benchmarks on a real energy-harvesting platform.
-Our evaluation not only shows that \sys reduces run time by up to 54\%, and by 26\% on average, but also is able to continue the execution where static task-based systems fail.
+Our evaluation shows that \sys reduces run time overhead by up to 54\% and solves task non-termination problem where existing static task-based systems fail.
 
@@
 Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
-The dynamic page privatization is explained in Section~\ref{sec:memory_virtulaization}.
+The page privatization is explained in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are given in Section~\ref{sec:implementation}.
-Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results.
-Section~\ref{sec:related_work} positions \sys in the context of related work and Section~\ref{sec:conclusions} concludes and discusses future work.
+Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results, respectively.
+Section~\ref{sec:related_work} positions \sys in the context of related work and Section~\ref{sec:conclusions} concludes our work.

### 2018-08-07 paper/asplos19/background.tex
revisions of background

@@
     \end{tabular}
-    \caption{Non-volatile memory access for data consistency; \emph{PC}: program counter, \emph{WAR variables}: variables involved in Write after Read (WAR) dependences, \emph{NV}: non-volatile.}
+    \caption{Non-volatile memory access for data consistency; {PC}: program counter, {WAR variables}: variables involved in Write-after-Read (WAR) dependencies, {NV}: non-volatile.}
     \label{table:chechpoint_comparison}
@@
 Moreover, when \emph{heterogeneous devices} with different energy buffers (e.g., 20 \si{\micro\farad}~\cite{rodriguez_tbcs_2015} to 0.1 \si{\farad}~\cite{moo}) are considered, the problem of porting a program becomes much worse because a large task on one device may be a relatively short task on a device with a larger buffer.
-\textbf{\sys's execution model} solves this dilemma by introducing {\em task coalescing}, which merges multiple tasks to amortize the overhead when tasks are too small, and {\em task downscaling}, which breaks up a task when it is too large to complete.
+\sys's \emph{execution model} solves this dilemma by introducing {\em task coalescing}, which merges multiple tasks to amortize the overhead when tasks are too small, and {\em task downscaling}, which breaks up a task when it is too large to complete.

### 2018-08-07 paper/asplos19/evaluation.tex
Evaluation revised
Final pass on section 8 (Carlo)

@@
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/coalEfficiency}
-    \caption{Breakdown of overhead for the proposed coalescing strategies.
-NC: no coalescing, EO: energy-oblivious coalescing, EG: Energy-guided coalescing, WEG:
-energy- and task-aware coalescing. All coalescing strategies reduce total
-overhead and maximize useful work. EG and WEG perform better than EO.}
-	\label{fig:overallOverheadBreakdown}
-\end{figure}
 % \end{wrapfigure}
 %
-Building on top of the existing work, we have selected the most efficient task-based run-time library~\cite{alpaca} and compared \sys against it.
-Our evaluation quantitatively demonstrates that \sys (i) \emph{reduces the memory
-protection overhead} by coalescing, (ii) \emph{improves execution speed}, in most cases; %
-and (iii) \sys is able to progress where static systems suffer from a task non-termination loop. 
-We have also broken down \sys's overhead to explain the source of \sys's improved performance.
+%Building on top of the existing work, we have selected the most efficient task-based run-time library~\cite{alpaca} and compared \sys against it.
+Our evaluation quantitatively demonstrates that \sys (i) \emph{reduces memory
+protection overhead}, (ii) \emph{improves execution speed} in most cases,
+and (iii) \emph{is able to progress} where static systems suffer from a task non-termination loop. 
 %
@@
 
-\subsubsection{Overhead Reduced by Coalescing}
+To characterize Coala's overhead we experimented with WISP positioned at
+15\,cm away from the signal generator antenna.
+We have broken down \sys's overhead to explain the source of its improved performance.
+
+\paragraph{Overhead Reduced by Coalescing.}
 \label{sec:overhead-coalescing}
+
+\begin{figure}
+    \centering
+    \includegraphics[width=\columnwidth]{figures/coalEfficiency}
+    \caption{Breakdown of overhead for the proposed coalescing strategies:
+    NC (no coalescing), EO (energy-oblivious), EG (energy-guided), WEG
+    (weighted energy-guided). All coalescing strategies reduce total
+    overhead and maximize useful work. EG and WEG perform better than EO.}
+    \label{fig:overallOverheadBreakdown}
+\end{figure}
 
@@
 and WEG) and for a baseline without coalescing (NC), we have measured the time
-spent on (i) executing task code (ii) re-executing task code after a power failure, and
-(iii) committing state to non-volatile memory at the end of each (coalesced)
-task.  In this experiment the WISP was 15\,cm away from the signal generator
-antenna.
-
+spent on executing (i) useful task code, (ii) task code wasted due to a power failure, and
+(iii) commits to non-volatile memory at the end of each (coalesced)
+task.
+%
 The overhead incurred by each coalescing strategy is broken down in
@@
 re-execution penalty is smallest, because the amount of work that can happen
-between commits and may have to be re-executed if interrupted is smallest when
+between commits and may have to be re-executed if interrupted is reduced when
 work from multiple static tasks is not combined.
@@
 to the EO strategy. We attribute this discrepancy to EO's slow adjustment
-of the target task size without regard to the energy conditions.
+of the coalescing target without regard to the energy conditions.
 %
-In the subsequent experiments, we focus on the better-performing EG and WEG
-strategies.
+In the subsequent experiments, we focus on the better-performing EG and WEG.
+
+\paragraph{\sys's Kernel Overhead.}
 %
-\begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/overallOverhead}
-	\includegraphics[width=\columnwidth]{figures/memAccess}
-	\caption{\sys kernel overhead: a breakdown of time spent in
-the application and in the runtime (top) and a breakdown of protected memory
-accesses into three types with different costs. NC: no coalescing, EG:
-Energy-guided coalescing, WEG: weighted Energy-guided coalescing.}
-	\label{fig:coalEfficiency}
-\end{figure}
+Fig.~\ref{fig:kernel-overhead} breaks down the time spent on executing 
+user task code versus the time spent on kernel operations.
+When coalescing is not enabled the commit overhead is highest.
 %
-\subsubsection{\sys Kernel Overhead Breakdown}
-%
-Fig.~\ref{fig:coalEfficiency} (top) breaks down the time spent on executing 
-task code and committing further into time spent by the application and
-time spent by \sys's runtime. When coalescing is not enabled the commit overhead is 
-highest.
-%
-However, coalescing increases the cost of each access to a protected variable,
-since VMM has to spend more on average time in searching for the requested page.
+The overhead for memory accesses increases in percentage when enabling coalescing,
+but not in absolute terms.
 %
@@
 necessary on each protected access is the most critical bottleneck for \sys.
+
+\begin{figure}
+    \begin{subfigure}{\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/overallOverhead.pdf}
+        \caption{Kernel overhead breakdown}
+        \label{fig:kernel-overhead}
+    \end{subfigure}
+    \begin{subfigure}{\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/memAccess.pdf}
+        \caption{Protected memory accesses breakdown}
+        \label{fig:mem-accesses}
+    \end{subfigure}
+    \caption{\sys's internal overhead. NC: no coalescing, EG:
+    energy-guided, WEG: weighted energy-guided.}
+    \label{fig:war}
+\end{figure}
+
+\paragraph{Protected Memory Accesses Breakdown.}
 %
-\subsubsection{Protected Memory Accesses Breakdown}
-%
-Fig.~\ref{fig:coalEfficiency} (bottom) breaks down protected memory accesses
+Fig.~\ref{fig:mem-accesses} breaks down protected memory accesses
 into three categories.
 %
-Each type of protected access incurs a different amount of overhead: access to
-the same page in SRAM as the previous access is cheapest, access to a different
-SRAM page has a larger cost, and access to a page that needs to be swapped in
-from FRAM into SRAM is most expensive.
+Each type of protected access incurs a different.
+Accessing the most recently used SRAM page is of the cheapest kind.
+Accessing a different page in SRAM has a slightly higher cost.
+Finally, accessing a page that needs to be swapped in
+from FRAM into SRAM is the most expensive.
 %
 The results in the figure show that the overwhelming majority of accesses are
-of the cheapest kind, which motivated us to optimize this access type in our
-implementation (see Section~\ref{sec:impl:paging}).
+of the cheapest kind, which motivated us to optimize this accesses in our
+implementation.
 %
@@
 %
-\subsection{Execution time}
+\subsection{Execution Time}
 \label{sec:result_coalescing}
@@
 overhead, we now investigate the outcome of this reduction on the total
-execution time. We first investigate difference variants of \sys and
+execution time. We first investigate different variants of \sys and
 then compare the best variant to Alpaca~\cite{alpaca}.
-%
-\subsubsection{\sys Speedup with Coalescing}
 
-Fig.~\ref{fig:coalescing} shows the \sys run time with two coalescing
+\paragraph{Speedup with Coalescing.}
+
+Fig.~\ref{fig:coalescing} shows \sys's run time of two coalescing
 strategies (EG, WEG) normalized to the run time without coalescing (NC).
 %
-The results in Fig.~\ref{fig:coalescing} show that all benchmarks
-complete faster with coalescing than without coalescing: from 25\% faster
-(\textit{ar}) up to 70\% faster (\textit{sort}).
+The results show that all benchmarks
+complete faster with coalescing than without coalescing: from 25\%
+(\textit{ar}) up to 70\% (\textit{sort}).
 %
@@
 %
-In some cases (\textit{bc}, \textit{cuckoo}, \textit{sort}) the Weight Energy-guided
-feature in the coalescing algorithm is counter-productive.
+In some cases (\textit{bc}, \textit{cuckoo}, \textit{sort}) WEG's task weighting
+system is counter-productive.
 %
-This occurs in task decompositions with a uniform task energy, where counting
-tasks disregarding their energy provides an equal amount of information with a
-smaller effort.
+This occurs in task decompositions with energy-uniform tasks, where counting
+tasks disregarding their energy consumption provides an equal amount of information
+with a smaller effort.
 %
-In \textit{fft}, tasks are not uniform, and accounting for task energies is
+In \textit{fft}, tasks are not uniform, and accounting for their different weights
 beneficial. In fact, the lack of task energy awareness is detrimental: with EG
@@
 The speedup is highest for \textit{bc}, \textit{cuckoo}, \textit{dijkstra} and
-\textit{sort}, because their tasks are relatively small and are easily coalesced,
-eliminating many unnecessary commits.
+\textit{sort}, because their tasks are relatively small and are easily coalesced.
 %
@@
     \caption{\sys's coalescing performance. Application execution time
-with coalescing strategies (EG and WEG) normalized to the execution time
-without coalescing.}
+with coalescing (EG and WEG) normalized to the execution time
+without coalescing (NC).}
 \label{fig:coalescing}
@@
     \includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
-    \caption{\sys's execution time, normalized, where possible, to the execution time 
-    with Alpaca, as a function of distance to the energy source 
-    measured in meter.}
+    \caption{\sys's execution time normalized to Alpaca's,
+    for three distances from the energy source (15, 30 and 50\,cm).}
     \label{fig:runtime}
@@
 
-\subsubsection{Benefits of Adaptive Tasks}
+\paragraph{Benefits of Adaptive Tasks.}
 %
 We now compare \sys's performance to Alpaca~\cite{alpaca}---a
-\emph{non-adaptive} task-based system with tasks fixed at compile time.
+\emph{non-adaptive} task-based system with tasks fixed at compile-time.
 Fig.~\ref{fig:runtime} shows the average execution time of each
-application for \sys and Alpaca, normalized, where possible, to the latter. The results show
-that \sys provides a performance benefit compared to Alpaca for most
-applications. The speedup is greatest for applications with repeated write after
-read dependencies throughout their code, particularly involving arrays
+application for \sys and Alpaca, normalized, where possible, to the latter.
+\sys provides a performance benefit compared to Alpaca for most
+applications. The speedup is greatest for applications with repeated WAR
+dependencies throughout their code, particularly involving arrays
 (\textit{dijkstra}, \textit{fft} and \textit{sort}). \sys's VMM
@@
 
-Due to its static progressing behavior Alpaca was unable to finish the \textit{fft} benchmark on a distance larger than 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
+Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
 execution is negligible and stored energy is insufficient to complete some of
@@
 This is marked with $\infty$ signs in Fig.~\ref{fig:coalescing}.
-\sys, however, manged to complete \textit{fft} application by enabling its task downscaling on the distances 30\,cm and 50\,cm.
+\sys, however, manged to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.
 
@@
 We characterize the performance of \sys's virtual memory sub-system in an
-experiment on a continuously-powered evaluation board described in
-Section~\ref{sec:methodology}.  Fig.~\ref{fig:page_size} quantifies the
-effects of page size.
-%
+experiment on a continuously-powered evaluation board, as described in
+Section~\ref{sec:results_hardware_software}.
+% Fig.~\ref{fig:page_size} quantifies the effects of page size.
+
 \begin{figure}
-	\centering
-	\includegraphics[width=\columnwidth]{figures/page_exec-time}
-	\includegraphics[width=\columnwidth]{figures/pagePulls}
-    \caption{Effect of page size on execution time (top) and on the number of
-page faults (bottom). The execution time is normalized to the lowest in each set
-of page sizes.}
-	\label{fig:page_size}
+    \begin{subfigure}{\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/page_exec-time.pdf}
+        \caption{Execution time normalized to lowest per-application}
+        \label{fig:page-exec-time}
+    \end{subfigure}
+    \begin{subfigure}{\columnwidth}
+        \includegraphics[width=\columnwidth]{figures/pagePulls.pdf}
+        \caption{Number of page faults per application run}
+        \label{fig:page-pulls}
+    \end{subfigure}
+    \caption{Effect of page size (in bytes).}
+    \label{fig:war}
 \end{figure}
-%
-\subsubsection{Effect of Page Size on Runtime}
-%
-Fig.~\ref{fig:page_size} (top) shows the execution time as a function
-of page size, normalized to the lowest per-application performance among the
+
+\paragraph{Effect of Page Size on Runtime.}
+
+Fig.~\ref{fig:page-exec-time} shows the execution time as a function
+of page size (in bytes), normalized to the lowest per-application performance among the
 set of page sizes.
-The data suggest that there is a page size, distinct from the largest or
-smallest, that minimizes execution time. The best page size is not the same for
+The data suggest that there is a page size that minimizes execution time.
+The best page size is not the same for
 each application. Nevertheless, if a choice must be made for all applications,
-128\,B pages perform well for most of them.
-%
-\subsubsection{Effect of Page Size on Page Faults}
-%
-Fig.~\ref{fig:page_size} (bottom) shows the number of page faults as a
-function of page size.
+128-B pages are the best option.
+
+\paragraph{Effect of Page Size on Page Faults.}
+
+Fig.~\ref{fig:page-pulls} reports the number of page faults, per application
+run, as a function of page size (in bytes).
 %
 The smaller the page the more likely that a memory access will land
-outside that page and that a new page will need to be swapped in.
+outside that page and that a new page will have to be swapped in.
 %
-This trend is visible for all applications, exception \textit{bc}. The total
+This trend is visible for all applications, except for \textit{bc}. The total
 amount of data accessed by \textit{bc}, as well as its working set, is small.
 Even with the smallest page, all accesses are contained within that page, and
-no page faults occur. Without any page faults to begin with, increasing the
-page size result only in overhead.
+no page fault occurs. Without any page faults to begin with, increasing the
+page size only yields overhead.

### 2018-08-07 paper/asplos19/systemdescription.tex
revisions of system description

@@
 	\includegraphics[width=\columnwidth]{figures/system-overview.pdf}
-	\caption{\sys top-level view. ATS: adaptive task scheduler, PP: Page Privatizer, PM: physical memory.}
+	\caption{\sys top-level view. APP: application, ATS: adaptive task scheduler, VMM: Virtual Memory Manager, PM: physical memory.}
 	\label{fig:system_overview}
@@
 
-\sys is a new programming and execution model for intermittent operation on energy-harvesting devices.
+\sys is a new programming and execution model for intermittent operation on \emph{small embedded energy-harvesting devices}.
 It addresses the challenges given in Section~\ref{sec:background} making task-based intermittent applications {\em programmable}, {\em efficient} and {\em portable}, introducing a new programming model and runtime software system that supports adaptive task-based execution.

### 2018-08-07 paper/asplos19/systemdescription.tex
revisions of system description

@@
 
-\sys is a new programming and execution model for intermittent operation on \emph{small embedded energy-harvesting devices}.
-It addresses the challenges given in Section~\ref{sec:background} making task-based intermittent applications {\em programmable}, {\em efficient} and {\em portable}, introducing a new programming model and runtime software system that supports adaptive task-based execution.
+\sys is a new programming and execution model that supports adaptive task-based execution and eliminates restructuring and re-compilation of applications considering device's energy buffer.
+\sys addresses the challenges given in Section~\ref{sec:background} by making task-based intermittent applications portable in the sense of different energy storage sizes, while keeping their execution efficient.
 Fig.~\ref{fig:system_overview} shows an overview of \sys.

### 2018-08-07 paper/asplos19/systemdescription.tex
revisions of system description

@@
 The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
-Facilitating efficient task adaptation requires dynamic memory protection which is what \sys's \emph{page privatizer} does.
+Facilitating efficient task adaptation requires dynamic memory protection which \sys's \emph{virtual memory manager} handles through a process we call \emph{dynamic page privatization}.
 
@@
 
-\paragraph{Page Privatizer.}
-\sys can coalesce tasks thanks to its pager privatization (PP) technique.
-PP privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.
-PP keeps all page modifications in non-volatile memory on a coalesced task transition.
-Section~\ref{sec:memory_virtulaization} outlines the PP mechanism.
+\paragraph{Virtual Memory Manager.}
+\sys can coalesce tasks thanks to its virtual memory manager (VMM).
+VMM privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.
+VMM keeps all page modifications in non-volatile memory on a coalesced task transition.
+Section~\ref{sec:memory_virtulaization} outlines VMM.

### 2018-08-07 paper/asplos19/implementation.tex
VMM -> PP
Implementation revised

@@
 Equations (\ref{eq:eo}) and (\ref{eq:eg}) parametrize the behavior of the coalescing strategies in terms of $x$, $\rho$ and $\gamma$.
-We experimented with a range of values, and then used the ones that yielded the best performance: $x = 1$ (for EO) and $\rho = \gamma = 1$ (for EG/WEG).
+We experimented with a range of values, and then used the ones that yielded the best performance: $x = 1$ (for EO) and $\rho = \gamma = 1$ (for EG and WEG).
 
@@
 WEG could use the result of an arbitrarily sophisticated profiling procedure.
-To produce a concrete result in this paper, we give WEG access to a simple profile of task run time collected offline using a single fixed input.
+To produce a concrete result in this paper, we give WEG access to a simple profile of task run time (collected offline) using a single fixed input.
 WEG stores the profile in a lookup table that maps a task's identifier to its weight, making the information available to \sys's scheduler at run time.
@@
 %
-Partial commit relies on the existence of a hardware timer, which is a very
-common feature in a typical MCU and consumes very little energy (i.e.,
-$\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).
+%Partial commit relies on the existence of a hardware timer, which is a very common feature in a typical MCU and consumes very little energy (i.e., $\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).
 
@@
 
-% As described in Section~\ref{sec:memory_virtulaization}, \sys's memory
-% virtualization scheme uses three buffers: a volatile \texttt{working} buffer, a
-% non-volatile \texttt{\underline{private}}, and a non-volatile
-% \texttt{\underline{shadow}} buffer. 
-%
-% The main memory \texttt{\underline{private}} holds the last committed copy of all 
-% pages. 
-%
-% After reboot, the \texttt{working} buffer is empty. Execution fills 
-% the \texttt{working} buffer with privatized pages as the program accesses them. 
-%
-% The \texttt{\underline{shadow}} buffer serves as an intermediate,
-% non-volatile backing \texttt{\underline{private}} for dirty pages evicted from the space-constrained
-% \texttt{working} buffer.
-
-% To access a protected variable with either \texttt{RP}
-% or \texttt{WP}, \sys searches for the variable in its \texttt{working} buffer first.
- % If the page is not, the VMM fetches it from non-volatile memory. 
-%
 \paragraph{Efficiency.}
@@
 %
-Furthermore, the VMM moves pages of data efficiently using hardware accelerated DMA support.
+Furthermore, the PP moves pages of data efficiently using hardware accelerated DMA support.
 
@@
 When a page in \texttt{working} has to be swapped out to make room for a newly requested one, an eviction policy has to be chosen.
-While we opted for the simplicity of FIFO, any other replacement policy, such as Least Recently Used (LRU), could work in its place.
+While we opted for the simplicity of FIFO, any other replacement policy, such as Least Recently Used, could work in its place.

### 2018-08-07 paper/asplos19/hwswsetup.tex
Methodology revised revised

@@
 		\hline
-		App.&tasks&SLOC&Description\\
+		App.&Tasks&SLOC&Description\\
 		\hline\hline
-        \textbf{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
+        \emph{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
 		\hline
-        \textbf{bc} &10 &371 & several bitcount algorithms\\
+        \emph{bc} &10 &371 & several bitcount algorithms\\
         % \textbf{bc} &10 &371 & Counts bits in bytes sequences
@@
 		\hline
-                \textbf{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
+        \emph{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
 		\hline
-        \textbf{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
+        \emph{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
 		\hline
-        \textbf{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
+        \emph{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
 		\hline
-		\textbf{sort} &4 &167 & selection sort algorithm\\
+		\emph{sort} &4 &167 & selection sort algorithm\\
 		\hline

### 2018-08-07 paper/asplos19/partialcommit.tex
Task downscale revised
Partial commit revised

@@
 The exponential decrease of the timer value converges rapidly to a usable one.
+
+\subsection{Task Downscaling}
 \label{sec:task_downsizing}
-\subsection{Task Downscaling}
+
 \sys uses {\em task downscaling} to make progress through a task that is too large to complete using the energy that the device can buffer.
 Task downscaling executes part of the long task and interrupts its execution with a {\em partial commit}, after which the long task continues.
+This timer-based solution is similar to the one proposed in~\cite{ratchet}.
 If power fails, the partial commit preserves the progress through the task, and execution resumes from the point of the partial commit, rather than from the start of the task.
@@
 The key design issue for task downscaling is deciding when to partially commit.
-\sys only applies task downscaling to a task that is likely to be non-terminating.
 A task is likely to be non-terminating if it fails to run to completion twice consecutively.
 The second incomplete run executes after a power failure, when the device will have fully recharged its energy buffer.
-If a task cannot complete when executing with a full energy buffer, the task is marked as non-terminating.
+If a task cannot complete when executing with a full energy buffer, \sys marks the task as non-terminating.
+
+%timer, similarly used in~\cite{ratchet}.

### 2018-08-07 paper/asplos19/paging.tex
Paging revised

@@
 	\includegraphics[width=\columnwidth]{figures/coalesced-war.pdf}
-	\caption{A WAR dependency (at variable $x$) induced by task coalescing at runtime. Such dynamic dependencies cannot be resolved at compile time and they break memory consistency.
-%		induction of WAR dependency by coalescing: without extra protection, which static runtime cannot reason about, task coalescing induces WAR dependency (at variable $x$) even though individual tasks are WAR protected.
-}
+	\caption{Task coalescing-induced WAR dependency (in the case of this example, at variable $x$). Such dynamic dependencies cannot be resolved at compile time and they break memory consistency.}
 	\label{figure:coalescedWar}
@@
 
-Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when task are coalesced (see Fig.~\ref{figure:coalescedWar}).
-Therefore, \sys introduces its {\em page privatizer} (PP) which enables efficient task coalescing.
+Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when task are coalesced.
+Not handing WAR dependencies with task coalescing introduces a new problem, which we denote as \emph{task coalescing-induced WAR dependency}.
+This problem is illustrated in Fig.~\ref{figure:coalescedWar}.
+Therefore, \sys introduces its novel {\em Virtual Memory Manager} (WMM) which enables task coalescing guaranteeing data integrity, which is at the same time efficient.
 
-PP overview is given in Fig.~\ref{figure:mem-man}.
-PP abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
+VMM overview is given in Fig.~\ref{figure:mem-man}.
+VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
 \texttt{\underline{private}} holds the consistent version of pages after each commit (and on reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
-The PP prohibits applications from directly accessing these buffers.
+The VMM prohibits applications from directly accessing these buffers.
 Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
 It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
-When a coalesced task completes, the PP ensures that {\em dirty} (i.e., modified) pages are committed to FRAM and visible to subsequent tasks.
+When a coalesced task completes, the VMM ensures that {\em dirty}, i.e., modified, pages are committed to FRAM and visible to subsequent tasks.
 If a power failure interrupts a task, all temporary modifications to pages in \texttt{working} (and \texttt{\underline{shadow}}) occurred after the last commit are discarded in order to keep data consistency.
-On a commit, the PP atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
+On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
 % SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM.
@@
     \includegraphics[width=\columnwidth]{figures/mem-man.pdf}
-    \caption{Interaction between task, page privatizer and memory buffers upon \texttt{RP} and \texttt{WP}, and order of occurrence. Steps 2, 3, 4 and 5 are conditional.}
+    \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP}, and order of occurrence. Steps 2, 3, 4 and 5 are conditional.}
     \label{figure:mem-man}
@@
         \State $i \gets \{ j\ |\ \Call{GetTag}{\texttt{working}[j]} = t \}$ \Comment{Search page}
-		\If { $i = \emptyset$ } \Comment{Var in a resident page?}
+		\If { $i = \emptyset$ } \Comment{Variable in a resident page?}
 		\State	$i \gets \Call{PageFault}{t}$ \Comment{Swap page in}
@@
 
-When accessing a protected variable with \texttt{RP} or \texttt{WP}, the PP first searches the variable's page in the \texttt{working} buffer (Fig.~\ref{figure:mem-man}, Step 1).
+When accessing a protected variable with \texttt{RP} or \texttt{WP}, the memory manager first searches the variable's page in the \texttt{working} buffer (Fig.~\ref{figure:mem-man}, Step 1).
 If the page is not found there, a page fault is incurred and a new page needs to be swapped in.
-If the \texttt{working} buffer is full, a page fault on a memory access requires the PP to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
+If the \texttt{working} buffer is full, a page fault on a memory access requires the VMM to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
 If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes have to be persisted to non-volatile memory (Fig.~\ref{figure:mem-man}, Step 2).
@@
 The \texttt{\underline{committing}} bit indicates that a commit is in progress and is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16).
-The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to commit and the PP clears the counter when commit completes (Line 14).
-The \texttt{\underline{commitIndex}} indexes the next page to be committed (Line 11) and the PP clears the index at the end of the phase (Line 15).
+The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to commit and the VMM clears the counter when commit completes (Line 14).
+The \texttt{\underline{commitIndex}} indexes the next page to be committed (Line 11) and the VMM clears the index at the end of the phase (Line 15).
 
@@
 % \sys's implementation is also efficient because
-the PP does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.
+the VMM does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.
 
@@
 Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14) ensures that all accesses to protected variables in subsequent tasks correctly access their consistent memory locations in \texttt{\underline{private}}.
+This solves coalescing-induced WAR dependency problem (see Fig.~\ref{figure:coalescedWar} again).
 

### 2018-08-07 paper/asplos19/paging.tex
Paging revised

@@
 This problem is illustrated in Fig.~\ref{figure:coalescedWar}.
-Therefore, \sys introduces its novel {\em Virtual Memory Manager} (WMM) which enables task coalescing guaranteeing data integrity, which is at the same time efficient.
+Therefore, \sys implements novel {\em Virtual Memory Manager} (WMM) which enables task coalescing guaranteeing data integrity, which is at the same time efficient.
 

### 2018-08-07 paper/asplos19/intro.tex
minor changes

@@
 This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts?
-\sys addresses this challenge by using a novel approach called \emph{dynamic page privatization}---performing real-time dependency tracking to enable protection on a task transition.
+\sys addresses this challenge by relying on its novel \emph{Virtual Memory Manager}(VMM).
+The VMM performs real-time dependency tracking to enable protection on a task transition.
 Individual variables tracking, however, slows down a system dramatically.
-Therefore, \sys keeps memory consistent through \emph{memory virtualization} optimized for bulk accesses to task-shared data with high locality.
+Therefore, the VMM keeps memory consistent through \emph{privatizing pages} and optimizes optimize bulk shared-data transfer through Direct Memory Access (DMA).
 
@@
 Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
-The page privatization is explained in Section~\ref{sec:memory_virtulaization}.
+The the prower interrupt immune virtual memory manager is explained in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are given in Section~\ref{sec:implementation}.

### 2018-08-07 paper/asplos19/systemdescription.tex
revisions of system description
minor change

@@
 The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
-Facilitating efficient task adaptation requires dynamic memory protection which \sys's \emph{virtual memory manager} handles through a process we call \emph{dynamic page privatization}.
+Facilitating efficient task adaptation requires dynamic memory protection which \sys's \emph{virtual memory manager} handles through \emph{page privatization}.
 
@@
 VMM keeps all page modifications in non-volatile memory on a coalesced task transition.
+VMM also solves a novel data consistency problem we call \emph{coalescing-induced WAR dependency}.
 Section~\ref{sec:memory_virtulaization} outlines VMM.

### 2018-08-07 paper/asplos19/paging.tex
minor changes

@@
 
-Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when task are coalesced.
-Not handing WAR dependencies with task coalescing introduces a new problem, which we denote as \emph{task coalescing-induced WAR dependency}.
+Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when tasks are coalesced.
+Not handling WAR dependencies with task coalescing introduces a new problem, which we denote as \emph{task coalescing-induced WAR dependency}.
 This problem is illustrated in Fig.~\ref{figure:coalescedWar}.
-Therefore, \sys implements novel {\em Virtual Memory Manager} (WMM) which enables task coalescing guaranteeing data integrity, which is at the same time efficient.
+Therefore, \sys implements novel {\em Virtual Memory Manager} (VMM) which enables task coalescing guaranteeing data integrity, which is at the same time efficient.
 
@@
 After address translation (Fig.~\ref{figure:mem-man}, Step 6), a task accesses the protected variable's location in the volatile \texttt{working} buffer (Step 7).
-The PP keeps track of the page tags for the pages currently resident in the \texttt{working} buffer.
+The VMM keeps track of the page tags for the pages currently resident in the \texttt{working} buffer.
 When a task accesses a variable, it compares the variable's page tag to tags of the pages in {\tt working} (Algorithm~\ref{algo:rwar}, Line 2).
@@
 Clearing \texttt{\underline{shadowCount}} as part of the second phase commit (Line 14) ensures that all accesses to protected variables in subsequent tasks correctly access their consistent memory locations in \texttt{\underline{private}}.
-This solves coalescing-induced WAR dependency problem (see Fig.~\ref{figure:coalescedWar} again).
+This solves the coalescing-induced WAR dependency problem (see Fig.~\ref{figure:coalescedWar} again).
 
@@
 For example, in the presence of arbitrary pointer operations, a function call using a function pointer, or an interrupt within a task, the system cannot statically analyze the memory behavior.
-Second, a static approach \textit{cannot handle task coalescing}, because a protected variable's lifetime, i.e., from first use to commit, is unknown at compile time.
+Second, a static approach \textit{cannot handle task coalescing}, because a protected variable's lifetime, i.e., from first use to commit, is unknown at compile-time.
 \sys's dynamic, per-access instrumentation supports arbitrary use of pointers and enables task coalescing.

### 2018-08-07 paper/asplos19/intro.tex
Intro revised

@@
 This leads to the second challenge: how to dynamically detect inter-coalesced-task data dependencies and ensure efficient protection against power interrupts?
-\sys addresses this challenge by relying on its novel \emph{Virtual Memory Manager}(VMM).
+\sys addresses this challenge by relying on its novel \emph{Virtual Memory Manager} (VMM).
 The VMM performs real-time dependency tracking to enable protection on a task transition.
@@
 Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism.
-The the prower interrupt immune virtual memory manager is explained in Section~\ref{sec:memory_virtulaization}.
+The power interrupt immune virtual memory manager is explained in Section~\ref{sec:memory_virtulaization}.
 Implementation details of \sys are given in Section~\ref{sec:implementation}.

### 2018-08-07 paper/asplos19/abstract.tex
Abstract revised

@@
 \sys's specialized memory virtualization mechanism ensures that power failures do not leave the program state in non-volatile memory inconsistent.
-Our evaluation on a real energy-harvesting platform not only shows that \sys reduces run time by up to 54\% as compared to a state-of-the-art system, but also it is able to \emph{progress} where static systems fail.
+Our evaluation on a real energy-harvesting platform not only shows that \sys reduces run time by up to 54\% as compared to a state-of-the-art system, but also it is able to progress where static systems fail.

### 2018-08-07 paper/asplos19/background.tex
Background revised

@@
 
-A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists.
-Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
-Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes requiring hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}, Table~\ref{table:chechpoint_comparison} summarizes the data that gets copied to and from the non-volatile memory in service of memory consistency and progress preservation.
-Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}.
-Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory.
-Fig.~\ref{fig:war} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
-The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation.
-If a power failure occurs right before updating the non-volatile index \texttt{i} (\ref{fig:war-execution}, Line 6), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (\ref{fig:war-execution}, Lines 5 and 8).
-
 \begin{figure}
@@
 	\end{subfigure}
-    \caption{WAR dependency example. \texttt{NV} marks non-volatile variables, \texttt{checkpoint()} is a checkpoint of volatile state.}
-    \label{fig:war}
+	\caption{WAR dependency example. \texttt{NV} marks non-volatile variables, \texttt{checkpoint()} is a checkpoint of volatile state.	\label{fig:w-a-r}}
 \end{figure}
+
+A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists.
+Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
+Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes requiring hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}, Table~\ref{table:chechpoint_comparison} summarizes the data that gets copied to and from the non-volatile memory in service of memory consistency and progress preservation.
+Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}.
+Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory.
+Fig.~\ref{fig:w-a-r} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
+The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation.
+If a power failure occurs right before updating the non-volatile index \texttt{i} (Fig.~\ref{fig:war-execution}, Line 6), \texttt{sum} gets erroneously incremented twice consecutively by the same array element (Fig.~\ref{fig:war-execution}, Lines 5 and 8).
 

### 2018-08-07 paper/asplos19/paging.tex
Paging revised

@@
 	\includegraphics[width=\columnwidth]{figures/coalesced-war.pdf}
-	\caption{Task coalescing-induced WAR dependency (in the case of this example, at variable $x$). Such dynamic dependencies cannot be resolved at compile time and they break memory consistency.}
+	\caption{Task coalescing-induced WAR dependency (in the case of this example, at variable $x$). Such dynamic dependencies cannot be resolved at compile-time and they break memory consistency.}
 	\label{figure:coalescedWar}
@@
 Resolving WAR (Write-After-Read) dependencies by considering static task boundaries is not sufficient to keep non-volatile memory consistent when tasks are coalesced.
-Not handling WAR dependencies with task coalescing introduces a new problem, which we denote as \emph{task coalescing-induced WAR dependency}.
+Not handling WAR dependencies on a coalesced task scope introduces a new problem, which we denote as \emph{task coalescing-induced WAR dependency}.
 This problem is illustrated in Fig.~\ref{figure:coalescedWar}.

### 2018-08-07 paper/asplos19/systemdescription.tex
System description revised

@@
 \paragraph{Virtual Memory Manager.}
-\sys can coalesce tasks thanks to its virtual memory manager (VMM).
-VMM privatizes memory pages demanded by a coalesced task and limits application accesses only to the fast volatile memory.
-VMM keeps all page modifications in non-volatile memory on a coalesced task transition.
-VMM also solves a novel data consistency problem we call \emph{coalescing-induced WAR dependency}.
+\sys virtual memory manager (VMM) is the key enabler to ensure memory consistency while coalescing tasks.
+VMM allows applications to interface with only fast volatile memory pages and privatizes the modified pages demanded by a coalesced task in order to solve a novel data consistency problem; namely \emph{task coalescing-induced WAR dependency}.
+VMM achieves page privatization by keeping all page modifications in non-volatile memory on a coalesced task transition.
 Section~\ref{sec:memory_virtulaization} outlines VMM.

### 2018-08-07 paper/asplos19/systemdescription.tex
System description revised

@@
 \sys virtual memory manager (VMM) is the key enabler to ensure memory consistency while coalescing tasks.
-VMM allows applications to interface with only fast volatile memory pages and privatizes the modified pages demanded by a coalesced task in order to solve a novel data consistency problem; namely \emph{task coalescing-induced WAR dependency}.
+VMM allows applications to interface with only fast volatile memory pages and privatizes the pages demanded by a coalesced task in order to solve a novel data consistency problem; namely \emph{task coalescing-induced WAR dependency}.
 VMM achieves page privatization by keeping all page modifications in non-volatile memory on a coalesced task transition.

### 2018-12-27 paper/TOSN/background.tex
TOSN coala version

--- 2018-08-07 paper/asplos19/background.tex
@@
 
-\begin{table}
-    \centering
-    \footnotesize
-    \begin{tabular}{|c|c|}
+% Table
+\begin{table}%
+\caption{Simulation Configuration}
+\label{tab:one}
+\begin{minipage}{\columnwidth}
+\begin{center}
+\begin{tabular}{ll}
+  \toprule
+        Model & Data Copied to/from NVRAM \\
         \hline
-        Model & Data Copied to/from NVRAM \\
-        \hline\hline
         Mementos~\cite{mementos}    & Registers + Stack     \\
@@
         Region Formation~\cite{baghsorkhi_cgo_2018} & Registers + Updated variables in task \\
-        \hline
+  \bottomrule
 \end{tabular}
-    \caption{Non-volatile memory access for data consistency; {PC}: program counter, {WAR variables}: variables involved in Write-after-Read (WAR) dependencies, {NV}: non-volatile.}
+\end{center}
+\end{minipage}
   \label{table:chechpoint_comparison}
-\end{table}
+\end{table}%
+
+%\begin{table}
+%    \centering
+%    \begin{tabular}{|c|c|}
+%        \hline
+%        Model & Data Copied to/from NVRAM \\
+%        \hline
+%        Mementos~\cite{mementos}    & Registers + Stack     \\
+%        DINO~\cite{dino}    & Registers + Stack + WAR NV variables \\%used in task\\
+%        Chain~\cite{chain}  & PC + NV variables used in task\\
+%        Alpaca~\cite{alpaca}    & PC + WAR NV variables used in task\\
+%        Ratchet~\cite{ratchet}, Clank~\cite{hicks_isca_2017} & Registers (requires NV main memory) \\
+%        Region Formation~\cite{baghsorkhi_cgo_2018} & Registers + Updated variables in task \\
+%        \hline
+%    \end{tabular}
+%    \caption{Non-volatile memory access for data consistency; {PC}: program counter, {WAR variables}: variables involved in Write-after-Read (WAR) dependencies, {NV}: non-volatile.}
+%    \label{table:chechpoint_comparison}
+%\end{table}
 
@@
 \begin{figure}
-	\begin{subfigure}{0.49\columnwidth}
+	\begin{subfigure}{0.4\columnwidth}
+		\centering
 		\includegraphics[width=\columnwidth]{figures/war-example.pdf}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.49\columnwidth}
+	\begin{subfigure}{0.4\columnwidth}
+		\centering
 		\includegraphics[width=\columnwidth]{figures/war-execution.pdf}

### 2018-12-27 paper/TOSN/compiler.tex
TOSN coala version

--- 2017-11-15 paper/compiler.tex
@@
 
-After completing program cost analysis, \sys inserts task boundaries into the
-CFG in two steps. The first step is to insert task boundaries on the back-edges
-of all loops that have an energy cost greater than the device's energy buffer
-capacity, producing an initial boundary placement. The second step is to insert
-task boundaries at additional points in the code, sub-dividing tasks that have
-an estimated cost that exceeds the energy buffer's capacity.
+After completing program cost analysis, \sys inserts task boundaries into the CFG in two steps.
+The first step is to insert task boundaries on the back-edges of all loops that have an energy cost greater than the device's energy buffer capacity, producing an initial boundary placement.
+The second step is to insert task boundaries at additional points in the code, sub-dividing tasks that have an estimated cost that exceeds the energy buffer's capacity.
 
-To sub-divide these tasks, the compiler prototype does depth-first traversals
-of all paths in the CFG, ignoring back-edges. Traversals start from the
-beginning of the main function, or from an existing task boundary. The
-prototype accumulates the cost of basic blocks during the traversal until
-reaching another task boundary or an exit node in the CFG. At the end of a
-traversal, if the accumulated cost exceeds to estimated device energy capacity the
-compiler prototype must split the traversed path with a task boundary. Each
-edge on the path is a candidate location for a boundary.
+To sub-divide these tasks, the compiler prototype does depth-first traversals of all paths in the CFG, ignoring back-edges.
+Traversals start from the beginning of the main function, or from an existing task boundary.
+Our prototype accumulates the cost of basic blocks during the traversal until reaching another task boundary or an exit node in the CFG.
+At the end of a traversal, if the accumulated cost exceeds to estimated device energy capacity the compiler prototype must split the traversed path with a task boundary.
+Each edge on the path is a candidate location for a boundary.
 
-\sys uses a heuristic to compute a score for each candidate and places a
-boundary at the candidate with the highest score. The heuristic tries to avoid
-placing boundaries on loops, tries to minimize the number of live variables 
-protected across the new boundary, and tries to maximize the number of operations
-along the path. A boundary candidate's score is computed as
+\textbf{Boundary Placement Heuristic.} \sys uses a heuristic to compute a score for each candidate and places a boundary at the candidate with the highest score.
+The heuristic tries to avoid placing boundaries on loops, tries to minimize the number of live variables protected across the new boundary, and tries to maximize the number of operations along the path.
+A boundary candidate's score is computed as
 %

### 2018-12-27 paper/TOSN/implementation.tex
TOSN coala version

--- 2018-08-07 paper/asplos19/implementation.tex
@@
 %
-\begin{table}
-\footnotesize
-\centering
-\begin{tabular}{| r | p{0.7\columnwidth} |}
+\begin{table}%
+\caption{Simulation Configuration}
+\label{tab:one}
+\begin{minipage}{\columnwidth}
+\begin{center}
+\begin{tabular}{ll}
+  \toprule
+	{Method} & {Arguments} \\
 	\hline
-	{Method} & {Arguments} \\
-	\hline\hline
 	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
-	\hline
 	\texttt{RUN}() & --- \\
-	\hline
 	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
-	\hline
 	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
-	\hline
 	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
-	\hline
 	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
-	\hline	
 	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
-	\hline
 	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
-	\hline
 	\texttt{DISABLE\_PC}() & --- \\
-	\hline
 	\texttt{ENABLE\_PC}() & --- \\
-	\hline
+  \bottomrule
 \end{tabular}
@@
 \label{table:coala_api}
-\end{table}
+\end{center}
+\end{minipage}
+\end{table}%
+%%
+%\begin{table}
+%\footnotesize
+%\centering
+%\begin{tabular}{| r | p{0.7\columnwidth} |}
+%	\hline
+%	{Method} & {Arguments} \\
+%	\hline\hline
+%	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
+%	\hline
+%	\texttt{RUN}() & --- \\
+%	\hline
+%	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
+%	\hline
+%	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
+%	\hline
+%	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
+%	\hline
+%	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
+%	\hline	
+%	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
+%	\hline
+%	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
+%	\hline
+%	\texttt{DISABLE\_PC}() & --- \\
+%	\hline
+%	\texttt{ENABLE\_PC}() & --- \\
+%	\hline
+%\end{tabular}
+%\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
+%\label{table:coala_api}
+%\end{table}
 
-\paragraph{New Tasks.} The \texttt{TASK} annotation on a function
+\noindent\textit{New Tasks.} The \texttt{TASK} annotation on a function
 declaration statically allocates a non-volatile constant variable holding a
@@
 
-\paragraph{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed at the 
+\noindent\textit{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed at the 
 end of the current one task and it can be invoked along
@@
 
-\paragraph{Protected Variables.} The \texttt{PV} annotation on a
+\noindent\textit{Protected Variables.} The \texttt{PV} annotation on a
 variable statically allocates a protected non-volatile variable. The
@@
 
-\paragraph{Initialization.} The behavior of the API method
+\noindent\textit{Initialization.} The behavior of the API method
 \texttt{INIT} is very similar to \texttt{NEXT\_TASK}, with the
@@
 
-\paragraph{Execution.} The programmer passes control to \sys's task scheduler 
+\noindent\textit{Execution.} The programmer passes control to \sys's task scheduler 
 by calling \texttt{RUN} after device initialization.
-%
+
 First the scheduler updates the coalescing target using the coalescing
@@
 %
-Furthermore, the PP moves pages of data efficiently using hardware accelerated DMA support.
+Furthermore, the VMM moves pages of data efficiently using hardware-accelerated DMA support.
 

### 2018-12-27 paper/TOSN/relatedwork.tex
TOSN coala version

--- 2018-08-06 paper/asplos19/relatedwork.tex
@@
-\textbf{Intermittently-powered Devices.} There is a large body of research on energy-harvesting and ambient-powered embedded devices summarized by~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
+\paragraph{Intermittently-powered Devices.}
+There is a large body of research on energy-harvesting and ambient-powered embedded devices summarized by~\cite{prasad_comst_2014,sample_procieee_2013,huang:commag:2015,visser_procieee_2013,kamalinejad_commag_2015,ku_cst_2016}.
 For instance, a new wave of embedded systems powered by radio waves is emerging~\cite{patel_pervasive_2017,rf_powered_computing_gollakota_2014,wisp5,moo,zhao_rfid_2015,holleman_biocas_2008,thomas_jbcs_2012,naderiparizi_rfid_2015,rodriguez_tbcs_2015,liu_sigcomm_2013,kicksat,nadeau_naturebio_2017}.
@@
 
-\textbf{Checkpointing-based Systems.} Early work on energy-harvesting runtimes, such as Dewdrop~\cite{dewdrop}, assumed simple computations that complete on a predictable burst of energy.
+\paragraph{Checkpointing-based Systems.}
+Early work on energy-harvesting runtimes, such as Dewdrop~\cite{dewdrop}, assumed simple computations that complete on a predictable burst of energy.
 %
@@
 Ratchet~\cite{ratchet} ensured consistency by placing a checkpoint at the beginning of each idempotent region in the code.
-Ratchet has a similar technique to task splitting with the core difference: the checkpoint is of a fixed size, while \sys privatize variant number of memory pages.
+Ratchet has a similar technique to task splitting with a core difference: the checkpoint is of a fixed size, while \sys privatizes a varying number of memory pages.
 %
@@
 %
-Clank~\cite{hicks_isca_2017} ensured consistency with custom hardware that dynamically tracks read-after-write dependencies in memory accesses and checkpoints on demand.
-The consistency problem was also approached with a combination of undo and redo logging in software~\cite{baghsorkhi_cgo_2018}.
+Clank~\cite{hicks_isca_2017} ensured consistency with custom hardware that dynamically tracks WAR dependencies in memory accesses and checkpoints on demand.
+The consistency problem was also approached with a combination of undo- and redo-logging in software~\cite{baghsorkhi_cgo_2018}.
 Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all volatile state immediately before power failure and halting the execution.
@@
 
-In all of the above systems, with the exception of DINO~\cite{dino}, checkpoints are dynamic, i.e.
-the programmer does not have explicit control over the point at which the code may be resumed after power failure.
+In all of the above systems, with the exception of DINO~\cite{dino}, checkpoints are dynamic, i.e., the programmer does not have explicit control over the point at which the code may be resumed after power failure.
 Dynamic checkpointing systems make it difficult for the programmer to respect application-level atomicity constraints, such as correlating sensor readings.
@@
 
-\textbf{Task-based Systems.} An alternative to checkpointing are recent systems based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and Mayfly~\cite{hester_sensys_2017}.
+\paragraph{Task-based Systems.}
+Alternatives to checkpointing are recent systems based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and Mayfly~\cite{hester_sensys_2017}.
 Using static tasks, they eliminate the need to checkpoint volatile state.
@@
 
-\textbf{Task Decomposition.} In contrast to \sys's construction of coalesced tasks at runtime, prior work has proposed to optimize task size at compile time.
-CleanCut~\cite{cleancut_2018} program analysis statically estimated energy consumption and split the program into tasks until all tasks consumed less energy than the device can store.
-An alternative program analysis~\cite{baghsorkhi_cgo_2018} generated different versions of a program with different task sizes and empirically selected the best among them~\cite{baghsorkhi_cgo_2018}.
-HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program analysis to place a minimal number of conditional checkpoints that test the energy level at runtime before copying state, like Mementos~\cite{mementos}.
-Unlike \sys, such compiler-based approaches, face the challenge of statically predicting energy consumption of arbitrary input-dependent code with peripheral access, which is an problem without a general solution.
+\paragraph{Task Decomposition.}
+In contrast to \sys's construction of coalesced tasks at run-time, prior work has proposed to optimize task size at compile-time.
+CleanCut~\cite{cleancut_2018} program analysis statically estimates energy consumption and splits the program into tasks until all tasks consume less energy than the device can store.
+An alternative program analysis generates different versions of a program with different task sizes and empirically selects the best among them~\cite{baghsorkhi_cgo_2018}.
+HarvOS~\cite{mottola2017harvos} takes a hybrid approach that uses a program analysis to place a minimal number of conditional checkpoints that test the energy level at run-time before copying state, like Mementos~\cite{mementos}.
+Unlike \sys, such compiler-based approaches face the challenge of statically predicting energy consumption of arbitrary input-dependent code with peripheral access, which is a problem without a general solution.
 Furthermore, a static decomposition approach prevents portability across devices with different storage capacitors.
-In contrast, \sys avoids forcing any assumptions at compile time and adapts to energy storage capacity and incoming energy conditions at runtime.
+In contrast, \sys avoids forcing any assumptions at compile time and adapts to energy storage capacity and incoming energy conditions at run-time.
 
-\textbf{Memory Virtualization.} Prior work on embedded systems has studied a variety of memory virtualization strategies relating to \sys.
+\paragraph{Memory Virtualization.}
+Prior work on embedded systems has studied a variety of memory virtualization strategies relating to \sys.
 TinyOS~\cite{levis2005tinyos} and nesC~\cite{nesc} support dynamic memory management.
@@
 SOS~\cite{sos}, Contiki~\cite{contiki}, and t-kernel~\cite{tkernel} also developed memory management abstractions that virtualize memory size and provide safe and indirect access.
-Mat\'e~\cite{mate} developed full virtual machines support for sensor nodes, virtualizing not just memory resources, but other states and peripherals.
+Mat\'e~\cite{mate} developed full virtual machine support for sensor nodes, virtualizing not just memory resources, but other states and peripherals.
 The goal of \sys is to provide consistent, intermittent execution, leveraging the benefits of efficient bulk copying.
-In contrast, prior efforts focused more on programmability and runtime reliability properties provided by virtual memory.
+In contrast, prior efforts focused more on programmability and run-time reliability properties provided by virtual memory.
 
@@
 Their similarity with \sys is in managing state to ensure consistency at the granularity of pages to amortize checking and tracking costs.
-\sys's paging implementation, which keeps a shadow page for each page to use during commit, is similar to the shadow paging scheme use for transactional commit~\cite{pagebasedtm}.
+\sys's paging implementation, which keeps a shadow page for each page to use during commit, is similar to the shadow paging scheme used for transactional commit~\cite{pagebasedtm}.

### 2019-06-14 paper/asplos19/rebuttal.txt
working on the rebuttal

ASPLOS '19 Paper #336 Reviews and Comments
===========================================================================
Paper #336 Dynamic Task-Based Intermittent Execution for Energy-Harvesting
Devices

Review #336A
===========================================================================

Overall merit
-------------
3. Weak accept

Reviewer expertise
------------------
3. Knowledgeable

Reviewer confidence
-------------------
3. High

Paper summary
-------------
This paper proposes Coala, a task-based runtime system for intermittent computation on energy-harvesting systems.  To reduce checkpointing overhead and ensure forward progress, they propose mechanisms to do task coalescing and task splitting.  They propose a virtual memory manager to ensure consistency of updates to non-volatile memory.

Strengths
---------
+ Shows runtime improvements and in particular overcame limitation of recent prior art that some applications might be non-terminating due to tasks that exceed the available energy budet.  
+ Ran on real hardware so believable results. 
+ Good fit for ASPLOS

Weaknesses
----------
- Reading through the paper, I had some concerns about introducing the overhead of virtual memory management to solve this problem.  Those overhead do not seem to be a problem for 4 out of the 6 benchmarks (Figure 10).  It would have been nice to have more benchmarks to increase some confidence in the results.

Comments for author
-------------------
Programmer effort in the implementation is unclear.  The programmer has to specifically mark protected variables and use some other APIs.  Its unclear how difficult this might be or how error prone it might be.  Can this effort be automated by the compiler?

The weights used for WEG is an aspect of the paper that could have been explored in greater detail.  In particular, it is noted that in some cases WEG is counter-productive.  Is this just due to poor profiling to obtain weights? Would different weight values have been more productive?

In many cases the performance of EG and WEG is almost identical but this isn't really explored or discussed in the results section.

What version of Coala does Figure 10 use? Is it EG or WEG?

Minor: Section 7.1 mentions capturing a diverse power trace.  This was unclear to me.  From the results, there seem to be 3 different power setups (based on distance from RF source) is this what is meant by diverse?

References are parenthetical.  For example, it is incorrect to say "in [1]."

I don't think including the energy oblivious technique is worthwhile -- its only briefly evaluated because the other proposed techniques are better.  Its flaws are fairly obvious and so its not clear it adds anything to the paper.  Not a significant weakness but if you are looking to make space for other material, I think the paper stands fine without any mention of this approach.

Questions for authors’ response
---------------------------------
Can you comment on the programmer effort required to use Coala?

### Response@ReviewerA
1)** Programmer effort** Coala requires, in particular, less intellectual involvement as compared to the prior work. Therefore, its compiler would not require anything new in comparison with existing compilers, such as Alpaca's compiler, and hence we did not introduce yet another front-end compiler.

2)** Number of benchmarks**
We agree with the reviewer that the more benchmarks the better. However, the number of benchmarks that are introduced in this work is not less than what the typical number of benchmarks in related works, for example, Alpaca has 6, Chain 4, and DINO 3 benchmarks.

3)** EG vs WEG**
Coala is about enabling software to understand its energy, charging and discharging, conditions. Task's weight enables Coala's runtime to better reason about the energy cost of a particular task. For example, firing a sensor may require a single line of code, yet from the energy perspective, this task equals 10 other tasks. We highlighted the benefit of using a task's weight in the _fft_ app and show that if tasks are uniformly distributed from the energy perspective then checking task weight will only introduce overhead, and that is why it is an optional strategy.

Review #336B
===========================================================================

Overall merit
-------------
2. Weak reject

Reviewer expertise
------------------
2. Some familiarity

Reviewer confidence
-------------------
1. Low

Paper summary
-------------
The paper presents Coala, which is a programming system for intermittently powered computing systems.  Coala is based on a task model (like other approaches) but the key difference is that Coala supports coalescing and splitting user-defined tasks.  Coalescing allows tasks to reduce overhead of commiting non-volatile state to volatile memory.  Splitting allows large tasks to make forward progress.

Strengths
---------
I like the idea of automating task management (through splitting and coalescing).  The paper does a good job of making the case that the proposed dynamic approach is better than a static approach.

Weaknesses
----------
The Cambrian explosion of work in programming intermittent computing systems makes it difficult for those of us who don't work in the area to track when a paper is proposing a major breakthrough versus an incremental improvement.  This paper's presentation left me unsure of where it stands on this continuum.

Comments for author
-------------------
The paper does a good job of describing problems with static tasking systems and giving readers the intuition for why a dynamic tasking system would be beneficial.

My score is not higher because after reading the paper I am not confident about where the fundamental contribution of this approach falls in relation to the many other recent works on intermittent computing systems.  To be clear, there is a large list of references and the related work section provides some high-level guidance along these lines, but I found myself wanting much more detail from the presentation.  To try to make this constructive feedback, I will list some of the questions I had and how the presentation might be altered to help me understand these issues better.

First, the paper proposes a user-specified task model that the Coala system will then automatically coalesce or split.  This is a very intriguing idea.  Left unaddressed (to my satisfaction anyway) is that this approach seems to trade more programming burden (not just specifying tasks, but also manually labeling protected variables and manually reading and writing to these protected variables.  That seems like a much larger programmer burden compared to a system like Alpaca, but this burden does not appear to really be addressed by the paper.  How many lines of code change when modifying a C program to use Alpaca vs Coala?  Doesn't this approach make bugs more likely (where a programmer incorrectly labeled something)?  While the evaluation shows performance improvements over Alpaca, my intuition is that this comes at a cost of increased programmer burden, which does not appear to be addressed in the paper.

Second, the intro to the paper makes the claim that Coala supports both coalescing and splitting.  I found this intriguing for several reasons.  First is, perhaps a dumb question but I'll ask anyway: why do you need both?  If the system supports task splitting automatically, why shouldn't I just write one task and let the system split it for me automatically?  That seems like the easiest path for me as a programmer, but there must be a reason not to do that.  I could not figure out the reason from the paper (maybe it is obvious to the people who work on this stuff).  Related, what is this paper's contribution to Ratchet ([70])?  Section 4.2 on downscaling was surprisingly short and left me with the impression that it is just the same approach as [70].  The later Related Work section implies that there are fundamental differences, but I am still not clear on what they are.  The most impressive part of the results to me was getting the FFT to complete with Coala when it failed with Alpaca, but that result is based on downscaling the FFT tasks.  It would have been great to have Ratchet as another point of comparison to help illustrate whether that result is really a fundamental contribution of Coala or an incremental improvement over prior work.  Similarly, the related work section describes just-in-time checkpointing and states some high-level issues with those approaches, but there is no direct comparison.  Those JIT approaches seem like a natural comparison as they are also attack the problem of overhead in intermittent computing systems, same as Coala's coalescing.

Related to the evaluation is my final comment. In these results, it was hard for me to evaluate how much the dynamic aspect is buying versus just comparing to a bad implementation of a program in Alpaca.  I understand that programmers might make a bad choice when statically sizing tasks and that Coala could alleviate that problem (at the cost of much more invasive programming).  However, I would like to know if there are reasons to believe that no static task assignment will be reasonable.  That would be a much stronger argument for the proposed approach.

In summary, there are some good things about the paper.  My hesitation is that it is very hard for me to put it into the quickly evolving ecosystem of intermittent programming systems.

Questions for authors’ response
---------------------------------
I have a number of questions above.  Please answer any of those for which you have the space.

### Response@ReviewerB
1)**Programming burden** 
Coala's app size is similar to that of Alpaca with similar annotation requirements. Moreover, all the code transformation can be done by a front-end compiler, please see the response@ReviewerA.1. In addition, prior task-based systems require, due to their static nature, accurate reasoning about energy conditions to offer efficient execution, which is a significant intellectual programming effort. Whereas, Coala enables on-demand software adaptation to offer efficient and safe execution.

2)**Is Coala more prone to Bug?**
Due to Coala's dynamic nature, it is less prone to bug than the static systems, i.e. static systems suffer from large task problem while Coala does not. 
**TODO**, mentioned the benefit of using address level protection mechanism as compared to the naming annotation.

3)**Why splitting  is not default**
Task splitting requires stack protection (saving the stack in the non-volatile memory) while tasks coalescing does not, so it is more efficient. Moreover, stack growth and task size have a positive correlation.

4)**Downscaling and Rachet**
To be answered shortly

5)**Comparing against other systems**
In contrast to the vast majority of the presented systems, Rachet targets a completely different hardware (ARM Cortex). As consequence, porting it to MSP430FRxxxx (ultra-low energy microcontroller) requires a significant effort. Furthermore, Alpaca, the state-of-the-art, has already demonstrated its superiority to other systems like Chain (task-based), DINO (checkpoint), and mementos (checkpoint).

6)**Why not static-task**
Static tasks divisions require intermittent application design and implementation for each hardware configuration; otherwise, they are inefficient. Even worse, a solar-powered intermittent device, for example, indoor has completely different energy harvesting conditions than outdoor, yet the static division can only consider the worst case scenario.

Review #336C
===========================================================================

Overall merit
-------------
3. Weak accept

Reviewer expertise
------------------
2. Some familiarity

Reviewer confidence
-------------------
3. High

Paper summary
-------------
Energy-harvesting devices (battery-free) exhibit intermittent execution based on unreliable energy sources. The execution can stop any time, so regular checkpointing is necessary. Another issue is memory consistency if some writes are executed to non-volatile memory after checkpoint, and the program execution stops before the next checkpoint, the memory state cannot be rolled back leading to inconsistency. Prior works[12,42]  have proposed task-based programming model were program execution is divided into tasks, and all updates of tasks are committed at the end of the task.

The paper proposes a dynamic task management scheme: Coala. The key idea is dynamically coalacse tasks to avoid the overhead of numerous commits. Some simple history based policies have been explored to determine how many tasks can be combined before the commit. Also, to ensure atomic commits at the end of the dynamically coalesced task, a  software-only virtual memory technique is proposed. Write to NV memory locations from a task go through a level of indirection, buffered in a shadow memory region and committed to FRAM in a two-phase commit protocol. A timer based approach is used to break tasks which are non-terminating and keep restarting.

Strengths
---------
Dynamic task coalaescing can reduce checkpointing overheads

Evaluation on WISP intermittently powered platform,  Coala in its entirety is implemented and evaluated.

Weaknesses
----------
Builds upon and improves prior work on task-based intermittent execution

Virtualization causes overheads.  Some workloads show worse performance than static tasks.

Comments for author
-------------------
The paper is well-written and easy to read.

I see that you have started from a task-based execution model, and then made it more adaptive for reducing commit overheads. But I think virtualization is independent of task model and can reduce checkpointing overheads. For instance, the dirty pages can be periodically committed to NV memory based on some heuristics. The period can be thought of as the "checkpointing interval".
Is there a need for task wrapper?  On that note, if the microcontroller already comes with MMU support, then the overheads can be reduced significantly.

Questions for authors’ response
---------------------------------
Can you clarify the need for a two-phase protocol? why is it not sufficient to do just copy from shadow to private?

How are task boundaries exactly determined in prior works?

How difficult is to estimate time before next power-off based on energy buffer?

### Response@ReviewerC
1)**Two phase commit**
A consistent set of data need to be guaranteed under any execution circumstances, and the two-phase protocol offers this guarantee. Moreover, Coala does not copy the modified content of the buffer twice. It, instead, set flags to exchange buffers' cells rules.

2)**Task boundaries**
All prior works require the programmer to reason about the energy availability and set the STATIC task size accordingly. Coala, however, enables the runtime to reason about the energy conditions and adapt the coalesced/split task accordingly.

3)**Next power-off**
This discharging time (on-time) of an intermittent device depends on the incoming power and discharge rate. It is very complicated to determine in general accurately the next power interrupt. Especially, if the device drive sensors and its execution is environment dependent.

Review #336D
===========================================================================

Overall merit
-------------
2. Weak reject

Reviewer expertise
------------------
2. Some familiarity

Reviewer confidence
-------------------
2. Medium

Paper summary
-------------
This paper presents an approach to handle efficient checkpointing in intermittent computing based on task graphs. In existing task based designs each task is atomically executed on an intermittent computing platform. All the non-volatile state is preserved at the end of the task and a new task may start with a clean architecture state and inter-task data dependencies are communicated through  data stored in a byte addressable NVM.  The challenge in this scenario is to determine the size of a task so that the overhead of committing all the inter-task dependence data into NVM can be minimized. Smaller tasks are worse in this regard. But larger tasks may repeatedly fail to progress leading to deadlocks. So the goal of this work is to combine smaller tasks or break up larger tasks while taking into consideration the amount of battery power that is left.

Strengths
---------
The main novelty of the work in my view is that the paper proposes a virtual memory system for NVM that essentially allows task coalescing to occur efficiently. The proposed design essentially remaps the read (or write) NVM address to a location in volatile memory so that all updates within a coalesced set of tasks are directed to the volatile memory, through a paging scheme.

The second novelty, although I consider it to be a smaller contribution, is an approach to determine how to coalesce (or split) tasks based on energy availability using the historical task completion information.

Weaknesses
----------
The weakness of the work is the fact that the the VMM system designed for NVM is a straight forward and simple approach that is limited in novelty. The work applies the notion of shadow and private paging to protect intermediate updates from corrupting NVM. It comes at the cost of keeping two NVM copies always and a third copy in memory (called the working page). Hence, the benefits of byte addressable NVM is mostly lost in the proposed design.

Comments for author
-------------------
I enjoyed your full system implementation details and I think the work has some value. But in the current form I would like to see the following concerns addressed:

(1) Do you have any comments on the critique that the proposed design essentially triples memory usage and thereby reduces the size of the programs that can be run on Coala? Did this memory limitation cause significant issues in running any of your workloads (I did not see any reference to this aspect in the experimental section).

(2) Can you also  please comment on the difficulty of having to re-write each application to use the RD and WR API provided by your system? Is it possible to automate this RD and WR APIs to be handled by the underlying runtime without explicit user annotation? If so what are the pros and cons of such an approach?

(3) The improvements shown over Alpaca (Fig 10) seem to be highly dependent on the type of task creation a programmer may chose. I understand the argument that task size is a challenge but at least for the purposes of some empirical upper bounds I expect that Alpaca will do much better if the task creation is not completely oblivious to the underlying hardware.

Questions for authors’ response
---------------------------------
See my three comments above and please address them in your rebuttal.

### Response@ReviewerD
1) On the volatile memory (RAM) data cannot be saved even with normal devices on a power interrupt. The working buffer of Coala, which is much smaller than the other buffers, can be seen as a cache, and thus Coala requires double the size of an application's data. Coala requires that to reach its maximum performance from the energy perspective. We did not encounter any issue with benchmarks memory requirements. Please have a look at  Response@ReviewerC.
2) Please have a look at Response@ReviewerA.1.
3) We agree with the reviewer and that is what Coala trying to prove. However, Please have a look at Response@ReviewerC.3 to see why static approach cannot be optimal.

Review #336E
===========================================================================

Overall merit
-------------
3. Weak accept

Reviewer expertise
------------------
3. Knowledgeable

Reviewer confidence
-------------------
3. High

Paper summary
-------------
This paper presents an interesting systems support to solve a well-known dilemma in intermittent execution of energy-harvesting systems: the overhead of context savings vs. the cost of aborted execution. Using small tasks incurs high overhead of context savings; using large tasks incurs high cost when a task is aborted (without context saving). The key idea is to have a runtime that decide the size of tasks dynamically based on available energy. The runtime support has two key parts: scheduler and virtual memory manager. The scheduler determines the size of tasks; the virtual memory manager ensures consistency of non-volatile memory.

Strengths
---------
A refreshing approach toward a relatively niche/understudied system.

Weaknesses
----------
Important details of how the system work are missing so it is difficult to gauge how well it may work in real deployment. Specifically how do you manage the WAR consistency of knowledge needed for VMM and scheduler? How do you determine the granularity of tasks?

The system is ideally coupled with programming support, which is largely missing in the paper.

The evaluation setup is contrived.

Comments for author
-------------------
This reviewer found the scheduler part missing a lot of details. For example, your scheduler must have some knowledge about the tasks under question (e.g., an ordered list of tasks to be executed, or more generally a directed graph of tasks). As a task is committed, this knowledge must be updated. Where does it obtain this knowledge? How do you ensure the WAR consistency of this knowledge (i.e., the list or graph)?

The part of the scheduler left much to be desired for the compile time analysis & optimization (Maybe the authors are planning another paper?). For example, the weight for a task must be derived at compile time/static analysis. This paper simply assumes they are given. Neither was the granularity of tasks  discussed, which is likely to be dependent on the system’s configuration and operational environment.

The virtual memory management design is quite interesting, a clever realization of two-phase commitment using the virtual memory indirection. I hope the authors could discuss the intellectual link to two-phase commitment protocols in their camera ready. Like the scheduler, the VMM also needs some knowledge about the memory usage and this knowledge likely needs to be updated when memory is used. How do you ensure the WAR consistency of this knowledge?

The evaluation is conducted in a quite contrived environment, especially how energy is delivered to your system. To ask rhetorically, how could you be sure that your manual choice of task granularity is not simply optimized for the cycle of energy delivery to your system?

Please provide some background for the architectural/systems aspect of your intermittent systems. For example, no multitasking? no out of order execution? what is the interrupt model? Would a task be interrupted? If so, how do you deal with the nondeterministic behaviors introduced by interrupts?

Questions for authors’ response
---------------------------------
Is there any knowledge for the scheduler or VMM that needs to be updated over the lifetime of an app and saved before power-off? If so, how do you ensure their WAR consistency?

Are you writing another paper about the programming support for dynamic tasks?

### Response@ReviewerE

1)**Data consistency VMM and scheduler** 
When the scheduler _unlocks_ a coalesced task for execution, the VMM _locks_ the private buffer for any write access.
When The VMM _unlocks_ the private buffer to update its data (setting flags instead copying data between the NV-buffers), the scheduler locks task execution. This strategy guarantees the atomicity of data updates and consequently data consistency.

2)**Tasks granularity**
Static task sizes are specified by the programmer (or a compiler). Coalesced task size is determinate on-the-fly by Coala's runtime based on the recent execution history.

3)**Programming support**
See Response@ReviewerA.1

4)**Scheduler**
Coala's scheduler is a supervised, by the tasks, round-robin scheduler. It updates a minimalist amount of non-volatile data to latch to the under-execution task until the task execution is finished despite power interrupts.

5)**Another paper**
We are targeting a real-life application demonstration.

6)**Are the task sizes optimized for Coala** 
The design of our applications is based on the design of the benchmarks that are used Alpaca system (see Coala Table 3 and Alpaca application source code on the owners' repo).

7)**Handling interrupts**
Permanent data modification happens only across coalesced task boundaries and all task-shared data are protected. Therefore,  
Coala, in contrast to the other task-based systems, guarantees data consistency even when ISRs are considered.
However, Coala history tracker does not count the ISR; therefore, small ISRs are required for optimal performance.

### 2019-06-14 paper/asplos19/rebuttal_answers.txt
working on the rebuttal

We thank the reviewers for their time and the raised points and hope our answers are satisfactory.

### Response@ReviewerA

[A1]**Programmer effort** Coala requires less programmer effort as compared to other task-based systems for the following reasons: (i) Chain/Alpaca requires WAR-dependent variables identification, a prone to error process, while Coala does not. (ii) Coala does not require a special technique to handle arrays as Alpaca. (iii) Static systems require pre-deployment accurate reasoning about energy conditions for efficiency while Coala optimizes it execution on-demand. (iv) Coala applications have similar lines of code to Alpaca’s applications and LoCs around ⅓ of Chain’s applications (see Table3 Coala and Table4 Alpaca).

[A2]**Can a compiler automatically annotate Coala’s applications?** Yes. We did not introduce a compiler in the paper because it would require nothing new or novel compared to the introduced compilers by prior work. Therefore, we considered it as a solved problem.

[A3]**EG/WEG**: From the paper (page 9): "In some cases WEG’s task weighting system is counter-productive. This occurs in task decompositions with energy-uniform tasks, where counting tasks disregarding their energy consumption provides an equal amount of information with a smaller effort. In fft, tasks are not uniform, and accounting for their different weights beneficial”. Additionally, task weights can be extremely important---a task that triggers an I/O operation may consume several times more energy than a non-I/O task.

### Response@ReviewerB
[B1]**Where does Coala stands on the continuum of intermittent computing?** Coala is the first intermittent computing system that combines the flexibility of checkpointing with the efficiency of task-based models.

[B2]**Programming burden** See [A1]

[B3]**Why splitting is not default?**
Task splitting requires stack protection while tasks coalescing does not; therefore, it is more efficient.

[B4]**Downscaling and Ratchet**
Ratchet checkpoints the system on each encounter of a WAR-dependency. To reduce the size of the checkpoint, Ratchet operates on non-volatile memory which imposes _at least_ 2x energy consumption as compared to accessing volatile memory. Coala, however, strives to reduce the number of task splits, operates on the energy-efficient volatile memory, and tolerates checkpoint size. This could be made more clear in the paper, to give more focus to Section9.

[B5]**Comparing against other systems**: Alpaca, the state-of-the-art, had already demonstrated its superiority to other systems like Chain (task-based), DINO, and Mementos (checkpoint). Therefore, by comparing to Alpaca we implicitly compared Coala to all aforementioned systems. Ratchet, in constrast to the majority of intermittent systems, considers a completely different microcontroller architecture that makes this comparison out of scope.

[B5]**Why not static-task**: 
Static tasks require _significant_ code changes to port between devices. Additionally, choosing the approperiate static task size skyrokets the programmer effort.

### Response@ReviewerC
[C1]**Why two-phase commit**:
A form of two-phase commit is a must to ensure data and system states consistencies. Coala VMM must _protect_ the individual memory pages before they being accessed and _ensure_ the consistency of the states of the buffers.

[C2]**Why task wrapper?** To avoid the need for stack protection and increase the performance.

[C3]**Task boundaries in prior work**: See [A1 point (iii)]

[C4]**Next power-off**: Given the varying nature of harvested energy and the variety in a load power consumption, especially when I/O operations are considered, it is very challenging to accurately determine the next power failure.

### Response@ReviewerD
[D1] Coala doubles applications NVM requirements to reach its maximum performance and to reduce energy consumption. Additionally, we did not encounter any memory problems.

[D2] See [A1]

[D3] We have used the exact tasks devision the Alpaca’s owner shared online, and we designed Coala’s applications accordingly.

### Response@ReviewerE

[E1]**Scheduler&VMM consistency**:  Coala uses two-phase commit process to ensure the consistency of its states, also see [C1].

[E2]**Coala’s scheduler** is a persistent and supervised, by applications tasks, Round-Robin scheduler. It, despite power failures, remembers the address of the under-execution task until it finishes. Additionally, it prohibits the commit process from being interrupted by task-execution or vice versa. However, if a task is split the entire execution context is saved and retrieved to preserve correct execution.

[E3]**Optimizing task size for Coala**: See [D3]

[E4]**Handling interrupts**: Permanent data modification happens only across coalesced task boundaries and all task-shared data are protected. Therefore, Coala, in contrast to the other task-based systems, guarantees data consistency even when ISRs are considered. However, the history tracker used in EG/WEG does not account for the weight of an ISR, that could fire at any point in time, thus ISRs should be kept small in order for their energy cost to be negligible with respect to one of the tasks they are interrupting.

### 2019-06-14 paper/TOSN/cover-letter/letter.tex
working on the rebuttal

\documentclass[11pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{vmargin}
\usepackage{graphicx}
\usepackage{parskip}
\usepackage{url}
\usepackage{hyperref}

\setmarginsrb{2.25cm}{0.5cm}{2.25cm}{0.5cm}
{12pt}{20pt}{12pt}{36pt}

%----------------------------------------------------
% Sender info
%----------------------------------------------------
\def\Who{Amjad Yousef Majid} 
\def\What{}
\def\Where{Embedded Software Group}
\def\Affiliation{Delft University of Technology}
\def\Address{Mekelweg 4, 2628 CD, Delft, The Netherlands}
\def\Email{E-mail:  a.y.majid@tudelft.nl}
\def\Telephone{Phone: +31 61 694 4225}
\def\Webpage{\url{http://www.st.ewi.tudelft.nl/amjad}}

\def\Sender{

	\begin{flushright}
		\Who \What \\
		\Where \\
		\Affiliation \\
		\Address \\
		\Email \\
		\Telephone \\
		\Webpage
	\end{flushright}	
}

\begin{document}

	\thispagestyle{empty}

	\Sender

	\bigskip

	Dear Editor-in-Chief,

	I am sending to you a manuscript entitled "Coala: Dynamic Task-Based Intermittent Execution for Energy-Harvesting Devices" to be considered for publication in ACM Transactions on Sensor Networks.

	Coala advances intermittent software execution, which drives battery-free tiny sensors, by predicting the amount of ambient energy arrival and adapts the execution accordingly. Battery-free sensors and actuators are expected to play an important role in health-care, building monitoring, and space applications. Our experiments show that Coala optimizes applications execution, running on energy harvesting devices, by up to 54\% as compared to the state-of-the-art intermittent system.

	We believe that this manuscript is appropriate for publication by \emph{ACM Transactions on Sensor Network}, as its scope falls within the scope of the journal. This work has not been published or simultaneously submitted elsewhere. Thank you in advance for considering our submission.

	\vspace{0.3cm}

	Sincerely,

	\Who \\
	Corresponding Author

\end{document}

### 2019-06-22 paper/TOSN/response-letter/responseLetter.tex
minor revision TOSN

\documentclass[10pt]{article}
\usepackage{amsmath,amsfonts,amssymb,graphicx,color,psfrag}
\usepackage[utf8]{inputenc}

\usepackage{vmargin}
\setmarginsrb{2.25cm}{2.25cm}{2.25cm}{2.25cm}
         {12pt}{20pt}{12pt}{36pt}

\usepackage{soul}
\usepackage{hyperref}

\newcommand{\referee}[1]{
	{\item \color{OliveGreen} \emph{{#1}}}
	\label{R\therefereeCounter:\arabic{enumi}}
}

\definecolor{OliveGreen}{rgb}{0.1,0.2,0.5}

\newcommand{\response}[1]{{\color{blue} #1}}

\newenvironment{editor}{%
\textbf{\large Additional changes}
\begin{enumerate}%
\renewcommand{\labelenumi}{\textbf{[E:\,\arabic{enumi}]}} %
}{\pagebreak\end{enumerate}}

\newcounter{refereeCounter}

\newenvironment{responses}{%
\refstepcounter{refereeCounter}%
\textbf{\large Comments by Referee \therefereeCounter}
\begin{enumerate}%
\renewcommand{\labelenumi}{\textbf{[R\therefereeCounter :\,\arabic{enumi}]}} %
}{\end{enumerate}}

\setlength{\parindent}{0pt}

\begin{document}

\pagestyle{myheadings}
\thispagestyle{empty}

\markright{\footnotesize TOSN-2019-0002: {\sl Coala: Dynamic Task-Based Intermittent Execution for Energy-Harvesting Devices}}

\headsep 0.5cm

\bigskip\bigskip

%\noindent{Marwan Krunz, EIC, and Kevin Almeroth, Associate EIC \\
%	IEEE Transactions on Mobile Computing
%}

\bigskip\bigskip

\begin{flushright}
Amjad Yousef Majid \\
Embedded and Networked Systems Group \\ 
Delft University of Technology \\ 
Mekelweg 4\\
2628\,CD Delft, The Netherlands \\
E-mail: a.y.majid@tudelft.nl\\
Phone: +31 61 695 5224\\
\end{flushright}

\vspace*{2cm}

\today
\medskip

\textbf{TOSN-2019-0002:} {\sl Coala: Dynamic Task-Based Intermittent Execution for Energy-Harvesting Devices}

\bigskip

Dear Editor,\\

We would like to thank you for handling our submission and providing valuable review results. 
We are also grateful to the reviewers for their substantial effort spent for evaluating our work. \\

With this letter we address all comments of the reviewers and point to the modified sections of our revised paper. \\

With best regards, \\

in the name of all authors---Amjad Yousef Majid

\pagebreak

\begin{responses}

\referee{In section 2.3, authors mention "these models do not rely on capturing an expensive checkpoint, they are usually faster than the checkpoint-based solution [12, 42]". I don't fully agree with this. It depends on the underlying checkpointing mechanism. If it copies complete stack every time during the checkpoint then yes it's expensive. But checkpointing mechanism like incremental checkpointing [2] (also see the paper by Faycal Ait Aouda), where you only update modified variables, is much better in terms of energy consumption (even better than the task-based solution in my opinion). Also, [12] and [42] are not checkpoint-based solutions.}

\response{As indicated by the reviewer, [12, 42] are not checkpoint systems. They are task-based systems and their results show that task-based systems are faster than checkpoint-based ones. Therefore, we restated their claim and provided the supporting references, which are [12, 42].\\
 Coala addresses deficiencies in the \emph{static} task-based systems and provides a more efficient alternative. Therefore, we compared its performance against the performance of the state-of-the-art task-based runtime.}

\referee{In section 7.1, authors mention "produce a high number of complete runs". Please mention the exact number iterations. Also please mention the capacitor size for the WIPS 5.1.}

\response{The number of complete runs and WISP capacitor size are added to Section 7.1.}

\referee{In figure 10, it is not very clear how authors normalized the values for Coala in case of fft for 30cm and 50cm. Also in the text authors state that "This is marked with $\infty$ signs in Fig. 9", its figure 10. }

\response{We normalized these two special cases on 1 second. This infromation is added to Section 8.2, and the figure reference is corrected. }

\referee{However, I was hoping to see more evaluations to understand the effect of capacitor size. In the introduction, authors state that "Coala is the only system that eliminates restructuring and re-compilation of applications considering device’s energy buffer". However, in the evaluation, I didn't find any graph related to the performance of Coala w.r.t to different capacitor sizes.}

\response{An additional experiment regarding the capacitor size has been conducted and the results are added to Section 8.2.}

\referee{I would have also liked to see a comparison between Coala and Region Formation [2]. Both try to reduce checkpointing overhead.}

\response{As it is indicated in \hyperref[R1:1]{[R1:1]}, Coala addresses deficiencies in the \emph{static} task-based systems and provides a more efficient alternative. Therefore, we compared its performance against the performance of the state-of-the-art task-based runtime.}

\referee{Also, I would suggest incorporating error bars in all the graphs as your "results were averaged across the duration of the experiment"}
\response{We chose to normalize the results to focus on the speedup (or slowing down) and that is why we omitted the error-bars.}

\referee{In the abstract (and also in the introduction), authors claim that "Coala reduces run time by up to 54\%", however, I didn't see this number again anywhere in the evaluation.}
\response{This number is highlighted in Section 8.2.}

\referee{typos}
\response{are corrected}

\end{responses}

\begin{editor}
\referee{}
\response{Dear editor we added a new figure to the paper, Fig. 5. The figure's sole purpose is to aid explaining the concept of \textit{task coalescing}.
We hope that this modification is approved by you.}
\end{editor}

\end{document}

### 2019-06-22 paper/TOSN/coalescing.tex
TOSN coala version
minor revision TOSN

--- 2018-08-07 paper/asplos19/coalescing.tex
@@
 
-% Coalescing has a double benefit.
-First, coalescing reduces the total number of relatively high-latency, high-energy non-volatile memory accesses because tasks primarily access variables in relatively low-latency, low-energy volatile memory.
-Second, coalescing eliminates the unnecessary instructions that implement task commit, which constitute a substantial run-time overhead, as our measurements in Section~\ref{sec:evaluation} show.
-
 \begin{algorithm}[t]
@@
 	\label{algo:genCoalescing}
-	\scriptsize
+	\small
+%	\scriptsize
 	\begin{algorithmic}[1]
@@
 Different coalescing strategies may apply different weights to a task, e.g., $W_\text{i} = 1\ \forall i$, or $W_\text{i} = \alpha E(T_\text{i})$ where $E(T_\text{i})$ is the average execution time of $T_\text{i}$ and $\alpha$ is a constant.
-% For example, a uniform weighting scheme sets the weight of each task to one, making them all effectively equal in cost to the coalescing algorithm.
-A different policy might use a weighting scheme that sets a task's weight equal to its run time relative to the task that executes for the shortest duration, which has unity weight.
 
@@
 
-A very simple {\em energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly.
+{\em Energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly.
 In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits, and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure.
@@
 
+\begin{figure}
+\centering
+\includegraphics[width=.8\columnwidth]{figures/dif_cap}
+    \caption{\add{Anatomy of \sys coalescing---A snapshot of measured data showing how \sys coalesces static tasks.
+\sys takes advantage of the guarantee that an energy buffer offers after a reboot: code execution resumes after energy accumulation reaches the operational threshold (a maximum threshold).
+Thus, \sys starts with a big coalesced task and shrinks it gradually as the risk of a power failure increases.}}
+    \label{fig:coal_anatomy}
+\end{figure}
+
 The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions.
@@
 In fact, at each reboot, EG will {\em conservatively} try to coalesce only a fraction ($\rho$) of tasks that had successfully completed during the last power cycle.
-% In fact EG can reduce its target to one from any coalesced target in at most two successive power failures.
-For instance, if the execution is interrupted after finishing a single task the history will be one ($H = 1$) and the target will be set to one also.
-While a history of $H$ suggests that there is sufficient energy to coalesce and successfully commit $H$ tasks (i.e., to set $C = H$).
-EG conservatively sets $C$ equal to half of $H$ moderating the risk of a very long coalesced task.
-We experimented with a range of fraction values, settling on one half as the default for EG, because one half had the highest performance.
 Fig.~\ref{fig:hg-coal} illustrates a snapshot of the operation of EG, with $\rho = \gamma = 0.5$.
@@
 %
-% The effectiveness of the WEG hinges on correctly identifying the weight of each task, which WEG assumes is {\em statically} available.
-Profiling the time and energy cost of tasks in a program is a difficult, orthogonal problem~\cite{cleancut_2018,baghsorkhi_cgo_2018}.
-WEG could use the result of an arbitrarily sophisticated profiling procedure.
-To produce a concrete result in this paper, we give WEG access to a simple profile of task run time collected offline using a single fixed input.
-WEG stores the profile in a lookup table that maps a task's identifier to its weight, making the information available to \sys's scheduler at run time.
+
+\add{
+Fig.~\ref{fig:coal_anatomy} shows \sys in action.
+It shows how \sys coalesces aggressively when the execution is just re-started, taking advantage of the full energy buffer, and how it reduces the coalesced task gradually to minimize the risk of significant progress loss when the device fails.
+The average measured static task size is $\approx$\,0.28 $ms$ while the commit time (time needed to save the data into the non-volatile memory) averages at  $\approx$\,0.64 $ms$.
+This highlights the importance of coalescing.
+}

### 2019-06-22 paper/TOSN/evaluation.tex
TOSN coala version
minor revision TOSN

--- 2018-08-07 paper/asplos19/evaluation.tex
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/coalEfficiency}
+    \includegraphics[width=.8\columnwidth]{figures/coalEfficiency}
     \caption{Breakdown of overhead for the proposed coalescing strategies:
@@
     \begin{subfigure}{\columnwidth}
-        \includegraphics[width=\columnwidth]{figures/overallOverhead.pdf}
+			\centering
+        \includegraphics[width=.8\columnwidth]{figures/overallOverhead.pdf}
         \caption{Kernel overhead breakdown}
@@
     \begin{subfigure}{\columnwidth}
-        \includegraphics[width=\columnwidth]{figures/memAccess.pdf}
+			\centering
+        \includegraphics[width=.8\columnwidth]{figures/memAccess}
         \caption{Protected memory accesses breakdown}
@@
 \paragraph{Speedup with Coalescing.}
+%
+\begin{figure}
+    \includegraphics[width=.8\columnwidth]{figures/coalStrategies}%
+    \caption{\sys's coalescing performance. Application execution time
+with coalescing (EG and WEG) normalized to the execution time
+without coalescing (NC).}
+\label{fig:coalescing}
+\end{figure}
+
+\begin{table}[t]%
+
+\caption{\add{Comparison of \sys performance on two different capacitor sizes. We see that \sys optimizes its coalesced task size based on the energy buffer size.
+\textbf{coalesced task} refers to the first coalesced task after a reboot, \textbf{exp. time} shows the experiment duration, and the \textbf{runs} column lists 
+the number of complete runs of the \textit{sort} application during the experiments.}
+}
+\label{tab:dif_cap}
+\add{
+\begin{minipage}{\columnwidth}
+\begin{center}
+\begin{tabular}{llllll}
+  \toprule
+    \textbf{cap. size} ($\mu\text{f}$)& \textbf{exp. time} ($s$)& \textbf{duty cycle} & \textbf{coalesced task} ($ms$)   & \textbf{runs} & \textbf{app run-time} ($ms$) \\
+        \hline
+    47                 &1205             & 12.93\%    & 33                        & 85       & 183      \\
+    470                &1205             & 12.79\%    & 88                        & 87       & 177 \\
+  \bottomrule
+\end{tabular}
+\end{center}
+\end{minipage}  
+}
+\end{table}%
 
@@
 %
+
 \begin{figure}
-    \includegraphics[width=\columnwidth]{figures/coalStrategies}%
-    \caption{\sys's coalescing performance. Application execution time
-with coalescing (EG and WEG) normalized to the execution time
-without coalescing (NC).}
-\label{fig:coalescing}
+    \includegraphics[width=.8\columnwidth]{figures/coala_alpaca_gcc}
+    \caption{\sys's execution time normalized to Alpaca's,
+    for three distances from the energy source (15, 30 and 50\,cm). }
+    \label{fig:runtime}
 \end{figure}
 
-\begin{figure}
-    \includegraphics[width=\columnwidth]{figures/coala_alpaca_gcc}
-    \caption{\sys's execution time normalized to Alpaca's,
-    for three distances from the energy source (15, 30 and 50\,cm).}
-    \label{fig:runtime}
-\end{figure}
+\paragraph{\add{Different Capacitor Sizes}}
+\add{
+Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy.
+We see that \sys scales up its coalesced task size with a bigger energy buffer and vise versa. This allows it to reduce the time-to-completion of the applications. For example, 
+the \textit{sort} run-time is reduced from 183\,$ms$ to 177\,$ms$ when the capacitor is changed from 47\,$\mu$f to 470\,$\mu$f.
+It should be emphasized that a device with a bigger energy buffer suffers less from power failures (but requires longer charging time).
+However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.
+}
 
@@
 Fig.~\ref{fig:runtime} shows the average execution time of each
-application for \sys and Alpaca, normalized, where possible, to the latter.
+application for \sys and Alpaca, \add{normalized to the latter or, when not possible, to one second}.
 \sys provides a performance benefit compared to Alpaca for most
-applications. The speedup is greatest for applications with repeated WAR
+applications. \add{For example, it is 54\% faster than Alpaca when executing the \textit{sort} application. In general,}  the speedup is greatest for applications with repeated WAR
 dependencies throughout their code, particularly involving arrays
@@
 the static tasks.}.
-This is marked with $\infty$ signs in Fig.~\ref{fig:coalescing}.
-\sys, however, manged to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.
+This is marked with $\infty$ signs in Fig.~\ref{fig:runtime}.
+\sys, however, managed to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.
 
@@
 \begin{figure}
+	\centering
     \begin{subfigure}{\columnwidth}
-        \includegraphics[width=\columnwidth]{figures/page_exec-time.pdf}
+			\centering
+        \includegraphics[width=.8\columnwidth]{figures/page_exec-time.pdf}
         \caption{Execution time normalized to lowest per-application}
@@
     \begin{subfigure}{\columnwidth}
-        \includegraphics[width=\columnwidth]{figures/pagePulls.pdf}
+			\centering
+        \includegraphics[width=.8\columnwidth]{figures/pagePulls.pdf}
         \caption{Number of page faults per application run}

### 2019-06-22 paper/TOSN/hwswsetup.tex
TOSN coala version
minor revision TOSN

--- 2018-08-07 paper/asplos19/hwswsetup.tex
@@
 \label{sec:results_hardware_software}
+\add{
+We used three different setups to evaluate \sys: 
+(i) an RF-powered energy-harvesting device, WISP\,5.1~\cite{wisp5,wisp}, with a fixed capacitor size of 47\,$\mu$f; 
+(ii) an MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} powered by a BQ25570~\cite{BQ25570EVM-206_website} solar power harvester that is connected to an IXYS SLMD121H04L solar cell~\cite{SLMD121H04L_website} for experimenting with different energy buffers; and
+(iii) an MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} that is continuously powered.
+}
 
-We used two platforms to evaluate \sys: (i) an \emph{intermittently-powered}
-energy-harvesting device (WISP\,5.1~\cite{wisp5,wisp}), and (ii) a \emph{continuously-powered} evaluation
-board (MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad}).
-%
+\add{
 Every benchmark used in Section~\ref{sec:evaluation} was run repeatedly on
 each platform for a few minutes to ensure capturing a diverse power 
-trace and produce a high number of complete runs of all applications.
-The results were averaged across the duration of the experiment.
+trace. The exact number of iterations depends on the ambient power intensity and the application itself.
+In our experiments, the number of complete runs ranges from 4 to 125.
+}
 
@@
 We powered WISP using an RF signal generator emitting a 20\,dBm sinusoidal wave at 915\,MHz.
-The signal generator was connected to the Liard RFMAX S9028PCRJ 8\,dBic
+The signal generator was connected to the Laird RFMAX S9028PCRJ 8\,dBic
 antenna~\cite{atlas2015}.
@@
 %
-We affixed WISP with a paper harness at the edge of a table at a height, from the table suface, of 10\,cm.
+We affixed WISP with a paper harness at the edge of a table at a height, from the table surface, of 10\,cm.
 %
@@
 under profile, and the Saleae~\cite{saleae} logic analyzer measured
-intervals between edges in the signal. For continous power experiments, the 
+intervals between edges in the signal. For continuous power experiments, the 
 execution time was measured using the clock features in TI Code Composer Studio IDE version 7.1.
@@
 
-\begin{table}
-	\centering
-	\footnotesize
-	\begin{tabular}{| r|r|r | p{0.54\columnwidth} |}
+\begin{table}%
+\caption{Simulation Configuration}
+\label{tab:one}
+\begin{minipage}{\columnwidth}
+\begin{center}
+\begin{tabular}{llll}
+  \toprule
+		App.&Tasks&SLOC&Description\\
 		\hline
-		App.&Tasks&SLOC&Description\\
-		\hline\hline
         \emph{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
-		\hline
         \emph{bc} &10 &371 & several bitcount algorithms\\
@@
 % using several algorithms and check for correctness\\
-		\hline
         \emph{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
-		\hline
         \emph{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
-		\hline
         \emph{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
-		\hline
 		\emph{sort} &4 &167 & selection sort algorithm\\
-		\hline
+  \bottomrule
 \end{tabular}
+\end{center}
+\end{minipage}
 \caption{Characteristics of benchmarks used for evaluation.}
 \label{table:benchmark_table}
-\end{table}
+\end{table}%
+
+%\begin{table}
+%	\centering
+%	\footnotesize
+%	\begin{tabular}{| r|r|r | p{0.54\columnwidth} |}
+%		\hline
+%		App.&Tasks&SLOC&Description\\
+%		\hline\hline
+%        \emph{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
+%		\hline
+%        \emph{bc} &10 &371 & several bitcount algorithms\\
+%        % \textbf{bc} &10 &371 & Counts bits in bytes sequences
+%% using several algorithms and check for correctness\\
+%		\hline
+%        \emph{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
+%		\hline
+%        \emph{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
+%		\hline
+%        \emph{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
+%		\hline
+%		\emph{sort} &4 &167 & selection sort algorithm\\
+%		\hline
+%	\end{tabular}
+%\caption{Characteristics of benchmarks used for evaluation.}
+%\label{table:benchmark_table}
+%\end{table}
 

### 2019-06-22 paper/TOSN/intro.tex
TOSN coala version
minor revision TOSN

--- 2018-08-07 paper/asplos19/intro.tex
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/intro-figure-v5.pdf}
+    \includegraphics[width=0.8\columnwidth]{figures/intro-figure-v5.pdf}
     \caption{Task coalescing reduces commit overhead (left), while task splitting enables termination of large tasks (right).
@@
 Individual variables tracking, however, slows down a system dramatically.
-Therefore, the VMM keeps memory consistent through \emph{privatizing pages} and optimizes optimize bulk shared-data transfer through Direct Memory Access (DMA).
+Therefore, the VMM keeps memory consistent through \emph{privatizing pages} and optimizes bulk shared-data transfer through Direct Memory Access (DMA).
 
@@
 
-The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
+The rest of this paper is organized as follows.
+Section~\ref{sec:background} provides background on intermittent computing.
 Section~\ref{sec:systemdescription} illustrates an overview of \sys.

### 2019-06-24 paper/TOSN/response-letter/responseLetter.tex
First iteration

@@
 
+\newenvironment{additionalQuestions}{%
+\textbf{\large Additional questions}
+\begin{enumerate}%
+\renewcommand{\labelenumi}{\textbf{[Q:\,\arabic{enumi}]}} %
+}{\end{enumerate}}
+
 \newenvironment{editor}{%
-\textbf{\large Additional changes}
+\textbf{\large Additional remarks}
 \begin{enumerate}%
-\renewcommand{\labelenumi}{\textbf{[E:\,\arabic{enumi}]}} %
-}{\pagebreak\end{enumerate}}
+\renewcommand{\labelenumi}{\textbf{[R:\,\arabic{enumi}]}} %
+}{\end{enumerate}}
 
@@
 \begin{enumerate}%
-\renewcommand{\labelenumi}{\textbf{[R\therefereeCounter :\,\arabic{enumi}]}} %
+\renewcommand{\labelenumi}{\textbf{[C:\,\arabic{enumi}]}} %
 }{\end{enumerate}}
@@
 
-We would like to thank you for handling our submission and providing valuable review results. 
+We would like to thank for handling our submission and providing valuable review results. 
 We are also grateful to the reviewers for their substantial effort spent for evaluating our work. \\
@@
 Therefore, we restated their claim and provided the supporting references, which are [12, 42].\\
- Coala addresses deficiencies in the \emph{static} task-based systems and provides a more efficient alternative.
-Therefore, we compared its performance against the performance of the state-of-the-art task-based runtime.}
+ Coala addresses deficiencies in the \emph{static} task-based systems and provides more efficient alternative. Therefore, we compared its performance against the state-of-the-art task-based runtime.}
 
@@
 
-\response{The number of complete runs and WISP capacitor size are added to Section 7.1.}
+\response{The number of complete runs, which ranges between 4 and 125, and WISP capacitor size are added to Section 7.1.}
 
@@
 
-\response{We normalized these two special cases on 1 second. This infromation is added to Section 8.2, and the figure reference is corrected. }
+\response{We normalized these two special cases on 1 second. This infromation is added to Section 8.2; additionally, the figure reference is corrected. }
 
@@
 
-\response{As it is indicated in \hyperref[R1:1]{[R1:1]}, Coala addresses deficiencies in the \emph{static} task-based systems and provides a more efficient alternative.
+\response{As it is indicated in response \hyperref[R:1]{[R:1]}, Coala addresses deficiencies in the \emph{static} task-based systems and provides a more efficient alternative.
 Therefore, we compared its performance against the performance of the state-of-the-art task-based runtime.}
@@
 \referee{Also, I would suggest incorporating error bars in all the graphs as your "results were averaged across the duration of the experiment"}
+
 \response{We chose to normalize the results to focus on the speedup (or slowing down) and that is why we omitted the error-bars.}
@@
 \referee{In the abstract (and also in the introduction), authors claim that "Coala reduces run time by up to 54\%", however, I didn't see this number again anywhere in the evaluation.}
+
 \response{This number is highlighted in Section 8.2.}
 
-\referee{typos}
-\response{are corrected}
+\referee{mwntioned typos}
+
+\response{These typos are corrected.}
 
@@
 
+\begin{additionalQuestions}
+\referee{Does this paper cite and use appropriate references?: No\\
+If not, what important references are missing?: 
+Ratchet and Clank were the first to propose using a timer for breaking up a long task so that it can make forward progress during short power cycles.
+The authors propose the idea as if they were the first to use it.}
+
+\response{In section 4.2, we explicitly mentioned that the timer-based approach is similar to [72].
+[72] references the Ratchet paper.
+Table 1 mentioned Ratchet and Clank by names and they are referenced in our original paper.}
+\end{additionalQuestions}
+
 \begin{editor}
 \referee{}
-\response{Dear editor we added a new figure to the paper, Fig. 5. The figure's sole purpose is to aid explaining the concept of \textit{task coalescing}.
-We hope that this modification is approved by you.}
+\response{We have added a new figure to the paper, i.e. Fig. 5. The figure's sole purpose is to aid explaining the concept of \textit{task coalescing}.
+We hope that this modification further helps in understanding the core contribution of our paper.}
 \end{editor}

### 2019-06-24 paper/TOSN/background.tex
center figure and add label
minor revision TOSN
First iteration

@@
   \toprule
-        Model & Data Copied to/from NVRAM \\
+        \textbf{Model} & \textbf{Data Copied to/from NVRAM} \\
         \hline
@@
 An intermittent execution is composed of operating periods interspersed with power failures~\cite{dino,chain,alpaca,ratchet}.
-The frequency of failures depends on the size of the device's energy storage buffer (a larger buffer allows longer operating periods), current consumption and incoming energy.
+The frequency of failures depends on the size of the device's energy storage buffer (a larger buffer allows longer operating periods), current consumption and incoming power.
 
 \begin{figure}		
-	\begin{subfigure}{0.4\columnwidth}
 	\centering
-		\includegraphics[width=\columnwidth]{figures/war-example.pdf}
+	\begin{minipage}{.8\textwidth}
+	\begin{subfigure}{.5\columnwidth}
+			\includegraphics[width=\columnwidth, trim={.3cm 0 0 0}]{figures/war-example.pdf}
 		\caption{WAR-affected code}
@@
 	\end{subfigure}
-	\begin{subfigure}{0.4\columnwidth}
-		\centering
-		\includegraphics[width=\columnwidth]{figures/war-execution.pdf}
+	\begin{subfigure}{.5\columnwidth}
+			\includegraphics[width=\columnwidth,trim={.3cm 0 0 0 }  ]{figures/war-execution.pdf}
 		\caption{WAR-affected execution}
@@
 	\caption{WAR dependency example. \texttt{NV} marks non-volatile variables, \texttt{checkpoint()} is a checkpoint of volatile state.	\label{fig:w-a-r}}
+	\end{minipage}
 \end{figure}
@@
 Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
-Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes requiring hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}, Table~\ref{table:chechpoint_comparison} summarizes the data that gets copied to and from the non-volatile memory in service of memory consistency and progress preservation.
+Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes requiring hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}, Table~\ref{table:chechpoint_comparison} summarizes the data that get copied to and from the non-volatile memory in service of memory consistency and progress preservation.
 Checkpointing volatile state alone does not ensure data consistency when the system can directly manipulate non-volatile memory~\cite{mspcdino}.

### 2019-06-24 paper/TOSN/coalescing.tex
First iteration

@@
 {\em Energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly.
-In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits, and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure.
+In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure.
 The characteristic operations of EO are
@@
 \sys takes advantage of the guarantee that an energy buffer offers after a reboot: code execution resumes after energy accumulation reaches the operational threshold (a maximum threshold).
-Thus, \sys starts with a big coalesced task and shrinks it gradually as the risk of a power failure increases.}}
+Thus, \sys starts with a big coalesced task and shrinks it gradually as the risk of a power failure increases.
+T stands for task code execution and C for commit or power cut.}}
     \label{fig:coal_anatomy}
@@
 \add{
-Fig.~\ref{fig:coal_anatomy} shows \sys in action.
+Fig.~\ref{fig:coal_anatomy} shows a real measurement of \sys coalesced tasks.
 It shows how \sys coalesces aggressively when the execution is just re-started, taking advantage of the full energy buffer, and how it reduces the coalesced task gradually to minimize the risk of significant progress loss when the device fails.
-The average measured static task size is $\approx$\,0.28 $ms$ while the commit time (time needed to save the data into the non-volatile memory) averages at  $\approx$\,0.64 $ms$.
+The average measured static task size is $\approx$\,0.28\,ms while the commit time (time needed to save the data into the non-volatile memory) averages at  $\approx$\,0.64\,ms.
 This highlights the importance of coalescing.

### 2019-06-24 paper/TOSN/evaluation.tex
First iteration

@@
-% \end{wrapfigure}
-%
-%Building on top of the existing work, we have selected the most efficient task-based run-time library~\cite{alpaca} and compared \sys against it.
 Our evaluation quantitatively demonstrates that \sys (i) \emph{reduces memory
@@
 %
-Each type of protected access incurs a different.
+Each type of protected access incurs a different overhead.
 Accessing the most recently used SRAM page is of the cheapest kind.
@@
 execution time. We first investigate different variants of \sys and
-then compare the best variant to Alpaca~\cite{alpaca}.
+then compare the best variant to Alpaca~\cite{alpaca}. Additionally, we compare \sys
+performance running with different energy buffer sizes.
 
@@
 
-\caption{\add{Comparison of \sys performance on two different capacitor sizes. We see that \sys optimizes its coalesced task size based on the energy buffer size.
-\textbf{coalesced task} refers to the first coalesced task after a reboot, \textbf{exp. time} shows the experiment duration, and the \textbf{runs} column lists 
-the number of complete runs of the \textit{sort} application during the experiments.}
+\caption{\add{Comparison of \sys performance on two different capacitor sizes using the \textit{sort} application. We see that \sys optimizes its coalesced task size based on the energy buffer size.
+\textbf{Coalesced task} refers to the first coalesced task after a reboot, \textbf{Exp.
+time} shows the experiment duration, the \textbf{Runs} column lists the number of complete runs of the application during the experiments.
+\textbf{Run time} is the device collective uptime needed to finish a single iteration of the \textit{sort} application.}
 }
@@
   \toprule
-    \textbf{cap. size} ($\mu\text{f}$)& \textbf{exp. time} ($s$)& \textbf{duty cycle} & \textbf{coalesced task} ($ms$)   & \textbf{runs} & \textbf{app run-time} ($ms$) \\
+    \textbf{Cap. size} ($\mu\text{f}$)& \textbf{Exp. time} ($s$)& \textbf{Duty cycle} & \textbf{Coalesced task} ($ms$)   & \textbf{Runs} & \textbf{Run time} ($ms$) \\
         \hline
@@
 %
-In \textit{fft}, tasks are not uniform, and accounting for their different weights
+In \textit{fft}, tasks are not uniform, and accounting for their different weights is
 beneficial. In fact, the lack of task energy awareness is detrimental: with EG
@@
 
-\paragraph{\add{Different Capacitor Sizes}}
-\add{
-Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy.
-We see that \sys scales up its coalesced task size with a bigger energy buffer and vise versa. This allows it to reduce the time-to-completion of the applications. For example, 
-the \textit{sort} run-time is reduced from 183\,$ms$ to 177\,$ms$ when the capacitor is changed from 47\,$\mu$f to 470\,$\mu$f.
-It should be emphasized that a device with a bigger energy buffer suffers less from power failures (but requires longer charging time).
-However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.
-}
-
 \paragraph{Benefits of Adaptive Tasks.}
@@
 such patterns.  In applications without locality among accesses to protected
-variables, including array elements, \sys incurs overhead from memory
+variables \sys incurs overhead from memory
 virtualization that causes its performance to be comparable to (or worse than)
@@
 Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
-execution is negligible and stored energy is insufficient to complete some of
+execution is negligible and the stored energy is insufficient to complete some of
 the static tasks.}.
@@
 \sys, however, managed to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.
+
+\paragraph{\add{Different Capacitor Sizes}}
+\add{
+Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy.
+We see that \sys scales up its coalesced task size with a bigger energy buffer and vice versa. This allows it to reduce the time-to-completion of the applications. For example, 
+the \textit{sort} run-time is reduced from 183\,ms to 177\,ms when the capacitor is changed from 47\,$\mu$F to 470\,$\mu$F.
+It should be emphasized that a device with a bigger energy buffer suffers less from power failures (but requires longer charging time).
+However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.
+}
 

### 2019-06-24 paper/TOSN/hwswsetup.tex
First iteration

@@
 We used three different setups to evaluate \sys: 
-(i) an RF-powered energy-harvesting device, WISP\,5.1~\cite{wisp5,wisp}, with a fixed capacitor size of 47\,$\mu$f; 
+(i) an RF-powered energy-harvesting device, WISP\,5.1~\cite{wisp5,wisp}, with a fixed capacitor size of 47\,$\mu$F; 
 (ii) an MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} powered by a BQ25570~\cite{BQ25570EVM-206_website} solar power harvester that is connected to an IXYS SLMD121H04L solar cell~\cite{SLMD121H04L_website} for experimenting with different energy buffers; and
@@
 configured to 1\,MHz clock speed.
-%
 We powered WISP using an RF signal generator emitting a 20\,dBm sinusoidal wave at 915\,MHz.
@@
 antenna~\cite{atlas2015}.
-%
 The antenna was oriented towards and in parallel with WISP's antenna, and
 no objects obstructed the path.
-%
 We affixed WISP with a paper harness at the edge of a table at a height, from the table surface, of 10\,cm.
-%
 For distance-controlled experiments we positioned WISP at $d=\{15, 30,
 50\}$\,cm from the exciter antenna.
-%
-% We chose relatively shorter distances than in a deployment scenario to reduce
-% charge time and run experiments faster, after having confirmed ourselves a
-% result also reported in prior work~\cite{alpaca} that the amount of energy
-% available in one charge cycle---the sum of stored energy and incoming energy
-% during execution---is the same at distances beyond 50\,cm where incoming energy
-% becomes negligible.
-%
 To obtain execution time, the software toggled GPIO pins at sections of the code
@@
         \emph{bc} &10 &371 & several bitcount algorithms\\
-        % \textbf{bc} &10 &371 & Counts bits in bytes sequences
-% using several algorithms and check for correctness\\
         \emph{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
@@
 
-%\begin{table}
-%	\centering
-%	\footnotesize
-%	\begin{tabular}{| r|r|r | p{0.54\columnwidth} |}
-%		\hline
-%		App.&Tasks&SLOC&Description\\
-%		\hline\hline
-%        \emph{ar} &10 &428 & activity recognition using a KNN\\ % with a deterministic input from a randomly-generated sequence of accelerometer data\\
-%		\hline
-%        \emph{bc} &10 &371 & several bitcount algorithms\\
-%        % \textbf{bc} &10 &371 & Counts bits in bytes sequences
-%% using several algorithms and check for correctness\\
-%		\hline
-%        \emph{cuckoo} &14 &426 & Cuckoo Filter with pseudo-random values\\
-%		\hline
-%        \emph{dijkstra} &5 &198 & Dijkstra shortest path algorithm \\
-%		\hline
-%        \emph{fft} &8 &449 & Fast Fourier Transform\\ % on three pre-generated 128 sample vectors\\
-%		\hline
-%		\emph{sort} &4 &167 & selection sort algorithm\\
-%		\hline
-%	\end{tabular}
-%\caption{Characteristics of benchmarks used for evaluation.}
-%\label{table:benchmark_table}
-%\end{table}
-
 We evaluated \sys using six benchmarks that are often used in embedded systems (summarized in Table~\ref{table:benchmark_table}).
@@
 We compare \sys against Alpaca~\cite{alpaca}, the state-of-the-art task-based system for intermittent computing.
-For a fair comparison the task decomposition of the benchmarks are ensured to be the same for both systems.
+For a fair comparison, the task decomposition of the benchmarks is ensured to be the same for both systems.
 %
@@
 The instrumentation consists of identifying WAR dependencies  and adding code and memory allocations to make a private copy of the affected variables.
-% For array variables, instrumentation also includes adding code for dynamically tracking the array access. 
-%
-%Manual instrumentation of Alpaca code is not a general approach, but only served as a practical alternative to porting the compiler pass from LLVM to GCC.
-%
-%Our experiments show that manually-instrumented Alpaca code compiled with GCC is about 30\%-40\% faster than Alpaca code instrumented automatically by the Alpaca LLVM compiler pass and compiled with LLVM.
-This performance difference is due to the early experimental state of the MSP430 backend for LLVM~\cite{baghsorkhi_cgo_2018}.
-%
 By compiling both systems with the same compiler (GCC), we ensure that our comparison in Section~\ref{sec:evaluation} is fair.

### 2019-06-24 paper/TOSN/implementation.tex
First iteration

@@
 
-\noindent\textit{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed at the 
-end of the current one task and it can be invoked along
+\noindent\textit{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed after the current one task and it can be invoked along
 any control path to dynamically determine the next task.
@@
 
-First the scheduler updates the coalescing target using the coalescing
+First, the scheduler updates the coalescing target using the coalescing
 strategy.  Second, the scheduler resumes a commit, if one is in progress. Third,
@@
 Before executing the task, \sys checks whether there is a partially committed
-task to resume, which requires \sys to restore volatile state, including the program
+task to resume, which requires \sys to restore the volatile state, including the program
 counter.  

### 2019-06-24 paper/TOSN/paging.tex
TOSN coala version
First iteration

--- 2018-08-07 paper/asplos19/paging.tex
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/coalesced-war.pdf}
+	\includegraphics[width=0.8\columnwidth]{figures/coalesced-war.pdf}
 	\caption{Task coalescing-induced WAR dependency (in the case of this example, at variable $x$). Such dynamic dependencies cannot be resolved at compile-time and they break memory consistency.}
@@
 VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
-\texttt{\underline{private}} holds the consistent version of pages after each commit (and on reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
+\texttt{\underline{private}} holds the consistent version of pages after each commit (and on a reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
 The VMM prohibits applications from directly accessing these buffers.
@@
     \centering
-    \includegraphics[width=\columnwidth]{figures/mem-man.pdf}
+    \includegraphics[width=0.8\columnwidth]{figures/mem-man.pdf}
     \caption{Interaction between task, memory manager and memory buffers upon \texttt{RP} and \texttt{WP}, and order of occurrence. Steps 2, 3, 4 and 5 are conditional.}
@@
     \label{algo:rwar}
-    \scriptsize
+    \small
     \begin{algorithmic}[1]
@@
 The \texttt{\underline{committing}} bit indicates that a commit is in progress and is set before the first page is committed (Line 9) and cleared after the last page is committed (Line 16).
-The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to commit and the VMM clears the counter when commit completes (Line 14).
+The \texttt{\underline{shadowCount}} records the number of dirty shadow pages to be committed and the VMM clears the counter when commit completes (Line 14).
 The \texttt{\underline{commitIndex}} indexes the next page to be committed (Line 11) and the VMM clears the index at the end of the phase (Line 15).
@@
 	\label{algo:commit}
-	\scriptsize
+	\small
 	\begin{algorithmic}[1]
@@
 
-\sys asks the programmer to use its {\tt RP} and {\tt WP} API methods on every access to a protected variable (cf.
-Section~\ref{sec:coala_api}).
+\sys asks the programmer to use its {\tt RP} and {\tt WP} API methods on every access to a protected variable (Section~\ref{sec:coala_api}).
 These API invocations present a risk of high overhead because there is a dynamic check on every read and write.

### 2019-06-24 paper/TOSN/systemdescription.tex
TOSN coala version
First iteration

--- 2018-08-07 paper/asplos19/systemdescription.tex
@@
 	\centering
-	\includegraphics[width=\columnwidth]{figures/system-overview.pdf}
+	\includegraphics[width=.8\columnwidth]{figures/system-overview.pdf}
 	\caption{\sys top-level view. APP: application, ATS: adaptive task scheduler, VMM: Virtual Memory Manager, PM: physical memory.}
@@
 
-\sys is a new programming and execution model that supports adaptive task-based execution and eliminates restructuring and re-compilation of applications considering device's energy buffer.
-\sys addresses the challenges given in Section~\ref{sec:background} by making task-based intermittent applications portable in the sense of different energy storage sizes, while keeping their execution efficient.
+\sys is a new programming and execution model that supports adaptive task-based execution and eliminates restructuring and re-compilation of applications considering the device's energy buffer.
+\sys addresses the challenges given in Section~\ref{sec:background} by making task-based intermittent applications portable in the sense of different energy storage sizes while keeping their execution efficient.
 Fig.~\ref{fig:system_overview} shows an overview of \sys.
@@
 VMM achieves page privatization by keeping all page modifications in non-volatile memory on a coalesced task transition.
-Section~\ref{sec:memory_virtulaization} outlines VMM.
+Section~\ref{sec:memory_virtulaization} explains VMM.

### 2019-06-25 paper/TOSN/response-letter/responseLetter.tex
related work and minor revision

@@
 
-\response{In section 4.2, we explicitly mentioned that the timer-based approach is similar to [72].
-[72] references the Ratchet paper.
-Table 1 mentioned Ratchet and Clank by names and they are referenced in our original paper.}
+\response{Ratchet and Clank are mentioned and referenced in our original paper. 
+In Section 4.2, we explicitly mentioned that the timer-based approach is similar to [72]; [72] references the ``Ratchet'' paper. 
+In section 9 we explained the core difference between our timer-based approach and the approach explained  in ``Ratchet'' paper.
+Furthermore, In section 9 and in Table 1 we mentioned Ratchet and Clank by names.}
 \end{additionalQuestions}

### 2019-06-25 paper/TOSN/relatedwork.tex
related work and minor revision

@@
 The consistency problem was also approached with a combination of undo- and redo-logging in software~\cite{baghsorkhi_cgo_2018}.
-Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all volatile state immediately before power failure and halting the execution.
+Just-in-time checkpointing, such as Quickrecall~\cite{quickrecall} and Hibernus++~\cite{hibernusplusplus}, eschews inconsistency by saving all volatile state immediately before a power failure and halting the execution.
 Unlike \sys, such systems rely on introspection hardware to monitor supply voltage and on accurate worst-case bounds on checkpoint cost.
@@
 Checkpointing systems that copy most volatile state scale poorly as the size of volatile memory increases.
-QuickRecall, Clank and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory---requires more time and energy to access than volatile memory (cf.
+QuickRecall, Clank, and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory---requires more time and energy to access than volatile memory (cf.
 Section~\ref{sec:cost_task-based}) and not a viable option for off-chip non-volatile memory.
@@
 \paragraph{Task-based Systems.}
-Alternatives to checkpointing are recent systems based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca} and Mayfly~\cite{hester_sensys_2017}.
+Alternatives to checkpointing are recent systems based on static tasks, such as Chain~\cite{chain}, Alpaca~\cite{alpaca}, inK~\cite{yildirim2018ink}, and Mayfly~\cite{hester_sensys_2017}.
 Using static tasks, they eliminate the need to checkpoint volatile state.

### 2019-06-25 paper/TOSN/response-letter/responseLetter.tex
tiny correction

@@
 In Section 4.2, we explicitly mentioned that the timer-based approach is similar to [72]; [72] references the ``Ratchet'' paper. 
-In section 9 we explained the core difference between our timer-based approach and the approach explained  in ``Ratchet'' paper.
-Furthermore, In section 9 and in Table 1 we mentioned Ratchet and Clank by names.}
+In Section 9 we explained the core difference between our timer-based approach and the approach explained  in ``Ratchet'' paper.
+Furthermore, In Section 9 and Table 1 we mentioned Ratchet and Clank by names.}
 \end{additionalQuestions}

### 2019-06-27 paper/TOSN/response-letter/responseLetter.tex
first revised version

@@
 
-\response{As it is indicated in response \hyperref[R:1]{[R:1]}, Coala addresses deficiencies in the \emph{static} task-based systems and provides a more efficient alternative.
+\response{As it is indicated in response \hyperref[C:1]{[C:1]}, Coala addresses deficiencies in the \emph{static} task-based systems and provides a more efficient alternative.
 Therefore, we compared its performance against the performance of the state-of-the-art task-based runtime.}
@@
 
-\referee{mwntioned typos}
+\referee{mentioned typos}
 

### 2019-06-27 paper/TOSN/evaluation.tex
first revised version

@@
 
-\caption{\add{Comparison of \sys performance on two different capacitor sizes using the \textit{sort} application. We see that \sys optimizes its coalesced task size based on the energy buffer size.
+\caption{\add{Comparison of \sys performance running the \textit{sort} application on two different capacitor sizes. We see that \sys optimizes its coalesced task size based on the energy buffer size.
 \textbf{Coalesced task} refers to the first coalesced task after a reboot, \textbf{Exp.
@@
 \sys provides a performance benefit compared to Alpaca for most
-applications. \add{For example, it is 54\% faster than Alpaca when executing the \textit{sort} application. In general,}  the speedup is greatest for applications with repeated WAR
+applications. \add{For example, it is 54\% faster than Alpaca when executing the \textit{bc} application. In general,}  the speedup is greatest for applications with repeated WAR
 dependencies throughout their code, particularly involving arrays

### 2019-06-28 paper/TOSN/response-letter/responseLetter.tex
minor corrections

@@
 We would like to thank for handling our submission and providing valuable review results. 
-We are also grateful to the reviewers for their substantial effort spent for evaluating our work. \\
+We are also grateful to the reviewers for their substantial effort spent on evaluating our work. \\
 
-With this letter we address all comments of the reviewers and point to the modified sections of our revised paper. \\
+With this letter, we address all comments of the reviewers and point to the modified sections of our revised paper. \\
 
@@
 Therefore, we restated their claim and provided the supporting references, which are [12, 42].\\
- Coala addresses deficiencies in the \emph{static} task-based systems and provides more efficient alternative. Therefore, we compared its performance against the state-of-the-art task-based runtime.}
+ Coala addresses deficiencies in the \emph{static} task-based systems and provides a more efficient alternative.
+Therefore, we compared its performance against the state-of-the-art task-based runtime.}
 
@@
 
-\response{We normalized these two special cases on 1 second. This infromation is added to Section 8.2; additionally, the figure reference is corrected. }
+\response{We normalized these two special cases on 1 second. This information is added to Section 8.2; additionally, the figure reference is corrected. }
 
@@
 \referee{}
-\response{We have added a new figure to the paper, i.e. Fig. 5. The figure's sole purpose is to aid explaining the concept of \textit{task coalescing}.
+\response{We have added a new figure to the paper, i.e. Fig. 5. The figure's sole purpose is to aid in explaining the concept of \textit{task coalescing}.
 We hope that this modification further helps in understanding the core contribution of our paper.}

### 2019-06-28 paper/TOSN/abstract.tex
working on the rebuttal
minor corrections

--- 2018-08-07 paper/asplos19/abstract.tex
@@
 Therefore, the paradigm of intermittent execution is to save the program state into non-volatile memory frequently to preserve the execution progress.
-In task-based intermittent programming the state is saved at task transition.
+In task-based intermittent programming, the state is saved at task transition.
 Tasks are fixed at compile-time and agnostic to energy conditions.
 Thus, the state may be saved either more often than necessary or not often enough for the program to progress and terminate.
-To address this challenges, we propose \sys, an adaptive and efficient task-based execution model.
+To address these challenges, we propose \sys, an adaptive and efficient task-based execution model.
 \sys progresses on a multi-task scale when energy permits and preserves the computation progress on a sub-task scale if necessary.

### 2019-06-28 paper/TOSN/background.tex
related work and minor revision
minor corrections

@@
 For instance, computational RFIDs---open-source TI MSP430-based~\cite{wolverine} WISP~\cite{wisp5} (with its variants such as WISPCam~\cite{naderiparizi_rfid_2015}, NFC-WISP~\cite{zhao_rfid_2015} or NeuralWISP~\cite{holleman_biocas_2008}), Moo~\cite{moo}, and commercial ones such as~\cite{medusa_farsens_2017}.
-Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014} or battery-less phone~\cite{talla_imwut_2017}.
+Other platforms include ambient backscatter tag~\cite{liu_sigcomm_2013,parks_sigcomm_2014,majid2019multi} or battery-less phone~\cite{talla_imwut_2017}.
 %
@@
 A power failure clears volatile state (e.g., registers and SRAM) while non-volatile memory (e.g., FRAM) persists.
-Upon a power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
+Upon power failure, control flows to a prior point in the execution: by default, to the beginning of {\tt main()}.
 Early intermittent systems preserved progress by periodically checkpointing volatile execution context to non-volatile memory~\cite{mementos}, sometimes requiring hardware support~\cite{mottola2017harvos,hibernusplusplus,hibernus,idetic,quickrecall}, Table~\ref{table:chechpoint_comparison} summarizes the data that get copied to and from the non-volatile memory in service of memory consistency and progress preservation.
@@
 Precisely, data can get inconsistent when code includes a \emph{write-after-read} (WAR) dependency between operations that manipulate non-volatile memory.
-Fig.~\ref{fig:w-a-r} illustrates how state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
+Fig.~\ref{fig:w-a-r} illustrates how a proram state can become inconsistent in an intermittent execution using a simple example of an average operation over an array of integers.
 The non-volatile variable \texttt{sum} introduces the WAR, being read and written sequentially by the increment operation.

### 2019-06-28 paper/TOSN/evaluation.tex
minor corrections

@@
 Fig.~\ref{fig:overallOverheadBreakdown}. Without coalescing enabled (NC), the
-re-execution penalty is smallest, because the amount of work that can happen
+re-execution penalty is the smallest, because the amount of work that can happen
 between commits and may have to be re-executed if interrupted is reduced when

### 2019-06-28 paper/TOSN/intro.tex
minor corrections

@@
 Therefore, this work focuses on software approaches.
-In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} and \emph{energy-unawareness} of statically decomposing a program into tasks.
+In particular, we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} and \emph{energy-unawareness} of statically decomposing a program into tasks.
 %
@@
 Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks.
-\emph{Tasks}, top level functions, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.
+\emph{Tasks}, top-level functions, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.
 The programmer (or a compiler) explicitly expresses task-to-task control flow.
@@
 \sys accepts any static decomposition and it coalesces (groups) tasks or splits them (see again Fig.~\ref{fig:coalesce}) based on the estimated energy availability without demanding any hardware support.
-To the best of our knowledge, \sys is the only system that eliminates restructuring and re-compilation of applications considering device's energy buffer---enabling energy-storage independency for the applications, while keeping execution efficient.
+To the best of our knowledge, \sys is the only system that eliminates restructuring and re-compilation of applications considering a device's energy buffer---enabling energy-storage independency for the applications, while keeping execution efficient.
 

### 2019-06-28 paper/TOSN/paging.tex
minor corrections

@@
 On a commit, the VMM atomically copies, through a two-phase commit, all dirty pages back into their location in the non-volatile memory.
-% SRAM has a limited size and there is an extended working memory containing {\em shadow pages} in the larger FRAM.
-A shadow page stores the data from an updated page that needs to be committed back to non-volatile memory, but that exceeded the size of the volatile working memory.
 
@@
 If the page is not found there, a page fault is incurred and a new page needs to be swapped in.
-If the \texttt{working} buffer is full, a page fault on a memory access requires the VMM to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
+If the \texttt{working} buffer is full, a page fault on memory access requires the VMM to swap out one of the pages in the \texttt{working} buffer (a \emph{victim page}) preserving updates made to that page.
 If a task modified any byte in the victim page (i.e., using \texttt{WP}), then the page is dirty and its changes have to be persisted to non-volatile memory (Fig.~\ref{figure:mem-man}, Step 2).

### 2019-06-28 paper/TOSN/coalescing.tex
remove highlights

@@
 \includegraphics[width=.8\columnwidth]{figures/dif_cap}
-    \caption{\add{Anatomy of \sys coalescing---A snapshot of measured data showing how \sys coalesces static tasks.
+    \caption{Anatomy of \sys coalescing---A snapshot of measured data showing how \sys coalesces static tasks.
 \sys takes advantage of the guarantee that an energy buffer offers after a reboot: code execution resumes after energy accumulation reaches the operational threshold (a maximum threshold).
 Thus, \sys starts with a big coalesced task and shrinks it gradually as the risk of a power failure increases.
-T stands for task code execution and C for commit or power cut.}}
+T stands for task code execution and C for commit or power cut.}
     \label{fig:coal_anatomy}
@@
 
-\add{
 Fig.~\ref{fig:coal_anatomy} shows a real measurement of \sys coalesced tasks.
@@
 This highlights the importance of coalescing.
-}

### 2019-06-28 paper/TOSN/evaluation.tex
remove highlights

@@
 
-\caption{\add{Comparison of \sys performance running the \textit{sort} application on two different capacitor sizes. We see that \sys optimizes its coalesced task size based on the energy buffer size.
+\caption{Comparison of \sys performance running the \textit{sort} application on two different capacitor sizes. We see that \sys optimizes its coalesced task size based on the energy buffer size.
 \textbf{Coalesced task} refers to the first coalesced task after a reboot, \textbf{Exp.
@@
 \textbf{Run time} is the device collective uptime needed to finish a single iteration of the \textit{sort} application.}
-}
 \label{tab:dif_cap}
-\add{
+
 \begin{minipage}{\columnwidth}
@@
 \end{minipage}  
-}
 \end{table}%
@@
 Fig.~\ref{fig:runtime} shows the average execution time of each
-application for \sys and Alpaca, \add{normalized to the latter or, when not possible, to one second}.
+application for \sys and Alpaca, normalized to the latter or, when not possible, to one second.
 \sys provides a performance benefit compared to Alpaca for most
-applications. \add{For example, it is 54\% faster than Alpaca when executing the \textit{bc} application. In general,}  the speedup is greatest for applications with repeated WAR
+applications. For example, it is 54\% faster than Alpaca when executing the \textit{bc} application. In general, the speedup is greatest for applications with repeated WAR
 dependencies throughout their code, particularly involving arrays
@@
 
-\paragraph{\add{Different Capacitor Sizes}}
-\add{
+\paragraph{Different Capacitor Sizes}
 Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy.
@@
 However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.
-}
 

### 2019-06-28 paper/TOSN/hwswsetup.tex
remove highlights

@@
 \label{sec:results_hardware_software}
-\add{
+%
 We used three different setups to evaluate \sys: 
@@
 (iii) an MSP-EXP430FR5969 launchpad~\cite{MSP-EXP430FR5969_launchpad} that is continuously powered.
-}
 
-\add{
 Every benchmark used in Section~\ref{sec:evaluation} was run repeatedly on
@@
 In our experiments, the number of complete runs ranges from 4 to 125.
-}
 

### 2019-07-02 paper/TOSN/coalescing.tex
Alexei feedback

@@
         \While{$\texttt{true}$}       
-	        \State $ i \gets 0$
+	        \State $ j \gets 0$
 	        %
-	        \While{$ i <= C$}
-		        \State \Call{execute\_task}{$T_\texttt{i}$}
+	        \While{$ j \leq C$}
+		        \State \Call{execute\_task}{$T_\texttt{i}$} \Comment{$T_\texttt{i}$ is the $i$th task executed since the last power failure}
 		        \State $W \leftarrow $ \Call{$f_\text{weight}$}{$T_\text{i}$} \Comment{Assign new task-dependent weight}
-		        \State $i \gets i + W$
+		        \State $j \gets j + W$
 				\State $H \gets H + W$
@@
 
-{\em Energy-oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly.
+{\em Energy-Oblivious} (EO) coalescing strategy treats all tasks as having unity weight and varies the size of the coalescing target linearly.
 In such a scheme, the number of tasks to coalesce, $C$, increases by a constant $x$ when a coalesced sequence of tasks commits and decreases by the same constant $x$ when a coalesced sequence of tasks fails to complete, i.e., after a power failure.
@@
 Let us assume $x=1$.
-If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive power failures, before being able to progress.
+If, after $k$ successful commits, the target must decrease to its original value $C-k$ due to an energy drop, EO requires $k$ successive reboots to progress.
 
@@
 
-The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, addressing a key limitation of the EO, to adhere to changes in energy conditions.
+The \emph{Energy-Guided} (EG) coalescing strategy adapts its coalescing target more quickly, to adhere to changes in energy conditions,  addressing a key limitation of the EO.
 It uses its recent execution history, H, to estimate energy availability and alters its target accordingly.

### 2019-07-02 paper/TOSN/evaluation.tex
Alexei feedback

@@
 %
-Across all benchmarks, all coalescing strategies reduce more commit overhead
+Across all benchmarks, will all coalescing strategies \sys reduces more commit overhead
 than the re-execution overhead they add.
@@
   \toprule
-    \textbf{Cap. size} ($\mu\text{f}$)& \textbf{Exp. time} ($s$)& \textbf{Duty cycle} & \textbf{Coalesced task} ($ms$)   & \textbf{Runs} & \textbf{Run time} ($ms$) \\
+    \textbf{Cap. size} ($\mu\text{f}$)& \textbf{Exp. time} ($s$)& \textbf{on/off cycle} & \textbf{Coalesced task} ($ms$)   & \textbf{Runs} & \textbf{Run time} ($ms$) \\
         \hline
@@
 
-Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm\footnote{At distances greater than 15\,cm, energy incoming during
-execution is negligible and the stored energy is insufficient to complete some of
-the static tasks.}.
-This is marked with $\infty$ signs in Fig.~\ref{fig:runtime}.
-\sys, however, managed to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.
+Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm
+% \footnote{At distances greater than 15\,cm, energy incoming during
+% execution is negligible and the stored energy is insufficient to complete some of the static tasks.}. 
+\footnote{At distances less than 15\,cm the total amount of energy available for task execution includes significant amount of harvested-while-executing energy in addition to the stored energy.}.
+
+This is marked with $\infty$ signs in Fig.~\ref{fig:runtime}. \sys, however, managed to complete \textit{fft} by enabling its task downscaling at 30\,cm and 50\,cm from the energy source.
 
 \paragraph{Different Capacitor Sizes}
-Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy.
-We see that \sys scales up its coalesced task size with a bigger energy buffer and vice versa. This allows it to reduce the time-to-completion of the applications. For example, 
+Table~\ref{tab:dif_cap} shows how \sys optimizes its Coalescing task size based on the amount of buffered energy at runtime.
+This means that applications implemented in \sys are portable across devices with different capacitor sizes without recompilation; they are also more resilient to degradation in capacitor size due to temperature and device lifetime.
+
+We see that \sys scales up its coalesced task size with a bigger energy buffer and vice versa. 
+This allows it to reduce the time-to-completion of the applications. For example, 
 the \textit{sort} run-time is reduced from 183\,ms to 177\,ms when the capacitor is changed from 47\,$\mu$F to 470\,$\mu$F.
 It should be emphasized that a device with a bigger energy buffer suffers less from power failures (but requires longer charging time).
-However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.
+%However, as has been measured in Fig.~\ref{fig:overallOverheadBreakdown} and explained in Fig.~\ref{fig:coal_anatomy}, the re-execution penalty for the \textit{sort} application is negligible.
 

### 2019-07-02 paper/TOSN/implementation.tex
minor corrections
Alexei feedback

@@
 \begin{table}%
-\caption{Simulation Configuration}
-\label{tab:one}
+\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
 \begin{minipage}{\columnwidth}
@@
 \end{tabular}
-\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
 \label{table:coala_api}
@@
 \end{table}%
-%%
-%\begin{table}
-%\footnotesize
-%\centering
-%\begin{tabular}{| r | p{0.7\columnwidth} |}
-%	\hline
-%	{Method} & {Arguments} \\
-%	\hline\hline
-%	\texttt{INIT}($t$) & $t \in \mathbf{T}$: scheduled task on first boot \\
-%	\hline
-%	\texttt{RUN}() & --- \\
-%	\hline
-%	\texttt{TASK}($t$, $w_t$) & $t \in \mathbf{T}$: task name, $w_t$: weight of task $t$ \\
-%	\hline
-%	\texttt{NEXT\_TASK}($t$) & $t \in \mathbf{T}$ : task to be run next \\
-%	\hline
-%	\texttt{PV}($p$, $v$ [, $s$]) & $p$: type, $v \in \mathbf{V}$: name, $s$: array size \\
-%	\hline
-%	$u$ := \texttt{RP}($v$) & $v \in \mathbf{V}$: protected variable to read, $u$: dest.\ operand \\
-%	\hline	
-%	\texttt{WP}($v$) := $u$ &  $v \in \mathbf{V}$: protected variable to write, $u$: source operand \\
-%	\hline
-%	\texttt{SM}($p$, $m$ [, $s$]) & $p$: type, $m$: name, $s$: array size \\
-%	\hline
-%	\texttt{DISABLE\_PC}() & --- \\
-%	\hline
-%	\texttt{ENABLE\_PC}() & --- \\
-%	\hline
-%\end{tabular}
-%\caption{API summary; $\mathbf{T}$: set of all tasks, $\mathbf{V}$: set of all protected variables, $[, s]$: optional argument.}
-%\label{table:coala_api}
-%\end{table}
 
@@
 
-\noindent\textit{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed after the current one task and it can be invoked along
+\noindent\textit{Task Transitions.} \texttt{NEXT\_TASK} marks the task to be executed after the current one and it can be invoked along
 any control path to dynamically determine the next task.
@@
 variable statically allocates a protected non-volatile variable. The
-variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at
-run-time to ensure correct operation.
+variable must then be accessed with the \texttt{RP} and \texttt{WP} API methods at run-time to ensure correct operation.
+The \texttt{SM} annotation is used to align C structure data type within a page.
 
@@
 by calling \texttt{RUN} after device initialization. 
+\noindent\textit{Partial Commit.} The \texttt{DISABLE\_PC}() and \texttt{ENABLE\_PC}() allow a programmer to disable partial commit around certain code.
 
-First, the scheduler updates the coalescing target using the coalescing
-strategy.  Second, the scheduler resumes a commit, if one is in progress. Third,
-the scheduler clears the list of dirty shadow pages. The scheduler then 
-sets the program counter to the next task to run, which \sys tracks in  
+On a reboot, \sys's scheduler does a number of system level operations before executing a tast. 
+ First, it updates the coalescing target according to the applied coalescing
+strategy. Then, it finishes any interrupted commit and clears the list of dirty shasow pages.
+% Second, the scheduler resumes a commit, if one is in progress. 
+% Third, the scheduler clears the list of dirty shadow pages. 
+After that, the scheduler sets the program counter to the next task to run, which \sys tracks in  
 non-volatile memory.  
@@
 The exponential decrease of the timer value converges rapidly to a usable one.
-%
-%Partial commit relies on the existence of a hardware timer, which is a very common feature in a typical MCU and consumes very little energy (i.e., $\approx$3\,$\mu$A/MHz~\cite{msp430datasheet}).
 
@@
 %
-We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and
-an 8\,KB non-volatile \texttt{\underline{private}} buffer, and a \texttt{working} buffer
-of 1\,KB.
 
@@
 While we opted for the simplicity of FIFO, any other replacement policy, such as Least Recently Used, could work in its place.
+
+We experimented with 32-, 64-, 128- and 256\,B pages, an 8\,KB non-volatile \texttt{\underline{shadow}} buffer and an 8\,KB non-volatile \texttt{\underline{private}} buffer, and a \texttt{working} buffer
+of 1\,KB.

### 2019-07-02 paper/TOSN/paging.tex
Alexei feedback

@@
 This problem is illustrated in Fig.~\ref{figure:coalescedWar}.
-Therefore, \sys implements novel {\em Virtual Memory Manager} (VMM) which enables task coalescing guaranteeing data integrity, which is at the same time efficient.
+Therefore, \sys implements novel {\em Virtual Memory Manager} (VMM) that enables safe task coalescing and ensures efficient data manipulation.
 
@@
 VMM abstracts the physical address space of the non-volatile memory (FRAM) and divides it into \texttt{\underline{private}} and \texttt{\underline{shadow}} (underlined locations are non-volatile) buffers, and each buffer is divided into pages.
-\texttt{\underline{private}} holds the consistent version of pages after each commit (and on a reboot), while \texttt{\underline{shadow}} serves as support during the execution of coalesced tasks.
+\texttt{\underline{private}} holds the consistent version of pages after each commit (and on a reboot), while \texttt{\underline{shadow}} enables atomic two-phase commit: it allows \sys to ensure a persistent and consistent set of non-volatile pages before copying them to \texttt{\underline{private}} (Section~\ref{sec:two_phase_commit}).
+
 The VMM prohibits applications from directly accessing these buffers.
-Instead, it redirects any request to the \texttt{working} buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
+Instead, it redirects any request to the \texttt{working} buffer: a relatively small buffer located in the volatile memory (SRAM), which has a lower latency and energy cost to access than FRAM.
 It populates the \texttt{working} buffer with the privatized pages from non-volatile memory requested by tasks.
+With the help of \texttt{\underline{shadow}}, the \texttt{working} buffer can serve more pages than its own capacity.
+
 When a coalesced task completes, the VMM ensures that {\em dirty}, i.e., modified, pages are committed to FRAM and visible to subsequent tasks.
@@
 \subsection{Atomic Two-Phase Commit of Dirty Pages}
+\label{sec:two_phase_commit}
 
@@
 % \sys's implementation is also efficient because
-the VMM does not copy pages from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.
+the VMM does not copy page content from \texttt{\underline{shadow}} to \texttt{\underline{private}}, instead it swaps pointers in an indirection table that maintains the pages as a double buffer.
 
@@
 	\begin{algorithmic}[1]
-        \Procedure{CommitPhase1}{} \Comment{On last coalesced task}
+        \Procedure{CommitPhase1}{} \Comment{On completion of a coalesced task}
             \For {$i \in 0..|\texttt{working}|-1$}
-                \State $t \gets \Call{Tag}{\texttt{working}[i]}$
+                \State $f_\text{dirty}, t \gets \Call{TagFlag}{\texttt{working}[i]}$
+                % \State $f_\text{dirty} \gets \Call{DirtyFlag}{t}$
+                \If {$f_\text{dirty}$}   \Comment{Is the page content modified?}
                     \State $\texttt{\underline{shadow}}[\Call{Tag}{\texttt{working}[i]}] \xleftarrow{\text{DMA}} \texttt{working}[i]$
@@
                     \State $\texttt{\underline{shadowCount}} \gets \texttt{\underline{shadowCount}} + 1$
+                \EndIf
             \EndFor
@@
                 \State $t \gets \texttt{\underline{shadowList}}[\texttt{\underline{commitIndex}}]$
-                \State \Call{commitToPrivate}{\texttt{\underline{shadow}}[t]}
+                \State \Call{commitToPrivate}{\texttt{\underline{shadow}}[t]} \Comment{Copy \texttt{\underline{shadow}} to \texttt{\underline{private}} by swapping their pointers}
                 \State $\texttt{\underline{commitIndex}} \gets \texttt{\underline{commitIndex}} + 1$
@@
             \EndIf
-            \State $\texttt{\underline{shadowCount}} \gets 0$, $i_\text{victim} \gets 0$
+            \State $\texttt{\underline{shadowCount}} \gets 0$ %, $i_\text{victim} \gets 0$
         \EndProcedure

### 2019-07-02 paper/TOSN/partialcommit.tex
First iteration
Alexei feedback

--- 2018-08-07 paper/asplos19/partialcommit.tex
@@
 
-\sys uses {\em task downscaling} to make progress through a task that is too large to complete using the energy that the device can buffer.
+\sys uses {\em task downscaling} to make progress through a task that is too large to complete using the buffered energy.
 Task downscaling executes part of the long task and interrupts its execution with a {\em partial commit}, after which the long task continues.
@@
 Eventually, after some number of partial commits and power failures, the task will complete and execution will proceed.
-Task downscaling violates task atomicity in a non-terminating execution, favoring continued progress over atomicity.
-If a programmer requires a task's atomicity to be preserved, they can disable task downscaling for that task.
-The key design issue for task downscaling is deciding when to partially commit.
+
+\textit{Detection of Non-terminating Tasks.} The key design issue for task downscaling is deciding when to partially commit.
 A task is likely to be non-terminating if it fails to run to completion twice consecutively.
@@
 If a task cannot complete when executing with a full energy buffer, \sys marks the task as non-terminating.
-
-%timer, similarly used in~\cite{ratchet}.
+Task downscaling violates task atomicity in a non-terminating execution, favoring continued progress over atomicity.
+If a programmer requires a task's atomicity to be preserved, they can disable task downscaling for that task.

### 2019-07-02 paper/TOSN/relatedwork.tex
Alexei feedback

@@
 Checkpointing systems that copy most volatile state scale poorly as the size of volatile memory increases.
-QuickRecall, Clank, and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory---requires more time and energy to access than volatile memory (cf.
+QuickRecall, Clank, and Ratchet reduce the copying overhead by allocating the stack in non-volatile memory, whichrequires more time and energy to access than volatile memory (cf.
 Section~\ref{sec:cost_task-based}) and not a viable option for off-chip non-volatile memory.

### 2019-07-02 paper/TOSN/systemdescription.tex
Alexei feedback

@@
 Then, they compile and link the code against \sys's runtime, producing a \sys-enabled binary.
-The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution with the energy conditions.
+The runtime relies on \sys's novel {\em adaptive task scheduler} to adapt its execution to the energy conditions.
 Facilitating efficient task adaptation requires dynamic memory protection which \sys's \emph{virtual memory manager} handles through \emph{page privatization}.
@@
 \sys virtual memory manager (VMM) is the key enabler to ensure memory consistency while coalescing tasks.
-VMM allows applications to interface with only fast volatile memory pages and privatizes the pages demanded by a coalesced task in order to solve a novel data consistency problem; namely \emph{task coalescing-induced WAR dependency}.
+VMM allows applications to interface with only fast volatile memory pages and privatizes the pages demanded by a coalesced task in order to solve a novel data consistency problem; namely \emph{task coalescing-induced WAR dependencies}.
 VMM achieves page privatization by keeping all page modifications in non-volatile memory on a coalesced task transition.

### 2019-09-03 paper/TOSN/intro.tex
update

@@
 \paragraph{Challenges and Contributions.}
-We introduce \sys: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
+We introduce \sys~\footnote{\sys's source code is accessible via \url{https://github.com/TUDSSL/Coala} }: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
 By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overhead \emph{dynamically}, meanwhile avoiding the risk of non-termination.

### 2019-09-03 TaskCoalescingPlotting/csf.py

@@ -1,0 +1,234 @@
+"""
+This module is supporting library for other scripts
+"""
+
+import matplotlib.pyplot as plt
+import numpy as np
+import matplotlib.mlab as mlab
+import matplotlib.ticker as ticker
+import os
+import glob
+import random
+# import colors
+from collections import defaultdict
+import sys
+
+#plt.style.use('seaborn-darkgrid')
+
+# to reset the original style use `matplotlib.rcdefaults()` after the plot
+
+colorCnts=0
+
+def generateRandomColor():
+	global colorCnts
+	c  = colors.colors[colorCnts]
+	colorCnts+=1
+	colorCnts %=len(colors.colors)
+	return (c)
+
+def unified_input(aStr=""):
+	user_input =None
+	if( sys.version_info[0] ) < 3 :
+		user_input = raw_input(aStr)
+	else:
+		user_input = input(aStr)
+	return user_input
+
+
+dataUser = '../data/'
+scriptsDir = '../scripts/'
+
+def appsRootDirs():
+	os.chdir(dataUser)
+	return glob.glob('*')  # grep all apps folders
+
+
+def appsSelector():
+	apps = appsRootDirs()
+
+	disp_apps = apps[:]
+	# print(disp_apps)
+	user_choice=[]
+
+	while(1):
+
+		user_input = unified_input('to choose from '+str(disp_apps)+' or (done):')
+
+		if(user_input == 'done'):
+			break;
+		disp_apps.remove(user_input)
+		user_choice.append(user_input)
+
+	if user_choice !=[]:
+		apps = user_choice[:]
+	os.chdir(scriptsDir)
+
+	return apps
+
+def filesFinder(apps, path):
+	filesPaths = []
+	appsMainFolders = apps
+	os.chdir(dataUser)
+	for appFolder in appsMainFolders:
+		filesPaths.append( glob.glob( os.path.join(appFolder,path)) )
+
+	os.chdir(scriptsDir)
+	validPaths=[]
+	validApps=[]
+	for i, p in enumerate(filesPaths):
+		if p != []:
+			validPaths.append(p)
+			validApps.append(appsMainFolders[i])
+
+	return (validPaths)
+
+
+def dataMiner(fileName, memStart, memEnd, wordSize):
+	"""
+		This function extract data from a file formated in intel hex format
+		@ inputs:
+			filename
+			The start/end of the memory section in hex
[…]

### 2019-09-03 TaskCoalescingPlotting/logicAnalyzerData.py

@@ -1,0 +1,123 @@
+#
+# @Author: Amjad Majid
+# @Date  : March 14, 2019
+#
+import csv
+import numpy as np
+
+class DataException(Exception):
+    def __init__(self, msg):
+        super().__init__(msg)
+
+class LogicAnalyzerData:
+    """This class requires a csv file of time stamped binary states,
+    like files generated from Saleae logic analyzer. It extracts the 
+ i   rows within the given time interval, and it selects the states 
+    columns based on the given list of columns' indices. 
+    """
+
+    def __init__(self,path):
+        self.path = path
+        self.timestamps=[]
+        self.states=[]
+
+        self.__readData()
+
+    def __readData(self):
+        with open(self.path) as csvFile:
+            csvReader = csv.reader(csvFile, delimiter=',')
+
+            # If there is a header line drop it, otherwise rewind your file pointer 
+            line = csvReader.__next__()
+            if 'Time' not in line[0]:
+                csvFile.seek(0,0)
+
+            for line in csvReader:
+                self.states.append(list(map(int, line[1:])))
+                self.timestamps.append(float(line[0])) 
+            print(len(self.states))
+
+    def __timeRange(self,minTimestamp, maxTimestamp):
+        """ __timeRane(...) time range selector [...)
+
+            @param `minTimestamp` is the lower bound of the time interval. This bound is included  
+            @param `maxTimestamp` is the upper bound of the time interval. This bound is exculded 
+            @return `rowsIndices` is a list of row indices of the data within the given time interval
+        """
+        #TODO to optimize this method find the beginning of the time interval using binary search
+        rowsIndices = [self.timestamps.index(ts) for ts in self.timestamps \
+                if ts >= minTimestamp and ts <= maxTimestamp]
+        #print(rowsIndices)
+        if not rowsIndices:
+            raise DataException("Not data in the found between {} and {}".format(minTimestamp, maxTimestamp))
+        return rowsIndices 
+
+    def statesSelector(self, cols, states):
+        s = np.array(states)
+        # print(s)
+        # print("cols",cols)
+        return s[:,cols]
+
+    def getNumOfNodes(self):
+        return len(self.states[0])
+
+    def getTotalExperimentTime(self):
+        return self.timestamps[-1]
+
+    def getData(self, minTime, maxTime, cols):
+        # print("minMax",minTime,maxTime)
+        indices = self.__timeRange(minTime, maxTime)
+        # print("indices",indices)
[…]

### 2019-09-03 TaskCoalescingPlotting/main.py

@@ -1,0 +1,111 @@
+import numpy as np 
+import matplotlib.pyplot as plt 
+import matplotlib.patches as patches 
+
+from logicAnalyzerData import LogicAnalyzerData
+from plotter import Plotter
+from csf import *
+
+lad_47 = LogicAnalyzerData('sort_47uf_duty_cycle_12-9_extra-long.csv')
+lad_470 = LogicAnalyzerData('sort_470uf_duty_cycle_12-79_extra-long.csv')
+plt.figure(figsize=(8,2.5))
+
+base = 600.9
+signal_length = 1.6
+zoom_in_controll = 0.055
+shift = .9+zoom_in_controll  # for alignment 
+coal_tasks_470 = lad_470.getData(base,base+signal_length,[0])
+lad_470.intervalDataInterpolation(0.001) # in-place data manipulation 
+
+coal_tasks = lad_47.getData(base-shift,base-shift+signal_length,[0])
+lad_47.intervalDataInterpolation(0.001) # in-place data manipulation 
+static_tasks =    lad_47.getData(base-shift+zoom_in_controll+.192452,base-shift+zoom_in_controll+.23,[2])
+lad_47.intervalDataInterpolation(0.001) # in-place data manipulation 
+coalTask_detail = lad_47.getData(base-shift+zoom_in_controll+.192452,base-shift+zoom_in_controll+.23,[0])
+lad_47.intervalDataInterpolation(0.001) # in-place data manipulation 
+
+coal_tasks_470  = Plotter(coal_tasks_470 )
+
+coal_tasks = Plotter(coal_tasks)
+static_tasks = Plotter(static_tasks)
+coalTask_detail = Plotter(coalTask_detail)
+
+coal_tasks_470_x,coal_tasks_470_y = coal_tasks_470.getData()
+
+coal_tasks_x,coal_tasks_y = coal_tasks.getData()
+static_tasks_x,static_tasks_y = static_tasks.getData()
+coalTask_detail_x,coalTask_detail_y = coalTask_detail.getData()
+
+# plt.plot(coal_tasks_470_x,coal_tasks_470_y * .2)
+
+# plt.plot(coal_tasks_x+shift,coal_tasks_y * .2 + 0.21)
+
+figureSetting()
+
+
+plt.ylim([-0.1,.7])
+
+# sub_axes = plt.axes([.14,.6,.74,.3])
+# sub_axes.get_xaxis().set_visible(False)
+# sub_axes.get_yaxis().set_visible(False)
[…]

### 2019-09-03 TaskCoalescingPlotting/plotter.py

@@ -1,0 +1,42 @@
+#
+# @Author: Amjad Majid
+# @Date  : March 14, 2019
+#
+import logicAnalyzerData as lad
+import numpy as np
+import matplotlib.pyplot as plt
+
+class Plotter:
+
+    def __init__(self, data):
+        self.data = data
+        self.interpolatedStates=[]
+        self.interpolatedTimestamps=[] 
+        self.__dataInterpolation()
+
+    def __dataInterpolation(self):
+        """This method interpolates the logic analyzer data to enable
+        square-like graph plotting."""
+        clustersTimestamps=[]
+        timestamps=self.data[0]
+        states=self.data[1]
+        for idx in range(len(timestamps)-1):
+            self.interpolatedTimestamps.append(timestamps[idx])
+            self.interpolatedTimestamps.append(timestamps[idx+1])
+            self.interpolatedStates.append(states[idx])
+            self.interpolatedStates.append(states[idx])
+        self.interpolatedTimestamps.append(timestamps[-1])
+        self.interpolatedStates.append(states[-1])
+
+    def getData(self):
+        return np.array(self.interpolatedTimestamps), np.array(self.interpolatedStates)
+
+    def plotOnTime(self, lineWidth=1, posY=0, scaleY=1):
+        y = np.array(self.interpolatedStates) * scaleY+ posY
+        # print(y[:20])
+        # print(y[:10])
+        # plt.figure()
+        #print(np.sum(self.interpolatedStates, axis=1))
+        plt.plot(self.interpolatedTimestamps,y, lw = lineWidth)
+
+

### 2019-09-03 apps/sort/main.c

@@ -4,8 +4,8 @@
 #include "mspProfiler.h"
 #include "mspDebugger.h"
 
-#define TSK_SIZ 1
-//#define RAISE_PIN 1
+//#define TSK_SIZ 1
+#define RAISE_PIN 1
 //#define PWR_INT_SIM 1
 
 /*
@@ -69,7 +69,6 @@ void task_inner_loop()
 
     if( RP(j) < (arr_len-1) ) os_jump(0);
 
-
     WP( arr[RP(i)] ) = arr_i;
     WP( arr[RP(j)] ) = arr_j;
     WP(j)++;
@@ -107,11 +106,8 @@ void task_finish()
 #endif
 
 #if RAISE_PIN
-    if(protect){
     P3OUT |=BIT5;
     P3OUT &=~BIT5;
-    }
-    protect=0;
 #endif
 
     unsigned cct;
@@ -130,7 +126,7 @@ void task_finish()
 
 //cp_sendRes("\sort_100 \0");
 
-while(1);
+//while(1);
 }
 
 void init()
@@ -139,12 +135,11 @@ void init()
   // Disable the GPIO power-on default high-impedance mode to activate previously configured port settings.
   PM5CTL0 &= ~LOCKLPM5;       // Lock LPM5.
 
-#if RAISE_PIN
   P3OUT &=~BIT5;
   P3DIR |=BIT5;
-#endif
 
-#if 1
+
+#if 0
     CSCTL0_H = CSKEY >> 8;                // Unlock CS registers
     //    CSCTL1 = DCOFSEL_4 |  DCORSEL;                   // Set DCO to 16MHz
     CSCTL1 = DCOFSEL_6;                   // Set DCO to 8MHz
@@ -158,15 +153,15 @@ void init()
     cp_init();
 #endif
 
-//    uart_init();
-
-#ifdef LOG_INFO
     uart_init();
-#endif
 
-#ifdef AUTO_RST
-    mr_auto_rand_reseter(13000); // every 12 msec the MCU will be reseted
-#endif
+//#ifdef LOG_INFO
+//    uart_init();
+//#endif
+
+//#ifdef AUTO_RST
+    mr_auto_rand_reseter(20000); // every 12 msec the MCU will be reseted
+//#endif
 
 
 }

### 2019-09-03 coala/src/psched.c

@@ -1,17 +1,17 @@
 #include <psched.h>
 #include <msp430fr5969.h>
-//#include <mspDebugger.h>
+#include <mspDebugger.h>
 
 // PLEASE SELECT AN ALGORITHIN
-#define NO_COALESCING 1
+#define NO_COALESCING 0
 #define SLO_CHNG_ALGO 0
-#define FST_CHNG_ALGO 0
+#define FST_CHNG_ALGO 1
 #define FST_CHNG_TSK_AWAR_ALGO 0
 
 /***********************************
  *    Common Debugging flags
  **********************************/
-#define COAL_TSK_SIZ_SEND 0
+#define COAL_TSK_SIZ_SEND 1
 #define NUM_PWR_INTR__MEM_DUMP 0
 #define __numbPwrInt (*(uint16_t*)(0x1990))
 #define __PI_overflow (*(uint16_t*)(0x1992))
@@ -324,6 +324,8 @@ void os_scheduler()
     while (1)
     {
 
+//        //For Testing MUST Be removed
+//        __coalTaskCntr = 8;
 
 #if COAL_TSK_SIZ_SEND
         uart_sendHex8(__coalTaskCntr);

### 2019-09-13 paper/TOSN/intro.tex
update figure and add reference

@@
 \paragraph{Drawbacks of Static Task Decomposition.}
-Task-based programming and execution models require a programmer~\cite{alpaca,chain} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks.
+Task-based programming and execution models require a programmer~\cite{alpaca,chain,patoukas2018feasibility} or a compiler~\cite{baghsorkhi_cgo_2018} to statically decompose a program into a collection of tasks.
 \emph{Tasks}, top-level functions, can include arbitrary computation that should be executed despite arbitrarily-timed power failures.

### 2019-09-13 TaskCoalescingPlotting/main.py

@@ -53,8 +53,8 @@ plt.plot(coalTask_detail_x-base+shift-.245,coalTask_detail_y* .51, lw=2.5, color
 
 plt.annotate("", xytext = (0.0089, .598),  xy=(.01225,.3), 
 	arrowprops={'color':'.4','arrowstyle':"->", 'linewidth':2, "connectionstyle": "arc,angleA=-60,angleB=90,armA=0,armB=20,rad=5"}, )
-plt.annotate('static tasks', 
-             xytext = (0.0034, .56),  # text location
+plt.annotate('static task', 
+             xytext = (0.0033, .56),  # text location
              fontsize = 14,
              xy=(.01,.3),        # arrows points to 
              ha = 'left',    # horizontal alignment 
@@ -64,7 +64,7 @@ plt.annotate('static tasks',
 plt.annotate("", xytext = (0.0157, .598),  xy=(.0195,.4), 
 	arrowprops={'color':'.4', 'arrowstyle':"->", 'linewidth':2, "connectionstyle": "arc,angleA=-10,angleB=96,armA=0,armB=20,rad=5"}, )
 plt.annotate('commit', 
-             xytext = (0.012, .56),  # text location
+             xytext = (0.0116, .56),  # text location
              fontsize = 14,
              xy=(.01,.3),        # arrows points to 
              ha = 'left',    # horizontal alignment 
@@ -75,7 +75,7 @@ plt.annotate('commit',
 plt.annotate("", xytext = (0.027, .598),  xy=(.0306,.488), 
 	arrowprops={'color':'.4', 'arrowstyle':"->", 'linewidth':2, "connectionstyle": "arc,angleA=10,angleB=140,armA=0,armB=20,rad=5"}, )
 plt.annotate('coalesced task', 
-             xytext = (0.02, .56),  # text location
+             xytext = (0.019, .56),  # text location
              fontsize = 14,
              xy=(.01,.3),        # arrows points to 
              ha = 'left',    # horizontal alignment 
@@ -85,7 +85,7 @@ plt.annotate('coalesced task',
 plt.annotate("", xytext = (0.0355, .598),  xy=(.0375,.55), 
 	arrowprops={'color':".4" ,'arrowstyle':"->", 'linewidth':2}, )
 plt.annotate('lost tasks', 
-             xytext = (0.0308, .56),  # text location
+             xytext = (0.0304, .56),  # text location
              fontsize = 14,
              xy=(.01,.3),        # arrows points to 
              ha = 'left',    # horizontal alignment 

### 2019-10-01 paper/TOSN/systemdescription.tex
comma removed after i.e.

@@
 By coalescing tasks \sys \emph{amortizes commit and transition costs}, and by splitting a task, after it repeatedly failed to complete, it \emph{avoids the task non-termination problem}.
-The scheduler uses its recent execution history---i.e., it is hardware independent---as a metric to estimate energy availability and eventually to decide on the coalesced task size.
+The scheduler uses its recent execution history---i.e.
+it is hardware independent---as a metric to estimate energy availability and eventually to decide on the coalesced task size.
 Section~\ref{sec:task_adaptation} describes ATS.

### 2019-10-01 paper/TOSN/intro.tex
space from footnote removed

@@
 \paragraph{Challenges and Contributions.}
-We introduce \sys~\footnote{\sys's source code is accessible via \url{https://github.com/TUDSSL/Coala} }: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
+We introduce \sys\footnote{\sys's source code is accessible via \url{https://github.com/TUDSSL/Coala} }: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}.
 By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overhead \emph{dynamically}, meanwhile avoiding the risk of non-termination.

### 2019-10-02 paper/TOSN/evaluation.tex
revision of Table 4
remove space for footnote 2

@@
 
-\begin{table}[t]%
-
-\caption{Comparison of \sys performance running the \textit{sort} application on two different capacitor sizes. We see that \sys optimizes its coalesced task size based on the energy buffer size.
-\textbf{Coalesced task} refers to the first coalesced task after a reboot, \textbf{Exp.
-time} shows the experiment duration, the \textbf{Runs} column lists the number of complete runs of the application during the experiments.
-\textbf{Run time} is the device collective uptime needed to finish a single iteration of the \textit{sort} application.}
-\label{tab:dif_cap}
-
-\begin{minipage}{\columnwidth}
+\begin{table}
 \begin{center}
-\begin{tabular}{llllll}
+\begin{tabular}{c c c c c c}
   \toprule
-    \textbf{Cap. size} ($\mu\text{f}$)& \textbf{Exp. time} ($s$)& \textbf{on/off cycle} & \textbf{Coalesced task} ($ms$)   & \textbf{Runs} & \textbf{Run time} ($ms$) \\
+    \textbf{Cap. size} ($\mu\text{F}$)& \textbf{Exp. time} ($s$)& \textbf{on/off cycle} & \textbf{Coalesced task} ($ms$)   & \textbf{Runs} & \textbf{Run time} ($ms$) \\
         \hline
@@
 \end{center}
-\end{minipage}  
+\caption{Comparison of \sys performance running the \textit{sort} application on two different capacitor sizes. We see that \sys optimizes its coalesced task size based on the energy buffer size.
+	\textbf{Coalesced task} refers to the first coalesced task after a reboot, \textbf{Exp.
+time} shows the experiment duration, the \textbf{Runs} column lists the number of complete runs of the application during the experiments.
+\textbf{Run time} is the device collective uptime needed to finish a single iteration of the \textit{sort} application.}
+\label{tab:dif_cap}
 \end{table}%
@@
 
-Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm
-% \footnote{At distances greater than 15\,cm, energy incoming during
-% execution is negligible and the stored energy is insufficient to complete some of the static tasks.}. 
-\footnote{At distances less than 15\,cm the total amount of energy available for task execution includes significant amount of harvested-while-executing energy in addition to the stored energy.}.
+Due to its static progressing behavior, Alpaca was unable to complete the \textit{fft} benchmark on distances larger than 15\,cm\footnote{At distances less than 15\,cm the total amount of energy available for task execution includes significant amount of harvested-while-executing energy in addition to the stored energy.}.
 

### 2019-10-02 paper/TOSN/relatedwork.tex
correction of bin file (inlcuding removal of duplicate reference to Wisent

@@
 has led to the development of instrumentation, debugging, and prototyping tools for such systems
-~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,wisent,flicker}.
+~\cite{hester_sensys_2014,hester_sensys_2015,edb,capybara,stork,tan_infocom_2016,flicker}.
 
